### 2015-12-04 inst/doc/genetic.surrogates.Rmd
progress

---
title: "Are environmental and geographic effective surrogates for genetic variation in conservation planning?"
author: |
  | Jeffrey O. Hanson$^1$, Jonathan R. Rhodes$^2$, Cynthia Riginos$^2$, Hugh P. Possingham$^1$, Richard A. Fuller$^1$
  | $^1$School of Biological Sciences, The University of Queensland, Brisbane, QLD, Australia
  | $^2$School of Geography, Planning and Environmental Management, The University of Queensland, Brisbane, QLD, Australia
  | Correspondance should be addressed to jeffrey.hanson@uqconnect.edu.au
date: "`r format(Sys.time(), '%d %B %Y')`"
tags: ["conservation", "biodiversity", "AFLP", "reserve-selection"]
abstract: "Insert abstract here."
output:
  rmarkdown::pdf_document:
    toc: true
    toc_depth: 5
    keep_tex: yes
    fig_caption: yes
    includes:
      in_header: preamble-latex.tex
fontsize: 11pt
documentclass: article
bibliography: Endnote_lib.bib
csl: reference-style.csl
vignette: >
  %\VignetteIndexEntry{rapr}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, include=FALSE}
# set global cache
knitr::opts_chunk$set(cache=FALSE)

# load packages
library(genetic.surrogates)
library(data.table)
library(rgeos)
library(ggplot2)
library(grid)
library(gridExtra)
library(plyr)
library(dplyr)
library(pander)

# set pander options
panderOptions('knitr.auto.asis', FALSE)

# set default select method
select <- dplyr::select

### set parameters
## debugging parameters
bs.threshold <- 0.5
bs.threads <- 1
bs.n <- 50
bs.thin <- 1
bs.nbp <- 10
bs.pilot <- 50
bs.burn <- 10
mds.k <- 2
mds.trymax <- 2
n.spp <- 3

## analysis parameters
# bs.threshold <- 0.95
# bs.threads <- 10
# bs.n <- 5000
# bs.thin <- 10
# bs.nbp <- 20
# bs.pilot <- 5000
# bs.burn <- 50000
# mds.k <- 2
# mds.trymax <- 100

# set seed for reproducibility
set.seed(500)
```

## Introduction

## Methods

### Study area
To address the aims of this study, we obtained species distribution and genomic (AFLP) from [@r451]. This dataset was chosen because it provides genomic data for a multitude of species at a high spatial resolution (approx. $20$km$^{2} \times 22$km$^{2}$).

```{r}
## compile spatial grid data
# load grid cell centroids
grid.DF <- fread(
	'extdata/Data_Meirmans_et_al_IntrabioDiv/ReadMe.txt',
	data.table=FALSE,
	skip='cell\tLong\tLat'
) %>% rename(
		grid.longitude=Long,
		grid.latitude=Lat
) %>% mutate(
	id=seq_along(grid.latitude)
)
```

```{r}
# load in aflp data
spp.aflp.paths <- dir(
	'extdata/Data_Meirmans_et_al_IntrabioDiv',
	'^.*AFLP\\.dat$',
	full.names=TRUE
)[seq_len(n.spp)]
spp.BayeScanData.LST <- llply(
	spp.aflp.paths,
	read.BayeScanData
)
```

```{r}
## compile species occurence data
# load in data
spp.loc.paths <- dir(
	'extdata/Data_Meirmans_et_al_IntrabioDiv',
	'^.*locations\\.txt$',
	full.names=TRUE
)[seq_len(n.spp)]
spp.samples.DF <- ldply(
	seq_along(spp.loc.paths),
	.fun=function(i) {
		x <- mutate(
			fread(spp.loc.paths[i], data.table=FALSE),
			species=gsub('_locations.txt', '', basename(spp.loc.paths[i]), fixed=TRUE)
		) %>% rename(
			cell=population,
			sample.longitude=longitude,
			sample.latitude=latitude
		)
		return(x[as.numeric(spp.BayeScanData.LST[[i]]@populations),])
	}
) %>% left_join(
		grid.DF,
		by='cell'
)
# append species data to grid data.frame (wide-format)
for (i in unique(spp.samples.DF$species))
	grid.DF[[i]] <- replace(
		rep(0, nrow(grid.DF)),
		which(grid.DF$cell %in% filter(spp.samples.DF, species==i)$cell),
		1
	)
```

### Genomic data
Loci in the AFLP were classified as adaptive or neutral using BayeScan (version 2.1, using a proability threshold of `r `).

```{r}
# assign cells as populations
spp.BayeScanData.LST <- llply(
	seq_along(unique(spp.samples.DF$species)),
	function(i) {
		bd <- spp.BayeScanData.LST[[i]]
		bd@populations <- filter(spp.samples.DF, species==unique(spp.samples.DF$species)[i])$cell
		return(bd)
	}
)
# run BayeScan
spp.BayeScan.LST <- llply(
	spp.BayeScanData.LST,
	run.BayeScan,
	threshold=bs.threshold,
	threads=bs.threads,
	n=bs.n,
	thin=bs.thin,
	nbp=bs.nbp,
	pilot=bs.pilot,
	burn=bs.burn
)
# run MDS
spp.mds.LST <- llply(
	spp.BayeScan.LST,
	function(i) {
		`names<-`(llply(c('adaptive', 'neutral'), function(j) {
			if (sum(i@results@fst==j)==0)
				return(NULL)
			return(
				mds(
					i,
					metric='gower',
					type=j,
					k=mds.k,
					trymax=mds.trymax
				)
			)
		}), c('adaptive','neutral'))
	}
)
# store mds rotations for each sample
spp.samples.DF <- ldply(seq_along(unique(spp.samples.DF$species)), .fun=function(i) {
	x <- filter(spp.samples.DF, species==unique(spp.samples.DF$species)[i])
	for (j in c('adaptive', 'neutral')) {
		if (!is.null(spp.mds.LST[[i]][[j]])) {
			x <- cbind(
				x,
				`names<-`(
					as.data.frame(spp.mds.LST[[i]][[j]]$points),
					paste0(j,'_d',seq_len(mds.k))
				)
			)
		}
	}
	return(x)
})
# store mds average rotation for each grid
for (i in seq_along(unique(spp.samples.DF$species))) {
	for (j in c('adaptive', 'neutral')) {
		if(!is.null(spp.mds.LST[[i]][[j]])) {
			curr.sub <- filter(spp.samples.DF, species==unique(spp.samples.DF$species)[i])
			for (k in seq_len(mds.k)) {
				curr.vals <- tapply(
					curr.sub[[paste0(j,'_d',k)]],
					curr.sub$cell,
					FUN=mean
				)
				curr.pos <- match(names(curr.vals), grid.DF$cell)
				grid.DF[curr.pos,paste0(unique(spp.samples.DF$species)[i],'_',j,'_d',k)] <- curr.vals
			}
		}
	}
}
```
### Surrogate data

```{r}
## create spatial data
# grid data as SpatialPolygonsDataFrame
grid.PTS <- SpatialPoints(as.matrix(grid.DF[,2:3]))
grid.PLY <- grid.PTS %>%
	points2grid(tolerance=0.05) %>%
	as('SpatialPolygons')
grid.PLY <- grid.PLY[sapply(gIntersects(grid.PTS, grid.PLY, byid=TRUE, returnDense=FALSE), `[[`, 1),] %>%
	spChFIDs(
		as.character(seq_len(nrow(grid.DF)))
	) %>% 
	SpatialPolygonsDataFrame(
		data=grid.DF
	)
grid.PLY@proj4string <- wgs1984
grid.PPLY <- spTransform(grid.PLY, europeEA)
# sample data as SpatialPoints
spp.sample.PTS <- SpatialPointsDataFrame(
	coords=as.matrix(spp.samples.DF[,5:6]),
	data=spp.samples.DF,
	proj4string=wgs1984
)
spp.sample.PPTS <- spTransform(spp.sample.PTS, europeEA)
## extract geographic data
centroids.DF <- gCentroid(grid.PPLY, byid=TRUE) %>% slot('coords') %>%
	as.data.frame() %>% `names<-`(paste0('geo_d',1:2))
grid.DF <- cbind(grid.DF, centroids.DF)
## extract climatic data
# load climatic data
bioclim.STK <- stack('extdata/BioClim_variables/bioclim_pca.tif')
# extract mean for each cell for each principle component
extract.DF <- grid.PPLY %>% rasterize(bioclim.STK, field='id') %>% 
	zonal(x=bioclim.STK) %>% as.data.frame() %>% select(-1) %>%
	`names<-`(paste0('env_d',seq_len(nlayers(bioclim.STK))))
# merge with grid.DF
grid.DF <- cbind(grid.DF, extract.DF)
## update spatial objects
grid.PLY@data <- grid.DF
grid.PPLY@data <- grid.DF
```

### Prioritisations

## Results

## Discussion

## Acknowledgements
JOH is funded by an Australian Postgraduate Award (APA) scholarship. RAF has an Australian Research Council Future Fellowship. This work was supported by the Centre of Excellence for Environmental Decisions (CEED) and the Landscape Ecology and Conservation Group (LEC) at The University of Queensland.

## References

## Supporting Information
### Appendix S1: Species distributions

```{r, fig.width=9.5, fig.height=10, out.width='7.5in', out.height='8.5in', fig.cap='Species distributions. Squares represent planning units. For a given species, planning units that were found to be inhabited are denoted with bright blue.'}
## plot map of species distributions
# download basemap
data(countriesHigh)
countries.FPLY <- countriesHigh[
	countriesHigh$ADMIN %in% c(
		'Italy', 'Switzerland', 'France', 'Austria',
		'Germany', 'Slovenia', 'Croatia', 'Hungary',
		'Monaco', 'Germany'
	)
,] %>% spFortify
# fortify data
grid.FPLY <- spFortify(grid.PLY)
spp.grid.FPLY <- ldply(unique(spp.samples.DF$species), function(x) {
		z <- grid.FPLY[,c('long', 'lat', 'group', x),drop=FALSE]
		names(z)[4] <- 'presence'
		z$species <- gsub('\\_', ' ', x)
		return(z)
	}
)
# plot species data
ggplot() +
	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
		fill='grey20', color='grey80') +
	geom_polygon(data=spp.grid.FPLY, aes(x=long, y=lat, 
		group=group, fill=presence), alpha=0.8, color='grey10') +
	theme_classic() +
	guides(fill=guide_legend(title='Presence')) +
	theme(axis.ticks=element_blank(), axis.text=element_blank()) +
	coord_cartesian(
		xlim=buffered.range(grid.FPLY$long, 0.05),
		ylim=buffered.range(grid.FPLY$lat, 0.05)
	) +
	xlab('') +
	ylab('') +
	facet_wrap(~ species, ncol=4)
```

```{r, fig.width=4, fig.height=3.5, fig.cap="Species richness. Squares denote planning units. Planning units with a brighter color are inhabited by more species."}
# calculate species richness
grid.PLY$Species_richness <- grid.PLY@data %>%
	select(5:(4+n.spp)) %>% as.matrix() %>% rowSums()

# plot species richness
ggplot() +
	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
		fill='grey20', color='grey80') +
	geom_polygon(data=spFortify(grid.PLY), aes(x=long, y=lat, 
		group=group, fill=Species_richness), alpha=0.8, color='grey10') +
	guides(fill=guide_legend(title='Count (#)')) +
	theme_classic() +
	theme(axis.ticks=element_blank(), axis.text=element_blank()) +
	coord_cartesian(
		xlim=buffered.range(grid.FPLY$long, 0.05),
		ylim=buffered.range(grid.FPLY$lat, 0.05)
	) +
	xlab('') +
	ylab('') +
	ggtitle('Species richness')
```

### Appendix S2: Genomic MDS 
```{r}
knitr::kable(
	ldply(
		seq_along(unique(spp.samples.DF$species)), 
		function(i) {
			ldply(
				seq_along(spp.mds.LST[[i]]),
				function(j) {
				data.frame(
					Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
					Loci=names(spp.mds.LST[[i]])[j],
					Stress=spp.mds.LST[[i]][[j]]$stress,
					Converged=spp.mds.LST[[i]][[j]]$converged
				)
			})
		}
	),
	digits=2,
	caption='Summary of nonmetric-dimensional scaling (MDS) analyses on genetic variation for each species.'
)
```

### Appendix S3: Distribution maps of intra-specific variation
```{r}
plot.spp.mds <- function(i) {
	do.call(
		grid.arrange,
		append(
			unlist(llply(c('adaptive','neutral'), function(g) {
				llply(seq_len(mds.k), function(k) {
					ggplot() +
						geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
							fill='grey20', color='grey80') +
						geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
							group='group', fill=paste0(unique(spp.samples.DF$species)[i], '_', j, '_d',k)),
							alpha=0.8, color='grey10') +
						guides(fill=guide_legend(title=' ')) +
						theme_classic() +
						theme(axis.ticks=element_blank(), axis.text=element_blank(),
							plot.margin=unit(c(0,0,0,0),'cm')) +
						coord_cartesian(
							xlim=buffered.range(grid.FPLY$long, 0.05),
							ylim=buffered.range(grid.FPLY$lat, 0.05)
						) +
						xlab('') +
						ylab('') +
						ggtitle(paste0(g,' (',k,')'))
				})
			}),recursive=FALSE),
			list(ncol=2)
		)
	)
}
```

```{r, fig.width=6.5, fig.height=4.5, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[1]),'}. Each square represents a planning unit. The color of each planning unit panel corresponds to ordination values. Planning units with similar colors contain individiduals with similar genetic variation.')}
plot.spp.mds(1)
```

```{r, fig.width=6.5, fig.height=4.5, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[2]),'}. See Figure XX caption for conventions.')}
plot.spp.mds(2)
```

```{r, fig.width=6.5, fig.height=4.5, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[3]),'}. See Figure XX caption for conventions.')}
plot.spp.mds(3)
```

### Appendix S4: Principle components analysis on climatic variation
```{r}
## load pca summary
pca.DF <- read.table('extdata/BioClim_variables/pca.TXT', skip=80) %>% `names<-`(
	c('Principle Component', 'Eigen Value', 'Variation explained (%)',
	'Accumulative variation explained (%)')
)
## make results table showing Eigen values
knitr::kable(
	pca.DF,
	digits=2,
	caption='Summary of priciniple components analysis (PCA) on bioclimatic variation across the study area. The first two principle components (PCs) were used for subsequent analysis.'
)
```

### Appendix S5: Maps of climatic variation
```{r, width=6.5, height=4.5, fig.cap='Climatic variation. Each panel depicts variation based on a different principle component (PC). Sqaures represent planning units. The color of each planning unit denotes the average priciniple component value of pixels inside it. Planning units with more similar colors have more similar climates regimes.'}
do.call(
	grid.arrange,
		append(
		llply(grep('^env\\_.*$', names(grid.DF), value=TRUE), function(x) {
			ggplot() +
				geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
					fill='grey20', color='grey80') +
				geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
					group='group', fill=x),
					alpha=0.8, color='grey10') +
				guides(fill=guide_legend(title=' ')) +
				theme_classic() +
				theme(axis.ticks=element_blank(), axis.text=element_blank(),
					plot.margin=unit(c(0,0,0,0),'cm')) +
				coord_cartesian(
					xlim=buffered.range(grid.FPLY$long, 0.05),
					ylim=buffered.range(grid.FPLY$lat, 0.05)
				) +
				xlab('') +
				ylab('') +
				ggtitle(paste0('PC ', substr(x, nchar(x), nchar(x))))
		}),
		list(ncol=2)
	)
)
```

### 2015-12-04 R/data.R

@@ -7,11 +7,11 @@
 #' wgs1984 <- CRS('+init=epsg:4326')
 "wgs1984"
 
-#' Europe Albers Equal Area Conic CRS
+#' Europe Lambert Conformal Conic CRS
 #'
-#' This object stores the coordinate reference system (CRS) for latitude/longitude data (WGS1984).
+#' This object stores the coordinate reference system (CRS) for the Europe Lambert Conformal Conic projection.
 #'
 #' @seealso \code{\link[sp]{CRS}}
 #' @examples
-#' europeEA <- CRS('+init=esri:102013')
+#' europeEA <- CRS('+init=esri:102014')
 "europeEA"

### 2015-12-04 R/make.single.species.AttributeSpace.R

@@ -1,0 +1,38 @@
+#' Make AttributeSpace using data for one species
+#'
+#' This function creates an \code{AttributeSpace} object with multiple speciess. It
+#' has a valid \code{DemandPoints} object for a single species,
+#' and has dummy \code{DemandPoints} objects for the remaining species.
+#'
+#' @param site.data \code{data.frame} with coordinates for planning units.
+#' @param species.data \code{data.frame} with coordinates of demand points for the species.
+#' @param n.species \code{numeric} number of species to output.
+#' @param spp.pos \code{numeric} index of this species.
+#' @return \code{AttributeSpace}.
+#' @seealso \code{\link[rapr]{AttributeSpace}}, \code{\link[rapr]{AttributeSpace}}.
+#' @export
+make.single.species.AttributeSpace <- function(site.data,species.data,n.species,spp.pos) {
+	# replace NA values in site.data with -9999
+	site.data <- apply(site.data, 2, function(x) {
+		replace(x, which(is.na(x)), -9999)
+	}) %>% as.data.frame()
+	return(
+		AttributeSpace(
+			pu=SimplePoints(as.matrix(site.data)),
+			demand.points=replace(
+				replicate(
+					n.species,
+					DemandPoints(
+						points=SimplePoints(coords=matrix(-9999, ncol=ncol(site.data))),
+						weights=rep(-9999)
+					)
+				),
+				spp.pos,
+				DemandPoints(
+					points=SimplePoints(coords=as.matrix(species.data)),
+					weights=rep(1, nrow(species.data))
+				)
+			)
+		)
+	)
+}

### 2015-12-04 R/make.targets.R

@@ -1,0 +1,84 @@
+#' Make target data using attribute spaces
+#'
+#' This function takes a set of species names and 
+#' attribute spaces and generates a table with targets.
+#'
+#' @param species \code{character} speces names.
+#' @param environmental.space \code{AttributeSpace} object with environmental space data.
+#' @param geographic.space \code{AttributeSpace} object with geographic space data.
+#' @param adaptive.spaces \code{list} of \code{AttributeSpace} objects with adaptive space data for species with loci undergoing adaptation.
+#' @param neutral.spaces \code{list} of \code{AttributeSpace} objects with netrual space data for species with loci not undergoing adaptation. 
+#' @details The \code{AttributeSpace} objects in the arguments to \code{adaptive.spaces} and \code{neutral.spaces} are special, such that the species they are not associated with will have negative weights.
+#' @seealso \code{\link[rapr]{AttributeSpace}}.
+#' @export
+make.targets <- function(
+	species, 
+	environmental.space, geographic.space, 
+	adaptive.spaces, neutral.spaces,
+	amount.target=0.2, space.target=0.2
+) {
+	# amount targets
+	amount.targets.DF <- data.frame(
+		species=seq_along(species),
+		target=rep(0, length(species)),
+		proportion=rep(amount.target, length(species))
+	)
+	row.names(amount.targets.DF) <- paste0('amount_',species)
+	# surrogate.spaces
+	surrogate.targets.DF <- data.frame(
+		species=rep(seq_along(species), 2),
+		target=c(rep(1, length(species)), rep(2, length(species))),
+		proportion=rep(space.target, length(species)*2)
+	)
+	row.names(surrogate.targets.DF) <- c(paste0('env_',species), paste0('geo_',species))
+	# adaptive spaces
+	adaptive.targets.DF <- ldply(seq_along(adaptive.spaces), .fun=function(i) {
+		# get species position
+		spp.pos <- which(sapply(adaptive.spaces[[i]]@demand.points, function(y) {y@weights[1]})>0)
+		# return data.frame
+		spp.DF <- data.frame(
+			species=seq_along(species),
+			target=rep(i+2, length(species)),
+			proportion=replace(
+				rep(0, length(species)),
+				spp.pos,
+				space.target
+			),
+			name=replace(
+				paste0('nulla_',i,'_',species),
[…]

### 2015-12-08 inst/doc/genetic.surrogates.Rmd
assorted progress

@@
 # set global cache
-knitr::opts_chunk$set(cache=FALSE)
+knitr::opts_chunk$set(cache=TRUE)
 
@@
 library(gridExtra)
+library(plotrix)
 library(plyr)
 library(dplyr)
+library(tidyr)
 library(pander)
+library(multcomp)
 
@@
 ## debugging parameters
+# number of species
+n.spp <- 3
+# BayeScan parameters
 bs.threshold <- 0.5
@@
 bs.burn <- 10
+# MDS parameters
 mds.k <- 2
 mds.trymax <- 2
-n.spp <- 3
+# Gurobi parameters
+gb.Threads <- 1L
+gb.MIPGap <- 0.9
+# targets
+rapr.amount.target <- 0.2
+rapr.surrogate.target <- 0.9
+rapr.genetic.target <- 0.9
+rapr.pareto.surrogate.targets <- seq(0.01, 1, length.out=3)
 
 ## analysis parameters
+## number of species
+## n.spp <- 27
+## BayeScan parameters
 # bs.threshold <- 0.95
@@
 # bs.nbp <- 20
-# bs.pilot <- 5000
+# bs.pilot <- 50000
 # bs.burn <- 50000
+# MDS parameters
 # mds.k <- 2
 # mds.trymax <- 100
+## Gurobi parameters
+# gb.Threads <- 10
+# gb.MIPGap <- 0.05
 
@@
 ### Genomic data
-Loci in the AFLP were classified as adaptive or neutral using BayeScan (version 2.1, using a proability threshold of `r `).
+Loci in the AFLP were classified as adaptive or neutral using BayeScan (version 2.1) using a proability threshold of `r bs.threshold`).
 
@@
 
+```{r}
+# generate attribute spaces for geographic and environmental data
+surrogate.ASL <- llply(
+	list(grep('^env.*$', names(grid.DF)),grep('^geo.*$', names(grid.DF))),
+	.fun=function(x) {
+		make.multi.species.AttributeSpace(
+			site.data=grid.DF[,x,drop=FALSE],
+			species.data=grid.DF[,unique(spp.samples.DF$species),drop=FALSE]
+		)
+	}
+)
+# generate attribute spaces for genetic data
+adaptive.ASL <- llply(
+	seq_along(unique(spp.samples.DF$species)),
+	function(i) {
+		make.single.species.AttributeSpace(
+			site.data=select(grid.DF, contains(paste0(unique(spp.samples.DF$species)[i], '_adaptive'))),
+			species.data=select(filter(spp.samples.DF, species==unique(spp.samples.DF$species)[i]), contains('adaptive')),
+			spp.pos=i,
+			n.species=n_distinct(spp.samples.DF$species)
+		)
+	}
+)
+neutral.ASL <- llply(
+	seq_along(unique(spp.samples.DF$species)),
+	function(i) {
+		make.single.species.AttributeSpace(
+			site.data=select(grid.DF, contains(paste0(unique(spp.samples.DF$species)[i], '_neutral'))),
+			species.data=select(filter(spp.samples.DF, species==unique(spp.samples.DF$species)[i]), contains('neutral')),
+			spp.pos=i,
+			n.species=n_distinct(spp.samples.DF$species)
+		)
+	}
+)
+# make table with targets
+target.DF <- make.targets(
+	species=unique(spp.samples.DF$species),
+	environmental.space=as1[[1]], geographic.space=as1[[2]],
+	adaptive.spaces=adaptive.ASL, neutral.spaces=neutral.ASL,
+	amount.target=0.2, space.target=0.2
+)
+# make Rap objects
+rd <- RapData(
+	polygon=SpatialPolygons2PolySet(grid.PLY),
+	pu=data.frame(
+		cost=rep(1, nrow(grid.DF)),
+		area=rep(1, nrow(grid.DF)),
+		status=rep(0L, nrow(grid.DF))
+	),
+	species=data.frame(name=unique(spp.samples.DF$species)),
+	target=target.DF,
+	attribute.spaces=append(append(surrogate.ASL, adaptive.ASL), neutral.ASL),
+	pu.species.probabilities=ldply(
+		seq_along(unique(spp.samples.DF$species)),
+		.fun=function(i) {
+			data.frame(
+				species=i,
+				pu=which(grid.DF[[unique(spp.samples.DF$species)[i]]]==1),
+				value=1
+			)
+		}
+	),
+	boundary=calcBoundaryData(grid.PLY)
+)
+ru <- RapUnsolved(RapUnreliableOpts(), rd)
+```
+
+```{r}
+## single species analysis
+# generate RapSolved objects
+single.spp.prioritisations <- llply(
+	seq_along(unique(spp.samples.DF$species)),
+	function(x) {
+		llply(
+			list(
+				c(rapr.amount.target,0,0), c(rapr.amount.target,rapr.surrogate.target,0), 
+					c(rapr.amount.target,0,rapr.genetic.target,0)
+			), 
+			function(y) {
+				species.prioritisation(
+					x=spp.subset(ru, x),
+					amount.targets=y[1],
+					env.surrogate.targets=y[2],
+					geo.surrogate.targets=y[2],
+					adaptive.genetic.targets=y[3],
+					neutral.genetic.targets=y[3],
+					Threads=gb.Threads,
+					MIPGap=gb.MIPGap
+				)
+			}
+		)
+	}
+)
+# generate results table
+single.spp.DF <- ldply(
+	single.spp.prioritisations,
+	function(x) {
+		mutate(
+			ldply(x, extractResults),
+			Prioritisation=c('Amount', 'Surrogate', 'Genetic')
+		)
+	}
+)
+```
+
+```{r}
+## multispecies analysis
+# make prioritisations
+multi.spp.prioritisations <- llply(
+	list(
+		c(rapr.amount.target,0,0), c(rapr.amount.target,rapr.surrogate.target,0), 
+			c(rapr.amount.target,0,rapr.genetic.target,0)
+	), 
+	function(y) {
+		species.prioritisation(
+			x=ru,
+			amount.targets=y[1],
+			env.surrogate.targets=y[2],
+			geo.surrogate.targets=y[2],
+			adaptive.genetic.targets=y[3],
+			neutral.genetic.targets=y[3],
+			Threads=gb.Threads,
+			MIPGap=gb.MIPGap
+		)
+	}
+)
+# generate results table
+multi.spp.DF <- ldply(seq_along(multi.spp.prioritisations), function(i) {
+	mutate(
+		extractResults(multi.spp.prioritisations[[i]]),
+		Prioritisation=c('Amount', 'Surrogate', 'Genetic')[i]
+	)
+})
+```
+
+```{r}
+## pareto frontier analysis
+# generate prioritistions
+env.pareto.prioritisations <- llply(
+	rapr.pareto.surrogate.targets,
+	species.prioritisation,
+	x=ru,
+	amount.targets=0,
+	geo.surrogate.targets=0,
+	adaptive.genetic.targets=0,
+	neutral.genetic.targets=0
+)
+geo.pareto.prioritisations <- llply(
+	rapr.pareto.surrogate.targets,
+	species.prioritisation,
+	x=ru,
+	amount.targets=0,
+	env.surrogate.targets=0,
+	adaptive.genetic.targets=0,
+	neutral.genetic.targets=0
+)
+# extract results
+env.pareto.DF <- ldply(seq_along(env.pareto.prioritisations), function(i) {
+	mutate(
+		extractResults(env.pareto.prioritisations[[i]]),
+		Surrogate.target=rapr.pareto.surrogate.targets[i]
+	)
+})
+geo.pareto.DF <- ldply(seq_along(geo.pareto.prioritisations), function(i) {
+	mutate(
+		extractResults(geo.pareto.prioritisations[[i]]),
+		Surrogate.target=rapr.pareto.surrogate.targets[i]
+	)
+})
+```
+
 ## Results
+### Single species prioritisations
+```{r}
+## statistical analysis
+# prepare data
+single.spp.SDF <- single.spp.DF %>%
+	gather(Metric, value, amount.held:neutral.held) %>%
+	filter(Metric %in% c('adaptive.held', 'neutral.held')) %>%
+	mutate(Metric=revalue(Metric, c('adaptive.held'='Adaptive variation', 
+		'neutral.held'='Neutral variation'))) %>%
+	mutate(Prioritisation.Metric=interaction(Prioritisation,Metric)) 
+# model
+single.spp.GLM <- suppressWarnings(glm(value ~ Prioritisation * Metric,
+	family='binomial', data=single.spp.SDF))
+single.spp.AOV <- suppressWarnings(anova(single.spp.GLM))
+# post-hoc
+single.spp.GLM2 <- suppressWarnings(glm(value ~ Prioritisation.Metric,
+	family='binomial', data=single.spp.SDF))
+single.spp.MCP <- summary(
+	glht(single.spp.GLM2,
+		linfct=mcp(Prioritisation.Metric='Tukey')),
+	adjusted('bonferroni'))
+```
+
+```{r, fig.width=6, fig.height=3, fig.cap='Summary of single species prioritisations.
+Single species prioritisations were generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species.
+Data shows the performance of prioritisations generated using these three sets of targets.
+Bars denote means and standard errors.'}
+# prepare data for plotting
+single.spp.PDF <- expand.grid(
+	Prioritisation=unique(single.spp.SDF$Prioritisation),
+	Metric=unique(single.spp.SDF$Metric))
+single.spp.PDF <- cbind(single.spp.PDF,
+	as.data.frame(predict(single.spp.GLM, single.spp.PDF,
+		type='response', se.fit=TRUE))) %>%
+	mutate(lower=fit-se.fit, upper=fit+se.fit,
+		letters=toupper(cld(single.spp.MCP)$mcletters$Letters),
+		letter_pos=upper+0.05)
+# make plot
+ggplot(aes(x=Metric,y=fit,fill=Prioritisation),
+	data=single.spp.PDF) +
+	geom_bar(position=position_dodge(0.9),
+		stat='identity') +
+	geom_errorbar(
+		aes(ymin=lower,ymax=upper),
+		position=position_dodge(0.9), width=0.6) +
+	geom_text(aes(x=Metric, y=letter_pos,
+		label=letters), position=position_dodge(0.9)) +
+	scale_fill_manual(name='Prioritisation',
+		values=c('grey80','grey50','grey20')) +
+	ylab('Proportion genetic\nvariation secured (%)') +
+	xlab('') +
+	theme_classic()
+```
+
+### Multi-species prioritisations
+```{r}
+## statistical analysis
+# prepare data
+multi.spp.SDF <- multi.spp.DF %>%
+	gather(Metric, value, amount.held:neutral.held) %>%
+	filter(Metric %in% c('adaptive.held', 'neutral.held')) %>%
+	mutate(Metric=revalue(Metric, c('adaptive.held'='Adaptive variation', 
+		'neutral.held'='Neutral variation'))) %>%
+	mutate(Prioritisation.Metric=interaction(Prioritisation,Metric)) 
+# model
+multi.spp.GLM <- suppressWarnings(glm(value ~ Prioritisation * Metric,
+	family='binomial', data=multi.spp.SDF))
+multi.spp.AOV <- suppressWarnings(anova(multi.spp.GLM))
+# post-hoc
+multi.spp.GLM2 <- suppressWarnings(glm(value ~ Prioritisation.Metric,
+	family='binomial', data=multi.spp.SDF))
+multi.spp.MCP <- summary(
+	glht(multi.spp.GLM2,
+		linfct=mcp(Prioritisation.Metric='Tukey')),
+	adjusted('bonferroni'))
+```
+
+```{r, fig.width=8, fig.height=3.5, fig.cap='Multi-species prioritisations.
+Panel (a) shows the prioritisation generated for using just amount-based targets.
+Panel (b) shows the prioritisation generated using amount-based and surrogate based targets.
+Panel (c) shows the prioritisation generated using amount-based and genetic-based targets'}
+# download basemap
+data(countriesHigh)
+countries.FPLY <- countriesHigh[
+	countriesHigh$ADMIN %in% c(
+		'Italy', 'Switzerland', 'France', 'Austria',
+		'Germany', 'Slovenia', 'Croatia', 'Hungary',
+		'Monaco', 'Germany'
+	)
+,] %>% spFortify
+# prepare data for plotting
+multi.spp.grid.FPLY <- grid.PLY
+for (i in seq_along(multi.spp.prioritisations))
+	multi.spp.grid.FPLY@data[[paste0('v',i)]] <- selections(multi.spp.prioritisations[[i]])
+multi.spp.grid.FPLY <- spFortify(multi.spp.grid.FPLY)
+# make maps
+do.call(
+	grid.arrange,
+	append(
+		llply(
+			seq_along(multi.spp.prioritisations),
+			function(i) {
+				ggplot() +
+					geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
+						fill='grey20', color='grey80') +
+					geom_polygon(data=multi.spp.grid.FPLY, aes_string(x='long', y='lat', 
+						group='group', fill=paste0('v',i)),
+						alpha=0.8, color='grey10') +
+					guides(fill=guide_legend(title=' ')) +
+					theme_classic() +
+					theme(axis.ticks=element_blank(), axis.text=element_blank(),
+						plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(),
+						legend.position='none') +
+					coord_cartesian(
+						xlim=buffered.range(multi.spp.grid.FPLY$long, 0.05),
+						ylim=buffered.range(multi.spp.grid.FPLY$lat, 0.05)
+					) +
+					xlab('') +
+					ylab('') +
+					annotate('text',
+						x=min(multi.spp.grid.FPLY$long)+diff(range(multi.spp.grid.FPLY$long)*0.1),
+						y=min(multi.spp.grid.FPLY$lat)+diff(range(multi.spp.grid.FPLY$lat)*1.01),
+						label=letters[i], hjust=1, vjust=1, color='white', size=8)
+			}
+		),
+		list(nrow=1)
+	)
+)
+```
+
+```{r, fig.width=6, fig.height=3, fig.cap='Summary of multi-species prioritisations.
+Three prioritisations were generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species.
+Data shows the performance of these prioritisations based on how much genetic variation they explain.
+Bars denote means and standard errors.'}
+# prepare data for plotting
+multi.spp.PDF <- expand.grid(
+	Prioritisation=unique(multi.spp.SDF$Prioritisation),
+	Metric=unique(multi.spp.SDF$Metric))
+multi.spp.PDF <- cbind(multi.spp.PDF,
+	as.data.frame(predict(multi.spp.GLM, multi.spp.PDF,
+		type='response', se.fit=TRUE))) %>%
+	mutate(lower=fit-se.fit, upper=fit+se.fit,
+		letters=toupper(cld(multi.spp.MCP)$mcletters$Letters),
+		letter_pos=upper+0.05)
+# make plot
+ggplot(aes(x=Metric,y=fit,fill=Prioritisation),
+	data=multi.spp.PDF) +
+	geom_bar(position=position_dodge(0.9),
+		stat='identity') +
+	geom_errorbar(
+		aes(ymin=lower,ymax=upper),
+		position=position_dodge(0.9), width=0.6) +
+	geom_text(aes(x=Metric, y=letter_pos,
+		label=letters), position=position_dodge(0.9)) +
+	scale_fill_manual(name='Prioritisation',
+		values=c('grey80','grey50','grey20')) +
+	ylab('Proportion genetic\nvariation secured (%)') +
+	xlab('') +
+	theme_classic()
+```
+
+### Pareo-frontier analysis
+
+```{r, fig.width=8, fig.height=3.5, fig.cap='The relationship between surrogates and genetic variation secured in prioritisations.'}
+# make plots
+p1 <- ggplot(data=env.pareto.DF) +
+	geom_line(aes(x=Surrogate.target,y=adaptive.held,group=Species),
+		alpha=0.5) +
+	xlab('Environmental variation secured (%)') +
+	ylab('Adaptive genetic\nvariation secured (%)') +
+	theme_classic()
+p2 <- ggplot(data=geo.pareto.DF) +
+	geom_line(aes(x=Surrogate.target,y=neutral.held,group=Species),
+		alpha=0.5) +
+	xlab('Geographic variation secured (%)') +
+	ylab('Neutral genetic\nvariation secured (%)') +
+	theme_classic()
+grid.arrange(p1, p2, nrow=1)
+```
 
@@
 ## plot map of species distributions
-# download basemap
-data(countriesHigh)
-countries.FPLY <- countriesHigh[
-	countriesHigh$ADMIN %in% c(
-		'Italy', 'Switzerland', 'France', 'Austria',
-		'Germany', 'Slovenia', 'Croatia', 'Hungary',
-		'Monaco', 'Germany'
-	)
-,] %>% spFortify
 # fortify data
@@
 	guides(fill=guide_legend(title='Presence')) +
-	theme(axis.ticks=element_blank(), axis.text=element_blank()) +
+	theme(axis.ticks=element_blank(), axis.text=element_blank(),
+		axis.line=element_blank()) +
 	coord_cartesian(
@@
 	theme_classic() +
-	theme(axis.ticks=element_blank(), axis.text=element_blank()) +
+	theme(axis.ticks=element_blank(), axis.text=element_blank(),
+		axis.line=element_blank()) +
 	coord_cartesian(
@@
 
-### Appendix S2: Genomic MDS 
+### Appendix S2: BayeScan Results
+
 ```{r}
 knitr::kable(
+	format.table(
+		ldply(
+			seq_along(unique(spp.samples.DF$species)), 
+			function(i) {
+				data.frame(
+					Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
+					Individuals=nrow(spp.BayeScanData.LST[[i]]@matrix),
+					Total_Loci=ncol(spp.BayeScanData.LST[[i]]@matrix),
+					Neutral_Loci=sum(spp.BayeScan.LST[[i]]@results@fst[[7]]=='neutral'),
+					Adpative_Loci=sum(spp.BayeScan.LST[[i]]@results@fst[[7]]=='adaptive'),
+					Proportion_adaptive=sum(spp.BayeScan.LST[[i]]@results@fst[[7]]=='adaptive')/ nrow(spp.BayeScan.LST[[i]]@results@fst)
+				)
+			}
+		)
+	),
+	digits=2,
+	colnames=c('Species','Individuals (#)', 'Total Loci (#)', 'Neutral Loci (#)', 'Adaptive Loci (#)', 'Proportion Adaptive (%)'),
+	align=c('l', 'c', 'c', 'c', 'c', 'c')
+)
+```
+
+```{r}
+knitr::kable(
+	format.table(
+		ldply(
+			seq_along(unique(spp.samples.DF$species)), 
+			function(i) {
+				data.frame(
+					Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
+					Primer=spp.BayeScanData.LST[[i]]@primers,
+					Probability=spp.BayeScan.LST[[i]]@results@fst[[2]],
+					qval=spp.BayeScan.LST[[i]]@results@fst[[4]],
+					alpha=spp.BayeScan.LST[[i]]@results@fst[[5]],
+					fst=spp.BayeScan.LST[[i]]@results@fst[[6]],
+					Type=spp.BayeScan.LST[[i]]@results@fst[[7]]
+				)
+			}
+		),
+		omit='Type'
+	),
+	digits=2,
+	colnames=c('Species', 'Primer', 'Probability', 'q-value', '$\\alpha$', '$F_{ST}$', 'Type'),
+	align=c('l', 'c', 'c', 'c', 'c', 'c', 'c')
+)
+```
+
+### Appendix S3: Genomic MDS 
+```{r}
+knitr::kable(
+	format.table(
 		ldply(
@@
 		),
+		omit='Converged'
+	),
 	digits=2,
-	caption='Summary of nonmetric-dimensional scaling (MDS) analyses on genetic variation for each species.'
+	caption='Summary of non-metric multi-dimensional scaling (MDS) analyses on genetic variation for each species.',
+	colnames=c('Species', 'Loci Type', 'NMDS Stress', 'Converged'),
+	align=c('l', 'c', 'c', 'c')
 )
@@
 
-### Appendix S3: Distribution maps of intra-specific variation
+### Appendix S4: Distribution maps of intra-specific variation
 ```{r}
@@
 						theme(axis.ticks=element_blank(), axis.text=element_blank(),
-							plot.margin=unit(c(0,0,0,0),'cm')) +
+							plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank()) +
 						coord_cartesian(
@@
 
-### Appendix S4: Principle components analysis on climatic variation
+### Appendix S5: Principle components analysis on climatic variation
 ```{r}
@@
 	digits=2,
-	caption='Summary of priciniple components analysis (PCA) on bioclimatic variation across the study area. The first two principle components (PCs) were used for subsequent analysis.'
+	caption='Summary of priciniple components analysis (PCA) on bioclimatic variation across the study area. The first two principle components (PCs) were used for subsequent analysis.',
+	align=c('l', 'c', 'c', 'c')
 )
@@
 
-### Appendix S5: Maps of climatic variation
+### Appendix S6: Maps of climatic variation
 ```{r, width=6.5, height=4.5, fig.cap='Climatic variation.
@@
 				theme(axis.ticks=element_blank(), axis.text=element_blank(),
-					plot.margin=unit(c(0,0,0,0),'cm')) +
+					plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank()) +
 				coord_cartesian(
@@
 ```
+
+```{r, include=FALSE}
+# save vignette's global.env 
+save.image('vignette.rda')
+```

### 2015-12-08 R/extractResults.R

@@ -1,0 +1,37 @@
+#' Extract results from RapSolved object
+#'
+#' This function extracts results from a RapSolved object
+#'
+#' @param x \code{RapSolved} object.
+#' @export
+extractResults <- function(x) {
+	# init
+	extract.held <- function(x, spp.name, space.name) {
+		pos <- which(x@data@targets$name==paste0(space.name,'_',spp.name))
+		space.index <- x@data@targets$target[pos]
+		spp.index <- x@data@targets$species[pos]
+		return(c(space.held(x, species=spp.index, space=space.index)))
+	}
+	# extract amount held
+	amount.held.DF <- data.frame(
+		Species=x@data@species$name,
+		amount.held=c(amount.held(x))
+	)
+	# extract space held
+	space.held.MTX <- space.held(x)
+	space.held.DF <- ldply(x@data@species$name, function(y) {
+		data.frame(
+			environmental.held = extract.held(x, y, 'env'),
+			geographic.held = extract.held(x, y, 'geo'),
+			adaptive.held = extract.held(x, y, 'adaptive'),
+			neutral.held = extract.held(x, y, 'neutral')
+		)
+	})
+	# return data.frame
+	return(
+		cbind(
+			amount.held.DF,
+			space.held.DF
+		)
+	)
+}

### 2015-12-08 R/make.targets.R

@@ -21,16 +21,16 @@ make.targets <- function(
 	amount.targets.DF <- data.frame(
 		species=seq_along(species),
 		target=rep(0, length(species)),
-		proportion=rep(amount.target, length(species))
+		proportion=rep(amount.target, length(species)),
+		name=paste0('amount_',species)
 	)
-	row.names(amount.targets.DF) <- paste0('amount_',species)
 	# surrogate.spaces
 	surrogate.targets.DF <- data.frame(
 		species=rep(seq_along(species), 2),
 		target=c(rep(1, length(species)), rep(2, length(species))),
-		proportion=rep(space.target, length(species)*2)
+		proportion=rep(space.target, length(species)*2),
+		name=c(paste0('env_',species), paste0('geo_',species))
 	)
-	row.names(surrogate.targets.DF) <- c(paste0('env_',species), paste0('geo_',species))
 	# adaptive spaces
 	adaptive.targets.DF <- ldply(seq_along(adaptive.spaces), .fun=function(i) {
 		# get species position
@@ -45,7 +45,7 @@ make.targets <- function(
 				space.target
 			),
 			name=replace(
-				paste0('nulla_',i,'_',species),
+				rep('null',length(species)),
 				spp.pos,
 				paste0('adaptive_',species[spp.pos])
 			)
@@ -66,19 +66,16 @@ make.targets <- function(
 				space.target
 			),
 			name=replace(
-				paste0('nulln_',i,'_',species),
+				rep('null',length(species)),
 				spp.pos,
 				paste0('neutral_',species[spp.pos])
 			)
 		)
 		return(spp.DF)
 	})
-	# assign rownames
-	row.names(adaptive.targets.DF) <- as.character(adaptive.targets.DF$name)
-	row.names(neutral.targets.DF) <- as.character(neutral.targets.DF$name)
 	# return
-	mutate(do.call(rbind, list(
+	return(mutate(do.call(rbind, list(
 		amount.targets.DF, surrogate.targets.DF,
-		select(adaptive.targets.DF, -name), select(neutral.targets.DF, -name)
-	)), target=as.integer(target), species=as.integer(species))
+		adaptive.targets.DF, neutral.targets.DF)
+	), target=as.integer(target), species=as.integer(species)))
 }

### 2015-12-11 article/article.Rmd
complete rewrite using standard ropensci article format

--- 2015-12-08 inst/doc/genetic.surrogates.Rmd
@@
 ```{r, include=FALSE}
-# set global cache
-knitr::opts_chunk$set(cache=TRUE)
+# set wd
+setwd('..')
 
 # load packages
-library(genetic.surrogates)
-library(data.table)
-library(rgeos)
-library(ggplot2)
-library(grid)
-library(gridExtra)
-library(plotrix)
-library(plyr)
-library(dplyr)
-library(tidyr)
-library(pander)
-library(multcomp)
+source('R/analysis/0-initialization.R')
 
-# set pander options
-panderOptions('knitr.auto.asis', FALSE)
-
-# set default select method
-select <- dplyr::select
-
-### set parameters
-## debugging parameters
-# number of species
-n.spp <- 3
-# BayeScan parameters
-bs.threshold <- 0.5
-bs.threads <- 1
-bs.n <- 50
-bs.thin <- 1
-bs.nbp <- 10
-bs.pilot <- 50
-bs.burn <- 10
-# MDS parameters
-mds.k <- 2
-mds.trymax <- 2
-# Gurobi parameters
-gb.Threads <- 1L
-gb.MIPGap <- 0.9
-# targets
-rapr.amount.target <- 0.2
-rapr.surrogate.target <- 0.9
-rapr.genetic.target <- 0.9
-rapr.pareto.surrogate.targets <- seq(0.01, 1, length.out=3)
-
-## analysis parameters
-## number of species
-## n.spp <- 27
-## BayeScan parameters
-# bs.threshold <- 0.95
-# bs.threads <- 10
-# bs.n <- 5000
-# bs.thin <- 10
-# bs.nbp <- 20
-# bs.pilot <- 50000
-# bs.burn <- 50000
-# MDS parameters
-# mds.k <- 2
-# mds.trymax <- 100
-## Gurobi parameters
-# gb.Threads <- 10
-# gb.MIPGap <- 0.05
-
-# set seed for reproducibility
-set.seed(500)
+# load .rda files
+load('results/results.rda')
 ```
@@
 ## Methods
+### Study area
+To address the aims of this study, we utilized species distribution and genomic data collected by the IntraBioDiv project [@r452].
+As part of this project, data for 27 Alpine plant species were collected across the European Alps using a regular grid (22.3 km \times 25 km; see @452 for methodological details).
+In every second cell, plant samples were collected from three individiduals for each species present inside the cell .
+Samples were genotyped using amplified fragment length polymorphisms (AFLP; r453).
+Matrices denoting the presence/absence of polymorphisms at loci were consutrcted independently for each species.
+This dataset was ideal for our purposes, because it contains genomic data for a multitude of species collected at a high spatial resolution.
+Additionally, the data were collected across broad-scale environmental gradients that are suitable for exploring broad-scale surrogates for genetic variation.
 
-### Study area
-To address the aims of this study, we obtained species distribution and genomic (AFLP) from [@r451].
-This dataset was chosen because it provides genomic data for a multitude of species at a high spatial resolution (approx.
-$20$km$^{2} \times 22$km$^{2}$).
+### Surrogate data
+We used explored the effectiveness of environmental and geographic surrogates.
+To describe variation in the location of each grid, we projected the grids to ESRI:XXXX and calculated the centroid of each cell.
+To describe the climatic variation among each cell, we obtained 21 bioclimatic variables (Hijmans et al.
+XXXX).
+These layers were clipped to the extent of the grid, and layers were subject to a principle components analysis.
+The first two principle components were used to generate two new layers that described broad-scale climatic variation among the layers.
+The average value of each PC component in each grid was used to characterise a different dimension of climatic variation in the grid cell.
 
-```{r}
-## compile spatial grid data
-# load grid cell centroids
-grid.DF <- fread(
-	'extdata/Data_Meirmans_et_al_IntrabioDiv/ReadMe.txt',
-	data.table=FALSE,
-	skip='cell\tLong\tLat'
-) %>% rename(
-		grid.longitude=Long,
-		grid.latitude=Lat
-) %>% mutate(
-	id=seq_along(grid.latitude)
-)
-```
+### Outlier locus detection
+The environmental and geographic surrogates are expected to have different levels of effectiveness for adaptive and neutral genetic variation.
+To investigate the effectiveness of these surrogates, we identified the loci under selection for each species.
+We used outlier locus detection methods [reviewed in @r453] to avoid circularity issues.
+The basic premise underpinning these methods is that neutral loci are expected to have a specific level of variation among populations ($F_{st}$), and loci that deviate from this expectation are under selection.
+Previously, XXXX _et al._ (XXXX) utilised this dataset and applied such detection methods to _Getiana nivalis_.
+Here, we applied their methods to each of the 27 species in the dataset.
 
-```{r}
-# load in aflp data
-spp.aflp.paths <- dir(
-	'extdata/Data_Meirmans_et_al_IntrabioDiv',
-	'^.*AFLP\\.dat$',
-	full.names=TRUE
-)[seq_len(n.spp)]
-spp.BayeScanData.LST <- llply(
-	spp.aflp.paths,
-	read.BayeScanData
-)
-```
+Breifly, we grouped individuals into genetic lineages (Figure XXX).
+The purposes of this step was to avoid confounding effects of genetic history between populations when identifying loci under selection.
+We used the Bayesian clustering method implemented in STRUCTURE (version XXXX) assuming admixture and an independent alleles model.
+For each species, we identified the most plausible number of populations (K = 1--10; five replicates for each K; model with lowest mean posterior negative log likelihood), and estimated the probability that each individual belonged to each population (200,000 burn-in; 1,000,000 iterations).
+Individuals with $\geq 0.75$ probability of belonging to a single population were assigned accordingly, and used for subsequent analysis (see Appendix XXX for STRUCTURE results for each species).
 
-```{r}
-## compile species occurence data
-# load in data
-spp.loc.paths <- dir(
-	'extdata/Data_Meirmans_et_al_IntrabioDiv',
-	'^.*locations\\.txt$',
-	full.names=TRUE
-)[seq_len(n.spp)]
-spp.samples.DF <- ldply(
-	seq_along(spp.loc.paths),
-	.fun=function(i) {
-		x <- mutate(
-			fread(spp.loc.paths[i], data.table=FALSE),
-			species=gsub('_locations.txt', '', basename(spp.loc.paths[i]), fixed=TRUE)
-		) %>% rename(
-			cell=population,
-			sample.longitude=longitude,
-			sample.latitude=latitude
-		)
-		return(x[as.numeric(spp.BayeScanData.LST[[i]]@populations),])
-	}
-) %>% left_join(
-		grid.DF,
-		by='cell'
-)
-# append species data to grid data.frame (wide-format)
-for (i in unique(spp.samples.DF$species))
-	grid.DF[[i]] <- replace(
-		rep(0, nrow(grid.DF)),
-		which(grid.DF$cell %in% filter(spp.samples.DF, species==i)$cell),
-		1
-	)
-```
+We used BAYESCAN [@rXX] to identify loci under selection for each species (Figure XXX).
+For each species, BAYESCAN was run using the population memberships identied in the STRUCTURE analysis (using 1:1 prior odds, `r bs.nbp` pilot runs, and `r bs.n` iterations thinned by `r bs.thin` iterations with a `bs.burn` burn-in).
+Following recomendations in the BAYESCAN manual, we omitted loci where the global frequency of the minor allele was <0.05 to avoid false-positives.
+To ensure convergence of results, we ran four replicates of BAYESCAN per species using a false discovery rate (FDR) < 0.1 (see Appendix XXX for BAYESCAN results for each species).
 
-### Genomic data
-Loci in the AFLP were classified as adaptive or neutral using BayeScan (version 2.1) using a proability threshold of `r bs.threshold`).
-
-```{r}
-# assign cells as populations
-spp.BayeScanData.LST <- llply(
-	seq_along(unique(spp.samples.DF$species)),
-	function(i) {
-		bd <- spp.BayeScanData.LST[[i]]
-		bd@populations <- filter(spp.samples.DF, species==unique(spp.samples.DF$species)[i])$cell
-		return(bd)
-	}
-)
-# run BayeScan
-spp.BayeScan.LST <- llply(
-	spp.BayeScanData.LST,
-	run.BayeScan,
-	threshold=bs.threshold,
-	threads=bs.threads,
-	n=bs.n,
-	thin=bs.thin,
-	nbp=bs.nbp,
-	pilot=bs.pilot,
-	burn=bs.burn
-)
-# run MDS
-spp.mds.LST <- llply(
-	spp.BayeScan.LST,
-	function(i) {
-		`names<-`(llply(c('adaptive', 'neutral'), function(j) {
-			if (sum(i@results@fst==j)==0)
-				return(NULL)
-			return(
-				mds(
-					i,
-					metric='gower',
-					type=j,
-					k=mds.k,
-					trymax=mds.trymax
-				)
-			)
-		}), c('adaptive','neutral'))
-	}
-)
-# store mds rotations for each sample
-spp.samples.DF <- ldply(seq_along(unique(spp.samples.DF$species)), .fun=function(i) {
-	x <- filter(spp.samples.DF, species==unique(spp.samples.DF$species)[i])
-	for (j in c('adaptive', 'neutral')) {
-		if (!is.null(spp.mds.LST[[i]][[j]])) {
-			x <- cbind(
-				x,
-				`names<-`(
-					as.data.frame(spp.mds.LST[[i]][[j]]$points),
-					paste0(j,'_d',seq_len(mds.k))
-				)
-			)
-		}
-	}
-	return(x)
-})
-# store mds average rotation for each grid
-for (i in seq_along(unique(spp.samples.DF$species))) {
-	for (j in c('adaptive', 'neutral')) {
-		if(!is.null(spp.mds.LST[[i]][[j]])) {
-			curr.sub <- filter(spp.samples.DF, species==unique(spp.samples.DF$species)[i])
-			for (k in seq_len(mds.k)) {
-				curr.vals <- tapply(
-					curr.sub[[paste0(j,'_d',k)]],
-					curr.sub$cell,
-					FUN=mean
-				)
-				curr.pos <- match(names(curr.vals), grid.DF$cell)
-				grid.DF[curr.pos,paste0(unique(spp.samples.DF$species)[i],'_',j,'_d',k)] <- curr.vals
-			}
-		}
-	}
-}
-```
-### Surrogate data
-
-```{r}
-## create spatial data
-# grid data as SpatialPolygonsDataFrame
-grid.PTS <- SpatialPoints(as.matrix(grid.DF[,2:3]))
-grid.PLY <- grid.PTS %>%
-	points2grid(tolerance=0.05) %>%
-	as('SpatialPolygons')
-grid.PLY <- grid.PLY[sapply(gIntersects(grid.PTS, grid.PLY, byid=TRUE, returnDense=FALSE), `[[`, 1),] %>%
-	spChFIDs(
-		as.character(seq_len(nrow(grid.DF)))
-	) %>% 
-	SpatialPolygonsDataFrame(
-		data=grid.DF
-	)
-grid.PLY@proj4string <- wgs1984
-grid.PPLY <- spTransform(grid.PLY, europeEA)
-# sample data as SpatialPoints
-spp.sample.PTS <- SpatialPointsDataFrame(
-	coords=as.matrix(spp.samples.DF[,5:6]),
-	data=spp.samples.DF,
-	proj4string=wgs1984
-)
-spp.sample.PPTS <- spTransform(spp.sample.PTS, europeEA)
-## extract geographic data
-centroids.DF <- gCentroid(grid.PPLY, byid=TRUE) %>% slot('coords') %>%
-	as.data.frame() %>% `names<-`(paste0('geo_d',1:2))
-grid.DF <- cbind(grid.DF, centroids.DF)
-## extract climatic data
-# load climatic data
-bioclim.STK <- stack('extdata/BioClim_variables/bioclim_pca.tif')
-# extract mean for each cell for each principle component
-extract.DF <- grid.PPLY %>% rasterize(bioclim.STK, field='id') %>% 
-	zonal(x=bioclim.STK) %>% as.data.frame() %>% select(-1) %>%
-	`names<-`(paste0('env_d',seq_len(nlayers(bioclim.STK))))
-# merge with grid.DF
-grid.DF <- cbind(grid.DF, extract.DF)
-## update spatial objects
-grid.PLY@data <- grid.DF
-grid.PPLY@data <- grid.DF
-```
+After classifying the loci as adaptive or neutral for each species, we used non-metric multi-dimensional scaling (NMDS) to derive new variables to describe the main themes of variation in the adaptive and/or neutral loci for each species.
+Specifically, we calculated Gower distances between the adaptive and/or neutral polymorphisms using the cluster R package (XXXX) to accomodate missing data.
+These distance matrices were then subject to XXXX as implemented in the vegan R package (XXX; K=2, 100 runs).
+The resulting ordinations were then associated with the grid cells such that each grid cell where as species was detected with individuals that were assigned to a population was associated a coordinate in adaptive and neutral space.
 
@@
 
-```{r}
-# generate attribute spaces for geographic and environmental data
-surrogate.ASL <- llply(
-	list(grep('^env.*$', names(grid.DF)),grep('^geo.*$', names(grid.DF))),
-	.fun=function(x) {
-		make.multi.species.AttributeSpace(
-			site.data=grid.DF[,x,drop=FALSE],
-			species.data=grid.DF[,unique(spp.samples.DF$species),drop=FALSE]
-		)
-	}
-)
-# generate attribute spaces for genetic data
-adaptive.ASL <- llply(
-	seq_along(unique(spp.samples.DF$species)),
-	function(i) {
-		make.single.species.AttributeSpace(
-			site.data=select(grid.DF, contains(paste0(unique(spp.samples.DF$species)[i], '_adaptive'))),
-			species.data=select(filter(spp.samples.DF, species==unique(spp.samples.DF$species)[i]), contains('adaptive')),
-			spp.pos=i,
-			n.species=n_distinct(spp.samples.DF$species)
-		)
-	}
-)
-neutral.ASL <- llply(
-	seq_along(unique(spp.samples.DF$species)),
-	function(i) {
-		make.single.species.AttributeSpace(
-			site.data=select(grid.DF, contains(paste0(unique(spp.samples.DF$species)[i], '_neutral'))),
-			species.data=select(filter(spp.samples.DF, species==unique(spp.samples.DF$species)[i]), contains('neutral')),
-			spp.pos=i,
-			n.species=n_distinct(spp.samples.DF$species)
-		)
-	}
-)
-# make table with targets
-target.DF <- make.targets(
-	species=unique(spp.samples.DF$species),
-	environmental.space=as1[[1]], geographic.space=as1[[2]],
-	adaptive.spaces=adaptive.ASL, neutral.spaces=neutral.ASL,
-	amount.target=0.2, space.target=0.2
-)
-# make Rap objects
-rd <- RapData(
-	polygon=SpatialPolygons2PolySet(grid.PLY),
-	pu=data.frame(
-		cost=rep(1, nrow(grid.DF)),
-		area=rep(1, nrow(grid.DF)),
-		status=rep(0L, nrow(grid.DF))
-	),
-	species=data.frame(name=unique(spp.samples.DF$species)),
-	target=target.DF,
-	attribute.spaces=append(append(surrogate.ASL, adaptive.ASL), neutral.ASL),
-	pu.species.probabilities=ldply(
-		seq_along(unique(spp.samples.DF$species)),
-		.fun=function(i) {
-			data.frame(
-				species=i,
-				pu=which(grid.DF[[unique(spp.samples.DF$species)[i]]]==1),
-				value=1
-			)
-		}
-	),
-	boundary=calcBoundaryData(grid.PLY)
-)
-ru <- RapUnsolved(RapUnreliableOpts(), rd)
-```
+## Results
 
-```{r}
-## single species analysis
-# generate RapSolved objects
-single.spp.prioritisations <- llply(
-	seq_along(unique(spp.samples.DF$species)),
-	function(x) {
-		llply(
-			list(
-				c(rapr.amount.target,0,0), c(rapr.amount.target,rapr.surrogate.target,0), 
-					c(rapr.amount.target,0,rapr.genetic.target,0)
-			), 
-			function(y) {
-				species.prioritisation(
-					x=spp.subset(ru, x),
-					amount.targets=y[1],
-					env.surrogate.targets=y[2],
-					geo.surrogate.targets=y[2],
-					adaptive.genetic.targets=y[3],
-					neutral.genetic.targets=y[3],
-					Threads=gb.Threads,
-					MIPGap=gb.MIPGap
-				)
-			}
-		)
-	}
-)
-# generate results table
-single.spp.DF <- ldply(
-	single.spp.prioritisations,
-	function(x) {
-		mutate(
-			ldply(x, extractResults),
-			Prioritisation=c('Amount', 'Surrogate', 'Genetic')
-		)
-	}
-)
-```
-
-```{r}
-## multispecies analysis
-# make prioritisations
-multi.spp.prioritisations <- llply(
-	list(
-		c(rapr.amount.target,0,0), c(rapr.amount.target,rapr.surrogate.target,0), 
-			c(rapr.amount.target,0,rapr.genetic.target,0)
-	), 
-	function(y) {
-		species.prioritisation(
-			x=ru,
-			amount.targets=y[1],
-			env.surrogate.targets=y[2],
-			geo.surrogate.targets=y[2],
-			adaptive.genetic.targets=y[3],
-			neutral.genetic.targets=y[3],
-			Threads=gb.Threads,
-			MIPGap=gb.MIPGap
-		)
-	}
-)
-# generate results table
-multi.spp.DF <- ldply(seq_along(multi.spp.prioritisations), function(i) {
-	mutate(
-		extractResults(multi.spp.prioritisations[[i]]),
-		Prioritisation=c('Amount', 'Surrogate', 'Genetic')[i]
-	)
-})
-```
-
-```{r}
-## pareto frontier analysis
-# generate prioritistions
-env.pareto.prioritisations <- llply(
-	rapr.pareto.surrogate.targets,
-	species.prioritisation,
-	x=ru,
-	amount.targets=0,
-	geo.surrogate.targets=0,
-	adaptive.genetic.targets=0,
-	neutral.genetic.targets=0
-)
-geo.pareto.prioritisations <- llply(
-	rapr.pareto.surrogate.targets,
-	species.prioritisation,
-	x=ru,
-	amount.targets=0,
-	env.surrogate.targets=0,
-	adaptive.genetic.targets=0,
-	neutral.genetic.targets=0
-)
-# extract results
-env.pareto.DF <- ldply(seq_along(env.pareto.prioritisations), function(i) {
-	mutate(
-		extractResults(env.pareto.prioritisations[[i]]),
-		Surrogate.target=rapr.pareto.surrogate.targets[i]
-	)
-})
-geo.pareto.DF <- ldply(seq_along(geo.pareto.prioritisations), function(i) {
-	mutate(
-		extractResults(geo.pareto.prioritisations[[i]]),
-		Surrogate.target=rapr.pareto.surrogate.targets[i]
-	)
-})
-```
-
-## Results
 ### Single species prioritisations
-```{r}
-## statistical analysis
-# prepare data
-single.spp.SDF <- single.spp.DF %>%
-	gather(Metric, value, amount.held:neutral.held) %>%
-	filter(Metric %in% c('adaptive.held', 'neutral.held')) %>%
-	mutate(Metric=revalue(Metric, c('adaptive.held'='Adaptive variation', 
-		'neutral.held'='Neutral variation'))) %>%
-	mutate(Prioritisation.Metric=interaction(Prioritisation,Metric)) 
-# model
-single.spp.GLM <- suppressWarnings(glm(value ~ Prioritisation * Metric,
-	family='binomial', data=single.spp.SDF))
-single.spp.AOV <- suppressWarnings(anova(single.spp.GLM))
-# post-hoc
-single.spp.GLM2 <- suppressWarnings(glm(value ~ Prioritisation.Metric,
-	family='binomial', data=single.spp.SDF))
-single.spp.MCP <- summary(
-	glht(single.spp.GLM2,
-		linfct=mcp(Prioritisation.Metric='Tukey')),
-	adjusted('bonferroni'))
-```
 
@@
 ### Multi-species prioritisations
-```{r}
-## statistical analysis
-# prepare data
-multi.spp.SDF <- multi.spp.DF %>%
-	gather(Metric, value, amount.held:neutral.held) %>%
-	filter(Metric %in% c('adaptive.held', 'neutral.held')) %>%
-	mutate(Metric=revalue(Metric, c('adaptive.held'='Adaptive variation', 
-		'neutral.held'='Neutral variation'))) %>%
-	mutate(Prioritisation.Metric=interaction(Prioritisation,Metric)) 
-# model
-multi.spp.GLM <- suppressWarnings(glm(value ~ Prioritisation * Metric,
-	family='binomial', data=multi.spp.SDF))
-multi.spp.AOV <- suppressWarnings(anova(multi.spp.GLM))
-# post-hoc
-multi.spp.GLM2 <- suppressWarnings(glm(value ~ Prioritisation.Metric,
-	family='binomial', data=multi.spp.SDF))
-multi.spp.MCP <- summary(
-	glht(multi.spp.GLM2,
-		linfct=mcp(Prioritisation.Metric='Tukey')),
-	adjusted('bonferroni'))
-```
 
@@
 	digits=2,
-	colnames=c('Species','Individuals (#)', 'Total Loci (#)', 'Neutral Loci (#)', 'Adaptive Loci (#)', 'Proportion Adaptive (%)'),
+	col.names=c('Species','Individuals (#)', 'Total Loci (#)', 'Neutral Loci (#)', 'Adaptive Loci (#)', 'Proportion Adaptive (%)'),
 	align=c('l', 'c', 'c', 'c', 'c', 'c')
@@
 	digits=2,
-	colnames=c('Species', 'Primer', 'Probability', 'q-value', '$\\alpha$', '$F_{ST}$', 'Type'),
+	col.names=c('Species', 'Primer', 'Probability', 'q-value', '$\\alpha$', '$F_{ST}$', 'Type'),
 	align=c('l', 'c', 'c', 'c', 'c', 'c', 'c')
@@
 	caption='Summary of non-metric multi-dimensional scaling (MDS) analyses on genetic variation for each species.',
-	colnames=c('Species', 'Loci Type', 'NMDS Stress', 'Converged'),
+	col.names=c('Species', 'Loci Type', 'NMDS Stress', 'Converged'),
 	align=c('l', 'c', 'c', 'c')
@@
 ```{r}
-## load pca summary
-pca.DF <- read.table('extdata/BioClim_variables/pca.TXT', skip=80) %>% `names<-`(
-	c('Principle Component', 'Eigen Value', 'Variation explained (%)',
-	'Accumulative variation explained (%)')
-)
 ## make results table showing Eigen values
@@
 ```
-
-```{r, include=FALSE}
-# save vignette's global.env 
-save.image('vignette.rda')
-```

### 2015-12-11 article/preamble-latex.tex
complete rewrite using standard ropensci article format

% load packages
\usepackage{amsmath,amsfonts,float,makecell,titletoc,tocloft,titlesec}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[utf8]{inputenc}

% format captions
\usepackage[labelfont={small,bf}, labelsep=space, font={small}]{caption}

% format headings
\titleformat{\subsection}
{\normalfont\Large\bfseries}{\thesubsection}{1em}{}
\titleformat{\subsubsection}
{\normalfont\large\bfseries}{\thesubsubsection}{1em}{}
\titleformat{\paragraph}
{\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
\titleformat{\subparagraph}
{\normalfont\normalsize\itshape}{\thesubparagraph}{1em}{}

\titlespacing\paragraph{0pt}{12pt plus 4pt minus 2pt}{+0pt plus 2pt minus 2pt}
\titlespacing{\subparagraph}{0pt}{6pt plus 4pt minus 2pt}{+0pt plus 2pt minus 2pt}

% format toc
\renewcommand\cftsubsecfont{\normalfont\normalsize\bfseries}
% \renewcommand\cftsubsubsecfont{\normalfont\normalsize\bfseries}
% \renewcommand\cftparafont{\normalfont\normalsize\bfseries}
% \renewcommand\cftsubparafont{\normalfont\normalsize\itshape}

% make figures static
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
	\expandafter\origfigure\expandafter[H]
} {
	\endorigfigure
}

### 2015-12-11 R/BayeScan.R

@@ -1,189 +1,0 @@
-#' @include genetic.surrogates-internal.R misc.R BayeScanOpts.R BayeScanData.R BayeScanResults.R
-
-#' BayeScan: An S4 class to represent inputs and outputs from BayeScan
-#'
-#' This class stores input data and associated output results from the BayeScan program.
-#'
-#' @slot opts \code{BayeScanOpts} object with parameters used to run BayeScan .
-#' @slot data \code{BayeScanData} object with input data used for analysis.
-#' @slot results \code{BayeScanResults} object with results from analysis.
-#' @seealso \code{\link{BayeScan}}.
-#' @export
-setClass(
-	"BayeScan",
-	representation(
-		opts='BayeScanOpts',
-		data='BayeScanData',
-		results='BayeScanResults'
-	),
-	validity=function(object) {
-		## opts
-		# checks are internal
-		## data
-		# checks are internal
-		## results
-		# checks are internal
-		## cross-object checks
-		expect_equal(n.pop(object@data), n.pop(object@results))
-		expect_equal(n.loci(object@data), n.loci(object@results))
-		return(TRUE)
-	}
-)
-
-#' Create BayeScan object
-#'
-#' This function creates a new \code{BayeScan} object.
-#'
-#' @param opts \code{BayeScanOpts} object with parameters used to run BayeScan .
-#' @param data \code{BayeScanData} object with input data used for analysis.
-#' @param results \code{BayeScanResults} object with results from analysis.
-#' @seealso \code{\link{BayeScan-class}}, \code{\link{BayeScanData}}, \code{\link{BayeScanData}}, \code{\link{BayeScanResults}}.
-#' @export
-BayeScan<-function(opts, data, results) {
-	x<-new("BayeScan", opts=opts, data=data, results=results)
-	validObject(x, test=FALSE)
-	return(x)
-}
-
-#' @rdname n.loci
-#' @method n.loci BayeScan
-#' @export
-n.loci.BayeScan <- function(x) {
-	return(n.loci(x@data))
-}
-
-#' @rdname n.pop
-#' @method n.pop BayeScan
-#' @export
-n.pop.BayeScan <- function(x) {
-	return(n.pop(x@data))
-}
-
-#' @rdname n.samples
-#' @method n.samples BayeScan
[…]

### 2015-12-11 R/BayeScanOpts.R

@@ -1,92 +1,0 @@
-#' @include genetic.surrogates-internal.R misc.R
-
-#' BayeScanOpts: An S4 class to represent BayeScan parameters
-#'
-#' This class stores input parameters for the BayeScan program.
-#'
-#' @slot threads \code{numeric} number of threads used for computation. Defaults to 1.
-#' @slot n \code{numeric} number of outputted iterations. Defaults to 5000.
-#' @slot thin \code{numeric} thinning interval size. Defaults to 10.
-#' @slot nbp \code{numeric} number of pilot runs. Defaults to 20.
-#' @slot pilot \code{numeric} length of pilot runs. Defaults to 5000.
-#' @slot burn \code{numeric} burn-in length. Defaults to 50000.
-#' @slot threshold \code{numeric} threshold probability used to classify loci as adaptive. Defaults to 0.95.
-#' @seealso \code{\link{BayeScanOpts}}.
-#' @export
-setClass(
-	"BayeScanOpts",
-	representation(
-		threads="numeric",
-		n="numeric",
-		thin="numeric",
-		nbp="numeric",
-		pilot="numeric",
-		burn="numeric",
-		threshold="numeric"
-	),
-	prototype=list(
-		threads=1,
-		n=5000,
-		thin=10,
-		nbp=20,
-		pilot=5000,
-		burn=50000,
-		threshold=0.95
-	),
-	validity=function(object) {
-		# check that parameters are greater than zero and not NA
-		sapply(
-			slotNames(object),
-			function(x) {
-				expect_true(is.finite(slot(object, x)))
-				expect_true(slot(object, x)>0)
-				return(invisible())
-		})
-		# check that threads is not higher than the number of threads on cpu
-		expect_true(object@threads<detectCores())
-		return(TRUE)
-	}
-)
-
-#' Create BayeScanOpts object
-#'
-#' This function creates a new \code{BayeScanOpts} object.
-#'
-#' @param threads \code{numeric} number of threads used for computation. Defaults to 1.
-#' @param n \code{numeric} number of outputted iterations. Defaults to 5000.
-#' @param thin \code{numeric} thinning interval size. Defaults to 10.
-#' @param nbp \code{numeric} number of pilot runs. Defaults to 20.
[…]

### 2015-12-11 R/BayeScanResults.R

@@ -1,145 +1,0 @@
-#' @include genetic.surrogates-internal.R misc.R
-
-#' BayeScanResults: An S4 class to results from BayeScan
-#'
-#' This class stores results from the BayeScan program.
-#'
-#' @slot fst \code{data.frame}
-#' @slot mcmc \code{data.frame}
-#' @slot acceptance.rate \code{data.frame}
-#' @slot verification \code{data.frame}
-#' @seealso \code{\link{BayeScanResults}}.
-#' @export
-setClass(
-	"BayeScanResults",
-	representation(
-		fst='data.frame',
-		mcmc='data.frame',
-		acceptance.rate='data.frame',
-		verification='character'
-	),
-	validity=function(object) {
-		# fst
-		expect_is(object@fst, 'data.frame')
-		expect_equal(names(object@fst), c("loci", "prob", "log10_PO", "qval", "alpha", "fst", "type"))
-		for (i in c("loci", "prob", "log10_PO", "qval", "alpha", "fst"))
-			expect_is(object@fst[[i]], c('numeric','integer'))
-		expect_is(object@fst[['type']], 'factor')
-		# mcmc
-		expect_is(object@mcmc, 'data.frame')
-		for (i in names(object@mcmc))
-			expect_is(object@mcmc[[i]], c('numeric','integer'))
-		# acceptance.rate
-		expect_is(object@acceptance.rate, 'data.frame')
-		expect_equal(names(object@acceptance.rate), c("beta", "ances", "freq"))
-		for (i in names(object@acceptance.rate))
-			expect_is(object@acceptance.rate[[i]], c('numeric','integer'))
-		# verification
-		expect_is(object@verification, 'character')
-		expect_equal(length(object@verification), 1L)
-		return(TRUE)
-	}
-)
-
-#' Create BayeScanResults object
-#'
-#' This function creates a new \code{BayeScanResults} object.
-#'
-
-#' @seealso \code{\link{BayeScanResults-class}}.
-#' @return \code{\link{BayeScanResults}}.
-#' @export
-BayeScanResults<-function(fst, mcmc, acceptance.rate, verification) {
-	x<-new("BayeScanResults", fst=fst, mcmc=mcmc, acceptance.rate=acceptance.rate, verification=verification)
-	validObject(x, test=FALSE)
-	return(x)
-}
-
-#' @rdname n.loci
-#' @method n.loci BayeScanResults
-#' @export
[…]

### 2015-12-11 R/analysis/04-bayescan-analysis.R

@@ -1,0 +1,89 @@
+# assign cells as labels
+spp.BayeScanData.LST <- llply(
+	seq_along(unique(spp.samples.DF$species)),
+	function(i) {
+		bd <- spp.BayeScanData.LST[[i]]
+		sample.labels(bd) <- filter(spp.samples.DF, species==unique(spp.samples.DF$species)[i])$cell
+		return(bd)
+	}
+)
+
+# subset out polymorphisms with really low or really high frequency
+spp.BayeScanData.LST <- llply(
+	spp.BayeScanData.LST,
+	function(x) {
+		freqs <- colMeans(x@matrix)
+		valid.loci <- which(freqs <= 1-(bs.freq) | freqs >= bs.freq)
+		return(subset.loci(x, valid.loci))
+	}
+)
+
+
+# run BayeScan
+spp.BayeScan.LST <- llply(
+	spp.BayeScanData.LST,
+	run.BayeScan,
+	threshold=bs.threshold,
+	threads=bs.threads,
+	n=bs.n,
+	thin=bs.thin,
+	nbp=bs.nbp,
+	pilot=bs.pilot,
+	burn=bs.burn
+)
+
+# run MDS
+spp.mds.LST <- llply(
+	spp.BayeScan.LST,
+	function(i) {
+		`names<-`(llply(c('adaptive', 'neutral'), function(j) {
+			if (sum(i@results@fst==j)==0)
+				return(NULL)
+			return(
+				mds(
+					i,
+					metric='gower',
+					type=j,
+					k=mds.k,
+					trymax=mds.trymax
+				)
+			)
+		}), c('adaptive','neutral'))
+	}
+)
+
+# store mds rotations for each sample
+spp.samples.DF <- ldply(seq_along(unique(spp.samples.DF$species)), .fun=function(i) {
+	x <- filter(spp.samples.DF, species==unique(spp.samples.DF$species)[i])
+	for (j in c('adaptive', 'neutral')) {
+		if (!is.null(spp.mds.LST[[i]][[j]])) {
+			x <- cbind(
+				x,
+				`names<-`(
+					as.data.frame(spp.mds.LST[[i]][[j]]$points),
+					paste0(j,'_d',seq_len(mds.k))
+				)
+			)
+		}
+	}
+	return(x)
+})
+
+# store mds average rotation for each grid
+for (i in seq_along(unique(spp.samples.DF$species))) {
[…]

### 2015-12-11 R/analysis/08-pareto-frontier-analysis.R

@@ -1,0 +1,35 @@
+## pareto frontier analysis
+# generate prioritistions
+env.pareto.prioritisations <- llply(
+	rapr.pareto.surrogate.targets,
+	species.prioritisation,
+	x=ru,
+	amount.targets=0,
+	geo.surrogate.targets=0,
+	adaptive.genetic.targets=0,
+	neutral.genetic.targets=0
+)
+geo.pareto.prioritisations <- llply(
+	rapr.pareto.surrogate.targets,
+	species.prioritisation,
+	x=ru,
+	amount.targets=0,
+	env.surrogate.targets=0,
+	adaptive.genetic.targets=0,
+	neutral.genetic.targets=0
+)
+
+# extract results
+env.pareto.DF <- ldply(seq_along(env.pareto.prioritisations), function(i) {
+	mutate(
+		extractResults(env.pareto.prioritisations[[i]]),
+		Surrogate.target=rapr.pareto.surrogate.targets[i]
+	)
+})
+geo.pareto.DF <- ldply(seq_along(geo.pareto.prioritisations), function(i) {
+	mutate(
+		extractResults(geo.pareto.prioritisations[[i]]),
+		Surrogate.target=rapr.pareto.surrogate.targets[i]
+	)
+})
+  

### 2015-12-11 vignettes/genetic.surrogates.R

@@ -1,69 +1,0 @@
-## ---- include=FALSE------------------------------------------------------
-# set global cache
-knitr::opts_chunk$set(cache=TRUE)
-
-# load packages
-library(genetic.surrogates)
-library(data.table)
-library(rgeos)
-library(ggplot2)
-library(grid)
-library(gridExtra)
-library(plotrix)
-library(plyr)
-library(dplyr)
-library(tidyr)
-library(pander)
-library(multcomp)
-
-# set pander options
-panderOptions('knitr.auto.asis', FALSE)
-
-# set default select method
-select <- dplyr::select
-
-### set parameters
-## debugging parameters
-# number of species
-n.spp <- 3
-# BayeScan parameters
-bs.threshold <- 0.5
-bs.threads <- 1
-bs.n <- 50
-bs.thin <- 1
-bs.nbp <- 10
-bs.pilot <- 50
-bs.burn <- 10
-# MDS parameters
-mds.k <- 2
-mds.trymax <- 2
-# Gurobi parameters
-gb.Threads <- 1L
-gb.MIPGap <- 0.9
-# targets
-rapr.amount.target <- 0.2
-rapr.surrogate.target <- 0.9
-rapr.genetic.target <- 0.9
-rapr.pareto.surrogate.targets <- seq(0.01, 1, length.out=3)
-
-## analysis parameters
-## number of species
-## n.spp <- 27
-## BayeScan parameters
-# bs.threshold <- 0.95
-# bs.threads <- 10
-# bs.n <- 5000
-# bs.thin <- 10
-# bs.nbp <- 20
-# bs.pilot <- 50000
-# bs.burn <- 50000
-# MDS parameters
-# mds.k <- 2
-# mds.trymax <- 100
-## Gurobi parameters
-# gb.Threads <- 10
-# gb.MIPGap <- 0.05
-
-# set seed for reproducibility
-set.seed(500)
-

### 2015-12-13 article/article.Rmd
analysis seems to be working

@@
 # load packages
-source('R/analysis/0-initialization.R')
+source('R/analysis/00-initialization.R')
 
@@
 ### Study area
-To address the aims of this study, we utilized species distribution and genomic data collected by the IntraBioDiv project [@r452].
-As part of this project, data for 27 Alpine plant species were collected across the European Alps using a regular grid (22.3 km \times 25 km; see @452 for methodological details).
-In every second cell, plant samples were collected from three individiduals for each species present inside the cell .
-Samples were genotyped using amplified fragment length polymorphisms (AFLP; r453).
+To address the aims of this study, we utilized species distribution and genomic data collected by the IntraBioDiv project [Figure S1; @r451].
+As part of this project, data for 27 alpine plant species were collected across the European Alps using a regular grid [22.3 km $\times$ 25 km; see @r452 for methodological details].
+In every second cell, plant samples were collected from three individiduals for each species present inside the cell.
+Samples were genotyped using amplified fragment length polymorphisms [AFLP; @r453].
 Matrices denoting the presence/absence of polymorphisms at loci were consutrcted independently for each species.
-This dataset was ideal for our purposes, because it contains genomic data for a multitude of species collected at a high spatial resolution.
-Additionally, the data were collected across broad-scale environmental gradients that are suitable for exploring broad-scale surrogates for genetic variation.
+This dataset was well suited for the purposes of this study, because it provides spatially explicit genomic data for a multitude of species.
+Additionally, the study area spans across broad-scale environmental gradients--ranging from lowlands to snow-caped peaks--that have been found to correlate with genetic variation for some of these species [].
 
 ### Surrogate data
-We used explored the effectiveness of environmental and geographic surrogates.
-To describe variation in the location of each grid, we projected the grids to ESRI:XXXX and calculated the centroid of each cell.
-To describe the climatic variation among each cell, we obtained 21 bioclimatic variables (Hijmans et al.
-XXXX).
-These layers were clipped to the extent of the grid, and layers were subject to a principle components analysis.
-The first two principle components were used to generate two new layers that described broad-scale climatic variation among the layers.
-The average value of each PC component in each grid was used to characterise a different dimension of climatic variation in the grid cell.
+We explored the effectiveness of environmental and geographic surrogates.
+To describe variation in the location of each grid, we projected the grids to an equal-area coordinate system (ESRI:102014) and determined the centroid of each cell.
+To describe the climatic variation among each cell, we obtained twenty-one bioclimatic variables [@r33].
+These layers were clipped to the extent of the grid and subject to a principle components analysis (PCA; Table S1).
+The first two principle components (PCs) cumulatively explained `r round(pca.DF[[4]][2],1)`% of the variation and were used to generate two new layers.
+The average value of each PC layer in each grid was used to characterise the climatic conditions in the grid cell (Figure S2).
 
 ### Outlier locus detection
-The environmental and geographic surrogates are expected to have different levels of effectiveness for adaptive and neutral genetic variation.
-To investigate the effectiveness of these surrogates, we identified the loci under selection for each species.
-We used outlier locus detection methods [reviewed in @r453] to avoid circularity issues.
-The basic premise underpinning these methods is that neutral loci are expected to have a specific level of variation among populations ($F_{st}$), and loci that deviate from this expectation are under selection.
-Previously, XXXX _et al._ (XXXX) utilised this dataset and applied such detection methods to _Getiana nivalis_.
-Here, we applied their methods to each of the 27 species in the dataset.
+To investigate the effectiveness of the surrogates for adaptive and neutral genetic variation, we first identifed which of the sampled loci for each species for each species were under selection.
+We used an outlier locus detection method [evaluated in @r454] to avoid circularity issues.
+The basic premise underpinning this method is that neutral loci are expected to have a specific level of variation among populations ($F_{st}$), and loci that deviate from this expectation are under selection.
+Previously, Bothwell _et al._ [-@r455] applied such detection methods to one of the species (_Getiana nivalis_) in this dataset.
+Here, we applied their methods to each of the twenty-seven species in the dataset.
 
-Breifly, we grouped individuals into genetic lineages (Figure XXX).
-The purposes of this step was to avoid confounding effects of genetic history between populations when identifying loci under selection.
-We used the Bayesian clustering method implemented in STRUCTURE (version XXXX) assuming admixture and an independent alleles model.
-For each species, we identified the most plausible number of populations (K = 1--10; five replicates for each K; model with lowest mean posterior negative log likelihood), and estimated the probability that each individual belonged to each population (200,000 burn-in; 1,000,000 iterations).
-Individuals with $\geq 0.75$ probability of belonging to a single population were assigned accordingly, and used for subsequent analysis (see Appendix XXX for STRUCTURE results for each species).
+Breifly, we grouped individuals into genetic lineages (Figure S2).
+The purpose of this step was to avoid confounding variation between populations with due to genetic history with variation attributed to selection pressures.
+We used the Bayesian clustering method implemented in Structure (version 2.3.4) assuming admixture and an independent alleles model.
+For each species, we identified the most plausible number of populations (K = `r st.k[1]`--`r st.k[2]`; `r st.numruns` runs for each K; model with lowest mean posterior negative log likelihood; Figure S3), and estimated the probability that each individual belonged to each population (`r st.burnin` burn-in; `r st.numreps` iterations).
+Individuals with $\geq `r st.probthresh`$ probability of belonging to a single population were assigned accordingly (Figure S4), and used for subseqent detection methods.
 
-We used BAYESCAN [@rXX] to identify loci under selection for each species (Figure XXX).
-For each species, BAYESCAN was run using the population memberships identied in the STRUCTURE analysis (using 1:1 prior odds, `r bs.nbp` pilot runs, and `r bs.n` iterations thinned by `r bs.thin` iterations with a `bs.burn` burn-in).
-Following recomendations in the BAYESCAN manual, we omitted loci where the global frequency of the minor allele was <0.05 to avoid false-positives.
-To ensure convergence of results, we ran four replicates of BAYESCAN per species using a false discovery rate (FDR) < 0.1 (see Appendix XXX for BAYESCAN results for each species).
+We used Bayescan [version 2.1; @r456] to identify the sampled loci under selection for each species (Table S2).
+For each species, Bayescan was run using the population memberships identified using Structure (1:1 prior odds, `r bs.nbp` pilot runs, and `r bs.n` iterations thinned by `r bs.thin` iterations with a `r bs.burn` burn-in).
+Following guidelines in the Bayescan manual, we omitted loci where the global frequency of the minor allele was $\geq$ 0.05 to avoid false-positives.
+To ensure convergence, we ran `r as.english(bs.reps)` replicates per species using a false discovery rate (FDR) $\leq `r bs.fdr`$.
 
-After classifying the loci as adaptive or neutral for each species, we used non-metric multi-dimensional scaling (NMDS) to derive new variables to describe the main themes of variation in the adaptive and/or neutral loci for each species.
-Specifically, we calculated Gower distances between the adaptive and/or neutral polymorphisms using the cluster R package (XXXX) to accomodate missing data.
-These distance matrices were then subject to XXXX as implemented in the vegan R package (XXX; K=2, 100 runs).
-The resulting ordinations were then associated with the grid cells such that each grid cell where as species was detected with individuals that were assigned to a population was associated a coordinate in adaptive and neutral space.
+After classifying the loci as adaptive or neutral for each species, we mapped the main gradients of adaptive and/or neutral genetic variation for each species (Figures S4--S29).
+For each species, we used Gower distances to express the difference in adaptive and/or neutral polymorphisms between individuals [using the cluster R package; @r457].
+These distances were subject to a non metric multi-dimensional scaling analysis [implemented in the vegan R package using K=`r mds.k` and `r mds.trymax` random starts; Table S3; @r458].
+The ordinated variables were then associated with the grid cells were each species was sampled.
 
@@
 ## Results
-
 ### Single species prioritisations
@@
 
+## Figures
+
+## Tables
+
 ## Supporting Information
-### Appendix S1: Species distributions
 
+### Figure S1: Species distributions
 ```{r, fig.width=9.5, fig.height=10, out.width='7.5in', out.height='8.5in', fig.cap='Species distributions.
@@
 
-### Appendix S2: BayeScan Results
-
-```{r}
-knitr::kable(
-	format.table(
-		ldply(
-			seq_along(unique(spp.samples.DF$species)), 
-			function(i) {
-				data.frame(
-					Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
-					Individuals=nrow(spp.BayeScanData.LST[[i]]@matrix),
-					Total_Loci=ncol(spp.BayeScanData.LST[[i]]@matrix),
-					Neutral_Loci=sum(spp.BayeScan.LST[[i]]@results@fst[[7]]=='neutral'),
-					Adpative_Loci=sum(spp.BayeScan.LST[[i]]@results@fst[[7]]=='adaptive'),
-					Proportion_adaptive=sum(spp.BayeScan.LST[[i]]@results@fst[[7]]=='adaptive')/ nrow(spp.BayeScan.LST[[i]]@results@fst)
-				)
-			}
-		)
-	),
-	digits=2,
-	col.names=c('Species','Individuals (#)', 'Total Loci (#)', 'Neutral Loci (#)', 'Adaptive Loci (#)', 'Proportion Adaptive (%)'),
-	align=c('l', 'c', 'c', 'c', 'c', 'c')
-)
-```
-
-```{r}
-knitr::kable(
-	format.table(
-		ldply(
-			seq_along(unique(spp.samples.DF$species)), 
-			function(i) {
-				data.frame(
-					Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
-					Primer=spp.BayeScanData.LST[[i]]@primers,
-					Probability=spp.BayeScan.LST[[i]]@results@fst[[2]],
-					qval=spp.BayeScan.LST[[i]]@results@fst[[4]],
-					alpha=spp.BayeScan.LST[[i]]@results@fst[[5]],
-					fst=spp.BayeScan.LST[[i]]@results@fst[[6]],
-					Type=spp.BayeScan.LST[[i]]@results@fst[[7]]
-				)
-			}
-		),
-		omit='Type'
-	),
-	digits=2,
-	col.names=c('Species', 'Primer', 'Probability', 'q-value', '$\\alpha$', '$F_{ST}$', 'Type'),
-	align=c('l', 'c', 'c', 'c', 'c', 'c', 'c')
-)
-```
-
-### Appendix S3: Genomic MDS 
-```{r}
-knitr::kable(
-	format.table(
-		ldply(
-			seq_along(unique(spp.samples.DF$species)), 
-			function(i) {
-				ldply(
-					seq_along(spp.mds.LST[[i]]),
-					function(j) {
-					data.frame(
-						Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
-						Loci=names(spp.mds.LST[[i]])[j],
-						Stress=spp.mds.LST[[i]][[j]]$stress,
-						Converged=spp.mds.LST[[i]][[j]]$converged
-					)
-				})
-			}
-		),
-		omit='Converged'
-	),
-	digits=2,
-	caption='Summary of non-metric multi-dimensional scaling (MDS) analyses on genetic variation for each species.',
-	col.names=c('Species', 'Loci Type', 'NMDS Stress', 'Converged'),
-	align=c('l', 'c', 'c', 'c')
-)
-```
-
-### Appendix S4: Distribution maps of intra-specific variation
-```{r}
-plot.spp.mds <- function(i) {
-	do.call(
-		grid.arrange,
-		append(
-			unlist(llply(c('adaptive','neutral'), function(g) {
-				llply(seq_len(mds.k), function(k) {
-					ggplot() +
-						geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
-							fill='grey20', color='grey80') +
-						geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
-							group='group', fill=paste0(unique(spp.samples.DF$species)[i], '_', j, '_d',k)),
-							alpha=0.8, color='grey10') +
-						guides(fill=guide_legend(title=' ')) +
-						theme_classic() +
-						theme(axis.ticks=element_blank(), axis.text=element_blank(),
-							plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank()) +
-						coord_cartesian(
-							xlim=buffered.range(grid.FPLY$long, 0.05),
-							ylim=buffered.range(grid.FPLY$lat, 0.05)
-						) +
-						xlab('') +
-						ylab('') +
-						ggtitle(paste0(g,' (',k,')'))
-				})
-			}),recursive=FALSE),
-			list(ncol=2)
-		)
-	)
-}
-```
-
-```{r, fig.width=6.5, fig.height=4.5, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[1]),'}.
-Each square represents a planning unit.
-The color of each planning unit panel corresponds to ordination values.
-Planning units with similar colors contain individiduals with similar genetic variation.')}
-plot.spp.mds(1)
-```
-
-```{r, fig.width=6.5, fig.height=4.5, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[2]),'}.
-See Figure XX caption for conventions.')}
-plot.spp.mds(2)
-```
-
-```{r, fig.width=6.5, fig.height=4.5, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[3]),'}.
-See Figure XX caption for conventions.')}
-plot.spp.mds(3)
-```
-
-### Appendix S5: Principle components analysis on climatic variation
-```{r}
-## make results table showing Eigen values
-knitr::kable(
-	pca.DF,
-	digits=2,
-	caption='Summary of priciniple components analysis (PCA) on bioclimatic variation across the study area. The first two principle components (PCs) were used for subsequent analysis.',
-	align=c('l', 'c', 'c', 'c')
-)
-```
-
-### Appendix S6: Maps of climatic variation
+### Figure S2: Maps of climatic variation
 ```{r, width=6.5, height=4.5, fig.cap='Climatic variation.
@@
 ```
+
+### Figure S3. Number populations in each species.
+```{r}
+```
+
+### Figure S4. Distribution of populations for each species.
+```{r}
+```
+
+```{r}
+plot.spp.mds <- function(i) {
+	# define function
+	make.mds.plot <- function(j,k) {
+		ggplot() +
+			geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
+				fill='grey20', color='grey80') +
+			geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
+				group='group', fill=paste0(unique(spp.samples.DF$species)[i], '_', j, '_d',k)),
+				alpha=0.8, color='grey10') +
+			guides(fill=guide_legend(title=' ')) +
+			theme_classic() +
+			theme(axis.ticks=element_blank(), axis.text=element_blank(),
+				plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank()) +
+			coord_cartesian(
+				xlim=buffered.range(grid.FPLY$long, 0.05),
+				ylim=buffered.range(grid.FPLY$lat, 0.05)
+			) +
+			xlab('') +
+			ylab('') +
+			ggtitle(paste0(j,' (',k,')'))
+	}
+	# init
+	adapt.col <- paste0(unique(spp.samples.DF$species)[i], '_adaptive_d1')
+	neutral.col <- paste0(unique(spp.samples.DF$species)[i], '_neutral_d1')
+	plotLST <- list()
+	# make adaptive plot
+	if (adapt.col %in% names(grid.FPLY)) {
+		for (k in seq_len(mds.k))
+			plotLST <- append(plotLST, list(make.mds.plot('adaptive', k)))
+	} 
+	# make neutral plot
+	if (neutral.col %in% names(grid.FPLY)) {
+		for (k in seq_len(mds.k))
+			plotLST <- append(plotLST, list(make.mds.plot('neutral', k)))
+	} 
+	# make plot
+	do.call(
+		grid.arrange,
+		append(
+			plotLST,
+			list(ncol=6)
+		)
+	)
+}
+```
+
+### Figures S
+```{r, fig.width=6.5, fig.height=4.5, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[1]),'}.
+Each square represents a planning unit.
+The color of each planning unit panel corresponds to ordination values.
+Planning units with similar colors contain individiduals with similar genetic variation.')}
+plot.spp.mds(1)
+```
+
+```{r, fig.width=6.5, fig.height=4.5, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[2]),'}.
+See Figure XX caption for conventions.')}
+plot.spp.mds(2)
+```
+
+```{r, fig.width=6.5, fig.height=4.5, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[3]),'}.
+See Figure XX caption for conventions.')}
+plot.spp.mds(3)
+```
+
+### Table S1: Principle components analysis on climatic variation
+```{r}
+## make results table showing Eigen values
+knitr::kable(
+	pca.DF,
+	digits=2,
+	caption='Summary of principle components analysis (PCA) on bioclimatic variation across the study area. The first two principle components (PCs) were used for subsequent analysis.',
+	align=c('l', 'c', 'c', 'c')
+)
+```
+
+### Table S2: BayeScan Results
+```{r}
+knitr::kable(
+	format.table(
+		ldply(
+			seq_along(unique(spp.samples.DF$species)), 
+			function(i) {
+				data.frame(
+					Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
+					Primer=spp.BayeScan.sample.loci.subset.LST[[i]]@data@primers,
+					Probability=spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[2]],
+					qval=spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[4]],
+					alpha=spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[5]],
+					fst=spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[6]],
+					Type=spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[7]]
+				)
+			}
+		),
+		omit='Type'
+	),
+	digits=2,
+	col.names=c('Species', 'Primer', 'Probability', 'q-value', '$\\alpha$', '$F_{ST}$', 'Type'),
+	align=c('l', 'c', 'c', 'c', 'c', 'c', 'c')
+)
+```
+### Table S3: Genomic MDS results
+```{r}
+knitr::kable(
+	format.table(
+		filter(
+			ldply(
+				seq_along(unique(spp.samples.DF$species)), 
+				function(i) {
+					ldply(
+						seq_along(spp.mds.LST[[i]]),
+						function(j) {
+						data.frame(
+							Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
+							Loci=names(spp.mds.LST[[i]])[j],
+							Stress=ifelse(is.null(spp.mds.LST[[i]][[j]]), NA_real_, spp.mds.LST[[i]][[j]]$stress),
+							Converged=ifelse(is.null(spp.mds.LST[[i]][[j]]), NA_real_, spp.mds.LST[[i]][[j]]$converged)
+						)
+					})
+				}
+			),
+			!is.na(Stress)
+		),
+		omit=c('Loci','Converged')
+	),
+	digits=2,
+	caption='Summary of non-metric multi-dimensional scaling (MDS) analyses on genetic variation for each species.',
+	col.names=c('Species', 'Loci Type', 'NMDS Stress', 'Converged'),
+	align=c('l', 'c', 'c', 'c')
+)
+```

### 2015-12-13 article/article.tex
analysis seems to be working

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\ErrorTok}[1]{\textbf{{#1}}}
\newcommand{\NormalTok}[1]{{#1}}
\usepackage{longtable,booktabs}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Are environmental and geographic effective surrogates for genetic variation in conservation planning?},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{Are environmental and geographic effective surrogates for genetic
variation in conservation planning?}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Jeffrey O. Hanson$^1$, Jonathan R. Rhodes$^2$, Cynthia Riginos$^2$, Hugh
P. Possingham$^1$, Richard A. Fuller$^1$\\$^1$School of Biological
Sciences, The University of Queensland, Brisbane, QLD,
Australia\\$^2$School of Geography, Planning and Environmental
Management, The University of Queensland, Brisbane, QLD,
Australia\\Correspondance should be addressed to
\href{mailto:jeffrey.hanson@uqconnect.edu.au}{jeffrey.hanson@uqconnect.edu.au}}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{13 December 2015}

% load packages
\usepackage{amsmath,amsfonts,float,makecell,titletoc,tocloft,titlesec,natbib}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[utf8]{inputenc}

% format captions
\usepackage[labelfont={small,bf}, labelsep=space, font={small}]{caption}

% format headings
\titleformat{\subsection}
{\normalfont\Large\bfseries}{\thesubsection}{1em}{}
\titleformat{\subsubsection}
{\normalfont\large\bfseries}{\thesubsubsection}{1em}{}
\titleformat{\paragraph}
{\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
\titleformat{\subparagraph}
{\normalfont\normalsize\itshape}{\thesubparagraph}{1em}{}

\titlespacing\paragraph{0pt}{12pt plus 4pt minus 2pt}{+0pt plus 2pt minus 2pt}
\titlespacing{\subparagraph}{0pt}{6pt plus 4pt minus 2pt}{+0pt plus 2pt minus 2pt}

% format toc
\renewcommand\cftsubsecfont{\normalfont\normalsize\bfseries}
% \renewcommand\cftsubsubsecfont{\normalfont\normalsize\bfseries}
% \renewcommand\cftparafont{\normalfont\normalsize\bfseries}
% \renewcommand\cftsubparafont{\normalfont\normalsize\itshape}

% make figures static
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
	\expandafter\origfigure\expandafter[H]
} {
	\endorigfigure
}

\begin{document}

\maketitle

\begin{abstract}
Insert abstract here.
\end{abstract}

{
\hypersetup{linkcolor=black}
\setcounter{tocdepth}{5}
\tableofcontents
}
\subsection{Introduction}\label{introduction}

\subsection{Methods}\label{methods}

\subsubsection{Study area}\label{study-area}

To address the aims of this study, we utilized species distribution and
genomic data collected by the IntraBioDiv project (Figure S1; Meirmans
\emph{et al.} 2011). As part of this project, data for 27 alpine plant
species were collected across the European Alps using a regular grid
(22.3 km $\times$ 25 km; see Gugerli \emph{et al.} 2008 for
methodological details). In every second cell, plant samples were
collected from three individiduals for each species present inside the
cell. Samples were genotyped using amplified fragment length
polymorphisms (AFLP; Vos \emph{et al.} 1995). Matrices denoting the
presence/absence of polymorphisms at loci were consutrcted independently
for each species. This dataset was well suited for the purposes of this
study, because it provides spatially explicit genomic data for a
multitude of species. Additionally, the study area spans across
broad-scale environmental gradients--ranging from lowlands to snow-caped
peaks--that have been found to correlate with genetic variation for some
of these species {[}{]}.

\subsubsection{Surrogate data}\label{surrogate-data}

We explored the effectiveness of environmental and geographic
surrogates. To describe variation in the location of each grid, we
projected the grids to an equal-area coordinate system (ESRI:102014) and
determined the centroid of each cell. To describe the climatic variation
among each cell, we obtained twenty-one bioclimatic variables (Hijmans
\emph{et al.} 2005). These layers were clipped to the extent of the grid
and subject to a principle components analysis (PCA; Table S1). The
first two principle components (PCs) cumulatively explained 97.2\% of
the variation and were used to generate two new layers. The average
value of each PC layer in each grid was used to characterise the
climatic conditions in the grid cell (Figure S2).

\subsubsection{Outlier locus detection}\label{outlier-locus-detection}

To investigate the effectiveness of the surrogates for adaptive and
neutral genetic variation, we first identifed which of the sampled loci
for each species for each species were under selection. We used an
outlier locus detection method (evaluated in P{é}rez-Figueroa \emph{et
al.} 2010) to avoid circularity issues. The basic premise underpinning
this method is that neutral loci are expected to have a specific level
of variation among populations ($F_{st}$), and loci that deviate from
this expectation are under selection. Previously, Bothwell \emph{et al.}
(2013) applied such detection methods to one of the species
(\emph{Getiana nivalis}) in this dataset. Here, we applied their methods
to each of the twenty-seven species in the dataset.

Breifly, we grouped individuals into genetic lineages (Figure S2). The
purpose of this step was to avoid confounding variation between
populations with due to genetic history with variation attributed to
selection pressures. We used the Bayesian clustering method implemented
in Structure (version 2.3.4) assuming admixture and an independent
alleles model. For each species, we identified the most plausible number
of populations (K = 1--10; 4 runs for each K; model with lowest mean
posterior negative log likelihood; Figure S3), and estimated the
probability that each individual belonged to each population (10
burn-in; 10 iterations). Individuals with $\geq 0.75$ probability of
belonging to a single population were assigned accordingly (Figure S4),
and used for subseqent detection methods.

We used Bayescan (version 2.1; Foll and Gaggiotti 2008) to identify the
sampled loci under selection for each species (Table S2). For each
species, Bayescan was run using the population memberships identified
using Structure (1:1 prior odds, 3 pilot runs, and 10 iterations thinned
by 1 iterations with a 40 burn-in). Following guidelines in the Bayescan
manual, we omitted loci where the global frequency of the minor allele
was $\geq$ 0.05 to avoid false-positives. To ensure convergence, we ran
2 replicates per species using a false discovery rate (FDR) $\leq 0.5$.

After classifying the loci as adaptive or neutral for each species, we
mapped the main gradients of adaptive and/or neutral genetic variation
for each species (Figures S4--S29). For each species, we used Gower
distances to express the difference in adaptive and/or neutral
polymorphisms between individuals (using the cluster R package; Maechler
\emph{et al.} 2015). These distances were subject to a non metric
multi-dimensional scaling analysis (implemented in the vegan R package
using K=2 and 2 random starts; Table S3; Oksanen \emph{et al.} 2015).
The ordinated variables were then associated with the grid cells were
each species was sampled.

\subsubsection{Prioritisations}\label{prioritisations}

\subsection{Results}\label{results}

\subsubsection{Single species
prioritisations}\label{single-species-prioritisations}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# prepare data for plotting}
\NormalTok{single.spp.PDF <-}\StringTok{ }\KeywordTok{expand.grid}\NormalTok{(}
    \DataTypeTok{Prioritisation=}\KeywordTok{unique}\NormalTok{(single.spp.SDF$Prioritisation),}
    \DataTypeTok{Metric=}\KeywordTok{unique}\NormalTok{(single.spp.SDF$Metric))}
\NormalTok{single.spp.PDF <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(single.spp.PDF,}
    \KeywordTok{as.data.frame}\NormalTok{(}\KeywordTok{predict}\NormalTok{(single.spp.GLM, single.spp.PDF,}
        \DataTypeTok{type=}\StringTok{'response'}\NormalTok{, }\DataTypeTok{se.fit=}\OtherTok{TRUE}\NormalTok{))) %>%}
\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{lower=}\NormalTok{fit-se.fit, }\DataTypeTok{upper=}\NormalTok{fit+se.fit,}
        \DataTypeTok{letters=}\KeywordTok{toupper}\NormalTok{(}\KeywordTok{cld}\NormalTok{(single.spp.MCP)$mcletters$Letters),}
        \DataTypeTok{letter_pos=}\NormalTok{upper}\FloatTok{+0.05}\NormalTok{)}
\CommentTok{# make plot}
\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{Metric,}\DataTypeTok{y=}\NormalTok{fit,}\DataTypeTok{fill=}\NormalTok{Prioritisation),}
    \DataTypeTok{data=}\NormalTok{single.spp.PDF) +}
\StringTok{    }\KeywordTok{geom_bar}\NormalTok{(}\DataTypeTok{position=}\KeywordTok{position_dodge}\NormalTok{(}\FloatTok{0.9}\NormalTok{),}
        \DataTypeTok{stat=}\StringTok{'identity'}\NormalTok{) +}
\StringTok{    }\KeywordTok{geom_errorbar}\NormalTok{(}
        \KeywordTok{aes}\NormalTok{(}\DataTypeTok{ymin=}\NormalTok{lower,}\DataTypeTok{ymax=}\NormalTok{upper),}
        \DataTypeTok{position=}\KeywordTok{position_dodge}\NormalTok{(}\FloatTok{0.9}\NormalTok{), }\DataTypeTok{width=}\FloatTok{0.6}\NormalTok{) +}
\StringTok{    }\KeywordTok{geom_text}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{Metric, }\DataTypeTok{y=}\NormalTok{letter_pos,}
        \DataTypeTok{label=}\NormalTok{letters), }\DataTypeTok{position=}\KeywordTok{position_dodge}\NormalTok{(}\FloatTok{0.9}\NormalTok{)) +}
\StringTok{    }\KeywordTok{scale_fill_manual}\NormalTok{(}\DataTypeTok{name=}\StringTok{'Prioritisation'}\NormalTok{,}
        \DataTypeTok{values=}\KeywordTok{c}\NormalTok{(}\StringTok{'grey80'}\NormalTok{,}\StringTok{'grey50'}\NormalTok{,}\StringTok{'grey20'}\NormalTok{)) +}
\StringTok{    }\KeywordTok{ylab}\NormalTok{(}\StringTok{'Proportion genetic}\CharTok{\textbackslash{}n}\StringTok{variation secured (%)'}\NormalTok{) +}
\StringTok{    }\KeywordTok{xlab}\NormalTok{(}\StringTok{''}\NormalTok{) +}
\StringTok{    }\KeywordTok{theme_classic}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## ymax not defined: adjusting position using y instead
\end{verbatim}

\begin{figure}[htbp]
\centering
\includegraphics{article_files/figure-latex/unnamed-chunk-2-1.pdf}
\caption{Summary of single species prioritisations. Single species
prioritisations were generated using amount-based targets, amount-based
and surrogate-based targets, and amount-based and genetic-based targets
for each species. Data shows the performance of prioritisations
generated using these three sets of targets. Bars denote means and
standard errors.}
\end{figure}

\subsubsection{Multi-species
prioritisations}\label{multi-species-prioritisations}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# download basemap}
\KeywordTok{data}\NormalTok{(countriesHigh)}
\NormalTok{countries.FPLY <-}\StringTok{ }\NormalTok{countriesHigh[}
    \NormalTok{countriesHigh$ADMIN %in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}
        \StringTok{'Italy'}\NormalTok{, }\StringTok{'Switzerland'}\NormalTok{, }\StringTok{'France'}\NormalTok{, }\StringTok{'Austria'}\NormalTok{,}
        \StringTok{'Germany'}\NormalTok{, }\StringTok{'Slovenia'}\NormalTok{, }\StringTok{'Croatia'}\NormalTok{, }\StringTok{'Hungary'}\NormalTok{,}
        \StringTok{'Monaco'}\NormalTok{, }\StringTok{'Germany'}
    \NormalTok{)}
\NormalTok{,] %>%}\StringTok{ }\NormalTok{spFortify}
\CommentTok{# prepare data for plotting}
\NormalTok{multi.spp.grid.FPLY <-}\StringTok{ }\NormalTok{grid.PLY}
\NormalTok{for (i in }\KeywordTok{seq_along}\NormalTok{(multi.spp.prioritisations))}
    \NormalTok{multi.spp.grid.FPLY@data[[}\KeywordTok{paste0}\NormalTok{(}\StringTok{'v'}\NormalTok{,i)]] <-}\StringTok{ }\KeywordTok{selections}\NormalTok{(multi.spp.prioritisations[[i]])}
\NormalTok{multi.spp.grid.FPLY <-}\StringTok{ }\KeywordTok{spFortify}\NormalTok{(multi.spp.grid.FPLY)}
\CommentTok{# make maps}
\KeywordTok{do.call}\NormalTok{(}
    \NormalTok{grid.arrange,}
    \KeywordTok{append}\NormalTok{(}
        \KeywordTok{llply}\NormalTok{(}
            \KeywordTok{seq_along}\NormalTok{(multi.spp.prioritisations),}
            \NormalTok{function(i) \{}
                \KeywordTok{ggplot}\NormalTok{() +}
\StringTok{                    }\KeywordTok{geom_polygon}\NormalTok{(}\DataTypeTok{data=}\NormalTok{countries.FPLY, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{long, }\DataTypeTok{y=}\NormalTok{lat, }\DataTypeTok{group=}\NormalTok{group),}
                        \DataTypeTok{fill=}\StringTok{'grey20'}\NormalTok{, }\DataTypeTok{color=}\StringTok{'grey80'}\NormalTok{) +}
\StringTok{                    }\KeywordTok{geom_polygon}\NormalTok{(}\DataTypeTok{data=}\NormalTok{multi.spp.grid.FPLY, }\KeywordTok{aes_string}\NormalTok{(}\DataTypeTok{x=}\StringTok{'long'}\NormalTok{, }\DataTypeTok{y=}\StringTok{'lat'}\NormalTok{, }
                        \DataTypeTok{group=}\StringTok{'group'}\NormalTok{, }\DataTypeTok{fill=}\KeywordTok{paste0}\NormalTok{(}\StringTok{'v'}\NormalTok{,i)),}
                        \DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{, }\DataTypeTok{color=}\StringTok{'grey10'}\NormalTok{) +}
\StringTok{                    }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{fill=}\KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{title=}\StringTok{' '}\NormalTok{)) +}
\StringTok{                    }\KeywordTok{theme_classic}\NormalTok{() +}
\StringTok{                    }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.ticks=}\KeywordTok{element_blank}\NormalTok{(), }\DataTypeTok{axis.text=}\KeywordTok{element_blank}\NormalTok{(),}
                        \DataTypeTok{plot.margin=}\KeywordTok{unit}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{),}\StringTok{'cm'}\NormalTok{), }\DataTypeTok{axis.line=}\KeywordTok{element_blank}\NormalTok{(),}
                        \DataTypeTok{legend.position=}\StringTok{'none'}\NormalTok{) +}
\StringTok{                    }\KeywordTok{coord_cartesian}\NormalTok{(}
                        \DataTypeTok{xlim=}\KeywordTok{buffered.range}\NormalTok{(multi.spp.grid.FPLY$long, }\FloatTok{0.05}\NormalTok{),}
                        \DataTypeTok{ylim=}\KeywordTok{buffered.range}\NormalTok{(multi.spp.grid.FPLY$lat, }\FloatTok{0.05}\NormalTok{)}
                    \NormalTok{) +}
\StringTok{                    }\KeywordTok{xlab}\NormalTok{(}\StringTok{''}\NormalTok{) +}
\StringTok{                    }\KeywordTok{ylab}\NormalTok{(}\StringTok{''}\NormalTok{) +}
\StringTok{                    }\KeywordTok{annotate}\NormalTok{(}\StringTok{'text'}\NormalTok{,}
                        \DataTypeTok{x=}\KeywordTok{min}\NormalTok{(multi.spp.grid.FPLY$long)+}\KeywordTok{diff}\NormalTok{(}\KeywordTok{range}\NormalTok{(multi.spp.grid.FPLY$long)*}\FloatTok{0.1}\NormalTok{),}
                        \DataTypeTok{y=}\KeywordTok{min}\NormalTok{(multi.spp.grid.FPLY$lat)+}\KeywordTok{diff}\NormalTok{(}\KeywordTok{range}\NormalTok{(multi.spp.grid.FPLY$lat)*}\FloatTok{1.01}\NormalTok{),}
                        \DataTypeTok{label=}\NormalTok{letters[i], }\DataTypeTok{hjust=}\DecValTok{1}\NormalTok{, }\DataTypeTok{vjust=}\DecValTok{1}\NormalTok{, }\DataTypeTok{color=}\StringTok{'white'}\NormalTok{, }\DataTypeTok{size=}\DecValTok{8}\NormalTok{)}
            \NormalTok{\}}
        \NormalTok{),}
        \KeywordTok{list}\NormalTok{(}\DataTypeTok{nrow=}\DecValTok{1}\NormalTok{)}
    \NormalTok{)}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{article_files/figure-latex/unnamed-chunk-3-1.pdf}
\caption{Multi-species prioritisations. Panel (a) shows the
prioritisation generated for using just amount-based targets. Panel (b)
shows the prioritisation generated using amount-based and surrogate
based targets. Panel (c) shows the prioritisation generated using
amount-based and genetic-based targets}
\end{figure}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# prepare data for plotting}
\NormalTok{multi.spp.PDF <-}\StringTok{ }\KeywordTok{expand.grid}\NormalTok{(}
    \DataTypeTok{Prioritisation=}\KeywordTok{unique}\NormalTok{(multi.spp.SDF$Prioritisation),}
    \DataTypeTok{Metric=}\KeywordTok{unique}\NormalTok{(multi.spp.SDF$Metric))}
\NormalTok{multi.spp.PDF <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(multi.spp.PDF,}
    \KeywordTok{as.data.frame}\NormalTok{(}\KeywordTok{predict}\NormalTok{(multi.spp.GLM, multi.spp.PDF,}
        \DataTypeTok{type=}\StringTok{'response'}\NormalTok{, }\DataTypeTok{se.fit=}\OtherTok{TRUE}\NormalTok{))) %>%}
\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{lower=}\NormalTok{fit-se.fit, }\DataTypeTok{upper=}\NormalTok{fit+se.fit,}
        \DataTypeTok{letters=}\KeywordTok{toupper}\NormalTok{(}\KeywordTok{cld}\NormalTok{(multi.spp.MCP)$mcletters$Letters),}
        \DataTypeTok{letter_pos=}\NormalTok{upper}\FloatTok{+0.05}\NormalTok{)}
\CommentTok{# make plot}
\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{Metric,}\DataTypeTok{y=}\NormalTok{fit,}\DataTypeTok{fill=}\NormalTok{Prioritisation),}
    \DataTypeTok{data=}\NormalTok{multi.spp.PDF) +}
\StringTok{    }\KeywordTok{geom_bar}\NormalTok{(}\DataTypeTok{position=}\KeywordTok{position_dodge}\NormalTok{(}\FloatTok{0.9}\NormalTok{),}
        \DataTypeTok{stat=}\StringTok{'identity'}\NormalTok{) +}
\StringTok{    }\KeywordTok{geom_errorbar}\NormalTok{(}
        \KeywordTok{aes}\NormalTok{(}\DataTypeTok{ymin=}\NormalTok{lower,}\DataTypeTok{ymax=}\NormalTok{upper),}
        \DataTypeTok{position=}\KeywordTok{position_dodge}\NormalTok{(}\FloatTok{0.9}\NormalTok{), }\DataTypeTok{width=}\FloatTok{0.6}\NormalTok{) +}
\StringTok{    }\KeywordTok{geom_text}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{Metric, }\DataTypeTok{y=}\NormalTok{letter_pos,}
        \DataTypeTok{label=}\NormalTok{letters), }\DataTypeTok{position=}\KeywordTok{position_dodge}\NormalTok{(}\FloatTok{0.9}\NormalTok{)) +}
\StringTok{    }\KeywordTok{scale_fill_manual}\NormalTok{(}\DataTypeTok{name=}\StringTok{'Prioritisation'}\NormalTok{,}
        \DataTypeTok{values=}\KeywordTok{c}\NormalTok{(}\StringTok{'grey80'}\NormalTok{,}\StringTok{'grey50'}\NormalTok{,}\StringTok{'grey20'}\NormalTok{)) +}
\StringTok{    }\KeywordTok{ylab}\NormalTok{(}\StringTok{'Proportion genetic}\CharTok{\textbackslash{}n}\StringTok{variation secured (%)'}\NormalTok{) +}
\StringTok{    }\KeywordTok{xlab}\NormalTok{(}\StringTok{''}\NormalTok{) +}
\StringTok{    }\KeywordTok{theme_classic}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## ymax not defined: adjusting position using y instead
\end{verbatim}

\begin{figure}[htbp]
\centering
\includegraphics{article_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{Summary of multi-species prioritisations. Three prioritisations
were generated using amount-based targets, amount-based and
surrogate-based targets, and amount-based and genetic-based targets for
each species. Data shows the performance of these prioritisations based
on how much genetic variation they explain. Bars denote means and
standard errors.}
\end{figure}

\subsubsection{Pareo-frontier analysis}\label{pareo-frontier-analysis}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# make plots}
\NormalTok{p1 <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(}\DataTypeTok{data=}\NormalTok{env.pareto.DF) +}
\StringTok{    }\KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{Surrogate.target,}\DataTypeTok{y=}\NormalTok{adaptive.held,}\DataTypeTok{group=}\NormalTok{Species),}
        \DataTypeTok{alpha=}\FloatTok{0.5}\NormalTok{) +}
\StringTok{    }\KeywordTok{xlab}\NormalTok{(}\StringTok{'Environmental variation secured (%)'}\NormalTok{) +}
\StringTok{    }\KeywordTok{ylab}\NormalTok{(}\StringTok{'Adaptive genetic}\CharTok{\textbackslash{}n}\StringTok{variation secured (%)'}\NormalTok{) +}
\StringTok{    }\KeywordTok{theme_classic}\NormalTok{()}
\NormalTok{p2 <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(}\DataTypeTok{data=}\NormalTok{geo.pareto.DF) +}
\StringTok{    }\KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{Surrogate.target,}\DataTypeTok{y=}\NormalTok{neutral.held,}\DataTypeTok{group=}\NormalTok{Species),}
        \DataTypeTok{alpha=}\FloatTok{0.5}\NormalTok{) +}
\StringTok{    }\KeywordTok{xlab}\NormalTok{(}\StringTok{'Geographic variation secured (%)'}\NormalTok{) +}
\StringTok{    }\KeywordTok{ylab}\NormalTok{(}\StringTok{'Neutral genetic}\CharTok{\textbackslash{}n}\StringTok{variation secured (%)'}\NormalTok{) +}
\StringTok{    }\KeywordTok{theme_classic}\NormalTok{()}
\KeywordTok{grid.arrange}\NormalTok{(p1, p2, }\DataTypeTok{nrow=}\DecValTok{1}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{article_files/figure-latex/unnamed-chunk-5-1.pdf}
\caption{The relationship between surrogates and genetic variation
secured in prioritisations.}
\end{figure}

\subsection{Discussion}\label{discussion}

\subsection{Acknowledgements}\label{acknowledgements}

JOH is funded by an Australian Postgraduate Award (APA) scholarship. RAF
has an Australian Research Council Future Fellowship. This work was
supported by the Centre of Excellence for Environmental Decisions (CEED)
and the Landscape Ecology and Conservation Group (LEC) at The University
of Queensland.

\subsection{References}\label{references}

\subsection{Figures}\label{figures}

\subsection{Tables}\label{tables}

\subsection{Supporting Information}\label{supporting-information}

\subsubsection{Figure S1: Species
distributions}\label{figure-s1-species-distributions}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## plot map of species distributions}
\CommentTok{# fortify data}
\NormalTok{grid.FPLY <-}\StringTok{ }\KeywordTok{spFortify}\NormalTok{(grid.PLY)}
\NormalTok{spp.grid.FPLY <-}\StringTok{ }\KeywordTok{ldply}\NormalTok{(}\KeywordTok{unique}\NormalTok{(spp.samples.DF$species), function(x) \{}
        \NormalTok{z <-}\StringTok{ }\NormalTok{grid.FPLY[,}\KeywordTok{c}\NormalTok{(}\StringTok{'long'}\NormalTok{, }\StringTok{'lat'}\NormalTok{, }\StringTok{'group'}\NormalTok{, x),drop=}\OtherTok{FALSE}\NormalTok{]}
        \KeywordTok{names}\NormalTok{(z)[}\DecValTok{4}\NormalTok{] <-}\StringTok{ 'presence'}
        \NormalTok{z$species <-}\StringTok{ }\KeywordTok{gsub}\NormalTok{(}\StringTok{'}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{_'}\NormalTok{, }\StringTok{' '}\NormalTok{, x)}
        \KeywordTok{return}\NormalTok{(z)}
    \NormalTok{\}}
\NormalTok{)}
\CommentTok{# plot species data}
\KeywordTok{ggplot}\NormalTok{() +}
\StringTok{    }\KeywordTok{geom_polygon}\NormalTok{(}\DataTypeTok{data=}\NormalTok{countries.FPLY, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{long, }\DataTypeTok{y=}\NormalTok{lat, }\DataTypeTok{group=}\NormalTok{group),}
        \DataTypeTok{fill=}\StringTok{'grey20'}\NormalTok{, }\DataTypeTok{color=}\StringTok{'grey80'}\NormalTok{) +}
\StringTok{    }\KeywordTok{geom_polygon}\NormalTok{(}\DataTypeTok{data=}\NormalTok{spp.grid.FPLY, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{long, }\DataTypeTok{y=}\NormalTok{lat, }
        \DataTypeTok{group=}\NormalTok{group, }\DataTypeTok{fill=}\NormalTok{presence), }\DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{, }\DataTypeTok{color=}\StringTok{'grey10'}\NormalTok{) +}
\StringTok{    }\KeywordTok{theme_classic}\NormalTok{() +}
\StringTok{    }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{fill=}\KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{title=}\StringTok{'Presence'}\NormalTok{)) +}
\StringTok{    }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.ticks=}\KeywordTok{element_blank}\NormalTok{(), }\DataTypeTok{axis.text=}\KeywordTok{element_blank}\NormalTok{(),}
        \DataTypeTok{axis.line=}\KeywordTok{element_blank}\NormalTok{()) +}
\StringTok{    }\KeywordTok{coord_cartesian}\NormalTok{(}
        \DataTypeTok{xlim=}\KeywordTok{buffered.range}\NormalTok{(grid.FPLY$long, }\FloatTok{0.05}\NormalTok{),}
        \DataTypeTok{ylim=}\KeywordTok{buffered.range}\NormalTok{(grid.FPLY$lat, }\FloatTok{0.05}\NormalTok{)}
    \NormalTok{) +}
\StringTok{    }\KeywordTok{xlab}\NormalTok{(}\StringTok{''}\NormalTok{) +}
\StringTok{    }\KeywordTok{ylab}\NormalTok{(}\StringTok{''}\NormalTok{) +}
\StringTok{    }\KeywordTok{facet_wrap}\NormalTok{(~}\StringTok{ }\NormalTok{species, }\DataTypeTok{ncol=}\DecValTok{4}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}
\includegraphics[width=7.5in,height=8.5in]{article_files/figure-latex/unnamed-chunk-6-1} \caption{Species distributions. Squares represent planning units. For a given species, planning units that were found to be inhabited are denoted with bright blue.}\label{fig:unnamed-chunk-6}
\end{figure}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# calculate species richness}
\NormalTok{grid.PLY$Species_richness <-}\StringTok{ }\NormalTok{grid.PLY@data %>%}
\StringTok{    }\KeywordTok{select}\NormalTok{(}\DecValTok{5}\NormalTok{:(}\DecValTok{4}\NormalTok{+n.spp)) %>%}\StringTok{ }\KeywordTok{as.matrix}\NormalTok{() %>%}\StringTok{ }\KeywordTok{rowSums}\NormalTok{()}

\CommentTok{# plot species richness}
\KeywordTok{ggplot}\NormalTok{() +}
\StringTok{    }\KeywordTok{geom_polygon}\NormalTok{(}\DataTypeTok{data=}\NormalTok{countries.FPLY, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{long, }\DataTypeTok{y=}\NormalTok{lat, }\DataTypeTok{group=}\NormalTok{group),}
        \DataTypeTok{fill=}\StringTok{'grey20'}\NormalTok{, }\DataTypeTok{color=}\StringTok{'grey80'}\NormalTok{) +}
\StringTok{    }\KeywordTok{geom_polygon}\NormalTok{(}\DataTypeTok{data=}\KeywordTok{spFortify}\NormalTok{(grid.PLY), }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{long, }\DataTypeTok{y=}\NormalTok{lat, }
        \DataTypeTok{group=}\NormalTok{group, }\DataTypeTok{fill=}\NormalTok{Species_richness), }\DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{, }\DataTypeTok{color=}\StringTok{'grey10'}\NormalTok{) +}
\StringTok{    }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{fill=}\KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{title=}\StringTok{'Count (#)'}\NormalTok{)) +}
\StringTok{    }\KeywordTok{theme_classic}\NormalTok{() +}
\StringTok{    }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.ticks=}\KeywordTok{element_blank}\NormalTok{(), }\DataTypeTok{axis.text=}\KeywordTok{element_blank}\NormalTok{(),}
        \DataTypeTok{axis.line=}\KeywordTok{element_blank}\NormalTok{()) +}
\StringTok{    }\KeywordTok{coord_cartesian}\NormalTok{(}
        \DataTypeTok{xlim=}\KeywordTok{buffered.range}\NormalTok{(grid.FPLY$long, }\FloatTok{0.05}\NormalTok{),}
        \DataTypeTok{ylim=}\KeywordTok{buffered.range}\NormalTok{(grid.FPLY$lat, }\FloatTok{0.05}\NormalTok{)}
    \NormalTok{) +}
\StringTok{    }\KeywordTok{xlab}\NormalTok{(}\StringTok{''}\NormalTok{) +}
\StringTok{    }\KeywordTok{ylab}\NormalTok{(}\StringTok{''}\NormalTok{) +}
\StringTok{    }\KeywordTok{ggtitle}\NormalTok{(}\StringTok{'Species richness'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{article_files/figure-latex/unnamed-chunk-7-1.pdf}
\caption{Species richness. Squares denote planning units. Planning units
with a brighter color are inhabited by more species.}
\end{figure}

\subsubsection{Figure S2: Maps of climatic
variation}\label{figure-s2-maps-of-climatic-variation}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{do.call}\NormalTok{(}
    \NormalTok{grid.arrange,}
        \KeywordTok{append}\NormalTok{(}
        \KeywordTok{llply}\NormalTok{(}\KeywordTok{grep}\NormalTok{(}\StringTok{'^env}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{_.*$'}\NormalTok{, }\KeywordTok{names}\NormalTok{(grid.DF), }\DataTypeTok{value=}\OtherTok{TRUE}\NormalTok{), function(x) \{}
            \KeywordTok{ggplot}\NormalTok{() +}
\StringTok{                }\KeywordTok{geom_polygon}\NormalTok{(}\DataTypeTok{data=}\NormalTok{countries.FPLY, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{long, }\DataTypeTok{y=}\NormalTok{lat, }\DataTypeTok{group=}\NormalTok{group),}
                    \DataTypeTok{fill=}\StringTok{'grey20'}\NormalTok{, }\DataTypeTok{color=}\StringTok{'grey80'}\NormalTok{) +}
\StringTok{                }\KeywordTok{geom_polygon}\NormalTok{(}\DataTypeTok{data=}\NormalTok{grid.FPLY, }\KeywordTok{aes_string}\NormalTok{(}\DataTypeTok{x=}\StringTok{'long'}\NormalTok{, }\DataTypeTok{y=}\StringTok{'lat'}\NormalTok{, }
                    \DataTypeTok{group=}\StringTok{'group'}\NormalTok{, }\DataTypeTok{fill=}\NormalTok{x),}
                    \DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{, }\DataTypeTok{color=}\StringTok{'grey10'}\NormalTok{) +}
\StringTok{                }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{fill=}\KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{title=}\StringTok{' '}\NormalTok{)) +}
\StringTok{                }\KeywordTok{theme_classic}\NormalTok{() +}
\StringTok{                }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.ticks=}\KeywordTok{element_blank}\NormalTok{(), }\DataTypeTok{axis.text=}\KeywordTok{element_blank}\NormalTok{(),}
                    \DataTypeTok{plot.margin=}\KeywordTok{unit}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{),}\StringTok{'cm'}\NormalTok{), }\DataTypeTok{axis.line=}\KeywordTok{element_blank}\NormalTok{()) +}
\StringTok{                }\KeywordTok{coord_cartesian}\NormalTok{(}
                    \DataTypeTok{xlim=}\KeywordTok{buffered.range}\NormalTok{(grid.FPLY$long, }\FloatTok{0.05}\NormalTok{),}
                    \DataTypeTok{ylim=}\KeywordTok{buffered.range}\NormalTok{(grid.FPLY$lat, }\FloatTok{0.05}\NormalTok{)}
                \NormalTok{) +}
\StringTok{                }\KeywordTok{xlab}\NormalTok{(}\StringTok{''}\NormalTok{) +}
\StringTok{                }\KeywordTok{ylab}\NormalTok{(}\StringTok{''}\NormalTok{) +}
\StringTok{                }\KeywordTok{ggtitle}\NormalTok{(}\KeywordTok{paste0}\NormalTok{(}\StringTok{'PC '}\NormalTok{, }\KeywordTok{substr}\NormalTok{(x, }\KeywordTok{nchar}\NormalTok{(x), }\KeywordTok{nchar}\NormalTok{(x))))}
        \NormalTok{\}),}
        \KeywordTok{list}\NormalTok{(}\DataTypeTok{ncol=}\DecValTok{2}\NormalTok{)}
    \NormalTok{)}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{article_files/figure-latex/unnamed-chunk-8-1.pdf}
\caption{Climatic variation. Each panel depicts variation based on a
different principle component (PC). Sqaures represent planning units.
The color of each planning unit denotes the average priciniple component
value of pixels inside it. Planning units with more similar colors have
more similar climates regimes.}
\end{figure}

\subsubsection{Figure S3. Number populations in each
species.}\label{figure-s3.-number-populations-in-each-species.}

\subsubsection{Figure S4. Distribution of populations for each
species.}\label{figure-s4.-distribution-of-populations-for-each-species.}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{plot.spp.mds <-}\StringTok{ }\NormalTok{function(i) \{}
    \CommentTok{# define function}
    \NormalTok{make.mds.plot <-}\StringTok{ }\NormalTok{function(j,k) \{}
        \KeywordTok{ggplot}\NormalTok{() +}
\StringTok{            }\KeywordTok{geom_polygon}\NormalTok{(}\DataTypeTok{data=}\NormalTok{countries.FPLY, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{long, }\DataTypeTok{y=}\NormalTok{lat, }\DataTypeTok{group=}\NormalTok{group),}
                \DataTypeTok{fill=}\StringTok{'grey20'}\NormalTok{, }\DataTypeTok{color=}\StringTok{'grey80'}\NormalTok{) +}
\StringTok{            }\KeywordTok{geom_polygon}\NormalTok{(}\DataTypeTok{data=}\NormalTok{grid.FPLY, }\KeywordTok{aes_string}\NormalTok{(}\DataTypeTok{x=}\StringTok{'long'}\NormalTok{, }\DataTypeTok{y=}\StringTok{'lat'}\NormalTok{, }
                \DataTypeTok{group=}\StringTok{'group'}\NormalTok{, }\DataTypeTok{fill=}\KeywordTok{paste0}\NormalTok{(}\KeywordTok{unique}\NormalTok{(spp.samples.DF$species)[i], }\StringTok{'_'}\NormalTok{, j, }\StringTok{'_d'}\NormalTok{,k)),}
                \DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{, }\DataTypeTok{color=}\StringTok{'grey10'}\NormalTok{) +}
\StringTok{            }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{fill=}\KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{title=}\StringTok{' '}\NormalTok{)) +}
\StringTok{            }\KeywordTok{theme_classic}\NormalTok{() +}
\StringTok{            }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.ticks=}\KeywordTok{element_blank}\NormalTok{(), }\DataTypeTok{axis.text=}\KeywordTok{element_blank}\NormalTok{(),}
                \DataTypeTok{plot.margin=}\KeywordTok{unit}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{),}\StringTok{'cm'}\NormalTok{), }\DataTypeTok{axis.line=}\KeywordTok{element_blank}\NormalTok{()) +}
\StringTok{            }\KeywordTok{coord_cartesian}\NormalTok{(}
                \DataTypeTok{xlim=}\KeywordTok{buffered.range}\NormalTok{(grid.FPLY$long, }\FloatTok{0.05}\NormalTok{),}
                \DataTypeTok{ylim=}\KeywordTok{buffered.range}\NormalTok{(grid.FPLY$lat, }\FloatTok{0.05}\NormalTok{)}
            \NormalTok{) +}
\StringTok{            }\KeywordTok{xlab}\NormalTok{(}\StringTok{''}\NormalTok{) +}
\StringTok{            }\KeywordTok{ylab}\NormalTok{(}\StringTok{''}\NormalTok{) +}
\StringTok{            }\KeywordTok{ggtitle}\NormalTok{(}\KeywordTok{paste0}\NormalTok{(j,}\StringTok{' ('}\NormalTok{,k,}\StringTok{')'}\NormalTok{))}
    \NormalTok{\}}
    \CommentTok{# init}
    \NormalTok{adapt.col <-}\StringTok{ }\KeywordTok{paste0}\NormalTok{(}\KeywordTok{unique}\NormalTok{(spp.samples.DF$species)[i], }\StringTok{'_adaptive_d1'}\NormalTok{)}
    \NormalTok{neutral.col <-}\StringTok{ }\KeywordTok{paste0}\NormalTok{(}\KeywordTok{unique}\NormalTok{(spp.samples.DF$species)[i], }\StringTok{'_neutral_d1'}\NormalTok{)}
    \NormalTok{plotLST <-}\StringTok{ }\KeywordTok{list}\NormalTok{()}
    \CommentTok{# make adaptive plot}
    \NormalTok{if (adapt.col %in%}\StringTok{ }\KeywordTok{names}\NormalTok{(grid.FPLY)) \{}
        \NormalTok{for (k in }\KeywordTok{seq_len}\NormalTok{(mds.k))}
            \NormalTok{plotLST <-}\StringTok{ }\KeywordTok{append}\NormalTok{(plotLST, }\KeywordTok{list}\NormalTok{(}\KeywordTok{make.mds.plot}\NormalTok{(}\StringTok{'adaptive'}\NormalTok{, k)))}
    \NormalTok{\} }
    \CommentTok{# make neutral plot}
    \NormalTok{if (neutral.col %in%}\StringTok{ }\KeywordTok{names}\NormalTok{(grid.FPLY)) \{}
        \NormalTok{for (k in }\KeywordTok{seq_len}\NormalTok{(mds.k))}
            \NormalTok{plotLST <-}\StringTok{ }\KeywordTok{append}\NormalTok{(plotLST, }\KeywordTok{list}\NormalTok{(}\KeywordTok{make.mds.plot}\NormalTok{(}\StringTok{'neutral'}\NormalTok{, k)))}
    \NormalTok{\} }
    \CommentTok{# make plot}
    \KeywordTok{do.call}\NormalTok{(}
        \NormalTok{grid.arrange,}
        \KeywordTok{append}\NormalTok{(}
            \NormalTok{plotLST,}
            \KeywordTok{list}\NormalTok{(}\DataTypeTok{ncol=}\DecValTok{6}\NormalTok{)}
        \NormalTok{)}
    \NormalTok{)}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\subsubsection{Figures S}\label{figures-s}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot.spp.mds}\NormalTok{(}\DecValTok{1}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{article_files/figure-latex/unnamed-chunk-12-1.pdf}
\caption{Distribution of adaptive and neutral genetic variation in
\textit{Androsace obtusifolia}. Each square represents a planning unit.
The color of each planning unit panel corresponds to ordination values.
Planning units with similar colors contain individiduals with similar
genetic variation.}
\end{figure}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot.spp.mds}\NormalTok{(}\DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{article_files/figure-latex/unnamed-chunk-13-1.pdf}
\caption{Distribution of adaptive and neutral genetic variation in
\textit{Arabis alpina}. See Figure XX caption for conventions.}
\end{figure}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot.spp.mds}\NormalTok{(}\DecValTok{3}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{article_files/figure-latex/unnamed-chunk-14-1.pdf}
\caption{Distribution of adaptive and neutral genetic variation in
\textit{Campanula barbata}. See Figure XX caption for conventions.}
\end{figure}

\subsubsection{Table S1: Principle components analysis on climatic
variation}\label{table-s1-principle-components-analysis-on-climatic-variation}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## make results table showing Eigen values}
\NormalTok{knitr::}\KeywordTok{kable}\NormalTok{(}
    \NormalTok{pca.DF,}
    \DataTypeTok{digits=}\DecValTok{2}\NormalTok{,}
    \DataTypeTok{caption=}\StringTok{'Summary of principle components analysis (PCA) on bioclimatic variation across the study area. The first two principle components (PCs) were used for subsequent analysis.'}\NormalTok{,}
    \DataTypeTok{align=}\KeywordTok{c}\NormalTok{(}\StringTok{'l'}\NormalTok{, }\StringTok{'c'}\NormalTok{, }\StringTok{'c'}\NormalTok{, }\StringTok{'c'}\NormalTok{)}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[c]{@{}lccc@{}}
\toprule\addlinespace
Principle Component & Eigen Value & Variation explained (\%) &
Accumulative variation explained (\%)
\\\addlinespace
\midrule\endhead
1 & 216765.14 & 82.67 & 82.67
\\\addlinespace
2 & 38177.84 & 14.56 & 97.23
\\\addlinespace
3 & 5356.75 & 2.04 & 99.27
\\\addlinespace
4 & 1216.67 & 0.46 & 99.73
\\\addlinespace
5 & 700.39 & 0.27 & 100.00
\\\addlinespace
\bottomrule
\addlinespace
\caption{Summary of principle components analysis (PCA) on bioclimatic
variation across the study area. The first two principle components
(PCs) were used for subsequent analysis.}
\end{longtable}

\subsubsection{Table S2: BayeScan
Results}\label{table-s2-bayescan-results}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{knitr::}\KeywordTok{kable}\NormalTok{(}
    \KeywordTok{format.table}\NormalTok{(}
        \KeywordTok{ldply}\NormalTok{(}
            \KeywordTok{seq_along}\NormalTok{(}\KeywordTok{unique}\NormalTok{(spp.samples.DF$species)), }
            \NormalTok{function(i) \{}
                \KeywordTok{data.frame}\NormalTok{(}
                    \DataTypeTok{Species=}\KeywordTok{paste0}\NormalTok{(}\StringTok{'}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{textit\{'}\NormalTok{,}\KeywordTok{gsub}\NormalTok{(}\StringTok{'}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{_'}\NormalTok{, }\StringTok{' '}\NormalTok{, }\KeywordTok{unique}\NormalTok{(spp.samples.DF$species)[i]),}\StringTok{'\}'}\NormalTok{),}
                    \DataTypeTok{Primer=}\NormalTok{spp.BayeScan.sample.loci.subset.LST[[i]]@data@primers,}
                    \DataTypeTok{Probability=}\NormalTok{spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[}\DecValTok{2}\NormalTok{]],}
                    \DataTypeTok{qval=}\NormalTok{spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[}\DecValTok{4}\NormalTok{]],}
                    \DataTypeTok{alpha=}\NormalTok{spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[}\DecValTok{5}\NormalTok{]],}
                    \DataTypeTok{fst=}\NormalTok{spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[}\DecValTok{6}\NormalTok{]],}
                    \DataTypeTok{Type=}\NormalTok{spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[}\DecValTok{7}\NormalTok{]]}
                \NormalTok{)}
            \NormalTok{\}}
        \NormalTok{),}
        \DataTypeTok{omit=}\StringTok{'Type'}
    \NormalTok{),}
    \DataTypeTok{digits=}\DecValTok{2}\NormalTok{,}
    \DataTypeTok{col.names=}\KeywordTok{c}\NormalTok{(}\StringTok{'Species'}\NormalTok{, }\StringTok{'Primer'}\NormalTok{, }\StringTok{'Probability'}\NormalTok{, }\StringTok{'q-value'}\NormalTok{, }\StringTok{'$}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{alpha$'}\NormalTok{, }\StringTok{'$F_\{ST\}$'}\NormalTok{, }\StringTok{'Type'}\NormalTok{),}
    \DataTypeTok{align=}\KeywordTok{c}\NormalTok{(}\StringTok{'l'}\NormalTok{, }\StringTok{'c'}\NormalTok{, }\StringTok{'c'}\NormalTok{, }\StringTok{'c'}\NormalTok{, }\StringTok{'c'}\NormalTok{, }\StringTok{'c'}\NormalTok{, }\StringTok{'c'}\NormalTok{)}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[c]{@{}lcccccc@{}}
\toprule\addlinespace
Species & Primer & Probability & q-value & $\alpha$ & $F_{ST}$ & Type
\\\addlinespace
\midrule\endhead
\textit{Androsace obtusifolia} & AAC\_CAN\_83.0 & 0.00 & 0.87 & 0.00 &
0.14 & neutral
\\\addlinespace
& AAC\_CAN\_85.0 & 0.00 & 0.87 & 0.00 & 0.14 & neutral
\\\addlinespace
& AAC\_CAN\_89.0 & 0.04 & 0.83 & 0.04 & 0.14 & neutral
\\\addlinespace
& AAC\_CAN\_91.0 & 0.07 & 0.77 & -0.04 & 0.13 & neutral
\\\addlinespace
& AAC\_CAN\_100.0 & 0.07 & 0.77 & 0.01 & 0.14 & neutral
\\\addlinespace
& AAC\_CAN\_102.0 & 0.30 & 0.55 & 0.20 & 0.16 & neutral
\\\addlinespace
& AAC\_CAN\_108.0 & 0.04 & 0.83 & 0.04 & 0.14 & neutral
\\\addlinespace
& AAC\_CAN\_124.0 & 0.33 & 0.57 & -0.46 & 0.10 & neutral
\\\addlinespace
& AAC\_CAN\_125.0 & 0.07 & 0.78 & 0.00 & 0.14 & neutral
\\\addlinespace
& AAC\_CAN\_128.0 & 0.33 & 0.58 & -0.11 & 0.13 & neutral
\\\addlinespace
& AAC\_CAN\_130.0 & 0.19 & 0.59 & NaN & NaN & neutral
\\\addlinespace
& AAC\_CAN\_132.0 & 0.30 & 0.55 & 0.05 & 0.14 & neutral
\\\addlinespace
& AAC\_CAN\_133.0 & 0.00 & 0.87 & 0.00 & 0.14 & neutral
\\\addlinespace
& AAC\_CAN\_136.0 & 0.07 & 0.77 & 0.00 & 0.14 & neutral
\\\addlinespace
& AAC\_CAN\_137.0 & 0.19 & 0.59 & 0.14 & 0.16 & neutral
\\\addlinespace
& AAC\_CAN\_146.0 & 0.22 & 0.61 & 0.15 & 0.16 & neutral
\\\addlinespace
& AAC\_CAN\_151.0 & 0.56 & 0.17 & 0.33 & 0.24 & adaptive
\\\addlinespace
& AAC\_CAN\_152.0 & 0.19 & 0.59 & NaN & NaN & neutral
\\\addlinespace
& AAC\_CAN\_153.0 & 0.11 & 0.68 & 0.02 & 0.14 & neutral
\\\addlinespace
& AAC\_CAN\_182.0 & 0.15 & 0.65 & -0.26 & 0.12 & neutral
\\\addlinespace
& AAC\_CAN\_195.0 & 0.07 & 0.78 & 0.02 & 0.14 & neutral
\\\addlinespace
& AAC\_CAN\_211.0 & 0.04 & 0.83 & 0.01 & 0.14 & neutral
\\\addlinespace
& AAC\_CAN\_220.0 & 0.00 & 0.87 & 0.00 & 0.14 & neutral
\\\addlinespace
& AAC\_CAN\_231.0 & 0.48 & 0.37 & -0.18 & 0.12 & neutral
\\\addlinespace
& AAC\_CAN\_239.0 & 0.37 & 0.54 & -0.56 & 0.10 & neutral
\\\addlinespace
& AAC\_CAN\_272.0 & 0.04 & 0.82 & -0.01 & 0.14 & neutral
\\\addlinespace
& AAC\_CAN\_319.0 & 0.07 & 0.77 & 0.05 & 0.14 & neutral
\\\addlinespace
& ACA\_CAT\_81.0 & 0.04 & 0.82 & -0.01 & 0.14 & neutral
\\\addlinespace
& ACA\_CAT\_85.0 & 0.37 & 0.52 & 0.18 & 0.17 & neutral
\\\addlinespace
& ACA\_CAT\_90.0 & 0.15 & 0.61 & 0.03 & 0.14 & neutral
\\\addlinespace
& ACA\_CAT\_97.0 & 0.07 & 0.73 & -0.04 & 0.13 & neutral
\\\addlinespace
& ACA\_CAT\_99.0 & 0.15 & 0.65 & 0.16 & 0.17 & neutral
\\\addlinespace
& ACA\_CAT\_100.0 & 0.07 & 0.76 & 0.02 & 0.14 & neutral
\\\addlinespace
& ACA\_CAT\_102.0 & 0.07 & 0.76 & -0.04 & 0.13 & neutral
\\\addlinespace
& ACA\_CAT\_103.0 & 0.00 & 0.87 & 0.00 & 0.14 & neutral
\\\addlinespace
& ACA\_CAT\_108.0 & 0.15 & 0.64 & -0.10 & 0.13 & neutral
\\\addlinespace
& ACA\_CAT\_120.0 & 0.11 & 0.67 & -0.05 & 0.13 & neutral
\\\addlinespace
& ACA\_CAT\_124.0 & 0.04 & 0.82 & -0.04 & 0.13 & neutral
\\\addlinespace
& ACA\_CAT\_126.0 & 0.04 & 0.82 & 0.05 & 0.15 & neutral
\\\addlinespace
& ACA\_CAT\_129.0 & 0.33 & 0.58 & 0.00 & 0.14 & neutral
\\\addlinespace
& ACA\_CAT\_131.0 & 0.19 & 0.59 & -0.12 & 0.13 & neutral
\\\addlinespace
& ACA\_CAT\_133.0 & 0.70 & 0.24 & -0.35 & 0.12 & neutral
\\\addlinespace
& ACA\_CAT\_135.0 & 0.07 & 0.73 & -0.04 & 0.13 & neutral
\\\addlinespace
& ACA\_CAT\_140.0 & 0.04 & 0.82 & 0.02 & 0.14 & neutral
\\\addlinespace
& ACA\_CAT\_148.0 & 0.04 & 0.82 & 0.03 & 0.14 & neutral
\\\addlinespace
& ACA\_CAT\_152.0 & 0.07 & 0.78 & -0.05 & 0.13 & neutral
\\\addlinespace
& ACA\_CAT\_153.0 & 0.04 & 0.82 & 0.00 & 0.14 & neutral
\\\addlinespace
& ACA\_CAT\_155.0 & 0.07 & 0.76 & 0.01 & 0.14 & neutral
\\\addlinespace
& ACA\_CAT\_157.0 & 0.37 & 0.54 & -0.16 & 0.12 & neutral
\\\addlinespace
& ACA\_CAT\_159.0 & 0.11 & 0.70 & 0.18 & 0.17 & neutral
\\\addlinespace
& ACA\_CAT\_162.0 & 0.07 & 0.73 & 0.01 & 0.14 & neutral
\\\addlinespace
& ACA\_CAT\_168.0 & 0.19 & 0.60 & -0.05 & 0.13 & neutral
\\\addlinespace
& ACA\_CAT\_173.0 & 0.33 & 0.58 & 0.20 & 0.17 & neutral
\\\addlinespace
& ACA\_CAT\_177.0 & 0.00 & 0.87 & 0.00 & 0.14 & neutral
\\\addlinespace
& ACA\_CAT\_178.0 & 0.19 & 0.58 & 0.07 & 0.14 & neutral
\\\addlinespace
& ACA\_CAT\_187.0 & 0.04 & 0.81 & -0.02 & 0.14 & neutral
\\\addlinespace
& ACA\_CAT\_192.0 & 0.19 & 0.59 & 0.09 & 0.15 & neutral
\\\addlinespace
& ACA\_CAT\_196.0 & 0.00 & 0.87 & 0.00 & 0.14 & neutral
\\\addlinespace
& ACA\_CAT\_199.0 & 0.04 & 0.83 & 0.00 & 0.14 & neutral
\\\addlinespace
& ACA\_CAT\_200.0 & 0.04 & 0.83 & -0.02 & 0.14 & neutral
\\\addlinespace
& ACA\_CAT\_204.0 & 0.00 & 0.87 & 0.00 & 0.14 & neutral
\\\addlinespace
& ACA\_CAT\_205.0 & 0.07 & 0.77 & -0.02 & 0.13 & neutral
\\\addlinespace
& ACA\_CAT\_210.0 & 0.15 & 0.61 & -0.03 & 0.13 & neutral
\\\addlinespace
& ACA\_CAT\_214.0 & 0.33 & 0.58 & 0.03 & 0.14 & neutral
\\\addlinespace
& ACA\_CAT\_219.0 & 0.11 & 0.68 & -0.03 & 0.13 & neutral
\\\addlinespace
& ACA\_CAT\_229.0 & 0.15 & 0.61 & -0.04 & 0.13 & neutral
\\\addlinespace
& ACA\_CAT\_237.0 & 0.15 & 0.65 & -0.07 & 0.13 & neutral
\\\addlinespace
& ACA\_CAT\_243.0 & 0.04 & 0.82 & -0.02 & 0.14 & neutral
\\\addlinespace
& ACA\_CAT\_246.0 & 0.04 & 0.82 & -0.01 & 0.14 & neutral
\\\addlinespace
& ACA\_CAT\_248.0 & 0.04 & 0.83 & -0.02 & 0.13 & neutral
\\\addlinespace
& ACA\_CAT\_282.0 & 0.00 & 0.87 & 0.00 & 0.14 & neutral
\\\addlinespace
& ACA\_CAT\_300.0 & 0.04 & 0.81 & -0.02 & 0.14 & neutral
\\\addlinespace
& ACA\_CAT\_390.0 & 0.41 & 0.49 & -0.02 & 0.15 & neutral
\\\addlinespace
& ACA\_CAT\_391.0 & 0.04 & 0.81 & 0.03 & 0.14 & neutral
\\\addlinespace
& ACA\_CAT\_393.0 & 0.04 & 0.83 & -0.01 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_82.3 & 0.04 & 0.83 & -0.03 & 0.13 & neutral
\\\addlinespace
& AGG\_CAA\_83.0 & 0.15 & 0.65 & -0.02 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_84.2 & 0.41 & 0.45 & -0.05 & 0.13 & neutral
\\\addlinespace
& AGG\_CAA\_86.9 & 0.07 & 0.77 & 0.00 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_90.9 & 0.11 & 0.65 & -0.02 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_94.8 & 0.04 & 0.81 & 0.01 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_95.6 & 0.07 & 0.71 & 0.05 & 0.15 & neutral
\\\addlinespace
& AGG\_CAA\_100.0 & 0.07 & 0.78 & -0.01 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_101.0 & 0.07 & 0.73 & -0.13 & 0.13 & neutral
\\\addlinespace
& AGG\_CAA\_109.3 & 0.07 & 0.78 & 0.04 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_110.2 & 0.00 & 0.87 & 0.00 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_113.5 & 0.04 & 0.81 & 0.04 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_115.9 & 0.11 & 0.68 & -0.02 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_117.4 & 0.33 & 0.58 & -0.31 & 0.12 & neutral
\\\addlinespace
& AGG\_CAA\_118.2 & 0.04 & 0.82 & -0.02 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_122.9 & 0.33 & 0.58 & -0.11 & 0.13 & neutral
\\\addlinespace
& AGG\_CAA\_129.5 & 0.37 & 0.42 & 0.43 & 0.23 & neutral
\\\addlinespace
& AGG\_CAA\_130.1 & 0.48 & 0.29 & 0.17 & 0.16 & adaptive
\\\addlinespace
& AGG\_CAA\_135.5 & 0.37 & 0.52 & -0.10 & 0.13 & neutral
\\\addlinespace
& AGG\_CAA\_137.9 & 0.26 & 0.57 & -0.12 & 0.13 & neutral
\\\addlinespace
& AGG\_CAA\_144.9 & 0.37 & 0.53 & -0.03 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_145.8 & 0.37 & 0.52 & -0.36 & 0.11 & neutral
\\\addlinespace
& AGG\_CAA\_150.9 & 0.00 & 0.87 & 0.00 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_152.8 & 0.07 & 0.74 & 0.04 & 0.15 & neutral
\\\addlinespace
& AGG\_CAA\_155.4 & 0.04 & 0.83 & 0.02 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_164.0 & 0.22 & 0.61 & -0.08 & 0.13 & neutral
\\\addlinespace
& AGG\_CAA\_175.9 & 0.41 & 0.49 & -0.17 & 0.13 & neutral
\\\addlinespace
& AGG\_CAA\_181.6 & 0.04 & 0.83 & 0.00 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_182.1 & 0.00 & 0.87 & 0.00 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_188.1 & 0.07 & 0.73 & 0.01 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_191.3 & 0.07 & 0.77 & 0.01 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_193.9 & 0.04 & 0.83 & 0.02 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_196.8 & 0.11 & 0.70 & 0.02 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_201.7 & 0.07 & 0.77 & -0.04 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_212.4 & 0.07 & 0.76 & -0.02 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_215.6 & 0.15 & 0.65 & -0.05 & 0.13 & neutral
\\\addlinespace
& AGG\_CAA\_219.0 & 0.04 & 0.83 & 0.00 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_225.2 & 0.04 & 0.83 & -0.02 & 0.13 & neutral
\\\addlinespace
& AGG\_CAA\_253.2 & 0.07 & 0.76 & 0.01 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_256.7 & 0.04 & 0.82 & 0.02 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_262.5 & 0.30 & 0.54 & -0.02 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_263.6 & 0.11 & 0.67 & -0.09 & 0.13 & neutral
\\\addlinespace
& AGG\_CAA\_264.8 & 0.04 & 0.81 & -0.01 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_267.9 & 0.41 & 0.49 & 0.28 & 0.17 & neutral
\\\addlinespace
& AGG\_CAA\_269.8 & 0.04 & 0.83 & 0.03 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_270.8 & 0.11 & 0.72 & 0.05 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_276.8 & 0.07 & 0.78 & -0.06 & 0.13 & neutral
\\\addlinespace
& AGG\_CAA\_299.0 & 0.26 & 0.55 & 0.11 & 0.15 & neutral
\\\addlinespace
& AGG\_CAA\_305.1 & 0.11 & 0.69 & -0.06 & 0.13 & neutral
\\\addlinespace
& AGG\_CAA\_312.1 & 0.04 & 0.82 & 0.01 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_313.2 & 0.04 & 0.81 & -0.03 & 0.13 & neutral
\\\addlinespace
& AGG\_CAA\_316.0 & 0.00 & 0.87 & 0.00 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_319.0 & 0.00 & 0.87 & 0.00 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_324.1 & 0.04 & 0.81 & -0.06 & 0.13 & neutral
\\\addlinespace
& AGG\_CAA\_359.4 & 0.11 & 0.70 & -0.02 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_360.5 & 0.19 & 0.60 & -0.05 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_363.2 & 0.07 & 0.73 & -0.05 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_364.1 & 0.07 & 0.78 & 0.03 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_376.2 & 0.11 & 0.68 & 0.05 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_376.7 & 0.00 & 0.87 & 0.00 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_396.0 & 0.30 & 0.54 & -0.25 & 0.11 & neutral
\\\addlinespace
& AGG\_CAA\_403.4 & 0.07 & 0.73 & 0.05 & 0.14 & neutral
\\\addlinespace
& AGG\_CAA\_420.5 & 0.07 & 0.78 & -0.06 & 0.13 & neutral
\\\addlinespace
\textit{Arabis alpina} & AAT\_CAC\_51.9 & 0.30 & 0.57 & 0.51 & 0.26 &
neutral
\\\addlinespace
& AAT\_CAC\_54.7 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_69.0 & 0.22 & 0.52 & 0.03 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_73.6 & 0.04 & 0.81 & -0.03 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_77.9 & 0.37 & 0.53 & 0.25 & 0.21 & neutral
\\\addlinespace
& AAT\_CAC\_86.9 & 0.15 & 0.64 & 0.08 & 0.18 & neutral
\\\addlinespace
& AAT\_CAC\_90.5 & 0.04 & 0.80 & 0.01 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_95.3 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_97.0 & 0.07 & 0.76 & -0.08 & 0.16 & neutral
\\\addlinespace
& AAT\_CAC\_97.1 & 0.07 & 0.76 & 0.02 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_100.3 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_105.4 & 0.33 & 0.57 & 0.13 & 0.19 & neutral
\\\addlinespace
& AAT\_CAC\_118.4 & 0.04 & 0.82 & 0.01 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_121.3 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_128.0 & 0.37 & 0.50 & -0.45 & 0.16 & neutral
\\\addlinespace
& AAT\_CAC\_130.0 & 0.15 & 0.62 & 0.09 & 0.18 & neutral
\\\addlinespace
& AAT\_CAC\_147.4 & 0.33 & 0.57 & 0.07 & 0.19 & neutral
\\\addlinespace
& AAT\_CAC\_156.9 & 0.11 & 0.70 & 0.03 & 0.18 & neutral
\\\addlinespace
& AAT\_CAC\_175.1 & 0.07 & 0.78 & -0.02 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_177.9 & 0.04 & 0.81 & -0.02 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_179.6 & 0.11 & 0.72 & 0.03 & 0.18 & neutral
\\\addlinespace
& AAT\_CAC\_188.6 & 0.07 & 0.75 & -0.01 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_190.0 & 0.37 & 0.54 & 0.22 & 0.21 & neutral
\\\addlinespace
& AAT\_CAC\_195.5 & 0.26 & 0.43 & -0.04 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_197.1 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_200.6 & 0.04 & 0.81 & -0.02 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_201.8 & 0.11 & 0.66 & 0.05 & 0.18 & neutral
\\\addlinespace
& AAT\_CAC\_209.2 & 0.11 & 0.68 & 0.01 & 0.18 & neutral
\\\addlinespace
& AAT\_CAC\_213.1 & 0.04 & 0.80 & 0.00 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_215.0 & 0.07 & 0.78 & -0.01 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_216.2 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_217.0 & 0.30 & 0.55 & 0.13 & 0.20 & neutral
\\\addlinespace
& AAT\_CAC\_218.1 & 0.11 & 0.68 & -0.06 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_219.1 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_225.5 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_227.1 & 0.37 & 0.54 & -0.15 & 0.16 & neutral
\\\addlinespace
& AAT\_CAC\_229.1 & 0.44 & 0.40 & 0.39 & 0.24 & neutral
\\\addlinespace
& AAT\_CAC\_231.4 & 0.04 & 0.82 & -0.03 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_249.7 & 0.07 & 0.76 & 0.01 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_259.3 & 0.33 & 0.58 & -0.05 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_277.3 & 0.04 & 0.80 & -0.01 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_298.1 & 0.11 & 0.71 & -0.01 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_315.9 & 0.19 & 0.58 & NaN & NaN & neutral
\\\addlinespace
& AAT\_CAC\_330.4 & 0.07 & 0.76 & 0.02 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_334.8 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_336.6 & 0.11 & 0.72 & -0.07 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_353.0 & 0.04 & 0.82 & 0.00 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_359.2 & 0.07 & 0.76 & -0.08 & 0.16 & neutral
\\\addlinespace
& AAT\_CAC\_399.9 & 0.04 & 0.82 & 0.02 & 0.18 & neutral
\\\addlinespace
& AAT\_CAC\_410.5 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_412.4 & 0.33 & 0.58 & 0.12 & 0.19 & neutral
\\\addlinespace
& AAT\_CAC\_458.1 & 0.15 & 0.59 & -0.03 & 0.17 & neutral
\\\addlinespace
& AAT\_CAC\_488.6 & 0.04 & 0.82 & 0.03 & 0.18 & neutral
\\\addlinespace
& AGT\_CAC\_53.8 & 0.07 & 0.72 & 0.01 & 0.17 & neutral
\\\addlinespace
& AGT\_CAC\_56.3 & 0.04 & 0.80 & -0.01 & 0.17 & neutral
\\\addlinespace
& AGT\_CAC\_98.2 & 0.44 & 0.43 & -0.44 & 0.13 & neutral
\\\addlinespace
& AGT\_CAC\_104.9 & 0.41 & 0.47 & -0.16 & 0.16 & neutral
\\\addlinespace
& AGT\_CAC\_114.8 & 0.07 & 0.75 & -0.05 & 0.16 & neutral
\\\addlinespace
& AGT\_CAC\_145.8 & 0.07 & 0.75 & 0.00 & 0.17 & neutral
\\\addlinespace
& AGT\_CAC\_154.2 & 0.04 & 0.80 & 0.01 & 0.17 & neutral
\\\addlinespace
& AGT\_CAC\_158.1 & 0.07 & 0.76 & 0.01 & 0.17 & neutral
\\\addlinespace
& AGT\_CAC\_169.0 & 0.04 & 0.81 & -0.02 & 0.17 & neutral
\\\addlinespace
& AGT\_CAC\_171.1 & 0.11 & 0.68 & 0.03 & 0.18 & neutral
\\\addlinespace
& AGT\_CAC\_181.1 & 0.15 & 0.62 & -0.04 & 0.17 & neutral
\\\addlinespace
& AGT\_CAC\_183.0 & 0.33 & 0.57 & 0.21 & 0.20 & neutral
\\\addlinespace
& AGT\_CAC\_184.4 & 0.11 & 0.72 & 0.00 & 0.17 & neutral
\\\addlinespace
& AGT\_CAC\_191.2 & 0.04 & 0.81 & 0.03 & 0.18 & neutral
\\\addlinespace
& AGT\_CAC\_195.0 & 0.15 & 0.61 & -0.05 & 0.17 & neutral
\\\addlinespace
& AGT\_CAC\_200.9 & 0.19 & 0.58 & 0.08 & 0.19 & neutral
\\\addlinespace
& AGT\_CAC\_203.8 & 0.07 & 0.75 & 0.01 & 0.17 & neutral
\\\addlinespace
& AGT\_CAC\_205.8 & 0.11 & 0.67 & -0.02 & 0.17 & neutral
\\\addlinespace
& AGT\_CAC\_210.0 & 0.04 & 0.82 & 0.01 & 0.17 & neutral
\\\addlinespace
& AGT\_CAC\_230.8 & 0.41 & 0.44 & -0.20 & 0.15 & neutral
\\\addlinespace
& AGT\_CAC\_241.0 & 0.41 & 0.44 & 0.00 & 0.17 & neutral
\\\addlinespace
& AGT\_CAC\_245.6 & 0.07 & 0.76 & -0.05 & 0.17 & neutral
\\\addlinespace
& AGT\_CAC\_264.7 & 0.04 & 0.82 & -0.04 & 0.17 & neutral
\\\addlinespace
& AGT\_CAC\_266.9 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
\\\addlinespace
& AGT\_CAC\_269.4 & 0.04 & 0.82 & 0.04 & 0.18 & neutral
\\\addlinespace
& AGT\_CAC\_274.0 & 0.11 & 0.68 & 0.04 & 0.18 & neutral
\\\addlinespace
& AGT\_CAC\_285.6 & 0.07 & 0.78 & 0.02 & 0.18 & neutral
\\\addlinespace
& AGT\_CAC\_291.5 & 0.07 & 0.71 & -0.01 & 0.17 & neutral
\\\addlinespace
& AGT\_CAC\_295.6 & 0.07 & 0.78 & 0.02 & 0.17 & neutral
\\\addlinespace
& AGT\_CAC\_315.2 & 0.44 & 0.38 & 0.29 & 0.23 & neutral
\\\addlinespace
& AGT\_CAC\_332.1 & 0.48 & 0.31 & 0.72 & 0.31 & neutral
\\\addlinespace
& AGT\_CAC\_347.8 & 0.11 & 0.72 & -0.07 & 0.16 & neutral
\\\addlinespace
& AGT\_CAC\_355.2 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
\\\addlinespace
& AGT\_CAC\_360.2 & 0.15 & 0.65 & 0.08 & 0.19 & neutral
\\\addlinespace
& AGT\_CAC\_386.5 & 0.04 & 0.82 & 0.02 & 0.17 & neutral
\\\addlinespace
& AGT\_CAC\_418.5 & 0.15 & 0.64 & 0.05 & 0.19 & neutral
\\\addlinespace
& AGT\_CAC\_420.2 & 0.37 & 0.47 & -0.08 & 0.16 & neutral
\\\addlinespace
& AGT\_CAC\_444.3 & 0.04 & 0.82 & -0.04 & 0.17 & neutral
\\\addlinespace
& AGT\_CAC\_453.4 & 0.11 & 0.68 & -0.04 & 0.17 & neutral
\\\addlinespace
& AGT\_CAC\_458.5 & 0.15 & 0.57 & -0.01 & 0.17 & neutral
\\\addlinespace
& AGT\_CAC\_489.1 & 0.07 & 0.75 & -0.06 & 0.16 & neutral
\\\addlinespace
& ATC\_CAC\_52.4 & 0.15 & 0.65 & 0.00 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_56.7 & 0.15 & 0.65 & 0.06 & 0.18 & neutral
\\\addlinespace
& ATC\_CAC\_61.5 & 0.15 & 0.65 & -0.06 & 0.16 & neutral
\\\addlinespace
& ATC\_CAC\_64.3 & 0.07 & 0.75 & 0.04 & 0.18 & neutral
\\\addlinespace
& ATC\_CAC\_91.9 & 0.07 & 0.76 & -0.02 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_96.2 & 0.41 & 0.42 & -0.54 & 0.13 & neutral
\\\addlinespace
& ATC\_CAC\_99.5 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_100.6 & 0.11 & 0.68 & -0.02 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_102.0 & 0.11 & 0.66 & 0.02 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_111.9 & 0.37 & 0.53 & -0.35 & 0.15 & neutral
\\\addlinespace
& ATC\_CAC\_113.5 & 0.41 & 0.48 & 0.30 & 0.23 & neutral
\\\addlinespace
& ATC\_CAC\_123.5 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_139.7 & 0.07 & 0.71 & -0.05 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_140.8 & 0.26 & 0.55 & -0.15 & 0.16 & neutral
\\\addlinespace
& ATC\_CAC\_142.9 & 0.41 & 0.47 & -0.07 & 0.16 & neutral
\\\addlinespace
& ATC\_CAC\_144.1 & 0.15 & 0.64 & 0.21 & 0.21 & neutral
\\\addlinespace
& ATC\_CAC\_148.6 & 0.07 & 0.76 & 0.02 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_149.8 & 0.07 & 0.72 & -0.06 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_151.8 & 0.07 & 0.75 & 0.03 & 0.18 & neutral
\\\addlinespace
& ATC\_CAC\_156.1 & 0.04 & 0.80 & 0.00 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_162.5 & 0.19 & 0.60 & 0.31 & 0.23 & neutral
\\\addlinespace
& ATC\_CAC\_181.9 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_186.4 & 0.07 & 0.76 & -0.02 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_189.9 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_194.8 & 0.07 & 0.75 & 0.00 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_198.3 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_199.4 & 0.22 & 0.55 & -0.15 & 0.15 & neutral
\\\addlinespace
& ATC\_CAC\_204.2 & 0.37 & 0.53 & 0.14 & 0.20 & neutral
\\\addlinespace
& ATC\_CAC\_207.3 & 0.04 & 0.81 & -0.02 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_215.8 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_220.7 & 0.41 & 0.44 & 0.18 & 0.21 & neutral
\\\addlinespace
& ATC\_CAC\_223.7 & 0.07 & 0.76 & 0.02 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_229.1 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_230.7 & 0.04 & 0.80 & 0.00 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_233.7 & 0.56 & 0.23 & -0.14 & 0.16 & neutral
\\\addlinespace
& ATC\_CAC\_235.8 & 0.37 & 0.54 & -0.05 & 0.16 & neutral
\\\addlinespace
& ATC\_CAC\_258.7 & 0.07 & 0.78 & -0.03 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_266.2 & 0.04 & 0.82 & -0.04 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_270.8 & 0.41 & 0.44 & -0.51 & 0.12 & neutral
\\\addlinespace
& ATC\_CAC\_273.6 & 0.07 & 0.78 & -0.01 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_274.9 & 0.07 & 0.76 & -0.05 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_276.4 & 0.33 & 0.58 & 0.44 & 0.26 & neutral
\\\addlinespace
& ATC\_CAC\_277.8 & 0.11 & 0.67 & -0.08 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_287.8 & 0.07 & 0.75 & 0.06 & 0.18 & neutral
\\\addlinespace
& ATC\_CAC\_288.7 & 0.04 & 0.82 & 0.00 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_332.4 & 0.52 & 0.30 & 0.26 & 0.22 & neutral
\\\addlinespace
& ATC\_CAC\_347.9 & 0.19 & 0.58 & NaN & NaN & neutral
\\\addlinespace
& ATC\_CAC\_370.4 & 0.07 & 0.75 & 0.01 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_373.3 & 0.07 & 0.78 & 0.05 & 0.18 & neutral
\\\addlinespace
& ATC\_CAC\_378.0 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_387.7 & 0.33 & 0.58 & 0.15 & 0.20 & neutral
\\\addlinespace
& ATC\_CAC\_401.5 & 0.22 & 0.53 & 0.27 & 0.22 & neutral
\\\addlinespace
& ATC\_CAC\_405.7 & 0.04 & 0.82 & 0.00 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_430.5 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_442.2 & 0.37 & 0.53 & -0.19 & 0.16 & neutral
\\\addlinespace
& ATC\_CAC\_445.2 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
\\\addlinespace
& ATC\_CAC\_456.3 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
\\\addlinespace
\textit{Campanula barbata} & ACA\_CTA\_55.8 & 0.07 & 0.78 & -0.03 & 0.12
& neutral
\\\addlinespace
& ACA\_CTA\_69.2 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
\\\addlinespace
& ACA\_CTA\_101.6 & 0.19 & 0.55 & 0.03 & 0.13 & neutral
\\\addlinespace
& ACA\_CTA\_114.7 & 0.33 & 0.54 & -0.14 & 0.11 & neutral
\\\addlinespace
& ACA\_CTA\_122.8 & 0.04 & 0.78 & -0.01 & 0.12 & neutral
\\\addlinespace
& ACA\_CTA\_132.9 & 0.07 & 0.78 & 0.01 & 0.12 & neutral
\\\addlinespace
& ACA\_CTA\_153.4 & 0.33 & 0.55 & -0.24 & 0.11 & neutral
\\\addlinespace
& ACA\_CTA\_155.4 & 0.04 & 0.83 & 0.00 & 0.12 & neutral
\\\addlinespace
& ACA\_CTA\_164.8 & 0.41 & 0.48 & -0.45 & 0.09 & neutral
\\\addlinespace
& ACA\_CTA\_169.4 & 0.07 & 0.78 & -0.09 & 0.12 & neutral
\\\addlinespace
& ACA\_CTA\_174.3 & 0.07 & 0.77 & -0.06 & 0.12 & neutral
\\\addlinespace
& ACA\_CTA\_178.0 & 0.04 & 0.78 & -0.05 & 0.12 & neutral
\\\addlinespace
& ACA\_CTA\_179.5 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
\\\addlinespace
& ACA\_CTA\_183.6 & 0.15 & 0.61 & -0.09 & 0.11 & neutral
\\\addlinespace
& ACA\_CTA\_186.8 & 0.19 & 0.59 & -0.15 & 0.11 & neutral
\\\addlinespace
& ACA\_CTA\_187.9 & 0.07 & 0.78 & -0.02 & 0.12 & neutral
\\\addlinespace
& ACA\_CTA\_194.5 & 0.15 & 0.60 & 0.07 & 0.13 & neutral
\\\addlinespace
& ACA\_CTA\_195.9 & 0.11 & 0.72 & 0.04 & 0.13 & neutral
\\\addlinespace
& ACA\_CTA\_197.7 & 0.04 & 0.82 & 0.01 & 0.12 & neutral
\\\addlinespace
& ACA\_CTA\_203.8 & 0.56 & 0.20 & -0.30 & 0.13 & adaptive
\\\addlinespace
& ACA\_CTA\_213.5 & 0.15 & 0.64 & 0.02 & 0.13 & neutral
\\\addlinespace
& ACA\_CTA\_254.3 & 0.04 & 0.83 & 0.02 & 0.12 & neutral
\\\addlinespace
& ACA\_CTA\_284.5 & 0.04 & 0.82 & -0.01 & 0.12 & neutral
\\\addlinespace
& ACA\_CTA\_289.0 & 0.07 & 0.77 & -0.08 & 0.12 & neutral
\\\addlinespace
& ACA\_CTA\_296.4 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
\\\addlinespace
& ACA\_CTA\_311.1 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
\\\addlinespace
& ACA\_CTA\_347.7 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
\\\addlinespace
& ACA\_CTA\_368.6 & 0.07 & 0.75 & -0.01 & 0.12 & neutral
\\\addlinespace
& ACA\_CTA\_378.8 & 0.04 & 0.78 & 0.00 & 0.12 & neutral
\\\addlinespace
& ACA\_CTA\_382.7 & 0.19 & 0.56 & -0.08 & 0.12 & neutral
\\\addlinespace
& ACA\_CTA\_393.5 & 0.07 & 0.75 & 0.03 & 0.12 & neutral
\\\addlinespace
& ACA\_CTA\_415.5 & 0.11 & 0.70 & -0.01 & 0.12 & neutral
\\\addlinespace
& ACA\_CTA\_489.5 & 0.37 & 0.55 & 0.57 & 0.22 & neutral
\\\addlinespace
& ACA\_CTA\_491.1 & 0.07 & 0.78 & 0.01 & 0.13 & neutral
\\\addlinespace
& AGA\_CAC\_84.2 & 0.11 & 0.68 & -0.10 & 0.11 & neutral
\\\addlinespace
& AGA\_CAC\_89.4 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_91.6 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_102.6 & 0.04 & 0.82 & -0.02 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_106.2 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_107.0 & 0.22 & 0.50 & -0.05 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_109.0 & 0.11 & 0.70 & 0.01 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_116.9 & 0.04 & 0.78 & -0.04 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_130.4 & 0.07 & 0.75 & -0.07 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_133.3 & 0.37 & 0.55 & -0.11 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_135.7 & 0.07 & 0.77 & 0.03 & 0.13 & neutral
\\\addlinespace
& AGA\_CAC\_141.4 & 0.04 & 0.83 & -0.09 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_168.9 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_170.2 & 0.41 & 0.48 & -0.09 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_180.6 & 0.04 & 0.82 & 0.01 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_183.6 & 0.41 & 0.48 & 0.21 & 0.15 & neutral
\\\addlinespace
& AGA\_CAC\_196.1 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_202.0 & 0.52 & 0.26 & 0.10 & 0.14 & adaptive
\\\addlinespace
& AGA\_CAC\_204.8 & 0.04 & 0.83 & -0.02 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_214.4 & 0.07 & 0.74 & -0.06 & 0.11 & neutral
\\\addlinespace
& AGA\_CAC\_218.3 & 0.11 & 0.68 & -0.18 & 0.11 & neutral
\\\addlinespace
& AGA\_CAC\_227.1 & 0.07 & 0.74 & 0.06 & 0.13 & neutral
\\\addlinespace
& AGA\_CAC\_231.4 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_245.4 & 0.07 & 0.78 & -0.03 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_247.3 & 0.11 & 0.68 & 0.01 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_250.2 & 0.07 & 0.75 & 0.04 & 0.13 & neutral
\\\addlinespace
& AGA\_CAC\_251.1 & 0.04 & 0.83 & 0.00 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_269.0 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_279.6 & 0.04 & 0.82 & 0.04 & 0.13 & neutral
\\\addlinespace
& AGA\_CAC\_283.4 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_285.4 & 0.15 & 0.64 & -0.09 & 0.11 & neutral
\\\addlinespace
& AGA\_CAC\_286.8 & 0.11 & 0.69 & 0.03 & 0.13 & neutral
\\\addlinespace
& AGA\_CAC\_294.8 & 0.04 & 0.83 & 0.00 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_299.4 & 0.41 & 0.46 & -0.11 & 0.11 & neutral
\\\addlinespace
& AGA\_CAC\_308.2 & 0.33 & 0.57 & -0.23 & 0.11 & neutral
\\\addlinespace
& AGA\_CAC\_314.3 & 0.19 & 0.54 & -0.13 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_316.2 & 0.04 & 0.82 & 0.04 & 0.13 & neutral
\\\addlinespace
& AGA\_CAC\_318.3 & 0.37 & 0.54 & 0.13 & 0.14 & neutral
\\\addlinespace
& AGA\_CAC\_321.0 & 0.11 & 0.65 & 0.01 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_324.2 & 0.33 & 0.58 & 0.12 & 0.14 & neutral
\\\addlinespace
& AGA\_CAC\_326.2 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_338.4 & 0.04 & 0.82 & -0.05 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_356.5 & 0.04 & 0.78 & 0.00 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_441.7 & 0.04 & 0.82 & -0.03 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_445.1 & 0.07 & 0.77 & -0.01 & 0.12 & neutral
\\\addlinespace
& AGA\_CAC\_475.7 & 0.37 & 0.54 & 0.06 & 0.13 & neutral
\\\addlinespace
& AGA\_CAC\_477.9 & 0.67 & 0.29 & -0.25 & 0.13 & neutral
\\\addlinespace
& AGA\_CAC\_487.7 & 0.04 & 0.82 & -0.01 & 0.12 & neutral
\\\addlinespace
& AGT\_CTG\_85.2 & 0.33 & 0.57 & -0.13 & 0.11 & neutral
\\\addlinespace
& AGT\_CTG\_109.9 & 0.11 & 0.66 & 0.01 & 0.13 & neutral
\\\addlinespace
& AGT\_CTG\_127.9 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
\\\addlinespace
& AGT\_CTG\_130.9 & 0.11 & 0.66 & -0.03 & 0.12 & neutral
\\\addlinespace
& AGT\_CTG\_135.5 & 0.07 & 0.75 & -0.06 & 0.12 & neutral
\\\addlinespace
& AGT\_CTG\_144.4 & 0.19 & 0.56 & 0.06 & 0.13 & neutral
\\\addlinespace
& AGT\_CTG\_151.4 & 0.07 & 0.70 & -0.02 & 0.12 & neutral
\\\addlinespace
& AGT\_CTG\_152.8 & 0.15 & 0.64 & -0.05 & 0.12 & neutral
\\\addlinespace
& AGT\_CTG\_181.2 & 0.07 & 0.78 & -0.01 & 0.12 & neutral
\\\addlinespace
& AGT\_CTG\_191.2 & 0.33 & 0.58 & -1.05 & 0.09 & neutral
\\\addlinespace
& AGT\_CTG\_196.4 & 0.37 & 0.55 & -0.28 & 0.10 & neutral
\\\addlinespace
& AGT\_CTG\_202.3 & 0.04 & 0.82 & 0.00 & 0.12 & neutral
\\\addlinespace
& AGT\_CTG\_218.5 & 0.07 & 0.78 & 0.00 & 0.12 & neutral
\\\addlinespace
& AGT\_CTG\_221.0 & 0.04 & 0.82 & 0.03 & 0.13 & neutral
\\\addlinespace
& AGT\_CTG\_226.4 & 0.11 & 0.69 & -0.08 & 0.12 & neutral
\\\addlinespace
& AGT\_CTG\_228.9 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
\\\addlinespace
& AGT\_CTG\_230.8 & 0.11 & 0.68 & 0.13 & 0.14 & neutral
\\\addlinespace
& AGT\_CTG\_234.9 & 0.04 & 0.82 & 0.02 & 0.13 & neutral
\\\addlinespace
& AGT\_CTG\_245.4 & 0.07 & 0.75 & -0.01 & 0.12 & neutral
\\\addlinespace
& AGT\_CTG\_262.1 & 0.04 & 0.82 & -0.01 & 0.12 & neutral
\\\addlinespace
& AGT\_CTG\_266.2 & 0.48 & 0.29 & -0.21 & 0.12 & neutral
\\\addlinespace
& AGT\_CTG\_297.5 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
\\\addlinespace
& AGT\_CTG\_336.5 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
\\\addlinespace
& AGT\_CTG\_344.2 & 0.33 & 0.58 & -0.23 & 0.10 & neutral
\\\addlinespace
& AGT\_CTG\_359.0 & 0.44 & 0.44 & -0.47 & 0.10 & neutral
\\\addlinespace
& AGT\_CTG\_363.4 & 0.04 & 0.82 & 0.00 & 0.12 & neutral
\\\addlinespace
& AGT\_CTG\_392.1 & 0.37 & 0.49 & -0.28 & 0.10 & neutral
\\\addlinespace
& AGT\_CTG\_415.3 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
\\\addlinespace
& AGT\_CTG\_443.5 & 0.04 & 0.83 & 0.02 & 0.12 & neutral
\\\addlinespace
& AGT\_CTG\_452.7 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
\\\addlinespace
& AGT\_CTG\_459.7 & 0.37 & 0.55 & 0.25 & 0.16 & neutral
\\\addlinespace
& AGT\_CTG\_488.9 & 0.07 & 0.70 & 0.02 & 0.12 & neutral
\\\addlinespace
\#\#\# Table S3: Genomic MDS results & & & & & &
\\\addlinespace
\bottomrule
\end{longtable}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{knitr::}\KeywordTok{kable}\NormalTok{(}
    \KeywordTok{format.table}\NormalTok{(}
        \KeywordTok{filter}\NormalTok{(}
            \KeywordTok{ldply}\NormalTok{(}
                \KeywordTok{seq_along}\NormalTok{(}\KeywordTok{unique}\NormalTok{(spp.samples.DF$species)), }
                \NormalTok{function(i) \{}
                    \KeywordTok{ldply}\NormalTok{(}
                        \KeywordTok{seq_along}\NormalTok{(spp.mds.LST[[i]]),}
                        \NormalTok{function(j) \{}
                        \KeywordTok{data.frame}\NormalTok{(}
                            \DataTypeTok{Species=}\KeywordTok{paste0}\NormalTok{(}\StringTok{'}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{textit\{'}\NormalTok{,}\KeywordTok{gsub}\NormalTok{(}\StringTok{'}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{_'}\NormalTok{, }\StringTok{' '}\NormalTok{, }\KeywordTok{unique}\NormalTok{(spp.samples.DF$species)[i]),}\StringTok{'\}'}\NormalTok{),}
                            \DataTypeTok{Loci=}\KeywordTok{names}\NormalTok{(spp.mds.LST[[i]])[j],}
                            \DataTypeTok{Stress=}\KeywordTok{ifelse}\NormalTok{(}\KeywordTok{is.null}\NormalTok{(spp.mds.LST[[i]][[j]]), }\OtherTok{NA_real_}\NormalTok{, spp.mds.LST[[i]][[j]]$stress),}
                            \DataTypeTok{Converged=}\KeywordTok{ifelse}\NormalTok{(}\KeywordTok{is.null}\NormalTok{(spp.mds.LST[[i]][[j]]), }\OtherTok{NA_real_}\NormalTok{, spp.mds.LST[[i]][[j]]$converged)}
                        \NormalTok{)}
                    \NormalTok{\})}
                \NormalTok{\}}
            \NormalTok{),}
            \NormalTok{!}\KeywordTok{is.na}\NormalTok{(Stress)}
        \NormalTok{),}
        \DataTypeTok{omit=}\KeywordTok{c}\NormalTok{(}\StringTok{'Loci'}\NormalTok{,}\StringTok{'Converged'}\NormalTok{)}
    \NormalTok{),}
    \DataTypeTok{digits=}\DecValTok{2}\NormalTok{,}
    \DataTypeTok{caption=}\StringTok{'Summary of non-metric multi-dimensional scaling (MDS) analyses on genetic variation for each species.'}\NormalTok{,}
    \DataTypeTok{col.names=}\KeywordTok{c}\NormalTok{(}\StringTok{'Species'}\NormalTok{, }\StringTok{'Loci Type'}\NormalTok{, }\StringTok{'NMDS Stress'}\NormalTok{, }\StringTok{'Converged'}\NormalTok{),}
    \DataTypeTok{align=}\KeywordTok{c}\NormalTok{(}\StringTok{'l'}\NormalTok{, }\StringTok{'c'}\NormalTok{, }\StringTok{'c'}\NormalTok{, }\StringTok{'c'}\NormalTok{)}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[c]{@{}lccc@{}}
\toprule\addlinespace
Species & Loci Type & NMDS Stress & Converged
\\\addlinespace
\midrule\endhead
\textit{Androsace obtusifolia} & adaptive & 0.00 & 0
\\\addlinespace
& neutral & 0.22 & 0
\\\addlinespace
\textit{Arabis alpina} & neutral & 0.24 & 0
\\\addlinespace
\textit{Campanula barbata} & adaptive & 0.00 & 0
\\\addlinespace
& neutral & 0.20 & 0
\\\addlinespace
\bottomrule
\addlinespace
\caption{Summary of non-metric multi-dimensional scaling (MDS) analyses
on genetic variation for each species.}
\end{longtable}

Bothwell, H., Bisbing, S., Therkildsen, N., Crawford, L., Alvarez, N.,
Holderegger, R., Manel, S. (2013) Identifying genetic signatures of
selection in a non-model species, alpine gentian (gentiana nivalis l.),
using a landscape genetic approach. \emph{Conservation Genetics}.
\textbf{14}, 467--481.

Foll, M., Gaggiotti, O. (2008) A genome-scan method to identify selected
loci appropriate for both dominant and codominant markers: A bayesian
perspective. \emph{Genetics}. \textbf{180}, 977--993.

Gugerli, F., Englisch, T., Niklfeld, H., Tribsch, A., Mirek, Z.,
Ronikier, M., Zimmermann, N., Holderegger, R., Taberlet, P. (2008)
Relationships among levels of biodiversity and the relevance of
intraspecific diversity in conservation - a project synopsis.
\emph{Perspectives in Plant Ecology, Evolution and Systematics}.
\textbf{10}, 259--281.

Hijmans, R. J., Cameron, S. E., Parra, J. L., Jones, P. G., Jarvis, A.
(2005) Very high resolution interpolated climate surfaces for global
land areas. \emph{International Journal of Climatology}. \textbf{25},
1965--1978.

Maechler, M., Rousseeuw, P., Struyf, A., Hubert, M., Hornik, K. (2015)
\emph{cluster: Cluster analysis basics and extensions}.

Meirmans, P., Goudet, J., IntraBioDiv Consortium, Gaggiotti, O. (2011)
Ecology and life history affect different aspects of the population
structure of 27 high-alpine plants. \emph{Molecular Ecology}.
\textbf{20}, 3144--3155.

Oksanen, J., Blanchet, F. G., Kindt, R., Legendre, P., Minchin, P. R.,
O'Hara, R. B., Simpson, G. L., Solymos, P., Stevens, M. H. H., Wagner,
H. (2015) \emph{vegan: Community ecology package}.

P{é}rez-Figueroa, A., Garc{í}a-Pereira, M. J., Saura, M.,
Rol{á}n-{Á}lvarez, E., Caballero, A. (2010) Comparing three different
methods to detect selective loci using dominant markers. \emph{Journal
of Evolutionary Biology}. \textbf{23}, 2267--2276.

Vos, P., Hogers, R., Bleeker, M., Reijans, M., Lee, T. van de, Hornes,
M., Friters, A., Pot, J., Paleman, J., Kuiper, M., Zabeau, M. (1995)
AFLP: a new technique for dNA fingerprinting. \emph{Nucleic Acids
Research}. \textbf{23}, 4407--4414.

\end{document}

### 2015-12-15 article/article.Rmd
inclusion of clumpp for post-processing of structure replicate analyses

@@
 Previously, Bothwell _et al._ [-@r455] applied such detection methods to one of the species (_Getiana nivalis_) in this dataset.
-Here, we applied their methods to each of the twenty-seven species in the dataset.
+Here, we applied similar methods to each of the twenty-seven species in the dataset.
 
@@
 We used the Bayesian clustering method implemented in Structure (version 2.3.4) assuming admixture and an independent alleles model.
-For each species, we identified the most plausible number of populations (K = `r st.k[1]`--`r st.k[2]`; `r st.numruns` runs for each K; model with lowest mean posterior negative log likelihood; Figure S3), and estimated the probability that each individual belonged to each population (`r st.burnin` burn-in; `r st.numreps` iterations).
-Individuals with $\geq `r st.probthresh`$ probability of belonging to a single population were assigned accordingly (Figure S4), and used for subseqent detection methods.
+For each species, we identified the most plausible number of populations (K = `r st.k[1]`--`r st.k[2]`; `r st.numruns` runs per K; `r st.burnin` burn-in; `r st.numreps` iterations) using the delta-K method [Figure S3; @r460].
+Population membership probabilities from replicate runs were aligned using CLUMPP [version 1.1.2; `r cl.numreps` runs of the greedy search method with random starts per K per species; @r459], and averaged to estimate the probability that each individual belonged to each population (Figure S4).
+Individuals with a reasonably high probability of belonging to a single population ($\geq `r st.probthresh`$) were assigned accordingly (Figure S5), and used for subseqent locus detection methods.
 
@@
 ```{r}
+# format data
+deltak.DF <- ldply(
+	seq_along(unique(spp.samples.DF$species)),
+	.fun=function(i) {
+		mutate(
+			spp.StructureCollection.LST[[i]]@summary,
+			Species=gsub('_', ' ', unique(spp.samples.DF$species), fixed=TRUE)
+		)
+	}
+)
+# make plot
+ggplot() +
+	geom_point(aes(y=delta.k, x=k)) +
+	geom_line(aes(y=delta.k, x=k)) +
+	theme_classic() +
+	xlab('Number of populations') +
+	ylab('Relative support (delta-K)') +
+	facet_wrap(~ Species, ncol=6)
 ```
 
-### Figure S4. Distribution of populations for each species.
+### Figure S4. Population memberships for each species.
 ```{r}
+# prepare data
+
+# make plot
+
+```
+
+### Figure S5. Distribution of populations for each species.
+```{r}
+# prepare data
+
+# make plot
+
 ```

### 2015-12-15 R/analysis/00-initialization.R

@@ -72,6 +72,10 @@ if (MODE=='debug') {
 	st.noadmix  <- FALSE
 	st.admburnin  <- 10
 	st.probthresh <- 0.75
+	# clumpp parameters
+	cl.repeats <- 1000
+	cl.s <- FALSE
+	cl.m <- 'LargeKGreedy'
 	# BayeScan parameters
 	bs.reps <- 2
 	bs.fdr <- 0.5
@@ -108,6 +112,10 @@ if (MODE=='release') {
 	st.noadmix  <- FALSE
 	st.admburnin  <- 500
 	st.probthresh <- 0.75
+	# clumpp parameters
+	cl.repeats <- 1000
+	cl.s <- FALSE
+	cl.m <- 'LargeKGreedy'
 	# BayeScan parameters
 	bs.reps <- 4
 	bs.fdr <- 0.1

### 2015-12-24 article/article.Rmd
+ updated analysis to use equi-distant crs for attribute space + updated methods section in ms + attempting to set the supp info to be correct

--- 2015-12-13 article/article.Rmd
@@
 author: |
-  | Jeffrey O. Hanson$^1$, Jonathan R. Rhodes$^2$, Cynthia Riginos$^2$, Hugh P. Possingham$^1$, Richard A. Fuller$^1$
+  | Jeffrey O. Hanson$^1$, Jonathan R. Rhodes$^2$, Cynthia Riginos$^1$, Hugh P. Possingham$^1$, Richard A. Fuller$^1$
   | $^1$School of Biological Sciences, The University of Queensland, Brisbane, QLD, Australia
@@
   rmarkdown::pdf_document:
-    toc: true
+    toc: false
     toc_depth: 5
@@
       in_header: preamble-latex.tex
+      after_body: postamble-latex.tex
 fontsize: 11pt
@@
 csl: reference-style.csl
-vignette: >
-  %\VignetteIndexEntry{rapr}
-  %\VignetteEngine{knitr::rmarkdown}
-  \usepackage[utf8]{inputenc}
 ---
+
+<!-- NOTES -->
+<!-- 6 page limit w/o incurring costs -->
 
@@
 
+# load .rda files
+load('results/results.rda')
+
 # load packages
@@
 
-# load .rda files
-load('results/results.rda')
+# set options
+options(error=NULL)
 ```
@@
 ## Introduction
+<!-- 200 word limit -->
 
@@
 ### Study area
-To address the aims of this study, we utilized species distribution and genomic data collected by the IntraBioDiv project [Figure S1; @r451].
-As part of this project, data for 27 alpine plant species were collected across the European Alps using a regular grid [22.3 km $\times$ 25 km; see @r452 for methodological details].
-In every second cell, plant samples were collected from three individiduals for each species present inside the cell.
+To address the aims of the study, we utilized species distribution and genomic data collected by the IntraBioDiv project [Figure S1; @r451; see @r452 for methodological details].
+This dataset was well suited for this study, because it provides spatially explicit genomic data for a multitude of species across a broad geographic area and over large environmental gradients.
+As part of this project, data for 27 alpine plant species were collected across the European Alps according to a grid (22.3 km $\times$ 25 km).
+Briefly, in every second cell, plant samples were collected from three individuals for each species present inside the cell.
 Samples were genotyped using amplified fragment length polymorphisms [AFLP; @r453].
-Matrices denoting the presence/absence of polymorphisms at loci were consutrcted independently for each species.
-This dataset was well suited for the purposes of this study, because it provides spatially explicit genomic data for a multitude of species.
-Additionally, the study area spans across broad-scale environmental gradients--ranging from lowlands to snow-caped peaks--that have been found to correlate with genetic variation for some of these species [].
+Matrices denoting the presence/absence of polymorphisms at loci were constructed independently for each species.
+Thus for each second grid cell, for each of the species in the grid cell, the dataset contains information describing the genomic variation of individuals inside the grid cell.
+We used these grid cells as planning units to develop prioritizations to investigate the effectiveness of surrogates for the (AFLP) genetic variation.
 
 ### Surrogate data
-We explored the effectiveness of environmental and geographic surrogates.
-To describe variation in the location of each grid, we projected the grids to an equal-area coordinate system (ESRI:102014) and determined the centroid of each cell.
-To describe the climatic variation among each cell, we obtained twenty-one bioclimatic variables [@r33].
-These layers were clipped to the extent of the grid and subject to a principle components analysis (PCA; Table S1).
+We explored the effectiveness of climatic and geographic surrogates.
+To describe variation in the location of each grid cell, we projected the grid into an equi-distant coordinate system (ESRI:102031) and used the centroid of each cell.
+To describe the climatic variation among each cell, we obtained twenty-one bioclimatic variables [approx.
+1km resolution; @r33].
+These layers and the grid cells were projected into an equal-area coordinate system (ESRI:102014).
+The bioclimatic variables were clipped to the extent of the grid and subject to a principle components analysis (PCA; Table S1).
 The first two principle components (PCs) cumulatively explained `r round(pca.DF[[4]][2],1)`% of the variation and were used to generate two new layers.
-The average value of each PC layer in each grid was used to characterise the climatic conditions in the grid cell (Figure S2).
+We calculated the average value for each PC layer in each grid cell, and used these values to characterize the main themes of climatic variation among the grid cells (Figure S2).
 
 ### Outlier locus detection
-To investigate the effectiveness of the surrogates for adaptive and neutral genetic variation, we first identifed which of the sampled loci for each species for each species were under selection.
-We used an outlier locus detection method [evaluated in @r454] to avoid circularity issues.
-The basic premise underpinning this method is that neutral loci are expected to have a specific level of variation among populations ($F_{st}$), and loci that deviate from this expectation are under selection.
-Previously, Bothwell _et al._ [-@r455] applied such detection methods to one of the species (_Getiana nivalis_) in this dataset.
-Here, we applied their methods to each of the twenty-seven species in the dataset.
+To investigate the effectiveness of the surrogates for adaptive and neutral genetic variation, we first identified which of the sampled loci for each species were under selection.
+We used outlier locus detection methods [evaluated in @r454] to avoid circularity issues.
+The basic premise underpinning these analyses is that neutral loci are expected to have a specific level of variation among populations ($F_{st}$), and loci that deviate from this expectation should be under selection.
+Thus loci can be identified as being under selection without using environmental data.
+In an earlier study, Bothwell _et al._ [-@r455] applied such outlier detection methods to one of the species in this dataset (_Getiana nivalis_).
+Here, we adopted a similar methodology and applied it to each of the 27 species in the dataset.
 
-Breifly, we grouped individuals into genetic lineages (Figure S2).
+First, we grouped individuals into genetic lineages for each species (Figure S2).
 The purpose of this step was to avoid confounding variation between populations with due to genetic history with variation attributed to selection pressures.
-We used the Bayesian clustering method implemented in Structure (version 2.3.4) assuming admixture and an independent alleles model.
-For each species, we identified the most plausible number of populations (K = `r st.k[1]`--`r st.k[2]`; `r st.numruns` runs for each K; model with lowest mean posterior negative log likelihood; Figure S3), and estimated the probability that each individual belonged to each population (`r st.burnin` burn-in; `r st.numreps` iterations).
-Individuals with $\geq `r st.probthresh`$ probability of belonging to a single population were assigned accordingly (Figure S4), and used for subseqent detection methods.
+We used the Bayesian clustering method implemented in $\texttt{Structure}$ [version 2.3.4; @r461; @r462] assuming admixture and an independent alleles model.
+For each species, we identified the most plausible number of populations (K = `r st.k[1]`--`r st.k[2]`; `r st.numruns` runs per K; `r st.burnin` burn-in iterations; `r st.numreps` post-burn-in iterations) using the $\Delta$K method [Figure S3; @r460].
+We aggregated the population membership probabilities from replicate runs using $\texttt{CLUMPP}$ [version 1.1.2; large K greedy search method with `r cl.repeats` random starts per K per species; Figure S4; @r459].
+Individuals with a moderate probability of belonging to a single population ($\geq `r st.probthresh`$) were assigned accordingly (Figure S5), and subject to subsequent outlier locus detection methods.
 
-We used Bayescan [version 2.1; @r456] to identify the sampled loci under selection for each species (Table S2).
-For each species, Bayescan was run using the population memberships identified using Structure (1:1 prior odds, `r bs.nbp` pilot runs, and `r bs.n` iterations thinned by `r bs.thin` iterations with a `r bs.burn` burn-in).
-Following guidelines in the Bayescan manual, we omitted loci where the global frequency of the minor allele was $\geq$ 0.05 to avoid false-positives.
-To ensure convergence, we ran `r as.english(bs.reps)` replicates per species using a false discovery rate (FDR) $\leq `r bs.fdr`$.
+We used $\texttt{Bayescan}$ [version 2.1; @r456] to identify loci under selection for each species (Table S2).
+For each species, $\texttt{Bayescan}$ was run using the population memberships identified in the previous analysis (1:1 prior odds; `r bs.nbp` pilot runs; `r bs.burn` burn-in iterations; `r bs.n` post-burn-in iterations thinned by `r bs.thin` iterations).
+Following guidelines in the $\texttt{Bayescan}$ manual, we omitted loci where the global frequency of the minor allele was $\geq 0.05$ to avoid false-positives.
+To ensure convergence, we ran `r bs.reps` replicates per species using a suitable false discovery rate [FDR; $\leq `r bs.fdr`$; @r466].
 
-After classifying the loci as adaptive or neutral for each species, we mapped the main gradients of adaptive and/or neutral genetic variation for each species (Figures S4--S29).
-For each species, we used Gower distances to express the difference in adaptive and/or neutral polymorphisms between individuals [using the cluster R package; @r457].
+After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if present) and neutral genetic variation for each species (Figures S4--S29).
+Note that while individuals not assigned to populations were omitted from the $\texttt{Bayescan}$ analysis, here we used all the individuals.
+For each species, for each type of genetic variation present, we used Gower distances [@r464] to express the genomic differences between individuals [using the cluster R package to accommodate sparsity; @r457].
 These distances were subject to a non metric multi-dimensional scaling analysis [implemented in the vegan R package using K=`r mds.k` and `r mds.trymax` random starts; Table S3; @r458].
-The ordinated variables were then associated with the grid cells were each species was sampled.
+The average of the ordinated values associated with individuals in each grid cell we used to characterize the genomic variation in the grid cell.
 
-### Prioritisations
+### Prioritizations
+We used the unreliable representative and adequate prioritization (URAP) formulation of the reserve selection problem [@r465].
+We treated the genomic and surrogate variables as distinct attribute spaces.
+We set the demand points for each species and its attribute space using the values associated with the planning units.
+First, we generated a set of single species prioritizations.
+For each species, we generated a prioritization using a `r (rapr.amount.target*100)`% amount-based target, a prioritization using a `r (rapr.amount.target*100)`% amount-based target and `r (rapr.surrogate.target*100)`% space-based targets for the surrogate variation, and a prioritization using a `r (rapr.amount.target*100)`% amount-based target and `r (rapr.genetic.target*100) `% space-based targets for the genomic variation.
+Next, we generated three multi-species prioritizations using the same sets of targets used in the single species prioritizations.
+Finally, to understand the relationship between the surrogate-based and genomic-based targets, we generated a suite of prioritizations using increasing space-based targets between `r (min(rapr.pareto.surrogate.targets)*100)`% and `r (max(rapr.pareto.surrogate.targets)*100)`% in increments of `r diff(rapr.pareto.surrogate.targets[1:2])*100`%.
+For each prioritization we calculated the proportion of the adaptive (if present) and neutral genetic variation that it secured.
+We solved the reserve selection problems to within `r (gb.MIPGap*100)`% of optimality using the $\texttt{rapr}$ R package [@r465] and $\texttt{Gurobi}$ [@r463].
 
 ## Results
-### Single species prioritisations
-
-```{r, fig.width=6, fig.height=3, fig.cap='Summary of single species prioritisations.
-Single species prioritisations were generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species.
-Data shows the performance of prioritisations generated using these three sets of targets.
-Bars denote means and standard errors.'}
-# prepare data for plotting
-single.spp.PDF <- expand.grid(
-	Prioritisation=unique(single.spp.SDF$Prioritisation),
-	Metric=unique(single.spp.SDF$Metric))
-single.spp.PDF <- cbind(single.spp.PDF,
-	as.data.frame(predict(single.spp.GLM, single.spp.PDF,
-		type='response', se.fit=TRUE))) %>%
-	mutate(lower=fit-se.fit, upper=fit+se.fit,
-		letters=toupper(cld(single.spp.MCP)$mcletters$Letters),
-		letter_pos=upper+0.05)
-# make plot
-ggplot(aes(x=Metric,y=fit,fill=Prioritisation),
-	data=single.spp.PDF) +
-	geom_bar(position=position_dodge(0.9),
-		stat='identity') +
-	geom_errorbar(
-		aes(ymin=lower,ymax=upper),
-		position=position_dodge(0.9), width=0.6) +
-	geom_text(aes(x=Metric, y=letter_pos,
-		label=letters), position=position_dodge(0.9)) +
-	scale_fill_manual(name='Prioritisation',
-		values=c('grey80','grey50','grey20')) +
-	ylab('Proportion genetic\nvariation secured (%)') +
-	xlab('') +
-	theme_classic()
-```
+### Single species prioritizations
 
@@
 
-```{r, fig.width=8, fig.height=3.5, fig.cap='Multi-species prioritisations.
-Panel (a) shows the prioritisation generated for using just amount-based targets.
-Panel (b) shows the prioritisation generated using amount-based and surrogate based targets.
-Panel (c) shows the prioritisation generated using amount-based and genetic-based targets'}
-# download basemap
-data(countriesHigh)
-countries.FPLY <- countriesHigh[
-	countriesHigh$ADMIN %in% c(
-		'Italy', 'Switzerland', 'France', 'Austria',
-		'Germany', 'Slovenia', 'Croatia', 'Hungary',
-		'Monaco', 'Germany'
-	)
-,] %>% spFortify
-# prepare data for plotting
-multi.spp.grid.FPLY <- grid.PLY
-for (i in seq_along(multi.spp.prioritisations))
-	multi.spp.grid.FPLY@data[[paste0('v',i)]] <- selections(multi.spp.prioritisations[[i]])
-multi.spp.grid.FPLY <- spFortify(multi.spp.grid.FPLY)
-# make maps
-do.call(
-	grid.arrange,
-	append(
-		llply(
-			seq_along(multi.spp.prioritisations),
-			function(i) {
-				ggplot() +
-					geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
-						fill='grey20', color='grey80') +
-					geom_polygon(data=multi.spp.grid.FPLY, aes_string(x='long', y='lat', 
-						group='group', fill=paste0('v',i)),
-						alpha=0.8, color='grey10') +
-					guides(fill=guide_legend(title=' ')) +
-					theme_classic() +
-					theme(axis.ticks=element_blank(), axis.text=element_blank(),
-						plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(),
-						legend.position='none') +
-					coord_cartesian(
-						xlim=buffered.range(multi.spp.grid.FPLY$long, 0.05),
-						ylim=buffered.range(multi.spp.grid.FPLY$lat, 0.05)
-					) +
-					xlab('') +
-					ylab('') +
-					annotate('text',
-						x=min(multi.spp.grid.FPLY$long)+diff(range(multi.spp.grid.FPLY$long)*0.1),
-						y=min(multi.spp.grid.FPLY$lat)+diff(range(multi.spp.grid.FPLY$lat)*1.01),
-						label=letters[i], hjust=1, vjust=1, color='white', size=8)
-			}
-		),
-		list(nrow=1)
-	)
-)
-```
-
-```{r, fig.width=6, fig.height=3, fig.cap='Summary of multi-species prioritisations.
-Three prioritisations were generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species.
-Data shows the performance of these prioritisations based on how much genetic variation they explain.
-Bars denote means and standard errors.'}
-# prepare data for plotting
-multi.spp.PDF <- expand.grid(
-	Prioritisation=unique(multi.spp.SDF$Prioritisation),
-	Metric=unique(multi.spp.SDF$Metric))
-multi.spp.PDF <- cbind(multi.spp.PDF,
-	as.data.frame(predict(multi.spp.GLM, multi.spp.PDF,
-		type='response', se.fit=TRUE))) %>%
-	mutate(lower=fit-se.fit, upper=fit+se.fit,
-		letters=toupper(cld(multi.spp.MCP)$mcletters$Letters),
-		letter_pos=upper+0.05)
-# make plot
-ggplot(aes(x=Metric,y=fit,fill=Prioritisation),
-	data=multi.spp.PDF) +
-	geom_bar(position=position_dodge(0.9),
-		stat='identity') +
-	geom_errorbar(
-		aes(ymin=lower,ymax=upper),
-		position=position_dodge(0.9), width=0.6) +
-	geom_text(aes(x=Metric, y=letter_pos,
-		label=letters), position=position_dodge(0.9)) +
-	scale_fill_manual(name='Prioritisation',
-		values=c('grey80','grey50','grey20')) +
-	ylab('Proportion genetic\nvariation secured (%)') +
-	xlab('') +
-	theme_classic()
-```
-
 ### Pareo-frontier analysis
-
-```{r, fig.width=8, fig.height=3.5, fig.cap='The relationship between surrogates and genetic variation secured in prioritisations.'}
-# make plots
-p1 <- ggplot(data=env.pareto.DF) +
-	geom_line(aes(x=Surrogate.target,y=adaptive.held,group=Species),
-		alpha=0.5) +
-	xlab('Environmental variation secured (%)') +
-	ylab('Adaptive genetic\nvariation secured (%)') +
-	theme_classic()
-p2 <- ggplot(data=geo.pareto.DF) +
-	geom_line(aes(x=Surrogate.target,y=neutral.held,group=Species),
-		alpha=0.5) +
-	xlab('Geographic variation secured (%)') +
-	ylab('Neutral genetic\nvariation secured (%)') +
-	theme_classic()
-grid.arrange(p1, p2, nrow=1)
-```
 
@@
 ## References
-
-## Figures
-
-## Tables
-
-## Supporting Information
-
-### Figure S1: Species distributions
-```{r, fig.width=9.5, fig.height=10, out.width='7.5in', out.height='8.5in', fig.cap='Species distributions.
-Squares represent planning units.
-For a given species, planning units that were found to be inhabited are denoted with bright blue.'}
-## plot map of species distributions
-# fortify data
-grid.FPLY <- spFortify(grid.PLY)
-spp.grid.FPLY <- ldply(unique(spp.samples.DF$species), function(x) {
-		z <- grid.FPLY[,c('long', 'lat', 'group', x),drop=FALSE]
-		names(z)[4] <- 'presence'
-		z$species <- gsub('\\_', ' ', x)
-		return(z)
-	}
-)
-# plot species data
-ggplot() +
-	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
-		fill='grey20', color='grey80') +
-	geom_polygon(data=spp.grid.FPLY, aes(x=long, y=lat, 
-		group=group, fill=presence), alpha=0.8, color='grey10') +
-	theme_classic() +
-	guides(fill=guide_legend(title='Presence')) +
-	theme(axis.ticks=element_blank(), axis.text=element_blank(),
-		axis.line=element_blank()) +
-	coord_cartesian(
-		xlim=buffered.range(grid.FPLY$long, 0.05),
-		ylim=buffered.range(grid.FPLY$lat, 0.05)
-	) +
-	xlab('') +
-	ylab('') +
-	facet_wrap(~ species, ncol=4)
-```
-
-```{r, fig.width=4, fig.height=3.5, fig.cap="Species richness. Squares denote planning units. Planning units with a brighter color are inhabited by more species."}
-# calculate species richness
-grid.PLY$Species_richness <- grid.PLY@data %>%
-	select(5:(4+n.spp)) %>% as.matrix() %>% rowSums()
-
-# plot species richness
-ggplot() +
-	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
-		fill='grey20', color='grey80') +
-	geom_polygon(data=spFortify(grid.PLY), aes(x=long, y=lat, 
-		group=group, fill=Species_richness), alpha=0.8, color='grey10') +
-	guides(fill=guide_legend(title='Count (#)')) +
-	theme_classic() +
-	theme(axis.ticks=element_blank(), axis.text=element_blank(),
-		axis.line=element_blank()) +
-	coord_cartesian(
-		xlim=buffered.range(grid.FPLY$long, 0.05),
-		ylim=buffered.range(grid.FPLY$lat, 0.05)
-	) +
-	xlab('') +
-	ylab('') +
-	ggtitle('Species richness')
-```
-
-### Figure S2: Maps of climatic variation
-```{r, width=6.5, height=4.5, fig.cap='Climatic variation.
-Each panel depicts variation based on a different principle component (PC).
-Sqaures represent planning units.
-The color of each planning unit denotes the average priciniple component value of pixels inside it.
-Planning units with more similar colors have more similar climates regimes.'}
-do.call(
-	grid.arrange,
-		append(
-		llply(grep('^env\\_.*$', names(grid.DF), value=TRUE), function(x) {
-			ggplot() +
-				geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
-					fill='grey20', color='grey80') +
-				geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
-					group='group', fill=x),
-					alpha=0.8, color='grey10') +
-				guides(fill=guide_legend(title=' ')) +
-				theme_classic() +
-				theme(axis.ticks=element_blank(), axis.text=element_blank(),
-					plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank()) +
-				coord_cartesian(
-					xlim=buffered.range(grid.FPLY$long, 0.05),
-					ylim=buffered.range(grid.FPLY$lat, 0.05)
-				) +
-				xlab('') +
-				ylab('') +
-				ggtitle(paste0('PC ', substr(x, nchar(x), nchar(x))))
-		}),
-		list(ncol=2)
-	)
-)
-```
-
-### Figure S3. Number populations in each species.
-```{r}
-```
-
-### Figure S4. Distribution of populations for each species.
-```{r}
-```
-
-```{r}
-plot.spp.mds <- function(i) {
-	# define function
-	make.mds.plot <- function(j,k) {
-		ggplot() +
-			geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
-				fill='grey20', color='grey80') +
-			geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
-				group='group', fill=paste0(unique(spp.samples.DF$species)[i], '_', j, '_d',k)),
-				alpha=0.8, color='grey10') +
-			guides(fill=guide_legend(title=' ')) +
-			theme_classic() +
-			theme(axis.ticks=element_blank(), axis.text=element_blank(),
-				plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank()) +
-			coord_cartesian(
-				xlim=buffered.range(grid.FPLY$long, 0.05),
-				ylim=buffered.range(grid.FPLY$lat, 0.05)
-			) +
-			xlab('') +
-			ylab('') +
-			ggtitle(paste0(j,' (',k,')'))
-	}
-	# init
-	adapt.col <- paste0(unique(spp.samples.DF$species)[i], '_adaptive_d1')
-	neutral.col <- paste0(unique(spp.samples.DF$species)[i], '_neutral_d1')
-	plotLST <- list()
-	# make adaptive plot
-	if (adapt.col %in% names(grid.FPLY)) {
-		for (k in seq_len(mds.k))
-			plotLST <- append(plotLST, list(make.mds.plot('adaptive', k)))
-	} 
-	# make neutral plot
-	if (neutral.col %in% names(grid.FPLY)) {
-		for (k in seq_len(mds.k))
-			plotLST <- append(plotLST, list(make.mds.plot('neutral', k)))
-	} 
-	# make plot
-	do.call(
-		grid.arrange,
-		append(
-			plotLST,
-			list(ncol=6)
-		)
-	)
-}
-```
-
-### Figures S
-```{r, fig.width=6.5, fig.height=4.5, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[1]),'}.
-Each square represents a planning unit.
-The color of each planning unit panel corresponds to ordination values.
-Planning units with similar colors contain individiduals with similar genetic variation.')}
-plot.spp.mds(1)
-```
-
-```{r, fig.width=6.5, fig.height=4.5, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[2]),'}.
-See Figure XX caption for conventions.')}
-plot.spp.mds(2)
-```
-
-```{r, fig.width=6.5, fig.height=4.5, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[3]),'}.
-See Figure XX caption for conventions.')}
-plot.spp.mds(3)
-```
-
-### Table S1: Principle components analysis on climatic variation
-```{r}
-## make results table showing Eigen values
-knitr::kable(
-	pca.DF,
-	digits=2,
-	caption='Summary of principle components analysis (PCA) on bioclimatic variation across the study area. The first two principle components (PCs) were used for subsequent analysis.',
-	align=c('l', 'c', 'c', 'c')
-)
-```
-
-### Table S2: BayeScan Results
-```{r}
-knitr::kable(
-	format.table(
-		ldply(
-			seq_along(unique(spp.samples.DF$species)), 
-			function(i) {
-				data.frame(
-					Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
-					Primer=spp.BayeScan.sample.loci.subset.LST[[i]]@data@primers,
-					Probability=spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[2]],
-					qval=spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[4]],
-					alpha=spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[5]],
-					fst=spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[6]],
-					Type=spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[7]]
-				)
-			}
-		),
-		omit='Type'
-	),
-	digits=2,
-	col.names=c('Species', 'Primer', 'Probability', 'q-value', '$\\alpha$', '$F_{ST}$', 'Type'),
-	align=c('l', 'c', 'c', 'c', 'c', 'c', 'c')
-)
-```
-### Table S3: Genomic MDS results
-```{r}
-knitr::kable(
-	format.table(
-		filter(
-			ldply(
-				seq_along(unique(spp.samples.DF$species)), 
-				function(i) {
-					ldply(
-						seq_along(spp.mds.LST[[i]]),
-						function(j) {
-						data.frame(
-							Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
-							Loci=names(spp.mds.LST[[i]])[j],
-							Stress=ifelse(is.null(spp.mds.LST[[i]][[j]]), NA_real_, spp.mds.LST[[i]][[j]]$stress),
-							Converged=ifelse(is.null(spp.mds.LST[[i]][[j]]), NA_real_, spp.mds.LST[[i]][[j]]$converged)
-						)
-					})
-				}
-			),
-			!is.na(Stress)
-		),
-		omit=c('Loci','Converged')
-	),
-	digits=2,
-	caption='Summary of non-metric multi-dimensional scaling (MDS) analyses on genetic variation for each species.',
-	col.names=c('Species', 'Loci Type', 'NMDS Stress', 'Converged'),
-	align=c('l', 'c', 'c', 'c')
-)
-```

### 2015-12-24 article/article.tex
+ updated analysis to use equi-distant crs for attribute space + updated methods section in ms + attempting to set the supp info to be correct

@@
 \usepackage[margin=1in]{geometry}
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
-\newcommand{\RegionMarkerTok}[1]{{#1}}
-\newcommand{\ErrorTok}[1]{\textbf{{#1}}}
-\newcommand{\NormalTok}[1]{{#1}}
-\usepackage{longtable,booktabs}
-\usepackage{graphicx}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
 \ifxetex
@@
   \posttitle{\par}
-  \author{Jeffrey O. Hanson$^1$, Jonathan R. Rhodes$^2$, Cynthia Riginos$^2$, Hugh
+  \author{Jeffrey O. Hanson$^1$, Jonathan R. Rhodes$^2$, Cynthia Riginos$^1$, Hugh
 P. Possingham$^1$, Richard A. Fuller$^1$\\$^1$School of Biological
@@
   \postdate{\par}
-  \date{13 December 2015}
+  \date{24 December 2015}
 
@@
 
-{
-\hypersetup{linkcolor=black}
-\setcounter{tocdepth}{5}
-\tableofcontents
-}
 \subsection{Introduction}\label{introduction}
@@
 
-To address the aims of this study, we utilized species distribution and
+To address the aims of the study, we utilized species distribution and
 genomic data collected by the IntraBioDiv project (Figure S1; Meirmans
-\emph{et al.} 2011). As part of this project, data for 27 alpine plant
-species were collected across the European Alps using a regular grid
-(22.3 km $\times$ 25 km; see Gugerli \emph{et al.} 2008 for
-methodological details). In every second cell, plant samples were
-collected from three individiduals for each species present inside the
-cell. Samples were genotyped using amplified fragment length
-polymorphisms (AFLP; Vos \emph{et al.} 1995). Matrices denoting the
-presence/absence of polymorphisms at loci were consutrcted independently
-for each species. This dataset was well suited for the purposes of this
-study, because it provides spatially explicit genomic data for a
-multitude of species. Additionally, the study area spans across
-broad-scale environmental gradients--ranging from lowlands to snow-caped
-peaks--that have been found to correlate with genetic variation for some
-of these species {[}{]}.
+\emph{et al.} 2011; see Gugerli \emph{et al.} 2008 for methodological
+details). This dataset was well suited for this study, because it
+provides spatially explicit genomic data for a multitude of species
+across a broad geographic area and over large environmental gradients.
+As part of this project, data for 27 alpine plant species were collected
+across the European Alps according to a grid (22.3 km $\times$ 25 km).
+Briefly, in every second cell, plant samples were collected from three
+individuals for each species present inside the cell. Samples were
+genotyped using amplified fragment length polymorphisms (AFLP; Vos
+\emph{et al.} 1995). Matrices denoting the presence/absence of
+polymorphisms at loci were constructed independently for each species.
+Thus for each second grid cell, for each of the species in the grid
+cell, the dataset contains information describing the genomic variation
+of individuals inside the grid cell. We used these grid cells as
+planning units to develop prioritizations to investigate the
+effectiveness of surrogates for the (AFLP) genetic variation.
 
@@
 
-We explored the effectiveness of environmental and geographic
-surrogates. To describe variation in the location of each grid, we
-projected the grids to an equal-area coordinate system (ESRI:102014) and
-determined the centroid of each cell. To describe the climatic variation
-among each cell, we obtained twenty-one bioclimatic variables (Hijmans
-\emph{et al.} 2005). These layers were clipped to the extent of the grid
-and subject to a principle components analysis (PCA; Table S1). The
-first two principle components (PCs) cumulatively explained 97.2\% of
-the variation and were used to generate two new layers. The average
-value of each PC layer in each grid was used to characterise the
-climatic conditions in the grid cell (Figure S2).
+We explored the effectiveness of climatic and geographic surrogates. To
+describe variation in the location of each grid cell, we projected the
+grid into an equi-distant coordinate system (ESRI:102031) and used the
+centroid of each cell. To describe the climatic variation among each
+cell, we obtained twenty-one bioclimatic variables (approx. 1km
+resolution; Hijmans \emph{et al.} 2005). These layers and the grid cells
+were projected into an equal-area coordinate system (ESRI:102014). The
+bioclimatic variables were clipped to the extent of the grid and subject
+to a principle components analysis (PCA; Table S1). The first two
+principle components (PCs) cumulatively explained 97.2\% of the
+variation and were used to generate two new layers. We calculated the
+average value for each PC layer in each grid cell, and used these values
+to characterize the main themes of climatic variation among the grid
+cells (Figure S2).
 
@@
 To investigate the effectiveness of the surrogates for adaptive and
-neutral genetic variation, we first identifed which of the sampled loci
-for each species for each species were under selection. We used an
-outlier locus detection method (evaluated in P{é}rez-Figueroa \emph{et
-al.} 2010) to avoid circularity issues. The basic premise underpinning
-this method is that neutral loci are expected to have a specific level
-of variation among populations ($F_{st}$), and loci that deviate from
-this expectation are under selection. Previously, Bothwell \emph{et al.}
-(2013) applied such detection methods to one of the species
-(\emph{Getiana nivalis}) in this dataset. Here, we applied their methods
-to each of the twenty-seven species in the dataset.
+neutral genetic variation, we first identified which of the sampled loci
+for each species were under selection. We used outlier locus detection
+methods (evaluated in P{é}rez-Figueroa \emph{et al.} 2010) to avoid
+circularity issues. The basic premise underpinning these analyses is
+that neutral loci are expected to have a specific level of variation
+among populations ($F_{st}$), and loci that deviate from this
+expectation should be under selection. Thus loci can be identified as
+being under selection without using environmental data. In an earlier
+study, Bothwell \emph{et al.} (2013) applied such outlier detection
+methods to one of the species in this dataset (\emph{Getiana nivalis}).
+Here, we adopted a similar methodology and applied it to each of the 27
+species in the dataset.
 
-Breifly, we grouped individuals into genetic lineages (Figure S2). The
-purpose of this step was to avoid confounding variation between
-populations with due to genetic history with variation attributed to
-selection pressures. We used the Bayesian clustering method implemented
-in Structure (version 2.3.4) assuming admixture and an independent
-alleles model. For each species, we identified the most plausible number
-of populations (K = 1--10; 4 runs for each K; model with lowest mean
-posterior negative log likelihood; Figure S3), and estimated the
-probability that each individual belonged to each population (10
-burn-in; 10 iterations). Individuals with $\geq 0.75$ probability of
-belonging to a single population were assigned accordingly (Figure S4),
-and used for subseqent detection methods.
+First, we grouped individuals into genetic lineages for each species
+(Figure S2). The purpose of this step was to avoid confounding variation
+between populations with due to genetic history with variation
+attributed to selection pressures. We used the Bayesian clustering
+method implemented in $\texttt{Structure}$ (version 2.3.4; Pritchard
+\emph{et al.} 2000; Falush \emph{et al.} 2007) assuming admixture and an
+independent alleles model. For each species, we identified the most
+plausible number of populations (K = 1--2; 4 runs per K; 10 burn-in
+iterations; 10 post-burn-in iterations) using the $\Delta$K method
+(Figure S3; Evanno \emph{et al.} 2005). We aggregated the population
+membership probabilities from replicate runs using $\texttt{CLUMPP}$
+(version 1.1.2; large K greedy search method with 1000 random starts per
+K per species; Figure S4; Jakobsson and Rosenberg 2007). Individuals
+with a moderate probability of belonging to a single population
+($\geq 0.75$) were assigned accordingly (Figure S5), and subject to
+subsequent outlier locus detection methods.
 
-We used Bayescan (version 2.1; Foll and Gaggiotti 2008) to identify the
-sampled loci under selection for each species (Table S2). For each
-species, Bayescan was run using the population memberships identified
-using Structure (1:1 prior odds, 3 pilot runs, and 10 iterations thinned
-by 1 iterations with a 40 burn-in). Following guidelines in the Bayescan
-manual, we omitted loci where the global frequency of the minor allele
-was $\geq$ 0.05 to avoid false-positives. To ensure convergence, we ran
-2 replicates per species using a false discovery rate (FDR) $\leq 0.5$.
+We used $\texttt{Bayescan}$ (version 2.1; Foll and Gaggiotti 2008) to
+identify loci under selection for each species (Table S2). For each
+species, $\texttt{Bayescan}$ was run using the population memberships
+identified in the previous analysis (1:1 prior odds; 3 pilot runs; 40
+burn-in iterations; 10 post-burn-in iterations thinned by 1 iterations).
+Following guidelines in the $\texttt{Bayescan}$ manual, we omitted loci
+where the global frequency of the minor allele was $\geq 0.05$ to avoid
+false-positives. To ensure convergence, we ran 2 replicates per species
+using a suitable false discovery rate (FDR; $\leq 0.8$; Benjamini and
+Hochberg 1995).
 
-After classifying the loci as adaptive or neutral for each species, we
-mapped the main gradients of adaptive and/or neutral genetic variation
-for each species (Figures S4--S29). For each species, we used Gower
-distances to express the difference in adaptive and/or neutral
-polymorphisms between individuals (using the cluster R package; Maechler
-\emph{et al.} 2015). These distances were subject to a non metric
-multi-dimensional scaling analysis (implemented in the vegan R package
-using K=2 and 2 random starts; Table S3; Oksanen \emph{et al.} 2015).
-The ordinated variables were then associated with the grid cells were
-each species was sampled.
+After classifying loci as adaptive or neutral, we mapped the main
+gradients of the adaptive (if present) and neutral genetic variation for
+each species (Figures S4--S29). Note that while individuals not assigned
+to populations were omitted from the $\texttt{Bayescan}$ analysis, here
+we used all the individuals. For each species, for each type of genetic
+variation present, we used Gower distances (Gower 1971) to express the
+genomic differences between individuals (using the cluster R package to
+accommodate sparsity; Maechler \emph{et al.} 2015). These distances were
+subject to a non metric multi-dimensional scaling analysis (implemented
+in the vegan R package using K=2 and 2 random starts; Table S3; Oksanen
+\emph{et al.} 2015). The average of the ordinated values associated with
+individuals in each grid cell we used to characterize the genomic
+variation in the grid cell.
 
-\subsubsection{Prioritisations}\label{prioritisations}
+\subsubsection{Prioritizations}\label{prioritizations}
+
+We used the unreliable representative and adequate prioritization (URAP)
+formulation of the reserve selection problem (Hanson \emph{et al.}
+2016). We treated the genomic and surrogate variables as distinct
+attribute spaces. We set the demand points for each species and its
+attribute space using the values associated with the planning units.
+First, we generated a set of single species prioritizations. For each
+species, we generated a prioritization using a 20\% amount-based target,
+a prioritization using a 20\% amount-based target and 90\% space-based
+targets for the surrogate variation, and a prioritization using a 20\%
+amount-based target and 90\% space-based targets for the genomic
+variation. Next, we generated three multi-species prioritizations using
+the same sets of targets used in the single species prioritizations.
+Finally, to understand the relationship between the surrogate-based and
+genomic-based targets, we generated a suite of prioritizations using
+increasing space-based targets between 1\% and 100\% in increments of
+49.5\%. For each prioritization we calculated the proportion of the
+adaptive (if present) and neutral genetic variation that it secured. We
+solved the reserve selection problems to within 90\% of optimality using
+the $\texttt{rapr}$ R package (Hanson \emph{et al.} 2016) and
+$\texttt{Gurobi}$ (Gurobi Optimization 2015).
 
@@
 \subsubsection{Single species
-prioritisations}\label{single-species-prioritisations}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# prepare data for plotting}
-\NormalTok{single.spp.PDF <-}\StringTok{ }\KeywordTok{expand.grid}\NormalTok{(}
-    \DataTypeTok{Prioritisation=}\KeywordTok{unique}\NormalTok{(single.spp.SDF$Prioritisation),}
-    \DataTypeTok{Metric=}\KeywordTok{unique}\NormalTok{(single.spp.SDF$Metric))}
-\NormalTok{single.spp.PDF <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(single.spp.PDF,}
-    \KeywordTok{as.data.frame}\NormalTok{(}\KeywordTok{predict}\NormalTok{(single.spp.GLM, single.spp.PDF,}
-        \DataTypeTok{type=}\StringTok{'response'}\NormalTok{, }\DataTypeTok{se.fit=}\OtherTok{TRUE}\NormalTok{))) %>%}
-\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{lower=}\NormalTok{fit-se.fit, }\DataTypeTok{upper=}\NormalTok{fit+se.fit,}
-        \DataTypeTok{letters=}\KeywordTok{toupper}\NormalTok{(}\KeywordTok{cld}\NormalTok{(single.spp.MCP)$mcletters$Letters),}
-        \DataTypeTok{letter_pos=}\NormalTok{upper}\FloatTok{+0.05}\NormalTok{)}
-\CommentTok{# make plot}
-\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{Metric,}\DataTypeTok{y=}\NormalTok{fit,}\DataTypeTok{fill=}\NormalTok{Prioritisation),}
-    \DataTypeTok{data=}\NormalTok{single.spp.PDF) +}
-\StringTok{    }\KeywordTok{geom_bar}\NormalTok{(}\DataTypeTok{position=}\KeywordTok{position_dodge}\NormalTok{(}\FloatTok{0.9}\NormalTok{),}
-        \DataTypeTok{stat=}\StringTok{'identity'}\NormalTok{) +}
-\StringTok{    }\KeywordTok{geom_errorbar}\NormalTok{(}
-        \KeywordTok{aes}\NormalTok{(}\DataTypeTok{ymin=}\NormalTok{lower,}\DataTypeTok{ymax=}\NormalTok{upper),}
-        \DataTypeTok{position=}\KeywordTok{position_dodge}\NormalTok{(}\FloatTok{0.9}\NormalTok{), }\DataTypeTok{width=}\FloatTok{0.6}\NormalTok{) +}
-\StringTok{    }\KeywordTok{geom_text}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{Metric, }\DataTypeTok{y=}\NormalTok{letter_pos,}
-        \DataTypeTok{label=}\NormalTok{letters), }\DataTypeTok{position=}\KeywordTok{position_dodge}\NormalTok{(}\FloatTok{0.9}\NormalTok{)) +}
-\StringTok{    }\KeywordTok{scale_fill_manual}\NormalTok{(}\DataTypeTok{name=}\StringTok{'Prioritisation'}\NormalTok{,}
-        \DataTypeTok{values=}\KeywordTok{c}\NormalTok{(}\StringTok{'grey80'}\NormalTok{,}\StringTok{'grey50'}\NormalTok{,}\StringTok{'grey20'}\NormalTok{)) +}
-\StringTok{    }\KeywordTok{ylab}\NormalTok{(}\StringTok{'Proportion genetic}\CharTok{\textbackslash{}n}\StringTok{variation secured (%)'}\NormalTok{) +}
-\StringTok{    }\KeywordTok{xlab}\NormalTok{(}\StringTok{''}\NormalTok{) +}
-\StringTok{    }\KeywordTok{theme_classic}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## ymax not defined: adjusting position using y instead
-\end{verbatim}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{article_files/figure-latex/unnamed-chunk-2-1.pdf}
-\caption{Summary of single species prioritisations. Single species
-prioritisations were generated using amount-based targets, amount-based
-and surrogate-based targets, and amount-based and genetic-based targets
-for each species. Data shows the performance of prioritisations
-generated using these three sets of targets. Bars denote means and
-standard errors.}
-\end{figure}
+prioritizations}\label{single-species-prioritizations}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# download basemap}
-\KeywordTok{data}\NormalTok{(countriesHigh)}
-\NormalTok{countries.FPLY <-}\StringTok{ }\NormalTok{countriesHigh[}
-    \NormalTok{countriesHigh$ADMIN %in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}
-        \StringTok{'Italy'}\NormalTok{, }\StringTok{'Switzerland'}\NormalTok{, }\StringTok{'France'}\NormalTok{, }\StringTok{'Austria'}\NormalTok{,}
-        \StringTok{'Germany'}\NormalTok{, }\StringTok{'Slovenia'}\NormalTok{, }\StringTok{'Croatia'}\NormalTok{, }\StringTok{'Hungary'}\NormalTok{,}
-        \StringTok{'Monaco'}\NormalTok{, }\StringTok{'Germany'}
-    \NormalTok{)}
-\NormalTok{,] %>%}\StringTok{ }\NormalTok{spFortify}
-\CommentTok{# prepare data for plotting}
-\NormalTok{multi.spp.grid.FPLY <-}\StringTok{ }\NormalTok{grid.PLY}
-\NormalTok{for (i in }\KeywordTok{seq_along}\NormalTok{(multi.spp.prioritisations))}
-    \NormalTok{multi.spp.grid.FPLY@data[[}\KeywordTok{paste0}\NormalTok{(}\StringTok{'v'}\NormalTok{,i)]] <-}\StringTok{ }\KeywordTok{selections}\NormalTok{(multi.spp.prioritisations[[i]])}
-\NormalTok{multi.spp.grid.FPLY <-}\StringTok{ }\KeywordTok{spFortify}\NormalTok{(multi.spp.grid.FPLY)}
-\CommentTok{# make maps}
-\KeywordTok{do.call}\NormalTok{(}
-    \NormalTok{grid.arrange,}
-    \KeywordTok{append}\NormalTok{(}
-        \KeywordTok{llply}\NormalTok{(}
-            \KeywordTok{seq_along}\NormalTok{(multi.spp.prioritisations),}
-            \NormalTok{function(i) \{}
-                \KeywordTok{ggplot}\NormalTok{() +}
-\StringTok{                    }\KeywordTok{geom_polygon}\NormalTok{(}\DataTypeTok{data=}\NormalTok{countries.FPLY, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{long, }\DataTypeTok{y=}\NormalTok{lat, }\DataTypeTok{group=}\NormalTok{group),}
-                        \DataTypeTok{fill=}\StringTok{'grey20'}\NormalTok{, }\DataTypeTok{color=}\StringTok{'grey80'}\NormalTok{) +}
-\StringTok{                    }\KeywordTok{geom_polygon}\NormalTok{(}\DataTypeTok{data=}\NormalTok{multi.spp.grid.FPLY, }\KeywordTok{aes_string}\NormalTok{(}\DataTypeTok{x=}\StringTok{'long'}\NormalTok{, }\DataTypeTok{y=}\StringTok{'lat'}\NormalTok{, }
-                        \DataTypeTok{group=}\StringTok{'group'}\NormalTok{, }\DataTypeTok{fill=}\KeywordTok{paste0}\NormalTok{(}\StringTok{'v'}\NormalTok{,i)),}
-                        \DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{, }\DataTypeTok{color=}\StringTok{'grey10'}\NormalTok{) +}
-\StringTok{                    }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{fill=}\KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{title=}\StringTok{' '}\NormalTok{)) +}
-\StringTok{                    }\KeywordTok{theme_classic}\NormalTok{() +}
-\StringTok{                    }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.ticks=}\KeywordTok{element_blank}\NormalTok{(), }\DataTypeTok{axis.text=}\KeywordTok{element_blank}\NormalTok{(),}
-                        \DataTypeTok{plot.margin=}\KeywordTok{unit}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{),}\StringTok{'cm'}\NormalTok{), }\DataTypeTok{axis.line=}\KeywordTok{element_blank}\NormalTok{(),}
-                        \DataTypeTok{legend.position=}\StringTok{'none'}\NormalTok{) +}
-\StringTok{                    }\KeywordTok{coord_cartesian}\NormalTok{(}
-                        \DataTypeTok{xlim=}\KeywordTok{buffered.range}\NormalTok{(multi.spp.grid.FPLY$long, }\FloatTok{0.05}\NormalTok{),}
-                        \DataTypeTok{ylim=}\KeywordTok{buffered.range}\NormalTok{(multi.spp.grid.FPLY$lat, }\FloatTok{0.05}\NormalTok{)}
-                    \NormalTok{) +}
-\StringTok{                    }\KeywordTok{xlab}\NormalTok{(}\StringTok{''}\NormalTok{) +}
-\StringTok{                    }\KeywordTok{ylab}\NormalTok{(}\StringTok{''}\NormalTok{) +}
-\StringTok{                    }\KeywordTok{annotate}\NormalTok{(}\StringTok{'text'}\NormalTok{,}
-                        \DataTypeTok{x=}\KeywordTok{min}\NormalTok{(multi.spp.grid.FPLY$long)+}\KeywordTok{diff}\NormalTok{(}\KeywordTok{range}\NormalTok{(multi.spp.grid.FPLY$long)*}\FloatTok{0.1}\NormalTok{),}
-                        \DataTypeTok{y=}\KeywordTok{min}\NormalTok{(multi.spp.grid.FPLY$lat)+}\KeywordTok{diff}\NormalTok{(}\KeywordTok{range}\NormalTok{(multi.spp.grid.FPLY$lat)*}\FloatTok{1.01}\NormalTok{),}
-                        \DataTypeTok{label=}\NormalTok{letters[i], }\DataTypeTok{hjust=}\DecValTok{1}\NormalTok{, }\DataTypeTok{vjust=}\DecValTok{1}\NormalTok{, }\DataTypeTok{color=}\StringTok{'white'}\NormalTok{, }\DataTypeTok{size=}\DecValTok{8}\NormalTok{)}
-            \NormalTok{\}}
-        \NormalTok{),}
-        \KeywordTok{list}\NormalTok{(}\DataTypeTok{nrow=}\DecValTok{1}\NormalTok{)}
-    \NormalTok{)}
-\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{article_files/figure-latex/unnamed-chunk-3-1.pdf}
-\caption{Multi-species prioritisations. Panel (a) shows the
-prioritisation generated for using just amount-based targets. Panel (b)
-shows the prioritisation generated using amount-based and surrogate
-based targets. Panel (c) shows the prioritisation generated using
-amount-based and genetic-based targets}
-\end{figure}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# prepare data for plotting}
-\NormalTok{multi.spp.PDF <-}\StringTok{ }\KeywordTok{expand.grid}\NormalTok{(}
-    \DataTypeTok{Prioritisation=}\KeywordTok{unique}\NormalTok{(multi.spp.SDF$Prioritisation),}
-    \DataTypeTok{Metric=}\KeywordTok{unique}\NormalTok{(multi.spp.SDF$Metric))}
-\NormalTok{multi.spp.PDF <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(multi.spp.PDF,}
-    \KeywordTok{as.data.frame}\NormalTok{(}\KeywordTok{predict}\NormalTok{(multi.spp.GLM, multi.spp.PDF,}
-        \DataTypeTok{type=}\StringTok{'response'}\NormalTok{, }\DataTypeTok{se.fit=}\OtherTok{TRUE}\NormalTok{))) %>%}
-\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{lower=}\NormalTok{fit-se.fit, }\DataTypeTok{upper=}\NormalTok{fit+se.fit,}
-        \DataTypeTok{letters=}\KeywordTok{toupper}\NormalTok{(}\KeywordTok{cld}\NormalTok{(multi.spp.MCP)$mcletters$Letters),}
-        \DataTypeTok{letter_pos=}\NormalTok{upper}\FloatTok{+0.05}\NormalTok{)}
-\CommentTok{# make plot}
-\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{Metric,}\DataTypeTok{y=}\NormalTok{fit,}\DataTypeTok{fill=}\NormalTok{Prioritisation),}
-    \DataTypeTok{data=}\NormalTok{multi.spp.PDF) +}
-\StringTok{    }\KeywordTok{geom_bar}\NormalTok{(}\DataTypeTok{position=}\KeywordTok{position_dodge}\NormalTok{(}\FloatTok{0.9}\NormalTok{),}
-        \DataTypeTok{stat=}\StringTok{'identity'}\NormalTok{) +}
-\StringTok{    }\KeywordTok{geom_errorbar}\NormalTok{(}
-        \KeywordTok{aes}\NormalTok{(}\DataTypeTok{ymin=}\NormalTok{lower,}\DataTypeTok{ymax=}\NormalTok{upper),}
-        \DataTypeTok{position=}\KeywordTok{position_dodge}\NormalTok{(}\FloatTok{0.9}\NormalTok{), }\DataTypeTok{width=}\FloatTok{0.6}\NormalTok{) +}
-\StringTok{    }\KeywordTok{geom_text}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{Metric, }\DataTypeTok{y=}\NormalTok{letter_pos,}
-        \DataTypeTok{label=}\NormalTok{letters), }\DataTypeTok{position=}\KeywordTok{position_dodge}\NormalTok{(}\FloatTok{0.9}\NormalTok{)) +}
-\StringTok{    }\KeywordTok{scale_fill_manual}\NormalTok{(}\DataTypeTok{name=}\StringTok{'Prioritisation'}\NormalTok{,}
-        \DataTypeTok{values=}\KeywordTok{c}\NormalTok{(}\StringTok{'grey80'}\NormalTok{,}\StringTok{'grey50'}\NormalTok{,}\StringTok{'grey20'}\NormalTok{)) +}
-\StringTok{    }\KeywordTok{ylab}\NormalTok{(}\StringTok{'Proportion genetic}\CharTok{\textbackslash{}n}\StringTok{variation secured (%)'}\NormalTok{) +}
-\StringTok{    }\KeywordTok{xlab}\NormalTok{(}\StringTok{''}\NormalTok{) +}
-\StringTok{    }\KeywordTok{theme_classic}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## ymax not defined: adjusting position using y instead
-\end{verbatim}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{article_files/figure-latex/unnamed-chunk-4-1.pdf}
-\caption{Summary of multi-species prioritisations. Three prioritisations
-were generated using amount-based targets, amount-based and
-surrogate-based targets, and amount-based and genetic-based targets for
-each species. Data shows the performance of these prioritisations based
-on how much genetic variation they explain. Bars denote means and
-standard errors.}
-\end{figure}
-
 \subsubsection{Pareo-frontier analysis}\label{pareo-frontier-analysis}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# make plots}
-\NormalTok{p1 <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(}\DataTypeTok{data=}\NormalTok{env.pareto.DF) +}
-\StringTok{    }\KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{Surrogate.target,}\DataTypeTok{y=}\NormalTok{adaptive.held,}\DataTypeTok{group=}\NormalTok{Species),}
-        \DataTypeTok{alpha=}\FloatTok{0.5}\NormalTok{) +}
-\StringTok{    }\KeywordTok{xlab}\NormalTok{(}\StringTok{'Environmental variation secured (%)'}\NormalTok{) +}
-\StringTok{    }\KeywordTok{ylab}\NormalTok{(}\StringTok{'Adaptive genetic}\CharTok{\textbackslash{}n}\StringTok{variation secured (%)'}\NormalTok{) +}
-\StringTok{    }\KeywordTok{theme_classic}\NormalTok{()}
-\NormalTok{p2 <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(}\DataTypeTok{data=}\NormalTok{geo.pareto.DF) +}
-\StringTok{    }\KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{Surrogate.target,}\DataTypeTok{y=}\NormalTok{neutral.held,}\DataTypeTok{group=}\NormalTok{Species),}
-        \DataTypeTok{alpha=}\FloatTok{0.5}\NormalTok{) +}
-\StringTok{    }\KeywordTok{xlab}\NormalTok{(}\StringTok{'Geographic variation secured (%)'}\NormalTok{) +}
-\StringTok{    }\KeywordTok{ylab}\NormalTok{(}\StringTok{'Neutral genetic}\CharTok{\textbackslash{}n}\StringTok{variation secured (%)'}\NormalTok{) +}
-\StringTok{    }\KeywordTok{theme_classic}\NormalTok{()}
-\KeywordTok{grid.arrange}\NormalTok{(p1, p2, }\DataTypeTok{nrow=}\DecValTok{1}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{article_files/figure-latex/unnamed-chunk-5-1.pdf}
-\caption{The relationship between surrogates and genetic variation
-secured in prioritisations.}
-\end{figure}
 
@@
 
-\subsection{Figures}\label{figures}
-
-\subsection{Tables}\label{tables}
-
-\subsection{Supporting Information}\label{supporting-information}
-
-\subsubsection{Figure S1: Species
-distributions}\label{figure-s1-species-distributions}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## plot map of species distributions}
-\CommentTok{# fortify data}
-\NormalTok{grid.FPLY <-}\StringTok{ }\KeywordTok{spFortify}\NormalTok{(grid.PLY)}
-\NormalTok{spp.grid.FPLY <-}\StringTok{ }\KeywordTok{ldply}\NormalTok{(}\KeywordTok{unique}\NormalTok{(spp.samples.DF$species), function(x) \{}
-        \NormalTok{z <-}\StringTok{ }\NormalTok{grid.FPLY[,}\KeywordTok{c}\NormalTok{(}\StringTok{'long'}\NormalTok{, }\StringTok{'lat'}\NormalTok{, }\StringTok{'group'}\NormalTok{, x),drop=}\OtherTok{FALSE}\NormalTok{]}
-        \KeywordTok{names}\NormalTok{(z)[}\DecValTok{4}\NormalTok{] <-}\StringTok{ 'presence'}
-        \NormalTok{z$species <-}\StringTok{ }\KeywordTok{gsub}\NormalTok{(}\StringTok{'}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{_'}\NormalTok{, }\StringTok{' '}\NormalTok{, x)}
-        \KeywordTok{return}\NormalTok{(z)}
-    \NormalTok{\}}
-\NormalTok{)}
-\CommentTok{# plot species data}
-\KeywordTok{ggplot}\NormalTok{() +}
-\StringTok{    }\KeywordTok{geom_polygon}\NormalTok{(}\DataTypeTok{data=}\NormalTok{countries.FPLY, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{long, }\DataTypeTok{y=}\NormalTok{lat, }\DataTypeTok{group=}\NormalTok{group),}
-        \DataTypeTok{fill=}\StringTok{'grey20'}\NormalTok{, }\DataTypeTok{color=}\StringTok{'grey80'}\NormalTok{) +}
-\StringTok{    }\KeywordTok{geom_polygon}\NormalTok{(}\DataTypeTok{data=}\NormalTok{spp.grid.FPLY, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{long, }\DataTypeTok{y=}\NormalTok{lat, }
-        \DataTypeTok{group=}\NormalTok{group, }\DataTypeTok{fill=}\NormalTok{presence), }\DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{, }\DataTypeTok{color=}\StringTok{'grey10'}\NormalTok{) +}
-\StringTok{    }\KeywordTok{theme_classic}\NormalTok{() +}
-\StringTok{    }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{fill=}\KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{title=}\StringTok{'Presence'}\NormalTok{)) +}
-\StringTok{    }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.ticks=}\KeywordTok{element_blank}\NormalTok{(), }\DataTypeTok{axis.text=}\KeywordTok{element_blank}\NormalTok{(),}
-        \DataTypeTok{axis.line=}\KeywordTok{element_blank}\NormalTok{()) +}
-\StringTok{    }\KeywordTok{coord_cartesian}\NormalTok{(}
-        \DataTypeTok{xlim=}\KeywordTok{buffered.range}\NormalTok{(grid.FPLY$long, }\FloatTok{0.05}\NormalTok{),}
-        \DataTypeTok{ylim=}\KeywordTok{buffered.range}\NormalTok{(grid.FPLY$lat, }\FloatTok{0.05}\NormalTok{)}
-    \NormalTok{) +}
-\StringTok{    }\KeywordTok{xlab}\NormalTok{(}\StringTok{''}\NormalTok{) +}
-\StringTok{    }\KeywordTok{ylab}\NormalTok{(}\StringTok{''}\NormalTok{) +}
-\StringTok{    }\KeywordTok{facet_wrap}\NormalTok{(~}\StringTok{ }\NormalTok{species, }\DataTypeTok{ncol=}\DecValTok{4}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}
-\includegraphics[width=7.5in,height=8.5in]{article_files/figure-latex/unnamed-chunk-6-1} \caption{Species distributions.
-Squares represent planning units.
-For a given species, planning units that were found to be inhabited are denoted with bright blue.}\label{fig:unnamed-chunk-6}
-\end{figure}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# calculate species richness}
-\NormalTok{grid.PLY$Species_richness <-}\StringTok{ }\NormalTok{grid.PLY@data %>%}
-\StringTok{    }\KeywordTok{select}\NormalTok{(}\DecValTok{5}\NormalTok{:(}\DecValTok{4}\NormalTok{+n.spp)) %>%}\StringTok{ }\KeywordTok{as.matrix}\NormalTok{() %>%}\StringTok{ }\KeywordTok{rowSums}\NormalTok{()}
-
-\CommentTok{# plot species richness}
-\KeywordTok{ggplot}\NormalTok{() +}
-\StringTok{    }\KeywordTok{geom_polygon}\NormalTok{(}\DataTypeTok{data=}\NormalTok{countries.FPLY, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{long, }\DataTypeTok{y=}\NormalTok{lat, }\DataTypeTok{group=}\NormalTok{group),}
-        \DataTypeTok{fill=}\StringTok{'grey20'}\NormalTok{, }\DataTypeTok{color=}\StringTok{'grey80'}\NormalTok{) +}
-\StringTok{    }\KeywordTok{geom_polygon}\NormalTok{(}\DataTypeTok{data=}\KeywordTok{spFortify}\NormalTok{(grid.PLY), }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{long, }\DataTypeTok{y=}\NormalTok{lat, }
-        \DataTypeTok{group=}\NormalTok{group, }\DataTypeTok{fill=}\NormalTok{Species_richness), }\DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{, }\DataTypeTok{color=}\StringTok{'grey10'}\NormalTok{) +}
-\StringTok{    }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{fill=}\KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{title=}\StringTok{'Count (#)'}\NormalTok{)) +}
-\StringTok{    }\KeywordTok{theme_classic}\NormalTok{() +}
-\StringTok{    }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.ticks=}\KeywordTok{element_blank}\NormalTok{(), }\DataTypeTok{axis.text=}\KeywordTok{element_blank}\NormalTok{(),}
-        \DataTypeTok{axis.line=}\KeywordTok{element_blank}\NormalTok{()) +}
-\StringTok{    }\KeywordTok{coord_cartesian}\NormalTok{(}
-        \DataTypeTok{xlim=}\KeywordTok{buffered.range}\NormalTok{(grid.FPLY$long, }\FloatTok{0.05}\NormalTok{),}
-        \DataTypeTok{ylim=}\KeywordTok{buffered.range}\NormalTok{(grid.FPLY$lat, }\FloatTok{0.05}\NormalTok{)}
-    \NormalTok{) +}
-\StringTok{    }\KeywordTok{xlab}\NormalTok{(}\StringTok{''}\NormalTok{) +}
-\StringTok{    }\KeywordTok{ylab}\NormalTok{(}\StringTok{''}\NormalTok{) +}
-\StringTok{    }\KeywordTok{ggtitle}\NormalTok{(}\StringTok{'Species richness'}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{article_files/figure-latex/unnamed-chunk-7-1.pdf}
-\caption{Species richness. Squares denote planning units. Planning units
-with a brighter color are inhabited by more species.}
-\end{figure}
-
-\subsubsection{Figure S2: Maps of climatic
-variation}\label{figure-s2-maps-of-climatic-variation}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{do.call}\NormalTok{(}
-    \NormalTok{grid.arrange,}
-        \KeywordTok{append}\NormalTok{(}
-        \KeywordTok{llply}\NormalTok{(}\KeywordTok{grep}\NormalTok{(}\StringTok{'^env}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{_.*$'}\NormalTok{, }\KeywordTok{names}\NormalTok{(grid.DF), }\DataTypeTok{value=}\OtherTok{TRUE}\NormalTok{), function(x) \{}
-            \KeywordTok{ggplot}\NormalTok{() +}
-\StringTok{                }\KeywordTok{geom_polygon}\NormalTok{(}\DataTypeTok{data=}\NormalTok{countries.FPLY, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{long, }\DataTypeTok{y=}\NormalTok{lat, }\DataTypeTok{group=}\NormalTok{group),}
-                    \DataTypeTok{fill=}\StringTok{'grey20'}\NormalTok{, }\DataTypeTok{color=}\StringTok{'grey80'}\NormalTok{) +}
-\StringTok{                }\KeywordTok{geom_polygon}\NormalTok{(}\DataTypeTok{data=}\NormalTok{grid.FPLY, }\KeywordTok{aes_string}\NormalTok{(}\DataTypeTok{x=}\StringTok{'long'}\NormalTok{, }\DataTypeTok{y=}\StringTok{'lat'}\NormalTok{, }
-                    \DataTypeTok{group=}\StringTok{'group'}\NormalTok{, }\DataTypeTok{fill=}\NormalTok{x),}
-                    \DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{, }\DataTypeTok{color=}\StringTok{'grey10'}\NormalTok{) +}
-\StringTok{                }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{fill=}\KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{title=}\StringTok{' '}\NormalTok{)) +}
-\StringTok{                }\KeywordTok{theme_classic}\NormalTok{() +}
-\StringTok{                }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.ticks=}\KeywordTok{element_blank}\NormalTok{(), }\DataTypeTok{axis.text=}\KeywordTok{element_blank}\NormalTok{(),}
-                    \DataTypeTok{plot.margin=}\KeywordTok{unit}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{),}\StringTok{'cm'}\NormalTok{), }\DataTypeTok{axis.line=}\KeywordTok{element_blank}\NormalTok{()) +}
-\StringTok{                }\KeywordTok{coord_cartesian}\NormalTok{(}
-                    \DataTypeTok{xlim=}\KeywordTok{buffered.range}\NormalTok{(grid.FPLY$long, }\FloatTok{0.05}\NormalTok{),}
-                    \DataTypeTok{ylim=}\KeywordTok{buffered.range}\NormalTok{(grid.FPLY$lat, }\FloatTok{0.05}\NormalTok{)}
-                \NormalTok{) +}
-\StringTok{                }\KeywordTok{xlab}\NormalTok{(}\StringTok{''}\NormalTok{) +}
-\StringTok{                }\KeywordTok{ylab}\NormalTok{(}\StringTok{''}\NormalTok{) +}
-\StringTok{                }\KeywordTok{ggtitle}\NormalTok{(}\KeywordTok{paste0}\NormalTok{(}\StringTok{'PC '}\NormalTok{, }\KeywordTok{substr}\NormalTok{(x, }\KeywordTok{nchar}\NormalTok{(x), }\KeywordTok{nchar}\NormalTok{(x))))}
-        \NormalTok{\}),}
-        \KeywordTok{list}\NormalTok{(}\DataTypeTok{ncol=}\DecValTok{2}\NormalTok{)}
-    \NormalTok{)}
-\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{article_files/figure-latex/unnamed-chunk-8-1.pdf}
-\caption{Climatic variation. Each panel depicts variation based on a
-different principle component (PC). Sqaures represent planning units.
-The color of each planning unit denotes the average priciniple component
-value of pixels inside it. Planning units with more similar colors have
-more similar climates regimes.}
-\end{figure}
-
-\subsubsection{Figure S3. Number populations in each
-species.}\label{figure-s3.-number-populations-in-each-species.}
-
-\subsubsection{Figure S4. Distribution of populations for each
-species.}\label{figure-s4.-distribution-of-populations-for-each-species.}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{plot.spp.mds <-}\StringTok{ }\NormalTok{function(i) \{}
-    \CommentTok{# define function}
-    \NormalTok{make.mds.plot <-}\StringTok{ }\NormalTok{function(j,k) \{}
-        \KeywordTok{ggplot}\NormalTok{() +}
-\StringTok{            }\KeywordTok{geom_polygon}\NormalTok{(}\DataTypeTok{data=}\NormalTok{countries.FPLY, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{long, }\DataTypeTok{y=}\NormalTok{lat, }\DataTypeTok{group=}\NormalTok{group),}
-                \DataTypeTok{fill=}\StringTok{'grey20'}\NormalTok{, }\DataTypeTok{color=}\StringTok{'grey80'}\NormalTok{) +}
-\StringTok{            }\KeywordTok{geom_polygon}\NormalTok{(}\DataTypeTok{data=}\NormalTok{grid.FPLY, }\KeywordTok{aes_string}\NormalTok{(}\DataTypeTok{x=}\StringTok{'long'}\NormalTok{, }\DataTypeTok{y=}\StringTok{'lat'}\NormalTok{, }
-                \DataTypeTok{group=}\StringTok{'group'}\NormalTok{, }\DataTypeTok{fill=}\KeywordTok{paste0}\NormalTok{(}\KeywordTok{unique}\NormalTok{(spp.samples.DF$species)[i], }\StringTok{'_'}\NormalTok{, j, }\StringTok{'_d'}\NormalTok{,k)),}
-                \DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{, }\DataTypeTok{color=}\StringTok{'grey10'}\NormalTok{) +}
-\StringTok{            }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{fill=}\KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{title=}\StringTok{' '}\NormalTok{)) +}
-\StringTok{            }\KeywordTok{theme_classic}\NormalTok{() +}
-\StringTok{            }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.ticks=}\KeywordTok{element_blank}\NormalTok{(), }\DataTypeTok{axis.text=}\KeywordTok{element_blank}\NormalTok{(),}
-                \DataTypeTok{plot.margin=}\KeywordTok{unit}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{),}\StringTok{'cm'}\NormalTok{), }\DataTypeTok{axis.line=}\KeywordTok{element_blank}\NormalTok{()) +}
-\StringTok{            }\KeywordTok{coord_cartesian}\NormalTok{(}
-                \DataTypeTok{xlim=}\KeywordTok{buffered.range}\NormalTok{(grid.FPLY$long, }\FloatTok{0.05}\NormalTok{),}
-                \DataTypeTok{ylim=}\KeywordTok{buffered.range}\NormalTok{(grid.FPLY$lat, }\FloatTok{0.05}\NormalTok{)}
-            \NormalTok{) +}
-\StringTok{            }\KeywordTok{xlab}\NormalTok{(}\StringTok{''}\NormalTok{) +}
-\StringTok{            }\KeywordTok{ylab}\NormalTok{(}\StringTok{''}\NormalTok{) +}
-\StringTok{            }\KeywordTok{ggtitle}\NormalTok{(}\KeywordTok{paste0}\NormalTok{(j,}\StringTok{' ('}\NormalTok{,k,}\StringTok{')'}\NormalTok{))}
-    \NormalTok{\}}
-    \CommentTok{# init}
-    \NormalTok{adapt.col <-}\StringTok{ }\KeywordTok{paste0}\NormalTok{(}\KeywordTok{unique}\NormalTok{(spp.samples.DF$species)[i], }\StringTok{'_adaptive_d1'}\NormalTok{)}
-    \NormalTok{neutral.col <-}\StringTok{ }\KeywordTok{paste0}\NormalTok{(}\KeywordTok{unique}\NormalTok{(spp.samples.DF$species)[i], }\StringTok{'_neutral_d1'}\NormalTok{)}
-    \NormalTok{plotLST <-}\StringTok{ }\KeywordTok{list}\NormalTok{()}
-    \CommentTok{# make adaptive plot}
-    \NormalTok{if (adapt.col %in%}\StringTok{ }\KeywordTok{names}\NormalTok{(grid.FPLY)) \{}
-        \NormalTok{for (k in }\KeywordTok{seq_len}\NormalTok{(mds.k))}
-            \NormalTok{plotLST <-}\StringTok{ }\KeywordTok{append}\NormalTok{(plotLST, }\KeywordTok{list}\NormalTok{(}\KeywordTok{make.mds.plot}\NormalTok{(}\StringTok{'adaptive'}\NormalTok{, k)))}
-    \NormalTok{\} }
-    \CommentTok{# make neutral plot}
-    \NormalTok{if (neutral.col %in%}\StringTok{ }\KeywordTok{names}\NormalTok{(grid.FPLY)) \{}
-        \NormalTok{for (k in }\KeywordTok{seq_len}\NormalTok{(mds.k))}
-            \NormalTok{plotLST <-}\StringTok{ }\KeywordTok{append}\NormalTok{(plotLST, }\KeywordTok{list}\NormalTok{(}\KeywordTok{make.mds.plot}\NormalTok{(}\StringTok{'neutral'}\NormalTok{, k)))}
-    \NormalTok{\} }
-    \CommentTok{# make plot}
-    \KeywordTok{do.call}\NormalTok{(}
-        \NormalTok{grid.arrange,}
-        \KeywordTok{append}\NormalTok{(}
-            \NormalTok{plotLST,}
-            \KeywordTok{list}\NormalTok{(}\DataTypeTok{ncol=}\DecValTok{6}\NormalTok{)}
-        \NormalTok{)}
-    \NormalTok{)}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-\subsubsection{Figures S}\label{figures-s}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{plot.spp.mds}\NormalTok{(}\DecValTok{1}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{article_files/figure-latex/unnamed-chunk-12-1.pdf}
-\caption{Distribution of adaptive and neutral genetic variation in
-\textit{Androsace obtusifolia}. Each square represents a planning unit.
-The color of each planning unit panel corresponds to ordination values.
-Planning units with similar colors contain individiduals with similar
-genetic variation.}
-\end{figure}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{plot.spp.mds}\NormalTok{(}\DecValTok{2}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{article_files/figure-latex/unnamed-chunk-13-1.pdf}
-\caption{Distribution of adaptive and neutral genetic variation in
-\textit{Arabis alpina}. See Figure XX caption for conventions.}
-\end{figure}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{plot.spp.mds}\NormalTok{(}\DecValTok{3}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{article_files/figure-latex/unnamed-chunk-14-1.pdf}
-\caption{Distribution of adaptive and neutral genetic variation in
-\textit{Campanula barbata}. See Figure XX caption for conventions.}
-\end{figure}
-
-\subsubsection{Table S1: Principle components analysis on climatic
-variation}\label{table-s1-principle-components-analysis-on-climatic-variation}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## make results table showing Eigen values}
-\NormalTok{knitr::}\KeywordTok{kable}\NormalTok{(}
-    \NormalTok{pca.DF,}
-    \DataTypeTok{digits=}\DecValTok{2}\NormalTok{,}
-    \DataTypeTok{caption=}\StringTok{'Summary of principle components analysis (PCA) on bioclimatic variation across the study area.
-The first two principle components (PCs) were used for subsequent analysis.'}\NormalTok{,}
-    \DataTypeTok{align=}\KeywordTok{c}\NormalTok{(}\StringTok{'l'}\NormalTok{, }\StringTok{'c'}\NormalTok{, }\StringTok{'c'}\NormalTok{, }\StringTok{'c'}\NormalTok{)}
-\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{longtable}[c]{@{}lccc@{}}
-\toprule\addlinespace
-Principle Component & Eigen Value & Variation explained (\%) &
-Accumulative variation explained (\%)
-\\\addlinespace
-\midrule\endhead
-1 & 216765.14 & 82.67 & 82.67
-\\\addlinespace
-2 & 38177.84 & 14.56 & 97.23
-\\\addlinespace
-3 & 5356.75 & 2.04 & 99.27
-\\\addlinespace
-4 & 1216.67 & 0.46 & 99.73
-\\\addlinespace
-5 & 700.39 & 0.27 & 100.00
-\\\addlinespace
-\bottomrule
-\addlinespace
-\caption{Summary of principle components analysis (PCA) on bioclimatic
-variation across the study area. The first two principle components
-(PCs) were used for subsequent analysis.}
-\end{longtable}
-
-\subsubsection{Table S2: BayeScan
-Results}\label{table-s2-bayescan-results}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{knitr::}\KeywordTok{kable}\NormalTok{(}
-    \KeywordTok{format.table}\NormalTok{(}
-        \KeywordTok{ldply}\NormalTok{(}
-            \KeywordTok{seq_along}\NormalTok{(}\KeywordTok{unique}\NormalTok{(spp.samples.DF$species)), }
-            \NormalTok{function(i) \{}
-                \KeywordTok{data.frame}\NormalTok{(}
-                    \DataTypeTok{Species=}\KeywordTok{paste0}\NormalTok{(}\StringTok{'}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{textit\{'}\NormalTok{,}\KeywordTok{gsub}\NormalTok{(}\StringTok{'}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{_'}\NormalTok{, }\StringTok{' '}\NormalTok{, }\KeywordTok{unique}\NormalTok{(spp.samples.DF$species)[i]),}\StringTok{'\}'}\NormalTok{),}
-                    \DataTypeTok{Primer=}\NormalTok{spp.BayeScan.sample.loci.subset.LST[[i]]@data@primers,}
-                    \DataTypeTok{Probability=}\NormalTok{spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[}\DecValTok{2}\NormalTok{]],}
-                    \DataTypeTok{qval=}\NormalTok{spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[}\DecValTok{4}\NormalTok{]],}
-                    \DataTypeTok{alpha=}\NormalTok{spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[}\DecValTok{5}\NormalTok{]],}
-                    \DataTypeTok{fst=}\NormalTok{spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[}\DecValTok{6}\NormalTok{]],}
-                    \DataTypeTok{Type=}\NormalTok{spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[}\DecValTok{7}\NormalTok{]]}
-                \NormalTok{)}
-            \NormalTok{\}}
-        \NormalTok{),}
-        \DataTypeTok{omit=}\StringTok{'Type'}
-    \NormalTok{),}
-    \DataTypeTok{digits=}\DecValTok{2}\NormalTok{,}
-    \DataTypeTok{col.names=}\KeywordTok{c}\NormalTok{(}\StringTok{'Species'}\NormalTok{, }\StringTok{'Primer'}\NormalTok{, }\StringTok{'Probability'}\NormalTok{, }\StringTok{'q-value'}\NormalTok{, }\StringTok{'$}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{alpha$'}\NormalTok{, }\StringTok{'$F_\{ST\}$'}\NormalTok{, }\StringTok{'Type'}\NormalTok{),}
-    \DataTypeTok{align=}\KeywordTok{c}\NormalTok{(}\StringTok{'l'}\NormalTok{, }\StringTok{'c'}\NormalTok{, }\StringTok{'c'}\NormalTok{, }\StringTok{'c'}\NormalTok{, }\StringTok{'c'}\NormalTok{, }\StringTok{'c'}\NormalTok{, }\StringTok{'c'}\NormalTok{)}
-\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{longtable}[c]{@{}lcccccc@{}}
-\toprule\addlinespace
-Species & Primer & Probability & q-value & $\alpha$ & $F_{ST}$ & Type
-\\\addlinespace
-\midrule\endhead
-\textit{Androsace obtusifolia} & AAC\_CAN\_83.0 & 0.00 & 0.87 & 0.00 &
-0.14 & neutral
-\\\addlinespace
-& AAC\_CAN\_85.0 & 0.00 & 0.87 & 0.00 & 0.14 & neutral
-\\\addlinespace
-& AAC\_CAN\_89.0 & 0.04 & 0.83 & 0.04 & 0.14 & neutral
-\\\addlinespace
-& AAC\_CAN\_91.0 & 0.07 & 0.77 & -0.04 & 0.13 & neutral
-\\\addlinespace
-& AAC\_CAN\_100.0 & 0.07 & 0.77 & 0.01 & 0.14 & neutral
-\\\addlinespace
-& AAC\_CAN\_102.0 & 0.30 & 0.55 & 0.20 & 0.16 & neutral
-\\\addlinespace
-& AAC\_CAN\_108.0 & 0.04 & 0.83 & 0.04 & 0.14 & neutral
-\\\addlinespace
-& AAC\_CAN\_124.0 & 0.33 & 0.57 & -0.46 & 0.10 & neutral
-\\\addlinespace
-& AAC\_CAN\_125.0 & 0.07 & 0.78 & 0.00 & 0.14 & neutral
-\\\addlinespace
-& AAC\_CAN\_128.0 & 0.33 & 0.58 & -0.11 & 0.13 & neutral
-\\\addlinespace
-& AAC\_CAN\_130.0 & 0.19 & 0.59 & NaN & NaN & neutral
-\\\addlinespace
-& AAC\_CAN\_132.0 & 0.30 & 0.55 & 0.05 & 0.14 & neutral
-\\\addlinespace
-& AAC\_CAN\_133.0 & 0.00 & 0.87 & 0.00 & 0.14 & neutral
-\\\addlinespace
-& AAC\_CAN\_136.0 & 0.07 & 0.77 & 0.00 & 0.14 & neutral
-\\\addlinespace
-& AAC\_CAN\_137.0 & 0.19 & 0.59 & 0.14 & 0.16 & neutral
-\\\addlinespace
-& AAC\_CAN\_146.0 & 0.22 & 0.61 & 0.15 & 0.16 & neutral
-\\\addlinespace
-& AAC\_CAN\_151.0 & 0.56 & 0.17 & 0.33 & 0.24 & adaptive
-\\\addlinespace
-& AAC\_CAN\_152.0 & 0.19 & 0.59 & NaN & NaN & neutral
-\\\addlinespace
-& AAC\_CAN\_153.0 & 0.11 & 0.68 & 0.02 & 0.14 & neutral
-\\\addlinespace
-& AAC\_CAN\_182.0 & 0.15 & 0.65 & -0.26 & 0.12 & neutral
-\\\addlinespace
-& AAC\_CAN\_195.0 & 0.07 & 0.78 & 0.02 & 0.14 & neutral
-\\\addlinespace
-& AAC\_CAN\_211.0 & 0.04 & 0.83 & 0.01 & 0.14 & neutral
-\\\addlinespace
-& AAC\_CAN\_220.0 & 0.00 & 0.87 & 0.00 & 0.14 & neutral
-\\\addlinespace
-& AAC\_CAN\_231.0 & 0.48 & 0.37 & -0.18 & 0.12 & neutral
-\\\addlinespace
-& AAC\_CAN\_239.0 & 0.37 & 0.54 & -0.56 & 0.10 & neutral
-\\\addlinespace
-& AAC\_CAN\_272.0 & 0.04 & 0.82 & -0.01 & 0.14 & neutral
-\\\addlinespace
-& AAC\_CAN\_319.0 & 0.07 & 0.77 & 0.05 & 0.14 & neutral
-\\\addlinespace
-& ACA\_CAT\_81.0 & 0.04 & 0.82 & -0.01 & 0.14 & neutral
-\\\addlinespace
-& ACA\_CAT\_85.0 & 0.37 & 0.52 & 0.18 & 0.17 & neutral
-\\\addlinespace
-& ACA\_CAT\_90.0 & 0.15 & 0.61 & 0.03 & 0.14 & neutral
-\\\addlinespace
-& ACA\_CAT\_97.0 & 0.07 & 0.73 & -0.04 & 0.13 & neutral
-\\\addlinespace
-& ACA\_CAT\_99.0 & 0.15 & 0.65 & 0.16 & 0.17 & neutral
-\\\addlinespace
-& ACA\_CAT\_100.0 & 0.07 & 0.76 & 0.02 & 0.14 & neutral
-\\\addlinespace
-& ACA\_CAT\_102.0 & 0.07 & 0.76 & -0.04 & 0.13 & neutral
-\\\addlinespace
-& ACA\_CAT\_103.0 & 0.00 & 0.87 & 0.00 & 0.14 & neutral
-\\\addlinespace
-& ACA\_CAT\_108.0 & 0.15 & 0.64 & -0.10 & 0.13 & neutral
-\\\addlinespace
-& ACA\_CAT\_120.0 & 0.11 & 0.67 & -0.05 & 0.13 & neutral
-\\\addlinespace
-& ACA\_CAT\_124.0 & 0.04 & 0.82 & -0.04 & 0.13 & neutral
-\\\addlinespace
-& ACA\_CAT\_126.0 & 0.04 & 0.82 & 0.05 & 0.15 & neutral
-\\\addlinespace
-& ACA\_CAT\_129.0 & 0.33 & 0.58 & 0.00 & 0.14 & neutral
-\\\addlinespace
-& ACA\_CAT\_131.0 & 0.19 & 0.59 & -0.12 & 0.13 & neutral
-\\\addlinespace
-& ACA\_CAT\_133.0 & 0.70 & 0.24 & -0.35 & 0.12 & neutral
-\\\addlinespace
-& ACA\_CAT\_135.0 & 0.07 & 0.73 & -0.04 & 0.13 & neutral
-\\\addlinespace
-& ACA\_CAT\_140.0 & 0.04 & 0.82 & 0.02 & 0.14 & neutral
-\\\addlinespace
-& ACA\_CAT\_148.0 & 0.04 & 0.82 & 0.03 & 0.14 & neutral
-\\\addlinespace
-& ACA\_CAT\_152.0 & 0.07 & 0.78 & -0.05 & 0.13 & neutral
-\\\addlinespace
-& ACA\_CAT\_153.0 & 0.04 & 0.82 & 0.00 & 0.14 & neutral
-\\\addlinespace
-& ACA\_CAT\_155.0 & 0.07 & 0.76 & 0.01 & 0.14 & neutral
-\\\addlinespace
-& ACA\_CAT\_157.0 & 0.37 & 0.54 & -0.16 & 0.12 & neutral
-\\\addlinespace
-& ACA\_CAT\_159.0 & 0.11 & 0.70 & 0.18 & 0.17 & neutral
-\\\addlinespace
-& ACA\_CAT\_162.0 & 0.07 & 0.73 & 0.01 & 0.14 & neutral
-\\\addlinespace
-& ACA\_CAT\_168.0 & 0.19 & 0.60 & -0.05 & 0.13 & neutral
-\\\addlinespace
-& ACA\_CAT\_173.0 & 0.33 & 0.58 & 0.20 & 0.17 & neutral
-\\\addlinespace
-& ACA\_CAT\_177.0 & 0.00 & 0.87 & 0.00 & 0.14 & neutral
-\\\addlinespace
-& ACA\_CAT\_178.0 & 0.19 & 0.58 & 0.07 & 0.14 & neutral
-\\\addlinespace
-& ACA\_CAT\_187.0 & 0.04 & 0.81 & -0.02 & 0.14 & neutral
-\\\addlinespace
-& ACA\_CAT\_192.0 & 0.19 & 0.59 & 0.09 & 0.15 & neutral
-\\\addlinespace
-& ACA\_CAT\_196.0 & 0.00 & 0.87 & 0.00 & 0.14 & neutral
-\\\addlinespace
-& ACA\_CAT\_199.0 & 0.04 & 0.83 & 0.00 & 0.14 & neutral
-\\\addlinespace
-& ACA\_CAT\_200.0 & 0.04 & 0.83 & -0.02 & 0.14 & neutral
-\\\addlinespace
-& ACA\_CAT\_204.0 & 0.00 & 0.87 & 0.00 & 0.14 & neutral
-\\\addlinespace
-& ACA\_CAT\_205.0 & 0.07 & 0.77 & -0.02 & 0.13 & neutral
-\\\addlinespace
-& ACA\_CAT\_210.0 & 0.15 & 0.61 & -0.03 & 0.13 & neutral
-\\\addlinespace
-& ACA\_CAT\_214.0 & 0.33 & 0.58 & 0.03 & 0.14 & neutral
-\\\addlinespace
-& ACA\_CAT\_219.0 & 0.11 & 0.68 & -0.03 & 0.13 & neutral
-\\\addlinespace
-& ACA\_CAT\_229.0 & 0.15 & 0.61 & -0.04 & 0.13 & neutral
-\\\addlinespace
-& ACA\_CAT\_237.0 & 0.15 & 0.65 & -0.07 & 0.13 & neutral
-\\\addlinespace
-& ACA\_CAT\_243.0 & 0.04 & 0.82 & -0.02 & 0.14 & neutral
-\\\addlinespace
-& ACA\_CAT\_246.0 & 0.04 & 0.82 & -0.01 & 0.14 & neutral
-\\\addlinespace
-& ACA\_CAT\_248.0 & 0.04 & 0.83 & -0.02 & 0.13 & neutral
-\\\addlinespace
-& ACA\_CAT\_282.0 & 0.00 & 0.87 & 0.00 & 0.14 & neutral
-\\\addlinespace
-& ACA\_CAT\_300.0 & 0.04 & 0.81 & -0.02 & 0.14 & neutral
-\\\addlinespace
-& ACA\_CAT\_390.0 & 0.41 & 0.49 & -0.02 & 0.15 & neutral
-\\\addlinespace
-& ACA\_CAT\_391.0 & 0.04 & 0.81 & 0.03 & 0.14 & neutral
-\\\addlinespace
-& ACA\_CAT\_393.0 & 0.04 & 0.83 & -0.01 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_82.3 & 0.04 & 0.83 & -0.03 & 0.13 & neutral
-\\\addlinespace
-& AGG\_CAA\_83.0 & 0.15 & 0.65 & -0.02 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_84.2 & 0.41 & 0.45 & -0.05 & 0.13 & neutral
-\\\addlinespace
-& AGG\_CAA\_86.9 & 0.07 & 0.77 & 0.00 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_90.9 & 0.11 & 0.65 & -0.02 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_94.8 & 0.04 & 0.81 & 0.01 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_95.6 & 0.07 & 0.71 & 0.05 & 0.15 & neutral
-\\\addlinespace
-& AGG\_CAA\_100.0 & 0.07 & 0.78 & -0.01 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_101.0 & 0.07 & 0.73 & -0.13 & 0.13 & neutral
-\\\addlinespace
-& AGG\_CAA\_109.3 & 0.07 & 0.78 & 0.04 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_110.2 & 0.00 & 0.87 & 0.00 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_113.5 & 0.04 & 0.81 & 0.04 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_115.9 & 0.11 & 0.68 & -0.02 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_117.4 & 0.33 & 0.58 & -0.31 & 0.12 & neutral
-\\\addlinespace
-& AGG\_CAA\_118.2 & 0.04 & 0.82 & -0.02 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_122.9 & 0.33 & 0.58 & -0.11 & 0.13 & neutral
-\\\addlinespace
-& AGG\_CAA\_129.5 & 0.37 & 0.42 & 0.43 & 0.23 & neutral
-\\\addlinespace
-& AGG\_CAA\_130.1 & 0.48 & 0.29 & 0.17 & 0.16 & adaptive
-\\\addlinespace
-& AGG\_CAA\_135.5 & 0.37 & 0.52 & -0.10 & 0.13 & neutral
-\\\addlinespace
-& AGG\_CAA\_137.9 & 0.26 & 0.57 & -0.12 & 0.13 & neutral
-\\\addlinespace
-& AGG\_CAA\_144.9 & 0.37 & 0.53 & -0.03 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_145.8 & 0.37 & 0.52 & -0.36 & 0.11 & neutral
-\\\addlinespace
-& AGG\_CAA\_150.9 & 0.00 & 0.87 & 0.00 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_152.8 & 0.07 & 0.74 & 0.04 & 0.15 & neutral
-\\\addlinespace
-& AGG\_CAA\_155.4 & 0.04 & 0.83 & 0.02 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_164.0 & 0.22 & 0.61 & -0.08 & 0.13 & neutral
-\\\addlinespace
-& AGG\_CAA\_175.9 & 0.41 & 0.49 & -0.17 & 0.13 & neutral
-\\\addlinespace
-& AGG\_CAA\_181.6 & 0.04 & 0.83 & 0.00 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_182.1 & 0.00 & 0.87 & 0.00 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_188.1 & 0.07 & 0.73 & 0.01 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_191.3 & 0.07 & 0.77 & 0.01 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_193.9 & 0.04 & 0.83 & 0.02 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_196.8 & 0.11 & 0.70 & 0.02 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_201.7 & 0.07 & 0.77 & -0.04 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_212.4 & 0.07 & 0.76 & -0.02 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_215.6 & 0.15 & 0.65 & -0.05 & 0.13 & neutral
-\\\addlinespace
-& AGG\_CAA\_219.0 & 0.04 & 0.83 & 0.00 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_225.2 & 0.04 & 0.83 & -0.02 & 0.13 & neutral
-\\\addlinespace
-& AGG\_CAA\_253.2 & 0.07 & 0.76 & 0.01 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_256.7 & 0.04 & 0.82 & 0.02 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_262.5 & 0.30 & 0.54 & -0.02 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_263.6 & 0.11 & 0.67 & -0.09 & 0.13 & neutral
-\\\addlinespace
-& AGG\_CAA\_264.8 & 0.04 & 0.81 & -0.01 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_267.9 & 0.41 & 0.49 & 0.28 & 0.17 & neutral
-\\\addlinespace
-& AGG\_CAA\_269.8 & 0.04 & 0.83 & 0.03 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_270.8 & 0.11 & 0.72 & 0.05 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_276.8 & 0.07 & 0.78 & -0.06 & 0.13 & neutral
-\\\addlinespace
-& AGG\_CAA\_299.0 & 0.26 & 0.55 & 0.11 & 0.15 & neutral
-\\\addlinespace
-& AGG\_CAA\_305.1 & 0.11 & 0.69 & -0.06 & 0.13 & neutral
-\\\addlinespace
-& AGG\_CAA\_312.1 & 0.04 & 0.82 & 0.01 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_313.2 & 0.04 & 0.81 & -0.03 & 0.13 & neutral
-\\\addlinespace
-& AGG\_CAA\_316.0 & 0.00 & 0.87 & 0.00 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_319.0 & 0.00 & 0.87 & 0.00 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_324.1 & 0.04 & 0.81 & -0.06 & 0.13 & neutral
-\\\addlinespace
-& AGG\_CAA\_359.4 & 0.11 & 0.70 & -0.02 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_360.5 & 0.19 & 0.60 & -0.05 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_363.2 & 0.07 & 0.73 & -0.05 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_364.1 & 0.07 & 0.78 & 0.03 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_376.2 & 0.11 & 0.68 & 0.05 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_376.7 & 0.00 & 0.87 & 0.00 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_396.0 & 0.30 & 0.54 & -0.25 & 0.11 & neutral
-\\\addlinespace
-& AGG\_CAA\_403.4 & 0.07 & 0.73 & 0.05 & 0.14 & neutral
-\\\addlinespace
-& AGG\_CAA\_420.5 & 0.07 & 0.78 & -0.06 & 0.13 & neutral
-\\\addlinespace
-\textit{Arabis alpina} & AAT\_CAC\_51.9 & 0.30 & 0.57 & 0.51 & 0.26 &
-neutral
-\\\addlinespace
-& AAT\_CAC\_54.7 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_69.0 & 0.22 & 0.52 & 0.03 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_73.6 & 0.04 & 0.81 & -0.03 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_77.9 & 0.37 & 0.53 & 0.25 & 0.21 & neutral
-\\\addlinespace
-& AAT\_CAC\_86.9 & 0.15 & 0.64 & 0.08 & 0.18 & neutral
-\\\addlinespace
-& AAT\_CAC\_90.5 & 0.04 & 0.80 & 0.01 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_95.3 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_97.0 & 0.07 & 0.76 & -0.08 & 0.16 & neutral
-\\\addlinespace
-& AAT\_CAC\_97.1 & 0.07 & 0.76 & 0.02 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_100.3 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_105.4 & 0.33 & 0.57 & 0.13 & 0.19 & neutral
-\\\addlinespace
-& AAT\_CAC\_118.4 & 0.04 & 0.82 & 0.01 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_121.3 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_128.0 & 0.37 & 0.50 & -0.45 & 0.16 & neutral
-\\\addlinespace
-& AAT\_CAC\_130.0 & 0.15 & 0.62 & 0.09 & 0.18 & neutral
-\\\addlinespace
-& AAT\_CAC\_147.4 & 0.33 & 0.57 & 0.07 & 0.19 & neutral
-\\\addlinespace
-& AAT\_CAC\_156.9 & 0.11 & 0.70 & 0.03 & 0.18 & neutral
-\\\addlinespace
-& AAT\_CAC\_175.1 & 0.07 & 0.78 & -0.02 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_177.9 & 0.04 & 0.81 & -0.02 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_179.6 & 0.11 & 0.72 & 0.03 & 0.18 & neutral
-\\\addlinespace
-& AAT\_CAC\_188.6 & 0.07 & 0.75 & -0.01 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_190.0 & 0.37 & 0.54 & 0.22 & 0.21 & neutral
-\\\addlinespace
-& AAT\_CAC\_195.5 & 0.26 & 0.43 & -0.04 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_197.1 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_200.6 & 0.04 & 0.81 & -0.02 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_201.8 & 0.11 & 0.66 & 0.05 & 0.18 & neutral
-\\\addlinespace
-& AAT\_CAC\_209.2 & 0.11 & 0.68 & 0.01 & 0.18 & neutral
-\\\addlinespace
-& AAT\_CAC\_213.1 & 0.04 & 0.80 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_215.0 & 0.07 & 0.78 & -0.01 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_216.2 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_217.0 & 0.30 & 0.55 & 0.13 & 0.20 & neutral
-\\\addlinespace
-& AAT\_CAC\_218.1 & 0.11 & 0.68 & -0.06 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_219.1 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_225.5 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_227.1 & 0.37 & 0.54 & -0.15 & 0.16 & neutral
-\\\addlinespace
-& AAT\_CAC\_229.1 & 0.44 & 0.40 & 0.39 & 0.24 & neutral
-\\\addlinespace
-& AAT\_CAC\_231.4 & 0.04 & 0.82 & -0.03 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_249.7 & 0.07 & 0.76 & 0.01 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_259.3 & 0.33 & 0.58 & -0.05 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_277.3 & 0.04 & 0.80 & -0.01 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_298.1 & 0.11 & 0.71 & -0.01 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_315.9 & 0.19 & 0.58 & NaN & NaN & neutral
-\\\addlinespace
-& AAT\_CAC\_330.4 & 0.07 & 0.76 & 0.02 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_334.8 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_336.6 & 0.11 & 0.72 & -0.07 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_353.0 & 0.04 & 0.82 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_359.2 & 0.07 & 0.76 & -0.08 & 0.16 & neutral
-\\\addlinespace
-& AAT\_CAC\_399.9 & 0.04 & 0.82 & 0.02 & 0.18 & neutral
-\\\addlinespace
-& AAT\_CAC\_410.5 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_412.4 & 0.33 & 0.58 & 0.12 & 0.19 & neutral
-\\\addlinespace
-& AAT\_CAC\_458.1 & 0.15 & 0.59 & -0.03 & 0.17 & neutral
-\\\addlinespace
-& AAT\_CAC\_488.6 & 0.04 & 0.82 & 0.03 & 0.18 & neutral
-\\\addlinespace
-& AGT\_CAC\_53.8 & 0.07 & 0.72 & 0.01 & 0.17 & neutral
-\\\addlinespace
-& AGT\_CAC\_56.3 & 0.04 & 0.80 & -0.01 & 0.17 & neutral
-\\\addlinespace
-& AGT\_CAC\_98.2 & 0.44 & 0.43 & -0.44 & 0.13 & neutral
-\\\addlinespace
-& AGT\_CAC\_104.9 & 0.41 & 0.47 & -0.16 & 0.16 & neutral
-\\\addlinespace
-& AGT\_CAC\_114.8 & 0.07 & 0.75 & -0.05 & 0.16 & neutral
-\\\addlinespace
-& AGT\_CAC\_145.8 & 0.07 & 0.75 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& AGT\_CAC\_154.2 & 0.04 & 0.80 & 0.01 & 0.17 & neutral
-\\\addlinespace
-& AGT\_CAC\_158.1 & 0.07 & 0.76 & 0.01 & 0.17 & neutral
-\\\addlinespace
-& AGT\_CAC\_169.0 & 0.04 & 0.81 & -0.02 & 0.17 & neutral
-\\\addlinespace
-& AGT\_CAC\_171.1 & 0.11 & 0.68 & 0.03 & 0.18 & neutral
-\\\addlinespace
-& AGT\_CAC\_181.1 & 0.15 & 0.62 & -0.04 & 0.17 & neutral
-\\\addlinespace
-& AGT\_CAC\_183.0 & 0.33 & 0.57 & 0.21 & 0.20 & neutral
-\\\addlinespace
-& AGT\_CAC\_184.4 & 0.11 & 0.72 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& AGT\_CAC\_191.2 & 0.04 & 0.81 & 0.03 & 0.18 & neutral
-\\\addlinespace
-& AGT\_CAC\_195.0 & 0.15 & 0.61 & -0.05 & 0.17 & neutral
-\\\addlinespace
-& AGT\_CAC\_200.9 & 0.19 & 0.58 & 0.08 & 0.19 & neutral
-\\\addlinespace
-& AGT\_CAC\_203.8 & 0.07 & 0.75 & 0.01 & 0.17 & neutral
-\\\addlinespace
-& AGT\_CAC\_205.8 & 0.11 & 0.67 & -0.02 & 0.17 & neutral
-\\\addlinespace
-& AGT\_CAC\_210.0 & 0.04 & 0.82 & 0.01 & 0.17 & neutral
-\\\addlinespace
-& AGT\_CAC\_230.8 & 0.41 & 0.44 & -0.20 & 0.15 & neutral
-\\\addlinespace
-& AGT\_CAC\_241.0 & 0.41 & 0.44 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& AGT\_CAC\_245.6 & 0.07 & 0.76 & -0.05 & 0.17 & neutral
-\\\addlinespace
-& AGT\_CAC\_264.7 & 0.04 & 0.82 & -0.04 & 0.17 & neutral
-\\\addlinespace
-& AGT\_CAC\_266.9 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& AGT\_CAC\_269.4 & 0.04 & 0.82 & 0.04 & 0.18 & neutral
-\\\addlinespace
-& AGT\_CAC\_274.0 & 0.11 & 0.68 & 0.04 & 0.18 & neutral
-\\\addlinespace
-& AGT\_CAC\_285.6 & 0.07 & 0.78 & 0.02 & 0.18 & neutral
-\\\addlinespace
-& AGT\_CAC\_291.5 & 0.07 & 0.71 & -0.01 & 0.17 & neutral
-\\\addlinespace
-& AGT\_CAC\_295.6 & 0.07 & 0.78 & 0.02 & 0.17 & neutral
-\\\addlinespace
-& AGT\_CAC\_315.2 & 0.44 & 0.38 & 0.29 & 0.23 & neutral
-\\\addlinespace
-& AGT\_CAC\_332.1 & 0.48 & 0.31 & 0.72 & 0.31 & neutral
-\\\addlinespace
-& AGT\_CAC\_347.8 & 0.11 & 0.72 & -0.07 & 0.16 & neutral
-\\\addlinespace
-& AGT\_CAC\_355.2 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& AGT\_CAC\_360.2 & 0.15 & 0.65 & 0.08 & 0.19 & neutral
-\\\addlinespace
-& AGT\_CAC\_386.5 & 0.04 & 0.82 & 0.02 & 0.17 & neutral
-\\\addlinespace
-& AGT\_CAC\_418.5 & 0.15 & 0.64 & 0.05 & 0.19 & neutral
-\\\addlinespace
-& AGT\_CAC\_420.2 & 0.37 & 0.47 & -0.08 & 0.16 & neutral
-\\\addlinespace
-& AGT\_CAC\_444.3 & 0.04 & 0.82 & -0.04 & 0.17 & neutral
-\\\addlinespace
-& AGT\_CAC\_453.4 & 0.11 & 0.68 & -0.04 & 0.17 & neutral
-\\\addlinespace
-& AGT\_CAC\_458.5 & 0.15 & 0.57 & -0.01 & 0.17 & neutral
-\\\addlinespace
-& AGT\_CAC\_489.1 & 0.07 & 0.75 & -0.06 & 0.16 & neutral
-\\\addlinespace
-& ATC\_CAC\_52.4 & 0.15 & 0.65 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_56.7 & 0.15 & 0.65 & 0.06 & 0.18 & neutral
-\\\addlinespace
-& ATC\_CAC\_61.5 & 0.15 & 0.65 & -0.06 & 0.16 & neutral
-\\\addlinespace
-& ATC\_CAC\_64.3 & 0.07 & 0.75 & 0.04 & 0.18 & neutral
-\\\addlinespace
-& ATC\_CAC\_91.9 & 0.07 & 0.76 & -0.02 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_96.2 & 0.41 & 0.42 & -0.54 & 0.13 & neutral
-\\\addlinespace
-& ATC\_CAC\_99.5 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_100.6 & 0.11 & 0.68 & -0.02 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_102.0 & 0.11 & 0.66 & 0.02 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_111.9 & 0.37 & 0.53 & -0.35 & 0.15 & neutral
-\\\addlinespace
-& ATC\_CAC\_113.5 & 0.41 & 0.48 & 0.30 & 0.23 & neutral
-\\\addlinespace
-& ATC\_CAC\_123.5 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_139.7 & 0.07 & 0.71 & -0.05 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_140.8 & 0.26 & 0.55 & -0.15 & 0.16 & neutral
-\\\addlinespace
-& ATC\_CAC\_142.9 & 0.41 & 0.47 & -0.07 & 0.16 & neutral
-\\\addlinespace
-& ATC\_CAC\_144.1 & 0.15 & 0.64 & 0.21 & 0.21 & neutral
-\\\addlinespace
-& ATC\_CAC\_148.6 & 0.07 & 0.76 & 0.02 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_149.8 & 0.07 & 0.72 & -0.06 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_151.8 & 0.07 & 0.75 & 0.03 & 0.18 & neutral
-\\\addlinespace
-& ATC\_CAC\_156.1 & 0.04 & 0.80 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_162.5 & 0.19 & 0.60 & 0.31 & 0.23 & neutral
-\\\addlinespace
-& ATC\_CAC\_181.9 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_186.4 & 0.07 & 0.76 & -0.02 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_189.9 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_194.8 & 0.07 & 0.75 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_198.3 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_199.4 & 0.22 & 0.55 & -0.15 & 0.15 & neutral
-\\\addlinespace
-& ATC\_CAC\_204.2 & 0.37 & 0.53 & 0.14 & 0.20 & neutral
-\\\addlinespace
-& ATC\_CAC\_207.3 & 0.04 & 0.81 & -0.02 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_215.8 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_220.7 & 0.41 & 0.44 & 0.18 & 0.21 & neutral
-\\\addlinespace
-& ATC\_CAC\_223.7 & 0.07 & 0.76 & 0.02 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_229.1 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_230.7 & 0.04 & 0.80 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_233.7 & 0.56 & 0.23 & -0.14 & 0.16 & neutral
-\\\addlinespace
-& ATC\_CAC\_235.8 & 0.37 & 0.54 & -0.05 & 0.16 & neutral
-\\\addlinespace
-& ATC\_CAC\_258.7 & 0.07 & 0.78 & -0.03 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_266.2 & 0.04 & 0.82 & -0.04 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_270.8 & 0.41 & 0.44 & -0.51 & 0.12 & neutral
-\\\addlinespace
-& ATC\_CAC\_273.6 & 0.07 & 0.78 & -0.01 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_274.9 & 0.07 & 0.76 & -0.05 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_276.4 & 0.33 & 0.58 & 0.44 & 0.26 & neutral
-\\\addlinespace
-& ATC\_CAC\_277.8 & 0.11 & 0.67 & -0.08 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_287.8 & 0.07 & 0.75 & 0.06 & 0.18 & neutral
-\\\addlinespace
-& ATC\_CAC\_288.7 & 0.04 & 0.82 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_332.4 & 0.52 & 0.30 & 0.26 & 0.22 & neutral
-\\\addlinespace
-& ATC\_CAC\_347.9 & 0.19 & 0.58 & NaN & NaN & neutral
-\\\addlinespace
-& ATC\_CAC\_370.4 & 0.07 & 0.75 & 0.01 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_373.3 & 0.07 & 0.78 & 0.05 & 0.18 & neutral
-\\\addlinespace
-& ATC\_CAC\_378.0 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_387.7 & 0.33 & 0.58 & 0.15 & 0.20 & neutral
-\\\addlinespace
-& ATC\_CAC\_401.5 & 0.22 & 0.53 & 0.27 & 0.22 & neutral
-\\\addlinespace
-& ATC\_CAC\_405.7 & 0.04 & 0.82 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_430.5 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_442.2 & 0.37 & 0.53 & -0.19 & 0.16 & neutral
-\\\addlinespace
-& ATC\_CAC\_445.2 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
-\\\addlinespace
-& ATC\_CAC\_456.3 & 0.00 & 0.86 & 0.00 & 0.17 & neutral
-\\\addlinespace
-\textit{Campanula barbata} & ACA\_CTA\_55.8 & 0.07 & 0.78 & -0.03 & 0.12
-& neutral
-\\\addlinespace
-& ACA\_CTA\_69.2 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
-\\\addlinespace
-& ACA\_CTA\_101.6 & 0.19 & 0.55 & 0.03 & 0.13 & neutral
-\\\addlinespace
-& ACA\_CTA\_114.7 & 0.33 & 0.54 & -0.14 & 0.11 & neutral
-\\\addlinespace
-& ACA\_CTA\_122.8 & 0.04 & 0.78 & -0.01 & 0.12 & neutral
-\\\addlinespace
-& ACA\_CTA\_132.9 & 0.07 & 0.78 & 0.01 & 0.12 & neutral
-\\\addlinespace
-& ACA\_CTA\_153.4 & 0.33 & 0.55 & -0.24 & 0.11 & neutral
-\\\addlinespace
-& ACA\_CTA\_155.4 & 0.04 & 0.83 & 0.00 & 0.12 & neutral
-\\\addlinespace
-& ACA\_CTA\_164.8 & 0.41 & 0.48 & -0.45 & 0.09 & neutral
-\\\addlinespace
-& ACA\_CTA\_169.4 & 0.07 & 0.78 & -0.09 & 0.12 & neutral
-\\\addlinespace
-& ACA\_CTA\_174.3 & 0.07 & 0.77 & -0.06 & 0.12 & neutral
-\\\addlinespace
-& ACA\_CTA\_178.0 & 0.04 & 0.78 & -0.05 & 0.12 & neutral
-\\\addlinespace
-& ACA\_CTA\_179.5 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
-\\\addlinespace
-& ACA\_CTA\_183.6 & 0.15 & 0.61 & -0.09 & 0.11 & neutral
-\\\addlinespace
-& ACA\_CTA\_186.8 & 0.19 & 0.59 & -0.15 & 0.11 & neutral
-\\\addlinespace
-& ACA\_CTA\_187.9 & 0.07 & 0.78 & -0.02 & 0.12 & neutral
-\\\addlinespace
-& ACA\_CTA\_194.5 & 0.15 & 0.60 & 0.07 & 0.13 & neutral
-\\\addlinespace
-& ACA\_CTA\_195.9 & 0.11 & 0.72 & 0.04 & 0.13 & neutral
-\\\addlinespace
-& ACA\_CTA\_197.7 & 0.04 & 0.82 & 0.01 & 0.12 & neutral
-\\\addlinespace
-& ACA\_CTA\_203.8 & 0.56 & 0.20 & -0.30 & 0.13 & adaptive
-\\\addlinespace
-& ACA\_CTA\_213.5 & 0.15 & 0.64 & 0.02 & 0.13 & neutral
-\\\addlinespace
-& ACA\_CTA\_254.3 & 0.04 & 0.83 & 0.02 & 0.12 & neutral
-\\\addlinespace
-& ACA\_CTA\_284.5 & 0.04 & 0.82 & -0.01 & 0.12 & neutral
-\\\addlinespace
-& ACA\_CTA\_289.0 & 0.07 & 0.77 & -0.08 & 0.12 & neutral
-\\\addlinespace
-& ACA\_CTA\_296.4 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
-\\\addlinespace
-& ACA\_CTA\_311.1 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
-\\\addlinespace
-& ACA\_CTA\_347.7 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
-\\\addlinespace
-& ACA\_CTA\_368.6 & 0.07 & 0.75 & -0.01 & 0.12 & neutral
-\\\addlinespace
-& ACA\_CTA\_378.8 & 0.04 & 0.78 & 0.00 & 0.12 & neutral
-\\\addlinespace
-& ACA\_CTA\_382.7 & 0.19 & 0.56 & -0.08 & 0.12 & neutral
-\\\addlinespace
-& ACA\_CTA\_393.5 & 0.07 & 0.75 & 0.03 & 0.12 & neutral
-\\\addlinespace
-& ACA\_CTA\_415.5 & 0.11 & 0.70 & -0.01 & 0.12 & neutral
-\\\addlinespace
-& ACA\_CTA\_489.5 & 0.37 & 0.55 & 0.57 & 0.22 & neutral
-\\\addlinespace
-& ACA\_CTA\_491.1 & 0.07 & 0.78 & 0.01 & 0.13 & neutral
-\\\addlinespace
-& AGA\_CAC\_84.2 & 0.11 & 0.68 & -0.10 & 0.11 & neutral
-\\\addlinespace
-& AGA\_CAC\_89.4 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_91.6 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_102.6 & 0.04 & 0.82 & -0.02 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_106.2 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_107.0 & 0.22 & 0.50 & -0.05 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_109.0 & 0.11 & 0.70 & 0.01 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_116.9 & 0.04 & 0.78 & -0.04 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_130.4 & 0.07 & 0.75 & -0.07 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_133.3 & 0.37 & 0.55 & -0.11 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_135.7 & 0.07 & 0.77 & 0.03 & 0.13 & neutral
-\\\addlinespace
-& AGA\_CAC\_141.4 & 0.04 & 0.83 & -0.09 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_168.9 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_170.2 & 0.41 & 0.48 & -0.09 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_180.6 & 0.04 & 0.82 & 0.01 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_183.6 & 0.41 & 0.48 & 0.21 & 0.15 & neutral
-\\\addlinespace
-& AGA\_CAC\_196.1 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_202.0 & 0.52 & 0.26 & 0.10 & 0.14 & adaptive
-\\\addlinespace
-& AGA\_CAC\_204.8 & 0.04 & 0.83 & -0.02 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_214.4 & 0.07 & 0.74 & -0.06 & 0.11 & neutral
-\\\addlinespace
-& AGA\_CAC\_218.3 & 0.11 & 0.68 & -0.18 & 0.11 & neutral
-\\\addlinespace
-& AGA\_CAC\_227.1 & 0.07 & 0.74 & 0.06 & 0.13 & neutral
-\\\addlinespace
-& AGA\_CAC\_231.4 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_245.4 & 0.07 & 0.78 & -0.03 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_247.3 & 0.11 & 0.68 & 0.01 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_250.2 & 0.07 & 0.75 & 0.04 & 0.13 & neutral
-\\\addlinespace
-& AGA\_CAC\_251.1 & 0.04 & 0.83 & 0.00 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_269.0 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_279.6 & 0.04 & 0.82 & 0.04 & 0.13 & neutral
-\\\addlinespace
-& AGA\_CAC\_283.4 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_285.4 & 0.15 & 0.64 & -0.09 & 0.11 & neutral
-\\\addlinespace
-& AGA\_CAC\_286.8 & 0.11 & 0.69 & 0.03 & 0.13 & neutral
-\\\addlinespace
-& AGA\_CAC\_294.8 & 0.04 & 0.83 & 0.00 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_299.4 & 0.41 & 0.46 & -0.11 & 0.11 & neutral
-\\\addlinespace
-& AGA\_CAC\_308.2 & 0.33 & 0.57 & -0.23 & 0.11 & neutral
-\\\addlinespace
-& AGA\_CAC\_314.3 & 0.19 & 0.54 & -0.13 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_316.2 & 0.04 & 0.82 & 0.04 & 0.13 & neutral
-\\\addlinespace
-& AGA\_CAC\_318.3 & 0.37 & 0.54 & 0.13 & 0.14 & neutral
-\\\addlinespace
-& AGA\_CAC\_321.0 & 0.11 & 0.65 & 0.01 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_324.2 & 0.33 & 0.58 & 0.12 & 0.14 & neutral
-\\\addlinespace
-& AGA\_CAC\_326.2 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_338.4 & 0.04 & 0.82 & -0.05 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_356.5 & 0.04 & 0.78 & 0.00 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_441.7 & 0.04 & 0.82 & -0.03 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_445.1 & 0.07 & 0.77 & -0.01 & 0.12 & neutral
-\\\addlinespace
-& AGA\_CAC\_475.7 & 0.37 & 0.54 & 0.06 & 0.13 & neutral
-\\\addlinespace
-& AGA\_CAC\_477.9 & 0.67 & 0.29 & -0.25 & 0.13 & neutral
-\\\addlinespace
-& AGA\_CAC\_487.7 & 0.04 & 0.82 & -0.01 & 0.12 & neutral
-\\\addlinespace
-& AGT\_CTG\_85.2 & 0.33 & 0.57 & -0.13 & 0.11 & neutral
-\\\addlinespace
-& AGT\_CTG\_109.9 & 0.11 & 0.66 & 0.01 & 0.13 & neutral
-\\\addlinespace
-& AGT\_CTG\_127.9 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
-\\\addlinespace
-& AGT\_CTG\_130.9 & 0.11 & 0.66 & -0.03 & 0.12 & neutral
-\\\addlinespace
-& AGT\_CTG\_135.5 & 0.07 & 0.75 & -0.06 & 0.12 & neutral
-\\\addlinespace
-& AGT\_CTG\_144.4 & 0.19 & 0.56 & 0.06 & 0.13 & neutral
-\\\addlinespace
-& AGT\_CTG\_151.4 & 0.07 & 0.70 & -0.02 & 0.12 & neutral
-\\\addlinespace
-& AGT\_CTG\_152.8 & 0.15 & 0.64 & -0.05 & 0.12 & neutral
-\\\addlinespace
-& AGT\_CTG\_181.2 & 0.07 & 0.78 & -0.01 & 0.12 & neutral
-\\\addlinespace
-& AGT\_CTG\_191.2 & 0.33 & 0.58 & -1.05 & 0.09 & neutral
-\\\addlinespace
-& AGT\_CTG\_196.4 & 0.37 & 0.55 & -0.28 & 0.10 & neutral
-\\\addlinespace
-& AGT\_CTG\_202.3 & 0.04 & 0.82 & 0.00 & 0.12 & neutral
-\\\addlinespace
-& AGT\_CTG\_218.5 & 0.07 & 0.78 & 0.00 & 0.12 & neutral
-\\\addlinespace
-& AGT\_CTG\_221.0 & 0.04 & 0.82 & 0.03 & 0.13 & neutral
-\\\addlinespace
-& AGT\_CTG\_226.4 & 0.11 & 0.69 & -0.08 & 0.12 & neutral
-\\\addlinespace
-& AGT\_CTG\_228.9 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
-\\\addlinespace
-& AGT\_CTG\_230.8 & 0.11 & 0.68 & 0.13 & 0.14 & neutral
-\\\addlinespace
-& AGT\_CTG\_234.9 & 0.04 & 0.82 & 0.02 & 0.13 & neutral
-\\\addlinespace
-& AGT\_CTG\_245.4 & 0.07 & 0.75 & -0.01 & 0.12 & neutral
-\\\addlinespace
-& AGT\_CTG\_262.1 & 0.04 & 0.82 & -0.01 & 0.12 & neutral
-\\\addlinespace
-& AGT\_CTG\_266.2 & 0.48 & 0.29 & -0.21 & 0.12 & neutral
-\\\addlinespace
-& AGT\_CTG\_297.5 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
-\\\addlinespace
-& AGT\_CTG\_336.5 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
-\\\addlinespace
-& AGT\_CTG\_344.2 & 0.33 & 0.58 & -0.23 & 0.10 & neutral
-\\\addlinespace
-& AGT\_CTG\_359.0 & 0.44 & 0.44 & -0.47 & 0.10 & neutral
-\\\addlinespace
-& AGT\_CTG\_363.4 & 0.04 & 0.82 & 0.00 & 0.12 & neutral
-\\\addlinespace
-& AGT\_CTG\_392.1 & 0.37 & 0.49 & -0.28 & 0.10 & neutral
-\\\addlinespace
-& AGT\_CTG\_415.3 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
-\\\addlinespace
-& AGT\_CTG\_443.5 & 0.04 & 0.83 & 0.02 & 0.12 & neutral
-\\\addlinespace
-& AGT\_CTG\_452.7 & 0.00 & 0.87 & 0.00 & 0.12 & neutral
-\\\addlinespace
-& AGT\_CTG\_459.7 & 0.37 & 0.55 & 0.25 & 0.16 & neutral
-\\\addlinespace
-& AGT\_CTG\_488.9 & 0.07 & 0.70 & 0.02 & 0.12 & neutral
-\\\addlinespace
-\#\#\# Table S3: Genomic MDS results & & & & & &
-\\\addlinespace
-\bottomrule
-\end{longtable}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{knitr::}\KeywordTok{kable}\NormalTok{(}
-    \KeywordTok{format.table}\NormalTok{(}
-        \KeywordTok{filter}\NormalTok{(}
-            \KeywordTok{ldply}\NormalTok{(}
-                \KeywordTok{seq_along}\NormalTok{(}\KeywordTok{unique}\NormalTok{(spp.samples.DF$species)), }
-                \NormalTok{function(i) \{}
-                    \KeywordTok{ldply}\NormalTok{(}
-                        \KeywordTok{seq_along}\NormalTok{(spp.mds.LST[[i]]),}
-                        \NormalTok{function(j) \{}
-                        \KeywordTok{data.frame}\NormalTok{(}
-                            \DataTypeTok{Species=}\KeywordTok{paste0}\NormalTok{(}\StringTok{'}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{textit\{'}\NormalTok{,}\KeywordTok{gsub}\NormalTok{(}\StringTok{'}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{_'}\NormalTok{, }\StringTok{' '}\NormalTok{, }\KeywordTok{unique}\NormalTok{(spp.samples.DF$species)[i]),}\StringTok{'\}'}\NormalTok{),}
-                            \DataTypeTok{Loci=}\KeywordTok{names}\NormalTok{(spp.mds.LST[[i]])[j],}
-                            \DataTypeTok{Stress=}\KeywordTok{ifelse}\NormalTok{(}\KeywordTok{is.null}\NormalTok{(spp.mds.LST[[i]][[j]]), }\OtherTok{NA_real_}\NormalTok{, spp.mds.LST[[i]][[j]]$stress),}
-                            \DataTypeTok{Converged=}\KeywordTok{ifelse}\NormalTok{(}\KeywordTok{is.null}\NormalTok{(spp.mds.LST[[i]][[j]]), }\OtherTok{NA_real_}\NormalTok{, spp.mds.LST[[i]][[j]]$converged)}
-                        \NormalTok{)}
-                    \NormalTok{\})}
-                \NormalTok{\}}
-            \NormalTok{),}
-            \NormalTok{!}\KeywordTok{is.na}\NormalTok{(Stress)}
-        \NormalTok{),}
-        \DataTypeTok{omit=}\KeywordTok{c}\NormalTok{(}\StringTok{'Loci'}\NormalTok{,}\StringTok{'Converged'}\NormalTok{)}
-    \NormalTok{),}
-    \DataTypeTok{digits=}\DecValTok{2}\NormalTok{,}
-    \DataTypeTok{caption=}\StringTok{'Summary of non-metric multi-dimensional scaling (MDS) analyses on genetic variation for each species.'}\NormalTok{,}
-    \DataTypeTok{col.names=}\KeywordTok{c}\NormalTok{(}\StringTok{'Species'}\NormalTok{, }\StringTok{'Loci Type'}\NormalTok{, }\StringTok{'NMDS Stress'}\NormalTok{, }\StringTok{'Converged'}\NormalTok{),}
-    \DataTypeTok{align=}\KeywordTok{c}\NormalTok{(}\StringTok{'l'}\NormalTok{, }\StringTok{'c'}\NormalTok{, }\StringTok{'c'}\NormalTok{, }\StringTok{'c'}\NormalTok{)}
-\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{longtable}[c]{@{}lccc@{}}
-\toprule\addlinespace
-Species & Loci Type & NMDS Stress & Converged
-\\\addlinespace
-\midrule\endhead
-\textit{Androsace obtusifolia} & adaptive & 0.00 & 0
-\\\addlinespace
-& neutral & 0.22 & 0
-\\\addlinespace
-\textit{Arabis alpina} & neutral & 0.24 & 0
-\\\addlinespace
-\textit{Campanula barbata} & adaptive & 0.00 & 0
-\\\addlinespace
-& neutral & 0.20 & 0
-\\\addlinespace
-\bottomrule
-\addlinespace
-\caption{Summary of non-metric multi-dimensional scaling (MDS) analyses
-on genetic variation for each species.}
-\end{longtable}
+Benjamini, Y., Hochberg, Y. (1995) Controlling the false discovery rate:
+a practical and powerful approach to multiple testing. \emph{Journal of
+the Royal Statistical Society. Series B (Methodological)}. \textbf{57},
+289--300.
 
@@
 
+Evanno, G., Regnaut, S., Goudet, J. (2005) Detecting the number of
+clusters of individuals using the software structure: a simulation
+study. \emph{Molecular Ecology}. \textbf{14}, 2611--2620.
+
+Falush, D., Stephens, M., Pritchard, J. K. (2007) Inference of
+population structure using multilocus genotype data: dominant markers
+and null alleles. \emph{Molecular Ecology Notes}. \textbf{7}, 574--578.
+
 Foll, M., Gaggiotti, O. (2008) A genome-scan method to identify selected
@@
 perspective. \emph{Genetics}. \textbf{180}, 977--993.
+
+Gower, J. C. (1971) A general coefficient of similarity and some of its
+properties. \emph{Biometrics}. \textbf{27}, 857--871.
 
@@
 
+Gurobi Optimization, I. (2015) Gurobi optimizer reference manual.
+
+Hanson, J. O., Rhodes, J. R., Possingham, H. P., Fuller, R. A. (2016)
+rapr: representative and adequate prioritisations in r.
+\emph{Unpublished}. \textbf{X}, XX--XX.
+
 Hijmans, R. J., Cameron, S. E., Parra, J. L., Jones, P. G., Jarvis, A.
@@
 1965--1978.
+
+Jakobsson, M., Rosenberg, N. A. (2007) CLUMPP: a cluster matching and
+permutation program for dealing with label switching and multimodality
+in analysis of population structure. \emph{Bioinformatics}. \textbf{23},
+1801--1806.
 
@@
 
+Pritchard, J. K., Stephens, M., Donnelly, P. (2000) Inference of
+population structure using multilocus genotype data. \emph{Genetics}.
+\textbf{155}, 945--959.
+
 Vos, P., Hogers, R., Bleeker, M., Reijans, M., Lee, T. van de, Hornes,
@@
 
+\documentclass[11pt,]{article}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+    \usepackage{xltxtra,xunicode}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
+  \newcommand{\euro}{€}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\usepackage[margin=1in]{geometry}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\ifxetex
+  \usepackage[setpagesize=false, % page size defined by xetex
+              unicode=false, % unicode breaks when used with xetex
+              xetex]{hyperref}
+\else
+  \usepackage[unicode=true]{hyperref}
+\fi
+\hypersetup{breaklinks=true,
+            bookmarks=true,
+            pdfauthor={},
+            pdftitle={},
+            colorlinks=true,
+            citecolor=blue,
+            urlcolor=blue,
+            linkcolor=magenta,
+            pdfborder={0 0 0}}
+\urlstyle{same}  % don't use monospace font for urls
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\setcounter{secnumdepth}{0}
+
+%%% Use protect on footnotes to avoid problems with footnotes in titles
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
+
+%%% Change title format to be more compact
+\usepackage{titling}
+
+% Create subtitle command for use in maketitle
+\newcommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
+
+\setlength{\droptitle}{-2em}
+  \title{}
+  \pretitle{\vspace{\droptitle}}
+  \posttitle{}
+  \author{}
+  \preauthor{}\postauthor{}
+  \date{}
+  \predate{}\postdate{}
+
+\begin{document}
+
+\maketitle
+
+\subsection{Figures}\label{figures}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures_files/figure-latex/unnamed-chunk-2-1.pdf}
+\caption{Summary of single species prioritisations. Single species
+prioritisations were generated using amount-based targets, amount-based
+and surrogate-based targets, and amount-based and genetic-based targets
+for each species. Data shows the performance of prioritisations
+generated using these three sets of targets. Bars denote means and
+standard errors.}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures_files/figure-latex/unnamed-chunk-3-1.pdf}
+\caption{Multi-species prioritisations. Panel (a) shows the
+prioritisation generated for using just amount-based targets. Panel (b)
+shows the prioritisation generated using amount-based and surrogate
+based targets. Panel (c) shows the prioritisation generated using
+amount-based and genetic-based targets}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures_files/figure-latex/unnamed-chunk-4-1.pdf}
+\caption{Summary of multi-species prioritisations. Three prioritisations
+were generated using amount-based targets, amount-based and
+surrogate-based targets, and amount-based and genetic-based targets for
+each species. Data shows the performance of these prioritisations based
+on how much genetic variation they explain. Bars denote means and
+standard errors.}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures_files/figure-latex/unnamed-chunk-5-1.pdf}
+\caption{The relationship between surrogates and genetic variation
+secured in prioritizations.}
+\end{figure}
+
 \end{document}
+
+\end{document}

### 2015-12-24 article/figures.Rmd
+ updated analysis to use equi-distant crs for attribute space + updated methods section in ms + attempting to set the supp info to be correct

--- 2015-12-04 inst/doc/genetic.surrogates.Rmd
@@
 ---
-title: "Are environmental and geographic effective surrogates for genetic variation in conservation planning?"
-author: |
-  | Jeffrey O. Hanson$^1$, Jonathan R. Rhodes$^2$, Cynthia Riginos$^2$, Hugh P. Possingham$^1$, Richard A. Fuller$^1$
-  | $^1$School of Biological Sciences, The University of Queensland, Brisbane, QLD, Australia
-  | $^2$School of Geography, Planning and Environmental Management, The University of Queensland, Brisbane, QLD, Australia
-  | Correspondance should be addressed to jeffrey.hanson@uqconnect.edu.au
-date: "`r format(Sys.time(), '%d %B %Y')`"
-tags: ["conservation", "biodiversity", "AFLP", "reserve-selection"]
-abstract: "Insert abstract here."
-output:
-  rmarkdown::pdf_document:
-    toc: true
-    toc_depth: 5
-    keep_tex: yes
-    fig_caption: yes
-    includes:
-      in_header: preamble-latex.tex
 fontsize: 11pt
@@
 csl: reference-style.csl
-vignette: >
-  %\VignetteIndexEntry{rapr}
-  %\VignetteEngine{knitr::rmarkdown}
-  \usepackage[utf8]{inputenc}
+output:
+  rmarkdown::pdf_document:
+    toc: false
+    keep_tex: no
+    fig_caption: yes
+    includes:
+      in_header: preamble-latex.tex
 ---
@@
 ```{r, include=FALSE}
-# set global cache
-knitr::opts_chunk$set(cache=FALSE)
+# set wd
+setwd('..')
+
+# load .rda files
+load('results/results.rda')
 
 # load packages
-library(genetic.surrogates)
-library(data.table)
-library(rgeos)
-library(ggplot2)
-library(grid)
-library(gridExtra)
-library(plyr)
-library(dplyr)
-library(pander)
+source('R/analysis/00-initialization.R')
 
-# set pander options
-panderOptions('knitr.auto.asis', FALSE)
-
-# set default select method
-select <- dplyr::select
-
-### set parameters
-## debugging parameters
-bs.threshold <- 0.5
-bs.threads <- 1
-bs.n <- 50
-bs.thin <- 1
-bs.nbp <- 10
-bs.pilot <- 50
-bs.burn <- 10
-mds.k <- 2
-mds.trymax <- 2
-n.spp <- 3
-
-## analysis parameters
-# bs.threshold <- 0.95
-# bs.threads <- 10
-# bs.n <- 5000
-# bs.thin <- 10
-# bs.nbp <- 20
-# bs.pilot <- 5000
-# bs.burn <- 50000
-# mds.k <- 2
-# mds.trymax <- 100
-
-# set seed for reproducibility
-set.seed(500)
+# set options
+options(error=NULL)
 ```
 
-## Introduction
+## Figures
 
-## Methods
-
-### Study area
-To address the aims of this study, we obtained species distribution and genomic (AFLP) from [@r451].
-This dataset was chosen because it provides genomic data for a multitude of species at a high spatial resolution (approx.
-$20$km$^{2} \times 22$km$^{2}$).
-
-```{r}
-## compile spatial grid data
-# load grid cell centroids
-grid.DF <- fread(
-	'extdata/Data_Meirmans_et_al_IntrabioDiv/ReadMe.txt',
-	data.table=FALSE,
-	skip='cell\tLong\tLat'
-) %>% rename(
-		grid.longitude=Long,
-		grid.latitude=Lat
-) %>% mutate(
-	id=seq_along(grid.latitude)
-)
+```{r, echo=FALSE, fig.width=6, fig.height=3, fig.cap='Summary of single species prioritisations.
+Single species prioritisations were generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species.
+Data shows the performance of prioritisations generated using these three sets of targets.
+Bars denote means and standard errors.'}
+# prepare data for plotting
+single.spp.PDF <- expand.grid(
+	Prioritisation=unique(single.spp.SDF$Prioritisation),
+	Metric=unique(single.spp.SDF$Metric))
+single.spp.PDF <- cbind(single.spp.PDF,
+	as.data.frame(predict(single.spp.GLM, single.spp.PDF,
+		type='response', se.fit=TRUE))) %>%
+	mutate(lower=fit-se.fit, upper=fit+se.fit,
+		letters=toupper(cld(single.spp.MCP)$mcletters$Letters),
+		letter_pos=upper+0.05)
+# make plot
+ggplot(aes(x=Metric,y=fit,fill=Prioritisation),
+	data=single.spp.PDF) +
+	geom_bar(position=position_dodge(0.9),
+		stat='identity') +
+	geom_errorbar(
+		aes(ymin=lower,ymax=upper),
+		position=position_dodge(0.9), width=0.6) +
+	geom_text(aes(x=Metric, y=letter_pos,
+		label=letters), position=position_dodge(0.9)) +
+	scale_fill_manual(name='Prioritisation',
+		values=c('grey80','grey50','grey20')) +
+	ylab('Proportion genetic\nvariation secured (%)') +
+	xlab('') +
+	theme_classic()
 ```
 
-```{r}
-# load in aflp data
-spp.aflp.paths <- dir(
-	'extdata/Data_Meirmans_et_al_IntrabioDiv',
-	'^.*AFLP\\.dat$',
-	full.names=TRUE
-)[seq_len(n.spp)]
-spp.BayeScanData.LST <- llply(
-	spp.aflp.paths,
-	read.BayeScanData
-)
-```
-
-```{r}
-## compile species occurence data
-# load in data
-spp.loc.paths <- dir(
-	'extdata/Data_Meirmans_et_al_IntrabioDiv',
-	'^.*locations\\.txt$',
-	full.names=TRUE
-)[seq_len(n.spp)]
-spp.samples.DF <- ldply(
-	seq_along(spp.loc.paths),
-	.fun=function(i) {
-		x <- mutate(
-			fread(spp.loc.paths[i], data.table=FALSE),
-			species=gsub('_locations.txt', '', basename(spp.loc.paths[i]), fixed=TRUE)
-		) %>% rename(
-			cell=population,
-			sample.longitude=longitude,
-			sample.latitude=latitude
-		)
-		return(x[as.numeric(spp.BayeScanData.LST[[i]]@populations),])
-	}
-) %>% left_join(
-		grid.DF,
-		by='cell'
-)
-# append species data to grid data.frame (wide-format)
-for (i in unique(spp.samples.DF$species))
-	grid.DF[[i]] <- replace(
-		rep(0, nrow(grid.DF)),
-		which(grid.DF$cell %in% filter(spp.samples.DF, species==i)$cell),
-		1
-	)
-```
-
-### Genomic data
-Loci in the AFLP were classified as adaptive or neutral using BayeScan (version 2.1, using a proability threshold of `r `).
-
-```{r}
-# assign cells as populations
-spp.BayeScanData.LST <- llply(
-	seq_along(unique(spp.samples.DF$species)),
-	function(i) {
-		bd <- spp.BayeScanData.LST[[i]]
-		bd@populations <- filter(spp.samples.DF, species==unique(spp.samples.DF$species)[i])$cell
-		return(bd)
-	}
-)
-# run BayeScan
-spp.BayeScan.LST <- llply(
-	spp.BayeScanData.LST,
-	run.BayeScan,
-	threshold=bs.threshold,
-	threads=bs.threads,
-	n=bs.n,
-	thin=bs.thin,
-	nbp=bs.nbp,
-	pilot=bs.pilot,
-	burn=bs.burn
-)
-# run MDS
-spp.mds.LST <- llply(
-	spp.BayeScan.LST,
-	function(i) {
-		`names<-`(llply(c('adaptive', 'neutral'), function(j) {
-			if (sum(i@results@fst==j)==0)
-				return(NULL)
-			return(
-				mds(
-					i,
-					metric='gower',
-					type=j,
-					k=mds.k,
-					trymax=mds.trymax
-				)
-			)
-		}), c('adaptive','neutral'))
-	}
-)
-# store mds rotations for each sample
-spp.samples.DF <- ldply(seq_along(unique(spp.samples.DF$species)), .fun=function(i) {
-	x <- filter(spp.samples.DF, species==unique(spp.samples.DF$species)[i])
-	for (j in c('adaptive', 'neutral')) {
-		if (!is.null(spp.mds.LST[[i]][[j]])) {
-			x <- cbind(
-				x,
-				`names<-`(
-					as.data.frame(spp.mds.LST[[i]][[j]]$points),
-					paste0(j,'_d',seq_len(mds.k))
-				)
-			)
-		}
-	}
-	return(x)
-})
-# store mds average rotation for each grid
-for (i in seq_along(unique(spp.samples.DF$species))) {
-	for (j in c('adaptive', 'neutral')) {
-		if(!is.null(spp.mds.LST[[i]][[j]])) {
-			curr.sub <- filter(spp.samples.DF, species==unique(spp.samples.DF$species)[i])
-			for (k in seq_len(mds.k)) {
-				curr.vals <- tapply(
-					curr.sub[[paste0(j,'_d',k)]],
-					curr.sub$cell,
-					FUN=mean
-				)
-				curr.pos <- match(names(curr.vals), grid.DF$cell)
-				grid.DF[curr.pos,paste0(unique(spp.samples.DF$species)[i],'_',j,'_d',k)] <- curr.vals
-			}
-		}
-	}
-}
-```
-### Surrogate data
-
-```{r}
-## create spatial data
-# grid data as SpatialPolygonsDataFrame
-grid.PTS <- SpatialPoints(as.matrix(grid.DF[,2:3]))
-grid.PLY <- grid.PTS %>%
-	points2grid(tolerance=0.05) %>%
-	as('SpatialPolygons')
-grid.PLY <- grid.PLY[sapply(gIntersects(grid.PTS, grid.PLY, byid=TRUE, returnDense=FALSE), `[[`, 1),] %>%
-	spChFIDs(
-		as.character(seq_len(nrow(grid.DF)))
-	) %>% 
-	SpatialPolygonsDataFrame(
-		data=grid.DF
-	)
-grid.PLY@proj4string <- wgs1984
-grid.PPLY <- spTransform(grid.PLY, europeEA)
-# sample data as SpatialPoints
-spp.sample.PTS <- SpatialPointsDataFrame(
-	coords=as.matrix(spp.samples.DF[,5:6]),
-	data=spp.samples.DF,
-	proj4string=wgs1984
-)
-spp.sample.PPTS <- spTransform(spp.sample.PTS, europeEA)
-## extract geographic data
-centroids.DF <- gCentroid(grid.PPLY, byid=TRUE) %>% slot('coords') %>%
-	as.data.frame() %>% `names<-`(paste0('geo_d',1:2))
-grid.DF <- cbind(grid.DF, centroids.DF)
-## extract climatic data
-# load climatic data
-bioclim.STK <- stack('extdata/BioClim_variables/bioclim_pca.tif')
-# extract mean for each cell for each principle component
-extract.DF <- grid.PPLY %>% rasterize(bioclim.STK, field='id') %>% 
-	zonal(x=bioclim.STK) %>% as.data.frame() %>% select(-1) %>%
-	`names<-`(paste0('env_d',seq_len(nlayers(bioclim.STK))))
-# merge with grid.DF
-grid.DF <- cbind(grid.DF, extract.DF)
-## update spatial objects
-grid.PLY@data <- grid.DF
-grid.PPLY@data <- grid.DF
-```
-
-### Prioritisations
-
-## Results
-
-## Discussion
-
-## Acknowledgements
-JOH is funded by an Australian Postgraduate Award (APA) scholarship.
-RAF has an Australian Research Council Future Fellowship.
-This work was supported by the Centre of Excellence for Environmental Decisions (CEED) and the Landscape Ecology and Conservation Group (LEC) at The University of Queensland.
-
-## References
-
-## Supporting Information
-### Appendix S1: Species distributions
-
-```{r, fig.width=9.5, fig.height=10, out.width='7.5in', out.height='8.5in', fig.cap='Species distributions.
-Squares represent planning units.
-For a given species, planning units that were found to be inhabited are denoted with bright blue.'}
-## plot map of species distributions
+```{r, echo=FALSE, fig.width=8, fig.height=3.5, fig.cap='Multi-species prioritisations.
+Panel (a) shows the prioritisation generated for using just amount-based targets.
+Panel (b) shows the prioritisation generated using amount-based and surrogate based targets.
+Panel (c) shows the prioritisation generated using amount-based and genetic-based targets'}
 # download basemap
@@
 ,] %>% spFortify
-# fortify data
-grid.FPLY <- spFortify(grid.PLY)
-spp.grid.FPLY <- ldply(unique(spp.samples.DF$species), function(x) {
-		z <- grid.FPLY[,c('long', 'lat', 'group', x),drop=FALSE]
-		names(z)[4] <- 'presence'
-		z$species <- gsub('\\_', ' ', x)
-		return(z)
-	}
-)
-# plot species data
+# prepare data for plotting
+multi.spp.grid.FPLY <- grid.PLY
+for (i in seq_along(multi.spp.prioritisations))
+	multi.spp.grid.FPLY@data[[paste0('v',i)]] <- selections(multi.spp.prioritisations[[i]])
+multi.spp.grid.FPLY <- spFortify(multi.spp.grid.FPLY)
+# make maps
+do.call(
+	grid.arrange,
+	append(
+		llply(
+			seq_along(multi.spp.prioritisations),
+			function(i) {
 				ggplot() +
@@
 						fill='grey20', color='grey80') +
-	geom_polygon(data=spp.grid.FPLY, aes(x=long, y=lat, 
-		group=group, fill=presence), alpha=0.8, color='grey10') +
-	theme_classic() +
-	guides(fill=guide_legend(title='Presence')) +
-	theme(axis.ticks=element_blank(), axis.text=element_blank()) +
-	coord_cartesian(
-		xlim=buffered.range(grid.FPLY$long, 0.05),
-		ylim=buffered.range(grid.FPLY$lat, 0.05)
-	) +
-	xlab('') +
-	ylab('') +
-	facet_wrap(~ species, ncol=4)
-```
-
-```{r, fig.width=4, fig.height=3.5, fig.cap="Species richness. Squares denote planning units. Planning units with a brighter color are inhabited by more species."}
-# calculate species richness
-grid.PLY$Species_richness <- grid.PLY@data %>%
-	select(5:(4+n.spp)) %>% as.matrix() %>% rowSums()
-
-# plot species richness
-ggplot() +
-	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
-		fill='grey20', color='grey80') +
-	geom_polygon(data=spFortify(grid.PLY), aes(x=long, y=lat, 
-		group=group, fill=Species_richness), alpha=0.8, color='grey10') +
-	guides(fill=guide_legend(title='Count (#)')) +
-	theme_classic() +
-	theme(axis.ticks=element_blank(), axis.text=element_blank()) +
-	coord_cartesian(
-		xlim=buffered.range(grid.FPLY$long, 0.05),
-		ylim=buffered.range(grid.FPLY$lat, 0.05)
-	) +
-	xlab('') +
-	ylab('') +
-	ggtitle('Species richness')
-```
-
-### Appendix S2: Genomic MDS 
-```{r}
-knitr::kable(
-	ldply(
-		seq_along(unique(spp.samples.DF$species)), 
-		function(i) {
-			ldply(
-				seq_along(spp.mds.LST[[i]]),
-				function(j) {
-				data.frame(
-					Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
-					Loci=names(spp.mds.LST[[i]])[j],
-					Stress=spp.mds.LST[[i]][[j]]$stress,
-					Converged=spp.mds.LST[[i]][[j]]$converged
-				)
-			})
-		}
-	),
-	digits=2,
-	caption='Summary of nonmetric-dimensional scaling (MDS) analyses on genetic variation for each species.'
-)
-```
-
-### Appendix S3: Distribution maps of intra-specific variation
-```{r}
-plot.spp.mds <- function(i) {
-	do.call(
-		grid.arrange,
-		append(
-			unlist(llply(c('adaptive','neutral'), function(g) {
-				llply(seq_len(mds.k), function(k) {
-					ggplot() +
-						geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
-							fill='grey20', color='grey80') +
-						geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
-							group='group', fill=paste0(unique(spp.samples.DF$species)[i], '_', j, '_d',k)),
+					geom_polygon(data=multi.spp.grid.FPLY, aes_string(x='long', y='lat', 
+						group='group', fill=paste0('v',i)),
 						alpha=0.8, color='grey10') +
@@
 					theme(axis.ticks=element_blank(), axis.text=element_blank(),
-							plot.margin=unit(c(0,0,0,0),'cm')) +
+						plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(),
+						legend.position='none') +
 					coord_cartesian(
-							xlim=buffered.range(grid.FPLY$long, 0.05),
-							ylim=buffered.range(grid.FPLY$lat, 0.05)
+						xlim=buffered.range(multi.spp.grid.FPLY$long, 0.05),
+						ylim=buffered.range(multi.spp.grid.FPLY$lat, 0.05)
 					) +
@@
 					ylab('') +
-						ggtitle(paste0(g,' (',k,')'))
-				})
-			}),recursive=FALSE),
-			list(ncol=2)
+					annotate('text',
+						x=min(multi.spp.grid.FPLY$long)+diff(range(multi.spp.grid.FPLY$long)*0.1),
+						y=min(multi.spp.grid.FPLY$lat)+diff(range(multi.spp.grid.FPLY$lat)*1.01),
+						label=letters[i], hjust=1, vjust=1, color='white', size=8)
+			}
+		),
+		list(nrow=1)
 	)
-	)
-}
-```
-
-```{r, fig.width=6.5, fig.height=4.5, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[1]),'}.
-Each square represents a planning unit.
-The color of each planning unit panel corresponds to ordination values.
-Planning units with similar colors contain individiduals with similar genetic variation.')}
-plot.spp.mds(1)
-```
-
-```{r, fig.width=6.5, fig.height=4.5, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[2]),'}.
-See Figure XX caption for conventions.')}
-plot.spp.mds(2)
-```
-
-```{r, fig.width=6.5, fig.height=4.5, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[3]),'}.
-See Figure XX caption for conventions.')}
-plot.spp.mds(3)
-```
-
-### Appendix S4: Principle components analysis on climatic variation
-```{r}
-## load pca summary
-pca.DF <- read.table('extdata/BioClim_variables/pca.TXT', skip=80) %>% `names<-`(
-	c('Principle Component', 'Eigen Value', 'Variation explained (%)',
-	'Accumulative variation explained (%)')
-)
-## make results table showing Eigen values
-knitr::kable(
-	pca.DF,
-	digits=2,
-	caption='Summary of priciniple components analysis (PCA) on bioclimatic variation across the study area. The first two principle components (PCs) were used for subsequent analysis.'
 )
@@
 
-### Appendix S5: Maps of climatic variation
-```{r, width=6.5, height=4.5, fig.cap='Climatic variation.
-Each panel depicts variation based on a different principle component (PC).
-Sqaures represent planning units.
-The color of each planning unit denotes the average priciniple component value of pixels inside it.
-Planning units with more similar colors have more similar climates regimes.'}
-do.call(
-	grid.arrange,
-		append(
-		llply(grep('^env\\_.*$', names(grid.DF), value=TRUE), function(x) {
-			ggplot() +
-				geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
-					fill='grey20', color='grey80') +
-				geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
-					group='group', fill=x),
-					alpha=0.8, color='grey10') +
-				guides(fill=guide_legend(title=' ')) +
-				theme_classic() +
-				theme(axis.ticks=element_blank(), axis.text=element_blank(),
-					plot.margin=unit(c(0,0,0,0),'cm')) +
-				coord_cartesian(
-					xlim=buffered.range(grid.FPLY$long, 0.05),
-					ylim=buffered.range(grid.FPLY$lat, 0.05)
-				) +
+```{r, echo=FALSE, fig.width=6, fig.height=3, fig.cap='Summary of multi-species prioritisations.
+Three prioritisations were generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species.
+Data shows the performance of these prioritisations based on how much genetic variation they explain.
+Bars denote means and standard errors.'}
+# prepare data for plotting
+multi.spp.PDF <- expand.grid(
+	Prioritisation=unique(multi.spp.SDF$Prioritisation),
+	Metric=unique(multi.spp.SDF$Metric))
+multi.spp.PDF <- cbind(multi.spp.PDF,
+	as.data.frame(predict(multi.spp.GLM, multi.spp.PDF,
+		type='response', se.fit=TRUE))) %>%
+	mutate(lower=fit-se.fit, upper=fit+se.fit,
+		letters=toupper(cld(multi.spp.MCP)$mcletters$Letters),
+		letter_pos=upper+0.05)
+# make plot
+ggplot(aes(x=Metric,y=fit,fill=Prioritisation),
+	data=multi.spp.PDF) +
+	geom_bar(position=position_dodge(0.9),
+		stat='identity') +
+	geom_errorbar(
+		aes(ymin=lower,ymax=upper),
+		position=position_dodge(0.9), width=0.6) +
+	geom_text(aes(x=Metric, y=letter_pos,
+		label=letters), position=position_dodge(0.9)) +
+	scale_fill_manual(name='Prioritisation',
+		values=c('grey80','grey50','grey20')) +
+	ylab('Proportion genetic\nvariation secured (%)') +
 	xlab('') +
-				ylab('') +
-				ggtitle(paste0('PC ', substr(x, nchar(x), nchar(x))))
-		}),
-		list(ncol=2)
-	)
-)
+	theme_classic()
 ```
+
+```{r, echo=FALSE, fig.width=8, fig.height=3.5, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations.'}
+# make plots
+p1 <- ggplot(data=env.pareto.DF) +
+	geom_line(aes(x=Surrogate.target,y=adaptive.held,group=Species),
+		alpha=0.5) +
+	xlab('Environmental variation secured (%)') +
+	ylab('Adaptive genetic\nvariation secured (%)') +
+	theme_classic()
+p2 <- ggplot(data=geo.pareto.DF) +
+	geom_line(aes(x=Surrogate.target,y=neutral.held,group=Species),
+		alpha=0.5) +
+	xlab('Geographic variation secured (%)') +
+	ylab('Neutral genetic\nvariation secured (%)') +
+	theme_classic()
+grid.arrange(p1, p2, nrow=1)
+```

### 2015-12-24 article/figures.tex
+ updated analysis to use equi-distant crs for attribute space + updated methods section in ms + attempting to set the supp info to be correct

--- 2015-12-24 article/article.tex
@@
-\documentclass[11pt,]{article}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-    \usepackage{xltxtra,xunicode}
-  \else
-    \usepackage{fontspec}
-  \fi
-  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
-  \newcommand{\euro}{€}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
-\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\usepackage[margin=1in]{geometry}
-\ifxetex
-  \usepackage[setpagesize=false, % page size defined by xetex
-              unicode=false, % unicode breaks when used with xetex
-              xetex]{hyperref}
-\else
-  \usepackage[unicode=true]{hyperref}
-\fi
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={},
-            pdftitle={Are environmental and geographic effective surrogates for genetic variation in conservation planning?},
-            colorlinks=true,
-            citecolor=blue,
-            urlcolor=blue,
-            linkcolor=magenta,
-            pdfborder={0 0 0}}
-\urlstyle{same}  % don't use monospace font for urls
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\setcounter{secnumdepth}{0}
-
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-%%% Change title format to be more compact
-\usepackage{titling}
-
-% Create subtitle command for use in maketitle
-\newcommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
-
-\setlength{\droptitle}{-2em}
-  \title{Are environmental and geographic effective surrogates for genetic
-variation in conservation planning?}
-  \pretitle{\vspace{\droptitle}\centering\huge}
-  \posttitle{\par}
-  \author{Jeffrey O. Hanson$^1$, Jonathan R. Rhodes$^2$, Cynthia Riginos$^1$, Hugh
-P. Possingham$^1$, Richard A. Fuller$^1$\\$^1$School of Biological
-Sciences, The University of Queensland, Brisbane, QLD,
-Australia\\$^2$School of Geography, Planning and Environmental
-Management, The University of Queensland, Brisbane, QLD,
-Australia\\Correspondance should be addressed to
-\href{mailto:jeffrey.hanson@uqconnect.edu.au}{jeffrey.hanson@uqconnect.edu.au}}
-  \preauthor{\centering\large\emph}
-  \postauthor{\par}
-  \predate{\centering\large\emph}
-  \postdate{\par}
-  \date{24 December 2015}
-
-% load packages
-\usepackage{amsmath,amsfonts,float,makecell,titletoc,tocloft,titlesec,natbib}
-\usepackage[T1]{fontenc}
-\usepackage{lmodern}
-\usepackage[utf8]{inputenc}
-
-% format captions
-\usepackage[labelfont={small,bf}, labelsep=space, font={small}]{caption}
-
-% format headings
-\titleformat{\subsection}
-{\normalfont\Large\bfseries}{\thesubsection}{1em}{}
-\titleformat{\subsubsection}
-{\normalfont\large\bfseries}{\thesubsubsection}{1em}{}
-\titleformat{\paragraph}
-{\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
-\titleformat{\subparagraph}
-{\normalfont\normalsize\itshape}{\thesubparagraph}{1em}{}
-
-\titlespacing\paragraph{0pt}{12pt plus 4pt minus 2pt}{+0pt plus 2pt minus 2pt}
-\titlespacing{\subparagraph}{0pt}{6pt plus 4pt minus 2pt}{+0pt plus 2pt minus 2pt}
-
-% format toc
-\renewcommand\cftsubsecfont{\normalfont\normalsize\bfseries}
-% \renewcommand\cftsubsubsecfont{\normalfont\normalsize\bfseries}
-% \renewcommand\cftparafont{\normalfont\normalsize\bfseries}
-% \renewcommand\cftsubparafont{\normalfont\normalsize\itshape}
-
-% make figures static
-\let\origfigure\figure
-\let\endorigfigure\endfigure
-\renewenvironment{figure}[1][2] {
-	\expandafter\origfigure\expandafter[H]
-} {
-	\endorigfigure
-}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-Insert abstract here.
-\end{abstract}
-
-\subsection{Introduction}\label{introduction}
-
-\subsection{Methods}\label{methods}
-
-\subsubsection{Study area}\label{study-area}
-
-To address the aims of the study, we utilized species distribution and
-genomic data collected by the IntraBioDiv project (Figure S1; Meirmans
-\emph{et al.} 2011; see Gugerli \emph{et al.} 2008 for methodological
-details). This dataset was well suited for this study, because it
-provides spatially explicit genomic data for a multitude of species
-across a broad geographic area and over large environmental gradients.
-As part of this project, data for 27 alpine plant species were collected
-across the European Alps according to a grid (22.3 km $\times$ 25 km).
-Briefly, in every second cell, plant samples were collected from three
-individuals for each species present inside the cell. Samples were
-genotyped using amplified fragment length polymorphisms (AFLP; Vos
-\emph{et al.} 1995). Matrices denoting the presence/absence of
-polymorphisms at loci were constructed independently for each species.
-Thus for each second grid cell, for each of the species in the grid
-cell, the dataset contains information describing the genomic variation
-of individuals inside the grid cell. We used these grid cells as
-planning units to develop prioritizations to investigate the
-effectiveness of surrogates for the (AFLP) genetic variation.
-
-\subsubsection{Surrogate data}\label{surrogate-data}
-
-We explored the effectiveness of climatic and geographic surrogates. To
-describe variation in the location of each grid cell, we projected the
-grid into an equi-distant coordinate system (ESRI:102031) and used the
-centroid of each cell. To describe the climatic variation among each
-cell, we obtained twenty-one bioclimatic variables (approx. 1km
-resolution; Hijmans \emph{et al.} 2005). These layers and the grid cells
-were projected into an equal-area coordinate system (ESRI:102014). The
-bioclimatic variables were clipped to the extent of the grid and subject
-to a principle components analysis (PCA; Table S1). The first two
-principle components (PCs) cumulatively explained 97.2\% of the
-variation and were used to generate two new layers. We calculated the
-average value for each PC layer in each grid cell, and used these values
-to characterize the main themes of climatic variation among the grid
-cells (Figure S2).
-
-\subsubsection{Outlier locus detection}\label{outlier-locus-detection}
-
-To investigate the effectiveness of the surrogates for adaptive and
-neutral genetic variation, we first identified which of the sampled loci
-for each species were under selection. We used outlier locus detection
-methods (evaluated in P{é}rez-Figueroa \emph{et al.} 2010) to avoid
-circularity issues. The basic premise underpinning these analyses is
-that neutral loci are expected to have a specific level of variation
-among populations ($F_{st}$), and loci that deviate from this
-expectation should be under selection. Thus loci can be identified as
-being under selection without using environmental data. In an earlier
-study, Bothwell \emph{et al.} (2013) applied such outlier detection
-methods to one of the species in this dataset (\emph{Getiana nivalis}).
-Here, we adopted a similar methodology and applied it to each of the 27
-species in the dataset.
-
-First, we grouped individuals into genetic lineages for each species
-(Figure S2). The purpose of this step was to avoid confounding variation
-between populations with due to genetic history with variation
-attributed to selection pressures. We used the Bayesian clustering
-method implemented in $\texttt{Structure}$ (version 2.3.4; Pritchard
-\emph{et al.} 2000; Falush \emph{et al.} 2007) assuming admixture and an
-independent alleles model. For each species, we identified the most
-plausible number of populations (K = 1--2; 4 runs per K; 10 burn-in
-iterations; 10 post-burn-in iterations) using the $\Delta$K method
-(Figure S3; Evanno \emph{et al.} 2005). We aggregated the population
-membership probabilities from replicate runs using $\texttt{CLUMPP}$
-(version 1.1.2; large K greedy search method with 1000 random starts per
-K per species; Figure S4; Jakobsson and Rosenberg 2007). Individuals
-with a moderate probability of belonging to a single population
-($\geq 0.75$) were assigned accordingly (Figure S5), and subject to
-subsequent outlier locus detection methods.
-
-We used $\texttt{Bayescan}$ (version 2.1; Foll and Gaggiotti 2008) to
-identify loci under selection for each species (Table S2). For each
-species, $\texttt{Bayescan}$ was run using the population memberships
-identified in the previous analysis (1:1 prior odds; 3 pilot runs; 40
-burn-in iterations; 10 post-burn-in iterations thinned by 1 iterations).
-Following guidelines in the $\texttt{Bayescan}$ manual, we omitted loci
-where the global frequency of the minor allele was $\geq 0.05$ to avoid
-false-positives. To ensure convergence, we ran 2 replicates per species
-using a suitable false discovery rate (FDR; $\leq 0.8$; Benjamini and
-Hochberg 1995).
-
-After classifying loci as adaptive or neutral, we mapped the main
-gradients of the adaptive (if present) and neutral genetic variation for
-each species (Figures S4--S29). Note that while individuals not assigned
-to populations were omitted from the $\texttt{Bayescan}$ analysis, here
-we used all the individuals. For each species, for each type of genetic
-variation present, we used Gower distances (Gower 1971) to express the
-genomic differences between individuals (using the cluster R package to
-accommodate sparsity; Maechler \emph{et al.} 2015). These distances were
-subject to a non metric multi-dimensional scaling analysis (implemented
-in the vegan R package using K=2 and 2 random starts; Table S3; Oksanen
-\emph{et al.} 2015). The average of the ordinated values associated with
-individuals in each grid cell we used to characterize the genomic
-variation in the grid cell.
-
-\subsubsection{Prioritizations}\label{prioritizations}
-
-We used the unreliable representative and adequate prioritization (URAP)
-formulation of the reserve selection problem (Hanson \emph{et al.}
-2016). We treated the genomic and surrogate variables as distinct
-attribute spaces. We set the demand points for each species and its
-attribute space using the values associated with the planning units.
-First, we generated a set of single species prioritizations. For each
-species, we generated a prioritization using a 20\% amount-based target,
-a prioritization using a 20\% amount-based target and 90\% space-based
-targets for the surrogate variation, and a prioritization using a 20\%
-amount-based target and 90\% space-based targets for the genomic
-variation. Next, we generated three multi-species prioritizations using
-the same sets of targets used in the single species prioritizations.
-Finally, to understand the relationship between the surrogate-based and
-genomic-based targets, we generated a suite of prioritizations using
-increasing space-based targets between 1\% and 100\% in increments of
-49.5\%. For each prioritization we calculated the proportion of the
-adaptive (if present) and neutral genetic variation that it secured. We
-solved the reserve selection problems to within 90\% of optimality using
-the $\texttt{rapr}$ R package (Hanson \emph{et al.} 2016) and
-$\texttt{Gurobi}$ (Gurobi Optimization 2015).
-
-\subsection{Results}\label{results}
-
-\subsubsection{Single species
-prioritizations}\label{single-species-prioritizations}
-
-\subsubsection{Multi-species
-prioritisations}\label{multi-species-prioritisations}
-
-\subsubsection{Pareo-frontier analysis}\label{pareo-frontier-analysis}
-
-\subsection{Discussion}\label{discussion}
-
-\subsection{Acknowledgements}\label{acknowledgements}
-
-JOH is funded by an Australian Postgraduate Award (APA) scholarship. RAF
-has an Australian Research Council Future Fellowship. This work was
-supported by the Centre of Excellence for Environmental Decisions (CEED)
-and the Landscape Ecology and Conservation Group (LEC) at The University
-of Queensland.
-
-\subsection{References}\label{references}
-
-Benjamini, Y., Hochberg, Y. (1995) Controlling the false discovery rate:
-a practical and powerful approach to multiple testing. \emph{Journal of
-the Royal Statistical Society. Series B (Methodological)}. \textbf{57},
-289--300.
-
-Bothwell, H., Bisbing, S., Therkildsen, N., Crawford, L., Alvarez, N.,
-Holderegger, R., Manel, S. (2013) Identifying genetic signatures of
-selection in a non-model species, alpine gentian (gentiana nivalis l.),
-using a landscape genetic approach. \emph{Conservation Genetics}.
-\textbf{14}, 467--481.
-
-Evanno, G., Regnaut, S., Goudet, J. (2005) Detecting the number of
-clusters of individuals using the software structure: a simulation
-study. \emph{Molecular Ecology}. \textbf{14}, 2611--2620.
-
-Falush, D., Stephens, M., Pritchard, J. K. (2007) Inference of
-population structure using multilocus genotype data: dominant markers
-and null alleles. \emph{Molecular Ecology Notes}. \textbf{7}, 574--578.
-
-Foll, M., Gaggiotti, O. (2008) A genome-scan method to identify selected
-loci appropriate for both dominant and codominant markers: A bayesian
-perspective. \emph{Genetics}. \textbf{180}, 977--993.
-
-Gower, J. C. (1971) A general coefficient of similarity and some of its
-properties. \emph{Biometrics}. \textbf{27}, 857--871.
-
-Gugerli, F., Englisch, T., Niklfeld, H., Tribsch, A., Mirek, Z.,
-Ronikier, M., Zimmermann, N., Holderegger, R., Taberlet, P. (2008)
-Relationships among levels of biodiversity and the relevance of
-intraspecific diversity in conservation - a project synopsis.
-\emph{Perspectives in Plant Ecology, Evolution and Systematics}.
-\textbf{10}, 259--281.
-
-Gurobi Optimization, I. (2015) Gurobi optimizer reference manual.
-
-Hanson, J. O., Rhodes, J. R., Possingham, H. P., Fuller, R. A. (2016)
-rapr: representative and adequate prioritisations in r.
-\emph{Unpublished}. \textbf{X}, XX--XX.
-
-Hijmans, R. J., Cameron, S. E., Parra, J. L., Jones, P. G., Jarvis, A.
-(2005) Very high resolution interpolated climate surfaces for global
-land areas. \emph{International Journal of Climatology}. \textbf{25},
-1965--1978.
-
-Jakobsson, M., Rosenberg, N. A. (2007) CLUMPP: a cluster matching and
-permutation program for dealing with label switching and multimodality
-in analysis of population structure. \emph{Bioinformatics}. \textbf{23},
-1801--1806.
-
-Maechler, M., Rousseeuw, P., Struyf, A., Hubert, M., Hornik, K. (2015)
-\emph{cluster: Cluster analysis basics and extensions}.
-
-Meirmans, P., Goudet, J., IntraBioDiv Consortium, Gaggiotti, O. (2011)
-Ecology and life history affect different aspects of the population
-structure of 27 high-alpine plants. \emph{Molecular Ecology}.
-\textbf{20}, 3144--3155.
-
-Oksanen, J., Blanchet, F. G., Kindt, R., Legendre, P., Minchin, P. R.,
-O'Hara, R. B., Simpson, G. L., Solymos, P., Stevens, M. H. H., Wagner,
-H. (2015) \emph{vegan: Community ecology package}.
-
-P{é}rez-Figueroa, A., Garc{í}a-Pereira, M. J., Saura, M.,
-Rol{á}n-{Á}lvarez, E., Caballero, A. (2010) Comparing three different
-methods to detect selective loci using dominant markers. \emph{Journal
-of Evolutionary Biology}. \textbf{23}, 2267--2276.
-
-Pritchard, J. K., Stephens, M., Donnelly, P. (2000) Inference of
-population structure using multilocus genotype data. \emph{Genetics}.
-\textbf{155}, 945--959.
-
-Vos, P., Hogers, R., Bleeker, M., Reijans, M., Lee, T. van de, Hornes,
-M., Friters, A., Pot, J., Paleman, J., Kuiper, M., Zabeau, M. (1995)
-AFLP: a new technique for dNA fingerprinting. \emph{Nucleic Acids
-Research}. \textbf{23}, 4407--4414.
-
 \documentclass[11pt,]{article}
@@
 \end{document}
-
-\end{document}

### 2015-12-24 article/supporting_information.Rmd
+ updated analysis to use equi-distant crs for attribute space + updated methods section in ms + attempting to set the supp info to be correct

---
fontsize: 11pt
documentclass: article
bibliography: Endnote_lib.bib
csl: reference-style.csl
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    includes:
      in_header: [preamble-latex.tex, preamble-latex2.tex]
---
## Supporting Information

```{r, include=FALSE}
# set wd
setwd('..')

# load .rda files
load('results/results.rda')

# load packages
source('R/analysis/00-initialization.R')

# set options
options(error=NULL)

# download basemap
data(countriesHigh)
countries.FPLY <- countriesHigh[
	countriesHigh$ADMIN %in% c(
		'Italy', 'Switzerland', 'France', 'Austria',
		'Germany', 'Slovenia', 'Croatia', 'Hungary',
		'Monaco', 'Germany'
	)
,] %>% spFortify
```

```{r, echo=FALSE, fig.width=9.5, fig.height=10, out.width='7.5in', out.height='8.5in', fig.cap='Species distributions. Squares represent planning units. For a given species, planning units that were found to be inhabited are denoted with bright blue.'}
## plot map of species distributions
# fortify data
grid.FPLY <- spFortify(grid.PLY)
spp.grid.FPLY <- ldply(unique(spp.samples.DF$species), function(x) {
		z <- grid.FPLY[,c('long', 'lat', 'group', x),drop=FALSE]
		names(z)[4] <- 'presence'
		z$species <- gsub('\\_', ' ', x)
		return(z)
	}
)
# plot species data
ggplot() +
	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
		fill='grey20', color='grey80') +
	geom_polygon(data=spp.grid.FPLY, aes(x=long, y=lat, 
		group=group, fill=presence), alpha=0.8, color='grey10') +
	theme_classic() +
	guides(fill=guide_legend(title='Presence')) +
	theme(axis.ticks=element_blank(), axis.text=element_blank(),
		axis.line=element_blank()) +
	coord_cartesian(
		xlim=buffered.range(grid.FPLY$long, 0.05),
		ylim=buffered.range(grid.FPLY$lat, 0.05)
	) +
	xlab('') +
	ylab('') +
	facet_wrap(~ species, ncol=4)
```

```{r, echo=FALSE, fig.width=4, fig.height=3.5, fig.cap="Species richness. Squares denote planning units. Planning units with a brighter color are inhabited by more species."}
# calculate species richness
grid.PLY$Species_richness <- grid.PLY@data %>%
	select(5:(4+n_distinct(spp.samples.DF$species))) %>% as.matrix() %>% rowSums()

# plot species richness
ggplot() +
	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
		fill='grey20', color='grey80') +
	geom_polygon(data=spFortify(grid.PLY), aes(x=long, y=lat, 
		group=group, fill=Species_richness), alpha=0.8, color='grey10') +
	guides(fill=guide_legend(title='Count (#)')) +
	theme_classic() +
	theme(axis.ticks=element_blank(), axis.text=element_blank(),
		axis.line=element_blank()) +
	coord_cartesian(
		xlim=buffered.range(grid.FPLY$long, 0.05),
		ylim=buffered.range(grid.FPLY$lat, 0.05)
	) +
	xlab('') +
	ylab('') +
	ggtitle('Species richness')
```

```{r, fig.width=6.5, fig.height=3.25, fig.cap='Climatic variation. Each panel depicts variation based on a different principle component (PC). Squares represent planning units. The color of each planning unit denotes the average PC value of pixels inside it. Planning units with more similar colors have more similar climate regimes.', echo=FALSE}
do.call(
	grid.arrange,
		append(
		llply(grep('^env\\_.*$', names(grid.DF), value=TRUE), function(x) {
			ggplot() +
				geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
					fill='grey20', color='grey80') +
				geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
					group='group', fill=x),
					alpha=0.8, color='grey10') +
				guides(fill=guide_legend(title=' ')) +
				theme_classic() +
				theme(axis.ticks=element_blank(), axis.text=element_blank(),
					plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank()) +
				coord_cartesian(
					xlim=buffered.range(grid.FPLY$long, 0.05),
					ylim=buffered.range(grid.FPLY$lat, 0.05)
				) +
				xlab('') +
				ylab('') +
				ggtitle(paste0('PC ', substr(x, nchar(x), nchar(x))))
		}),
		list(ncol=2)
	)
)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=6.5, fig.cap='Relative support ($\\Delta$K) for the number of populations in each species. Each panel denotes a different species. The number of populations with the highest $Delta$K value is has the most support.'}
# format data
deltak.DF <- ldply(
	seq_along(unique(spp.samples.DF$species)),
	.fun=function(i) {
		mutate(
			spp.StructureCollection.LST[[i]]@summary,
			Species=gsub('_', ' ', unique(spp.samples.DF$species)[i], fixed=TRUE)
		)
	}
)
# make plot
suppressWarnings({
	ggplot(data=deltak.DF) +
		geom_point(aes(y=delta.k, x=k)) +
		geom_line(aes(y=delta.k, x=k)) +
		theme_classic() +
		xlab('Number of populations') +
		ylab(expression(Relative~support~(Delta*K))) +
		facet_wrap(~ Species, ncol=6)
})
```

```{r, echo=FALSE, fig.width=9, fig.height=6.5, fig.cap='Population membership plots. Each panel depicts a different species. Each vertical bar represents a different individual. The colors in each bar denote the probability that the individual belongs to different populations.'}
# prepare data
popm.DF <- ldply(
	seq_along(unique(spp.samples.DF$species)),
	.fun=function(i) {
		# init
		raw.popm <- spp.StructureCollection.LST[[i]] %>% 
			structurer:::sample.membership()
		raw.popi <- apply(raw.popm, 1, which.max)
		# compile data
		curr.popm <- raw.popm %>%
			as.data.frame %>%
			mutate(id=as.character(seq_along(V2))) %>%
			gather(id, percent) %>%
			`names<-`(c('id','pop.id','percent')) %>%
			mutate(Species=gsub('\\_', ' ',
				unique(spp.samples.DF$species)[i]),
				pop.id=as.character(as.numeric(gsub('V', '', pop.id))-1))
		# sort data
		sorted.popm <- curr.popm[0,]
		for (i in seq_len(n_distinct(curr.popm$pop.id))) {
				# get individuals with greatest membership
				curr.i <- which(raw.popi==i)
				curr.DF <- filter(curr.popm,
					id %in% as.character(curr.i))
				# sort these inidividuals by membership probability
				curr.i <- curr.i[order(raw.popm[curr.i,i],
					decreasing=TRUE)]
				# rbind these individuals to sorted.popm
				for (j in curr.i)
					sorted.popm <- rbind(
						sorted.popm,
						filter(curr.DF, id==as.character(j))
					)
		}
		sorted.popm$id <- factor(as.character(sorted.popm$id), 
			levels=unique(as.character(sorted.popm$id)))
		return(sorted.popm)
	}
)
# make plot
grobLST <- dlply(popm.DF, .(Species), .fun=function(x) {
	ggplot(data=x, aes(x=id, y=percent, fill=pop.id)) +
		geom_bar(stat='identity', width=1) +
		theme_classic() +
		theme(axis.ticks=element_blank(),
			axis.text=element_blank(),
			axis.title=element_blank(),
			axis.line=element_blank(),
			legend.position='none') +
		ggtitle(x$Species[1])
})
# render plot
do.call(grid.arrange, append(grobLST, list(ncol=4,nrow=6)))
```

```{r, echo=FALSE, fig.width=6.5, fig.height=6.5, fig.cap='Distribution of populations. Each panel depicts a different species. Each pie chart represents a grid cell where individuals belonging to a species were sampled. The colors of the pie chart show the frequency of individuals belonging to different populations.'}
# prepare data
# make plot
plot(1, main='not done')
```

```{r, echo=FALSE}
plot.spp.mds <- function(i) {
	# define function
	make.mds.plot <- function(j,k) {
		ggplot() +
			geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
				fill='grey20', color='grey80') +
			geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
				group='group', fill=paste0(unique(spp.samples.DF$species)[i], '_', j, '_d',k)),
				alpha=0.8, color='grey10') +
			guides(fill=guide_legend(title=' ')) +
			theme_classic() +
			theme(axis.ticks=element_blank(), axis.text=element_blank(),
				plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank()) +
			coord_cartesian(
				xlim=buffered.range(grid.FPLY$long, 0.05),
				ylim=buffered.range(grid.FPLY$lat, 0.05)
			) +
			xlab('') +
			ylab('') +
			ggtitle(paste0(j,' (',k,')'))
	}
	# init
	adapt.col <- paste0(unique(spp.samples.DF$species)[i], '_adaptive_d1')
	neutral.col <- paste0(unique(spp.samples.DF$species)[i], '_neutral_d1')
	plotLST <- list()
	# make adaptive plot
	if (adapt.col %in% names(grid.FPLY)) {
		for (k in seq_len(mds.k))
			plotLST <- append(plotLST, list(make.mds.plot('adaptive', k)))
	} 
	# make neutral plot
	if (neutral.col %in% names(grid.FPLY)) {
		for (k in seq_len(mds.k))
			plotLST <- append(plotLST, list(make.mds.plot('neutral', k)))
	} 
	# make plot
	do.call(
		grid.arrange,
		append(
			plotLST,
			list(ncol=6)
		)
	)
}
```

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[1]),'}. Each square represents a planning unit. The color of each planning unit panel corresponds to ordination values. Planning units with similar colors contain individiduals with similar genetic variation.')}
plot.spp.mds(1)
```

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[2]),'}. See Figure S6 caption for conventions.')}
plot.spp.mds(2)
```

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[3]),'}. See Figure S6 caption for conventions.')}
plot.spp.mds(3)
```

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[4]),'}. See Figure S6 caption for conventions.')}
# plot.spp.mds(4)
```

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[5]),'}. See Figure S6 caption for conventions.')}
# plot.spp.mds(5)
```

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[6]),'}. See Figure S6 caption for conventions.')}
# plot.spp.mds(6)
```

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[7]),'}. See Figure S6 caption for conventions.')}
# plot.spp.mds(7)
```

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[8]),'}. See Figure S6 caption for conventions.')}
# plot.spp.mds(8)
```

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[9]),'}. See Figure S6 caption for conventions.')}
# plot.spp.mds(9)
```

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[10]),'}. See Figure S6 caption for conventions.')}
# plot.spp.mds(10)
```

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[11]),'}. See Figure S6 caption for conventions.')}
# plot.spp.mds(11)
```

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[12]),'}. See Figure S6 caption for conventions.')}
# plot.spp.mds(12)
```

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[13]),'}. See Figure S6 caption for conventions.')}
# plot.spp.mds(13)
```

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[14]),'}. See Figure S6 caption for conventions.')}
# plot.spp.mds(14)
```

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[15]),'}. See Figure S6 caption for conventions.')}
# plot.spp.mds(15)
```

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[16]),'}. See Figure S6 caption for conventions.')}
# plot.spp.mds(16)
```

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[17]),'}. See Figure S6 caption for conventions.')}
# plot.spp.mds(17)
```

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[18]),'}. See Figure S6 caption for conventions.')}
# plot.spp.mds(18)
```

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[19]),'}. See Figure S6 caption for conventions.')}
# plot.spp.mds(19)
```

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[20]),'}. See Figure S6 caption for conventions.')}
# plot.spp.mds(20)
```

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[21]),'}. See Figure S6 caption for conventions.')}
# plot.spp.mds(21)
```

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[22]),'}. See Figure S6 caption for conventions.')}
# plot.spp.mds(22)
```

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[23]),'}. See Figure S6 caption for conventions.')}
# plot.spp.mds(23)
```

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[24]),'}. See Figure S6 caption for conventions.')}
# plot.spp.mds(24)
```

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[25]),'}. See Figure S6 caption for conventions.')}
# plot.spp.mds(25)
```
```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[26]),'}. See Figure S6 caption for conventions.')}
# plot.spp.mds(26)
```

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[27]),'}. See Figure S6 caption for conventions.')}
# plot.spp.mds(27)
```

```{r, echo=FALSE}
## make results table showing Eigen values
knitr::kable(
	pca.DF,
	digits=2,
	caption='Summary of principle components analysis (PCA) on bioclimatic variation across the study area. The first two principle components (PCs) were used for subsequent analysis.',
	align=c('l', 'c', 'c', 'c')
)
```

```{r, echo=FALSE}
knitr::kable(
	format.table(
		ldply(
			seq_along(unique(spp.samples.DF$species)), 
			function(i) {
				data.frame(
					Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
					Primer=spp.BayeScan.sample.loci.subset.LST[[i]]@data@primers,
					Probability=spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[2]],
					qval=spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[4]],
					alpha=spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[5]],
					fst=spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[6]],
					Type=spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[7]]
				)
			}
		),
		omit='Type'
	),
	digits=2,
	col.names=c('Species', 'Primer', 'Probability', 'q-value', '$\\alpha$', '$F_{ST}$', 'Type'),
	align=c('l', 'c', 'c', 'c', 'c', 'c', 'c')
)
```

```{r, echo=FALSE}
knitr::kable(
	format.table(
		filter(
			ldply(
				seq_along(unique(spp.samples.DF$species)), 
				function(i) {
					ldply(
						seq_along(spp.mds.LST[[i]]),
						function(j) {
						data.frame(
							Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
							Loci=names(spp.mds.LST[[i]])[j],
							Stress=ifelse(is.null(spp.mds.LST[[i]][[j]]), NA_real_, spp.mds.LST[[i]][[j]]$stress),
							Converged=ifelse(is.null(spp.mds.LST[[i]][[j]]), NA_real_, spp.mds.LST[[i]][[j]]$converged)
						)
					})
				}
			),
			!is.na(Stress)
		),
		omit=c('Loci','Converged')
	),
	digits=2,
	caption='Summary of non-metric multi-dimensional scaling (MDS) analyses on genetic variation for each species.',
	col.names=c('Species', 'Loci Type', 'NMDS Stress', 'Converged'),
	align=c('l', 'c', 'c', 'c')
)
```

### 2015-12-24 R/analysis/02-surrogate-data.R

@@ -25,8 +25,12 @@ spp.sample.PTS <- SpatialPointsDataFrame(
 spp.sample.PPTS <- spTransform(spp.sample.PTS, europeEA)
 
 ## extract geographic data
-centroids.DF <- gCentroid(grid.PPLY, byid=TRUE) %>% slot('coords') %>%
-	as.data.frame() %>% `names<-`(paste0('geo_d',1:2))
+centroids.DF <- grid.PPLY %>%
+	spTransform(europeED) %>%
+	gCentroid(byid=TRUE) %>%
+	slot('coords') %>%
+	as.data.frame() %>%
+	`names<-`(paste0('geo_d',1:2))
 grid.DF <- cbind(grid.DF, centroids.DF)
 
 ## extract climatic data

### 2015-12-24 R/functions/data.R

@@ -9,9 +9,19 @@ wgs1984 <- CRS('+init=epsg:4326')
 
 #' Europe Lambert Conformal Conic CRS
 #'
-#' This object stores the coordinate reference system (CRS) for the Europe Lambert Conformal Conic projection.
+#' This object stores the equal-area coordinate reference system (CRS) for the Europe Lambert Conformal Conic projection.
 #'
 #' @seealso \code{\link[sp]{CRS}}
 #' @examples
 #' europeEA <- CRS('+init=esri:102014')
 europeEA <- CRS('+init=esri:102014')
+
+#' Europe Equidistant Conic
+#'
+#' This object stores the equi-distant coordinate reference system (CRS) for the  projection.
+#'
+#' @seealso \code{\link[sp]{CRS}}
+#' @examples
+#' europeED <- CRS('+init=esri:102031')
+europeED <- CRS('+init=esri:102031')
+

### 2016-01-01 article/article.Rmd
- set parameters to TOML format - make parameters loaded only in when needed to reduce overheads if parameters are changed

@@
 We used the Bayesian clustering method implemented in $\texttt{Structure}$ [version 2.3.4; @r461; @r462] assuming admixture and an independent alleles model.
-For each species, we identified the most plausible number of populations (K = `r st.k[1]`--`r st.k[2]`; `r st.numruns` runs per K; `r st.burnin` burn-in iterations; `r st.numreps` post-burn-in iterations) using the $\Delta$K method [Figure S3; @r460].
-We aggregated the population membership probabilities from replicate runs using $\texttt{CLUMPP}$ [version 1.1.2; large K greedy search method with `r cl.repeats` random starts per K per species; Figure S4; @r459].
-Individuals with a moderate probability of belonging to a single population ($\geq `r st.probthresh`$) were assigned accordingly (Figure S5), and subject to subsequent outlier locus detection methods.
+For each species, we identified the most plausible number of populations (K = `r min(structure.params.LST[[MODE]]$k)`--`r max(structure.params.LST[[MODE]]$k)`; `r structure.params.LST[[MODE]]$numruns` runs per K; `r structure.params.LST[[MODE]]$burnin` burn-in iterations; `r structure.params.LST[[MODE]]$numreps` post-burn-in iterations) using the $\Delta$K method [Figure S3; @r460].
+We aggregated the population membership probabilities from replicate runs using $\texttt{CLUMPP}$ [version 1.1.2; large K greedy search method with `r clumpp.params.LST[[MODE]]$repeats` random starts per K per species; Figure S4; @r459].
+Individuals with a moderate probability of belonging to a single population ($\geq `r structure.params.LST[[MODE]]$probthresh`$) were assigned accordingly (Figure S5), and subject to subsequent outlier locus detection methods.
 
 We used $\texttt{Bayescan}$ [version 2.1; @r456] to identify loci under selection for each species (Table S2).
-For each species, $\texttt{Bayescan}$ was run using the population memberships identified in the previous analysis (1:1 prior odds; `r bs.nbp` pilot runs; `r bs.burn` burn-in iterations; `r bs.n` post-burn-in iterations thinned by `r bs.thin` iterations).
+For each species, $\texttt{Bayescan}$ was run using the population memberships identified in the previous analysis (1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations).
 Following guidelines in the $\texttt{Bayescan}$ manual, we omitted loci where the global frequency of the minor allele was $\geq 0.05$ to avoid false-positives.
-To ensure convergence, we ran `r bs.reps` replicates per species using a suitable false discovery rate [FDR; $\leq `r bs.fdr`$; @r466].
+To ensure convergence, we ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate [FDR; $\leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].
 
@@
 For each species, for each type of genetic variation present, we used Gower distances [@r464] to express the genomic differences between individuals [using the cluster R package to accommodate sparsity; @r457].
-These distances were subject to a non metric multi-dimensional scaling analysis [implemented in the vegan R package using K=`r mds.k` and `r mds.trymax` random starts; Table S3; @r458].
+These distances were subject to a non metric multi-dimensional scaling analysis [implemented in the vegan R package using K=`r mds.params.LST[[MODE]]$k` and `r mds.params.LST[[MODE]]$trymax` random starts; Table S3; @r458].
 The average of the ordinated values associated with individuals in each grid cell we used to characterize the genomic variation in the grid cell.
@@
 First, we generated a set of single species prioritizations.
-For each species, we generated a prioritization using a `r (rapr.amount.target*100)`% amount-based target, a prioritization using a `r (rapr.amount.target*100)`% amount-based target and `r (rapr.surrogate.target*100)`% space-based targets for the surrogate variation, and a prioritization using a `r (rapr.amount.target*100)`% amount-based target and `r (rapr.genetic.target*100) `% space-based targets for the genomic variation.
+For each species, we generated a prioritization using a `r (rapr.params.LST[[MODE]]$amount.target*100)`% amount-based target, a prioritization using a `r (rapr.params.LST[[MODE]]$amount.target*100)`% amount-based target and `r (rapr.params.LST[[MODE]]$surrogate.target*100)`% space-based targets for the surrogate variation, and a prioritization using a `r (rapr.params.LST[[MODE]]$amount.target*100)`% amount-based target and `r (rapr.params.LST[[MODE]]$genetic.target*100) `% space-based targets for the genomic variation.
 Next, we generated three multi-species prioritizations using the same sets of targets used in the single species prioritizations.
-Finally, to understand the relationship between the surrogate-based and genomic-based targets, we generated a suite of prioritizations using increasing space-based targets between `r (min(rapr.pareto.surrogate.targets)*100)`% and `r (max(rapr.pareto.surrogate.targets)*100)`% in increments of `r diff(rapr.pareto.surrogate.targets[1:2])*100`%.
+Finally, to understand the relationship between the surrogate-based and genomic-based targets, we generated a suite of prioritizations using increasing space-based targets between `r (min(rapr.params.LST[[MODE]]$pareto.surrogate.targets)*100)`% and `r (max(rapr.params.LST[[MODE]]$pareto.surrogate.targets)*100)`% in increments of `r diff(rapr.params.LST[[MODE]]$pareto.surrogate.targets[1:2])*100`%.
 For each prioritization we calculated the proportion of the adaptive (if present) and neutral genetic variation that it secured.
-We solved the reserve selection problems to within `r (gb.MIPGap*100)`% of optimality using the $\texttt{rapr}$ R package [@r465] and $\texttt{Gurobi}$ [@r463].
+We solved the reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)`% of optimality using the $\texttt{rapr}$ R package [@r465] and $\texttt{Gurobi}$ [@r463].
 

### 2016-01-01 article/article.tex
- set parameters to TOML format - make parameters loaded only in when needed to reduce overheads if parameters are changed

@@
   \postdate{\par}
-  \date{24 December 2015}
+  \date{01 January 2016}
 
 % load packages
-\usepackage{amsmath,amsfonts,float,makecell,titletoc,tocloft,titlesec,natbib}
+\usepackage{amsmath,amsfonts,float,makecell,titletoc,tocloft,titlesec,natbib,pdfpages}
 \usepackage[T1]{fontenc}
@@
 independent alleles model. For each species, we identified the most
-plausible number of populations (K = 1--2; 4 runs per K; 10 burn-in
+plausible number of populations (K = 1--3; 4 runs per K; 10 burn-in
 iterations; 10 post-burn-in iterations) using the $\Delta$K method
@@
 
-\documentclass[11pt,]{article}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-    \usepackage{xltxtra,xunicode}
-  \else
-    \usepackage{fontspec}
-  \fi
-  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
-  \newcommand{\euro}{€}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
-\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\usepackage[margin=1in]{geometry}
-\usepackage{graphicx}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\ifxetex
-  \usepackage[setpagesize=false, % page size defined by xetex
-              unicode=false, % unicode breaks when used with xetex
-              xetex]{hyperref}
-\else
-  \usepackage[unicode=true]{hyperref}
-\fi
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={},
-            pdftitle={},
-            colorlinks=true,
-            citecolor=blue,
-            urlcolor=blue,
-            linkcolor=magenta,
-            pdfborder={0 0 0}}
-\urlstyle{same}  % don't use monospace font for urls
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\setcounter{secnumdepth}{0}
-
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-%%% Change title format to be more compact
-\usepackage{titling}
-
-% Create subtitle command for use in maketitle
-\newcommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
-
-\setlength{\droptitle}{-2em}
-  \title{}
-  \pretitle{\vspace{\droptitle}}
-  \posttitle{}
-  \author{}
-  \preauthor{}\postauthor{}
-  \date{}
-  \predate{}\postdate{}
-
-\begin{document}
-
-\maketitle
-
-\subsection{Figures}\label{figures}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures_files/figure-latex/unnamed-chunk-2-1.pdf}
-\caption{Summary of single species prioritisations. Single species
-prioritisations were generated using amount-based targets, amount-based
-and surrogate-based targets, and amount-based and genetic-based targets
-for each species. Data shows the performance of prioritisations
-generated using these three sets of targets. Bars denote means and
-standard errors.}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures_files/figure-latex/unnamed-chunk-3-1.pdf}
-\caption{Multi-species prioritisations. Panel (a) shows the
-prioritisation generated for using just amount-based targets. Panel (b)
-shows the prioritisation generated using amount-based and surrogate
-based targets. Panel (c) shows the prioritisation generated using
-amount-based and genetic-based targets}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures_files/figure-latex/unnamed-chunk-4-1.pdf}
-\caption{Summary of multi-species prioritisations. Three prioritisations
-were generated using amount-based targets, amount-based and
-surrogate-based targets, and amount-based and genetic-based targets for
-each species. Data shows the performance of these prioritisations based
-on how much genetic variation they explain. Bars denote means and
-standard errors.}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures_files/figure-latex/unnamed-chunk-5-1.pdf}
-\caption{The relationship between surrogates and genetic variation
-secured in prioritizations.}
-\end{figure}
+\includepdf[pages=-]{figures.pdf}
+\includepdf[pages=-]{supporting_information.pdf}
 
 \end{document}
-
-\end{document}

### 2016-01-01 article/supporting_information.Rmd
- set parameters to TOML format - make parameters loaded only in when needed to reduce overheads if parameters are changed

@@
 	if (adapt.col %in% names(grid.FPLY)) {
-		for (k in seq_len(mds.k))
+		for (k in seq_len(mds.params.LST[[MODE]]$k))
 			plotLST <- append(plotLST, list(make.mds.plot('adaptive', k)))
@@
 	if (neutral.col %in% names(grid.FPLY)) {
-		for (k in seq_len(mds.k))
+		for (k in seq_len(mds.params.LST[[MODE]]$k))
 			plotLST <- append(plotLST, list(make.mds.plot('neutral', k)))

### 2016-01-01 R/analysis/08-pareto-frontier-analysis.R

@@ -4,7 +4,7 @@ session::restore.session('results/.cache/07-multi-species-prioritisations.rda')
 ## pareto frontier analysis
 # generate prioritistions
 env.pareto.prioritisations <- llply(
-	rapr.pareto.surrogate.targets,
+	rapr.params.LST[[MODE]]$pareto.surrogate.targets,
 	species.prioritisation,
 	x=ru,
 	amount.targets=0,
@@ -13,7 +13,7 @@ env.pareto.prioritisations <- llply(
 	neutral.genetic.targets=0
 )
 geo.pareto.prioritisations <- llply(
-	rapr.pareto.surrogate.targets,
+	rapr.params.LST[[MODE]]$pareto.surrogate.targets,
 	species.prioritisation,
 	x=ru,
 	amount.targets=0,
@@ -26,13 +26,13 @@ geo.pareto.prioritisations <- llply(
 env.pareto.DF <- ldply(seq_along(env.pareto.prioritisations), function(i) {
 	mutate(
 		extractResults(env.pareto.prioritisations[[i]]),
-		Surrogate.target=rapr.pareto.surrogate.targets[i]
+		Surrogate.target=rapr.params.LST[[MODE]]$pareto.surrogate.targets[i]
 	)
 })
 geo.pareto.DF <- ldply(seq_along(geo.pareto.prioritisations), function(i) {
 	mutate(
 		extractResults(geo.pareto.prioritisations[[i]]),
-		Surrogate.target=rapr.pareto.surrogate.targets[i]
+		Surrogate.target=rapr.params.LST[[MODE]]$pareto.surrogate.targets[i]
 	)
 })
 

### 2016-01-16 article/supporting_information.Rmd
use adegenet for population structure analysis instead of structure + clumpp

@@
 ,] %>% spFortify
+grid.FPLY <- spFortify(grid.PLY)
 ```
@@
 # fortify data
-grid.FPLY <- spFortify(grid.PLY)
 spp.grid.FPLY <- ldply(unique(spp.samples.DF$species), function(x) {
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=6.5, fig.cap='Relative support ($\\Delta$K) for the number of populations in each species.
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=6.5, fig.cap='Relative support ($BIC$) for the number of populations in each species.
 Each panel denotes a different species.
-The number of populations with the highest $Delta$K value is has the most support.'}
+Asterisks indicate the number of populations with the most support.
+See text for the method used to select these populations.'}
 # format data
-deltak.DF <- ldply(
+bic.buff <- laply(spp.Adegenet.LST, `[[`, 'Kstat') %>%
+	c() %>% range() %>% diff %>% `*`(0.05)
+bic.DF <- ldply(
 	seq_along(unique(spp.samples.DF$species)),
 	.fun=function(i) {
-		mutate(
-			spp.StructureCollection.LST[[i]]@summary,
+		data.frame(
+			k=as.numeric(gsub('K=', '', names(spp.Adegenet.LST[[i]]$Kstat), fixed=TRUE)),
+			stat=unname(spp.Adegenet.LST[[i]]$Kstat),
+			best_k=as.numeric(gsub('K=', '', names(spp.Adegenet.LST[[i]]$stat), fixed=TRUE)),
+			best_stat=unname(spp.Adegenet.LST[[i]]$stat) + bic.buff,
+			asterisk_col=c('black',rep('transparent', length(spp.Adegenet.LST[[i]]$Kstat)-1)),
 			Species=gsub('_', ' ', unique(spp.samples.DF$species)[i], fixed=TRUE)
@@
 suppressWarnings({
-	ggplot(data=deltak.DF) +
-		geom_point(aes(y=delta.k, x=k)) +
-		geom_line(aes(y=delta.k, x=k)) +
+	ggplot(data=bic.DF) +
+		geom_point(aes(y=stat, x=k)) +
+		geom_point(aes(y=best_stat, x=best_k), size=4, shape=8) +
+		geom_line(aes(y=stat, x=k)) +
 		theme_classic() +
 		xlab('Number of populations') +
-		ylab(expression(Relative~support~(Delta*K))) +
+		ylab('BIC') +
 		facet_wrap(~ Species, ncol=6)
@@
 
-```{r, echo=FALSE, fig.width=9, fig.height=6.5, fig.cap='Population membership plots.
+```{r, echo=FALSE, fig.width=6.5, fig.height=6.5, fig.cap='Distribution of populations.
 Each panel depicts a different species.
-Each vertical bar represents a different individual.
-The colors in each bar denote the probability that the individual belongs to different populations.'}
-# prepare data
-popm.DF <- ldply(
+Squares denote planning units.
+Points denote the location of sampled individuals.
+The color of each point indicates the population to which each indiviudal was assigned.'}
+# define function
+plotLST <- llply(
 	seq_along(unique(spp.samples.DF$species)),
@@
 		# init
-		raw.popm <- spp.StructureCollection.LST[[i]] %>% 
-			structurer:::sample.membership()
-		raw.popi <- apply(raw.popm, 1, which.max)
-		# compile data
-		curr.popm <- raw.popm %>%
-			as.data.frame %>%
-			mutate(id=as.character(seq_along(V2))) %>%
-			gather(id, percent) %>%
-			`names<-`(c('id','pop.id','percent')) %>%
-			mutate(Species=gsub('\\_', ' ',
-				unique(spp.samples.DF$species)[i]),
-				pop.id=as.character(as.numeric(gsub('V', '', pop.id))-1))
-		# sort data
-		sorted.popm <- curr.popm[0,]
-		for (i in seq_len(n_distinct(curr.popm$pop.id))) {
-				# get individuals with greatest membership
-				curr.i <- which(raw.popi==i)
-				curr.DF <- filter(curr.popm,
-					id %in% as.character(curr.i))
-				# sort these inidividuals by membership probability
-				curr.i <- curr.i[order(raw.popm[curr.i,i],
-					decreasing=TRUE)]
-				# rbind these individuals to sorted.popm
-				for (j in curr.i)
-					sorted.popm <- rbind(
-						sorted.popm,
-						filter(curr.DF, id==as.character(j))
+		curr.pts <- filter(
+			spp.sample.PTS@data,
+			species==unique(spp.samples.DF$species)[i]
+		) %>% mutate(
+			population=spp.Adegenet.LST[[i]]$grp
 		)
-		}
-		sorted.popm$id <- factor(as.character(sorted.popm$id), 
-			levels=unique(as.character(sorted.popm$id)))
-		return(sorted.popm)
+		# create plot
+		ggplot() +
+			geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
+				fill='grey20', color='grey80') +
+			geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
+				group='group'), fill='transparent', color='grey10') +
+			geom_point(data=curr.pts, aes_string(x='sample.longitude', y='sample.latitude', color='population'),
+				alpha=0.3, size=3, position='jitter') +
+			guides(fill=guide_legend(title=' ')) +
+			theme_classic() +
+			theme(axis.ticks=element_blank(), axis.text=element_blank(),
+				plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank()) +
+			coord_cartesian(
+				xlim=buffered.range(grid.FPLY$long, 0.05),
+				ylim=buffered.range(grid.FPLY$lat, 0.05)
+			) +
+			xlab('') +
+			ylab('') +
+			ggtitle(gsub('_', ' ', unique(spp.samples.DF$species)[i], fixed=TRUE))
 	}
 )
-# make plot
-grobLST <- dlply(popm.DF, .(Species), .fun=function(x) {
-	ggplot(data=x, aes(x=id, y=percent, fill=pop.id)) +
-		geom_bar(stat='identity', width=1) +
-		theme_classic() +
-		theme(axis.ticks=element_blank(),
-			axis.text=element_blank(),
-			axis.title=element_blank(),
-			axis.line=element_blank(),
-			legend.position='none') +
-		ggtitle(x$Species[1])
-})
 # render plot
-do.call(grid.arrange, append(grobLST, list(ncol=4,nrow=6)))
-```
-
-```{r, echo=FALSE, fig.width=6.5, fig.height=6.5, fig.cap='Distribution of populations.
-Each panel depicts a different species.
-Each pie chart represents a grid cell where individuals belonging to a species were sampled.
-The colors of the pie chart show the frequency of individuals belonging to different populations.'}
-# prepare data
-# make plot
-plot(1, main='not done')
+do.call(
+	grid.arrange,
+	append(
+		plotLST,
+		list(ncol=6)
+	)
+)
 ```

### 2016-01-16 R/analysis/03-adegenet-analysis.R

@@ -1,0 +1,55 @@
+## load .rda
+session::restore.session('results/.cache/02-surrogate-data.rda')
+
+## load parameters
+adegenet.params.LST <- parseTOML('parameters/adegenet.toml')
+
+### adegenet analyses
+## run analyses
+clust <- makeCluster(adegenet.params.LST[[MODE]]$threads, type='SOCK')
+clusterEvalQ(clust, {library(structurer);library(adegenet)})
+clusterExport(clust, c('adegenet.params.LST','MODE', 'spp.StructureData.LST', 'spp.samples.DF'))
+registerDoParallel(clust)
+spp.Adegenet.LST <- llply(
+	seq_along(spp.StructureData.LST),
+	.fun=function(i) {
+		# run adegenet analysis
+		return(
+			find.clusters(
+				spp.StructureData.LST[[i]]@matrix,
+				stat = adegenet.params.LST[[MODE]]$stat,
+				choose.n.clust = adegenet.params.LST[[MODE]]$choose.n.clust,
+				criterion = adegenet.params.LST[[MODE]]$criterion,
+				max.n.clust = adegenet.params.LST[[MODE]]$max.n.clust,
+				n.iter = adegenet.params.LST[[MODE]]$n.iter,
+				n.start = adegenet.params.LST[[MODE]]$n.start,
+				center = adegenet.params.LST[[MODE]]$center,
+				scale = adegenet.params.LST[[MODE]]$scale,
+				pca.select = adegenet.params.LST[[MODE]]$pca.select,
+				perc.pca = adegenet.params.LST[[MODE]]$perc.pca
+			)
+		)
+	},
+	.parallel=TRUE
+)
+clust <- stopCluster(clust)
+
+## assign populations
+spp.BayeScanData.sample.subset.LST <- llply(
+	seq_along(spp.BayeScanData.LST),
+	.fun=function(i) {
+		print(i)
+		# get population identities
+		ids <- spp.Adegenet.LST[[i]]$grp
+		validPos <- which(!is.na(ids))
+		# remove individuals below threshold
+		curr.spp <- spp.BayeScanData.LST[[i]]
+		curr.spp <- bayescanr:::sample.subset.BayeScanData(curr.spp, validPos)
+		curr.spp@labels <- filter(spp.samples.DF, species==unique(spp.samples.DF$species)[i])$cell[validPos]
+		curr.spp@populations <- as.character(ids[validPos])
+		return(curr.spp)
+	}
[…]

### 2016-01-17 article/supporting_information.Rmd
- update genetic mds to use varying number of k-values till stress is suitable

@@
 	if (adapt.col %in% names(grid.FPLY)) {
-		for (k in seq_len(mds.params.LST[[MODE]]$k))
+		for (k in seq_len(spp.mds.LST[[i]][['adaptive']]$ndim))
 			plotLST <- append(plotLST, list(make.mds.plot('adaptive', k)))
@@
 	if (neutral.col %in% names(grid.FPLY)) {
-		for (k in seq_len(mds.params.LST[[MODE]]$k))
+		for (k in seq_len(spp.mds.LST[[i]][['neutral']]$ndim))
 			plotLST <- append(plotLST, list(make.mds.plot('neutral', k)))
@@
 							Loci=names(spp.mds.LST[[i]])[j],
+							k=ifelse(is.null(spp.mds.LST[[i]][[j]]), NA_integer_, spp.mds.LST[[i]][[j]]$ndim),
 							Stress=ifelse(is.null(spp.mds.LST[[i]][[j]]), NA_real_, spp.mds.LST[[i]][[j]]$stress),
-							Converged=ifelse(is.null(spp.mds.LST[[i]][[j]]), NA_real_, spp.mds.LST[[i]][[j]]$converged)
+							Converged=ifelse(is.null(spp.mds.LST[[i]][[j]]), NA, spp.mds.LST[[i]][[j]]$converged)
 						)
@@
 	caption='Summary of non-metric multi-dimensional scaling (MDS) analyses on genetic variation for each species.',
-	col.names=c('Species', 'Loci Type', 'NMDS Stress', 'Converged'),
-	align=c('l', 'c', 'c', 'c')
+	col.names=c('Species', 'Loci Type', 'k', 'NMDS Stress', 'Converged'),
+	align=c('l', 'c', 'c', 'c', 'c')
 )

### 2016-01-18 article/article.Rmd

@@
 <!-- NOTES -->
+<!-- 200 word limit for abstract -->
 <!-- 6 page limit w/o incurring costs -->
@@
 ## Introduction
-<!-- 200 word limit -->
 
@@
 First, we grouped individuals into genetic lineages for each species (Figure S2).
-The purpose of this step was to avoid confounding variation between populations with due to genetic history with variation attributed to selection pressures.
-We used the Bayesian clustering method implemented in $\texttt{Structure}$ [version 2.3.4; @r461; @r462] assuming admixture and an independent alleles model.
-For each species, we identified the most plausible number of populations (K = `r min(structure.params.LST[[MODE]]$k)`--`r max(structure.params.LST[[MODE]]$k)`; `r structure.params.LST[[MODE]]$numruns` runs per K; `r structure.params.LST[[MODE]]$burnin` burn-in iterations; `r structure.params.LST[[MODE]]$numreps` post-burn-in iterations) using the $\Delta$K method [Figure S3; @r460].
-We aggregated the population membership probabilities from replicate runs using $\texttt{CLUMPP}$ [version 1.1.2; large K greedy search method with `r clumpp.params.LST[[MODE]]$repeats` random starts per K per species; Figure S4; @r459].
-Individuals with a moderate probability of belonging to a single population ($\geq `r structure.params.LST[[MODE]]$probthresh`$) were assigned accordingly (Figure S5), and subject to subsequent outlier locus detection methods.
+The purpose of this step was to avoid confounding variation between populations attributed to genetic history instead of selection pressures.
+We used discriminant analyses of principle components [DAPC; @r467] to assign individuals to populations [implemented in \texttt{adegenet} R package; @r468].
+Briefly, for each species, we transformed the binary loci data using a principle components analysis (PCA) and obtained the principle components required to explain at least `r adegenet.params.LST[[MODE]]$perc.pca`% variation.
+We subject these components to multiple $k$-means clustering models with varying numbers of assumed populations (1--`r adegenet.params.LST[[MODE]]$max.n.clust`; `r adegenet.params.LST[[MODE]]$n.inter` iterations; `r adegenet.params.LST[[MODE]]$n.start` random starts).
+We then  applied Ward's clustering method to the Bayesian Information Criterion (BIC) statistics associated with each model to partition the models into two groups.
+We selected the model before the first group switch, and used the population memberships in that model for subsequent analysis.
 
@@
 After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if present) and neutral genetic variation for each species (Figures S4--S29).
-Note that while individuals not assigned to populations were omitted from the $\texttt{Bayescan}$ analysis, here we used all the individuals.
-For each species, for each type of genetic variation present, we used Gower distances [@r464] to express the genomic differences between individuals [using the cluster R package to accommodate sparsity; @r457].
-These distances were subject to a non metric multi-dimensional scaling analysis [implemented in the vegan R package using K=`r mds.params.LST[[MODE]]$k` and `r mds.params.LST[[MODE]]$trymax` random starts; Table S3; @r458].
-The average of the ordinated values associated with individuals in each grid cell we used to characterize the genomic variation in the grid cell.
+For each species, for each type of genetic variation present, we used Gower distances [@r464] to express the genomic differences between individuals [using the $\texttt{cluster}$ R package to accommodate sparsity; @r457].
+These distances were subject to multiple non-metric multi-dimensional scaling analyses using the minimum number of dimensions required to acheive an adequate stress value  [`r mds.params.LST[[MODE]]$max.stress`; `r mds.params.LST[[MODE]]$trymax` random starts; Table S3; implemented in the \texttt{vegan} R package; @r458].
+Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the samples in each grid cell to characterize the genomic variation among the grid cells.
 
@@
 We used the unreliable representative and adequate prioritization (URAP) formulation of the reserve selection problem [@r465].
+This formulation can identify the minimum number of planning units required to preserve a proportion of the species' range as well as a proportion of the species' variation [using an attribute space; _sensu_ @r465].
 We treated the genomic and surrogate variables as distinct attribute spaces.
 We set the demand points for each species and its attribute space using the values associated with the planning units.
+
+To fulfil the aims of the study, we generated three sets of prioritisations.
 First, we generated a set of single species prioritizations.
@@
 Finally, to understand the relationship between the surrogate-based and genomic-based targets, we generated a suite of prioritizations using increasing space-based targets between `r (min(rapr.params.LST[[MODE]]$pareto.surrogate.targets)*100)`% and `r (max(rapr.params.LST[[MODE]]$pareto.surrogate.targets)*100)`% in increments of `r diff(rapr.params.LST[[MODE]]$pareto.surrogate.targets[1:2])*100`%.
+All of these prioritisations had no amount-based targets.
 For each prioritization we calculated the proportion of the adaptive (if present) and neutral genetic variation that it secured.
 We solved the reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)`% of optimality using the $\texttt{rapr}$ R package [@r465] and $\texttt{Gurobi}$ [@r463].
+
+## Statistical analysis
+We tested for differences in the effectiveness of prioritisations at securing genomic variation using general linear models (GLMs).
+These models were fit to the proportion of genomic variation explained using a logit link.
+They contained a predictor variable indicating if the prioritisation was generated using amount, surrogate, or genetic targets, a second predictor variable indicating if the genomic variation was adaptive or neutral, and the interaction between these variables.
+These models were subject to stepwise term deletion routes using likelihood ratio tests (LRT) to identify the minimal adequate model (terms were retained when $P < 0.05$).
 

### 2016-01-18 article/supporting_information.Rmd

@@
 Planning units with similar colors contain individiduals with similar genetic variation.')}
-plot.spp.mds(1)
+if (1 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(1)
 ```
@@
 See Figure S6 caption for conventions.')}
-plot.spp.mds(2)
+if (2 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(2)
 ```
@@
 See Figure S6 caption for conventions.')}
-plot.spp.mds(3)
+if (3 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(3)
 ```
@@
 See Figure S6 caption for conventions.')}
-# plot.spp.mds(4)
+if (4 <= general.params.LST[[MODE]]$n.spp) if (4 <= ) plot.spp.mds(4)
 ```
@@
 See Figure S6 caption for conventions.')}
-# plot.spp.mds(5)
+if (5 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(5)
 ```
@@
 See Figure S6 caption for conventions.')}
-# plot.spp.mds(6)
+if (6 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(6)
 ```
@@
 See Figure S6 caption for conventions.')}
-# plot.spp.mds(7)
+if (7 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(7)
 ```
@@
 See Figure S6 caption for conventions.')}
-# plot.spp.mds(8)
+if (8 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(8)
 ```
@@
 See Figure S6 caption for conventions.')}
-# plot.spp.mds(9)
+if (9 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(9)
 ```
@@
 See Figure S6 caption for conventions.')}
-# plot.spp.mds(10)
+if (10 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(10)
 ```
@@
 See Figure S6 caption for conventions.')}
-# plot.spp.mds(11)
+if (11 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(11)
 ```
@@
 See Figure S6 caption for conventions.')}
-# plot.spp.mds(12)
+if (12 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(12)
 ```
@@
 See Figure S6 caption for conventions.')}
-# plot.spp.mds(13)
+if (13 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(13)
 ```
@@
 See Figure S6 caption for conventions.')}
-# plot.spp.mds(14)
+if (14 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(14)
 ```
@@
 See Figure S6 caption for conventions.')}
-# plot.spp.mds(15)
+if (15 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(15)
 ```
@@
 See Figure S6 caption for conventions.')}
-# plot.spp.mds(16)
+if (16 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(16)
 ```
@@
 See Figure S6 caption for conventions.')}
-# plot.spp.mds(17)
+if (17 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(17)
 ```
@@
 See Figure S6 caption for conventions.')}
-# plot.spp.mds(18)
+if (18 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(18)
 ```
@@
 See Figure S6 caption for conventions.')}
-# plot.spp.mds(19)
+if (19 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(19)
 ```
@@
 See Figure S6 caption for conventions.')}
-# plot.spp.mds(20)
+if (20 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(20)
 ```
@@
 See Figure S6 caption for conventions.')}
-# plot.spp.mds(21)
+if (21 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(21)
 ```
@@
 See Figure S6 caption for conventions.')}
-# plot.spp.mds(22)
+if (22 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(22)
 ```
@@
 See Figure S6 caption for conventions.')}
-# plot.spp.mds(23)
+if (23 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(23)
 ```
@@
 See Figure S6 caption for conventions.')}
-# plot.spp.mds(24)
+if (24 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(24)
 ```
@@
 See Figure S6 caption for conventions.')}
-# plot.spp.mds(25)
+if (25 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(25)
 ```
@@
 See Figure S6 caption for conventions.')}
-# plot.spp.mds(26)
+if (26 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(26)
 ```
@@
 See Figure S6 caption for conventions.')}
-# plot.spp.mds(27)
+if (27 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(27)
 ```

### 2016-01-19 article/article.Rmd
scaffold for intro section of ms

@@
 ## Introduction
+
+<!--
+paragraph 1
+topic sentence: biodiversity is in crisis (@rXX, @rXX).
+	+ in recent years many species have gone extinct (@rXX).
+	+ currently there are thousands of species listed as endangered or near threatened in the International Union for Conservation of Nature (IUCN) red list (www.iucnredlist.org).
+	+ futhermore, climate change is expected cause even greater loss to global biodiversity (@rXXX).
+linking sentence: in response to this crisis, the overarching aim of conservation is to ensure the persistance of biodiversity patterns and processes.
+-->
+
+Biodiversity is under threat.
+In recent years many species have gone extinct [@r272; @r270].
+currently there are thousands of species listed as endangered or near threatened in the International Union for Conservation of Nature (IUCN) red list (www.iucnredlist.org).
+Futhermore, climate change is expected cause even greater loss to global biodiversity [@r236].
+In response to this crisis, the overarching aim of conservation is to maximise the long-term persistance of conservation [@r255].
+
+<!--
+paragraph 2
+topic sentence: the act of setting aside areas for the preservation of biodiversity is one the hallmarks of conservation.
+	+ protected areas buffer species from anthropogenic impacts and provide areas for biodiversity processes to persist (eg. nutrient cycling and water purification).
+	+ additionally, protected areas set the stage for direct management actions (eg. habitat rehabitilation, translocation, and invasive species management).
+	+ since the resources avaliable for conservation are limited, the plans for protected areas (prioritizations) are often generated using biodiversity data (eg.
+species distribution data; @rXXX) , expressing the conservation planning problem as an optimisation problem, and solving it.
+linking sentence:
+-->
+
+The act of setting aside areas for the preservation of biodiversity is one the hallmarks of modern conservation [@r440].
+Protected areas buffer species from anthropogenic impacts and provide areas for biodiversity processes to persist (eg.
+nutrient cycling and water purification).
+Additionally, protected areas set the stage for direct management actions (eg.
+habitat rehabitilation, translocation, and invasive species management).
+Since the resources avaliable for conservation are limited, the plans for protected areas (prioritizations) are often generated by using biodiversity data and decision support tools [eg.
+\texttt{MARXAN}, @r22; \texttt{Zonation}, @r429] to identify cost-effective prioritizations.
+
+<!--
+paragraph 3
+topic sentence: intra-specific genetic variation is important for species persistance.
+	+ Genetic variation is classified as either adaptive or neutral. 
+	+ Adaptive genetic variation refers to loci currently under selectiom.
+	+ Neutral genetic variation to loci that are not under selection, and reflects the evolutionary history of populations.
+	+ By preserving existing patterns of adaptive genetic variation, protected areas can ensure that populations with particularly beneficial adaptations are not lost.
+	+ In constrast, by preserving existing patterns of neutral genetic variation, protected areas can secure populations belonging to genetic lineages and avoid inbreeding depresssion.
+	+ Protected areas must preserve a representative sample of both of these types of genetic variation to fulfil the overarching aims of conservation.
+linking sentence: However, genetic data is costly to obtain and requires expert knowledge to process for use in conservation planning.
+As a consequence, genetic data is not used to inform large-scale prioritisations, and so these prioritisations may fail to preserve the intra-specific genetic variation for the species they aim to preserve.
+-->
+
+<!--
+paragraph 5
+topic sentence: Cost-effective surrogate data could be used to identify prioritizations that preserve intra-specific genetic variation.
++ Adaptive genetic variation is driven by selection pressures.
++ Conservation planners have explored the use of data on the distribution of likely selection pressures to generate prioritizations to secure adaptive genetic variation [eg. @rXXX; @rXXX; @rXXXX].
++ On the other hand, neutral genetic variation arises due to the break down of gene flow between populations.
++ Conservation planners have explored the use of variables that predict a reduction in gene flow between populations.
++ The isolation by distance theory [IBD; @rXX] predicts that populations are further apart will have lower gene flow.
++ The use of geographical surrogate data has been found to improve prioritisations in an island network [@rXX].
++ However this remains untested in system with in a terrestrial study system where connectivity is affected additional factors (eg. elevation).
+linking sentence:
+-->
+
+<!--
+paragraph 6
+topic sentence: The aim of this study is to investigate the effectiveness of environmental and geographic surrogates for conservation planning.
+	+ First, we aim to test the effectiveness of these surrogates under the simplest of conditions in a set of single-species prioritisations.
+	+ Second, we aim to test the effectiveness of these surrogates in a more realistic multi-species context.
+	+ Third, we aim to understand the relationship between the proportion of surrogate variation secured in a multi-species prioritisation and the prorpotion of genetic variation secured in a prioritization.
+linking sentence:
+-->
 
@@
 ## Statistical analysis
-We tested for differences in the effectiveness of prioritisations at securing genomic variation using general linear models (GLMs).
+We tested for differences in the effectiveness of prioritisations at securing genomic variation using generalized linear mixed-effects models (GLMMs).
 These models were fit to the proportion of genomic variation explained using a logit link.
 They contained a predictor variable indicating if the prioritisation was generated using amount, surrogate, or genetic targets, a second predictor variable indicating if the genomic variation was adaptive or neutral, and the interaction between these variables.
-These models were subject to stepwise term deletion routes using likelihood ratio tests (LRT) to identify the minimal adequate model (terms were retained when $P < 0.05$).
+All models contained an intercept-only random effect to ensure that comparisons were made between species.
+These models were subject to stepwise term deletion routes using likelihood ratio tests (LRT) to assess term significance and identify the minimal adequate model (terms were retained when $P < 0.05$).
 

### 2016-01-20 article/article.Rmd
progress on manuscript

--- 2016-01-18 article/article.Rmd
@@
 ## Introduction
+Biodiversity is under threat.
+Many species have gone extinct over the last few decades [@r270; @r272].
+Currently there are thousands of species listed as endangered or near threatened in the International Union for Conservation of Nature (IUCN) red list (\url{www.iucnredlist.org}).
+Futhermore, climate change is expected cause even greater loss to global biodiversity [@r236; @r243].
+In response to this crisis, the overarching aim of conservation is to maximise the long-term persistance of biodiversity [@r255; @r67].
+
+The act of setting aside areas for preservation is the hallmark of modern conservation [@r440].
+Protected areas buffer species from anthropogenic impacts and provide areas for biodiversity processes to persist (eg.
+nutrient cycling and water purification).
+Additionally, protected areas set the stage for direct management actions (eg.
+habitat rehabitilation, translocation, and invasive species management).
+Since the resources avaliable for conservation are limited, the plans for protected areas (prioritizations) are often generated using spatially explicit biological data (eg.
+species distribution data) and decision support tools [eg.
+$\texttt{MARXAN}$; @r22; $\texttt{Zonation}$; @r429] that identify cost-effective solutions.
+
+Protected areas must preserve intra-specific genetic variation to ensure their persistance [@r67; @r3; @r61].
+This variation can be classified as either adaptive or neutral [reviewed in @r????].
+Adaptive genetic variation is associated with loci currently under selection, whereas neutral genetic variation is associated with loci that are not.
+By preserving existing patterns of adaptive variation, protected areas can ensure that populations with particularly beneficial adaptations are not lost [@r3].
+In constrast, neutral variation reflects the evolutionary history of different populations.
+By preserving existing patterns of neutral variation, protected areas can secure different genetic lineages and avoid the adverse effects of low genetic diversity [eg.
+inbreeding depression; @r67].
+However, genetic data requires considerable resources to obtain and analyse, and as a consequence, conservation planners do not typically use genetic data to inform prioritizations [@r61].
+Thus prioritisations may fail to capture species' intra-specific genetic variation, and ultimately fail to deliver desired outcomes.
+
+Recently, conservation scientists have begun investigating the use of surrogates for genetic data to generate more effective prioritisations without the overheads associated with genetic data.
+Since adaptive genetic variation is driven by selection pressures, environmental variables have been explored as potential surrogates for genetic variation [eg.
+@r247; @r2].
+On the other hand, neutral variation arises from a reduction in gene flow, and so variables that predict reduction in gene flow have been explored.
+For instance, the isolation by distance theory [IBD; @r259] predicts that populations further apart will have lower gene flow, and in turn have greater neutral variation among them.
+The use of distance-based surrogates has been found to improve prioritisations in an island network [@r200].
+However the efftiveness of this surrogate remains untested in systems where connectivity is complicated by additional factors (eg.
+elevation in terrestrial systems).
+But before conservation planners can reliably use surrogates for genetic data in conservation planning, the effectiveness of these surrogates needs to be verified.
+
+Our overarching aim is to investigate the effectiveness of environmental and geographic surrogates for conservation planning.
+First, we aim to test the effectiveness of these surrogates under the simplest of conditions by using a set of single-species prioritisations.
+Second, we aim to test the effectiveness of these surrogates in a more realistic multi-species context, and see how they compare to the single-species prioritisations.
+Third, we aim to understand the relationship between the proportion of surrogate variation secured in a multi-species prioritisation and the prorpotion of genetic variation secured in a prioritization.
+If these surrogates markedly improve the proportion of genetic variation secured in a prioritization, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.
 
@@
 ### Study area
-To address the aims of the study, we utilized species distribution and genomic data collected by the IntraBioDiv project [Figure S1; @r451; see @r452 for methodological details].
+To address the aims of the study, we utilized species distribution and genomic data collected by the IntraBioDiv project [@r451; see @r452 for methodological details].
 This dataset was well suited for this study, because it provides spatially explicit genomic data for a multitude of species across a broad geographic area and over large environmental gradients.
-As part of this project, data for 27 alpine plant species were collected across the European Alps according to a grid (22.3 km $\times$ 25 km).
+As part of this project, data for 27 alpine plant species were collected across the European Alps according to a grid (22.3 km $\times$ 25 km; Figure S1).
 Briefly, in every second cell, plant samples were collected from three individuals for each species present inside the cell.
@@
 We explored the effectiveness of climatic and geographic surrogates.
-To describe variation in the location of each grid cell, we projected the grid into an equi-distant coordinate system (ESRI:102031) and used the centroid of each cell.
+To describe variation in the location of each grid cell, we projected the grid into an equi-distant coordinate system ($\texttt{ESRI:102031}$) and used the centroid of each cell.
 To describe the climatic variation among each cell, we obtained twenty-one bioclimatic variables [approx.
-1km resolution; @r33].
-These layers and the grid cells were projected into an equal-area coordinate system (ESRI:102014).
+1 km resolution; @r33].
+These layers and the grid cells were projected into an equal-area coordinate system ($\texttt{ESRI:102014}$).
 The bioclimatic variables were clipped to the extent of the grid and subject to a principle components analysis (PCA; Table S1).
-The first two principle components (PCs) cumulatively explained `r round(pca.DF[[4]][2],1)`% of the variation and were used to generate two new layers.
+The first two principle components (PCs) cumulatively explained `r round(pca.DF[[4]][2],1)` % of the variation and were used to generate two new layers.
 We calculated the average value for each PC layer in each grid cell, and used these values to characterize the main themes of climatic variation among the grid cells (Figure S2).
@@
 The purpose of this step was to avoid confounding variation between populations attributed to genetic history instead of selection pressures.
-We used discriminant analyses of principle components [DAPC; @r467] to assign individuals to populations [implemented in \texttt{adegenet} R package; @r468].
-Briefly, for each species, we transformed the binary loci data using a principle components analysis (PCA) and obtained the principle components required to explain at least `r adegenet.params.LST[[MODE]]$perc.pca`% variation.
-We subject these components to multiple $k$-means clustering models with varying numbers of assumed populations (1--`r adegenet.params.LST[[MODE]]$max.n.clust`; `r adegenet.params.LST[[MODE]]$n.inter` iterations; `r adegenet.params.LST[[MODE]]$n.start` random starts).
-We then  applied Ward's clustering method to the Bayesian Information Criterion (BIC) statistics associated with each model to partition the models into two groups.
-We selected the model before the first group switch, and used the population memberships in that model for subsequent analysis.
+We used discriminant analyses of principle components [DAPC; @r467] to assign individuals to populations [implemented in $\texttt{adegenet}$ R package; @r468].
+Briefly, for each species, we transformed the binary loci data using a principle components analysis (PCA) and obtained the principle components required to explain at least `r adegenet.params.LST[[MODE]]$perc.pca` % variation.
+We subject these components to multiple $k$-means clustering models with varying numbers of assumed populations (\textit{k} = 1--`r adegenet.params.LST[[MODE]]$max.n.clust`; `r adegenet.params.LST[[MODE]]$n.iter` iterations; `r adegenet.params.LST[[MODE]]$n.start` random starts).
+We then  applied Ward's clustering method to the Bayesian Information Criterion (BIC) statistics associated with each model to partition the models into two groups, and selected the model before the first group switch (as recomended in the \texttt{adegenet} user manual).
+The populations members associated with these models were used for subsequent analysis.
 
@@
 For each species, for each type of genetic variation present, we used Gower distances [@r464] to express the genomic differences between individuals [using the $\texttt{cluster}$ R package to accommodate sparsity; @r457].
-These distances were subject to multiple non-metric multi-dimensional scaling analyses using the minimum number of dimensions required to acheive an adequate stress value  [`r mds.params.LST[[MODE]]$max.stress`; `r mds.params.LST[[MODE]]$trymax` random starts; Table S3; implemented in the \texttt{vegan} R package; @r458].
+These distances were subject to multiple non-metric multi-dimensional scaling analyses using the minimum number of dimensions required to acheive an adequate stress value  [$\leq$ `r mds.params.LST[[MODE]]$max.stress`; `r mds.params.LST[[MODE]]$trymax` random starts; Table S3; implemented in the $\texttt{vegan}$ R package; @r458].
 Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the samples in each grid cell to characterize the genomic variation among the grid cells.
@@
 We used the unreliable representative and adequate prioritization (URAP) formulation of the reserve selection problem [@r465].
-This formulation can identify the minimum number of planning units required to preserve a proportion of the species' range as well as a proportion of the species' variation [using an attribute space; _sensu_ @r465].
+This formulation can identify the minimum number of planning units required to preserve a proportion of the species' range as well as a proportion of the species' variation using an attribute space [_sensu_ @r465].
 We treated the genomic and surrogate variables as distinct attribute spaces.
@@
 First, we generated a set of single species prioritizations.
-For each species, we generated a prioritization using a `r (rapr.params.LST[[MODE]]$amount.target*100)`% amount-based target, a prioritization using a `r (rapr.params.LST[[MODE]]$amount.target*100)`% amount-based target and `r (rapr.params.LST[[MODE]]$surrogate.target*100)`% space-based targets for the surrogate variation, and a prioritization using a `r (rapr.params.LST[[MODE]]$amount.target*100)`% amount-based target and `r (rapr.params.LST[[MODE]]$genetic.target*100) `% space-based targets for the genomic variation.
+For each species, we generated `r rapr.params.LST[[MODE]]$single.species.amount.replicates` prioritizations using a `r (rapr.params.LST[[MODE]]$amount.target*100)` % amount-based target, a single prioritization using a `r (rapr.params.LST[[MODE]]$amount.target*100)` % amount-based target and `r (rapr.params.LST[[MODE]]$surrogate.target*100)` % space-based targets for the surrogate variation, and single a prioritization using a `r (rapr.params.LST[[MODE]]$amount.target*100)` % amount-based target and `r (rapr.params.LST[[MODE]]$genetic.target*100) ` % space-based targets for the genomic variation.
+We generated multiple solutions for the single-species prioritizations containing only amount-based targets because they were effectively a random selection of the planning units that the species inhabits, and so we obtained multiple solutions to account for stochasticity.
 Next, we generated three multi-species prioritizations using the same sets of targets used in the single species prioritizations.
-Finally, to understand the relationship between the surrogate-based and genomic-based targets, we generated a suite of prioritizations using increasing space-based targets between `r (min(rapr.params.LST[[MODE]]$pareto.surrogate.targets)*100)`% and `r (max(rapr.params.LST[[MODE]]$pareto.surrogate.targets)*100)`% in increments of `r diff(rapr.params.LST[[MODE]]$pareto.surrogate.targets[1:2])*100`%.
+Finally, to understand the relationship between the surrogate-based and genomic-based targets, we generated a suite of prioritizations using increasing space-based targets between `r (min(rapr.params.LST[[MODE]]$pareto.surrogate.targets)*100)` % and `r (max(rapr.params.LST[[MODE]]$pareto.surrogate.targets)*100)` % in increments of `r diff(rapr.params.LST[[MODE]]$pareto.surrogate.targets[1:2])*100` %.
 All of these prioritisations had no amount-based targets.
 For each prioritization we calculated the proportion of the adaptive (if present) and neutral genetic variation that it secured.
-We solved the reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)`% of optimality using the $\texttt{rapr}$ R package [@r465] and $\texttt{Gurobi}$ [@r463].
+We solved the reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality using the $\texttt{rapr}$ R package [@r465] and $\texttt{Gurobi}$ [@r463].
 
 ## Statistical analysis
-We tested for differences in the effectiveness of prioritisations at securing genomic variation using general linear models (GLMs).
-These models were fit to the proportion of genomic variation explained using a logit link.
-They contained a predictor variable indicating if the prioritisation was generated using amount, surrogate, or genetic targets, a second predictor variable indicating if the genomic variation was adaptive or neutral, and the interaction between these variables.
-These models were subject to stepwise term deletion routes using likelihood ratio tests (LRT) to identify the minimal adequate model (terms were retained when $P < 0.05$).
+We tested for differences in the effectiveness of prioritizations at securing genetic variation using generalized linear mixed-effects models [GLMMs; using the $\texttt{lme4}$ R package; @r469].
+We fit two full models to the proportion of genomic variation secured in the single- and multi-species prioritizations using a logit link.
+Both models contained a predictor variable indicating if the prioritisation was generated using amount, surrogate, or genetic targets, a second predictor variable indicating if the genomic variation was adaptive or neutral, and the interaction between these variables.
+Additionally, both models contained an intercept-only random effect to ensure that comparisons were made within species.
+Each full model was subject to a stepwise term deletion routine using likelihood ratio tests (LRT) to assess term significance (terms were retained when $P < 0.05$).
+Finally, we conducted a post-hoc analysis using Tukey constrasts with a Bonferroni correction [using the $\texttt{multcomp}$ R package; @r470].
 

### 2016-01-20 article/figures.Rmd
progress on manuscript

@@
   rmarkdown::pdf_document:
-    toc: false
-    keep_tex: no
     fig_caption: yes
     includes:
-      in_header: preamble-latex.tex
+      in_header: [preamble-latex.tex, preamble-latex2.tex]
 ---
@@
 
-```{r, echo=FALSE, fig.width=6, fig.height=3, fig.cap='Summary of single species prioritisations.
-Single species prioritisations were generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species.
-Data shows the performance of prioritisations generated using these three sets of targets.
-Bars denote means and standard errors.'}
+```{r, echo=FALSE, fig.width=6, fig.height=3, fig.cap='Summary of single species prioritizations.
+Single species prioritizations were generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species.
+Violin plots show the performance of prioritizations generated using these three sets of targets.
+Points denote the values associated with each prioritization.
+Letters denote significant differences ($P < 0.05$).'}
 # prepare data for plotting
@@
 	Prioritisation=unique(single.spp.SDF$Prioritisation),
-	Metric=unique(single.spp.SDF$Metric))
-single.spp.PDF <- cbind(single.spp.PDF,
-	as.data.frame(predict(single.spp.GLM, single.spp.PDF,
-		type='response', se.fit=TRUE))) %>%
-	mutate(lower=fit-se.fit, upper=fit+se.fit,
+	Metric=unique(single.spp.SDF$Metric)
+) %>% mutate(
 	letters=toupper(cld(single.spp.MCP)$mcletters$Letters),
-		letter_pos=upper+0.05)
+	letter.height=max(single.spp.SDF$value)+(diff(range(single.spp.SDF$value))*0.1)
+)
 # make plot
-ggplot(aes(x=Metric,y=fit,fill=Prioritisation),
-	data=single.spp.PDF) +
-	geom_bar(position=position_dodge(0.9),
-		stat='identity') +
-	geom_errorbar(
-		aes(ymin=lower,ymax=upper),
-		position=position_dodge(0.9), width=0.6) +
-	geom_text(aes(x=Metric, y=letter_pos,
-		label=letters), position=position_dodge(0.9)) +
+ggplot(aes(x=Metric,y=value,fill=Prioritisation),
+	data=single.spp.SDF) +
+	geom_violin(position=position_dodge(0.9)) +
+	geom_point(alpha=0.2,position=position_jitterdodge(dodge.width=0.9, jitter.width=0.25)) +
+	geom_text(aes(x=Metric, y=letter.height, label=letters),
+		data=single.spp.PDF, position=position_dodge(0.9)) +
 	scale_fill_manual(name='Prioritisation',
 		values=c('grey80','grey50','grey20')) +
-	ylab('Proportion genetic\nvariation secured (%)') +
+	ylab('Genetic variation secured (%)') +
 	xlab('') +
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=3.5, fig.cap='Multi-species prioritisations.
-Panel (a) shows the prioritisation generated for using just amount-based targets.
-Panel (b) shows the prioritisation generated using amount-based and surrogate based targets.
-Panel (c) shows the prioritisation generated using amount-based and genetic-based targets'}
+```{r, echo=FALSE, fig.width=8, fig.height=3, fig.cap='Multi-species prioritizations.
+Panel (a) shows the prioritization generated for using just amount-based targets.
+Panel (b) shows the prioritization generated using amount-based and surrogate based targets.
+Panel (c) shows the prioritization generated using amount-based and genetic-based targets.'}
 # download basemap
@@
 
-```{r, echo=FALSE, fig.width=6, fig.height=3, fig.cap='Summary of multi-species prioritisations.
-Three prioritisations were generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species.
-Data shows the performance of these prioritisations based on how much genetic variation they explain.
-Bars denote means and standard errors.'}
+```{r, echo=FALSE, fig.width=6, fig.height=3, fig.cap='Summary of multi-species prioritizations.
+Three prioritizations were generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species.
+Data shows the performance of these prioritizations based on how much genetic variation they explain.
+See Figure 1 caption for conventions.'}
 # prepare data for plotting
@@
 	Prioritisation=unique(multi.spp.SDF$Prioritisation),
-	Metric=unique(multi.spp.SDF$Metric))
-multi.spp.PDF <- cbind(multi.spp.PDF,
-	as.data.frame(predict(multi.spp.GLM, multi.spp.PDF,
-		type='response', se.fit=TRUE))) %>%
-	mutate(lower=fit-se.fit, upper=fit+se.fit,
+	Metric=unique(multi.spp.SDF$Metric)
+) %>% mutate(
 	letters=toupper(cld(multi.spp.MCP)$mcletters$Letters),
-		letter_pos=upper+0.05)
+	letter.height=max(multi.spp.SDF$value)+(diff(range(multi.spp.SDF$value))*0.1)
+)
 # make plot
-ggplot(aes(x=Metric,y=fit,fill=Prioritisation),
-	data=multi.spp.PDF) +
-	geom_bar(position=position_dodge(0.9),
-		stat='identity') +
-	geom_errorbar(
-		aes(ymin=lower,ymax=upper),
-		position=position_dodge(0.9), width=0.6) +
-	geom_text(aes(x=Metric, y=letter_pos,
-		label=letters), position=position_dodge(0.9)) +
+ggplot(aes(x=Metric,y=value,fill=Prioritisation),
+	data=multi.spp.SDF) +
+	geom_violin(position=position_dodge(0.9)) +
+	geom_point(alpha=0.2,position=position_jitterdodge(dodge.width=0.9, jitter.width=0.25)) +
+	geom_text(aes(x=Metric, y=letter.height,label=letters),
+		data=multi.spp.PDF, position=position_dodge(0.9)) +
 	scale_fill_manual(name='Prioritisation',
 		values=c('grey80','grey50','grey20')) +
-	ylab('Proportion genetic\nvariation secured (%)') +
+	ylab('Genetic variation secured (%)') +
 	xlab('') +
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=3.5, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations.'}
+```{r, echo=FALSE, fig.width=8, fig.height=3, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations.
+Each panel shows the relationship for a different type of genetic variation.
+Each line denotes a different species.
+Lines are transparent and so darker regions show areas where species have similar relationships.'}
 # make plots
@@
 	geom_line(aes(x=Surrogate.target,y=adaptive.held,group=Species),
-		alpha=0.5) +
+		alpha=0.3) +
 	xlab('Environmental variation secured (%)') +
@@
 	geom_line(aes(x=Surrogate.target,y=neutral.held,group=Species),
-		alpha=0.5) +
+		alpha=0.3) +
 	xlab('Geographic variation secured (%)') +

### 2016-01-20 article/supporting_information.Rmd
progress on manuscript

@@
 ## Supporting Information
+All data and code is stored in an online repository (\url{www.github.com/paleo13/genetic.surrogates}) to replicate analyses in this study. See the \textit{README} file for instructions.
 
@@
 
-```{r, fig.width=6.5, fig.height=3.25, fig.cap='Climatic variation.
+```{r, fig.width=8, fig.height=3.5, fig.cap='Climatic variation.
 Each panel depicts variation based on a different principle component (PC).
@@
 See Figure S6 caption for conventions.')}
-if (4 <= general.params.LST[[MODE]]$n.spp) if (4 <= ) plot.spp.mds(4)
+if (4 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(4)
 ```

### 2016-01-20 article/article.Rmd
fixes to ms + removed unsampled cells from analysis

@@
 Briefly, in every second cell, plant samples were collected from three individuals for each species present inside the cell.
+Cells that did not contain any plant samples were omitted from analysis.
 Samples were genotyped using amplified fragment length polymorphisms [AFLP; @r453].
@@
 To describe variation in the location of each grid cell, we projected the grid into an equi-distant coordinate system ($\texttt{ESRI:102031}$) and used the centroid of each cell.
-To describe the climatic variation among each cell, we obtained twenty-one bioclimatic variables [approx.
+To describe the climatic variation among each cell, we obtained 21 bioclimatic variables [approx.
 1 km resolution; @r33].
@@
 First, we grouped individuals into genetic lineages for each species (Figure S2).
-The purpose of this step was to avoid confounding variation between populations attributed to genetic history instead of selection pressures.
+The purpose of this step was to avoid misclassifying loci as adaptive.
 We used discriminant analyses of principle components [DAPC; @r467] to assign individuals to populations [implemented in $\texttt{adegenet}$ R package; @r468].
 Briefly, for each species, we transformed the binary loci data using a principle components analysis (PCA) and obtained the principle components required to explain at least `r adegenet.params.LST[[MODE]]$perc.pca` % variation.
-We subject these components to multiple $k$-means clustering models with varying numbers of assumed populations (\textit{k} = 1--`r adegenet.params.LST[[MODE]]$max.n.clust`; `r adegenet.params.LST[[MODE]]$n.iter` iterations; `r adegenet.params.LST[[MODE]]$n.start` random starts).
+We subject these components to multiple \textit{k}-means clustering models with varying numbers of assumed populations (\textit{k} = 1--`r adegenet.params.LST[[MODE]]$max.n.clust`; `r adegenet.params.LST[[MODE]]$n.iter` iterations; `r adegenet.params.LST[[MODE]]$n.start` random starts).
 We then  applied Ward's clustering method to the Bayesian Information Criterion (BIC) statistics associated with each model to partition the models into two groups, and selected the model before the first group switch (as recomended in the \texttt{adegenet} user manual).
@@
 For each species, $\texttt{Bayescan}$ was run using the population memberships identified in the previous analysis (1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations).
-Following guidelines in the $\texttt{Bayescan}$ manual, we omitted loci where the global frequency of the minor allele was $\geq 0.05$ to avoid false-positives.
+We omitted loci where the global frequency of the minor allele was $\geq 0.05$ to avoid false-positives (following guidelines in the $\texttt{Bayescan}$ manual).
 To ensure convergence, we ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate [FDR; $\leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].
 
-After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if present) and neutral genetic variation for each species (Figures S4--S29).
-For each species, for each type of genetic variation present, we used Gower distances [@r464] to express the genomic differences between individuals [using the $\texttt{cluster}$ R package to accommodate sparsity; @r457].
+After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species (Figures S4--S29).
+For each species, for each type of genetic variation detected, we used Gower distances [@r464] to express the genomic differences between individuals [using the $\texttt{cluster}$ R package to accommodate sparsity; @r457].
 These distances were subject to multiple non-metric multi-dimensional scaling analyses using the minimum number of dimensions required to acheive an adequate stress value  [$\leq$ `r mds.params.LST[[MODE]]$max.stress`; `r mds.params.LST[[MODE]]$trymax` random starts; Table S3; implemented in the $\texttt{vegan}$ R package; @r458].

### 2016-01-20 article/figures.Rmd
fixes to ms + removed unsampled cells from analysis

@@
 		'Germany', 'Slovenia', 'Croatia', 'Hungary',
-		'Monaco', 'Germany'
+		'Monaco', 'Germany', 'Slovakia', 'Czech Republic'
 	)

### 2016-01-20 article/supporting_information.Rmd
fixes to ms + removed unsampled cells from analysis

@@
     includes:
-      in_header: [preamble-latex.tex, preamble-latex2.tex]
+      in_header: [preamble-latex.tex, preamble-latex3.tex]
 ---

### 2016-01-20 R/analysis/02-surrogate-data.R

@@ -43,10 +43,23 @@ extract.DF <- grid.PPLY %>% rasterize(bioclim.STK, field='id') %>%
 # merge with grid.DF
 grid.DF <- cbind(grid.DF, extract.DF)
 
-## update spatial objects
+## update @data slot in spatial objects
 grid.PLY@data <- grid.DF
 grid.PPLY@data <- grid.DF
- 
+
+## remove grids without any species occurences
+# idenitfy pus with occurences
+valid.rows <- which(apply(grid.DF[,unique(spp.samples.DF$species),drop=FALSE], 1, sum)>0)
+# subset objects
+grid.DF <- grid.DF[valid.rows,,drop=FALSE]
+grid.PLY <- grid.PLY[valid.rows,]
+grid.PPLY <- grid.PPLY[valid.rows,]
+centroids.DF <- centroids.DF[valid.rows,,drop=FALSE]
+# reset ids
+grid.DF$id <- seq_len(nrow(grid.DF))
+grid.PLY$id <- seq_len(nrow(grid.DF))
+grid.PPLY$id <- seq_len(nrow(grid.DF))
+
 ## load and save pca summary
 pca.DF <- read.table('data/BioClim_variables/pca.TXT', skip=80) %>% `names<-`(
 	c('Principle Component', 'Eigen Value', 'Variation explained (%)',

### 2016-01-21 article/article.Rmd
* reduce disk usage of analysis * progress writing manuscript

@@
 ---
-title: "Are environmental and geographic effective surrogates for genetic variation in conservation planning?"
+title: "Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?"
 author: |
@@
 
-Our overarching aim is to investigate the effectiveness of environmental and geographic surrogates for conservation planning.
+Our overarching aim is to investigate the effectiveness of environmental and geographic surrogates for adaptive neutral genetic variation in the context of conservation planning.
 First, we aim to test the effectiveness of these surrogates under the simplest of conditions by using a set of single-species prioritisations.
@@
 Third, we aim to understand the relationship between the proportion of surrogate variation secured in a multi-species prioritisation and the prorpotion of genetic variation secured in a prioritization.
-If these surrogates markedly improve the proportion of genetic variation secured in a prioritization, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.
+If these surrogates can be shown to markedly improve the proportion of genetic variation secured in a prioritization, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.
 
@@
 We set the demand points for each species and its attribute space using the values associated with the planning units.
+Although Hanson _et al._ [-@r465] recomend against using the same coordinates for demand points and planning units for real-word conservation planning exercises, we did this to provide a more robust test and reduce uncertainty in the analysis.
+For every prioritization we generated, we calculated the proportion of the adaptive (if detected) and neutral genetic variation that every prioritization secured to assess its performance.
+We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality using the $\texttt{rapr}$ R package [@r465] and $\texttt{Gurobi}$ [@r463].
 
-To fulfil the aims of the study, we generated three sets of prioritisations.
-First, we generated a set of single species prioritizations.
-For each species, we generated `r rapr.params.LST[[MODE]]$single.species.amount.replicates` prioritizations using a `r (rapr.params.LST[[MODE]]$amount.target*100)` % amount-based target, a single prioritization using a `r (rapr.params.LST[[MODE]]$amount.target*100)` % amount-based target and `r (rapr.params.LST[[MODE]]$surrogate.target*100)` % space-based targets for the surrogate variation, and single a prioritization using a `r (rapr.params.LST[[MODE]]$amount.target*100)` % amount-based target and `r (rapr.params.LST[[MODE]]$genetic.target*100) ` % space-based targets for the genomic variation.
-We generated multiple solutions for the single-species prioritizations containing only amount-based targets because they were effectively a random selection of the planning units that the species inhabits, and so we obtained multiple solutions to account for stochasticity.
-Next, we generated three multi-species prioritizations using the same sets of targets used in the single species prioritizations.
-Finally, to understand the relationship between the surrogate-based and genomic-based targets, we generated a suite of prioritizations using increasing space-based targets between `r (min(rapr.params.LST[[MODE]]$pareto.surrogate.targets)*100)` % and `r (max(rapr.params.LST[[MODE]]$pareto.surrogate.targets)*100)` % in increments of `r diff(rapr.params.LST[[MODE]]$pareto.surrogate.targets[1:2])*100` %.
-All of these prioritisations had no amount-based targets.
-For each prioritization we calculated the proportion of the adaptive (if present) and neutral genetic variation that it secured.
-We solved the reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality using the $\texttt{rapr}$ R package [@r465] and $\texttt{Gurobi}$ [@r463].
+To understand how surrogate-based targets might perform under typical conservation planning scenarios, we generated a set of single-species and multi-species prioritizations.
+All of these prioritizations used `r (rapr.params.LST[[MODE]]$amount.target*100)` % amount-based targets to ensure an adequate proportion of habitat was secured for each species [@r197; @r255; @r471].
+We generated a set of single-species prioritizations for each species: `r rapr.params.LST[[MODE]]$single.species.amount.replicates` prioritizations using only the amount-based targets, a single prioritization using the amount-based target and `r (rapr.params.LST[[MODE]]$surrogate.target*100)` % surrogate-based targets, and a single prioritization using the amount-based target and `r (rapr.params.LST[[MODE]]$genetic.target*100)` % genetic-based targets.
+We also generated three multi-species prioritizations.
+These prioritizations used the three unique combinations of targets used in the single-species prioritizations.
+Note that we obtained multiple solutions for the single-species prioritisations that only used an amount-based target because, unlike the prioritizations, these targets yielded many optimal solutions with varying levels effectiveness.
+
+To understand the relationship between the surrogate-based targets and the proportion of genetic variation secured, we generated a set of multi-species prioritizations with increasing surrogate-based targets between `r (min(rapr.params.LST[[MODE]]$pareto.surrogate.targets)*100)` % and `r (max(rapr.params.LST[[MODE]]$pareto.surrogate.targets)*100)` % in increments of `r diff(rapr.params.LST[[MODE]]$pareto.surrogate.targets[1:2])*100` %.
+None of these priortizations used amount-based targets.
 
 ## Statistical analysis
-We tested for differences in the effectiveness of prioritizations at securing genetic variation using generalized linear mixed-effects models [GLMMs; using the $\texttt{lme4}$ R package; @r469].
+We tested for differences in the effectiveness of prioritizations based on typical scenarios using generalized linear mixed-effects models [GLMMs; using the $\texttt{lme4}$ R package; @r469].
 We fit two full models to the proportion of genomic variation secured in the single- and multi-species prioritizations using a logit link.
@@
 Finally, we conducted a post-hoc analysis using Tukey constrasts with a Bonferroni correction [using the $\texttt{multcomp}$ R package; @r470].
+
+We investigated the relationship between surrogate-based targets and the genetic variation secured in a prioritization using nonlinear mixed-effects models [NLMM; using the $\texttt{lme4}$ R package].
+Two models were fit to the adaptive and neutral genetic variation secured using a negative exponential formulation to estimate the rate of the accumulation of genetic variation ($y = 0 + (0 - 1) \times exp(r \times x)$; $r$ is the accumulate rate).
+Both models contained intercept and slope random effects to accomodate variation between species.
 

### 2016-01-21 R/analysis/06-single-species-prioritisations.R

@@ -97,5 +97,22 @@ single.spp.DF <- ldply(
 	}
 )
 
+# remove all but the first amount-based prioritisation for each species to reduce disk usage
+single.spp.prioritisations <- llply(
+	seq_along(unique(spp.samples.DF$species)),
+	function(i) {
+		list(
+			single.spp.amount.prioritisations[[i]][[1]],
+			single.spp.surrogate.prioritisations[[i]],
+			single.spp.genetic.prioritisations[[i]]
+		)
+	}
+)
+
+# remove objects to clear disk usage
+rm(single.spp.amount.prioritisations)
+rm(single.spp.surrogate.prioritisations)
+rm(single.spp.genetic.prioritisations)
+
 ## save .rda
 save.session('results/.cache/06-single-species-prioritisations.rda')

### 2016-01-22 article/article.Rmd
* progress to manuscript * change from pareto frontier to correct terminology of correlation * model trade-offs using nlmm

@@
 Biodiversity is under threat.
-Many species have gone extinct over the last few decades [@r270; @r272].
-Currently there are thousands of species listed as endangered or near threatened in the International Union for Conservation of Nature (IUCN) red list (\url{www.iucnredlist.org}).
-Futhermore, climate change is expected cause even greater loss to global biodiversity [@r236; @r243].
-In response to this crisis, the overarching aim of conservation is to maximise the long-term persistance of biodiversity [@r255; @r67].
+Over the last few decades, many species have gone extinct [@r270; @r272].
+Today, thousands of species are listed as endangered or near threatened in the International Union for Conservation of Nature (IUCN) red list (\url{www.iucnredlist.org}).
+In the future, species are expected to face even more threats [eg.
+climate change; @r236; @r243].
+In response to this crisis, the overarching aim of conservation is to maximize the long-term persistence of biodiversity [@r255; @r67].
 
 The act of setting aside areas for preservation is the hallmark of modern conservation [@r440].
-Protected areas buffer species from anthropogenic impacts and provide areas for biodiversity processes to persist (eg.
+Protected areas buffer species from anthropogenic impacts and provide places for biodiversity processes to persist (eg.
 nutrient cycling and water purification).
 Additionally, protected areas set the stage for direct management actions (eg.
-habitat rehabitilation, translocation, and invasive species management).
-Since the resources avaliable for conservation are limited, the plans for protected areas (prioritizations) are often generated using spatially explicit biological data (eg.
+habitat rehabilitation, translocation, and invasive species management).
+Since the resources available for conservation are limited, the plans for protected areas (prioritizations) are often generated using spatially explicit biological data (eg.
 species distribution data) and decision support tools [eg.
@@
 
-Protected areas must preserve intra-specific genetic variation to ensure their persistance [@r67; @r3; @r61].
+Protected areas must preserve intra-specific genetic variation to maximize species recovery [@r67; @r3; @r61].
 This variation can be classified as either adaptive or neutral [reviewed in @r????].
@@
 By preserving existing patterns of adaptive variation, protected areas can ensure that populations with particularly beneficial adaptations are not lost [@r3].
-In constrast, neutral variation reflects the evolutionary history of different populations.
+In contrast, neutral variation reflects the evolutionary history of different populations.
 By preserving existing patterns of neutral variation, protected areas can secure different genetic lineages and avoid the adverse effects of low genetic diversity [eg.
 inbreeding depression; @r67].
-However, genetic data requires considerable resources to obtain and analyse, and as a consequence, conservation planners do not typically use genetic data to inform prioritizations [@r61].
-Thus prioritisations may fail to capture species' intra-specific genetic variation, and ultimately fail to deliver desired outcomes.
+However, genetic data requires considerable investment to obtain and analyze, and as a consequence, conservation planners do not typically use genetic data to inform prioritizations [@r61].
+Thus many prioritizations may fail to capture species' intra-specific genetic variation, and ultimately fail to deliver conservation outcomes.
 
-Recently, conservation scientists have begun investigating the use of surrogates for genetic data to generate more effective prioritisations without the overheads associated with genetic data.
+Recently, conservation scientists have begun investigating the use of surrogates for genetic data to generate more effective prioritizations without the overheads associated with genetic data.
 Since adaptive genetic variation is driven by selection pressures, environmental variables have been explored as potential surrogates for genetic variation [eg.
 @r247; @r2].
-On the other hand, neutral variation arises from a reduction in gene flow, and so variables that predict reduction in gene flow have been explored.
+On the other hand, neutral variation arises from a reduction in gene flow, and so variables that predict reduction in gene flow between areas have been explored.
 For instance, the isolation by distance theory [IBD; @r259] predicts that populations further apart will have lower gene flow, and in turn have greater neutral variation among them.
-The use of distance-based surrogates has been found to improve prioritisations in an island network [@r200].
-However the efftiveness of this surrogate remains untested in systems where connectivity is complicated by additional factors (eg.
+The use of distance-based surrogates has been found to improve prioritizations in an island network [@r200].
+However the effectiveness of distance-based surrogates remains untested in systems where connectivity is complicated by additional factors (eg.
 elevation in terrestrial systems).
-But before conservation planners can reliably use surrogates for genetic data in conservation planning, the effectiveness of these surrogates needs to be verified.
+Before conservation planners can reliably use surrogates for genetic data in conservation planning, the effectiveness of these surrogates needs to be verified.
+Otherwise, conservation planners may waste precious resources on securing places that do little to benefit biodiversity.
 
 Our overarching aim is to investigate the effectiveness of environmental and geographic surrogates for adaptive neutral genetic variation in the context of conservation planning.
-First, we aim to test the effectiveness of these surrogates under the simplest of conditions by using a set of single-species prioritisations.
-Second, we aim to test the effectiveness of these surrogates in a more realistic multi-species context, and see how they compare to the single-species prioritisations.
-Third, we aim to understand the relationship between the proportion of surrogate variation secured in a multi-species prioritisation and the prorpotion of genetic variation secured in a prioritization.
+First, we aim to test the effectiveness of these surrogates under the simplest of conditions by using a set of single-species prioritizations.
+Second, we aim to test the effectiveness of these surrogates in a more realistic multi-species context, and see how they compare to the single-species prioritizations.
+Third, we aim to understand the relationship between the proportion of surrogate variation secured in a multi-species prioritization and the proportion of genetic variation secured in a prioritization.
 If these surrogates can be shown to markedly improve the proportion of genetic variation secured in a prioritization, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.
@@
 ## Methods
-### Study area
+### Study system
 To address the aims of the study, we utilized species distribution and genomic data collected by the IntraBioDiv project [@r451; see @r452 for methodological details].
@@
 Briefly, in every second cell, plant samples were collected from three individuals for each species present inside the cell.
-Cells that did not contain any plant samples were omitted from analysis.
+Cells that did not contain any plant samples were omitted from our analysis.
 Samples were genotyped using amplified fragment length polymorphisms [AFLP; @r453].
@@
 We used these grid cells as planning units to develop prioritizations to investigate the effectiveness of surrogates for the (AFLP) genetic variation.
+Most of our analyses described hereafter were performed using the \texttt{R} software package [version `r paste0(version$major, '.', version$minor)`; @r192].
 
@@
 To investigate the effectiveness of the surrogates for adaptive and neutral genetic variation, we first identified which of the sampled loci for each species were under selection.
-We used outlier locus detection methods [evaluated in @r454] to avoid circularity issues.
-The basic premise underpinning these analyses is that neutral loci are expected to have a specific level of variation among populations ($F_{st}$), and loci that deviate from this expectation should be under selection.
-Thus loci can be identified as being under selection without using environmental data.
+Specifically, we used outlier locus detection methods [evaluated in @r454] to avoid circularity issues.
+The basic premise underpinning these analyses is that neutral loci are expected to exhibit a specific level of variation among populations ($F_{st}$), and loci that deviate from this expectation should be under selection.
 In an earlier study, Bothwell _et al._ [-@r455] applied such outlier detection methods to one of the species in this dataset (_Getiana nivalis_).
@@
 
-First, we grouped individuals into genetic lineages for each species (Figure S2).
+First, we grouped individuals into genetic lineages for each species.
 The purpose of this step was to avoid misclassifying loci as adaptive.
@@
 We subject these components to multiple \textit{k}-means clustering models with varying numbers of assumed populations (\textit{k} = 1--`r adegenet.params.LST[[MODE]]$max.n.clust`; `r adegenet.params.LST[[MODE]]$n.iter` iterations; `r adegenet.params.LST[[MODE]]$n.start` random starts).
-We then  applied Ward's clustering method to the Bayesian Information Criterion (BIC) statistics associated with each model to partition the models into two groups, and selected the model before the first group switch (as recomended in the \texttt{adegenet} user manual).
-The populations members associated with these models were used for subsequent analysis.
+We then  applied Ward's clustering method to the Bayesian Information Criterion (BIC) statistics associated with each model to partition the models into two groups, and selected the model before the first group switch (as recommended in the \texttt{adegenet} manual; Figure S4).
+The populations memberships associated with these models were used for subsequent analysis (Figure S5).
 
-We used $\texttt{Bayescan}$ [version 2.1; @r456] to identify loci under selection for each species (Table S2).
-For each species, $\texttt{Bayescan}$ was run using the population memberships identified in the previous analysis (1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations).
+We identified loci under selection for each species (Table S2).
+For each species, we ran $\texttt{Bayescan}$ [version 2.1; @r456] using the population memberships identified in the previous analysis (1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations).
 We omitted loci where the global frequency of the minor allele was $\geq 0.05$ to avoid false-positives (following guidelines in the $\texttt{Bayescan}$ manual).
@@
 
-After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species (Figures S4--S29).
+After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species (Figures S6--S32).
 For each species, for each type of genetic variation detected, we used Gower distances [@r464] to express the genomic differences between individuals [using the $\texttt{cluster}$ R package to accommodate sparsity; @r457].
-These distances were subject to multiple non-metric multi-dimensional scaling analyses using the minimum number of dimensions required to acheive an adequate stress value  [$\leq$ `r mds.params.LST[[MODE]]$max.stress`; `r mds.params.LST[[MODE]]$trymax` random starts; Table S3; implemented in the $\texttt{vegan}$ R package; @r458].
-Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the samples in each grid cell to characterize the genomic variation among the grid cells.
+These distances were subject to multiple non-metric multi-dimensional scaling analyses [NMDS; implemented in the $\texttt{vegan}$ R package; @r458] using the minimum number of dimensions required to achieve an adequate stress value  ($\leq$ `r mds.params.LST[[MODE]]$max.stress`; `r mds.params.LST[[MODE]]$trymax` random starts for all models; Table S3).
+Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the samples in each grid cell to characterize their the genetic variation of individuals in the cells.
 
@@
 We set the demand points for each species and its attribute space using the values associated with the planning units.
-Although Hanson _et al._ [-@r465] recomend against using the same coordinates for demand points and planning units for real-word conservation planning exercises, we did this to provide a more robust test and reduce uncertainty in the analysis.
+Although Hanson _et al._ [-@r465] recommend against using the same coordinates for demand points and planning units for real-word conservation planning exercises, we did this to provide a more robust test and reduce uncertainty in the analysis.
 For every prioritization we generated, we calculated the proportion of the adaptive (if detected) and neutral genetic variation that every prioritization secured to assess its performance.
@@
 
-To understand how surrogate-based targets might perform under typical conservation planning scenarios, we generated a set of single-species and multi-species prioritizations.
+To understand how prioritizations generated using surrogate-based targets might perform under typical conservation planning scenarios, we generated a set of single-species and multi-species prioritizations.
 All of these prioritizations used `r (rapr.params.LST[[MODE]]$amount.target*100)` % amount-based targets to ensure an adequate proportion of habitat was secured for each species [@r197; @r255; @r471].
 We generated a set of single-species prioritizations for each species: `r rapr.params.LST[[MODE]]$single.species.amount.replicates` prioritizations using only the amount-based targets, a single prioritization using the amount-based target and `r (rapr.params.LST[[MODE]]$surrogate.target*100)` % surrogate-based targets, and a single prioritization using the amount-based target and `r (rapr.params.LST[[MODE]]$genetic.target*100)` % genetic-based targets.
-We also generated three multi-species prioritizations.
-These prioritizations used the three unique combinations of targets used in the single-species prioritizations.
-Note that we obtained multiple solutions for the single-species prioritisations that only used an amount-based target because, unlike the prioritizations, these targets yielded many optimal solutions with varying levels effectiveness.
+We also generated three multi-species prioritizations using the three unique combinations of targets used to generate the single-species prioritizations.
+Note that we obtained multiple solutions for the amount-based single-species prioritizations because, unlike the other prioritizations, these targets yielded many optimal solutions with varying levels effectiveness.
 
-To understand the relationship between the surrogate-based targets and the proportion of genetic variation secured, we generated a set of multi-species prioritizations with increasing surrogate-based targets between `r (min(rapr.params.LST[[MODE]]$pareto.surrogate.targets)*100)` % and `r (max(rapr.params.LST[[MODE]]$pareto.surrogate.targets)*100)` % in increments of `r diff(rapr.params.LST[[MODE]]$pareto.surrogate.targets[1:2])*100` %.
-None of these priortizations used amount-based targets.
+To understand the relationship between the surrogate-based targets and the proportion of genetic variation secured, we generated two sets of multi-species prioritizations with increasing surrogate-based targets between `r (min(rapr.params.LST[[MODE]]$correlation.surrogate.targets)*100)` % and `r (max(rapr.params.LST[[MODE]]$correlation.surrogate.targets)*100)` % in increments of `r diff(rapr.params.LST[[MODE]]$correlation.surrogate.targets[1:2])*100` %.
+The first set of prioritizations contained prioritizations with varying environmental surrogate-based targets with no specified geographic targets.
+The second contained prioritizations with varying geographic surrogate-based targets and no environmental targets.
+None of these prioritizations used amount-based targets.
 
 ## Statistical analysis
-We tested for differences in the effectiveness of prioritizations based on typical scenarios using generalized linear mixed-effects models [GLMMs; using the $\texttt{lme4}$ R package; @r469].
+We tested for differences in the effectiveness of prioritizations representing typical scenarios using generalized linear mixed-effects models [GLMMs; using the $\texttt{lme4}$ R package; @r469].
 We fit two full models to the proportion of genomic variation secured in the single- and multi-species prioritizations using a logit link.
-Both models contained a predictor variable indicating if the prioritisation was generated using amount, surrogate, or genetic targets, a second predictor variable indicating if the genomic variation was adaptive or neutral, and the interaction between these variables.
+Both models contained a predictor variable indicating if the prioritization was generated using amount, surrogate, or genetic targets, a second predictor variable indicating if the genetic variation was adaptive or neutral, and the interaction between these variables.
 Additionally, both models contained an intercept-only random effect to ensure that comparisons were made within species.
-Each full model was subject to a stepwise term deletion routine using likelihood ratio tests (LRT) to assess term significance (terms were retained when $P < 0.05$).
+Each full model was subject to a step-wise term deletion routine using likelihood ratio tests (LRT) to assess term significance (terms were retained when $P < 0.05$).
 Finally, we conducted a post-hoc analysis using Tukey constrasts with a Bonferroni correction [using the $\texttt{multcomp}$ R package; @r470].
 
-We investigated the relationship between surrogate-based targets and the genetic variation secured in a prioritization using nonlinear mixed-effects models [NLMM; using the $\texttt{lme4}$ R package].
-Two models were fit to the adaptive and neutral genetic variation secured using a negative exponential formulation to estimate the rate of the accumulation of genetic variation ($y = 0 + (0 - 1) \times exp(r \times x)$; $r$ is the accumulate rate).
-Both models contained intercept and slope random effects to accomodate variation between species.
+We investigated the relationship between surrogate-based targets and the genetic variation secured in a prioritization using nonlinear mixed-effects models (NLMMs; using the $\texttt{lme4}$ R package).
+Two models were fit to the adaptive and neutral genetic variation secured using a negative exponential formulation to estimate the rate of the accumulation of genetic variation ($y = 0 + (0 - 1) \times exp(-r \times x)$; $r$ is the accumulate rate).
+Both models contained intercept and slope random effects to accommodate variation between species.
 
 ## Results
-### Single species prioritizations
+<!-- Single species prioritizations -->
+Surrogates-targets improved the performance of single-species prioritizations (Figure 1).
+There were significant differences among the performance of prioritizations depending on the targets used to generate them and if they were assessed according to adaptive or neutral genetic variation (${\chi^{2}}_{`r single.spp.AOV1[['Df']][1]`,`r single.spp.AOV1[['Df']][2]`} = `r round(single.spp.AOV1[['Chisq']][2],2)`$, $P `r pretty.pval(single.spp.AOV1[['Pr(>Chisq)']][2])`$).
+Specifically, the amount-based prioritization captured much less adaptive and neutral genetic variation than either the surrogate-based ($Z = `r round(single.spp.MCP$test$tstat[['Surrogate.Adaptive variation - Amount.Adaptive variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Surrogate.Adaptive variation - Amount.Adaptive variation']])`$; $Z = `r round(single.spp.MCP$test$tstat[['Surrogate.Neutral variation - Amount.Neutral variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Surrogate.Neutral variation - Amount.Neutral variation']])`$) or genetic-based ($Z = `r round(single.spp.MCP$test$tstat[['Genetic.Adaptive variation - Amount.Adaptive variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Genetic.Adaptive variation - Amount.Adaptive variation']])`$; $Z = `r round(single.spp.MCP$test$tstat[['Genetic.Neutral variation - Amount.Neutral variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Genetic.Neutral variation - Amount.Neutral variation']])`$) prioritizations.
+No differences were detected in the proportion of adaptive ($Z = `r round(single.spp.MCP$test$tstat[['Surrogate.Adaptive variation - Genetic.Adaptive variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Surrogate.Adaptive variation - Genetic.Adaptive variation']])`$) or neutral genetic variation ($Z = `r round(single.spp.MCP$test$tstat[['Surrogate.Neutral variation - Genetic.Neutral variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Surrogate.Neutral variation - Genetic.Neutral variation']])`$) secured by the the surrogate-based and genetic-based prioritizations.
+However, these patterns were not reflected in the multi-species prioritizations.
 
-### Multi-species prioritisations
+<!-- Multi-species prioritizations -->
+Surrogate-targets were not found to improve the performance of multi-species prioritizations (Figure 2; Figure 3).
+No significant differences in the effectiveness of prioritizations were found depending on whether they were generated using different targets (${\chi^{2}}_{`r multi.spp.AOV3[['Df']][1]`,`r multi.spp.AOV3[['Df']][2]`} = `r round(multi.spp.AOV3[['Chisq']][2],2)`$; $P `r pretty.pval(multi.spp.AOV3[['Pr(>Chisq)']][2])`$), if they were assessed using adaptive or neutral genetic variation (${\chi^{2}}_{`r multi.spp.AOV2[['Df']][1]`,`r multi.spp.AOV2[['Df']][2]`} = `r round(multi.spp.AOV2[['Chisq']][2],2)`$; $P `r pretty.pval(multi.spp.AOV2[['Pr(>Chisq)']][2])`$), nor the interaction between these variables (${\chi^{2}}_{`r multi.spp.AOV1[['Df']][1]`,`r multi.spp.AOV1[['Df']][2]`} = `r round(multi.spp.AOV1[['Chisq']][2],2)`$; $P  `r pretty.pval(multi.spp.AOV1[['Pr(>Chisq)']][2])`$).
 
-### Pareo-frontier analysis
+<!-- Correlation analysis -->
+The surrogate variables showed a greater utility for adaptive genetic variation than neutral variation (Figure 4).
+The accumulation rate ($r$) was `r round(summary(env.correlation.NLMM)$coefficients[1],2)` $\pm$ `r round(summary(env.correlation.NLMM)$coefficients[2],2)` S.E.
+for the proportion of adaptive variation captured in a prioritization using environmental surrogates.
+On the other hand, the rate ($r$) was `r round(summary(geo.correlation.NLMM)$coefficients[1],2)` $\pm$ `r round(summary(geo.correlation.NLMM)$coefficients[2],2)` S.E.
+for neutral variation using geographic surrogates.
+Thus minor increases in the target for either surrogate results in major increases in the benefit of the prioritization.
+However, the coefficient was much larger for the adaptative genetic variation so high targets are not required to acheive substantial gains in adaptive genetic variaition.
 
 ## Discussion
+### Introduction
+* aim
+	+ we aimed to investigate the benefits of using surrogates for genetic variation in conservation planning.
+* innovation
+	+ first to test the effectiveness of environmental and geographic surrogates in a conservation planning context
+* key findings
+	+ surrogate variables are beneficial in single-species contexts.
+	+ but not in multi-species contexts when a comprehensive set of species is used.
+* consequences
+	+ highlights the benefit of using a comprehensive set of species
+	+ supports literature on using comprehensive set of species
+
+### Single + multi-species analysis
+
+### Correlation analysis
+
+### Limitations/Future directions
+* spatial/surrogate data
+	+ isolation by resistance: least-cost paths and circuit theory distances are becoming more population for modeling connectivity, and may offer better surrogates than linear geographic distances.
+However, the scaling for these variables requires an understanding of the often non-linear relationship, which is not available to conservation planners
+	+ cost data: here we used no cost data to provide a rigorous test for the benefits of surrogates in conservation planning.
+however in real conservation planning situations, prioritizations are largely driven by cost.
+future studies should look at including cost data to see if surrogates provide any benefit when including realistic cost data.
+
+* genetic data
+	+ AFLP data: next gen data would be better however the aim of a prioritization is to capture the main gradients of genetic variation--not all of the variation.
+	+ adegenet: whilst other software packages are more popular (eg.
+Structure), we used adegenet because information theory is more reliable than an ad-hoc second derivative process for comparing models.
+Additionally, it is computationally infeasible to analyze run 27 species in Structure in a suitable period of time.
+
+### Concluding remarks
+* take-home messages
+	+ surrogates are effective for single-species prioritizations in conservation planning
+	+ however, surrogates don't provide any noticeable benefits when using a comprehensive set of species in a multi-species prioritization
+	+ conservation planners are encouraged to use comprehensive sets of species
 

### 2016-01-22 article/figures.Rmd
* progress to manuscript * change from pareto frontier to correct terminology of correlation * model trade-offs using nlmm

@@
 ```{r, echo=FALSE, fig.width=8, fig.height=3, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations.
-Each panel shows the relationship for a different type of genetic variation.
-Each line denotes a different species.
-Lines are transparent and so darker regions show areas where species have similar relationships.'}
-# make plots
-p1 <- ggplot(data=env.pareto.DF) +
-	geom_line(aes(x=Surrogate.target,y=adaptive.held,group=Species),
-		alpha=0.3) +
+The left panel shows the relationship between environmental surrogates and adaptive genetic variation, and the right for geographic surrogates and neutral genetic variation.
+Transparent gray lines denote trneds for each species, and black lines show general trends across all species.'}
+### prepare data frames for plotting
+## environmental vs adaptive
+# random effects
+env.correlation.RPDF <- expand.grid(Surrogate.target=seq(min(env.correlation.DF$Surrogate.target), max(env.correlation.DF$Surrogate.target), length.out=100),
+	Species=unique(env.correlation.DF$Species))
+env.correlation.RPDF$r.ranef <- fixef(env.correlation.NLMM)[[1]] + ranef(env.correlation.NLMM)$Species$r[env.correlation.RPDF$Species]
+env.correlation.RPDF$fit <- 0 + (1 - 0) * (1 - exp(- env.correlation.RPDF$r.ranef * env.correlation.RPDF$Surrogate.target))
+# fixed effects
+env.correlation.FPDF <- data.frame(Surrogate.target=seq(min(env.correlation.DF$Surrogate.target), max(env.correlation.DF$Surrogate.target), length.out=100))
+env.correlation.FPDF$fit <- 0 + (1 - 0) * (1 - exp(-fixef(env.correlation.NLMM) * env.correlation.FPDF$Surrogate.target))
+
+## geographic vs neutral
+# random effects
+geo.correlation.RPDF <- expand.grid(Surrogate.target=seq(min(geo.correlation.DF$Surrogate.target), max(geo.correlation.DF$Surrogate.target), length.out=100),
+	Species=unique(geo.correlation.DF$Species))
+geo.correlation.RPDF$r.ranef <- fixef(geo.correlation.NLMM)[[1]] + ranef(geo.correlation.NLMM)$Species$r[geo.correlation.RPDF$Species]
+geo.correlation.RPDF$fit <- 0 + (1 - 0) * (1 - exp(- geo.correlation.RPDF$r.ranef * geo.correlation.RPDF$Surrogate.target))
+# fixed effects
+geo.correlation.FPDF <- data.frame(Surrogate.target=seq(min(geo.correlation.DF$Surrogate.target), max(geo.correlation.DF$Surrogate.target), length.out=100))
+geo.correlation.FPDF$fit <- 0 + (1 - 0) * (1 - exp(-fixef(geo.correlation.NLMM) * geo.correlation.FPDF$Surrogate.target))
+
+### make plots
+p1 <- ggplot() +
+	geom_point(aes(x=Surrogate.target, y=adaptive.held), data=env.correlation.DF) +
+	geom_line(aes(x=Surrogate.target,y=fit,group=Species), data=env.correlation.RPDF, alpha=0.3) +
+	geom_line(aes(x=Surrogate.target,y=fit), data=env.correlation.FPDF, size=1.5) +
 	xlab('Environmental variation secured (%)') +
@@
 	theme_classic()
-p2 <- ggplot(data=geo.pareto.DF) +
-	geom_line(aes(x=Surrogate.target,y=neutral.held,group=Species),
-		alpha=0.3) +
+p2 <- ggplot() +
+	geom_point(aes(x=Surrogate.target, y=neutral.held), data=geo.correlation.DF) +
+	geom_line(aes(x=Surrogate.target,y=fit,group=Species), data=geo.correlation.RPDF, alpha=0.3) +
+	geom_line(aes(x=Surrogate.target,y=fit), data=geo.correlation.FPDF, size=1.5) +
 	xlab('Geographic variation secured (%)') +

### 2016-01-22 article/supporting_information.Rmd
* progress to manuscript * change from pareto frontier to correct terminology of correlation * model trade-offs using nlmm

@@
 	col.names=c('Species', 'Primer', 'Probability', 'q-value', '$\\alpha$', '$F_{ST}$', 'Type'),
-	align=c('l', 'c', 'c', 'c', 'c', 'c', 'c')
+	align=c('l', 'c', 'c', 'c', 'c', 'c', 'c'),
+	caption='Loci classifications for each species.
+Each row denotes a sampled loci for a given species, and the evidence for it being under selection based on outputs from the program \texttt{Bayescan}.'
 )

### 2016-01-22 R/analysis/04-bayescan-analysis.R

@@ -3,7 +3,7 @@ session::restore.session('results/.cache/03-adegenet-analysis.rda')
 
 ## load parameters
 bayescan.params.LST <- parseTOML('parameters/bayescan.toml')
-mds.params.LST <- parseTOML('parameters/mds.toml')
+mds.params.LST <- parseTOML('parameters/nmds.toml')
 
 # subset out loci with polymorphisms that have really low or really high frequency
 spp.BayeScanData.sample.loci.subset.LST <- llply(

### 2016-01-22 R/analysis/08-pareto-frontier-analysis.R

@@ -1,45 +1,0 @@
-## load .rda
-session::restore.session('results/.cache/07-multi-species-prioritisations.rda')
-
-## pareto frontier analysis
-# generate prioritistions
-cat("starting environmental pareto frontier calculations\n")
-env.pareto.prioritisations <- llply(
-	rapr.params.LST[[MODE]]$pareto.surrogate.targets,
-	species.prioritisation,
-	x=ru,
-	amount.targets=0,
-	geo.surrogate.targets=0,
-	adaptive.genetic.targets=0,
-	neutral.genetic.targets=0,
-	.progress='text'
-)
-
-cat("starting geographic pareto frontier calculations\n")
-geo.pareto.prioritisations <- llply(
-	rapr.params.LST[[MODE]]$pareto.surrogate.targets,
-	species.prioritisation,
-	x=ru,
-	amount.targets=0,
-	env.surrogate.targets=0,
-	adaptive.genetic.targets=0,
-	neutral.genetic.targets=0,
-	.progress='text'
-)
-
-# extract results
-env.pareto.DF <- ldply(seq_along(env.pareto.prioritisations), function(i) {
-	mutate(
-		extractResults(env.pareto.prioritisations[[i]]),
-		Surrogate.target=rapr.params.LST[[MODE]]$pareto.surrogate.targets[i]
-	)
-})
-geo.pareto.DF <- ldply(seq_along(geo.pareto.prioritisations), function(i) {
-	mutate(
-		extractResults(geo.pareto.prioritisations[[i]]),
-		Surrogate.target=rapr.params.LST[[MODE]]$pareto.surrogate.targets[i]
-	)
-})
-
-## save .rda
-save.session('results/.cache/08-pareto-frontier-analysis.rda')

### 2016-01-22 R/analysis/09-statistical-analysis.R

@@ -1,5 +1,5 @@
 ## load .rda
-session::restore.session('results/.cache/08-pareto-frontier-analysis.rda')
+session::restore.session('results/.cache/08-correlation-analysis.rda')
 
 ### Single species prioritisations
 ## statistical analysis
@@ -12,7 +12,15 @@ single.spp.SDF <- single.spp.DF %>%
 	mutate(Prioritisation.Metric=interaction(Prioritisation,Metric)) 
 # model
 single.spp.GLMM <- suppressWarnings(glmer(value ~ Prioritisation * Metric + (1 | Species), family='binomial', data=single.spp.SDF, nAGQ=20))
+single.spp.GLMM1 <- suppressWarnings(glmer(value ~ Prioritisation + Metric + (1 | Species), family='binomial', data=single.spp.SDF, nAGQ=20))
+single.spp.GLMM2 <- suppressWarnings(glmer(value ~ Prioritisation + (1 | Species), family='binomial', data=single.spp.SDF, nAGQ=20))
+single.spp.GLMM3 <- suppressWarnings(glmer(value ~ 1 + (1 | Species), family='binomial', data=single.spp.SDF, nAGQ=20))
+
 single.spp.AOV <- suppressWarnings(anova(single.spp.GLMM, test="LRT"))
+single.spp.AOV1 <- suppressWarnings(anova(single.spp.GLMM, single.spp.GLMM1, test="LRT"))
+single.spp.AOV2 <- suppressWarnings(anova(single.spp.GLMM1, single.spp.GLMM2, test="LRT"))
+single.spp.AOV3 <- suppressWarnings(anova(single.spp.GLMM2, single.spp.GLMM3, test="LRT"))
+
 # post-hoc
 single.spp.GLMM2 <- suppressWarnings(glmer(value ~ Prioritisation.Metric + (1 | Species), family='binomial', data=single.spp.SDF, nAGQ=20))
 single.spp.MCP <- summary(
@@ -32,13 +40,27 @@ multi.spp.SDF <- multi.spp.DF %>%
 	mutate(Prioritisation.Metric=interaction(Prioritisation,Metric)) 
 # model
 multi.spp.GLMM <- suppressWarnings(glmer(value ~ Prioritisation * Metric + (1 | Species), family='binomial', data=multi.spp.SDF, nAGQ=20))
[…]

### 2016-01-23 article/figures.Rmd
+ fixed issues with order of bars in the violin plots + fixed issue with letters not showing in violin plots

@@
 Violin plots show the performance of prioritizations generated using these three sets of targets.
-Points denote the values associated with each prioritization.
 Letters denote significant differences ($P < 0.05$).'}
@@
 	letters=toupper(cld(single.spp.MCP)$mcletters$Letters),
-	letter.height=max(single.spp.SDF$value)+(diff(range(single.spp.SDF$value))*0.1)
+	letter.height=max(single.spp.SDF$value, na.rm=TRUE)+(diff(range(single.spp.SDF$value, na.rm=TRUE))*0.1)
 )
@@
 	geom_violin(position=position_dodge(0.9)) +
-	geom_point(alpha=0.2,position=position_jitterdodge(dodge.width=0.9, jitter.width=0.25)) +
 	geom_text(aes(x=Metric, y=letter.height, label=letters),
@@
 	letters=toupper(cld(multi.spp.MCP)$mcletters$Letters),
-	letter.height=max(multi.spp.SDF$value)+(diff(range(multi.spp.SDF$value))*0.1)
+	letter.height=max(multi.spp.SDF$value, na.rm=TRUE)+(diff(range(multi.spp.SDF$value, na.rm=TRUE))*0.1)
 )
@@
 	geom_violin(position=position_dodge(0.9)) +
-	geom_point(alpha=0.2,position=position_jitterdodge(dodge.width=0.9, jitter.width=0.25)) +
 	geom_text(aes(x=Metric, y=letter.height,label=letters),

### 2016-01-23 article/supporting_information.Rmd
+ fixed issues with order of bars in the violin plots + fixed issue with letters not showing in violin plots

@@
 			plotLST,
-			list(ncol=6)
+			list(ncol=4)
 		)

### 2016-01-24 article/supporting_information.Rmd
+ account for species with individuals with no data + fix captions in supplementary info

@@
 Each panel denotes a different species.
-Asterisks indicate the number of populations with the most support.
-See text for the method used to select these populations.'}
+Asterisks indicate the number of populations with the most support (see text for methodological details).'}
 # format data
@@
 Squares denote planning units.
-Points denote the location of sampled individuals.
-The color of each point indicates the population to which each indiviudal was assigned.'}
+Points denote the location of sampled individuals, and their colors denote the distribution of populations.
+Note that points have been slightly jittered so that individuals in close proximity are visible.'}
 # define function
@@
 ```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[1]),'}.
-Each square represents a planning unit.
-The color of each planning unit panel corresponds to ordination values.
-Planning units with similar colors contain individiduals with similar genetic variation.')}
+Each panel shows a different dimension of the variation as indicated by its title.
+Squares represent planning units.
+The color of each planning unit panel corresponds to the average ordinated values associated with individuals sampled inside it.
+No indiviudals of this species were found inside the gray planning units.')}
 if (1 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(1)

### 2016-01-24 R/analysis/01-load-data.R

@@ -2,6 +2,9 @@
 session::restore.session('results/.cache/00-initialization.rda')
 
 ## compile spatial grid data
+# all.spp <- seq_len(general.params.LST[[MODE]]$n.spp)
+all.spp <- c(1:3, 23)
+
 # load grid cell centroids
 grid.DF <- fread(
 	'data/Data_Meirmans_et_al_IntrabioDiv/ReadMe.txt',
@@ -19,7 +22,7 @@ spp.aflp.paths <- dir(
 	'data/Data_Meirmans_et_al_IntrabioDiv',
 	'^.*AFLP\\.dat$',
 	full.names=TRUE
-)[seq_len(general.params.LST[[MODE]]$n.spp)]
+)[all.spp]
 
 spp.BayeScanData.LST <- llply(
 	spp.aflp.paths,
@@ -37,7 +40,7 @@ spp.loc.paths <- dir(
 	'data/Data_Meirmans_et_al_IntrabioDiv',
 	'^.*locations\\.txt$',
 	full.names=TRUE
-)[seq_len(general.params.LST[[MODE]]$n.spp)]
+)[all.spp]
 spp.samples.DF <- ldply(
 	seq_along(spp.loc.paths),
 	.fun=function(i) {
@@ -56,13 +59,31 @@ spp.samples.DF <- ldply(
 		by='cell'
 )
 
-# append species data to grid data.frame (wide-format)
+## remove indidivduals that are all NAs
+for (i in seq_along(spp.StructureData.LST)) {
+	# find individuals that are all NAs
+	curr.invalid <- which(rowSums(is.na(spp.StructureData.LST[[i]]@matrix))==ncol(spp.StructureData.LST[[i]]@matrix))
+	# if any invalid then remove from objects
+	if (length(curr.invalid)>0) {
+		# get valid individuals
+		curr.valid <- which(rowSums(is.na(spp.StructureData.LST[[i]]@matrix))!=ncol(spp.StructureData.LST[[i]]@matrix))
+		# subset from StructureData
+		spp.StructureData.LST[[i]] <- structurer:::sample.subset.StructureData(spp.StructureData.LST[[i]], curr.valid)
+		# subset from BayescanData
+		spp.BayeScanData.LST[[i]] <- bayescanr:::sample.subset.BayeScanData(spp.BayeScanData.LST[[i]], curr.valid)
+		# remove from spp.samples.DF
[…]

### 2016-01-24 article/article.Rmd
+ fixes to article + fixes to stats + remove debugging that manually overrides species

@@
 Specifically, the amount-based prioritization captured much less adaptive and neutral genetic variation than either the surrogate-based ($Z = `r round(single.spp.MCP$test$tstat[['Surrogate.Adaptive variation - Amount.Adaptive variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Surrogate.Adaptive variation - Amount.Adaptive variation']])`$; $Z = `r round(single.spp.MCP$test$tstat[['Surrogate.Neutral variation - Amount.Neutral variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Surrogate.Neutral variation - Amount.Neutral variation']])`$) or genetic-based ($Z = `r round(single.spp.MCP$test$tstat[['Genetic.Adaptive variation - Amount.Adaptive variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Genetic.Adaptive variation - Amount.Adaptive variation']])`$; $Z = `r round(single.spp.MCP$test$tstat[['Genetic.Neutral variation - Amount.Neutral variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Genetic.Neutral variation - Amount.Neutral variation']])`$) prioritizations.
-No differences were detected in the proportion of adaptive ($Z = `r round(single.spp.MCP$test$tstat[['Surrogate.Adaptive variation - Genetic.Adaptive variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Surrogate.Adaptive variation - Genetic.Adaptive variation']])`$) or neutral genetic variation ($Z = `r round(single.spp.MCP$test$tstat[['Surrogate.Neutral variation - Genetic.Neutral variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Surrogate.Neutral variation - Genetic.Neutral variation']])`$) secured by the the surrogate-based and genetic-based prioritizations.
+No differences were detected in the proportion of adaptive ($Z = `r round(single.spp.MCP$test$tstat[['Genetic.Adaptive variation - Surrogate.Adaptive variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Genetic.Adaptive variation - Surrogate.Adaptive variation']])`$) or neutral genetic variation ($Z = `r round(single.spp.MCP$test$tstat[['Genetic.Neutral variation - Surrogate.Neutral variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Genetic.Neutral variation - Surrogate.Neutral variation']])`$) secured by the the surrogate-based and genetic-based prioritizations.
 However, these patterns were not reflected in the multi-species prioritizations.
@@
 <!-- Correlation analysis -->
-The surrogate variables showed a greater utility for adaptive genetic variation than neutral variation (Figure 4).
+The surrogate variables showed a greater utility for neutral genetic variation than adaptive variation (Figure 4).
 The accumulation rate ($r$) was `r round(summary(env.correlation.NLMM)$coefficients[1],2)` $\pm$ `r round(summary(env.correlation.NLMM)$coefficients[2],2)` S.E.
@@
 Thus minor increases in the target for either surrogate results in major increases in the benefit of the prioritization.
-However, the coefficient was much larger for the adaptative genetic variation so high targets are not required to acheive substantial gains in adaptive genetic variaition.
+However, the coefficient was much larger for neutral genetic variation so minor increases in the target can acheive substantial gains.
 

### 2016-01-24 article/figures.Rmd
+ fixes to article + fixes to stats + remove debugging that manually overrides species

--- 2016-01-20 article/figures.Rmd
@@
 	letters=toupper(cld(single.spp.MCP)$mcletters$Letters),
-	letter.height=max(single.spp.SDF$value)+(diff(range(single.spp.SDF$value))*0.1)
+	letter.height=max(single.spp.SDF$value, na.rm=TRUE)+(diff(range(single.spp.SDF$value, na.rm=TRUE))*0.1)
 )
 # make plot
+suppressWarnings({
 	ggplot(aes(x=Metric,y=value,fill=Prioritisation),
@@
 		geom_violin(position=position_dodge(0.9)) +
-	geom_point(alpha=0.2,position=position_jitterdodge(dodge.width=0.9, jitter.width=0.25)) +
 		geom_text(aes(x=Metric, y=letter.height, label=letters),
@@
 		theme_classic()
+})
 ```
@@
 	letters=toupper(cld(multi.spp.MCP)$mcletters$Letters),
-	letter.height=max(multi.spp.SDF$value)+(diff(range(multi.spp.SDF$value))*0.1)
+	letter.height=max(multi.spp.SDF$value, na.rm=TRUE)+(diff(range(multi.spp.SDF$value, na.rm=TRUE))*0.1)
 )
 # make plot
+suppressWarnings({
 	ggplot(aes(x=Metric,y=value,fill=Prioritisation),
@@
 		geom_violin(position=position_dodge(0.9)) +
-	geom_point(alpha=0.2,position=position_jitterdodge(dodge.width=0.9, jitter.width=0.25)) +
 		geom_text(aes(x=Metric, y=letter.height,label=letters),
@@
 ```{r, echo=FALSE, fig.width=8, fig.height=3, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations.
-Each panel shows the relationship for a different type of genetic variation.
-Each line denotes a different species.
-Lines are transparent and so darker regions show areas where species have similar relationships.'}
-# make plots
-p1 <- ggplot(data=env.pareto.DF) +
-	geom_line(aes(x=Surrogate.target,y=adaptive.held,group=Species),
-		alpha=0.3) +
+The left panel shows the relationship between environmental surrogates and adaptive genetic variation, and the right for geographic surrogates and neutral genetic variation.
+Transparent gray lines denote trneds for each species, and black lines show general trends across all species.'}
+### prepare data frames for plotting
+## environmental vs adaptive
+# random effects
+env.correlation.RPDF <- expand.grid(Surrogate.target=seq(min(env.correlation.DF$Surrogate.target), max(env.correlation.DF$Surrogate.target), length.out=100),
+	Species=unique(env.correlation.DF$Species))
+env.correlation.RPDF$r.ranef <- fixef(env.correlation.NLMM)[[1]] + ranef(env.correlation.NLMM)$Species$r[env.correlation.RPDF$Species]
+env.correlation.RPDF$fit <- 1 - (1 - 0) * exp(- env.correlation.RPDF$r.ranef * env.correlation.RPDF$Surrogate.target)
+# fixed effects
+env.correlation.FPDF <- data.frame(Surrogate.target=seq(min(env.correlation.DF$Surrogate.target), max(env.correlation.DF$Surrogate.target), length.out=100))
+env.correlation.FPDF$fit <- 1 - (1 - 0) * exp(- fixef(env.correlation.NLMM)[[1]] * env.correlation.FPDF$Surrogate.target)
+
+## geographic vs neutral
+# random effects
+geo.correlation.RPDF <- expand.grid(Surrogate.target=seq(min(geo.correlation.DF$Surrogate.target), max(geo.correlation.DF$Surrogate.target), length.out=100),
+	Species=unique(geo.correlation.DF$Species))
+geo.correlation.RPDF$r.ranef <- fixef(geo.correlation.NLMM)[[1]] + ranef(geo.correlation.NLMM)$Species$r[geo.correlation.RPDF$Species]
+geo.correlation.RPDF$fit <- 1 - (1 - 0) * exp(- geo.correlation.RPDF$r.ranef * geo.correlation.RPDF$Surrogate.target)
+# fixed effects
+geo.correlation.FPDF <- data.frame(Surrogate.target=seq(min(geo.correlation.DF$Surrogate.target), max(geo.correlation.DF$Surrogate.target), length.out=100))
+geo.correlation.FPDF$fit <- 1 - (1 - 0) * exp(- fixef(geo.correlation.NLMM)[[1]] * geo.correlation.FPDF$Surrogate.target)
+
+### make plots
+suppressWarnings({
+	p1 <- ggplot() +
+		geom_point(aes(x=Surrogate.target, y=adaptive.held), data=env.correlation.DF) +
+		geom_line(aes(x=Surrogate.target,y=fit,group=Species), data=env.correlation.RPDF, alpha=0.3) +
+		geom_line(aes(x=Surrogate.target,y=fit), data=env.correlation.FPDF, size=1.5) +
 		xlab('Environmental variation secured (%)') +
@@
 		theme_classic()
-p2 <- ggplot(data=geo.pareto.DF) +
-	geom_line(aes(x=Surrogate.target,y=neutral.held,group=Species),
-		alpha=0.3) +
+	p2 <- ggplot() +
+		geom_point(aes(x=Surrogate.target, y=neutral.held), data=geo.correlation.DF) +
+		geom_line(aes(x=Surrogate.target,y=fit,group=Species), data=geo.correlation.RPDF, alpha=0.3) +
+		geom_line(aes(x=Surrogate.target,y=fit), data=geo.correlation.FPDF, size=1.5) +
 		xlab('Geographic variation secured (%)') +
@@
 	grid.arrange(p1, p2, nrow=1)
+})
 ```

### 2016-01-24 article/supporting_information.Rmd
+ fixes to article + fixes to stats + remove debugging that manually overrides species

@@
 plotLST <- llply(
-	seq_along(unique(spp.samples.DF$species)),
+	# seq_along(unique(spp.samples.DF$species)),
+	seq_along(unique(spp.samples.DF$species)[-23]), # omit spp 23 from current version
 	.fun=function(i) {
@@
 See Figure S6 caption for conventions.')}
-if (23 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(23)
+# if (23 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(23)
 ```

### 2016-01-28 article/article.Rmd
+ generate replicate prioritisations for inference + fixes to article + switch to ecology letters ref style

@@
 <!-- NOTES -->
-<!-- 200 word limit for abstract -->
-<!-- 6 page limit w/o incurring costs -->
+<!-- 150 word limit for abstract -->
+<!-- 5000 word length limit -->
 
@@
 
-# load .rda files
-load('results/results.rda')
-
-# load packages
-source('R/analysis/00-initialization.R')
+# load .rda file
+session::restore.session('results/results.rda')
 
@@
 
+<!--  title page -->
+Short running title: Genetic surrogates in conservation \newline
+Word count: XXXXX \newline
+Number of references: XXXX \newline
+Number of figures, table and text boxes: 4 \newline
+
+\newpage
+
+<!-- add line numbers -->
+\begin{linenumbers}
+
 ## Introduction
+<!--  main text -->
 Biodiversity is under threat.
@@
 Two models were fit to the adaptive and neutral genetic variation secured using a negative exponential formulation to estimate the rate of the accumulation of genetic variation ($y = 0 + (0 - 1) \times exp(-r \times x)$; $r$ is the accumulate rate).
-Both models contained intercept and slope random effects to accommodate variation between species.
+Both models accomodated random effects for the accumulation rate ($r$) parameter.
 
@@
 ## References
+
+\end{linenumbers}

### 2016-01-28 article/figures.Rmd
+ generate replicate prioritisations for inference + fixes to article + switch to ecology letters ref style

@@
     fig_caption: yes
+    keep_tex: yes
     includes:
@@
 
-```{r, echo=FALSE, fig.width=6, fig.height=3, fig.cap='Summary of single species prioritizations.
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height=3, fig.cap='Summary of single species prioritizations.
 Single species prioritizations were generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species.
 Violin plots show the performance of prioritizations generated using these three sets of targets.
+Points indicate the average in each group.
 Letters denote significant differences ($P < 0.05$).'}
@@
 )
+
 # make plot
@@
 			data=single.spp.PDF, position=position_dodge(0.9)) +
+		stat_summary(fun.y=mean, geom='point', size=2,
+			position=position_dodge(0.9)) +
 		scale_fill_manual(name='Prioritisation',
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=3, fig.cap='Multi-species prioritizations.
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.height=3, fig.cap='Multi-species prioritizations.
 Panel (a) shows the prioritization generated for using just amount-based targets.
@@
 
-```{r, echo=FALSE, fig.width=6, fig.height=3, fig.cap='Summary of multi-species prioritizations.
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height=3, fig.cap='Summary of multi-species prioritizations.
 Three prioritizations were generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species.
@@
 			data=multi.spp.PDF, position=position_dodge(0.9)) +
+		stat_summary(fun.y=mean, geom='point', size=2,
+			position=position_dodge(0.9)) +
 		scale_fill_manual(name='Prioritisation',
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=3, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations.
+```{r, echo=FALSE,  warning=FALSE, message=FALSE, fig.width=8, fig.height=3, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations.
 The left panel shows the relationship between environmental surrogates and adaptive genetic variation, and the right for geographic surrogates and neutral genetic variation.

### 2016-01-28 article/supporting_information.Rmd
+ generate replicate prioritisations for inference + fixes to article + switch to ecology letters ref style

--- 2016-01-24 article/supporting_information.Rmd
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=6.5, fig.cap='Relative support ($BIC$) for the number of populations in each species.
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9.5, fig.height=10, fig.cap='Relative support ($BIC$) for the number of populations in each species.
 Each panel denotes a different species.
@@
 		ylab('BIC') +
-		facet_wrap(~ Species, ncol=6)
+		facet_wrap(~ Species, ncol=4)
 })
@@
 
-```{r, echo=FALSE, fig.width=6.5, fig.height=6.5, fig.cap='Distribution of populations.
+```{r, echo=FALSE, fig.width=15, fig.height=14, fig.cap='Distribution of populations.
 Each panel depicts a different species.
@@
 				alpha=0.3, size=3, position='jitter') +
-			guides(fill=guide_legend(title=' ')) +
+			guides(fill=guide_legend(title=''), color=guide_legend(title='')) +
 			theme_classic() +
@@
 		plotLST,
-		list(ncol=6)
+		list(ncol=4)
 	)
@@
 }
+
+calc.fig.height <- function(i) {
+	# init
+	adapt.col <- paste0(unique(spp.samples.DF$species)[i], '_adaptive_d1')
+	neutral.col <- paste0(unique(spp.samples.DF$species)[i], '_neutral_d1')
+	n.plots <- 0 
+	# count adaptive plot
+	if (adapt.col %in% names(grid.FPLY))
+		n.plots <- n.plots + spp.mds.LST[[i]][['adaptive']]$ndim
+	# count neutral plot
+	if (neutral.col %in% names(grid.FPLY))
+		n.plots <- n.plots + spp.mds.LST[[i]][['neutral']]$ndim
+	# exports
+	return(ceiling(n.plots/4) * 2.8)
+}
 ```
 
-```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[1]),'}.
+```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(1), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[1]),'}.
 Each panel shows a different dimension of the variation as indicated by its title.
@@
 
-```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[2]),'}.
+```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(2), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[2]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[3]),'}.
+```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(3), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[3]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[4]),'}.
+```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(4), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[4]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[5]),'}.
+```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(5), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[5]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[6]),'}.
+```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(6), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[6]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[7]),'}.
+```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(7), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[7]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[8]),'}.
+```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(8), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[8]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[9]),'}.
+```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(9), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[9]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[10]),'}.
+```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(10), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[10]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[11]),'}.
+```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(11), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[11]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[12]),'}.
+```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(12), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[12]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[13]),'}.
+```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(13), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[13]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[14]),'}.
+```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(14), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[14]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[15]),'}.
+```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(15), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[15]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[16]),'}.
+```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(16), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[16]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[17]),'}.
+```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(17), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[17]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[18]),'}.
+```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(18), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[18]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[19]),'}.
+```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(19), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[19]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[20]),'}.
+```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(20), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[20]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[21]),'}.
+```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(21), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[21]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[22]),'}.
+```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(22), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[22]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[23]),'}.
+```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(23), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[23]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[24]),'}.
+```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(24), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[24]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[25]),'}.
+```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(25), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[25]),'}.
 See Figure S6 caption for conventions.')}
@@
 ```
-```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[26]),'}.
+
+```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(26), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[26]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[27]),'}.
+```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(27), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[27]),'}.
 See Figure S6 caption for conventions.')}
@@
 	caption='Loci classifications for each species.
-Each row denotes a sampled loci for a given species, and the evidence for it being under selection based on outputs from the program \texttt{Bayescan}.'
+Each row denotes a sampled loci for a given species, and the evidence for it being under selection based on outputs from the program \\texttt{Bayescan}.'
 )

### 2016-01-28 R/analysis/07-multi-species-prioritisations.R

@@ -5,30 +5,40 @@ session::restore.session('results/.cache/06-single-species-prioritisations.rda')
 # make prioritisations
 multi.spp.prioritisations <- llply(
 	list(
-		c(rapr.params.LST[[MODE]]$amount.target,0,0), c(rapr.params.LST[[MODE]]$amount.target,rapr.params.LST[[MODE]]$surrogate.target,0), 
-			c(rapr.params.LST[[MODE]]$amount.target,0,rapr.params.LST[[MODE]]$genetic.target,0)
+		list(rapr.params.LST[[MODE]]$multi.species$amount.target,0,0,
+			rapr.params.LST[[MODE]]$multi.species$amount.replicates),
+		list(rapr.params.LST[[MODE]]$multi.species$amount.target,
+			rapr.params.LST[[MODE]]$multi.species$surrogate.target,0,
+			rapr.params.LST[[MODE]]$multi.species$surrogate.replicates), 
+		list(rapr.params.LST[[MODE]]$multi.species$amount.target,0,
+			rapr.params.LST[[MODE]]$multi.species$genetic.target,
+			rapr.params.LST[[MODE]]$multi.species$genetic.replicates)
 	), 
 	function(y) {
 		species.prioritisation(
 			x=ru,
-			amount.targets=y[1],
-			env.surrogate.targets=y[2],
-			geo.surrogate.targets=y[2],
-			adaptive.genetic.targets=y[3],
-			neutral.genetic.targets=y[3],
+			amount.targets=y[[1]],
+			env.surrogate.targets=y[[2]],
+			geo.surrogate.targets=y[[2]],
+			adaptive.genetic.targets=y[[3]],
+			neutral.genetic.targets=y[[3]],
 			Threads=gurobi.params.LST[[MODE]]$Threads,
-			MIPGap=gurobi.params.LST[[MODE]]$MIPGap
+			MIPGap=gurobi.params.LST[[MODE]]$MIPGap,
+			NumberSolutions=y[[4]]
 		)
 	}
 )
 
 # generate results table
-multi.spp.DF <- ldply(seq_along(multi.spp.prioritisations), function(i) {
-	mutate(
-		extractResults(multi.spp.prioritisations[[i]]),
-		Prioritisation=c('Amount', 'Surrogate', 'Genetic')[i]
-	)
-})
+multi.spp.DF <- ldply(
+	seq_along(multi.spp.prioritisations),
+	.fun=function(i) {
+		mutate(
[…]

### 2016-01-30 article/article.Rmd
+ cost data + create multi-species priortisation with cost

@@
 Biodiversity is under threat.
-Over the last few decades, many species have gone extinct [@r270; @r272].
+In recent history, many species have gone extinct [@r270; @r272].
 Today, thousands of species are listed as endangered or near threatened in the International Union for Conservation of Nature (IUCN) red list (\url{www.iucnredlist.org}).
-In the future, species are expected to face even more threats [eg.
+In the near future, species are expected to face even more threats [eg.
 climate change; @r236; @r243].
@@
 
-The act of setting aside areas for preservation is the hallmark of modern conservation [@r440].
-Protected areas buffer species from anthropogenic impacts and provide places for biodiversity processes to persist (eg.
+Protected areas are one of the hallmarks of modern conservation [@r440].
+They buffer species from anthropogenic impacts and provide places for biodiversity processes to persist (eg.
 nutrient cycling and water purification).
-Additionally, protected areas set the stage for direct management actions (eg.
+Additionally, protected areas set the stage for direct management actions to enhance species recovery (eg.
 habitat rehabilitation, translocation, and invasive species management).
-Since the resources available for conservation are limited, the plans for protected areas (prioritizations) are often generated using spatially explicit biological data (eg.
-species distribution data) and decision support tools [eg.
-$\texttt{MARXAN}$; @r22; $\texttt{Zonation}$; @r429] that identify cost-effective solutions.
+Since the resources available for conservation are limited, protected areas must be cost-effective.
+As a consequence, plans for protected areas (prioritizations) are often generated using decision support tools [eg.
+$\texttt{MARXAN}$; @r22; $\texttt{Zonation}$; @r429].
+These tools often formulate the benefit that a given species receives from a solution based on the proportion of its range that is protected.
+However, the true benefit that a species might receive is dependent on many other additional factors.
 
@@
 
-Recently, conservation scientists have begun investigating the use of surrogates for genetic data to generate more effective prioritizations without the overheads associated with genetic data.
+Recently, conservation scientists have begun investigating surrogates to generate more effective prioritizations without the overheads associated with genetic data.
 Since adaptive genetic variation is driven by selection pressures, environmental variables have been explored as potential surrogates for genetic variation [eg.
@@
 The use of distance-based surrogates has been found to improve prioritizations in an island network [@r200].
-However the effectiveness of distance-based surrogates remains untested in systems where connectivity is complicated by additional factors (eg.
+However the effectiveness of distance-based surrogates remains untested in contiguous systems conventionally used in conservation planning where gene flow is complicated by additional factors (eg.
 elevation in terrestrial systems).
 Before conservation planners can reliably use surrogates for genetic data in conservation planning, the effectiveness of these surrogates needs to be verified.
-Otherwise, conservation planners may waste precious resources on securing places that do little to benefit biodiversity.
+Otherwise, if these surrogates are ineffective, conservation planners may waste precious resources on places with little benefit.
 
@@
 To address the aims of the study, we utilized species distribution and genomic data collected by the IntraBioDiv project [@r451; see @r452 for methodological details].
-This dataset was well suited for this study, because it provides spatially explicit genomic data for a multitude of species across a broad geographic area and over large environmental gradients.
-As part of this project, data for 27 alpine plant species were collected across the European Alps according to a grid (22.3 km $\times$ 25 km; Figure S1).
+This dataset was particularly well suited; it provides spatially explicit genomic data for a multitude of species across a broad geographic area over large environmental gradients.
+As part of this project, data were collected for 27 alpine plant species were across the European Alps according to a grid (22.3 km $\times$ 25 km; Figure S1).
 Briefly, in every second cell, plant samples were collected from three individuals for each species present inside the cell.
@@
 
-### Surrogate data
-We explored the effectiveness of climatic and geographic surrogates.
-To describe variation in the location of each grid cell, we projected the grid into an equi-distant coordinate system ($\texttt{ESRI:102031}$) and used the centroid of each cell.
-To describe the climatic variation among each cell, we obtained 21 bioclimatic variables [approx.
+### Landscape data
+We aimed to investigate the effectiveness of climatic and geographic surrogates.
+To describe the location of each grid cell, we projected the grid into an equi-distant coordinate system ($\texttt{ESRI:102031}$), calculated their centroids, and extracted its two-dimensional coordinates.
+To describe the climatic variation among each cell, we obtained 19 bioclimatic variables [approx.
 1 km resolution; @r33].
@@
 We calculated the average value for each PC layer in each grid cell, and used these values to characterize the main themes of climatic variation among the grid cells (Figure S2).
+We used the total human population size inside each planning unit to represent its opportunity cost (obtained at 1 km resolution from the Global Rural-Urban Mapping Project; GRUMP V1; @rXXX).
 
 ### Outlier locus detection
-To investigate the effectiveness of the surrogates for adaptive and neutral genetic variation, we first identified which of the sampled loci for each species were under selection.
-Specifically, we used outlier locus detection methods [evaluated in @r454] to avoid circularity issues.
+To investigate the effectiveness of the surrogates for adaptive and neutral genetic variation, we first need to identify which of the sampled loci--if any--were under selection for each species.
+We used outlier locus detection methods [evaluated in @r454] to achieve this.
 The basic premise underpinning these analyses is that neutral loci are expected to exhibit a specific level of variation among populations ($F_{st}$), and loci that deviate from this expectation should be under selection.
+The advantage of these methods, in contrast with environmental association analyses [eg.
+r@XXX], is that they do not use environmental data to identify loci under selection, which would have biased our analysis.
 In an earlier study, Bothwell _et al._ [-@r455] applied such outlier detection methods to one of the species in this dataset (_Getiana nivalis_).
@@
 We set the demand points for each species and its attribute space using the values associated with the planning units.
-Although Hanson _et al._ [-@r465] recommend against using the same coordinates for demand points and planning units for real-word conservation planning exercises, we did this to provide a more robust test and reduce uncertainty in the analysis.
 For every prioritization we generated, we calculated the proportion of the adaptive (if detected) and neutral genetic variation that every prioritization secured to assess its performance.
@@
 
-To understand how prioritizations generated using surrogate-based targets might perform under typical conservation planning scenarios, we generated a set of single-species and multi-species prioritizations.
-All of these prioritizations used `r (rapr.params.LST[[MODE]]$amount.target*100)` % amount-based targets to ensure an adequate proportion of habitat was secured for each species [@r197; @r255; @r471].
-We generated a set of single-species prioritizations for each species: `r rapr.params.LST[[MODE]]$single.species.amount.replicates` prioritizations using only the amount-based targets, a single prioritization using the amount-based target and `r (rapr.params.LST[[MODE]]$surrogate.target*100)` % surrogate-based targets, and a single prioritization using the amount-based target and `r (rapr.params.LST[[MODE]]$genetic.target*100)` % genetic-based targets.
-We also generated three multi-species prioritizations using the three unique combinations of targets used to generate the single-species prioritizations.
-Note that we obtained multiple solutions for the amount-based single-species prioritizations because, unlike the other prioritizations, these targets yielded many optimal solutions with varying levels effectiveness.
+To understand how prioritizations generated using surrogate-based targets might perform in conservation planning scenarios, we generated three sets of prioritizations: a set of single-species prioritizations, a set of multi-species prioritizations, and a set of multi-species prioritizations that accommodated opportunity cost.
+For each of these sets, we generated prioritizations using only amount-based targets, prioritizations using both amount- and surrogate-based targets, and prioritizations using amount- and genetic-based targets.
+All amount-based targets were set as `r (rapr.params.LST[[MODE]]$amount.target*100)` % to ensure an adequate proportion of habitat was secured for each species [@rXXX].
+All surrogate- and genetic-based targets were set as `r (rapr.params.LST[[MODE]]$surrogate.target*100)`.
+For each set, for each combination of targets, we identified the `r rapr.params.LST[[MODE]]$single.species.surrogate.replicates` solutions nearest to optimality for replication.
+Except for the single-species amount-only prioritizations for which we generated `r rapr.params.LST[[MODE]]$single.species.amount.replicates` replicates because they were associated with many optimal solutions with varying levels of effectiveness.
 
@@
 We fit two full models to the proportion of genomic variation secured in the single- and multi-species prioritizations using a logit link.
-Both models contained a predictor variable indicating if the prioritization was generated using amount, surrogate, or genetic targets, a second predictor variable indicating if the genetic variation was adaptive or neutral, and the interaction between these variables.
+Both models contained a predictor variable indicating if the prioritization was generated using amount-only, surrogate, or genetic targets, a second predictor variable indicating if the genetic variation was adaptive or neutral, and the interaction between these variables.
 Additionally, both models contained an intercept-only random effect to ensure that comparisons were made within species.
 Each full model was subject to a step-wise term deletion routine using likelihood ratio tests (LRT) to assess term significance (terms were retained when $P < 0.05$).
-Finally, we conducted a post-hoc analysis using Tukey constrasts with a Bonferroni correction [using the $\texttt{multcomp}$ R package; @r470].
+Finally, we conducted a post-hoc analysis using Tukey contrasts with a Bonferroni correction [using the $\texttt{multcomp}$ R package; @r470].
 
 We investigated the relationship between surrogate-based targets and the genetic variation secured in a prioritization using nonlinear mixed-effects models (NLMMs; using the $\texttt{lme4}$ R package).
-Two models were fit to the adaptive and neutral genetic variation secured using a negative exponential formulation to estimate the rate of the accumulation of genetic variation ($y = 0 + (0 - 1) \times exp(-r \times x)$; $r$ is the accumulate rate).
-Both models accomodated random effects for the accumulation rate ($r$) parameter.
+Two models were fit to the adaptive and neutral genetic variation secured using a negative exponential formulation to estimate the rate of the accumulation of genetic variation ($y = a - (a - 0) \times exp(-r \times x)$; where $a$ is the asymptote and $r$ is the accumulate rate).
+Both models accommodated species-level random effects for the asymptote ($a$) and rate ($r$) parameters.
 
@@
 <!-- Multi-species prioritizations -->
-Surrogate-targets were not found to improve the performance of multi-species prioritizations (Figure 2; Figure 3).
+Environmental and geographic surrogates were not found to improve the performance of multi-species prioritizations (Figure 2; Figure 3).
 No significant differences in the effectiveness of prioritizations were found depending on whether they were generated using different targets (${\chi^{2}}_{`r multi.spp.AOV3[['Df']][1]`,`r multi.spp.AOV3[['Df']][2]`} = `r round(multi.spp.AOV3[['Chisq']][2],2)`$; $P `r pretty.pval(multi.spp.AOV3[['Pr(>Chisq)']][2])`$), if they were assessed using adaptive or neutral genetic variation (${\chi^{2}}_{`r multi.spp.AOV2[['Df']][1]`,`r multi.spp.AOV2[['Df']][2]`} = `r round(multi.spp.AOV2[['Chisq']][2],2)`$; $P `r pretty.pval(multi.spp.AOV2[['Pr(>Chisq)']][2])`$), nor the interaction between these variables (${\chi^{2}}_{`r multi.spp.AOV1[['Df']][1]`,`r multi.spp.AOV1[['Df']][2]`} = `r round(multi.spp.AOV1[['Chisq']][2],2)`$; $P  `r pretty.pval(multi.spp.AOV1[['Pr(>Chisq)']][2])`$).
@@
 Thus minor increases in the target for either surrogate results in major increases in the benefit of the prioritization.
-However, the coefficient was much larger for neutral genetic variation so minor increases in the target can acheive substantial gains.
+However, the coefficient was much larger for neutral genetic variation so minor increases in the target can achieve substantial gains.
 
 ## Discussion
-### Introduction
-* aim
-	+ we aimed to investigate the benefits of using surrogates for genetic variation in conservation planning.
-* innovation
-	+ first to test the effectiveness of environmental and geographic surrogates in a conservation planning context
-* key findings
-	+ surrogate variables are beneficial in single-species contexts.
-	+ but not in multi-species contexts when a comprehensive set of species is used.
-* consequences
-	+ highlights the benefit of using a comprehensive set of species
-	+ supports literature on using comprehensive set of species
+We aimed to investigate the effectiveness of environmental and geographic surrogates for genetic variation in conservation planning.
+This study is the first to empirically test this using a realistic system.
+We found that the inclusion of environmental and geographic surrogates in single-species prioritizations significantly improved the proportion of intra-specific genetic variation they secured.
+But we also found that the same surrogates did little to improve the multi-species prioritizations.
+Our results suggest that conservation planners can use a comprehensive set of species to secure a sufficient proportion of their intra-specific genetic variation.
 
 ### Single + multi-species analysis
+* summarize results
+
+* compare + contrast with other papers
+
+* implication of results
 
 ### Correlation analysis
+* summarize results
+
+* compare + contrast with other papers
+
+* implication of results
 
 ### Limitations/Future directions
-* spatial/surrogate data
-	+ isolation by resistance: least-cost paths and circuit theory distances are becoming more population for modeling connectivity, and may offer better surrogates than linear geographic distances.
-However, the scaling for these variables requires an understanding of the often non-linear relationship, which is not available to conservation planners
-	+ cost data: here we used no cost data to provide a rigorous test for the benefits of surrogates in conservation planning.
-however in real conservation planning situations, prioritizations are largely driven by cost.
-future studies should look at including cost data to see if surrogates provide any benefit when including realistic cost data.
-
-* genetic data
-	+ AFLP data: next gen data would be better however the aim of a prioritization is to capture the main gradients of genetic variation--not all of the variation.
-	+ adegenet: whilst other software packages are more popular (eg.
-Structure), we used adegenet because information theory is more reliable than an ad-hoc second derivative process for comparing models.
-Additionally, it is computationally infeasible to analyze run 27 species in Structure in a suitable period of time.
+We wish to alert practitioners to several limitations associated with our analysis.
+First, the size of the planning units we used (23 \times 25 km) is much larger than typically used in conservation planning scenarios (approx.
+1--5 km^2).
+We used this resolution because the genetic data was collected at this scale.
+Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis because the surrogate variables are spatially auto-correlated.
+Second, we used geographic distances as surrogates for connectivity between planning units.
+Although connectivity between areas can be modeled using isolation by resistance measures derived from landscape data [eg.
+elevation; reviewed in @rXX], they are unlikely to be of use to conservation planners since they require genetic data to be tuned.
+Third, we used AFLP data to characterize the main gradients of genetic variation.
+Whilst next-generation sequencing provides higher resolution genetic information than AFLP data [reviewed in @rXXX], our methodology would still have utilised only the main gradients of the genetic variation to generate prioritizations in a feasible amount of time.
+Overall, the limitations associated with this study are relatively minor.
 
 ### Concluding remarks
-* take-home messages
-	+ surrogates are effective for single-species prioritizations in conservation planning
-	+ however, surrogates don't provide any noticeable benefits when using a comprehensive set of species in a multi-species prioritization
-	+ conservation planners are encouraged to use comprehensive sets of species
+The overarching aim of our study was to investigate the effectiveness of surrogates in conservation planning.
+We found that environmental and geographic surrogates are effective.
+However, they seem to be largely unnecessary in conservation planning scenarios because a comprehensive set of species can be used to achieve a similar result.
+Conservation planners are therefore urged to use a comprehensive set of species in prioritizations--one of the fundamental principles of conservation planning.
 

### 2016-01-30 R/analysis/02-surrogate-data.R

@@ -38,11 +38,22 @@ grid.DF <- cbind(grid.DF, centroids.DF)
 bioclim.STK <- stack('data/BioClim_variables/bioclim_pca.tif')
 # extract mean for each cell for each principle component
 extract.DF <- grid.PPLY %>% rasterize(bioclim.STK, field='id') %>% 
-	zonal(x=bioclim.STK) %>% as.data.frame() %>% select(-1) %>%
+	zonal(x=bioclim.STK) %>%  as.data.frame() %>% select(-1) %>%
 	`names<-`(paste0('env_d',seq_len(nlayers(bioclim.STK))))
 # merge with grid.DF
 grid.DF <- cbind(grid.DF, extract.DF)
 
+## extract population density data
+# load pop data
+pop.RST <- raster('data/GRUMP_V1_Population_Density/grumpv1-popdensity.tif') %>% 
+	crop(grid.PLY) %>% projectRaster(crs=grid.PPLY@proj4string)
+# extract total for each cell
+extract2.DF <- grid.PPLY %>% rasterize(pop.RST, field='id') %>% 
+	zonal(x=pop.RST, fun='sum') %>% as.data.frame() %>%
+	select(-1) %>% `names<-`('pop.density') %>% mutate(pop.density=pop.density/1000)
+# merge with grid.DF
+grid.DF <- cbind(grid.DF, extract2.DF)
+	
 ## update @data slot in spatial objects
 grid.PLY@data <- grid.DF
 grid.PPLY@data <- grid.DF

### 2016-01-30 R/analysis/05-format-data-for-prioritisations.R

@@ -59,7 +59,7 @@ target.DF <- make.targets(
 rd <- RapData(
 	polygon=SpatialPolygons2PolySet(grid.PLY),
 	pu=data.frame(
-		cost=rep(1, nrow(grid.DF)),
+		cost=grid.DF$pop.density,
 		area=rep(1, nrow(grid.DF)),
 		status=rep(0L, nrow(grid.DF))
 	),
@@ -78,7 +78,13 @@ rd <- RapData(
 	),
 	boundary=calcBoundaryData(grid.PLY)
 )
+
+# create RapUnsolved without cost data
 ru <- RapUnsolved(RapUnreliableOpts(), rd)
- 
+ru@data@pu$cost <- 1
+
+# create RapUnsolved with  cost data
+ru_with_cost <- RapUnsolved(RapUnreliableOpts(), rd)
+
 ## save .rda
 save.session('results/.cache/05-format-data-for-prioritisations.rda')

### 2016-02-02 article/figures.Rmd
more changes

@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height=3, fig.cap='Summary of single species prioritizations.
-Single species prioritizations were generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species.
-Violin plots show the performance of prioritizations generated using these three sets of targets.
-Points indicate the average in each group.
-Letters denote significant differences ($P < 0.05$).'}
-# prepare data for plotting
-single.spp.PDF <- expand.grid(
-	Prioritisation=unique(single.spp.SDF$Prioritisation),
-	Metric=unique(single.spp.SDF$Metric)
-) %>% mutate(
-	letters=toupper(cld(single.spp.MCP)$mcletters$Letters),
-	letter.height=max(single.spp.SDF$value, na.rm=TRUE)+(diff(range(single.spp.SDF$value, na.rm=TRUE))*0.1)
-)
-
-# make plot
-suppressWarnings({
-	ggplot(aes(x=Metric,y=value,fill=Prioritisation),
-		data=single.spp.SDF) +
-		geom_violin(position=position_dodge(0.9)) +
-		geom_text(aes(x=Metric, y=letter.height, label=letters),
-			data=single.spp.PDF, position=position_dodge(0.9)) +
-		stat_summary(fun.y=mean, geom='point', size=2,
-			position=position_dodge(0.9)) +
-		scale_fill_manual(name='Prioritisation',
-			values=c('grey80','grey50','grey20')) +
-		ylab('Genetic variation secured (%)') +
-		xlab('') +
-		theme_classic()
-})
-```
-
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.height=3, fig.cap='Multi-species prioritizations.
-Panel (a) shows the prioritization generated for using just amount-based targets.
-Panel (b) shows the prioritization generated using amount-based and surrogate based targets.
-Panel (c) shows the prioritization generated using amount-based and genetic-based targets.'}
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=5, fig.cap='Planning unit selection frequencies.
+The top row of panels (a, b, c) show single-species prioritizations generated assuming equal costs.
+The middle row of panels (d, e, f) show multi-species prioritizations generated assuming equal costs.
+The bottom row of panels (g, h, i) show multi-species generated using opportunity costs.
+The left column of panels (a, d, g) show prioritizations generated using only amount-based targets.
+The middle column of panels (b, e, h) show prioritizations generated using amount- and surrogate-based targets.
+The right column of panels (c, f, i) show prioritizations generated using amount- and genetic-based targets.'}
 # download basemap
@@
 # prepare data for plotting
-multi.spp.grid.FPLY <- grid.PLY
+grid.FPLY <- grid.PLY
+for (i in seq_along(single.spp.prioritisations))
+	grid.FPLY@data[[paste0('Single-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations[[i]], function(x){colMeans(x@results@selections)}))
 for (i in seq_along(multi.spp.prioritisations))
-	multi.spp.grid.FPLY@data[[paste0('v',i)]] <- selections(multi.spp.prioritisations[[i]])
-multi.spp.grid.FPLY <- spFortify(multi.spp.grid.FPLY)
+	grid.FPLY@data[[paste0('Multi-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations[[i]]@results@selections)
+for (i in seq_along(multi.spp.prioritisations.with.cost))
+	grid.FPLY@data[[paste0('Multi-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations.with.cost[[i]]@results@selections)
+grid.FPLY <- spFortify(grid.FPLY)
+grid.FPLY <- grid.FPLY %>% gather(
+		Combination,
+		`Selection.Frequency`,
+		`Single-species (equal costs).Amount`:`Multi-species (opportunity costs).Genetic`
+	) %>% mutate(
+		Context=sapply(strsplit(as.character(Combination), '.', fixed=TRUE), `[[`, 1),
+		Prioritization=sapply(strsplit(as.character(Combination), '.', fixed=TRUE), `[[`, 2)
+	) %>% mutate(
+		Prioritization=factor(Prioritization, levels=c('Amount', 'Surrogate', 'Genetic')),
+		Context=factor(
+			gsub(' (', '\n(', Context, fixed=TRUE),
+			levels=c("Single-species\n(equal costs)", "Multi-species\n(equal costs)", 
+			"Multi-species\n(opportunity costs)"))
+	)
+panel.DF <- grid.FPLY %>% select(Context, Prioritization, long, lat) %>% 
+	mutate(
+		lat=min(lat)+(diff(range(lat))*0.98),
+		long=min(long)+(diff(range(long))*0.02)
+	) %>%
+	group_by(Context,Prioritization) %>% filter(c(TRUE, rep(FALSE, length(Context)-1))) %>%
+	mutate(
+		panel.letter=paste0('(',letters[(as.integer(Context)-1)*3 + as.integer(Prioritization)],')'),
+		panel.letter=c(as.character(panel.letter[1]), rep(NA_character_, length(Context)-1))
+	)
 # make maps
-do.call(
-	grid.arrange,
-	append(
-		llply(
-			seq_along(multi.spp.prioritisations),
-			function(i) {
 ggplot() +
 	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
-						fill='grey20', color='grey80') +
-					geom_polygon(data=multi.spp.grid.FPLY, aes_string(x='long', y='lat', 
-						group='group', fill=paste0('v',i)),
+		fill='grey85', color='grey70') +
+	geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
+		group='group', fill='Selection.Frequency'),
 		alpha=0.8, color='grey10') +
-					guides(fill=guide_legend(title=' ')) +
+	geom_text(aes_string(x='long',y='lat', label='panel.letter'), data=panel.DF) +
 	theme_classic() +
@@
 		plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(),
-						legend.position='none') +
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white'),
+		legend.position='bottom', legend.key.width=unit(1.8, 'cm'),
+		legend.title=element_text(vjust=-5),
+		legend.margin=unit(c(0.1), 'cm'),
+		axis.title=element_blank()) +
 	coord_cartesian(
-						xlim=buffered.range(multi.spp.grid.FPLY$long, 0.05),
-						ylim=buffered.range(multi.spp.grid.FPLY$lat, 0.05)
+		xlim=buffered.range(grid.FPLY$long, 0),
+		ylim=buffered.range(grid.FPLY$lat, 0)
 	) +
-					xlab('') +
-					ylab('') +
-					annotate('text',
-						x=min(multi.spp.grid.FPLY$long)+diff(range(multi.spp.grid.FPLY$long)*0.1),
-						y=min(multi.spp.grid.FPLY$lat)+diff(range(multi.spp.grid.FPLY$lat)*1.01),
-						label=letters[i], hjust=1, vjust=1, color='white', size=8)
-			}
-		),
-		list(nrow=1)
-	)
+	facet_grid(Context ~ Prioritization) +
+	scale_fill_gradientn(name='Selection Frequency (%)',
+		colors=brewer.pal(9, 'PuBu'),
+		guide=guide_colorbar(title.position='bottom', title.hjust=0.5,
+			ticks=element_line(color='black'), border=element_line(color='black'))
 	)
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height=3, fig.cap='Summary of multi-species prioritizations.
-Three prioritizations were generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species.
-Data shows the performance of these prioritizations based on how much genetic variation they explain.
-See Figure 1 caption for conventions.'}
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=5, fig.cap='The effectiveness of prioritizations at securing genetic variation.
+The top row of panels (a--b) show the performance of single-species prioritizations generated assuming equal costs among planning costs.
+The middle row of panels (c--d) show the performance of multi-species prioritizations generated assuming equal costs among planning units.
+The bottom row of panels (e--f) show the performance of multi-species prioritizations generated using opportunity costs.
+The left column of panels (a, c, d) and the right column of panels (b, d, f) show the proportion of adaptive and neutral genetic variation secured (respectively).
+Box plots show the effectiveness of prioritizations generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species.
+Letters denote significant differences ($P < 0.05$).'}
 # prepare data for plotting
-multi.spp.PDF <- expand.grid(
-	Prioritisation=unique(multi.spp.SDF$Prioritisation),
-	Metric=unique(multi.spp.SDF$Metric)
+results.SDF <- results.SDF %>% mutate(
+		Prioritisation=factor(as.character(Prioritisation), levels=c('Amount','Surrogate','Genetic')),
+		Context=factor(as.character(Context), levels=c("single-species (equal costs)", "multi-species (equal costs)", "multi-species (opportunity costs)")),
+		Metric=factor(as.character(Metric), c("Adaptive variation", "Neutral variation"))
+	)
+results.PDF <- results.SDF %>% group_by(
+		Prioritisation, Context, Metric
 	) %>% mutate(
-	letters=toupper(cld(multi.spp.MCP)$mcletters$Letters),
-	letter.height=max(multi.spp.SDF$value, na.rm=TRUE)+(diff(range(multi.spp.SDF$value, na.rm=TRUE))*0.1)
+		letter.height=max(value, na.rm=TRUE)+0.04,
+		Combination=paste0(Prioritisation,'.',Metric,'.',Context)
+	) %>% filter(
+		c(TRUE, rep(FALSE, length(Context)-1))
+	) %>% data.frame() %>% group_by(Context, Metric) %>% mutate(
+		panel.letter=paste0('(',letters[as.integer(Context) + (as.integer(Metric)-1)],')'),
+		panel.letter=c(as.character(panel.letter[1]), rep(NA_character_, length(Context)-1))
+	) %>% data.frame() %>% left_join(
+		data.frame(
+			letter=unname(toupper(cld(full.model.GLHT)$mcletters$Letters)),
+			Combination=as.character(names(toupper(cld(full.model.GLHT)$mcletters$Letters)))
+		) %>% mutate(Combination=as.character(Combination)),
+		by='Combination'
+	) %>% data.frame()
+results.SDF <- results.SDF %>% mutate(
+	Context=sapply(as.character(Context), function(x) {paste0(toupper(substr(x, 1, 1)), substr(x, 2, nchar(x)))}),
+	Context=factor(
+		gsub(' (', '\n(', Context, fixed=TRUE),
+		levels=c("Single-species\n(equal costs)",
+		"Multi-species\n(equal costs)", 
+		"Multi-species\n(opportunity costs)"))
+	)
+results.PDF <- results.PDF %>% mutate(
+	Context=sapply(as.character(Context), function(x) {paste0(toupper(substr(x, 1, 1)), substr(x, 2, nchar(x)))}),
+	Context=factor(
+		gsub(' (', '\n(', Context, fixed=TRUE),
+		levels=c("Single-species\n(equal costs)",
+		"Multi-species\n(equal costs)", 
+		"Multi-species\n(opportunity costs)"))
 	)
@@
 suppressWarnings({
-	ggplot(aes(x=Metric,y=value,fill=Prioritisation),
-		data=multi.spp.SDF) +
-		geom_violin(position=position_dodge(0.9)) +
-		geom_text(aes(x=Metric, y=letter.height,label=letters),
-			data=multi.spp.PDF, position=position_dodge(0.9)) +
-		stat_summary(fun.y=mean, geom='point', size=2,
-			position=position_dodge(0.9)) +
+	ggplot(aes(x=Prioritisation,y=value),
+		data=results.SDF) +
+		geom_boxplot(position=position_dodge(0.9)) +
+		geom_text(aes(x=Prioritisation, y=letter.height, label=letter),
+			data=results.PDF, position=position_dodge(0.9)) +
+		geom_text(aes(x=0.6, y=1.03, label=panel.letter),
+			data=results.PDF) +
 		scale_fill_manual(name='Prioritisation',
@@
 		xlab('') +
-		theme_classic()
+		ylim(c(NA, 1.05)) +
+		theme_classic() +
+		theme(strip.background = element_rect(fill='grey20'),
+			strip.text = element_text(color='white'),
+			panel.border = element_rect(color='black', fill=NA),
+			axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
+		facet_grid(Context ~ Metric)
 })
 ```
-
-```{r, echo=FALSE,  warning=FALSE, message=FALSE, fig.width=8, fig.height=3, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations.
-The left panel shows the relationship between environmental surrogates and adaptive genetic variation, and the right for geographic surrogates and neutral genetic variation.
-Transparent gray lines denote trneds for each species, and black lines show general trends across all species.'}
-### prepare data frames for plotting
-## environmental vs adaptive
-# random effects
-env.correlation.RPDF <- expand.grid(Surrogate.target=seq(min(env.correlation.DF$Surrogate.target), max(env.correlation.DF$Surrogate.target), length.out=100),
-	Species=unique(env.correlation.DF$Species))
-env.correlation.RPDF$r.ranef <- fixef(env.correlation.NLMM)[[1]] + ranef(env.correlation.NLMM)$Species$r[env.correlation.RPDF$Species]
-env.correlation.RPDF$fit <- 1 - (1 - 0) * exp(- env.correlation.RPDF$r.ranef * env.correlation.RPDF$Surrogate.target)
-# fixed effects
-env.correlation.FPDF <- data.frame(Surrogate.target=seq(min(env.correlation.DF$Surrogate.target), max(env.correlation.DF$Surrogate.target), length.out=100))
-env.correlation.FPDF$fit <- 1 - (1 - 0) * exp(- fixef(env.correlation.NLMM)[[1]] * env.correlation.FPDF$Surrogate.target)
-
-## geographic vs neutral
-# random effects
-geo.correlation.RPDF <- expand.grid(Surrogate.target=seq(min(geo.correlation.DF$Surrogate.target), max(geo.correlation.DF$Surrogate.target), length.out=100),
-	Species=unique(geo.correlation.DF$Species))
-geo.correlation.RPDF$r.ranef <- fixef(geo.correlation.NLMM)[[1]] + ranef(geo.correlation.NLMM)$Species$r[geo.correlation.RPDF$Species]
-geo.correlation.RPDF$fit <- 1 - (1 - 0) * exp(- geo.correlation.RPDF$r.ranef * geo.correlation.RPDF$Surrogate.target)
-# fixed effects
-geo.correlation.FPDF <- data.frame(Surrogate.target=seq(min(geo.correlation.DF$Surrogate.target), max(geo.correlation.DF$Surrogate.target), length.out=100))
-geo.correlation.FPDF$fit <- 1 - (1 - 0) * exp(- fixef(geo.correlation.NLMM)[[1]] * geo.correlation.FPDF$Surrogate.target)
-
-### make plots
-suppressWarnings({
-	p1 <- ggplot() +
-		geom_point(aes(x=Surrogate.target, y=adaptive.held), data=env.correlation.DF) +
-		geom_line(aes(x=Surrogate.target,y=fit,group=Species), data=env.correlation.RPDF, alpha=0.3) +
-		geom_line(aes(x=Surrogate.target,y=fit), data=env.correlation.FPDF, size=1.5) +
-		xlab('Environmental variation secured (%)') +
-		ylab('Adaptive genetic\nvariation secured (%)') +
-		theme_classic()
-	p2 <- ggplot() +
-		geom_point(aes(x=Surrogate.target, y=neutral.held), data=geo.correlation.DF) +
-		geom_line(aes(x=Surrogate.target,y=fit,group=Species), data=geo.correlation.RPDF, alpha=0.3) +
-		geom_line(aes(x=Surrogate.target,y=fit), data=geo.correlation.FPDF, size=1.5) +
-		xlab('Geographic variation secured (%)') +
-		ylab('Neutral genetic\nvariation secured (%)') +
-		theme_classic()
-	grid.arrange(p1, p2, nrow=1)
-})
-```

### 2016-02-02 R/analysis/09-statistical-analysis.R

@@ -1,0 +1,42 @@
+## load .rda
+session::restore.session('results/.cache/08-multi-species-prioritisations-with-cost.rda')
+
+### Statistical analysis
+## prepare data
+# single species prioritisations
+single.spp.SDF <- single.spp.DF %>%
+	gather(Metric, value, amount.held:neutral.held) %>%
+	filter(Metric %in% c('adaptive.held', 'neutral.held')) %>%
+	mutate(Context='single-species (equal costs)')
+	
+# multi-species prioritisations without cost
+multi.spp.SDF <- multi.spp.DF %>%
+	gather(Metric, value, amount.held:neutral.held) %>%
+	filter(Metric %in% c('adaptive.held', 'neutral.held')) %>%
+	mutate(Context='multi-species (equal costs)') 
+
+# multi-species prioritisations with cost
+multi.spp.with.cost.SDF <- multi.spp.with.cost.DF %>%
+	gather(Metric, value, amount.held:neutral.held) %>%
+	filter(Metric %in% c('adaptive.held', 'neutral.held')) %>%
+	mutate(Context='multi-species (opportunity costs)') 
+
+# compile data.frames
+results.SDF <- rbind.fill(list(single.spp.SDF, multi.spp.SDF, multi.spp.with.cost.SDF)) %>%
+	mutate(Metric=revalue(Metric, c('adaptive.held'='Adaptive variation', 'neutral.held'='Neutral variation'))) %>%
+	mutate(Prioritisation = factor(Prioritisation, levels=c('Amount','Surrogate','Genetic'))) %>%
+	mutate(Prioritisation.Metric.Context=interaction(Prioritisation,Metric,Context))
+
+## main analysis
+# model
+full.model.GLMM <- glmer(value ~ Prioritisation.Metric.Context + (1 | Species), family='binomial', data=results.SDF, nAGQ=20)
+null.model.GLMM <- glmer(value ~ 1 + (1 | Species), family='binomial', data=results.SDF, nAGQ=20)
+
+## post-hoc
+full.model.GLHT <- summary(
+	glht(full.model.GLMM,
+		linfct=mcp(Prioritisation.Metric.Context='Tukey')),
+	adjusted('bonferroni'))
+ 
+## save .rda
+save.session('results/.cache/09-statistical-analysis.rda')

### 2016-02-02 R/analysis/10-exports.R

@@ -1,0 +1,7 @@
+## load .rda
+session::restore.session('results/.cache/09-statistical-analysis.rda')
+
+#### save results
+session::save.session('results/results.rda')
+
+

### 2016-02-03 article/article.Rmd
* fix readmes * add in surrogacy tests * progress on article * added in figure in figure.Rmd

@@
 
-Our overarching aim is to investigate the effectiveness of environmental and geographic surrogates for adaptive neutral genetic variation in the context of conservation planning.
-First, we aim to test the effectiveness of these surrogates under the simplest of conditions by using a set of single-species prioritizations.
-Second, we aim to test the effectiveness of these surrogates in a more realistic multi-species context, and see how they compare to the single-species prioritizations.
-Third, we aim to understand the relationship between the proportion of surrogate variation secured in a multi-species prioritization and the proportion of genetic variation secured in a prioritization.
-If these surrogates can be shown to markedly improve the proportion of genetic variation secured in a prioritization, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.
+Our overarching aim was to investigate the effectiveness of environmental and geographic surrogates for adaptive neutral genetic variation in the context of conservation planning.
+First, we aimed to understand the relationship between the surrogate targets used to generate a prioritization and the proportion of genetic variation secured in the prioritisation.
+Second, we aimed to determine if the addition of surrogate-based targets in conservation planning exercises would result in prioritizations that secured substantially more genetic variation.
+If these surrogatess can be shown to improve the proportion of genetic variation secured in a prioritization, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.
 
@@
 We calculated the average value for each PC layer in each grid cell, and used these values to characterize the main themes of climatic variation among the grid cells (Figure S2).
-We used the total human population size inside each planning unit to represent its opportunity cost (obtained at 1 km resolution from the Global Rural-Urban Mapping Project; GRUMP V1; @rXXX).
+To examine if the influence of opportunity cost, we calculated the total human population density inside each planning unit (obtained at 1 km resolution from the Global Rural-Urban Mapping Project; GRUMP V1; @r472).
 
@@
 
-To understand how prioritizations generated using surrogate-based targets might perform in conservation planning scenarios, we generated three sets of prioritizations: a set of single-species prioritizations, a set of multi-species prioritizations, and a set of multi-species prioritizations that accommodated opportunity cost.
-For each of these sets, we generated prioritizations using only amount-based targets, prioritizations using both amount- and surrogate-based targets, and prioritizations using amount- and genetic-based targets.
-All amount-based targets were set as `r (rapr.params.LST[[MODE]]$amount.target*100)` % to ensure an adequate proportion of habitat was secured for each species [@rXXX].
-All surrogate- and genetic-based targets were set as `r (rapr.params.LST[[MODE]]$surrogate.target*100)`.
-For each set, for each combination of targets, we identified the `r rapr.params.LST[[MODE]]$single.species.surrogate.replicates` solutions nearest to optimality for replication.
-Except for the single-species amount-only prioritizations for which we generated `r rapr.params.LST[[MODE]]$single.species.amount.replicates` replicates because they were associated with many optimal solutions with varying levels of effectiveness.
+To understand the relationship between the surrogate-based targets used to generate a prioritizations and the amount of genetic variation secured, we generated a collection of single-species prioritizations.
+For each species, we generated two sets of prioritizations using increasing environmental and geographic-based targets (respectively).
+Prioritizations were generated using surrogate-based targets ranging from `r (min(rapr.params.LST[[MODE]]$correlation.surrogate.targets)*100)` % and `r (max(rapr.params.LST[[MODE]]$correlation.surrogate.targets)*100)` % in increments of `r diff(rapr.params.LST[[MODE]]$correlation.surrogate.targets[1:2])*100` %.
+None of these prioritizations contained any amount-based targets.
 
-To understand the relationship between the surrogate-based targets and the proportion of genetic variation secured, we generated two sets of multi-species prioritizations with increasing surrogate-based targets between `r (min(rapr.params.LST[[MODE]]$correlation.surrogate.targets)*100)` % and `r (max(rapr.params.LST[[MODE]]$correlation.surrogate.targets)*100)` % in increments of `r diff(rapr.params.LST[[MODE]]$correlation.surrogate.targets[1:2])*100` %.
-The first set of prioritizations contained prioritizations with varying environmental surrogate-based targets with no specified geographic targets.
-The second contained prioritizations with varying geographic surrogate-based targets and no environmental targets.
-None of these prioritizations used amount-based targets.
+To determine if the use of surrogate-based targets can actually improve prioritizations in real-world conservation planning scenarios, we used three scenarios with increasing levels of complexity.
+To assess the benefits of surrogate-based targets under the simplest of circumstances, our first scenario involved generating single-species prioritizations for each species.
+Most large-scale prioritizations involve a comprehensive set of species, and in turn, may secure a greater proportion of intra-specific genetic variation.
+Our second scenario involved generating multi-species prioritizations.
+Finally, most large-scale prioritizations also explicitly consider opportunity cost to ensure prioritizations are cost-effective.
+Our third scenario involved generating multi-species prioritizations that satisfied targets for minimal opportunity cost.
+
+For each scenario, we generated three types of prioritizations.
+We generated prioritizations using only amount-based, both amount- and surrogate-based targets, and both amount- and genetic-based targets.
+The purely amount-based prioritizations represented those typically generated in conservation planning exercises and serve as our baseline.
+We used `r (rapr.params.LST[[MODE]]$amount.target*100)` % amount-based targets in all prioritizations to ensure an adequate proportion of habitat was secured for each species [@rXXX].
+We set surrogate- and genetic-based targets as `r (rapr.params.LST[[MODE]]$surrogate.target*100)` % to ensure that most of the repsective variation was secured for each species.
+To accomodate variation in the high number of optimal solutions for the single-species amount-based prioritizations, we generated `r (rapr.parmas.LST[[MODE]]$amount.replicates)` replicate prioritizations.
+Since all other contexts were likely to be associated with a single optimal solution, we used only a single replicate.
 
 ## Statistical analysis
-We tested for differences in the effectiveness of prioritizations representing typical scenarios using generalized linear mixed-effects models [GLMMs; using the $\texttt{lme4}$ R package; @r469].
-We fit two full models to the proportion of genomic variation secured in the single- and multi-species prioritizations using a logit link.
-Both models contained a predictor variable indicating if the prioritization was generated using amount-only, surrogate, or genetic targets, a second predictor variable indicating if the genetic variation was adaptive or neutral, and the interaction between these variables.
-Additionally, both models contained an intercept-only random effect to ensure that comparisons were made within species.
-Each full model was subject to a step-wise term deletion routine using likelihood ratio tests (LRT) to assess term significance (terms were retained when $P < 0.05$).
+We investigated the relationship between surrogate-based targets and the genetic variation secured in a prioritization using nonlinear mixed-effects models [NLMMs; using the $\texttt{lme4}$ R package; @r469].
+A model was fit to the genetic variation secured using a negative exponential formulation to estimate the rate of the accumulation of genetic variation ($y = a - (a - 0) \times exp(-r \times x)$; where $a$ is the asymptote and $r$ is the accumulate rate).
+An interaction term was included to estimate parameters separately for adaptive and neutral variation.
+The models accommodated species-level variation using random effects for both parameters.
+
+We tested for differences in the effectiveness of prioritizations generated under different scenarios and targets using generalized linear mixed-effects models [GLMMs; using the $\texttt{lme4}$ R package].
+We fit a full model to the proportion of genomic variation secured in all prioritizations using a logit link.
+Both models contained variables indicating the scenario used to generate the prioritization, the targets used to generate the prioritizations, and the type of genetic variation measured.
+The model was subject to a step-wise term deletion routine using likelihood ratio tests (LRT) to assess term significance (terms were retained when $P < 0.05$).
+All models contained an intercept-only random effect to ensure that comparisons were made within species.
 Finally, we conducted a post-hoc analysis using Tukey contrasts with a Bonferroni correction [using the $\texttt{multcomp}$ R package; @r470].
-
-We investigated the relationship between surrogate-based targets and the genetic variation secured in a prioritization using nonlinear mixed-effects models (NLMMs; using the $\texttt{lme4}$ R package).
-Two models were fit to the adaptive and neutral genetic variation secured using a negative exponential formulation to estimate the rate of the accumulation of genetic variation ($y = a - (a - 0) \times exp(-r \times x)$; where $a$ is the asymptote and $r$ is the accumulate rate).
-Both models accommodated species-level random effects for the asymptote ($a$) and rate ($r$) parameters.
 

### 2016-02-03 article/figures.Rmd
* fix readmes * add in surrogacy tests * progress on article * added in figure in figure.Rmd

@@
 ## Figures
+
+```{r, echo=FALSE, fig.width=8, fig.height=3, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations.
+Panel (a) shows the relationship between environmental surrogates and adaptive genetic variation.
+Panel (b) shows that for geographic surrogates and neutral genetic variation.
+Transparent gray lines denote trends for each species, and black lines show general trends across all species.'}
+### prepare data frames for plotting
+## environmental vs adaptive
+# random effects
+env.correlation.RPDF <- expand.grid(
+	Surrogate.type='Environmental variation',
+	Surrogate.target=seq(min(env.correlation.DF$Surrogate.target), max(env.correlation.DF$Surrogate.target), length.out=100),
+	Species=unique(env.correlation.DF$Species)
+)
+env.correlation.RPDF$a.ranef <- fixef(env.correlation.NLMM)[[1]] + ranef(env.correlation.NLMM)$Species$a[env.correlation.RPDF$Species]
+env.correlation.RPDF$r.ranef <- fixef(env.correlation.NLMM)[[2]] + ranef(env.correlation.NLMM)$Species$r[env.correlation.RPDF$Species]
+env.correlation.RPDF$fit <- env.correlation.RPDF$a.ranef - (env.correlation.RPDF$a.ranef - 0) * exp(- env.correlation.RPDF$r.ranef * env.correlation.RPDF$Surrogate.target)
+# fixed effects
+env.correlation.FPDF <- data.frame(Surrogate.target=seq(min(env.correlation.DF$Surrogate.target), max(env.correlation.DF$Surrogate.target), length.out=100))
+env.correlation.FPDF$fit <- fixef(env.correlation.NLMM)[[1]] - (fixef(env.correlation.NLMM)[[1]] - 0) * exp(- fixef(env.correlation.NLMM)[[2]] * env.correlation.FPDF$Surrogate.target)
+## geographic vs neutral
+# random effects
+geo.correlation.RPDF <- expand.grid(
+	Surrogate.type='Geographic variation',
+	Surrogate.target=seq(min(geo.correlation.DF$Surrogate.target), max(geo.correlation.DF$Surrogate.target), length.out=100),
+	Species=unique(geo.correlation.DF$Species)
+)
+geo.correlation.RPDF$a.ranef <- fixef(geo.correlation.NLMM)[[1]] + ranef(geo.correlation.NLMM)$Species$a[geo.correlation.RPDF$Species]
+geo.correlation.RPDF$r.ranef <- fixef(geo.correlation.NLMM)[[2]] + ranef(geo.correlation.NLMM)$Species$r[geo.correlation.RPDF$Species]
+geo.correlation.RPDF$fit <- geo.correlation.RPDF$a.ranef - (geo.correlation.RPDF$a.ranef - 0) * exp(- geo.correlation.RPDF$r.ranef * geo.correlation.RPDF$Surrogate.target)
+# fixed effects
+geo.correlation.FPDF <- data.frame(Surrogate.target=seq(min(geo.correlation.DF$Surrogate.target), max(geo.correlation.DF$Surrogate.target), length.out=100))
+geo.correlation.FPDF$fit <- fixef(geo.correlation.NLMM)[[1]] - (fixef(geo.correlation.NLMM)[[1]] - 0) * exp(- fixef(geo.correlation.NLMM)[[2]] * geo.correlation.FPDF$Surrogate.target)
+
+### make plots
+suppressWarnings({
+	p1 <- ggplot() +
+		geom_point(aes(x=Surrogate.target, y=adaptive.held), data=env.correlation.DF) +
+		geom_line(aes(x=Surrogate.target,y=fit,group=Species), data=env.correlation.RPDF, alpha=0.3) +
+		geom_line(aes(x=Surrogate.target,y=fit), data=env.correlation.FPDF, size=1.5) +
+		geom_text(aes(x=0.6, y=1.03, label='(a)')) +
+		xlab('Environmental variation secured (%)') +
+		ylab('Adaptive genetic\nvariation secured (%)') +
+		theme_classic()
+	p2 <- ggplot() +
+		geom_point(aes(x=Surrogate.target, y=neutral.held), data=geo.correlation.DF) +
+		geom_line(aes(x=Surrogate.target,y=fit,group=Species), data=geo.correlation.RPDF, alpha=0.3) +
+		geom_line(aes(x=Surrogate.target,y=fit), data=geo.correlation.FPDF, size=1.5) +
+		geom_text(aes(x=0.6, y=1.03, label='(b)')) +
+		xlab('Geographic variation secured (%)') +
+		ylab('Neutral genetic\nvariation secured (%)') +
+		theme_classic()
+	grid.arrange(p1, p2, nrow=1)
+})
+```
 

### 2016-02-03 R/analysis/06-surrogacy-prioritizations.R

@@ -1,0 +1,79 @@
+## load .rda
+session::restore.session('results/.cache/05-format-data-for-prioritisations.rda')
+
+## load parameters
+rapr.params.LST <- parseTOML('parameters/rapr.toml')
+gurobi.params.LST <- parseTOML('parameters/gurobi.toml')
+
+## correlation analysis
+# generate prioritistions
+cat("starting environmental surrogacy prioritizations\n")
+env.correlation.prioritisations <- llply(
+	seq_along(unique(spp.samples.DF$species)),
+	function(i) {
+		llply(
+			rapr.params.LST[[MODE]]$surrogacy.analysis$surrogate.target,
+			function(j) {
+				species.prioritisation(
+					x=spp.subset(ru, i),
+					amount.targets=0,
+					env.surrogate.targets=j,
+					geo.surrogate.targets=0,
+					adaptive.genetic.targets=0,
+					neutral.genetic.targets=0,
+					Threads=gurobi.params.LST[[MODE]]$Threads,
+					MIPGap=gurobi.params.LST[[MODE]]$MIPGap
+				)
+			}
+		)
+	},
+	.progress='text'
+)
+
+cat("starting geographic surrogacy prioritizations\n")
+geo.correlation.prioritisations <- llply(
+	seq_along(unique(spp.samples.DF$species)),
+	function(i) {
+		llply(
+			rapr.params.LST[[MODE]]$surrogacy.analysis$surrogate.target,
+			function(j) {
+				species.prioritisation(
+					x=spp.subset(ru, i),
+					amount.targets=0,
+					env.surrogate.targets=0,
+					geo.surrogate.targets=j,
+					adaptive.genetic.targets=0,
+					neutral.genetic.targets=0,
+					Threads=gurobi.params.LST[[MODE]]$Threads,
+					MIPGap=gurobi.params.LST[[MODE]]$MIPGap
+				)
+			}
+		)
+	},
+	.progress='text'
+)
+
+# extract results
+env.correlation.DF <- ldply(
+	seq_along(env.correlation.prioritisations), 
+	function(i) {
+		mutate(
+			ldply(env.correlation.prioritisations[[i]], extractResults),
+			Surrogate.target=rapr.params.LST[[MODE]]$surrogacy.analysis$surrogate.target
+		)
+	}
+)
+
+geo.correlation.DF <- ldply(
+	seq_along(geo.correlation.prioritisations), 
+	function(i) {
+		mutate(
[…]

### 2016-02-03 R/analysis/08-multi-species-prioritisations-no-cost.R

@@ -1,5 +1,5 @@
 ## load .rda
-session::restore.session('results/.cache/06-single-species-prioritisations.rda')
+session::restore.session('results/.cache/07-single-species-prioritisations.rda')
 
 ## multispecies analysis
 # make prioritisations
@@ -41,4 +41,4 @@ multi.spp.DF <- ldply(
 )
 
 ## save .rda
-save.session('results/.cache/07-multi-species-prioritisations-no-cost.rda')
+save.session('results/.cache/08-multi-species-prioritisations-no-cost.rda')

### 2016-02-05 article/article.Rmd
+ replace adegenet with mclust

@@
 nutrient cycling and water purification).
-Additionally, protected areas set the stage for direct management actions to enhance species recovery (eg.
+Protected areas also set the stage for direct management actions to enhance species recovery (eg.
 habitat rehabilitation, translocation, and invasive species management).
-Since the resources available for conservation are limited, protected areas must be cost-effective.
-As a consequence, plans for protected areas (prioritizations) are often generated using decision support tools [eg.
-$\texttt{MARXAN}$; @r22; $\texttt{Zonation}$; @r429].
-These tools often formulate the benefit that a given species receives from a solution based on the proportion of its range that is protected.
-However, the true benefit that a species might receive is dependent on many other additional factors.
+But the resources available for conservation are limited, and so protected areas must be cost-effective.
+As a consequence, decision support tools [eg.
+$\texttt{MARXAN}$; @r22; $\texttt{Zonation}$; @r429] are used to identify cost-effective plans for protected areas (prioritizations).
+These tools typically estimate the benefit that a given species receives from a given prioritization based on the proportion of its range that is secured (eg.
+using targets).
+However, the true benefit of a given prioritization is dependent on many additional factors that are not used to inform the reserve selection process.
 
@@
 This variation can be classified as either adaptive or neutral [reviewed in @r????].
-Adaptive genetic variation is associated with loci currently under selection, whereas neutral genetic variation is associated with loci that are not.
+Adaptive genetic variation is associated with loci currently under selection.
 By preserving existing patterns of adaptive variation, protected areas can ensure that populations with particularly beneficial adaptations are not lost [@r3].
-In contrast, neutral variation reflects the evolutionary history of different populations.
+In contrast, neutral genetic variation is associated with loci that are not under natural selection.
+Neutral variation often reflects the evolutionary history of different populations.
 By preserving existing patterns of neutral variation, protected areas can secure different genetic lineages and avoid the adverse effects of low genetic diversity [eg.
 inbreeding depression; @r67].
-However, genetic data requires considerable investment to obtain and analyze, and as a consequence, conservation planners do not typically use genetic data to inform prioritizations [@r61].
+However, genetic data requires considerable investment to obtain and analyze.
+As a consequence, conservation planners do not typically use genetic data to inform prioritizations [@r61].
 Thus many prioritizations may fail to capture species' intra-specific genetic variation, and ultimately fail to deliver conservation outcomes.
 
-Recently, conservation scientists have begun investigating surrogates to generate more effective prioritizations without the overheads associated with genetic data.
-Since adaptive genetic variation is driven by selection pressures, environmental variables have been explored as potential surrogates for genetic variation [eg.
-@r247; @r2].
-On the other hand, neutral variation arises from a reduction in gene flow, and so variables that predict reduction in gene flow between areas have been explored.
-For instance, the isolation by distance theory [IBD; @r259] predicts that populations further apart will have lower gene flow, and in turn have greater neutral variation among them.
-The use of distance-based surrogates has been found to improve prioritizations in an island network [@r200].
-However the effectiveness of distance-based surrogates remains untested in contiguous systems conventionally used in conservation planning where gene flow is complicated by additional factors (eg.
-elevation in terrestrial systems).
+In light of this problem, conservation scientists have been investigating surrogates for genetic variation to generate better prioritizations--but without the overheads associated with genetic data.
+Since adaptive genetic variation is ultimately driven by selection pressures, conservation planners have used the distribution of likely selection pressures.
+For instance, environmental variation such as local habitat type [@r247] and climate [@r2] have been used to guide the reserve selection process.
+While such variables have been used to detect and verify adaptive genetic variation [eg.
+@r475; @r455], the effectiveness of prioritizations based on such surrogates remains unverified.
+
+On the other hand, neutral genetic variation arises a reduction in gene flow between individuals.
+By modeling the connectivity between areas, a function of gene flow, conservation scientists have explored the effectiveness of surrogates for neutral genetic variation.
+For instance, the isolation by distance theory [IBD; @r259] predicts that populations further apart will have reduced connectivity, and in turn exchange less genetic material.
+The use of geographic distance-based surrogates has been found to improve prioritizations in an island network [@r200].
+However this remains untested in spatially contiguous systems where connectivity is complicated by additional factors [eg.
+terrain; eg.
+@r474]--the systems commonly featured in real-world conservation planning exercises.
 Before conservation planners can reliably use surrogates for genetic data in conservation planning, the effectiveness of these surrogates needs to be verified.
@@
 Our overarching aim was to investigate the effectiveness of environmental and geographic surrogates for adaptive neutral genetic variation in the context of conservation planning.
-First, we aimed to understand the relationship between the surrogate targets used to generate a prioritization and the proportion of genetic variation secured in the prioritisation.
-Second, we aimed to determine if the addition of surrogate-based targets in conservation planning exercises would result in prioritizations that secured substantially more genetic variation.
-If these surrogatess can be shown to improve the proportion of genetic variation secured in a prioritization, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.
+First, we aimed to understand the relationship between the surrogate targets used to generate a prioritization and the proportion of genetic variation secured in the prioritization.
+Second, we aimed to determine if the addition of surrogate-based targets in more realistic conservation planning scenarios would result in prioritizations that secured substantially more genetic variation.
+If these surrogates can be shown to improve the intra-specific genetic variation captured in a prioritization, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.
 
@@
 We used these grid cells as planning units to develop prioritizations to investigate the effectiveness of surrogates for the (AFLP) genetic variation.
-Most of our analyses described hereafter were performed using the \texttt{R} software package [version `r paste0(version$major, '.', version$minor)`; @r192].
+The majority of our analyses described hereafter were performed using the \texttt{R} software package [version `r paste0(version$major, '.', version$minor)`; @r192].
 
@@
 These layers and the grid cells were projected into an equal-area coordinate system ($\texttt{ESRI:102014}$).
-The bioclimatic variables were clipped to the extent of the grid and subject to a principle components analysis (PCA; Table S1).
+The bioclimatic variables were clipped to the extent of the grid and subject to a principle components analysis (PCA; using ArcMap version 10.2.2; Table S1).
 The first two principle components (PCs) cumulatively explained `r round(pca.DF[[4]][2],1)` % of the variation and were used to generate two new layers.
@@
 The basic premise underpinning these analyses is that neutral loci are expected to exhibit a specific level of variation among populations ($F_{st}$), and loci that deviate from this expectation should be under selection.
-The advantage of these methods, in contrast with environmental association analyses [eg.
-r@XXX], is that they do not use environmental data to identify loci under selection, which would have biased our analysis.
+The advantage of these methods--in contrast with environmental association analyses [eg.
+r@475]--is that they do not use environmental data to identify loci under selection, which would have biased our analysis.
 In an earlier study, Bothwell _et al._ [-@r455] applied such outlier detection methods to one of the species in this dataset (_Getiana nivalis_).
@@
 The purpose of this step was to avoid misclassifying loci as adaptive.
-We used discriminant analyses of principle components [DAPC; @r467] to assign individuals to populations [implemented in $\texttt{adegenet}$ R package; @r468].
-Briefly, for each species, we transformed the binary loci data using a principle components analysis (PCA) and obtained the principle components required to explain at least `r adegenet.params.LST[[MODE]]$perc.pca` % variation.
-We subject these components to multiple \textit{k}-means clustering models with varying numbers of assumed populations (\textit{k} = 1--`r adegenet.params.LST[[MODE]]$max.n.clust`; `r adegenet.params.LST[[MODE]]$n.iter` iterations; `r adegenet.params.LST[[MODE]]$n.start` random starts).
-We then  applied Ward's clustering method to the Bayesian Information Criterion (BIC) statistics associated with each model to partition the models into two groups, and selected the model before the first group switch (as recommended in the \texttt{adegenet} manual; Figure S4).
-The populations memberships associated with these models were used for subsequent analysis (Figure S5).
+We used Gaussian mixture models to assign individuals to populations [implemented in $\texttt{mclust}$ R package; @r477; @r467].
+Briefly, for each species, we fit a suite of models to estimate the probability that individuals belonged different populations (using default priors and covariance structures).
+For each parametrization, we fit a set of models assuming varying numbers of total populations (\textit{G} = 1--`max(r mclust.params.LST[[MODE]]$G)`).
+We identified the most parsimonious models (lowest Bayesian Information Criterion $BIC$ statistic; Figures S4), and used these to group individuals into populations (Figure S5).
 
-We identified loci under selection for each species (Table S2).
-For each species, we ran $\texttt{Bayescan}$ [version 2.1; @r456] using the population memberships identified in the previous analysis (1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations).
-We omitted loci where the global frequency of the minor allele was $\geq 0.05$ to avoid false-positives (following guidelines in the $\texttt{Bayescan}$ manual).
-To ensure convergence, we ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate [FDR; $\leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].
+We identified loci under selection.
+For each species, we ran $\texttt{Bayescan}$ [version 2.1; @r456] using the population memberships identified in the previous analysis (1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations; Table S2).
+To avoid false-positive, we omitted loci where the global frequency of the minor allele was $\geq 0.05$ (following guidelines in the $\texttt{Bayescan}$ manual).
+Additionally, to prevent individuals with unclear ancestry from biasing this analysis, we omitted individuals if there was considerable uncertainty in their population membership (maximum probability $\leq `r mclust.params.LST[[MODE]]$probability.threshold`$ % ).
+We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate to ensure convergence [FDR $\leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].
 
@@
 For each species, for each type of genetic variation detected, we used Gower distances [@r464] to express the genomic differences between individuals [using the $\texttt{cluster}$ R package to accommodate sparsity; @r457].
-These distances were subject to multiple non-metric multi-dimensional scaling analyses [NMDS; implemented in the $\texttt{vegan}$ R package; @r458] using the minimum number of dimensions required to achieve an adequate stress value  ($\leq$ `r mds.params.LST[[MODE]]$max.stress`; `r mds.params.LST[[MODE]]$trymax` random starts for all models; Table S3).
-Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the samples in each grid cell to characterize their the genetic variation of individuals in the cells.
+These distances were subject to multiple non-metric multi-dimensional scaling analyses [NMDS; implemented in the $\texttt{vegan}$ R package; @r458] using the minimum number of dimensions required to achieve an adequate stress value  ($\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for all models; Table S3).
+Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the samples in each grid cell to summarize the typical genetic characteristics of individuals in the cells.
 
@@
 Second, we used geographic distances as surrogates for connectivity between planning units.
-Although connectivity between areas can be modeled using isolation by resistance measures derived from landscape data [eg.
-elevation; reviewed in @rXX], they are unlikely to be of use to conservation planners since they require genetic data to be tuned.
+Although connectivity between areas can be modeled using isolation by resistance measures [@r473] derived from landscape data [eg.
+@r474], they are unlikely to be of use in broad-scale conservation exercises since they need to be tuned using species-specific parameters.
 Third, we used AFLP data to characterize the main gradients of genetic variation.
-Whilst next-generation sequencing provides higher resolution genetic information than AFLP data [reviewed in @rXXX], our methodology would still have utilised only the main gradients of the genetic variation to generate prioritizations in a feasible amount of time.
+Whilst next-generation sequencing provides higher resolution genetic information than AFLP data [reviewed in @rXXX], our methodology would still have utilized only the main gradients of the genetic variation to generate prioritizations in a feasible amount of time.
 Overall, the limitations associated with this study are relatively minor.

### 2016-02-05 article/supporting_information.Rmd
+ replace adegenet with mclust

@@
 	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
-		fill='grey20', color='grey80') +
+		fill='grey85', color='grey70') +
 	geom_polygon(data=spp.grid.FPLY, aes(x=long, y=lat, 
@@
 	theme(axis.ticks=element_blank(), axis.text=element_blank(),
-		axis.line=element_blank()) +
+		axis.line=element_blank(), 
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white', face='italic'),
+		axis.title=element_blank()
+	) +
 	coord_cartesian(
-		xlim=buffered.range(grid.FPLY$long, 0.05),
-		ylim=buffered.range(grid.FPLY$lat, 0.05)
+		xlim=buffered.range(grid.FPLY$long, 0),
+		ylim=buffered.range(grid.FPLY$lat, 0)
 	) +
-	xlab('') +
-	ylab('') +
 	facet_wrap(~ species, ncol=4)
@@
 	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
-		fill='grey20', color='grey80') +
+		fill='grey85', color='grey70') +
 	geom_polygon(data=spFortify(grid.PLY), aes(x=long, y=lat, 
-		group=group, fill=Species_richness), alpha=0.8, color='grey10') +
-	guides(fill=guide_legend(title='Count (#)')) +
+		group=group, fill=Species_richness), color='grey10') +
+	scale_fill_gradientn(name='Count (#)',
+		colors=brewer.pal(9,'PuBu'),
+		guide=guide_colorbar(ticks=element_line(color='black'), border=element_line(color='black'))
+	) +
 	theme_classic() +
 	theme(axis.ticks=element_blank(), axis.text=element_blank(),
-		axis.line=element_blank()) +
+		axis.line=element_blank(), axis.title=element_blank(),
+		panel.border=element_rect(color='black', fill=NA, size=1)
+	) +
 	coord_cartesian(
-		xlim=buffered.range(grid.FPLY$long, 0.05),
-		ylim=buffered.range(grid.FPLY$lat, 0.05)
+		xlim=buffered.range(grid.FPLY$long, 0),
+		ylim=buffered.range(grid.FPLY$lat, 0)
 	) + 
-	xlab('') +
-	ylab('') +
-	ggtitle('Species richness')
+	ggtitle('Species Richness')
+```
+
+```{r, echo=FALSE, fig.width=4, fig.height=3.5, fig.cap="Opportunity cost.
+Squares denote planning units.
+The color of each planning unit represents its opportunity cost (computed as total population density)"}
+# plot opportunity costs
+ggplot() +
+	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
+		fill='grey85', color='grey70') +
+	geom_polygon(data=spFortify(grid.PLY), aes(x=long, y=lat, 
+		group=group, fill=pop.density), color='grey10') +
+	scale_fill_gradientn(name='Cost',
+		colors=brewer.pal(9, 'PuBu'),
+		guide=guide_colorbar(ticks=element_line(color='black'), border=element_line(color='black'))
+	) +
+	theme_classic() +
+	theme(axis.ticks=element_blank(), axis.text=element_blank(),
+		axis.line=element_blank(), axis.title=element_blank(),
+		panel.border=element_rect(color='black', fill=NA, size=1)
+	) +
+	coord_cartesian(
+		xlim=buffered.range(grid.FPLY$long, 0),
+		ylim=buffered.range(grid.FPLY$lat, 0)
+	) +
+	ggtitle('Opportunity cost')
 ```
@@
 		llply(grep('^env\\_.*$', names(grid.DF), value=TRUE), function(x) {
+			grid.FPLY$plot_title <- paste0('PC ', substr(x, nchar(x), nchar(x)))
 			ggplot() +
 				geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
-					fill='grey20', color='grey80') +
+					fill='grey85', color='grey70') +
 				geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
-					group='group', fill=x),
-					alpha=0.8, color='grey10') +
-				guides(fill=guide_legend(title=' ')) +
+					group='group', fill=x), color='grey10') +
 				theme_classic() +
 				theme(axis.ticks=element_blank(), axis.text=element_blank(),
-					plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank()) +
+					plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(), 
+					axis.title=element_blank(), 
+					panel.border=element_rect(color='black', fill=NA, size=1),
+					strip.background = element_rect(fill='grey20'),
+					strip.text = element_text(color='white')
+				) +
 				coord_cartesian(
-					xlim=buffered.range(grid.FPLY$long, 0.05),
-					ylim=buffered.range(grid.FPLY$lat, 0.05)
+					xlim=buffered.range(grid.FPLY$long, 0),
+					ylim=buffered.range(grid.FPLY$lat, 0)
 				) +
-				xlab('') +
-				ylab('') +
-				ggtitle(paste0('PC ', substr(x, nchar(x), nchar(x))))
+				scale_fill_gradientn(name='',
+					colors=brewer.pal(9, 'RdYlBu'),
+					guide=guide_colorbar(ticks=element_line(color='black'), border=element_line(color='black'))
+				) +
+				facet_wrap(~ plot_title)
 		}),
@@
 Each panel denotes a different species.
-Asterisks indicate the number of populations with the most support (see text for methodological details).'}
+Each line denotes a different model parametrization.
+Asterisks indicate the model and number of populations with the most support.'}
 # format data
-bic.buff <- laply(spp.Adegenet.LST, `[[`, 'Kstat') %>%
-	c() %>% range() %>% diff %>% `*`(0.05)
 bic.DF <- ldply(
@@
 	.fun=function(i) {
-		data.frame(
-			k=as.numeric(gsub('K=', '', names(spp.Adegenet.LST[[i]]$Kstat), fixed=TRUE)),
-			stat=unname(spp.Adegenet.LST[[i]]$Kstat),
-			best_k=as.numeric(gsub('K=', '', names(spp.Adegenet.LST[[i]]$stat), fixed=TRUE)),
-			best_stat=unname(spp.Adegenet.LST[[i]]$stat) + bic.buff,
-			asterisk_col=c('black',rep('transparent', length(spp.Adegenet.LST[[i]]$Kstat)-1)),
-			Species=gsub('_', ' ', unique(spp.samples.DF$species)[i], fixed=TRUE)
+		curr.MTX <- spp.Mclust.LST[[i]]$BIC
+		attr(curr.MTX, 'dim') <- c(nrow(curr.MTX), ncol(curr.MTX))
+		attr(curr.MTX, 'dimnames') <- list(
+			seq_len(nrow(curr.MTX)),
+			attr(curr.MTX, 'modelNames')
+		)
+		class(curr.MTX) <- 'matrix'
+		curr.DF <- curr.MTX %>% as.data.frame %>% 
+			mutate(G=attr(curr.MTX, 'G')) %>% 
+			gather(Model,BIC,-G)
+		curr.DF$BIC=curr.DF$BIC*-1
+		return(
+			mutate(
+				curr.DF,
+				Species=gsub('_', ' ', unique(spp.samples.DF$species)[i], fixed=TRUE),
+				best_G=curr.DF[which.min(curr.DF$BIC),'G'][[1]],
+				best_stat=min(curr.DF$BIC,na.rm=TRUE),
+				asterisk_col=c('black', rep(NA_character_, length(Species)-1))
+			)
 		)
@@
 	ggplot(data=bic.DF) +
-		geom_point(aes(y=stat, x=k)) +
-		geom_point(aes(y=best_stat, x=best_k), size=4, shape=8) +
-		geom_line(aes(y=stat, x=k)) +
+		geom_point(aes(y=BIC, x=G, color=Model)) +
+		geom_line(aes(y=BIC, x=G, color=Model)) +
+		geom_point(aes(y=best_stat, x=best_G), size=4, shape=8) +
 		theme_classic() +
+		scale_y_reverse()+
+		theme(
+			strip.background = element_rect(fill='grey20'),
+			strip.text = element_text(color='white', face='italic', size=20)
+		) +
 		xlab('Number of populations') +
 		ylab('BIC') +
-		facet_wrap(~ Species, ncol=4)
+		facet_wrap(~ Species, ncol=4, scales='free_y')
 })
@@
 Note that points have been slightly jittered so that individuals in close proximity are visible.'}
-# define function
-plotLST <- llply(
-	# seq_along(unique(spp.samples.DF$species)),
-	seq_along(unique(spp.samples.DF$species)[-23]), # omit spp 23 from current version
-	.fun=function(i) {
-		# init
+# prepare data
+pop.DF <- llply(
+	seq_along(unique(spp.samples.DF$species)),
+	function(i) {
 		curr.pts <- filter(
@@
 		) %>% mutate(
+			species=gsub('_', ' ', unique(spp.samples.DF$species)[i], fixed=TRUE),
 			population=spp.Adegenet.LST[[i]]$grp
 		)
-		# create plot
+	}
+) %>% do.call(what=rbind)
+# make plot
 ggplot() +
 	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
-				fill='grey20', color='grey80') +
+		fill='grey85', color='grey75') +
 	geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
 		group='group'), fill='transparent', color='grey10') +
-			geom_point(data=curr.pts, aes_string(x='sample.longitude', y='sample.latitude', color='population'),
-				alpha=0.3, size=3, position='jitter') +
+	geom_point(data=pop.DF, aes_string(x='sample.longitude', y='sample.latitude', 
+		color='population'), alpha=0.8, size=3, position='jitter') +
 	guides(fill=guide_legend(title=''), color=guide_legend(title='')) +
@@
 	theme(axis.ticks=element_blank(), axis.text=element_blank(),
-				plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank()) +
+		plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(), 
+		axis.title=element_blank(), 
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white', face='italic', size=20),
+		legend.text= element_text(size=20)
+	) +
 	coord_cartesian(
-				xlim=buffered.range(grid.FPLY$long, 0.05),
-				ylim=buffered.range(grid.FPLY$lat, 0.05)
+		xlim=buffered.range(grid.FPLY$long, 0),
+		ylim=buffered.range(grid.FPLY$lat, 0)
 	) +
-			xlab('') +
-			ylab('') +
-			ggtitle(gsub('_', ' ', unique(spp.samples.DF$species)[i], fixed=TRUE))
-	}
-)
-# render plot
-do.call(
-	grid.arrange,
-	append(
-		plotLST,
-		list(ncol=4)
-	)
-)
+	facet_wrap(~ species, ncol=4)
 ```
@@
 plot.spp.mds <- function(i) {
-	# define function
-	make.mds.plot <- function(j,k) {
+	## prepare data
+	# get vector of adaptive and neutral column names
+	adapt.col <- grep(
+		paste0(unique(spp.samples.DF$species)[i], '_adaptive_d'),
+		names(grid.FPLY), value=TRUE, fixed=TRUE
+	)
+	neutral.col <- grep(
+		paste0(unique(spp.samples.DF$species)[i], '_neutral_d'),
+		names(grid.FPLY), value=TRUE, fixed=TRUE
+	)
+	curr.FPLY <- grid.FPLY[,
+		c(adapt.col, neutral.col,
+		'long','lat','group'),drop=FALSE]
+	curr.FPLY <- curr.FPLY %>% gather(d, value, contains('adaptive'), contains('neutral')) %>% mutate(
+		d=gsub(paste0(unique(spp.samples.DF$species)[i], '_'), '', d, fixed=TRUE),
+		d=gsub('_', ' ', d, fixed=TRUE)
+	)
+	for (j in seq_len(length(adapt.col)+length(neutral.col)))
+		curr.FPLY[['d']] <- gsub(as.character(j), paste0('(', j, ')'), curr.FPLY[['d']], fixed=TRUE)
+	curr.FPLY <- curr.FPLY %>% mutate(d=factor(gsub(' d(', ' (', d, fixed=TRUE)))
+	# make plot
 	ggplot() +
 		geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
-				fill='grey20', color='grey80') +
-			geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
-				group='group', fill=paste0(unique(spp.samples.DF$species)[i], '_', j, '_d',k)),
-				alpha=0.8, color='grey10') +
-			guides(fill=guide_legend(title=' ')) +
+			fill='grey85', color='grey70') +
+		geom_polygon(data=curr.FPLY, aes_string(x='long', y='lat', 
+			group='group', fill='value'), color='grey10') +
 		theme_classic() +
 		theme(axis.ticks=element_blank(), axis.text=element_blank(),
-				plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank()) +
+			plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(),
+			axis.title=element_blank(), strip.background = element_rect(fill='grey20'),
+			strip.text = element_text(color='white', size=20),
+			legend.text = element_text(size=20)
+		) +
 		coord_cartesian(
-				xlim=buffered.range(grid.FPLY$long, 0.05),
-				ylim=buffered.range(grid.FPLY$lat, 0.05)
+			xlim=buffered.range(grid.FPLY$long, 0),
+			ylim=buffered.range(grid.FPLY$lat, 0)
 		) +
-			xlab('') +
-			ylab('') +
-			ggtitle(paste0(j,' (',k,')'))
-	}
-	# init
-	adapt.col <- paste0(unique(spp.samples.DF$species)[i], '_adaptive_d1')
-	neutral.col <- paste0(unique(spp.samples.DF$species)[i], '_neutral_d1')
-	plotLST <- list()
-	# make adaptive plot
-	if (adapt.col %in% names(grid.FPLY)) {
-		for (k in seq_len(spp.mds.LST[[i]][['adaptive']]$ndim))
-			plotLST <- append(plotLST, list(make.mds.plot('adaptive', k)))
-	} 
-	# make neutral plot
-	if (neutral.col %in% names(grid.FPLY)) {
-		for (k in seq_len(spp.mds.LST[[i]][['neutral']]$ndim))
-			plotLST <- append(plotLST, list(make.mds.plot('neutral', k)))
-	}
-	# make plot
-	do.call(
-		grid.arrange,
-		append(
-			plotLST,
-			list(ncol=4)
-		)
-	)
+		scale_fill_gradientn(name='',
+			colors=brewer.pal(9, 'RdYlBu'),
+			guide=guide_colorbar(ticks=element_line(color='black'),
+				border=element_line(color='black'), barheight=unit(7, 'cm'))
+		) +
+		facet_wrap(~ d, ncol=4)
 }
@@
 	# exports
-	return(ceiling(n.plots/4) * 2.8)
+	return(ceiling(n.plots/4) * 3.5)
 }

### 2016-02-05 R/analysis/00-initialization.R

@@ -37,9 +37,10 @@ suppressMessages(library(english))
 suppressMessages(library(session))
 suppressMessages(library(maptools))
 suppressMessages(library(RcppTOML))
-suppressMessages(library(adegenet))
+suppressMessages(library(mclust))
 suppressMessages(library(lme4))
 suppressMessages(library(RColorBrewer))
+suppressMessages(library(mix))
 
 ## load github packages
 # devtools::install_github('paleo13/raspr')

### 2016-02-05 R/analysis/03-population-clustering-analysis.R

@@ -1,0 +1,54 @@
+## load .rda
+session::restore.session('results/.cache/02-surrogate-data.rda')
+
+## load parameters
+mclust.params.LST <- parseTOML('parameters/mclust.toml')
+
+### mclust analyses
+# run clustering analyses
+clust <- makeCluster(mclust.params.LST[[MODE]]$threads, type='SOCK')
+clusterEvalQ(clust, {library(structurer);library(mclust)})
+clusterExport(clust, c('mclust.params.LST','MODE', 'spp.StructureData.LST', 'spp.samples.DF'))
+registerDoParallel(clust)
+spp.Mclust.LST <- llply(
+	seq_along(spp.StructureData.LST),
+	.fun=function(i) {
+		# init
+		curr.MTX <- spp.StructureData.LST[[i]]@matrix
+		# impute missing data if apropriate
+		if (any(is.na(curr.MTX)))
+			curr.MTX <- imputeData(curr.MTX)
+		# run cluster analysis
+		return(
+			Mclust(
+				curr.MTX,
+				G=mclust.params.LST[[MODE]]$G
+			)
+		)
+	},
+	.parallel=TRUE
+)
+clust <- stopCluster(clust)
+
+# assign individuals to populations if they have probability to a single population above threshold
+spp.BayeScanData.sample.subset.LST <- llply(
+	seq_along(spp.BayeScanData.LST),
+	.fun=function(i) {
+		print(i)
+		# get population identities
+		probs <- spp.Mclust.LST[[i]]$z
+		ids <- apply(probs, 1, which.max)
+		ids[which(apply(probs, 1, function(x) {max(x) < mclust.params.LST[[MODE]]$probability.threshold}))] <- NA_integer_
+		validPos <- which(!is.na(ids))
+		# remove individuals below threshold
+		curr.spp <- spp.BayeScanData.LST[[i]]
+		curr.spp <- bayescanr:::sample.subset.BayeScanData(curr.spp, validPos)
+		curr.spp@labels <- filter(spp.samples.DF, species==unique(spp.samples.DF$species)[i])$cell[validPos]
+		curr.spp@populations <- as.character(ids[validPos])
+		return(curr.spp)
+	}
+)
+
+
+## load .rda
+save.session('results/.cache/03-population-clustering-analysis.rda')

### 2016-02-05 R/analysis/05-format-data-for-prioritisations.R

@@ -1,5 +1,5 @@
 ## load .rda
-session::restore.session('results/.cache/04-bayescan-analysis.rda')
+session::restore.session('results/.cache/04-identify-adaptive-loci.rda')
 
 # generate attribute spaces for geographic and environmental data
 surrogate.ASL <- llply(

### 2016-02-05 article/article.Rmd
+ replace adegenet with mclust (reverted from commit 08f8bcca6079915141bf3f8b3b3510f1d1a4edb9)

@@
 We used these grid cells as planning units to develop prioritizations to investigate the effectiveness of surrogates for the (AFLP) genetic variation.
-The majority of our analyses described hereafter were performed using the \texttt{R} software package [version `r paste0(version$major, '.', version$minor)`; @r192].
+Most of our analyses described hereafter were performed using the \texttt{R} software package [version `r paste0(version$major, '.', version$minor)`; @r192].
 
@@
 These layers and the grid cells were projected into an equal-area coordinate system ($\texttt{ESRI:102014}$).
-The bioclimatic variables were clipped to the extent of the grid and subject to a principle components analysis (PCA; using ArcMap version 10.2.2; Table S1).
+The bioclimatic variables were clipped to the extent of the grid and subject to a principle components analysis (PCA; Table S1).
 The first two principle components (PCs) cumulatively explained `r round(pca.DF[[4]][2],1)` % of the variation and were used to generate two new layers.
@@
 The purpose of this step was to avoid misclassifying loci as adaptive.
-We used Gaussian mixture models to assign individuals to populations [implemented in $\texttt{mclust}$ R package; @r477; @r467].
-Briefly, for each species, we fit a suite of models to estimate the probability that individuals belonged different populations (using default priors and covariance structures).
-For each parametrization, we fit a set of models assuming varying numbers of total populations (\textit{G} = 1--`max(r mclust.params.LST[[MODE]]$G)`).
-We identified the most parsimonious models (lowest Bayesian Information Criterion $BIC$ statistic; Figures S4), and used these to group individuals into populations (Figure S5).
+We used discriminant analyses of principle components [DAPC; @r467] to assign individuals to populations [implemented in $\texttt{adegenet}$ R package; @r468].
+Briefly, for each species, we transformed the binary loci data using a principle components analysis (PCA) and obtained the principle components required to explain at least `r adegenet.params.LST[[MODE]]$perc.pca` % variation.
+We subject these components to multiple \textit{k}-means clustering models with varying numbers of assumed populations (\textit{k} = 1--`r adegenet.params.LST[[MODE]]$max.n.clust`; `r adegenet.params.LST[[MODE]]$n.iter` iterations; `r adegenet.params.LST[[MODE]]$n.start` random starts).
+We then  applied Ward's clustering method to the Bayesian Information Criterion (BIC) statistics associated with each model to partition the models into two groups, and selected the model before the first group switch (as recommended in the \texttt{adegenet} manual; Figure S4).
+The populations memberships associated with these models were used for subsequent analysis (Figure S5).
 
-We identified loci under selection.
-For each species, we ran $\texttt{Bayescan}$ [version 2.1; @r456] using the population memberships identified in the previous analysis (1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations; Table S2).
-To avoid false-positive, we omitted loci where the global frequency of the minor allele was $\geq 0.05$ (following guidelines in the $\texttt{Bayescan}$ manual).
-Additionally, to prevent individuals with unclear ancestry from biasing this analysis, we omitted individuals if there was considerable uncertainty in their population membership (maximum probability $\leq `r mclust.params.LST[[MODE]]$probability.threshold`$ % ).
-We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate to ensure convergence [FDR $\leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].
+We identified loci under selection for each species (Table S2).
+For each species, we ran $\texttt{Bayescan}$ [version 2.1; @r456] using the population memberships identified in the previous analysis (1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations).
+We omitted loci where the global frequency of the minor allele was $\geq 0.05$ to avoid false-positives (following guidelines in the $\texttt{Bayescan}$ manual).
+To ensure convergence, we ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate [FDR; $\leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].
 
@@
 For each species, for each type of genetic variation detected, we used Gower distances [@r464] to express the genomic differences between individuals [using the $\texttt{cluster}$ R package to accommodate sparsity; @r457].
-These distances were subject to multiple non-metric multi-dimensional scaling analyses [NMDS; implemented in the $\texttt{vegan}$ R package; @r458] using the minimum number of dimensions required to achieve an adequate stress value  ($\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for all models; Table S3).
-Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the samples in each grid cell to summarize the typical genetic characteristics of individuals in the cells.
+These distances were subject to multiple non-metric multi-dimensional scaling analyses [NMDS; implemented in the $\texttt{vegan}$ R package; @r458] using the minimum number of dimensions required to achieve an adequate stress value  ($\leq$ `r mds.params.LST[[MODE]]$max.stress`; `r mds.params.LST[[MODE]]$trymax` random starts for all models; Table S3).
+Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the samples in each grid cell to characterize their the genetic variation of individuals in the cells.
 

### 2016-02-05 article/supporting_information.Rmd
+ replace adegenet with mclust (reverted from commit 08f8bcca6079915141bf3f8b3b3510f1d1a4edb9)

@@
 Each panel denotes a different species.
-Each line denotes a different model parametrization.
-Asterisks indicate the model and number of populations with the most support.'}
+Asterisks indicate the number of populations with the most support (see text for methodological details).'}
 # format data
+bic.buff <- laply(spp.Adegenet.LST, `[[`, 'Kstat') %>%
+	c() %>% range() %>% diff %>% `*`(0.05)
 bic.DF <- ldply(
@@
 	.fun=function(i) {
-		curr.MTX <- spp.Mclust.LST[[i]]$BIC
-		attr(curr.MTX, 'dim') <- c(nrow(curr.MTX), ncol(curr.MTX))
-		attr(curr.MTX, 'dimnames') <- list(
-			seq_len(nrow(curr.MTX)),
-			attr(curr.MTX, 'modelNames')
-		)
-		class(curr.MTX) <- 'matrix'
-		curr.DF <- curr.MTX %>% as.data.frame %>% 
-			mutate(G=attr(curr.MTX, 'G')) %>% 
-			gather(Model,BIC,-G)
-		curr.DF$BIC=curr.DF$BIC*-1
-		return(
-			mutate(
-				curr.DF,
-				Species=gsub('_', ' ', unique(spp.samples.DF$species)[i], fixed=TRUE),
-				best_G=curr.DF[which.min(curr.DF$BIC),'G'][[1]],
-				best_stat=min(curr.DF$BIC,na.rm=TRUE),
-				asterisk_col=c('black', rep(NA_character_, length(Species)-1))
-			)
+		data.frame(
+			k=as.numeric(gsub('K=', '', names(spp.Adegenet.LST[[i]]$Kstat), fixed=TRUE)),
+			stat=unname(spp.Adegenet.LST[[i]]$Kstat),
+			best_k=as.numeric(gsub('K=', '', names(spp.Adegenet.LST[[i]]$stat), fixed=TRUE)),
+			best_stat=unname(spp.Adegenet.LST[[i]]$stat) + bic.buff,
+			asterisk_col=c('black',rep('transparent', length(spp.Adegenet.LST[[i]]$Kstat)-1)),
+			Species=gsub('_', ' ', unique(spp.samples.DF$species)[i], fixed=TRUE)
 		)
@@
 )
-
 # make plot
@@
 	ggplot(data=bic.DF) +
-		geom_point(aes(y=BIC, x=G, color=Model)) +
-		geom_line(aes(y=BIC, x=G, color=Model)) +
-		geom_point(aes(y=best_stat, x=best_G), size=4, shape=8) +
+		geom_point(aes(y=stat, x=k)) +
+		geom_point(aes(y=best_stat, x=best_k), size=4, shape=8) +
+		geom_line(aes(y=stat, x=k)) +
 		theme_classic() +
-		scale_y_reverse()+
 		theme(
@@
 		ylab('BIC') +
-		facet_wrap(~ Species, ncol=4, scales='free_y')
+		facet_wrap(~ Species, ncol=4)
 })

### 2016-02-05 R/analysis/00-initialization.R

@@ -37,10 +37,9 @@ suppressMessages(library(english))
 suppressMessages(library(session))
 suppressMessages(library(maptools))
 suppressMessages(library(RcppTOML))
-suppressMessages(library(mclust))
+suppressMessages(library(adegenet))
 suppressMessages(library(lme4))
 suppressMessages(library(RColorBrewer))
-suppressMessages(library(mix))
 
 ## load github packages
 # devtools::install_github('paleo13/raspr')

### 2016-02-05 R/analysis/03-population-clustering-analysis.R

@@ -1,54 +1,0 @@
-## load .rda
-session::restore.session('results/.cache/02-surrogate-data.rda')
-
-## load parameters
-mclust.params.LST <- parseTOML('parameters/mclust.toml')
-
-### mclust analyses
-# run clustering analyses
-clust <- makeCluster(mclust.params.LST[[MODE]]$threads, type='SOCK')
-clusterEvalQ(clust, {library(structurer);library(mclust)})
-clusterExport(clust, c('mclust.params.LST','MODE', 'spp.StructureData.LST', 'spp.samples.DF'))
-registerDoParallel(clust)
-spp.Mclust.LST <- llply(
-	seq_along(spp.StructureData.LST),
-	.fun=function(i) {
-		# init
-		curr.MTX <- spp.StructureData.LST[[i]]@matrix
-		# impute missing data if apropriate
-		if (any(is.na(curr.MTX)))
-			curr.MTX <- imputeData(curr.MTX)
-		# run cluster analysis
-		return(
-			Mclust(
-				curr.MTX,
-				G=mclust.params.LST[[MODE]]$G
-			)
-		)
-	},
-	.parallel=TRUE
-)
-clust <- stopCluster(clust)
-
-# assign individuals to populations if they have probability to a single population above threshold
-spp.BayeScanData.sample.subset.LST <- llply(
-	seq_along(spp.BayeScanData.LST),
-	.fun=function(i) {
-		print(i)
-		# get population identities
-		probs <- spp.Mclust.LST[[i]]$z
-		ids <- apply(probs, 1, which.max)
-		ids[which(apply(probs, 1, function(x) {max(x) < mclust.params.LST[[MODE]]$probability.threshold}))] <- NA_integer_
-		validPos <- which(!is.na(ids))
-		# remove individuals below threshold
-		curr.spp <- spp.BayeScanData.LST[[i]]
-		curr.spp <- bayescanr:::sample.subset.BayeScanData(curr.spp, validPos)
-		curr.spp@labels <- filter(spp.samples.DF, species==unique(spp.samples.DF$species)[i])$cell[validPos]
-		curr.spp@populations <- as.character(ids[validPos])
-		return(curr.spp)
-	}
-)
-
-
-## load .rda
-save.session('results/.cache/03-population-clustering-analysis.rda')

### 2016-02-06 article/article.Rmd
+ reversion to adegenet + but include updates to manuscript + change 03-*.R and 04-*.R to meaningful file names

--- 2016-02-05 article/article.Rmd
@@
 The purpose of this step was to avoid misclassifying loci as adaptive.
-We used Gaussian mixture models to assign individuals to populations [implemented in $\texttt{mclust}$ R package; @r477; @r467].
-Briefly, for each species, we fit a suite of models to estimate the probability that individuals belonged different populations (using default priors and covariance structures).
-For each parametrization, we fit a set of models assuming varying numbers of total populations (\textit{G} = 1--`max(r mclust.params.LST[[MODE]]$G)`).
-We identified the most parsimonious models (lowest Bayesian Information Criterion $BIC$ statistic; Figures S4), and used these to group individuals into populations (Figure S5).
+We used discriminant analyses of principle components [DAPC; @r467] to assign individuals to populations [implemented in $\texttt{adegenet}$ R package; @r468].
+Briefly, for each species, we transformed the binary loci data using a principle components analysis (PCA) and obtained the principle components required to explain at least `r adegenet.params.LST[[MODE]]$perc.pca` % variation.
+We subject these components to multiple \textit{k}-means clustering models with varying numbers of assumed populations (\textit{k} = 1--`r adegenet.params.LST[[MODE]]$max.n.clust`; `r adegenet.params.LST[[MODE]]$n.iter` iterations; `r adegenet.params.LST[[MODE]]$n.start` random starts).
+We then  applied Ward's clustering method to the Bayesian Information Criterion (BIC) statistics associated with each model to partition the models into two groups, and selected the model before the first group switch (as recommended in the \texttt{adegenet} manual; Figure S4).
+The populations memberships associated with these models were used for subsequent analysis (Figure S5).
 
@@
 To avoid false-positive, we omitted loci where the global frequency of the minor allele was $\geq 0.05$ (following guidelines in the $\texttt{Bayescan}$ manual).
-Additionally, to prevent individuals with unclear ancestry from biasing this analysis, we omitted individuals if there was considerable uncertainty in their population membership (maximum probability $\leq `r mclust.params.LST[[MODE]]$probability.threshold`$ % ).
 We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate to ensure convergence [FDR $\leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].
@@
 We set surrogate- and genetic-based targets as `r (rapr.params.LST[[MODE]]$surrogate.target*100)` % to ensure that most of the repsective variation was secured for each species.
-To accomodate variation in the high number of optimal solutions for the single-species amount-based prioritizations, we generated `r (rapr.parmas.LST[[MODE]]$amount.replicates)` replicate prioritizations.
+To accomodate variation in the high number of optimal solutions for the single-species amount-based prioritizations, we generated `r (rapr.params.LST[[MODE]]$amount.replicates)` replicate prioritizations.
 Since all other contexts were likely to be associated with a single optimal solution, we used only a single replicate.

### 2016-02-06 R/analysis/03-population-clustering-analysis.R

@@ -50,4 +50,4 @@ spp.BayeScanData.sample.subset.LST <- llply(
 )
 
 ## load .rda
-save.session('results/.cache/03-adegenet-analysis.rda')
+save.session('results/.cache/03-population-clustering-analysis.rda')

### 2016-02-06 R/analysis/05-format-data-for-prioritisations.R

@@ -1,5 +1,5 @@
 ## load .rda
-session::restore.session('results/.cache/04-bayescan-analysis.rda')
+session::restore.session('results/.cache/04-genetic-nmds.rda')
 
 # generate attribute spaces for geographic and environmental data
 surrogate.ASL <- llply(
@@ -22,7 +22,7 @@ adaptive.ASL <- llply(
 		# else return attribute space
 		make.single.species.AttributeSpace(
 			site.data=select(grid.DF, contains(paste0(unique(spp.samples.DF$species)[i], '_adaptive'))),
-			species.data=filter(spp.samples.DF, species==unique(spp.samples.DF$species)[i])[,paste0('adaptive_d',seq_len(spp.mds.LST[[i]][['adaptive']]$ndim)),drop=FALSE],
+			species.data=filter(spp.samples.DF, species==unique(spp.samples.DF$species)[i])[,paste0('adaptive_d',seq_len(spp.nmds.LST[[i]][['adaptive']]$ndim)),drop=FALSE],
 			spp.pos=i,
 			n.species=n_distinct(spp.samples.DF$species)
 		)
@@ -39,7 +39,7 @@ neutral.ASL <- llply(
 		# else return attribute space
 		make.single.species.AttributeSpace(
 			site.data=select(grid.DF, contains(paste0(unique(spp.samples.DF$species)[i], '_neutral'))),
-			species.data=filter(spp.samples.DF, species==unique(spp.samples.DF$species)[i])[,paste0('neutral_d',seq_len(spp.mds.LST[[i]][['neutral']]$ndim)),drop=FALSE],
+			species.data=filter(spp.samples.DF, species==unique(spp.samples.DF$species)[i])[,paste0('neutral_d',seq_len(spp.nmds.LST[[i]][['neutral']]$ndim)),drop=FALSE],
 			spp.pos=i,
 			n.species=n_distinct(spp.samples.DF$species)
 		)

### 2016-02-10 article/article.Rmd
+ include random prioritisations to verify surrogate performance + change dependnecies MAKEFILE to avoid re-running stuff + rewrite of results in article + changes to stats

@@
 ## Statistical analysis
-We investigated the relationship between surrogate-based targets and the genetic variation secured in a prioritization using nonlinear mixed-effects models [NLMMs; using the $\texttt{lme4}$ R package; @r469].
-A model was fit to the genetic variation secured using a negative exponential formulation to estimate the rate of the accumulation of genetic variation ($y = a - (a - 0) \times exp(-r \times x)$; where $a$ is the asymptote and $r$ is the accumulate rate).
-An interaction term was included to estimate parameters separately for adaptive and neutral variation.
-The models accommodated species-level variation using random effects for both parameters.
-
-We tested for differences in the effectiveness of prioritizations generated under different scenarios and targets using generalized linear mixed-effects models [GLMMs; using the $\texttt{lme4}$ R package].
-We fit a full model to the proportion of genomic variation secured in all prioritizations using a logit link.
-Both models contained variables indicating the scenario used to generate the prioritization, the targets used to generate the prioritizations, and the type of genetic variation measured.
-The model was subject to a step-wise term deletion routine using likelihood ratio tests (LRT) to assess term significance (terms were retained when $P < 0.05$).
-All models contained an intercept-only random effect to ensure that comparisons were made within species.
+We investigated the effectiveness of prioritizations generated using surrogate-based targets using generalized linear mixed-effects models [GLMMs; using the $\texttt{lme4}$ R package; @r469].
+All models were fit to the proportion of genetic variation that prioritizations secured using a logit link.
+To understand the relationship between the surrogate-based target used to generate a prioritization and how much genetic variation the resulting prioritization secured, we fit a full model containing a continuous quadratic predictor variable for the surrogate target used to generate the prioritization, a categorical variable indicating the type of genetic variation, and a variable for the interaction between the previous variables.
+This model contained random effects for both the intercept and the predictor variables to accommodate variation between species.
+To determine if surrogate-based targets improved the effectiveness of prioritizations under different scenarios, we fit a full model containing categorical variables indicating the scenario used to generate the prioritization, the combination of targets used to generate the prioritizations, and the type of genetic variation measured.
+This model contained intercept-only random effects.
+Both models were subject to step-wise term deletion routines using likelihood ratio tests (LRT) to assess significance (terms were retained when $P < 0.05$).
 Finally, we conducted a post-hoc analysis using Tukey contrasts with a Bonferroni correction [using the $\texttt{multcomp}$ R package; @r470].
@@
 ## Results
-<!-- Single species prioritizations -->
-Surrogates-targets improved the performance of single-species prioritizations (Figure 1).
-There were significant differences among the performance of prioritizations depending on the targets used to generate them and if they were assessed according to adaptive or neutral genetic variation (${\chi^{2}}_{`r single.spp.AOV1[['Df']][1]`,`r single.spp.AOV1[['Df']][2]`} = `r round(single.spp.AOV1[['Chisq']][2],2)`$, $P `r pretty.pval(single.spp.AOV1[['Pr(>Chisq)']][2])`$).
-Specifically, the amount-based prioritization captured much less adaptive and neutral genetic variation than either the surrogate-based ($Z = `r round(single.spp.MCP$test$tstat[['Surrogate.Adaptive variation - Amount.Adaptive variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Surrogate.Adaptive variation - Amount.Adaptive variation']])`$; $Z = `r round(single.spp.MCP$test$tstat[['Surrogate.Neutral variation - Amount.Neutral variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Surrogate.Neutral variation - Amount.Neutral variation']])`$) or genetic-based ($Z = `r round(single.spp.MCP$test$tstat[['Genetic.Adaptive variation - Amount.Adaptive variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Genetic.Adaptive variation - Amount.Adaptive variation']])`$; $Z = `r round(single.spp.MCP$test$tstat[['Genetic.Neutral variation - Amount.Neutral variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Genetic.Neutral variation - Amount.Neutral variation']])`$) prioritizations.
-No differences were detected in the proportion of adaptive ($Z = `r round(single.spp.MCP$test$tstat[['Genetic.Adaptive variation - Surrogate.Adaptive variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Genetic.Adaptive variation - Surrogate.Adaptive variation']])`$) or neutral genetic variation ($Z = `r round(single.spp.MCP$test$tstat[['Genetic.Neutral variation - Surrogate.Neutral variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Genetic.Neutral variation - Surrogate.Neutral variation']])`$) secured by the the surrogate-based and genetic-based prioritizations.
-However, these patterns were not reflected in the multi-species prioritizations.
+<!-- Surrogacy analysis -->
+Prioritizations generated using greater environmental and geographic surrogate-based targets were found to secure greater proportions of adaptive and neutral genetic variation (Figure 1).
+The proportion of genetic variation that these prioritizations secured was substantially more than the genetic variation secured by randomly generated prioritizations (`r xxx`).
+Generally, prioritizations needed to capture at least `r xxx` % of the environmental and geographic surrogates to secure 90 % of the species' genetic variation.
+Additionally, we detected differences in the effectiveness of the environmental and geographic surrogates (`r xxxx`).
+Increases in environmental surrogate-based targets resulted prioritizations that secured more adaptive genetic variation, than geographic surrogate-based targets for neutral variation.
+These results demonstrate that environmental and geographic variables are effective surrogates for genetic variation.
 
-<!-- Multi-species prioritizations -->
-Environmental and geographic surrogates were not found to improve the performance of multi-species prioritizations (Figure 2; Figure 3).
-No significant differences in the effectiveness of prioritizations were found depending on whether they were generated using different targets (${\chi^{2}}_{`r multi.spp.AOV3[['Df']][1]`,`r multi.spp.AOV3[['Df']][2]`} = `r round(multi.spp.AOV3[['Chisq']][2],2)`$; $P `r pretty.pval(multi.spp.AOV3[['Pr(>Chisq)']][2])`$), if they were assessed using adaptive or neutral genetic variation (${\chi^{2}}_{`r multi.spp.AOV2[['Df']][1]`,`r multi.spp.AOV2[['Df']][2]`} = `r round(multi.spp.AOV2[['Chisq']][2],2)`$; $P `r pretty.pval(multi.spp.AOV2[['Pr(>Chisq)']][2])`$), nor the interaction between these variables (${\chi^{2}}_{`r multi.spp.AOV1[['Df']][1]`,`r multi.spp.AOV1[['Df']][2]`} = `r round(multi.spp.AOV1[['Chisq']][2],2)`$; $P  `r pretty.pval(multi.spp.AOV1[['Pr(>Chisq)']][2])`$).
-
-<!-- Correlation analysis -->
-The surrogate variables showed a greater utility for neutral genetic variation than adaptive variation (Figure 4).
-The accumulation rate ($r$) was `r round(summary(env.correlation.NLMM)$coefficients[1],2)` $\pm$ `r round(summary(env.correlation.NLMM)$coefficients[2],2)` S.E.
-for the proportion of adaptive variation captured in a prioritization using environmental surrogates.
-On the other hand, the rate ($r$) was `r round(summary(geo.correlation.NLMM)$coefficients[1],2)` $\pm$ `r round(summary(geo.correlation.NLMM)$coefficients[2],2)` S.E.
-for neutral variation using geographic surrogates.
-Thus minor increases in the target for either surrogate results in major increases in the benefit of the prioritization.
-However, the coefficient was much larger for neutral genetic variation so minor increases in the target can achieve substantial gains.
+<!-- Scenario analysis -->
+The addition of environmental and geographic surrogate-based targets to amount-based targets was not found to substantially improve the effectiveness of prioritizations under any of the scenarios (Figure 2).
+No significant differences were detected between prioritizations generated using different combinations of targets (${\chi^{2}}_{`r `,`r `} = `r `$, $P `r `$), or under any of the different scenarios (${\chi^{2}}_{`r `,`r `} = `r `$, $P `r `$).
+The only difference was that the inclusion of surrogate-based targets tended to result in prioritizations that secured slightly more adaptive than neutral genetic variation.
+Under the simplest of circumstances, the single-species amount-based prioritizations secured most of the species' adaptive (mean `r ` \pm `r `) and neutral (mean `r ` \pm `r `) genetic variation.
+These proportions were not significantly different to those associated with prioritizations generated using additional surrogate-based targets ($Z = `r `$, $P `r `$) or genetic-based targets ($Z = `r `$, $P `r `$).
+We found similar patterns in the multi-species scenario () and the multi-species with opportunity cost scenario ().
+Although the prioritizations had similar levels of performance, there were different in the their spatial distributions (Figure 3).
+In particular, the inclusion of cost data in the multi-species prioritizations had the effect of concentrating conservation effort towards areas with higher elevation.
+Overall, these results suggest that while environmental and geographic surrogate variables are effective surrogates, they are not needed to generate prioritizations that secure most of the intra-specific genetic variation.
 

### 2016-02-10 article/supporting_information.Rmd
+ include random prioritisations to verify surrogate performance + change dependnecies MAKEFILE to avoid re-running stuff + rewrite of results in article + changes to stats

@@
 ```{r, echo=FALSE}
-plot.spp.mds <- function(i) {
+plot.spp.nmds <- function(i) {
 	## prepare data
@@
 	if (adapt.col %in% names(grid.FPLY))
-		n.plots <- n.plots + spp.mds.LST[[i]][['adaptive']]$ndim
+		n.plots <- n.plots + spp.nmds.LST[[i]][['adaptive']]$ndim
 	# count neutral plot
 	if (neutral.col %in% names(grid.FPLY))
-		n.plots <- n.plots + spp.mds.LST[[i]][['neutral']]$ndim
+		n.plots <- n.plots + spp.nmds.LST[[i]][['neutral']]$ndim
 	# exports
@@
 No indiviudals of this species were found inside the gray planning units.')}
-if (1 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(1)
+if (1 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(1)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (2 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(2)
+if (2 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(2)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (3 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(3)
+if (3 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(3)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (4 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(4)
+if (4 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(4)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (5 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(5)
+if (5 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(5)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (6 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(6)
+if (6 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(6)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (7 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(7)
+if (7 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(7)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (8 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(8)
+if (8 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(8)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (9 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(9)
+if (9 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(9)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (10 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(10)
+if (10 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(10)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (11 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(11)
+if (11 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(11)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (12 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(12)
+if (12 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(12)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (13 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(13)
+if (13 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(13)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (14 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(14)
+if (14 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(14)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (15 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(15)
+if (15 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(15)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (16 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(16)
+if (16 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(16)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (17 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(17)
+if (17 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(17)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (18 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(18)
+if (18 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(18)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (19 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(19)
+if (19 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(19)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (20 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(20)
+if (20 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(20)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (21 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(21)
+if (21 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(21)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (22 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(22)
+if (22 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(22)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (23 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(23)
+if (23 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(23)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (24 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(24)
+if (24 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(24)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (25 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(25)
+if (25 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(25)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (26 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(26)
+if (26 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(26)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (27 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(27)
+if (27 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(27)
 ```
@@
 					ldply(
-						seq_along(spp.mds.LST[[i]]),
+						seq_along(spp.nmds.LST[[i]]),
 						function(j) {
@@
 							Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
-							Loci=names(spp.mds.LST[[i]])[j],
-							k=ifelse(is.null(spp.mds.LST[[i]][[j]]), NA_integer_, spp.mds.LST[[i]][[j]]$ndim),
-							Stress=ifelse(is.null(spp.mds.LST[[i]][[j]]), NA_real_, spp.mds.LST[[i]][[j]]$stress),
-							Converged=ifelse(is.null(spp.mds.LST[[i]][[j]]), NA, spp.mds.LST[[i]][[j]]$converged)
+							Loci=names(spp.nmds.LST[[i]])[j],
+							k=ifelse(is.null(spp.nmds.LST[[i]][[j]]), NA_integer_, spp.nmds.LST[[i]][[j]]$ndim),
+							Stress=ifelse(is.null(spp.nmds.LST[[i]][[j]]), NA_real_, spp.nmds.LST[[i]][[j]]$stress),
+							Converged=ifelse(is.null(spp.nmds.LST[[i]][[j]]), NA, spp.nmds.LST[[i]][[j]]$converged)
 						)

### 2016-02-10 R/analysis/08-multi-species-prioritisations-no-cost.R

@@ -1,5 +1,9 @@
 ## load .rda
-session::restore.session('results/.cache/07-single-species-prioritisations.rda')
+session::restore.session('results/.cache/05-format-data-for-prioritisations.rda')
+
+## load parameters
+rapr.params.LST <- parseTOML('parameters/rapr.toml')
+gurobi.params.LST <- parseTOML('parameters/gurobi.toml')
 
 ## multispecies analysis
 # make prioritisations

### 2016-02-18 article/article.Rmd
+ manuscript edits + switch back to mclust + use less complicated stats

@@
 <!--  main text -->
-Biodiversity is under threat.
-In recent history, many species have gone extinct [@r270; @r272].
-Today, thousands of species are listed as endangered or near threatened in the International Union for Conservation of Nature (IUCN) red list (\url{www.iucnredlist.org}).
-In the near future, species are expected to face even more threats [eg.
-climate change; @r236; @r243].
-In response to this crisis, the overarching aim of conservation is to maximize the long-term persistence of biodiversity [@r255; @r67].
-
-Protected areas are one of the hallmarks of modern conservation [@r440].
-They buffer species from anthropogenic impacts and provide places for biodiversity processes to persist (eg.
-nutrient cycling and water purification).
-Protected areas also set the stage for direct management actions to enhance species recovery (eg.
-habitat rehabilitation, translocation, and invasive species management).
-But the resources available for conservation are limited, and so protected areas must be cost-effective.
-As a consequence, decision support tools [eg.
-$\texttt{MARXAN}$; @r22; $\texttt{Zonation}$; @r429] are used to identify cost-effective plans for protected areas (prioritizations).
-These tools typically estimate the benefit that a given species receives from a given prioritization based on the proportion of its range that is secured (eg.
-using targets).
-However, the true benefit of a given prioritization is dependent on many additional factors that are not used to inform the reserve selection process.
-
-Protected areas must preserve intra-specific genetic variation to maximize species recovery [@r67; @r3; @r61].
-This variation can be classified as either adaptive or neutral [reviewed in @r????].
+The act of setting aside areas for preservation spearheads modern conservation efforts [@r440].
+They buffer species from anthropogenic impacts and provide places for populations of species to persist.
+Intra-specific genetic variation underpins species persistence, and there has been increasing interest in understanding how to design protected areas networks in a way that maximizes this variation [@r67; @r3; @r61].
+Broadly speaking, genetic variation can be classified as either adaptive or neutral [reviewed in @r????].
 Adaptive genetic variation is associated with loci currently under selection.
@@
 By preserving existing patterns of neutral variation, protected areas can secure different genetic lineages and avoid the adverse effects of low genetic diversity [eg.
-inbreeding depression; @r67].
+inbreeding depression and reduced capacity to respond to environmental change; @r67].
 However, genetic data requires considerable investment to obtain and analyze.
 As a consequence, conservation planners do not typically use genetic data to inform prioritizations [@r61].
-Thus many prioritizations may fail to capture species' intra-specific genetic variation, and ultimately fail to deliver conservation outcomes.
+Thus many prioritizations may fail to capture species' intra-specific genetic variation, and ultimately fail to deliver long-term conservation outcomes.
 
-In light of this problem, conservation scientists have been investigating surrogates for genetic variation to generate better prioritizations--but without the overheads associated with genetic data.
-Since adaptive genetic variation is ultimately driven by selection pressures, conservation planners have used the distribution of likely selection pressures.
-For instance, environmental variation such as local habitat type [@r247] and climate [@r2] have been used to guide the reserve selection process.
-While such variables have been used to detect and verify adaptive genetic variation [eg.
-@r475; @r455], the effectiveness of prioritizations based on such surrogates remains unverified.
-
-On the other hand, neutral genetic variation arises a reduction in gene flow between individuals.
-By modeling the connectivity between areas, a function of gene flow, conservation scientists have explored the effectiveness of surrogates for neutral genetic variation.
-For instance, the isolation by distance theory [IBD; @r259] predicts that populations further apart will have reduced connectivity, and in turn exchange less genetic material.
-The use of geographic distance-based surrogates has been found to improve prioritizations in an island network [@r200].
+Recently, conservation scientists have been investigating surrogates for genetic variation to generate better prioritizations without the overheads associated with genetic data.
+Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been used to guide reserve selection [eg.
+habitat type and local climate; @r247; @r2].
+These variables are often used to detect adaptive variation [eg.
+@r475; @r455].
+However, their effectiveness as surrogates in conservation planning remains unverified.
+On the other hand, neutral genetic variation arises from a reduction in gene flow between individuals.
+Thus surrogates for neutral variation are based on variables that predict the level of connectivity between planning units.
+For instance, the isolation by distance theory [IBD; @259] predicts populations further apart will have less connectivity,  result in reduced gene flow, and in turn share less neutral genetic variation.
+Geographic distance-based surrogates have been found to improve prioritizations in an island network [@200].
 However this remains untested in spatially contiguous systems where connectivity is complicated by additional factors [eg.
@@
 Before conservation planners can reliably use surrogates for genetic data in conservation planning, the effectiveness of these surrogates needs to be verified.
-Otherwise, if these surrogates are ineffective, conservation planners may waste precious resources on places with little benefit.
+Otherwise, conservation planners may waste precious resources on places that do little to benefit biodiversity..
 
-Our overarching aim was to investigate the effectiveness of environmental and geographic surrogates for adaptive neutral genetic variation in the context of conservation planning.
-First, we aimed to understand the relationship between the surrogate targets used to generate a prioritization and the proportion of genetic variation secured in the prioritization.
-Second, we aimed to determine if the addition of surrogate-based targets in more realistic conservation planning scenarios would result in prioritizations that secured substantially more genetic variation.
+Here we investigate the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation in the context of conservation planning.
+First, we examine the relationship between surrogates and gene variation, and verify that these variables can be used as surrogates.
+Second, we determine if the inclusion of surrogate-based targets in a variety of conservation planning scenarios with increasingly levels of realism can result in prioritizations that secure more genetic variation.
 If these surrogates can be shown to improve the intra-specific genetic variation captured in a prioritization, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.
@@
 ### Study system
-To address the aims of the study, we utilized species distribution and genomic data collected by the IntraBioDiv project [@r451; see @r452 for methodological details].
-This dataset was particularly well suited; it provides spatially explicit genomic data for a multitude of species across a broad geographic area over large environmental gradients.
-As part of this project, data were collected for 27 alpine plant species were across the European Alps according to a grid (22.3 km $\times$ 25 km; Figure S1).
-Briefly, in every second cell, plant samples were collected from three individuals for each species present inside the cell.
-Cells that did not contain any plant samples were omitted from our analysis.
-Samples were genotyped using amplified fragment length polymorphisms [AFLP; @r453].
-Matrices denoting the presence/absence of polymorphisms at loci were constructed independently for each species.
-Thus for each second grid cell, for each of the species in the grid cell, the dataset contains information describing the genomic variation of individuals inside the grid cell.
-We used these grid cells as planning units to develop prioritizations to investigate the effectiveness of surrogates for the (AFLP) genetic variation.
+We used species distribution and genomic data collected by the IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for methodological details].
+This dataset was particularly well suited for this study, because it provides spatially explicit genomic data for multiple species with vastly different demographic characteristics and evolutionary histories across a broad geographic area with large environmental gradients.
+As part of this project, data was collected for 27 alpine plant species using a 20' longitude by 21' latitude grid (approx.
+21' 22.3 km $\times$ 25 km; Figure S1).
+They visited every second grid cell, and if a species was detected in a grid cell, samples were collected from three individuals.
+They genotyped samples using amplified fragment length polymorphisms [AFLP; @r453], and constructed matrices denoting the presence/absence of polymorphisms at loci for each species.
+Thus for each second grid cell, for each of the species present in the grid cell, the dataset contains information describing the genomic variation of individuals in that grid cell.
+We used these grid cells as planning units to develop prioritizations investigating the effectiveness of surrogates for the (AFLP) genetic variation.
+Note that we omitted grid cells that did not contain any individuals from our analysis.
 The majority of our analyses described hereafter were performed using the \texttt{R} software package [version `r paste0(version$major, '.', version$minor)`; @r192].
@@
 These layers and the grid cells were projected into an equal-area coordinate system ($\texttt{ESRI:102014}$).
-The bioclimatic variables were clipped to the extent of the grid and subject to a principle components analysis (PCA; using ArcMap version 10.2.2; Table S1).
-The first two principle components (PCs) cumulatively explained `r round(pca.DF[[4]][2],1)` % of the variation and were used to generate two new layers.
+The bioclimatic variables were clipped to the extent of the grid and subject to a principal components analysis (PCA; using ArcMap version 10.2.2; Table S1).
+The first two principal components (PCs) cumulatively explained `r round(pca.DF[[4]][2],1)` % of the variation and were used to generate two new layers.
 We calculated the average value for each PC layer in each grid cell, and used these values to characterize the main themes of climatic variation among the grid cells (Figure S2).
-To examine if the influence of opportunity cost, we calculated the total human population density inside each planning unit (obtained at 1 km resolution from the Global Rural-Urban Mapping Project; GRUMP V1; @r472).
+To examine if the influence of opportunity cost, we calculated the total human population density inside each planning unit (obtained at 1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472).
 
@@
 r@475]--is that they do not use environmental data to identify loci under selection, which would have biased our analysis.
-In an earlier study, Bothwell _et al._ [-@r455] applied such outlier detection methods to one of the species in this dataset (_Getiana nivalis_).
+In an earlier study, Bothwell _et al._ [-@r455] applied such outlier detection methods to one of the species in this dataset (_Gentiana nivalis_).
 Here, we adopted a similar methodology and applied it to each of the 27 species in the dataset.
@@
 The purpose of this step was to avoid misclassifying loci as adaptive.
-We used discriminant analyses of principle components [DAPC; @r467] to assign individuals to populations [implemented in $\texttt{adegenet}$ R package; @r468].
-Briefly, for each species, we transformed the binary loci data using a principle components analysis (PCA) and obtained the principle components required to explain at least `r adegenet.params.LST[[MODE]]$perc.pca` % variation.
-We subject these components to multiple \textit{k}-means clustering models with varying numbers of assumed populations (\textit{k} = 1--`r adegenet.params.LST[[MODE]]$max.n.clust`; `r adegenet.params.LST[[MODE]]$n.iter` iterations; `r adegenet.params.LST[[MODE]]$n.start` random starts).
-We then  applied Ward's clustering method to the Bayesian Information Criterion (BIC) statistics associated with each model to partition the models into two groups, and selected the model before the first group switch (as recommended in the \texttt{adegenet} manual; Figure S4).
-The populations memberships associated with these models were used for subsequent analysis (Figure S5).
+For each species, we used non-metric multi-dimensional scaling [NMDS; implemented in the $\texttt{vegan R}$ package] to transform the binary loci data into continuous variables (Table S2).
+We used Gower distances [@r464] to express the genomic differences between individuals [using the $\texttt{cluster R}$ package to accommodate sparsity; @r457].
+We ran successive scaling analyses with increasing dimensionality until ordinations were obtained with an adequate stress value ($\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for all models).
+We then fit a suite of Gaussian mixture to the ordinated values to estimate the probability that individuals belonged to different populations [using the default priors and covariance structures implemented in the $\texttt{mclust} R$ package; @r477; @r476].
+For each parametrization, we fit a set of models assuming varying numbers of total populations (\textit{G}= 1--`r max(mclust.params.LST[[MODE]]$G)`).
+We identified the most parsimonious models (lowest Bayesian Information Criterion $BIC$ statistic; Figures S4), and used these to group individuals into populations (Figure S5).
 
-We identified loci under selection.
-For each species, we ran $\texttt{Bayescan}$ [version 2.1; @r456] using the population memberships identified in the previous analysis (1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations; Table S2).
+We identified loci under selection (Table S3).
+For each species, we ran $\texttt{Bayescan}$ [version 2.1; @r456] using the population memberships identified in the previous analysis (1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations).
 To avoid false-positive, we omitted loci where the global frequency of the minor allele was $\geq 0.05$ (following guidelines in the $\texttt{Bayescan}$ manual).
@@
 After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species (Figures S6--S32).
-For each species, for each type of genetic variation detected, we used Gower distances [@r464] to express the genomic differences between individuals [using the $\texttt{cluster}$ R package to accommodate sparsity; @r457].
-These distances were subject to multiple non-metric multi-dimensional scaling analyses [NMDS; implemented in the $\texttt{vegan}$ R package; @r458] using the minimum number of dimensions required to achieve an adequate stress value  ($\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for all models; Table S3).
-Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the samples in each grid cell to summarize the typical genetic characteristics of individuals in the cells.
+For each species, we partitioned adaptive and neutral loci into separate datasets, and reran the nonmetric multidimensional scaling analysis (NMDS) to derive continuous variables that described the main gradients of adaptive and neutral genetic variation separately (Table S2).
+Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the individuals in each grid cell to summarize the typical genetic characteristics of individuals in the cells.
 
@@
 However, they seem to be largely unnecessary in conservation planning scenarios because a comprehensive set of species can be used to achieve a similar result.
-Conservation planners are therefore urged to use a comprehensive set of species in prioritizations--one of the fundamental principles of conservation planning.
+Conservation planners are therefore urged to use a comprehensive set of species in prioritizations--one of the fundamental principals of conservation planning.
 

### 2016-02-18 article/supporting_information.Rmd
+ manuscript edits + switch back to mclust + use less complicated stats

@@
 	.fun=function(i) {
-		data.frame(
-			k=as.numeric(gsub('K=', '', names(spp.Adegenet.LST[[i]]$Kstat), fixed=TRUE)),
-			stat=unname(spp.Adegenet.LST[[i]]$Kstat),
-			best_k=as.numeric(gsub('K=', '', names(spp.Adegenet.LST[[i]]$stat), fixed=TRUE)),
-			best_stat=unname(spp.Adegenet.LST[[i]]$stat) + bic.buff,
-			asterisk_col=c('black',rep('transparent', length(spp.Adegenet.LST[[i]]$Kstat)-1)),
-			Species=gsub('_', ' ', unique(spp.samples.DF$species)[i], fixed=TRUE)
+		curr.MTX <- spp.Mclust.LST[[i]]$mclust$BIC
+		attr(curr.MTX, 'dim') <- c(nrow(curr.MTX), ncol(curr.MTX))
+		attr(curr.MTX, 'dimnames') <- list(
+			seq_len(nrow(curr.MTX)),
+			attr(curr.MTX, 'modelNames')
+		)
+		class(curr.MTX) <- 'matrix'
+		curr.DF <- curr.MTX %>% as.data.frame %>% 
+			mutate(G=attr(curr.MTX, 'G')) %>% 
+			gather(Model,BIC,-G)
+		curr.DF$BIC=curr.DF$BIC*-1
+		return(
+			mutate(
+				curr.DF,
+				Species=gsub('_', ' ', unique(spp.samples.DF$species)[i], fixed=TRUE),
+				best_G=curr.DF[which.min(curr.DF$BIC),'G'][[1]],
+				best_stat=min(curr.DF$BIC,na.rm=TRUE),
+				asterisk_col=c('black', rep(NA_character_, length(Species)-1))
+			)
 		)
@@
 )
+# define ggplot2 color wheel function
+gg_color_hue <- function(n) {
+	hues <- seq(15, 375, length=n+1)
+	hcl(h=hues, l=65, c=100)[1:n]
+}
 # make plot
@@
 	ggplot(data=bic.DF) +
-		geom_point(aes(y=stat, x=k)) +
-		geom_point(aes(y=best_stat, x=best_k), size=4, shape=8) +
-		geom_line(aes(y=stat, x=k)) +
+		geom_point(aes(y=BIC, x=G, color=Model)) +
+		geom_line(aes(y=BIC, x=G, color=Model)) +
+		geom_point(aes(y=best_stat, x=best_G), size=4, shape=8) +
 		theme_classic() +
+		scale_y_reverse()+
 		theme(
@@
 		ylab('BIC') +
-		facet_wrap(~ Species, ncol=4)
+		scale_linetype_manual(value=rep(c('solid','dashed'),each=7)) +
+		scale_color_manual(value=rep(gg_color_hue(7),2) +
+		facet_wrap(~ Species, ncol=4, scales='free_y')
 })
@@
 	format.table(
+		filter(
+			ldply(
+				seq_along(unique(spp.samples.DF$species)), 
+				function(i) {
+					ldply(
+						seq_along(spp.nmds.LST[[i]]),
+						function(j) {
+						rbind(
+							data.frmae(
+								Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
+								Loci='all',
+								k=spp.Mclust.LST[[i]]$nmds$ndim,
+								Stress=spp.Mclust.LST[[i]]$nmds$stress,
+								Converged=spp.Mclust.LST[[i]]$nmds$converged
+							),
+							data.frame(
+								Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
+								Loci=names(spp.nmds.LST[[i]])[j],
+								k=ifelse(is.null(spp.nmds.LST[[i]][[j]]), NA_integer_, spp.nmds.LST[[i]][[j]]$ndim),
+								Stress=ifelse(is.null(spp.nmds.LST[[i]][[j]]), NA_real_, spp.nmds.LST[[i]][[j]]$stress),
+								Converged=ifelse(is.null(spp.nmds.LST[[i]][[j]]), NA, spp.nmds.LST[[i]][[j]]$converged)
+							)
+					})
+				}
+			),
+			!is.na(Stress)
+		),
+		omit=c('Loci','Converged')
+	),
+	digits=2,
+	caption='Summary of non-metric multi-dimensional scaling (MDS) analyses on genetic variation for each species.
+The analyses on all loci were used to transform the binary presence/absence polymorphism data to continuous variables to train Gaussian mixture models and ultimately identify adaptive loci.
+The analyses on adaptive and neutral loci were used to transform the polymorphism data to identify main gradients of this variation.',
+	col.names=c('Species', 'Loci Type', 'k', 'NMDS Stress', 'Converged'),
+	align=c('l', 'c', 'c', 'c', 'c')
+)
+```
+
+```{r, echo=FALSE}
+knitr::kable(
+	format.table(
 		ldply(
@@
 ```
-
-```{r, echo=FALSE}
-knitr::kable(
-	format.table(
-		filter(
-			ldply(
-				seq_along(unique(spp.samples.DF$species)), 
-				function(i) {
-					ldply(
-						seq_along(spp.nmds.LST[[i]]),
-						function(j) {
-						data.frame(
-							Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
-							Loci=names(spp.nmds.LST[[i]])[j],
-							k=ifelse(is.null(spp.nmds.LST[[i]][[j]]), NA_integer_, spp.nmds.LST[[i]][[j]]$ndim),
-							Stress=ifelse(is.null(spp.nmds.LST[[i]][[j]]), NA_real_, spp.nmds.LST[[i]][[j]]$stress),
-							Converged=ifelse(is.null(spp.nmds.LST[[i]][[j]]), NA, spp.nmds.LST[[i]][[j]]$converged)
-						)
-					})
-				}
-			),
-			!is.na(Stress)
-		),
-		omit=c('Loci','Converged')
-	),
-	digits=2,
-	caption='Summary of non-metric multi-dimensional scaling (MDS) analyses on genetic variation for each species.',
-	col.names=c('Species', 'Loci Type', 'k', 'NMDS Stress', 'Converged'),
-	align=c('l', 'c', 'c', 'c', 'c')
-)
-```

### 2016-02-18 R/analysis/06-surrogacy-prioritizations.R

@@ -73,7 +73,6 @@ env.random.prioritisations <- llply(
 			}
 		)
 	},
-	.progress='text',
 	.parallel=TRUE
 )
 # kill cluster
@@ -146,7 +145,6 @@ geo.random.prioritisations <- llply(
 			}
 		)
 	},
-	.progress='text',
 	.parallel=TRUE
 )
 # kill cluster
@@ -160,7 +158,10 @@ env.correlation.DF <- ldply(
 			rbind(
 				mutate(
 					ldply(env.correlation.prioritisations[[i]], extractResults),
-					Surrogate.target=rapr.params.LST[[MODE]]$surrogacy.analysis$surrogate.target,
+					Surrogate.target=rep(
+						rapr.params.LST[[MODE]]$surrogacy.analysis$surrogate.target,
+						each=rapr.params.LST[[MODE]]$surrogacy.analysis$surrogate.optimal.replicates
+					),
 					Method='Optimal'
 				),
 				mutate(
@@ -185,7 +186,10 @@ geo.correlation.DF <- ldply(
 			rbind(
 				mutate(
 					ldply(geo.correlation.prioritisations[[i]], extractResults),
-					Surrogate.target=rapr.params.LST[[MODE]]$surrogacy.analysis$surrogate.target,
+					Surrogate.target=rep(
+						rapr.params.LST[[MODE]]$surrogacy.analysis$surrogate.target,
+						each=rapr.params.LST[[MODE]]$surrogacy.analysis$surrogate.optimal.replicates
+					),
 					Method='Optimal'
 					),
 				mutate(

### 2016-02-21 article/article.Rmd
+ 4 species for debugging + fixes to article so it compiles + fixes to nmds analyses so they use sparse distances

@@
 # load .rda file
-session::restore.session('results/results.rda')
+try(session::restore.session('results/results.rda'))
 
@@
 For each species, we generated two sets of prioritizations using increasing environmental and geographic-based targets (respectively).
-Prioritizations were generated using surrogate-based targets ranging from `r (min(rapr.params.LST[[MODE]]$correlation.surrogate.targets)*100)` % and `r (max(rapr.params.LST[[MODE]]$correlation.surrogate.targets)*100)` % in increments of `r diff(rapr.params.LST[[MODE]]$correlation.surrogate.targets[1:2])*100` %.
-None of these prioritizations contained any amount-based targets.
+Prioritizations were generated using surrogate-based targets ranging from `r (min(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target)*100)` % and `r (max(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target)*100)` % in increments of `r diff(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target[1:2])*100` % ($n = `r rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.optimal.replicates`$ per target).
+To determine if these surrogate-based prioritizations performed better then random, we generated another two sets of prioritizations by randomly selecting planning units that were occupied by the species ($n = `r rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.random.replicates`$ per surrogate target).
+To account for variation in the number of selected planning units in the surrogate-based prioritizations, the number of selected of planning units in the randomly generated prioritizations were generated following their distribution.
 
@@
 The purely amount-based prioritizations represented those typically generated in conservation planning exercises and serve as our baseline.
-We used `r (rapr.params.LST[[MODE]]$amount.target*100)` % amount-based targets in all prioritizations to ensure an adequate proportion of habitat was secured for each species [@rXXX].
-We set surrogate- and genetic-based targets as `r (rapr.params.LST[[MODE]]$surrogate.target*100)` % to ensure that most of the repsective variation was secured for each species.
-To accomodate variation in the high number of optimal solutions for the single-species amount-based prioritizations, we generated `r (rapr.params.LST[[MODE]]$amount.replicates)` replicate prioritizations.
-Since all other contexts were likely to be associated with a single optimal solution, we used only a single replicate.
+We used `r (rapr.params.LST[[MODE]][['single.species']]$amount.target*100)` % amount-based targets in all prioritizations to ensure an adequate proportion of habitat was secured for each species [@rXXX].
+We set surrogate- and genetic-based targets as `r (rapr.params.LST[[MODE]][['single.species']]$amount.target*100)` % to ensure that most of the repsective variation was secured for each species.
+To accomodate variation in the high number of optimal solutions for the single-species amount-based prioritizations, we generated `r (rapr.params.LST[[MODE]][['single.species']]$amount.replicates)` replicate prioritizations.
+Since all other contexts were likely to be associated with few optimal solutions, we generated only `r rapr.params.LST[[MODE]][['multi.species']]$genetic.replicates` replicates.
 
@@
 Prioritizations generated using greater environmental and geographic surrogate-based targets were found to secure greater proportions of adaptive and neutral genetic variation (Figure 1).
-The proportion of genetic variation that these prioritizations secured was substantially more than the genetic variation secured by randomly generated prioritizations (`r xxx`).
-Generally, prioritizations needed to capture at least `r xxx` % of the environmental and geographic surrogates to secure 90 % of the species' genetic variation.
-Additionally, we detected differences in the effectiveness of the environmental and geographic surrogates (`r xxxx`).
+The proportion of genetic variation that these prioritizations secured was substantially more than the genetic variation secured by randomly generated prioritizations (${\chi^{2}}_{`r NA`,`r NA`} = `r  NA`$, $P < `r NA`$).
+Generally, prioritizations needed to capture at least `r NA` % of the environmental and geographic surrogates to secure 90 % of the species' genetic variation.
+Additionally, we detected differences in the effectiveness of the environmental and geographic surrogates (${\chi^{2}}_{`r NA`,`r NA`} = `r  NA`$, $P < `r NA`$).
 Increases in environmental surrogate-based targets resulted prioritizations that secured more adaptive genetic variation, than geographic surrogate-based targets for neutral variation.
@@
 The addition of environmental and geographic surrogate-based targets to amount-based targets was not found to substantially improve the effectiveness of prioritizations under any of the scenarios (Figure 2).
-No significant differences were detected between prioritizations generated using different combinations of targets (${\chi^{2}}_{`r `,`r `} = `r `$, $P `r `$), or under any of the different scenarios (${\chi^{2}}_{`r `,`r `} = `r `$, $P `r `$).
+No significant differences were detected between prioritizations generated using different combinations of targets (${\chi^{2}}_{`r NA`,`r NA`} = `r  NA`$, $P < `r NA`$), or under any of the different scenarios (${\chi^{2}}_{`r NA`,`r NA`} = `r NA`$, $P < `r NA`$).
 The only difference was that the inclusion of surrogate-based targets tended to result in prioritizations that secured slightly more adaptive than neutral genetic variation.
-Under the simplest of circumstances, the single-species amount-based prioritizations secured most of the species' adaptive (mean `r ` \pm `r `) and neutral (mean `r ` \pm `r `) genetic variation.
-These proportions were not significantly different to those associated with prioritizations generated using additional surrogate-based targets ($Z = `r `$, $P `r `$) or genetic-based targets ($Z = `r `$, $P `r `$).
-We found similar patterns in the multi-species scenario () and the multi-species with opportunity cost scenario ().
+Under the simplest of circumstances, the single-species amount-based prioritizations secured most of the species' adaptive (mean `r NA` $\pm$ `r NA`) and neutral (mean `r NA` $\pm$ `r NA`) genetic variation.
+These proportions were not significantly different to those associated with prioritizations generated using additional surrogate-based targets ($Z = `r NA`$, $P < `r NA`$) or genetic-based targets ($Z = `r NA`$, $P < `r NA`$).
+We found similar patterns in the multi-species scenario ${\chi^{2}}_{`r NA`,`r NA`} = `r NA`$, $P < `r NA`$) and the multi-species with opportunity cost scenario ${\chi^{2}}_{`r NA`,`r NA`} = `r NA`$, $P < `r NA`$).
 Although the prioritizations had similar levels of performance, there were different in the their spatial distributions (Figure 3).
@@
 We wish to alert practitioners to several limitations associated with our analysis.
-First, the size of the planning units we used (23 \times 25 km) is much larger than typically used in conservation planning scenarios (approx.
+First, the size of the planning units we used (23 $\times$ 25 km) is much larger than typically used in conservation planning scenarios (approx.
 1--5 km^2).

### 2016-02-21 article/figures.Rmd
+ 4 species for debugging + fixes to article so it compiles + fixes to nmds analyses so they use sparse distances

--- 2016-02-02 article/figures.Rmd
@@
 ## Figures
+
+```{r, echo=FALSE, fig.width=8, fig.height=3, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations.
+Panel (a) shows the relationship between environmental surrogates and adaptive genetic variation.
+Panel (b) shows that for geographic surrogates and neutral genetic variation.
+Transparent gray lines denote trends for each species, and black lines show general trends across all species.'}
+### prepare data frames for plotting
+#env.correlation.PDF <- filter(correlation.DF, Type='Environmental')
+#
+#%>% left_join(
+#		data.frame(
+#			letter=unname(toupper(cld(full.model.GLHT)$mcletters$Letters)),
+#			Combination=as.character(names(toupper(cld(full.model.GLHT)$mcletters$Letters)))
+#		) %>% mutate(Combination=as.character(Combination)),
+#		by='Combination'
+#	) %>% data.frame()
+
+#geo.correlation.PDF <- filter(correlation.DF, Type='Geographic')
+### make plots
+#suppressWarnings({
+#	p1 <- ggplot() +
+#		geom_boxplot(aes(x=Surrogate.target, y=genetic.held, color=Method, group=Method), data=env.correlation.DF) +
+#		geom_text(aes(x=Surrogate.target, y=letter.height, label=letter),
+#			data=results.PDF, position=position_dodge(0.9)) +
+#		geom_text(aes(x=0.6, y=1.03, label='(a)')) +
+#		xlab('Environmental variation secured (%)') +
+#		ylab('Adaptive genetic\nvariation secured (%)') +
+#		theme_classic()
+#	p2 <- ggplot() +
+#		geom_boxplot(aes(x=Surrogate.target, y=neutral.held, color=Method), data=geo.correlation.DF) +
+#		geom_text(aes(x=Surrogate.target, y=letter.height, label=letter),
+#			data=results.PDF, position=position_dodge(0.9)) +
+#		geom_text(aes(x=0.6, y=1.03, label='(b)')) +
+#		xlab('Geographic variation secured (%)') +
+#		ylab('Neutral genetic\nvariation secured (%)') +
+#		theme_classic()
+#	grid.arrange(p1, p2, nrow=1)
+# })
+
+plot(1)
+
+```
 
@@
 		data.frame(
-			letter=unname(toupper(cld(full.model.GLHT)$mcletters$Letters)),
-			Combination=as.character(names(toupper(cld(full.model.GLHT)$mcletters$Letters)))
+			letter=unname(toupper(cld(posthoc.model.GLHT)$mcletters$Letters)),
+			Combination=as.character(names(toupper(cld(posthoc.model.GLHT)$mcletters$Letters)))
 		) %>% mutate(Combination=as.character(Combination)),

### 2016-02-21 article/supporting_information.Rmd
+ 4 species for debugging + fixes to article so it compiles + fixes to nmds analyses so they use sparse distances

@@
 	ggplot(data=bic.DF) +
-		geom_point(aes(y=BIC, x=G, color=Model)) +
-		geom_line(aes(y=BIC, x=G, color=Model)) +
+		geom_line(aes(y=BIC, x=G, color=Model, linetype=Model)) +
 		geom_point(aes(y=best_stat, x=best_G), size=4, shape=8) +
@@
 		ylab('BIC') +
-		scale_linetype_manual(value=rep(c('solid','dashed'),each=7)) +
-		scale_color_manual(value=rep(gg_color_hue(7),2) +
+		scale_linetype_manual(values=rep(c('solid','dashed'),each=7)) +
+		scale_color_manual(values=rep(gg_color_hue(7),2)) +
 		facet_wrap(~ Species, ncol=4, scales='free_y')
@@
 			species=gsub('_', ' ', unique(spp.samples.DF$species)[i], fixed=TRUE),
-			population=spp.Adegenet.LST[[i]]$grp
+			population=apply(spp.Mclust.LST[[i]]$mclust$z, 1, which.max)
 		)
@@
 	facet_wrap(~ species, ncol=4)
+```
+
+```{r, echo=FALSE, fig.width=15, fig.height=14, fig.cap='Population membership probabilities.
+Each panel depicts a different species.
+Each vertical bar represents a different individual, and colors show the probability that it belongs to different populations.
+This type of plot is also known as a \texttt{Structure} plot.'}
+plot(1)
 ```
@@
 						rbind(
-							data.frmae(
+							data.frame(
 								Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
@@
 								Converged=ifelse(is.null(spp.nmds.LST[[i]][[j]]), NA, spp.nmds.LST[[i]][[j]]$converged)
+							)
 						)

### 2016-02-23 article/supporting_information.Rmd
+ assorted fixes + grob dumps for debugging

@@
 grid.FPLY <- spFortify(grid.PLY)
+
+# define ggplot2 color wheel function
+gg_color_hue <- function(n) {
+	hues <- seq(15, 375, length=n+1)
+	hcl(h=hues, l=65, c=100)[1:n]
+}
 ```
@@
 	guides(fill=guide_legend(title='Presence')) +
-	theme(axis.ticks=element_blank(), axis.text=element_blank(),
-		axis.line=element_blank(), 
+	theme(
+		line = element_blank(), rect = element_blank(), 
+		text = element_blank(), axis.ticks.length = unit(0, "cm"),
+		legend.position = "none", panel.margin = unit(0, "lines"),
+		plot.margin = unit(c(0, 0, 0, 0), "lines"),
 		strip.background = element_rect(fill='grey20'),
@@
 	theme_classic() +
-	theme(axis.ticks=element_blank(), axis.text=element_blank(),
-		axis.line=element_blank(), axis.title=element_blank(),
+	theme(
+		line = element_blank(), rect = element_blank(), 
+		text = element_blank(), axis.ticks.length = unit(0, "cm"),
+		legend.position = "none", panel.margin = unit(0, "lines"),
+		plot.margin = unit(c(0, 0, 0, 0), "lines"),
 		panel.border=element_rect(color='black', fill=NA, size=1)
@@
 				theme_classic() +
-				theme(axis.ticks=element_blank(), axis.text=element_blank(),
-					plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(), 
-					axis.title=element_blank(), 
+				theme(
+					line = element_blank(), rect = element_blank(), 
+					text = element_blank(), axis.ticks.length = unit(0, "cm"),
+					legend.position = "none", panel.margin = unit(0, "lines"),
+					plot.margin = unit(c(0, 0, 0, 0), "lines"),
 					panel.border=element_rect(color='black', fill=NA, size=1),
@@
 )
-# define ggplot2 color wheel function
-gg_color_hue <- function(n) {
-	hues <- seq(15, 375, length=n+1)
-	hcl(h=hues, l=65, c=100)[1:n]
-}
 # make plot
@@
 Squares denote planning units.
-Points denote the location of sampled individuals, and their colors denote the distribution of populations.
-Note that points have been slightly jittered so that individuals in close proximity are visible.'}
+Pie charts denote the location of sampled individuals, and colors denote the probability that they belong to different populations.
+Note that pie charts have been jittered to make individuals in close proximity visible.'}
 # prepare data
-pop.DF <- llply(
+pop.probs.DF <- llply(
 	seq_along(unique(spp.samples.DF$species)),
@@
 		) %>% mutate(
-			species=gsub('_', ' ', unique(spp.samples.DF$species)[i], fixed=TRUE),
-			population=apply(spp.Mclust.LST[[i]]$mclust$z, 1, which.max)
+			Individual=seq_along(species)
 		) 	
+		curr.probs <- spp.Mclust.LST[[i]]$mclust$z %>%
+			as.data.frame() %>%
+			mutate(
+				species=unique(spp.samples.DF$species)[i],
+				Individual=seq_along(species)) %>%
+			gather(Population,Probability,-species,-Individual) %>%
+			mutate(Population=as.numeric(gsub('V', '', Population, fixed=TRUE))) %>%
+			left_join(curr.pts, by=c('species','Individual')) %>%
+			mutate(species=gsub('_', ' ', species, fixed=TRUE)) %>%
+			arrange(species,Individual,Population)
+		return(curr.probs)
 	}
 ) %>% do.call(what=rbind)
-# make plot
-ggplot() +
+# create legend colors
+legend.CHR <- `names<-`(
+	gg_color_hue(n_distinct(pop.probs.DF$Population)),
+	as.character(unique(pop.probs.DF$Population))
+)
+# create grobs
+grobLST <- llply(
+	seq_along(unique(pop.probs.DF$species)),
+	function(i) {
+		# init
+		curr.spp.DF <- filter(pop.probs.DF, species==unique(pop.probs.DF$species)[i])
+		# create base mapmap
+		curr.GROB <- ggplot(data=curr.spp.DF) +
 			geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
@@
 			geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
-		group='group'), fill='transparent', color='grey10') +
-	geom_point(data=pop.DF, aes_string(x='sample.longitude', y='sample.latitude', 
-		color='population'), alpha=0.8, size=3, position='jitter') +
+				group='group'), fill='grey70', color='grey10') +
 			guides(fill=guide_legend(title=''), color=guide_legend(title='')) +
 			theme_classic() +
-	theme(axis.ticks=element_blank(), axis.text=element_blank(),
-		plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(), 
-		axis.title=element_blank(), 
+			theme(
+				line = element_blank(), rect = element_blank(), 
+				text = element_blank(), axis.ticks.length = unit(0, "cm"),
+				legend.position = "none", panel.margin = unit(0, "lines"),
+				plot.margin = unit(c(0, 0, 0, 0), "lines"),
 				strip.background = element_rect(fill='grey20'),
@@
 			) +
-	facet_wrap(~ species, ncol=4)
+			scale_fill_manual(
+				name='',
+				values=legend.CHR[seq_len(n_distinct(curr.spp.DF$Population))]
+			)
+		# create bbox object
+		bbox <- c(ggplot_build(curr.GROB)$panel$ranges[[1]]$x.range,
+			ggplot_build(curr.GROB)$panel$ranges[[1]]$y.range)
+		names(bbox) <- c('ll.lon', 'ur.lon', 'll.lat', 'ur.lat')
+		# render grob
+		grid.newpage()
+		pushViewport(viewport(layout=grid.layout(100,100)))
+		vplayout <- function(x,y) viewport(layout.pos.row=x, layout.pos.col=y)
+		print(curr.GROB, vp=vplayout(1:100,1:100))
+		# add in pies
+		for (j in unique(curr.spp.DF$cell)) {
+			# init
+			curr.cell.DF <- filter(curr.spp.DF, cell==j) %>% 
+				group_by(cell, Population) %>%
+				summarize(
+					grid.longitude=mean(grid.longitude), 
+					grid.latitude=mean(grid.latitude),
+					Probability=mean(Probability)
+				) %>% data.frame()
+			curr.lon <- (curr.cell.DF$grid.longitude[1]-bbox['ll.lon'])/abs(bbox['ur.lon']-bbox['ll.lon'])
+			curr.lat <- (curr.cell.DF$grid.latitude[1]-bbox['ll.lat'])/abs(bbox['ur.lat']-bbox['ll.lat'])
+			# create pie chart for individual
+			curr.pie <- ggplot(curr.cell.DF, aes(x="", y=Probability, fill=factor(Population))) +
+				geom_bar(width=0.75, color='grey20', stat='identity') +
+				coord_polar(theta='y') +
+				theme(
+					line = element_blank(), rect = element_blank(), 
+					text = element_blank(), axis.ticks.length = unit(0, "cm"),
+					legend.position = "none", panel.margin = unit(0, "lines"),
+					plot.margin = unit(c(0, 0, 0, 0), "lines"), complete = TRUE
+				) +
+				scale_fill_manual(
+					name='',
+					values=legend.CHR[seq_len(n_distinct(curr.spp.DF$Population))]
+				)
+			# render pie
+			print(curr.pie, vp=viewport(x=curr.lon, y=curr.lat, width=0.05, height=0.05))
+		}
+	}
+)
 ```
@@
 This type of plot is also known as a \texttt{Structure} plot.'}
-plot(1)
+# prepare data
+pop.probs.DF <- ldply(
+	seq_along(spp.StructureData.LST),
+	function(i) {
+		spp.Mclust.LST[[i]]$mclust$z
+			as.data.frame() %>%
+			mutate(Species=unique(spp.samples.DF$species)[i]) %>%
+			gather(Population,Probability,-Species) %>%
+			mutate(Population=as.numeric(gsub('V', '', Population, fixed=TRUE)),
+				Individual=seq_along(Population)
+				) %>%
+			return()
+	}
+)
+# make plot
+ggplot(
+		aes(x=Individual, y=Probability, fill=Population)
+		data=pop.probs.DF) +
+	geom_bar() +
+	theme_classic() +
+	theme(
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white', face='italic'),
+
+	) +
+	xlab('') +
+	ylab('Probability of population membership (%)') +
+	facet_wrap(~ Species)
 ```
@@
 		theme_classic() +
-		theme(axis.ticks=element_blank(), axis.text=element_blank(),
-			plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(),
+		theme(
+			line = element_blank(), rect = element_blank(), 
+			text = element_blank(), axis.ticks.length = unit(0, "cm"),
+			legend.position = "none", panel.margin = unit(0, "lines"),
+			plot.margin = unit(c(0, 0, 0, 0), "lines"),
 			axis.title=element_blank(), strip.background = element_rect(fill='grey20'),

### 2016-02-24 article/figures.Rmd
+ fixes to supp material and figures

@@
 Panel (b) shows that for geographic surrogates and neutral genetic variation.
-Transparent gray lines denote trends for each species, and black lines show general trends across all species.'}
-### prepare data frames for plotting
-#env.correlation.PDF <- filter(correlation.DF, Type='Environmental')
-#
-#%>% left_join(
-#		data.frame(
-#			letter=unname(toupper(cld(full.model.GLHT)$mcletters$Letters)),
-#			Combination=as.character(names(toupper(cld(full.model.GLHT)$mcletters$Letters)))
-#		) %>% mutate(Combination=as.character(Combination)),
-#		by='Combination'
-#	) %>% data.frame()
+Box-plots show variation in the proportion of genetic variation secured using different surrogate-based targets.
+Lines show the general trends.'}
+# make figure
+p1 <- ggplot(mapping=aes(x=factor(Surrogate.target),y=genetic.held, color=Method)) +
+	geom_boxplot(data=env.correlation.DF) +
+	xlab('Environmental surrogate target (%)') + 
+	ylab('Adaptive genetic variation secured (%)') +
+	geom_text(aes(x=0.05, y=0.9, label='(a)')) +
+	geom_line(data=data.frame(summarize(group_by(env.correlation.DF, Surrogate.target, Method), genetic.held=mean(genetic.held)))) +
+	scale_color_manual(name='Prioritization', values=c('Optimal'='black', 'Random'='grey50'))
 
-#geo.correlation.PDF <- filter(correlation.DF, Type='Geographic')
-### make plots
-#suppressWarnings({
-#	p1 <- ggplot() +
-#		geom_boxplot(aes(x=Surrogate.target, y=genetic.held, color=Method, group=Method), data=env.correlation.DF) +
-#		geom_text(aes(x=Surrogate.target, y=letter.height, label=letter),
-#			data=results.PDF, position=position_dodge(0.9)) +
-#		geom_text(aes(x=0.6, y=1.03, label='(a)')) +
-#		xlab('Environmental variation secured (%)') +
-#		ylab('Adaptive genetic\nvariation secured (%)') +
-#		theme_classic()
-#	p2 <- ggplot() +
-#		geom_boxplot(aes(x=Surrogate.target, y=neutral.held, color=Method), data=geo.correlation.DF) +
-#		geom_text(aes(x=Surrogate.target, y=letter.height, label=letter),
-#			data=results.PDF, position=position_dodge(0.9)) +
-#		geom_text(aes(x=0.6, y=1.03, label='(b)')) +
-#		xlab('Geographic variation secured (%)') +
-#		ylab('Neutral genetic\nvariation secured (%)') +
-#		theme_classic()
-#	grid.arrange(p1, p2, nrow=1)
-# })
+p2 <- ggplot2(mapping=aes(x=factor(Surrogate.target),y=genetic.held, color=Method)) +
+	geom_boxplot(data=geo.correlation.DF) +
+	xlab('Geographic surrogate target (%)') + 
+	ylab('Neutral genetic variation secured (%)') +
+	geom_text(aes(x=0.05, y=0.9, label='(b)')) +
+	geom_line(data=data.frame(summarize(group_by(geo.correlation.DF, Surrogate.target, Method), genetic.held=mean(genetic.held)))) +
+	theme_classic() +
+	scale_color_manual(name='Prioritization', values=c('Optimal'='black', 'Random'='grey50'))
 
-plot(1)
+# render figure
+grid.arrange(p1,p2,nrow=1)
 

### 2016-02-24 article/supporting_information.Rmd
+ fixes to supp material and figures

@@
 	theme(
-		line = element_blank(), rect = element_blank(), 
-		text = element_blank(), axis.ticks.length = unit(0, "cm"),
-		legend.position = "none", panel.margin = unit(0, "lines"),
-		plot.margin = unit(c(0, 0, 0, 0), "lines"),
+		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
 		strip.background = element_rect(fill='grey20'),
@@
 
-```{r, echo=FALSE, fig.width=4, fig.height=3.5, fig.cap="Species richness. Squares denote planning units. Planning units with a brighter color are inhabited by more species."}
+```{r, echo=FALSE, fig.width=4, fig.height=2.5, fig.cap="Species richness. Squares denote planning units. Planning units with a brighter color are inhabited by more species."}
 # calculate species richness
@@
 	select(5:(4+n_distinct(spp.samples.DF$species))) %>% as.matrix() %>% rowSums()
-
 # plot species richness
@@
 		group=group, fill=Species_richness), color='grey10') +
-	scale_fill_gradientn(name='Count (#)',
+	scale_fill_gradientn(name='Richness (#)',
 		colors=brewer.pal(9,'PuBu'),
@@
 	theme(
-		line = element_blank(), rect = element_blank(), 
-		text = element_blank(), axis.ticks.length = unit(0, "cm"),
-		legend.position = "none", panel.margin = unit(0, "lines"),
-		plot.margin = unit(c(0, 0, 0, 0), "lines"),
+		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"), axis.title=element_blank(),
 		panel.border=element_rect(color='black', fill=NA, size=1)
@@
 		ylim=buffered.range(grid.FPLY$lat, 0)
-	) + 
-	ggtitle('Species Richness')
+	)
 ```
 
-```{r, echo=FALSE, fig.width=4, fig.height=3.5, fig.cap="Opportunity cost.
+```{r, echo=FALSE, fig.width=4, fig.height=2.5, fig.cap="Opportunity cost.
 Squares denote planning units.
@@
 		ylim=buffered.range(grid.FPLY$lat, 0)
-	) +
-	ggtitle('Opportunity cost')
+	)
 ```
@@
 				theme(
-					line = element_blank(), rect = element_blank(), 
-					text = element_blank(), axis.ticks.length = unit(0, "cm"),
-					legend.position = "none", panel.margin = unit(0, "lines"),
-					plot.margin = unit(c(0, 0, 0, 0), "lines"),
+					axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"), axis.title=element_blank(),
 					panel.border=element_rect(color='black', fill=NA, size=1),
@@
 
-```{r, echo=FALSE, fig.width=15, fig.height=14, fig.cap='Distribution of populations.
-Each panel depicts a different species.
-Squares denote planning units.
-Pie charts denote the location of sampled individuals, and colors denote the probability that they belong to different populations.
-Note that pie charts have been jittered to make individuals in close proximity visible.'}
+```{r, echo=FALSE}
 # prepare data
@@
 )
-# create grobs
-grobLST <- llply(
-	seq_along(unique(pop.probs.DF$species)),
-	function(i) {
+# create function
+plot.spp.pop.probs <- function(i) {
 	# init
+	prop <- 0.85
 	curr.spp.DF <- filter(pop.probs.DF, species==unique(pop.probs.DF$species)[i])
-		# create base mapmap
+	# create base map grob
 	curr.GROB <- ggplot(data=curr.spp.DF) +
@@
 		theme(
-				line = element_blank(), rect = element_blank(), 
-				text = element_blank(), axis.ticks.length = unit(0, "cm"),
+			axis.title=element_blank(), axis.text=element_blank(),
+			axis.line=element_blank(), axis.ticks=element_blank(),
+			axis.ticks.length = unit(0, "cm"),
 			legend.position = "none", panel.margin = unit(0, "lines"),
@@
 		)
+	# create header grob
+	curr.GROBH <- ggplot() +
+			geom_text(aes(x=0,y=0,label=species),
+				data=curr.spp.DF[1,,drop=FALSE],
+				color='white', fontface='italic', size=4) +
+		theme_classic() +
+		theme(
+			axis.title=element_blank(), axis.text=element_blank(),
+			axis.line=element_blank(), axis.ticks=element_blank(),
+			axis.ticks.length = unit(0, "cm"),
+			legend.position = "none", panel.margin = unit(0, "lines"),
+			plot.margin = unit(c(0, 0, 0, 0), "lines"),
+			panel.background=element_rect(fill='grey20')
+		) +
+		coord_cartesian(xlim=c(-1,1), ylim=c(-1,1))
 	# create bbox object
@@
 	grid.newpage()
-		pushViewport(viewport(layout=grid.layout(100,100)))
-		vplayout <- function(x,y) viewport(layout.pos.row=x, layout.pos.col=y)
-		print(curr.GROB, vp=vplayout(1:100,1:100))
+	pushViewport(viewport(layout=grid.layout(2,1, heights=unit(c(1-prop, prop), "npc"))))
+	print(curr.GROBH, vp=viewport(layout.pos.row=1, layout.pos.col=1))
+	print(curr.GROB, vp=viewport(layout.pos.row=2, layout.pos.col=1))
 	# add in pies
@@
 		curr.lon <- (curr.cell.DF$grid.longitude[1]-bbox['ll.lon'])/abs(bbox['ur.lon']-bbox['ll.lon'])
-			curr.lat <- (curr.cell.DF$grid.latitude[1]-bbox['ll.lat'])/abs(bbox['ur.lat']-bbox['ll.lat'])
+		curr.lat <- ((curr.cell.DF$grid.latitude[1]-bbox['ll.lat'])/abs(bbox['ur.lat']-bbox['ll.lat']))*prop
 		# create pie chart for individual
@@
 		# render pie
-			print(curr.pie, vp=viewport(x=curr.lon, y=curr.lat, width=0.05, height=0.05))
+		print(curr.pie, vp=viewport(x=curr.lon, y=curr.lat, width=0.075, height=0.075))
 	}
 }
-)
 ```
 
-```{r, echo=FALSE, fig.width=15, fig.height=14, fig.cap='Population membership probabilities.
-Each panel depicts a different species.
-Each vertical bar represents a different individual, and colors show the probability that it belongs to different populations.
-This type of plot is also known as a \texttt{Structure} plot.'}
-# prepare data
-pop.probs.DF <- ldply(
-	seq_along(spp.StructureData.LST),
-	function(i) {
-		spp.Mclust.LST[[i]]$mclust$z
-			as.data.frame() %>%
-			mutate(Species=unique(spp.samples.DF$species)[i]) %>%
-			gather(Population,Probability,-Species) %>%
-			mutate(Population=as.numeric(gsub('V', '', Population, fixed=TRUE)),
-				Individual=seq_along(Population)
-				) %>%
-			return()
-	}
-)
-# make plot
-ggplot(
-		aes(x=Individual, y=Probability, fill=Population)
-		data=pop.probs.DF) +
-	geom_bar() +
-	theme_classic() +
-	theme(
-		strip.background = element_rect(fill='grey20'),
-		strip.text = element_text(color='white', face='italic'),
+```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_',' ',unique(spp.samples.DF$species)[1], fixed=TRUE),'} populations.
+Squares denote planning units.
+Pie charts denote quadrats where individuals were sampled, and colors denote the average probability that individuals in the quadrat belong to different populations on average.')}
+plot.spp.pop.probs(1)
+```
 
-	) +
-	xlab('') +
-	ylab('Probability of population membership (%)') +
-	facet_wrap(~ Species)
+```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[2],fixed=TRUE),'} populations.
+See Figure SX caption for conventions.')}
+if (2 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(2)
+```
+
+```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[3],fixed=TRUE),'} populations.
+See Figure SX caption for conventions.')}
+if (3 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(3)
+```
+
+```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[4],fixed=TRUE),'} populations.
+See Figure SX caption for conventions.')}
+if (4 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(4)
+```
+
+```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[5],fixed=TRUE),'} populations.
+See Figure SX caption for conventions.')}
+if (5 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(5)
+```
+
+```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[6],fixed=TRUE),'} populations.
+See Figure SX caption for conventions.')}
+if (6 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(6)
+```
+
+```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[7],fixed=TRUE),'} populations.
+See Figure SX caption for conventions.')}
+if (7 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(7)
+```
+
+```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[8],fixed=TRUE),'} populations.
+See Figure SX caption for conventions.')}
+if (8 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(8)
+```
+
+```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[9],fixed=TRUE),'} populations.
+See Figure SX caption for conventions.')}
+if (9 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(9)
+```
+
+```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[10],fixed=TRUE),'} populations.
+See Figure SX caption for conventions.')}
+if (10 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(10)
+```
+
+```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[11],fixed=TRUE),'} populations.
+See Figure SX caption for conventions.')}
+if (12 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(11)
+```
+
+```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[12],fixed=TRUE),'} populations.
+See Figure SX caption for conventions.')}
+if (12 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(12)
+```
+
+```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[13],fixed=TRUE),'} populations.
+See Figure SX caption for conventions.')}
+if (13 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(13)
+```
+
+```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[14],fixed=TRUE),'} populations.
+See Figure SX caption for conventions.')}
+if (14 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(14)
+```
+
+```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[15],fixed=TRUE),'} populations.
+See Figure SX caption for conventions.')}
+if (15 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(15)
+```
+
+```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[16],fixed=TRUE),'} populations.
+See Figure SX caption for conventions.')}
+if (16 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(16)
+```
+
+```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[17],fixed=TRUE),'} populations.
+See Figure SX caption for conventions.')}
+if (17 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(17)
+```
+
+```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[18],fixed=TRUE),'} populations.
+See Figure SX caption for conventions.')}
+if (18 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(18)
+```
+
+```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[19],fixed=TRUE),'} populations.
+See Figure SX caption for conventions.')}
+if (19 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(19)
+```
+
+```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[20],fixed=TRUE),'} populations.
+See Figure SX caption for conventions.')}
+if (20 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(20)
+```
+
+```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[21],fixed=TRUE),'} populations.
+See Figure SX caption for conventions.')}
+if (21 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(21)
+```
+
+```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[22],fixed=TRUE),'} populations.
+See Figure SX caption for conventions.')}
+if (22 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(22)
+```
+
+```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[23],fixed=TRUE),'} populations.
+See Figure SX caption for conventions.')}
+if (23 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(23)
+```
+
+```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[24],fixed=TRUE),'} populations.
+See Figure SX caption for conventions.')}
+if (24 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(24)
+```
+
+```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[25],fixed=TRUE),'} populations.
+See Figure SX caption for conventions.')}
+if (25 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(25)
+```
+
+```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[26],fixed=TRUE),'} populations.
+See Figure SX caption for conventions.')}
+if (26 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(26)
+```
+
+```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[27],fixed=TRUE),'} populations.
+See Figure SX caption for conventions.')}
+if (27 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(27)
 ```
@@
 		theme(
-			line = element_blank(), rect = element_blank(), 
-			text = element_blank(), axis.ticks.length = unit(0, "cm"),
-			legend.position = "none", panel.margin = unit(0, "lines"),
-			plot.margin = unit(c(0, 0, 0, 0), "lines"),
-			axis.title=element_blank(), strip.background = element_rect(fill='grey20'),
-			strip.text = element_text(color='white', size=20),
-			legend.text = element_text(size=20)
+			axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"), axis.title=element_blank(),
+			panel.border=element_rect(color='black', fill=NA, size=1),
+			strip.background = element_rect(fill='grey20'),
+			strip.text = element_text(color='white', size=20)
 		) +

### 2016-02-24 article/article.Rmd
+ update methods + replace % variation explained with 0 if less than 0

@@
 We used species distribution and genomic data collected by the IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for methodological details].
-This dataset was particularly well suited for this study, because it provides spatially explicit genomic data for multiple species with vastly different demographic characteristics and evolutionary histories across a broad geographic area with large environmental gradients.
+This data set was particularly well suited for this study, because it provides spatially explicit genomic data for multiple species with vastly different demographic characteristics and evolutionary histories across a broad geographic area with large environmental gradients.
 As part of this project, data was collected for 27 alpine plant species using a 20' longitude by 21' latitude grid (approx.
@@
 They genotyped samples using amplified fragment length polymorphisms [AFLP; @r453], and constructed matrices denoting the presence/absence of polymorphisms at loci for each species.
-Thus for each second grid cell, for each of the species present in the grid cell, the dataset contains information describing the genomic variation of individuals in that grid cell.
+Thus for each second grid cell, for each of the species present in the grid cell, the data set contains information describing the genomic variation of individuals in that grid cell.
 We used these grid cells as planning units to develop prioritizations investigating the effectiveness of surrogates for the (AFLP) genetic variation.
@@
 r@475]--is that they do not use environmental data to identify loci under selection, which would have biased our analysis.
-In an earlier study, Bothwell _et al._ [-@r455] applied such outlier detection methods to one of the species in this dataset (_Gentiana nivalis_).
-Here, we adopted a similar methodology and applied it to each of the 27 species in the dataset.
+In an earlier study, Bothwell _et al._ [-@r455] applied such outlier detection methods to one of the species in this data set (_Gentiana nivalis_).
+Here, we adopted a similar methodology and applied it to each of the 27 species in the data set.
 
-First, we grouped individuals into genetic lineages for each species.
+First, we grouped individuals into genetic lineages for each species (using a methodology similar to that implemented in \texttt{prabclust R} package; @r479).
 The purpose of this step was to avoid misclassifying loci as adaptive.
-For each species, we used non-metric multi-dimensional scaling [NMDS; implemented in the $\texttt{vegan R}$ package] to transform the binary loci data into continuous variables (Table S2).
+For each species, we ordinated the genomic data and fit models to the ordinations to group individuals into populations.
+Specifically, we used non-metric multi-dimensional scaling [NMDS; implemented in the $\texttt{vegan R}$ package] to transform the binary loci data into continuous variables (Table S2).
 We used Gower distances [@r464] to express the genomic differences between individuals [using the $\texttt{cluster R}$ package to accommodate sparsity; @r457].
-We ran successive scaling analyses with increasing dimensionality until ordinations were obtained with an adequate stress value ($\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for all models).
+To ensure that the ordinations described an adequate amount of the genomic variation, we ran successive scaling analyses with increasing dimensionality to obtained ordinations until this was achieved (stress value $\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for each analysis).
 We then fit a suite of Gaussian mixture to the ordinated values to estimate the probability that individuals belonged to different populations [using the default priors and covariance structures implemented in the $\texttt{mclust} R$ package; @r477; @r476].
@@
 We identified loci under selection (Table S3).
-For each species, we ran $\texttt{Bayescan}$ [version 2.1; @r456] using the population memberships identified in the previous analysis (1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations).
-To avoid false-positive, we omitted loci where the global frequency of the minor allele was $\geq 0.05$ (following guidelines in the $\texttt{Bayescan}$ manual).
+Since outlier locus detection methods compare allele frequencies between multiple populations, we were unable to investigate species that only contained a single population.
+For these species ($n = `r sum(sapply(spp.Mclust.LST, function(x) return(ncol(x$mclust$z)==1)))`$), we assumed that all genetic variation was neutral.
+For species with more than two populations, we ran $\texttt{Bayescan}$ [version 2.1; @r456; 1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations].
+To avoid false-positives, we omitted loci where the global frequency of the minor allele was $\geq 0.05$ (following guidelines in the $\texttt{Bayescan}$ manual).
 We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate to ensure convergence [FDR $\leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].
@@
 After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species (Figures S6--S32).
-For each species, we partitioned adaptive and neutral loci into separate datasets, and reran the nonmetric multidimensional scaling analysis (NMDS) to derive continuous variables that described the main gradients of adaptive and neutral genetic variation separately (Table S2).
+For each species, we partitioned adaptive and neutral loci into separate data sets, and reran the nonmetric multidimensional scaling analysis (NMDS) to derive continuous variables that described the main gradients of adaptive and neutral genetic variation separately (Table S2).
 Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the individuals in each grid cell to summarize the typical genetic characteristics of individuals in the cells.

### 2016-02-24 R/analysis/06-surrogacy-prioritizations.R

@@ -177,7 +177,12 @@ env.correlation.DF <- ldply(
 			genetic.held=adaptive.held
 		)
 	}
+) %>% mutate(
+	genetic.held=replace(genetic.held, which(genetic.held<0), 0),
+	adaptive.held=replace(adaptive.held, which(adaptive.held<0), 0),
+	neutral.held=replace(neutral.held, which(neutral.held<0),0)
 )
+
 
 geo.correlation.DF <- ldply(
 	seq_along(geo.correlation.prioritisations), 
@@ -205,7 +210,12 @@ geo.correlation.DF <- ldply(
 			genetic.held=neutral.held
 		)
 	}
+) %>% mutate(
+	genetic.held=replace(genetic.held, which(genetic.held<0), 0),
+	adaptive.held=replace(adaptive.held, which(adaptive.held<0), 0),
+	neutral.held=replace(neutral.held, which(neutral.held<0),0)
 )
+
 
 correlation.DF <- rbind(env.correlation.DF, geo.correlation.DF)
 

### 2016-02-24 R/analysis/08-multi-species-prioritisations-no-cost.R

@@ -42,7 +42,12 @@ multi.spp.DF <- ldply(
 			Prioritisation=c('Amount','Surrogate','Genetic')[i]
 		)
 	}
+)  %>% mutate(
+	genetic.held=replace(genetic.held, which(genetic.held<0), 0),
+	adaptive.held=replace(adaptive.held, which(adaptive.held<0), 0),
+	neutral.held=replace(neutral.held, which(neutral.held<0),0)
 )
+
 
 ## save .rda
 save.session('results/.cache/08-multi-species-prioritisations-no-cost.rda')

### 2016-02-24 R/analysis/09-multi-species-prioritisations-with-cost.R

@@ -42,7 +42,12 @@ multi.spp.with.cost.DF <- ldply(
 			Prioritisation=c('Amount','Surrogate','Genetic')[i]
 		)
 	}
+)  %>% mutate(
+	genetic.held=replace(genetic.held, which(genetic.held<0), 0),
+	adaptive.held=replace(adaptive.held, which(adaptive.held<0), 0),
+	neutral.held=replace(neutral.held, which(neutral.held<0),0)
 )
+
 
 ## save .rda
 save.session('results/.cache/09-multi-species-prioritisations-with-cost.rda')

### 2016-02-24 article/figures.Rmd
+ bug fix + fixes to article and caption

@@
 ```{r, echo=FALSE, fig.width=8, fig.height=3, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations.
-Panel (a) shows the relationship between environmental surrogates and adaptive genetic variation.
-Panel (b) shows that for geographic surrogates and neutral genetic variation.
-Box-plots show variation in the proportion of genetic variation secured using different surrogate-based targets.
-Lines show the general trends.'}
+Panel (a) shows the relationship between environmental surrogate targets used to generate a prioritization and the proportion of adaptive genetic variation it secures.
+Similarly, panel (b) shows that for geographic surrogate targets and neutral genetic variation.
+Box-plots show inter-specific variation in the proportion of genetic variation secured using different surrogate-based targets.
+Lines indicate the general trends in the data.'}
 # make figure
@@
 	geom_text(aes(x=0.05, y=0.9, label='(a)')) +
-	geom_line(data=data.frame(summarize(group_by(env.correlation.DF, Surrogate.target, Method), genetic.held=mean(genetic.held)))) +
+	geom_line(data=data.frame(summarize(group_by(env.correlation.DF, Surrogate.target, Method), genetic.held=median(genetic.held)))) +
 	scale_color_manual(name='Prioritization', values=c('Optimal'='black', 'Random'='grey50'))
@@
 	geom_text(aes(x=0.05, y=0.9, label='(b)')) +
-	geom_line(data=data.frame(summarize(group_by(geo.correlation.DF, Surrogate.target, Method), genetic.held=mean(genetic.held)))) +
+	geom_line(data=data.frame(summarize(group_by(geo.correlation.DF, Surrogate.target, Method), genetic.held=median(genetic.held)))) +
 	theme_classic() +

### 2016-02-25 article/article.Rmd
+ bug fix + fixes to article and caption
+ fixes to manuscript figures and main text

@@
 Intra-specific genetic variation underpins species persistence, and there has been increasing interest in understanding how to design protected areas networks in a way that maximizes this variation [@r67; @r3; @r61].
-Broadly speaking, genetic variation can be classified as either adaptive or neutral [reviewed in @r????].
-Adaptive genetic variation is associated with loci currently under selection.
+Broadly speaking, genetic variation can be classified as either adaptive or neutral [reviewed in @r48].
+Adaptive genetic variation is associated with loci that affect fitness.
 By preserving existing patterns of adaptive variation, protected areas can ensure that populations with particularly beneficial adaptations are not lost [@r3].
-In contrast, neutral genetic variation is associated with loci that are not under natural selection.
+In contrast, neutral genetic variation is associated with loci that do not affect fitness.
 Neutral variation often reflects the evolutionary history of different populations.
 By preserving existing patterns of neutral variation, protected areas can secure different genetic lineages and avoid the adverse effects of low genetic diversity [eg.
-inbreeding depression and reduced capacity to respond to environmental change; @r67].
+inbreeding depression; @r67].
 However, genetic data requires considerable investment to obtain and analyze.
 As a consequence, conservation planners do not typically use genetic data to inform prioritizations [@r61].
-Thus many prioritizations may fail to capture species' intra-specific genetic variation, and ultimately fail to deliver long-term conservation outcomes.
+Thus prioritizations may fail to capture species' intra-specific genetic variation, and ultimately fail to deliver long-term conservation outcomes.
 
@@
 Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been used to guide reserve selection [eg.
-habitat type and local climate; @r247; @r2].
-These variables are often used to detect adaptive variation [eg.
-@r475; @r455].
+habitat conditions; @r247; @r2].
 However, their effectiveness as surrogates in conservation planning remains unverified.
 On the other hand, neutral genetic variation arises from a reduction in gene flow between individuals.
-Thus surrogates for neutral variation are based on variables that predict the level of connectivity between planning units.
-For instance, the isolation by distance theory [IBD; @259] predicts populations further apart will have less connectivity,  result in reduced gene flow, and in turn share less neutral genetic variation.
-Geographic distance-based surrogates have been found to improve prioritizations in an island network [@200].
+Surrogates for neutral variation are based on variables that predict the level of connectivity between planning units.
+For instance, the isolation by distance theory [IBD; @259] predicts that populations further apart will have less connectivity, exchange less genetic material, and in turn have less genetic variation in common.
+Based on this theory, geographic distance-based surrogates have been found to improve prioritizations in an island network [@200].
 However this remains untested in spatially contiguous systems where connectivity is complicated by additional factors [eg.
@@
 Before conservation planners can reliably use surrogates for genetic data in conservation planning, the effectiveness of these surrogates needs to be verified.
-Otherwise, conservation planners may waste precious resources on places that do little to benefit biodiversity..
+Otherwise, conservation planners may waste precious resources on places that do little to benefit biodiversity.
 
@@
 We used species distribution and genomic data collected by the IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for methodological details].
-This data set was particularly well suited for this study, because it provides spatially explicit genomic data for multiple species with vastly different demographic characteristics and evolutionary histories across a broad geographic area with large environmental gradients.
+This data set was particularly well suited for our study, because it provides spatially explicit genomic data for multiple species with vastly different demographic characteristics and evolutionary histories across a broad geographic area with large environmental gradients.
 As part of this project, data was collected for 27 alpine plant species using a 20' longitude by 21' latitude grid (approx.
 21' 22.3 km $\times$ 25 km; Figure S1).
-They visited every second grid cell, and if a species was detected in a grid cell, samples were collected from three individuals.
+They visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals.
 They genotyped samples using amplified fragment length polymorphisms [AFLP; @r453], and constructed matrices denoting the presence/absence of polymorphisms at loci for each species.
-Thus for each second grid cell, for each of the species present in the grid cell, the data set contains information describing the genomic variation of individuals in that grid cell.
-We used these grid cells as planning units to develop prioritizations investigating the effectiveness of surrogates for the (AFLP) genetic variation.
-Note that we omitted grid cells that did not contain any individuals from our analysis.
-The majority of our analyses described hereafter were performed using the \texttt{R} software package [version `r paste0(version$major, '.', version$minor)`; @r192].
+Thus for each second grid cell, for each of the species present in the grid cell, the data set contains information describing the genomic variation of individuals in the cell.
+We used these grid cells as planning units to develop prioritizations.
+To reduce computational effort, we only used grid cells that contained at least a single individual ($n = `r nrow(grid.DF)`$).
+The majority of our analyses were performed using the \texttt{R} software package [version `r paste0(version$major, '.', version$minor)`; @r192].
 
 ### Landscape data
-We aimed to investigate the effectiveness of climatic and geographic surrogates.
-To describe the location of each grid cell, we projected the grid into an equi-distant coordinate system ($\texttt{ESRI:102031}$), calculated their centroids, and extracted its two-dimensional coordinates.
-To describe the climatic variation among each cell, we obtained 19 bioclimatic variables [approx.
+We examined the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation (respectively).
+To describe the geographic location of each grid cell, we projected the grid into an equi-distant coordinate system ($\texttt{ESRI:102031}$), calculated their centroids, and extracted their two-dimensional coordinates.
+To describe the environmental characteristics of each grid cell, we obtained 19 bioclimatic variables [approx.
 1 km resolution; @r33].
@@
 ### Outlier locus detection
-To investigate the effectiveness of the surrogates for adaptive and neutral genetic variation, we first need to identify which of the sampled loci--if any--were under selection for each species.
-We used outlier locus detection methods [evaluated in @r454] to achieve this.
+To investigate the effectiveness of the surrogates for adaptive and neutral genetic variation, we first needed to identify which of the sampled loci--if any--were under selection.
+We used outlier locus detection methods [compared in @r454; reviewed in @r480] to achieve this.
 The basic premise underpinning these analyses is that neutral loci are expected to exhibit a specific level of variation among populations ($F_{st}$), and loci that deviate from this expectation should be under selection.
@@
 First, we grouped individuals into genetic lineages for each species (using a methodology similar to that implemented in \texttt{prabclust R} package; @r479).
-The purpose of this step was to avoid misclassifying loci as adaptive.
-For each species, we ordinated the genomic data and fit models to the ordinations to group individuals into populations.
-Specifically, we used non-metric multi-dimensional scaling [NMDS; implemented in the $\texttt{vegan R}$ package] to transform the binary loci data into continuous variables (Table S2).
-We used Gower distances [@r464] to express the genomic differences between individuals [using the $\texttt{cluster R}$ package to accommodate sparsity; @r457].
-To ensure that the ordinations described an adequate amount of the genomic variation, we ran successive scaling analyses with increasing dimensionality to obtained ordinations until this was achieved (stress value $\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for each analysis).
-We then fit a suite of Gaussian mixture to the ordinated values to estimate the probability that individuals belonged to different populations [using the default priors and covariance structures implemented in the $\texttt{mclust} R$ package; @r477; @r476].
+The purpose of this step was to avoid falsely classifying loci as adaptive.
+For each species, we ordinated the genomic data and fit models to the ordinations to fit to turn group individuals into populations.
+Specifically, we used non-metric multi-dimensional scaling [NMDS; implemented in the $\texttt{vegan R}$ package] using Gower distances [@r464; using the $\texttt{cluster R}$ package to accommodate sparsity; @r457] to transform the binary loci data into continuous variables (Table S2).
+To ensure that the ordinations described a sufficient amount of the genomic variation, we ran successive scaling analyses with increasing dimensionality until this was achieved (maximum stress value $\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for each analysis).
+We then fit a suite of Gaussian mixture to the ordinations to estimate the probability that individuals belonged to different populations [assuming different covariance structures implemented in the $\texttt{mclust} R$ package; @r477; @r476].
 For each parametrization, we fit a set of models assuming varying numbers of total populations (\textit{G}= 1--`r max(mclust.params.LST[[MODE]]$G)`).
-We identified the most parsimonious models (lowest Bayesian Information Criterion $BIC$ statistic; Figures S4), and used these to group individuals into populations (Figure S5).
+We identified the most parsimonious model for each species (lowest Bayesian Information Criterion $BIC$ statistic; Figures S4), and used them to group individuals to populations (Figure S5).
 
-We identified loci under selection (Table S3).
-Since outlier locus detection methods compare allele frequencies between multiple populations, we were unable to investigate species that only contained a single population.
-For these species ($n = `r sum(sapply(spp.Mclust.LST, function(x) return(ncol(x$mclust$z)==1)))`$), we assumed that all genetic variation was neutral.
-For species with more than two populations, we ran $\texttt{Bayescan}$ [version 2.1; @r456; 1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations].
-To avoid false-positives, we omitted loci where the global frequency of the minor allele was $\geq 0.05$ (following guidelines in the $\texttt{Bayescan}$ manual).
+We identified the loci under selection (Table S3).
+Since outlier locus detection methods compare allele frequencies between populations, we were unable to examine species that contained a single population.
+For these species ($n = `r sum(sapply(spp.Mclust.LST, function(x) return(ncol(x$mclust$z)==1)))`$), we assumed that all loci were neutral.
+For species with more than a single population, we ran $\texttt{Bayescan}$ [version 2.1; @r456; 1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations].
+To avoid false-positives, we omitted loci where the global frequency of the minor allele was $\geq 0.05$ (following guidelines in the $\texttt{Bayescan}$ manual) and assumed they were neutral.
+Additionally, we omitted individuals from this analysis when their population membership was uncertain (maximum membership probability $\l `r mclust.params.LST[[MODE]]$probability.threshold`$).
 We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate to ensure convergence [FDR $\leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].
@@
 After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species (Figures S6--S32).
-For each species, we partitioned adaptive and neutral loci into separate data sets, and reran the nonmetric multidimensional scaling analysis (NMDS) to derive continuous variables that described the main gradients of adaptive and neutral genetic variation separately (Table S2).
-Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the individuals in each grid cell to summarize the typical genetic characteristics of individuals in the cells.
+For each species, we partitioned adaptive and neutral loci into separate data sets, and reran the successive nonmetric multidimensional scaling analyses (using parameters described above) to derive continuous variables that described the main gradients of adaptive and neutral genetic variation separately (Table S2).
+Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the individuals in each grid cell to express the typical genomic characteristics of individuals in the cells.
 
@@
 We used the unreliable representative and adequate prioritization (URAP) formulation of the reserve selection problem [@r465].
-This formulation can identify the minimum number of planning units required to preserve a proportion of the species' range as well as a proportion of the species' variation using an attribute space [_sensu_ @r465].
-We treated the genomic and surrogate variables as distinct attribute spaces.
-We set the demand points for each species and its attribute space using the values associated with the planning units.
+This formulation can identify the minimum number of planning units required to preserve both a proportion of the species' range (using amount-based targets) and a proportion of the species' intra-specific variation [using space-based targets and attribute spaces; _sensu_ @r465].
+We treated the environmental and geographic surrogate variables as separate attribute spaces.
+We also treated the adaptive (if detected) and neutral variables as separate attribute spaces for each species.
+We set the demand points for each species and each attribute space using the values associated with the planning units.
 For every prioritization we generated, we calculated the proportion of the adaptive (if detected) and neutral genetic variation that every prioritization secured to assess its performance.
@@
 
-To understand the relationship between the surrogate-based targets used to generate a prioritizations and the amount of genetic variation secured, we generated a collection of single-species prioritizations.
-For each species, we generated two sets of prioritizations using increasing environmental and geographic-based targets (respectively).
+To understand the relationship between the surrogate-based targets used to generate a prioritization and the proportion of genetic variation it secures, we generated a collection of single-species prioritizations.
+For each species, we first generated two sets of prioritizations using increasing environmental and geographic-based targets (separately).
 Prioritizations were generated using surrogate-based targets ranging from `r (min(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target)*100)` % and `r (max(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target)*100)` % in increments of `r diff(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target[1:2])*100` % ($n = `r rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.optimal.replicates`$ per target).
-To determine if these surrogate-based prioritizations performed better then random, we generated another two sets of prioritizations by randomly selecting planning units that were occupied by the species ($n = `r rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.random.replicates`$ per surrogate target).
-To account for variation in the number of selected planning units in the surrogate-based prioritizations, the number of selected of planning units in the randomly generated prioritizations were generated following their distribution.
+To determine if these surrogate-based prioritizations performed better then random, we then generated another two sets of prioritizations by randomly selecting planning units that were occupied by the species ($n = `r rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.random.replicates`$ per surrogate target).
+To account for variation in the number of selected planning units in the target-based prioritizations, the number of selected of planning units in the randomly generated prioritizations were generated following their distribution.
+Randomly generated prioritizations associated with negative values--because they secured only a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis.
 
-To determine if the use of surrogate-based targets can actually improve prioritizations in real-world conservation planning scenarios, we used three scenarios with increasing levels of complexity.
+To determine if the use of surrogate-based targets can actually improve prioritizations in real-world conservation planning scenarios, we investigated the effectiveness of surrogate-based targets using three scenarios with increasing levels of complexity.
 To assess the benefits of surrogate-based targets under the simplest of circumstances, our first scenario involved generating single-species prioritizations for each species.
-Most large-scale prioritizations involve a comprehensive set of species, and in turn, may secure a greater proportion of intra-specific genetic variation.
-Our second scenario involved generating multi-species prioritizations.
+However, most large-scale prioritizations involve a comprehensive set of species.
+This property may lead to prioritizations that secure a greater proportion of intra-specific genetic variation for each species by returning solutions that preserve the same species in multiple communities.
+To test for this effect, our second scenario involved generating multi-species prioritizations.
 Finally, most large-scale prioritizations also explicitly consider opportunity cost to ensure prioritizations are cost-effective.
@@
 For each scenario, we generated three types of prioritizations.
-We generated prioritizations using only amount-based, both amount- and surrogate-based targets, and both amount- and genetic-based targets.
-The purely amount-based prioritizations represented those typically generated in conservation planning exercises and serve as our baseline.
-We used `r (rapr.params.LST[[MODE]][['single.species']]$amount.target*100)` % amount-based targets in all prioritizations to ensure an adequate proportion of habitat was secured for each species [@rXXX].
-We set surrogate- and genetic-based targets as `r (rapr.params.LST[[MODE]][['single.species']]$amount.target*100)` % to ensure that most of the repsective variation was secured for each species.
+We generated prioritizations using only amount-based targets, both amount- and surrogate-based targets, and both amount- and genetic-based targets.
+The purely amount-based prioritizations serve as our baseline--they represent prioritizations typically generated in conservation planning exercises.
+We used `r (rapr.params.LST[[MODE]][['single.species']]$amount.target*100)` % amount-based targets in all prioritizations to ensure an adequate proportion of habitat was secured for each species.
+We set surrogate- and genetic-based targets as `r (rapr.params.LST[[MODE]][['single.species']]$amount.target*100)` % to ensure that most of the respective variation was secured for each species.
 To accomodate variation in the high number of optimal solutions for the single-species amount-based prioritizations, we generated `r (rapr.params.LST[[MODE]][['single.species']]$amount.replicates)` replicate prioritizations.
@@
 ## Statistical analysis
-We investigated the effectiveness of prioritizations generated using surrogate-based targets using generalized linear mixed-effects models [GLMMs; using the $\texttt{lme4}$ R package; @r469].
-All models were fit to the proportion of genetic variation that prioritizations secured using a logit link.
-To understand the relationship between the surrogate-based target used to generate a prioritization and how much genetic variation the resulting prioritization secured, we fit a full model containing a continuous quadratic predictor variable for the surrogate target used to generate the prioritization, a categorical variable indicating the type of genetic variation, and a variable for the interaction between the previous variables.
-This model contained random effects for both the intercept and the predictor variables to accommodate variation between species.
-To determine if surrogate-based targets improved the effectiveness of prioritizations under different scenarios, we fit a full model containing categorical variables indicating the scenario used to generate the prioritization, the combination of targets used to generate the prioritizations, and the type of genetic variation measured.
-This model contained intercept-only random effects.
-Both models were subject to step-wise term deletion routines using likelihood ratio tests (LRT) to assess significance (terms were retained when $P < 0.05$).
-Finally, we conducted a post-hoc analysis using Tukey contrasts with a Bonferroni correction [using the $\texttt{multcomp}$ R package; @r470].
+We investigated the effectiveness of prioritizations generated using surrogate-based targets using generalized linear mixed-effects models [GLMMs; using a logit link and the $\texttt{lme4}$ R package; @r469].
+To determine  if the environmental and geographic variables were effective surrogates, we fit a full model to the proportion of genetic variation that the first collection of prioritizations secured using categorical variables to indicate the surrogate-based targets used to generate them, the type of genetic variation (adaptive or neutral), the method using the generate the prioritization (optimal or random), and the interactions between these variables.
+To determine if surrogate-based targets improved the effectiveness of prioritizations under various scenarios, we fit a full model to the proportion of genetic variation that second collection of prioritizations secured using categorical variables to indicate the scenario used to generate the prioritization (single-species, multi-species, multi-species with cost), the combination of targets used to generate the prioritizations (amount-based targtes, amount and surrogate-based targets, amount and genetic-based targets), the type of genetic variation measured (adaptive or neutral), and the interactions between them.
+All model contained intercept- and slope-level random effects to accommodate variation between species.
+Both full models were subject to a step-wise term deletion routine using $\chi^{2}$ tests to assess significance (terms were retained when $P < 0.05$).
+Finally, we conducted post-hoc analysis on each full model using Tukey contrasts with a Bonferroni correction [using the $\texttt{multcomp}$ R package; @r470].
 
@@
 Increases in environmental surrogate-based targets resulted prioritizations that secured more adaptive genetic variation, than geographic surrogate-based targets for neutral variation.
-These results demonstrate that environmental and geographic variables are effective surrogates for genetic variation.
+These results demonstrate that environmental and geographic variables are effective surrogates for genetic variation for these plant species.
 
@@
 We found similar patterns in the multi-species scenario ${\chi^{2}}_{`r NA`,`r NA`} = `r NA`$, $P < `r NA`$) and the multi-species with opportunity cost scenario ${\chi^{2}}_{`r NA`,`r NA`} = `r NA`$, $P < `r NA`$).
-Although the prioritizations had similar levels of performance, there were different in the their spatial distributions (Figure 3).
+Although the prioritizations had similar levels of performance, there were differences in their spatial distributions (Figure 3).
 In particular, the inclusion of cost data in the multi-species prioritizations had the effect of concentrating conservation effort towards areas with higher elevation.
-Overall, these results suggest that while environmental and geographic surrogate variables are effective surrogates, they are not needed to generate prioritizations that secure most of the intra-specific genetic variation.
+Overall, these results suggest that while environmental and geographic surrogate variables are effective surrogates, they are not needed to secure most of the intra-specific genetic variation.
 
@@
 We aimed to investigate the effectiveness of environmental and geographic surrogates for genetic variation in conservation planning.
-This study is the first to empirically test this using a realistic system.
-We found that the inclusion of environmental and geographic surrogates in single-species prioritizations significantly improved the proportion of intra-specific genetic variation they secured.
-But we also found that the same surrogates did little to improve the multi-species prioritizations.
-Our results suggest that conservation planners can use a comprehensive set of species to secure a sufficient proportion of their intra-specific genetic variation.
+Our study is the first to test this in a realistic context.
+We found that environmental and geographic variables are effective surrogates for genetic variation in conservation planning.
+But we also found that just using amount-based targets--typically used in conservation planning--result in prioritizations that secure most of the adaptive and neutral intra-specific genetic variation.
+Our results support existing guidelines on using a comprehensive set of species in conservation planning.
 
@@
 Third, we used AFLP data to characterize the main gradients of genetic variation.
-Whilst next-generation sequencing provides higher resolution genetic information than AFLP data [reviewed in @rXXX], our methodology would still have utilized only the main gradients of the genetic variation to generate prioritizations in a feasible amount of time.
-Overall, the limitations associated with this study are relatively minor.
+Whilst next-generation sequencing provides higher resolution genetic information than AFLP data [reviewed in @r480], our methodology would still have utilized only the main gradients of the genetic variation to generate prioritizations in a feasible amount of time.
+Overall, we are confident that the limitations associated with our study are relatively minor.
 

### 2016-02-25 article/figures.Rmd
+ fixes to manuscript figures and main text

@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=3, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations.
+```{r, echo=FALSE, fig.width=6, fig.height=4.5, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations.
 Panel (a) shows the relationship between environmental surrogate targets used to generate a prioritization and the proportion of adaptive genetic variation it secures.
@@
 Box-plots show inter-specific variation in the proportion of genetic variation secured using different surrogate-based targets.
-Lines indicate the general trends in the data.'}
+Letters indicate significant differences ($P < 0.05$).'}
+# prepare data
+env.correlation.DF$Surrogate.target <- factor(env.correlation.DF$Surrogate.target)
+env.correlation.DF2 <- env.correlation.DF %>%
+	group_by(Surrogate.target, Method, Type) %>% 
+	summarize(genetic.held=max(genetic.held)+0.1) %>%
+	ungroup() %>% mutate(Combination=as.character(paste0(Surrogate.target,'.',Method,'.',Type))) %>% 
+	left_join(
+		data.frame(
+			letter=unname(toupper(cld(posthoc.correlation.GLHT)$mcletters$Letters)),
+			Combination=as.character(names(toupper(cld(posthoc.correlation.GLHT)$mcletters$Letters))),
+			stringsAsFactors=FALSE
+		),
+		by='Combination'
+	) %>% data.frame()
+geo.correlation.DF$Surrogate.target <- factor(geo.correlation.DF$Surrogate.target)
+geo.correlation.DF2 <- geo.correlation.DF %>%
+	group_by(Surrogate.target, Method, Type) %>% 
+	summarize(genetic.held=max(genetic.held)+0.1) %>%
+	ungroup() %>% mutate(Combination=as.character(paste0(Surrogate.target,'.',Method,'.',Type))) %>% 
+	left_join(
+		data.frame(
+			letter=unname(toupper(cld(posthoc.correlation.GLHT)$mcletters$Letters)),
+			Combination=as.character(names(toupper(cld(posthoc.correlation.GLHT)$mcletters$Letters))),
+			stringsAsFactors=FALSE
+		),
+		by='Combination'
+	) %>% data.frame()
+
 # make figure
-p1 <- ggplot(mapping=aes(x=factor(Surrogate.target),y=genetic.held, color=Method)) +
-	geom_boxplot(data=env.correlation.DF) +
+p1 <- ggplot(aes(x=Surrogate.target,y=genetic.held, color=Method),
+		data=env.correlation.DF) +
+	geom_boxplot(position=position_dodge(0.9)) +
 	xlab('Environmental surrogate target (%)') + 
-	ylab('Adaptive genetic variation secured (%)') +
-	geom_text(aes(x=0.05, y=0.9, label='(a)')) +
-	geom_line(data=data.frame(summarize(group_by(env.correlation.DF, Surrogate.target, Method), genetic.held=median(genetic.held)))) +
-	scale_color_manual(name='Prioritization', values=c('Optimal'='black', 'Random'='grey50'))
+	ylab('Adaptive genetic\nvariation secured (%)') +
+	geom_text(aes(x=0.75, y=1, label='(a)'), color='black', size=5) +
+	theme_classic() +
+	theme(legend.position=c(0.225,0.8)) +
+	scale_color_manual(name='Prioritization', values=c('Optimal'='black', 'Random'='grey50')) +
+	geom_text(aes(x=Surrogate.target, y=genetic.held, color=Method, label=letter),
+		data=env.correlation.DF2, position=position_dodge(0.9))
 
-p2 <- ggplot2(mapping=aes(x=factor(Surrogate.target),y=genetic.held, color=Method)) +
-	geom_boxplot(data=geo.correlation.DF) +
+p2 <- ggplot(aes(x=Surrogate.target,y=genetic.held, color=Method),
+			data=geo.correlation.DF) +
+	geom_boxplot(position=position_dodge(0.9)) +
 	xlab('Geographic surrogate target (%)') + 
-	ylab('Neutral genetic variation secured (%)') +
-	geom_text(aes(x=0.05, y=0.9, label='(b)')) +
-	geom_line(data=data.frame(summarize(group_by(geo.correlation.DF, Surrogate.target, Method), genetic.held=median(genetic.held)))) +
+	ylab('Neutral genetic\nvariation secured (%)') +
+	geom_text(aes(x=0.75, y=1, label='(b)'), color='black', size=5) +
 	theme_classic() +
-	scale_color_manual(name='Prioritization', values=c('Optimal'='black', 'Random'='grey50'))
+	theme(legend.position='none') +
+	scale_color_manual(name='Prioritization', values=c('Optimal'='black', 'Random'='grey50')) +
+	geom_text(aes(x=Surrogate.target, y=genetic.held, color=Method, label=letter),
+		data=geo.correlation.DF2, position=position_dodge(0.9))
 
 # render figure
-grid.arrange(p1,p2,nrow=1)
+grid.arrange(p1,p2,nrow=2)
 
@@
 	) %>% mutate(
-		letter.height=max(value, na.rm=TRUE)+0.04,
+		letter.height=max(value, na.rm=TRUE)+0.1,
 		Combination=paste0(Prioritisation,'.',Metric,'.',Context)
@@
 	)
+
 # make plot
@@
 			data=results.PDF, position=position_dodge(0.9)) +
-		geom_text(aes(x=0.6, y=1.03, label=panel.letter),
+		geom_text(aes(x=0.7, y=1, label=panel.letter),
 			data=results.PDF) +
@@
 		xlab('') +
-		ylim(c(NA, 1.05)) +
+		scale_y_continuous(breaks=c(0,0.25,0.5,0.75,1), limits=c(NA, 1.2)) +
 		theme_classic() +

### 2016-03-02 R/analysis/00-initialization.R
+ retrieve replicate solutions from solution pool - as opposed to benders cuts

@@ -52,7 +52,8 @@ suppressMessages(library(bayescanr))
 suppressMessages(library(structurer))
 # devtools::install_github('paleo13/ggplot2')
 suppressMessages(library(ggplot2))
-
+# devtools::install_github('paleo13/rgurobi')
+suppressMessages(library(rgurobi))
 
 # set pander options
 panderOptions('knitr.auto.asis', FALSE)

### 2016-03-02 R/analysis/07-single-species-prioritisations.R

@@ -57,7 +57,8 @@ single.spp.surrogate.prioritisations <- llply(
 				neutral.genetic.targets=NA,
 				Threads=gurobi.params.LST[[MODE]]$Threads,
 				MIPGap=gurobi.params.LST[[MODE]]$MIPGap,
-				NumberSolutions=rapr.params.LST[[MODE]]$single.species$surrogate.replicates
+				NumberSolutions=rapr.params.LST[[MODE]]$single.species$surrogate.replicates,
+				MultipleSolutionsMethod='solution.pool'
 		)
 	}
 )
@@ -75,7 +76,8 @@ single.spp.genetic.prioritisations <- llply(
 				neutral.genetic.targets=rapr.params.LST[[MODE]]$single.species$genetic.target,
 				Threads=gurobi.params.LST[[MODE]]$Threads,
 				MIPGap=gurobi.params.LST[[MODE]]$MIPGap,
-				NumberSolutions=rapr.params.LST[[MODE]]$single.species$genetic.replicates
+				NumberSolutions=rapr.params.LST[[MODE]]$single.species$genetic.replicates,
+				MultipleSolutionsMethod='solution.pool'
 		)
 	}
 )

### 2016-03-02 R/analysis/08-multi-species-prioritisations-no-cost.R

@@ -28,7 +28,8 @@ multi.spp.prioritisations <- llply(
 			neutral.genetic.targets=y[[3]],
 			Threads=gurobi.params.LST[[MODE]]$Threads,
 			MIPGap=gurobi.params.LST[[MODE]]$MIPGap,
-			NumberSolutions=y[[4]]
+			NumberSolutions=y[[4]],
+			MultipleSolutionsMethod='solution.pool'
 		)
 	}
 )

### 2016-03-06 article/article.Rmd
+ retrieve replicate solutions from solution pool - as opposed to benders cuts
+ updates to article + set mclust threads to 21 + set gurobi MIPGap to 0.1

@@
 Number of figures, table and text boxes: 4 \newline
-
 \newpage
-
-<!-- add line numbers -->
-\begin{linenumbers}
 
@@
 Surrogates for neutral variation are based on variables that predict the level of connectivity between planning units.
-For instance, the isolation by distance theory [IBD; @259] predicts that populations further apart will have less connectivity, exchange less genetic material, and in turn have less genetic variation in common.
-Based on this theory, geographic distance-based surrogates have been found to improve prioritizations in an island network [@200].
+For instance, the isolation by distance theory [IBD; @r259] predicts that populations further apart will have less connectivity, exchange less genetic material, and in turn have less genetic variation in common.
+Based on this theory, geographic distance-based surrogates have been found to improve prioritizations in an island network [@r200].
 However this remains untested in spatially contiguous systems where connectivity is complicated by additional factors [eg.
@@
 These layers and the grid cells were projected into an equal-area coordinate system ($\texttt{ESRI:102014}$).
-The bioclimatic variables were clipped to the extent of the grid and subject to a principal components analysis (PCA; using ArcMap version 10.2.2; Table S1).
+The bioclimatic variables were clipped to the extent of the grid and subject to a principal components analysis (PCA; using \texttt{ArcMap} version 10.2.2; Table S1).
 The first two principal components (PCs) cumulatively explained `r round(pca.DF[[4]][2],1)` % of the variation and were used to generate two new layers.
@@
 The advantage of these methods--in contrast with environmental association analyses [eg.
-r@475]--is that they do not use environmental data to identify loci under selection, which would have biased our analysis.
+@r475]--is that they do not use environmental data to identify loci under selection, which would have biased our analysis.
 In an earlier study, Bothwell _et al._ [-@r455] applied such outlier detection methods to one of the species in this data set (_Gentiana nivalis_).
@@
 To ensure that the ordinations described a sufficient amount of the genomic variation, we ran successive scaling analyses with increasing dimensionality until this was achieved (maximum stress value $\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for each analysis).
-We then fit a suite of Gaussian mixture to the ordinations to estimate the probability that individuals belonged to different populations [assuming different covariance structures implemented in the $\texttt{mclust} R$ package; @r477; @r476].
+We then fit a suite of Gaussian mixture to the ordinations to estimate the probability that individuals belonged to different populations [assuming different covariance structures implemented in the \texttt{mclust} R package; @r477; @r476].
 For each parametrization, we fit a set of models assuming varying numbers of total populations (\textit{G}= 1--`r max(mclust.params.LST[[MODE]]$G)`).
@@
 For species with more than a single population, we ran $\texttt{Bayescan}$ [version 2.1; @r456; 1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations].
-To avoid false-positives, we omitted loci where the global frequency of the minor allele was $\geq 0.05$ (following guidelines in the $\texttt{Bayescan}$ manual) and assumed they were neutral.
+To avoid false-positives, we omitted loci where the global frequency of the minor allele was $\geq 0.05$ (following guidelines in the user manual) and assumed they were neutral.
 Additionally, we omitted individuals from this analysis when their population membership was uncertain (maximum membership probability $\l `r mclust.params.LST[[MODE]]$probability.threshold`$).
@@
 ## References
-
-\end{linenumbers}

### 2016-03-11 article/article.Rmd
+ fixes to article

@@
 As part of this project, data was collected for 27 alpine plant species using a 20' longitude by 21' latitude grid (approx.
-21' 22.3 km $\times$ 25 km; Figure S1).
+22.3 km $\times$ 25 km; Figure S1).
 They visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals.
@@
 We calculated the average value for each PC layer in each grid cell, and used these values to characterize the main themes of climatic variation among the grid cells (Figure S2).
-To examine if the influence of opportunity cost, we calculated the total human population density inside each planning unit (obtained at 1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472).
+To examine if the influence of opportunity cost, we calculated the total human population density inside each planning unit [obtained at 1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472].
 
@@
 
-First, we grouped individuals into genetic lineages for each species (using a methodology similar to that implemented in \texttt{prabclust R} package; @r479).
+First, we grouped individuals into genetic lineages for each species [using a methodology similar to that implemented in \texttt{prabclust R} package; @r479].
 The purpose of this step was to avoid falsely classifying loci as adaptive.
 For each species, we ordinated the genomic data and fit models to the ordinations to fit to turn group individuals into populations.
-Specifically, we used non-metric multi-dimensional scaling [NMDS; implemented in the $\texttt{vegan R}$ package] using Gower distances [@r464; using the $\texttt{cluster R}$ package to accommodate sparsity; @r457] to transform the binary loci data into continuous variables (Table S2).
+Specifically, we used non-metric multi-dimensional scaling [NMDS; implemented in the $\texttt{vegan R}$ package; @r458] using Gower distances [@r464; using the $\texttt{cluster R}$ package to accommodate sparsity; @r457] to transform the binary loci data into continuous variables (Table S2).
 To ensure that the ordinations described a sufficient amount of the genomic variation, we ran successive scaling analyses with increasing dimensionality until this was achieved (maximum stress value $\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for each analysis).
@@
 For these species ($n = `r sum(sapply(spp.Mclust.LST, function(x) return(ncol(x$mclust$z)==1)))`$), we assumed that all loci were neutral.
-For species with more than a single population, we ran $\texttt{Bayescan}$ [version 2.1; @r456; 1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations].
+For species with more than a single population, we ran $\texttt{Bayescan}$ [version 2.1; 1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations; @r456].
 To avoid false-positives, we omitted loci where the global frequency of the minor allele was $\geq 0.05$ (following guidelines in the user manual) and assumed they were neutral.

### 2016-03-11 article/figures.Rmd
+ fixes to article

@@
 	group_by(Surrogate.target, Method, Type) %>% 
-	summarize(genetic.held=max(genetic.held)+0.1) %>%
+	summarize(genetic.held=max(genetic.held, na.rm=TRUE)+0.1) %>%
 	ungroup() %>% mutate(Combination=as.character(paste0(Surrogate.target,'.',Method,'.',Type))) %>% 

### 2016-03-15 article/supporting_information.Rmd
+ change format.tabel to format.dataframe + change random replicates to 100 + fix bug in extractResults

@@
 knitr::kable(
-	format.table(
+	format.dataframe(
 		filter(
@@
 knitr::kable(
-	format.table(
+	format.dataframe(
 		ldply(

### 2016-03-17 article/article.Rmd
+ update surrogate-based targets + increase percentages + replace drop1 with manual term removal + use minkowski distances instead of euclid + assorted fixes in manuscript

@@
 To determine if these surrogate-based prioritizations performed better then random, we then generated another two sets of prioritizations by randomly selecting planning units that were occupied by the species ($n = `r rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.random.replicates`$ per surrogate target).
-To account for variation in the number of selected planning units in the target-based prioritizations, the number of selected of planning units in the randomly generated prioritizations were generated following their distribution.
+To account for variation in the number of selected planning units in the target-based prioritizations, the number of selected planning units in the randomly generated prioritizations was equal to the target-based prioritizations.
 Randomly generated prioritizations associated with negative values--because they secured only a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis.
@@
 We investigated the effectiveness of prioritizations generated using surrogate-based targets using generalized linear mixed-effects models [GLMMs; using a logit link and the $\texttt{lme4}$ R package; @r469].
-To determine  if the environmental and geographic variables were effective surrogates, we fit a full model to the proportion of genetic variation that the first collection of prioritizations secured using categorical variables to indicate the surrogate-based targets used to generate them, the type of genetic variation (adaptive or neutral), the method using the generate the prioritization (optimal or random), and the interactions between these variables.
-To determine if surrogate-based targets improved the effectiveness of prioritizations under various scenarios, we fit a full model to the proportion of genetic variation that second collection of prioritizations secured using categorical variables to indicate the scenario used to generate the prioritization (single-species, multi-species, multi-species with cost), the combination of targets used to generate the prioritizations (amount-based targtes, amount and surrogate-based targets, amount and genetic-based targets), the type of genetic variation measured (adaptive or neutral), and the interactions between them.
-All model contained intercept- and slope-level random effects to accommodate variation between species.
+To determine  if the environmental and geographic variables were effective surrogates, we fit a full model to the proportion of genetic variation that the first collection of prioritizations secured using categorical variables to indicate the surrogate-based targets used to generate them, the type of genetic variation (adaptive or neutral), the method used to generate the prioritization (optimal or random), and the interactions between these variables.
+To determine if surrogate-based targets improved the effectiveness of prioritizations under various scenarios, we fit a full model to the proportion of genetic variation that second collection of prioritizations secured using categorical variables to indicate the scenario (single-species, multi-species, multi-species with cost), the combination of targets used to generate the prioritizations (amount-based targets, amount and surrogate-based targets, amount and genetic-based targets), the type of genetic variation measured (adaptive or neutral), and the interactions between them.
+All model contained intercept-level random effects to accommodate variation between species.
 Both full models were subject to a step-wise term deletion routine using $\chi^{2}$ tests to assess significance (terms were retained when $P < 0.05$).
@@
 ### Limitations/Future directions
-We wish to alert practitioners to several limitations associated with our analysis.
+We wish to alert our readers to several limitations associated with our analysis.
 First, the size of the planning units we used (23 $\times$ 25 km) is much larger than typically used in conservation planning scenarios (approx.
@@
 We used this resolution because the genetic data was collected at this scale.
-Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis because the surrogate variables are spatially auto-correlated.
+Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis because this process would have introduced artificial spatial auto-correlation into the dataset.
 Second, we used geographic distances as surrogates for connectivity between planning units.
@@
 Third, we used AFLP data to characterize the main gradients of genetic variation.
-Whilst next-generation sequencing provides higher resolution genetic information than AFLP data [reviewed in @r480], our methodology would still have utilized only the main gradients of the genetic variation to generate prioritizations in a feasible amount of time.
+Whilst next-generation sequencing provides higher resolution genetic information than AFLP data [reviewed in @r480], our methodology would still have utilized only the main gradients of the genetic variation in order to generate prioritizations in a feasible amount of time.
 Overall, we are confident that the limitations associated with our study are relatively minor.

### 2016-03-18 article/article.Rmd
+ @Zongle comments incorperated to article + final dataset added to ms

@@
 options(error=NULL)
+
 ```
@@
 The act of setting aside areas for preservation spearheads modern conservation efforts [@r440].
-They buffer species from anthropogenic impacts and provide places for populations of species to persist.
-Intra-specific genetic variation underpins species persistence, and there has been increasing interest in understanding how to design protected areas networks in a way that maximizes this variation [@r67; @r3; @r61].
+Protected area networks buffer species from anthropogenic impacts and provide places for populations of species to persist.
+Intra-specific genetic variation underpins species persistence, and there has been increasing interest in understanding how to design protected area networks in a way that maximizes this variation [@r67; @r3; @r61].
 Broadly speaking, genetic variation can be classified as either adaptive or neutral [reviewed in @r48].
@@
 terrain; eg.
-@r474]--the systems commonly featured in real-world conservation planning exercises.
+@r474]--the very systems commonly featured in real-world conservation planning exercises [eg.
+@r197].
 Before conservation planners can reliably use surrogates for genetic data in conservation planning, the effectiveness of these surrogates needs to be verified.
-Otherwise, conservation planners may waste precious resources on places that do little to benefit biodiversity.
+Otherwise, conservation planners may waste precious resources on protecting places that provide little benefit to biodiversity.
 
 Here we investigate the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation in the context of conservation planning.
-First, we examine the relationship between surrogates and gene variation, and verify that these variables can be used as surrogates.
-Second, we determine if the inclusion of surrogate-based targets in a variety of conservation planning scenarios with increasingly levels of realism can result in prioritizations that secure more genetic variation.
-If these surrogates can be shown to improve the intra-specific genetic variation captured in a prioritization, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.
+First, we examine the relationship between surrogates and genetic variation, and verify whether these variables are effective surrogates.
+Second, we determine if the inclusion of surrogates in various conservation planning scenarios can improve the proportion of genetic variation they secure.
+If these surrogates can improve the intra-specific genetic variation captured in a prioritization, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.
 
@@
 ### Study system
-We used species distribution and genomic data collected by the IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for methodological details].
-This data set was particularly well suited for our study, because it provides spatially explicit genomic data for multiple species with vastly different demographic characteristics and evolutionary histories across a broad geographic area with large environmental gradients.
+We used species distribution and genomic data collected by the IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for a detailed explanation of data collection methods].
+This data set was particularly well suited for our study because it contains genetic data for a diverse range of species that were collected using the same sampling regime.
+Additionally, the study area covers a broad geographic domain that encompasses  attitudinal gradients which provide substantial variation to examine the surrogates.
 As part of this project, data was collected for 27 alpine plant species using a 20' longitude by 21' latitude grid (approx.
 22.3 km $\times$ 25 km; Figure S1).
-They visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals.
+Project members visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals.
 They genotyped samples using amplified fragment length polymorphisms [AFLP; @r453], and constructed matrices denoting the presence/absence of polymorphisms at loci for each species.
-Thus for each second grid cell, for each of the species present in the grid cell, the data set contains information describing the genomic variation of individuals in the cell.
+Thus the data set contains information describing the genomic variation of individuals for each species present in every second grid cell.
 We used these grid cells as planning units to develop prioritizations.
-To reduce computational effort, we only used grid cells that contained at least a single individual ($n = `r nrow(grid.DF)`$).
+To reduce computational effort, we used only grid cells that contained at least a single individual ($n = `r nrow(grid.DF)`$).
 The majority of our analyses were performed using the \texttt{R} software package [version `r paste0(version$major, '.', version$minor)`; @r192].
@@
 We calculated the average value for each PC layer in each grid cell, and used these values to characterize the main themes of climatic variation among the grid cells (Figure S2).
-To examine if the influence of opportunity cost, we calculated the total human population density inside each planning unit [obtained at 1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472].
+The cost of purchasing land for preservation often exerts a strong influence on which planning units are recommended for prioritization [reviewed in @r482].
+To examine the potential effects of cost, we calculated the total human population density inside each planning unit [obtained at 1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472] and used this to represent acquisition cost.
 
-### Outlier locus detection
+### Adaptive and neutral genetic data
 To investigate the effectiveness of the surrogates for adaptive and neutral genetic variation, we first needed to identify which of the sampled loci--if any--were under selection.
-We used outlier locus detection methods [compared in @r454; reviewed in @r480] to achieve this.
+We used outlier detection methods [compared in @r454; reviewed in @r480] to achieve this.
 The basic premise underpinning these analyses is that neutral loci are expected to exhibit a specific level of variation among populations ($F_{st}$), and loci that deviate from this expectation should be under selection.
@@
 The purpose of this step was to avoid falsely classifying loci as adaptive.
-For each species, we ordinated the genomic data and fit models to the ordinations to fit to turn group individuals into populations.
+For each species, we ordinated the genomic data, and then fit models to the ordinated data to group individuals into populations.
 Specifically, we used non-metric multi-dimensional scaling [NMDS; implemented in the $\texttt{vegan R}$ package; @r458] using Gower distances [@r464; using the $\texttt{cluster R}$ package to accommodate sparsity; @r457] to transform the binary loci data into continuous variables (Table S2).
@@
 
-We identified the loci under selection (Table S3).
-Since outlier locus detection methods compare allele frequencies between populations, we were unable to examine species that contained a single population.
+After grouping individuals into populations, we identified the loci under selection (Table S3).
+Since outlier detection methods compare allele frequencies between populations, we were unable to examine species that contained a single population.
 For these species ($n = `r sum(sapply(spp.Mclust.LST, function(x) return(ncol(x$mclust$z)==1)))`$), we assumed that all loci were neutral.
@@
 To avoid false-positives, we omitted loci where the global frequency of the minor allele was $\geq 0.05$ (following guidelines in the user manual) and assumed they were neutral.
-Additionally, we omitted individuals from this analysis when their population membership was uncertain (maximum membership probability $\l `r mclust.params.LST[[MODE]]$probability.threshold`$).
+Additionally, we omitted individuals from this analysis when their population membership was uncertain (maximum membership probability $< `r mclust.params.LST[[MODE]]$probability.threshold`$).
 We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate to ensure convergence [FDR $\leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].
@@
 However, most large-scale prioritizations involve a comprehensive set of species.
-This property may lead to prioritizations that secure a greater proportion of intra-specific genetic variation for each species by returning solutions that preserve the same species in multiple communities.
+By forcing solutions to secure individuals in multiple communities, this may result in prioritizations that secure a greater proportion of intra-specific genetic variation for each species.
 To test for this effect, our second scenario involved generating multi-species prioritizations.
@@
 
-For each scenario, we generated three types of prioritizations.
-We generated prioritizations using only amount-based targets, both amount- and surrogate-based targets, and both amount- and genetic-based targets.
-The purely amount-based prioritizations serve as our baseline--they represent prioritizations typically generated in conservation planning exercises.
+For each scenario, we generated three types of prioritizations using (1) only amount-based targets, (2) both amount- and surrogate-based targets, and (3) both amount- and genetic-based targets.
+The purely amount-based prioritizations served as our baseline--representing prioritizations typically generated in conservation planning exercises.
 We used `r (rapr.params.LST[[MODE]][['single.species']]$amount.target*100)` % amount-based targets in all prioritizations to ensure an adequate proportion of habitat was secured for each species.
@@
 We investigated the effectiveness of prioritizations generated using surrogate-based targets using generalized linear mixed-effects models [GLMMs; using a logit link and the $\texttt{lme4}$ R package; @r469].
-To determine  if the environmental and geographic variables were effective surrogates, we fit a full model to the proportion of genetic variation that the first collection of prioritizations secured using categorical variables to indicate the surrogate-based targets used to generate them, the type of genetic variation (adaptive or neutral), the method used to generate the prioritization (optimal or random), and the interactions between these variables.
-To determine if surrogate-based targets improved the effectiveness of prioritizations under various scenarios, we fit a full model to the proportion of genetic variation that second collection of prioritizations secured using categorical variables to indicate the scenario (single-species, multi-species, multi-species with cost), the combination of targets used to generate the prioritizations (amount-based targets, amount and surrogate-based targets, amount and genetic-based targets), the type of genetic variation measured (adaptive or neutral), and the interactions between them.
+To determine  if the environmental and geographic variables were effective surrogates, we fit a full model to the proportion of genetic variation that the first collection of prioritizations.
+This full model contained categorical predictor variables to indicate the surrogate-based targets used to generate them, the type of genetic variation (adaptive or neutral), the method used to generate the prioritization (optimal or random), and the interactions between these variables.
+To determine if surrogate-based targets improved the effectiveness of prioritizations under various scenarios, we fit a full model to the proportion of genetic variation that second collection of prioritizations secured.
+This second full model contained categorical predictor variables to indicate the scenario (single-species, multi-species, multi-species with cost), the combination of targets used to generate the prioritizations (amount-based targets, amount and surrogate-based targets, amount and genetic-based targets), the type of genetic variation measured (adaptive or neutral), and the interactions between them.
 All model contained intercept-level random effects to accommodate variation between species.
@@
 <!-- Surrogacy analysis -->
-Prioritizations generated using greater environmental and geographic surrogate-based targets were found to secure greater proportions of adaptive and neutral genetic variation (Figure 1).
+Prioritizations generated using higher environmental and geographic surrogate-based targets were found to secure greater proportions of adaptive and neutral genetic variation (Figure 1).
 The proportion of genetic variation that these prioritizations secured was substantially more than the genetic variation secured by randomly generated prioritizations (${\chi^{2}}_{`r NA`,`r NA`} = `r  NA`$, $P < `r NA`$).

### 2016-03-18 article/supporting_information.Rmd
+ bug fix in supp info

@@
 			function(i) {
+				if (!is.null(spp.BayeScan.sample.loci.subset.LST[[i]])) {
 					data.frame(
@@
 				}
+			}
 		),
@@
 	caption='Loci classifications for each species.
-Each row denotes a sampled loci for a given species, and the evidence for it being under selection based on outputs from the program \\texttt{Bayescan}.'
+Each row denotes a sampled loci for a given species, and the evidence for it being under selection based on outputs from the program \\texttt{Bayescan}.
+Note that loci in species with a single population were assumed to be neutral.
+Loci for these species do not appear in this table.'
 )

### 2016-03-18 article/figures.Rmd
+ @Zongle comments incorperated to article + final dataset added to ms
+ fix bug

@@
 		data.frame(
-			letter=unname(toupper(cld(posthoc.model.GLHT)$mcletters$Letters)),
-			Combination=as.character(names(toupper(cld(posthoc.model.GLHT)$mcletters$Letters)))
+			letter=unname(toupper(cld(posthoc.scenario.GLHT)$mcletters$Letters)),
+			Combination=as.character(names(toupper(cld(posthoc.scenario.GLHT)$mcletters$Letters)))
 		) %>% mutate(Combination=as.character(Combination)),

### 2016-03-23 article/figures.Rmd
+ new parameters + fixes with figures

@@
 
-```{r, echo=FALSE, fig.width=6, fig.height=4.5, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations.
+```{r, echo=FALSE, message=FALSE, fig.width=6, fig.height=4.5, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations.
 Panel (a) shows the relationship between environmental surrogate targets used to generate a prioritization and the proportion of adaptive genetic variation it secures.
@@
 	) %>% data.frame()
+
 geo.correlation.DF$Surrogate.target <- factor(geo.correlation.DF$Surrogate.target)
@@
 	ylab('Adaptive genetic\nvariation secured (%)') +
-	geom_text(aes(x=0.75, y=1, label='(a)'), color='black', size=5) +
+	geom_text(aes(x=0.55, y=1.125, label='(a)'), color='black', size=5) +
 	theme_classic() +
-	theme(legend.position=c(0.225,0.8)) +
+	theme(legend.position=c(0.9,0.4)) +
 	scale_color_manual(name='Prioritization', values=c('Optimal'='black', 'Random'='grey50')) +
@@
 	ylab('Neutral genetic\nvariation secured (%)') +
-	geom_text(aes(x=0.75, y=1, label='(b)'), color='black', size=5) +
+	geom_text(aes(x=0.55, y=1.125, label='(b)'), color='black', size=5) +
 	theme_classic() +
@@
 grid.FPLY <- grid.PLY
-for (i in seq_along(single.spp.prioritisations))
-	grid.FPLY@data[[paste0('Single-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations[[i]], function(x){colMeans(x@results@selections)}))
+for (i in seq_along(single.spp.prioritisations[[1]]))
+	grid.FPLY@data[[paste0('Single-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations, function(x){colMeans(x[[i]]@results@selections)}))
 for (i in seq_along(multi.spp.prioritisations))
@@
 			data=results.PDF, position=position_dodge(0.9)) +
-		geom_text(aes(x=0.7, y=1, label=panel.letter),
+		geom_text(aes(x=0.7, y=1.05, label=panel.letter),
 			data=results.PDF) +

### 2016-03-24 article/article.Rmd
+ article fixes + remove randomly generated prioritizations from surrogacy analysis

@@
 
-To understand the relationship between the surrogate-based targets used to generate a prioritization and the proportion of genetic variation it secures, we generated a collection of single-species prioritizations.
+To determine if the environmental and geographic variables are effective surrogates for adaptive and neutral genetic variation, we generated a collection of single-species prioritizations.
 For each species, we first generated two sets of prioritizations using increasing environmental and geographic-based targets (separately).
 Prioritizations were generated using surrogate-based targets ranging from `r (min(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target)*100)` % and `r (max(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target)*100)` % in increments of `r diff(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target[1:2])*100` % ($n = `r rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.optimal.replicates`$ per target).
-To determine if these surrogate-based prioritizations performed better then random, we then generated another two sets of prioritizations by randomly selecting planning units that were occupied by the species ($n = `r rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.random.replicates`$ per surrogate target).
-To account for variation in the number of selected planning units in the target-based prioritizations, the number of selected planning units in the randomly generated prioritizations was equal to the target-based prioritizations.
-Randomly generated prioritizations associated with negative values--because they secured only a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis.
+Prioritizations associated with negative values--because they secured only a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis.
 
@@
 To test for this effect, our second scenario involved generating multi-species prioritizations.
-Finally, most large-scale prioritizations also explicitly consider opportunity cost to ensure prioritizations are cost-effective.
+Finally, most large-scale prioritizations also explicitly consider opportunity cost to ensure that prioritizations are cost-effective.
 Our third scenario involved generating multi-species prioritizations that satisfied targets for minimal opportunity cost.
@@
 For each scenario, we generated three types of prioritizations using (1) only amount-based targets, (2) both amount- and surrogate-based targets, and (3) both amount- and genetic-based targets.
-The purely amount-based prioritizations served as our baseline--representing prioritizations typically generated in conservation planning exercises.
+The purely amount-based prioritizations served as our baseline--representing the prioritizations typically generated in conservation planning exercises.
 We used `r (rapr.params.LST[[MODE]][['single.species']]$amount.target*100)` % amount-based targets in all prioritizations to ensure an adequate proportion of habitat was secured for each species.
@@
 To determine  if the environmental and geographic variables were effective surrogates, we fit a full model to the proportion of genetic variation that the first collection of prioritizations.
-This full model contained categorical predictor variables to indicate the surrogate-based targets used to generate them, the type of genetic variation (adaptive or neutral), the method used to generate the prioritization (optimal or random), and the interactions between these variables.
+This full model contained categorical predictor variables to indicate the surrogate-based targets used to generate them, the type of genetic variation (adaptive or neutral), and the interactions between these variables.
 To determine if surrogate-based targets improved the effectiveness of prioritizations under various scenarios, we fit a full model to the proportion of genetic variation that second collection of prioritizations secured.
 This second full model contained categorical predictor variables to indicate the scenario (single-species, multi-species, multi-species with cost), the combination of targets used to generate the prioritizations (amount-based targets, amount and surrogate-based targets, amount and genetic-based targets), the type of genetic variation measured (adaptive or neutral), and the interactions between them.
-All model contained intercept-level random effects to accommodate variation between species.
-Both full models were subject to a step-wise term deletion routine using $\chi^{2}$ tests to assess significance (terms were retained when $P < 0.05$).
-Finally, we conducted post-hoc analysis on each full model using Tukey contrasts with a Bonferroni correction [using the $\texttt{multcomp}$ R package; @r470].
+All models contained intercept-level random effects to accommodate variation between species.
+We conducted a post-hoc analysis on each full model using Tukey contrasts with a Bonferroni correction [using the $\texttt{multcomp}$ R package; @r470].
 
@@
 <!-- Surrogacy analysis -->
-Prioritizations generated using higher environmental and geographic surrogate-based targets were found to secure greater proportions of adaptive and neutral genetic variation (Figure 1).
-The proportion of genetic variation that these prioritizations secured was substantially more than the genetic variation secured by randomly generated prioritizations (${\chi^{2}}_{`r NA`,`r NA`} = `r  NA`$, $P < `r NA`$).
-Generally, prioritizations needed to capture at least `r NA` % of the environmental and geographic surrogates to secure 90 % of the species' genetic variation.
-Additionally, we detected differences in the effectiveness of the environmental and geographic surrogates (${\chi^{2}}_{`r NA`,`r NA`} = `r  NA`$, $P < `r NA`$).
-Increases in environmental surrogate-based targets resulted prioritizations that secured more adaptive genetic variation, than geographic surrogate-based targets for neutral variation.
-These results demonstrate that environmental and geographic variables are effective surrogates for genetic variation for these plant species.
+The environmental and geographic-based variables were effective surrogates for adaptive and genetic variation (Figure 1).
+The prioritizations generated using greater surrogate-based targets secured more genetic variation (${\chi^{2}}_{`r NA`,`r NA`} = `r  NA`$, $P < `r NA`$).
+The environmental-based surrogates tended to secure a substantially higher proportion of genetic variation than the geographic-based prioritizations (${\chi^{2}}_{`r NA`,`r NA`} = `r  NA`$, $P < `r NA`$).
+Only particularly high geographic-based targets ($\geq 0.95 %$) were able to secure a moderate proportion of neutral variation.
+Overall, there appears to be a non-linear relationship between the surrogate-based target used to generate a prioritization and the proportion of genetic variation secured.
 
 <!-- Scenario analysis -->
-The addition of environmental and geographic surrogate-based targets to amount-based targets was not found to substantially improve the effectiveness of prioritizations under any of the scenarios (Figure 2).
-No significant differences were detected between prioritizations generated using different combinations of targets (${\chi^{2}}_{`r NA`,`r NA`} = `r  NA`$, $P < `r NA`$), or under any of the different scenarios (${\chi^{2}}_{`r NA`,`r NA`} = `r NA`$, $P < `r NA`$).
-The only difference was that the inclusion of surrogate-based targets tended to result in prioritizations that secured slightly more adaptive than neutral genetic variation.
-Under the simplest of circumstances, the single-species amount-based prioritizations secured most of the species' adaptive (mean `r NA` $\pm$ `r NA`) and neutral (mean `r NA` $\pm$ `r NA`) genetic variation.
-These proportions were not significantly different to those associated with prioritizations generated using additional surrogate-based targets ($Z = `r NA`$, $P < `r NA`$) or genetic-based targets ($Z = `r NA`$, $P < `r NA`$).
-We found similar patterns in the multi-species scenario ${\chi^{2}}_{`r NA`,`r NA`} = `r NA`$, $P < `r NA`$) and the multi-species with opportunity cost scenario ${\chi^{2}}_{`r NA`,`r NA`} = `r NA`$, $P < `r NA`$).
-Although the prioritizations had similar levels of performance, there were differences in their spatial distributions (Figure 3).
-In particular, the inclusion of cost data in the multi-species prioritizations had the effect of concentrating conservation effort towards areas with higher elevation.
-Overall, these results suggest that while environmental and geographic surrogate variables are effective surrogates, they are not needed to secure most of the intra-specific genetic variation.
+To determine if the addition of environmental and geographic surrogate-based targets could improve the effectiveness of prioritizations, we generated prioritizations according to three different scenarios (Figure 2).
+The single-species prioritizations scenarios showed similar selection frequencies across the study area, regardless of whether they were generated using amount-only targets, amount- and surrogate-based targets, or amount- and genetic-based targets.
+On the other hand, the multi-species prioritizations showed different selection frequencies depending on the targets used to generate them.
+The addition of surrogate- or genetic-based targets resulted in prioritizations that contained substantially more planning units.
+Furthermore, the inclusion of cost data in the multi-species prioritizations resulted in even more planning units being selected.
+These prioritizations also tended to concentrate conservation effort towards areas with higher elevation.
+The inclusion of surrogate- and genetic- based targets tended to result in prioritizations that captured more genetic variation.
+
+The addition of environmental and geographic surrogate-based targets had the greatest effect on the proportion of neutral variation secured in a prioritization.
+In both the single species and multi-species scenarios, the inclusion of surrogate- and genetic-based targets resulted in prioritizations that secured substantially more neutral variation.
+However, this effect was not present in the multi-species with cost scenario.
+Unlike neutral variation, the addition of environmental and geographic surrogate-based targets had little effect on the proportion of adaptive variation secured in a prioritization.
+Although the inclusion of surrogate- and genetic-based targets resulted in more adaptive variation being secured in the single species scenarios, this effect was not present in either the multi-species or the multi-species with cost scenarios.
+This is likely due to the fact that the multi-species prioritizations were constrained to secure each species in a variety of different communities, and therefore, potentially different habitats.
 
 ## Discussion
-We aimed to investigate the effectiveness of environmental and geographic surrogates for genetic variation in conservation planning.
-Our study is the first to test this in a realistic context.
-We found that environmental and geographic variables are effective surrogates for genetic variation in conservation planning.
-But we also found that just using amount-based targets--typically used in conservation planning--result in prioritizations that secure most of the adaptive and neutral intra-specific genetic variation.
-Our results support existing guidelines on using a comprehensive set of species in conservation planning.
+We aimed to investigate the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation in conservation planning.
+Our study is the first to test this in a multi-species context.
+Overall, we found that environmental and geographic variables are not necessarily effective surrogates for genetic variation in conservation planning--though they can be useful to for identifying suitable amount-based targets.
+Additionally, we found that surrogate-based targets can supplement amount-based targets and improve the proportion of neutral variation secured in a prioritization if the amount-based targets are too low.
+Our results suggest that in most cases--barring extremely unusual spatial patterns of intra-specific genetic variation--conservation planners can secure sufficient adaptive and neutral variation by following conventional guidelines.
 

### 2016-03-24 article/figures.Rmd
+ article fixes + remove randomly generated prioritizations from surrogacy analysis

@@
 env.correlation.DF$Surrogate.target <- factor(env.correlation.DF$Surrogate.target)
+env.correlation.DF <- env.correlation.DF %>%
+	group_by(Surrogate.target, Type, Species) %>% 
+	summarize(genetic.held=mean(genetic.held, na.rm=TRUE)) %>%
+	data.frame()
+
 env.correlation.DF2 <- env.correlation.DF %>%
-	group_by(Surrogate.target, Method, Type) %>% 
+	group_by(Surrogate.target, Type) %>% 
 	summarize(genetic.held=max(genetic.held, na.rm=TRUE)+0.1) %>%
-	ungroup() %>% mutate(Combination=as.character(paste0(Surrogate.target,'.',Method,'.',Type))) %>% 
+	ungroup() %>% mutate(Combination=as.character(paste0(Surrogate.target,'.',Type))) %>% 
 	left_join(
@@
 geo.correlation.DF$Surrogate.target <- factor(geo.correlation.DF$Surrogate.target)
+geo.correlation.DF <- geo.correlation.DF %>%
+	group_by(Surrogate.target, Type, Species) %>% 
+	summarize(genetic.held=mean(genetic.held, na.rm=TRUE)) %>%
+	data.frame()
+
 geo.correlation.DF2 <- geo.correlation.DF %>%
-	group_by(Surrogate.target, Method, Type) %>% 
+	group_by(Surrogate.target, Type) %>% 
 	summarize(genetic.held=max(genetic.held, na.rm=TRUE)+0.1) %>%
-	ungroup() %>% mutate(Combination=as.character(paste0(Surrogate.target,'.',Method,'.',Type))) %>% 
+	ungroup() %>% mutate(Combination=as.character(paste0(Surrogate.target,'.',Type))) %>% 
 	left_join(
@@
 # make figure
-p1 <- ggplot(aes(x=Surrogate.target,y=genetic.held, color=Method),
+p1 <- ggplot(aes(x=Surrogate.target,y=genetic.held),
 		data=env.correlation.DF) +
@@
 	theme_classic() +
-	theme(legend.position=c(0.9,0.4)) +
-	scale_color_manual(name='Prioritization', values=c('Optimal'='black', 'Random'='grey50')) +
-	geom_text(aes(x=Surrogate.target, y=genetic.held, color=Method, label=letter),
+	theme(legend.position='none') +
+	geom_text(aes(x=Surrogate.target, y=genetic.held, label=letter),
 		data=env.correlation.DF2, position=position_dodge(0.9))
 
-p2 <- ggplot(aes(x=Surrogate.target,y=genetic.held, color=Method),
+p2 <- ggplot(aes(x=Surrogate.target,y=genetic.held),
 			data=geo.correlation.DF) +
@@
 	theme(legend.position='none') +
-	scale_color_manual(name='Prioritization', values=c('Optimal'='black', 'Random'='grey50')) +
-	geom_text(aes(x=Surrogate.target, y=genetic.held, color=Method, label=letter),
+	geom_text(aes(x=Surrogate.target, y=genetic.held, label=letter),
 		data=geo.correlation.DF2, position=position_dodge(0.9))

### 2016-03-24 R/analysis/07-surrogacy-prioritizations.R

@@ -32,54 +32,6 @@ env.correlation.prioritisations <- llply(
 	.parallel=FALSE
 )
 
-cat("starting random prioritizations for the environmental surrogate prioritizations \n")
-clust <- makeCluster(general.params.LST[[MODE]]$threads, type='SOCK')
-clusterEvalQ(clust, {library(rapr)})
-clusterExport(clust, c('spp.samples.DF','grid.DF','rapr.params.LST',
-	'MODE', 'species.prioritisation', 'ru', 'env.correlation.prioritisations'))
-registerDoParallel(clust)
-env.random.prioritisations <- llply(
-	seq_along(unique(spp.samples.DF$species)),
-	function(i) {
-		llply(
-			seq_along(rapr.params.LST[[MODE]]$surrogacy.analysis$surrogate.target),
-			function(j) {
-				setMKLthreads(1)
-				# extract number of planning units
-				curr.optim_n <- table(rowSums(env.correlation.prioritisations[[i]][[j]]@results@selections))
-				curr.optim_p <- curr.optim_n / sum(curr.optim_n)
-				# identify which planning units are occupied by the species
-				curr.pu <- which(grid.DF[[unique(spp.samples.DF$species)[[i]]]]==1)
-				zeros <- rep(0, length=nrow(grid.DF))
-				# generate portfolio of random selections
-				curr.random_n <- as.numeric(sample(
-					names(curr.optim_n),
-					size=rapr.params.LST[[MODE]]$surrogacy.analysis$surrogate.random.replicates,
-					replace=TRUE, prob=as.numeric(curr.optim_p)
-				))
-				curr.sol.MTX<-t(sapply(
-					curr.random_n,
-					function(n) {replace(zeros, sample(x=curr.pu, size=n), rep(1, n))}
-				))
-				# generate RapSolved object
-				species.prioritisation(
-					x=spp.subset(ru, i),
-					amount.targets=0.0001,
-					env.surrogate.targets=NA,
-					geo.surrogate.targets=NA,
-					adaptive.genetic.targets=NA,
-					neutral.genetic.targets=NA,
-					b=curr.sol.MTX
-				)
-			}
-		)
-	},
-	.parallel=TRUE
-)
-# kill cluster
-clust <- stopCluster(clust)
-
-
 cat("starting geographic surrogacy prioritizations\n")
 geo.correlation.prioritisations <- llply(
[…]

### 2016-03-30 article/article.Rmd
+ use fisher tests to compare porportion of species with genetic targets fulfilled

@@
 The environmental and geographic-based variables were effective surrogates for adaptive and genetic variation (Figure 1).
-The prioritizations generated using greater surrogate-based targets secured more genetic variation (${\chi^{2}}_{`r NA`,`r NA`} = `r  NA`$, $P < `r NA`$).
-The environmental-based surrogates tended to secure a substantially higher proportion of genetic variation than the geographic-based prioritizations (${\chi^{2}}_{`r NA`,`r NA`} = `r  NA`$, $P < `r NA`$).
-Only particularly high geographic-based targets ($\geq 0.95 %$) were able to secure a moderate proportion of neutral variation.
-Overall, there appears to be a non-linear relationship between the surrogate-based target used to generate a prioritization and the proportion of genetic variation secured.
+The prioritizations generated using greater surrogate-based targets secured more genetic variation (${\chi^{2}}_{`r sub.correlation.3.ANOVA[[7]][2]`} = `r  sub.correlation.3.ANOVA[[6]][2]`$, $P  `pretty.pval(sub.correlation.3.ANOVA[[8]][2])`$).
+In addition, prioritizations generated using environmental targets often secured more adaptive genetic variation than prioritizations generated using geographic-based targets did neutral genetic variation (${\chi^{2}}_{`r sub.correlation.2.ANOVA[[7]][2]`} = `r  sub.correlation.2.ANOVA[[6]][2]`$, $P `r pretty.pval(sub.correlation.2.ANOVA[[8]][2])`$).
+Only particularly high geographic targets ($\geq 0.95$ %) were able to secure a moderate proportion of neutral variation ($\geq XXXX$ %).
+Overall, there was a non-linear relationship between the surrogate-based targets used to generate the prioritizations and the proportion of genetic variation they secured.
 
 <!-- Scenario analysis -->
-To determine if the addition of environmental and geographic surrogate-based targets could improve the effectiveness of prioritizations, we generated prioritizations according to three different scenarios (Figure 2).
-The single-species prioritizations scenarios showed similar selection frequencies across the study area, regardless of whether they were generated using amount-only targets, amount- and surrogate-based targets, or amount- and genetic-based targets.
-On the other hand, the multi-species prioritizations showed different selection frequencies depending on the targets used to generate them.
+To determine if the addition of environmental and geographic targets could improve the effectiveness of prioritizations in a more realistic context, we generated prioritizations according to three different scenarios (Figure 2).
+The prioritizations generated under the single-species scenario showed similar selection frequencies across the study area--regardless of the combination of targets used to generate them.
+On the other hand, the prioritizations generated under the multi-species scenario showed different selection frequencies depending on the targets used to generate them.
 The addition of surrogate- or genetic-based targets resulted in prioritizations that contained substantially more planning units.
-Furthermore, the inclusion of cost data in the multi-species prioritizations resulted in even more planning units being selected.
-These prioritizations also tended to concentrate conservation effort towards areas with higher elevation.
-The inclusion of surrogate- and genetic- based targets tended to result in prioritizations that captured more genetic variation.
+The prioritizations generated under the multi-species with cost scenario had different spatial distributions to the previous two scenarios.
+The prioritizations generated under this third scenario concentrated conservation effort towards areas with higher elevation.
 
-The addition of environmental and geographic surrogate-based targets had the greatest effect on the proportion of neutral variation secured in a prioritization.
-In both the single species and multi-species scenarios, the inclusion of surrogate- and genetic-based targets resulted in prioritizations that secured substantially more neutral variation.
-However, this effect was not present in the multi-species with cost scenario.
-Unlike neutral variation, the addition of environmental and geographic surrogate-based targets had little effect on the proportion of adaptive variation secured in a prioritization.
-Although the inclusion of surrogate- and genetic-based targets resulted in more adaptive variation being secured in the single species scenarios, this effect was not present in either the multi-species or the multi-species with cost scenarios.
-This is likely due to the fact that the multi-species prioritizations were constrained to secure each species in a variety of different communities, and therefore, potentially different habitats.
+The inclusion of surrogate-based targets caused an improvement in the proportion of neutral variation secured in a prioritization (Figure 3).
+Specifically, the proportion of genetic variation secured by the prioritizations differed depending on the type of genetic variation measured (${\chi^{2}}_{`r sub.scenario.6.ANOVA[[7]][2]`} = `r  sub.scenario.6.ANOVA[[6]][2]`$, $P  `pretty.pval(sub.scenario.6.ANOVA[[8]][2])`$) and combination of targets used to generate the prioritizations (${\chi^{2}}_{`r sub.scenario.7.ANOVA[[7]][2]`} = `r  sub.scenario.7.ANOVA[[6]][2]`$, $P  `pretty.pval(sub.scenario.7.ANOVA[[8]][2])`$).
+No significant differences were found among the single-species, multi-species, and multi-species with cost contexts (${\chi^{2}}_{`r sub.scenario.5.ANOVA[[7]][2]`} = `r  sub.scenario.5.ANOVA[[6]][2]`$, $P `pretty.pval(sub.scenario.5.ANOVA[[8]][2])`$).
 

### 2016-03-30 article/figures.Rmd
+ use fisher tests to compare porportion of species with genetic targets fulfilled

--- 2016-03-23 article/figures.Rmd
@@
 # prepare data
-env.correlation.DF$Surrogate.target <- factor(env.correlation.DF$Surrogate.target)
-env.correlation.DF2 <- env.correlation.DF %>%
-	group_by(Surrogate.target, Method, Type) %>% 
+correlation.PDF <- correlation.sub.DF %>%
+	group_by(Surrogate.target, Type, Species) %>% 
+	summarize(genetic.held=mean(genetic.held, na.rm=TRUE)) %>%
+	data.frame()
+
+correlation.PDF2 <- correlation.PDF %>%
+	group_by(Type, Surrogate.target, Type) %>% 
 	summarize(genetic.held=max(genetic.held, na.rm=TRUE)+0.1) %>%
-	ungroup() %>% mutate(Combination=as.character(paste0(Surrogate.target,'.',Method,'.',Type))) %>% 
+	ungroup() %>%
+	mutate(Combination=as.character(paste0(Surrogate.target,'.',Type))) %>% 
 	left_join(
 		data.frame(
-			letter=unname(toupper(cld(posthoc.correlation.GLHT)$mcletters$Letters)),
-			Combination=as.character(names(toupper(cld(posthoc.correlation.GLHT)$mcletters$Letters))),
-			stringsAsFactors=FALSE
-		),
-		by='Combination'
-	) %>% data.frame()
-
-geo.correlation.DF$Surrogate.target <- factor(geo.correlation.DF$Surrogate.target)
-geo.correlation.DF2 <- geo.correlation.DF %>%
-	group_by(Surrogate.target, Method, Type) %>% 
-	summarize(genetic.held=max(genetic.held, na.rm=TRUE)+0.1) %>%
-	ungroup() %>% mutate(Combination=as.character(paste0(Surrogate.target,'.',Method,'.',Type))) %>% 
-	left_join(
-		data.frame(
-			letter=unname(toupper(cld(posthoc.correlation.GLHT)$mcletters$Letters)),
-			Combination=as.character(names(toupper(cld(posthoc.correlation.GLHT)$mcletters$Letters))),
+			letter=unname(correlation.fisher.cld$Letters),
+			Combination=names(correlation.fisher.cld$Letters),
 			stringsAsFactors=FALSE
@@
 # make figure
-p1 <- ggplot(aes(x=Surrogate.target,y=genetic.held, color=Method),
-		data=env.correlation.DF) +
+p1 <- ggplot(aes(x=Surrogate.target,y=genetic.held),
+		data=filter(correlation.PDF, Type=='Environmental')) +
 	geom_boxplot(position=position_dodge(0.9)) +
@@
 	ylab('Adaptive genetic\nvariation secured (%)') +
-	geom_text(aes(x=0.55, y=1.125, label='(a)'), color='black', size=5) +
+	geom_abline(intercept=rapr.params.LST[[MODE]]$surrogacy.analysis$genetic.target,
+		slope=0, color='red', linetype='dashed') +
+	geom_text(aes(x=0.55, y=1.115, label='(a)'), color='black', size=5) +
 	theme_classic() +
-	theme(legend.position=c(0.9,0.4)) +
-	scale_color_manual(name='Prioritization', values=c('Optimal'='black', 'Random'='grey50')) +
-	geom_text(aes(x=Surrogate.target, y=genetic.held, color=Method, label=letter),
-		data=env.correlation.DF2, position=position_dodge(0.9))
+	theme(legend.position='none') +
+	geom_text(aes(x=Surrogate.target, y=genetic.held, label=letter),
+		data=filter(correlation.PDF2, Type=='Environmental'),
+		position=position_dodge(0.9))
 
-p2 <- ggplot(aes(x=Surrogate.target,y=genetic.held, color=Method),
-			data=geo.correlation.DF) +
+p2 <- ggplot(aes(x=Surrogate.target,y=genetic.held),
+			data=filter(correlation.PDF, Type=='Geographic')) +
 	geom_boxplot(position=position_dodge(0.9)) +
@@
 	ylab('Neutral genetic\nvariation secured (%)') +
-	geom_text(aes(x=0.55, y=1.125, label='(b)'), color='black', size=5) +
+	geom_abline(intercept=rapr.params.LST[[MODE]]$surrogacy.analysis$genetic.target, 
+		slope=0, color='red', linetype='dashed') +
+	geom_text(aes(x=0.55, y=1.115, label='(b)'), color='black', size=5) +
 	theme_classic() +
 	theme(legend.position='none') +
-	scale_color_manual(name='Prioritization', values=c('Optimal'='black', 'Random'='grey50')) +
-	geom_text(aes(x=Surrogate.target, y=genetic.held, color=Method, label=letter),
-		data=geo.correlation.DF2, position=position_dodge(0.9))
+	geom_text(aes(x=Surrogate.target, y=genetic.held, label=letter),
+		data=filter(correlation.PDF2, Type=='Geographic'),
+		position=position_dodge(0.9))
 
@@
 	) %>% mutate(
-		Prioritization=factor(Prioritization, levels=c('Amount', 'Surrogate', 'Genetic')),
+		Prioritization=revalue(Prioritization, c('Amount'='Amount\ntargets', 'Surrogate'='Amount & surrogate\ntargets', 'Genetic'='Amount & genetic\ntargets')),
+		Prioritization=factor(as.character(Prioritization), levels=c('Amount\ntargets', 'Amount & surrogate\ntargets', 'Amount & genetic\ntargets')),
 		Context=factor(
@@
 # prepare data for plotting
-results.SDF <- results.SDF %>% mutate(
-		Prioritisation=factor(as.character(Prioritisation), levels=c('Amount','Surrogate','Genetic')),
-		Context=factor(as.character(Context), levels=c("single-species (equal costs)", "multi-species (equal costs)", "multi-species (opportunity costs)")),
-		Metric=factor(as.character(Metric), c("Adaptive variation", "Neutral variation"))
-	)
-results.PDF <- results.SDF %>% group_by(
-		Prioritisation, Context, Metric
-	) %>% mutate(
+scenario.PDF <- scenario.DF %>% group_by(Prioritisation, Context, Metric) %>% 
+	mutate(
 		letter.height=max(value, na.rm=TRUE)+0.1,
-		Combination=paste0(Prioritisation,'.',Metric,'.',Context)
-	) %>% filter(
-		c(TRUE, rep(FALSE, length(Context)-1))
-	) %>% data.frame() %>% group_by(Context, Metric) %>% mutate(
-		panel.letter=paste0('(',letters[as.integer(Context) + (as.integer(Metric)-1)],')'),
+		Combination=paste0(Prioritisation,'.',Metric,'.',Context)) %>%
+	filter(c(TRUE, rep(FALSE, length(Context)-1))) %>% 
+	ungroup() %>%
+	group_by(Context, Metric) %>% 
+	mutate(
+		panel.letter=paste0(
+			'(',letters[(as.integer(Context)*2)-1 + (as.integer(Metric)-1)],')'
+		),
 		panel.letter=c(as.character(panel.letter[1]), rep(NA_character_, length(Context)-1))
-	) %>% data.frame() %>% left_join(
+	) %>% 
+	ungroup() %>% 
+	left_join(
 		data.frame(
-			letter=unname(toupper(cld(posthoc.scenario.GLHT)$mcletters$Letters)),
-			Combination=as.character(names(toupper(cld(posthoc.scenario.GLHT)$mcletters$Letters)))
+			letter=unname(scenario.fisher.cld$Letters),
+			Combination=names(scenario.fisher.cld$Letters)
 		) %>% mutate(Combination=as.character(Combination)),
@@
 	) %>% data.frame()
-results.SDF <- results.SDF %>% mutate(
-	Context=sapply(as.character(Context), function(x) {paste0(toupper(substr(x, 1, 1)), substr(x, 2, nchar(x)))}),
-	Context=factor(
-		gsub(' (', '\n(', Context, fixed=TRUE),
-		levels=c("Single-species\n(equal costs)",
-		"Multi-species\n(equal costs)", 
-		"Multi-species\n(opportunity costs)"))
-	)
-results.PDF <- results.PDF %>% mutate(
-	Context=sapply(as.character(Context), function(x) {paste0(toupper(substr(x, 1, 1)), substr(x, 2, nchar(x)))}),
-	Context=factor(
-		gsub(' (', '\n(', Context, fixed=TRUE),
-		levels=c("Single-species\n(equal costs)",
-		"Multi-species\n(equal costs)", 
-		"Multi-species\n(opportunity costs)"))
-	)
 
@@
 	ggplot(aes(x=Prioritisation,y=value),
-		data=results.SDF) +
+		data=scenario.DF) +
 		geom_boxplot(position=position_dodge(0.9)) +
+		geom_abline(
+			intercept=rapr.params.LST[[MODE]]$scenario.analysis$genetic.target,
+			slope=0, color='red',linetype='dashed') +
 		geom_text(aes(x=Prioritisation, y=letter.height, label=letter),
-			data=results.PDF, position=position_dodge(0.9)) +
-		geom_text(aes(x=0.7, y=1.05, label=panel.letter),
-			data=results.PDF) +
+			data=scenario.PDF, position=position_dodge(0.9)) +
+		geom_text(aes(x=0.5, y=1.075, label=panel.letter),
+			data=scenario.PDF) +
 		scale_fill_manual(name='Prioritisation',

### 2016-03-30 R/functions/cld.RV.multcomp.R

@@ -1,0 +1,19 @@
+#' Compact letter display
+#'
+#' This function takes a post-hoc analysis indicating differences between different groups and returns
+#' Ia vector of letters indicating which groups are different from each other.
+#' For instance, outputs from the \code{\link[RVAideMemoire]{fisher.multcomp}}
+#' function are a valid input. 
+#' @param x \code{RV.multcomp} object.
+#' @return \code{character} vector.
+#' @seealso \code{\link[multcomp]{cld}}
+cld.RV.multcomp <- function(x) {
+	multcompView::multcompLetters(
+		full.p.table(x$p.value),
+		compare="<",
+		threshold=0.05,
+		Letters=LETTERS,
+		reversed=FALSE
+	)
+}
+

### 2016-03-30 article/article.Rmd
+ big fixes

@@
 The environmental and geographic-based variables were effective surrogates for adaptive and genetic variation (Figure 1).
-The prioritizations generated using greater surrogate-based targets secured more genetic variation (${\chi^{2}}_{`r sub.correlation.3.ANOVA[[7]][2]`} = `r  sub.correlation.3.ANOVA[[6]][2]`$, $P  `pretty.pval(sub.correlation.3.ANOVA[[8]][2])`$).
-In addition, prioritizations generated using environmental targets often secured more adaptive genetic variation than prioritizations generated using geographic-based targets did neutral genetic variation (${\chi^{2}}_{`r sub.correlation.2.ANOVA[[7]][2]`} = `r  sub.correlation.2.ANOVA[[6]][2]`$, $P `r pretty.pval(sub.correlation.2.ANOVA[[8]][2])`$).
+The prioritizations generated using greater surrogate-based targets secured more genetic variation (XXXX).
+In addition, prioritizations generated using environmental targets often secured more adaptive genetic variation than prioritizations generated using geographic-based targets did neutral genetic variation (XXXX).
 Only particularly high geographic targets ($\geq 0.95$ %) were able to secure a moderate proportion of neutral variation ($\geq XXXX$ %).
@@
 The inclusion of surrogate-based targets caused an improvement in the proportion of neutral variation secured in a prioritization (Figure 3).
-Specifically, the proportion of genetic variation secured by the prioritizations differed depending on the type of genetic variation measured (${\chi^{2}}_{`r sub.scenario.6.ANOVA[[7]][2]`} = `r  sub.scenario.6.ANOVA[[6]][2]`$, $P  `pretty.pval(sub.scenario.6.ANOVA[[8]][2])`$) and combination of targets used to generate the prioritizations (${\chi^{2}}_{`r sub.scenario.7.ANOVA[[7]][2]`} = `r  sub.scenario.7.ANOVA[[6]][2]`$, $P  `pretty.pval(sub.scenario.7.ANOVA[[8]][2])`$).
-No significant differences were found among the single-species, multi-species, and multi-species with cost contexts (${\chi^{2}}_{`r sub.scenario.5.ANOVA[[7]][2]`} = `r  sub.scenario.5.ANOVA[[6]][2]`$, $P `pretty.pval(sub.scenario.5.ANOVA[[8]][2])`$).
+Specifically, the proportion of genetic variation secured by the prioritizations differed depending on the type of genetic variation measured (XXXX) and combination of targets used to generate the prioritizations (XXXX).
+No significant differences were found among the single-species, multi-species, and multi-species with cost contexts (XXXX).
 

### 2016-03-30 article/figures.Rmd
+ big fixes

@@
 	) %>% mutate(
-		Prioritization=revalue(Prioritization, c('Amount'='Amount\ntargets', 'Surrogate'='Amount & surrogate\ntargets', 'Genetic'='Amount & genetic\ntargets')),
-		Prioritization=factor(as.character(Prioritization), levels=c('Amount\ntargets', 'Amount & surrogate\ntargets', 'Amount & genetic\ntargets')),
+		Prioritization=revalue(Prioritization, c('Amount'='Amount targets', 'Surrogate'='Amount & surrogate\ntargets', 'Genetic'='Amount & genetic\ntargets')),
+		Prioritization=factor(as.character(Prioritization), levels=c('Amount targets', 'Amount & surrogate\ntargets', 'Amount & genetic\ntargets')),
 		Context=factor(
@@
 			data=scenario.PDF, position=position_dodge(0.9)) +
-		geom_text(aes(x=0.5, y=1.075, label=panel.letter),
+		geom_text(aes(x=0.6, y=1.09, label=panel.letter),
 			data=scenario.PDF) +

### 2016-03-31 article/article.Rmd
+ included chisq stats in article + revert to using glmm for analysis + set nmds max.stress to 0.25 to reduce dimensionality

@@
 ### Study system
-We used species distribution and genomic data collected by the IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for a detailed explanation of data collection methods].
+We used species distribution and genomic data collected by the IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for further explanation of data collection methods].
 This data set was particularly well suited for our study because it contains genetic data for a diverse range of species that were collected using the same sampling regime.
-Additionally, the study area covers a broad geographic domain that encompasses  attitudinal gradients which provide substantial variation to examine the surrogates.
-As part of this project, data was collected for 27 alpine plant species using a 20' longitude by 21' latitude grid (approx.
+Additionally, the study area covers a broad geographic domain that encompasses attitudinal gradients.
+As part of this project, data were collected for 27 alpine plant species using a 20' longitude by 21' latitude grid (approx.
 22.3 km $\times$ 25 km; Figure S1).
@@
 We set the demand points for each species and each attribute space using the values associated with the planning units.
-For every prioritization we generated, we calculated the proportion of the adaptive (if detected) and neutral genetic variation that every prioritization secured to assess its performance.
 We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality using the $\texttt{rapr}$ R package [@r465] and $\texttt{Gurobi}$ [@r463].
+Statistical analysis were conducted using the $\texttt{RVAideMemoire}$ R package [@rXXX].
 
-To determine if the environmental and geographic variables are effective surrogates for adaptive and neutral genetic variation, we generated a collection of single-species prioritizations.
-For each species, we first generated two sets of prioritizations using increasing environmental and geographic-based targets (separately).
-Prioritizations were generated using surrogate-based targets ranging from `r (min(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target)*100)` % and `r (max(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target)*100)` % in increments of `r diff(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target[1:2])*100` % ($n = `r rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.optimal.replicates`$ per target).
-Prioritizations associated with negative values--because they secured only a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis.
+To determine if the environmental and geographic variables are effective surrogates for adaptive and neutral genetic variation (respectively), we generated a collection of single-species prioritizations.
+For each species, we generated two sets of prioritizations generated using increasing environmental and geographic targets (targets $\in$ `r paste(rapr.params.LST[[MODE]]$surrogacy.analysis$surrogate.target, collapse=', ')`).
+We calculated the proportion of adaptive genetic variation secured by the environmental-based prioritizations.
+We also calculated the proportion of neutral genetic variation secured by the geographic-based prioritizations.
+For each target, we counted the number of species that had a sufficient proportion of their genetic variation secured ($\geq$ `r rapr.params.LST[[MODE]]$surrogacy.analysis$genetic.target`).
+We used a $\chi^2$ test with Yates' continuity correction to examine whether the number of species with their genetic variation sufficiently preserved depended on the targets used to generate the prioritizations, and if the number of species sufficiently preserved given the same target differed depending on the type of genetic variation (adaptive vs.
+neutral) assessed.
+We then conducted post-hoc pairwise comparison analysis using $chi^2$ tests with Yates' continuity and Bonferroni corrections.
 
-To determine if the use of surrogate-based targets can actually improve prioritizations in real-world conservation planning scenarios, we investigated the effectiveness of surrogate-based targets using three scenarios with increasing levels of complexity.
+To determine if the addition of surrogate-based targets to conventional reserve selection methods can yield more effective prioritizations, we investigated the effectiveness of prioritizations that were generated using different types of targets.
+We generated three types of prioritizations using (1) only amount-based targets, (2) both amount- and surrogate-based targets (environmental and geographic targets), and (3) both amount- and genetic-based targets (adaptive and neutral genetic variation targets).
+The purely amount-based prioritizations served as our baseline--representing the prioritizations typically generated in conservation planning exercises.
+We used `r (rapr.params.LST[[MODE]]$scenario.analysis$amount.target*100)` % amount-based targets in all prioritizations to ensure an adequate proportion of habitat was secured for each species.
+We set the surrogate-based targets as `r (rapr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100)` % since this was high enough to secure most of the genetic variation according to the surrogacy analysis.
+
+We generated prioritizations using these three combinations of targets under different scenarios to determine if including surrogate-based targets is beneficial across a range of different conservation planning contexts.
 To assess the benefits of surrogate-based targets under the simplest of circumstances, our first scenario involved generating single-species prioritizations for each species.
-However, most large-scale prioritizations involve a comprehensive set of species.
-By forcing solutions to secure individuals in multiple communities, this may result in prioritizations that secure a greater proportion of intra-specific genetic variation for each species.
-To test for this effect, our second scenario involved generating multi-species prioritizations.
-Finally, most large-scale prioritizations also explicitly consider opportunity cost to ensure that prioritizations are cost-effective.
-Our third scenario involved generating multi-species prioritizations that satisfied targets for minimal opportunity cost.
+To determine if the benefits of surrogate-based targets are still evident when using a comprehensive set of species in a multi-species context, our second scenario involved generating multi-species prioritizations.
+To understand the effects of opportunity cost on the benefits of surrogate-based targets, our third scenario involved generating multi-species prioritizations that satisfied targets for minimal opportunity cost.
+Thus three types of prioritizations were generated for each scenario.
+Since multiple optimal solutions were available for the amount-only prioritizations in the single-species context, we generated `r rapr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates` prioritizations per species to accommodate variation in the proportion of genetic variation they secured.
 
-For each scenario, we generated three types of prioritizations using (1) only amount-based targets, (2) both amount- and surrogate-based targets, and (3) both amount- and genetic-based targets.
-The purely amount-based prioritizations served as our baseline--representing the prioritizations typically generated in conservation planning exercises.
-We used `r (rapr.params.LST[[MODE]][['single.species']]$amount.target*100)` % amount-based targets in all prioritizations to ensure an adequate proportion of habitat was secured for each species.
-We set surrogate- and genetic-based targets as `r (rapr.params.LST[[MODE]][['single.species']]$amount.target*100)` % to ensure that most of the respective variation was secured for each species.
-To accomodate variation in the high number of optimal solutions for the single-species amount-based prioritizations, we generated `r (rapr.params.LST[[MODE]][['single.species']]$amount.replicates)` replicate prioritizations.
-Since all other contexts were likely to be associated with few optimal solutions, we generated only `r rapr.params.LST[[MODE]][['multi.species']]$genetic.replicates` replicates.
-
-## Statistical analysis
-We investigated the effectiveness of prioritizations generated using surrogate-based targets using generalized linear mixed-effects models [GLMMs; using a logit link and the $\texttt{lme4}$ R package; @r469].
-To determine  if the environmental and geographic variables were effective surrogates, we fit a full model to the proportion of genetic variation that the first collection of prioritizations.
-This full model contained categorical predictor variables to indicate the surrogate-based targets used to generate them, the type of genetic variation (adaptive or neutral), and the interactions between these variables.
-To determine if surrogate-based targets improved the effectiveness of prioritizations under various scenarios, we fit a full model to the proportion of genetic variation that second collection of prioritizations secured.
-This second full model contained categorical predictor variables to indicate the scenario (single-species, multi-species, multi-species with cost), the combination of targets used to generate the prioritizations (amount-based targets, amount and surrogate-based targets, amount and genetic-based targets), the type of genetic variation measured (adaptive or neutral), and the interactions between them.
-All models contained intercept-level random effects to accommodate variation between species.
-We conducted a post-hoc analysis on each full model using Tukey contrasts with a Bonferroni correction [using the $\texttt{multcomp}$ R package; @r470].
+After generating all the prioritizations, we then counted the number of species that had a sufficient proportion of their genetic variation secured ($\geq$ `r rapr.params.LST[[MODE]]$scenario.analysis$genetic.target`) in each of the prioritizations.
+We used a $\chi^2$ test with Yates' continuity correction to assess whether the number of species with a sufficient proportion of their genetic variation secured depended on the combination of targets used to generate the prioritization (amount-only, amount and surrogate-based targets, amount and genetic-based targets), the scenario under which the prioritization was generated (single species, multi-species, or multi-species with cost), and the type of genetic variation measured (adaptive or neutral).
+Finally, we conducted a post-hoc pairwise comparison analysis using $chi^2$ tests with Yates' continuity and Bonferroni corrections.
 

### 2016-03-31 article/figures.Rmd
+ included chisq stats in article + revert to using glmm for analysis + set nmds max.stress to 0.25 to reduce dimensionality

@@
 ```{r, echo=FALSE, message=FALSE, fig.width=6, fig.height=4.5, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations.
+White boxplots show the proportion of genetic variation secured by prioritizations generated using surrogate-based targets.
+Grey boxplots show the proportion of genetic variation secured by randomly generated prioritizations with the same number of planning units as the surrogate-based prioritizations.
 Panel (a) shows the relationship between environmental surrogate targets used to generate a prioritization and the proportion of adaptive genetic variation it secures.
@@
 correlation.PDF <- correlation.sub.DF %>%
-	group_by(Surrogate.target, Type, Species) %>% 
+	group_by(Type, Surrogate.target, Method, Species) %>% 
 	summarize(genetic.held=mean(genetic.held, na.rm=TRUE)) %>%
@@
 correlation.PDF2 <- correlation.PDF %>%
-	group_by(Type, Surrogate.target, Type) %>% 
+	group_by(Type, Surrogate.target, Method) %>% 
 	summarize(genetic.held=max(genetic.held, na.rm=TRUE)+0.1) %>%
 	ungroup() %>%
-	mutate(Combination=as.character(paste0(Surrogate.target,'.',Type))) %>% 
+	mutate(Combination=as.character(paste0(Type,'.',Surrogate.target,'.',Method))) %>% 
 	left_join(
 		data.frame(
-			letter=unname(correlation.fisher.cld$Letters),
-			Combination=names(correlation.fisher.cld$Letters),
+			letter=unname(toupper(cld(correlation.GLHT)$mcletters$Letters),
+			Combination=as.character(names(toupper(cld(correlation.GLHT)$mcletters$Letters))),
 			stringsAsFactors=FALSE
@@
 		data.frame(
-			letter=unname(scenario.fisher.cld$Letters),
-			Combination=names(scenario.fisher.cld$Letters)
+			letter=unname(toupper(cld(scenario.GLHT)$mcletters$Letters)),
+			Combination=as.character(names(toupper(cld(scenario.GLHT)$mcletters$Letters)))
 		) %>% mutate(Combination=as.character(Combination)),

### 2016-03-31 R/analysis/07-surrogacy-prioritizations.R

@@ -23,7 +23,7 @@ env.correlation.prioritisations <- llply(
 					neutral.genetic.targets=NA,
 					Threads=general.params.LST[[MODE]]$threads,
 					MIPGap=gurobi.params.LST[[MODE]]$MIPGap,
-					NumberSolutions=rapr.params.LST[[MODE]]$surrogacy.analysis$surrogate.optimal.replicates
+					NumberSolutions=rapr.params.LST[[MODE]]$surrogacy.analysis$optimal.replicates
 				)
 			}
 		)
@@ -31,6 +31,53 @@ env.correlation.prioritisations <- llply(
 	.progress='text',
 	.parallel=FALSE
 )
+
+cat("starting random prioritizations for the environmental surrogate prioritizations \n")
+clust <- makeCluster(general.params.LST[[MODE]]$threads, type='SOCK')
+clusterEvalQ(clust, {library(rapr)})
+clusterExport(clust, c('spp.samples.DF','grid.DF','rapr.params.LST',
+	'MODE', 'species.prioritisation', 'ru', 'env.correlation.prioritisations'))
+registerDoParallel(clust)
+env.random.prioritisations <- llply(
+	seq_along(unique(spp.samples.DF$species)),
+	function(i) {
+		llply(
+			seq_along(rapr.params.LST[[MODE]]$surrogacy.analysis$surrogate.target),
+			function(j) {
+				setMKLthreads(1)
+				# extract number of planning units
+				curr.optim_n <- table(rowSums(env.correlation.prioritisations[[i]][[j]]@results@selections))
+				curr.optim_p <- curr.optim_n / sum(curr.optim_n)
+				# identify which planning units are occupied by the species
+				curr.pu <- which(grid.DF[[unique(spp.samples.DF$species)[[i]]]]==1)
+				zeros <- rep(0, length=nrow(grid.DF))
+				# generate portfolio of random selections
+				curr.random_n <- as.numeric(sample(
+					names(curr.optim_n),
+					size=rapr.params.LST[[MODE]]$surrogacy.analysis$random.replicates,
+					replace=TRUE, prob=as.numeric(curr.optim_p)
+				))
+				curr.sol.MTX<-t(sapply(
+					curr.random_n,
+					function(n) {replace(zeros, sample(x=curr.pu, size=n), rep(1, n))}
+				))
+				# generate RapSolved object
+				species.prioritisation(
[…]

### 2016-03-31 R/analysis/11-statistical-analysis.R

@@ -8,38 +8,39 @@ load('results/.cache/07-surrogacy-prioritizations.rda')
 ### surrogacy analyses
 # prepare data
 correlation.DF$Surrogate.target <- as.factor(correlation.DF$Surrogate.target)
-correlation.DF$Surrogate.target.Type <- with(correlation.DF, interaction(Surrogate.target, Type))
-correlation.sub.DF <- correlation.DF[rowSums(apply(select(correlation.DF, Surrogate.target, Type), 2, is.na))==0,]
+correlation.DF$Type.Surrogate.target.Method <- with(correlation.DF, interaction(Type, Surrogate.target, Method))
+correlation.sub.DF <- correlation.DF[rowSums(apply(select(correlation.DF, Type, Surrogate.target, Method), 2, is.na))==0,]
 
-# prepare matrix with success vs. failures
-correlation.contingency.DF <- correlation.sub.DF %>% group_by(Surrogate.target,Type) %>% filter(!is.na(genetic.held)) %>% summarize(count=sum(genetic.held>=0.8), total=length(genetic.held)) %>% ungroup %>% mutate(prop=count/total, Surrogate.target2=as.numeric(as.character(Surrogate.target))) %>% mutate(secured=count, not.secured=total-count, name=paste0(Surrogate.target, '.', Type)) %>% select(secured, not.secured, name) %>% data.frame
-correlation.MTX <- as.matrix(select(correlation.contingency.DF, secured, not.secured))
-rownames(correlation.MTX) <- correlation.contingency.DF$name
+# run models
+full.correlation.GLMM <- glmer(genetic.held~Type.Surrogate.target.Method + (1|Species), data=correlation.sub.DF, family='binomial', nAGQ=10, control=glmerControl(optimizer=c('bobyqa'),optCtrl=list(maxfun=1e5)))
+null.correlation.GLMM <- glmer(genetic.held~ 1 + (1|Species), data=correlation.sub.DF, family='binomial', nAGQ=10, control=glmerControl(optimizer=c('bobyqa'),optCtrl=list(maxfun=1e5)))
 
-# run fisher test
-correlation.fisher.test <- fisher.test(correlation.MTX, workspace=1e8)
+# compare models
+correlation.ANOVA <- anova(full.correlation.GLMM, null.correlation.GLMM, test='Chisq')
 
-# run post-hoc fisher test
[…]

### 2016-03-31 article/figures.Rmd
+ assorted bug fixes

@@
 		data.frame(
-			letter=unname(toupper(cld(correlation.GLHT)$mcletters$Letters),
+			letter=unname(toupper(cld(correlation.GLHT)$mcletters$Letters)),
 			Combination=as.character(names(toupper(cld(correlation.GLHT)$mcletters$Letters))),
@@
 # make figure
-p1 <- ggplot(aes(x=Surrogate.target,y=genetic.held),
+p1 <- ggplot(aes(x=Surrogate.target,y=genetic.held, fill=Method),
 		data=filter(correlation.PDF, Type=='Environmental')) +
@@
 	ylab('Adaptive genetic\nvariation secured (%)') +
-	geom_abline(intercept=rapr.params.LST[[MODE]]$surrogacy.analysis$genetic.target,
-		slope=0, color='red', linetype='dashed') +
 	geom_text(aes(x=0.55, y=1.115, label='(a)'), color='black', size=5) +
 	theme_classic() +
-	theme(legend.position='none') +
-	geom_text(aes(x=Surrogate.target, y=genetic.held, label=letter),
+	theme(legend.position=c(0.9,0.4)) +
+	scale_fill_manual(name='Prioritization', values=c('Optimal'='white', 'Random'='grey50')) +
+	geom_text(aes(x=Surrogate.target, y=genetic.held, fill=Method, label=letter),
 		data=filter(correlation.PDF2, Type=='Environmental'),
@@
 
-p2 <- ggplot(aes(x=Surrogate.target,y=genetic.held),
+p2 <- ggplot(aes(x=Surrogate.target,y=genetic.held, fill=Method),
 			data=filter(correlation.PDF, Type=='Geographic')) +
@@
 	ylab('Neutral genetic\nvariation secured (%)') +
-	geom_abline(intercept=rapr.params.LST[[MODE]]$surrogacy.analysis$genetic.target, 
-		slope=0, color='red', linetype='dashed') +
 	geom_text(aes(x=0.55, y=1.115, label='(b)'), color='black', size=5) +
@@
 	theme(legend.position='none') +
-	geom_text(aes(x=Surrogate.target, y=genetic.held, label=letter),
+	scale_fill_manual(name='Prioritization', values=c('Optimal'='white', 'Random'='grey50')) +
+	geom_text(aes(x=Surrogate.target, y=genetic.held, fill=Method, label=letter),
 		data=filter(correlation.PDF2, Type=='Geographic'),
@@
 	mutate(
-		letter.height=max(value, na.rm=TRUE)+0.1,
+		letter.height=max(genetic.held, na.rm=TRUE)+0.1,
 		Combination=paste0(Prioritisation,'.',Metric,'.',Context)) %>%
@@
 suppressWarnings({
-	ggplot(aes(x=Prioritisation,y=value),
+	ggplot(aes(x=Prioritisation,y=genetic.held),
 		data=scenario.DF) +
 		geom_boxplot(position=position_dodge(0.9)) +
-		geom_abline(
-			intercept=rapr.params.LST[[MODE]]$scenario.analysis$genetic.target,
-			slope=0, color='red',linetype='dashed') +
 		geom_text(aes(x=Prioritisation, y=letter.height, label=letter),

### 2016-03-31 article/supporting_information.Rmd
+ assorted bug fixes

@@
 		'long','lat','group'),drop=FALSE]
+	curr.FPLY <- curr.FPLY[which(!is.na(curr.FPLY[[1]])),]
 	curr.FPLY <- curr.FPLY %>% gather(d, value, contains('adaptive'), contains('neutral')) %>% mutate(

### 2016-03-31 article/article.Rmd
+ manuscript fixes + si adaptive + neutral space graphs

--- 2016-03-30 article/article.Rmd
@@
 For every prioritization we generated, we calculated the proportion of the adaptive (if detected) and neutral genetic variation that every prioritization secured to assess its performance.
+Prioritizations associated with negative values--because they secured only a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis.
 We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality using the $\texttt{rapr}$ R package [@r465] and $\texttt{Gurobi}$ [@r463].
@@
 For each species, we first generated two sets of prioritizations using increasing environmental and geographic-based targets (separately).
-Prioritizations were generated using surrogate-based targets ranging from `r (min(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target)*100)` % and `r (max(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target)*100)` % in increments of `r diff(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target[1:2])*100` % ($n = `r rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.optimal.replicates`$ per target).
-Prioritizations associated with negative values--because they secured only a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis.
+Prioritizations were generated using the following surrogate-based targets: `r paste(rapr.params.LST[[MODE]]$surrogate.target,collapse=', ')`.
+We then create a set of null expectations with which to compare the surrogate-based prioritizations.
+These null expectations represented possible solutions obtained using conventional conservation planning methods that just use amount (eg.
+Marxan).
+For each species, for each target, we generated `r ` prioritizations by randomly selecting planning units occupied by the species.
+The number of randomly selected planning units was set to equal the number of planning units in the optimal prioritizations.
 
@@
 The purely amount-based prioritizations served as our baseline--representing the prioritizations typically generated in conservation planning exercises.
-We used `r (rapr.params.LST[[MODE]][['single.species']]$amount.target*100)` % amount-based targets in all prioritizations to ensure an adequate proportion of habitat was secured for each species.
-We set surrogate- and genetic-based targets as `r (rapr.params.LST[[MODE]][['single.species']]$amount.target*100)` % to ensure that most of the respective variation was secured for each species.
-To accomodate variation in the high number of optimal solutions for the single-species amount-based prioritizations, we generated `r (rapr.params.LST[[MODE]][['single.species']]$amount.replicates)` replicate prioritizations.
-Since all other contexts were likely to be associated with few optimal solutions, we generated only `r rapr.params.LST[[MODE]][['multi.species']]$genetic.replicates` replicates.
+We used `r (rapr.params.LST[[MODE]]$scenario.analysis$amount.target*100)` % amount-based targets in all prioritizations to ensure an adequate proportion of habitat was secured for each species.
+Based on the previous analysis, we set surrogate-based targets as `r (rapr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100)` % and genetic-based targets as `r (rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)` % to ensure that most of the variation was secured for each species.
+To accomodate variation in the high number of optimal solutions for the single-species amount-based prioritizations, we generated `r (rapr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates)` replicate prioritizations.
 

### 2016-03-31 article/supporting_information.Rmd
+ manuscript fixes + si adaptive + neutral space graphs

--- 2016-03-18 article/supporting_information.Rmd
@@
 	# make plot
+	curr.GROBS <- llply(unique(curr.FPLY$d), function(x) {
 		ggplot() +
@@
 				fill='grey85', color='grey70') +
-		geom_polygon(data=curr.FPLY, aes_string(x='long', y='lat', 
+			geom_polygon(data=filter(curr.FPLY, d==x), aes_string(x='long', y='lat', 
 				group='group', fill='value'), color='grey10') +
@@
 					border=element_line(color='black'), barheight=unit(7, 'cm'))
-		) +
-		facet_wrap(~ d, ncol=4)
+			)
+	})
+	do.call(grid.arrange, append(curr.GROBS, list(ncol=4)))
+	invisible()
 }

### 2016-04-01 article/supporting_information.Rmd
+ include faceting in si genetic maps

--- 2016-03-18 article/supporting_information.Rmd
@@
 					border=element_line(color='black'), barheight=unit(7, 'cm'))
-		) +
-		facet_wrap(~ d, ncol=4)
+			) + facet_wrap(~d)
+	})
+	do.call(grid.arrange, append(curr.GROBS, list(ncol=4)))
+	invisible()
 }

### 2016-04-01 article/article.Rmd
+ edits to article

@@
 Protected area networks buffer species from anthropogenic impacts and provide places for populations of species to persist.
-Intra-specific genetic variation underpins species persistence, and there has been increasing interest in understanding how to design protected area networks in a way that maximizes this variation [@r67; @r3; @r61].
+Intra-specific genetic variation underpins species persistence, and there has been increasing interest in understanding how to design protected area networks in a way that maximizes this variation for minimal cost [@r67; @r3; @r61].
 Broadly speaking, genetic variation can be classified as either adaptive or neutral [reviewed in @r48].
@@
 
-Recently, conservation scientists have been investigating surrogates for genetic variation to generate better prioritizations without the overheads associated with genetic data.
+Recently, conservation scientists have been investigating surrogates for intra-specific genetic variation to generate prioritizations that secure itra-specific genetic variation without the overheads associated with using it.
 Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been used to guide reserve selection [eg.
@@
 Here we investigate the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation in the context of conservation planning.
-First, we examine the relationship between surrogates and genetic variation, and verify whether these variables are effective surrogates.
-Second, we determine if the inclusion of surrogates in various conservation planning scenarios can improve the proportion of genetic variation they secure.
+Conventionally, prioritizations are generated by setting targets on the total amount of habitat that needs to be preserved for each species (amount-based targets), and finding the combination of planning units that fulfills all the targets for minimal cost [eg.
+using \texttt{Marxan}; @r22].
+First, we aim to verify that environmental and geographic variables are effective surrogates for adaptive and neutral variation (respectively).
+We expect that prioritizations generated using surrogate-based targets will secure more genetic variation than prioritizations generated using only amount-based targets with the same number of planning units.
+Second, we aim to determine if the addition of surrogate-based targets to conservation planning exercises can improve the proportion of genetic variation they secure.
+We expect that prioritizations generated using amount- and surrogate-based targets will secure more genetic variation than prioritizations generated using only amount-based targets, and that they will secure a similar amount of genetic variation to prioritizations generated using amount- and genetic based targets.
+Additionally, we expect that prioritizations generated using amount- and surrogate-based targets will share many planning units with prioritizations generated using amount- and genetic-based targets.
 If these surrogates can improve the intra-specific genetic variation captured in a prioritization, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.
@@
 We also treated the adaptive (if detected) and neutral variables as separate attribute spaces for each species.
-We set the demand points for each species and each attribute space using the values associated with the planning units.
+We set the demand points for each species and attribute space using the values associated with the planning units.
 For every prioritization we generated, we calculated the proportion of the adaptive (if detected) and neutral genetic variation that every prioritization secured to assess its performance.
@@
 
-To determine if the environmental and geographic variables are effective surrogates for adaptive and neutral genetic variation, we generated a collection of single-species prioritizations.
+To address the first aim and determine if the environmental and geographic variables are effective surrogates for adaptive and neutral genetic variation, we generated a collection of single-species prioritizations.
 For each species, we first generated two sets of prioritizations using increasing environmental and geographic-based targets (separately).
-Prioritizations were generated using the following surrogate-based targets: `r paste(rapr.params.LST[[MODE]]$surrogate.target,collapse=', ')`.
-We then create a set of null expectations with which to compare the surrogate-based prioritizations.
-These null expectations represented possible solutions obtained using conventional conservation planning methods that just use amount (eg.
-Marxan).
-For each species, for each target, we generated `r ` prioritizations by randomly selecting planning units occupied by the species.
-The number of randomly selected planning units was set to equal the number of planning units in the optimal prioritizations.
+Prioritizations were generated using the following surrogate-based targets: `r paste(rapr.params.LST[[MODE]]$surrogacy.analysis$surrogate.target[-length(rapr.params.LST[[MODE]]$surrogacy.analysis$surrogate.target)],collapse=', ')`, and `r rapr.params.LST[[MODE]]$surrogacy.analysis$surrogate.target[length(rapr.params.LST[[MODE]]$surrogacy.analysis$surrogate.target)]`.
+We then created two corresponding sets of amount-based prioritizations with which to compare the surrogate-based prioritizations.
+For each species, for each surrogate-based target, we generated `r rapr.params.LST[[MODE]]$surrogacy.analysis$random.replicates` replicate prioritizations by randomly selecting planning units occupied by the species.
+The number of randomly selected planning units in each prioritization was set to equal the number of planning units in the surrogate-based prioritization.
 
-To determine if the use of surrogate-based targets can actually improve prioritizations in real-world conservation planning scenarios, we investigated the effectiveness of surrogate-based targets using three scenarios with increasing levels of complexity.
+To address the second aim and determine if the addition of surrogate-based targets can result in more effective prioritizations, we generated a suite of prioritizations using different types of targets.
 To assess the benefits of surrogate-based targets under the simplest of circumstances, our first scenario involved generating single-species prioritizations for each species.
-However, most large-scale prioritizations involve a comprehensive set of species.
-By forcing solutions to secure individuals in multiple communities, this may result in prioritizations that secure a greater proportion of intra-specific genetic variation for each species.
-To test for this effect, our second scenario involved generating multi-species prioritizations.
-Finally, most large-scale prioritizations also explicitly consider opportunity cost to ensure that prioritizations are cost-effective.
+To assess the benefits in a more complex situation, our second scenario involved generating multi-species prioritizations.
+Since most large-scale prioritizations also explicitly consider opportunity cost to ensure that prioritizations are cost-effective.
 Our third scenario involved generating multi-species prioritizations that satisfied targets for minimal opportunity cost.
-
 For each scenario, we generated three types of prioritizations using (1) only amount-based targets, (2) both amount- and surrogate-based targets, and (3) both amount- and genetic-based targets.
@@
 We investigated the effectiveness of prioritizations generated using surrogate-based targets using generalized linear mixed-effects models [GLMMs; using a logit link and the $\texttt{lme4}$ R package; @r469].
-To determine  if the environmental and geographic variables were effective surrogates, we fit a full model to the proportion of genetic variation that the first collection of prioritizations.
-This full model contained categorical predictor variables to indicate the surrogate-based targets used to generate them, the type of genetic variation (adaptive or neutral), and the interactions between these variables.
-To determine if surrogate-based targets improved the effectiveness of prioritizations under various scenarios, we fit a full model to the proportion of genetic variation that second collection of prioritizations secured.
-This second full model contained categorical predictor variables to indicate the scenario (single-species, multi-species, multi-species with cost), the combination of targets used to generate the prioritizations (amount-based targets, amount and surrogate-based targets, amount and genetic-based targets), the type of genetic variation measured (adaptive or neutral), and the interactions between them.
-All models contained intercept-level random effects to accommodate variation between species.
-We conducted a post-hoc analysis on each full model using Tukey contrasts with a Bonferroni correction [using the $\texttt{multcomp}$ R package; @r470].
+To address the first aim, we fit a full model to the proportion of genetic variation that the first collection of prioritizations.
+This full model contained a categorical predictor variable indicating the surrogate-based targets used to generate them, the type of targets used to generate the prioritization (amount or surrogate-based targets), and the type of genetic variation (adaptive or neutral).
+To address the second aim, we fit a full model to the proportion of genetic variation that second collection of prioritizations secured.
+This second full model contained a categorical predictor variable indicating the scenario (single-species, multi-species, multi-species with cost), the combination of targets used to generate the prioritizations (amount-based targets, amount and surrogate-based targets, amount and genetic-based targets), and the type of genetic variation measured (adaptive or neutral).
+Both models contained intercept-level random effects to accommodate variation between species.
+We also conducted a post-hoc analysis on both models using Tukey contrasts with a Bonferroni correction [using the $\texttt{multcomp}$ R package; @r470].
 
@@
 <!-- Surrogacy analysis -->
-The environmental and geographic-based variables were effective surrogates for adaptive and genetic variation (Figure 1).
-The prioritizations generated using greater surrogate-based targets secured more genetic variation (XXXX).
-In addition, prioritizations generated using environmental targets often secured more adaptive genetic variation than prioritizations generated using geographic-based targets did neutral genetic variation (XXXX).
-Only particularly high geographic targets ($\geq 0.95$ %) were able to secure a moderate proportion of neutral variation ($\geq XXXX$ %).
-Overall, there was a non-linear relationship between the surrogate-based targets used to generate the prioritizations and the proportion of genetic variation they secured.
+The environmental and geographic-based variables were not effective surrogates for adaptive and genetic variation (Figure 1).
+Although prioritizations generated using greater surrogate-based targets secured more genetic variation (), they did not secure more genetic variation than prioritizations generated using just amount-based targets ().
+Moreover, the geographic variables were particularly poor surrogates because they often performed worse than just using amount-based targets on average.
+Overall, these results suggest that environmental and geographic surrogates are poor predictor for adaptive and neutral genetic variation in the context of conservation planning.
 

### 2016-04-04 article/article.Rmd
+ omit species with 1 population from analysis + use nonparametric g tests in scenario analysis + use no stats in surrogacy analysis + increase random reps to 1k

--- 2016-03-31 article/article.Rmd
@@
 Specifically, we used non-metric multi-dimensional scaling [NMDS; implemented in the $\texttt{vegan R}$ package; @r458] using Gower distances [@r464; using the $\texttt{cluster R}$ package to accommodate sparsity; @r457] to transform the binary loci data into continuous variables (Table S2).
-To ensure that the ordinations described a sufficient amount of the genomic variation, we ran successive scaling analyses with increasing dimensionality until this was achieved (maximum stress value $\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for each analysis).
-We then fit a suite of Gaussian mixture to the ordinations to estimate the probability that individuals belonged to different populations [assuming different covariance structures implemented in the \texttt{mclust} R package; @r477; @r476].
+To ensure that the ordinations described a sufficient amount of the genomic variation, we ran successive scaling analyses with increasing dimensionality until this was achieved (maximum stress value $\leq$ `r nmds.params.LST[[MODE]]$population.clustering$max.stress`; `r nmds.params.LST[[MODE]]$population.clustering$trymax` random starts for each analysis).
+We then fit a suite of Gaussian mixture models to the ordinations to estimate the probability that individuals belonged to different populations [assuming different covariance structures implemented in the \texttt{mclust R} package; @r477; @r476].
 For each parametrization, we fit a set of models assuming varying numbers of total populations (\textit{G}= 1--`r max(mclust.params.LST[[MODE]]$G)`).
@@
 Since outlier detection methods compare allele frequencies between populations, we were unable to examine species that contained a single population.
-For these species ($n = `r sum(sapply(spp.Mclust.LST, function(x) return(ncol(x$mclust$z)==1)))`$), we assumed that all loci were neutral.
-For species with more than a single population, we ran $\texttt{Bayescan}$ [version 2.1; 1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations; @r456].
+These species were omitted from subsequent analysis ($n = `r sum(sapply(spp.Mclust.LST, function(x) return(ncol(x$mclust$z)==1)))`$).
+For the remaning species, we ran $\texttt{Bayescan}$ [version 2.1; 1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations; @r456].
 To avoid false-positives, we omitted loci where the global frequency of the minor allele was $\geq 0.05$ (following guidelines in the user manual) and assumed they were neutral.
@@
 After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species (Figures S6--S32).
-For each species, we partitioned adaptive and neutral loci into separate data sets, and reran the successive nonmetric multidimensional scaling analyses (using parameters described above) to derive continuous variables that described the main gradients of adaptive and neutral genetic variation separately (Table S2).
+For each species, we partitioned adaptive and neutral loci into separate data sets, and reran the successive nonmetric multidimensional scaling analyses (using parameters described above but with maximum stress values $\leq$ `r nmds.params.LST[[MODE]]$genetic.space$max.stress`) to derive continuous variables that described the main gradients of adaptive and neutral genetic variation separately (Table S2).
 Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the individuals in each grid cell to express the typical genomic characteristics of individuals in the cells.

### 2016-04-04 article/figures.Rmd
+ omit species with 1 population from analysis + use nonparametric g tests in scenario analysis + use no stats in surrogacy analysis + increase random reps to 1k

@@
 Similarly, panel (b) shows that for geographic surrogate targets and neutral genetic variation.
-Box-plots show inter-specific variation in the proportion of genetic variation secured using different surrogate-based targets.
-Letters indicate significant differences ($P < 0.05$).'}
+Box-plots show inter-specific variation in the proportion of genetic variation secured using different surrogate-based targets.'}
 # prepare data
@@
 	data.frame()
-
-correlation.PDF2 <- correlation.PDF %>%
-	group_by(Type, Surrogate.target, Method) %>% 
-	summarize(genetic.held=max(genetic.held, na.rm=TRUE)+0.1) %>%
-	ungroup() %>%
-	mutate(Combination=as.character(paste0(Type,'.',Surrogate.target,'.',Method))) %>% 
-	left_join(
-		data.frame(
-			letter=unname(toupper(cld(correlation.GLHT)$mcletters$Letters)),
-			Combination=as.character(names(toupper(cld(correlation.GLHT)$mcletters$Letters))),
-			stringsAsFactors=FALSE
-		),
-		by='Combination'
-	) %>% data.frame()
 
@@
 	theme(legend.position=c(0.9,0.4)) +
-	scale_fill_manual(name='Prioritization', values=c('Optimal'='white', 'Random'='grey50')) +
-	geom_text(aes(x=Surrogate.target, y=genetic.held, fill=Method, label=letter),
-		data=filter(correlation.PDF2, Type=='Environmental'),
-		position=position_dodge(0.9))
+	scale_fill_manual(name='Prioritization', values=c('Optimal'='white', 'Random'='grey50'))
 
@@
 	scale_fill_manual(name='Prioritization', values=c('Optimal'='white', 'Random'='grey50')) +
-	geom_text(aes(x=Surrogate.target, y=genetic.held, fill=Method, label=letter),
-		data=filter(correlation.PDF2, Type=='Geographic'),
-		position=position_dodge(0.9))
 
@@
 # prepare data for plotting
-grid.FPLY <- grid.PLY
+grid.sub.FPLY <- grid.sub.PLY
 for (i in seq_along(single.spp.prioritisations[[1]]))
-	grid.FPLY@data[[paste0('Single-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations, function(x){colMeans(x[[i]]@results@selections)}))
+	grid.sub.FPLY@data[[paste0('Single-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations, function(x){colMeans(x[[i]]@results@selections)}))
 for (i in seq_along(multi.spp.prioritisations))
-	grid.FPLY@data[[paste0('Multi-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations[[i]]@results@selections)
+	grid.sub.FPLY@data[[paste0('Multi-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations[[i]]@results@selections)
 for (i in seq_along(multi.spp.prioritisations.with.cost))
-	grid.FPLY@data[[paste0('Multi-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations.with.cost[[i]]@results@selections)
-grid.FPLY <- spFortify(grid.FPLY)
-grid.FPLY <- grid.FPLY %>% gather(
+	grid.sub.FPLY@data[[paste0('Multi-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations.with.cost[[i]]@results@selections)
+grid.sub.FPLY <- spFortify(grid.sub.FPLY)
+grid.sub.FPLY <- grid.sub.FPLY %>% gather(
 		Combination,
@@
 	)
-panel.DF <- grid.FPLY %>% select(Context, Prioritization, long, lat) %>% 
+panel.DF <- grid.sub.FPLY %>% select(Context, Prioritization, long, lat) %>% 
 	mutate(
@@
 		fill='grey85', color='grey70') +
-	geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
+	geom_polygon(data=grid.sub.FPLY, aes_string(x='long', y='lat', 
 		group='group', fill='Selection.Frequency'),
@@
 	coord_cartesian(
-		xlim=buffered.range(grid.FPLY$long, 0),
-		ylim=buffered.range(grid.FPLY$lat, 0)
+		xlim=buffered.range(grid.sub.FPLY$long, 0),
+		ylim=buffered.range(grid.sub.FPLY$lat, 0)
 	) +

### 2016-04-04 article/supporting_information.Rmd
+ omit species with 1 population from analysis + use nonparametric g tests in scenario analysis + use no stats in surrogacy analysis + increase random reps to 1k

--- 2016-03-31 article/supporting_information.Rmd
@@
 grid.FPLY <- spFortify(grid.PLY)
+grid.sub.FPLY <- spFortify(grid.sub.PLY)
 
@@
 
-```{r, echo=FALSE, fig.width=9.5, fig.height=10, out.width='7.5in', out.height='8.5in', fig.cap='Species distributions.
+```{r, echo=FALSE, fig.width=9.5, fig.height=10, out.width='7.5in', out.height='8.5in', fig.cap=paste0('Species distributions.
 Squares represent planning units.
-For a given species, planning units that were found to be inhabited are denoted with bright blue.'}
+For a given species, planning units that were found to be inhabited are denoted with bright blue.
+Note that \textit{',missing.species,'} was omitted from analysis.')}
 ## plot map of species distributions
@@
 			strip.background = element_rect(fill='grey20'),
-			strip.text = element_text(color='white', face='italic', size=20)
+			strip.text = element_text(color='white', face='italic', size=1)
 		) +
@@
 plot.spp.pop.probs <- function(i) {
-	# init
+	# init spp.samples
 	prop <- 0.85
@@
 		paste0(unique(spp.samples.DF$species)[i], '_adaptive_d'),
-		names(grid.FPLY), value=TRUE, fixed=TRUE
+		names(grid.sub.FPLY), value=TRUE, fixed=TRUE
 	)
@@
 		paste0(unique(spp.samples.DF$species)[i], '_neutral_d'),
-		names(grid.FPLY), value=TRUE, fixed=TRUE
+		names(grid.sub.FPLY), value=TRUE, fixed=TRUE
 	)
+	# skip species if not used in analysis
+	if (length(c(adapt.col, neutral.col)==0)
+		return(NULL)
 	curr.FPLY <- grid.FPLY[,
@@
 	curr.FPLY <- curr.FPLY %>% gather(d, value, contains('adaptive'), contains('neutral')) %>% mutate(
-		d=gsub(paste0(unique(spp.samples.DF$species)[i], '_'), '', d, fixed=TRUE),
+		d=gsub(paste0(unique(spp.samples.sub.DF$species)[i], '_'), '', d, fixed=TRUE),
 		d=gsub('_', ' ', d, fixed=TRUE)
@@
 	# init
-	adapt.col <- paste0(unique(spp.samples.DF$species)[i], '_adaptive_d1')
-	neutral.col <- paste0(unique(spp.samples.DF$species)[i], '_neutral_d1')
+	adapt.col <- paste0(unique(spp.samples.sub.DF$species)[i], '_adaptive_d1')
+	neutral.col <- paste0(unique(spp.samples.sub.DF$species)[i], '_neutral_d1')
 	n.plots <- 0 
@@
 				function(i) {
-					ldply(
-						seq_along(spp.nmds.LST[[i]]),
-						function(j) {
+
 				rbind(
@@
 					),
+					ldply(
+						seq_along(spp.nmds.LST[[i]]),
+						function(j) {
 							data.frame(
@@
 								Converged=ifelse(is.null(spp.nmds.LST[[i]][[j]]), NA, spp.nmds.LST[[i]][[j]]$converged)
-							)
 							)

### 2016-04-04 R/analysis/05-genetic-nmds.R

@@ -18,10 +18,8 @@ spp.nmds.LST <- llply(
 		setMKLthreads(1)
 		cat('starting species',i,'\n')
 		if (is.null(spp.BayeScan.sample.loci.subset.LST[[i]])) {
-			# all loci are neutral since only 1 population
-			curr.spp <- spp.BayeScanData.LST[[i]]
-			# manually classify loci as neutral or adaptive
-			curr.spp.type <- rep('neutral', bayescanr:::n.loci(curr.spp))
+			# only 1 population in this species
+			return(list('adaptive'=NULL, 'neutral'=NULL))
 		} else {
 			# extract loci
 			curr.spp <- bayescanr:::loci.subset(
@@ -44,10 +42,10 @@ spp.nmds.LST <- llply(
 				curr.nmds <- bayescanr::nmds(
 					bayescanr:::loci.subset(curr.spp, curr.spp.type==j),
 					metric='gower',
-					max.stress=nmds.params.LST[[MODE]]$max.stress,
-					min.k=nmds.params.LST[[MODE]]$min.k,
-					max.k=nmds.params.LST[[MODE]]$max.k,
-					trymax=nmds.params.LST[[MODE]]$trymax
+					max.stress=nmds.params.LST[[MODE]]$genetic.space$max.stress,
+					min.k=nmds.params.LST[[MODE]]$genetic.space$min.k,
+					max.k=nmds.params.LST[[MODE]]$genetic.space$max.k,
+					trymax=nmds.params.LST[[MODE]]$genetic.space$trymax
 				)
 			)
 		}), c('adaptive','neutral')))

### 2016-04-04 R/analysis/11-statistical-analysis.R

@@ -6,23 +6,7 @@ load('results/.cache/07-surrogacy-prioritizations.rda')
 
 #### Statistical analyses
 ### surrogacy analyses
-# prepare data
-correlation.DF$Surrogate.target <- as.factor(correlation.DF$Surrogate.target)
-correlation.DF$Type.Surrogate.target.Method <- with(correlation.DF, interaction(Type, Surrogate.target, Method))
-correlation.sub.DF <- correlation.DF[rowSums(apply(select(correlation.DF, Type, Surrogate.target, Method), 2, is.na))==0,]
-
-# run models
-full.correlation.GLMM <- glmer(genetic.held~Type.Surrogate.target.Method + (1|Species), data=correlation.sub.DF, family='binomial', nAGQ=10, control=glmerControl(optimizer=c('bobyqa'),optCtrl=list(maxfun=1e5)))
-null.correlation.GLMM <- glmer(genetic.held~ 1 + (1|Species), data=correlation.sub.DF, family='binomial', nAGQ=10, control=glmerControl(optimizer=c('bobyqa'),optCtrl=list(maxfun=1e5)))
-
-# compare models
-correlation.ANOVA <- anova(full.correlation.GLMM, null.correlation.GLMM, test='Chisq')
-
-# posthoc analysis
-correlation.GLHT <- summary(
-	glht(full.correlation.GLMM, linfct=mcp(Type.Surrogate.target.Method='Tukey')),
-	adjusted('bonferroni')
-)
+# none because significance can be acheived by generating sufficient random reps 
 
 ### scenario analyses
 ## prepare data
@@ -66,18 +50,26 @@ scenario.DF <- rbind.fill(list(single.spp.SDF, multi.spp.SDF, multi.spp.with.cos
 # remove NA values
 scenario.sub.DF <- scenario.DF[rowSums(apply(select(scenario.DF, Prioritisation, Metric, Context), 2, is.na))==0,]
 
-# run models
-full.scenario.GLMM <- glmer(genetic.held~Prioritisation.Metric.Context + (1|Species), data=scenario.sub.DF, family='binomial', nAGQ=10, control=glmerControl(optimizer=c('bobyqa'),optCtrl=list(maxfun=1e5)))
[…]

### 2016-04-04 article/figures.Rmd
+ fixes to si and figures

@@
 			axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
+		geom_abline(intercept=intercept=rapr.params.LST[[MODE]]$surrogacy.analysis$genetic.target,
+			slope=0 color='red', linetype='dashed') +
 		facet_grid(Context ~ Metric)

### 2016-04-04 article/supporting_information.Rmd
+ fixes to si and figures

@@
 For a given species, planning units that were found to be inhabited are denoted with bright blue.
-Note that \textit{',missing.species,'} was omitted from analysis.')}
+Note that \\textit{',gsub('\\_', '\\ ', missing.species),'} was omitted from analysis.')}
 ## plot map of species distributions
@@
 		paste0(unique(spp.samples.DF$species)[i], '_adaptive_d'),
-		names(grid.sub.FPLY), value=TRUE, fixed=TRUE
+		names(grid.FPLY), value=TRUE, fixed=TRUE
 	)
@@
 		paste0(unique(spp.samples.DF$species)[i], '_neutral_d'),
-		names(grid.sub.FPLY), value=TRUE, fixed=TRUE
+		names(grid.FPLY), value=TRUE, fixed=TRUE
 	)
 	# skip species if not used in analysis
-	if (length(c(adapt.col, neutral.col)==0)
-		return(NULL)
 	curr.FPLY <- grid.FPLY[,
@@
 	curr.FPLY <- curr.FPLY %>% gather(d, value, contains('adaptive'), contains('neutral')) %>% mutate(
-		d=gsub(paste0(unique(spp.samples.sub.DF$species)[i], '_'), '', d, fixed=TRUE),
+		d=gsub(paste0(unique(spp.samples.DF$species)[i], '_'), '', d, fixed=TRUE),
 		d=gsub('_', ' ', d, fixed=TRUE)
@@
 	# init
-	adapt.col <- paste0(unique(spp.samples.sub.DF$species)[i], '_adaptive_d1')
-	neutral.col <- paste0(unique(spp.samples.sub.DF$species)[i], '_neutral_d1')
+	adapt.col <- paste0(unique(spp.samples.DF$species)[i], '_adaptive_d1')
+	neutral.col <- paste0(unique(spp.samples.DF$species)[i], '_neutral_d1')
 	n.plots <- 0 
@@
 	# exports
-	return(ceiling(n.plots/4) * 3.5)
+	return(((n.plots %/% 5)+1) * 2.5)
 }
@@
 No indiviudals of this species were found inside the gray planning units.')}
-if (1 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(1)
+if (1 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[1] %in% missing.species)) plot.spp.nmds(1)
  ```
@@
 See Figure S6 caption for conventions.')}
-if (2 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(2)
+if (2 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[2] %in% missing.species)) plot.spp.nmds(2)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (3 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(3)
+if (3 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[3] %in% missing.species)) plot.spp.nmds(3)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (4 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(4)
+if (4 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[4] %in% missing.species)) plot.spp.nmds(4)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (5 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(5)
+if (5 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[5] %in% missing.species)) plot.spp.nmds(5)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (6 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(6)
+if (6 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[6] %in% missing.species)) plot.spp.nmds(6)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (7 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(7)
+if (7 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[7] %in% missing.species)) plot.spp.nmds(7)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (8 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(8)
+if (8 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[8] %in% missing.species)) plot.spp.nmds(8)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (9 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(9)
+if (9 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[9] %in% missing.species)) plot.spp.nmds(9)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (10 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(10)
+if (10 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[10] %in% missing.species)) plot.spp.nmds(10)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (11 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(11)
+if (11 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[11] %in% missing.species)) plot.spp.nmds(11)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (12 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(12)
+if (12 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[12] %in% missing.species)) plot.spp.nmds(12)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (13 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(13)
+if (13 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[13] %in% missing.species)) plot.spp.nmds(13)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (14 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(14)
+if (14 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[14] %in% missing.species)) plot.spp.nmds(14)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (15 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(15)
+if (15 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[15] %in% missing.species)) plot.spp.nmds(15)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (16 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(16)
+if (16 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[16] %in% missing.species)) plot.spp.nmds(16)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (17 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(17)
+if (17 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[17] %in% missing.species)) plot.spp.nmds(17)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (18 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(18)
+if (18 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[18] %in% missing.species)) plot.spp.nmds(18)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (19 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(19)
+if (19 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[19] %in% missing.species)) plot.spp.nmds(19)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (20 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(20)
+if (20 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[20] %in% missing.species)) plot.spp.nmds(20)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (21 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(21)
+if (21 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[21] %in% missing.species)) plot.spp.nmds(21)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (22 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(22)
+if (22 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[22] %in% missing.species)) plot.spp.nmds(22)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (23 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(23)
+if (23 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[23] %in% missing.species)) plot.spp.nmds(23)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (24 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(24)
+if (24 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[24] %in% missing.species)) plot.spp.nmds(24)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (25 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(25)
+if (25 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[25] %in% missing.species)) plot.spp.nmds(25)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (26 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(26)
+if (26 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[26] %in% missing.species)) plot.spp.nmds(26)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (27 <= general.params.LST[[MODE]]$n.spp) plot.spp.nmds(27)
+if (27 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[27] %in% missing.species)) plot.spp.nmds(27)
 ```
@@
 				function(i) {
-
 					rbind(
@@
 						})
+					)
 				}

### 2016-04-05 article/article.Rmd
+ fixes to article/si/figures

@@
 
-To determine if the environmental and geographic variables are effective surrogates for adaptive and neutral genetic variation, we generated a collection of single-species prioritizations.
+To address the first aim and determine if the environmental and geographic variables are effective surrogates for adaptive and neutral genetic variation, we generated a collection of single-species prioritizations.
 For each species, we first generated two sets of prioritizations using increasing environmental and geographic-based targets (separately).
@@
 
-To determine if the use of surrogate-based targets can actually improve prioritizations in real-world conservation planning scenarios, we investigated the effectiveness of surrogate-based targets using three scenarios with increasing levels of complexity.
+To address the second aim and determine if the use of surrogate-based targets can actually improve prioritizations in real-world conservation planning scenarios, we investigated the effectiveness of surrogate-based targets using three scenarios with increasing levels of complexity.
 To assess the benefits of surrogate-based targets under the simplest of circumstances, our first scenario involved generating single-species prioritizations for each species.
@@
 
-## Statistical analysis
-We investigated the effectiveness of prioritizations generated using surrogate-based targets using generalized linear mixed-effects models [GLMMs; using a logit link and the $\texttt{lme4}$ R package; @r469].
-To determine  if the environmental and geographic variables were effective surrogates, we fit a full model to the proportion of genetic variation that the first collection of prioritizations.
-This full model contained categorical predictor variables to indicate the surrogate-based targets used to generate them, the type of genetic variation (adaptive or neutral), and the interactions between these variables.
-To determine if surrogate-based targets improved the effectiveness of prioritizations under various scenarios, we fit a full model to the proportion of genetic variation that second collection of prioritizations secured.
-This second full model contained categorical predictor variables to indicate the scenario (single-species, multi-species, multi-species with cost), the combination of targets used to generate the prioritizations (amount-based targets, amount and surrogate-based targets, amount and genetic-based targets), the type of genetic variation measured (adaptive or neutral), and the interactions between them.
-All models contained intercept-level random effects to accommodate variation between species.
-We conducted a post-hoc analysis on each full model using Tukey contrasts with a Bonferroni correction [using the $\texttt{multcomp}$ R package; @r470].
+We then counted the number of species that were sufficiently represented in terms of their adaptive and neutral genetic variation (genetic variation secured $\geq$ `r (rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)` %) for each prioritization.
+We used a global $G$ test to determine the proportion of species that met their genetic target in a given prioritization depended on the combination of targets used to generate the planning units (amount-based targets, amount and surrogate-based targets, amount and genetic-based targets), the planning scenario (single-species, multi-species, multi-species with cost), or the type of genetic variation measured (adaptive or neutral).
+We also conducted a post-hoc analysis involving pairwise $G$ tests (using the \texttt{RVAideMemoire R} package; r@XXX).
 

### 2016-04-05 article/figures.Rmd
+ fixes to article/si/figures

--- 2016-04-04 article/figures.Rmd
@@
 	theme(legend.position='none') +
-	scale_fill_manual(name='Prioritization', values=c('Optimal'='white', 'Random'='grey50')) +
+	scale_fill_manual(name='Prioritization', values=c('Optimal'='white', 'Random'='grey50'))
 
@@
 		data.frame(
-			letter=unname(toupper(cld(scenario.GLHT)$mcletters$Letters)),
-			Combination=as.character(names(toupper(cld(scenario.GLHT)$mcletters$Letters)))
+			letter=unname(scenario.G.cld$Letters),
+			Combination=as.character(names(scenario.G.cld$Letters))
 		) %>% mutate(Combination=as.character(Combination)),
@@
 			axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
+		geom_abline(intercept=rapr.params.LST[[MODE]]$scenario.analysis$genetic.target,
+			slope=0, color='red', linetype='dashed') +
 		facet_grid(Context ~ Metric)

### 2016-04-05 article/supporting_information.Rmd
+ fixes to article/si/figures

@@
 	# exports
-	return(((n.plots %/% 5)+1) * 2.5)
+	return(((n.plots %/% 5)+1) * 3)
 }

### 2016-04-05 article/article.Rmd
+ ref fix + redid pca using only pixels over the grid cells to avoid surrounding environmental conditions from biasing the analysis

@@
 We expect that prioritizations generated using surrogate-based targets will secure more genetic variation than prioritizations generated using only amount-based targets with the same number of planning units.
-Second, we aim to determine if the addition of surrogate-based targets to conservation planning exercises can improve the proportion of genetic variation they secure.
-We expect that prioritizations generated using amount- and surrogate-based targets will secure more genetic variation than prioritizations generated using only amount-based targets, and that they will secure a similar amount of genetic variation to prioritizations generated using amount- and genetic based targets.
-Additionally, we expect that prioritizations generated using amount- and surrogate-based targets will share many planning units with prioritizations generated using amount- and genetic-based targets.
+Second, we aim to determine if the addition of surrogate-based targets to conservation planning exercises can result in more effective prioritizations.
+We expect that prioritizations generated using amount- and surrogate-based targets will sufficiently preserve the genetic variation of more species than those generated using only amount-based targets.
 If these surrogates can improve the intra-specific genetic variation captured in a prioritization, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.
@@
 The bioclimatic variables were clipped to the extent of the grid and subject to a principal components analysis (PCA; using \texttt{ArcMap} version 10.2.2; Table S1).
-The first two principal components (PCs) cumulatively explained `r round(pca.DF[[4]][2],1)` % of the variation and were used to generate two new layers.
+The first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components (PCs) cumulatively explained `r round(pca.DF[[4]][surrogate.params.LST[[MODE]]$number.components],1)` % of the variation and were used to generate two new layers.
 We calculated the average value for each PC layer in each grid cell, and used these values to characterize the main themes of climatic variation among the grid cells (Figure S2).
@@
 These species were omitted from subsequent analysis ($n = `r sum(sapply(spp.Mclust.LST, function(x) return(ncol(x$mclust$z)==1)))`$).
-For the remaning species, we ran $\texttt{Bayescan}$ [version 2.1; 1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations; @r456].
+For the remaining species, we ran $\texttt{Bayescan}$ [version 2.1; 1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations; @r456].
 To avoid false-positives, we omitted loci where the global frequency of the minor allele was $\geq 0.05$ (following guidelines in the user manual) and assumed they were neutral.
@@
 
-To address the first aim and determine if the environmental and geographic variables are effective surrogates for adaptive and neutral genetic variation, we generated a collection of single-species prioritizations.
+To address the first aim--to determine if the environmental and geographic variables are effective surrogates for adaptive and neutral genetic variation--we generated a collection of single-species prioritizations.
 For each species, we first generated two sets of prioritizations using increasing environmental and geographic-based targets (separately).
@@
 
-To address the second aim and determine if the use of surrogate-based targets can actually improve prioritizations in real-world conservation planning scenarios, we investigated the effectiveness of surrogate-based targets using three scenarios with increasing levels of complexity.
+To address the second aim--to determine if the use of surrogate-based targets can actually improve prioritizations in real-world conservation planning scenarios--we investigated the effectiveness of surrogate-based targets using three scenarios with increasing levels of complexity.
 To assess the benefits of surrogate-based targets under the simplest of circumstances, our first scenario involved generating single-species prioritizations for each species.
@@
 We then counted the number of species that were sufficiently represented in terms of their adaptive and neutral genetic variation (genetic variation secured $\geq$ `r (rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)` %) for each prioritization.
-We used a global $G$ test to determine the proportion of species that met their genetic target in a given prioritization depended on the combination of targets used to generate the planning units (amount-based targets, amount and surrogate-based targets, amount and genetic-based targets), the planning scenario (single-species, multi-species, multi-species with cost), or the type of genetic variation measured (adaptive or neutral).
-We also conducted a post-hoc analysis involving pairwise $G$ tests (using the \texttt{RVAideMemoire R} package; r@XXX).
+We used a global $G$ test to assess whether the proportion of species that met their genetic target in a given prioritization depended on the combination of targets used to generate the planning units (amount-based targets, amount and surrogate-based targets, amount and genetic-based targets), the planning scenario (single-species, multi-species, multi-species with cost), or the type of genetic variation measured (adaptive or neutral).
+We also conducted a post-hoc analysis involving pairwise $G$ tests (using the \texttt{RVAideMemoire R} package; r@483).
 

### 2016-04-05 article/supporting_information.Rmd
+ ref fix + redid pca using only pixels over the grid cells to avoid surrounding environmental conditions from biasing the analysis

@@
 	digits=2,
-	caption='Summary of principle components analysis (PCA) on bioclimatic variation across the study area. The first two principle components (PCs) were used for subsequent analysis.',
+	caption=paste0('Summary of principle components analysis (PCA) on bioclimatic variation across the study area.
+The first ',as.character(as.english(surrogate.params.LST[[MODE]]$number.components)),' principle components (PCs) were used for subsequent analysis.'),
 	align=c('l', 'c', 'c', 'c')

### 2016-04-05 article/supporting_information.Rmd
+ adjustments to figure sizes

@@
 
-```{r, fig.width=8, fig.height=3.5, fig.cap='Climatic variation.
+```{r, fig.width=15, fig.height=3, fig.cap='Climatic variation.
 Each panel depicts variation based on a different principle component (PC).
@@
 		}),
-		list(ncol=2)
+		list(ncol=4)
 	)
@@
 The first ',as.character(as.english(surrogate.params.LST[[MODE]]$number.components)),' principle components (PCs) were used for subsequent analysis.'),
-	align=c('l', 'c', 'c', 'c')
+	align=c('c', 'c', 'c', 'c')
 )
@@
 	col.names=c('Species', 'Loci Type', 'k', 'NMDS Stress', 'Converged'),
-	align=c('l', 'c', 'c', 'c', 'c')
+	align=c('c', 'c', 'c', 'c', 'c')
 )
@@
 	col.names=c('Species', 'Primer', 'Probability', 'q-value', '$\\alpha$', '$F_{ST}$', 'Type'),
-	align=c('l', 'c', 'c', 'c', 'c', 'c', 'c'),
+	align=c('c', 'c', 'c', 'c', 'c', 'c', 'c'),
 	caption='Loci classifications for each species.

### 2016-04-06 article/article.Rmd
+ discussion edits + add in tests to surrogacy stuff

@@
 ## Discussion
-We aimed to investigate the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation in conservation planning.
-Our study is the first to test this in a multi-species context.
-Overall, we found that environmental and geographic variables are not necessarily effective surrogates for genetic variation in conservation planning--though they can be useful to for identifying suitable amount-based targets.
-Additionally, we found that surrogate-based targets can supplement amount-based targets and improve the proportion of neutral variation secured in a prioritization if the amount-based targets are too low.
-Our results suggest that in most cases--barring extremely unusual spatial patterns of intra-specific genetic variation--conservation planners can secure sufficient adaptive and neutral variation by following conventional guidelines.
+<!-- introduction -->
+We aimed to investigate the effectiveness of broad-scale environmental and geographic variables as surrogates for adaptive and neutral genetic variation in conservation planning.
+Our study is the first to address this hypothesis using moderately realistic planning scenarios.
+Overall, we found that environmental and geographic variables are effective surrogates for a few species when developing prioritizations.
+Furthermore, when developing prioritizations under various planning scenarios, we found that the addition of environmental and geographic targets resulted in prioritizations that sufficiently represented the genetic variation of a greater number of species.
+Our results suggest that while the environmental and geographic variables have little benefit for most species, they can make a large difference for some species.
 
-### Single + multi-species analysis
-* summarize results
+<!-- consequences for conservation planning -->
++ Typically, conservation planning exercises do not consider intra-specific variation when designing prioritizations. 
++ our results show that for most species, a sufficient proportion of their intra-specific variation can be secured using suitable amount-based targets alone.
++ in particular, we found that using suitable amount-based targets in a multi-species context with comprehensive set of target species resulted in a prioritization that adequately represented the adaptive genetic variation for nearly all the target species (panel c; Figure 3).
++ however, for some species, such a prioritization secures only a small proportion of their genetic variation.
++ we found that for some species, prioritizations generated using amount-based targets only secured a small amount of their neutral genetic variation (panels b, d, e; Figure 3).
++ in planning exercises where the decision maker does not have access to genetic data on their target species across the extent of the study area--ie.
+nearly all real world planning scenarios--the decision maker will not know which species are poorly represented in terms of their inta-specific variation.
++ the poorly represented species may be those of particular significant or they may not.
++ the decision maker has to weigh the increased costs of a prioritization that includes environmental and geographic surrogates against the risk that a prioritization that does not include these surrogates fails to represent species of conservation significance.
++ here we found that prioritizations generated using environmental and geographic surrogates have only minor increased costs (compare panels c and d; panels e and f; Figure 3).
++ overall, these results suggest that although surrogates are unnecessary for most species, they can be highly beneficial for a few species, the increased costs associated with using these surrogates is small, so generally using surrogates is a good idea.
 
-* compare + contrast with other papers
+<!-- comparison to previous studies that have investigated genetic studies -->
++ Previous studies on this data set have found correlations between environmental and genetic variation.
++ Yet, for most species, prioritizations generated using environmental surrogates did not secure substantially more genetic variation than randomly generated prioritizations with the same number of planning units.
++ Here we discuss some potential reasons for this disagreement.
++
 
-* implication of results
++ This apparent disagreement can be attributed to multiple reasons.
 
-### Correlation analysis
-* summarize results
++ These results highlight the importance of using prioritizations to assess the effectiveness of potential surrogates for conservation and not simply measuring the strength of correlation between potential surrogates and genetic variation.
 
-* compare + contrast with other papers
-
-* implication of results
-
-### Limitations/Future directions
+<!-- limitations/future directions -->
 We wish to alert our readers to several limitations associated with our analysis.
-First, the size of the planning units we used (23 $\times$ 25 km) is much larger than typically used in conservation planning scenarios (approx.
-1--5 km^2).
+First, the size of the planning units we used (23 $\times$ 25 km) is much larger than typically used in conservation planning scenarios (between 1--10 km^2).
 We used this resolution because the genetic data was collected at this scale.
-Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis because this process would have introduced artificial spatial auto-correlation into the dataset.
+Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis because this process would have  artificially introduced additional spatial auto-correlation into the dataset.
 Second, we used geographic distances as surrogates for connectivity between planning units.

### 2016-04-06 R/analysis/11-statistical-analysis.R

@@ -8,11 +8,26 @@ load('results/.cache/07-surrogacy-prioritizations.rda')
 ### surrogacy analyses
 ## prepare data
 correlation.DF$Surrogate.target <- as.factor(correlation.DF$Surrogate.target)
-correlation.DF$Surrogate.target.Type <- with(correlation.DF, interaction(Surrogate.target, Type))
-correlation.sub.DF <- correlation.DF[rowSums(apply(select(correlation.DF, Surrogate.target, Type), 2, is.na))==0,]
+surrogacy.DF <- correlation.DF %>% ddply(.(Species, Surrogate.target, Type), .fun=function(x) {
+	df1 <- filter(x, Method=='Optimal')
+	df2 <- filter(x, Method=='Random')
+	df3 <- mutate(df2, genetic.held = mean(df1$genetic.held)-genetic.held) %>% select(-Method)
+	return(df3)
+})
 
-## models
-# none
+surrogacy.test.DF <- surrogacy.DF %>% ddply(.(Species, Surrogate.target, Type), .fun=function(x) {
+	return(
+		data.frame(
+			Species=x$Species[[1]], Surrogate.target=x$Surrogate.target[[1]], Type=x$Type[[1]],
+			raw.P=prop.test(x=sum(x$genetic.held>0), n=nrow(x))$p.value
+		)
+	)
+})
+surrogacy.test.DF$P=p.adjust(surrogacy.test$raw.P, 'bonferroni')
+
+surrogacy.counts.DF <- surrogacy.test.DF %>% 
+	group_by(Surrogate.target,Type) %>%
+	summarize(number.species=sum(P<0.05))
 
 ### scenario analyses
 ## prepare data

### 2016-04-07 article/figures.Rmd
+ first commit for incliding alps + carpathians data
Former-commit-id: 5d139b2018f497062d147d39cbcc8faab7af29e9

@@
 options(error=NULL)
+```
 
+```{r global_options, include=FALSE}
+knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
 ```
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=5, fig.cap='The effectiveness of prioritizations at securing genetic variation.
+```{r, results='hide', echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=5, fig.cap='The effectiveness of prioritizations at securing genetic variation.
 The top row of panels (a--b) show the performance of single-species prioritizations generated assuming equal costs among planning costs.

### 2016-04-07 R/analysis/01-load-data.R

@@ -1,62 +1,103 @@
 ## load .rda
 session::restore.session('results/.cache/00-initialization.rda')
 
-## compile spatial grid data
+# compile spatial grid data
 all.spp <- seq_len(general.params.LST[[MODE]]$n.spp)
 
-# load grid cell centroids
-grid.DF <- fread(
-	'data/Data_Meirmans_et_al_IntrabioDiv/ReadMe.txt',
-	data.table=FALSE,
-	skip='cell\tLong\tLat'
-) %>% rename(
-		grid.longitude=Long,
-		grid.latitude=Lat
-) %>% mutate(
-	id=seq_along(grid.latitude)
-)
+### load data
+## load grid cell centroids
+grid.DF <- 'data/Taberlet_etal_EcologyLetters2012/FloristicData/IBDcoordinates.csv' %>% 	
+	fread(,data.table=FALSE) %>%
+	select(GRID_CELL, LON, LAT) %>%
+	rename(cell=GRID_CELL, grid.longitude=LON, grid.latitude=LAT) %>%
+	mutate(id=seq_along(grid.longitude))
 
-# load in aflp data
-spp.aflp.paths <- dir(
-	'data/Data_Meirmans_et_al_IntrabioDiv',
-	'^.*AFLP\\.dat$',
-	full.names=TRUE
-)[all.spp]
+## load species names
+spp.names.DF <- read.table('data/Taberlet_etal_EcologyLetters2012/Taberlet_etal_EcologyLetters2012-ReadMe.txt', skip=118, nrows=27, sep='-', stringsAsFactors=FALSE) %>% 
+	rename(abbreviation=V1, full=V2) %>%
+	mutate(abbreviation=str_trim(abbreviation), full=str_trim(full))
+spp.names.DF$binomial <- laply(spp.names.DF[[2]], function(x) {
+	return(paste(strsplit(x, '\\ ')[[1]][1:2], collapse='_'))
+})
 
-spp.BayeScanData.LST <- llply(
-	spp.aflp.paths,
-	read.BayeScanData
-)
+# limit species for analysis if specified
+spp.names.DF <- spp.names.DF[all.spp,,drop=FALSE]
 
+## load species AFLP data
+spp.PTHS <- 'data/Taberlet_etal_EcologyLetters2012/GeneticData'  %>%
+	dir(pattern='^.*\\.csv$', full.names=TRUE, recursive=TRUE)
+
+spp.DF.LST <- spp.PTHS %>%
+	llply(fread, data.table=FALSE) %>%
[…]

### 2016-04-07 R/analysis/05-genetic-nmds.R

@@ -12,7 +12,7 @@ registerDoParallel(clust)
 
 # run mds
 spp.nmds.LST <- llply(
-	seq_along(spp.BayeScanData.LST),
+	seq_along(spp.StructureData.LST),
 	.fun=function(i) {
 		# subset loci for species
 		setMKLthreads(1)
@@ -22,10 +22,12 @@ spp.nmds.LST <- llply(
 			return(list('adaptive'=NULL, 'neutral'=NULL))
 		} else {
 			# extract loci
-			curr.spp <- bayescanr:::loci.subset(
-				spp.BayeScanData.LST[[i]],
-				which(spp.BayeScanData.LST[[i]]@primers %in% spp.BayeScan.sample.loci.subset.LST[[i]]@data@primers)
-			)
+			curr.spp <- bayescanr:::BayeScanData(spp.StructureData.LST[[i]]@matrix,
+				spp.StructureData.LST[[i]]@loci.names, rep('A', nrow(spp.StructureData.LST[[i]]@matrix)),
+				spp.StructureData.LST[[i]]@sample.names) %>%
+				bayescanr:::loci.subset(
+					which(spp.BayeScanData.LST[[i]]@primers %in% spp.BayeScan.sample.loci.subset.LST[[i]]@data@primers)
+				)
 			# manually classify loci as neutral or adaptive
 			curr.spp.type <- replace(
 				rep('neutral', bayescanr:::n.loci(curr.spp)),

### 2016-04-07 article/figures.Rmd
+ additional fixes
Former-commit-id: cfcc1b7cbf19826af41725f06f349826ed877e39

@@
 # prepare data
-correlation.PDF <- correlation.sub.DF %>%
+correlation.PDF <- correlation.DF %>%
 	group_by(Type, Surrogate.target, Method, Species) %>% 

### 2016-04-11 article/article.Rmd
+ article fixes + projected bioclim pca layer
Former-commit-id: eda1d553c0e5173c94f928353a61c1d1f9fbeb70

@@
 <!--  main text -->
-The act of setting aside areas for preservation spearheads modern conservation efforts [@r440].
+The act of setting aside areas for preservation is one of the hallmarks of modern conservation [@r440].
 Protected area networks buffer species from anthropogenic impacts and provide places for populations of species to persist.
-Intra-specific genetic variation underpins species persistence, and there has been increasing interest in understanding how to design protected area networks in a way that maximizes this variation [@r67; @r3; @r61].
+Intra-specific genetic variation is an important component of species persistence, and there has been increasing interest in understanding how to design protected area networks that preserve this variation [@r67; @r3; @r61].
 Broadly speaking, genetic variation can be classified as either adaptive or neutral [reviewed in @r48].

### 2016-04-12 article/supporting_information.Rmd
+ fixes to stats + fixes to si + final results dataset
Former-commit-id: 6cd2439655495a88edeb6a7710229a4671514244

@@
 }
+
+pretty.paste <- function(x) {
+	if (length(x)==1)
+		return(x)
+	x <- sort(x)
+	x1 <- paste(x[-length(x)], collapse',')
+	x1 <- paste0(x1, ' and ', x[length(x)])
+	return(x1)
+}
 ```
@@
 For a given species, planning units that were found to be inhabited are denoted with bright blue.
-Note that \\textit{',gsub('\\_', '\\ ', missing.species),'} was omitted from analysis.')}
+Note that \\textit{',pretty.paste(gsub('\\_', '\\ ', missing.species)),'} were omitted from analysis.')}
 ## plot map of species distributions
@@
 
-```{r, echo=FALSE, fig.width=4, fig.height=2.5, fig.cap="Species richness. Squares denote planning units. Planning units with a brighter color are inhabited by more species."}
+```{r, echo=FALSE, fig.width=4, fig.height=2.5, fig.cap="Species richness. Squares denote planning units. Planning units with a darker color are inhabited by more species."}
 # calculate species richness
@@
 			strip.background = element_rect(fill='grey20'),
-			strip.text = element_text(color='white', face='italic', size=1)
+			strip.text = element_text(color='white', face='italic', size=3)
 		) +
@@
 }
+
+valid.species <- function(i) {
+	return((i <= n_distinct(spp.samples.DF$species)) & (!unique(spp.samples.DF$species)[i] %in% missing.species))
+}
 ```
@@
 No indiviudals of this species were found inside the gray planning units.')}
-if (1 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[1] %in% missing.species)) plot.spp.nmds(1)
+if (valid.species(1)) plot.spp.nmds(1)
  ```
@@
 See Figure S6 caption for conventions.')}
-if (2 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[2] %in% missing.species)) plot.spp.nmds(2)
+if (valid.species(3)) plot.spp.nmds(2)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (3 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[3] %in% missing.species)) plot.spp.nmds(3)
+if (valid.species(3)) plot.spp.nmds(3)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (4 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[4] %in% missing.species)) plot.spp.nmds(4)
+if (valid.species(4)) plot.spp.nmds(4)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (5 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[5] %in% missing.species)) plot.spp.nmds(5)
+if (valid.species(5)) plot.spp.nmds(5)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (6 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[6] %in% missing.species)) plot.spp.nmds(6)
+if (valid.species(6)) plot.spp.nmds(6)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (7 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[7] %in% missing.species)) plot.spp.nmds(7)
+if (valid.species(7)) plot.spp.nmds(7)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (8 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[8] %in% missing.species)) plot.spp.nmds(8)
+if (valid.species(8)) plot.spp.nmds(8)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (9 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[9] %in% missing.species)) plot.spp.nmds(9)
+if (valid.species(9)) plot.spp.nmds(9)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (10 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[10] %in% missing.species)) plot.spp.nmds(10)
+if (valid.species(10)) plot.spp.nmds(10)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (11 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[11] %in% missing.species)) plot.spp.nmds(11)
+if (valid.species(11)) plot.spp.nmds(11)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (12 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[12] %in% missing.species)) plot.spp.nmds(12)
+if (valid.species(12)) plot.spp.nmds(12)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (13 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[13] %in% missing.species)) plot.spp.nmds(13)
+if (valid.species(13)) plot.spp.nmds(13)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (14 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[14] %in% missing.species)) plot.spp.nmds(14)
+if (valid.species(14)) plot.spp.nmds(14)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (15 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[15] %in% missing.species)) plot.spp.nmds(15)
+if (valid.species(15)) plot.spp.nmds(15)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (16 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[16] %in% missing.species)) plot.spp.nmds(16)
+if (valid.species(16)) plot.spp.nmds(16)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (17 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[17] %in% missing.species)) plot.spp.nmds(17)
+if (valid.species(17)) plot.spp.nmds(17)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (18 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[18] %in% missing.species)) plot.spp.nmds(18)
+if (valid.species(18)) plot.spp.nmds(18)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (19 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[19] %in% missing.species)) plot.spp.nmds(19)
+if (valid.species(19)) plot.spp.nmds(19)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (20 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[20] %in% missing.species)) plot.spp.nmds(20)
+if (valid.species(20)) plot.spp.nmds(20)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (21 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[21] %in% missing.species)) plot.spp.nmds(21)
+if (valid.species(21)) plot.spp.nmds(21)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (22 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[22] %in% missing.species)) plot.spp.nmds(22)
+if (valid.species(22)) plot.spp.nmds(22)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (23 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[23] %in% missing.species)) plot.spp.nmds(23)
+if (valid.species(23)) plot.spp.nmds(23)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (24 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[24] %in% missing.species)) plot.spp.nmds(24)
+if (valid.species(24)) plot.spp.nmds(24)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (25 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[25] %in% missing.species)) plot.spp.nmds(25)
+if (valid.species(25)) plot.spp.nmds(25)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (26 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[26] %in% missing.species)) plot.spp.nmds(26)
+if (valid.species(26)) plot.spp.nmds(26)
 ```
@@
 See Figure S6 caption for conventions.')}
-if (27 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[27] %in% missing.species)) plot.spp.nmds(27)
+if (valid.species(27)) plot.spp.nmds(27)
 ```

### 2016-04-12 R/analysis/11-statistical-analysis.R

@@ -10,26 +10,26 @@ load('results/.cache/07-surrogacy-prioritizations.rda')
 correlation.DF$Surrogate.target <- as.factor(correlation.DF$Surrogate.target)
 correlation.DF$Type.Surrogate.target.Method=with(correlation.DF, interaction(Type, Surrogate.target, Method))
 
-surrogacy.DF <- correlation.DF %>% ddply(.(Species, Surrogate.target, Type), .fun=function(x) {
-	df1 <- filter(x, Method=='Optimal')
-	df2 <- filter(x, Method=='Random')
-	df3 <- mutate(df2, genetic.held = mean(df1$genetic.held)-genetic.held) %>% select(-Method)
-	return(df3)
-})
-
-surrogacy.test.DF <- surrogacy.DF %>% ddply(.(Species, Surrogate.target, Type), .fun=function(x) {
-	return(
-		data.frame(
-			Species=x$Species[[1]], Surrogate.target=x$Surrogate.target[[1]], Type=x$Type[[1]],
-			raw.P=prop.test(x=sum(x$genetic.held>0), n=nrow(x))$p.value
-		)
-	)
-})
-surrogacy.test.DF$P=p.adjust(surrogacy.test.DF$raw.P, 'bonferroni')
-
-surrogacy.counts.DF <- surrogacy.test.DF %>% 
-	group_by(Surrogate.target,Type) %>%
-	summarize(number.species=sum(P<0.05))
+# surrogacy.DF <- correlation.DF %>% ddply(.(Species, Surrogate.target, Type), .fun=function(x) {
+# 	df1 <- filter(x, Method=='Optimal')
+# 	df2 <- filter(x, Method=='Random')
+# 	df3 <- mutate(df2, genetic.held = mean(df1$genetic.held)-genetic.held) %>% select(-Method)
+# 	return(df3)
+# })
+# 
+# surrogacy.test.DF <- surrogacy.DF %>% ddply(.(Species, Surrogate.target, Type), .fun=function(x) {
+# 	return(
+# 		data.frame(
+# 			Species=x$Species[[1]], Surrogate.target=x$Surrogate.target[[1]], Type=x$Type[[1]],
+# 			raw.P=prop.test(x=sum(x$genetic.held>0), n=nrow(x))$p.value
+# 		)
+# 	)
+# })
+# surrogacy.test.DF$P=p.adjust(surrogacy.test.DF$raw.P, 'bonferroni')
+# 
[…]

### 2016-04-13 code/rmarkdown/supporting_information.Rmd
+ restructure folders + include pcadapt for outlier loci analysis

--- 2016-04-05 article/supporting_information.Rmd
@@
 ```{r, echo=FALSE}
-knitr::kable(
-	format.dataframe(
-		ldply(
+loci_classification.DF <- ldply(
 	seq_along(unique(spp.samples.DF$species)), 
@@
 				Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
-						Primer=spp.BayeScan.sample.loci.subset.LST[[i]]@data@primers,
-						Probability=spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[2]],
-						qval=spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[4]],
-						alpha=spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[5]],
-						fst=spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[6]],
-						Type=spp.BayeScan.sample.loci.subset.LST[[i]]@results@summary[[7]]
+				Primer=spp.BayeScan.LST[[i]]$bayescan@data@primers,
+
+				Bayescan_Probability=spp.BayeScan.LST[[i]]$bayescan@results@summary[[2]],
+				Bayescan_qval=spp.BayeScan.LST[[i]]$bayescan@results@summary[[4]],
+				Bayescan_alpha=spp.BayeScan.LST[[i]]$bayescan@results@summary[[5]],
+				Bayescan_fst=spp.BayeScan.LST[[i]]$bayescan@results@summary[[6]],
+				Bayescan_Type=spp.BayeScan.LST[[i]]$bayescan@results@summary[[7]],
+
+				PCadapt_K=spp.pcadapt.LST[[i]]$K,
+				PCadapt_chi2_statistic=spp.pcadapt.LST[[i]]$chi2.stat,
+				PCadapt_pvalue=spp.pcadapt.LST[[i]]$pvalues,
+				PCadapt_Type=replace(rep('neutral, length(spp.pcadapt.LST[[i]]$pvalues)), spp.pcadapt.LST[[i]]$adaptive.loci, 'adaptive')
+
+				Overall_Type=replace(rep('neutral, length(spp.pcadapt.LST[[i]]$pvalues)), spp.OutlierDetectionResults.LST[[i]], 'adaptive')
 			)
@@
 	}
-		),
-		omit='Type'
-	),
-	digits=2,
-	col.names=c('Species', 'Primer', 'Probability', 'q-value', '$\\alpha$', '$F_{ST}$', 'Type'),
-	align=c('c', 'c', 'c', 'c', 'c', 'c', 'c'),
-	caption='Loci classifications for each species.
-Each row denotes a sampled loci for a given species, and the evidence for it being under selection based on outputs from the program \\texttt{Bayescan}.
-Note that loci in species with a single population were assumed to be neutral.
-Loci for these species do not appear in this table.'
 )
+
+# save to file
+write.table(loci_classification.DF, 'article/Table_S2', sep=',', row.names=FALSE)
 ```

### 2016-04-13 R/analysis/04-outlier-locus-analysis.R

@@ -1,49 +1,0 @@
-## load .rda
-session::restore.session('results/.cache/03-population-clustering-analysis.rda')
-
-## load parameters
-bayescan.params.LST <- parseTOML('parameters/bayescan.toml')
-nmds.params.LST <- parseTOML('parameters/nmds.toml')
-
-# subset out loci with polymorphisms that have really low or really high frequency
-spp.BayeScanData.sample.loci.subset.LST <- llply(
-	spp.BayeScanData.sample.subset.LST,
-	function(x) {
-		# skip if null
-		if (is.null(x))
-			return(NULL)
-		# else remove abundant or rare loci
-		freqs <- colMeans(x@matrix, na.rm=TRUE)
-		valid.loci <- which(freqs <= 1-(bayescan.params.LST[[MODE]]$freq) | freqs >= bayescan.params.LST[[MODE]]$freq)
-		return(bayescanr:::loci.subset(x, valid.loci))
-	}
-)
-
-# run BayeScan over subsetted objects
-dir.create('results/.cache/bayescan')
-spp.BayeScan.sample.loci.subset.LST <- llply(
-	seq_along(spp.BayeScanData.sample.loci.subset.LST),
-	.fun=function(i) {
-		# skip if null
-		if (is.null(spp.BayeScanData.sample.loci.subset.LST[[i]]))
-			return(NULL)
-		# else run bayescan
-		curr.dir <- paste0('results/.cache/bayescan/',unique(spp.samples.DF$species)[i])
-		dir.create(curr.dir)
-		run.BayeScan(
-			spp.BayeScanData.sample.loci.subset.LST[[i]],
-			fdr=bayescan.params.LST[[MODE]]$fdr,
-			threads=general.params.LST[[MODE]]$threads,
-			n=bayescan.params.LST[[MODE]]$n,
-			thin=bayescan.params.LST[[MODE]]$thin,
-			nbp=bayescan.params.LST[[MODE]]$nbp,
-			pilot=bayescan.params.LST[[MODE]]$pilot,
-			burn=bayescan.params.LST[[MODE]]$burn,
-			dir=curr.dir,
-			clean=FALSE
-		)
-	}
-)
-
-## save .rda
-save.session('results/.cache/04-outlier-locus-analysis.rda', compress='xz')

### 2016-04-13 code/R/analysis/04-outlier-locus-analysis.R

@@ -1,0 +1,108 @@
+## load .rda
+session::restore.session('data/results/03-population-clustering-analysis.rda')
+
+## load parameters
+bayescan.params.LST <- parseTOML('code/parameters/bayescan.toml')
+pcadapt.params.LST <- parseTOML('code/parameters/pcadapt.toml')
+nmds.params.LST <- parseTOML('code/parameters/nmds.toml')
+
+# subset out loci with polymorphisms that have really low or really high frequency
+spp.OutlierDetectionData.LST <- llply(
+	spp.OutlierDetectionData.LST,
+	function(x) {
+		# skip if null
+		if (is.null(x))
+			return(NULL)
+		# else remove abundant or rare loci
+		freqs <- colMeans(x@matrix, na.rm=TRUE)
+		valid.loci <- which(freqs <= 1-(bayescan.params.LST[[MODE]]$freq) | freqs >= bayescan.params.LST[[MODE]]$freq)
+		return(bayescanr:::loci.subset(x, valid.loci))
+	}
+)
+
+# run BayeScan over subsetted objects
+dir.create('data/results/bayescan')
+spp.BayeScan.LST <- llply(
+	seq_along(spp.OutlierDetectionData.LST),
+	.fun=function(i) {
+		# skip if null
+		if (is.null(spp.OutlierDetectionData.LST[[i]]))
+			return(NULL)
+		# else run bayescan
+		curr.dir <- paste0('data/results/bayescan/',unique(spp.samples.DF$species)[i])
+		dir.create(curr.dir)
+		curr.spp <- run.BayeScan(
+			spp.OutlierDetectionData.LST[[i]],
+			fdr=bayescan.params.LST[[MODE]]$fdr,
+			threads=general.params.LST[[MODE]]$threads,
+			n=bayescan.params.LST[[MODE]]$n,
+			thin=bayescan.params.LST[[MODE]]$thin,
+			nbp=bayescan.params.LST[[MODE]]$nbp,
+			pilot=bayescan.params.LST[[MODE]]$pilot,
+			burn=bayescan.params.LST[[MODE]]$burn,
+			dir=curr.dir,
+			clean=FALSE
+		)
+		return(
+			list(
+				bayescan=curr.spp,
+				adaptive.loci=which(curr.spp@results@summary$type=='adaptive')
+			)
+		)
+	}
+)
+
+# run pcadapt over subsetted objects
+spp.pcadapt.LST <- llply(
+	seq_along(spp.OutlierDetectionData.LST),
[…]

### 2016-04-13 code/R/analysis/07-surrogacy-prioritizations.R

@@ -1,9 +1,9 @@
 ## load .rda
-session::restore.session('results/.cache/06-format-data-for-prioritisations.rda')
+session::restore.session('data/results/06-format-data-for-prioritisations.rda')
 
 ## load parameters
-rapr.params.LST <- parseTOML('parameters/rapr.toml')
-gurobi.params.LST <- parseTOML('parameters/gurobi.toml')
+rapr.params.LST <- parseTOML('code/parameters/rapr.toml')
+gurobi.params.LST <- parseTOML('code/parameters/gurobi.toml')
 
 ## correlation analysis
 # generate prioritistions
@@ -219,5 +219,5 @@ geo.correlation.DF <- ldply(
 correlation.DF <- rbind(env.correlation.DF, geo.correlation.DF)
 
 ## save .rda
-save.session('results/.cache/07-surrogacy-prioritizations.rda', compress='xz')
+save.session('data/results/07-surrogacy-prioritizations.rda', compress='xz')
 

### 2016-04-13 code/R/analysis/09-multi-species-prioritisations-no-cost.R

@@ -1,13 +1,13 @@
 ## load .rda
-session::restore.session('results/.cache/06-format-data-for-prioritisations.rda')
+session::restore.session('data/results/06-format-data-for-prioritisations.rda')
 
 ## load parameters
-rapr.params.LST <- parseTOML('parameters/rapr.toml')
-gurobi.params.LST <- parseTOML('parameters/gurobi.toml')
+rapr.params.LST <- parseTOML('code/parameters/rapr.toml')
+gurobi.params.LST <- parseTOML('code/parameters/gurobi.toml')
 
 ## multispecies analysis
 # make prioritisations
-multi.spp.prioritisations.with.cost <- llply(
+multi.spp.prioritisations <- llply(
 	list(
 		list(rapr.params.LST[[MODE]]$scenario.analysis$amount.target,NA,NA,
 			rapr.params.LST[[MODE]]$scenario.analysis$other.replicates),
@@ -20,7 +20,7 @@ multi.spp.prioritisations.with.cost <- llply(
 	), 
 	function(y) {
 		species.prioritisation(
-			x=ru_with_cost,
+			x=ru,
 			amount.targets=y[[1]],
 			env.surrogate.targets=y[[2]],
 			geo.surrogate.targets=y[[2]],
@@ -34,11 +34,11 @@ multi.spp.prioritisations.with.cost <- llply(
 )
 
 # generate results table
-multi.spp.with.cost.DF <- ldply(
-	seq_along(multi.spp.prioritisations.with.cost),
+multi.spp.DF <- ldply(
+	seq_along(multi.spp.prioritisations),
 	.fun=function(i) {
 		mutate(
-			extractResults(multi.spp.prioritisations.with.cost[[i]]),
+			extractResults(multi.spp.prioritisations[[i]]),
 			Prioritisation=c('Amount','Surrogate','Genetic')[i]
 		)
 	}
@@ -49,4 +49,4 @@ multi.spp.with.cost.DF <- ldply(
 
 
 ## save .rda
-save.session('results/.cache/10-multi-species-prioritisations-with-cost.rda', compress='xz')
+save.session('data/results/09-multi-species-prioritisations-no-cost.rda', compress='xz')

### 2016-04-13 code/R/analysis/01-load-data.R
+ updates to directory structure + use structure for analysis

@@ -1,8 +1,15 @@
 ## load .rda
-session::restore.session('data/results/00-initialization.rda')
+session::restore.session('data/intermediate/00-initialization.rda')
 
 ## compile spatial grid data
-all.spp <- seq_len(general.params.LST[[MODE]]$n.spp)
+# load in species population numbers
+spp.populations.DF <- fread('data/raw/Data_Meirmans_et_al_IntrabioDiv/NumberPopulations.csv', data.table=FALSE)
+
+# subset to species that have >1 population
+all.spp <- which(spp.populations.DF[[2]] > 1)
+all.spp <- all.spp[seq_len(min(general.params.LST[[MODE]]$n.spp, length(all.spp)))]
+missing.species <- as.character(spp.populations.DF[[1]][spp.populations.DF[[2]]==1])
+spp.populations.DF <- spp.populations.DF[all.spp,]
 
 # load grid cell centroids
 grid.DF <- fread(
@@ -78,4 +85,4 @@ for (i in unique(spp.samples.DF$species))
 
 
 ## save .rda
-save.session('data/results/01-load-data.rda', compress='xz')
+save.session('data/intermediate/01-load-data.rda', compress='xz')

### 2016-04-13 code/R/analysis/05-genetic-nmds.R

@@ -1,5 +1,5 @@
 ## load .rda
-session::restore.session('data/results/04-outlier-locus-analysis.rda')
+session::restore.session('data/intermediate/04-outlier-locus-analysis.rda')
 
 ## load parameters
 nmds.params.LST <- parseTOML('code/parameters/nmds.toml')
@@ -44,10 +44,10 @@ spp.nmds.LST <- llply(
 				curr.nmds <- bayescanr::nmds(
 					bayescanr:::loci.subset(curr.spp, curr.spp.type==j),
 					metric='gower',
-					max.stress=nmds.params.LST[[MODE]]$genetic.space$max.stress,
-					min.k=nmds.params.LST[[MODE]]$genetic.space$min.k,
-					max.k=nmds.params.LST[[MODE]]$genetic.space$max.k,
-					trymax=nmds.params.LST[[MODE]]$genetic.space$trymax
+					max.stress=nmds.params.LST[[MODE]]$max.stress,
+					min.k=nmds.params.LST[[MODE]]$min.k,
+					max.k=nmds.params.LST[[MODE]]$max.k,
+					trymax=nmds.params.LST[[MODE]]$trymax
 				)
 			)
 		}), c('adaptive','neutral')))
@@ -97,4 +97,4 @@ for (i in seq_along(unique(spp.samples.DF$species))) {
 grid.PLY@data <- grid.DF
 
 ## save .rda
-save.session('data/results/05-genetic-nmds.rda', compress='xz')
+save.session('data/intermediate/05-genetic-nmds.rda', compress='xz')

### 2016-04-13 code/R/analysis/12-exports.R

@@ -1,7 +1,7 @@
 ## load .rda
-session::restore.session('data/results/11-statistical-analysis.rda')
+session::restore.session('data/intermediate/11-statistical-analysis.rda')
 
 #### save results
-session::save.session('results/results.rda', compress='xz')
+session::save.session('data/final/results.rda', compress='xz')
 
 

### 2016-04-14 code/rmarkdown/article.Rmd
+ updates to directory structure + use structure for analysis
+ article fixes + extra refs

--- 2016-04-06 article/article.Rmd
@@
 setwd('..')
+setwd('..')
 
 # load .rda file
-try(session::restore.session('results/results.rda'))
+try(session::restore.session('data/results/results.rda'))
 
@@
 Here we investigate the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation in the context of conservation planning.
-Conventionally, prioritizations are generated by setting targets on the total amount of habitat that needs to be preserved for each species (amount-based targets), and finding the combination of planning units that fulfills all the targets for minimal cost [eg.
-using \texttt{Marxan}; @r22].
-First, we aim to verify that environmental and geographic variables are effective surrogates for adaptive and neutral variation (respectively).
+First, we aim to verify if environmental and geographic variables correlate with broad-scale patterns in adaptive and neutral variation (respectively).
+Although these variables may correlate with genetic variation, they may not be effective surrogates if the level of correlation is not high enough or due to spatial auto-correlative effects.
+Second, we aim to verify that environmental and geographic variables are effective surrogates for adaptive and neutral variation (respectively).
 We expect that prioritizations generated using surrogate-based targets will secure more genetic variation than prioritizations generated using only amount-based targets with the same number of planning units.
-Second, we aim to determine if the addition of surrogate-based targets to conservation planning exercises can result in more effective prioritizations.
+Third, we aim to determine if the addition of surrogate-based targets to conservation planning exercises can result in more effective prioritizations.
 We expect that prioritizations generated using amount- and surrogate-based targets will sufficiently preserve the genetic variation of more species than those generated using only amount-based targets.
@@
 We used species distribution and genomic data collected by the IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for further explanation of data collection methods].
-This data set was particularly well suited for our study because it contains genetic data for a diverse range of species that were collected using the same sampling regime.
-Additionally, the study area covers a broad geographic domain that encompasses attitudinal gradients.
-As part of this project, data were collected for 27 alpine plant species using a 20' longitude by 21' latitude grid (approx.
+This data set contains genetic data for a diverse range of species that were collected using a standardized sampling regime.
+Additionally, the data set covers a large geographic domain that spans across a range of attitudinal gradients.
+As a consequence, there is large environmental variation among the grid cell cells which is desirable to assess effectiveness of environmental variables as surrogates for adaptive genetic variation.
+Thus this data set was particularly well suited for our study.
+
+The IntroBioDiv project data set contains data for 27 alpine plant species that was collected using a 20' longitude by 21' latitude grid (approx.
 22.3 km $\times$ 25 km; Figure S1).
 Project members visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals.
-They genotyped samples using amplified fragment length polymorphisms [AFLP; @r453], and constructed matrices denoting the presence/absence of polymorphisms at loci for each species.
-Thus the data set contains information describing the genomic variation of individuals for each species present in every second grid cell.
-We used these grid cells as planning units to develop prioritizations.
-To reduce computational effort, we used only grid cells that contained at least a single individual ($n = `r nrow(grid.DF)`$).
-The majority of our analyses were performed using the \texttt{R} software package [version `r paste0(version$major, '.', version$minor)`; @r192].
+They genotyped samples using amplified fragment length polymorphisms [AFLP; @r453], and constructed matrices denoting the presence/absence of polymorphisms at loci for each species (mean `r mean(sapply(spp.StructureData.LST, n.loci))` \pm `r sd(sapply(spp.StructureData.LST, n.loci))` SD markers genotyped per species; see Table 1 in r@478 for the number markers for each species).
+Although AFLPs are not ideal as genomic representatives (cf.
+single nucleotide polymorphisms; SNPs), no multi-species genomic data set with similar properties currently exists.
+Thus the data set contains information describing the genomic properties of individuals for each species present in every second grid cell.
+
+This data set has previously been used to address a range of different questions.
+Many studies have used this data set to explore patterns of adaptive [@r484; @r475; @r455; @r451; r@485] and neutral genetic variation [@r451; @r478; @r486].
+However, to our knowledge, only a single study has used this data set to explore the effectiveness of a surrogate in the context of conservation planning.
+Taberlet et al.
+[-@r486] investigated whether prioritizations generated using the spatial distribution of a comprehensive set species (> 300) could adequately secure the intra-specific genetic variation for 27 species.
+They found that the prioritization based on species distributions was substantially different to that based on genomic data.
+Their results suggest that prioritizations based on species distribution data alone will be unable to secure intra-specific genetic variation.
 
@@
 We examined the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation (respectively).
+We used the grid cells used to sample genetic data as planning units to develop prioritizations.
+We used only planning units that contained at least a single individual ($n = `r nrow(grid.DF)`$) to reduce computational burden.
+To examine the potential effects of cost, we calculated the total human population density inside each planning unit [obtained at 1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472] and used this to represent acquisition cost.
+
 To describe the geographic location of each grid cell, we projected the grid into an equi-distant coordinate system ($\texttt{ESRI:102031}$), calculated their centroids, and extracted their two-dimensional coordinates.
-To describe the environmental characteristics of each grid cell, we obtained 19 bioclimatic variables [approx.
+Thus we constructed two-dimensional geographic space-wherein each planning unit occupies a single point--as a potential surrogate for neutral genetic variation.
+
+To describe the environmental characteristics of each grid cell, we obtained data for 19 bioclimatic variables across the study area [approx.
 1 km resolution; @r33].
-These layers and the grid cells were projected into an equal-area coordinate system ($\texttt{ESRI:102014}$).
-The bioclimatic variables were clipped to the extent of the grid and subject to a principal components analysis (PCA; using \texttt{ArcMap} version 10.2.2; Table S1).
-The first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components (PCs) cumulatively explained `r round(pca.DF[[4]][surrogate.params.LST[[MODE]]$number.components],1)` % of the variation and were used to generate two new layers.
-We calculated the average value for each PC layer in each grid cell, and used these values to characterize the main themes of climatic variation among the grid cells (Figure S2).
-The cost of purchasing land for preservation often exerts a strong influence on which planning units are recommended for prioritization [reviewed in @r482].
-To examine the potential effects of cost, we calculated the total human population density inside each planning unit [obtained at 1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472] and used this to represent acquisition cost.
+We used this data set in particular because it is commonly used in conservation planning to characterize environmental variation in conservation planning.
+These layers and the planning units were projected into an equal-area coordinate system ($\texttt{ESRI:102014}$).
+The 19 bioclimatic variables were clipped to the grid and subject to a principal components analysis (PCA; using \texttt{ArcMap} version 10.2.2; Table S1) to extract the main themes of environmental variation among the grid cells.
+The first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components (PCs) cumulatively explained `r round(pca.DF[[4]][surrogate.params.LST[[MODE]]$number.components],1)` % of the variation and were used to generate `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` new layers.
+We calculated the average value for each PC layer in each planning unit, and used these values to characterize climatic variation among the planning units (Figure S2).
+Thus we constructed a three-dimensional environmental space--wherein each planning unit occupies a single point--as a potential surrogate for adaptive variation.
 
@@
 To investigate the effectiveness of the surrogates for adaptive and neutral genetic variation, we first needed to identify which of the sampled loci--if any--were under selection.
-We used outlier detection methods [compared in @r454; reviewed in @r480] to achieve this.
-The basic premise underpinning these analyses is that neutral loci are expected to exhibit a specific level of variation among populations ($F_{st}$), and loci that deviate from this expectation should be under selection.
+We used two outlier detection methods to achieve this.
+The basic premise underpinning outlier detection methods is that neutral loci are expected to exhibit a specific level of variation among populations, and loci that deviate from this expectation should be under selection [reviewed in @r480].
 The advantage of these methods--in contrast with environmental association analyses [eg.
 @r475]--is that they do not use environmental data to identify loci under selection, which would have biased our analysis.
-In an earlier study, Bothwell _et al._ [-@r455] applied such outlier detection methods to one of the species in this data set (_Gentiana nivalis_).
-Here, we adopted a similar methodology and applied it to each of the 27 species in the data set.
+We omitted the species \textit{`r missing.species`} from all subsequent analysis because it was only associated with one genetic lineage and thus not compatible with one of the outlier detection methods.
+Loci identified as outliers in both analyses were treated as adaptive, and the remaining loci were treated as neutral.
 
-First, we grouped individuals into genetic lineages for each species [using a methodology similar to that implemented in \texttt{prabclust R} package; @r479].
-The purpose of this step was to avoid falsely classifying loci as adaptive.
-For each species, we ordinated the genomic data, and then fit models to the ordinated data to group individuals into populations.
-Specifically, we used non-metric multi-dimensional scaling [NMDS; implemented in the $\texttt{vegan R}$ package; @r458] using Gower distances [@r464; using the $\texttt{cluster R}$ package to accommodate sparsity; @r457] to transform the binary loci data into continuous variables (Table S2).
-To ensure that the ordinations described a sufficient amount of the genomic variation, we ran successive scaling analyses with increasing dimensionality until this was achieved (maximum stress value $\leq$ `r nmds.params.LST[[MODE]]$population.clustering$max.stress`; `r nmds.params.LST[[MODE]]$population.clustering$trymax` random starts for each analysis).
-We then fit a suite of Gaussian mixture models to the ordinations to estimate the probability that individuals belonged to different populations [assuming different covariance structures implemented in the \texttt{mclust R} package; @r477; @r476].
-For each parametrization, we fit a set of models assuming varying numbers of total populations (\textit{G}= 1--`r max(mclust.params.LST[[MODE]]$G)`).
-We identified the most parsimonious model for each species (lowest Bayesian Information Criterion $BIC$ statistic; Figures S4), and used them to group individuals to populations (Figure S5).
+First, we used multinominal-Dirichlet models implemented in \texttt{BayeScan} [version 2.1; @r456] to identify loci under selection.
+We adopted a similar methodology to Bothwell _et al._ [-@r455] and applied it to each of the species in the data set.
+To avoid falsely classifying loci as adaptive, we grouped individuals into genetic lineages for each species (Figures SXXX--SXXX).
+We used the number of genetic lineages in each species previously identified by Alvarez _et al._ [-@r478].
+For each species, we used \texttt{Structure} (version 2.3.1; using same parameters as @r478 per run with `r structure.params.LST[[MODE]]$numruns` runs per species) and \texttt{ClumPP} (version 1.1.2; @r459) to group individuals into populations.
+We then ran $\texttt{Bayescan}$ for each species using these groupings [1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations; @r456].
+To avoid false-positives, we omitted loci where the global frequency of the minor allele was $\geq 0.05$ (following guidelines in the user manual) and assumed they were neutral.
+Additionally, we omitted individuals from this analysis when their population membership was uncertain (maximum membership probability $< `r structure.params.LST[[MODE]]$probability.threshold`$).
+We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate [FDR $\leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].
 
-After grouping individuals into populations, we identified the loci under selection (Table S3).
-Since outlier detection methods compare allele frequencies between populations, we were unable to examine species that contained a single population.
-These species were omitted from subsequent analysis ($n = `r sum(sapply(spp.Mclust.LST, function(x) return(ncol(x$mclust$z)==1)))`$).
-For the remaining species, we ran $\texttt{Bayescan}$ [version 2.1; 1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations; @r456].
-To avoid false-positives, we omitted loci where the global frequency of the minor allele was $\geq 0.05$ (following guidelines in the user manual) and assumed they were neutral.
-Additionally, we omitted individuals from this analysis when their population membership was uncertain (maximum membership probability $< `r mclust.params.LST[[MODE]]$probability.threshold`$).
-We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate to ensure convergence [FDR $\leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].
+Second, we used PCAs to identify outlier loci using methods implemented in the \texttt{pcadapt R} package [version 2.2; @r487].
+For each of the 26 species, we ran a PCA and extracted the minimum number of components needed to secure `r pcadapt.params.LST[[MODE]]$min.variance.explained` % of the variation in loci.
+We then computed Mahalanobis distances [@r398] to detect outlier loci ($P < `r pcadapt.params.LST[[MODE]]$alpha.level`$).
+To enable comparisons between the two outlier detection analyses, we used the same individuals in this analysis as used in the $BayeScan$ analysis.
 
 After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species (Figures S6--S32).
-For each species, we partitioned adaptive and neutral loci into separate data sets, and reran the successive nonmetric multidimensional scaling analyses (using parameters described above but with maximum stress values $\leq$ `r nmds.params.LST[[MODE]]$genetic.space$max.stress`) to derive continuous variables that described the main gradients of adaptive and neutral genetic variation separately (Table S2).
+For each species, we discarded the population groupings, and partitioned adaptive and neutral loci into separate data sets.
+We used non-metric multi-dimensional scaling [NMDS; implemented in the \texttt{vegan R} pacakge; @r458] using Gower distances [@r464; using the \texttt{cluster R} package to accommodate sparsity; @r457] to derive continuous variables that described the main gradients of adaptive and neutral genetic variation for each species (Table S2).
+To ensure that the ordinations described a sufficient amount of the genomic variation, we ran successive scaling analyses with increasing dimensionality until this was achieved (maximum stress value $\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for each analysis).
 Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the individuals in each grid cell to express the typical genomic characteristics of individuals in the cells.
 
-### Prioritizations
+The previous analyses resulted in an adaptive (if detected) and neutral genetic space for each species.
+For a given species, each planning unit occupied by the species was associated with a multi-dimensional point in the species' neutral genetic space.
+Planning units that were closer together in this space are occupied by individuals with similar neutral genetic loci.
+Similarly, if the species was associated with adaptive loci, each planning unit occupied by the species was associated with a multi-dimensional point in the species' adaptive genetic space.
+By spreading out conservation effort across a given genetic space for a given species, prioritizations can secure more genetic variation [see @r465 for discussion on attribute spaces].
+
+### Prioritizations and statistical analyses
 We used the unreliable representative and adequate prioritization (URAP) formulation of the reserve selection problem [@r465].
-This formulation can identify the minimum number of planning units required to preserve both a proportion of the species' range (using amount-based targets) and a proportion of the species' intra-specific variation [using space-based targets and attribute spaces; _sensu_ @r465].
+This formulation can identify the minimum number of planning units required to preserve both a proportion of the species' range (using amount-based targets) and a proportion of variation found across the species' range [using space-based targets and attribute spaces; _sensu_ @r465].
 We treated the environmental and geographic surrogate variables as separate attribute spaces.
@@
 
-To address the first aim--to determine if the environmental and geographic variables are effective surrogates for adaptive and neutral genetic variation--we generated a collection of single-species prioritizations.
+To address the first aim--to determine if the environmental and geographic variables correlate with adaptive and neutral genetic variation--we fit maximum likelihood population-effects (MLPE) models [implemented in the \texttt{ResistanceGA R} package; @r488].
+We computed differences between the grid cells in terms of their environmental conditions and  geographic position.
+Additionally, for each species, we computed the differences between the grid cells in terms of the adaptive and neutral genetic characteristics of individuals that inhabited them (if detected).
+All differences were calculated using Euclidean distances.
+We then fit a set of MLPE models to the differences in the adaptive genetic characteristics among the grid cells and the differences in the environmental conditions for each species associated with adaptive genetic variation.
+We also fit a set of similar models using the differences in neutral genetic characteristics and the geographic distances for each species.
+For each model we fit an intercept-only null model, and tested surrogate variables whether they explained substantially more variation using $\chi^2$.
+We then applied Bonferroni corrections to each set of models (separately).
+Note that this analysis does not take into account spatial auto-correlation.
+The subsequent analyses explicitly control for this effect.
+
+To address the second aim--to determine if the environmental and geographic variables are effect surrogates for adaptive and neutral genetic variation--we generated a collection of single-species prioritizations.
 For each species, we first generated two sets of prioritizations using increasing environmental and geographic-based targets (separately).
@@
 
-To address the second aim--to determine if the use of surrogate-based targets can actually improve prioritizations in real-world conservation planning scenarios--we investigated the effectiveness of surrogate-based targets using three scenarios with increasing levels of complexity.
+To address the third aim--to determine if the use of surrogate-based targets can actually improve prioritizations in real-world conservation planning scenarios--we investigated the effectiveness of surrogate-based targets using three scenarios with increasing levels of complexity.
 To assess the benefits of surrogate-based targets under the simplest of circumstances, our first scenario involved generating single-species prioritizations for each species.
@@
 Based on the previous analysis, we set surrogate-based targets as `r (rapr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100)` % and genetic-based targets as `r (rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)` % to ensure that most of the variation was secured for each species.
-To accomodate variation in the high number of optimal solutions for the single-species amount-based prioritizations, we generated `r (rapr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates)` replicate prioritizations.
+To accommodate variation in the high number of optimal solutions for the single-species amount-based prioritizations, we generated `r (rapr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates)` replicate prioritizations.
 
@@
 We used a global $G$ test to assess whether the proportion of species that met their genetic target in a given prioritization depended on the combination of targets used to generate the planning units (amount-based targets, amount and surrogate-based targets, amount and genetic-based targets), the planning scenario (single-species, multi-species, multi-species with cost), or the type of genetic variation measured (adaptive or neutral).
-We also conducted a post-hoc analysis involving pairwise $G$ tests (using the \texttt{RVAideMemoire R} package; r@483).
+We also conducted a post-hoc analysis involving pairwise $G$ tests [using the \texttt{RVAideMemoire R} package; r@483].
 

### 2016-04-19 code/rmarkdown/article.Rmd
++ some article fixes + use PSOCK clusters (not SOCK) + use MLPE to correlate distances + use pcadapt + fix article + remove auto sorting from refs

@@
 # set wd
-setwd('..')
-setwd('..')
+knitr::opts_knit$set(root.dir=normalizePath('../..'))
+```
 
+```{r,include=FALSE}
 # load .rda file
-try(session::restore.session('data/results/results.rda'))
+try(session::restore.session('data/final/results.rda'))
 
@@
 <!--  main text -->
-The act of setting aside areas for preservation spearheads modern conservation efforts [@r440].
-Protected area networks buffer species from anthropogenic impacts and provide places for populations of species to persist.
-Intra-specific genetic variation underpins species persistence, and there has been increasing interest in understanding how to design protected area networks in a way that maximizes this variation [@r67; @r3; @r61].
-Broadly speaking, genetic variation can be classified as either adaptive or neutral [reviewed in @r48].
-Adaptive genetic variation is associated with loci that affect fitness.
+Protected areas spearhead conservation efforts [@r440].
+These places buffer species from anthropogenic impacts and provide places for species to persist.
+However, the resources available for conservation are limited.
+Thus protected areas need to be sited in places that achieve conservation goals for minimal cost.
+To ensure that protected areas are cost-effective, plans for protected areas (prioritizations) are often generated using decision support tools to identify near-optimal solutions.
+Additionally, because reserves cannot preserve the entire range for a given species, the plans for multiple reserves are often designed simultaneously to ensure that the network of protected areas can facilitate biological processes (eg.
+gene flow between populations, seasonal migration).
+For instance, \texttt{Marxan} [@r22] uses amount-based targets to identify protected area networks that preserve a suitable amount of habitat for each species (eg.
+100 km $^2$ of land occupied by the species needs to be secured in a reserve) and weights to penalize fragmented solutions.
+However, conservation planning exercises typically assume that all individuals in the same species are equivalent.
+
+Intra-specific genetic variation is an important component for species persistence.
+As a consequence, there has been increasing interest in understanding how to design prioritizations that preserve and cultivate this variation [@r67; @r3; @r61].
+Although the strength of natural selection is a continuous force; broadly speaking, genetic variation can be classified as either adaptive or neutral [reviewed in @r480].
+Adaptive genetic variation is associated with loci that (significantly) affect fitness.
+Typically, such "adaptive" loci are detected due to anomalous patterns among individuals [eg.
+@r456; @r487; @r480; @r490; but see @r489 for alternative methods].
 By preserving existing patterns of adaptive variation, protected areas can ensure that populations with particularly beneficial adaptations are not lost [@r3].
-In contrast, neutral genetic variation is associated with loci that do not affect fitness.
+In contrast, neutral genetic variation is associated with loci that do not (significantly) affect fitness.
 Neutral variation often reflects the evolutionary history of different populations.
@@
 inbreeding depression; @r67].
-However, genetic data requires considerable investment to obtain and analyze.
-As a consequence, conservation planners do not typically use genetic data to inform prioritizations [@r61].
+Thus protected areas networks should preserve both adaptive and neutral patterns of genetic variation for species of conservation interest [@r67; @r3].
+Despite this, genetic data are not often used to inform prioritizations because they require considerable investment to obtain and analyze [@r61; but see @r5 and @r157 for developments on using genetic data in conservation planning].
 Thus prioritizations may fail to capture species' intra-specific genetic variation, and ultimately fail to deliver long-term conservation outcomes.
 
-Recently, conservation scientists have been investigating surrogates for genetic variation to generate better prioritizations without the overheads associated with genetic data.
+Recently, conservation scientists have been investigating surrogates to generate prioritizations that adequately preserve genetic variation--but without actually needing to utilize genetic data to do this.
 Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been used to guide reserve selection [eg.
-habitat conditions; @r247; @r2].
-However, their effectiveness as surrogates in conservation planning remains unverified.
+environmental conditions; @r247; @r2].
+By preserving a more representative sampled of environmental conditions across the species' ranges, they aimed to generate prioritizations that preserved more adaptive variation.
+However, the effectiveness of this approach remains unverified.
 On the other hand, neutral genetic variation arises from a reduction in gene flow between individuals.
-Surrogates for neutral variation are based on variables that predict the level of connectivity between planning units.
+Surrogates for neutral variation have been based on variables that predict the level of connectivity between planning units.
 For instance, the isolation by distance theory [IBD; @r259] predicts that populations further apart will have less connectivity, exchange less genetic material, and in turn have less genetic variation in common.
@@
 Project members visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals.
-They genotyped samples using amplified fragment length polymorphisms [AFLP; @r453], and constructed matrices denoting the presence/absence of polymorphisms at loci for each species (mean `r mean(sapply(spp.StructureData.LST, n.loci))` \pm `r sd(sapply(spp.StructureData.LST, n.loci))` SD markers genotyped per species; see Table 1 in r@478 for the number markers for each species).
+They genotyped samples using amplified fragment length polymorphisms [AFLP; @r453], and constructed matrices denoting the presence/absence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; see Table 1 in @r478 for the number markers for each species).
 Although AFLPs are not ideal as genomic representatives (cf.
@@
 This data set has previously been used to address a range of different questions.
-Many studies have used this data set to explore patterns of adaptive [@r484; @r475; @r455; @r451; r@485] and neutral genetic variation [@r451; @r478; @r486].
+Many studies have used this data set to explore patterns of adaptive [@r484; @r451; @r475; @r455; @r485] and neutral genetic variation [@r478; @r451; @r486].
 However, to our knowledge, only a single study has used this data set to explore the effectiveness of a surrogate in the context of conservation planning.
-Taberlet et al.
-[-@r486] investigated whether prioritizations generated using the spatial distribution of a comprehensive set species (> 300) could adequately secure the intra-specific genetic variation for 27 species.
+Taberlet _et al._ [-@r486] investigated whether prioritizations generated using the spatial distribution of a comprehensive set species (> 300) could adequately secure the intra-specific genetic variation for 27 species.
 They found that the prioritization based on species distributions was substantially different to that based on genomic data.
@@
 To describe the geographic location of each grid cell, we projected the grid into an equi-distant coordinate system ($\texttt{ESRI:102031}$), calculated their centroids, and extracted their two-dimensional coordinates.
-Thus we constructed two-dimensional geographic space-wherein each planning unit occupies a single point--as a potential surrogate for neutral genetic variation.
+Thus we constructed two-dimensional geographic space--wherein each planning unit occupies a single point--as a potential surrogate for neutral genetic variation.
 
@@
 We used the number of genetic lineages in each species previously identified by Alvarez _et al._ [-@r478].
-For each species, we used \texttt{Structure} (version 2.3.1; using same parameters as @r478 per run with `r structure.params.LST[[MODE]]$numruns` runs per species) and \texttt{ClumPP} (version 1.1.2; @r459) to group individuals into populations.
+For each species, we used \texttt{Structure} [version 2.3.1; using same parameters as @r478 per run with `r structure.params.LST[[MODE]]$numruns` runs per species] and \texttt{ClumPP} [version 1.1.2; @r459] to group individuals into populations.
 We then ran $\texttt{Bayescan}$ for each species using these groupings [1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations; @r456].
@@
 For each species, we first generated two sets of prioritizations using increasing environmental and geographic-based targets (separately).
-Prioritizations were generated using the following surrogate-based targets: `r paste(rapr.params.LST[[MODE]]$surrogate.target,collapse=', ')`.
+Prioritizations were generated using the following surrogate-based targets: `r paste(rapr.params.LST[[MODE]]$surrogacy.analysis$surrogate.target,collapse=', ')`.
 We then create a set of null expectations with which to compare the surrogate-based prioritizations.
@@
 Marxan).
-For each species, for each target, we generated `r ` prioritizations by randomly selecting planning units occupied by the species.
+For each species, for each target, we generated `r rapr.params.LST[[MODE]]$surrogacy.analysis$random.replicates` prioritizations by randomly selecting planning units occupied by the species.
 The number of randomly selected planning units was set to equal the number of planning units in the optimal prioritizations.
@@
 We used a global $G$ test to assess whether the proportion of species that met their genetic target in a given prioritization depended on the combination of targets used to generate the planning units (amount-based targets, amount and surrogate-based targets, amount and genetic-based targets), the planning scenario (single-species, multi-species, multi-species with cost), or the type of genetic variation measured (adaptive or neutral).
-We also conducted a post-hoc analysis involving pairwise $G$ tests [using the \texttt{RVAideMemoire R} package; r@483].
+We also conducted a post-hoc analysis involving pairwise $G$ tests [using the \texttt{RVAideMemoire R} package; @r483].
 

### 2016-04-19 code/rmarkdown/figures.Rmd
+ updates to directory structure + use structure for analysis
++ some article fixes + use PSOCK clusters (not SOCK) + use MLPE to correlate distances + use pcadapt + fix article + remove auto sorting from refs

--- 2016-04-05 article/figures.Rmd
@@
 # set wd
-setwd('..')
+knitr::opts_knit$set(root.dir=normalizePath('../..'))
+```
 
+```{r, include=FALSE}
 # load .rda file
-try(session::restore.session('results/results.rda'))
+try(session::restore.session('data/final/results.rda'))
 
@@
 # prepare data
-correlation.PDF <- correlation.sub.DF %>%
+surrogacy.PDF <- surrogacy.DF %>%
 	group_by(Type, Surrogate.target, Method, Species) %>% 
@@
 p1 <- ggplot(aes(x=Surrogate.target,y=genetic.held, fill=Method),
-		data=filter(correlation.PDF, Type=='Environmental')) +
+		data=filter(surrogacy.PDF, Type=='Environmental')) +
 	geom_boxplot(position=position_dodge(0.9)) +
@@
 p2 <- ggplot(aes(x=Surrogate.target,y=genetic.held, fill=Method),
-			data=filter(correlation.PDF, Type=='Geographic')) +
+			data=filter(surrogacy.PDF, Type=='Geographic')) +
 	geom_boxplot(position=position_dodge(0.9)) +

### 2016-04-19 code/rmarkdown/supporting_information.Rmd
+ updates to directory structure + use structure for analysis
++ some article fixes + use PSOCK clusters (not SOCK) + use MLPE to correlate distances + use pcadapt + fix article + remove auto sorting from refs

@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9.5, fig.height=10, fig.cap='Relative support ($BIC$) for the number of populations in each species.
-Each panel denotes a different species.
-Each line denotes a different model parametrization.
-Asterisks indicate the model and number of populations with the most support.'}
-# format data
-bic.DF <- ldply(
-	seq_along(unique(spp.samples.DF$species)),
-	.fun=function(i) {
-		curr.MTX <- spp.Mclust.LST[[i]]$mclust$BIC
-		attr(curr.MTX, 'dim') <- c(nrow(curr.MTX), ncol(curr.MTX))
-		attr(curr.MTX, 'dimnames') <- list(
-			seq_len(nrow(curr.MTX)),
-			attr(curr.MTX, 'modelNames')
-		)
-		class(curr.MTX) <- 'matrix'
-		curr.DF <- curr.MTX %>% as.data.frame %>% 
-			mutate(G=attr(curr.MTX, 'G')) %>% 
-			gather(Model,BIC,-G)
-		curr.DF$BIC=curr.DF$BIC*-1
-		return(
-			mutate(
-				curr.DF,
-				Species=gsub('_', ' ', unique(spp.samples.DF$species)[i], fixed=TRUE),
-				best_G=curr.DF[which.min(curr.DF$BIC),'G'][[1]],
-				best_stat=min(curr.DF$BIC,na.rm=TRUE),
-				asterisk_col=c('black', rep(NA_character_, length(Species)-1))
-			)
-		)
-	}
-)
-# make plot
-suppressWarnings({
-	ggplot(data=bic.DF) +
-		geom_line(aes(y=BIC, x=G, color=Model, linetype=Model)) +
-		geom_point(aes(y=best_stat, x=best_G), size=4, shape=8) +
-		theme_classic() +
-		scale_y_reverse()+
-		theme(
-			strip.background = element_rect(fill='grey20'),
-			strip.text = element_text(color='white', face='italic', size=1)
-		) +
-		xlab('Number of populations') +
-		ylab('BIC') +
-		scale_linetype_manual(values=rep(c('solid','dashed'),each=7)) +
-		scale_color_manual(values=rep(gg_color_hue(7),2)) +
-		facet_wrap(~ Species, ncol=4, scales='free_y')
-})
-```
-
 ```{r, echo=FALSE}
@@
 		) 	
-		curr.probs <- spp.Mclust.LST[[i]]$mclust$z %>%
+		curr.probs <- spp.StructureAnalysis.LST[[i]]@results@summary %>%
 			as.data.frame() %>%
@@
 ```{r, echo=FALSE}
-knitr::kable(
-	format.dataframe(
-		filter(
-			ldply(
-				seq_along(unique(spp.samples.DF$species)), 
-				function(i) {
-					rbind(
-						data.frame(
-							Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
-							Loci='all',
-							k=spp.Mclust.LST[[i]]$nmds$ndim,
-							Stress=spp.Mclust.LST[[i]]$nmds$stress,
-							Converged=spp.Mclust.LST[[i]]$nmds$converged
-						),
-						ldply(
-							seq_along(spp.nmds.LST[[i]]),
-							function(j) {
-								data.frame(
-									Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
-									Loci=names(spp.nmds.LST[[i]])[j],
-									k=ifelse(is.null(spp.nmds.LST[[i]][[j]]), NA_integer_, spp.nmds.LST[[i]][[j]]$ndim),
-									Stress=ifelse(is.null(spp.nmds.LST[[i]][[j]]), NA_real_, spp.nmds.LST[[i]][[j]]$stress),
-									Converged=ifelse(is.null(spp.nmds.LST[[i]][[j]]), NA, spp.nmds.LST[[i]][[j]]$converged)
-								)
-						})
-					)
-				}
-			),
-			!is.na(Stress)
-		),
-		omit=c('Loci','Converged')
-	),
-	digits=2,
-	caption='Summary of non-metric multi-dimensional scaling (MDS) analyses on genetic variation for each species.
-The analyses on all loci were used to transform the binary presence/absence polymorphism data to continuous variables to train Gaussian mixture models and ultimately identify adaptive loci.
-The analyses on adaptive and neutral loci were used to transform the polymorphism data to identify main gradients of this variation.',
-	col.names=c('Species', 'Loci Type', 'k', 'NMDS Stress', 'Converged'),
-	align=c('c', 'c', 'c', 'c', 'c')
-)
-```
-
-```{r, echo=FALSE}
 loci_classification.DF <- ldply(
@@
 	function(i) {
-		if (!is.null(spp.BayeScan.sample.loci.subset.LST[[i]])) {
+		if (!is.null(spp.BayeScan.LST[[i]])) {
 			data.frame(
@@
 
+				Bayescan_fst=spp.BayeScan.LST[[i]]$bayescan@results@summary[[6]],
 				Bayescan_Probability=spp.BayeScan.LST[[i]]$bayescan@results@summary[[2]],
-				Bayescan_qval=spp.BayeScan.LST[[i]]$bayescan@results@summary[[4]],
 				Bayescan_alpha=spp.BayeScan.LST[[i]]$bayescan@results@summary[[5]],
-				Bayescan_fst=spp.BayeScan.LST[[i]]$bayescan@results@summary[[6]],
+				Bayescan_qvalue=spp.BayeScan.LST[[i]]$bayescan@results@summary[[4]],
 				Bayescan_Type=spp.BayeScan.LST[[i]]$bayescan@results@summary[[7]],
@@
 				PCadapt_K=spp.pcadapt.LST[[i]]$K,
-				PCadapt_chi2_statistic=spp.pcadapt.LST[[i]]$chi2.stat,
+				PCadapt_chi2_statistic=spp.pcadapt.LST[[i]]$stats$chi2.stat,
 				PCadapt_pvalue=spp.pcadapt.LST[[i]]$pvalues,
-				PCadapt_Type=replace(rep('neutral, length(spp.pcadapt.LST[[i]]$pvalues)), spp.pcadapt.LST[[i]]$adaptive.loci, 'adaptive')
+				PCadapt_qvalue=spp.pcadapt.LST[[i]]$qvalues$qvalues,
+				PCadapt_Type=replace(rep('neutral', length(spp.pcadapt.LST[[i]]$pvalues)), spp.pcadapt.LST[[i]]$adaptive.loci, 'adaptive'),
 
-				Overall_Type=replace(rep('neutral, length(spp.pcadapt.LST[[i]]$pvalues)), spp.OutlierDetectionResults.LST[[i]], 'adaptive')
+				Overall_Type=replace(rep('neutral', length(spp.pcadapt.LST[[i]]$pvalues)), spp.OutlierDetectionResults.LST[[i]], 'adaptive')
 			)
@@
 
-# save to file
-write.table(loci_classification.DF, 'article/Table_S2', sep=',', row.names=FALSE)
+write.table(loci_classification.DF, 'article/Table_S2.csv', sep=',', row.names=FALSE)
 ```

### 2016-04-19 code/rmarkdown/tables.Rmd
++ some article fixes + use PSOCK clusters (not SOCK) + use MLPE to correlate distances + use pcadapt + fix article + remove auto sorting from refs

---
fontsize: 11pt
documentclass: article
bibliography: Endnote_lib.bib
csl: reference-style.csl
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    keep_tex: yes
    includes:
      in_header: [preamble-latex.tex, preamble-latex2.tex]
---

```{r, include=FALSE}
# set wd
knitr::opts_knit$set(root.dir=normalizePath('../..'))
```

```{r, include=FALSE}
# load .rda file
try(session::restore.session('data/final/results.rda'))

# set options
options(error=NULL)

```

## Tables

```{r, echo=FALSE}
knitr::kable(
	format.dataframe(correlation.DF),
	digits=2,
	caption='Correlation between surrogate variables and genetic variation. The $R^2$ columns show the proportion of genetic variation explained by surrogates. Specifically, marginal $R^2$ represents the proportion of variation explained by the fixed effects alone. The conditional $R^2$ represents the proportion of variation explained by the fixed and random effects. All $\\chi^2$ tests were based on one degree of freedom. Significant $P$ values ($P$ < 0.05) indicate that surrogate variables correlate with genetic variation.',
	col.names=c('Species', 'Test', 'Marginal $R^2$', 'Conditional $R^2$', '$\\chi^2$', '\\textit{P}'),
	align=rep('c', ncol(correlation.DF)),
)
```

### 2016-04-19 code/R/analysis/00-initialization.R

@@ -18,6 +18,7 @@ if (length(args)>0) {
 # install packages
 # source('http://bioconductor.org/biocLite.R')
 # biocLite('qvalue')
+suppressMessages(library(qvalue))
 
 # load CRAN packages
 suppressMessages(library(stats))
@@ -35,7 +36,6 @@ suppressMessages(library(testthat))
 suppressMessages(library(parallel))
 suppressMessages(library(cluster))
 suppressMessages(library(rworldxtra))
-suppressMessages(library(snow))
 suppressMessages(library(doParallel))
 suppressMessages(library(english))
 suppressMessages(library(session))
@@ -47,6 +47,9 @@ suppressMessages(library(mclust))
 suppressMessages(library(cluster))
 suppressMessages(library(RVAideMemoire))
 suppressMessages(library(pcadapt))
+suppressMessages(library(ResistanceGA))
+suppressMessages(library(MuMIn))
+suppressMessages(library(knitr))
 
 ## load github packages
 # devtools::install_github('paleo13/raspr')

### 2016-04-19 code/R/analysis/04-outlier-locus-analysis.R

@@ -68,25 +68,36 @@ spp.pcadapt.LST <- llply(
 			v[which(is.na(v))] <- mean(v, na.rm=TRUE)
 			return(v)
 		})
-		
-		# run pcadapt with all principle components and
-		# determine the number needed to secure a acceptable
-		# level of variation
-		curr.first.pca <- pcadapt(curr.spp, K=pcadapt.params.LST[[MODE]]$initial.K, ploidy=1, min.maf=0) # super rare/abundant loci already omitted from analysis
-		prop.var.explained <- ((curr.first.pca$singular_values)^2)/length(curr.first.pca$pvalues) # adapted from code in pcadapt:::screePlot
+
+		# run first principle components to determine the number needed to secure a acceptable level of variation
+		initial.pca <- corpca(curr.spp, K=pcadapt.params.LST[[MODE]]$initial.K, ploidy=1, scale=FALSE)
+		prop.var.explained <- initial.pca$singular_values/sum(initial.pca$singular_values)
 		curr.spp.K <- which(cumsum(prop.var.explained) > pcadapt.params.LST[[MODE]]$min.variance.explained)[1]
 		
-		# run pcadapt with minimum number of principle components
-		# needed to secure a acceptable level of variation
-		curr.second.pca <- pcadapt(curr.spp, K=curr.spp.K, ploidy=1, min.maf=0)
+		# run second pica with minimum number of principle components needed to secure a acceptable level of variation
+		inference.pca <- corpca(curr.spp, K=curr.spp.K, ploidy=1, scale=FALSE)
 		
+		# compute statistics from pcadapt
+		stats.LST <- computeStats(
+			data=curr.spp, res=inference.pca, method='mahalanobis', 
+			nSNP=ncol(curr.spp), K=curr.spp.K, data.type='genotype', min.maf=0
+		)
+		
+		# calculate pvalues
+		pvalues.DBL <- pval(stats.LST, method='mahalanobis', K=curr.spp.K)
+		qvalues.LST <- qvalue(pvalues.DBL)
+		outliers <- which(qvalues.LST$qvalues < pcadapt.params.LST[[MODE]]$alpha.level)
+
 		# return list 
 		return(
 			list(
 				K=curr.spp.K,
-				first.pca=curr.first.pca,
-				second.pca=curr.second.pca,
[…]

### 2016-04-19 code/R/functions/MLPE.R

@@ -1,0 +1,41 @@
+#' MLPE model
+#'
+#' This function runs a maximum likelihood population effects mixed effects model given two pairwise distance matrices.
+#' The code for this function is based on \code{\link[ResistanceGA]{MLPE.lmm}}.
+#' @param y \code{matrix} object with pairwise genetic distances
+#' @param x \code{matrix} object with pairwise surrogate distances
+#' @param REML \code{logical} should model be fitted using REML?
+#' @example
+#' x <- as.matrix(dist(matrix(rnorm(10),ncol=2)))
+#' y <- as.matrix(dist(matrix(rnorm(10),ncol=2)))
+#' model <- MLPE(y,x,REML=FALSE)
+#' print(model)
+#' @export
+MLPE <- function(y,x,REML=TRUE) {
+	# init
+	if (inherits(y, 'dist')) y <- as.matrix(y)
+	if (inherits(x, 'dist')) x <- as.matrix(x)
+	ID <- ResistanceGA:::To.From.ID(POPS = nrow(x))
+	ZZ <- ResistanceGA:::ZZ.mat(ID = ID)
+	y <- ResistanceGA:::lower(y)
+	x <- ResistanceGA:::lower(x)
+	# prepare data
+	dat <- data.frame(ID, x = x, y = y)
+	dat <- dplyr::mutate(dat, x = (mean(x)-x)/sd(x), y = (mean(y)-y)/sd(y))
+	colnames(dat) <- c("id1", "id2", "x", "y")
+	# prepare model
+	mod_list <- list('full'=lme4:::lFormula(y ~ x + (1 | id1), data = dat, REML = REML),
+		"null"=lme4:::lFormula(y ~ 1 + (1 | id1), data = dat, REML = REML))
+	MODs <- lapply(mod_list, function(x) {
+		x$reTrms$Zt <- ZZ
+		# fit model
+		dfun <- do.call(lme4:::mkLmerDevfun, x)
+		opt <- lme4:::optimizeLmer(dfun)
+		MOD <- (lme4:::mkMerMod(environment(dfun), opt, x$reTrms, fr = x$fr))
+	})
+	# return model
+	return(MODs)
+}
+
+#' Calculate R^2 model from 
+

### 2016-04-26 code/rmarkdown/article.Rmd
+ article fixes + bug fix in pcadapt analysis

@@
 
-To describe the environmental characteristics of each grid cell, we obtained data for 19 bioclimatic variables across the study area [approx.
+To describe the environmental characteristics of each grid cell, we used broad-scale climatic variables.
+We obtained data for 19 bioclimatic variables across the extent of the study area [approx.
 1 km resolution; @r33].
-We used this data set in particular because it is commonly used in conservation planning to characterize environmental variation in conservation planning.
+We used this data set because it is commonly to map habitat suitability in conservation planning.
 These layers and the planning units were projected into an equal-area coordinate system ($\texttt{ESRI:102014}$).
-The 19 bioclimatic variables were clipped to the grid and subject to a principal components analysis (PCA; using \texttt{ArcMap} version 10.2.2; Table S1) to extract the main themes of environmental variation among the grid cells.
+To reduce dimensionality, the 19 bioclimatic variables were clipped to the planning units and subject to a principal components analysis (PCA; using \texttt{ArcMap} version 10.2.2; Table S1).
 The first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components (PCs) cumulatively explained `r round(pca.DF[[4]][surrogate.params.LST[[MODE]]$number.components],1)` % of the variation and were used to generate `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` new layers.
@@
 We used two outlier detection methods to achieve this.
-The basic premise underpinning outlier detection methods is that neutral loci are expected to exhibit a specific level of variation among populations, and loci that deviate from this expectation should be under selection [reviewed in @r480].
-The advantage of these methods--in contrast with environmental association analyses [eg.
-@r475]--is that they do not use environmental data to identify loci under selection, which would have biased our analysis.
+The basic premise underpinning such methods is that neutral loci are expected to exhibit a specific level of variation among populations, and loci that deviate from this expectation should be under selection [reviewed in @r480].
+The advantage of these methods--in contrast with environmental association analyses [XXX]--is that they do not use environmental data to identify loci under selection, which would have biased our analysis.
 We omitted the species \textit{`r missing.species`} from all subsequent analysis because it was only associated with one genetic lineage and thus not compatible with one of the outlier detection methods.
@@
 We used the number of genetic lineages in each species previously identified by Alvarez _et al._ [-@r478].
-For each species, we used \texttt{Structure} [version 2.3.1; using same parameters as @r478 per run with `r structure.params.LST[[MODE]]$numruns` runs per species] and \texttt{ClumPP} [version 1.1.2; @r459] to group individuals into populations.
+For each species, we used \texttt{Structure} [version 2.3.4; using same parameters as @r478 per run with `r structure.params.LST[[MODE]]$numruns` runs per species] and \texttt{ClumPP} [version 1.1.2; @r459] to group individuals into populations.
 We then ran $\texttt{Bayescan}$ for each species using these groupings [1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations; @r456].
-To avoid false-positives, we omitted loci where the global frequency of the minor allele was $\geq 0.05$ (following guidelines in the user manual) and assumed they were neutral.
+To avoid false-positives, we omitted loci where the global frequency of the minor allele was $\geq 0.1$ (following guidelines in the user manual) and assumed they were neutral.
 Additionally, we omitted individuals from this analysis when their population membership was uncertain (maximum membership probability $< `r structure.params.LST[[MODE]]$probability.threshold`$).

### 2016-04-29 code/rmarkdown/article.Rmd
+ changes to stats + surrogacy tests + figures, tables, SI

@@
 Number of references: XXXX \newline
-Number of figures, table and text boxes: 4 \newline
+Number of figures: XXXX \newline
+Number of tables: XXXX \newline
+Number of text boxes: XXXX \newline
 \newpage
@@
 
-Intra-specific genetic variation is an important component for species persistence.
-As a consequence, there has been increasing interest in understanding how to design prioritizations that preserve and cultivate this variation [@r67; @r3; @r61].
+Intra-specific genetic variation is an important for long-term species persistence.
+As a consequence, there has been increasing interest in designing prioritizations that preserve and cultivate this variation [@r67; @r3; @r61].
 Although the strength of natural selection is a continuous force; broadly speaking, genetic variation can be classified as either adaptive or neutral [reviewed in @r480].
@@
 inbreeding depression; @r67].
-Thus protected areas networks should preserve both adaptive and neutral patterns of genetic variation for species of conservation interest [@r67; @r3].
+Thus protected area networks should preserve both adaptive and neutral patterns of genetic variation for species of conservation interest [@r67; @r3].
 Despite this, genetic data are not often used to inform prioritizations because they require considerable investment to obtain and analyze [@r61; but see @r5 and @r157 for developments on using genetic data in conservation planning].
@@
 We used species distribution and genomic data collected by the IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for further explanation of data collection methods].
-This data set contains genetic data for a diverse range of species that were collected using a standardized sampling regime.
-Additionally, the data set covers a large geographic domain that spans across a range of attitudinal gradients.
-As a consequence, there is large environmental variation among the grid cell cells which is desirable to assess effectiveness of environmental variables as surrogates for adaptive genetic variation.
-Thus this data set was particularly well suited for our study.
-
-The IntroBioDiv project data set contains data for 27 alpine plant species that was collected using a 20' longitude by 21' latitude grid (approx.
+This data set contains information for 27 alpine plant species (Figure 1a).
+It was collected using a 20' longitude by 21' latitude grid (approx.
 22.3 km $\times$ 25 km; Figure S1).
@@
 Thus the data set contains information describing the genomic properties of individuals for each species present in every second grid cell.
+
+We chose this data set for several reasons.
+First, conservation planning exercises typically consider tens to hundreds of species with different evolutionary histories and life histories.
+To ensure that our results were relevant for conservation planners, we required a similar sized data set.
+This data set contains genetic data for a diverse range of plant species.
+Second, we required genetic data for all species to be collected under a standardized sampling scheme, and this data set used a grid to collect samples.
+Third, we required broad-scale environmental variation across the study area to test the effectiveness of environmental variables.
+This data set covers a large geographic domain that spans across attitudinal gradients.
+Thus this data set was particularly well suited for our study.
 
@@
 We examined the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation (respectively).
-We used the grid cells used to sample genetic data as planning units to develop prioritizations.
+We used the grid cells used to collect samples by the IntraBioDiv project as planning units to develop prioritizations.
 We used only planning units that contained at least a single individual ($n = `r nrow(grid.DF)`$) to reduce computational burden.
-To examine the potential effects of cost, we calculated the total human population density inside each planning unit [obtained at 1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472] and used this to represent acquisition cost.
+To examine the potential effects of cost, we calculated the total human population density inside each planning unit [obtained at 1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472] and used this to represent acquisition cost (Figure 1b).
 
@@
 1 km resolution; @r33].
-We used this data set because it is commonly to map habitat suitability in conservation planning.
+We used this data set because it is commonly to map habitat suitability in conservation planning [r@XXXX].
 These layers and the planning units were projected into an equal-area coordinate system ($\texttt{ESRI:102014}$).
 To reduce dimensionality, the 19 bioclimatic variables were clipped to the planning units and subject to a principal components analysis (PCA; using \texttt{ArcMap} version 10.2.2; Table S1).
-The first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components (PCs) cumulatively explained `r round(pca.DF[[4]][surrogate.params.LST[[MODE]]$number.components],1)` % of the variation and were used to generate `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` new layers.
+The first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components (PCs) cumulatively explained `r round(pca.DF[[4]][surrogate.params.LST[[MODE]]$number.components],1)` % of the variation and were used to generate new layers.
 We calculated the average value for each PC layer in each planning unit, and used these values to characterize climatic variation among the planning units (Figure S2).
@@
 We used two outlier detection methods to achieve this.
+We omitted the species \textit{`r missing.species`} from all subsequent analyses because it was not compatible with one of the outlier detection methods (see below).
 The basic premise underpinning such methods is that neutral loci are expected to exhibit a specific level of variation among populations, and loci that deviate from this expectation should be under selection [reviewed in @r480].
-The advantage of these methods--in contrast with environmental association analyses [XXX]--is that they do not use environmental data to identify loci under selection, which would have biased our analysis.
-We omitted the species \textit{`r missing.species`} from all subsequent analysis because it was only associated with one genetic lineage and thus not compatible with one of the outlier detection methods.
-Loci identified as outliers in both analyses were treated as adaptive, and the remaining loci were treated as neutral.
+The advantage of these methods--in contrast with environmental association analyses [@rXXX]--is that they do not use environmental data to identify loci under selection, which would have biased our analysis.
+Loci identified as outliers in both analyses were treated as adaptive and the rest were treated as neutral.
+To avoid false-positives, we omitted loci from both outlier detection methods where the global frequency of the minor allele was $\geq 0.1$ and assumed they were neutral.
 
@@
 We used the number of genetic lineages in each species previously identified by Alvarez _et al._ [-@r478].
+Since \textit{`r missing.species`} was previously identified as containing only one population, it was not compatible with \texttt{BayeScan}.
 For each species, we used \texttt{Structure} [version 2.3.4; using same parameters as @r478 per run with `r structure.params.LST[[MODE]]$numruns` runs per species] and \texttt{ClumPP} [version 1.1.2; @r459] to group individuals into populations.
 We then ran $\texttt{Bayescan}$ for each species using these groupings [1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations; @r456].
-To avoid false-positives, we omitted loci where the global frequency of the minor allele was $\geq 0.1$ (following guidelines in the user manual) and assumed they were neutral.
 Additionally, we omitted individuals from this analysis when their population membership was uncertain (maximum membership probability $< `r structure.params.LST[[MODE]]$probability.threshold`$).
-We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate [FDR $\leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].
+We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate [q $\leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].
 
 Second, we used PCAs to identify outlier loci using methods implemented in the \texttt{pcadapt R} package [version 2.2; @r487].
+To enable comparisons between the two outlier detection analyses, we used the same individuals in this analysis as used in the $BayeScan$ analysis.
 For each of the 26 species, we ran a PCA and extracted the minimum number of components needed to secure `r pcadapt.params.LST[[MODE]]$min.variance.explained` % of the variation in loci.
-We then computed Mahalanobis distances [@r398] to detect outlier loci ($P < `r pcadapt.params.LST[[MODE]]$alpha.level`$).
-To enable comparisons between the two outlier detection analyses, we used the same individuals in this analysis as used in the $BayeScan$ analysis.
+We then used Mahalanobis distances [@r398] to compute _P_ values and then q-values [using the \texttt{qvalue R} package; @r491].
+Outlier loci were detected using a suitable FDR ($q < `r pcadapt.params.LST[[MODE]]$alpha.level`$).
 
 After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species (Figures S6--S32).
-For each species, we discarded the population groupings, and partitioned adaptive and neutral loci into separate data sets.
-We used non-metric multi-dimensional scaling [NMDS; implemented in the \texttt{vegan R} pacakge; @r458] using Gower distances [@r464; using the \texttt{cluster R} package to accommodate sparsity; @r457] to derive continuous variables that described the main gradients of adaptive and neutral genetic variation for each species (Table S2).
+For each species, we discarded the population groupings, and partitioned adaptive and neutral loci into separate matrices.
+We used non-metric multi-dimensional scaling [NMDS; implemented in the \texttt{vegan R} package; @r458] using Gower distances [@r464; using the \texttt{cluster R} package to accommodate sparsity; @r457] to derive continuous variables that described the main gradients of adaptive and neutral genetic variation separately for each species (Table S2).
 To ensure that the ordinations described a sufficient amount of the genomic variation, we ran successive scaling analyses with increasing dimensionality until this was achieved (maximum stress value $\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for each analysis).
@@
 
-### Prioritizations and statistical analyses
+## Prioritizations and statistical analyses
 We used the unreliable representative and adequate prioritization (URAP) formulation of the reserve selection problem [@r465].
@@
 
-To address the first aim--to determine if the environmental and geographic variables correlate with adaptive and neutral genetic variation--we fit maximum likelihood population-effects (MLPE) models [implemented in the \texttt{ResistanceGA R} package; @r488].
-We computed differences between the grid cells in terms of their environmental conditions and  geographic position.
-Additionally, for each species, we computed the differences between the grid cells in terms of the adaptive and neutral genetic characteristics of individuals that inhabited them (if detected).
-All differences were calculated using Euclidean distances.
-We then fit a set of MLPE models to the differences in the adaptive genetic characteristics among the grid cells and the differences in the environmental conditions for each species associated with adaptive genetic variation.
-We also fit a set of similar models using the differences in neutral genetic characteristics and the geographic distances for each species.
-For each model we fit an intercept-only null model, and tested surrogate variables whether they explained substantially more variation using $\chi^2$.
-We then applied Bonferroni corrections to each set of models (separately).
-Note that this analysis does not take into account spatial auto-correlation.
-The subsequent analyses explicitly control for this effect.
+To address the second aim--to determine if the environmental and geographic variables are effective surrogates for adaptive and neutral genetic variation--we generated a collection of single-species prioritizations.
+For each species, we generated `r rapr.params.LST[[MODE]]$surrogacy.analysis$replicates` prioritizations by randomly selecting different combinations of planning units that were occupied by the species.
+We then computed the proportion of adaptive genetic variation and environmental variation secured in each prioritization.
+We then generated another set of random prioritizations for each species, and computed the proportion of neutral genetic variation and geographic variation secured in each of these prioritizations.
 
-To address the second aim--to determine if the environmental and geographic variables are effect surrogates for adaptive and neutral genetic variation--we generated a collection of single-species prioritizations.
-For each species, we first generated two sets of prioritizations using increasing environmental and geographic-based targets (separately).
-Prioritizations were generated using the following surrogate-based targets: `r paste(rapr.params.LST[[MODE]]$surrogacy.analysis$surrogate.target,collapse=', ')`.
-We then create a set of null expectations with which to compare the surrogate-based prioritizations.
-These null expectations represented possible solutions obtained using conventional conservation planning methods that just use amount (eg.
-Marxan).
-For each species, for each target, we generated `r rapr.params.LST[[MODE]]$surrogacy.analysis$random.replicates` prioritizations by randomly selecting planning units occupied by the species.
-The number of randomly selected planning units was set to equal the number of planning units in the optimal prioritizations.
+We then fit two sets of generalized linear model (GLM) with logit links.
+The first set of models were fit to the proportion of adaptive genetic variation secured by the prioritizations using the environmental variation secured secured in the prioritization for each species.
+Similarly, the second set of models assessed the relationship between the proportion of neutral genetic variation and the geographic variation secured in each prioritization for each species.
+We conducted likelihood ratio tests between each model and its corresponding null model, and applied Bonferroni corrections to assess significance.
+To assess the effectiveness of these variables as surrogates, we computed the Cragg and Uhler's pseudo-R$^2$ value for each model.
 
@@
 ## Results
-<!-- Surrogacy analysis -->
-The environmental and geographic-based variables were effective surrogates for adaptive and genetic variation (Figure 1).
-The prioritizations generated using greater surrogate-based targets secured more genetic variation (XXXX).
-In addition, prioritizations generated using environmental targets often secured more adaptive genetic variation than prioritizations generated using geographic-based targets did neutral genetic variation (XXXX).
-Only particularly high geographic targets ($\geq 0.95$ %) were able to secure a moderate proportion of neutral variation ($\geq XXXX$ %).
-Overall, there was a non-linear relationship between the surrogate-based targets used to generate the prioritizations and the proportion of genetic variation they secured.
+<!-- summary of methods -->
+We successfully mapped the main gradients of adaptive (if present) and neutral variation for each of the 26 species.
+The outlier detection methods detected adaptive genetic variation in `r length(env.surrogacy.LST)` species.
+Generally, the two outlier detection methods gave consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD).
+Of the species that were associated with adaptive genetic variation, only a few loci were detected as under selection (mean `r round(mean(spp.outlier.loci.props, na.rm=TRUE),2)` % loci per species `r round(sd(spp.outlier.loci.props, na.rm=TRUE),2)` SD).
+After identifying the loci under selection, we used NMDS analyses to identify the main gradients of adaptive (if detected) and neutral variation for each species.
+Generally, only a couple of continuous dimensions were needed to sufficiently describe the adaptive genetic variation for each species (minimum $K$ `r min(adaptive.k)`; mean $K$ `r mean(adaptive.k)`; max $K$ `r max(adaptive.k)`; Table SXXX).
+On the other hand, many continuous dimensions were often needed to sufficiently describe the neutral genetic variation for each species (minimum $K$ `r min(neutral.k)`; mean $K$ `r mean(neutral.k)`; max $K$ `r max(neutral.k)`; Table SXXX).
 
-<!-- Scenario analysis -->
+<!-- surrogacy analysis -->
+The surrogate variables correlated with the genetic characteristics of individuals found in the planning units for most species (Table SXXX).
+Specifically, environmental variation was found to significantly correlate with adaptive genetic variation for `r ` % of the species associated with loci under selection (_P_ < 0.05; mean `r ` $R^2$ $\pm$ `r ` SD).
+Additionally, geographic distances between planning units were found to significantly correlate with neutral genetic variation among planning units for `r ` % of the species (_P_ < 0.05; mean `r ` $R^2$ $\pm$ `r ` SD).
+These results suggest that prioritizations with different environmental characteristics tend to have individuals with different adaptive genetic characteristics.
+Similarly, planning units that are further apart tend to have individuals with different neutral genetic characteristics.
+
+<!-- scenario analysis -->
 To determine if the addition of environmental and geographic targets could improve the effectiveness of prioritizations in a more realistic context, we generated prioritizations according to three different scenarios (Figure 2).

### 2016-04-29 code/rmarkdown/figures.Rmd
+ changes to stats + surrogacy tests + figures, tables, SI

@@
 
+# download basemap
+data(countriesHigh)
+countries.FPLY <- countriesHigh[
+	countriesHigh$ADMIN %in% c(
+		'Italy', 'Switzerland', 'France', 'Austria',
+		'Germany', 'Slovenia', 'Croatia', 'Hungary',
+		'Monaco', 'Germany', 'Slovakia', 'Czech Republic'
+	)
+,] %>% spFortify
+grid.FPLY <- spFortify(grid.PLY)
+grid.sub.FPLY <- spFortify(grid.sub.PLY)
+
+# define ggplot2 color wheel function
+gg_color_hue <- function(n) {
+	hues <- seq(15, 375, length=n+1)
+	hcl(h=hues, l=65, c=100)[1:n]
+}
+
 ```
@@
 
-```{r, echo=FALSE, message=FALSE, fig.width=6, fig.height=4.5, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations.
-White boxplots show the proportion of genetic variation secured by prioritizations generated using surrogate-based targets.
-Grey boxplots show the proportion of genetic variation secured by randomly generated prioritizations with the same number of planning units as the surrogate-based prioritizations.
-Panel (a) shows the relationship between environmental surrogate targets used to generate a prioritization and the proportion of adaptive genetic variation it secures.
-Similarly, panel (b) shows that for geographic surrogate targets and neutral genetic variation.
-Box-plots show inter-specific variation in the proportion of genetic variation secured using different surrogate-based targets.'}
-# prepare data
-surrogacy.PDF <- surrogacy.DF %>%
-	group_by(Type, Surrogate.target, Method, Species) %>% 
-	summarize(genetic.held=mean(genetic.held, na.rm=TRUE)) %>%
-	data.frame()
+```{r, echo=FALSE, fig.width=8, fig.height=2.5, fig.cap="Map of the study area.
+Squares denote planning units.
+Panel (a) depicts patterns of species riches.
+Planning units with a brighter color are inhabited by more species.
+Panel (b) shows the distribution of opportunity cost across the study area (computed as total population density)."}
 
+# calculate species richness
+grid.PLY$Species_richness <- grid.PLY@data %>%
+	select(5:(4+n_distinct(spp.samples.DF$species))) %>% as.matrix() %>% rowSums()
+
+panel.DF <- grid.sub.FPLY %>% 
+	summarize(lat=min(lat)+(diff(range(lat))*0.98),
+		long=min(long)+(diff(range(long))*0.02))
+
+# plot species richness
+p1 <- ggplot() +
+	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
+		fill='grey85', color='grey70') +
+	geom_polygon(data=spFortify(grid.PLY), aes(x=long, y=lat, 
+		group=group, fill=Species_richness), color='grey10') +
+	scale_fill_gradientn(name='Richness',
+		colors=brewer.pal(9,'PuBu'),
+		guide=guide_colorbar(ticks=element_line(color='black'), border=element_line(color='black'))
+	) +
+	theme_classic() +
+	theme(
+		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"), axis.title=element_blank(),
+		panel.border=element_rect(color='black', fill=NA, size=1)
+	) +
+	coord_cartesian(
+		xlim=buffered.range(grid.FPLY$long, 0),
+		ylim=buffered.range(grid.FPLY$lat, 0)
+	) + geom_text(aes_string(x='long',y='lat'), data=panel.DF, label='(a)')
+
+qn01 <- scales::rescale(quantile(grid.PLY@data$pop.density, seq(0,1,length.out=9), na.rm=TRUE))
+p2 <- ggplot() +
+	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
+		fill='grey85', color='grey70') +
+	geom_polygon(data=spFortify(grid.PLY), aes(x=long, y=lat, 
+		group=group, fill=pop.density), color='grey10') +
+	scale_fill_gradientn(name='Cost',
+		colors=brewer.pal(9, 'PuBu'),
+		values=qn01,
+		guide=guide_colorbar(ticks=element_line(color='black'), border=element_line(color='black'))
+	) +
+	theme_classic() +
+	theme(
+		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"), axis.title=element_blank(),
+		panel.border=element_rect(color='black', fill=NA, size=1)
+	) +
+	coord_cartesian(
+		xlim=buffered.range(grid.FPLY$long, 0),
+		ylim=buffered.range(grid.FPLY$lat, 0)
+	)  + geom_text(aes_string(x='long',y='lat'), data=panel.DF, label='(b)')
+
+grid.arrange(p1,p2,nrow=1)
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=4.5, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations.
+Lines indicate the average trend for each species.
+Grey sillhouettes indicate their uncertainty in these trends.
+Histogram inserts show the distribution of $R^2$ values.
+Panel (a) shows the relationship between the proportion of environmental variation and adaptive genetic variation secured in a prioritization.
+Similarly, panel (b) shows that for geographic surrogate targets and neutral genetic variation.'}
 # make figure
-p1 <- ggplot(aes(x=Surrogate.target,y=genetic.held, fill=Method),
-		data=filter(surrogacy.PDF, Type=='Environmental')) +
-	geom_boxplot(position=position_dodge(0.9)) +
-	xlab('Environmental surrogate target (%)') + 
+p1 <- ggplot(aes(x=surrogate.held,y=genetic.held), data=env.surrogacy.DF) +
+	xlab(expression(atop('Environmental variation','secured (%)'))) + 
 	ylab('Adaptive genetic\nvariation secured (%)') +
-	geom_text(aes(x=0.55, y=1.115, label='(a)'), color='black', size=5) +
-	theme_classic() +
-	theme(legend.position=c(0.9,0.4)) +
-	scale_fill_manual(name='Prioritization', values=c('Optimal'='white', 'Random'='grey50'))
+	theme_classic() + theme(axis.line.x=element_line(color='black'),
+		axis.line.y=element_line(color='black')) +
+	geom_text(x=0.05, y=0.95, label='(a)') +
+	coord_cartesian(xlim=c(0, 1))
+for (i in seq_along(env.surrogacy.LST))
+	p1 <- p1 + stat_smooth(method='glm', method.args=list(family='binomial'),
+		data=filter(env.surrogacy.DF, Species==unique(Species)[i]),
+		fullrange=TRUE, se=FALSE, colour='black')
 
-p2 <- ggplot(aes(x=Surrogate.target,y=genetic.held, fill=Method),
-			data=filter(surrogacy.PDF, Type=='Geographic')) +
-	geom_boxplot(position=position_dodge(0.9)) +
-	xlab('Geographic surrogate target (%)') + 
+p2 <- ggplot(aes(x=surrogate.held,y=genetic.held), data=geo.surrogacy.DF) +
+	xlab(expression(atop('Geographic variation','secured (%)'))) + 
 	ylab('Neutral genetic\nvariation secured (%)') +
-	geom_text(aes(x=0.55, y=1.115, label='(b)'), color='black', size=5) +
-	theme_classic() +
-	theme(legend.position='none') +
-	scale_fill_manual(name='Prioritization', values=c('Optimal'='white', 'Random'='grey50'))
+	theme_classic() + theme(axis.line.x=element_line(color='black'),
+		axis.line.y=element_line(color='black', colour='transparent')) +
+	geom_text(x=0.05, y=0.95, label='(b)') +
+	coord_cartesian(xlim=c(0, 1))
+for (i in seq_along(geo.surrogacy.LST))
+	p2 <- p2 + stat_smooth(method='glm', method.args=list(family='binomial'),
+		data=filter(geo.surrogacy.DF, Species==unique(Species)[i]),
+		fullrange=TRUE, se=FALSE, colour='black')
+
+p1.hist <- ggplot(aes(R2),
+		data=filter(surrogacy.model.summary.DF, Type=='Environmental')) +
+	geom_histogram(fill='grey10') +
+	ylab('\nFrequency') + xlab(expression(atop('Effectiveness of environmental', variables~(R^2~coefficient)))) +
+	theme_classic() + theme(axis.line.x=element_line(color='black'),
+		axis.line.y=element_line(color='black'),
+		plot.background=element_rect(fill='transparent', colour='transparent')) +
+		geom_text(x=-1.375, y=0.95, label='(c)') +
+		coord_cartesian(xlim=c(-1.5, 1))
+
+p2.hist <- ggplot(aes(R2),
+		data=filter(surrogacy.model.summary.DF, Type=='Geographic')) +
+	geom_histogram(fill='grey10') +
+	ylab('\nFrequency') + xlab(expression(atop('Effectiveness of geographic', variables~(R^2~coefficient)))) +
+	theme_classic() + theme(axis.line.x=element_line(color='black'),
+		axis.line.y=element_line(color='black'), 
+		plot.background=element_rect(fill='transparent')) +
+		geom_text(x=-1.375, y=0.95, label='(d)') +
+		coord_cartesian(xlim=c(-1.5, 1))
 
 # render figure
-grid.arrange(p1,p2,nrow=2)
-
+grid.arrange(p1,p2,p1.hist,p2.hist,nrow=2)
 ```
@@
 The right column of panels (c, f, i) show prioritizations generated using amount- and genetic-based targets.'}
-# download basemap
-data(countriesHigh)
-countries.FPLY <- countriesHigh[
-	countriesHigh$ADMIN %in% c(
-		'Italy', 'Switzerland', 'France', 'Austria',
-		'Germany', 'Slovenia', 'Croatia', 'Hungary',
-		'Monaco', 'Germany', 'Slovakia', 'Czech Republic'
-	)
-,] %>% spFortify
 # prepare data for plotting

### 2016-04-29 code/rmarkdown/supporting_information.Rmd
+ changes to stats + surrogacy tests + figures, tables, SI

@@
 	facet_wrap(~ species, ncol=4)
-```
-
-```{r, echo=FALSE, fig.width=4, fig.height=2.5, fig.cap="Species richness. Squares denote planning units. Planning units with a brighter color are inhabited by more species."}
-# calculate species richness
-grid.PLY$Species_richness <- grid.PLY@data %>%
-	select(5:(4+n_distinct(spp.samples.DF$species))) %>% as.matrix() %>% rowSums()
-# plot species richness
-ggplot() +
-	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
-		fill='grey85', color='grey70') +
-	geom_polygon(data=spFortify(grid.PLY), aes(x=long, y=lat, 
-		group=group, fill=Species_richness), color='grey10') +
-	scale_fill_gradientn(name='Richness (#)',
-		colors=brewer.pal(9,'PuBu'),
-		guide=guide_colorbar(ticks=element_line(color='black'), border=element_line(color='black'))
-	) +
-	theme_classic() +
-	theme(
-		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"), axis.title=element_blank(),
-		panel.border=element_rect(color='black', fill=NA, size=1)
-	) +
-	coord_cartesian(
-		xlim=buffered.range(grid.FPLY$long, 0),
-		ylim=buffered.range(grid.FPLY$lat, 0)
-	)
-```
-
-```{r, echo=FALSE, fig.width=4, fig.height=2.5, fig.cap="Opportunity cost.
-Squares denote planning units.
-The color of each planning unit represents its opportunity cost (computed as total population density)"}
-# plot opportunity costs
-ggplot() +
-	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
-		fill='grey85', color='grey70') +
-	geom_polygon(data=spFortify(grid.PLY), aes(x=long, y=lat, 
-		group=group, fill=pop.density), color='grey10') +
-	scale_fill_gradientn(name='Cost',
-		colors=brewer.pal(9, 'PuBu'),
-		guide=guide_colorbar(ticks=element_line(color='black'), border=element_line(color='black'))
-	) +
-	theme_classic() +
-	theme(
-		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"), axis.title=element_blank(),
-		panel.border=element_rect(color='black', fill=NA, size=1)
-	) +
-	coord_cartesian(
-		xlim=buffered.range(grid.FPLY$long, 0),
-		ylim=buffered.range(grid.FPLY$lat, 0)
-	)
 ```
@@
 ```{r, echo=FALSE}
-loci_classification.DF <- ldply(
-	seq_along(unique(spp.samples.DF$species)), 
-	function(i) {
-		if (!is.null(spp.BayeScan.LST[[i]])) {
-			data.frame(
-				Species=paste0('\\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[i]),'}'),
-				Primer=spp.BayeScan.LST[[i]]$bayescan@data@primers,
+knitr::kable(
+	format.dataframe(
+		filter(spp.nmds.summary.DF,!is.na(Stress)),
+		omit=c('Loci','Converged')
+	),
+	digits=2,
+	caption='Summary of non-metric multi-dimensional scaling (MDS) analyses on genetic variation for each species.
+The analyses were used to identify the main gradients of variation in binary binary adaptive and neutral loci data.',
+	col.names=c('Species', 'Loci Type', 'K', 'NMDS Stress', 'Converged'),
+	align=c('c', 'c', 'c', 'c', 'c')
+)
+```
 
-				Bayescan_fst=spp.BayeScan.LST[[i]]$bayescan@results@summary[[6]],
-				Bayescan_Probability=spp.BayeScan.LST[[i]]$bayescan@results@summary[[2]],
-				Bayescan_alpha=spp.BayeScan.LST[[i]]$bayescan@results@summary[[5]],
-				Bayescan_qvalue=spp.BayeScan.LST[[i]]$bayescan@results@summary[[4]],
-				Bayescan_Type=spp.BayeScan.LST[[i]]$bayescan@results@summary[[7]],
-
-				PCadapt_K=spp.pcadapt.LST[[i]]$K,
-				PCadapt_chi2_statistic=spp.pcadapt.LST[[i]]$stats$chi2.stat,
-				PCadapt_pvalue=spp.pcadapt.LST[[i]]$pvalues,
-				PCadapt_qvalue=spp.pcadapt.LST[[i]]$qvalues$qvalues,
-				PCadapt_Type=replace(rep('neutral', length(spp.pcadapt.LST[[i]]$pvalues)), spp.pcadapt.LST[[i]]$adaptive.loci, 'adaptive'),
-
-				Overall_Type=replace(rep('neutral', length(spp.pcadapt.LST[[i]]$pvalues)), spp.OutlierDetectionResults.LST[[i]], 'adaptive')
-			)
-		}
-	}
-)
-
+```{r, echo=FALSE}
 write.table(loci_classification.DF, 'article/Table_S2.csv', sep=',', row.names=FALSE)

### 2016-04-29 code/rmarkdown/tables.Rmd
+ changes to stats + surrogacy tests + figures, tables, SI

@@
 ```{r, echo=FALSE}
-knitr::kable(
-	format.dataframe(correlation.DF),
+curr.DF <- correlation.model.summary.DF %>%
+	select(-contains('raw')) %>%
+	mutate(Environmental.corr.P=pvalString(Environmental.corr.P),
+		Geographic.corr.P=pvalString(Geographic.corr.P), blank=NA, Species=gsub('\\ ', '\\n', Species)) %>%
+	select(Species:Environmental.corr.P, blank, Geographic.Marginal.R2:Geographic.corr.P) %>%
+	format.dataframe 
+emphasize.strong.cells(which(curr.DF$Environmental.corr.P < 0.05, arr.ind=TRUE))
+emphasize.strong.cells(which(curr.DF$Geographic.corr.P < 0.05, arr.ind=TRUE))
+curr.txt=knitr::kable(
+	curr.DF,
 	digits=2,
@@
 Significant $P$ values ($P$ < 0.05) indicate that surrogate variables correlate with genetic variation.',
-	col.names=c('Species', 'Test', 'Marginal $R^2$', 'Conditional $R^2$', '$\\chi^2$', '\\textit{P}'),
-	align=rep('c', ncol(correlation.DF)),
-)
+	col.names=c('\\textbf{Species}', '\\textbf{Marginal $R^2$}', '\\textbf{Conditional $R^2$}', '\\textbf{$\\chi^2$}', '\\textbf{\\textit{P}}', '', '\\textbf{Marginal $R^2$}', '\\textbf{Conditional $R^2$}', '\\textbf{$\\chi^2$}', '\\textbf{\\textit{P}}'),
+	align=rep('c', ncol(curr.DF)),
+	format='latex'
+) %>% as.character()
+
+curr.txt <- gsub('\\hline\nSpecies',
+'
+\\hline
+& \\multicolumn{4}{c}{Environmental} & & \\multicolumn{4}{c}{Geographic} \\\\ \\cmidrule{2-5} \\cmidrule{7-10}
+Species', curr.txt, fixed=TRUE)
+
+class(curr.txt) <- 'knitr_kable'
+attr(curr.txt, 'format') <- 'latex'
+curr.txt
 ```

### 2016-05-02 code/rmarkdown/figures.Rmd
+ fix bug with missing species variable being overwritten + assorted figures and tables fixes

@@
 		data=filter(surrogacy.model.summary.DF, Type=='Environmental')) +
-	geom_histogram(fill='grey10') +
+	geom_histogram(fill='grey10', bins=10) +
 	ylab('\nFrequency') + xlab(expression(atop('Effectiveness of environmental', variables~(R^2~coefficient)))) +
@@
 		plot.background=element_rect(fill='transparent', colour='transparent')) +
-		geom_text(x=-1.375, y=0.95, label='(c)') +
-		coord_cartesian(xlim=c(-1.5, 1))
+		geom_text(x=0.05, y=0.95, label='(c)') +
+		coord_cartesian(xlim=c(0, 1))
 
@@
 		data=filter(surrogacy.model.summary.DF, Type=='Geographic')) +
-	geom_histogram(fill='grey10') +
+	geom_histogram(fill='grey10', bins=10) +
 	ylab('\nFrequency') + xlab(expression(atop('Effectiveness of geographic', variables~(R^2~coefficient)))) +
@@
 		plot.background=element_rect(fill='transparent')) +
-		geom_text(x=-1.375, y=0.95, label='(d)') +
-		coord_cartesian(xlim=c(-1.5, 1))
+		geom_text(x=0.05, y=0.95, label='(d)') +
+		coord_cartesian(xlim=c(0, 1))
 

### 2016-05-02 code/rmarkdown/tables.Rmd
+ fix bug with missing species variable being overwritten + assorted figures and tables fixes

@@
 	mutate(Environmental.corr.P=pvalString(Environmental.corr.P),
-		Geographic.corr.P=pvalString(Geographic.corr.P), blank=NA, Species=gsub('\\ ', '\\n', Species)) %>%
+		Geographic.corr.P=pvalString(Geographic.corr.P), blank='', Species=gsub('\\ ', '\\n', Species)) %>%
 	select(Species:Environmental.corr.P, blank, Geographic.Marginal.R2:Geographic.corr.P) %>%
-	format.dataframe 
+	format.dataframe(omit=c('Environmental.corr.P', 'blank', 'Geographic.corr.P')) 
 emphasize.strong.cells(which(curr.DF$Environmental.corr.P < 0.05, arr.ind=TRUE))

### 2016-05-03 code/rmarkdown/article.Rmd
+ changes to rapr params + manuscript edits + figure fixes

@@
 1 km resolution; @r33].
-We used this data set because it is commonly to map habitat suitability in conservation planning [r@XXXX].
+We used this data set because it is commonly to map habitat suitability in conservation planning [r@r493].
 These layers and the planning units were projected into an equal-area coordinate system ($\texttt{ESRI:102014}$).
@@
 The basic premise underpinning such methods is that neutral loci are expected to exhibit a specific level of variation among populations, and loci that deviate from this expectation should be under selection [reviewed in @r480].
-The advantage of these methods--in contrast with environmental association analyses [@rXXX]--is that they do not use environmental data to identify loci under selection, which would have biased our analysis.
+The advantage of these methods--in contrast with environmental association analyses [@r489]--is that they do not use environmental data to identify loci under selection, which would have biased our analysis.
 Loci identified as outliers in both analyses were treated as adaptive and the rest were treated as neutral.
@@
 We adopted a similar methodology to Bothwell _et al._ [-@r455] and applied it to each of the species in the data set.
-To avoid falsely classifying loci as adaptive, we grouped individuals into genetic lineages for each species (Figures SXXX--SXXX).
+To avoid falsely classifying loci as adaptive, we grouped individuals into genetic lineages for each species (Figures S3--S28).
 We used the number of genetic lineages in each species previously identified by Alvarez _et al._ [-@r478].
@@
 Additionally, we omitted individuals from this analysis when their population membership was uncertain (maximum membership probability $< `r structure.params.LST[[MODE]]$probability.threshold`$).
-We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate [q $\leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].
+We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate [$q \leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].
 
 Second, we used PCAs to identify outlier loci using methods implemented in the \texttt{pcadapt R} package [version 2.2; @r487].
-To enable comparisons between the two outlier detection analyses, we used the same individuals in this analysis as used in the $BayeScan$ analysis.
+To enable comparisons between the two outlier detection analyses, we used the same individuals in this analysis as used in the \texttt{BayeScan} analysis.
 For each of the 26 species, we ran a PCA and extracted the minimum number of components needed to secure `r pcadapt.params.LST[[MODE]]$min.variance.explained` % of the variation in loci.
@@
 
-After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species (Figures S6--S32).
+After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species (Figures S29--S54).
 For each species, we discarded the population groupings, and partitioned adaptive and neutral loci into separate matrices.
@@
 
-## Prioritizations and statistical analyses
+## Prioritization method
 We used the unreliable representative and adequate prioritization (URAP) formulation of the reserve selection problem [@r465].
@@
 
+## Statistical analyses
+To address the first aim--to determine if the environmental and geographic variables correlate with spatial patterns of adaptive and neutral genetic variation--we computed dissimilarity matrices between the planning units, and tested for correlations between matrices based on surrogate and genetic data.
+For each species, we computed two dissimilarity matrices between the planning units occupied by the species in terms of their environmental conditions (the PC values from the bioclimatc data) and geographic position (the z-scaled XY coordinates of the units' centroids).
+We then computed another dissimilarity matrix between the planning units using their values in the species' neutral genetic space.
+If the species was associated with adaptive loci, we also computed another dissimilarity matrix between the planning units using the species' adaptive genetic space.
+All dissimilarity matrices calculated using Euclidean distances.
+
+We then fit maximum likelihood population-effects (MLPE) models [implemented in the \texttt{ResistanceGA R} package; @r488].
+These models explicitly accommodate the structure of dissimilarity matrices using random effects [@rXXX].
+For each species, we fit a MLPE model to correlate the species' dissimilarity matrices based on geographic position and neutral genetic variation.
+Additionally, if the species was associated with adaptive loci, we also fit a MLPE model to correlate the species' dissimilarity matrices calculated using enviuronmental conditions and adaptive genetic variation.
+To test if the surrogate variables explained a substantial amount of genetic variation, we conducted $\chi^2$ tests between each model and its corresponding null model, and Bonferroni corrections.
+
 To address the second aim--to determine if the environmental and geographic variables are effective surrogates for adaptive and neutral genetic variation--we generated a collection of single-species prioritizations.
@@
 We then computed the proportion of adaptive genetic variation and environmental variation secured in each prioritization.
-We then generated another set of random prioritizations for each species, and computed the proportion of neutral genetic variation and geographic variation secured in each of these prioritizations.
+We then generated another random set of prioritizations, and computed the proportion of neutral genetic variation and geographic variation secured in each prioritization.
 
-We then fit two sets of generalized linear model (GLM) with logit links.
-The first set of models were fit to the proportion of adaptive genetic variation secured by the prioritizations using the environmental variation secured secured in the prioritization for each species.
+We fit two sets of generalized linear models (GLM) with logit links.
+The first set of models were fit to the proportion of adaptive genetic variation and environmental variation secured in each prioritization for each species associated with adaptive loci.
 Similarly, the second set of models assessed the relationship between the proportion of neutral genetic variation and the geographic variation secured in each prioritization for each species.
 We conducted likelihood ratio tests between each model and its corresponding null model, and applied Bonferroni corrections to assess significance.
-To assess the effectiveness of these variables as surrogates, we computed the Cragg and Uhler's pseudo-R$^2$ value for each model.
+To assess the effectiveness of these variables as surrogates, we computed the Cragg and Uhler's pseudo-R$^2$ value for each model [using the \texttt{pscl} R package; @r492].
 
@@
 To accommodate variation in the high number of optimal solutions for the single-species amount-based prioritizations, we generated `r (rapr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates)` replicate prioritizations.
-
-We then counted the number of species that were sufficiently represented in terms of their adaptive and neutral genetic variation (genetic variation secured $\geq$ `r (rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)` %) for each prioritization.
-We used a global $G$ test to assess whether the proportion of species that met their genetic target in a given prioritization depended on the combination of targets used to generate the planning units (amount-based targets, amount and surrogate-based targets, amount and genetic-based targets), the planning scenario (single-species, multi-species, multi-species with cost), or the type of genetic variation measured (adaptive or neutral).
-We also conducted a post-hoc analysis involving pairwise $G$ tests [using the \texttt{RVAideMemoire R} package; @r483].
+We then counted the proportion of species that were sufficiently represented in terms of their adaptive and neutral genetic variation (genetic variation secured $\geq$ `r (rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)` %) for each prioritization.
 
@@
 <!-- summary of methods -->
-We successfully mapped the main gradients of adaptive (if present) and neutral variation for each of the 26 species.
+We successfully mapped the main spatial patterns of adaptive (if present) and neutral variation for each of the 26 species.
 The outlier detection methods detected adaptive genetic variation in `r length(env.surrogacy.LST)` species.
@@
 After identifying the loci under selection, we used NMDS analyses to identify the main gradients of adaptive (if detected) and neutral variation for each species.
-Generally, only a couple of continuous dimensions were needed to sufficiently describe the adaptive genetic variation for each species (minimum $K$ `r min(adaptive.k)`; mean $K$ `r mean(adaptive.k)`; max $K$ `r max(adaptive.k)`; Table SXXX).
-On the other hand, many continuous dimensions were often needed to sufficiently describe the neutral genetic variation for each species (minimum $K$ `r min(neutral.k)`; mean $K$ `r mean(neutral.k)`; max $K$ `r max(neutral.k)`; Table SXXX).
+Generally, only a couple of continuous dimensions were needed to sufficiently describe the adaptive genetic variation for each species (minimum $K$ `r min(adaptive.k)`; mean $K$ `r mean(adaptive.k)`; max $K$ `r max(adaptive.k)`; Table S2).
+On the other hand, a few more continuous dimensions were often needed for neutral genetic variation for each species (minimum $K$ `r min(neutral.k)`; mean $K$ `r mean(neutral.k)`; max $K$ `r max(neutral.k)`; Table S2).
+The ordinations resulting from the NMDS analyses were used to construct an adaptive (if present) and neutral genetic space for each species.
+The spatial distribution of these genetic spaces (Figures S29--S54) generally show patterns of spatial autocorrelation--especially the adaptive genetic spaces (eg.
+_Phyteuma hemisphaericum_; S49).
+Additionally, most species show east-west patterns in their intra-specific genetic variation (eg.
+_Dryas octopetala_; S36).
+
+<!-- correlation analysis -->
+The dissimilarity matrices generated using the surrogate variables correlated with the dissimilarity matrices based on the genetic characteristics of individuals for most species (Table 1).
+Specifically, environmental variation was found to significantly correlate with adaptive genetic variation for `r sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)` species (`r round((sum(correlation.model.summary.DF$Environmental.corr.P<0.05)/length(na.omit(correlation.model.summary.DF$Environmental.corr.P)))*100,2)` % associated with loci under selection; _P_ < 0.05; mean marginal `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models).
+Additionally, geographic distances between planning units were found to significantly correlate with neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` species (`r round((sum(correlation.model.summary.DF$Geographic.corr.P<0.05)/length(na.omit(correlation.model.summary.DF$Geographic.corr.P)))*100,2)` % of all the species investigated;_P_ < 0.05; mean marginal `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)Geographic.Marginal.R2),2)` $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models).
+These results suggest that prioritizations with different environmental characteristics tend to have individuals with different adaptive genetic characteristics.
+Similarly, planning units that are further apart tend to have individuals with different neutral genetic characteristics.
+After verifying that the geographic and environmental variables correlate with genetic variation, the next logical step was to determine how effective these variables were as surrogates.
 
 <!-- surrogacy analysis -->
-The surrogate variables correlated with the genetic characteristics of individuals found in the planning units for most species (Table SXXX).
-Specifically, environmental variation was found to significantly correlate with adaptive genetic variation for `r ` % of the species associated with loci under selection (_P_ < 0.05; mean `r ` $R^2$ $\pm$ `r ` SD).
-Additionally, geographic distances between planning units were found to significantly correlate with neutral genetic variation among planning units for `r ` % of the species (_P_ < 0.05; mean `r ` $R^2$ $\pm$ `r ` SD).
-These results suggest that prioritizations with different environmental characteristics tend to have individuals with different adaptive genetic characteristics.
-Similarly, planning units that are further apart tend to have individuals with different neutral genetic characteristics.
 
 <!-- scenario analysis -->
-To determine if the addition of environmental and geographic targets could improve the effectiveness of prioritizations in a more realistic context, we generated prioritizations according to three different scenarios (Figure 2).
-The prioritizations generated under the single-species scenario showed similar selection frequencies across the study area--regardless of the combination of targets used to generate them.
+To determine if the addition of environmental and geographic targets could improve the effectiveness of prioritizations in a more realistic context, we generated prioritizations according to three different scenarios (Figure 3).
+The prioritizations generated under the single-species scenario showed relatively homogeneous selection frequencies across the study area--regardless of the combination of targets used to generate them (Figures 3a, 3b, 3c).
 On the other hand, the prioritizations generated under the multi-species scenario showed different selection frequencies depending on the targets used to generate them.
-The addition of surrogate- or genetic-based targets resulted in prioritizations that contained substantially more planning units.
+The prioritization generated using only amount-based targets contained fewer planning units (Figure 3d; $n = `r sum(selections(multi.spp.prioritisations[[1]]))`$) than the prioritizations generated using additional of surrogate-based targets (Figure 3e; $n = `r sum(selections(multi.spp.prioritisations[[2]]))`$) or genetic-based targets (Figure 3f; $n = `r sum(selections(multi.spp.prioritisations[[3]]))`$).
 The prioritizations generated under the multi-species with cost scenario had different spatial distributions to the previous two scenarios.
-The prioritizations generated under this third scenario concentrated conservation effort towards areas with higher elevation.
+These prioritizations tended to concentrate conservation effort towards areas with higher elevation.
+Similar to the second scenario, the amount-based prioritization contained less panning units (Figure 3g; $n = `r sum(selections(multi.spp.prioritisations.with.cost[[1]]))`$) than either the surrogate-based (Figure 3h; $n = `r sum(selections(multi.spp.prioritisations.with.cost[[2]]))`$) or the genetic-based prioritizations ((Figure 3i; $n = `r sum(selections(multi.spp.prioritisations.with.cost[[3]]))`$)).
 
-The inclusion of surrogate-based targets caused an improvement in the proportion of neutral variation secured in a prioritization (Figure 3).
-Specifically, the proportion of genetic variation secured by the prioritizations differed depending on the type of genetic variation measured (XXXX) and combination of targets used to generate the prioritizations (XXXX).
-No significant differences were found among the single-species, multi-species, and multi-species with cost contexts (XXXX).
+Generally, the prioritizations generated using amount-based targets met the targets for adaptive genetic variation for most species associated with adaptive loci (Figure 3).
+Among the three scenarios, the amount-based prioritization in the single-species scenario had the lowest proportion of species with targets met (`r ` % of species with adaptive loci; Figure 3a).
+In the other two scenarios--the multi-species (`r ` % species preserved with adaptive loci; Figure 3c) and multi-species with cost (`r ` % species preserved with adaptive loci; Figure 3e) scenarios--the amount-based prioritizations successfully preserved the adaptive variation for most of the species.
+The addition of surrogate- and genetic-based targets did little to improve the proportion of species with their targets met for adaptive genetic variation (Figures 3c and 3e).
+These results suggest that the main gradients of intra-specific adaptive variation can be sufficiently preserved using suitable amount-based targets.
+
+The prioritizations generated using additional surrogate-based targets met the neutral genetic variation targets for more species than the prioritizations generated using just amount-based targets (Figure 3).
+The biggest difference between the prioritizations generated using additional surrogate-based targets and just the amount-based targets was in the single-species scenarios (`r ` % and `r ` % of species meeting genetic variation targets respectively).
+In the other two scenarios, the amount-based prioritizations tended to perform slightly better (`r ` % of species with targets met), but were outperformed by prioritizations with additional surrogate-based targets.
+Although the surrogate-based targets tended to fulfill the neutral genetic variation targets for most species, they captured only a very small proportion of neutral genetic variation for two species in particular (_Gypsophila_repens_ and _Luzula_alpinopilosa_).
+These results suggest that surrogate-based targets can substantially improve the proportion of neutral genetic variation secured in a prioritization.
 

### 2016-05-03 code/rmarkdown/figures.Rmd
+ changes to rapr params + manuscript edits + figure fixes

@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=4.5, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations.
-Lines indicate the average trend for each species.
-Grey sillhouettes indicate their uncertainty in these trends.
-Histogram inserts show the distribution of $R^2$ values.
-Panel (a) shows the relationship between the proportion of environmental variation and adaptive genetic variation secured in a prioritization.
-Similarly, panel (b) shows that for geographic surrogate targets and neutral genetic variation.'}
+Panel (a) shows the relationship between the proportion of intra-specific adaptive genetic variation and environmental variation secured in a prioritization.
+Similarly, panel (b) shows the relationship between the proportion of intra-specific neutral variaiton and geograpgic variation secured in a prioritization.
+Panels (c) and (d) show the distribution of $R^2$ values for these models.'}
 # make figure

### 2016-05-03 code/rmarkdown/article.Rmd
+ bug fix in article.Rmd + added pull_ms command to makefile

@@
 Specifically, environmental variation was found to significantly correlate with adaptive genetic variation for `r sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)` species (`r round((sum(correlation.model.summary.DF$Environmental.corr.P<0.05)/length(na.omit(correlation.model.summary.DF$Environmental.corr.P)))*100,2)` % associated with loci under selection; _P_ < 0.05; mean marginal `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models).
-Additionally, geographic distances between planning units were found to significantly correlate with neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` species (`r round((sum(correlation.model.summary.DF$Geographic.corr.P<0.05)/length(na.omit(correlation.model.summary.DF$Geographic.corr.P)))*100,2)` % of all the species investigated;_P_ < 0.05; mean marginal `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)Geographic.Marginal.R2),2)` $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models).
+Additionally, geographic distances between planning units were found to significantly correlate with neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` species (`r round((sum(correlation.model.summary.DF$Geographic.corr.P<0.05)/length(na.omit(correlation.model.summary.DF$Geographic.corr.P)))*100,2)` % of all the species investigated;_P_ < 0.05; mean marginal `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models).
 These results suggest that prioritizations with different environmental characteristics tend to have individuals with different adaptive genetic characteristics.

### 2016-05-03 code/rmarkdown/article.Rmd
+ included missing ref

@@
 We used two outlier detection methods to achieve this.
-We omitted the species \textit{`r missing.species`} from all subsequent analyses because it was not compatible with one of the outlier detection methods (see below).
+We omitted the species \textit{`r gsub('_', ' ' , missing.species, fixed=TRUE)`} from all subsequent analyses because it was not compatible with one of the outlier detection methods (see below).
 The basic premise underpinning such methods is that neutral loci are expected to exhibit a specific level of variation among populations, and loci that deviate from this expectation should be under selection [reviewed in @r480].
@@
 We then fit maximum likelihood population-effects (MLPE) models [implemented in the \texttt{ResistanceGA R} package; @r488].
-These models explicitly accommodate the structure of dissimilarity matrices using random effects [@rXXX].
+These models explicitly accommodate the structure of dissimilarity matrices using random effects [@r494].
 For each species, we fit a MLPE model to correlate the species' dissimilarity matrices based on geographic position and neutral genetic variation.

### 2016-05-04 code/rmarkdown/article.Rmd
+ make file fix + article fix

@@
 We used the number of genetic lineages in each species previously identified by Alvarez _et al._ [-@r478].
-Since \textit{`r missing.species`} was previously identified as containing only one population, it was not compatible with \texttt{BayeScan}.
+Since \textit{`r gsub('_', ' ' , missing.species, fixed=TRUE)`} was previously identified as containing only one population, it was not compatible with \texttt{BayeScan}.
 For each species, we used \texttt{Structure} [version 2.3.4; using same parameters as @r478 per run with `r structure.params.LST[[MODE]]$numruns` runs per species] and \texttt{ClumPP} [version 1.1.2; @r459] to group individuals into populations.

### 2016-05-04 code/rmarkdown/supporting_information.Rmd
+ changes to rapr params + manuscript edits + figure fixes
+ fix adaptive + neutral plots

@@
 }
+
+is.speciesPlottable <- function(i) {
+	return(i <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[i] %in% missing.species) & !is.na(unique(spp.samples.DF$species)[i]))
+}
 ```
 
-```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(1), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[1]),'}.
+```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(1), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[1]),'}.
 Each panel shows a different dimension of the variation as indicated by its title.
@@
 No indiviudals of this species were found inside the gray planning units.')}
-if (1 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[1] %in% missing.species)) plot.spp.nmds(1)
+if (is.speciesPlottable(1)) plot.spp.nmds(1)
  ```
 
-```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(2), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[2]),'}.
+```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(2), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[2]),'}.
 See Figure S6 caption for conventions.')}
-if (2 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[2] %in% missing.species)) plot.spp.nmds(2)
+if (is.speciesPlottable(2)) plot.spp.nmds(2)
 ```
 
-```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(3), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[3]),'}.
+```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(3), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[3]),'}.
 See Figure S6 caption for conventions.')}
-if (3 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[3] %in% missing.species)) plot.spp.nmds(3)
+if (is.speciesPlottable(3)) plot.spp.nmds(3)
 ```
 
-```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(4), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[4]),'}.
+```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(4), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[4]),'}.
 See Figure S6 caption for conventions.')}
-if (4 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[4] %in% missing.species)) plot.spp.nmds(4)
+if (is.speciesPlottable(4)) plot.spp.nmds(4)
 ```
 
-```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(5), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[5]),'}.
+```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(5), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[5]),'}.
 See Figure S6 caption for conventions.')}
-if (5 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[5] %in% missing.species)) plot.spp.nmds(5)
+if (is.speciesPlottable(5)) plot.spp.nmds(5)
 ```
 
-```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(6), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[6]),'}.
+```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(6), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[6]),'}.
 See Figure S6 caption for conventions.')}
-if (6 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[6] %in% missing.species)) plot.spp.nmds(6)
+if (is.speciesPlottable(6)) plot.spp.nmds(6)
 ```
 
-```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(7), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[7]),'}.
+```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(7), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[7]),'}.
 See Figure S6 caption for conventions.')}
-if (7 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[7] %in% missing.species)) plot.spp.nmds(7)
+if (is.speciesPlottable(7)) plot.spp.nmds(7)
 ```
 
-```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(8), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[8]),'}.
+```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(8), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[8]),'}.
 See Figure S6 caption for conventions.')}
-if (8 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[8] %in% missing.species)) plot.spp.nmds(8)
+if (is.speciesPlottable(8)) plot.spp.nmds(8)
 ```
 
-```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(9), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[9]),'}.
+```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(9), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[9]),'}.
 See Figure S6 caption for conventions.')}
-if (9 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[9] %in% missing.species)) plot.spp.nmds(9)
+if (is.speciesPlottable(9)) plot.spp.nmds(9)
 ```
 
-```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(10), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[10]),'}.
+```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(10), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[10]),'}.
 See Figure S6 caption for conventions.')}
-if (10 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[10] %in% missing.species)) plot.spp.nmds(10)
+if (is.speciesPlottable(10)) plot.spp.nmds(10)
 ```
 
-```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(11), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[11]),'}.
+```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(11), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[11]),'}.
 See Figure S6 caption for conventions.')}
-if (11 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[11] %in% missing.species)) plot.spp.nmds(11)
+if (is.speciesPlottable(11)) plot.spp.nmds(11)
 ```
 
-```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(12), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[12]),'}.
+```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(12), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[12]),'}.
 See Figure S6 caption for conventions.')}
-if (12 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[12] %in% missing.species)) plot.spp.nmds(12)
+if (is.speciesPlottable(12)) plot.spp.nmds(12)
 ```
 
-```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(13), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[13]),'}.
+```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(13), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[13]),'}.
 See Figure S6 caption for conventions.')}
-if (13 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[13] %in% missing.species)) plot.spp.nmds(13)
+if (is.speciesPlottable(13)) plot.spp.nmds(13)
 ```
 
-```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(14), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[14]),'}.
+```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(14), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[14]),'}.
 See Figure S6 caption for conventions.')}
-if (14 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[14] %in% missing.species)) plot.spp.nmds(14)
+if (is.speciesPlottable(14)) plot.spp.nmds(14)
 ```
 
-```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(15), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[15]),'}.
+```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(15), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[15]),'}.
 See Figure S6 caption for conventions.')}
-if (15 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[15] %in% missing.species)) plot.spp.nmds(15)
+if (is.speciesPlottable(15)) plot.spp.nmds(15)
 ```
 
-```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(16), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[16]),'}.
+```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(16), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[16]),'}.
 See Figure S6 caption for conventions.')}
-if (16 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[16] %in% missing.species)) plot.spp.nmds(16)
+if (is.speciesPlottable(16)) plot.spp.nmds(16)
 ```
 
-```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(17), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[17]),'}.
+```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(17), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[17]),'}.
 See Figure S6 caption for conventions.')}
-if (17 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[17] %in% missing.species)) plot.spp.nmds(17)
+if (is.speciesPlottable(17)) plot.spp.nmds(17)
 ```
 
-```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(18), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[18]),'}.
+```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(18), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[18]),'}.
 See Figure S6 caption for conventions.')}
-if (18 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[18] %in% missing.species)) plot.spp.nmds(18)
+if (is.speciesPlottable(18)) plot.spp.nmds(18)
 ```
 
-```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(19), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[19]),'}.
+```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(19), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[19]),'}.
 See Figure S6 caption for conventions.')}
-if (19 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[19] %in% missing.species)) plot.spp.nmds(19)
+if (is.speciesPlottable(19)) plot.spp.nmds(19)
 ```
 
-```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(20), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[20]),'}.
+```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(20), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[20]),'}.
 See Figure S6 caption for conventions.')}
-if (20 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[20] %in% missing.species)) plot.spp.nmds(20)
+if (is.speciesPlottable(20)) plot.spp.nmds(20)
 ```
 
-```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(21), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[21]),'}.
+```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(21), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[21]),'}.
 See Figure S6 caption for conventions.')}
-if (21 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[21] %in% missing.species)) plot.spp.nmds(21)
+if (is.speciesPlottable(21)) plot.spp.nmds(21)
 ```
 
-```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(22), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[22]),'}.
+```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(22), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[22]),'}.
 See Figure S6 caption for conventions.')}
-if (22 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[22] %in% missing.species)) plot.spp.nmds(22)
+if (is.speciesPlottable(22)) plot.spp.nmds(22)
 ```
 
-```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(23), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[23]),'}.
+```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(23), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[23]),'}.
 See Figure S6 caption for conventions.')}
-if (23 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[23] %in% missing.species)) plot.spp.nmds(23)
+if (is.speciesPlottable(23)) plot.spp.nmds(23)
 ```
 
-```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(24), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[24]),'}.
+```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(24), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[24]),'}.
 See Figure S6 caption for conventions.')}
-if (24 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[24] %in% missing.species)) plot.spp.nmds(24)
+if (is.speciesPlottable(24)) plot.spp.nmds(24)
 ```
 
-```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(25), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[25]),'}.
+```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(25), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[25]),'}.
 See Figure S6 caption for conventions.')}
-if (25 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[25] %in% missing.species)) plot.spp.nmds(25)
+if (is.speciesPlottable(25)) plot.spp.nmds(25)
 ```
 
-```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(26), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[26]),'}.
+```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(26), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[26]),'}.
 See Figure S6 caption for conventions.')}
-if (26 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[26] %in% missing.species)) plot.spp.nmds(26)
+if (is.speciesPlottable(26)) plot.spp.nmds(26)
 ```
 
-```{r, echo=FALSE, fig.width=15, fig.height=calc.fig.height(27), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[27]),'}.
+```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(27), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[27]),'}.
 See Figure S6 caption for conventions.')}
-if (27 <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[27] %in% missing.species)) plot.spp.nmds(27)
+if (is.speciesPlottable(27)) plot.spp.nmds(27)
 ```
@@
 ```{r, echo=FALSE}
-write.table(loci_classification.DF, 'article/Table_S2.csv', sep=',', row.names=FALSE)
+write.table(loci_classification.DF, 'article/Table_S3.csv', sep=',', row.names=FALSE)
 ```

### 2016-05-04 code/rmarkdown/article.Rmd
+ assorted manuscript fixes + use version 3 of pcadapt + added extra figures to SI

@@
 1 km resolution; @r33].
-We used this data set because it is commonly to map habitat suitability in conservation planning [r@r493].
+We used this data set because it is commonly to map habitat suitability in conservation planning [@r493].
 These layers and the planning units were projected into an equal-area coordinate system ($\texttt{ESRI:102014}$).
@@
 To investigate the effectiveness of the surrogates for adaptive and neutral genetic variation, we first needed to identify which of the sampled loci--if any--were under selection.
-We used two outlier detection methods to achieve this.
+We used two outlier detection methods to achieve this [as recomended by @r495].
 We omitted the species \textit{`r gsub('_', ' ' , missing.species, fixed=TRUE)`} from all subsequent analyses because it was not compatible with one of the outlier detection methods (see below).
@@
 
-Second, we used PCAs to identify outlier loci using methods implemented in the \texttt{pcadapt R} package [version 2.2; @r487].
+Second, we used PCAs to identify outlier loci using methods implemented in the \texttt{pcadapt R} package [version 3.0; @r487; @r496].
 To enable comparisons between the two outlier detection analyses, we used the same individuals in this analysis as used in the \texttt{BayeScan} analysis.
-For each of the 26 species, we ran a PCA and extracted the minimum number of components needed to secure `r pcadapt.params.LST[[MODE]]$min.variance.explained` % of the variation in loci.
+For each of the 26 species, we first imputed missing loci data to permit further analysis.
+Specifically, the value for a given individual at a given locus was replaced with the mean frequency of that locus among all individuals.
+Next, we ran a PCA and extracted the minimum number of components needed to secure `r pcadapt.params.LST[[MODE]]$min.variance.explained` % of the variation in loci.
 We then used Mahalanobis distances [@r398] to compute _P_ values and then q-values [using the \texttt{qvalue R} package; @r491].
@@
 The outlier detection methods detected adaptive genetic variation in `r length(env.surrogacy.LST)` species.
-Generally, the two outlier detection methods gave consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD).
+Generally, two outlier detection methods gave fairly consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD).
 Of the species that were associated with adaptive genetic variation, only a few loci were detected as under selection (mean `r round(mean(spp.outlier.loci.props, na.rm=TRUE),2)` % loci per species `r round(sd(spp.outlier.loci.props, na.rm=TRUE),2)` SD).
 After identifying the loci under selection, we used NMDS analyses to identify the main gradients of adaptive (if detected) and neutral variation for each species.
-Generally, only a couple of continuous dimensions were needed to sufficiently describe the adaptive genetic variation for each species (minimum $K$ `r min(adaptive.k)`; mean $K$ `r mean(adaptive.k)`; max $K$ `r max(adaptive.k)`; Table S2).
+Generally, only a couple of continuous dimensions were needed to sufficiently describe the adaptive genetic variation for each species (minimum $K$ `r min(adaptive.k,ma.rm=TRUE)`; mean $K$ `r mean(adaptive.k,na.rm=TRUE)`; max $K$ `r max(adaptive.k,na.rm=TRUE)`; Table S2).
 On the other hand, a few more continuous dimensions were often needed for neutral genetic variation for each species (minimum $K$ `r min(neutral.k)`; mean $K$ `r mean(neutral.k)`; max $K$ `r max(neutral.k)`; Table S2).

### 2016-05-04 code/rmarkdown/figures.Rmd
+ assorted manuscript fixes + use version 3 of pcadapt + added extra figures to SI

@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=2.5, fig.cap="Map of the study area.
+```{r, echo=FALSE, fig.width=6, fig.height=1.5, fig.cap="Map of the study area.
 Squares denote planning units.
@@
 Planning units with a brighter color are inhabited by more species.
-Panel (b) shows the distribution of opportunity cost across the study area (computed as total population density)."}
+Panel (b) shows the distribution of opportunity cost across the study area (computed as total population density).
+Note that a quantile-based color ramp is used in this panel."}
 
@@
 panel.DF <- grid.sub.FPLY %>% 
-	summarize(lat=min(lat)+(diff(range(lat))*0.98),
-		long=min(long)+(diff(range(long))*0.02))
+	summarize(lat=min(lat)+(diff(range(lat))*0.96),
+		long=min(long)+(diff(range(long))*0.04))
 
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=5, fig.cap='The effectiveness of prioritizations at securing genetic variation.
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=5.5, fig.cap='The effectiveness of prioritizations at securing genetic variation.
 The top row of panels (a--b) show the performance of single-species prioritizations generated assuming equal costs among planning costs.
@@
 			data=scenario.PDF, position=position_dodge(0.9)) +
-		geom_text(aes(x=0.6, y=1.09, label=panel.letter),
+		geom_text(aes(x=0.6, y=1.15, label=panel.letter),
 			data=scenario.PDF) +

### 2016-05-04 code/rmarkdown/supporting_information.Rmd
+ assorted manuscript fixes + use version 3 of pcadapt + added extra figures to SI

@@
 
+```{r, echo=FALSE, message=FALSE, warnings=FALSE, fig.width=8, fig.height=8, fig.cap='The relationship between the proportion of adaptive genetic variation secured in a prioritizaiton and the proportion of environmental variation it also secures.
+Points represent values associated with randomly generated prioritizations.
+Black lines indicate average trends computed using generalised linear models with logit links.
+Each panel corresponds to different species.'}
+# format data
+curr.DF <- env.surrogacy.DF %>% filter(!is.na(genetic.held))
+
+# extract R2 values
+curr.DF2 <- surrogacy.model.summary.DF %>% filter(Type=='Environmental', !is.na(R2))
+curr.DF2$R2_repr <- round(curr.DF2$R2,2)
+curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
+curr.DF2$surrogate.held <- 0.1
+curr.DF2$genetic.held <- 0.9
+
+# make plot
+ggplot(aes(x=surrogate.held, y=genetic.held), data=curr.DF) +
+	theme_classic() +
+	theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=1)) +
+	xlab('Environmental variation secured (%)') +
+	ylab('Adaptive genetic variation secured (%)') +
+	stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
+		se=FALSE, color='black') +
+	geom_point() +
+	geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0) +
+	facet_wrap(~ Species, ncol=4)
+```
+
+```{r, echo=FALSE, message=FALSE, warnings=FALSE, fig.width=8, fig.height=8, fig.cap='The relationship between the proportion of neutral genetic variation secured in a prioritizaiton and the proportion of geographic variation it also secures.
+Conventions are detailed in Figure SXXX.'}
+# format data
+curr.DF <- geo.surrogacy.DF %>% filter(!is.na(genetic.held))
+
+# extract R2 values
+curr.DF2 <- surrogacy.model.summary.DF %>% filter(Type=='Geographic', !is.na(R2))
+curr.DF2$R2_repr <- round(curr.DF2$R2,2)
+curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
+curr.DF2$surrogate.held <- 0.1
+curr.DF2$genetic.held <- 0.9
+
+# make plot
+ggplot(aes(x=surrogate.held, y=genetic.held), data=curr.DF) +
+	theme_classic() +
+	theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=1)) +
+	xlab('Geographic variation secured (%)') +
+	ylab('Neutral genetic variation secured (%)') +
+	stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
+		se=FALSE, color='black') +
+	geom_point() +
+	geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0) +
+	facet_wrap(~ Species, ncol=4)
+```
+
 ```{r, echo=FALSE}

### 2016-05-04 code/R/analysis/00-initialization.R

@@ -51,6 +51,7 @@ suppressMessages(library(MuMIn))
 suppressMessages(library(knitr))
 suppressMessages(library(pscl))
 suppressMessages(library(lazyWeave))
+suppressMessages(library(pcadapt))
 
 ## load github packages
 # devtools::install_github('paleo13/raspr')
@@ -65,8 +66,6 @@ suppressMessages(library(ggplot2))
 suppressMessages(library(rgurobi))
 # devtools::install_github('wpeterman/ResistanceGA')
 suppressMessages(library(ResistanceGA))
-# devtools::install_github('cran/pcadapt@2.2') # install version 2.2 to manually control scaling
-suppressMessages(library(pcadapt))
 
 # set pander options
 panderOptions('knitr.auto.asis', FALSE)

### 2016-05-04 code/R/analysis/11-statistical-analysis.R

@@ -21,7 +21,7 @@ loci_classification.DF <- ldply(
 				Bayescan_Type=spp.BayeScan.LST[[i]]$bayescan@results@summary[[7]],
 				
 				PCadapt_K=spp.pcadapt.LST[[i]]$K,
-				PCadapt_chi2_statistic=spp.pcadapt.LST[[i]]$stats$chi2.stat,
+				PCadapt_chi2_statistic=spp.pcadapt.LST[[i]]$inference.run$chi2.stat,
 				PCadapt_pvalue=spp.pcadapt.LST[[i]]$pvalues,
 				PCadapt_qvalue=spp.pcadapt.LST[[i]]$qvalues$qvalues,
 				PCadapt_Type=replace(rep('neutral', length(spp.pcadapt.LST[[i]]$pvalues)), spp.pcadapt.LST[[i]]$adaptive.loci, 'adaptive'),

### 2016-05-06 code/rmarkdown/article.Rmd
+ fixes to stats + written more of results section + fixed all tables + in processing of fixing figures

@@
 
-We fit two sets of generalized linear models (GLM) with logit links.
-The first set of models were fit to the proportion of adaptive genetic variation and environmental variation secured in each prioritization for each species associated with adaptive loci.
-Similarly, the second set of models assessed the relationship between the proportion of neutral genetic variation and the geographic variation secured in each prioritization for each species.
-We conducted likelihood ratio tests between each model and its corresponding null model, and applied Bonferroni corrections to assess significance.
-To assess the effectiveness of these variables as surrogates, we computed the Cragg and Uhler's pseudo-R$^2$ value for each model [using the \texttt{pscl} R package; @r492].
+We fit two full generalized linear models (GLM) with a logit links.
+The first model was fit to the proportion of adaptive genetic variation secured in a prioritization, and contained a continuous predictor variable measuring the proportion of environmental variation also secured in the prioritization, a categorical variable indicating the species for which the prioritization was generated, and an interaction term between these predictor variables.
+The second model was fit to the proportion of neutral genetic variation secured in a prioritization, and contained a continuous predictor variable assessing the proportion of geographic variation also secured in the prioritization.
+Similar to the previous model, this model also contained a predictor variable indicating the species for which the prioritizations were generated, and an interaction term.
+We subject these models to a backwards step-wise term deletion routine to assess term significance.
+To assess the performance of these surrogates, we fit a suite of models correlating the proportion of adaptive and neutral genetic variation held in each prioritization and the environmental and geographic variation they also secured (respectively).
+We then computed the Cragg and Uhler's pseudo-R$^2$ value for each model [using the \texttt{pscl} R package; @r492].
 
@@
 Similarly, planning units that are further apart tend to have individuals with different neutral genetic characteristics.
-After verifying that the geographic and environmental variables correlate with genetic variation, the next logical step was to determine how effective these variables were as surrogates.
+After verifying that the geographic and environmental variables correlate with genetic variation, the next step was to determine how effective these variables were as surrogates in the context of conservation planning.
+The dissimilarity matrices generated using the surrogate variables correlated with the dissimilarity matrices based on the genetic characteristics of individuals for most species (Table 1).
+Specifically, environmental variation was found to significantly correlate with adaptive genetic variation for `r sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)` species (`r round((sum(correlation.model.summary.DF$Environmental.corr.P<0.05)/length(na.omit(correlation.model.summary.DF$Environmental.corr.P)))*100,2)` % associated with loci under selection; _P_ < 0.05; mean marginal `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models).
+Additionally, geographic distances between planning units were found to significantly correlate with neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` species (`r round((sum(correlation.model.summary.DF$Geographic.corr.P<0.05)/length(na.omit(correlation.model.summary.DF$Geographic.corr.P)))*100,2)` % of all the species investigated;_P_ < 0.05; mean marginal `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models).
+These results suggest that prioritizations with different environmental characteristics tend to have individuals with different adaptive genetic characteristics.
+Similarly, planning units that are further apart tend to have individuals with different neutral genetic characteristics.
+After verifying that the geographic and environmental variables correlate with genetic variation, the next step was to determine how effective these variables were as surrogates in the context of conservation planning.
 
 <!-- surrogacy analysis -->
+The environmental and geographic variables were moderately effective surrogates for genetic variation in most species (Figure 2).
+The relationship between the proportion of adaptive genetic variation secured in a prioritization and the proportion environment variation it also secured differed depending on the species for which the prioritization was generated ($\chi^2_{`r env.surrogacy.AOV$Resid..Dev[[4]]`} P = `r env.surrogacy.AOV$Pr..Chi.[[4]]`$).
+The slope for each species were all significantly different to zero (minimum $z value = `r min(env.surrogate.spp.DF$statistic)`$; maximum $P = r `max(env.surrogate.spp.DF$p.value)`$; Table S3).
+The explanatory power of the environmental surrogates were also generally quite high for most species (mean `r mean(env.surrogate.spp.DF$r2CU)` $R^2$ `r sd(env.surrogate.spp.DF$r2CU)` SD; Figure S56), with exception to _Carex sempervirens_.
+The nature of these relationships suggest that most of the adaptive variation for a given species can be secured in a prioritization using moderately high environmental-based targets.
+Similarly, the proportion of neutral genetic variation in a prioritization was affected by the interaction between species and the proportion of neutral variation secured in the prioritization ($\chi^2_{`r geo.surrogacy.AOV$Resid..Dev[[4]]`} P = `r geo.surrogacy.AOV$Pr..Chi.[[4]]`$).
+The slope for each species were also significantly different to zero (minimum $z value = `r min(geo.surrogate.spp.DF$statistic)`$; maximum $P = `r max(env.surrogate.spp.DF$p.value)`; Table S3).
+Generally, the explanatory power of geographic variables were higher than that of environmental variables (mean `r mean(geo.surrogate.spp.DF$r2CU)` $R^2$ `r sd(geo.surrogate.spp.DF$r2CU)` SD; Figure S57).
+Compared to the environmental variables, the slopes of the relationships suggest that very high geographic-based targets are required to generate prioritizations that secure a sufficient proportion of neutral genetic variation for most species.
+Additionally, all species were associated with prioritizations that secured a large proportion of environmental or geographic variation (> 70 %), yet preserved only a very small proportion of the genetic variation (< 10 %).
+These results suggest while prioritizations generated using surrogate-based targets tend to secure more genetic variation, prioritizations generated using just these targets can result in poor solutions.
 

### 2016-05-06 code/rmarkdown/figures.Rmd
+ fixes to stats + written more of results section + fixed all tables + in processing of fixing figures

@@
 
-p1.hist <- ggplot(aes(R2),
-		data=filter(surrogacy.model.summary.DF, Type=='Environmental')) +
+p1.hist <- ggplot(aes(r2CU),
+		data=filter(env.surrogate.spp.DF, !is.na(r2CU))) +
 	geom_histogram(fill='grey10', bins=10) +
@@
 
-p2.hist <- ggplot(aes(R2),
-		data=filter(surrogacy.model.summary.DF, Type=='Geographic')) +
+p2.hist <- ggplot(aes(r2CU),
+		data=filter(geo.surrogate.spp.DF, !is.na(r2CU))) +
 	geom_histogram(fill='grey10', bins=10) +
@@
 ```
+
+#### Supplementary info figures
+
+```{r, echo=FALSE, message=FALSE, warnings=FALSE, fig.width=8, fig.height=8, fig.cap='The relationship between the proportion of adaptive genetic variation secured in a prioritizaiton and the proportion of environmental variation it also secures.
+Points represent values associated with randomly generated prioritizations.
+Black lines indicate average trends computed using generalised linear models with logit links.
+Each panel corresponds to different species.'}
+# format data
+curr.DF <- env.surrogacy.DF %>% filter(!is.na(genetic.held))
+curr.DF2 <- env.surrogate.spp.DF %>% filter(!is.na(r2CU))
+curr.DF2$R2_repr <- round(curr.DF2$r2CU,2)
+curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
+curr.DF2$surrogate.held <- 0.05
+curr.DF2$genetic.held <- 0.95
+
+# make plot
+suppressWarnings(
+	ggplot(aes(x=surrogate.held, y=genetic.held), data=curr.DF) +
+		theme_classic() +
+		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=1)) +
+		xlab('Environmental variation secured (%)') +
+		ylab('Adaptive genetic variation secured (%)') +
+		stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
+			se=FALSE, color='blue') +
+		geom_point(fill='transparent') +
+		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=1, color='blue') +
+		facet_wrap(~ Species, ncol=4)
+)
+```
+
+```{r, echo=FALSE, message=FALSE, warnings=FALSE, fig.width=8, fig.height=8, fig.cap='The relationship between the proportion of neutral genetic variation secured in a prioritizaiton and the proportion of geographic variation it also secures.
+Conventions are detailed in Figure SXXX.'}
+# format data
+curr.DF <- geo.surrogacy.DF %>% filter(!is.na(genetic.held))
+curr.DF2 <- geo.surrogate.spp.DF %>% filter(!is.na(r2CU))
+curr.DF2$R2_repr <- round(curr.DF2$r2CU,2)
+curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
+curr.DF2$surrogate.held <- 0.05
+curr.DF2$genetic.held <- 0.95
+
+# make plot
+suppressWarnings(
+	ggplot(aes(x=surrogate.held, y=genetic.held), data=curr.DF) +
+		theme_classic() +
+		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=1)) +
+		xlab('Geographic variation secured (%)') +
+		ylab('Neutral genetic variation secured (%)') +
+		stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
+			se=FALSE, color='blue') +
+		geom_point(fill='transparent') +
+		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=1, color='blue') +
+		facet_wrap(~ Species, ncol=4)
+)
+```

### 2016-05-06 code/rmarkdown/supporting_information.Rmd
+ fixes to stats + written more of results section + fixed all tables + in processing of fixing figures

--- 2016-05-04 code/rmarkdown/supporting_information.Rmd
@@
 
-```{r, echo=FALSE}
+### insert figures here
+
+```{r, echo=FALSE, results='asis'}
 ## make results table showing Eigen values
-knitr::kable(
-	pca.DF,
-	digits=2,
+curr.file <- tempfile(fileext='.tex')
+pca.DF[['Eigen Value']] <- c(round(pca.DF[['Eigen Value']][-19],2), pvalString(pca.DF[['Eigen Value']][19], digits=2))
+pca.DF[['Variation explained (%)']] <- c(round(pca.DF[['Variation explained (%)']][1:3],2), pvalString(pca.DF[['Variation explained (%)']][-1:-3], digits=2))
+pca.DF[['Accumulative variation explained (%)']] <- as.character(round(pca.DF[['Accumulative variation explained (%)']],2))
+x <- pca.DF %>%
+	latex(file=curr.file, 
+		rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, col.just=rep('c', ncol(pca.DF)), here=TRUE,
+		colheads=c("\\parbox[t]{2.5cm}{\\centering \\bfseries{Principle Component}}", "\\parbox[t]{2cm}{\\centering \\bfseries{Eigen value}}",
+			"\\parbox[t]{3cm}{\\centering \\bfseries{Variation explained (\\%)}}", "\\parbox[t]{4.5cm}{\\centering \\bfseries{Cumulative variation explained (\\%)}}"),
 		caption=paste0('Summary of principle components analysis (PCA) on bioclimatic variation across the study area.
-The first ',as.character(as.english(surrogate.params.LST[[MODE]]$number.components)),' principle components (PCs) were used for subsequent analysis.'),
-	align=c('c', 'c', 'c', 'c')
+The first ',as.character(as.english(surrogate.params.LST[[MODE]]$number.components)),' principle components (PCs) were used for subsequent analysis.')
 	)
+readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
+	gsub(pattern='toprule', replacement='toprule[1pt]', fixed=TRUE) %>%
+	gsub(pattern='bottomrule', replacement='bottomrule[1pt]', fixed=TRUE) %>%
+	cat()
 ```
 
-```{r, echo=FALSE}
-knitr::kable(
-	format.dataframe(
-		filter(spp.nmds.summary.DF,!is.na(Stress)),
-		omit=c('Loci','Converged')
-	),
-	digits=2,
-	caption='Summary of non-metric multi-dimensional scaling (MDS) analyses on genetic variation for each species.
-The analyses were used to identify the main gradients of variation in binary binary adaptive and neutral loci data.',
-	col.names=c('Species', 'Loci Type', 'K', 'NMDS Stress', 'Converged'),
-	align=c('c', 'c', 'c', 'c', 'c')
+```{r, echo=FALSE, results='asis'}
+adaptive.spp.nmds.summary.DF <- spp.nmds.summary.DF %>% filter(Loci=='adaptive') %>% mutate(Stress=pvalString(Stress))
+neutral.spp.nmds.summary.DF <- spp.nmds.summary.DF %>% filter(Loci=='neutral') %>% mutate(Stress=pvalString(Stress))
+names(adaptive.spp.nmds.summary.DF)[2:5] <- paste0('adaptive.', names(adaptive.spp.nmds.summary.DF)[2:5])
+names(neutral.spp.nmds.summary.DF)[2:5] <- paste0('neutral.', names(neutral.spp.nmds.summary.DF)[2:5])
+spp.nmds.summary.DF2 <- full_join(adaptive.spp.nmds.summary.DF, neutral.spp.nmds.summary.DF, by='Species')
+spp.nmds.summary.DF2$adaptive.Stress[which(spp.nmds.summary.DF2$adaptive.Converged)] <- paste0('\\bfseries{', spp.nmds.summary.DF2$adaptive.Stress[which(spp.nmds.summary.DF2$adaptive.Converged)],'}')
+spp.nmds.summary.DF2$neutral.Stress[which(spp.nmds.summary.DF2$neutral.Converged)] <- paste0('\\bfseries{', spp.nmds.summary.DF2$neutral.Stress[which(spp.nmds.summary.DF2$neutral.Converged)],'}')
+spp.nmds.summary.DF2 <- spp.nmds.summary.DF2 %>%
+	mutate(blank='') %>%
+	select(Species, adaptive.k:adaptive.Stress, neutral.k:neutral.Stress)
+curr.file <- tempfile(fileext='.tex')
+x <- spp.nmds.summary.DF2 %>%
+	latex(
+		file=curr.file,
+		rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, col.just=rep('c', 5),
+		cgroup=c('', 'Adaptive', 'Neutral'), n.cgroup=c(1, 2, 2),
+		colheads=c('\\bfseries{Species}', 'K', 'Stress', 'K', 'Stress'), here=TRUE,
+		caption='Summary of non-metric multi-dimensional scaling (NMDS) analyses on genetic variation for each species.
+The analyses were used to identify the main graondients of variation in binary adaptive and neutral loci data.
+Bold stress values indicate NMDS analyses that have converged.'
 	)
+readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
+	gsub(pattern='\\cline{1-', replacement='\\cline{3-4} \\cline{6-', fixed=TRUE) %>% 
+	gsub(pattern='toprule', replacement='toprule[1pt]', fixed=TRUE) %>%
+	gsub(pattern='bottomrule', replacement='bottomrule[1pt]', fixed=TRUE) %>%
+	cat()
+```
+
+```{r, echo=FALSE, results='asis'}
+curr.file <- tempfile(fileext='.tex')
+x <- surrogate.spp.DF %>%
+	mutate(
+		Environmental.z = paste(round(Environmental.estimate,2), '$\\pm$', round(Environmental.std.error,2)),
+		Geographic.z = paste(round(Geographic.estimate,2), '$\\pm$', round(Geographic.std.error,2)),
+		Species = paste0('\\textit{', gsub('_', ' ', Species, fixed=TRUE), '}')) %>%
+	select(Species, Environmental.z, Environmental.p.value, Environmental.r2CU,
+		Geographic.z, Geographic.p.value, Geographic.r2CU) %>%
+	latex(file=curr.file, 
+		cgroup=c('', 'Environmental', 'Geographic'), n.cgroup=c(1, 3, 3), col.just=rep('c', 7),
+		digits=2, rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, here=TRUE,
+		colheads=c('\\bfseries{Species}', '$\\textit{Z}$', '\\textit{P}', '$R^2$', '$\\textit{Z}$', '\\textit{P}', '$R^2$'),
+		caption='Summary of post-hoc analyses on generalized linear models.
+Briefly, two generalized linear models were fit to correlate the proportion of adaptive genetic variation and environmental variation secured, and the proportion of neutral genetic variation and geographic variation secured in a prioritization.
+These models also included a predictor variable indicating the species for which the prioritizations were generated, and an interaction term.
+A post-hoc analysis was conducted to determine if the slopes for each species were significantly different to zero (detailed in the \\textit{Z} and \\textit{P} columns).
+All tests used a single degree of freedom.
+Cragg and Uhler\'s pseudo-R$^2$ values were calculated to show the proportion of variation explained by the slopes for each species ($R^2$ column).'
+	)
+readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
+	gsub(pattern='\\cline{1-', replacement='\\cline{2-5} \\cline{7-', fixed=TRUE) %>% 
+	gsub(pattern='ulticolumn{1}{c}{$R^2$}\\tabularnewline', replacement='ulticolumn{1}{c}{$R^2$}\\T\\B\\tabularnewline', fixed=TRUE) %>% 
+	gsub(pattern='toprule', replacement='toprule[1pt]', fixed=TRUE) %>%
+	gsub(pattern='bottomrule', replacement='bottomrule[1pt]', fixed=TRUE) %>%
+	cat()
 ```

### 2016-05-06 code/rmarkdown/tables.Rmd
+ fixes to stats + written more of results section + fixed all tables + in processing of fixing figures

--- 2016-04-19 code/rmarkdown/tables.Rmd
@@
 options(error=NULL)
-
 ```
-
 ## Tables
 
-```{r, echo=FALSE}
-knitr::kable(
-	format.dataframe(correlation.DF),
-	digits=2,
-	caption='Correlation between surrogate variables and genetic variation.
-The $R^2$ columns show the proportion of genetic variation explained by surrogates.
-Specifically, marginal $R^2$ represents the proportion of variation explained by the fixed effects alone.
-The conditional $R^2$ represents the proportion of variation explained by the fixed and random effects.
-All $\\chi^2$ tests were based on one degree of freedom.
-Significant $P$ values ($P$ < 0.05) indicate that surrogate variables correlate with genetic variation.',
-	col.names=c('Species', 'Test', 'Marginal $R^2$', 'Conditional $R^2$', '$\\chi^2$', '\\textit{P}'),
-	align=rep('c', ncol(correlation.DF)),
+```{r, echo=FALSE, results='asis'}
+curr.file <- tempfile(fileext='.tex')
+x <- correlation.model.summary.DF %>%
+	select(-contains('raw')) %>%
+	mutate(
+		Environmental.corr.P=pvalString(Environmental.corr.P),
+		Geographic.corr.P=pvalString(Geographic.corr.P), 
+		blank=rep('', length(Species)),
+		Species=paste0('\\textit{', gsub('\\ ', ' ', Species), '}')) %>%
+	select(Species:Environmental.corr.P, Geographic.Marginal.R2:Geographic.corr.P) %>%
+	latex(file=curr.file, 
+		cgroup=c('', 'Environmental', 'Geographic'), n.cgroup=c(1, 4, 4), col.just=rep('c', 9),
+		digits=2, rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, here=TRUE,
+		colheads=c('\\bfseries{Species}', '$MR^2$', '$CR^2$', '$\\chi^2$', '\\textit{P}', '$MR^2$', '$CR^2$', '$\\chi^2$', '\\textit{P}'),
+		caption='Summary of maximum likelihood population-effects (MLPE) models.
+Briefly, for each species associated with adaptive loci, a model was fit to correlate dissimilarities among planning units where the species in terms of the adaptive genetic characteristics of individuals in them with dissimilarities in terms of the environmental conditions inside the planning units (denoted in the environmental column group).
+Similarly, for each species, a model was fit to correlate dissimilarities among the planning units in terms of the neutral genetic characteristics of indiviudals in them with dissimilarities in terms of the geographic position of the planning units (denoted in the geographic column group).
+Data shows the marginal $R^2$ ($MR^2$) and conditional $R^2$ ($CR^2$) for each model, and also the results from comparing the models to their corresponding null models.
+All $\\chi^2$ tests had 1 degree of freedom.'
 	)
+readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
+	gsub(pattern='\\cline{1-', replacement='\\cline{2-6} \\cline{8-', fixed=TRUE) %>% 
+	gsub(pattern='{c}{\\textit{P}}\\tabularnewline', replacement='{c}{\\textit{P}}\\T\\B\\tabularnewline', fixed=TRUE) %>% 
+	gsub(pattern='toprule', replacement='toprule[1pt]', fixed=TRUE) %>%
+	gsub(pattern='bottomrule', replacement='bottomrule[1pt]', fixed=TRUE) %>%
+	cat()
  ```

### 2016-05-06 code/R/analysis/00-initialization.R

@@ -22,6 +22,8 @@ suppressMessages(library(qvalue))
 
 # load CRAN packages
 suppressMessages(library(stats))
+suppressMessages(library(Hmisc))
+suppressMessages(library(multcomp))
 suppressMessages(library(data.table))
 suppressMessages(library(grid))
 suppressMessages(library(gridExtra))
@@ -52,6 +54,7 @@ suppressMessages(library(knitr))
 suppressMessages(library(pscl))
 suppressMessages(library(lazyWeave))
 suppressMessages(library(pcadapt))
+suppressMessages(library(broom))
 
 ## load github packages
 # devtools::install_github('paleo13/raspr')

### 2016-05-06 code/R/analysis/11-statistical-analysis.R

@@ -153,47 +153,53 @@ correlation.model.summary.DF <- correlation.model.summary.DF %>%
 	Geographic.corr.P=corr.pvalues[nrow(correlation.model.summary.DF)+seq_len(nrow(correlation.model.summary.DF))])
 
 ### surrogacy analyses
-## run models
+## full model analysis
+# run models
+env.surrogacy.GLM <- glm(genetic.held~surrogate.held*Species, data=env.surrogacy.DF, family='binomial')
+geo.surrogacy.GLM <- glm(genetic.held~surrogate.held*Species, data=geo.surrogacy.DF, family='binomial')
+
+# assess term significance
+env.surrogacy.AOV <- anova(env.surrogacy.GLM, test='Chisq') %>% tidy()
+geo.surrogacy.AOV <- anova(geo.surrogacy.GLM, test='Chisq') %>% tidy()
+
+## species-level analysis
+# assess slope significance for each species
+env.surrogacy.spp.slopes.DF <- tidy(summary(glht(env.surrogacy.GLM, linfct=c('surrogate.held == 0', paste0('surrogate.held + surrogate.held:Species',unique(env.surrogacy.DF$Species)[-1],' == 0'))), adjusted('bonferroni')))
+geo.surrogacy.spp.slopes.DF <- tidy(summary(glht(geo.surrogacy.GLM, linfct=c('surrogate.held == 0', paste0('surrogate.held + surrogate.held:Species',unique(geo.surrogacy.DF$Species)[-1],' == 0'))), adjusted('bonferroni')))
+
+# compute R^2 values for each species
 env.surrogacy.LST <- list()
 for (x in unique(env.surrogacy.DF$Species)) {
-	curr.model=glm(genetic.held~surrogate.held, data=filter(env.surrogacy.DF, Species==x), family='binomial')
+	curr.model <- glm(genetic.held~surrogate.held, data=filter(env.surrogacy.DF, Species==x), family='binomial')
 	env.surrogacy.LST[[x]] <- list(model=curr.model, r2=pR2(curr.model))
 }
-
 geo.surrogacy.LST <- list()
-for (x in  unique(geo.surrogacy.DF$Species)) {
[…]

### 2016-05-06 code/rmarkdown/figures.Rmd
+ figure fixes

@@
 ## Figures
+
+<!--
 
@@
 
+-->
+
 #### Supplementary info figures
@@
 		theme_classic() +
-		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=1)) +
+		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=2.5)) +
 		xlab('Environmental variation secured (%)') +
 		ylab('Adaptive genetic variation secured (%)') +
+		geom_point(fill='transparent') +
 		stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
 			se=FALSE, color='blue') +
-		geom_point(fill='transparent') +
-		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=1, color='blue') +
+		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=2.5, color='blue') +
 		facet_wrap(~ Species, ncol=4)
@@
 		theme_classic() +
-		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=1)) +
+		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=5)) +
 		xlab('Geographic variation secured (%)') +
 		ylab('Neutral genetic variation secured (%)') +
+		geom_point(fill='transparent') +
 		stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
 			se=FALSE, color='blue') +
-		geom_point(fill='transparent') +
-		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=1, color='blue') +
+		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=5, color='blue') +
 		facet_wrap(~ Species, ncol=4)

### 2016-05-06 code/rmarkdown/figures.Rmd
+ set knitr exit for debugging
+ parameter changes

--- 2016-05-06 code/rmarkdown/figures.Rmd
@@
 
+```
+
+#### Supplementary info figures
+
+```{r, echo=FALSE, message=FALSE, warnings=FALSE, fig.width=8, fig.height=8, fig.cap='The relationship between the proportion of adaptive genetic variation secured in a prioritizaiton and the proportion of environmental variation it also secures.
+Points represent values associated with randomly generated prioritizations.
+Black lines indicate average trends computed using generalised linear models with logit links.
+Each panel corresponds to different species.'}
+# format data
+curr.DF <- env.surrogacy.DF %>% filter(!is.na(genetic.held))
+curr.DF2 <- env.surrogate.spp.DF %>% filter(!is.na(r2CU))
+curr.DF2$R2_repr <- round(curr.DF2$r2CU,2)
+curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
+curr.DF2$surrogate.held <- 0.05
+curr.DF2$genetic.held <- 0.95
+
+# make plot
+suppressWarnings(
+	ggplot(aes(x=surrogate.held, y=genetic.held), data=curr.DF) +
+		theme_classic() +
+		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=6)) +
+		xlab('Environmental variation secured (%)') +
+		ylab('Adaptive genetic variation secured (%)') +
+		geom_point(fill='transparent') +
+		suppressWarnings(stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
+			se=FALSE, fill='blue', alpha=0.8)) +
+		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=3, fill='blue', alpha=0.8) +
+		facet_wrap(~ Species, ncol=4)
+)
+```
+
+```{r, echo=FALSE, message=FALSE, warnings=FALSE, fig.width=8, fig.height=8, fig.cap='The relationship between the proportion of neutral genetic variation secured in a prioritizaiton and the proportion of geographic variation it also secures.
+Conventions are detailed in Figure SXXX.'}
+# format data
+curr.DF <- geo.surrogacy.DF %>% filter(!is.na(genetic.held))
+curr.DF2 <- geo.surrogate.spp.DF %>% filter(!is.na(r2CU))
+curr.DF2$R2_repr <- round(curr.DF2$r2CU,2)
+curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
+curr.DF2$surrogate.held <- 0.05
+curr.DF2$genetic.held <- 0.95
+
+# make plot
+suppressWarnings(
+	ggplot(aes(x=surrogate.held, y=genetic.held), data=curr.DF) +
+		theme_classic() +
+		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=8),
+			panel.margin.x.units=unit(1, 'lines')) +
+		xlab('Geographic variation secured (%)') +
+		ylab('Neutral genetic variation secured (%)') +
+		geom_point(fill='transparent') +
+		suppressWarnings(stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
+			se=FALSE, fill='blue', alpha=0.8)) +
+		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=4, fill='blue', alpha=0.8) +
+		facet_wrap(~ Species, ncol=4)
+)
+```
+
+```{r}
+knitr::knit_exit()
 ```
@@
 ```
-
-#### Supplementary info figures
-
-```{r, echo=FALSE, message=FALSE, warnings=FALSE, fig.width=8, fig.height=8, fig.cap='The relationship between the proportion of adaptive genetic variation secured in a prioritizaiton and the proportion of environmental variation it also secures.
-Points represent values associated with randomly generated prioritizations.
-Black lines indicate average trends computed using generalised linear models with logit links.
-Each panel corresponds to different species.'}
-# format data
-curr.DF <- env.surrogacy.DF %>% filter(!is.na(genetic.held))
-curr.DF2 <- env.surrogate.spp.DF %>% filter(!is.na(r2CU))
-curr.DF2$R2_repr <- round(curr.DF2$r2CU,2)
-curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
-curr.DF2$surrogate.held <- 0.05
-curr.DF2$genetic.held <- 0.95
-
-# make plot
-suppressWarnings(
-	ggplot(aes(x=surrogate.held, y=genetic.held), data=curr.DF) +
-		theme_classic() +
-		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=1)) +
-		xlab('Environmental variation secured (%)') +
-		ylab('Adaptive genetic variation secured (%)') +
-		stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
-			se=FALSE, color='blue') +
-		geom_point(fill='transparent') +
-		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=1, color='blue') +
-		facet_wrap(~ Species, ncol=4)
-)
-```
-
-```{r, echo=FALSE, message=FALSE, warnings=FALSE, fig.width=8, fig.height=8, fig.cap='The relationship between the proportion of neutral genetic variation secured in a prioritizaiton and the proportion of geographic variation it also secures.
-Conventions are detailed in Figure SXXX.'}
-# format data
-curr.DF <- geo.surrogacy.DF %>% filter(!is.na(genetic.held))
-curr.DF2 <- geo.surrogate.spp.DF %>% filter(!is.na(r2CU))
-curr.DF2$R2_repr <- round(curr.DF2$r2CU,2)
-curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
-curr.DF2$surrogate.held <- 0.05
-curr.DF2$genetic.held <- 0.95
-
-# make plot
-suppressWarnings(
-	ggplot(aes(x=surrogate.held, y=genetic.held), data=curr.DF) +
-		theme_classic() +
-		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=1)) +
-		xlab('Geographic variation secured (%)') +
-		ylab('Neutral genetic variation secured (%)') +
-		stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
-			se=FALSE, color='blue') +
-		geom_point(fill='transparent') +
-		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=1, color='blue') +
-		facet_wrap(~ Species, ncol=4)
-)
-```

### 2016-05-06 code/rmarkdown/figures.Rmd
+ fixes

--- 2016-05-06 code/rmarkdown/figures.Rmd
@@
 
+```
+
+#### Supplementary info figures
+
+```{r, echo=FALSE, message=FALSE, warnings=FALSE, fig.width=8, fig.height=8, fig.cap='The relationship between the proportion of adaptive genetic variation secured in a prioritizaiton and the proportion of environmental variation it also secures.
+Points represent values associated with randomly generated prioritizations.
+Black lines indicate average trends computed using generalised linear models with logit links.
+Each panel corresponds to different species.'}
+# format data
+curr.DF <- env.surrogacy.DF %>% filter(!is.na(genetic.held))
+curr.DF2 <- env.surrogate.spp.DF %>% filter(!is.na(r2CU))
+curr.DF2$R2_repr <- round(curr.DF2$r2CU,2)
+curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
+curr.DF2$surrogate.held <- 0.05
+curr.DF2$genetic.held <- 0.95
+
+# make plot
+suppressWarnings(
+	ggplot(aes(x=surrogate.held, y=genetic.held), data=curr.DF) +
+		theme_classic() +
+		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=6)) +
+		xlab('Environmental variation secured (%)') +
+		ylab('Adaptive genetic variation secured (%)') +
+		geom_point(fill='transparent') +
+		suppressWarnings(stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
+			se=FALSE, fill='blue', alpha=0.8)) +
+		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=3, color='blue', alpha=0.8) +
+		facet_wrap(~ Species, ncol=4)
+)
+```
+
+```{r, echo=FALSE, message=FALSE, warnings=FALSE, fig.width=8, fig.height=8, fig.cap='The relationship between the proportion of neutral genetic variation secured in a prioritizaiton and the proportion of geographic variation it also secures.
+Conventions are detailed in Figure SXXX.'}
+# format data
+curr.DF <- geo.surrogacy.DF %>% filter(!is.na(genetic.held))
+curr.DF2 <- geo.surrogate.spp.DF %>% filter(!is.na(r2CU))
+curr.DF2$R2_repr <- round(curr.DF2$r2CU,2)
+curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
+curr.DF2$surrogate.held <- 0.05
+curr.DF2$genetic.held <- 0.95
+
+# make plot
+suppressWarnings(
+	ggplot(aes(x=surrogate.held, y=genetic.held), data=curr.DF) +
+		theme_classic() +
+		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=8),
+			panel.margin.x.units=unit(1, 'lines')) +
+		xlab('Geographic variation secured (%)') +
+		ylab('Neutral genetic variation secured (%)') +
+		geom_point(fill='transparent') +
+		suppressWarnings(stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
+			se=FALSE, fill='blue', alpha=0.8)) +
+		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=4, color='blue', alpha=0.8) +
+		facet_wrap(~ Species, ncol=4)
+)
+```
+
+```{r}
+knitr::knit_exit()
 ```

### 2016-05-06 code/rmarkdown/figures.Rmd
+ fix2

@@
 		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=8),
-			panel.margin.x.units=unit(1, 'lines')) +
+			panel.margin.x.unit=unit(1, 'lines')) +
 		xlab('Geographic variation secured (%)') +

### 2016-05-06 code/rmarkdown/figures.Rmd
+ fix3

@@
 		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=8),
-			panel.margin.x.unit=unit(1, 'lines')) +
+			panel.margin.x=unit(1, 'lines')) +
 		xlab('Geographic variation secured (%)') +

### 2016-05-06 code/rmarkdown/figures.Rmd
+ fix figure heights

@@
 
-```{r, echo=FALSE, message=FALSE, warnings=FALSE, fig.width=8, fig.height=8, fig.cap='The relationship between the proportion of adaptive genetic variation secured in a prioritizaiton and the proportion of environmental variation it also secures.
+```{r}
+calc.fig.height2 <- function(x) {
+	return(((x-1) %/% 4)*3)
+}
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=calc.fig.height2(n_distinct(geo.surrogacy.DF$Species)), fig.cap='The relationship between the proportion of adaptive genetic variation secured in a prioritizaiton and the proportion of environmental variation it also secures.
 Points represent values associated with randomly generated prioritizations.
@@
 # format data
-curr.DF <- env.surrogacy.DF %>% filter(!is.na(genetic.held))
+curr.DF <- env.surrogacy.DF %>% filter(!is.na(genetic.held)) %>% mutate(Species=gsub('_', ' ', Species))
 curr.DF2 <- env.surrogate.spp.DF %>% filter(!is.na(r2CU))
@@
 		theme_classic() +
-		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=6)) +
+		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=10)) +
 		xlab('Environmental variation secured (%)') +
@@
 
-```{r, echo=FALSE, message=FALSE, warnings=FALSE, fig.width=8, fig.height=8, fig.cap='The relationship between the proportion of neutral genetic variation secured in a prioritizaiton and the proportion of geographic variation it also secures.
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=calc.fig.height2(n_distinct(geo.surrogacy.DF$Species)), fig.cap='The relationship between the proportion of neutral genetic variation secured in a prioritizaiton and the proportion of geographic variation it also secures.
 Conventions are detailed in Figure SXXX.'}
 # format data
-curr.DF <- geo.surrogacy.DF %>% filter(!is.na(genetic.held))
+curr.DF <- geo.surrogacy.DF %>% filter(!is.na(genetic.held)) %>% mutate(Species=gsub('_', ' ', Species))
 curr.DF2 <- geo.surrogate.spp.DF %>% filter(!is.na(r2CU))
@@
 		theme_classic() +
-		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=8),
+		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=10),
 			panel.margin.x=unit(1, 'lines')) +

### 2016-05-06 code/rmarkdown/figures.Rmd
+ fixes4

@@
 calc.fig.height2 <- function(x) {
-	return(((x-1) %/% 4)*3)
+	return(((x-1) %/% 4)*2.5)
 }
@@
 curr.DF <- env.surrogacy.DF %>% filter(!is.na(genetic.held)) %>% mutate(Species=gsub('_', ' ', Species))
-curr.DF2 <- env.surrogate.spp.DF %>% filter(!is.na(r2CU))
+curr.DF2 <- env.surrogate.spp.DF %>% filter(!is.na(r2CU)) %>% mutate(Species=gsub('_', ' ', Species))
 curr.DF2$R2_repr <- round(curr.DF2$r2CU,2)
@@
 curr.DF <- geo.surrogacy.DF %>% filter(!is.na(genetic.held)) %>% mutate(Species=gsub('_', ' ', Species))
-curr.DF2 <- geo.surrogate.spp.DF %>% filter(!is.na(r2CU))
+curr.DF2 <- geo.surrogate.spp.DF %>% filter(!is.na(r2CU)) %>% mutate(Species=gsub('_', ' ', Species))
 curr.DF2$R2_repr <- round(curr.DF2$r2CU,2)

### 2016-05-06 code/rmarkdown/figures.Rmd
+ more fixes

--- 2016-05-06 code/rmarkdown/figures.Rmd
@@
 
-```{r}
+```{r, include=FALSE}
 calc.fig.height2 <- function(x) {
-	return(((x-1) %/% 4)*3)
+	return(((x-1) %/% 4)*2.5)
 }
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=calc.fig.height2(n_distinct(geo.surrogacy.DF$Species)), fig.cap='The relationship between the proportion of adaptive genetic variation secured in a prioritizaiton and the proportion of environmental variation it also secures.
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=calc.fig.height2(n_distinct(geo.surrogacy.DF$Species)), fig.cap='The relationship between the proportion of adaptive genetic variation secured in a prioritizaiton and the proportion of environmental variation it also secures.
 Points represent values associated with randomly generated prioritizations.
@@
 curr.DF <- env.surrogacy.DF %>% filter(!is.na(genetic.held)) %>% mutate(Species=gsub('_', ' ', Species))
-curr.DF2 <- env.surrogate.spp.DF %>% filter(!is.na(r2CU))
+curr.DF2 <- env.surrogate.spp.DF %>% filter(!is.na(r2CU)) %>% mutate(Species=gsub('_', ' ', Species))
 curr.DF2$R2_repr <- round(curr.DF2$r2CU,2)
 curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
-curr.DF2$surrogate.held <- 0.05
+curr.DF2$surrogate.held <- 0.025
 curr.DF2$genetic.held <- 0.95
@@
 		theme_classic() +
-		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=10)) +
+		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=10),
+			panel.margin.x=unit(1, 'lines')) +
 		xlab('Environmental variation secured (%)') +
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=calc.fig.height2(n_distinct(geo.surrogacy.DF$Species)), fig.cap='The relationship between the proportion of neutral genetic variation secured in a prioritizaiton and the proportion of geographic variation it also secures.
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=calc.fig.height2(n_distinct(geo.surrogacy.DF$Species)), fig.cap='The relationship between the proportion of neutral genetic variation secured in a prioritizaiton and the proportion of geographic variation it also secures.
 Conventions are detailed in Figure SXXX.'}
@@
 curr.DF <- geo.surrogacy.DF %>% filter(!is.na(genetic.held)) %>% mutate(Species=gsub('_', ' ', Species))
-curr.DF2 <- geo.surrogate.spp.DF %>% filter(!is.na(r2CU))
+curr.DF2 <- geo.surrogate.spp.DF %>% filter(!is.na(r2CU)) %>% mutate(Species=gsub('_', ' ', Species))
 curr.DF2$R2_repr <- round(curr.DF2$r2CU,2)
 curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
-curr.DF2$surrogate.held <- 0.05
+curr.DF2$surrogate.held <- 0.025
 curr.DF2$genetic.held <- 0.95

### 2016-05-06 code/rmarkdown/figures.Rmd
+ more fig fixes
+ more fixes

@@
 calc.fig.height2 <- function(x) {
-	return(((x-1) %/% 4)*2.5)
+	return((((x-1) %/% 4)+1)*1)
 }
+
+calc.fig.height3 <- function(x) {
+	return((((x-1) %/% 4)+1)*1.5)
+}
+
 ```
@@
 		theme_classic() +
-		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=10),
+		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=14),
 			panel.margin.x=unit(1, 'lines')) +
@@
 			se=FALSE, fill='blue', alpha=0.8)) +
-		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=3, color='blue', alpha=0.8) +
+		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=6, color='blue', alpha=0.8) +
 		facet_wrap(~ Species, ncol=4)
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=calc.fig.height2(n_distinct(geo.surrogacy.DF$Species)), fig.cap='The relationship between the proportion of neutral genetic variation secured in a prioritizaiton and the proportion of geographic variation it also secures.
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=calc.fig.height3(n_distinct(geo.surrogacy.DF$Species)), fig.cap='The relationship between the proportion of neutral genetic variation secured in a prioritizaiton and the proportion of geographic variation it also secures.
 Conventions are detailed in Figure SXXX.'}
@@
 			se=FALSE, fill='blue', alpha=0.8)) +
-		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=4, color='blue', alpha=0.8) +
+		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=8, color='blue', alpha=0.8) +
 		facet_wrap(~ Species, ncol=4)

### 2016-05-06 code/rmarkdown/figures.Rmd
+ parameter fiddling

@@
 curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
-curr.DF2$surrogate.held <- 0.025
-curr.DF2$genetic.held <- 0.95
+curr.DF2$surrogate.held <- 0.01
+curr.DF2$genetic.held <- 0.9
 
@@
 		theme_classic() +
-		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=14),
+		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=12),
 			panel.margin.x=unit(1, 'lines')) +
@@
 			se=FALSE, fill='blue', alpha=0.8)) +
-		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=6, color='blue', alpha=0.8) +
+		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=6, color='blue') +
 		facet_wrap(~ Species, ncol=4)
@@
 			se=FALSE, fill='blue', alpha=0.8)) +
-		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=8, color='blue', alpha=0.8) +
+		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=4, color='blue') +
 		facet_wrap(~ Species, ncol=4)

### 2016-05-06 code/rmarkdown/figures.Rmd
+ more param fidddling

@@
 curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
-curr.DF2$surrogate.held <- 0.01
+curr.DF2$surrogate.held <- 0
 curr.DF2$genetic.held <- 0.9
@@
 			se=FALSE, fill='blue', alpha=0.8)) +
-		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=6, color='blue') +
+		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=5, color='blue') +
 		facet_wrap(~ Species, ncol=4)
@@
 			se=FALSE, fill='blue', alpha=0.8)) +
-		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=4, color='blue') +
+		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=5, color='blue') +
 		facet_wrap(~ Species, ncol=4)

### 2016-05-06 code/rmarkdown/figures.Rmd
+ fixed all figures except nmds maps

@@
 
-```
-
-#### Supplementary info figures
-
-```{r, include=FALSE}
-calc.fig.height2 <- function(x) {
-	return((((x-1) %/% 4)+1)*1)
-}
-
-calc.fig.height3 <- function(x) {
-	return((((x-1) %/% 4)+1)*1.5)
-}
-
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=calc.fig.height2(n_distinct(geo.surrogacy.DF$Species)), fig.cap='The relationship between the proportion of adaptive genetic variation secured in a prioritizaiton and the proportion of environmental variation it also secures.
-Points represent values associated with randomly generated prioritizations.
-Black lines indicate average trends computed using generalised linear models with logit links.
-Each panel corresponds to different species.'}
-# format data
-curr.DF <- env.surrogacy.DF %>% filter(!is.na(genetic.held)) %>% mutate(Species=gsub('_', ' ', Species))
-curr.DF2 <- env.surrogate.spp.DF %>% filter(!is.na(r2CU)) %>% mutate(Species=gsub('_', ' ', Species))
-curr.DF2$R2_repr <- round(curr.DF2$r2CU,2)
-curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
-curr.DF2$surrogate.held <- 0
-curr.DF2$genetic.held <- 0.9
-
-# make plot
-suppressWarnings(
-	ggplot(aes(x=surrogate.held, y=genetic.held), data=curr.DF) +
-		theme_classic() +
-		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=12),
-			panel.margin.x=unit(1, 'lines')) +
-		xlab('Environmental variation secured (%)') +
-		ylab('Adaptive genetic variation secured (%)') +
-		geom_point(fill='transparent') +
-		suppressWarnings(stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
-			se=FALSE, fill='blue', alpha=0.8)) +
-		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=5, color='blue') +
-		facet_wrap(~ Species, ncol=4)
-)
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=calc.fig.height3(n_distinct(geo.surrogacy.DF$Species)), fig.cap='The relationship between the proportion of neutral genetic variation secured in a prioritizaiton and the proportion of geographic variation it also secures.
-Conventions are detailed in Figure SXXX.'}
-# format data
-curr.DF <- geo.surrogacy.DF %>% filter(!is.na(genetic.held)) %>% mutate(Species=gsub('_', ' ', Species))
-curr.DF2 <- geo.surrogate.spp.DF %>% filter(!is.na(r2CU)) %>% mutate(Species=gsub('_', ' ', Species))
-curr.DF2$R2_repr <- round(curr.DF2$r2CU,2)
-curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
-curr.DF2$surrogate.held <- 0
-curr.DF2$genetic.held <- 0.9
-
-# make plot
-suppressWarnings(
-	ggplot(aes(x=surrogate.held, y=genetic.held), data=curr.DF) +
-		theme_classic() +
-		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=12),
-			panel.margin.x=unit(1, 'lines')) +
-		xlab('Geographic variation secured (%)') +
-		ylab('Neutral genetic variation secured (%)') +
-		geom_point(fill='transparent') +
-		suppressWarnings(stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
-			se=FALSE, fill='blue', alpha=0.8)) +
-		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=5, color='blue') +
-		facet_wrap(~ Species, ncol=4)
-)
-```
-
-```{r}
-knitr::knit_exit()
 ```

### 2016-05-06 code/rmarkdown/supporting_information.Rmd
+ fixed all figures except nmds maps

@@
 
-### insert figures here
+```{r, include=FALSE}
+calc.fig.height2 <- function(x) {
+	return((((x-1) %/% 4)+1)*1)
+}
+
+calc.fig.height3 <- function(x) {
+	return((((x-1) %/% 4)+1)*1.5)
+}
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=calc.fig.height2(n_distinct(geo.surrogacy.DF$Species)), fig.cap='The relationship between the proportion of adaptive genetic variation secured in a prioritizaiton and the proportion of environmental variation it also secures.
+Points represent values associated with randomly generated prioritizations.
+Black lines indicate average trends computed using generalised linear models with logit links.
+Each panel corresponds to different species.'}
+# format data
+curr.DF <- env.surrogacy.DF %>% filter(!is.na(genetic.held)) %>% mutate(Species=gsub('_', ' ', Species))
+curr.DF2 <- env.surrogate.spp.DF %>% filter(!is.na(r2CU)) %>% mutate(Species=gsub('_', ' ', Species))
+curr.DF2$R2_repr <- round(curr.DF2$r2CU,2)
+curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
+curr.DF2$surrogate.held <- 0
+curr.DF2$genetic.held <- 0.9
+
+# make plot
+suppressWarnings(
+	ggplot(aes(x=surrogate.held, y=genetic.held), data=curr.DF) +
+		theme_classic() +
+		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=12),
+			panel.margin.x=unit(1, 'lines')) +
+		xlab('Environmental variation secured (%)') +
+		ylab('Adaptive genetic variation secured (%)') +
+		geom_point(fill='transparent') +
+		suppressWarnings(stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
+			se=FALSE, fill='blue', alpha=0.8)) +
+		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=5, color='blue') +
+		facet_wrap(~ Species, ncol=4)
+)
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=calc.fig.height3(n_distinct(geo.surrogacy.DF$Species)), fig.cap='The relationship between the proportion of neutral genetic variation secured in a prioritizaiton and the proportion of geographic variation it also secures.
+Conventions are detailed in Figure SXXX.'}
+# format data
+curr.DF <- geo.surrogacy.DF %>% filter(!is.na(genetic.held)) %>% mutate(Species=gsub('_', ' ', Species))
+curr.DF2 <- geo.surrogate.spp.DF %>% filter(!is.na(r2CU)) %>% mutate(Species=gsub('_', ' ', Species))
+curr.DF2$R2_repr <- round(curr.DF2$r2CU,2)
+curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
+curr.DF2$surrogate.held <- 0
+curr.DF2$genetic.held <- 0.9
+
+# make plot
+suppressWarnings(
+	ggplot(aes(x=surrogate.held, y=genetic.held), data=curr.DF) +
+		theme_classic() +
+		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=12),
+			panel.margin.x=unit(1, 'lines')) +
+		xlab('Geographic variation secured (%)') +
+		ylab('Neutral genetic variation secured (%)') +
+		geom_point(fill='transparent') +
+		suppressWarnings(stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
+			se=FALSE, fill='blue', alpha=0.8)) +
+		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=5, color='blue') +
+		facet_wrap(~ Species, ncol=4)
+)
+```
 

### 2016-05-07 code/rmarkdown/supporting_information.Rmd
+ fixes to SI

@@
 					strip.background = element_rect(fill='grey20'),
-					strip.text = element_text(color='white')
+					strip.text = element_text(color='white', size=5)
 				) +
@@
 
-```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_',' ',unique(spp.samples.DF$species)[1], fixed=TRUE),'} populations.
+```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_',' ',unique(spp.samples.DF$species)[1], fixed=TRUE),'} populations.
 Squares denote planning units.
@@
 
-```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[2],fixed=TRUE),'} populations.
-See Figure SX caption for conventions.')}
+```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[2],fixed=TRUE),'} populations.
+See Figure S3 caption for conventions.')}
 if (2 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(2)
@@
 
-```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[3],fixed=TRUE),'} populations.
-See Figure SX caption for conventions.')}
+```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[3],fixed=TRUE),'} populations.
+See Figure S3 caption for conventions.')}
 if (3 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(3)
@@
 
-```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[4],fixed=TRUE),'} populations.
-See Figure SX caption for conventions.')}
+```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[4],fixed=TRUE),'} populations.
+See Figure S3 caption for conventions.')}
 if (4 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(4)
@@
 
-```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[5],fixed=TRUE),'} populations.
-See Figure SX caption for conventions.')}
+```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[5],fixed=TRUE),'} populations.
+See Figure S3 caption for conventions.')}
 if (5 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(5)
@@
 
-```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[6],fixed=TRUE),'} populations.
-See Figure SX caption for conventions.')}
+```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[6],fixed=TRUE),'} populations.
+See Figure S3 caption for conventions.')}
 if (6 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(6)
@@
 
-```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[7],fixed=TRUE),'} populations.
-See Figure SX caption for conventions.')}
+```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[7],fixed=TRUE),'} populations.
+See Figure S3 caption for conventions.')}
 if (7 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(7)
@@
 
-```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[8],fixed=TRUE),'} populations.
-See Figure SX caption for conventions.')}
+```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[8],fixed=TRUE),'} populations.
+See Figure S3 caption for conventions.')}
 if (8 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(8)
@@
 
-```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[9],fixed=TRUE),'} populations.
-See Figure SX caption for conventions.')}
+```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[9],fixed=TRUE),'} populations.
+See Figure S3 caption for conventions.')}
 if (9 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(9)
@@
 
-```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[10],fixed=TRUE),'} populations.
-See Figure SX caption for conventions.')}
+```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[10],fixed=TRUE),'} populations.
+See Figure S3 caption for conventions.')}
 if (10 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(10)
@@
 
-```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[11],fixed=TRUE),'} populations.
-See Figure SX caption for conventions.')}
+```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[11],fixed=TRUE),'} populations.
+See Figure S3 caption for conventions.')}
 if (12 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(11)
@@
 
-```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[12],fixed=TRUE),'} populations.
-See Figure SX caption for conventions.')}
+```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[12],fixed=TRUE),'} populations.
+See Figure S3 caption for conventions.')}
 if (12 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(12)
@@
 
-```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[13],fixed=TRUE),'} populations.
-See Figure SX caption for conventions.')}
+```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[13],fixed=TRUE),'} populations.
+See Figure S3 caption for conventions.')}
 if (13 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(13)
@@
 
-```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[14],fixed=TRUE),'} populations.
-See Figure SX caption for conventions.')}
+```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[14],fixed=TRUE),'} populations.
+See Figure S3 caption for conventions.')}
 if (14 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(14)
@@
 
-```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[15],fixed=TRUE),'} populations.
-See Figure SX caption for conventions.')}
+```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[15],fixed=TRUE),'} populations.
+See Figure S3 caption for conventions.')}
 if (15 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(15)
@@
 
-```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[16],fixed=TRUE),'} populations.
-See Figure SX caption for conventions.')}
+```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[16],fixed=TRUE),'} populations.
+See Figure S3 caption for conventions.')}
 if (16 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(16)
@@
 
-```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[17],fixed=TRUE),'} populations.
-See Figure SX caption for conventions.')}
+```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[17],fixed=TRUE),'} populations.
+See Figure S3 caption for conventions.')}
 if (17 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(17)
@@
 
-```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[18],fixed=TRUE),'} populations.
-See Figure SX caption for conventions.')}
+```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[18],fixed=TRUE),'} populations.
+See Figure S3 caption for conventions.')}
 if (18 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(18)
@@
 
-```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[19],fixed=TRUE),'} populations.
-See Figure SX caption for conventions.')}
+```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[19],fixed=TRUE),'} populations.
+See Figure S3 caption for conventions.')}
 if (19 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(19)
@@
 
-```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[20],fixed=TRUE),'} populations.
-See Figure SX caption for conventions.')}
+```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[20],fixed=TRUE),'} populations.
+See Figure S3 caption for conventions.')}
 if (20 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(20)
@@
 
-```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[21],fixed=TRUE),'} populations.
-See Figure SX caption for conventions.')}
+```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[21],fixed=TRUE),'} populations.
+See Figure S3 caption for conventions.')}
 if (21 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(21)
@@
 
-```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[22],fixed=TRUE),'} populations.
-See Figure SX caption for conventions.')}
+```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[22],fixed=TRUE),'} populations.
+See Figure S3 caption for conventions.')}
 if (22 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(22)
@@
 
-```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[23],fixed=TRUE),'} populations.
-See Figure SX caption for conventions.')}
+```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[23],fixed=TRUE),'} populations.
+See Figure S3 caption for conventions.')}
 if (23 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(23)
@@
 
-```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[24],fixed=TRUE),'} populations.
-See Figure SX caption for conventions.')}
+```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[24],fixed=TRUE),'} populations.
+See Figure S3 caption for conventions.')}
 if (24 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(24)
@@
 
-```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[25],fixed=TRUE),'} populations.
-See Figure SX caption for conventions.')}
+```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[25],fixed=TRUE),'} populations.
+See Figure S3 caption for conventions.')}
 if (25 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(25)
@@
 
-```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[26],fixed=TRUE),'} populations.
-See Figure SX caption for conventions.')}
+```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[26],fixed=TRUE),'} populations.
+See Figure S3 caption for conventions.')}
 if (26 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(26)
-```
-
-```{r, echo=FALSE, fig.width=3.5, fig.height=2.5, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[27],fixed=TRUE),'} populations.
-See Figure SX caption for conventions.')}
-if (27 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(27)
 ```
@@
 	# exports
-	return(((n.plots %/% 5)+1) * 3)
+	return((((n.plots-1) %/% 4)+1) * 1)
 }
@@
 
-```{r, include=FALSE}
-calc.fig.height2 <- function(x) {
-	return((((x-1) %/% 4)+1)*1)
-}
-
-calc.fig.height3 <- function(x) {
-	return((((x-1) %/% 4)+1)*1.5)
-}
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=calc.fig.height2(n_distinct(geo.surrogacy.DF$Species)), fig.cap='The relationship between the proportion of adaptive genetic variation secured in a prioritizaiton and the proportion of environmental variation it also secures.
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=3.5, fig.cap='The relationship between the proportion of adaptive genetic variation secured in a prioritizaiton and the proportion of environmental variation it also secures.
 Points represent values associated with randomly generated prioritizations.
-Black lines indicate average trends computed using generalised linear models with logit links.
-Each panel corresponds to different species.'}
+Blue lines indicate average trends computed using a generalised linear model with a logit link.
+Each panel shows data for a different species.'}
 # format data
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=calc.fig.height3(n_distinct(geo.surrogacy.DF$Species)), fig.cap='The relationship between the proportion of neutral genetic variation secured in a prioritizaiton and the proportion of geographic variation it also secures.
-Conventions are detailed in Figure SXXX.'}
+Conventions are detailed in Figure S55.'}
 # format data
@@
 		Species = paste0('\\textit{', gsub('_', ' ', Species, fixed=TRUE), '}')) %>%
+	mutate(
+		Environmental.z=replace(Environmental.z, grepl('NA', Environmental.z, fixed=TRUE),'')
+		Geographic.z=replace(Geographic.z, grepl('NA', Geographic.z, fixed=TRUE),'')
+	) %>%
 	select(Species, Environmental.z, Environmental.p.value, Environmental.r2CU,
@@
 ```{r, echo=FALSE}
-write.table(loci_classification.DF, 'article/Table_S3.csv', sep=',', row.names=FALSE)
+write.table(loci_classification.DF, 'article/Table_S4.csv', sep=',', row.names=FALSE)
 ```

### 2016-05-07 code/rmarkdown/supporting_information.Rmd
+ fix bug in SI
+ bug fixes

@@
 
+```{r, include=FALSE}
+calc.fig.height3 <- function(x) {
+	return((((x-1) %/% 4)+1)*1.5)
+}
+
+```
+
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=calc.fig.height3(n_distinct(geo.surrogacy.DF$Species)), fig.cap='The relationship between the proportion of neutral genetic variation secured in a prioritizaiton and the proportion of geographic variation it also secures.
@@
 	mutate(
-		Environmental.z=replace(Environmental.z, grepl('NA', Environmental.z, fixed=TRUE),'')
+		Environmental.z=replace(Environmental.z, grepl('NA', Environmental.z, fixed=TRUE),''),
 		Geographic.z=replace(Geographic.z, grepl('NA', Geographic.z, fixed=TRUE),'')

### 2016-05-07 code/rmarkdown/article.tex
new article files
Former-commit-id: d01d61b9aa005f5ce6841ff877818216573b21b0

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{unicode=true,
            pdftitle={Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{Are environmental and geographic variables effective surrogates for
genetic variation in conservation planning?}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Jeffrey O. Hanson\(^1\), Jonathan R. Rhodes\(^2\), Cynthia
Riginos\(^1\), Hugh P. Possingham\(^1\), Richard A. Fuller\(^1\)\\
\(^1\)School of Biological Sciences, The University of Queensland,
Brisbane, QLD, Australia\\
\(^2\)School of Geography, Planning and Environmental Management, The
University of Queensland, Brisbane, QLD, Australia\\
Correspondance should be addressed to
\href{mailto:jeffrey.hanson@uqconnect.edu.au}{\nolinkurl{jeffrey.hanson@uqconnect.edu.au}}}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{07 May 2016}

% load packages
\usepackage{amsmath,amsfonts,float,makecell,titletoc,tocloft,titlesec,natbib,pdfpages,lineno,booktabs}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[utf8]{inputenc}
\usepackage[doublespacing]{setspace}

% format captions
\usepackage[labelfont={small,bf}, labelsep=space, font={small}]{caption}

% format headings
\titleformat{\subsection}
{\normalfont\Large\bfseries}{\thesubsection}{1em}{}
\titleformat{\subsubsection}
{\normalfont\large\bfseries}{\thesubsubsection}{1em}{}
\titleformat{\paragraph}
{\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
\titleformat{\subparagraph}
{\normalfont\normalsize\itshape}{\thesubparagraph}{1em}{}

\titlespacing\paragraph{0pt}{12pt plus 4pt minus 2pt}{+0pt plus 2pt minus 2pt}
\titlespacing{\subparagraph}{0pt}{6pt plus 4pt minus 2pt}{+0pt plus 2pt minus 2pt}

% format toc
\renewcommand\cftsubsecfont{\normalfont\normalsize\bfseries}
% \renewcommand\cftsubsubsecfont{\normalfont\normalsize\bfseries}
% \renewcommand\cftparafont{\normalfont\normalsize\bfseries}
% \renewcommand\cftsubparafont{\normalfont\normalsize\itshape}

% make figures static
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
	\expandafter\origfigure\expandafter[H]
} {
	\endorigfigure
}

% define struts for tables
\newcommand\T{\rule{0pt}{2.6ex}} % top strut
\newcommand\B{\rule[-1.2ex]{0pt}{0pt}} % bottom strut

% define command to put new lines in table cells
%\newcommand{\specialcell}[2][c]{%
%  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\begin{document}
\maketitle
\begin{abstract}
Insert abstract here.
\end{abstract}

Short running title: Genetic surrogates in conservation \newline
Word count: XXXXX \newline
Number of references: XXXX \newline
Number of figures: XXXX \newline
Number of tables: XXXX \newline
Number of text boxes: XXXX \newline
\newpage

\subsection{Introduction}\label{introduction}

Protected areas spearhead conservation efforts (Sanderson \emph{et al.}
2015). These places buffer species from anthropogenic impacts and
provide places for species to persist. However, the resources available
for conservation are limited. Thus protected areas need to be sited in
places that achieve conservation goals for minimal cost. To ensure that
protected areas are cost-effective, plans for protected areas
(prioritizations) are often generated using decision support tools to
identify near-optimal solutions. Additionally, because reserves cannot
preserve the entire range for a given species, the plans for multiple
reserves are often designed simultaneously to ensure that the network of
protected areas can facilitate biological processes (eg. gene flow
between populations, seasonal migration). For instance, \texttt{Marxan}
(Ball \emph{et al.} 2009) uses amount-based targets to identify
protected area networks that preserve a suitable amount of habitat for
each species (eg. 100 km \(^2\) of land occupied by the species needs to
be secured in a reserve) and weights to penalize fragmented solutions.
However, conservation planning exercises typically assume that all
individuals in the same species are equivalent.

Intra-specific genetic variation is an important for long-term species
persistence. As a consequence, there has been increasing interest in
designing prioritizations that preserve and cultivate this variation
(Moritz 1999; Crandall \emph{et al.} 2000; Hendry \emph{et al.} 2010).
Although the strength of natural selection is a continuous force;
broadly speaking, genetic variation can be classified as either adaptive
or neutral (reviewed in Schoville \emph{et al.} 2012). Adaptive genetic
variation is associated with loci that (significantly) affect fitness.
Typically, such ``adaptive'' loci are detected due to anomalous patterns
among individuals (eg. Foll \& Gaggiotti 2008; Duforet-Frebourg \emph{et
al.} 2014; Schoville \emph{et al.} 2012; Whitlock 2015; but see Rellstab
\emph{et al.} 2015 for alternative methods). By preserving existing
patterns of adaptive variation, protected areas can ensure that
populations with particularly beneficial adaptations are not lost
(Crandall \emph{et al.} 2000). In contrast, neutral genetic variation is
associated with loci that do not (significantly) affect fitness. Neutral
variation often reflects the evolutionary history of different
populations. By preserving existing patterns of neutral variation,
protected areas can secure different genetic lineages and avoid the
adverse effects of low genetic diversity (eg. inbreeding depression;
Moritz 1999). Thus protected area networks should preserve both adaptive
and neutral patterns of genetic variation for species of conservation
interest (Moritz 1999; Crandall \emph{et al.} 2000). Despite this,
genetic data are not often used to inform prioritizations because they
require considerable investment to obtain and analyze (Hendry \emph{et
al.} 2010; but see Moritz 1994 for developments on using genetic data in
conservation planning). Thus prioritizations may fail to capture
species' intra-specific genetic variation, and ultimately fail to
deliver long-term conservation outcomes.

Recently, conservation scientists have been investigating surrogates to
generate prioritizations that adequately preserve genetic variation--but
without actually needing to utilize genetic data to do this. Since
adaptive genetic variation is ultimately driven by extrinsic selection
pressures, environmental variables have been used to guide reserve
selection (eg. environmental conditions; Pyke \& Fischer 2005; Carvalho
\emph{et al.} 2011). By preserving a more representative sampled of
environmental conditions across the species' ranges, they aimed to
generate prioritizations that preserved more adaptive variation.
However, the effectiveness of this approach remains unverified. On the
other hand, neutral genetic variation arises from a reduction in gene
flow between individuals. Surrogates for neutral variation have been
based on variables that predict the level of connectivity between
planning units. For instance, the isolation by distance theory (IBD;
Wright 1943) predicts that populations further apart will have less
connectivity, exchange less genetic material, and in turn have less
genetic variation in common. Based on this theory, geographic
distance-based surrogates have been found to improve prioritizations in
an island network (Ponce-Reyes \emph{et al.} 2014). However this remains
untested in spatially contiguous systems where connectivity is
complicated by additional factors (eg. terrain; eg. Peterman \emph{et
al.} 2014a)--the very systems commonly featured in real-world
conservation planning exercises (eg. Cowling \emph{et al.} 2003). Before
conservation planners can reliably use surrogates for genetic data in
conservation planning, the effectiveness of these surrogates needs to be
verified. Otherwise, conservation planners may waste precious resources
on protecting places that provide little benefit to biodiversity.

Here we investigate the effectiveness of environmental and geographic
surrogates for adaptive and neutral genetic variation in the context of
conservation planning. First, we aim to verify if environmental and
geographic variables correlate with broad-scale patterns in adaptive and
neutral variation (respectively). Although these variables may correlate
with genetic variation, they may not be effective surrogates if the
level of correlation is not high enough or due to spatial
auto-correlative effects. Second, we aim to verify that environmental
and geographic variables are effective surrogates for adaptive and
neutral variation (respectively). We expect that prioritizations
generated using surrogate-based targets will secure more genetic
variation than prioritizations generated using only amount-based targets
with the same number of planning units. Third, we aim to determine if
the addition of surrogate-based targets to conservation planning
exercises can result in more effective prioritizations. We expect that
prioritizations generated using amount- and surrogate-based targets will
sufficiently preserve the genetic variation of more species than those
generated using only amount-based targets. If these surrogates can
improve the intra-specific genetic variation captured in a
prioritization, then conservation planners may use these surrogates when
genetic data is not available, and in turn deliver more effective
protected areas.

\subsection{Methods}\label{methods}

\subsubsection{Study system}\label{study-system}

We used species distribution and genomic data collected by the
IntraBioDiv project in the European Alps (Meirmans \emph{et al.} 2011;
see Gugerli \emph{et al.} 2008 for further explanation of data
collection methods). This data set contains information for 27 alpine
plant species (Figure 1a). It was collected using a 20' longitude by 21'
latitude grid (approx. 22.3 km \(\times\) 25 km; Figure S1). Project
members visited every second grid cell, and if a species was detected in
a cell, samples were collected from three individuals. They genotyped
samples using amplified fragment length polymorphisms (AFLP; Vos
\emph{et al.} 1995), and constructed matrices denoting the
presence/absence of polymorphisms at loci for each species (mean 132
\(\pm\) 55.6 SD markers genotyped per species; see Table 1 in Alvarez
\emph{et al.} (2009) for the number markers for each species). Although
AFLPs are not ideal as genomic representatives (cf.~single nucleotide
polymorphisms; SNPs), no multi-species genomic data set with similar
properties currently exists. Thus the data set contains information
describing the genomic properties of individuals for each species
present in every second grid cell.

We chose this data set for several reasons. First, conservation planning
exercises typically consider tens to hundreds of species with different
evolutionary histories and life histories. To ensure that our results
were relevant for conservation planners, we required a similar sized
data set. This data set contains genetic data for a diverse range of
plant species. Second, we required genetic data for all species to be
collected under a standardized sampling scheme, and this data set used a
grid to collect samples. Third, we required broad-scale environmental
variation across the study area to test the effectiveness of
environmental variables. This data set covers a large geographic domain
that spans across attitudinal gradients. Thus this data set was
particularly well suited for our study.

This data set has previously been used to address a range of different
questions. Many studies have used this data set to explore patterns of
adaptive (Manel \emph{et al.} 2010, 2012; Meirmans \emph{et al.} 2011;
Bothwell \emph{et al.} 2013; Rolland \emph{et al.} 2015) and neutral
genetic variation (Alvarez \emph{et al.} 2009; Meirmans \emph{et al.}
2011; Taberlet \emph{et al.} 2012). However, to our knowledge, only a
single study has used this data set to explore the effectiveness of a
surrogate in the context of conservation planning. Taberlet \emph{et
al.} (2012) investigated whether prioritizations generated using the
spatial distribution of a comprehensive set species (\textgreater{} 300)
could adequately secure the intra-specific genetic variation for 27
species. They found that the prioritization based on species
distributions was substantially different to that based on genomic data.
Their results suggest that prioritizations based on species distribution
data alone will be unable to secure intra-specific genetic variation.

\subsubsection{Landscape data}\label{landscape-data}

We examined the effectiveness of environmental and geographic surrogates
for adaptive and neutral genetic variation (respectively). We used the
grid cells used to collect samples by the IntraBioDiv project as
planning units to develop prioritizations. We used only planning units
that contained at least a single individual (\(n = 149\)) to reduce
computational burden. To examine the potential effects of cost, we
calculated the total human population density inside each planning unit
(obtained at 1 km resolution from the Global Rural-Urban Mapping
Project; \texttt{GRUMP V1}; CIESEN, Columbia University \emph{et al.}
2011) and used this to represent acquisition cost (Figure 1b).

To describe the geographic location of each grid cell, we projected the
grid into an equi-distant coordinate system (\(\texttt{ESRI:102031}\)),
calculated their centroids, and extracted their two-dimensional
coordinates. Thus we constructed two-dimensional geographic
space--wherein each planning unit occupies a single point--as a
potential surrogate for neutral genetic variation.

To describe the environmental characteristics of each grid cell, we used
broad-scale climatic variables. We obtained data for 19 bioclimatic
variables across the extent of the study area (approx. 1 km resolution;
Hijmans \emph{et al.} 2005). We used this data set because it is
commonly to map habitat suitability in conservation planning (Rodríguez
\emph{et al.} 2007). These layers and the planning units were projected
into an equal-area coordinate system (\(\texttt{ESRI:102014}\)). To
reduce dimensionality, the 19 bioclimatic variables were clipped to the
planning units and subject to a principal components analysis (PCA;
using \texttt{ArcMap} version 10.2.2; Table S1). The first three
principal components (PCs) cumulatively explained 99.2 \% of the
variation and were used to generate new layers. We calculated the
average value for each PC layer in each planning unit, and used these
values to characterize climatic variation among the planning units
(Figure S2). Thus we constructed a three-dimensional environmental
space--wherein each planning unit occupies a single point--as a
potential surrogate for adaptive variation.

\subsubsection{Adaptive and neutral genetic
data}\label{adaptive-and-neutral-genetic-data}

To investigate the effectiveness of the surrogates for adaptive and
neutral genetic variation, we first needed to identify which of the
sampled loci--if any--were under selection. We used two outlier
detection methods to achieve this (as recomended by Villemereuil
\emph{et al.} 2014). We omitted the species
\textit{Ligusticum mutellinoides} from all subsequent analyses because
it was not compatible with one of the outlier detection methods (see
below). The basic premise underpinning such methods is that neutral loci
are expected to exhibit a specific level of variation among populations,
and loci that deviate from this expectation should be under selection
(reviewed in Schoville \emph{et al.} 2012). The advantage of these
methods--in contrast with environmental association analyses (Rellstab
\emph{et al.} 2015)--is that they do not use environmental data to
identify loci under selection, which would have biased our analysis.
Loci identified as outliers in both analyses were treated as adaptive
and the rest were treated as neutral. To avoid false-positives, we
omitted loci from both outlier detection methods where the global
frequency of the minor allele was \(\geq 0.1\) and assumed they were
neutral.

First, we used multinominal-Dirichlet models implemented in
\texttt{BayeScan} (version 2.1; Foll \& Gaggiotti 2008) to identify loci
under selection. We adopted a similar methodology to Bothwell \emph{et
al.} (2013) and applied it to each of the species in the data set. To
avoid falsely classifying loci as adaptive, we grouped individuals into
genetic lineages for each species (Figures S3--S28). We used the number
of genetic lineages in each species previously identified by Alvarez
\emph{et al.} (2009). Since \textit{Ligusticum mutellinoides} was
previously identified as containing only one population, it was not
compatible with \texttt{BayeScan}. For each species, we used
\texttt{Structure} (version 2.3.4; using same parameters as Alvarez
\emph{et al.} 2009 per run with 5 runs per species) and \texttt{ClumPP}
(version 1.1.2; Jakobsson \& Rosenberg 2007) to group individuals into
populations. We then ran \(\texttt{Bayescan}\) for each species using
these groupings (1:1 prior odds; 20 pilot runs; 50000 burn-in
iterations; 5000 post-burn-in iterations thinned by 10 iterations; Foll
\& Gaggiotti 2008). Additionally, we omitted individuals from this
analysis when their population membership was uncertain (maximum
membership probability \(< 0.75\)). We ran 4 replicates per species
using a suitable false discovery rate (\(q \leq 0.1\); Benjamini \&
Hochberg 1995).

Second, we used PCAs to identify outlier loci using methods implemented
in the \texttt{pcadapt R} package (version 3.0; Duforet-Frebourg
\emph{et al.} 2014; Luu \emph{et al.} 2016). To enable comparisons
between the two outlier detection analyses, we used the same individuals
in this analysis as used in the \texttt{BayeScan} analysis. For each of
the 26 species, we first imputed missing loci data to permit further
analysis. Specifically, the value for a given individual at a given
locus was replaced with the mean frequency of that locus among all
individuals. Next, we ran a PCA and extracted the minimum number of
components needed to secure 0.75 \% of the variation in loci. We then
used Mahalanobis distances (Mahalanobis 1936) to compute \emph{P} values
and then q-values (using the \texttt{qvalue R} package; Andrew J. Bass
\emph{et al.} 2015). Outlier loci were detected using a suitable FDR
(\(q < 0.1\)).

After classifying loci as adaptive or neutral, we mapped the main
gradients of the adaptive (if detected) and neutral genetic variation
for each species (Figures S29--S54). For each species, we discarded the
population groupings, and partitioned adaptive and neutral loci into
separate matrices. We used non-metric multi-dimensional scaling (NMDS;
implemented in the \texttt{vegan R} package; Oksanen \emph{et al.} 2015)
using Gower distances (Gower 1971; using the \texttt{cluster R} package
to accommodate sparsity; Maechler \emph{et al.} 2015) to derive
continuous variables that described the main gradients of adaptive and
neutral genetic variation separately for each species (Table S2). To
ensure that the ordinations described a sufficient amount of the genomic
variation, we ran successive scaling analyses with increasing
dimensionality until this was achieved (maximum stress value \(\leq\)
0.25; 100 random starts for each analysis). Since each grid cell had up
to three samples per species, we used the average of the ordinated
values associated with the individuals in each grid cell to express the
typical genomic characteristics of individuals in the cells.

The previous analyses resulted in an adaptive (if detected) and neutral
genetic space for each species. For a given species, each planning unit
occupied by the species was associated with a multi-dimensional point in
the species' neutral genetic space. Planning units that were closer
together in this space are occupied by individuals with similar neutral
genetic loci. Similarly, if the species was associated with adaptive
loci, each planning unit occupied by the species was associated with a
multi-dimensional point in the species' adaptive genetic space. By
spreading out conservation effort across a given genetic space for a
given species, prioritizations can secure more genetic variation (see
Hanson \emph{et al.} 2016 for discussion on attribute spaces).

\subsection{Prioritization method}\label{prioritization-method}

We used the unreliable representative and adequate prioritization (URAP)
formulation of the reserve selection problem (Hanson \emph{et al.}
2016). This formulation can identify the minimum number of planning
units required to preserve both a proportion of the species' range
(using amount-based targets) and a proportion of variation found across
the species' range (using space-based targets and attribute spaces;
\emph{sensu} Hanson \emph{et al.} 2016). We treated the environmental
and geographic surrogate variables as separate attribute spaces. We also
treated the adaptive (if detected) and neutral variables as separate
attribute spaces for each species. We set the demand points for each
species and each attribute space using the values associated with the
planning units. For every prioritization we generated, we calculated the
proportion of the adaptive (if detected) and neutral genetic variation
that every prioritization secured to assess its performance.
Prioritizations associated with negative values--because they secured
only a very small amount of genetic variation--were replaced with zeros
to facilitate statistical analysis. We solved all reserve selection
problems to within 10 \% of optimality using the \(\texttt{rapr}\) R
package (Hanson \emph{et al.} 2016) and \(\texttt{Gurobi}\) (Gurobi
Optimization 2015).

\subsection{Statistical analyses}\label{statistical-analyses}

To address the first aim--to determine if the environmental and
geographic variables correlate with spatial patterns of adaptive and
neutral genetic variation--we computed dissimilarity matrices between
the planning units, and tested for correlations between matrices based
on surrogate and genetic data. For each species, we computed two
dissimilarity matrices between the planning units occupied by the
species in terms of their environmental conditions (the PC values from
the bioclimatc data) and geographic position (the z-scaled XY
coordinates of the units' centroids). We then computed another
dissimilarity matrix between the planning units using their values in
the species' neutral genetic space. If the species was associated with
adaptive loci, we also computed another dissimilarity matrix between the
planning units using the species' adaptive genetic space. All
dissimilarity matrices calculated using Euclidean distances.

We then fit maximum likelihood population-effects (MLPE) models
(implemented in the \texttt{ResistanceGA R} package; Peterman \emph{et
al.} 2014b). These models explicitly accommodate the structure of
dissimilarity matrices using random effects (Clarke \emph{et al.} 2002).
For each species, we fit a MLPE model to correlate the species'
dissimilarity matrices based on geographic position and neutral genetic
variation. Additionally, if the species was associated with adaptive
loci, we also fit a MLPE model to correlate the species' dissimilarity
matrices calculated using enviuronmental conditions and adaptive genetic
variation. To test if the surrogate variables explained a substantial
amount of genetic variation, we conducted \(\chi^2\) tests between each
model and its corresponding null model, and Bonferroni corrections.

To address the second aim--to determine if the environmental and
geographic variables are effective surrogates for adaptive and neutral
genetic variation--we generated a collection of single-species
prioritizations. For each species, we generated prioritizations by
randomly selecting different combinations of planning units that were
occupied by the species. We then computed the proportion of adaptive
genetic variation and environmental variation secured in each
prioritization. We then generated another random set of prioritizations,
and computed the proportion of neutral genetic variation and geographic
variation secured in each prioritization.

We fit two full generalized linear models (GLM) with a logit links. The
first model was fit to the proportion of adaptive genetic variation
secured in a prioritization, and contained a continuous predictor
variable measuring the proportion of environmental variation also
secured in the prioritization, a categorical variable indicating the
species for which the prioritization was generated, and an interaction
term between these predictor variables. The second model was fit to the
proportion of neutral genetic variation secured in a prioritization, and
contained a continuous predictor variable assessing the proportion of
geographic variation also secured in the prioritization. Similar to the
previous model, this model also contained a predictor variable
indicating the species for which the prioritizations were generated, and
an interaction term. We subject these models to a backwards step-wise
term deletion routine to assess term significance. To assess the
performance of these surrogates, we fit a suite of models correlating
the proportion of adaptive and neutral genetic variation held in each
prioritization and the environmental and geographic variation they also
secured (respectively). We then computed the Cragg and Uhler's
pseudo-R\(^2\) value for each model (using the \texttt{pscl} R package;
Jackman 2015).

To address the third aim--to determine if the use of surrogate-based
targets can actually improve prioritizations in real-world conservation
planning scenarios--we investigated the effectiveness of surrogate-based
targets using three scenarios with increasing levels of complexity. To
assess the benefits of surrogate-based targets under the simplest of
circumstances, our first scenario involved generating single-species
prioritizations for each species. However, most large-scale
prioritizations involve a comprehensive set of species. By forcing
solutions to secure individuals in multiple communities, this may result
in prioritizations that secure a greater proportion of intra-specific
genetic variation for each species. To test for this effect, our second
scenario involved generating multi-species prioritizations. Finally,
most large-scale prioritizations also explicitly consider opportunity
cost to ensure that prioritizations are cost-effective. Our third
scenario involved generating multi-species prioritizations that
satisfied targets for minimal opportunity cost.

For each scenario, we generated three types of prioritizations using (1)
only amount-based targets, (2) both amount- and surrogate-based targets,
and (3) both amount- and genetic-based targets. The purely amount-based
prioritizations served as our baseline--representing the prioritizations
typically generated in conservation planning exercises. We used 10 \%
amount-based targets in all prioritizations to ensure an adequate
proportion of habitat was secured for each species. Based on the
previous analysis, we set surrogate-based targets as 95 \% and
genetic-based targets as 80 \% to ensure that most of the variation was
secured for each species. To accommodate variation in the high number of
optimal solutions for the single-species amount-based prioritizations,
we generated 1000 replicate prioritizations. We then counted the
proportion of species that were sufficiently represented in terms of
their adaptive and neutral genetic variation (genetic variation secured
\(\geq\) 80 \%) for each prioritization.

\subsection{Results}\label{results}

We successfully mapped the main spatial patterns of adaptive (if
present) and neutral variation for each of the 26 species. The outlier
detection methods detected adaptive genetic variation in 4 species.
Generally, two outlier detection methods gave fairly consistent results
(mean 87.26 \% loci per species 4.97 SD). Of the species that were
associated with adaptive genetic variation, only a few loci were
detected as under selection (mean 1.77 \% loci per species 1.77 SD).
After identifying the loci under selection, we used NMDS analyses to
identify the main gradients of adaptive (if detected) and neutral
variation for each species. Generally, only a couple of continuous
dimensions were needed to sufficiently describe the adaptive genetic
variation for each species (minimum \(K\) NA; mean \(K\) 2; max \(K\) 2;
Table S2). On the other hand, a few more continuous dimensions were
often needed for neutral genetic variation for each species (minimum
\(K\) 2; mean \(K\) 2.4230769; max \(K\) 4; Table S2). The ordinations
resulting from the NMDS analyses were used to construct an adaptive (if
present) and neutral genetic space for each species. The spatial
distribution of these genetic spaces (Figures S29--S54) generally show
patterns of spatial autocorrelation--especially the adaptive genetic
spaces (eg. \emph{Phyteuma hemisphaericum}; S49). Additionally, most
species show east-west patterns in their intra-specific genetic
variation (eg. \emph{Dryas octopetala}; S36).

The dissimilarity matrices generated using the surrogate variables
correlated with the dissimilarity matrices based on the genetic
characteristics of individuals for most species (Table 1). Specifically,
environmental variation was found to significantly correlate with
adaptive genetic variation for 1 species (NA \% associated with loci
under selection; \emph{P} \textless{} 0.05; mean marginal 0.02 \(R^2\)
\(\pm\) NA SD for significant models). Additionally, geographic
distances between planning units were found to significantly correlate
with neutral genetic variation among planning units for 25 species
(96.15 \% of all the species investigated;\emph{P} \textless{} 0.05;
mean marginal 0.21 \(R^2\) \(\pm\) 0.17 SD for significant models).
These results suggest that prioritizations with different environmental
characteristics tend to have individuals with different adaptive genetic
characteristics. Similarly, planning units that are further apart tend
to have individuals with different neutral genetic characteristics.
After verifying that the geographic and environmental variables
correlate with genetic variation, the next step was to determine how
effective these variables were as surrogates in the context of
conservation planning. The dissimilarity matrices generated using the
surrogate variables correlated with the dissimilarity matrices based on
the genetic characteristics of individuals for most species (Table 1).
Specifically, environmental variation was found to significantly
correlate with adaptive genetic variation for 1 species (NA \%
associated with loci under selection; \emph{P} \textless{} 0.05; mean
marginal 0.02 \(R^2\) \(\pm\) NA SD for significant models).
Additionally, geographic distances between planning units were found to
significantly correlate with neutral genetic variation among planning
units for 25 species (96.15 \% of all the species investigated;\emph{P}
\textless{} 0.05; mean marginal 0.21 \(R^2\) \(\pm\) 0.17 SD for
significant models). These results suggest that prioritizations with
different environmental characteristics tend to have individuals with
different adaptive genetic characteristics. Similarly, planning units
that are further apart tend to have individuals with different neutral
genetic characteristics. After verifying that the geographic and
environmental variables correlate with genetic variation, the next step
was to determine how effective these variables were as surrogates in the
context of conservation planning.

The environmental and geographic variables were moderately effective
surrogates for genetic variation in most species (Figure 2). The
relationship between the proportion of adaptive genetic variation
secured in a prioritization and the proportion environment variation it
also secured differed depending on the species for which the
prioritization was generated
(\(\chi^2_{4933.615872} P = 2.7275469\times 10^{-11}\)). The slope for
each species were all significantly different to zero (minimum
\(z value = 31.6562515\); maximum
\(P = r `max(env.surrogate.spp.DF\)p.value)`\$; Table S3). The
explanatory power of the environmental surrogates were also generally
quite high for most species (mean 0.5746939 \(R^2\) 0.0365087 SD; Figure
S56), with exception to \emph{Carex sempervirens}. The nature of these
relationships suggest that most of the adaptive variation for a given
species can be secured in a prioritization using moderately high
environmental-based targets. Similarly, the proportion of neutral
genetic variation in a prioritization was affected by the interaction
between species and the proportion of neutral variation secured in the
prioritization (\(\chi^2_{4.462023\times 10^{4}} P = 0\)). The slope for
each species were also significantly different to zero (minimum
\(z value = 23.3822268\); maximum \$P = 0; Table S3). Generally, the
explanatory power of geographic variables were higher than that of
environmental variables (mean 0.5495176 \(R^2\) 0.1272203 SD; Figure
S57). Compared to the environmental variables, the slopes of the
relationships suggest that very high geographic-based targets are
required to generate prioritizations that secure a sufficient proportion
of neutral genetic variation for most species. Additionally, all species
were associated with prioritizations that secured a large proportion of
environmental or geographic variation (\textgreater{} 70 \%), yet
preserved only a very small proportion of the genetic variation
(\textless{} 10 \%). These results suggest while prioritizations
generated using surrogate-based targets tend to secure more genetic
variation, prioritizations generated using just these targets can result
in poor solutions.

To determine if the addition of environmental and geographic targets
could improve the effectiveness of prioritizations in a more realistic
context, we generated prioritizations according to three different
scenarios (Figure 3). The prioritizations generated under the
single-species scenario showed relatively homogeneous selection
frequencies across the study area--regardless of the combination of
targets used to generate them (Figures 3a, 3b, 3c). On the other hand,
the prioritizations generated under the multi-species scenario showed
different selection frequencies depending on the targets used to
generate them. The prioritization generated using only amount-based
targets contained fewer planning units (Figure 3d; \(n = 14\)) than the
prioritizations generated using additional of surrogate-based targets
(Figure 3e; \(n = 33\)) or genetic-based targets (Figure 3f;
\(n = 27\)). The prioritizations generated under the multi-species with
cost scenario had different spatial distributions to the previous two
scenarios. These prioritizations tended to concentrate conservation
effort towards areas with higher elevation. Similar to the second
scenario, the amount-based prioritization contained less panning units
(Figure 3g; \(n = 14\)) than either the surrogate-based (Figure 3h;
\(n = 48\)) or the genetic-based prioritizations ((Figure 3i;
\(n = 34\))).

Generally, the prioritizations generated using amount-based targets met
the targets for adaptive genetic variation for most species associated
with adaptive loci (Figure 3). Among the three scenarios, the
amount-based prioritization in the single-species scenario had the
lowest proportion of species with targets met (\texttt{r} \% of species
with adaptive loci; Figure 3a). In the other two scenarios--the
multi-species (\texttt{r} \% species preserved with adaptive loci;
Figure 3c) and multi-species with cost (\texttt{r} \% species preserved
with adaptive loci; Figure 3e) scenarios--the amount-based
prioritizations successfully preserved the adaptive variation for most
of the species. The addition of surrogate- and genetic-based targets did
little to improve the proportion of species with their targets met for
adaptive genetic variation (Figures 3c and 3e). These results suggest
that the main gradients of intra-specific adaptive variation can be
sufficiently preserved using suitable amount-based targets.

The prioritizations generated using additional surrogate-based targets
met the neutral genetic variation targets for more species than the
prioritizations generated using just amount-based targets (Figure 3).
The biggest difference between the prioritizations generated using
additional surrogate-based targets and just the amount-based targets was
in the single-species scenarios (\texttt{r} \% and \texttt{r} \% of
species meeting genetic variation targets respectively). In the other
two scenarios, the amount-based prioritizations tended to perform
slightly better (\texttt{r} \% of species with targets met), but were
outperformed by prioritizations with additional surrogate-based targets.
Although the surrogate-based targets tended to fulfill the neutral
genetic variation targets for most species, they captured only a very
small proportion of neutral genetic variation for two species in
particular (\emph{Gypsophila\_repens} and \emph{Luzula\_alpinopilosa}).
These results suggest that surrogate-based targets can substantially
improve the proportion of neutral genetic variation secured in a
prioritization.

\subsection{Discussion}\label{discussion}

We aimed to investigate the effectiveness of broad-scale environmental
and geographic variables as surrogates for adaptive and neutral genetic
variation in conservation planning. Our study is the first to address
this hypothesis using moderately realistic planning scenarios. Overall,
we found that environmental and geographic variables are effective
surrogates for a few species when developing prioritizations.
Furthermore, when developing prioritizations under various planning
scenarios, we found that the addition of environmental and geographic
targets resulted in prioritizations that sufficiently represented the
genetic variation of a greater number of species. Our results suggest
that while the environmental and geographic variables have little
benefit for most species, they can make a large difference for some
species.

\begin{itemize}
\tightlist
\item
  Typically, conservation planning exercises do not consider
  intra-specific variation when designing prioritizations.
\item
  our results show that for most species, a sufficient proportion of
  their intra-specific variation can be secured using suitable
  amount-based targets alone.
\item
  in particular, we found that using suitable amount-based targets in a
  multi-species context with comprehensive set of target species
  resulted in a prioritization that adequately represented the adaptive
  genetic variation for nearly all the target species (panel c; Figure
  3).
\item
  however, for some species, such a prioritization secures only a small
  proportion of their genetic variation.
\item
  we found that for some species, prioritizations generated using
  amount-based targets only secured a small amount of their neutral
  genetic variation (panels b, d, e; Figure 3).
\item
  in planning exercises where the decision maker does not have access to
  genetic data on their target species across the extent of the study
  area--ie. nearly all real world planning scenarios--the decision maker
  will not know which species are poorly represented in terms of their
  inta-specific variation.
\item
  the poorly represented species may be those of particular significant
  or they may not.
\item
  the decision maker has to weigh the increased costs of a
  prioritization that includes environmental and geographic surrogates
  against the risk that a prioritization that does not include these
  surrogates fails to represent species of conservation significance.
\item
  here we found that prioritizations generated using environmental and
  geographic surrogates have only minor increased costs (compare panels
  c and d; panels e and f; Figure 3).
\item
  overall, these results suggest that although surrogates are
  unnecessary for most species, they can be highly beneficial for a few
  species, the increased costs associated with using these surrogates is
  small, so generally using surrogates is a good idea.
\end{itemize}

\begin{itemize}
\item
  Previous studies on this data set have found correlations between
  environmental and genetic variation.
\item
  Yet, for most species, prioritizations generated using environmental
  surrogates did not secure substantially more genetic variation than
  randomly generated prioritizations with the same number of planning
  units.
\item
  Here we discuss some potential reasons for this disagreement.
\item
\item
  This apparent disagreement can be attributed to multiple reasons.
\item
  These results highlight the importance of using prioritizations to
  assess the effectiveness of potential surrogates for conservation and
  not simply measuring the strength of correlation between potential
  surrogates and genetic variation.
\end{itemize}

We wish to alert our readers to several limitations associated with our
analysis. First, the size of the planning units we used (23 \(\times\)
25 km) is much larger than typically used in conservation planning
scenarios (between 1--10 km\^{}2). We used this resolution because the
genetic data was collected at this scale. Whilst we could have
interpolated the genetic data to a finer resolution and used smaller
planning units, this would have biased our analysis because this process
would have artificially introduced additional spatial auto-correlation
into the dataset. Second, we used geographic distances as surrogates for
connectivity between planning units. Although connectivity between areas
can be modeled using isolation by resistance measures (McRae 2006)
derived from landscape data (eg. Peterman \emph{et al.} 2014a), they are
unlikely to be of use in broad-scale conservation exercises since they
need to be tuned using species-specific parameters. Third, we used AFLP
data to characterize the main gradients of genetic variation. Whilst
next-generation sequencing provides higher resolution genetic
information than AFLP data (reviewed in Schoville \emph{et al.} 2012),
our methodology would still have utilized only the main gradients of the
genetic variation in order to generate prioritizations in a feasible
amount of time. Overall, we are confident that the limitations
associated with our study are relatively minor.

\subsubsection{Concluding remarks}\label{concluding-remarks}

The overarching aim of our study was to investigate the effectiveness of
surrogates in conservation planning. We found that environmental and
geographic surrogates are effective. However, they seem to be largely
unnecessary in conservation planning scenarios because a comprehensive
set of species can be used to achieve a similar result. Conservation
planners are therefore urged to use a comprehensive set of species in
prioritizations--one of the fundamental principals of conservation
planning.

\subsection{Acknowledgements}\label{acknowledgements}

JOH is funded by an Australian Postgraduate Award (APA) scholarship. RAF
has an Australian Research Council Future Fellowship. This work was
supported by the Centre of Excellence for Environmental Decisions (CEED)
and the Landscape Ecology and Conservation Group (LEC) at The University
of Queensland.

\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}

\hypertarget{refs}{}
\hypertarget{ref-r478}{}
1. Alvarez, N., Thiel-Egenter, C., Tribsch, A., Holderegger, R., Manel,
S., \& Schönswetter, P. \emph{et al.} (2009). History or ecology?
Substrate type as a major driver of patial genetic structure in alpine
plants. \emph{Ecology Letters}, 12, 632--640.

\hypertarget{ref-r491}{}
2. Andrew J. Bass, J. D. S. with contributions from, Dabney, A., \&
Robinson, D. (2015). \emph{Qvalue: Q-value estimation for false
discovery rate control}.

\hypertarget{ref-r22}{}
3. Ball, I., Possingham, H., \& Watts, M. E. (2009). Marxan and
relatives: software for spatial conservation prioritisation. In:
\emph{Spatial conservation prioritisation: Quantitative methods \&
computational tools} (eds. Moilanen, A., Wilson, K.A. \& Possingham,
H.). Book Section. Oxford University Press, Oxford, UK, pp. 185--189.

\hypertarget{ref-r466}{}
4. Benjamini, Y. \& Hochberg, Y. (1995). Controlling the false discovery
rate: A practical and powerful approach to multiple testing.
\emph{Journal of the Royal Statistical Society. Series B
(Methodological)}, 57, 289--300.

\hypertarget{ref-r157}{}
5. Bonin, A., Nicole, F., Pompanon, F., Miaud, C., \& Taberlet, P.
(2007). Population adaptive index: A new method to help measure
intraspecific genetic diversity and prioritize populations for
conservation. \emph{Conservation Biology}, 21, 697--708.

\hypertarget{ref-r455}{}
6. Bothwell, H., Bisbing, S., Therkildsen, N., Crawford, L., Alvarez,
N., \& Holderegger, R. \emph{et al.} (2013). Identifying genetic
signatures of selection in a non-model species, alpine gentian (gentiana
nivalis l.), using a landscape genetic approach. \emph{Conservation
Genetics}, 14, 467--481.

\hypertarget{ref-r2}{}
7. Carvalho, S. B., Brito, J. C., Crespo, E. J., \& Possingham, H. P.
(2011). Incorporating evolutionary processes into conservation planning
using species distribution data: A case study with the western
mediterranean herpetofauna. \emph{Diversity \& Distributions}, 17,
408--421.

\hypertarget{ref-r472}{}
8. CIESEN, Columbia University, International Food Policy Research
Institute, The World Bank, \& Centro Internacional de Agricultura
Tropical. (2011). Global rural-urban mapping project, version 1 (gRUMP
v1): Urban extents grid.

\hypertarget{ref-r494}{}
9. Clarke, R. T., Rothery, P., \& Raybould, A. F. (2002). Confidence
limits for regression relationships between distance matrices:
Estimating gene flow with distance. \emph{Journal of Agricultural,
Biological, and Environmental Statistics}, 7, 361--372.

\hypertarget{ref-r197}{}
10. Cowling, R. M., Pressey, R. L., Rouget, M., \& Lombard, A. T.
(2003). A conservation plan for a global biodiversity hotspot - the cape
floristic region, south africa. \emph{Biological Conservation}, 112,
191--216.

\hypertarget{ref-r3}{}
11. Crandall, K. A., Bininda-Emonds, O. R. P., Mace, G. M., \& Wayne, R.
K. (2000). Considering evolutionary processes in conservation biology.
\emph{Trends in Ecology \& Evolution}, 15, 290--295.

\hypertarget{ref-r487}{}
12. Duforet-Frebourg, N., Bazin, E., \& Blum, M. G. (2014). Genome scans
for detecting footprints of local adaptation using a bayesian factor
model. \emph{Molecular Biology and Evolution}, 31, 2483--2495.

\hypertarget{ref-r456}{}
13. Foll, M. \& Gaggiotti, O. (2008). A genome-scan method to identify
selected loci appropriate for both dominant and codominant markers: A
bayesian perspective. \emph{Genetics}, 180, 977--993.

\hypertarget{ref-r464}{}
14. Gower, J. C. (1971). A general coefficient of similarity and some of
its properties. \emph{Biometrics}, 27, 857--871.

\hypertarget{ref-r452}{}
15. Gugerli, F., Englisch, T., Niklfeld, H., Tribsch, A., Mirek, Z., \&
Ronikier, M. \emph{et al.} (2008). Relationships among levels of
biodiversity and the relevance of intraspecific diversity in
conservation - a project synopsis. \emph{Perspectives in Plant Ecology,
Evolution and Systematics}, 10, 259--281.

\hypertarget{ref-r463}{}
16. Gurobi Optimization, I. (2015). Gurobi optimizer reference manual.

\hypertarget{ref-r465}{}
17. Hanson, J. O., Rhodes, J. R., Possingham, H. P., \& Fuller, R. A.
(2016). RAPR: representative and adequate prioritisations in R.
\emph{Unpublished}, X, XX--XX.

\hypertarget{ref-r61}{}
18. Hendry, A. P., Lohmann, L. G., Conti, E., Cracraft, J., Crandall, K.
A., \& Faith, D. P. \emph{et al.} (2010). Evolutionary biology in
biodiversity science, conservation, and policy: A call to action.
\emph{Evolution}, 64, 1517--1528.

\hypertarget{ref-r33}{}
19. Hijmans, R. J., Cameron, S. E., Parra, J. L., Jones, P. G., \&
Jarvis, A. (2005). Very high resolution interpolated climate surfaces
for global land areas. \emph{International Journal of Climatology}, 25,
1965--1978.

\hypertarget{ref-r492}{}
20. Jackman, S. (2015). \emph{pscl: Classes and methods for R developed
in the political science computational laboratory, stanford university}.
Department of Political Science, Stanford University, Stanford,
California.

\hypertarget{ref-r459}{}
21. Jakobsson, M. \& Rosenberg, N. A. (2007). CLUMPP: A cluster matching
and permutation program for dealing with label switching and
multimodality in analysis of population structure.
\emph{Bioinformatics}, 23, 1801--1806.

\hypertarget{ref-r496}{}
22. Luu, K., Blum, M. G., \& Duforet-Frebourg, N. (2016). \emph{Pcadapt:
Fast principal component analysis for outlier detection}.

\hypertarget{ref-r457}{}
23. Maechler, M., Rousseeuw, P., Struyf, A., Hubert, M., \& Hornik, K.
(2015). \emph{Cluster: Cluster analysis basics and extensions}.

\hypertarget{ref-r398}{}
24. Mahalanobis, P. C. (1936). On the generalized distance in
statistics. \emph{Proceedings of the National Institute of Sciences,
Calcutta}, 2, 49--55.

\hypertarget{ref-r475}{}
25. Manel, S., Gugerli, F., Thuiller, W., Alvarez, N., Legendre, P., \&
Holderegger, R. \emph{et al.} (2012). Broad-scale adaptive genetic
variation in alpine plants is driven by temperature and precipitation.
\emph{Molecular Ecology}, 21, 3729--3738.

\hypertarget{ref-r484}{}
26. Manel, S., Poncet, B. N., Legendre, P., Gugerli, F., \& Holderegger,
R. (2010). Common factors drive adaptive genetic variation at different
spatial scales in arabis alpina. \emph{Molecular Ecology}, 19,
3824--3835.

\hypertarget{ref-r473}{}
27. McRae, B. H. (2006). Isolation by resistance. \emph{Evolution}, 60,
1551--1561.

\hypertarget{ref-r451}{}
28. Meirmans, P., Goudet, J., IntraBioDiv Consortium, \& Gaggiotti, O.
(2011). Ecology and life history affect different aspects of the
population structure of 27 high-alpine plants. \emph{Molecular Ecology},
20, 3144--3155.

\hypertarget{ref-r5}{}
29. Moritz, C. (1994). Defining evolutionarily significant units for
conservation. \emph{Trends in Ecology \& Evolution}, 9, 373--375.

\hypertarget{ref-r67}{}
30. Moritz, C. (1999). Conservation units and translocations: Strategies
for conserving evolutionary processes. \emph{Hereditas}, 130, 217--228.

\hypertarget{ref-r458}{}
31. Oksanen, J., Blanchet, F. G., Kindt, R., Legendre, P., Minchin, P.
R., \& O'Hara, R. B. \emph{et al.} (2015). \emph{Vegan: Community
ecology package}.

\hypertarget{ref-r474}{}
32. Peterman, W. E., Anderson, T. L., Ousterhout, B. H., Drake, D. L.,
Semlitsch, R. D., \& Eggert, L. S. (2014a). Differential dispersal
shapes population structure and patterns of genetic differentiation in
two sympatric pond breeding salamanders. \emph{Conservation Genetics},
16, 59--69.

\hypertarget{ref-r488}{}
33. Peterman, W. E., Connette, G. M., Semlitsch, R. D., \& Eggert, L. S.
(2014b). Ecological resistance surfaces predict fine-scale genetic
differentiation in a terrestrial woodland salamander. \emph{Molecular
Ecology}, 23, 2402--2413.

\hypertarget{ref-r200}{}
34. Ponce-Reyes, R., Clegg, S. M., Carvalho, S. B., McDonald-Madden, E.,
\& Possingham, H. P. (2014). Geographical surrogates of genetic
variation for selecting island populations for conservation.
\emph{Diversity \& Distributions}, 20, 640--651.

\hypertarget{ref-r247}{}
35. Pyke, C. R. \& Fischer, D. T. (2005). Selection of bioclimatically
representative biological reserve systems under climate change.
\emph{Biological Conservation}, 121, 429--441.

\hypertarget{ref-r489}{}
36. Rellstab, C., Gugerli, F., Eckert, A. J., Hancock, A. M., \&
Holderegger, R. (2015). A practical guide to environmental association
analysis in landscape genomics. \emph{Molecular Ecology}, 24,
4348--4370.

\hypertarget{ref-r493}{}
37. Rodríguez, J. P., Brotons, L., Bustamante, J., \& Seoane, J. (2007).
The application of predictive modelling of species distribution to
biodiversity conservation. \emph{Diversity and Distributions}, 13,
243--251.

\hypertarget{ref-r485}{}
38. Rolland, J., Lavergne, S., \& Manel, S. (2015). Combining niche
modelling and landscape genetics to study local adaptation: A novel
approach illustrated using alpine plants. \emph{Perspectives in Plant
Ecology, Evolution and Systematics}, 17, 491--499.

\hypertarget{ref-r440}{}
39. Sanderson, E. W., Segan, D. B., \& Watson, J. E. M. (2015). Global
status of and prospects for protection of terrestrial geophysical
diversity. \emph{Conservation Biology}, 29, 649--656.

\hypertarget{ref-r480}{}
40. Schoville, S. D., Bonin, A., Francois, O., Lobreaux, S., Melodelima,
C., \& Manel, S. (2012). Adaptive genetic variation on the landscape:
Methods and cases. In: \emph{Annaul review of ecology, evolution, and
systematics}, Annual review of ecology evolution and systematics (ed.
Futuyma, D.). pp. 23--43.

\hypertarget{ref-r486}{}
41. Taberlet, P., Zimmermann, N. E., Englisch, T., Tribsch, A.,
Holderegger, R., \& Alvarez, N. \emph{et al.} (2012). Genetic diversity
in widespread species is not congruent with species richness in alpine
plant communities. \emph{Ecology Letters}, 15, 1439--1448.

\hypertarget{ref-r495}{}
42. Villemereuil, P. de, Frichot, É., Bazin, É., François, O., \&
Gaggiotti, O. E. (2014). Genome scan methods against more complex
models: When and how much should we trust them? \emph{Molecular
Ecology}, 23, 2006--2019.

\hypertarget{ref-r453}{}
43. Vos, P., Hogers, R., Bleeker, M., Reijans, M., Lee, T. van de, \&
Hornes, M. \emph{et al.} (1995). AFLP: A new technique for dNA
fingerprinting. \emph{Nucleic Acids Research}, 23, 4407--4414.

\hypertarget{ref-r490}{}
44. Whitlock, M. C. (2015). Modern approaches to local adaptation.
\emph{The American Naturalist}, 186, S1--S4.

\hypertarget{ref-r259}{}
45. Wright, S. (1943). Isolation by distance. \emph{Genetics}, 28,
114--138.

% \includepdf[pages=-, pagecommand={}]{figures.pdf}
% \includepdf[pages=-, pagecommand={}]{supporting_information.pdf}

\end{document}

### 2016-05-07 code/rmarkdown/figures.tex
new article files
Former-commit-id: d01d61b9aa005f5ce6841ff877818216573b21b0

--- 2016-05-07 code/rmarkdown/article.tex
@@
 \hypersetup{unicode=true,
-            pdftitle={Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?},
             pdfborder={0 0 0},
@@
 \setlength{\droptitle}{-2em}
-  \title{Are environmental and geographic variables effective surrogates for
-genetic variation in conservation planning?}
-  \pretitle{\vspace{\droptitle}\centering\huge}
-  \posttitle{\par}
-  \author{Jeffrey O. Hanson\(^1\), Jonathan R. Rhodes\(^2\), Cynthia
-Riginos\(^1\), Hugh P. Possingham\(^1\), Richard A. Fuller\(^1\)\\
-\(^1\)School of Biological Sciences, The University of Queensland,
-Brisbane, QLD, Australia\\
-\(^2\)School of Geography, Planning and Environmental Management, The
-University of Queensland, Brisbane, QLD, Australia\\
-Correspondance should be addressed to
-\href{mailto:jeffrey.hanson@uqconnect.edu.au}{\nolinkurl{jeffrey.hanson@uqconnect.edu.au}}}
-  \preauthor{\centering\large\emph}
-  \postauthor{\par}
-  \predate{\centering\large\emph}
-  \postdate{\par}
-  \date{07 May 2016}
+  \title{}
+  \pretitle{\vspace{\droptitle}}
+  \posttitle{}
+  \author{}
+  \preauthor{}\postauthor{}
+  \date{}
+  \predate{}\postdate{}
 
@@
 %  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
+% insert S before figure and table caption
+\pagenumbering{gobble}
 
@@
 \begin{document}
-\maketitle
-\begin{abstract}
-Insert abstract here.
-\end{abstract}
 
-Short running title: Genetic surrogates in conservation \newline
-Word count: XXXXX \newline
-Number of references: XXXX \newline
-Number of figures: XXXX \newline
-Number of tables: XXXX \newline
-Number of text boxes: XXXX \newline
-\newpage
+\subsection{Figures}\label{figures}
 
-\subsection{Introduction}\label{introduction}
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures_files/figure-latex/unnamed-chunk-3-1.pdf}
+\caption{Map of the study area. Squares denote planning units. Panel (a)
+depicts patterns of species riches. Planning units with a brighter color
+are inhabited by more species. Panel (b) shows the distribution of
+opportunity cost across the study area (computed as total population
+density). Note that a quantile-based color ramp is used in this panel.}
+\end{figure}
 
-Protected areas spearhead conservation efforts (Sanderson \emph{et al.}
-2015). These places buffer species from anthropogenic impacts and
-provide places for species to persist. However, the resources available
-for conservation are limited. Thus protected areas need to be sited in
-places that achieve conservation goals for minimal cost. To ensure that
-protected areas are cost-effective, plans for protected areas
-(prioritizations) are often generated using decision support tools to
-identify near-optimal solutions. Additionally, because reserves cannot
-preserve the entire range for a given species, the plans for multiple
-reserves are often designed simultaneously to ensure that the network of
-protected areas can facilitate biological processes (eg. gene flow
-between populations, seasonal migration). For instance, \texttt{Marxan}
-(Ball \emph{et al.} 2009) uses amount-based targets to identify
-protected area networks that preserve a suitable amount of habitat for
-each species (eg. 100 km \(^2\) of land occupied by the species needs to
-be secured in a reserve) and weights to penalize fragmented solutions.
-However, conservation planning exercises typically assume that all
-individuals in the same species are equivalent.
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures_files/figure-latex/unnamed-chunk-4-1.pdf}
+\caption{The relationship between surrogates and genetic variation
+secured in prioritizations. Panel (a) shows the relationship between the
+proportion of intra-specific adaptive genetic variation and
+environmental variation secured in a prioritization. Similarly, panel
+(b) shows the relationship between the proportion of intra-specific
+neutral variaiton and geograpgic variation secured in a prioritization.
+Panels (c) and (d) show the distribution of \(R^2\) values for these
+models.}
+\end{figure}
 
-Intra-specific genetic variation is an important for long-term species
-persistence. As a consequence, there has been increasing interest in
-designing prioritizations that preserve and cultivate this variation
-(Moritz 1999; Crandall \emph{et al.} 2000; Hendry \emph{et al.} 2010).
-Although the strength of natural selection is a continuous force;
-broadly speaking, genetic variation can be classified as either adaptive
-or neutral (reviewed in Schoville \emph{et al.} 2012). Adaptive genetic
-variation is associated with loci that (significantly) affect fitness.
-Typically, such ``adaptive'' loci are detected due to anomalous patterns
-among individuals (eg. Foll \& Gaggiotti 2008; Duforet-Frebourg \emph{et
-al.} 2014; Schoville \emph{et al.} 2012; Whitlock 2015; but see Rellstab
-\emph{et al.} 2015 for alternative methods). By preserving existing
-patterns of adaptive variation, protected areas can ensure that
-populations with particularly beneficial adaptations are not lost
-(Crandall \emph{et al.} 2000). In contrast, neutral genetic variation is
-associated with loci that do not (significantly) affect fitness. Neutral
-variation often reflects the evolutionary history of different
-populations. By preserving existing patterns of neutral variation,
-protected areas can secure different genetic lineages and avoid the
-adverse effects of low genetic diversity (eg. inbreeding depression;
-Moritz 1999). Thus protected area networks should preserve both adaptive
-and neutral patterns of genetic variation for species of conservation
-interest (Moritz 1999; Crandall \emph{et al.} 2000). Despite this,
-genetic data are not often used to inform prioritizations because they
-require considerable investment to obtain and analyze (Hendry \emph{et
-al.} 2010; but see Moritz 1994 for developments on using genetic data in
-conservation planning). Thus prioritizations may fail to capture
-species' intra-specific genetic variation, and ultimately fail to
-deliver long-term conservation outcomes.
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures_files/figure-latex/unnamed-chunk-5-1.pdf}
+\caption{Planning unit selection frequencies. The top row of panels (a,
+b, c) show single-species prioritizations generated assuming equal
+costs. The middle row of panels (d, e, f) show multi-species
+prioritizations generated assuming equal costs. The bottom row of panels
+(g, h, i) show multi-species generated using opportunity costs. The left
+column of panels (a, d, g) show prioritizations generated using only
+amount-based targets. The middle column of panels (b, e, h) show
+prioritizations generated using amount- and surrogate-based targets. The
+right column of panels (c, f, i) show prioritizations generated using
+amount- and genetic-based targets.}
+\end{figure}
 
-Recently, conservation scientists have been investigating surrogates to
-generate prioritizations that adequately preserve genetic variation--but
-without actually needing to utilize genetic data to do this. Since
-adaptive genetic variation is ultimately driven by extrinsic selection
-pressures, environmental variables have been used to guide reserve
-selection (eg. environmental conditions; Pyke \& Fischer 2005; Carvalho
-\emph{et al.} 2011). By preserving a more representative sampled of
-environmental conditions across the species' ranges, they aimed to
-generate prioritizations that preserved more adaptive variation.
-However, the effectiveness of this approach remains unverified. On the
-other hand, neutral genetic variation arises from a reduction in gene
-flow between individuals. Surrogates for neutral variation have been
-based on variables that predict the level of connectivity between
-planning units. For instance, the isolation by distance theory (IBD;
-Wright 1943) predicts that populations further apart will have less
-connectivity, exchange less genetic material, and in turn have less
-genetic variation in common. Based on this theory, geographic
-distance-based surrogates have been found to improve prioritizations in
-an island network (Ponce-Reyes \emph{et al.} 2014). However this remains
-untested in spatially contiguous systems where connectivity is
-complicated by additional factors (eg. terrain; eg. Peterman \emph{et
-al.} 2014a)--the very systems commonly featured in real-world
-conservation planning exercises (eg. Cowling \emph{et al.} 2003). Before
-conservation planners can reliably use surrogates for genetic data in
-conservation planning, the effectiveness of these surrogates needs to be
-verified. Otherwise, conservation planners may waste precious resources
-on protecting places that provide little benefit to biodiversity.
-
-Here we investigate the effectiveness of environmental and geographic
-surrogates for adaptive and neutral genetic variation in the context of
-conservation planning. First, we aim to verify if environmental and
-geographic variables correlate with broad-scale patterns in adaptive and
-neutral variation (respectively). Although these variables may correlate
-with genetic variation, they may not be effective surrogates if the
-level of correlation is not high enough or due to spatial
-auto-correlative effects. Second, we aim to verify that environmental
-and geographic variables are effective surrogates for adaptive and
-neutral variation (respectively). We expect that prioritizations
-generated using surrogate-based targets will secure more genetic
-variation than prioritizations generated using only amount-based targets
-with the same number of planning units. Third, we aim to determine if
-the addition of surrogate-based targets to conservation planning
-exercises can result in more effective prioritizations. We expect that
-prioritizations generated using amount- and surrogate-based targets will
-sufficiently preserve the genetic variation of more species than those
-generated using only amount-based targets. If these surrogates can
-improve the intra-specific genetic variation captured in a
-prioritization, then conservation planners may use these surrogates when
-genetic data is not available, and in turn deliver more effective
-protected areas.
-
-\subsection{Methods}\label{methods}
-
-\subsubsection{Study system}\label{study-system}
-
-We used species distribution and genomic data collected by the
-IntraBioDiv project in the European Alps (Meirmans \emph{et al.} 2011;
-see Gugerli \emph{et al.} 2008 for further explanation of data
-collection methods). This data set contains information for 27 alpine
-plant species (Figure 1a). It was collected using a 20' longitude by 21'
-latitude grid (approx. 22.3 km \(\times\) 25 km; Figure S1). Project
-members visited every second grid cell, and if a species was detected in
-a cell, samples were collected from three individuals. They genotyped
-samples using amplified fragment length polymorphisms (AFLP; Vos
-\emph{et al.} 1995), and constructed matrices denoting the
-presence/absence of polymorphisms at loci for each species (mean 132
-\(\pm\) 55.6 SD markers genotyped per species; see Table 1 in Alvarez
-\emph{et al.} (2009) for the number markers for each species). Although
-AFLPs are not ideal as genomic representatives (cf.~single nucleotide
-polymorphisms; SNPs), no multi-species genomic data set with similar
-properties currently exists. Thus the data set contains information
-describing the genomic properties of individuals for each species
-present in every second grid cell.
-
-We chose this data set for several reasons. First, conservation planning
-exercises typically consider tens to hundreds of species with different
-evolutionary histories and life histories. To ensure that our results
-were relevant for conservation planners, we required a similar sized
-data set. This data set contains genetic data for a diverse range of
-plant species. Second, we required genetic data for all species to be
-collected under a standardized sampling scheme, and this data set used a
-grid to collect samples. Third, we required broad-scale environmental
-variation across the study area to test the effectiveness of
-environmental variables. This data set covers a large geographic domain
-that spans across attitudinal gradients. Thus this data set was
-particularly well suited for our study.
-
-This data set has previously been used to address a range of different
-questions. Many studies have used this data set to explore patterns of
-adaptive (Manel \emph{et al.} 2010, 2012; Meirmans \emph{et al.} 2011;
-Bothwell \emph{et al.} 2013; Rolland \emph{et al.} 2015) and neutral
-genetic variation (Alvarez \emph{et al.} 2009; Meirmans \emph{et al.}
-2011; Taberlet \emph{et al.} 2012). However, to our knowledge, only a
-single study has used this data set to explore the effectiveness of a
-surrogate in the context of conservation planning. Taberlet \emph{et
-al.} (2012) investigated whether prioritizations generated using the
-spatial distribution of a comprehensive set species (\textgreater{} 300)
-could adequately secure the intra-specific genetic variation for 27
-species. They found that the prioritization based on species
-distributions was substantially different to that based on genomic data.
-Their results suggest that prioritizations based on species distribution
-data alone will be unable to secure intra-specific genetic variation.
-
-\subsubsection{Landscape data}\label{landscape-data}
-
-We examined the effectiveness of environmental and geographic surrogates
-for adaptive and neutral genetic variation (respectively). We used the
-grid cells used to collect samples by the IntraBioDiv project as
-planning units to develop prioritizations. We used only planning units
-that contained at least a single individual (\(n = 149\)) to reduce
-computational burden. To examine the potential effects of cost, we
-calculated the total human population density inside each planning unit
-(obtained at 1 km resolution from the Global Rural-Urban Mapping
-Project; \texttt{GRUMP V1}; CIESEN, Columbia University \emph{et al.}
-2011) and used this to represent acquisition cost (Figure 1b).
-
-To describe the geographic location of each grid cell, we projected the
-grid into an equi-distant coordinate system (\(\texttt{ESRI:102031}\)),
-calculated their centroids, and extracted their two-dimensional
-coordinates. Thus we constructed two-dimensional geographic
-space--wherein each planning unit occupies a single point--as a
-potential surrogate for neutral genetic variation.
-
-To describe the environmental characteristics of each grid cell, we used
-broad-scale climatic variables. We obtained data for 19 bioclimatic
-variables across the extent of the study area (approx. 1 km resolution;
-Hijmans \emph{et al.} 2005). We used this data set because it is
-commonly to map habitat suitability in conservation planning (Rodríguez
-\emph{et al.} 2007). These layers and the planning units were projected
-into an equal-area coordinate system (\(\texttt{ESRI:102014}\)). To
-reduce dimensionality, the 19 bioclimatic variables were clipped to the
-planning units and subject to a principal components analysis (PCA;
-using \texttt{ArcMap} version 10.2.2; Table S1). The first three
-principal components (PCs) cumulatively explained 99.2 \% of the
-variation and were used to generate new layers. We calculated the
-average value for each PC layer in each planning unit, and used these
-values to characterize climatic variation among the planning units
-(Figure S2). Thus we constructed a three-dimensional environmental
-space--wherein each planning unit occupies a single point--as a
-potential surrogate for adaptive variation.
-
-\subsubsection{Adaptive and neutral genetic
-data}\label{adaptive-and-neutral-genetic-data}
-
-To investigate the effectiveness of the surrogates for adaptive and
-neutral genetic variation, we first needed to identify which of the
-sampled loci--if any--were under selection. We used two outlier
-detection methods to achieve this (as recomended by Villemereuil
-\emph{et al.} 2014). We omitted the species
-\textit{Ligusticum mutellinoides} from all subsequent analyses because
-it was not compatible with one of the outlier detection methods (see
-below). The basic premise underpinning such methods is that neutral loci
-are expected to exhibit a specific level of variation among populations,
-and loci that deviate from this expectation should be under selection
-(reviewed in Schoville \emph{et al.} 2012). The advantage of these
-methods--in contrast with environmental association analyses (Rellstab
-\emph{et al.} 2015)--is that they do not use environmental data to
-identify loci under selection, which would have biased our analysis.
-Loci identified as outliers in both analyses were treated as adaptive
-and the rest were treated as neutral. To avoid false-positives, we
-omitted loci from both outlier detection methods where the global
-frequency of the minor allele was \(\geq 0.1\) and assumed they were
-neutral.
-
-First, we used multinominal-Dirichlet models implemented in
-\texttt{BayeScan} (version 2.1; Foll \& Gaggiotti 2008) to identify loci
-under selection. We adopted a similar methodology to Bothwell \emph{et
-al.} (2013) and applied it to each of the species in the data set. To
-avoid falsely classifying loci as adaptive, we grouped individuals into
-genetic lineages for each species (Figures S3--S28). We used the number
-of genetic lineages in each species previously identified by Alvarez
-\emph{et al.} (2009). Since \textit{Ligusticum mutellinoides} was
-previously identified as containing only one population, it was not
-compatible with \texttt{BayeScan}. For each species, we used
-\texttt{Structure} (version 2.3.4; using same parameters as Alvarez
-\emph{et al.} 2009 per run with 5 runs per species) and \texttt{ClumPP}
-(version 1.1.2; Jakobsson \& Rosenberg 2007) to group individuals into
-populations. We then ran \(\texttt{Bayescan}\) for each species using
-these groupings (1:1 prior odds; 20 pilot runs; 50000 burn-in
-iterations; 5000 post-burn-in iterations thinned by 10 iterations; Foll
-\& Gaggiotti 2008). Additionally, we omitted individuals from this
-analysis when their population membership was uncertain (maximum
-membership probability \(< 0.75\)). We ran 4 replicates per species
-using a suitable false discovery rate (\(q \leq 0.1\); Benjamini \&
-Hochberg 1995).
-
-Second, we used PCAs to identify outlier loci using methods implemented
-in the \texttt{pcadapt R} package (version 3.0; Duforet-Frebourg
-\emph{et al.} 2014; Luu \emph{et al.} 2016). To enable comparisons
-between the two outlier detection analyses, we used the same individuals
-in this analysis as used in the \texttt{BayeScan} analysis. For each of
-the 26 species, we first imputed missing loci data to permit further
-analysis. Specifically, the value for a given individual at a given
-locus was replaced with the mean frequency of that locus among all
-individuals. Next, we ran a PCA and extracted the minimum number of
-components needed to secure 0.75 \% of the variation in loci. We then
-used Mahalanobis distances (Mahalanobis 1936) to compute \emph{P} values
-and then q-values (using the \texttt{qvalue R} package; Andrew J. Bass
-\emph{et al.} 2015). Outlier loci were detected using a suitable FDR
-(\(q < 0.1\)).
-
-After classifying loci as adaptive or neutral, we mapped the main
-gradients of the adaptive (if detected) and neutral genetic variation
-for each species (Figures S29--S54). For each species, we discarded the
-population groupings, and partitioned adaptive and neutral loci into
-separate matrices. We used non-metric multi-dimensional scaling (NMDS;
-implemented in the \texttt{vegan R} package; Oksanen \emph{et al.} 2015)
-using Gower distances (Gower 1971; using the \texttt{cluster R} package
-to accommodate sparsity; Maechler \emph{et al.} 2015) to derive
-continuous variables that described the main gradients of adaptive and
-neutral genetic variation separately for each species (Table S2). To
-ensure that the ordinations described a sufficient amount of the genomic
-variation, we ran successive scaling analyses with increasing
-dimensionality until this was achieved (maximum stress value \(\leq\)
-0.25; 100 random starts for each analysis). Since each grid cell had up
-to three samples per species, we used the average of the ordinated
-values associated with the individuals in each grid cell to express the
-typical genomic characteristics of individuals in the cells.
-
-The previous analyses resulted in an adaptive (if detected) and neutral
-genetic space for each species. For a given species, each planning unit
-occupied by the species was associated with a multi-dimensional point in
-the species' neutral genetic space. Planning units that were closer
-together in this space are occupied by individuals with similar neutral
-genetic loci. Similarly, if the species was associated with adaptive
-loci, each planning unit occupied by the species was associated with a
-multi-dimensional point in the species' adaptive genetic space. By
-spreading out conservation effort across a given genetic space for a
-given species, prioritizations can secure more genetic variation (see
-Hanson \emph{et al.} 2016 for discussion on attribute spaces).
-
-\subsection{Prioritization method}\label{prioritization-method}
-
-We used the unreliable representative and adequate prioritization (URAP)
-formulation of the reserve selection problem (Hanson \emph{et al.}
-2016). This formulation can identify the minimum number of planning
-units required to preserve both a proportion of the species' range
-(using amount-based targets) and a proportion of variation found across
-the species' range (using space-based targets and attribute spaces;
-\emph{sensu} Hanson \emph{et al.} 2016). We treated the environmental
-and geographic surrogate variables as separate attribute spaces. We also
-treated the adaptive (if detected) and neutral variables as separate
-attribute spaces for each species. We set the demand points for each
-species and each attribute space using the values associated with the
-planning units. For every prioritization we generated, we calculated the
-proportion of the adaptive (if detected) and neutral genetic variation
-that every prioritization secured to assess its performance.
-Prioritizations associated with negative values--because they secured
-only a very small amount of genetic variation--were replaced with zeros
-to facilitate statistical analysis. We solved all reserve selection
-problems to within 10 \% of optimality using the \(\texttt{rapr}\) R
-package (Hanson \emph{et al.} 2016) and \(\texttt{Gurobi}\) (Gurobi
-Optimization 2015).
-
-\subsection{Statistical analyses}\label{statistical-analyses}
-
-To address the first aim--to determine if the environmental and
-geographic variables correlate with spatial patterns of adaptive and
-neutral genetic variation--we computed dissimilarity matrices between
-the planning units, and tested for correlations between matrices based
-on surrogate and genetic data. For each species, we computed two
-dissimilarity matrices between the planning units occupied by the
-species in terms of their environmental conditions (the PC values from
-the bioclimatc data) and geographic position (the z-scaled XY
-coordinates of the units' centroids). We then computed another
-dissimilarity matrix between the planning units using their values in
-the species' neutral genetic space. If the species was associated with
-adaptive loci, we also computed another dissimilarity matrix between the
-planning units using the species' adaptive genetic space. All
-dissimilarity matrices calculated using Euclidean distances.
-
-We then fit maximum likelihood population-effects (MLPE) models
-(implemented in the \texttt{ResistanceGA R} package; Peterman \emph{et
-al.} 2014b). These models explicitly accommodate the structure of
-dissimilarity matrices using random effects (Clarke \emph{et al.} 2002).
-For each species, we fit a MLPE model to correlate the species'
-dissimilarity matrices based on geographic position and neutral genetic
-variation. Additionally, if the species was associated with adaptive
-loci, we also fit a MLPE model to correlate the species' dissimilarity
-matrices calculated using enviuronmental conditions and adaptive genetic
-variation. To test if the surrogate variables explained a substantial
-amount of genetic variation, we conducted \(\chi^2\) tests between each
-model and its corresponding null model, and Bonferroni corrections.
-
-To address the second aim--to determine if the environmental and
-geographic variables are effective surrogates for adaptive and neutral
-genetic variation--we generated a collection of single-species
-prioritizations. For each species, we generated prioritizations by
-randomly selecting different combinations of planning units that were
-occupied by the species. We then computed the proportion of adaptive
-genetic variation and environmental variation secured in each
-prioritization. We then generated another random set of prioritizations,
-and computed the proportion of neutral genetic variation and geographic
-variation secured in each prioritization.
-
-We fit two full generalized linear models (GLM) with a logit links. The
-first model was fit to the proportion of adaptive genetic variation
-secured in a prioritization, and contained a continuous predictor
-variable measuring the proportion of environmental variation also
-secured in the prioritization, a categorical variable indicating the
-species for which the prioritization was generated, and an interaction
-term between these predictor variables. The second model was fit to the
-proportion of neutral genetic variation secured in a prioritization, and
-contained a continuous predictor variable assessing the proportion of
-geographic variation also secured in the prioritization. Similar to the
-previous model, this model also contained a predictor variable
-indicating the species for which the prioritizations were generated, and
-an interaction term. We subject these models to a backwards step-wise
-term deletion routine to assess term significance. To assess the
-performance of these surrogates, we fit a suite of models correlating
-the proportion of adaptive and neutral genetic variation held in each
-prioritization and the environmental and geographic variation they also
-secured (respectively). We then computed the Cragg and Uhler's
-pseudo-R\(^2\) value for each model (using the \texttt{pscl} R package;
-Jackman 2015).
-
-To address the third aim--to determine if the use of surrogate-based
-targets can actually improve prioritizations in real-world conservation
-planning scenarios--we investigated the effectiveness of surrogate-based
-targets using three scenarios with increasing levels of complexity. To
-assess the benefits of surrogate-based targets under the simplest of
-circumstances, our first scenario involved generating single-species
-prioritizations for each species. However, most large-scale
-prioritizations involve a comprehensive set of species. By forcing
-solutions to secure individuals in multiple communities, this may result
-in prioritizations that secure a greater proportion of intra-specific
-genetic variation for each species. To test for this effect, our second
-scenario involved generating multi-species prioritizations. Finally,
-most large-scale prioritizations also explicitly consider opportunity
-cost to ensure that prioritizations are cost-effective. Our third
-scenario involved generating multi-species prioritizations that
-satisfied targets for minimal opportunity cost.
-
-For each scenario, we generated three types of prioritizations using (1)
-only amount-based targets, (2) both amount- and surrogate-based targets,
-and (3) both amount- and genetic-based targets. The purely amount-based
-prioritizations served as our baseline--representing the prioritizations
-typically generated in conservation planning exercises. We used 10 \%
-amount-based targets in all prioritizations to ensure an adequate
-proportion of habitat was secured for each species. Based on the
-previous analysis, we set surrogate-based targets as 95 \% and
-genetic-based targets as 80 \% to ensure that most of the variation was
-secured for each species. To accommodate variation in the high number of
-optimal solutions for the single-species amount-based prioritizations,
-we generated 1000 replicate prioritizations. We then counted the
-proportion of species that were sufficiently represented in terms of
-their adaptive and neutral genetic variation (genetic variation secured
-\(\geq\) 80 \%) for each prioritization.
-
-\subsection{Results}\label{results}
-
-We successfully mapped the main spatial patterns of adaptive (if
-present) and neutral variation for each of the 26 species. The outlier
-detection methods detected adaptive genetic variation in 4 species.
-Generally, two outlier detection methods gave fairly consistent results
-(mean 87.26 \% loci per species 4.97 SD). Of the species that were
-associated with adaptive genetic variation, only a few loci were
-detected as under selection (mean 1.77 \% loci per species 1.77 SD).
-After identifying the loci under selection, we used NMDS analyses to
-identify the main gradients of adaptive (if detected) and neutral
-variation for each species. Generally, only a couple of continuous
-dimensions were needed to sufficiently describe the adaptive genetic
-variation for each species (minimum \(K\) NA; mean \(K\) 2; max \(K\) 2;
-Table S2). On the other hand, a few more continuous dimensions were
-often needed for neutral genetic variation for each species (minimum
-\(K\) 2; mean \(K\) 2.4230769; max \(K\) 4; Table S2). The ordinations
-resulting from the NMDS analyses were used to construct an adaptive (if
-present) and neutral genetic space for each species. The spatial
-distribution of these genetic spaces (Figures S29--S54) generally show
-patterns of spatial autocorrelation--especially the adaptive genetic
-spaces (eg. \emph{Phyteuma hemisphaericum}; S49). Additionally, most
-species show east-west patterns in their intra-specific genetic
-variation (eg. \emph{Dryas octopetala}; S36).
-
-The dissimilarity matrices generated using the surrogate variables
-correlated with the dissimilarity matrices based on the genetic
-characteristics of individuals for most species (Table 1). Specifically,
-environmental variation was found to significantly correlate with
-adaptive genetic variation for 1 species (NA \% associated with loci
-under selection; \emph{P} \textless{} 0.05; mean marginal 0.02 \(R^2\)
-\(\pm\) NA SD for significant models). Additionally, geographic
-distances between planning units were found to significantly correlate
-with neutral genetic variation among planning units for 25 species
-(96.15 \% of all the species investigated;\emph{P} \textless{} 0.05;
-mean marginal 0.21 \(R^2\) \(\pm\) 0.17 SD for significant models).
-These results suggest that prioritizations with different environmental
-characteristics tend to have individuals with different adaptive genetic
-characteristics. Similarly, planning units that are further apart tend
-to have individuals with different neutral genetic characteristics.
-After verifying that the geographic and environmental variables
-correlate with genetic variation, the next step was to determine how
-effective these variables were as surrogates in the context of
-conservation planning. The dissimilarity matrices generated using the
-surrogate variables correlated with the dissimilarity matrices based on
-the genetic characteristics of individuals for most species (Table 1).
-Specifically, environmental variation was found to significantly
-correlate with adaptive genetic variation for 1 species (NA \%
-associated with loci under selection; \emph{P} \textless{} 0.05; mean
-marginal 0.02 \(R^2\) \(\pm\) NA SD for significant models).
-Additionally, geographic distances between planning units were found to
-significantly correlate with neutral genetic variation among planning
-units for 25 species (96.15 \% of all the species investigated;\emph{P}
-\textless{} 0.05; mean marginal 0.21 \(R^2\) \(\pm\) 0.17 SD for
-significant models). These results suggest that prioritizations with
-different environmental characteristics tend to have individuals with
-different adaptive genetic characteristics. Similarly, planning units
-that are further apart tend to have individuals with different neutral
-genetic characteristics. After verifying that the geographic and
-environmental variables correlate with genetic variation, the next step
-was to determine how effective these variables were as surrogates in the
-context of conservation planning.
-
-The environmental and geographic variables were moderately effective
-surrogates for genetic variation in most species (Figure 2). The
-relationship between the proportion of adaptive genetic variation
-secured in a prioritization and the proportion environment variation it
-also secured differed depending on the species for which the
-prioritization was generated
-(\(\chi^2_{4933.615872} P = 2.7275469\times 10^{-11}\)). The slope for
-each species were all significantly different to zero (minimum
-\(z value = 31.6562515\); maximum
-\(P = r `max(env.surrogate.spp.DF\)p.value)`\$; Table S3). The
-explanatory power of the environmental surrogates were also generally
-quite high for most species (mean 0.5746939 \(R^2\) 0.0365087 SD; Figure
-S56), with exception to \emph{Carex sempervirens}. The nature of these
-relationships suggest that most of the adaptive variation for a given
-species can be secured in a prioritization using moderately high
-environmental-based targets. Similarly, the proportion of neutral
-genetic variation in a prioritization was affected by the interaction
-between species and the proportion of neutral variation secured in the
-prioritization (\(\chi^2_{4.462023\times 10^{4}} P = 0\)). The slope for
-each species were also significantly different to zero (minimum
-\(z value = 23.3822268\); maximum \$P = 0; Table S3). Generally, the
-explanatory power of geographic variables were higher than that of
-environmental variables (mean 0.5495176 \(R^2\) 0.1272203 SD; Figure
-S57). Compared to the environmental variables, the slopes of the
-relationships suggest that very high geographic-based targets are
-required to generate prioritizations that secure a sufficient proportion
-of neutral genetic variation for most species. Additionally, all species
-were associated with prioritizations that secured a large proportion of
-environmental or geographic variation (\textgreater{} 70 \%), yet
-preserved only a very small proportion of the genetic variation
-(\textless{} 10 \%). These results suggest while prioritizations
-generated using surrogate-based targets tend to secure more genetic
-variation, prioritizations generated using just these targets can result
-in poor solutions.
-
-To determine if the addition of environmental and geographic targets
-could improve the effectiveness of prioritizations in a more realistic
-context, we generated prioritizations according to three different
-scenarios (Figure 3). The prioritizations generated under the
-single-species scenario showed relatively homogeneous selection
-frequencies across the study area--regardless of the combination of
-targets used to generate them (Figures 3a, 3b, 3c). On the other hand,
-the prioritizations generated under the multi-species scenario showed
-different selection frequencies depending on the targets used to
-generate them. The prioritization generated using only amount-based
-targets contained fewer planning units (Figure 3d; \(n = 14\)) than the
-prioritizations generated using additional of surrogate-based targets
-(Figure 3e; \(n = 33\)) or genetic-based targets (Figure 3f;
-\(n = 27\)). The prioritizations generated under the multi-species with
-cost scenario had different spatial distributions to the previous two
-scenarios. These prioritizations tended to concentrate conservation
-effort towards areas with higher elevation. Similar to the second
-scenario, the amount-based prioritization contained less panning units
-(Figure 3g; \(n = 14\)) than either the surrogate-based (Figure 3h;
-\(n = 48\)) or the genetic-based prioritizations ((Figure 3i;
-\(n = 34\))).
-
-Generally, the prioritizations generated using amount-based targets met
-the targets for adaptive genetic variation for most species associated
-with adaptive loci (Figure 3). Among the three scenarios, the
-amount-based prioritization in the single-species scenario had the
-lowest proportion of species with targets met (\texttt{r} \% of species
-with adaptive loci; Figure 3a). In the other two scenarios--the
-multi-species (\texttt{r} \% species preserved with adaptive loci;
-Figure 3c) and multi-species with cost (\texttt{r} \% species preserved
-with adaptive loci; Figure 3e) scenarios--the amount-based
-prioritizations successfully preserved the adaptive variation for most
-of the species. The addition of surrogate- and genetic-based targets did
-little to improve the proportion of species with their targets met for
-adaptive genetic variation (Figures 3c and 3e). These results suggest
-that the main gradients of intra-specific adaptive variation can be
-sufficiently preserved using suitable amount-based targets.
-
-The prioritizations generated using additional surrogate-based targets
-met the neutral genetic variation targets for more species than the
-prioritizations generated using just amount-based targets (Figure 3).
-The biggest difference between the prioritizations generated using
-additional surrogate-based targets and just the amount-based targets was
-in the single-species scenarios (\texttt{r} \% and \texttt{r} \% of
-species meeting genetic variation targets respectively). In the other
-two scenarios, the amount-based prioritizations tended to perform
-slightly better (\texttt{r} \% of species with targets met), but were
-outperformed by prioritizations with additional surrogate-based targets.
-Although the surrogate-based targets tended to fulfill the neutral
-genetic variation targets for most species, they captured only a very
-small proportion of neutral genetic variation for two species in
-particular (\emph{Gypsophila\_repens} and \emph{Luzula\_alpinopilosa}).
-These results suggest that surrogate-based targets can substantially
-improve the proportion of neutral genetic variation secured in a
-prioritization.
-
-\subsection{Discussion}\label{discussion}
-
-We aimed to investigate the effectiveness of broad-scale environmental
-and geographic variables as surrogates for adaptive and neutral genetic
-variation in conservation planning. Our study is the first to address
-this hypothesis using moderately realistic planning scenarios. Overall,
-we found that environmental and geographic variables are effective
-surrogates for a few species when developing prioritizations.
-Furthermore, when developing prioritizations under various planning
-scenarios, we found that the addition of environmental and geographic
-targets resulted in prioritizations that sufficiently represented the
-genetic variation of a greater number of species. Our results suggest
-that while the environmental and geographic variables have little
-benefit for most species, they can make a large difference for some
-species.
-
-\begin{itemize}
-\tightlist
-\item
-  Typically, conservation planning exercises do not consider
-  intra-specific variation when designing prioritizations.
-\item
-  our results show that for most species, a sufficient proportion of
-  their intra-specific variation can be secured using suitable
-  amount-based targets alone.
-\item
-  in particular, we found that using suitable amount-based targets in a
-  multi-species context with comprehensive set of target species
-  resulted in a prioritization that adequately represented the adaptive
-  genetic variation for nearly all the target species (panel c; Figure
-  3).
-\item
-  however, for some species, such a prioritization secures only a small
-  proportion of their genetic variation.
-\item
-  we found that for some species, prioritizations generated using
-  amount-based targets only secured a small amount of their neutral
-  genetic variation (panels b, d, e; Figure 3).
-\item
-  in planning exercises where the decision maker does not have access to
-  genetic data on their target species across the extent of the study
-  area--ie. nearly all real world planning scenarios--the decision maker
-  will not know which species are poorly represented in terms of their
-  inta-specific variation.
-\item
-  the poorly represented species may be those of particular significant
-  or they may not.
-\item
-  the decision maker has to weigh the increased costs of a
-  prioritization that includes environmental and geographic surrogates
-  against the risk that a prioritization that does not include these
-  surrogates fails to represent species of conservation significance.
-\item
-  here we found that prioritizations generated using environmental and
-  geographic surrogates have only minor increased costs (compare panels
-  c and d; panels e and f; Figure 3).
-\item
-  overall, these results suggest that although surrogates are
-  unnecessary for most species, they can be highly beneficial for a few
-  species, the increased costs associated with using these surrogates is
-  small, so generally using surrogates is a good idea.
-\end{itemize}
-
-\begin{itemize}
-\item
-  Previous studies on this data set have found correlations between
-  environmental and genetic variation.
-\item
-  Yet, for most species, prioritizations generated using environmental
-  surrogates did not secure substantially more genetic variation than
-  randomly generated prioritizations with the same number of planning
-  units.
-\item
-  Here we discuss some potential reasons for this disagreement.
-\item
-\item
-  This apparent disagreement can be attributed to multiple reasons.
-\item
-  These results highlight the importance of using prioritizations to
-  assess the effectiveness of potential surrogates for conservation and
-  not simply measuring the strength of correlation between potential
-  surrogates and genetic variation.
-\end{itemize}
-
-We wish to alert our readers to several limitations associated with our
-analysis. First, the size of the planning units we used (23 \(\times\)
-25 km) is much larger than typically used in conservation planning
-scenarios (between 1--10 km\^{}2). We used this resolution because the
-genetic data was collected at this scale. Whilst we could have
-interpolated the genetic data to a finer resolution and used smaller
-planning units, this would have biased our analysis because this process
-would have artificially introduced additional spatial auto-correlation
-into the dataset. Second, we used geographic distances as surrogates for
-connectivity between planning units. Although connectivity between areas
-can be modeled using isolation by resistance measures (McRae 2006)
-derived from landscape data (eg. Peterman \emph{et al.} 2014a), they are
-unlikely to be of use in broad-scale conservation exercises since they
-need to be tuned using species-specific parameters. Third, we used AFLP
-data to characterize the main gradients of genetic variation. Whilst
-next-generation sequencing provides higher resolution genetic
-information than AFLP data (reviewed in Schoville \emph{et al.} 2012),
-our methodology would still have utilized only the main gradients of the
-genetic variation in order to generate prioritizations in a feasible
-amount of time. Overall, we are confident that the limitations
-associated with our study are relatively minor.
-
-\subsubsection{Concluding remarks}\label{concluding-remarks}
-
-The overarching aim of our study was to investigate the effectiveness of
-surrogates in conservation planning. We found that environmental and
-geographic surrogates are effective. However, they seem to be largely
-unnecessary in conservation planning scenarios because a comprehensive
-set of species can be used to achieve a similar result. Conservation
-planners are therefore urged to use a comprehensive set of species in
-prioritizations--one of the fundamental principals of conservation
-planning.
-
-\subsection{Acknowledgements}\label{acknowledgements}
-
-JOH is funded by an Australian Postgraduate Award (APA) scholarship. RAF
-has an Australian Research Council Future Fellowship. This work was
-supported by the Centre of Excellence for Environmental Decisions (CEED)
-and the Landscape Ecology and Conservation Group (LEC) at The University
-of Queensland.
-
-\subsection*{References}\label{references}
-\addcontentsline{toc}{subsection}{References}
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures_files/figure-latex/unnamed-chunk-6-1.pdf}
+\caption{The effectiveness of prioritizations at securing genetic
+variation. The top row of panels (a--b) show the performance of
+single-species prioritizations generated assuming equal costs among
+planning costs. The middle row of panels (c--d) show the performance of
+multi-species prioritizations generated assuming equal costs among
+planning units. The bottom row of panels (e--f) show the performance of
+multi-species prioritizations generated using opportunity costs. The
+left column of panels (a, c, d) and the right column of panels (b, d, f)
+show the proportion of adaptive and neutral genetic variation secured
+(respectively). Box plots show the effectiveness of prioritizations
+generated using amount-based targets, amount-based and surrogate-based
+targets, and amount-based and genetic-based targets for each species.
+Letters denote significant differences (\(P < 0.05\)).}
+\end{figure}
 
 \hypertarget{refs}{}
-\hypertarget{ref-r478}{}
-1. Alvarez, N., Thiel-Egenter, C., Tribsch, A., Holderegger, R., Manel,
-S., \& Schönswetter, P. \emph{et al.} (2009). History or ecology?
-Substrate type as a major driver of patial genetic structure in alpine
-plants. \emph{Ecology Letters}, 12, 632--640.
-
-\hypertarget{ref-r491}{}
-2. Andrew J. Bass, J. D. S. with contributions from, Dabney, A., \&
-Robinson, D. (2015). \emph{Qvalue: Q-value estimation for false
-discovery rate control}.
-
-\hypertarget{ref-r22}{}
-3. Ball, I., Possingham, H., \& Watts, M. E. (2009). Marxan and
-relatives: software for spatial conservation prioritisation. In:
-\emph{Spatial conservation prioritisation: Quantitative methods \&
-computational tools} (eds. Moilanen, A., Wilson, K.A. \& Possingham,
-H.). Book Section. Oxford University Press, Oxford, UK, pp. 185--189.
-
-\hypertarget{ref-r466}{}
-4. Benjamini, Y. \& Hochberg, Y. (1995). Controlling the false discovery
-rate: A practical and powerful approach to multiple testing.
-\emph{Journal of the Royal Statistical Society. Series B
-(Methodological)}, 57, 289--300.
-
-\hypertarget{ref-r157}{}
-5. Bonin, A., Nicole, F., Pompanon, F., Miaud, C., \& Taberlet, P.
-(2007). Population adaptive index: A new method to help measure
-intraspecific genetic diversity and prioritize populations for
-conservation. \emph{Conservation Biology}, 21, 697--708.
-
-\hypertarget{ref-r455}{}
-6. Bothwell, H., Bisbing, S., Therkildsen, N., Crawford, L., Alvarez,
-N., \& Holderegger, R. \emph{et al.} (2013). Identifying genetic
-signatures of selection in a non-model species, alpine gentian (gentiana
-nivalis l.), using a landscape genetic approach. \emph{Conservation
-Genetics}, 14, 467--481.
-
-\hypertarget{ref-r2}{}
-7. Carvalho, S. B., Brito, J. C., Crespo, E. J., \& Possingham, H. P.
-(2011). Incorporating evolutionary processes into conservation planning
-using species distribution data: A case study with the western
-mediterranean herpetofauna. \emph{Diversity \& Distributions}, 17,
-408--421.
-
-\hypertarget{ref-r472}{}
-8. CIESEN, Columbia University, International Food Policy Research
-Institute, The World Bank, \& Centro Internacional de Agricultura
-Tropical. (2011). Global rural-urban mapping project, version 1 (gRUMP
-v1): Urban extents grid.
-
-\hypertarget{ref-r494}{}
-9. Clarke, R. T., Rothery, P., \& Raybould, A. F. (2002). Confidence
-limits for regression relationships between distance matrices:
-Estimating gene flow with distance. \emph{Journal of Agricultural,
-Biological, and Environmental Statistics}, 7, 361--372.
-
-\hypertarget{ref-r197}{}
-10. Cowling, R. M., Pressey, R. L., Rouget, M., \& Lombard, A. T.
-(2003). A conservation plan for a global biodiversity hotspot - the cape
-floristic region, south africa. \emph{Biological Conservation}, 112,
-191--216.
-
-\hypertarget{ref-r3}{}
-11. Crandall, K. A., Bininda-Emonds, O. R. P., Mace, G. M., \& Wayne, R.
-K. (2000). Considering evolutionary processes in conservation biology.
-\emph{Trends in Ecology \& Evolution}, 15, 290--295.
-
-\hypertarget{ref-r487}{}
-12. Duforet-Frebourg, N., Bazin, E., \& Blum, M. G. (2014). Genome scans
-for detecting footprints of local adaptation using a bayesian factor
-model. \emph{Molecular Biology and Evolution}, 31, 2483--2495.
-
-\hypertarget{ref-r456}{}
-13. Foll, M. \& Gaggiotti, O. (2008). A genome-scan method to identify
-selected loci appropriate for both dominant and codominant markers: A
-bayesian perspective. \emph{Genetics}, 180, 977--993.
-
-\hypertarget{ref-r464}{}
-14. Gower, J. C. (1971). A general coefficient of similarity and some of
-its properties. \emph{Biometrics}, 27, 857--871.
-
-\hypertarget{ref-r452}{}
-15. Gugerli, F., Englisch, T., Niklfeld, H., Tribsch, A., Mirek, Z., \&
-Ronikier, M. \emph{et al.} (2008). Relationships among levels of
-biodiversity and the relevance of intraspecific diversity in
-conservation - a project synopsis. \emph{Perspectives in Plant Ecology,
-Evolution and Systematics}, 10, 259--281.
-
-\hypertarget{ref-r463}{}
-16. Gurobi Optimization, I. (2015). Gurobi optimizer reference manual.
-
-\hypertarget{ref-r465}{}
-17. Hanson, J. O., Rhodes, J. R., Possingham, H. P., \& Fuller, R. A.
-(2016). RAPR: representative and adequate prioritisations in R.
-\emph{Unpublished}, X, XX--XX.
-
-\hypertarget{ref-r61}{}
-18. Hendry, A. P., Lohmann, L. G., Conti, E., Cracraft, J., Crandall, K.
-A., \& Faith, D. P. \emph{et al.} (2010). Evolutionary biology in
-biodiversity science, conservation, and policy: A call to action.
-\emph{Evolution}, 64, 1517--1528.
-
-\hypertarget{ref-r33}{}
-19. Hijmans, R. J., Cameron, S. E., Parra, J. L., Jones, P. G., \&
-Jarvis, A. (2005). Very high resolution interpolated climate surfaces
-for global land areas. \emph{International Journal of Climatology}, 25,
-1965--1978.
-
-\hypertarget{ref-r492}{}
-20. Jackman, S. (2015). \emph{pscl: Classes and methods for R developed
-in the political science computational laboratory, stanford university}.
-Department of Political Science, Stanford University, Stanford,
-California.
-
-\hypertarget{ref-r459}{}
-21. Jakobsson, M. \& Rosenberg, N. A. (2007). CLUMPP: A cluster matching
-and permutation program for dealing with label switching and
-multimodality in analysis of population structure.
-\emph{Bioinformatics}, 23, 1801--1806.
-
-\hypertarget{ref-r496}{}
-22. Luu, K., Blum, M. G., \& Duforet-Frebourg, N. (2016). \emph{Pcadapt:
-Fast principal component analysis for outlier detection}.
-
-\hypertarget{ref-r457}{}
-23. Maechler, M., Rousseeuw, P., Struyf, A., Hubert, M., \& Hornik, K.
-(2015). \emph{Cluster: Cluster analysis basics and extensions}.
-
-\hypertarget{ref-r398}{}
-24. Mahalanobis, P. C. (1936). On the generalized distance in
-statistics. \emph{Proceedings of the National Institute of Sciences,
-Calcutta}, 2, 49--55.
-
-\hypertarget{ref-r475}{}
-25. Manel, S., Gugerli, F., Thuiller, W., Alvarez, N., Legendre, P., \&
-Holderegger, R. \emph{et al.} (2012). Broad-scale adaptive genetic
-variation in alpine plants is driven by temperature and precipitation.
-\emph{Molecular Ecology}, 21, 3729--3738.
-
-\hypertarget{ref-r484}{}
-26. Manel, S., Poncet, B. N., Legendre, P., Gugerli, F., \& Holderegger,
-R. (2010). Common factors drive adaptive genetic variation at different
-spatial scales in arabis alpina. \emph{Molecular Ecology}, 19,
-3824--3835.
-
-\hypertarget{ref-r473}{}
-27. McRae, B. H. (2006). Isolation by resistance. \emph{Evolution}, 60,
-1551--1561.
-
-\hypertarget{ref-r451}{}
-28. Meirmans, P., Goudet, J., IntraBioDiv Consortium, \& Gaggiotti, O.
-(2011). Ecology and life history affect different aspects of the
-population structure of 27 high-alpine plants. \emph{Molecular Ecology},
-20, 3144--3155.
-
-\hypertarget{ref-r5}{}
-29. Moritz, C. (1994). Defining evolutionarily significant units for
-conservation. \emph{Trends in Ecology \& Evolution}, 9, 373--375.
-
-\hypertarget{ref-r67}{}
-30. Moritz, C. (1999). Conservation units and translocations: Strategies
-for conserving evolutionary processes. \emph{Hereditas}, 130, 217--228.
-
-\hypertarget{ref-r458}{}
-31. Oksanen, J., Blanchet, F. G., Kindt, R., Legendre, P., Minchin, P.
-R., \& O'Hara, R. B. \emph{et al.} (2015). \emph{Vegan: Community
-ecology package}.
-
-\hypertarget{ref-r474}{}
-32. Peterman, W. E., Anderson, T. L., Ousterhout, B. H., Drake, D. L.,
-Semlitsch, R. D., \& Eggert, L. S. (2014a). Differential dispersal
-shapes population structure and patterns of genetic differentiation in
-two sympatric pond breeding salamanders. \emph{Conservation Genetics},
-16, 59--69.
-
-\hypertarget{ref-r488}{}
-33. Peterman, W. E., Connette, G. M., Semlitsch, R. D., \& Eggert, L. S.
-(2014b). Ecological resistance surfaces predict fine-scale genetic
-differentiation in a terrestrial woodland salamander. \emph{Molecular
-Ecology}, 23, 2402--2413.
-
-\hypertarget{ref-r200}{}
-34. Ponce-Reyes, R., Clegg, S. M., Carvalho, S. B., McDonald-Madden, E.,
-\& Possingham, H. P. (2014). Geographical surrogates of genetic
-variation for selecting island populations for conservation.
-\emph{Diversity \& Distributions}, 20, 640--651.
-
-\hypertarget{ref-r247}{}
-35. Pyke, C. R. \& Fischer, D. T. (2005). Selection of bioclimatically
-representative biological reserve systems under climate change.
-\emph{Biological Conservation}, 121, 429--441.
-
-\hypertarget{ref-r489}{}
-36. Rellstab, C., Gugerli, F., Eckert, A. J., Hancock, A. M., \&
-Holderegger, R. (2015). A practical guide to environmental association
-analysis in landscape genomics. \emph{Molecular Ecology}, 24,
-4348--4370.
-
-\hypertarget{ref-r493}{}
-37. Rodríguez, J. P., Brotons, L., Bustamante, J., \& Seoane, J. (2007).
-The application of predictive modelling of species distribution to
-biodiversity conservation. \emph{Diversity and Distributions}, 13,
-243--251.
-
-\hypertarget{ref-r485}{}
-38. Rolland, J., Lavergne, S., \& Manel, S. (2015). Combining niche
-modelling and landscape genetics to study local adaptation: A novel
-approach illustrated using alpine plants. \emph{Perspectives in Plant
-Ecology, Evolution and Systematics}, 17, 491--499.
-
-\hypertarget{ref-r440}{}
-39. Sanderson, E. W., Segan, D. B., \& Watson, J. E. M. (2015). Global
-status of and prospects for protection of terrestrial geophysical
-diversity. \emph{Conservation Biology}, 29, 649--656.
-
-\hypertarget{ref-r480}{}
-40. Schoville, S. D., Bonin, A., Francois, O., Lobreaux, S., Melodelima,
-C., \& Manel, S. (2012). Adaptive genetic variation on the landscape:
-Methods and cases. In: \emph{Annaul review of ecology, evolution, and
-systematics}, Annual review of ecology evolution and systematics (ed.
-Futuyma, D.). pp. 23--43.
-
-\hypertarget{ref-r486}{}
-41. Taberlet, P., Zimmermann, N. E., Englisch, T., Tribsch, A.,
-Holderegger, R., \& Alvarez, N. \emph{et al.} (2012). Genetic diversity
-in widespread species is not congruent with species richness in alpine
-plant communities. \emph{Ecology Letters}, 15, 1439--1448.
-
-\hypertarget{ref-r495}{}
-42. Villemereuil, P. de, Frichot, É., Bazin, É., François, O., \&
-Gaggiotti, O. E. (2014). Genome scan methods against more complex
-models: When and how much should we trust them? \emph{Molecular
-Ecology}, 23, 2006--2019.
-
-\hypertarget{ref-r453}{}
-43. Vos, P., Hogers, R., Bleeker, M., Reijans, M., Lee, T. van de, \&
-Hornes, M. \emph{et al.} (1995). AFLP: A new technique for dNA
-fingerprinting. \emph{Nucleic Acids Research}, 23, 4407--4414.
-
-\hypertarget{ref-r490}{}
-44. Whitlock, M. C. (2015). Modern approaches to local adaptation.
-\emph{The American Naturalist}, 186, S1--S4.
-
-\hypertarget{ref-r259}{}
-45. Wright, S. (1943). Isolation by distance. \emph{Genetics}, 28,
-114--138.
-
-% \includepdf[pages=-, pagecommand={}]{figures.pdf}
-% \includepdf[pages=-, pagecommand={}]{supporting_information.pdf}
 

### 2016-05-07 code/rmarkdown/article.Rmd
+ fixes to article
Former-commit-id: b60c0fbeff1f1c52aff2c2fee15295295b8f438c

@@
 We successfully mapped the main spatial patterns of adaptive (if present) and neutral variation for each of the 26 species.
-The outlier detection methods detected adaptive genetic variation in `r length(env.surrogacy.LST)` species.
-Generally, two outlier detection methods gave fairly consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD).
-Of the species that were associated with adaptive genetic variation, only a few loci were detected as under selection (mean `r round(mean(spp.outlier.loci.props, na.rm=TRUE),2)` % loci per species `r round(sd(spp.outlier.loci.props, na.rm=TRUE),2)` SD).
+The outlier detection methods detected adaptive genetic variation in `r as.character(as.english(length(env.surrogacy.LST)))` species.
+Generally, two outlier detection methods gave fairly consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD).
+Of the species that were associated with adaptive genetic variation, only a few loci were detected as under selection (mean `r round(mean(spp.outlier.loci.props, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.loci.props, na.rm=TRUE),2)` SD).
 After identifying the loci under selection, we used NMDS analyses to identify the main gradients of adaptive (if detected) and neutral variation for each species.
-Generally, only a couple of continuous dimensions were needed to sufficiently describe the adaptive genetic variation for each species (minimum $K$ `r min(adaptive.k,ma.rm=TRUE)`; mean $K$ `r mean(adaptive.k,na.rm=TRUE)`; max $K$ `r max(adaptive.k,na.rm=TRUE)`; Table S2).
-On the other hand, a few more continuous dimensions were often needed for neutral genetic variation for each species (minimum $K$ `r min(neutral.k)`; mean $K$ `r mean(neutral.k)`; max $K$ `r max(neutral.k)`; Table S2).
+Generally, only a couple of continuous dimensions were needed to sufficiently describe the adaptive genetic variation for each species (all species $K = 2$; Table S2).
+On the other hand, a few more continuous dimensions were often needed for neutral genetic variation for each species (minimum $K = `r min(neutral.k)`$; mean $K = `r mean(neutral.k)`$; max $K = `r max(neutral.k)`$; Table S2).
 The ordinations resulting from the NMDS analyses were used to construct an adaptive (if present) and neutral genetic space for each species.
@@
 The dissimilarity matrices generated using the surrogate variables correlated with the dissimilarity matrices based on the genetic characteristics of individuals for most species (Table 1).
-Specifically, environmental variation was found to significantly correlate with adaptive genetic variation for `r sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)` species (`r round((sum(correlation.model.summary.DF$Environmental.corr.P<0.05)/length(na.omit(correlation.model.summary.DF$Environmental.corr.P)))*100,2)` % associated with loci under selection; _P_ < 0.05; mean marginal `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models).
-Additionally, geographic distances between planning units were found to significantly correlate with neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` species (`r round((sum(correlation.model.summary.DF$Geographic.corr.P<0.05)/length(na.omit(correlation.model.summary.DF$Geographic.corr.P)))*100,2)` % of all the species investigated;_P_ < 0.05; mean marginal `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models).
-These results suggest that prioritizations with different environmental characteristics tend to have individuals with different adaptive genetic characteristics.
-Similarly, planning units that are further apart tend to have individuals with different neutral genetic characteristics.
-After verifying that the geographic and environmental variables correlate with genetic variation, the next step was to determine how effective these variables were as surrogates in the context of conservation planning.
-The dissimilarity matrices generated using the surrogate variables correlated with the dissimilarity matrices based on the genetic characteristics of individuals for most species (Table 1).
-Specifically, environmental variation was found to significantly correlate with adaptive genetic variation for `r sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)` species (`r round((sum(correlation.model.summary.DF$Environmental.corr.P<0.05)/length(na.omit(correlation.model.summary.DF$Environmental.corr.P)))*100,2)` % associated with loci under selection; _P_ < 0.05; mean marginal `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models).
-Additionally, geographic distances between planning units were found to significantly correlate with neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` species (`r round((sum(correlation.model.summary.DF$Geographic.corr.P<0.05)/length(na.omit(correlation.model.summary.DF$Geographic.corr.P)))*100,2)` % of all the species investigated;_P_ < 0.05; mean marginal `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models).
+Specifically, environmental variation was found to significantly correlate with adaptive genetic variation for `r as.character(as.english(sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)))` species (`r round((sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)/length(na.omit(correlation.model.summary.DF$Environmental.corr.P)))*100,2)` % associated with loci under selection; _P_ < 0.05; mean marginal `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models).
+Additionally, geographic distances between planning units were found to significantly correlate with neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` species (`r round((sum(correlation.model.summary.DF$Geographic.corr.P<0.05)/length(na.omit(correlation.model.summary.DF$Geographic.corr.P)))*100,2)` % of all the species investigated; _P_ < 0.05; mean marginal `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models).
 These results suggest that prioritizations with different environmental characteristics tend to have individuals with different adaptive genetic characteristics.
@@
 The environmental and geographic variables were moderately effective surrogates for genetic variation in most species (Figure 2).
-The relationship between the proportion of adaptive genetic variation secured in a prioritization and the proportion environment variation it also secured differed depending on the species for which the prioritization was generated ($\chi^2_{`r env.surrogacy.AOV$Resid..Dev[[4]]`} P = `r env.surrogacy.AOV$Pr..Chi.[[4]]`$).
-The slope for each species were all significantly different to zero (minimum $z value = `r min(env.surrogate.spp.DF$statistic)`$; maximum $P = r `max(env.surrogate.spp.DF$p.value)`$; Table S3).
-The explanatory power of the environmental surrogates were also generally quite high for most species (mean `r mean(env.surrogate.spp.DF$r2CU)` $R^2$ `r sd(env.surrogate.spp.DF$r2CU)` SD; Figure S56), with exception to _Carex sempervirens_.
+The relationship between the proportion of adaptive genetic variation secured in a prioritization and the proportion environment variation it also secured differed depending on the species for which the prioritization was generated ($\chi^2_{r env.surrogacy.AOV$df[[4]]} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`} P = pvalString(`r env.surrogacy.AOV$Pr..Chi.[[4]]`)$).
+The slope for each species were all significantly different to zero (minimum \textit{z} = `r round(min(env.surrogate.spp.DF$statistic),2)`; maximum $P = `r  pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3).
+The explanatory power of the environmental surrogates were also generally quite high for most species (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2 \pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figure S56), with exception to _Carex sempervirens_.
 The nature of these relationships suggest that most of the adaptive variation for a given species can be secured in a prioritization using moderately high environmental-based targets.
-Similarly, the proportion of neutral genetic variation in a prioritization was affected by the interaction between species and the proportion of neutral variation secured in the prioritization ($\chi^2_{`r geo.surrogacy.AOV$Resid..Dev[[4]]`} P = `r geo.surrogacy.AOV$Pr..Chi.[[4]]`$).
+Similarly, the proportion of neutral genetic variation in a prioritization was affected by the interaction between species and the proportion of neutral variation secured in the prioritization ($\chi^2_{r geo.surrogacy.AOV$df[[4]]} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ = `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`).
 The slope for each species were also significantly different to zero (minimum $z value = `r min(geo.surrogate.spp.DF$statistic)`$; maximum $P = `r max(env.surrogate.spp.DF$p.value)`; Table S3).
-Generally, the explanatory power of geographic variables were higher than that of environmental variables (mean `r mean(geo.surrogate.spp.DF$r2CU)` $R^2$ `r sd(geo.surrogate.spp.DF$r2CU)` SD; Figure S57).
+Generally, the explanatory power of geographic variables were higher than that of environmental variables (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2 \pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figure S57).
 Compared to the environmental variables, the slopes of the relationships suggest that very high geographic-based targets are required to generate prioritizations that secure a sufficient proportion of neutral genetic variation for most species.
@@
 These prioritizations tended to concentrate conservation effort towards areas with higher elevation.
-Similar to the second scenario, the amount-based prioritization contained less panning units (Figure 3g; $n = `r sum(selections(multi.spp.prioritisations.with.cost[[1]]))`$) than either the surrogate-based (Figure 3h; $n = `r sum(selections(multi.spp.prioritisations.with.cost[[2]]))`$) or the genetic-based prioritizations ((Figure 3i; $n = `r sum(selections(multi.spp.prioritisations.with.cost[[3]]))`$)).
+Similar to the second scenario, the amount-based prioritization contained less panning units (Figure 3g; $n = `r sum(selections(multi.spp.prioritisations.with.cost[[1]]))`$) than either the surrogate-based (Figure 3h; $n = `r sum(selections(multi.spp.prioritisations.with.cost[[2]]))`$) or the genetic-based prioritizations (Figure 3i; $n = `r sum(selections(multi.spp.prioritisations.with.cost[[3]]))`$).
 
@@
 In the other two scenarios, the amount-based prioritizations tended to perform slightly better (`r ` % of species with targets met), but were outperformed by prioritizations with additional surrogate-based targets.
-Although the surrogate-based targets tended to fulfill the neutral genetic variation targets for most species, they captured only a very small proportion of neutral genetic variation for two species in particular (_Gypsophila_repens_ and _Luzula_alpinopilosa_).
+Although the surrogate-based targets tended to fulfill the neutral genetic variation targets for most species, they captured only a very small proportion of neutral genetic variation for two species in particular (_Gypsophila repens_ and _Luzula alpinopilosa_).
 These results suggest that surrogate-based targets can substantially improve the proportion of neutral genetic variation secured in a prioritization.

### 2016-05-07 code/rmarkdown/supporting_information.Rmd
+ fixes to article
Former-commit-id: b60c0fbeff1f1c52aff2c2fee15295295b8f438c

@@
 	# exports
-	return((((n.plots-1) %/% 4)+1) * 1)
+	return((((n.plots-1) %/% 4)+1) * 2)
 }
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=3.5, fig.cap='The relationship between the proportion of adaptive genetic variation secured in a prioritizaiton and the proportion of environmental variation it also secures.
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=2.5, fig.cap='The relationship between the proportion of adaptive genetic variation secured in a prioritizaiton and the proportion of environmental variation it also secures.
 Points represent values associated with randomly generated prioritizations.
@@
 ```{r, echo=FALSE, results='asis'}
-adaptive.spp.nmds.summary.DF <- spp.nmds.summary.DF %>% filter(Loci=='adaptive') %>% mutate(Stress=pvalString(Stress))
-neutral.spp.nmds.summary.DF <- spp.nmds.summary.DF %>% filter(Loci=='neutral') %>% mutate(Stress=pvalString(Stress))
+adaptive.spp.nmds.summary.DF <- spp.nmds.summary.DF %>% filter(Loci=='adaptive') %>% mutate(Stress=as.character(round(Stress,3))) %>% mutate(Stress=replace(Stress, Stress=='0', '< 0.001'))
+neutral.spp.nmds.summary.DF <- spp.nmds.summary.DF %>% filter(Loci=='neutral') %>% mutate(Stress=as.character(round(Stress,3))) %>% mutate(Stress=replace(Stress, Stress=='0', '< 0.001'))
 names(adaptive.spp.nmds.summary.DF)[2:5] <- paste0('adaptive.', names(adaptive.spp.nmds.summary.DF)[2:5])

### 2016-05-07 code/rmarkdown/tables.Rmd
+ fixes to article
Former-commit-id: b60c0fbeff1f1c52aff2c2fee15295295b8f438c

@@
 		Geographic.corr.P=pvalString(Geographic.corr.P),
+		Geographic.Marginal.R2=pvalString(Geographic.Marginal.R2),
+		Geographic.Conditional.R2=pvalString(Geographic.Conditional.R2),
+		Environmental.Marginal.R2=pvalString(Environmental.Marginal.R2),
+		Environmental.Conditional.R2=pvalString(Environmental.Conditional.R2),
 		blank=rep('', length(Species)),

### 2016-05-07 code/rmarkdown/article.Rmd
+ article fixes
Former-commit-id: c2182da5b3d06da50c4bf34c2905e8b170c11b6f

@@
 The environmental and geographic variables were moderately effective surrogates for genetic variation in most species (Figure 2).
-The relationship between the proportion of adaptive genetic variation secured in a prioritization and the proportion environment variation it also secured differed depending on the species for which the prioritization was generated ($\chi^2_{r env.surrogacy.AOV$df[[4]]} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`} P = pvalString(`r env.surrogacy.AOV$Pr..Chi.[[4]]`)$).
+The relationship between the proportion of adaptive genetic variation secured in a prioritization and the proportion environment variation it also secured differed depending on the species for which the prioritization was generated ($\chi^2_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`} P = pvalString(`r env.surrogacy.AOV$Pr..Chi.[[4]]`)$).
 The slope for each species were all significantly different to zero (minimum \textit{z} = `r round(min(env.surrogate.spp.DF$statistic),2)`; maximum $P = `r  pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3).
@@
 The nature of these relationships suggest that most of the adaptive variation for a given species can be secured in a prioritization using moderately high environmental-based targets.
-Similarly, the proportion of neutral genetic variation in a prioritization was affected by the interaction between species and the proportion of neutral variation secured in the prioritization ($\chi^2_{r geo.surrogacy.AOV$df[[4]]} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ = `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`).
+Similarly, the proportion of neutral genetic variation in a prioritization was affected by the interaction between species and the proportion of neutral variation secured in the prioritization ($\chi^2_{`r geo.surrogacy.AOV$df[[4]]`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ = `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`).
 The slope for each species were also significantly different to zero (minimum $z value = `r min(geo.surrogate.spp.DF$statistic)`$; maximum $P = `r max(env.surrogate.spp.DF$p.value)`; Table S3).

### 2016-05-07 code/rmarkdown/article.tex
article fixes applied
Former-commit-id: 3a40abac5ca3a8656f5c5f1ed1fde5f7d46f5b73

@@
 present) and neutral variation for each of the 26 species. The outlier
-detection methods detected adaptive genetic variation in 4 species.
+detection methods detected adaptive genetic variation in four species.
 Generally, two outlier detection methods gave fairly consistent results
-(mean 87.26 \% loci per species 4.97 SD). Of the species that were
-associated with adaptive genetic variation, only a few loci were
-detected as under selection (mean 1.77 \% loci per species 1.77 SD).
-After identifying the loci under selection, we used NMDS analyses to
-identify the main gradients of adaptive (if detected) and neutral
+(mean 87.26 \% loci per species \(\pm\) 4.97 SD). Of the species that
+were associated with adaptive genetic variation, only a few loci were
+detected as under selection (mean 1.77 \% loci per species \(\pm\) 1.77
+SD). After identifying the loci under selection, we used NMDS analyses
+to identify the main gradients of adaptive (if detected) and neutral
 variation for each species. Generally, only a couple of continuous
 dimensions were needed to sufficiently describe the adaptive genetic
-variation for each species (minimum \(K\) NA; mean \(K\) 2; max \(K\) 2;
-Table S2). On the other hand, a few more continuous dimensions were
-often needed for neutral genetic variation for each species (minimum
-\(K\) 2; mean \(K\) 2.4230769; max \(K\) 4; Table S2). The ordinations
-resulting from the NMDS analyses were used to construct an adaptive (if
-present) and neutral genetic space for each species. The spatial
-distribution of these genetic spaces (Figures S29--S54) generally show
-patterns of spatial autocorrelation--especially the adaptive genetic
-spaces (eg. \emph{Phyteuma hemisphaericum}; S49). Additionally, most
-species show east-west patterns in their intra-specific genetic
-variation (eg. \emph{Dryas octopetala}; S36).
+variation for each species (all species \(K = 2\); Table S2). On the
+other hand, a few more continuous dimensions were often needed for
+neutral genetic variation for each species (minimum \(K = 2\); mean
+\(K = 2.4230769\); max \(K = 4\); Table S2). The ordinations resulting
+from the NMDS analyses were used to construct an adaptive (if present)
+and neutral genetic space for each species. The spatial distribution of
+these genetic spaces (Figures S29--S54) generally show patterns of
+spatial autocorrelation--especially the adaptive genetic spaces (eg.
+\emph{Phyteuma hemisphaericum}; S49). Additionally, most species show
+east-west patterns in their intra-specific genetic variation (eg.
+\emph{Dryas octopetala}; S36).
 
@@
 environmental variation was found to significantly correlate with
-adaptive genetic variation for 1 species (NA \% associated with loci
+adaptive genetic variation for one species (25 \% associated with loci
 under selection; \emph{P} \textless{} 0.05; mean marginal 0.02 \(R^2\)
@@
 with neutral genetic variation among planning units for 25 species
-(96.15 \% of all the species investigated;\emph{P} \textless{} 0.05;
+(96.15 \% of all the species investigated; \emph{P} \textless{} 0.05;
 mean marginal 0.21 \(R^2\) \(\pm\) 0.17 SD for significant models).
@@
 effective these variables were as surrogates in the context of
-conservation planning. The dissimilarity matrices generated using the
-surrogate variables correlated with the dissimilarity matrices based on
-the genetic characteristics of individuals for most species (Table 1).
-Specifically, environmental variation was found to significantly
-correlate with adaptive genetic variation for 1 species (NA \%
-associated with loci under selection; \emph{P} \textless{} 0.05; mean
-marginal 0.02 \(R^2\) \(\pm\) NA SD for significant models).
-Additionally, geographic distances between planning units were found to
-significantly correlate with neutral genetic variation among planning
-units for 25 species (96.15 \% of all the species investigated;\emph{P}
-\textless{} 0.05; mean marginal 0.21 \(R^2\) \(\pm\) 0.17 SD for
-significant models). These results suggest that prioritizations with
-different environmental characteristics tend to have individuals with
-different adaptive genetic characteristics. Similarly, planning units
-that are further apart tend to have individuals with different neutral
-genetic characteristics. After verifying that the geographic and
-environmental variables correlate with genetic variation, the next step
-was to determine how effective these variables were as surrogates in the
-context of conservation planning.
+conservation planning.
 
@@
 prioritization was generated
-(\(\chi^2_{4933.615872} P = 2.7275469\times 10^{-11}\)). The slope for
-each species were all significantly different to zero (minimum
-\(z value = 31.6562515\); maximum
-\(P = r `max(env.surrogate.spp.DF\)p.value)`\$; Table S3). The
+(\(\chi^2_{3} = {4933.62} P = pvalString(2.7275469\times 10^{-11})\)).
+The slope for each species were all significantly different to zero
+(minimum \textit{z} = 31.66; maximum \(P = < 0.001\); Table S3). The
 explanatory power of the environmental surrogates were also generally
-quite high for most species (mean 0.5746939 \(R^2\) 0.0365087 SD; Figure
-S56), with exception to \emph{Carex sempervirens}. The nature of these
+quite high for most species (mean 0.57 \(R^2 \pm\) 0.04 SD; Figure S56),
+with exception to \emph{Carex sempervirens}. The nature of these
 relationships suggest that most of the adaptive variation for a given
@@
 between species and the proportion of neutral variation secured in the
-prioritization (\(\chi^2_{4.462023\times 10^{4}} P = 0\)). The slope for
-each species were also significantly different to zero (minimum
-\(z value = 23.3822268\); maximum \$P = 0; Table S3). Generally, the
-explanatory power of geographic variables were higher than that of
-environmental variables (mean 0.5495176 \(R^2\) 0.1272203 SD; Figure
-S57). Compared to the environmental variables, the slopes of the
-relationships suggest that very high geographic-based targets are
+prioritization (\(\chi^2_{25} = 4.462023\times 10^{4}\); \(P\) =
+\textless{} 0.001). The slope for each species were also significantly
+different to zero (minimum \(z value = 23.3822268\); maximum \$P = 0;
+Table S3). Generally, the explanatory power of geographic variables were
+higher than that of environmental variables (mean 0.55 \(R^2 \pm\) 0.13
+SD; Figure S57). Compared to the environmental variables, the slopes of
+the relationships suggest that very high geographic-based targets are
 required to generate prioritizations that secure a sufficient proportion
@@
 (Figure 3g; \(n = 14\)) than either the surrogate-based (Figure 3h;
-\(n = 48\)) or the genetic-based prioritizations ((Figure 3i;
-\(n = 34\))).
+\(n = 48\)) or the genetic-based prioritizations (Figure 3i;
+\(n = 34\)).
 
@@
 small proportion of neutral genetic variation for two species in
-particular (\emph{Gypsophila\_repens} and \emph{Luzula\_alpinopilosa}).
+particular (\emph{Gypsophila repens} and \emph{Luzula alpinopilosa}).
 These results suggest that surrogate-based targets can substantially

### 2016-05-07 code/rmarkdown/article.Rmd
fixes2
Former-commit-id: c8fb1056728ff43f6dd503cbc9789512c03ba016

--- 2016-05-07 code/rmarkdown/article.Rmd
@@
 Generally, only a couple of continuous dimensions were needed to sufficiently describe the adaptive genetic variation for each species (all species $K = 2$; Table S2).
-On the other hand, a few more continuous dimensions were often needed for neutral genetic variation for each species (minimum $K = `r min(neutral.k)`$; mean $K = `r mean(neutral.k)`$; max $K = `r max(neutral.k)`$; Table S2).
+On the other hand, a few more continuous dimensions were often needed for neutral genetic variation for each species (minimum $K = `r min(neutral.k)`$; mean $K = `r round(mean(neutral.k),2)`$; max $K = `r max(neutral.k)`$; Table S2).
 The ordinations resulting from the NMDS analyses were used to construct an adaptive (if present) and neutral genetic space for each species.
-The spatial distribution of these genetic spaces (Figures S29--S54) generally show patterns of spatial autocorrelation--especially the adaptive genetic spaces (eg.
-_Phyteuma hemisphaericum_; S49).
+The spatial distribution of these genetic spaces (Figures S29--S54) generally show patterns of spatial autocorrelation--especially the adaptive genetic spaces (eg.XXXX; SXX).
 Additionally, most species show east-west patterns in their intra-specific genetic variation (eg.
-_Dryas octopetala_; S36).
+XXXX; SXXX).
 
@@
 The dissimilarity matrices generated using the surrogate variables correlated with the dissimilarity matrices based on the genetic characteristics of individuals for most species (Table 1).
-Specifically, environmental variation was found to significantly correlate with adaptive genetic variation for `r as.character(as.english(sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)))` species (`r round((sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)/length(na.omit(correlation.model.summary.DF$Environmental.corr.P)))*100,2)` % associated with loci under selection; _P_ < 0.05; mean marginal `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models).
-Additionally, geographic distances between planning units were found to significantly correlate with neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` species (`r round((sum(correlation.model.summary.DF$Geographic.corr.P<0.05)/length(na.omit(correlation.model.summary.DF$Geographic.corr.P)))*100,2)` % of all the species investigated; _P_ < 0.05; mean marginal `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models).
+Specifically, environmental variation was found to significantly correlate with adaptive genetic variation for `r as.character(as.english(sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)))` species (`r round((sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)/length(na.omit(correlation.model.summary.DF$Environmental.corr.P)))*100,2)` % associated with loci under selection; _P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models).
+Additionally, geographic distances between planning units were found to significantly correlate with neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` species (`r round((sum(correlation.model.summary.DF$Geographic.corr.P<0.05)/length(na.omit(correlation.model.summary.DF$Geographic.corr.P)))*100,2)` % of all the species investigated; _P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models).
 These results suggest that prioritizations with different environmental characteristics tend to have individuals with different adaptive genetic characteristics.
@@
 The environmental and geographic variables were moderately effective surrogates for genetic variation in most species (Figure 2).
-The relationship between the proportion of adaptive genetic variation secured in a prioritization and the proportion environment variation it also secured differed depending on the species for which the prioritization was generated ($\chi^2_{r env.surrogacy.AOV$df[[4]]} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`} P = pvalString(`r env.surrogacy.AOV$Pr..Chi.[[4]]`)$).
-The slope for each species were all significantly different to zero (minimum \textit{z} = `r round(min(env.surrogate.spp.DF$statistic),2)`; maximum $P = `r  pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3).
+The relationship between the proportion of adaptive genetic variation secured in a prioritization and the proportion environment variation it also secured differed depending on the species for which the prioritization was generated (${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`} P = `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$).
+The slope for each species were all significantly different to zero (minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`; maximum $P = `r  pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3).
 The explanatory power of the environmental surrogates were also generally quite high for most species (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2 \pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figure S56), with exception to _Carex sempervirens_.
 The nature of these relationships suggest that most of the adaptive variation for a given species can be secured in a prioritization using moderately high environmental-based targets.
-Similarly, the proportion of neutral genetic variation in a prioritization was affected by the interaction between species and the proportion of neutral variation secured in the prioritization ($\chi^2_{r geo.surrogacy.AOV$df[[4]]} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ = `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`).
+Similarly, the proportion of neutral genetic variation in a prioritization was affected by the interaction between species and the proportion of neutral variation secured in the prioritization (${\chi^2}_{`r geo.surrogacy.AOV$df[[4]]`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ = `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`).
 The slope for each species were also significantly different to zero (minimum $z value = `r min(geo.surrogate.spp.DF$statistic)`$; maximum $P = `r max(env.surrogate.spp.DF$p.value)`; Table S3).

### 2016-05-07 code/rmarkdown/figures.Rmd
fixes2
Former-commit-id: c8fb1056728ff43f6dd503cbc9789512c03ba016

@@
 		data=filter(env.surrogate.spp.DF, !is.na(r2CU))) +
-	geom_histogram(fill='grey10', bins=10) +
+	geom_histogram(fill='grey10', bins=5) +
 	ylab('\nFrequency') + xlab(expression(atop('Effectiveness of environmental', variables~(R^2~coefficient)))) +

### 2016-05-07 code/rmarkdown/supporting_information.Rmd
fixes2
Former-commit-id: c8fb1056728ff43f6dd503cbc9789512c03ba016

@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(1), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[1]),'}.
+```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(1), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[1]),'}.
 Each panel shows a different dimension of the variation as indicated by its title.
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(2), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[2]),'}.
+```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(2), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[2]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(3), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[3]),'}.
+```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(3), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[3]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(4), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[4]),'}.
+```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(4), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[4]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(5), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[5]),'}.
+```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(5), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[5]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(6), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[6]),'}.
+```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(6), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[6]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(7), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[7]),'}.
+```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(7), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[7]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(8), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[8]),'}.
+```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(8), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[8]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(9), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[9]),'}.
+```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(9), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[9]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(10), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[10]),'}.
+```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(10), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[10]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(11), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[11]),'}.
+```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(11), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[11]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(12), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[12]),'}.
+```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(12), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[12]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(13), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[13]),'}.
+```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(13), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[13]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(14), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[14]),'}.
+```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(14), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[14]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(15), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[15]),'}.
+```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(15), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[15]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(16), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[16]),'}.
+```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(16), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[16]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(17), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[17]),'}.
+```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(17), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[17]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(18), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[18]),'}.
+```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(18), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[18]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(19), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[19]),'}.
+```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(19), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[19]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(20), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[20]),'}.
+```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(20), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[20]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(21), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[21]),'}.
+```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(21), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[21]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(22), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[22]),'}.
+```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(22), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[22]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(23), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[23]),'}.
+```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(23), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[23]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(24), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[24]),'}.
+```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(24), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[24]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(25), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[25]),'}.
+```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(25), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[25]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(26), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[26]),'}.
+```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(26), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[26]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=calc.fig.height(27), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[27]),'}.
+```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(27), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[27]),'}.
 See Figure S6 caption for conventions.')}

### 2016-05-07 code/rmarkdown/article.tex
article fixes applied2
Former-commit-id: be7e21cc9f6a8768cf2d4cf4d35b361c28b0201e

@@
 neutral genetic variation for each species (minimum \(K = 2\); mean
-\(K = 2.4230769\); max \(K = 4\); Table S2). The ordinations resulting
-from the NMDS analyses were used to construct an adaptive (if present)
-and neutral genetic space for each species. The spatial distribution of
+\(K = 2.42\); max \(K = 4\); Table S2). The ordinations resulting from
+the NMDS analyses were used to construct an adaptive (if present) and
+neutral genetic space for each species. The spatial distribution of
 these genetic spaces (Figures S29--S54) generally show patterns of
-spatial autocorrelation--especially the adaptive genetic spaces (eg.
-\emph{Phyteuma hemisphaericum}; S49). Additionally, most species show
-east-west patterns in their intra-specific genetic variation (eg.
-\emph{Dryas octopetala}; S36).
+spatial autocorrelation--especially the adaptive genetic spaces
+(eg.XXXX; SXX). Additionally, most species show east-west patterns in
+their intra-specific genetic variation (eg. XXXX; SXXX).
 
@@
 adaptive genetic variation for one species (25 \% associated with loci
-under selection; \emph{P} \textless{} 0.05; mean marginal 0.02 \(R^2\)
+under selection; \emph{P} \textless{} 0.05; mean 0.02 marginal \(R^2\)
 \(\pm\) NA SD for significant models). Additionally, geographic
@@
 (96.15 \% of all the species investigated; \emph{P} \textless{} 0.05;
-mean marginal 0.21 \(R^2\) \(\pm\) 0.17 SD for significant models).
+mean 0.21 marginal \(R^2\) \(\pm\) 0.17 SD for significant models).
 These results suggest that prioritizations with different environmental
@@
 also secured differed depending on the species for which the
-prioritization was generated
-(\(\chi^2_{3} = {4933.62} P = pvalString(2.7275469\times 10^{-11})\)).
+prioritization was generated (\({\chi^2}_{3} = {4933.62} P = < 0.001\)).
 The slope for each species were all significantly different to zero
-(minimum \textit{z} = 31.66; maximum \(P = < 0.001\); Table S3). The
+(minimum \textit{z} = 40; maximum \(P = < 0.001\); Table S3). The
 explanatory power of the environmental surrogates were also generally
@@
 between species and the proportion of neutral variation secured in the
-prioritization (\(\chi^2_{25} = 4.462023\times 10^{4}\); \(P\) =
+prioritization (\({\chi^2}_{25} = 4.462023\times 10^{4}\); \(P\) =
 \textless{} 0.001). The slope for each species were also significantly

### 2016-05-09 code/rmarkdown/article.Rmd
+ article fixes + remove unneeded comments from MLPE.R
Former-commit-id: ee0043520642de57e7be9a2a46e563d35abc1e75

@@
 To enable comparisons between the two outlier detection analyses, we used the same individuals in this analysis as used in the \texttt{BayeScan} analysis.
-For each of the 26 species, we first imputed missing loci data to permit further analysis.
-Specifically, the value for a given individual at a given locus was replaced with the mean frequency of that locus among all individuals.
-Next, we ran a PCA and extracted the minimum number of components needed to secure `r pcadapt.params.LST[[MODE]]$min.variance.explained` % of the variation in loci.
+For each of the 26 species, we first imputed missing data to permit further analysis.
+Specifically, the value for a given individual at a given locus was replaced with the mean frequency of that locus among all conspecifics.
+Next, we ran a PCA and extracted the minimum number of components needed to secure the over-arching population level variation among the loci (`r pcadapt.params.LST[[MODE]]$min.variance.explained*10` % of the variation in loci).
 We then used Mahalanobis distances [@r398] to compute _P_ values and then q-values [using the \texttt{qvalue R} package; @r491].
@@
 
-After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species (Figures S29--S54).
+After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species.
 For each species, we discarded the population groupings, and partitioned adaptive and neutral loci into separate matrices.
@@
 After identifying the loci under selection, we used NMDS analyses to identify the main gradients of adaptive (if detected) and neutral variation for each species.
-Generally, only a couple of continuous dimensions were needed to sufficiently describe the adaptive genetic variation for each species (all species $K = 2$; Table S2).
-On the other hand, a few more continuous dimensions were often needed for neutral genetic variation for each species (minimum $K = `r min(neutral.k)`$; mean $K = `r round(mean(neutral.k),2)`$; max $K = `r max(neutral.k)`$; Table S2).
+For all species in which we detected adaptive loci, only a couple of continuous dimensions were needed to sufficiently describe their patterns of adaptive genetic variation (Table S2).
+On the other hand, a few additional continuous dimensions were often needed to describe their patterns of neutral genetic variation (minimum $K = `r min(neutral.k)`$; mean $K = `r round(mean(neutral.k),2)`$; max $K = `r max(neutral.k)`$; Table S2).
 The ordinations resulting from the NMDS analyses were used to construct an adaptive (if present) and neutral genetic space for each species.
-The spatial distribution of these genetic spaces (Figures S29--S54) generally show patterns of spatial autocorrelation--especially the adaptive genetic spaces (eg.XXXX; SXX).
+The spatial distribution of these genetic spaces (Figures S29--S54) generally show patterns of spatial auto-correlation--especially the adaptive genetic spaces (eg.XXXX; SXX).
 Additionally, most species show east-west patterns in their intra-specific genetic variation (eg.
@@
 <!-- correlation analysis -->
-The dissimilarity matrices generated using the surrogate variables correlated with the dissimilarity matrices based on the genetic characteristics of individuals for most species (Table 1).
+The dissimilarity matrices generated using the environmental and geographic variables correlated with those matrices based on the genetic characteristics for most species (Table 1).
 Specifically, environmental variation was found to significantly correlate with adaptive genetic variation for `r as.character(as.english(sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)))` species (`r round((sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)/length(na.omit(correlation.model.summary.DF$Environmental.corr.P)))*100,2)` % associated with loci under selection; _P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models).
@@
 Similarly, planning units that are further apart tend to have individuals with different neutral genetic characteristics.
-After verifying that the geographic and environmental variables correlate with genetic variation, the next step was to determine how effective these variables were as surrogates in the context of conservation planning.
+After verifying that the geographic and environmental variables generally correlated with genetic variation, our next step was to determine how effective these variables were as surrogates in the context of conservation planning.
 
@@
 The environmental and geographic variables were moderately effective surrogates for genetic variation in most species (Figure 2).
-The relationship between the proportion of adaptive genetic variation secured in a prioritization and the proportion environment variation it also secured differed depending on the species for which the prioritization was generated (${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`} P = `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$).
-The slope for each species were all significantly different to zero (minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`; maximum $P = `r  pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3).
-The explanatory power of the environmental surrogates were also generally quite high for most species (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2 \pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figure S56), with exception to _Carex sempervirens_.
-The nature of these relationships suggest that most of the adaptive variation for a given species can be secured in a prioritization using moderately high environmental-based targets.
+The relationship between the proportion of adaptive genetic variation secured in a prioritization and the proportion of environment variation it also secured was found to depend on the species for which the prioritization was generated (${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$).
+The slope for each species were all significantly different to zero (minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`; maximum $P `r  pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3).
+The explanatory power of the environmental surrogates were also generally quite high for most species (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figure S56).
+The nature of these relationships suggest that most of the adaptive variation for a given species can be secured in a prioritization using moderately high environmental-based targets ($geq 80$ %).
 Similarly, the proportion of neutral genetic variation in a prioritization was affected by the interaction between species and the proportion of neutral variation secured in the prioritization (${\chi^2}_{`r geo.surrogacy.AOV$df[[4]]`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ = `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`).
-The slope for each species were also significantly different to zero (minimum $z value = `r min(geo.surrogate.spp.DF$statistic)`$; maximum $P = `r max(env.surrogate.spp.DF$p.value)`; Table S3).
+The slope for each species were also significantly different to zero (minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`$; maximum $P = `r pvalString(max(env.surrogate.spp.DF$p.value))`; Table S3).
 Generally, the explanatory power of geographic variables were higher than that of environmental variables (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2 \pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figure S57).
-Compared to the environmental variables, the slopes of the relationships suggest that very high geographic-based targets are required to generate prioritizations that secure a sufficient proportion of neutral genetic variation for most species.
-Additionally, all species were associated with prioritizations that secured a large proportion of environmental or geographic variation (> 70 %), yet preserved only a very small proportion of the genetic variation (< 10 %).
+Compared to the environmental variables, the slopes of the relationships suggest that very high geographic-based targets are required to generate prioritizations that secure a sufficient proportion of neutral genetic variation for most species ($geq 95$ %).
+Additionally, all species were associated with prioritizations that secured a large proportion of environmental or geographic variation (> 70 %), yet preserved only a very small proportion of the genetic variation (< 10 %; Figures SX and SX; such as _Gypsophila repens_ and _Luzula alpinopilosa_).
 These results suggest while prioritizations generated using surrogate-based targets tend to secure more genetic variation, prioritizations generated using just these targets can result in poor solutions.

### 2016-05-09 code/rmarkdown/figures.Rmd
+ article fixes + remove unneeded comments from MLPE.R
Former-commit-id: ee0043520642de57e7be9a2a46e563d35abc1e75

@@
 p1 <- ggplot(aes(x=surrogate.held,y=genetic.held), data=env.surrogacy.DF) +
-	xlab(expression(atop('Environmental variation','secured (%)'))) + 
+	xlab('Environmental variation secured (%)') + 
 	ylab('Adaptive genetic\nvariation secured (%)') +
@@
 p2 <- ggplot(aes(x=surrogate.held,y=genetic.held), data=geo.surrogacy.DF) +
-	xlab(expression(atop('Geographic variation','secured (%)'))) + 
+	xlab('Geographic variation secured (%)') + 
 	ylab('Neutral genetic\nvariation secured (%)') +
@@
 	geom_histogram(fill='grey10', bins=5) +
-	ylab('\nFrequency') + xlab(expression(atop('Effectiveness of environmental', variables~(R^2~coefficient)))) +
+	ylab('\nFrequency') + xlab(expression(Environmental~variables~(R^2))) +
 	theme_classic() + theme(axis.line.x=element_line(color='black'),
@@
 		plot.background=element_rect(fill='transparent', colour='transparent')) +
-		geom_text(x=0.05, y=0.95, label='(c)') +
+		geom_text(x=0.05, y=6, label='(c)') +
 		coord_cartesian(xlim=c(0, 1))
@@
 	geom_histogram(fill='grey10', bins=10) +
-	ylab('\nFrequency') + xlab(expression(atop('Effectiveness of geographic', variables~(R^2~coefficient)))) +
+	ylab('\nFrequency') + xlab(expression(Geographic~variables~(R^2))) +
 	theme_classic() + theme(axis.line.x=element_line(color='black'),
@@
 		plot.background=element_rect(fill='transparent')) +
-		geom_text(x=0.05, y=0.95, label='(d)') +
+		geom_text(x=0.05, y=8, label='(d)') +
 		coord_cartesian(xlim=c(0, 1))

### 2016-05-09 code/rmarkdown/supporting_information.Rmd
+ article fixes + remove unneeded comments from MLPE.R
Former-commit-id: ee0043520642de57e7be9a2a46e563d35abc1e75

@@
 	# exports
-	return((((n.plots-1) %/% 4)+1) * 2)
+	return((((n.plots-1) %/% 4)+1) * 2.5)
 }
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=2.5, fig.cap='The relationship between the proportion of adaptive genetic variation secured in a prioritizaiton and the proportion of environmental variation it also secures.
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=(2.5*4), fig.cap='The relationship between the proportion of adaptive genetic variation secured in a prioritizaiton and the proportion of environmental variation it also secures.
 Points represent values associated with randomly generated prioritizations.

### 2016-05-09 code/rmarkdown/article.Rmd
+ aritcle + figure fixes + use glmms for scenario analyses
Former-commit-id: f0708b88545153350a42f98c90def8abd858deee

@@
 
+We investigated the benefits of using surrogate-based targets in prioritizations using generalized linear mixed-effects models fit with logit liknks [GLMMs; using the \texttt{lme4 R} package; @469].
+We fit a full model to the proportion of genetic variation secured in a given prioritization.
+This model contained categorical predictor variables to indicate the combination of targets used to generate the prioritizations (amount-only, amount and surrogate targets, and amount and genetic targets), the scenario that was used to generate the prioritization (single-species, multi-species, and multi-species with cost), the type of genetic variation measured (adaptive and neutral), and all interactions between these variables.
+We subject the full model to a backwards stepwise term deletion routine to assess term significance.
+We conducted a post-hoc analysis on the minimal adquate model using Tukey contrasts with a Bonferroni correction [using the \texttt{multcomp R} package; @r470].
+
 ## Results
@@
 Compared to the environmental variables, the slopes of the relationships suggest that very high geographic-based targets are required to generate prioritizations that secure a sufficient proportion of neutral genetic variation for most species ($geq 95$ %).
-Additionally, all species were associated with prioritizations that secured a large proportion of environmental or geographic variation (> 70 %), yet preserved only a very small proportion of the genetic variation (< 10 %; Figures SX and SX; such as _Gypsophila repens_ and _Luzula alpinopilosa_).
+Additionally, all species were associated with prioritizations that secured a large proportion of environmental or geographic variation (> 70 %), yet preserved only a very small proportion of the genetic variation (< 10 %; Figures S55 and S56; such as _Gypsophila repens_ and _Luzula alpinopilosa_).
 These results suggest while prioritizations generated using surrogate-based targets tend to secure more genetic variation, prioritizations generated using just these targets can result in poor solutions.
@@
 <!-- scenario analysis -->
-To determine if the addition of environmental and geographic targets could improve the effectiveness of prioritizations in a more realistic context, we generated prioritizations according to three different scenarios (Figure 3).
-The prioritizations generated under the single-species scenario showed relatively homogeneous selection frequencies across the study area--regardless of the combination of targets used to generate them (Figures 3a, 3b, 3c).
-On the other hand, the prioritizations generated under the multi-species scenario showed different selection frequencies depending on the targets used to generate them.
-The prioritization generated using only amount-based targets contained fewer planning units (Figure 3d; $n = `r sum(selections(multi.spp.prioritisations[[1]]))`$) than the prioritizations generated using additional of surrogate-based targets (Figure 3e; $n = `r sum(selections(multi.spp.prioritisations[[2]]))`$) or genetic-based targets (Figure 3f; $n = `r sum(selections(multi.spp.prioritisations[[3]]))`$).
-The prioritizations generated under the multi-species with cost scenario had different spatial distributions to the previous two scenarios.
-These prioritizations tended to concentrate conservation effort towards areas with higher elevation.
-Similar to the second scenario, the amount-based prioritization contained less panning units (Figure 3g; $n = `r sum(selections(multi.spp.prioritisations.with.cost[[1]]))`$) than either the surrogate-based (Figure 3h; $n = `r sum(selections(multi.spp.prioritisations.with.cost[[2]]))`$) or the genetic-based prioritizations (Figure 3i; $n = `r sum(selections(multi.spp.prioritisations.with.cost[[3]]))`$).
-
-Generally, the prioritizations generated using amount-based targets met the targets for adaptive genetic variation for most species associated with adaptive loci (Figure 3).
-Among the three scenarios, the amount-based prioritization in the single-species scenario had the lowest proportion of species with targets met (`r ` % of species with adaptive loci; Figure 3a).
-In the other two scenarios--the multi-species (`r ` % species preserved with adaptive loci; Figure 3c) and multi-species with cost (`r ` % species preserved with adaptive loci; Figure 3e) scenarios--the amount-based prioritizations successfully preserved the adaptive variation for most of the species.
-The addition of surrogate- and genetic-based targets did little to improve the proportion of species with their targets met for adaptive genetic variation (Figures 3c and 3e).
-These results suggest that the main gradients of intra-specific adaptive variation can be sufficiently preserved using suitable amount-based targets.
-
-The prioritizations generated using additional surrogate-based targets met the neutral genetic variation targets for more species than the prioritizations generated using just amount-based targets (Figure 3).
-The biggest difference between the prioritizations generated using additional surrogate-based targets and just the amount-based targets was in the single-species scenarios (`r ` % and `r ` % of species meeting genetic variation targets respectively).
-In the other two scenarios, the amount-based prioritizations tended to perform slightly better (`r ` % of species with targets met), but were outperformed by prioritizations with additional surrogate-based targets.
-Although the surrogate-based targets tended to fulfill the neutral genetic variation targets for most species, they captured only a very small proportion of neutral genetic variation for two species in particular (_Gypsophila repens_ and _Luzula alpinopilosa_).
-These results suggest that surrogate-based targets can substantially improve the proportion of neutral genetic variation secured in a prioritization.
+To determine if the addition of environmental and geographic targets could improve the effectiveness of prioritizations in a more realistic context, we generated prioritizations according according to three scenarios using different combinations of targets (Figure 3).
+We then computed the proportion of adaptive and neutral genetic variation secured in each prioritization (Figure 4).
+The proportion of genetic variation secured in a prioritization was found to depend on the combination of targets used to generate the prioritization (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r scenario.ANOVA.4.DF$statistic[[2]]`; $P$ `r pvalString(scenario.ANOVA.4.DF$Pr..Chisq.[[2]])`).
+Specifically, the amount-based prioritizations tended to secure less variation than the prioritizations generated using additional surrogate ($Z_1 = `r scenario.GLHT.2.DF$statistic[[1]]`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$) or genetic variables ($Z_1 = `r scenario.GLHT.2.DF$statistic[[2]]`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[2]])`$).
+No evidence was found to suggest that the prioritizations that included additional surrogate- or genetic-based targets secured a different proportion of genetic variation ($Z_1 = `r scenario.GLHT.2.DF$statistic[[3]]`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).
+We also found that the prioritizations tended to secure more adaptive than neutral genetic variation (${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r scenario.ANOVA.3.DF$statistic[[2]]`; $P$ `r pvalString(scenario.ANOVA.3.DF$Pr..Chisq.[[2]])`).
+We also found significant differences in the average proportion of genetic variation secured by prioritizations in each of the scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r scenario.ANOVA.2.DF$statistic[[2]]`; $P$ `r pvalString(scenario.ANOVA.2.DF$Pr..Chisq.[[2]])`).
+Specifically, both the multi-species ($Z_1 = `r scenario.GLHT.1.DF$statistic[[1]]`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$) and multi-species with cost ($Z_1 = `r scenario.GLHT.1.DF$statistic[[2]]`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[2]])`$) scenarios were found to secure more genetic variation than the single species scenario.
+Neither the multi-species scenarios were found to secure a different proportion of genetic variation ($Z_1 = `r scenario.GLHT.1.DF$statistic[[3]]`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[3]])`$).
+These results suggest that surrogate-based targets can substantially improve the proportion of genetic variation secured in a prioritization.
 
 ## Discussion
-<!-- introduction -->
-We aimed to investigate the effectiveness of broad-scale environmental and geographic variables as surrogates for adaptive and neutral genetic variation in conservation planning.
-Our study is the first to address this hypothesis using moderately realistic planning scenarios.
-Overall, we found that environmental and geographic variables are effective surrogates for a few species when developing prioritizations.
-Furthermore, when developing prioritizations under various planning scenarios, we found that the addition of environmental and geographic targets resulted in prioritizations that sufficiently represented the genetic variation of a greater number of species.
-Our results suggest that while the environmental and geographic variables have little benefit for most species, they can make a large difference for some species.
+* introduction
 
-<!-- consequences for conservation planning -->
-+ Typically, conservation planning exercises do not consider intra-specific variation when designing prioritizations. 
-+ our results show that for most species, a sufficient proportion of their intra-specific variation can be secured using suitable amount-based targets alone.
-+ in particular, we found that using suitable amount-based targets in a multi-species context with comprehensive set of target species resulted in a prioritization that adequately represented the adaptive genetic variation for nearly all the target species (panel c; Figure 3).
-+ however, for some species, such a prioritization secures only a small proportion of their genetic variation.
-+ we found that for some species, prioritizations generated using amount-based targets only secured a small amount of their neutral genetic variation (panels b, d, e; Figure 3).
-+ in planning exercises where the decision maker does not have access to genetic data on their target species across the extent of the study area--ie.
-nearly all real world planning scenarios--the decision maker will not know which species are poorly represented in terms of their inta-specific variation.
-+ the poorly represented species may be those of particular significant or they may not.
-+ the decision maker has to weigh the increased costs of a prioritization that includes environmental and geographic surrogates against the risk that a prioritization that does not include these surrogates fails to represent species of conservation significance.
-+ here we found that prioritizations generated using environmental and geographic surrogates have only minor increased costs (compare panels c and d; panels e and f; Figure 3).
-+ overall, these results suggest that although surrogates are unnecessary for most species, they can be highly beneficial for a few species, the increased costs associated with using these surrogates is small, so generally using surrogates is a good idea.
+* consequences for conservation planning
 
-<!-- comparison to previous studies that have investigated genetic studies -->
-+ Previous studies on this data set have found correlations between environmental and genetic variation.
-+ Yet, for most species, prioritizations generated using environmental surrogates did not secure substantially more genetic variation than randomly generated prioritizations with the same number of planning units.
-+ Here we discuss some potential reasons for this disagreement.
-+
+* outlier detection methods - how does our findings agree with other papers?
 
-+ This apparent disagreement can be attributed to multiple reasons.
+* mapping genetic variation - compare with GDMS
 
-+ These results highlight the importance of using prioritizations to assess the effectiveness of potential surrogates for conservation and not simply measuring the strength of correlation between potential surrogates and genetic variation.
-
-<!-- limitations/future directions -->
+* limitations/future directions
 We wish to alert our readers to several limitations associated with our analysis.
@@
 
-### Concluding remarks
+* Concluding remarks
 The overarching aim of our study was to investigate the effectiveness of surrogates in conservation planning.

### 2016-05-09 code/rmarkdown/figures.Rmd
+ aritcle + figure fixes + use glmms for scenario analyses
Former-commit-id: f0708b88545153350a42f98c90def8abd858deee

@@
 	ungroup() %>% 
-	left_join(
-		data.frame(
-			letter=unname(scenario.G.cld$Letters),
-			Combination=as.character(names(scenario.G.cld$Letters))
-		) %>% mutate(Combination=as.character(Combination)),
-		by='Combination'
-	) %>% data.frame()
+	data.frame()
 
@@
 		geom_boxplot(position=position_dodge(0.9)) +
-		geom_text(aes(x=Prioritisation, y=letter.height, label=letter),
-			data=scenario.PDF, position=position_dodge(0.9)) +
 		geom_text(aes(x=0.6, y=1.15, label=panel.letter),
@@
 			axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
-		geom_abline(intercept=rapr.params.LST[[MODE]]$scenario.analysis$genetic.target,
-			slope=0, color='red', linetype='dashed') +
 		facet_grid(Context ~ Metric)

### 2016-05-10 code/rmarkdown/article.Rmd
+ added lme4 deps + article eqn fixes
Former-commit-id: b5a36f0ef9e082c351f81972e5c50d983ff49e4f

@@
 This model contained categorical predictor variables to indicate the combination of targets used to generate the prioritizations (amount-only, amount and surrogate targets, and amount and genetic targets), the scenario that was used to generate the prioritization (single-species, multi-species, and multi-species with cost), the type of genetic variation measured (adaptive and neutral), and all interactions between these variables.
+This model accomodated prioritizations generated for the same species using a random intercept term.
 We subject the full model to a backwards stepwise term deletion routine to assess term significance.
@@
 We then computed the proportion of adaptive and neutral genetic variation secured in each prioritization (Figure 4).
-The proportion of genetic variation secured in a prioritization was found to depend on the combination of targets used to generate the prioritization (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r scenario.ANOVA.4.DF$statistic[[2]]`; $P$ `r pvalString(scenario.ANOVA.4.DF$Pr..Chisq.[[2]])`).
+The proportion of genetic variation secured in a prioritization was found to depend on the combination of targets used to generate the prioritization (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r scenario.ANOVA.4.DF$statistic[[2]]`$; $P$ `r pvalString(scenario.ANOVA.4.DF$Pr..Chisq.[[2]])`).
 Specifically, the amount-based prioritizations tended to secure less variation than the prioritizations generated using additional surrogate ($Z_1 = `r scenario.GLHT.2.DF$statistic[[1]]`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$) or genetic variables ($Z_1 = `r scenario.GLHT.2.DF$statistic[[2]]`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[2]])`$).
 No evidence was found to suggest that the prioritizations that included additional surrogate- or genetic-based targets secured a different proportion of genetic variation ($Z_1 = `r scenario.GLHT.2.DF$statistic[[3]]`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).
-We also found that the prioritizations tended to secure more adaptive than neutral genetic variation (${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r scenario.ANOVA.3.DF$statistic[[2]]`; $P$ `r pvalString(scenario.ANOVA.3.DF$Pr..Chisq.[[2]])`).
-We also found significant differences in the average proportion of genetic variation secured by prioritizations in each of the scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r scenario.ANOVA.2.DF$statistic[[2]]`; $P$ `r pvalString(scenario.ANOVA.2.DF$Pr..Chisq.[[2]])`).
+We also found that the prioritizations tended to secure more adaptive than neutral genetic variation (${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r scenario.ANOVA.3.DF$statistic[[2]]`$; $P$ `r pvalString(scenario.ANOVA.3.DF$Pr..Chisq.[[2]])`).
+We also found significant differences in the average proportion of genetic variation secured by prioritizations in each of the scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r scenario.ANOVA.2.DF$statistic[[2]]`$; $P$ `r pvalString(scenario.ANOVA.2.DF$Pr..Chisq.[[2]])`).
 Specifically, both the multi-species ($Z_1 = `r scenario.GLHT.1.DF$statistic[[1]]`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$) and multi-species with cost ($Z_1 = `r scenario.GLHT.1.DF$statistic[[2]]`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[2]])`$) scenarios were found to secure more genetic variation than the single species scenario.

### 2016-05-10 code/rmarkdown/article.Rmd
+ article fixes
Former-commit-id: abe5044ede72702a2b9c395e88bc9c6b2afd43d9

@@
 Loci identified as outliers in both analyses were treated as adaptive and the rest were treated as neutral.
-To avoid false-positives, we omitted loci from both outlier detection methods where the global frequency of the minor allele was $\geq 0.1$ and assumed they were neutral.
+To avoid false-positives, we omitted loci from both outlier detection methods where the global frequency of the minor allele was $> 0.1$ and assumed they were neutral.
 
@@
 To accommodate variation in the high number of optimal solutions for the single-species amount-based prioritizations, we generated `r (rapr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates)` replicate prioritizations.
-We then counted the proportion of species that were sufficiently represented in terms of their adaptive and neutral genetic variation (genetic variation secured $\geq$ `r (rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)` %) for each prioritization.
+We then counted the proportion of species that were sufficiently represented in terms of their adaptive and neutral genetic variation (genetic variation secured $> `r (rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)`$ %) for each prioritization.
 
@@
 The explanatory power of the environmental surrogates were also generally quite high for most species (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figure S56).
-The nature of these relationships suggest that most of the adaptive variation for a given species can be secured in a prioritization using moderately high environmental-based targets ($geq 80$ %).
+The nature of these relationships suggest that most of the adaptive variation for a given species can be secured in a prioritization using moderately high environmental-based targets ($> 80$ %).
 Similarly, the proportion of neutral genetic variation in a prioritization was affected by the interaction between species and the proportion of neutral variation secured in the prioritization (${\chi^2}_{`r geo.surrogacy.AOV$df[[4]]`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ = `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`).
-The slope for each species were also significantly different to zero (minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`$; maximum $P = `r pvalString(max(env.surrogate.spp.DF$p.value))`; Table S3).
+The slope for each species were also significantly different to zero (minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`; maximum $P = `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3).
 Generally, the explanatory power of geographic variables were higher than that of environmental variables (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2 \pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figure S57).
-Compared to the environmental variables, the slopes of the relationships suggest that very high geographic-based targets are required to generate prioritizations that secure a sufficient proportion of neutral genetic variation for most species ($geq 95$ %).
+Compared to the environmental variables, the slopes of the relationships suggest that very high geographic-based targets are required to generate prioritizations that secure a sufficient proportion of neutral genetic variation for most species ($> 95$ %).
 Additionally, all species were associated with prioritizations that secured a large proportion of environmental or geographic variation (> 70 %), yet preserved only a very small proportion of the genetic variation (< 10 %; Figures S55 and S56; such as _Gypsophila repens_ and _Luzula alpinopilosa_).
@@
 We then computed the proportion of adaptive and neutral genetic variation secured in each prioritization (Figure 4).
-The proportion of genetic variation secured in a prioritization was found to depend on the combination of targets used to generate the prioritization (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r scenario.ANOVA.4.DF$statistic[[2]]`$; $P$ `r pvalString(scenario.ANOVA.4.DF$Pr..Chisq.[[2]])`).
-Specifically, the amount-based prioritizations tended to secure less variation than the prioritizations generated using additional surrogate ($Z_1 = `r scenario.GLHT.2.DF$statistic[[1]]`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$) or genetic variables ($Z_1 = `r scenario.GLHT.2.DF$statistic[[2]]`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[2]])`$).
-No evidence was found to suggest that the prioritizations that included additional surrogate- or genetic-based targets secured a different proportion of genetic variation ($Z_1 = `r scenario.GLHT.2.DF$statistic[[3]]`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).
-We also found that the prioritizations tended to secure more adaptive than neutral genetic variation (${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r scenario.ANOVA.3.DF$statistic[[2]]`$; $P$ `r pvalString(scenario.ANOVA.3.DF$Pr..Chisq.[[2]])`).
-We also found significant differences in the average proportion of genetic variation secured by prioritizations in each of the scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r scenario.ANOVA.2.DF$statistic[[2]]`$; $P$ `r pvalString(scenario.ANOVA.2.DF$Pr..Chisq.[[2]])`).
-Specifically, both the multi-species ($Z_1 = `r scenario.GLHT.1.DF$statistic[[1]]`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$) and multi-species with cost ($Z_1 = `r scenario.GLHT.1.DF$statistic[[2]]`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[2]])`$) scenarios were found to secure more genetic variation than the single species scenario.
-Neither the multi-species scenarios were found to secure a different proportion of genetic variation ($Z_1 = `r scenario.GLHT.1.DF$statistic[[3]]`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[3]])`$).
+The proportion of genetic variation secured in a prioritization was found to depend on the combination of targets used to generate the prioritization (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$Pr..Chisq.[[2]])`)$.
+Specifically, the amount-based prioritizations tended to secure less variation than the prioritizations generated using additional surrogate ($Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]]`,2)$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$) or genetic variables ($Z_1 = `r round(scenario.GLHT.2.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[2]])`$).
+No evidence was found to suggest that the prioritizations that included additional surrogate- or genetic-based targets secured a different proportion of genetic variation ($Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).
+We also found that the prioritizations tended to secure more adaptive than neutral genetic variation (${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.3.DF$Pr..Chisq.[[2]])`$).
+We also found significant differences in the average proportion of genetic variation secured by prioritizations in each of the scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.2.DF$statistic[[2]],2)`$; $P$ `r pvalString(scenario.ANOVA.2.DF$Pr..Chisq.[[2]])`).
+Specifically, both the multi-species ($Z_1 = `r round(scenario.GLHT.1.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$) and multi-species with cost ($Z_1 = `r round(scenario.GLHT.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[2]])`$) scenarios were found to secure more genetic variation than the single species scenario.
+Neither the multi-species scenarios were found to secure a different proportion of genetic variation ($Z_1 = `r round(scenario.GLHT.1.DF$statistic[[3]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[3]])`$).
 These results suggest that surrogate-based targets can substantially improve the proportion of genetic variation secured in a prioritization.

### 2016-05-10 code/rmarkdown/article.Rmd
+ fix
Former-commit-id: 941f2a1c73f6a27e0e2c4150c9e8240e771f5f69

@@
 The proportion of genetic variation secured in a prioritization was found to depend on the combination of targets used to generate the prioritization (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$Pr..Chisq.[[2]])`)$.
-Specifically, the amount-based prioritizations tended to secure less variation than the prioritizations generated using additional surrogate ($Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]]`,2)$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$) or genetic variables ($Z_1 = `r round(scenario.GLHT.2.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[2]])`$).
+Specifically, the amount-based prioritizations tended to secure less variation than the prioritizations generated using additional surrogate ($Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$) or genetic variables ($Z_1 = `r round(scenario.GLHT.2.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[2]])`$).
 No evidence was found to suggest that the prioritizations that included additional surrogate- or genetic-based targets secured a different proportion of genetic variation ($Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).

### 2016-05-10 code/rmarkdown/article.Rmd
+ manuscript fixes
Former-commit-id: 88df6830484ad38e28a669bf3a5581b30a3eba7c

@@
 The nature of these relationships suggest that most of the adaptive variation for a given species can be secured in a prioritization using moderately high environmental-based targets ($> 80$ %).
-Similarly, the proportion of neutral genetic variation in a prioritization was affected by the interaction between species and the proportion of neutral variation secured in the prioritization (${\chi^2}_{`r geo.surrogacy.AOV$df[[4]]`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ = `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`).
-The slope for each species were also significantly different to zero (minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`; maximum $P = `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3).
+Similarly, the proportion of neutral genetic variation in a prioritization was affected by the interaction between species and the proportion of neutral variation secured in the prioritization (${\chi^2}_{`r geo.surrogacy.AOV$df[[4]]`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`).
+The slope for each species were also significantly different to zero (minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`; maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3).
 Generally, the explanatory power of geographic variables were higher than that of environmental variables (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2 \pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figure S57).

### 2016-05-10 code/rmarkdown/figures.Rmd
+ manuscript fixes
Former-commit-id: 88df6830484ad38e28a669bf3a5581b30a3eba7c

@@
 p1 <- ggplot(aes(x=surrogate.held,y=genetic.held), data=env.surrogacy.DF) +
-	xlab('Environmental variation secured (%)') + 
+	xlab('Environmental variation\nsecured (%)') + 
 	ylab('Adaptive genetic\nvariation secured (%)') +
@@
 p2 <- ggplot(aes(x=surrogate.held,y=genetic.held), data=geo.surrogacy.DF) +
-	xlab('Geographic variation secured (%)') + 
+	xlab('Geographic variation\nsecured (%)') + 
 	ylab('Neutral genetic\nvariation secured (%)') +
@@
 The left column of panels (a, c, d) and the right column of panels (b, d, f) show the proportion of adaptive and neutral genetic variation secured (respectively).
-Box plots show the effectiveness of prioritizations generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species.
-Letters denote significant differences ($P < 0.05$).'}
+Box plots show the effectiveness of prioritizations generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species.'}
 # prepare data for plotting

### 2016-05-10 code/rmarkdown/supporting_information.Rmd
+ manuscript fixes
Former-commit-id: 88df6830484ad38e28a669bf3a5581b30a3eba7c

@@
 	}
-)
+) %>% mutate(presence=factor(c('absent', 'present')[presence+1]))
+
 # plot species data
@@
 	theme_classic() +
-	guides(fill=guide_legend(title='Presence')) +
 	theme(
@@
 	) +
+		scale_fill_manual(name='Occupancy', values=c('absent'='#FC8D59', 'present'='#91CF60')) +
 	coord_cartesian(
@@
 	})
-	do.call(grid.arrange, append(curr.GROBS, list(ncol=4)))
+	do.call(grid.arrange, append(curr.GROBS, list(ncol=3)))
 	invisible()
@@
 	# exports
-	return((((n.plots-1) %/% 4)+1) * 2.5)
+	return((((n.plots-1) %/% 4)+1) * 5)
 }

### 2016-05-10 code/rmarkdown/supporting_information.Rmd
+ fixes in tables and figures
Former-commit-id: 1a858c86cb117571ebed1391396f5b82135d58bc

@@
 	# exports
-	return((((n.plots-1) %/% 4)+1) * 5)
+	return((((n.plots-1) %/% 3)+1) * 5)
 }

### 2016-05-10 code/rmarkdown/tables.Rmd
+ fixes in tables and figures
Former-commit-id: 1a858c86cb117571ebed1391396f5b82135d58bc

--- 2016-05-06 code/rmarkdown/tables.Rmd
@@
 		Geographic.corr.P=pvalString(Geographic.corr.P),
+		Geographic.Marginal.R2=pvalString(Geographic.Marginal.R2),
+		Geographic.Conditional.R2=pvalString(Geographic.Conditional.R2),
+		Environmental.Marginal.R2=pvalString(Environmental.Marginal.R2),
+		Environmental.Conditional.R2=pvalString(Environmental.Conditional.R2),
 		blank=rep('', length(Species)),
-		Species=paste0('\\textit{', gsub('\\ ', ' ', Species), '}')) %>%
+		Species=paste0('\\textit{', gsub('\\ ', ' ', Species), '}'),
+		Environmental.Chisq=sapply(round(Environmental.Chisq,2), as.character),
+		Geographic.Chisq=sapply(round(Environmental.Chisq,2), as.character)
+		) %>%
 	select(Species:Environmental.corr.P, Geographic.Marginal.R2:Geographic.corr.P) %>%

### 2016-05-10 code/rmarkdown/article.Rmd
+ fixes in tables and figures
Former-commit-id: 1a858c86cb117571ebed1391396f5b82135d58bc
+ include jonathons comments on ms + fix rounding issue in table
Former-commit-id: ab8e1be37ee244ac4365e131e96194020bacc0b1

@@
 options(error=NULL)
-
 ```
@@
 Recently, conservation scientists have been investigating surrogates to generate prioritizations that adequately preserve genetic variation--but without actually needing to utilize genetic data to do this.
-Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been used to guide reserve selection [eg.
+Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been used to guide reserve selection as surrogates for adaptive genetic variation [eg.
 environmental conditions; @r247; @r2].
@@
 However this remains untested in spatially contiguous systems where connectivity is complicated by additional factors [eg.
-terrain; eg.
-@r474]--the very systems commonly featured in real-world conservation planning exercises [eg.
+terrain or habitat type; @r474; @r497]--the very systems commonly featured in real-world conservation planning exercises [eg.
 @r197].
-Before conservation planners can reliably use surrogates for genetic data in conservation planning, the effectiveness of these surrogates needs to be verified.
+Before conservation planners can reliably use surrogates for genetic data in conservation planning, the effectiveness of these surrogates need to be verified.
 Otherwise, conservation planners may waste precious resources on protecting places that provide little benefit to biodiversity.
@@
 
-To describe the geographic location of each grid cell, we projected the grid into an equi-distant coordinate system ($\texttt{ESRI:102031}$), calculated their centroids, and extracted their two-dimensional coordinates.
+To describe the geographic location of each grid cell, we projected the grid into an equi-distant coordinate system (Europe Equidistant Conic: $\texttt{ESRI:102031}$), calculated their centroids, and extracted their two-dimensional coordinates.
 Thus we constructed two-dimensional geographic space--wherein each planning unit occupies a single point--as a potential surrogate for neutral genetic variation.
@@
 We used this data set because it is commonly to map habitat suitability in conservation planning [@r493].
-These layers and the planning units were projected into an equal-area coordinate system ($\texttt{ESRI:102014}$).
-To reduce dimensionality, the 19 bioclimatic variables were clipped to the planning units and subject to a principal components analysis (PCA; using \texttt{ArcMap} version 10.2.2; Table S1).
+These layers and the planning units were projected into an equal-area coordinate system (Europe Lambert Conformal Conic: $\texttt{ESRI:102014}$).
+To reduce dimensionality, the 19 bioclimatic variables were clipped to the planning units and subjected to a principal components analysis (PCA; using \texttt{ArcMap} version 10.2.2; Table S1).
 The first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components (PCs) cumulatively explained `r round(pca.DF[[4]][surrogate.params.LST[[MODE]]$number.components],1)` % of the variation and were used to generate new layers.
@@
 
-To address the second aim--to determine if the environmental and geographic variables are effective surrogates for adaptive and neutral genetic variation--we generated a collection of single-species prioritizations.
+To address the second aim--to determine if the environmental and geographic variables are effective surrogates for adaptive and neutral genetic variation--we generated a collection of single-species prioritizations and computed the proportion of intra-specific genetic variation they secured.
 For each species, we generated `r rapr.params.LST[[MODE]]$surrogacy.analysis$replicates` prioritizations by randomly selecting different combinations of planning units that were occupied by the species.
@@
 First, the size of the planning units we used (23 $\times$ 25 km) is much larger than typically used in conservation planning scenarios (between 1--10 km^2).
-We used this resolution because the genetic data was collected at this scale.
+We used this resolution because the genetic data were collected at this scale.
 Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis because this process would have  artificially introduced additional spatial auto-correlation into the dataset.

### 2016-05-10 code/rmarkdown/tables.Rmd
+ include jonathons comments on ms + fix rounding issue in table
Former-commit-id: ab8e1be37ee244ac4365e131e96194020bacc0b1

--- 2016-05-06 code/rmarkdown/tables.Rmd
@@
 		Geographic.corr.P=pvalString(Geographic.corr.P),
+		Geographic.Marginal.R2=pvalString(Geographic.Marginal.R2),
+		Geographic.Conditional.R2=pvalString(Geographic.Conditional.R2),
+		Environmental.Marginal.R2=pvalString(Environmental.Marginal.R2),
+		Environmental.Conditional.R2=pvalString(Environmental.Conditional.R2),
 		blank=rep('', length(Species)),
 		Species=paste0('\\textit{', gsub('\\ ', ' ', Species), '}')) %>%
+	mutate(Environmental.Chisq=sapply(round(Environmental.Chisq,2), as.character)) %>%
+	mutate(Geographic.Chisq=sapply(round(Geographic.Chisq,2), as.character)) %>%
 	select(Species:Environmental.corr.P, Geographic.Marginal.R2:Geographic.corr.P) %>%

### 2016-05-10 code/rmarkdown/supporting_information.Rmd
+ bug fixes
Former-commit-id: 4f9d4d26086f4c1ab7bc13ccb3ec3a7b8c050629

@@
 	# exports
-	return((((n.plots-1) %/% 3)+1) * 5)
+	return((((n.plots-1) %/% 3)+1) * 2.5)
 }

### 2016-05-10 code/rmarkdown/supporting_information.Rmd
+ fix nmds plots in SI
Former-commit-id: d37e8033bac4b94194105497c2e6a294c8b0e0aa

@@
 
+calc.fig.width <- function(i) {
+	# init
+	adapt.col <- paste0(unique(spp.samples.DF$species)[i], '_adaptive_d1')
+	neutral.col <- paste0(unique(spp.samples.DF$species)[i], '_neutral_d1')
+	n.plots <- 0 
+	# count adaptive plot
+	if (adapt.col %in% names(grid.FPLY))
+		n.plots <- n.plots + spp.nmds.LST[[i]][['adaptive']]$ndim
+	# count neutral plot
+	if (neutral.col %in% names(grid.FPLY))
+		n.plots <- n.plots + spp.nmds.LST[[i]][['neutral']]$ndim
+	# exports
+	if (n.plots > 3) n.plots <- 3
+	return((10/3) * n.plots)
+}
+
 is.speciesPlottable <- function(i) {
@@
 
-```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(1), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[1]),'}.
+```{r, echo=FALSE, fig.width=calc.fig.width(1), fig.height=calc.fig.height(1), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[1]),'}.
 Each panel shows a different dimension of the variation as indicated by its title.
@@
 
-```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(2), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[2]),'}.
+```{r, echo=FALSE, fig.width=calc.fig.width(2), fig.height=calc.fig.height(2), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[2]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(3), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[3]),'}.
+```{r, echo=FALSE, fig.width=calc.fig.width(3), fig.height=calc.fig.height(3), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[3]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(4), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[4]),'}.
+```{r, echo=FALSE, fig.width=calc.fig.width(4), fig.height=calc.fig.height(4), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[4]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(5), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[5]),'}.
+```{r, echo=FALSE, fig.width=calc.fig.width(5), fig.height=calc.fig.height(5), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[5]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(6), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[6]),'}.
+```{r, echo=FALSE, fig.width=calc.fig.width(6), fig.height=calc.fig.height(6), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[6]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(7), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[7]),'}.
+```{r, echo=FALSE, fig.width=calc.fig.width(7), fig.height=calc.fig.height(7), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[7]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(8), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[8]),'}.
+```{r, echo=FALSE, fig.width=calc.fig.width(8), fig.height=calc.fig.height(8), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[8]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(9), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[9]),'}.
+```{r, echo=FALSE, fig.width=calc.fig.width(9), fig.height=calc.fig.height(9), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[9]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(10), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[10]),'}.
+```{r, echo=FALSE, fig.width=calc.fig.width(10), fig.height=calc.fig.height(10), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[10]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(11), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[11]),'}.
+```{r, echo=FALSE, fig.width=calc.fig.width(11), fig.height=calc.fig.height(11), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[11]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(12), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[12]),'}.
+```{r, echo=FALSE, fig.width=calc.fig.width(12), fig.height=calc.fig.height(12), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[12]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(13), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[13]),'}.
+```{r, echo=FALSE, fig.width=calc.fig.width(13), fig.height=calc.fig.height(13), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[13]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(14), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[14]),'}.
+```{r, echo=FALSE, fig.width=calc.fig.width(14), fig.height=calc.fig.height(14), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[14]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(15), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[15]),'}.
+```{r, echo=FALSE, fig.width=calc.fig.width(15), fig.height=calc.fig.height(15), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[15]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(16), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[16]),'}.
+```{r, echo=FALSE, fig.width=calc.fig.width(16), fig.height=calc.fig.height(16), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[16]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(17), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[17]),'}.
+```{r, echo=FALSE, fig.width=calc.fig.width(17), fig.height=calc.fig.height(17), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[17]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(18), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[18]),'}.
+```{r, echo=FALSE, fig.width=calc.fig.width(18), fig.height=calc.fig.height(18), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[18]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(19), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[19]),'}.
+```{r, echo=FALSE, fig.width=calc.fig.width(19), fig.height=calc.fig.height(19), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[19]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(20), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[20]),'}.
+```{r, echo=FALSE, fig.width=calc.fig.width(20), fig.height=calc.fig.height(20), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[20]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(21), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[21]),'}.
+```{r, echo=FALSE, fig.width=calc.fig.width(21), fig.height=calc.fig.height(21), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[21]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(22), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[22]),'}.
+```{r, echo=FALSE, fig.width=calc.fig.width(22), fig.height=calc.fig.height(22), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[22]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(23), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[23]),'}.
+```{r, echo=FALSE, fig.width=calc.fig.width(23), fig.height=calc.fig.height(23), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[23]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(24), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[24]),'}.
+```{r, echo=FALSE, fig.width=calc.fig.width(24), fig.height=calc.fig.height(24), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[24]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(25), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[25]),'}.
+```{r, echo=FALSE, fig.width=calc.fig.width(25), fig.height=calc.fig.height(25), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[25]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(26), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[26]),'}.
+```{r, echo=FALSE, fig.width=calc.fig.width(26), fig.height=calc.fig.height(26), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[26]),'}.
 See Figure S6 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=10, fig.height=calc.fig.height(27), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[27]),'}.
+```{r, echo=FALSE, fig.width=calc.fig.width(27), fig.height=calc.fig.height(27), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[27]),'}.
 See Figure S6 caption for conventions.')}

### 2016-05-10 code/rmarkdown/article.Rmd
+ article fixes
Former-commit-id: 68074f643424c679603091286768045e7e9f377b

@@
 The ordinations resulting from the NMDS analyses were used to construct an adaptive (if present) and neutral genetic space for each species.
-The spatial distribution of these genetic spaces (Figures S29--S54) generally show patterns of spatial auto-correlation--especially the adaptive genetic spaces (eg.XXXX; SXX).
+The spatial distribution of these genetic spaces (Figures S29--S54) generally show patterns of spatial auto-correlation--especially the adaptive genetic spaces (eg.
+_Cerastium uniflorum_ and _Dryas octopetala_).
 Additionally, most species show east-west patterns in their intra-specific genetic variation (eg.
-XXXX; SXXX).
+_Peucedanum ostruthium_ and _Phyteuma betonicifolium_).
 

### 2016-05-11 code/rmarkdown/article.Rmd
+ improvements to main text - typos and rewriting sections
Former-commit-id: f9b83a03b9c3a961f850aecf61292aa5ec2ec647

@@
 Thus protected areas need to be sited in places that achieve conservation goals for minimal cost.
-To ensure that protected areas are cost-effective, plans for protected areas (prioritizations) are often generated using decision support tools to identify near-optimal solutions.
-Additionally, because reserves cannot preserve the entire range for a given species, the plans for multiple reserves are often designed simultaneously to ensure that the network of protected areas can facilitate biological processes (eg.
-gene flow between populations, seasonal migration).
+To ensure that the plans for protected areas (prioritizations) are cost-effective, they are often generated using decision support tools to identify near-optimal solutions.
+These prioritizations often contain plans for a network of multiple protected areas so they can preserve biological processes (eg.
+connectivity between populations).
 For instance, \texttt{Marxan} [@r22] uses amount-based targets to identify protected area networks that preserve a suitable amount of habitat for each species (eg.
-100 km $^2$ of land occupied by the species needs to be secured in a reserve) and weights to penalize fragmented solutions.
+100 km $^2$ of land occupied by the species needs to be secured in a reserve) and weights to penalize fragmented solutions (ie.
+the \textit{BLM} parameter).
 However, conservation planning exercises typically assume that all individuals in the same species are equivalent.
 
-Intra-specific genetic variation is an important for long-term species persistence.
+Intra-specific genetic variation is important for long-term species persistence.
 As a consequence, there has been increasing interest in designing prioritizations that preserve and cultivate this variation [@r67; @r3; @r61].
@@
 Neutral variation often reflects the evolutionary history of different populations.
-By preserving existing patterns of neutral variation, protected areas can secure different genetic lineages and avoid the adverse effects of low genetic diversity [eg.
+By preserving existing patterns of neutral variation, protected areas can safeguard against the adverse effects of low genetic diversity [eg.
 inbreeding depression; @r67].
-Thus protected area networks should preserve both adaptive and neutral patterns of genetic variation for species of conservation interest [@r67; @r3].
-Despite this, genetic data are not often used to inform prioritizations because they require considerable investment to obtain and analyze [@r61; but see @r5 and @r157 for developments on using genetic data in conservation planning].
-Thus prioritizations may fail to capture species' intra-specific genetic variation, and ultimately fail to deliver long-term conservation outcomes.
+Thus protected area networks should aim to preserve both adaptive and neutral patterns of genetic variation for species of conservation interest [@r67; @r3].
+Despite this, genetic data are not often used to inform conservation planning exercises because such data require considerable investment to obtain and analyze [@r61; but see @r5 and @r157 for developments on using genetic data in conservation planning]--especially since most large-scale exercises involve hundreds of species.
+Thus prioritizations may fail to secure a sufficient proportion of the intra-specific genetic variation of  species, and ultimately fail to deliver long-term conservation outcomes.
 
-Recently, conservation scientists have been investigating surrogates to generate prioritizations that adequately preserve genetic variation--but without actually needing to utilize genetic data to do this.
+Recently, conservation scientists have been investigating the use of surrogate data to generate prioritizations that adequately preserve intra-specific genetic variation--but without actually needing to utilize genetic data to acheive this.
 Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been used to guide reserve selection as surrogates for adaptive genetic variation [eg.
 environmental conditions; @r247; @r2].
-By preserving a more representative sampled of environmental conditions across the species' ranges, they aimed to generate prioritizations that preserved more adaptive variation.
+By preserving a more representative sample of environmental conditions across the species' ranges, these researchers aimed to generate prioritizations that would preserve more adaptive variation.
 However, the effectiveness of this approach remains unverified.
@@
 First, we aim to verify if environmental and geographic variables correlate with broad-scale patterns in adaptive and neutral variation (respectively).
-Although these variables may correlate with genetic variation, they may not be effective surrogates if the level of correlation is not high enough or due to spatial auto-correlative effects.
+Although these variables may correlate with genetic variation, they may not be effective surrogates if they have poor predictive ability or if these correlations are driven by spatial auto-correlation.
 Second, we aim to verify that environmental and geographic variables are effective surrogates for adaptive and neutral variation (respectively).
@@
 We expect that prioritizations generated using amount- and surrogate-based targets will sufficiently preserve the genetic variation of more species than those generated using only amount-based targets.
+Furthermore, we aim to test this hypothesis under several of different scenarios to see whether the additional constraints of real-world planning exercises exacerbate or nullify this effect.
 If these surrogates can improve the intra-specific genetic variation captured in a prioritization, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.
@@
 ### Study system
-We used species distribution and genomic data collected by the IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for further explanation of data collection methods].
+We used species distribution and genomic data collected by the international IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for further explanation of data collection methods].
 This data set contains information for 27 alpine plant species (Figure 1a).
@@
 They genotyped samples using amplified fragment length polymorphisms [AFLP; @r453], and constructed matrices denoting the presence/absence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; see Table 1 in @r478 for the number markers for each species).
-Although AFLPs are not ideal as genomic representatives (cf.
-single nucleotide polymorphisms; SNPs), no multi-species genomic data set with similar properties currently exists.
 Thus the data set contains information describing the genomic properties of individuals for each species present in every second grid cell.
@@
 First, conservation planning exercises typically consider tens to hundreds of species with different evolutionary histories and life histories.
-To ensure that our results were relevant for conservation planners, we required a similar sized data set.
-This data set contains genetic data for a diverse range of plant species.
-Second, we required genetic data for all species to be collected under a standardized sampling scheme, and this data set used a grid to collect samples.
-Third, we required broad-scale environmental variation across the study area to test the effectiveness of environmental variables.
-This data set covers a large geographic domain that spans across attitudinal gradients.
-Thus this data set was particularly well suited for our study.
+To ensure that our results were relevant for conservation planners, we required a data set with a diverse set of species and a comparable number of species.
+The IntraBioDiv data set contains genetic data for 27 species with different evolutionary histories and life-history strategies.
+Second, we required genetic data for all species to be collected using a standardized sampling scheme, and this data set used a grid to collect samples.
+Third, we required the study area to contain broad-scale environmental variation to test the effectiveness of environmental variables.
+This data set spans across large attitudinal gradients.
+Although this data set contains AFLPs  which not ideal as genomic representatives (cf.
+single nucleotide polymorphisms; SNPs), no multi-species genomic data set with similar properties currently exists.
+Thus this data set was well suited for our study.
 
@@
 Taberlet _et al._ [-@r486] investigated whether prioritizations generated using the spatial distribution of a comprehensive set species (> 300) could adequately secure the intra-specific genetic variation for 27 species.
-They found that the prioritization based on species distributions was substantially different to that based on genomic data.
+They found that the prioritization based on species distributions was substantially different to the prioritization based on genomic data.
 Their results suggest that prioritizations based on species distribution data alone will be unable to secure intra-specific genetic variation.
+However, whether or not environmental and geographic variables can augment prioritizations remains untested.
 
@@
 We examined the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation (respectively).
-We used the grid cells used to collect samples by the IntraBioDiv project as planning units to develop prioritizations.
+We utilized the sampling grid used by the IntraBioDiv project as planning units to develop prioritizations.
 We used only planning units that contained at least a single individual ($n = `r nrow(grid.DF)`$) to reduce computational burden.
-To examine the potential effects of cost, we calculated the total human population density inside each planning unit [obtained at 1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472] and used this to represent acquisition cost (Figure 1b).
+To examine the potential effects of variation in cost (see below), we calculated the total human population density inside each planning unit [obtained at 1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472] and used this to represent acquisition cost (Figure 1b).
 
-To describe the geographic location of each grid cell, we projected the grid into an equi-distant coordinate system (Europe Equidistant Conic: $\texttt{ESRI:102031}$), calculated their centroids, and extracted their two-dimensional coordinates.
+To describe the geographic location of each planning unit, we projected the grid into an equi-distant coordinate system (Europe Equidistant Conic: $\texttt{ESRI:102031}$), calculated their centroids, and extracted their two-dimensional coordinates.
 Thus we constructed two-dimensional geographic space--wherein each planning unit occupies a single point--as a potential surrogate for neutral genetic variation.
+We expected planning units further apart from each other in this geographic space to contain individuals with increasingly different patterns in their neutral loci.
 
-To describe the environmental characteristics of each grid cell, we used broad-scale climatic variables.
+To describe the environmental characteristics of each planning unit, we used broad-scale climatic variation.
 We obtained data for 19 bioclimatic variables across the extent of the study area [approx.
@@
 Thus we constructed a three-dimensional environmental space--wherein each planning unit occupies a single point--as a potential surrogate for adaptive variation.
+We expect planning units further apart from each other in this environmental space to contain individuals with increasingly different patterns in their adaptive loci.
 
 ### Adaptive and neutral genetic data
-To investigate the effectiveness of the surrogates for adaptive and neutral genetic variation, we first needed to identify which of the sampled loci--if any--were under selection.
-We used two outlier detection methods to achieve this [as recomended by @r495].
+To investigate the effectiveness of our supposed surrogates for adaptive and neutral genetic variation, we first needed to identify which of the sampled loci--if any--were adaptive.
+We used two outlier detection methods to achieve this [an individual- and a population-level method as recommended by @r495].
 We omitted the species \textit{`r gsub('_', ' ' , missing.species, fixed=TRUE)`} from all subsequent analyses because it was not compatible with one of the outlier detection methods (see below).
@@
 We adopted a similar methodology to Bothwell _et al._ [-@r455] and applied it to each of the species in the data set.
-To avoid falsely classifying loci as adaptive, we grouped individuals into genetic lineages for each species (Figures S3--S28).
-We used the number of genetic lineages in each species previously identified by Alvarez _et al._ [-@r478].
-Since \textit{`r gsub('_', ' ' , missing.species, fixed=TRUE)`} was previously identified as containing only one population, it was not compatible with \texttt{BayeScan}.
+Following this methodology, for each species, we first grouped individuals into genetic lineages to avoid falsely classifying loci as adaptive.
+We used the number of genetic lineages identified in a previous analysis of the data set [@r478].
+Since \textit{`r gsub('_', ' ' , missing.species, fixed=TRUE)`} was found to only contain a single lineage/population, it was not compatible with these models.
 For each species, we used \texttt{Structure} [version 2.3.4; using same parameters as @r478 per run with `r structure.params.LST[[MODE]]$numruns` runs per species] and \texttt{ClumPP} [version 1.1.2; @r459] to group individuals into populations.
@@
 
-Second, we used PCAs to identify outlier loci using methods implemented in the \texttt{pcadapt R} package [version 3.0; @r487; @r496].
+Second, we used PCAs to identify outlier loci [using methods implemented in the \texttt{pcadapt R} package; version 3.0; @r487; @r496].
 To enable comparisons between the two outlier detection analyses, we used the same individuals in this analysis as used in the \texttt{BayeScan} analysis.
 For each of the 26 species, we first imputed missing data to permit further analysis.
-Specifically, the value for a given individual at a given locus was replaced with the mean frequency of that locus among all conspecifics.
+Specifically, the missing value for a given locus for a given individual was replaced with the average frequency of that locus among all conspecifics.
 Next, we ran a PCA and extracted the minimum number of components needed to secure the over-arching population level variation among the loci (`r pcadapt.params.LST[[MODE]]$min.variance.explained*10` % of the variation in loci).
 We then used Mahalanobis distances [@r398] to compute _P_ values and then q-values [using the \texttt{qvalue R} package; @r491].
-Outlier loci were detected using a suitable FDR ($q < `r pcadapt.params.LST[[MODE]]$alpha.level`$).
+Outlier loci were detected using a suitable false discovery rate ($q < `r pcadapt.params.LST[[MODE]]$alpha.level`$).
 
@@
 For each species, we discarded the population groupings, and partitioned adaptive and neutral loci into separate matrices.
-We used non-metric multi-dimensional scaling [NMDS; implemented in the \texttt{vegan R} package; @r458] using Gower distances [@r464; using the \texttt{cluster R} package to accommodate sparsity; @r457] to derive continuous variables that described the main gradients of adaptive and neutral genetic variation separately for each species (Table S2).
-To ensure that the ordinations described a sufficient amount of the genomic variation, we ran successive scaling analyses with increasing dimensionality until this was achieved (maximum stress value $\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for each analysis).
+We applied separate non-metric multi-dimensional scaling analyses to each matrix [NMDS; implemented in the \texttt{vegan R} package; @r458] using Gower distances [@r464; using the \texttt{cluster R} package to accommodate sparsity; @r457] to derive continuous variables that described the main gradients of adaptive and neutral genetic variation (separately) for each species (Table S2).
+To ensure that the ordinations described a sufficient amount of the genomic variation, we ran successive scaling analyses with increasing dimensionality until a sufficient stress value was obtained (maximum stress value $\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for each analysis).
 Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the individuals in each grid cell to express the typical genomic characteristics of individuals in the cells.
 
-The previous analyses resulted in an adaptive (if detected) and neutral genetic space for each species.
+The previous analysis resulted in an adaptive (if detected) and neutral genetic space for each species.
 For a given species, each planning unit occupied by the species was associated with a multi-dimensional point in the species' neutral genetic space.
-Planning units that were closer together in this space are occupied by individuals with similar neutral genetic loci.
+Planning units that were closer together in this space are occupied by individuals with similar polymorphisms in their neutral loci.
 Similarly, if the species was associated with adaptive loci, each planning unit occupied by the species was associated with a multi-dimensional point in the species' adaptive genetic space.
-By spreading out conservation effort across a given genetic space for a given species, prioritizations can secure more genetic variation [see @r465 for discussion on attribute spaces].
+By spreading out conservation effort across a given genetic space, and in turn securing planning units occupied by individuals with increasingly different polymorphisms, prioritizations can secure more genetic variation [see @r465 for discussion on attribute spaces].
 
@@
 To address the first aim--to determine if the environmental and geographic variables correlate with spatial patterns of adaptive and neutral genetic variation--we computed dissimilarity matrices between the planning units, and tested for correlations between matrices based on surrogate and genetic data.
-For each species, we computed two dissimilarity matrices between the planning units occupied by the species in terms of their environmental conditions (the PC values from the bioclimatc data) and geographic position (the z-scaled XY coordinates of the units' centroids).
+For each species, we computed two dissimilarity matrices between the planning units occupied by the species in terms of their environmental conditions (their coordinates in the environmental space) and geographic position (their coordinates in the geographic space).
 We then computed another dissimilarity matrix between the planning units using their values in the species' neutral genetic space.
@@
 
-We then fit maximum likelihood population-effects (MLPE) models [implemented in the \texttt{ResistanceGA R} package; @r488].
-These models explicitly accommodate the structure of dissimilarity matrices using random effects [@r494].
+We then fit maximum likelihood population-effects (MLPE) models [@r494; implemented in the \texttt{ResistanceGA R} package; @r488].
+These models explicitly accommodate the structure of dissimilarity matrices using random effects.
 For each species, we fit a MLPE model to correlate the species' dissimilarity matrices based on geographic position and neutral genetic variation.
-Additionally, if the species was associated with adaptive loci, we also fit a MLPE model to correlate the species' dissimilarity matrices calculated using enviuronmental conditions and adaptive genetic variation.
-To test if the surrogate variables explained a substantial amount of genetic variation, we conducted $\chi^2$ tests between each model and its corresponding null model, and Bonferroni corrections.
+Additionally, if the species was associated with adaptive loci, we also fit a MLPE model to correlate the species' dissimilarity matrices calculated using environmental conditions and adaptive genetic variation.
+All predictor variables were z-transformed prior to model fitting to improve convergence.
+To test if the surrogate variables explained a substantial amount of genetic variation, we conducted $\chi^2$ tests between each model and its corresponding null model, and applied Bonferroni corrections.
 
@@
 Similar to the previous model, this model also contained a predictor variable indicating the species for which the prioritizations were generated, and an interaction term.
-We subject these models to a backwards step-wise term deletion routine to assess term significance.
-To assess the performance of these surrogates, we fit a suite of models correlating the proportion of adaptive and neutral genetic variation held in each prioritization and the environmental and geographic variation they also secured (respectively).
+We subjected these models to a backwards step-wise term deletion routine to assess term significance.
+To gauge the performance of these surrogates for each species, we fit a suite of models correlating the proportion of adaptive and neutral genetic variation held in each prioritization and the environmental and geographic variation they also secured (respectively).
 We then computed the Cragg and Uhler's pseudo-R$^2$ value for each model [using the \texttt{pscl} R package; @r492].
 
-To address the third aim--to determine if the use of surrogate-based targets can actually improve prioritizations in real-world conservation planning scenarios--we investigated the effectiveness of surrogate-based targets using three scenarios with increasing levels of complexity.
+To address the third aim--to determine if the use of surrogate-based targets can actually improve prioritizations--we investigated the effectiveness of surrogate-based targets under three scenarios with increasing levels of complexity.
 To assess the benefits of surrogate-based targets under the simplest of circumstances, our first scenario involved generating single-species prioritizations for each species.
@@
 For each scenario, we generated three types of prioritizations using (1) only amount-based targets, (2) both amount- and surrogate-based targets, and (3) both amount- and genetic-based targets.
-The purely amount-based prioritizations served as our baseline--representing the prioritizations typically generated in conservation planning exercises.
+The purely amount-based prioritizations served as our baseline--representing the prioritizations typically generated in conservation planning exercises (eg.
+by \texttt{Marxan}).
 We used `r (rapr.params.LST[[MODE]]$scenario.analysis$amount.target*100)` % amount-based targets in all prioritizations to ensure an adequate proportion of habitat was secured for each species.
-Based on the previous analysis, we set surrogate-based targets as `r (rapr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100)` % and genetic-based targets as `r (rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)` % to ensure that most of the variation was secured for each species.
-To accommodate variation in the high number of optimal solutions for the single-species amount-based prioritizations, we generated `r (rapr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates)` replicate prioritizations.
-We then counted the proportion of species that were sufficiently represented in terms of their adaptive and neutral genetic variation (genetic variation secured $> `r (rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)`$ %) for each prioritization.
+Based on the results from the analyses used to address our second aim, we set surrogate-based targets as `r (rapr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100)` % and genetic-based targets as `r (rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)` % to ensure that most of the genetic variation was secured for each species.
+To accommodate the fact that there were a large number of optimal solutions for the single-species amount-based prioritizations, we generated `r (rapr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates)` replicate prioritizations.
+We computed the proportion of adaptive (if present) and neutral genetic variation secured for each species in the prioritizations.
 
@@
 This model contained categorical predictor variables to indicate the combination of targets used to generate the prioritizations (amount-only, amount and surrogate targets, and amount and genetic targets), the scenario that was used to generate the prioritization (single-species, multi-species, and multi-species with cost), the type of genetic variation measured (adaptive and neutral), and all interactions between these variables.
-This model accomodated prioritizations generated for the same species using a random intercept term.
-We subject the full model to a backwards stepwise term deletion routine to assess term significance.
-We conducted a post-hoc analysis on the minimal adquate model using Tukey contrasts with a Bonferroni correction [using the \texttt{multcomp R} package; @r470].
+This model accommodated prioritizations generated for the same species using a random intercept term.
+We subject the full model to a backwards step-wise term deletion routine to assess term significance.
+We conducted a post-hoc analysis on the minimal adequate model using Tukey contrasts with a Bonferroni correction [using the \texttt{multcomp R} package; @r470].
 
@@
 The outlier detection methods detected adaptive genetic variation in `r as.character(as.english(length(env.surrogacy.LST)))` species.
-Generally, two outlier detection methods gave fairly consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD).
+Generally, the two outlier detection methods gave fairly consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD).
 Of the species that were associated with adaptive genetic variation, only a few loci were detected as under selection (mean `r round(mean(spp.outlier.loci.props, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.loci.props, na.rm=TRUE),2)` SD).
@@
 For all species in which we detected adaptive loci, only a couple of continuous dimensions were needed to sufficiently describe their patterns of adaptive genetic variation (Table S2).
-On the other hand, a few additional continuous dimensions were often needed to describe their patterns of neutral genetic variation (minimum $K = `r min(neutral.k)`$; mean $K = `r round(mean(neutral.k),2)`$; max $K = `r max(neutral.k)`$; Table S2).
+On the other hand, a few additional continuous dimensions were occasionally needed to describe their patterns of neutral genetic variation (minimum $K = `r min(neutral.k)`$; mean $K = `r round(mean(neutral.k),2)`$; max $K = `r max(neutral.k)`$; Table S2).
 The ordinations resulting from the NMDS analyses were used to construct an adaptive (if present) and neutral genetic space for each species.
@@
 _Cerastium uniflorum_ and _Dryas octopetala_).
-Additionally, most species show east-west patterns in their intra-specific genetic variation (eg.
+Additionally, several species show east-west patterns in their intra-specific genetic variation (eg.
 _Peucedanum ostruthium_ and _Phyteuma betonicifolium_).
@@
 Generally, the explanatory power of geographic variables were higher than that of environmental variables (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2 \pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figure S57).
-Compared to the environmental variables, the slopes of the relationships suggest that very high geographic-based targets are required to generate prioritizations that secure a sufficient proportion of neutral genetic variation for most species ($> 95$ %).
-Additionally, all species were associated with prioritizations that secured a large proportion of environmental or geographic variation (> 70 %), yet preserved only a very small proportion of the genetic variation (< 10 %; Figures S55 and S56; such as _Gypsophila repens_ and _Luzula alpinopilosa_).
-These results suggest while prioritizations generated using surrogate-based targets tend to secure more genetic variation, prioritizations generated using just these targets can result in poor solutions.
+Compared to the environmental variables, the slopes of these relationships suggest that very high geographic-based targets are required to generate prioritizations that secure a sufficient proportion of neutral genetic variation for most species ($> 95$ %).
+Additionally, all species were associated with prioritizations that secured a large proportion of environmental or geographic variation (> 70 %), yet preserved only a very small proportion of their intra-specific genetic variation (< 10 %; Figures S55 and S56; such as _Gypsophila repens_ and _Luzula alpinopilosa_).
+These results suggest while prioritizations generated using environmental and geographic targets tend to secure more genetic variation, prioritizations generated using only these targets can sometimes secure only a small fraction of intra-specific genetic variation.
 
 <!-- scenario analysis -->
-To determine if the addition of environmental and geographic targets could improve the effectiveness of prioritizations in a more realistic context, we generated prioritizations according according to three scenarios using different combinations of targets (Figure 3).
+To determine if the addition of environmental and geographic targets could improve the effectiveness of prioritizations in a more realistic context, we generated prioritizations according to three scenarios using different combinations of targets (Figure 3).
 We then computed the proportion of adaptive and neutral genetic variation secured in each prioritization (Figure 4).
@@
 Neither the multi-species scenarios were found to secure a different proportion of genetic variation ($Z_1 = `r round(scenario.GLHT.1.DF$statistic[[3]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[3]])`$).
-These results suggest that surrogate-based targets can substantially improve the proportion of genetic variation secured in a prioritization.
+These results suggest that surrogate-based targets can substantially improve the proportion of intra-specific genetic variation secured in a prioritization for most species.
 
@@
 The overarching aim of our study was to investigate the effectiveness of surrogates in conservation planning.
-We found that environmental and geographic surrogates are effective.
-However, they seem to be largely unnecessary in conservation planning scenarios because a comprehensive set of species can be used to achieve a similar result.
-Conservation planners are therefore urged to use a comprehensive set of species in prioritizations--one of the fundamental principals of conservation planning.
+We found that environmental and geographic are generally effective surrogates for most species.
+Although prioritizations generated using amount-only targets tended to secure most of the adaptive variation for most species, the addition of surrogate targets to the planning process substantially improved the proportion of intra-specific neutral genetic variation.
+Conservation planners are therefore urged to include these surrogates in conservation planning exercises.
 

### 2016-05-11 code/rmarkdown/article.Rmd
+ fix for stats in article
Former-commit-id: ca2b26947e16991e13f5023b668df5178cf80b76

@@
 The proportion of genetic variation secured in a prioritization was found to depend on the combination of targets used to generate the prioritization (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$Pr..Chisq.[[2]])`)$.
-Specifically, the amount-based prioritizations tended to secure less variation than the prioritizations generated using additional surrogate ($Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$) or genetic variables ($Z_1 = `r round(scenario.GLHT.2.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[2]])`$).
-No evidence was found to suggest that the prioritizations that included additional surrogate- or genetic-based targets secured a different proportion of genetic variation ($Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).
-We also found that the prioritizations tended to secure more adaptive than neutral genetic variation (${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.3.DF$Pr..Chisq.[[2]])`$).
+Specifically, the amount-based prioritizations tended to secure less variation than the prioritizations generated using additional surrogate (`r round(plogis(scenario.GLHT.2.DF$estimate[[1]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.2.DF$std.error[[1]])*100,2)` SE ; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$) or genetic variables (`r round(plogis(scenario.GLHT.2.DF$estimate[[2]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.2.DF$std.error[[2]])*100,2)` SE; `$Z_1 = `r round(scenario.GLHT.2.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[2]])`$).
+No evidence was found to suggest that the prioritizations that included additional surrogate- or genetic-based targets secured a different proportion of genetic variation (`r round(plogis(scenario.GLHT.2.DF$statistic[[3]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.2.DF$std.error[[3]])*100,2)` SE; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).
+We also found that the prioritizations tended to secure more adaptive than neutral genetic variation (`r round(plogis(scenario.GLHT.3.DF$statistic[[1]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.3.DF$std.error[[1]])*100,2)` SE; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.3.DF$Pr..Chisq.[[2]])`$).
 We also found significant differences in the average proportion of genetic variation secured by prioritizations in each of the scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.2.DF$statistic[[2]],2)`$; $P$ `r pvalString(scenario.ANOVA.2.DF$Pr..Chisq.[[2]])`).
-Specifically, both the multi-species ($Z_1 = `r round(scenario.GLHT.1.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$) and multi-species with cost ($Z_1 = `r round(scenario.GLHT.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[2]])`$) scenarios were found to secure more genetic variation than the single species scenario.
-Neither the multi-species scenarios were found to secure a different proportion of genetic variation ($Z_1 = `r round(scenario.GLHT.1.DF$statistic[[3]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[3]])`$).
+Specifically, both the multi-species (`r round(plogis(scenario.GLHT.1.DF$estimate[[1]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.1.DF$std.error[[1]])*100,2)` SE; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$) and multi-species with cost (`r round(plogis(scenario.GLHT.1.DF$estimate[[2]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.1.DF$std.error[[2]])*100,2)` SE; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[2]])`$) scenarios were found to secure more genetic variation than the single species scenario.
+Neither the multi-species scenarios were found to secure a different proportion of genetic variation (`r round(plogis(scenario.GLHT.1.DF$estimate[[3]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.1.DF$std.error[[3]])*100,2)` SE; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[3]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[3]])`$).
 These results suggest that surrogate-based targets can substantially improve the proportion of intra-specific genetic variation secured in a prioritization for most species.
@@
 * introduction
+We aimed  to investigate the effectiveness of broad-scale environmental and geographic variables as surrogates for adaptive and neutral genetic variation in conservation planning.
+Our study is the first to address this hypothesis using moderately realistic planning scenarios.
+Overall, we found that the spatial patterns of environmental and geographic variables correlate with those of adaptive and neutral patterns for most species.
+Furthermore, we found that the addition of environmental and geographic targets to the planning process resulted in prioritizations that generally secured a greater proportion of adaptive and neutral genetic variation.
 
@@
 
-* outlier detection methods - how does our findings agree with other papers?
+* outlier detection methods
 
-* mapping genetic variation - compare with GDMS
+* mapping genetic variation
 

### 2016-05-12 code/rmarkdown/figures.Rmd
+ fixes to caching with addition of new script
Former-commit-id: 4ed8f5a1338f6279f7cfa2d2b6449e471c179a26

@@
 ```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=5, fig.cap='Planning unit selection frequencies.
-The top row of panels (a, b, c) show single-species prioritizations generated assuming equal costs.
-The middle row of panels (d, e, f) show multi-species prioritizations generated assuming equal costs.
-The bottom row of panels (g, h, i) show multi-species generated using opportunity costs.
-The left column of panels (a, d, g) show prioritizations generated using only amount-based targets.
-The middle column of panels (b, e, h) show prioritizations generated using amount- and surrogate-based targets.
-The right column of panels (c, f, i) show prioritizations generated using amount- and genetic-based targets.'}
+The first row of panels (a--c) show single-species prioritizations generated assuming equal costs.
+The second row of panels (d--f) show single-species prioritizations generated using opportunity costs.
+The third row of panels (g--i) show multi-species prioritizations generated assuming equal costs.
+The fourth row of panels (j--l) show multi-species generated using opportunity costs.
+The left column of panels (a, d, g, j) show prioritizations generated using only amount-based targets.
+The middle column of panels (b, e, h, k) show prioritizations generated using amount- and surrogate-based targets.
+The right column of panels (c, f, i, l) show prioritizations generated using amount- and genetic-based targets.
+The color of each planning unit indicates its selection frequency.
+Note that the selection frequencies for the single-species prioritizations (panels a--f) are based on their average frequency among multiple prioritizations generated for each species.
+The selection frequences for the multi-species prioritizations (pabels g--l) are based on the optimal pioritization (thus they are all either zero or one).'}
 # prepare data for plotting
@@
 	grid.sub.FPLY@data[[paste0('Single-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations, function(x){colMeans(x[[i]]@results@selections)}))
+for (i in seq_along(single.spp.prioritisations.cost[[1]]))
+	grid.sub.FPLY@data[[paste0('Single-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations.with.cost, function(x){colMeans(x[[i]]@results@selections)}))
 for (i in seq_along(multi.spp.prioritisations))
@@
 			gsub(' (', '\n(', Context, fixed=TRUE),
-			levels=c("Single-species\n(equal costs)", "Multi-species\n(equal costs)", 
-			"Multi-species\n(opportunity costs)"))
+			levels=c("Single-species\n(equal costs)", "Single-species\n(opportunity costs)"
+				"Multi-species\n(equal costs)", "Multi-species\n(opportunity costs)"))
 	)
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=5.5, fig.cap='The effectiveness of prioritizations at securing genetic variation.
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=7.4, fig.cap='The effectiveness of prioritizations at securing genetic variation.
 The top row of panels (a--b) show the performance of single-species prioritizations generated assuming equal costs among planning costs.

### 2016-05-12 code/R/analysis/10-multi-species-prioritisations-no-cost.R

@@ -49,4 +49,4 @@ multi.spp.DF <- ldply(
 
 
 ## save .rda
-save.session('data/intermediate/09-multi-species-prioritisations-no-cost.rda', compress='xz')
+save.session('data/intermediate/10-multi-species-prioritisations-no-cost.rda', compress='xz')

### 2016-05-12 code/R/analysis/11-multi-species-prioritisations-with-cost.R

@@ -49,4 +49,4 @@ multi.spp.with.cost.DF <- ldply(
 
 
 ## save .rda
-save.session('data/intermediate/10-multi-species-prioritisations-with-cost.rda', compress='xz')
+save.session('data/intermediate/11-multi-species-prioritisations-with-cost.rda', compress='xz')

### 2016-05-14 code/rmarkdown/figures.Rmd
+ bug fix in figures
Former-commit-id: 18a72705e0a71d4a5aa07b57025495f472f8c877
+ another bug fix
Former-commit-id: c3da4e8802ccbdf0377c42e3e47fa22c14a116aa
+ fixes to docs + increased iterations for bayescan + fixed bug in pcadapt analysis + assorted changes to main and SI figures + added funcionality to push and pull ms from cb

@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=5, fig.cap='Planning unit selection frequencies.
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=6.5, fig.cap='Planning unit selection frequencies.
 The first row of panels (a--c) show single-species prioritizations generated assuming equal costs.
@@
 	grid.sub.FPLY@data[[paste0('Single-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations, function(x){colMeans(x[[i]]@results@selections)}))
-for (i in seq_along(single.spp.prioritisations.cost[[1]]))
+for (i in seq_along(single.spp.prioritisations.with.cost[[1]]))
 	grid.sub.FPLY@data[[paste0('Single-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations.with.cost, function(x){colMeans(x[[i]]@results@selections)}))
@@
 			gsub(' (', '\n(', Context, fixed=TRUE),
-			levels=c("Single-species\n(equal costs)", "Single-species\n(opportunity costs)"
+			levels=c("Single-species\n(equal costs)", "Single-species\n(opportunity costs)",
 				"Multi-species\n(equal costs)", "Multi-species\n(opportunity costs)"))
@@
 
+scenario.SDF <- scenario.DF %>% 
+	group_by(Prioritisation, Context, Metric, Species) %>% 
+	summarize(genetic.held=mean(genetic.held, na.rm=TRUE)) %>%
+	data.frame()
+
 # make plot
@@
 	ggplot(aes(x=Prioritisation,y=genetic.held),
-		data=scenario.DF) +
+		data=scenario.SDF) +
 		geom_boxplot(position=position_dodge(0.9)) +
@@
 			values=c('grey80','grey50','grey20')) +
+		geom_abline(slope=0, color='red', linetype='dashed',
+			intercept=rapr.params.LST[[MODE]]$scenario.analysis$genetic.target) +
 		ylab('Genetic variation secured (%)') +

### 2016-05-14 code/rmarkdown/supporting_information.Rmd
+ fixes to docs + increased iterations for bayescan + fixed bug in pcadapt analysis + assorted changes to main and SI figures + added funcionality to push and pull ms from cb

@@
 			se=FALSE, fill='blue', alpha=0.8)) +
-		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=5, color='blue') +
+		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=8, color='darkred') +
 		facet_wrap(~ Species, ncol=4)

### 2016-06-06 code/rmarkdown/article.Rmd
+ use new version of rapr + fix bug where numreps in bayescan is ignored + use correct clumpp parameters for data

@@
 We used two outlier detection methods to achieve this [an individual- and a population-level method as recommended by @r495].
-We omitted the species \textit{`r gsub('_', ' ' , missing.species, fixed=TRUE)`} from all subsequent analyses because it was not compatible with one of the outlier detection methods (see below).
 The basic premise underpinning such methods is that neutral loci are expected to exhibit a specific level of variation among populations, and loci that deviate from this expectation should be under selection [reviewed in @r480].
@@
 
-First, we used multinominal-Dirichlet models implemented in \texttt{BayeScan} [version 2.1; @r456] to identify loci under selection.
+The first outlier detection method we used involved fitting multinominal-Dirichlet models implemented in \texttt{BayeScan} [version 2.1; @r456].
 We adopted a similar methodology to Bothwell _et al._ [-@r455] and applied it to each of the species in the data set.
-Following this methodology, for each species, we first grouped individuals into genetic lineages to avoid falsely classifying loci as adaptive.
-We used the number of genetic lineages identified in a previous analysis of the data set [@r478].
-Since \textit{`r gsub('_', ' ' , missing.species, fixed=TRUE)`} was found to only contain a single lineage/population, it was not compatible with these models.
-For each species, we used \texttt{Structure} [version 2.3.4; using same parameters as @r478 per run with `r structure.params.LST[[MODE]]$numruns` runs per species] and \texttt{ClumPP} [version 1.1.2; @r459] to group individuals into populations.
-We then ran $\texttt{Bayescan}$ for each species using these groupings [1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations; @r456].
-Additionally, we omitted individuals from this analysis when their population membership was uncertain (maximum membership probability $< `r structure.params.LST[[MODE]]$probability.threshold`$).
+Following their methodology, we first grouped conspecifics into genetic lineages to avoid falsely classifying loci as adaptive.
+To complete this first step, for each species, we fit the admixture and correlated alleles model implemented in \texttt{Structure} [version 2.3.4; XXXX] assuming various numbers of populations (K = `r min(structure.params.LST[[MODE]]$maxpops)`--`r max(structure.params.LST[[MODE]]$maxpops)`; `structure.params.LST[[MODE]]$numruns` replicates per K; `r structure.params.LST[[MODE]]$admburnin` admixture burnin iterations; `r structure.params.LST[[MODE]]$burnin` burnin iterations; `r structure.params.LST[[MODE]]$numreps` iterations used for inference).
+We used the variant of Evanno's K method used in \texttt{Structure Harvester} [XXX] to identify the most suitable number of populations separately for each species.
+We used \texttt{ClumPP} [version 1.1.2; @r459] to combine replicate runs to group individuals into populations (greedy algorithm using the _G_ static; `r structure.params.LST[[MODE]]$repeats` iterations).
+After the \texttt{ClumPP} analysis, if most of the individuals for a given species (> 50 %) had uncertain population memberships (< 75 % probability for most likely population) then we treated this species as containing only a single population.
+We then ran $\texttt{Bayescan}$ for each species using these populations [1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations; @r456].
+Additionally, we omitted individuals from the \texttt{Bayescan} analysis if their population membership was uncertain (maximum membership probability $< `r structure.params.LST[[MODE]]$probability.threshold`$).
 We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate [$q \leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].
 
-Second, we used PCAs to identify outlier loci [using methods implemented in the \texttt{pcadapt R} package; version 3.0; @r487; @r496].
+The second outlier detection method involved fitting PCAs to identify outlier loci [using methods implemented in the \texttt{pcadapt R} package; version 3.0; @r487; @r496].
 To enable comparisons between the two outlier detection analyses, we used the same individuals in this analysis as used in the \texttt{BayeScan} analysis.
-For each of the 26 species, we first imputed missing data to permit further analysis.
+For each of the 27 species, we first imputed missing data to permit further analysis.
 Specifically, the missing value for a given locus for a given individual was replaced with the average frequency of that locus among all conspecifics.
 Next, we ran a PCA and extracted the minimum number of components needed to secure the over-arching population level variation among the loci (`r pcadapt.params.LST[[MODE]]$min.variance.explained*10` % of the variation in loci).
-We then used Mahalanobis distances [@r398] to compute _P_ values and then q-values [using the \texttt{qvalue R} package; @r491].
+We then used Mahalanobis distances [@r398] to compute _P_ values and then _q_-values [using the \texttt{qvalue R} package; @r491].
 Outlier loci were detected using a suitable false discovery rate ($q < `r pcadapt.params.LST[[MODE]]$alpha.level`$).
@@
 
-The previous analysis resulted in an adaptive (if detected) and neutral genetic space for each species.
+The previous set of analyses resulted in an adaptive (if detected) and neutral genetic space for each species.
 For a given species, each planning unit occupied by the species was associated with a multi-dimensional point in the species' neutral genetic space.

### 2016-06-06 code/R/functions/make.targets.R

@@ -13,7 +13,7 @@
 #' @export
 make.targets <- function(
 	species, 
-	environmental.space, geographic.space, 
+	environmental.space, geographic.space,
 	adaptive.spaces, neutral.spaces,
 	amount.target=0.2, space.target=0.2
 ) {
@@ -32,46 +32,28 @@ make.targets <- function(
 		name=c(paste0('env_',species), paste0('geo_',species))
 	)
 	# adaptive spaces
-	adaptive.targets.DF <- ldply(seq_along(adaptive.spaces), .fun=function(i) {
-		# get species position
-		spp.pos <- which(sapply(adaptive.spaces[[i]]@demand.points, function(y) {length(y@weights)})>2)
+	adaptive.targets.DF <- ldply(seq_along(adaptive.spaces@spaces), .fun=function(i) {
 		# return data.frame
-		spp.DF <- data.frame(
-			species=seq_along(species),
-			target=rep(i+2, length(species)),
-			proportion=replace(
-				rep(NA_real_, length(species)),
-				spp.pos,
-				space.target
-			),
-			name=replace(
-				rep('null',length(species)),
-				spp.pos,
-				paste0('adaptive_',species[spp.pos])
+		return(
+			data.frame(
+				species=adaptive.spaces@spaces[[i]]@species,
+				target=3,
+				proportion=space.target,
+				name=paste0('adaptive_',species[adaptive.spaces@spaces[[i]]@species])
 			)
 		)
-		return(spp.DF)
 	})
-	# neutral
-	neutral.targets.DF <- ldply(seq_along(neutral.spaces), .fun=function(i) {
-		# get species position
-		spp.pos <- which(sapply(neutral.spaces[[i]]@demand.points, function(y) {length(y@weights)})>2)
+	# neutral spaces
+	neutral.targets.DF <- ldply(seq_along(neutral.spaces@spaces), .fun=function(i) {
 		# return data.frame
-		spp.DF <- data.frame(
-			species=seq_along(species),
-			target=rep(i+length(adaptive.spaces)+2, length(species)),
-			proportion=replace(
-				rep(NA_real_, length(species)),
-				spp.pos,
-				space.target
-			),
-			name=replace(
-				rep('null',length(species)),
-				spp.pos,
-				paste0('neutral_',species[spp.pos])
+		return(
+			data.frame(
+				species=neutral.spaces@spaces[[i]]@species,
+				target=4,
[…]

### 2016-06-07 code/rmarkdown/article.Rmd
+ fixes to debugging parameters + added refs to article

@@
 Following their methodology, we first grouped conspecifics into genetic lineages to avoid falsely classifying loci as adaptive.
-To complete this first step, for each species, we fit the admixture and correlated alleles model implemented in \texttt{Structure} [version 2.3.4; XXXX] assuming various numbers of populations (K = `r min(structure.params.LST[[MODE]]$maxpops)`--`r max(structure.params.LST[[MODE]]$maxpops)`; `structure.params.LST[[MODE]]$numruns` replicates per K; `r structure.params.LST[[MODE]]$admburnin` admixture burnin iterations; `r structure.params.LST[[MODE]]$burnin` burnin iterations; `r structure.params.LST[[MODE]]$numreps` iterations used for inference).
-We used the variant of Evanno's K method used in \texttt{Structure Harvester} [XXX] to identify the most suitable number of populations separately for each species.
-We used \texttt{ClumPP} [version 1.1.2; @r459] to combine replicate runs to group individuals into populations (greedy algorithm using the _G_ static; `r structure.params.LST[[MODE]]$repeats` iterations).
+To complete this first step, for each species, we fit the admixture and correlated alleles model implemented in \texttt{Structure} [version 2.3.4; r@461; r@462] assuming various numbers of populations (K = `r min(structure.params.LST[[MODE]]$maxpops)`--`r max(structure.params.LST[[MODE]]$maxpops)`; `structure.params.LST[[MODE]]$numruns` replicates per K; `r structure.params.LST[[MODE]]$admburnin` admixture burnin iterations; `r structure.params.LST[[MODE]]$burnin` burnin iterations; `r structure.params.LST[[MODE]]$numreps` iterations used for inference).
+We applied the Evanno's K method variant used in \texttt{Structure Harvester} to identify the most suitable number of populations separately for each species [@r460; @r498].
+We used \texttt{ClumPP} [version 1.1.2; @r459] to combine replicate runs to group individuals into populations (greedy algorithm using the _G_ static; `r clumpp.params.LST[[MODE]]$repeats` iterations).
 After the \texttt{ClumPP} analysis, if most of the individuals for a given species (> 50 %) had uncertain population memberships (< 75 % probability for most likely population) then we treated this species as containing only a single population.

### 2016-06-07 code/rmarkdown/supporting_information.Rmd
+ find best K using evanno's method
+ fixes to debugging parameters + added refs to article

@@
 Squares represent planning units.
-For a given species, planning units that were found to be inhabited are denoted with bright blue.
-Note that \\textit{',gsub('\\_', '\\ ', missing.species),'} was omitted from analysis.')}
+For a given species, planning units that were found to be inhabited are denoted with bright blue.')}
 ## plot map of species distributions
@@
 		) 	
-		curr.probs <- spp.StructureAnalysis.LST[[i]]@results@summary %>%
+		curr.probs <- structurer:::sample.membership(spp.StructureCollection.LST[[i]]) %>%
 			as.data.frame() %>%
@@
 is.speciesPlottable <- function(i) {
-	return(i <= general.params.LST[[MODE]]$n.spp & (!unique(spp.samples.DF$species)[i] %in% missing.species) & !is.na(unique(spp.samples.DF$species)[i]))
+	if (i <= general.params.LST[[MODE]]$n.spp)
+		return(FALSE)
+	return(!is.na(unique(spp.samples.DF$species)[i]))
 }

### 2016-06-16 code/rmarkdown/article.Rmd
+ update article with population changes + change endnote_lib.bib to references.bib

--- 2016-06-06 code/rmarkdown/article.Rmd
@@
 documentclass: article
-bibliography: Endnote_lib.bib
+bibliography: references.bib
 csl: reference-style.csl
@@
 This data set contains information for 27 alpine plant species (Figure 1a).
-It was collected using a 20' longitude by 21' latitude grid (approx.
+It was collected using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx.
 22.3 km $\times$ 25 km; Figure S1).
@@
 We used only planning units that contained at least a single individual ($n = `r nrow(grid.DF)`$) to reduce computational burden.
-To examine the potential effects of variation in cost (see below), we calculated the total human population density inside each planning unit [obtained at 1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472] and used this to represent acquisition cost (Figure 1b).
+To examine the potential effects of variation in cost (see below), we calculated the total human population density inside each planning unit [obtained at 1 km$^2$ resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472] and used this to represent acquisition cost (Figure 1b).
 
@@
 To describe the environmental characteristics of each planning unit, we used broad-scale climatic variation.
-We obtained data for 19 bioclimatic variables across the extent of the study area [approx.
-1 km resolution; @r33].
+We obtained data for 19 bioclimatic variables across the extent of the study area [at $1^{\prime\prime}$ resolution; @r33].
 We used this data set because it is commonly to map habitat suitability in conservation planning [@r493].
@@
 Following their methodology, we first grouped conspecifics into genetic lineages to avoid falsely classifying loci as adaptive.
-To complete this first step, for each species, we fit the admixture and correlated alleles model implemented in \texttt{Structure} [version 2.3.4; XXXX] assuming various numbers of populations (K = `r min(structure.params.LST[[MODE]]$maxpops)`--`r max(structure.params.LST[[MODE]]$maxpops)`; `structure.params.LST[[MODE]]$numruns` replicates per K; `r structure.params.LST[[MODE]]$admburnin` admixture burnin iterations; `r structure.params.LST[[MODE]]$burnin` burnin iterations; `r structure.params.LST[[MODE]]$numreps` iterations used for inference).
-We used the variant of Evanno's K method used in \texttt{Structure Harvester} [XXX] to identify the most suitable number of populations separately for each species.
-We used \texttt{ClumPP} [version 1.1.2; @r459] to combine replicate runs to group individuals into populations (greedy algorithm using the _G_ static; `r structure.params.LST[[MODE]]$repeats` iterations).
-After the \texttt{ClumPP} analysis, if most of the individuals for a given species (> 50 %) had uncertain population memberships (< 75 % probability for most likely population) then we treated this species as containing only a single population.
-We then ran $\texttt{Bayescan}$ for each species using these populations [1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations; @r456].
+To complete this first step, for each species, we fit the admixture and correlated alleles model implemented in \texttt{Structure} [version 2.3.4; r@461; r@462] using the number of populations for each species previously determined by [@r478; `structure.params.LST[[MODE]]$numruns` replicates per K; `r structure.params.LST[[MODE]]$admburnin` admixture burnin iterations; `r structure.params.LST[[MODE]]$burnin` burnin iterations; `r structure.params.LST[[MODE]]$numreps` iterations used for inference].
+We used \texttt{ClumPP} [version 1.1.2; @r459] to combine replicate runs to group individuals into populations (greedy algorithm using the $G^{\prime}$ static; `r clumpp.params.LST[[MODE]]$repeats` iterations).
+We then ran $\texttt{Bayescan}$ [@r456] for each species using these populations [1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations].
 Additionally, we omitted individuals from the \texttt{Bayescan} analysis if their population membership was uncertain (maximum membership probability $< `r structure.params.LST[[MODE]]$probability.threshold`$).

### 2016-06-27 code/rmarkdown/article.Rmd
+ progress on the discussion + also included some stats in the discussion

@@
 The slope for each species were all significantly different to zero (minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`; maximum $P `r  pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3).
-The explanatory power of the environmental surrogates were also generally quite high for most species (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figure S56).
-The nature of these relationships suggest that most of the adaptive variation for a given species can be secured in a prioritization using moderately high environmental-based targets ($> 80$ %).
 Similarly, the proportion of neutral genetic variation in a prioritization was affected by the interaction between species and the proportion of neutral variation secured in the prioritization (${\chi^2}_{`r geo.surrogacy.AOV$df[[4]]`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`).
 The slope for each species were also significantly different to zero (minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`; maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3).
-Generally, the explanatory power of geographic variables were higher than that of environmental variables (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2 \pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figure S57).
-Compared to the environmental variables, the slopes of these relationships suggest that very high geographic-based targets are required to generate prioritizations that secure a sufficient proportion of neutral genetic variation for most species ($> 95$ %).
-Additionally, all species were associated with prioritizations that secured a large proportion of environmental or geographic variation (> 70 %), yet preserved only a very small proportion of their intra-specific genetic variation (< 10 %; Figures S55 and S56; such as _Gypsophila repens_ and _Luzula alpinopilosa_).
-These results suggest while prioritizations generated using environmental and geographic targets tend to secure more genetic variation, prioritizations generated using only these targets can sometimes secure only a small fraction of intra-specific genetic variation.
+After establishing  a relationship between the proportion of the surrogates secured in a prioritisation and proportion of genetic variation secured, the next step was to determine if prioritisations generated under various scenarios benefitted from the addition of surrogate-based targets.
 
@@
 ## Discussion
-* introduction
+<!--  introduction -->
 We aimed to investigate the effectiveness of broad-scale environmental and geographic variables as surrogates for adaptive and neutral genetic variation in conservation planning.
-Our study is the first to address this hypothesis using moderately realistic planning scenarios.
+Our study is the first to address this using a large-scale data set.
 Overall, we found that the spatial patterns of environmental and geographic variables correlate with those of adaptive and neutral patterns for most species.
-Furthermore, we found that the addition of environmental and geographic targets to the planning process resulted in prioritizations that generally secured a greater proportion of adaptive and neutral genetic variation.
+Furthermore, we found that the addition of environmental and geographic targets to the reserve selection process resulted in prioritizations that secured a greater proportion of adaptive and neutral genetic variation for most species.
 
-* consequences for conservation planning
+<!-- surrogate/genetic dissimilarity matrix correlations -->
+Differences in the the environmental conditions and geographic position among the planning units tended to correlate with genetic differences among the individuals in the planning units (Table 1).
+For both surrogates and their corresponding models, the conditional $R^2$ values (environmental vs.
+adaptive genetic variation: mean `r mean(correlation.model.summary.DF$Environmental.Conditional.R2,na.rm=TRUE)` conditional $R^2$ `r sd(correlation.model.summary.DF$Environmental.Conditional.R2,na.rm=TRUE)` $\pm$ SD; geographic vs.
+neutral genetic variation: mean `r mean(correlation.model.summary.DF$Geographic.Conditional.R2,na.rm=TRUE)` conditional $R^2$ `r sd(correlation.model.summary.DF$Geographic.Conditional.R2,na.rm=TRUE)` $\pm$ SD) were generally much higher than the marginal $R^2$ values (environmental vs.
+adaptive genetic variation: mean `r mean(correlation.model.summary.DF$Environmental.Marginal.R2,na.rm=TRUE)` conditional $R^2$ `r sd(correlation.model.summary.DF$Environmental.Marginal.R2,na.rm=TRUE)` $\pm$ SD; geographic vs.
+neutral genetic variation: mean `r mean(correlation.model.summary.DF$Geographic.Marginal.R2,na.rm=TRUE)` conditional $R^2$ `r sd(correlation.model.summary.DF$Geographic.Marginal.R2,na.rm=TRUE)` $\pm$ SD).
+This result suggests that while variation in the surrogate variables correlates with genetic variation, both forms of genetic variation contain additional spatially auto-correlated variation that is not explained by these surrogates.
+In particular, the large differences between the conditional and marginal $R^2$ values for the models examining the effectiveness of geographic distances suggest that this unexplained spatial auto-correlation may be non-linear or driven by a few neighboring planning units with genetically similar individuals.
+This finding is consonant with other studies that have found strong patterns of spatial auto-correlation in patterns of adaptive and neutral genetic variation [@rXXX; @rXXX; @rXXX].
+After verifying that the spatial pattern of variation in the surrogates generally correlated with patterns of intra-specific adaptive and neutral genetic variation, our next step was to determine if this pattern of correlation was strong enough to operationalise for conservation planning.
+Differences in the the environmental conditions and geographic position among the planning units tended to correlate with genetic differences among the individuals in the planning units (Table 1).
+For both surrogates and their corresponding models, the conditional $R^2$ values (environmental vs.
+adaptive genetic variation: mean `r mean(correlation.model.summary.DF$Environmental.Conditional.R2,na.rm=TRUE)` conditional $R^2$ `r sd(correlation.model.summary.DF$Environmental.Conditional.R2,na.rm=TRUE)` $\pm$ SD; geographic vs.
+neutral genetic variation: mean `r mean(correlation.model.summary.DF$Geographic.Conditional.R2,na.rm=TRUE)` conditional $R^2$ `r sd(correlation.model.summary.DF$Geographic.Conditional.R2,na.rm=TRUE)` $\pm$ SD) were generally much higher than the marginal $R^2$ values (environmental vs.
+adaptive genetic variation: mean `r mean(correlation.model.summary.DF$Environmental.Marginal.R2,na.rm=TRUE)` conditional $R^2$ `r sd(correlation.model.summary.DF$Environmental.Marginal.R2,na.rm=TRUE)` $\pm$ SD; geographic vs.
+neutral genetic variation: mean `r mean(correlation.model.summary.DF$Geographic.Marginal.R2,na.rm=TRUE)` conditional $R^2$ `r sd(correlation.model.summary.DF$Geographic.Marginal.R2,na.rm=TRUE)` $\pm$ SD).
+This result suggests that while variation in the surrogate variables correlates with genetic variation, both forms of genetic variation contain additional spatially auto-correlated variation that is not explained by these surrogates.
+In particular, the large differences between the conditional and marginal $R^2$ values for the models examining the effectiveness of geographic distances suggest that this unexplained spatial auto-correlation may be non-linear or driven by a few neighboring planning units with genetically similar individuals.
+This finding is consonant with other studies that have found strong patterns of spatial auto-correlation in patterns of adaptive and neutral genetic variation [@rXXX; @rXXX; @rXXX].
+After verifying that the spatial pattern of variation in the surrogates generally correlated with patterns of intra-specific adaptive and neutral genetic variation, our next step was to determine if this pattern of correlation was strong enough to operationalise for conservation planning.
 
-* outlier detection methods
+<!-- surrogate/genetic held correlations -->
+Prioritisations that secured a more representative sample of environmental conditions or geographic position tended to secure a greater proportion of adaptive and neutral genetic variation (Figure 2).
+Generally, the environmental and geographic surrogates variables were equally effective surrogates for adaptive (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figures 2c and S56) and neutral genetic variation (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2 \pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figure S57).
+But for a few species, the proportion of geographic variation secured in a prioritisation had a diminished capacity to explain the proportion of neutral variation secured in the prioritisation (eg.
+_Gypsophila repens_: $R^2 = `r round(filter(surrogate.spp.DF, Species=='Gypsophila_repens')$Geographic.r2CU,2)`; _Luzula alpinopilosa_: $R^2 = `r round(filter(surrogate.spp.DF, Species=='Luzula_alpinopilosa')$Geographic.r2CU,2)`; _Gentiana nivalis_ $R^2 = `r round(filter(surrogate.spp.DF, Species=='Gentiana_nivalis')$Geographic.r2CU,2)`).
+This bimodal distribution (Figure 2d) suggests that geographic position alone may not be a suitable surrogate for neutral variation in conservation planning for all species.
+The mechanism for this is two fold.
+In such alpine landscapes such as those investigated in this study, connectivity between areas can be strongly affected by local topology [@XXXX].
+Although  topographic-based variables describing the connectivity between areas may offer a better surrogate (eg.
+least cost paths based on elevation; @rXXX), such variables need to scaled to accommodate non-linear relationships between the variable and connectivity.
+and the appropriate scalings vary immensely between species--especially between plant species which can make use of remarkably different dispersal strategies (eg.
+wind dispersal, animal dispersal, offshoots from the mother plant).
+As a consequence, more complex surrogates of connectivity generally need to be tuned for each species--which requires either an in-depth understnading of the study species [eg.
+@XXX] and/or genetic data [eg.
+@XXX]--and so are not particularly applicable for broad-scale prioritisations that involve hundreds of species.
+Prioritisations that secured a more representative sample of environmental conditions or geographic position tended to secure a greater proportion of adaptive and neutral genetic variation (Figure 2).
+Generally, the environmental and geographic surrogates variables were equally effective surrogates for adaptive (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figures 2c and S56) and neutral genetic variation (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2 \pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figure S57).
+But for a few species, the proportion of geographic variation secured in a prioritisation had a diminished capacity to explain the proportion of neutral variation secured in the prioritisation (eg.
+_Gypsophila repens_: $R^2 = `r round(filter(surrogate.spp.DF, Species=='Gypsophila_repens')$Geographic.r2CU,2)`; _Luzula alpinopilosa_: $R^2 = `r round(filter(surrogate.spp.DF, Species=='Luzula_alpinopilosa')$Geographic.r2CU,2)`; _Gentiana nivalis_ $R^2 = `r round(filter(surrogate.spp.DF, Species=='Gentiana_nivalis')$Geographic.r2CU,2)`).
+This bimodal distribution (Figure 2d) suggests that geographic position alone may not be a suitable surrogate for neutral variation in conservation planning for all species.
+The mechanism for this is two-fold.
+First, the plant species investigated in this study utilise a diverse set of dispersal mechanisms, including anemochorous (eg.
+wind-based dispersal), zoochorous (animal-based dispersal), and boleochorous (gravity-mediated dispersal).
+Plant species utilising anemochorous and zoochorous strategies can exhibit remarkable long-distance dispersal capabilities [@rXXX], and geographic distance-based surrogates would fail to describe this.
+Second, in alpine landscapes such as those investigated in this study, connectivity between areas can be strongly affected by local topology [@XXXX].
+However, distance measures based on landscape features (eg.
+elevation) require non-linear scaling--based on an in-depth understanding of the study species [eg.
+@XXX] and/or genetic data [eg.
+@XXX]--to be effective surrogates for connectivity, and so are not applicable for large-scale prioritisations that involve hundreds of species.
+Therefore, we conclude that the surrogates generally performed well enough considering their lack of focus to any particular species.
 
-* mapping genetic variation
+<!-- suitable targets for surrogates -->
+Generally, the slopes correlating the proportion of geographic variation and neutral variation secured in a prioritisation were higher than those for the proportion of environmental variation and adaptive variation secured in a prioritisation (Figures 2a--2b and S56--S57).
+To secure a representative sample of each species' adaptive genetic variation ($>=$ `r rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %), the prioritisation would need to sample on average `r mean(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE)` % $\pm$ `r sd(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE)` SD of the environmental conditions found across the planning units occupied by the species.
+On the other hand, to secure a representative sample of the species' neutral genetic variation ($>=$ `r rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100`), the prioritisation would need to sample % `r mean(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE)` $\pm$ `r sd(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE)` SD of the variation in geographic position among the planning units occupied by the species.
+These results suggest that geographic surrogates may require higher targets than environmental surrogates to generate prioritisations that secure representative sample of genetic variation.
+Additionally, these results suggest using surrogate-based targets alone to generate prioritisations may be insufficient to secure a representative sample of species' genetic variation.
+For all species, prioritisations were generated that secured a high proportion of the surrogate variable (> 90 %) but also secured a minuscule proportion of adaptive or neutral genetic variation (< 20 %; Figures S56--S57).
+Thus we recommend that prioritisations be generated using a combination of amount-based targets--as is conventionally used in reserve selection processes--and also surrogate-based targets.
+After establishing a relationship between the proportion of surrogate variation and genetic variation secured in prioritisation, we than generated prioritisations under various planning scenarios to provide insight into how the surrogates may work in the real-world.
 
-* limitations/future directions
+<!-- adaptive genetic variation secured in prioritisations-->
+Almost all prioritisations secured a representative sample of species' adaptive genetic variation--regardless of whether surrogate-based or genetic targets were used or not (Figure 4).
+This result suggests that conservation planners can secure a representative sample of the intra-specific adaptive genetic variation using only amount targets.
+The spatial patterns of adaptive genetic variation tended to cluster into a few distinct groups for most species (eg.
+XXX; XXX; XXX; XXX), and so a prioritisation would only need to site a single planning unit in each cluster to secure a representative sample of the species' adaptive genetic variation.
+As a consequence, the single-species prioritisations generated using amount targets and without consideration to opportunity cost were often able to achieve high representation due to the fact they were essentially random selection (Figure 4a).
+To a lesser extent this is also true for the single-species prioritisations generated with amount targets and consideration to opportunity cost (Figure 4c), since the costs are fairly homogenous among planning units with exception to a few planning units lowland areas.
+
+Generally, all the multi-species prioritisations secured a representative sample of the species' intra-specific adaptive genetic variation (Figures 4e and 4g).
+Unlike the single-species prioritisations, the multi-species prioiritisations generated with only amount-targets did not secure a representative sample of adaptive variation due to stochastic processes.
+Since these prioritisations were forced to secure at least `r rapr.params.LST[[MODE]]$scenario.analysis$amount.target*100` % of land occupied each species, these prioritisations were forced to secured planning units a range of environmental conditions due to the presence of habitat specialists in the data set (eg.
+_Cerastium uniflorum_ and _Androsace obtusifolia_).
+Overall--contrary to our original hypotheses--it seems that conventional reserve selection methods may secure a representative sample of species' adaptive genetic variation if a comprehensive set of species is used.
+This outcome provides empirical support for one of the fundamental tenants of systematic conservation planning [@rXXX].
+
+<!-- neutral genetic variation secured in prioritisations-->
+Generally, the addition of environmental and geographic targets were needed to obtain prioritisations that secure a representative proportion of neutral genetic variation (Figure 4).
+With exception to the multi-species with cost planning scenario (Figure 4h), prioritisations generated using only amount targets failed to secure a representative proportion of neutral genetic variation ($>=$ `r rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %) for most species (Figures 4b, 4f, 4h).
+Whereas, in all scenarios, the prioritisations generated using amount and surrogate based targets secured a representative proportion of neutral genetic variation for most species.
+Although the the latter prioritisation did not secure a representative sample of the neutral genetic variation in all species, this result suggests that the use of environmental and geographic targets in conservation planning can result in a more effective prioritisation.
+Furthermore, this improvement was associated with only a minor increase in cost (XXXX)--emphasizing the utility of surrogates in conservation.
+
+<!--  limitations/ weaknesses -->
 We wish to alert our readers to several limitations associated with our analysis.
@@
 We used this resolution because the genetic data were collected at this scale.
-Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis because this process would have artificially introduced additional spatial auto-correlation into the dataset.
+Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis because this process would have artificially introduced additional spatial auto-correlation into the data set.
 Second, we used geographic distances as surrogates for connectivity between planning units.
-Although connectivity between areas can be modeled using isolation by resistance measures [@r473] derived from landscape data [eg.
-@r474], they are unlikely to be of use in broad-scale conservation exercises since they need to be tuned using species-specific parameters.
+But as previously, using more geographic distances were found to be adequate and more complex measures of connectivity cannot generally be used in more large-scale planning scenarios.
 Third, we used AFLP data to characterize the main gradients of genetic variation.
@@
 
-* Concluding remarks
+<!--  conclusion -->
 The overarching aim of our study was to investigate the effectiveness of surrogates in conservation planning.

### 2016-06-27 code/R/analysis/12-statistical-analysis.R

@@ -202,6 +202,35 @@ surrogate.spp.DF <- full_join(env.surrogate.spp.DF2, geo.surrogate.spp.DF2, by='
 		Geographic.p.value=pvalString(Geographic.p.value)
 	)
 
+# predict percentage of surrogate needed to secure a representative sample of genetic variation
+env.surrogate.repr.prop.DF <- ldply(
+	as.character(unique(env.surrogacy.DF$Species)),
+	function(x) {
+		curr.mod <- env.surrogacy.LST[[x]]$model
+		curr.DF <- data.frame(surrogate.held=seq(0,1,length.out=10000))
+		curr.DF$genetic.held <- predict(curr.mod, newdata=curr.DF, type='response')
+		min.prop <- min(filter(curr.DF, genetic.held >= rapr.params.LST[[MODE]]$scenario.analysis$genetic.target)$surrogate.held)
+		data.frame(Species=x, environmental.held=min.prop)
+	}
+)
+
+geo.surrogate.repr.prop.DF <- ldply(
+	as.character(unique(geo.surrogacy.DF$Species)),
+	function(x) {
+		curr.mod <- geo.surrogacy.LST[[x]]$model
+		curr.DF <- data.frame(surrogate.held=seq(0,1,length.out=10000))
+		curr.DF$genetic.held <- predict(curr.mod, newdata=curr.DF, type='response')
+		if (max(curr.DF$genetic.held) < rapr.params.LST[[MODE]]$scenario.analysis$genetic.target) {
+			min.prop <- 1
+		} else {
+			min.prop <- min(filter(curr.DF, genetic.held >= rapr.params.LST[[MODE]]$scenario.analysis$genetic.target)$surrogate.held)
+		}
+		data.frame(Species=x, geographic.held=min.prop)
+	}
+)
+
+surrogate.repr.prop.DF <- merge(geo.surrogate.repr.prop.DF, env.surrogate.repr.prop.DF, by=c('Species'), all=TRUE)
+
 ### scenario analyses
 ## prepare data
 # single species prioritisations without cost

### 2016-06-27 code/rmarkdown/article.Rmd
+ fix bug in si figures

@@
 Short running title: Genetic surrogates in conservation \newline
-Word count: XXXXX \newline
+Abstract word count: XXXXX \newline
+Main text word count: XXXXX \newline
 Number of references: XXXX \newline
+Number of equations: XXXX \newline
 Number of figures: XXXX \newline

### 2016-06-29 code/rmarkdown/article.Rmd
+ changes to article + added push and pulling to makefile + increased surrogate space targets

@@
 The proportion of genetic variation secured in a prioritization was found to depend on the combination of targets used to generate the prioritization (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$Pr..Chisq.[[2]])`)$.
-Specifically, the amount-based prioritizations tended to secure less variation than the prioritizations generated using additional surrogate (`r round(plogis(scenario.GLHT.2.DF$estimate[[1]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.2.DF$std.error[[1]])*100,2)` SE ; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$) or genetic variables (`r round(plogis(scenario.GLHT.2.DF$estimate[[2]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.2.DF$std.error[[2]])*100,2)` SE; `$Z_1 = `r round(scenario.GLHT.2.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[2]])`$).
+
+Specifically, the amount-based prioritizations tended to secure less variation than the prioritizations generated using additional surrogate (`r round(plogis(scenario.GLHT.2.DF$estimate[[1]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.2.DF$std.error[[1]])*100,2)` SE ; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$) or genetic variables (`r round(plogis(scenario.GLHT.2.DF$estimate[[2]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.2.DF$std.error[[2]])*100,2)` SE; `$Z_1 = `r round(scenario.GLHT.2.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[2]])`$).
 No evidence was found to suggest that the prioritizations that included additional surrogate- or genetic-based targets secured a different proportion of genetic variation (`r round(plogis(scenario.GLHT.2.DF$statistic[[3]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.2.DF$std.error[[3]])*100,2)` SE; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).
+
 We also found that the prioritizations tended to secure more adaptive than neutral genetic variation (`r round(plogis(scenario.GLHT.3.DF$statistic[[1]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.3.DF$std.error[[1]])*100,2)` SE; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.3.DF$Pr..Chisq.[[2]])`$).
 We also found significant differences in the average proportion of genetic variation secured by prioritizations in each of the scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.2.DF$statistic[[2]],2)`$; $P$ `r pvalString(scenario.ANOVA.2.DF$Pr..Chisq.[[2]])`).
+
 Specifically, both the multi-species (`r round(plogis(scenario.GLHT.1.DF$estimate[[1]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.1.DF$std.error[[1]])*100,2)` SE; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$) and multi-species with cost (`r round(plogis(scenario.GLHT.1.DF$estimate[[2]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.1.DF$std.error[[2]])*100,2)` SE; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[2]])`$) scenarios were found to secure more genetic variation than the single species scenario.
 Neither the multi-species scenarios were found to secure a different proportion of genetic variation (`r round(plogis(scenario.GLHT.1.DF$estimate[[3]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.1.DF$std.error[[3]])*100,2)` SE; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[3]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[3]])`$).
+
 These results suggest that surrogate-based targets can substantially improve the proportion of intra-specific genetic variation secured in a prioritization for most species.
@@
 Although the the latter prioritisation did not secure a representative sample of the neutral genetic variation in all species, this result suggests that the use of environmental and geographic targets in conservation planning can result in a more effective prioritisation.
-Furthermore, this improvement was associated with only a minor increase in cost (XXXX)--emphasizing the utility of surrogates in conservation.
+Furthermore, this improvement was associated with only a moderate increase in cost in both scenarios (scenario without opportunity: `r round((score(multi.spp.prioritisations[[2]])/score(multi.spp.prioritisations[[1]]))*100,2)` %; scenario with opportunity cost: `r  round((score(multi.spp.prioritisations.with.cost[[2]])/score(multi.spp.prioritisations.with.cost[[1]]))*100,2)` %)--emphasizing the utility of surrogates in conservation.
 

### 2016-06-29 code/rmarkdown/article.Rmd
+ article fixes

--- 2016-06-27 code/rmarkdown/article.Rmd
@@
 Thus protected areas need to be sited in places that achieve conservation goals for minimal cost.
-To ensure that the plans for protected areas (prioritizations) are cost-effective, they are often generated using decision support tools to identify near-optimal solutions.
-These prioritizations often contain plans for a network of multiple protected areas so they can preserve biological processes (eg.
+To ensure that the plans for protected areas (prioritisations) are cost-effective, they are often generated using decision support tools to identify near-optimal solutions.
+These prioritisations often contain plans for a network of multiple protected areas so they can preserve biological processes (eg.
 connectivity between populations).
 For instance, \texttt{Marxan} [@r22] uses amount-based targets to identify protected area networks that preserve a suitable amount of habitat for each species (eg.
-100 km $^2$ of land occupied by the species needs to be secured in a reserve) and weights to penalize fragmented solutions (ie.
+100 km $^2$ of land occupied by the species needs to be secured in a reserve) and weights to penalise fragmented solutions (ie.
 the \textit{BLM} parameter).
@@
 Intra-specific genetic variation is important for long-term species persistence.
-As a consequence, there has been increasing interest in designing prioritizations that preserve and cultivate this variation [@r67; @r3; @r61].
+As a consequence, there has been increasing interest in designing prioritisations that preserve and cultivate this variation [@r67; @r3; @r61].
 Although the strength of natural selection is a continuous force; broadly speaking, genetic variation can be classified as either adaptive or neutral [reviewed in @r480].
@@
 Despite this, genetic data are not often used to inform conservation planning exercises because such data require considerable investment to obtain and analyze [@r61; but see @r5 and @r157 for developments on using genetic data in conservation planning]--especially since most large-scale exercises involve hundreds of species.
-Thus prioritizations may fail to secure a sufficient proportion of the intra-specific genetic variation of  species, and ultimately fail to deliver long-term conservation outcomes.
+Thus prioritisations may fail to secure a sufficient proportion of the intra-specific genetic variation of  species, and ultimately fail to deliver long-term conservation outcomes.
 
-Recently, conservation scientists have been investigating the use of surrogate data to generate prioritizations that adequately preserve intra-specific genetic variation--but without actually needing to utilize genetic data to acheive this.
+Recently, conservation scientists have been investigating the use of surrogate data to generate prioritisations that adequately preserve intra-specific genetic variation--but without actually needing to utilise genetic data to acheive this.
 Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been used to guide reserve selection as surrogates for adaptive genetic variation [eg.
 environmental conditions; @r247; @r2].
-By preserving a more representative sample of environmental conditions across the species' ranges, these researchers aimed to generate prioritizations that would preserve more adaptive variation.
+By preserving a more representative sample of environmental conditions across the species' ranges, these researchers aimed to generate prioritisations that would preserve more adaptive variation.
 However, the effectiveness of this approach remains unverified.
@@
 For instance, the isolation by distance theory [IBD; @r259] predicts that populations further apart will have less connectivity, exchange less genetic material, and in turn have less genetic variation in common.
-Based on this theory, geographic distance-based surrogates have been found to improve prioritizations in an island network [@r200].
+Based on this theory, geographic distance-based surrogates have been found to improve prioritisations in an island network [@r200].
 However this remains untested in spatially contiguous systems where connectivity is complicated by additional factors [eg.
@@
 Second, we aim to verify that environmental and geographic variables are effective surrogates for adaptive and neutral variation (respectively).
-We expect that prioritizations generated using surrogate-based targets will secure more genetic variation than prioritizations generated using only amount-based targets with the same number of planning units.
-Third, we aim to determine if the addition of surrogate-based targets to conservation planning exercises can result in more effective prioritizations.
-We expect that prioritizations generated using amount- and surrogate-based targets will sufficiently preserve the genetic variation of more species than those generated using only amount-based targets.
+We expect that prioritisations generated using surrogate-based targets will secure more genetic variation than prioritisations generated using only amount-based targets with the same number of planning units.
+Third, we aim to determine if the addition of surrogate-based targets to conservation planning exercises can result in more effective prioritisations.
+We expect that prioritisations generated using amount- and surrogate-based targets will sufficiently preserve the genetic variation of more species than those generated using only amount-based targets.
 Furthermore, we aim to test this hypothesis under several of different scenarios to see whether the additional constraints of real-world planning exercises exacerbate or nullify this effect.
-If these surrogates can improve the intra-specific genetic variation captured in a prioritization, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.
+If these surrogates can improve the intra-specific genetic variation captured in a prioritisation, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.
 
@@
 The IntraBioDiv data set contains genetic data for 27 species with different evolutionary histories and life-history strategies.
-Second, we required genetic data for all species to be collected using a standardized sampling scheme, and this data set used a grid to collect samples.
+Second, we required genetic data for all species to be collected using a standardised sampling scheme, and this data set used a grid to collect samples.
 Third, we required the study area to contain broad-scale environmental variation to test the effectiveness of environmental variables.
@@
 However, to our knowledge, only a single study has used this data set to explore the effectiveness of a surrogate in the context of conservation planning.
-Taberlet _et al._ [-@r486] investigated whether prioritizations generated using the spatial distribution of a comprehensive set species (> 300) could adequately secure the intra-specific genetic variation for 27 species.
-They found that the prioritization based on species distributions was substantially different to the prioritization based on genomic data.
-Their results suggest that prioritizations based on species distribution data alone will be unable to secure intra-specific genetic variation.
-However, whether or not environmental and geographic variables can augment prioritizations remains untested.
+Taberlet _et al._ [-@r486] investigated whether prioritisations generated using the spatial distribution of a comprehensive set species (> 300) could adequately secure the intra-specific genetic variation for 27 species.
+They found that the prioritisation based on species distributions was substantially different to the prioritisation based on genomic data.
+Their results suggest that prioritisations based on species distribution data alone will be unable to secure intra-specific genetic variation.
+However, whether or not environmental and geographic variables can augment prioritisations remains untested.
 
@@
 We examined the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation (respectively).
-We utilized the sampling grid used by the IntraBioDiv project as planning units to develop prioritizations.
+We utilised the sampling grid used by the IntraBioDiv project as planning units to develop prioritisations.
 We used only planning units that contained at least a single individual ($n = `r nrow(grid.DF)`$) to reduce computational burden.
@@
 The first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components (PCs) cumulatively explained `r round(pca.DF[[4]][surrogate.params.LST[[MODE]]$number.components],1)` % of the variation and were used to generate new layers.
-We calculated the average value for each PC layer in each planning unit, and used these values to characterize climatic variation among the planning units (Figure S2).
+We calculated the average value for each PC layer in each planning unit, and used these values to characterise climatic variation among the planning units (Figure S2).
 Thus we constructed a three-dimensional environmental space--wherein each planning unit occupies a single point--as a potential surrogate for adaptive variation.
@@
 Similarly, if the species was associated with adaptive loci, each planning unit occupied by the species was associated with a multi-dimensional point in the species' adaptive genetic space.
-By spreading out conservation effort across a given genetic space, and in turn securing planning units occupied by individuals with increasingly different polymorphisms, prioritizations can secure more genetic variation [see @r465 for discussion on attribute spaces].
+By spreading out conservation effort across a given genetic space, and in turn securing planning units occupied by individuals with increasingly different polymorphisms, prioritisations can secure more genetic variation [see @r465 for discussion on attribute spaces].
 
-## Prioritization method
-We used the unreliable representative and adequate prioritization (URAP) formulation of the reserve selection problem [@r465].
+## Prioritisation method
+We used the unreliable representative and adequate prioritisation (URAP) formulation of the reserve selection problem [@r465].
 This formulation can identify the minimum number of planning units required to preserve both a proportion of the species' range (using amount-based targets) and a proportion of variation found across the species' range [using space-based targets and attribute spaces; _sensu_ @r465].
@@
 We set the demand points for each species and each attribute space using the values associated with the planning units.
-For every prioritization we generated, we calculated the proportion of the adaptive (if detected) and neutral genetic variation that every prioritization secured to assess its performance.
-Prioritizations associated with negative values--because they secured only a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis.
+For every prioritisation we generated, we calculated the proportion of the adaptive (if detected) and neutral genetic variation that every prioritisation secured to assess its performance.
+Prioritisations associated with negative values--because they secured only a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis.
 We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality using the $\texttt{rapr}$ R package [@r465] and $\texttt{Gurobi}$ [@r463].
@@
 
-To address the second aim--to determine if the environmental and geographic variables are effective surrogates for adaptive and neutral genetic variation--we generated a collection of single-species prioritizations and computed the proportion of intra-specific genetic variation they secured.
-For each species, we generated `r rapr.params.LST[[MODE]]$surrogacy.analysis$replicates` prioritizations by randomly selecting different combinations of planning units that were occupied by the species.
-We then computed the proportion of adaptive genetic variation and environmental variation secured in each prioritization.
-We then generated another random set of prioritizations, and computed the proportion of neutral genetic variation and geographic variation secured in each prioritization.
+To address the second aim--to determine if the environmental and geographic variables are effective surrogates for adaptive and neutral genetic variation--we generated a collection of single-species prioritisations and computed the proportion of intra-specific genetic variation they secured.
+For each species, we generated `r rapr.params.LST[[MODE]]$surrogacy.analysis$replicates` prioritisations by randomly selecting different combinations of planning units that were occupied by the species.
+We then computed the proportion of adaptive genetic variation and environmental variation secured in each prioritisation.
+We then generated another random set of prioritisations, and computed the proportion of neutral genetic variation and geographic variation secured in each prioritisation.
 
-We fit two full generalized linear models (GLM) with a logit links.
-The first model was fit to the proportion of adaptive genetic variation secured in a prioritization, and contained a continuous predictor variable measuring the proportion of environmental variation also secured in the prioritization, a categorical variable indicating the species for which the prioritization was generated, and an interaction term between these predictor variables.
-The second model was fit to the proportion of neutral genetic variation secured in a prioritization, and contained a continuous predictor variable assessing the proportion of geographic variation also secured in the prioritization.
-Similar to the previous model, this model also contained a predictor variable indicating the species for which the prioritizations were generated, and an interaction term.
+We fit two full generalised linear models (GLM) with a logit links.
+The first model was fit to the proportion of adaptive genetic variation secured in a prioritisation, and contained a continuous predictor variable measuring the proportion of environmental variation also secured in the prioritisation, a categorical variable indicating the species for which the prioritisation was generated, and an interaction term between these predictor variables.
+The second model was fit to the proportion of neutral genetic variation secured in a prioritisation, and contained a continuous predictor variable assessing the proportion of geographic variation also secured in the prioritisation.
+Similar to the previous model, this model also contained a predictor variable indicating the species for which the prioritisations were generated, and an interaction term.
 We subjected these models to a backwards step-wise term deletion routine to assess term significance.
-To gauge the performance of these surrogates for each species, we fit a suite of models correlating the proportion of adaptive and neutral genetic variation held in each prioritization and the environmental and geographic variation they also secured (respectively).
+To gauge the performance of these surrogates for each species, we fit a suite of models correlating the proportion of adaptive and neutral genetic variation held in each prioritisation and the environmental and geographic variation they also secured (respectively).
 We then computed the Cragg and Uhler's pseudo-R$^2$ value for each model [using the \texttt{pscl} R package; @r492].
 
-To address the third aim--to determine if the use of surrogate-based targets can actually improve prioritizations--we investigated the effectiveness of surrogate-based targets under three scenarios with increasing levels of complexity.
-To assess the benefits of surrogate-based targets under the simplest of circumstances, our first scenario involved generating single-species prioritizations for each species.
-However, most large-scale prioritizations involve a comprehensive set of species.
-By forcing solutions to secure individuals in multiple communities, this may result in prioritizations that secure a greater proportion of intra-specific genetic variation for each species.
-To test for this effect, our second scenario involved generating multi-species prioritizations.
-Finally, most large-scale prioritizations also explicitly consider opportunity cost to ensure that prioritizations are cost-effective.
-Our third scenario involved generating multi-species prioritizations that satisfied targets for minimal opportunity cost.
+To address the third aim--to determine if the use of surrogate-based targets can actually improve prioritisations--we investigated the effectiveness of surrogate-based targets under three scenarios with increasing levels of complexity.
+To assess the benefits of surrogate-based targets under the simplest of circumstances, our first scenario involved generating single-species prioritisations for each species.
+However, most large-scale prioritisations involve a comprehensive set of species.
+By forcing solutions to secure individuals in multiple communities, this may result in prioritisations that secure a greater proportion of intra-specific genetic variation for each species.
+To test for this effect, our second scenario involved generating multi-species prioritisations.
+Finally, most large-scale prioritisations also explicitly consider opportunity cost to ensure that prioritisations are cost-effective.
+Our third scenario involved generating multi-species prioritisations that satisfied targets for minimal opportunity cost.
 
-For each scenario, we generated three types of prioritizations using (1) only amount-based targets, (2) both amount- and surrogate-based targets, and (3) both amount- and genetic-based targets.
-The purely amount-based prioritizations served as our baseline--representing the prioritizations typically generated in conservation planning exercises (eg.
+For each scenario, we generated three types of prioritisations using (1) only amount-based targets, (2) both amount- and surrogate-based targets, and (3) both amount- and genetic-based targets.
+The purely amount-based prioritisations served as our baseline--representing the prioritisations typically generated in conservation planning exercises (eg.
 by \texttt{Marxan}).
-We used `r (rapr.params.LST[[MODE]]$scenario.analysis$amount.target*100)` % amount-based targets in all prioritizations to ensure an adequate proportion of habitat was secured for each species.
+We used `r (rapr.params.LST[[MODE]]$scenario.analysis$amount.target*100)` % amount-based targets in all prioritisations to ensure an adequate proportion of habitat was secured for each species.
 Based on the results from the analyses used to address our second aim, we set surrogate-based targets as `r (rapr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100)` % and genetic-based targets as `r (rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)` % to ensure that most of the genetic variation was secured for each species.
-To accommodate the fact that there were a large number of optimal solutions for the single-species amount-based prioritizations, we generated `r (rapr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates)` replicate prioritizations.
-We computed the proportion of adaptive (if present) and neutral genetic variation secured for each species in the prioritizations.
+To accommodate the fact that there were a large number of optimal solutions for the single-species amount-based prioritisations, we generated `r (rapr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates)` replicate prioritisations.
+We computed the proportion of adaptive (if present) and neutral genetic variation secured for each species in the prioritisations.
 
-We investigated the benefits of using surrogate-based targets in prioritizations using generalized linear mixed-effects models fit with logit liknks [GLMMs; using the \texttt{lme4 R} package; @469].
-We fit a full model to the proportion of genetic variation secured in a given prioritization.
-This model contained categorical predictor variables to indicate the combination of targets used to generate the prioritizations (amount-only, amount and surrogate targets, and amount and genetic targets), the scenario that was used to generate the prioritization (single-species, multi-species, and multi-species with cost), the type of genetic variation measured (adaptive and neutral), and all interactions between these variables.
-This model accommodated prioritizations generated for the same species using a random intercept term.
+We investigated the benefits of using surrogate-based targets in prioritisations using generalised linear mixed-effects models fit with logit liknks [GLMMs; using the \texttt{lme4 R} package; @469].
+We fit a full model to the proportion of genetic variation secured in a given prioritisation.
+This model contained categorical predictor variables to indicate the combination of targets used to generate the prioritisations (amount-only, amount and surrogate targets, and amount and genetic targets), the scenario that was used to generate the prioritisation (single-species, multi-species, and multi-species with cost), the type of genetic variation measured (adaptive and neutral), and all interactions between these variables.
+This model accommodated prioritisations generated for the same species using a random intercept term.
 We subject the full model to a backwards step-wise term deletion routine to assess term significance.
@@
 <!-- summary of methods -->
-We successfully mapped the main spatial patterns of adaptive (if present) and neutral variation for each of the 26 species.
+We successfully mapped the main spatial patterns of adaptive (if present) and neutral variation for each of the 27 species.
 The outlier detection methods detected adaptive genetic variation in `r as.character(as.english(length(env.surrogacy.LST)))` species.
@@
 After identifying the loci under selection, we used NMDS analyses to identify the main gradients of adaptive (if detected) and neutral variation for each species.
-For all species in which we detected adaptive loci, only a couple of continuous dimensions were needed to sufficiently describe their patterns of adaptive genetic variation (Table S2).
-On the other hand, a few additional continuous dimensions were occasionally needed to describe their patterns of neutral genetic variation (minimum $K = `r min(neutral.k)`$; mean $K = `r round(mean(neutral.k),2)`$; max $K = `r max(neutral.k)`$; Table S2).
+For all species in which we detected adaptive loci, only a couple of continuous dimensions were needed to sufficiently describe their patterns of adaptive genetic variation (all $K = `r unique(adaptive.k)`$; Table S2).
+On the other hand, a few additional continuous dimensions were occasionally needed to describe their patterns of neutral genetic variation (minimum $K = `r min(neutral.k)`$; mean $K = `r round(mean(neutral.k),2)`$; maximum $K = `r max(neutral.k)`$; Table S2).
 The ordinations resulting from the NMDS analyses were used to construct an adaptive (if present) and neutral genetic space for each species.
-The spatial distribution of these genetic spaces (Figures S29--S54) generally show patterns of spatial auto-correlation--especially the adaptive genetic spaces (eg.
-_Cerastium uniflorum_ and _Dryas octopetala_).
-Additionally, several species show east-west patterns in their intra-specific genetic variation (eg.
-_Peucedanum ostruthium_ and _Phyteuma betonicifolium_).
+The spatial distribution of these genetic spaces (Figures S29--S54) generally showed patterns of spatial auto-correlation (eg.
+_Cerastium uniflorum_, Figure SXX; _Dryas octopetala_, Figures SXXX)--a good sign that the environmental and geographic variables may prove effective surrogates.
 
 <!-- correlation analysis -->
-The dissimilarity matrices generated using the environmental and geographic variables correlated with those matrices based on the genetic characteristics for most species (Table 1).
+The dissimilarity matrices based on the environmental and geographic characteristics of the planning units correlated with the dissimilarity matrices based on the genetic characteristics of individuals in the planning units (Table 1).
 Specifically, environmental variation was found to significantly correlate with adaptive genetic variation for `r as.character(as.english(sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)))` species (`r round((sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)/length(na.omit(correlation.model.summary.DF$Environmental.corr.P)))*100,2)` % associated with loci under selection; _P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models).
 Additionally, geographic distances between planning units were found to significantly correlate with neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` species (`r round((sum(correlation.model.summary.DF$Geographic.corr.P<0.05)/length(na.omit(correlation.model.summary.DF$Geographic.corr.P)))*100,2)` % of all the species investigated; _P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models).
-These results suggest that prioritizations with different environmental characteristics tend to have individuals with different adaptive genetic characteristics.
+These results suggest that prioritisations with different environmental characteristics tend to have individuals with different adaptive genetic characteristics.
 Similarly, planning units that are further apart tend to have individuals with different neutral genetic characteristics.
@@
 The environmental and geographic variables were moderately effective surrogates for genetic variation in most species (Figure 2).
-The relationship between the proportion of adaptive genetic variation secured in a prioritization and the proportion of environment variation it also secured was found to depend on the species for which the prioritization was generated (${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$).
+The relationship between the proportion of adaptive genetic variation secured in a prioritisation and the proportion of environment variation it also secured was found to depend on the species for which the prioritisation was generated (${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$).
 The slope for each species were all significantly different to zero (minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`; maximum $P `r  pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3).
-Similarly, the proportion of neutral genetic variation in a prioritization was affected by the interaction between species and the proportion of neutral variation secured in the prioritization (${\chi^2}_{`r geo.surrogacy.AOV$df[[4]]`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`).
+Similarly, the proportion of neutral genetic variation in a prioritisation was affected by the interaction between species and the proportion of neutral variation secured in the prioritisation (${\chi^2}_{`r geo.surrogacy.AOV$df[[4]]`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`).
 The slope for each species were also significantly different to zero (minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`; maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3).
-After establishing  a relationship between the proportion of the surrogates secured in a prioritisation and proportion of genetic variation secured, the next step was to determine if prioritisations generated under various scenarios benefitted from the addition of surrogate-based targets.
+After establishing a relationship between the proportion of the surrogate variables and proportion of genetic variation secured in a prioritisation, the next step was to determine if prioritisations benefited from the addition of surrogate-based targets and verify if such benefits were evident across a variety of planning scenarios.
 
 <!-- scenario analysis -->
-To determine if the addition of environmental and geographic targets could improve the effectiveness of prioritizations in a more realistic context, we generated prioritizations according to three scenarios using different combinations of targets (Figure 3).
-We then computed the proportion of adaptive and neutral genetic variation secured in each prioritization (Figure 4).
-The proportion of genetic variation secured in a prioritization was found to depend on the combination of targets used to generate the prioritization (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$Pr..Chisq.[[2]])`)$.
-Specifically, the amount-based prioritizations tended to secure less variation than the prioritizations generated using additional surrogate (`r round(plogis(scenario.GLHT.2.DF$estimate[[1]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.2.DF$std.error[[1]])*100,2)` SE ; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$) or genetic variables (`r round(plogis(scenario.GLHT.2.DF$estimate[[2]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.2.DF$std.error[[2]])*100,2)` SE; `$Z_1 = `r round(scenario.GLHT.2.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[2]])`$).
-No evidence was found to suggest that the prioritizations that included additional surrogate- or genetic-based targets secured a different proportion of genetic variation (`r round(plogis(scenario.GLHT.2.DF$statistic[[3]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.2.DF$std.error[[3]])*100,2)` SE; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).
-We also found that the prioritizations tended to secure more adaptive than neutral genetic variation (`r round(plogis(scenario.GLHT.3.DF$statistic[[1]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.3.DF$std.error[[1]])*100,2)` SE; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.3.DF$Pr..Chisq.[[2]])`$).
-We also found significant differences in the average proportion of genetic variation secured by prioritizations in each of the scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.2.DF$statistic[[2]],2)`$; $P$ `r pvalString(scenario.ANOVA.2.DF$Pr..Chisq.[[2]])`).
+To determine if the addition of environmental and geographic targets could actually improve the effectiveness of prioritisations, we generated prioritisations according to three scenarios using different targets (Figure 3).
+We then computed the proportion of adaptive and neutral genetic variation secured in each prioritisation (Figure 4).
+We found no evidence to suggest that any of the interaction terms were significant (${\chi^2}_{`r scenario.ANOVA.1.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.1.DF$Pr..Chisq.[[2]])`$).
+We found that on average the prioritisations tended to secure more adaptive than neutral genetic variation (`r round(plogis(scenario.GLHT.3.DF$statistic[[1]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.3.DF$std.error[[1]])*100,2)` SE; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.3.DF$Pr..Chisq.[[2]])`$).
+
+The proportion of genetic variation secured in a prioritisation was found to depend on the combination of targets used to generate the prioritisation (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$Pr..Chisq.[[2]])`)$.
+Specifically, the amount-based prioritisations tended to secure less variation than the prioritisations generated using additional surrogate (`r round(plogis(scenario.GLHT.2.DF$estimate[[1]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.2.DF$std.error[[1]])*100,2)` SE ; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$) or genetic variables (`r round(plogis(scenario.GLHT.2.DF$estimate[[2]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.2.DF$std.error[[2]])*100,2)` SE; `$Z_1 = `r round(scenario.GLHT.2.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[2]])`$).
+No evidence was found to suggest that the prioritisations that included additional surrogate- or genetic-based targets secured a different proportion of genetic variation (`r round(plogis(scenario.GLHT.2.DF$statistic[[3]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.2.DF$std.error[[3]])*100,2)` SE; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).
+These results suggest that surrogate-based targets can substantially improve the proportion of intra-specific genetic variation secured in a prioritisation for most species.
+
+We also found significant differences in the average proportion of genetic variation secured by prioritisations in each of the scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.2.DF$statistic[[2]],2)`$; $P$ `r pvalString(scenario.ANOVA.2.DF$Pr..Chisq.[[2]])`).
 Specifically, both the multi-species (`r round(plogis(scenario.GLHT.1.DF$estimate[[1]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.1.DF$std.error[[1]])*100,2)` SE; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$) and multi-species with cost (`r round(plogis(scenario.GLHT.1.DF$estimate[[2]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.1.DF$std.error[[2]])*100,2)` SE; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[2]])`$) scenarios were found to secure more genetic variation than the single species scenario.
 Neither the multi-species scenarios were found to secure a different proportion of genetic variation (`r round(plogis(scenario.GLHT.1.DF$estimate[[3]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.1.DF$std.error[[3]])*100,2)` SE; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[3]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[3]])`$).
-These results suggest that surrogate-based targets can substantially improve the proportion of intra-specific genetic variation secured in a prioritization for most species.
+These results suggest that on average the multi-species prioritisations can secure more genetic variation than single-species prioritisations--regardless of targets used to generate them.
 
@@
 We aimed to investigate the effectiveness of broad-scale environmental and geographic variables as surrogates for adaptive and neutral genetic variation in conservation planning.
-Our study is the first to address this using a large-scale data set.
+Our study is the first to address this aim by measuring the proportion of genetic variation secured by prioritisations  generated for diverse set of species.
 Overall, we found that the spatial patterns of environmental and geographic variables correlate with those of adaptive and neutral patterns for most species.
-Furthermore, we found that the addition of environmental and geographic targets to the reserve selection process resulted in prioritizations that secured a greater proportion of adaptive and neutral genetic variation for most species.
+Furthermore, we found that the addition of environmental and geographic targets to the reserve selection process resulted in prioritisations that secured a greater proportion of adaptive and neutral genetic variation for most species.
 
@@
 For both surrogates and their corresponding models, the conditional $R^2$ values (environmental vs.
-adaptive genetic variation: mean `r mean(correlation.model.summary.DF$Environmental.Conditional.R2,na.rm=TRUE)` conditional $R^2$ `r sd(correlation.model.summary.DF$Environmental.Conditional.R2,na.rm=TRUE)` $\pm$ SD; geographic vs.
-neutral genetic variation: mean `r mean(correlation.model.summary.DF$Geographic.Conditional.R2,na.rm=TRUE)` conditional $R^2$ `r sd(correlation.model.summary.DF$Geographic.Conditional.R2,na.rm=TRUE)` $\pm$ SD) were generally much higher than the marginal $R^2$ values (environmental vs.
-adaptive genetic variation: mean `r mean(correlation.model.summary.DF$Environmental.Marginal.R2,na.rm=TRUE)` conditional $R^2$ `r sd(correlation.model.summary.DF$Environmental.Marginal.R2,na.rm=TRUE)` $\pm$ SD; geographic vs.
-neutral genetic variation: mean `r mean(correlation.model.summary.DF$Geographic.Marginal.R2,na.rm=TRUE)` conditional $R^2$ `r sd(correlation.model.summary.DF$Geographic.Marginal.R2,na.rm=TRUE)` $\pm$ SD).
+adaptive genetic variation: mean `r round(mean(correlation.model.summary.DF$Environmental.Conditional.R2,na.rm=TRUE),2)` conditional $R^2$ `r round(sd(correlation.model.summary.DF$Environmental.Conditional.R2,na.rm=TRUE),2)` $\pm$ SD; geographic vs.
+neutral genetic variation: mean `r round(mean(correlation.model.summary.DF$Geographic.Conditional.R2,na.rm=TRUE),2)` conditional $R^2$ `r round(sd(correlation.model.summary.DF$Geographic.Conditional.R2,na.rm=TRUE),2)` $\pm$ SD) were generally slightly higher than the marginal $R^2$ values (environmental vs.
+adaptive genetic variation: mean `r round(mean(correlation.model.summary.DF$Environmental.Marginal.R2,na.rm=TRUE),2)` marginal $R^2$ `r round(sd(correlation.model.summary.DF$Environmental.Marginal.R2,na.rm=TRUE),2)` $\pm$ SD; geographic vs.
+neutral genetic variation: mean `r round(mean(correlation.model.summary.DF$Geographic.Marginal.R2,na.rm=TRUE),2)` marginal $R^2$ `r round(sd(correlation.model.summary.DF$Geographic.Marginal.R2,na.rm=TRUE),2)` $\pm$ SD).
 This result suggests that while variation in the surrogate variables correlates with genetic variation, both forms of genetic variation contain additional spatially auto-correlated variation that is not explained by these surrogates.
-In particular, the large differences between the conditional and marginal $R^2$ values for the models examining the effectiveness of geographic distances suggest that this unexplained spatial auto-correlation may be non-linear or driven by a few neighboring planning units with genetically similar individuals.
+In particular, the differences between the conditional and marginal $R^2$ values for the models examining the effectiveness of geographic variables suggest that this unexplained spatial auto-correlation may be non-linear or driven by a few neighboring planning units with genetically similar individuals.
 This finding is consonant with other studies that have found strong patterns of spatial auto-correlation in patterns of adaptive and neutral genetic variation [@rXXX; @rXXX; @rXXX].
-After verifying that the spatial pattern of variation in the surrogates generally correlated with patterns of intra-specific adaptive and neutral genetic variation, our next step was to determine if this pattern of correlation was strong enough to operationalise for conservation planning.
-Differences in the the environmental conditions and geographic position among the planning units tended to correlate with genetic differences among the individuals in the planning units (Table 1).
-For both surrogates and their corresponding models, the conditional $R^2$ values (environmental vs.
-adaptive genetic variation: mean `r mean(correlation.model.summary.DF$Environmental.Conditional.R2,na.rm=TRUE)` conditional $R^2$ `r sd(correlation.model.summary.DF$Environmental.Conditional.R2,na.rm=TRUE)` $\pm$ SD; geographic vs.
-neutral genetic variation: mean `r mean(correlation.model.summary.DF$Geographic.Conditional.R2,na.rm=TRUE)` conditional $R^2$ `r sd(correlation.model.summary.DF$Geographic.Conditional.R2,na.rm=TRUE)` $\pm$ SD) were generally much higher than the marginal $R^2$ values (environmental vs.
-adaptive genetic variation: mean `r mean(correlation.model.summary.DF$Environmental.Marginal.R2,na.rm=TRUE)` conditional $R^2$ `r sd(correlation.model.summary.DF$Environmental.Marginal.R2,na.rm=TRUE)` $\pm$ SD; geographic vs.
-neutral genetic variation: mean `r mean(correlation.model.summary.DF$Geographic.Marginal.R2,na.rm=TRUE)` conditional $R^2$ `r sd(correlation.model.summary.DF$Geographic.Marginal.R2,na.rm=TRUE)` $\pm$ SD).
-This result suggests that while variation in the surrogate variables correlates with genetic variation, both forms of genetic variation contain additional spatially auto-correlated variation that is not explained by these surrogates.
-In particular, the large differences between the conditional and marginal $R^2$ values for the models examining the effectiveness of geographic distances suggest that this unexplained spatial auto-correlation may be non-linear or driven by a few neighboring planning units with genetically similar individuals.
-This finding is consonant with other studies that have found strong patterns of spatial auto-correlation in patterns of adaptive and neutral genetic variation [@rXXX; @rXXX; @rXXX].
-After verifying that the spatial pattern of variation in the surrogates generally correlated with patterns of intra-specific adaptive and neutral genetic variation, our next step was to determine if this pattern of correlation was strong enough to operationalise for conservation planning.
+After verifying that the spatial pattern of variation in the surrogates generally correlated with those of intra-specific adaptive and neutral genetic variation, our next step was to determine if this pattern of correlation was strong enough to operationalise for conservation planning.
 
@@
 Prioritisations that secured a more representative sample of environmental conditions or geographic position tended to secure a greater proportion of adaptive and neutral genetic variation (Figure 2).
-Generally, the environmental and geographic surrogates variables were equally effective surrogates for adaptive (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figures 2c and S56) and neutral genetic variation (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2 \pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figure S57).
+Generally, the environmental and geographic surrogates variables were equally effective surrogates for adaptive (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figures 2c and S56) and neutral genetic variation (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2$ \pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figure S57).
 But for a few species, the proportion of geographic variation secured in a prioritisation had a diminished capacity to explain the proportion of neutral variation secured in the prioritisation (eg.
-_Gypsophila repens_: $R^2 = `r round(filter(surrogate.spp.DF, Species=='Gypsophila_repens')$Geographic.r2CU,2)`; _Luzula alpinopilosa_: $R^2 = `r round(filter(surrogate.spp.DF, Species=='Luzula_alpinopilosa')$Geographic.r2CU,2)`; _Gentiana nivalis_ $R^2 = `r round(filter(surrogate.spp.DF, Species=='Gentiana_nivalis')$Geographic.r2CU,2)`).
+_Gypsophila repens_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gypsophila_repens')$Geographic.r2CU,2)`; _Luzula alpinopilosa_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Luzula_alpinopilosa')$Geographic.r2CU,2)`; _Gentiana nivalis_ $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gentiana_nivalis')$Geographic.r2CU,2)`).
 This bimodal distribution (Figure 2d) suggests that geographic position alone may not be a suitable surrogate for neutral variation in conservation planning for all species.
-The mechanism for this is two fold.
-In such alpine landscapes such as those investigated in this study, connectivity between areas can be strongly affected by local topology [@XXXX].
-Although  topographic-based variables describing the connectivity between areas may offer a better surrogate (eg.
-least cost paths based on elevation; @rXXX), such variables need to scaled to accommodate non-linear relationships between the variable and connectivity.
-and the appropriate scalings vary immensely between species--especially between plant species which can make use of remarkably different dispersal strategies (eg.
-wind dispersal, animal dispersal, offshoots from the mother plant).
-As a consequence, more complex surrogates of connectivity generally need to be tuned for each species--which requires either an in-depth understnading of the study species [eg.
-@XXX] and/or genetic data [eg.
-@XXX]--and so are not particularly applicable for broad-scale prioritisations that involve hundreds of species.
-Prioritisations that secured a more representative sample of environmental conditions or geographic position tended to secure a greater proportion of adaptive and neutral genetic variation (Figure 2).
-Generally, the environmental and geographic surrogates variables were equally effective surrogates for adaptive (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figures 2c and S56) and neutral genetic variation (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2 \pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figure S57).
-But for a few species, the proportion of geographic variation secured in a prioritisation had a diminished capacity to explain the proportion of neutral variation secured in the prioritisation (eg.
-_Gypsophila repens_: $R^2 = `r round(filter(surrogate.spp.DF, Species=='Gypsophila_repens')$Geographic.r2CU,2)`; _Luzula alpinopilosa_: $R^2 = `r round(filter(surrogate.spp.DF, Species=='Luzula_alpinopilosa')$Geographic.r2CU,2)`; _Gentiana nivalis_ $R^2 = `r round(filter(surrogate.spp.DF, Species=='Gentiana_nivalis')$Geographic.r2CU,2)`).
-This bimodal distribution (Figure 2d) suggests that geographic position alone may not be a suitable surrogate for neutral variation in conservation planning for all species.
-The mechanism for this is two-fold.
-First, the plant species investigated in this study utilise a diverse set of dispersal mechanisms, including anemochorous (eg.
-wind-based dispersal), zoochorous (animal-based dispersal), and boleochorous (gravity-mediated dispersal).
-Plant species utilising anemochorous and zoochorous strategies can exhibit remarkable long-distance dispersal capabilities [@rXXX], and geographic distance-based surrogates would fail to describe this.
-Second, in alpine landscapes such as those investigated in this study, connectivity between areas can be strongly affected by local topology [@XXXX].
-However, distance measures based on landscape features (eg.
-elevation) require non-linear scaling--based on an in-depth understanding of the study species [eg.
-@XXX] and/or genetic data [eg.
-@XXX]--to be effective surrogates for connectivity, and so are not applicable for large-scale prioritisations that involve hundreds of species.
-Therefore, we conclude that the surrogates generally performed well enough considering their lack of focus to any particular species.
+This result could be attributed to multiple sources.
+First, geographic distances will be a poor surrogate for genetic lineages in the process of sympatric speciation.
+Second, connectivity between areas can be affected by local topology.
+In particular, mountains can act as barriers to gene flow in alpine landscapes and cause relatively nearby populations to diverge genetically.
+Third, the species used in this dataset utilise different dispersal mechanisms [wind dispersal, animal dispersal, offshoots from the mother plant; @rXXX], and some of these species may have dispersal modes that are non-linearly related to geographic distance (eg.
+wind dispersing plants can have high connectivity between nearby and distant areas).
+Overall, however, these species were the exception, and both surrogates performed moderately well.
 
@@
 
-Generally, all the multi-species prioritisations secured a representative sample of the species' intra-specific adaptive genetic variation (Figures 4e and 4g).
+All the multi-species prioritisations secured a representative sample of most of the species' intra-specific adaptive genetic variation (Figures 4e and 4g).
 Unlike the single-species prioritisations, the multi-species prioiritisations generated with only amount-targets did not secure a representative sample of adaptive variation due to stochastic processes.
@@
 Although the the latter prioritisation did not secure a representative sample of the neutral genetic variation in all species, this result suggests that the use of environmental and geographic targets in conservation planning can result in a more effective prioritisation.
-Furthermore, this improvement was associated with only a minor increase in cost (XXXX)--emphasizing the utility of surrogates in conservation.
+Furthermore, this improvement was associated with only a moderate increase in cost in both scenarios (scenario without opportunity: `r round((score(multi.spp.prioritisations[[2]])/score(multi.spp.prioritisations[[1]]))*100,2)` %; scenario with opportunity cost: `r  round((score(multi.spp.prioritisations.with.cost[[2]])/score(multi.spp.prioritisations.with.cost[[1]]))*100,2)` %)--emphasising the utility of surrogates in conservation.
 
@@
 But as previously, using more geographic distances were found to be adequate and more complex measures of connectivity cannot generally be used in more large-scale planning scenarios.
-Third, we used AFLP data to characterize the main gradients of genetic variation.
-Whilst next-generation sequencing provides higher resolution genetic information than AFLP data [reviewed in @r480], our methodology would still have utilized only the main gradients of the genetic variation in order to generate prioritizations in a feasible amount of time.
+Third, we used AFLP data to characterise the main gradients of genetic variation.
+Whilst next-generation sequencing provides higher resolution genetic information than AFLP data [reviewed in @r480], our methodology would still have utilised only the main gradients of the genetic variation in order to generate prioritisations in a feasible amount of time.
 Overall, we are confident that the limitations associated with our study are relatively minor.
@@
 We found that environmental and geographic are generally effective surrogates for most species.
-Although prioritizations generated using amount-only targets tended to secure most of the adaptive variation for most species, the addition of surrogate targets to the planning process substantially improved the proportion of intra-specific neutral genetic variation.
+Although prioritisations generated using amount-only targets tended to secure most of the adaptive variation for most species, the addition of surrogate targets to the planning process substantially improved the proportion of intra-specific neutral genetic variation.
 Conservation planners are therefore urged to include these surrogates in conservation planning exercises.

### 2016-06-30 code/rmarkdown/article.Rmd
+ prepare for pnas submission

@@
 title: "Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?"
-author: |
-  | Jeffrey O. Hanson$^1$, Jonathan R. Rhodes$^2$, Cynthia Riginos$^1$, Hugh P. Possingham$^1$, Richard A. Fuller$^1$
-  | $^1$School of Biological Sciences, The University of Queensland, Brisbane, QLD, Australia
-  | $^2$School of Geography, Planning and Environmental Management, The University of Queensland, Brisbane, QLD, Australia
-  | Correspondance should be addressed to jeffrey.hanson@uqconnect.edu.au
-date: "`r format(Sys.time(), '%d %B %Y')`"
-tags: ["conservation", "biodiversity", "AFLP", "reserve-selection"]
-abstract: "Insert abstract here."
-output:
-  rmarkdown::pdf_document:
-    toc: false
-    toc_depth: 5
-    keep_tex: yes
-    fig_caption: yes
-    includes:
-      in_header: preamble-latex.tex
-      after_body: postamble-latex.tex
-fontsize: 11pt
-documentclass: article
+author:
+  - name: Jeffrey O. Hanson
+    affiliation: a,1
+  - name: Jonathan R. Rhodes
+    affiliation: b
+  - name: Cynthia Riginos
+    affiliation: a
+  - name: Hugh P. Possingham
+    affiliation: a
+  - name: Richard A. Fuller
+    affiliation: a
+address:
+  - code: a
+    address: School of Biological Sciences, The University of Queensland, Brisbane, QLD, Australia
+  - code: b
+    address: School of Geography, Planning and Environmental Management, The University of Queensland, Brisbane, QLD, Australia
+corresponding_author:
+  - code: 1  
+  - text: "To whom correspondence should be addresssed. E-mail: jeffrey.hanson@uqconnect.edu.au
+author_contributions: |
+  All authors developed the study and drafted the manuscript. J.O.H performed analyses.
+abstract: |
+  Please provide an abstract of no more than 250 words in a single paragraph.
+Abstracts should explain to the general reader the major contributions of the article.
+References in the abstract must be cited in full within the abstract itself and cited in the text.
+significance: |
+  Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality.
+The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership.
+The Significance Statement appears in the paper itself and is required for all research papers.
+acknowledgements: |
+  JOH is funded by an Australian Postgraduate Award (APA) scholarship.
+RAF has an Australian Research Council Future Fellowship.
+This work was supported by the Centre of Excellence for Environmental Decisions (CEED) and the Landscape Ecology and Conservation Group (LEC) at The University of Queensland.
+keywords:
+  - conservation
+  - biodiversity
+  - AFLP
+  - reserve-selection
+  - genetic diversity
+pnas_type: pnasresearcharticle
 bibliography: references.bib
 csl: reference-style.csl
+output: rticles::pnas_article
 ---
@@
 options(error=NULL)
+
+# download basemap
+data(countriesHigh)
+countries.FPLY <- countriesHigh[
+	countriesHigh$ADMIN %in% c(
+		'Italy', 'Switzerland', 'France', 'Austria',
+		'Germany', 'Slovenia', 'Croatia', 'Hungary',
+		'Monaco', 'Germany', 'Slovakia', 'Czech Republic'
+	)
+,] %>% spFortify
+grid.FPLY <- spFortify(grid.PLY)
+grid.sub.FPLY <- spFortify(grid.sub.PLY)
+
+# define ggplot2 color wheel function
+gg_color_hue <- function(n) {
+	hues <- seq(15, 375, length=n+1)
+	hcl(h=hues, l=65, c=100)[1:n]
+}
 ```
 
-<!--  title page -->
-Short running title: Genetic surrogates in conservation \newline
-Abstract word count: XXXXX \newline
-Main text word count: XXXXX \newline
-Number of references: XXXX \newline
-Number of equations: XXXX \newline
-Number of figures: XXXX \newline
-Number of tables: XXXX \newline
-Number of text boxes: XXXX \newline
-\newpage
-
-## Introduction
-<!--  main text -->
+<!--  Introduction -->
 Protected areas spearhead conservation efforts [@r440].
@@
 
-## Methods
-### Study system
-We used species distribution and genomic data collected by the international IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for further explanation of data collection methods].
-This data set contains information for 27 alpine plant species (Figure 1a).
-It was collected using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx.
-22.3 km $\times$ 25 km; Figure S1).
-Project members visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals.
-They genotyped samples using amplified fragment length polymorphisms [AFLP; @r453], and constructed matrices denoting the presence/absence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; see Table 1 in @r478 for the number markers for each species).
-Thus the data set contains information describing the genomic properties of individuals for each species present in every second grid cell.
+Results {.unnumbered}
+=====================
 
-We chose this data set for several reasons.
-First, conservation planning exercises typically consider tens to hundreds of species with different evolutionary histories and life histories.
-To ensure that our results were relevant for conservation planners, we required a data set with a diverse set of species and a comparable number of species.
-The IntraBioDiv data set contains genetic data for 27 species with different evolutionary histories and life-history strategies.
-Second, we required genetic data for all species to be collected using a standardised sampling scheme, and this data set used a grid to collect samples.
-Third, we required the study area to contain broad-scale environmental variation to test the effectiveness of environmental variables.
-This data set spans across large attitudinal gradients.
-Although this data set contains AFLPs  which not ideal as genomic representatives (cf.
-single nucleotide polymorphisms; SNPs), no multi-species genomic data set with similar properties currently exists.
-Thus this data set was well suited for our study.
-
-This data set has previously been used to address a range of different questions.
-Many studies have used this data set to explore patterns of adaptive [@r484; @r451; @r475; @r455; @r485] and neutral genetic variation [@r478; @r451; @r486].
-However, to our knowledge, only a single study has used this data set to explore the effectiveness of a surrogate in the context of conservation planning.
-Taberlet _et al._ [-@r486] investigated whether prioritisations generated using the spatial distribution of a comprehensive set species (> 300) could adequately secure the intra-specific genetic variation for 27 species.
-They found that the prioritisation based on species distributions was substantially different to the prioritisation based on genomic data.
-Their results suggest that prioritisations based on species distribution data alone will be unable to secure intra-specific genetic variation.
-However, whether or not environmental and geographic variables can augment prioritisations remains untested.
-
-### Landscape data
-We examined the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation (respectively).
-We utilised the sampling grid used by the IntraBioDiv project as planning units to develop prioritisations.
-We used only planning units that contained at least a single individual ($n = `r nrow(grid.DF)`$) to reduce computational burden.
-To examine the potential effects of variation in cost (see below), we calculated the total human population density inside each planning unit [obtained at 1 km$^2$ resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472] and used this to represent acquisition cost (Figure 1b).
-
-To describe the geographic location of each planning unit, we projected the grid into an equi-distant coordinate system (Europe Equidistant Conic: $\texttt{ESRI:102031}$), calculated their centroids, and extracted their two-dimensional coordinates.
-Thus we constructed two-dimensional geographic space--wherein each planning unit occupies a single point--as a potential surrogate for neutral genetic variation.
-We expected planning units further apart from each other in this geographic space to contain individuals with increasingly different patterns in their neutral loci.
-
-To describe the environmental characteristics of each planning unit, we used broad-scale climatic variation.
-We obtained data for 19 bioclimatic variables across the extent of the study area [at $1^{\prime\prime}$ resolution; @r33].
-We used this data set because it is commonly to map habitat suitability in conservation planning [@r493].
-These layers and the planning units were projected into an equal-area coordinate system (Europe Lambert Conformal Conic: $\texttt{ESRI:102014}$).
-To reduce dimensionality, the 19 bioclimatic variables were clipped to the planning units and subjected to a principal components analysis (PCA; using \texttt{ArcMap} version 10.2.2; Table S1).
-The first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components (PCs) cumulatively explained `r round(pca.DF[[4]][surrogate.params.LST[[MODE]]$number.components],1)` % of the variation and were used to generate new layers.
-We calculated the average value for each PC layer in each planning unit, and used these values to characterise climatic variation among the planning units (Figure S2).
-Thus we constructed a three-dimensional environmental space--wherein each planning unit occupies a single point--as a potential surrogate for adaptive variation.
-We expect planning units further apart from each other in this environmental space to contain individuals with increasingly different patterns in their adaptive loci.
-
-### Adaptive and neutral genetic data
-To investigate the effectiveness of our supposed surrogates for adaptive and neutral genetic variation, we first needed to identify which of the sampled loci--if any--were adaptive.
-We used two outlier detection methods to achieve this [an individual- and a population-level method as recommended by @r495].
-The basic premise underpinning such methods is that neutral loci are expected to exhibit a specific level of variation among populations, and loci that deviate from this expectation should be under selection [reviewed in @r480].
-The advantage of these methods--in contrast with environmental association analyses [@r489]--is that they do not use environmental data to identify loci under selection, which would have biased our analysis.
-Loci identified as outliers in both analyses were treated as adaptive and the rest were treated as neutral.
-To avoid false-positives, we omitted loci from both outlier detection methods where the global frequency of the minor allele was $> 0.1$ and assumed they were neutral.
-
-The first outlier detection method we used involved fitting multinominal-Dirichlet models implemented in \texttt{BayeScan} [version 2.1; @r456].
-We adopted a similar methodology to Bothwell _et al._ [-@r455] and applied it to each of the species in the data set.
-Following their methodology, we first grouped conspecifics into genetic lineages to avoid falsely classifying loci as adaptive.
-To complete this first step, for each species, we fit the admixture and correlated alleles model implemented in \texttt{Structure} [version 2.3.4; r@461; r@462] using the number of populations for each species previously determined by [@r478; `structure.params.LST[[MODE]]$numruns` replicates per K; `r structure.params.LST[[MODE]]$admburnin` admixture burnin iterations; `r structure.params.LST[[MODE]]$burnin` burnin iterations; `r structure.params.LST[[MODE]]$numreps` iterations used for inference].
-We used \texttt{ClumPP} [version 1.1.2; @r459] to combine replicate runs to group individuals into populations (greedy algorithm using the $G^{\prime}$ static; `r clumpp.params.LST[[MODE]]$repeats` iterations).
-We then ran $\texttt{Bayescan}$ [@r456] for each species using these populations [1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations].
-Additionally, we omitted individuals from the \texttt{Bayescan} analysis if their population membership was uncertain (maximum membership probability $< `r structure.params.LST[[MODE]]$probability.threshold`$).
-We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate [$q \leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].
-
-The second outlier detection method involved fitting PCAs to identify outlier loci [using methods implemented in the \texttt{pcadapt R} package; version 3.0; @r487; @r496].
-To enable comparisons between the two outlier detection analyses, we used the same individuals in this analysis as used in the \texttt{BayeScan} analysis.
-For each of the 27 species, we first imputed missing data to permit further analysis.
-Specifically, the missing value for a given locus for a given individual was replaced with the average frequency of that locus among all conspecifics.
-Next, we ran a PCA and extracted the minimum number of components needed to secure the over-arching population level variation among the loci (`r pcadapt.params.LST[[MODE]]$min.variance.explained*10` % of the variation in loci).
-We then used Mahalanobis distances [@r398] to compute _P_ values and then _q_-values [using the \texttt{qvalue R} package; @r491].
-Outlier loci were detected using a suitable false discovery rate ($q < `r pcadapt.params.LST[[MODE]]$alpha.level`$).
-
-After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species.
-For each species, we discarded the population groupings, and partitioned adaptive and neutral loci into separate matrices.
-We applied separate non-metric multi-dimensional scaling analyses to each matrix [NMDS; implemented in the \texttt{vegan R} package; @r458] using Gower distances [@r464; using the \texttt{cluster R} package to accommodate sparsity; @r457] to derive continuous variables that described the main gradients of adaptive and neutral genetic variation (separately) for each species (Table S2).
-To ensure that the ordinations described a sufficient amount of the genomic variation, we ran successive scaling analyses with increasing dimensionality until a sufficient stress value was obtained (maximum stress value $\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for each analysis).
-Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the individuals in each grid cell to express the typical genomic characteristics of individuals in the cells.
-
-The previous set of analyses resulted in an adaptive (if detected) and neutral genetic space for each species.
-For a given species, each planning unit occupied by the species was associated with a multi-dimensional point in the species' neutral genetic space.
-Planning units that were closer together in this space are occupied by individuals with similar polymorphisms in their neutral loci.
-Similarly, if the species was associated with adaptive loci, each planning unit occupied by the species was associated with a multi-dimensional point in the species' adaptive genetic space.
-By spreading out conservation effort across a given genetic space, and in turn securing planning units occupied by individuals with increasingly different polymorphisms, prioritisations can secure more genetic variation [see @r465 for discussion on attribute spaces].
-
-## Prioritisation method
-We used the unreliable representative and adequate prioritisation (URAP) formulation of the reserve selection problem [@r465].
-This formulation can identify the minimum number of planning units required to preserve both a proportion of the species' range (using amount-based targets) and a proportion of variation found across the species' range [using space-based targets and attribute spaces; _sensu_ @r465].
-We treated the environmental and geographic surrogate variables as separate attribute spaces.
-We also treated the adaptive (if detected) and neutral variables as separate attribute spaces for each species.
-We set the demand points for each species and each attribute space using the values associated with the planning units.
-For every prioritisation we generated, we calculated the proportion of the adaptive (if detected) and neutral genetic variation that every prioritisation secured to assess its performance.
-Prioritisations associated with negative values--because they secured only a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis.
-We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality using the $\texttt{rapr}$ R package [@r465] and $\texttt{Gurobi}$ [@r463].
-
-## Statistical analyses
-To address the first aim--to determine if the environmental and geographic variables correlate with spatial patterns of adaptive and neutral genetic variation--we computed dissimilarity matrices between the planning units, and tested for correlations between matrices based on surrogate and genetic data.
-For each species, we computed two dissimilarity matrices between the planning units occupied by the species in terms of their environmental conditions (their coordinates in the environmental space) and geographic position (their coordinates in the geographic space).
-We then computed another dissimilarity matrix between the planning units using their values in the species' neutral genetic space.
-If the species was associated with adaptive loci, we also computed another dissimilarity matrix between the planning units using the species' adaptive genetic space.
-All dissimilarity matrices calculated using Euclidean distances.
-
-We then fit maximum likelihood population-effects (MLPE) models [@r494; implemented in the \texttt{ResistanceGA R} package; @r488].
-These models explicitly accommodate the structure of dissimilarity matrices using random effects.
-For each species, we fit a MLPE model to correlate the species' dissimilarity matrices based on geographic position and neutral genetic variation.
-Additionally, if the species was associated with adaptive loci, we also fit a MLPE model to correlate the species' dissimilarity matrices calculated using environmental conditions and adaptive genetic variation.
-All predictor variables were z-transformed prior to model fitting to improve convergence.
-To test if the surrogate variables explained a substantial amount of genetic variation, we conducted $\chi^2$ tests between each model and its corresponding null model, and applied Bonferroni corrections.
-
-To address the second aim--to determine if the environmental and geographic variables are effective surrogates for adaptive and neutral genetic variation--we generated a collection of single-species prioritisations and computed the proportion of intra-specific genetic variation they secured.
-For each species, we generated `r rapr.params.LST[[MODE]]$surrogacy.analysis$replicates` prioritisations by randomly selecting different combinations of planning units that were occupied by the species.
-We then computed the proportion of adaptive genetic variation and environmental variation secured in each prioritisation.
-We then generated another random set of prioritisations, and computed the proportion of neutral genetic variation and geographic variation secured in each prioritisation.
-
-We fit two full generalised linear models (GLM) with a logit links.
-The first model was fit to the proportion of adaptive genetic variation secured in a prioritisation, and contained a continuous predictor variable measuring the proportion of environmental variation also secured in the prioritisation, a categorical variable indicating the species for which the prioritisation was generated, and an interaction term between these predictor variables.
-The second model was fit to the proportion of neutral genetic variation secured in a prioritisation, and contained a continuous predictor variable assessing the proportion of geographic variation also secured in the prioritisation.
-Similar to the previous model, this model also contained a predictor variable indicating the species for which the prioritisations were generated, and an interaction term.
-We subjected these models to a backwards step-wise term deletion routine to assess term significance.
-To gauge the performance of these surrogates for each species, we fit a suite of models correlating the proportion of adaptive and neutral genetic variation held in each prioritisation and the environmental and geographic variation they also secured (respectively).
-We then computed the Cragg and Uhler's pseudo-R$^2$ value for each model [using the \texttt{pscl} R package; @r492].
-
-To address the third aim--to determine if the use of surrogate-based targets can actually improve prioritisations--we investigated the effectiveness of surrogate-based targets under three scenarios with increasing levels of complexity.
-To assess the benefits of surrogate-based targets under the simplest of circumstances, our first scenario involved generating single-species prioritisations for each species.
-However, most large-scale prioritisations involve a comprehensive set of species.
-By forcing solutions to secure individuals in multiple communities, this may result in prioritisations that secure a greater proportion of intra-specific genetic variation for each species.
-To test for this effect, our second scenario involved generating multi-species prioritisations.
-Finally, most large-scale prioritisations also explicitly consider opportunity cost to ensure that prioritisations are cost-effective.
-Our third scenario involved generating multi-species prioritisations that satisfied targets for minimal opportunity cost.
-
-For each scenario, we generated three types of prioritisations using (1) only amount-based targets, (2) both amount- and surrogate-based targets, and (3) both amount- and genetic-based targets.
-The purely amount-based prioritisations served as our baseline--representing the prioritisations typically generated in conservation planning exercises (eg.
-by \texttt{Marxan}).
-We used `r (rapr.params.LST[[MODE]]$scenario.analysis$amount.target*100)` % amount-based targets in all prioritisations to ensure an adequate proportion of habitat was secured for each species.
-Based on the results from the analyses used to address our second aim, we set surrogate-based targets as `r (rapr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100)` % and genetic-based targets as `r (rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)` % to ensure that most of the genetic variation was secured for each species.
-To accommodate the fact that there were a large number of optimal solutions for the single-species amount-based prioritisations, we generated `r (rapr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates)` replicate prioritisations.
-We computed the proportion of adaptive (if present) and neutral genetic variation secured for each species in the prioritisations.
-
-We investigated the benefits of using surrogate-based targets in prioritisations using generalised linear mixed-effects models fit with logit liknks [GLMMs; using the \texttt{lme4 R} package; @469].
-We fit a full model to the proportion of genetic variation secured in a given prioritisation.
-This model contained categorical predictor variables to indicate the combination of targets used to generate the prioritisations (amount-only, amount and surrogate targets, and amount and genetic targets), the scenario that was used to generate the prioritisation (single-species, multi-species, and multi-species with cost), the type of genetic variation measured (adaptive and neutral), and all interactions between these variables.
-This model accommodated prioritisations generated for the same species using a random intercept term.
-We subject the full model to a backwards step-wise term deletion routine to assess term significance.
-We conducted a post-hoc analysis on the minimal adequate model using Tukey contrasts with a Bonferroni correction [using the \texttt{multcomp R} package; @r470].
-
-## Results
 <!-- summary of methods -->
@@
 
+```{r, echo=FALSE, results='asis'}
+curr.file <- tempfile(fileext='.tex')
+x <- correlation.model.summary.DF %>%
+	select(-contains('raw')) %>%
+	mutate(
+		Environmental.corr.P=pvalString(Environmental.corr.P),
+		Geographic.corr.P=pvalString(Geographic.corr.P),
+		Geographic.Marginal.R2=pvalString(Geographic.Marginal.R2),
+		Geographic.Conditional.R2=pvalString(Geographic.Conditional.R2),
+		Environmental.Marginal.R2=pvalString(Environmental.Marginal.R2),
+		Environmental.Conditional.R2=pvalString(Environmental.Conditional.R2),
+		blank=rep('', length(Species)),
+		Species=paste0('\\textit{', gsub('\\ ', ' ', Species), '}')) %>%
+	mutate(Environmental.Chisq=sapply(round(Environmental.Chisq,2), as.character)) %>%
+	mutate(Geographic.Chisq=sapply(round(Geographic.Chisq,2), as.character)) %>%
+	select(Species:Environmental.corr.P, Geographic.Marginal.R2:Geographic.corr.P) %>%
+	latex(file=curr.file, 
+		cgroup=c('', 'Environmental', 'Geographic'), n.cgroup=c(1, 4, 4), col.just=rep('c', 9),
+		digits=2, rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, here=TRUE,
+		colheads=c('\\bfseries{Species}', '$MR^2$', '$CR^2$', '$\\chi^2$', '\\textit{P}', '$MR^2$', '$CR^2$', '$\\chi^2$', '\\textit{P}'),
+		caption='Summary of maximum likelihood population-effects (MLPE) models.
+Briefly, for each species associated with adaptive loci, a model was fit to correlate dissimilarities among planning units where the species in terms of the adaptive genetic characteristics of individuals in them with dissimilarities in terms of the environmental conditions inside the planning units (denoted in the environmental column group).
+Similarly, for each species, a model was fit to correlate dissimilarities among the planning units in terms of the neutral genetic characteristics of indiviudals in them with dissimilarities in terms of the geographic position of the planning units (denoted in the geographic column group).
+Data shows the marginal $R^2$ ($MR^2$) and conditional $R^2$ ($CR^2$) for each model, and also the results from comparing the models to their corresponding null models.
+All $\\chi^2$ tests had 1 degree of freedom.'
+	)
+readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
+	gsub(pattern='\\cline{1-', replacement='\\cline{2-6} \\cline{8-', fixed=TRUE) %>% 
+	gsub(pattern='{c}{\\textit{P}}\\tabularnewline', replacement='{c}{\\textit{P}}\\T\\B\\tabularnewline', fixed=TRUE) %>% 
+	gsub(pattern='toprule', replacement='toprule[1pt]', fixed=TRUE) %>%
+	gsub(pattern='bottomrule', replacement='bottomrule[1pt]', fixed=TRUE) %>%
+	cat()
+```
+
 <!-- surrogacy analysis -->
@@
 
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=4.5, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations.
+Panel (a) shows the relationship between the proportion of intra-specific adaptive genetic variation and environmental variation secured in a prioritization.
+Similarly, panel (b) shows the relationship between the proportion of intra-specific neutral variaiton and geograpgic variation secured in a prioritization.
+Panels (c) and (d) show the distribution of $R^2$ values for these models.'}
+# make figure
+p1 <- ggplot(aes(x=surrogate.held,y=genetic.held), data=env.surrogacy.DF) +
+	xlab('Environmental variation\nsecured (%)') + 
+	ylab('Adaptive genetic\nvariation secured (%)') +
+	theme_classic() + theme(axis.line.x=element_line(color='black'),
+		axis.line.y=element_line(color='black')) +
+	geom_text(x=0.05, y=0.95, label='(a)') +
+	coord_cartesian(xlim=c(0, 1))
+for (i in seq_along(env.surrogacy.LST))
+	p1 <- p1 + stat_smooth(method='glm', method.args=list(family='binomial'),
+		data=filter(env.surrogacy.DF, Species==unique(Species)[i]),
+		fullrange=TRUE, se=FALSE, colour='black')
+
+p2 <- ggplot(aes(x=surrogate.held,y=genetic.held), data=geo.surrogacy.DF) +
+	xlab('Geographic variation\nsecured (%)') + 
+	ylab('Neutral genetic\nvariation secured (%)') +
+	theme_classic() + theme(axis.line.x=element_line(color='black'),
+		axis.line.y=element_line(color='black', colour='transparent')) +
+	geom_text(x=0.05, y=0.95, label='(b)') +
+	coord_cartesian(xlim=c(0, 1))
+for (i in seq_along(geo.surrogacy.LST))
+	p2 <- p2 + stat_smooth(method='glm', method.args=list(family='binomial'),
+		data=filter(geo.surrogacy.DF, Species==unique(Species)[i]),
+		fullrange=TRUE, se=FALSE, colour='black')
+
+p1.hist <- ggplot(aes(r2CU),
+		data=filter(env.surrogate.spp.DF, !is.na(r2CU))) +
+	geom_histogram(fill='grey10', bins=5) +
+	ylab('\nFrequency') + xlab(expression(Environmental~variables~(R^2))) +
+	theme_classic() + theme(axis.line.x=element_line(color='black'),
+		axis.line.y=element_line(color='black'),
+		plot.background=element_rect(fill='transparent', colour='transparent')) +
+		geom_text(x=0.05, y=6, label='(c)') +
+		coord_cartesian(xlim=c(0, 1))
+
+p2.hist <- ggplot(aes(r2CU),
+		data=filter(geo.surrogate.spp.DF, !is.na(r2CU))) +
+	geom_histogram(fill='grey10', bins=10) +
+	ylab('\nFrequency') + xlab(expression(Geographic~variables~(R^2))) +
+	theme_classic() + theme(axis.line.x=element_line(color='black'),
+		axis.line.y=element_line(color='black'), 
+		plot.background=element_rect(fill='transparent')) +
+		geom_text(x=0.05, y=8, label='(d)') +
+		coord_cartesian(xlim=c(0, 1))
+
+# render figure
+grid.arrange(p1,p2,p1.hist,p2.hist,nrow=2)
+```
+
 <!-- scenario analysis -->
@@
 We found that on average the prioritisations tended to secure more adaptive than neutral genetic variation (`r round(plogis(scenario.GLHT.3.DF$statistic[[1]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.3.DF$std.error[[1]])*100,2)` SE; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.3.DF$Pr..Chisq.[[2]])`$).
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=6.5, fig.cap='Planning unit selection frequencies.
+The first row of panels (a--c) show single-species prioritizations generated assuming equal costs.
+The second row of panels (d--f) show single-species prioritizations generated using opportunity costs.
+The third row of panels (g--i) show multi-species prioritizations generated assuming equal costs.
+The fourth row of panels (j--l) show multi-species generated using opportunity costs.
+The left column of panels (a, d, g, j) show prioritizations generated using only amount-based targets.
+The middle column of panels (b, e, h, k) show prioritizations generated using amount- and surrogate-based targets.
+The right column of panels (c, f, i, l) show prioritizations generated using amount- and genetic-based targets.
+The color of each planning unit indicates its selection frequency.
+Note that the selection frequencies for the single-species prioritizations (panels a--f) are based on their average frequency among multiple prioritizations generated for each species.
+The selection frequences for the multi-species prioritizations (pabels g--l) are based on the optimal pioritization (thus they are all either zero or one).'}
+# prepare data for plotting
+grid.sub.FPLY <- grid.sub.PLY
+for (i in seq_along(single.spp.prioritisations[[1]]))
+	grid.sub.FPLY@data[[paste0('Single-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations, function(x){colMeans(x[[i]]@results@selections)}))
+for (i in seq_along(single.spp.prioritisations.with.cost[[1]]))
+	grid.sub.FPLY@data[[paste0('Single-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations.with.cost, function(x){colMeans(x[[i]]@results@selections)}))
+for (i in seq_along(multi.spp.prioritisations))
+	grid.sub.FPLY@data[[paste0('Multi-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations[[i]]@results@selections)
+for (i in seq_along(multi.spp.prioritisations.with.cost))
+	grid.sub.FPLY@data[[paste0('Multi-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations.with.cost[[i]]@results@selections)
+grid.sub.FPLY <- spFortify(grid.sub.FPLY)
+grid.sub.FPLY <- grid.sub.FPLY %>% gather(
+		Combination,
+		`Selection.Frequency`,
+		`Single-species (equal costs).Amount`:`Multi-species (opportunity costs).Genetic`
+	) %>% mutate(
+		Context=sapply(strsplit(as.character(Combination), '.', fixed=TRUE), `[[`, 1),
+		Prioritization=sapply(strsplit(as.character(Combination), '.', fixed=TRUE), `[[`, 2)
+	) %>% mutate(
+		Prioritization=revalue(Prioritization, c('Amount'='Amount targets', 'Surrogate'='Amount & surrogate\ntargets', 'Genetic'='Amount & genetic\ntargets')),
+		Prioritization=factor(as.character(Prioritization), levels=c('Amount targets', 'Amount & surrogate\ntargets', 'Amount & genetic\ntargets')),
+		Context=factor(
+			gsub(' (', '\n(', Context, fixed=TRUE),
+			levels=c("Single-species\n(equal costs)", "Single-species\n(opportunity costs)",
+				"Multi-species\n(equal costs)", "Multi-species\n(opportunity costs)"))
+	)
+panel.DF <- grid.sub.FPLY %>% select(Context, Prioritization, long, lat) %>% 
+	mutate(
+		lat=min(lat)+(diff(range(lat))*0.98),
+		long=min(long)+(diff(range(long))*0.02)
+	) %>%
+	group_by(Context,Prioritization) %>% filter(c(TRUE, rep(FALSE, length(Context)-1))) %>%
+	mutate(
+		panel.letter=paste0('(',letters[(as.integer(Context)-1)*3 + as.integer(Prioritization)],')'),
+		panel.letter=c(as.character(panel.letter[1]), rep(NA_character_, length(Context)-1))
+	)
+# make maps
+ggplot() +
+	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
+		fill='grey85', color='grey70') +
+	geom_polygon(data=grid.sub.FPLY, aes_string(x='long', y='lat', 
+		group='group', fill='Selection.Frequency'),
+		alpha=0.8, color='grey10') +
+	geom_text(aes_string(x='long',y='lat', label='panel.letter'), data=panel.DF) +
+	theme_classic() +
+	theme(axis.ticks=element_blank(), axis.text=element_blank(),
+		plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(),
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white'),
+		legend.position='bottom', legend.key.width=unit(1.8, 'cm'),
+		legend.title=element_text(vjust=-5),
+		legend.margin=unit(c(0.1), 'cm'),
+		axis.title=element_blank()) +
+	coord_cartesian(
+		xlim=buffered.range(grid.sub.FPLY$long, 0),
+		ylim=buffered.range(grid.sub.FPLY$lat, 0)
+	) +
+	facet_grid(Context ~ Prioritization) +
+	scale_fill_gradientn(name='Selection Frequency (%)',
+		colors=brewer.pal(9, 'PuBu'),
+		guide=guide_colorbar(title.position='bottom', title.hjust=0.5,
+			ticks=element_line(color='black'), border=element_line(color='black'))
+	)
+```
 
@@
 
-## Discussion
+Discussion {.unnumbered}
+========================
+
 <!--  introduction -->
@@
 
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=7.4, fig.cap='The effectiveness of prioritizations at securing genetic variation.
+The top row of panels (a--b) show the performance of single-species prioritizations generated assuming equal costs among planning costs.
+The middle row of panels (c--d) show the performance of multi-species prioritizations generated assuming equal costs among planning units.
+The bottom row of panels (e--f) show the performance of multi-species prioritizations generated using opportunity costs.
+The left column of panels (a, c, d) and the right column of panels (b, d, f) show the proportion of adaptive and neutral genetic variation secured (respectively).
+Box plots show the effectiveness of prioritizations generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species.'}
+# prepare data for plotting
+scenario.PDF <- scenario.DF %>% group_by(Prioritisation, Context, Metric) %>% 
+	mutate(
+		letter.height=max(genetic.held, na.rm=TRUE)+0.1,
+		Combination=paste0(Prioritisation,'.',Metric,'.',Context)) %>%
+	filter(c(TRUE, rep(FALSE, length(Context)-1))) %>% 
+	ungroup() %>%
+	group_by(Context, Metric) %>% 
+	mutate(
+		panel.letter=paste0(
+			'(',letters[(as.integer(Context)*2)-1 + (as.integer(Metric)-1)],')'
+		),
+		panel.letter=c(as.character(panel.letter[1]), rep(NA_character_, length(Context)-1))
+	) %>% 
+	ungroup() %>% 
+	data.frame()
+
+scenario.SDF <- scenario.DF %>% 
+	group_by(Prioritisation, Context, Metric, Species) %>% 
+	summarize(genetic.held=mean(genetic.held, na.rm=TRUE)) %>%
+	data.frame()
+
+# make plot
+suppressWarnings({
+	ggplot(aes(x=Prioritisation,y=genetic.held),
+		data=scenario.SDF) +
+		geom_boxplot(position=position_dodge(0.9)) +
+		geom_text(aes(x=0.6, y=1.15, label=panel.letter),
+			data=scenario.PDF) +
+		scale_fill_manual(name='Prioritisation',
+			values=c('grey80','grey50','grey20')) +
+		geom_abline(slope=0, color='red', linetype='dashed',
+			intercept=rapr.params.LST[[MODE]]$scenario.analysis$genetic.target) +
+		ylab('Genetic variation secured (%)') +
+		xlab('') +
+		scale_y_continuous(breaks=c(0,0.25,0.5,0.75,1), limits=c(NA, 1.2)) +
+		theme_classic() +
+		theme(strip.background = element_rect(fill='grey20'),
+			strip.text = element_text(color='white'),
+			panel.border = element_rect(color='black', fill=NA),
+			axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
+		facet_grid(Context ~ Metric)
+})
+```
+
 All the multi-species prioritisations secured a representative sample of most of the species' intra-specific adaptive genetic variation (Figures 4e and 4g).
@@
 
-## Acknowledgements
-JOH is funded by an Australian Postgraduate Award (APA) scholarship.
-RAF has an Australian Research Council Future Fellowship.
-This work was supported by the Centre of Excellence for Environmental Decisions (CEED) and the Landscape Ecology and Conservation Group (LEC) at The University of Queensland.
+Materials and Methods {.unnumbered}
+===================================
 
-## References
+Study system {#study-system .unnumbered}
+----------------------------------------
+
+We used species distribution and genomic data collected by the international IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for further explanation of data collection methods].
+This data set contains information for 27 alpine plant species (Figure 1a).
+It was collected using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx.
+22.3 km $\times$ 25 km; Figure S1).
+Project members visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals.
+They genotyped samples using amplified fragment length polymorphisms [AFLP; @r453], and constructed matrices denoting the presence/absence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; see Table 1 in @r478 for the number markers for each species).
+Thus the data set contains information describing the genomic properties of individuals for each species present in every second grid cell.
+
+```{r, echo=FALSE, fig.width=6, fig.height=1.5, fig.cap="Map of the study area.
+Squares denote planning units.
+Panel (a) depicts patterns of species riches.
+Planning units with a brighter color are inhabited by more species.
+Panel (b) shows the distribution of opportunity cost across the study area (computed as total population density).
+Note that a quantile-based color ramp is used in this panel."}
+
+# calculate species richness
+grid.PLY$Species_richness <- grid.PLY@data %>%
+	select(5:(4+n_distinct(spp.samples.DF$species))) %>% as.matrix() %>% rowSums()
+
+panel.DF <- grid.sub.FPLY %>% 
+	summarize(lat=min(lat)+(diff(range(lat))*0.96),
+		long=min(long)+(diff(range(long))*0.04))
+
+# plot species richness
+p1 <- ggplot() +
+	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
+		fill='grey85', color='grey70') +
+	geom_polygon(data=spFortify(grid.PLY), aes(x=long, y=lat, 
+		group=group, fill=Species_richness), color='grey10') +
+	scale_fill_gradientn(name='Richness',
+		colors=brewer.pal(9,'PuBu'),
+		guide=guide_colorbar(ticks=element_line(color='black'), border=element_line(color='black'))
+	) +
+	theme_classic() +
+	theme(
+		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"), axis.title=element_blank(),
+		panel.border=element_rect(color='black', fill=NA, size=1)
+	) +
+	coord_cartesian(
+		xlim=buffered.range(grid.FPLY$long, 0),
+		ylim=buffered.range(grid.FPLY$lat, 0)
+	) + geom_text(aes_string(x='long',y='lat'), data=panel.DF, label='(a)')
+
+qn01 <- scales::rescale(quantile(grid.PLY@data$pop.density, seq(0,1,length.out=9), na.rm=TRUE))
+p2 <- ggplot() +
+	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
+		fill='grey85', color='grey70') +
+	geom_polygon(data=spFortify(grid.PLY), aes(x=long, y=lat, 
+		group=group, fill=pop.density), color='grey10') +
+	scale_fill_gradientn(name='Cost',
+		colors=brewer.pal(9, 'PuBu'),
+		values=qn01,
+		guide=guide_colorbar(ticks=element_line(color='black'), border=element_line(color='black'))
+	) +
+	theme_classic() +
+	theme(
+		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"), axis.title=element_blank(),
+		panel.border=element_rect(color='black', fill=NA, size=1)
+	) +
+	coord_cartesian(
+		xlim=buffered.range(grid.FPLY$long, 0),
+		ylim=buffered.range(grid.FPLY$lat, 0)
+	)  + geom_text(aes_string(x='long',y='lat'), data=panel.DF, label='(b)')
+
+grid.arrange(p1,p2,nrow=1)
+```
+
+We chose this data set for several reasons.
+First, conservation planning exercises typically consider tens to hundreds of species with different evolutionary histories and life histories.
+To ensure that our results were relevant for conservation planners, we required a data set with a diverse set of species and a comparable number of species.
+The IntraBioDiv data set contains genetic data for 27 species with different evolutionary histories and life-history strategies.
+Second, we required genetic data for all species to be collected using a standardised sampling scheme, and this data set used a grid to collect samples.
+Third, we required the study area to contain broad-scale environmental variation to test the effectiveness of environmental variables.
+This data set spans across large attitudinal gradients.
+Although this data set contains AFLPs  which not ideal as genomic representatives (cf.
+single nucleotide polymorphisms; SNPs), no multi-species genomic data set with similar properties currently exists.
+Thus this data set was well suited for our study.
+
+This data set has previously been used to address a range of different questions.
+Many studies have used this data set to explore patterns of adaptive [@r484; @r451; @r475; @r455; @r485] and neutral genetic variation [@r478; @r451; @r486].
+However, to our knowledge, only a single study has used this data set to explore the effectiveness of a surrogate in the context of conservation planning.
+Taberlet _et al._ [-@r486] investigated whether prioritisations generated using the spatial distribution of a comprehensive set species (> 300) could adequately secure the intra-specific genetic variation for 27 species.
+They found that the prioritisation based on species distributions was substantially different to the prioritisation based on genomic data.
+Their results suggest that prioritisations based on species distribution data alone will be unable to secure intra-specific genetic variation.
+However, whether or not environmental and geographic variables can augment prioritisations remains untested.
+
+Landscape data {#landscape-data .unnumbered}
+--------------------------------------------
+
+We examined the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation (respectively).
+We utilised the sampling grid used by the IntraBioDiv project as planning units to develop prioritisations.
+We used only planning units that contained at least a single individual ($n = `r nrow(grid.DF)`$) to reduce computational burden.
+To examine the potential effects of variation in cost (see below), we calculated the total human population density inside each planning unit [obtained at 1 km$^2$ resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472] and used this to represent acquisition cost (Figure 1b).
+
+To describe the geographic location of each planning unit, we projected the grid into an equi-distant coordinate system (Europe Equidistant Conic: $\texttt{ESRI:102031}$), calculated their centroids, and extracted their two-dimensional coordinates.
+Thus we constructed two-dimensional geographic space--wherein each planning unit occupies a single point--as a potential surrogate for neutral genetic variation.
+We expected planning units further apart from each other in this geographic space to contain individuals with increasingly different patterns in their neutral loci.
+
+To describe the environmental characteristics of each planning unit, we used broad-scale climatic variation.
+We obtained data for 19 bioclimatic variables across the extent of the study area [at $1^{\prime\prime}$ resolution; @r33].
+We used this data set because it is commonly to map habitat suitability in conservation planning [@r493].
+These layers and the planning units were projected into an equal-area coordinate system (Europe Lambert Conformal Conic: $\texttt{ESRI:102014}$).
+To reduce dimensionality, the 19 bioclimatic variables were clipped to the planning units and subjected to a principal components analysis (PCA; using \texttt{ArcMap} version 10.2.2; Table S1).
+The first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components (PCs) cumulatively explained `r round(pca.DF[[4]][surrogate.params.LST[[MODE]]$number.components],1)` % of the variation and were used to generate new layers.
+We calculated the average value for each PC layer in each planning unit, and used these values to characterise climatic variation among the planning units (Figure S2).
+Thus we constructed a three-dimensional environmental space--wherein each planning unit occupies a single point--as a potential surrogate for adaptive variation.
+We expect planning units further apart from each other in this environmental space to contain individuals with increasingly different patterns in their adaptive loci.
+
+Adaptive and neutral genetic data {#adaptive-and-neutral-genetic-data .unnumbered}
+----------------------------------------------------------------------------------
+
+To investigate the effectiveness of our supposed surrogates for adaptive and neutral genetic variation, we first needed to identify which of the sampled loci--if any--were adaptive.
+We used two outlier detection methods to achieve this [an individual- and a population-level method as recommended by @r495].
+The basic premise underpinning such methods is that neutral loci are expected to exhibit a specific level of variation among populations, and loci that deviate from this expectation should be under selection [reviewed in @r480].
+The advantage of these methods--in contrast with environmental association analyses [@r489]--is that they do not use environmental data to identify loci under selection, which would have biased our analysis.
+Loci identified as outliers in both analyses were treated as adaptive and the rest were treated as neutral.
+To avoid false-positives, we omitted loci from both outlier detection methods where the global frequency of the minor allele was $> 0.1$ and assumed they were neutral.
+
+The first outlier detection method we used involved fitting multinominal-Dirichlet models implemented in \texttt{BayeScan} [version 2.1; @r456].
+We adopted a similar methodology to Bothwell _et al._ [-@r455] and applied it to each of the species in the data set.
+Following their methodology, we first grouped conspecifics into genetic lineages to avoid falsely classifying loci as adaptive.
+To complete this first step, for each species, we fit the admixture and correlated alleles model implemented in \texttt{Structure} [version 2.3.4; r@461; r@462] using the number of populations for each species previously determined by [@r478; `structure.params.LST[[MODE]]$numruns` replicates per K; `r structure.params.LST[[MODE]]$admburnin` admixture burnin iterations; `r structure.params.LST[[MODE]]$burnin` burnin iterations; `r structure.params.LST[[MODE]]$numreps` iterations used for inference].
+We used \texttt{ClumPP} [version 1.1.2; @r459] to combine replicate runs to group individuals into populations (greedy algorithm using the $G^{\prime}$ static; `r clumpp.params.LST[[MODE]]$repeats` iterations).
+We then ran $\texttt{Bayescan}$ [@r456] for each species using these populations [1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations].
+Additionally, we omitted individuals from the \texttt{Bayescan} analysis if their population membership was uncertain (maximum membership probability $< `r structure.params.LST[[MODE]]$probability.threshold`$).
+We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate [$q \leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].
+
+The second outlier detection method involved fitting PCAs to identify outlier loci [using methods implemented in the \texttt{pcadapt R} package; version 3.0; @r487; @r496].
+To enable comparisons between the two outlier detection analyses, we used the same individuals in this analysis as used in the \texttt{BayeScan} analysis.
+For each of the 27 species, we first imputed missing data to permit further analysis.
+Specifically, the missing value for a given locus for a given individual was replaced with the average frequency of that locus among all conspecifics.
+Next, we ran a PCA and extracted the minimum number of components needed to secure the over-arching population level variation among the loci (`r pcadapt.params.LST[[MODE]]$min.variance.explained*10` % of the variation in loci).
+We then used Mahalanobis distances [@r398] to compute _P_ values and then _q_-values [using the \texttt{qvalue R} package; @r491].
+Outlier loci were detected using a suitable false discovery rate ($q < `r pcadapt.params.LST[[MODE]]$alpha.level`$).
+
+After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species.
+For each species, we discarded the population groupings, and partitioned adaptive and neutral loci into separate matrices.
+We applied separate non-metric multi-dimensional scaling analyses to each matrix [NMDS; implemented in the \texttt{vegan R} package; @r458] using Gower distances [@r464; using the \texttt{cluster R} package to accommodate sparsity; @r457] to derive continuous variables that described the main gradients of adaptive and neutral genetic variation (separately) for each species (Table S2).
+To ensure that the ordinations described a sufficient amount of the genomic variation, we ran successive scaling analyses with increasing dimensionality until a sufficient stress value was obtained (maximum stress value $\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for each analysis).
+Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the individuals in each grid cell to express the typical genomic characteristics of individuals in the cells.
+
+The previous set of analyses resulted in an adaptive (if detected) and neutral genetic space for each species.
+For a given species, each planning unit occupied by the species was associated with a multi-dimensional point in the species' neutral genetic space.
+Planning units that were closer together in this space are occupied by individuals with similar polymorphisms in their neutral loci.
+Similarly, if the species was associated with adaptive loci, each planning unit occupied by the species was associated with a multi-dimensional point in the species' adaptive genetic space.
+By spreading out conservation effort across a given genetic space, and in turn securing planning units occupied by individuals with increasingly different polymorphisms, prioritisations can secure more genetic variation [see @r465 for discussion on attribute spaces].
+
+Prioritisation method {#prioritisation-method .unnumbered}
+----------------------------------------------------------
+
+We used the unreliable representative and adequate prioritisation (URAP) formulation of the reserve selection problem [@r465].
+This formulation can identify the minimum number of planning units required to preserve both a proportion of the species' range (using amount-based targets) and a proportion of variation found across the species' range [using space-based targets and attribute spaces; _sensu_ @r465].
+We treated the environmental and geographic surrogate variables as separate attribute spaces.
+We also treated the adaptive (if detected) and neutral variables as separate attribute spaces for each species.
+We set the demand points for each species and each attribute space using the values associated with the planning units.
+For every prioritisation we generated, we calculated the proportion of the adaptive (if detected) and neutral genetic variation that every prioritisation secured to assess its performance.
+Prioritisations associated with negative values--because they secured only a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis.
+We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality using the $\texttt{rapr}$ R package [@r465] and $\texttt{Gurobi}$ [@r463].
+
+Statistical analyses {#statistical-analyses .unnumbered}
+----------------------------------------------------------
+
+To address the first aim--to determine if the environmental and geographic variables correlate with spatial patterns of adaptive and neutral genetic variation--we computed dissimilarity matrices between the planning units, and tested for correlations between matrices based on surrogate and genetic data.
+For each species, we computed two dissimilarity matrices between the planning units occupied by the species in terms of their environmental conditions (their coordinates in the environmental space) and geographic position (their coordinates in the geographic space).
+We then computed another dissimilarity matrix between the planning units using their values in the species' neutral genetic space.
+If the species was associated with adaptive loci, we also computed another dissimilarity matrix between the planning units using the species' adaptive genetic space.
+All dissimilarity matrices calculated using Euclidean distances.
+
+We then fit maximum likelihood population-effects (MLPE) models [@r494; implemented in the \texttt{ResistanceGA R} package; @r488].
+These models explicitly accommodate the structure of dissimilarity matrices using random effects.
+For each species, we fit a MLPE model to correlate the species' dissimilarity matrices based on geographic position and neutral genetic variation.
+Additionally, if the species was associated with adaptive loci, we also fit a MLPE model to correlate the species' dissimilarity matrices calculated using environmental conditions and adaptive genetic variation.
+All predictor variables were z-transformed prior to model fitting to improve convergence.
+To test if the surrogate variables explained a substantial amount of genetic variation, we conducted $\chi^2$ tests between each model and its corresponding null model, and applied Bonferroni corrections.
+
+To address the second aim--to determine if the environmental and geographic variables are effective surrogates for adaptive and neutral genetic variation--we generated a collection of single-species prioritisations and computed the proportion of intra-specific genetic variation they secured.
+For each species, we generated `r rapr.params.LST[[MODE]]$surrogacy.analysis$replicates` prioritisations by randomly selecting different combinations of planning units that were occupied by the species.
+We then computed the proportion of adaptive genetic variation and environmental variation secured in each prioritisation.
+We then generated another random set of prioritisations, and computed the proportion of neutral genetic variation and geographic variation secured in each prioritisation.
+
+We fit two full generalised linear models (GLM) with a logit links.
+The first model was fit to the proportion of adaptive genetic variation secured in a prioritisation, and contained a continuous predictor variable measuring the proportion of environmental variation also secured in the prioritisation, a categorical variable indicating the species for which the prioritisation was generated, and an interaction term between these predictor variables.
+The second model was fit to the proportion of neutral genetic variation secured in a prioritisation, and contained a continuous predictor variable assessing the proportion of geographic variation also secured in the prioritisation.
+Similar to the previous model, this model also contained a predictor variable indicating the species for which the prioritisations were generated, and an interaction term.
+We subjected these models to a backwards step-wise term deletion routine to assess term significance.
+To gauge the performance of these surrogates for each species, we fit a suite of models correlating the proportion of adaptive and neutral genetic variation held in each prioritisation and the environmental and geographic variation they also secured (respectively).
+We then computed the Cragg and Uhler's pseudo-R$^2$ value for each model [using the \texttt{pscl} R package; @r492].
+
+To address the third aim--to determine if the use of surrogate-based targets can actually improve prioritisations--we investigated the effectiveness of surrogate-based targets under three scenarios with increasing levels of complexity.
+To assess the benefits of surrogate-based targets under the simplest of circumstances, our first scenario involved generating single-species prioritisations for each species.
+However, most large-scale prioritisations involve a comprehensive set of species.
+By forcing solutions to secure individuals in multiple communities, this may result in prioritisations that secure a greater proportion of intra-specific genetic variation for each species.
+To test for this effect, our second scenario involved generating multi-species prioritisations.
+Finally, most large-scale prioritisations also explicitly consider opportunity cost to ensure that prioritisations are cost-effective.
+Our third scenario involved generating multi-species prioritisations that satisfied targets for minimal opportunity cost.
+
+For each scenario, we generated three types of prioritisations using (1) only amount-based targets, (2) both amount- and surrogate-based targets, and (3) both amount- and genetic-based targets.
+The purely amount-based prioritisations served as our baseline--representing the prioritisations typically generated in conservation planning exercises (eg.
+by \texttt{Marxan}).
+We used `r (rapr.params.LST[[MODE]]$scenario.analysis$amount.target*100)` % amount-based targets in all prioritisations to ensure an adequate proportion of habitat was secured for each species.
+Based on the results from the analyses used to address our second aim, we set surrogate-based targets as `r (rapr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100)` % and genetic-based targets as `r (rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)` % to ensure that most of the genetic variation was secured for each species.
+To accommodate the fact that there were a large number of optimal solutions for the single-species amount-based prioritisations, we generated `r (rapr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates)` replicate prioritisations.
+We computed the proportion of adaptive (if present) and neutral genetic variation secured for each species in the prioritisations.
+
+We investigated the benefits of using surrogate-based targets in prioritisations using generalised linear mixed-effects models fit with logit liknks [GLMMs; using the \texttt{lme4 R} package; @469].
+We fit a full model to the proportion of genetic variation secured in a given prioritisation.
+This model contained categorical predictor variables to indicate the combination of targets used to generate the prioritisations (amount-only, amount and surrogate targets, and amount and genetic targets), the scenario that was used to generate the prioritisation (single-species, multi-species, and multi-species with cost), the type of genetic variation measured (adaptive and neutral), and all interactions between these variables.
+This model accommodated prioritisations generated for the same species using a random intercept term.
+We subject the full model to a backwards step-wise term deletion routine to assess term significance.
+We conducted a post-hoc analysis on the minimal adequate model using Tukey contrasts with a Bonferroni correction [using the \texttt{multcomp R} package; @r470].
+
+\showmatmethods
+\showacknow
+\pnasbreak

### 2016-07-22 code/rmarkdown/article.Rmd
+ fixes to make compatiable with pnas + change to raptr

--- 2016-06-29 code/rmarkdown/article.Rmd
@@
 ---
-title: "Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?"
-author: |
-  | Jeffrey O. Hanson$^1$, Jonathan R. Rhodes$^2$, Cynthia Riginos$^1$, Hugh P. Possingham$^1$, Richard A. Fuller$^1$
-  | $^1$School of Biological Sciences, The University of Queensland, Brisbane, QLD, Australia
-  | $^2$School of Geography, Planning and Environmental Management, The University of Queensland, Brisbane, QLD, Australia
-  | Correspondance should be addressed to jeffrey.hanson@uqconnect.edu.au
-date: "`r format(Sys.time(), '%d %B %Y')`"
-tags: ["conservation", "biodiversity", "AFLP", "reserve-selection"]
-abstract: "Insert abstract here."
-output:
-  rmarkdown::pdf_document:
-    toc: false
-    toc_depth: 5
-    keep_tex: yes
-    fig_caption: yes
-    includes:
-      in_header: preamble-latex.tex
-      after_body: postamble-latex.tex
-fontsize: 11pt
-documentclass: article
+title: "Environmental and geographic variables are effective surrogates for genetic variation in conservation planning"
+author:
+  - name: Jeffrey O. Hanson
+    affiliation: a,1
+  - name: Jonathan R. Rhodes
+    affiliation: b
+  - name: Cynthia Riginos
+    affiliation: a
+  - name: Hugh P. Possingham
+    affiliation: a
+  - name: Richard A. Fuller
+    affiliation: a
+address:
+  - code: a
+    address: School of Biological Sciences, The University of Queensland, Brisbane, QLD, Australia
+  - code: b
+    address: School of Geography, Planning and Environmental Management, The University of Queensland, Brisbane, QLD, Australia
+corresponding_author:
+  - code: 1
+  - text: "To whom correspondence should be addresssed. E-mail: jeffrey.hanson@uqconnect.edu.au"
+author_contributions: |
+  All authors developed the study and drafted the manuscript. JOH performed all analyses.
+lead_author_surname: "Hanson"
+abstract: |
+  Please provide an abstract of no more than 250 words in a single paragraph.
+Abstracts should explain to the general reader the major contributions of the article.
+References in the abstract must be cited in full within the abstract itself and cited in the text.
+significance: |
+  Protected areas are used to preserve biodiversity patterns and processes.
+Since the resources available for conservation are limited, protected areas need to be sited in cost-effective locations.
+However, despite the stated emphasis on biodiversity processes, evolutionary processes are not often used to guide reserve selection.
+This is due to the fact that genetic data is often too costly to obtain and process.
+Here we found that environmental and geographic variables are cost-effective surrogates for genetic data.
+These variable can be used to identify candidate reserve systems that secure the adaptive and neutral evolutionary processes for a greater number of species.
+This research highlights the importance of obtaining a representative sample of species when generating broad-scale prioritisations.
+acknowledgements: |
+  JOH is funded by an Australian Postgraduate Award (APA) scholarship.
+RAF has an Australian Research Council Future Fellowship.
+This work was supported by the Centre of Excellence for Environmental Decisions (CEED) and the Landscape Ecology and Conservation Group (LEC) at The University of Queensland.
+keywords:
+  - conservation
+  - biodiversity
+  - AFLP
+  - genetic diversity
+pnas_type: pnasresearcharticle
 bibliography: references.bib
 csl: reference-style.csl
+output:
+  rticles::pnas_article:
+    includes:
+      in_header: [preamble.tex]
 ---
@@
 connectivity between populations).
-For instance, \texttt{Marxan} [@r22] uses amount-based targets to identify protected area networks that preserve a suitable amount of habitat for each species (eg.
-100 km $^2$ of land occupied by the species needs to be secured in a reserve) and weights to penalise fragmented solutions (ie.
-the \textit{BLM} parameter).
+For example, \texttt{Marxan} [@r22] uses amount-based targets that specify a minimum amount of habitat to preserve for each species to identify cost-effective solutions (eg.
+100 km $^2$ of land occupied by the species needs to be secured in a reserve).
+To promote connecitivty among multiple reserves, \texttt{Marxan} also contains the functionality to penalise fragmented solutions (using the boundary length modifier (\textit{BLM}) parameter).
 However, conservation planning exercises typically assume that all individuals in the same species are equivalent.
@@
 Intra-specific genetic variation is important for long-term species persistence.
-As a consequence, there has been increasing interest in designing prioritisations that preserve and cultivate this variation [@r67; @r3; @r61].
+As a consequence, there has been increasing interest in designing prioritisations that preserve and cultivate this aspect of biodiversity [@r67; @r3; @r61].
 Although the strength of natural selection is a continuous force; broadly speaking, genetic variation can be classified as either adaptive or neutral [reviewed in @r480].
@@
 Typically, such "adaptive" loci are detected due to anomalous patterns among individuals [eg.
-@r456; @r487; @r480; @r490; but see @r489 for alternative methods].
+@r456; @r487; @r480; @r490], though phenotypic-genetic associations such as quantitative trait analyses or genome-wide association measures are more likely to identify functionally important genomic regions for individual species [@rXXXX; @rXXXX].
 By preserving existing patterns of adaptive variation, protected areas can ensure that populations with particularly beneficial adaptations are not lost [@r3].
@@
 Thus protected area networks should aim to preserve both adaptive and neutral patterns of genetic variation for species of conservation interest [@r67; @r3].
-Despite this, genetic data are not often used to inform conservation planning exercises because such data require considerable investment to obtain and analyze [@r61; but see @r5 and @r157 for developments on using genetic data in conservation planning]--especially since most large-scale exercises involve hundreds of species.
+Despite this, genetic data are not often used to inform conservation planning exercises because such data require considerable investment to obtain and analyze for a single species [@r61; but see @r5 and @r157 for developments on using genetic data in conservation planning], and most large-scale planning exercises involve hundreds of species [eg.
+@rXXX].
 Thus prioritisations may fail to secure a sufficient proportion of the intra-specific genetic variation of species, and ultimately fail to deliver long-term conservation outcomes.
 
-Recently, conservation scientists have been investigating the use of surrogate data to generate prioritisations that adequately preserve intra-specific genetic variation--but without actually needing to utilise genetic data to acheive this.
-Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been used to guide reserve selection as surrogates for adaptive genetic variation [eg.
+Recently, conservation scientists have proposed using surrogate data to generate prioritisations that adequately preserve intra-specific genetic variation--without needing to utilise genetic data to acheive this--but have yet to comprehensively assess whether the proposed spatial surrogates adequately represent relevant aspects of intra-specific genetic variation.
+Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been proposed to guide reserve selection as surrogates for adaptive genetic variation [eg.
 environmental conditions; @r247; @r2].
@@
 On the other hand, neutral genetic variation arises from a reduction in gene flow between individuals.
-Surrogates for neutral variation have been based on variables that predict the level of connectivity between planning units.
+Over the last few years, the field of phylogeography has predominantly focused on describing neutral genetic diversity, and so the landscape factors that affect neutral variation are relatively well understood [@rXXXX].
+Potential surrogates for neutral variation have been based on variables that predict the levels of connectivity between planning units.
 For instance, the isolation by distance theory [IBD; @r259] predicts that populations further apart will have less connectivity, exchange less genetic material, and in turn have less genetic variation in common.
@@
 
-Here we investigate the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation in the context of conservation planning.
-First, we aim to verify if environmental and geographic variables correlate with broad-scale patterns in adaptive and neutral variation (respectively).
-Although these variables may correlate with genetic variation, they may not be effective surrogates if they have poor predictive ability or if these correlations are driven by spatial auto-correlation.
-Second, we aim to verify that environmental and geographic variables are effective surrogates for adaptive and neutral variation (respectively).
+Here we investigated the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation in the context of conservation planning.
+To achieve this, we developed a novel framework for incorporating adaptive and neutral genetic data into conservation planning.
+First, we aim to verify if the spatial patterns in environmental and geographic variation correlate with the broad-scale spatial patterns in adaptive and neutral variation (respectively).
+Second, we aim to test whether environmental and geographic variables are effective surrogates for adaptive and neutral variation (respectively) when generating prioritisations.
 We expect that prioritisations generated using surrogate-based targets will secure more genetic variation than prioritisations generated using only amount-based targets with the same number of planning units.
@@
 
-## Methods
-### Study system
+Results {.unnumbered}
+=====================
+
+<!-- summary of methods -->
+We successfully mapped the main spatial patterns of adaptive (if present) and neutral variation for each of the 27 species.
+The outlier detection methods detected putatively adaptive genetic variation in `r as.character(as.english(length(env.surrogacy.LST)))` species.
+Generally, the two outlier detection methods gave fairly consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD).
+Of the species that were associated with adaptive genetic variation, only a few loci were detected as under selection (mean `r round(mean(spp.outlier.loci.props, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.loci.props, na.rm=TRUE),2)` SD).
+After identifying the loci showing strong signals of selection, we used NMDS analyses to identify the main gradients of adaptive (if detected) and neutral variation for each species.
+For all species in which we detected outlier loci, only a couple of continuous dimensions were needed to sufficiently describe their patterns of adaptive genetic variation (all $K = `r unique(na.omit(adaptive.k))`$; Table S2).
+On the other hand, a few additional continuous dimensions were occasionally needed to describe their patterns of neutral genetic variation (minimum $K = `r min(neutral.k)`$; mean $K = `r round(mean(neutral.k),2)`$; maximum $K = `r max(neutral.k)`$; Table S2).
+The ordinations resulting from the NMDS analyses were used to construct an adaptive (if present) and neutral genetic space for each species.
+The spatial distribution of these genetic spaces (Figures S29--S54) generally showed patterns of spatial auto-correlation (eg.
+_Cerastium uniflorum_, Figure SXX; _Dryas octopetala_, Figures SXXX)--a good sign that the environmental and geographic variables may prove effective surrogates.
+
+<!-- correlation analysis -->
+The correlation between dissimilarity matrices based on the environmental and geographic characteristics of the planning units and dissimilarity matrices based on the adaptive and neutral genetic characteristics of individuals in the planning units (Table 1).
+Specifically, environmental variation was found to significantly correlate with adaptive genetic variation for `r as.character(as.english(sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)))` species of the `r as.character(as.english(sum(!is.na(correlation.model.summary.DF$Environmental.corr.P))))` species associated with adaptive variation (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models).
+These results suggest that prioritisations with different environmental characteristics tend to have individuals with different adaptive genetic characteristics.
+Additionally, geographic distances between planning units were found to significantly correlate with neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` species of the total `r sum(!is.na(correlation.model.summary.DF$Geographic.corr.P))` (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models).
+Thus planning units that were located further apart tended to have individuals with increasingly different neutral genetic characteristics.
+After verifying that the geographic and environmental variables generally correlated with genetic variation, our next step was to determine how effective these variables were as surrogates in the context of conservation planning.
+```{r, echo=FALSE, results='asis'}
+curr.file <- tempfile(fileext='.tex')
+x <- correlation.model.summary.DF %>%
+	select(-contains('raw')) %>%
+	mutate(
+		Environmental.corr.P=pvalString(Environmental.corr.P),
+		Geographic.corr.P=pvalString(Geographic.corr.P),
+		Geographic.Marginal.R2=pvalString(Geographic.Marginal.R2),
+		Geographic.Conditional.R2=pvalString(Geographic.Conditional.R2),
+		Environmental.Marginal.R2=pvalString(Environmental.Marginal.R2),
+		Environmental.Conditional.R2=pvalString(Environmental.Conditional.R2),
+		blank=rep('', length(Species)),
+		Species=paste0('\\textit{', gsub('\\ ', ' ', Species), '}')) %>%
+	mutate(Environmental.Chisq=sapply(round(Environmental.Chisq,2), as.character)) %>%
+	mutate(Geographic.Chisq=sapply(round(Geographic.Chisq,2), as.character)) %>%
+	select(Species:Environmental.corr.P, Geographic.Marginal.R2:Geographic.corr.P) %>%
+	latex(file=curr.file, 
+		cgroup=c('', 'Environmental', 'Geographic'), n.cgroup=c(1, 4, 4), col.just=rep('c', 9),
+		digits=2, rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, here=TRUE,
+		colheads=c('\\bfseries{Species}', '$MR^2$', '$CR^2$', '$\\chi^2$', '\\textit{P}', '$MR^2$', '$CR^2$', '$\\chi^2$', '\\textit{P}'),
+		caption='Summary of maximum likelihood population-effects (MLPE) models.
+Briefly, for each species associated with potentially adaptive loci, a model was fit to correlate dissimilarities among planning units where the species in terms of the adaptive genetic characteristics of individuals in them with dissimilarities in terms of the environmental conditions inside the planning units (denoted in the environmental column group).
+Similarly, for each species, a model was fit to correlate dissimilarities among the planning units in terms of the neutral genetic characteristics of indiviudals in them with dissimilarities in terms of the geographic position of the planning units (denoted in the geographic column group).
+Data shows the results from comparing each model to its corresponding null model using $\\mathbf{\\chi^2}$ tests (1 degree of freedom), and the $\\mathbf{R^2}$ ($\\mathbf{MR^2}$) and conditional $\\mathbf{R^2}$ ($\\mathbf{CR^2}$) for each model.'
+	)
+readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
+	gsub(pattern='\\cline{1-', replacement='\\cline{2-6} \\cline{8-', fixed=TRUE) %>% 
+	gsub(pattern='{c}{\\textit{P}}\\tabularnewline', replacement='{c}{\\textit{P}}\\T\\B\\tabularnewline', fixed=TRUE) %>% 
+	gsub(pattern='\\toprule', replacement='', fixed=TRUE) %>%
+	gsub(pattern='\\midrule', replacement='\\midrule', fixed=TRUE) %>%
+	gsub(pattern='\\bottomrule', replacement='\\midrule', fixed=TRUE) %>%
+	cat()
+```
+
+<!-- surrogacy analysis -->
+The environmental and geographic variables were moderately effective surrogates for genetic variation in most species (Figure 1).
+The relationship between the proportion of adaptive genetic variation and environmental variation secured in a given prioritisation was found to differ depending on the species for which the prioritisation was generated (species $\times$ proportion of environmental variation interaction term; ${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$).
+The slopes for each species were all significantly different to zero (minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`; maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3).
+Similarly, the relationship between the proportion of neutral genetic variation and geographic variation secured in a prioritisation was also found to affected by the interaction between species and the proportion of neutral variation secured in the prioritisation (${\chi^2}_{`r geo.surrogacy.AOV$df[[4]]`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`).
+The slope for each species were also significantly different to zero (minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`; maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3).
+After establishing a relationship between the proportion of the surrogate variables and proportion of genetic variation secured in a prioritisation, the next step was to determine if prioritisations benefited from the addition of surrogate-based targets and verify if such benefits were evident across a variety of planning scenarios.
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=6, fig.height=4.5, out.width='3.42in', out.height='2.5in', fig.cap='The relationship between surrogates and genetic variation secured in prioritizations.
+Panel (a) shows the relationship between the proportion of intra-specific adaptive genetic variation and environmental variation secured in a prioritization.
+Similarly, panel (b) shows the relationship between the proportion of intra-specific neutral variaiton and geograpgic variation secured in a prioritization.
+Panels (c) and (d) show the distribution of $R^2$ values for these models.'}
+# make figure
+p1 <- ggplot(aes(x=surrogate.held,y=genetic.held), data=env.surrogacy.DF) +
+	xlab('Environmental variation\nsecured (%)') + 
+	ylab('Adaptive genetic\nvariation secured (%)') +
+	theme_classic() + theme(axis.line.x=element_line(color='black'),
+		axis.line.y=element_line(color='black')) +
+	geom_text(x=0.05, y=0.95, label='(a)') +
+	coord_cartesian(xlim=c(0, 1))
+for (i in seq_along(env.surrogacy.LST))
+	p1 <- p1 + stat_smooth(method='glm', method.args=list(family='binomial'),
+		data=filter(env.surrogacy.DF, Species==unique(Species)[i]),
+		fullrange=TRUE, se=FALSE, colour='black')
+
+p2 <- ggplot(aes(x=surrogate.held,y=genetic.held), data=geo.surrogacy.DF) +
+	xlab('Geographic variation\nsecured (%)') + 
+	ylab('Neutral genetic\nvariation secured (%)') +
+	theme_classic() + theme(axis.line.x=element_line(color='black'),
+		axis.line.y=element_line(color='black', colour='transparent')) +
+	geom_text(x=0.05, y=0.95, label='(b)') +
+	coord_cartesian(xlim=c(0, 1))
+for (i in seq_along(geo.surrogacy.LST))
+	p2 <- p2 + stat_smooth(method='glm', method.args=list(family='binomial'),
+		data=filter(geo.surrogacy.DF, Species==unique(Species)[i]),
+		fullrange=TRUE, se=FALSE, colour='black')
+
+p1.hist <- ggplot(aes(r2CU),
+		data=filter(env.surrogate.spp.DF, !is.na(r2CU))) +
+	geom_histogram(fill='grey10', bins=5) +
+	ylab('\nFrequency') + xlab(expression(Environmental~variables~(R^2))) +
+	theme_classic() + theme(axis.line.x=element_line(color='black'),
+		axis.line.y=element_line(color='black'),
+		plot.background=element_rect(fill='transparent', colour='transparent')) +
+		geom_text(x=0.05, y=6, label='(c)') +
+		coord_cartesian(xlim=c(0, 1))
+
+p2.hist <- ggplot(aes(r2CU),
+		data=filter(geo.surrogate.spp.DF, !is.na(r2CU))) +
+	geom_histogram(fill='grey10', bins=10) +
+	ylab('\nFrequency') + xlab(expression(Geographic~variables~(R^2))) +
+	theme_classic() + theme(axis.line.x=element_line(color='black'),
+		axis.line.y=element_line(color='black'), 
+		plot.background=element_rect(fill='transparent')) +
+		geom_text(x=0.05, y=8, label='(d)') +
+		coord_cartesian(xlim=c(0, 1))
+
+# render figure
+grid.arrange(p1,p2,p1.hist,p2.hist,nrow=2)
+```
+
+<!-- scenario analysis -->
+To determine if the addition of environmental and geographic targets could actually improve the effectiveness of prioritisations, we generated prioritisations according to three scenarios using different targets (Figure 3).
+We then computed the proportion of adaptive and neutral genetic variation secured in each prioritisation (Figure 4).
+We found no evidence to suggest that any of the interaction terms were significant (${\chi^2}_{`r scenario.ANOVA.1.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.1.DF$Pr..Chisq.[[2]])`$).
+This result indicates the trends observed in each of the main effects were consistent across all other levels in the main effects.
+
+First, we found that on average the prioritisations tended to secure more adaptive (mean `r round(mean(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` % secured in all prioritisations $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)`) than neutral genetic variation (mean `r round(mean(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)` % secured in all prioritisations $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)`; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.3.DF$Pr..Chisq.[[2]])`$).
+This result indicates that regardless of the targets used to generate a prioritisation, or the planning context under which the prioritisation was generated, prioritisations tend to secure more adaptive then neutral genetic variation.
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=(5*0.9), fig.height=(6.5*0.8), out.width='3.42in', height='3.9in', fig.cap='Planning unit selection frequencies.
+Squares represent planning units, and their colors indicate their selection frequency.
+Each column of panels shows solutions generated using different targets.
+Each row of panels shows solutions generated under different scenarios.
+Note that the selection frequencies for the single-species prioritizations are based on their frequency among prioritizations generated for each species.
+The selection frequencies for the multi-species prioritizations are each based on a single optimal pioritization for all species and so are binary.'}
+# prepare data for plotting
+grid.sub.FPLY <- grid.sub.PLY
+for (i in seq_along(single.spp.prioritisations[[1]]))
+	grid.sub.FPLY@data[[paste0('Single-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations, function(x){colMeans(x[[i]]@results@selections)}))
+for (i in seq_along(single.spp.prioritisations.with.cost[[1]]))
+	grid.sub.FPLY@data[[paste0('Single-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations.with.cost, function(x){colMeans(x[[i]]@results@selections)}))
+for (i in seq_along(multi.spp.prioritisations))
+	grid.sub.FPLY@data[[paste0('Multi-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations[[i]]@results@selections)
+for (i in seq_along(multi.spp.prioritisations.with.cost))
+	grid.sub.FPLY@data[[paste0('Multi-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations.with.cost[[i]]@results@selections)
+grid.sub.FPLY <- spFortify(grid.sub.FPLY)
+grid.sub.FPLY <- grid.sub.FPLY %>% gather(
+		Combination,
+		`Selection.Frequency`,
+		`Single-species (equal costs).Amount`:`Multi-species (opportunity costs).Genetic`
+	) %>% mutate(
+		Context=sapply(strsplit(as.character(Combination), '.', fixed=TRUE), `[[`, 1),
+		Prioritization=sapply(strsplit(as.character(Combination), '.', fixed=TRUE), `[[`, 2)
+	) %>% mutate(
+		Prioritization=revalue(Prioritization, c('Amount'='Amount targets', 'Surrogate'='Amount & surrogate\ntargets', 'Genetic'='Amount & genetic\ntargets')),
+		Prioritization=factor(as.character(Prioritization), levels=c('Amount targets', 'Amount & surrogate\ntargets', 'Amount & genetic\ntargets')),
+		Context=factor(
+			gsub(' (', '\n(', Context, fixed=TRUE),
+			levels=c("Single-species\n(equal costs)", "Single-species\n(opportunity costs)",
+				"Multi-species\n(equal costs)", "Multi-species\n(opportunity costs)"))
+	)
+
+# make maps
+ggplot() +
+	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
+		fill='grey85', color='grey70') +
+	geom_polygon(data=grid.sub.FPLY, aes_string(x='long', y='lat', 
+		group='group', fill='Selection.Frequency'),
+		alpha=0.8, color=NA) +
+	theme_classic() +
+	theme(axis.ticks=element_blank(), axis.text=element_blank(),
+		plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(),
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white', size=8),
+		legend.position='bottom', legend.key.width=unit(1.8, 'cm'),
+		legend.title=element_text(vjust=-5),
+		legend.margin=unit(c(0.1), 'cm'),
+		axis.title=element_blank()) +
+	coord_cartesian(
+		xlim=buffered.range(grid.sub.FPLY$long, 0),
+		ylim=buffered.range(grid.sub.FPLY$lat, 0)
+	) +
+	facet_grid(Context ~ Prioritization) +
+	scale_fill_gradientn(name='Selection Frequency (%)',
+		colors=brewer.pal(9, 'PuBu'),
+		guide=guide_colorbar(title.position='bottom', title.hjust=0.5,
+			ticks=element_line(color='black'), border=element_line(color='black'))
+	)
+```
+
+Second, the proportion of genetic variation secured in a prioritisation was found to depend on the combination of targets used to generate the prioritisation (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$Pr..Chisq.[[2]])`)$.
+Specifically, the amount-based prioritisations (`r round(mean(scenario.sub.DF,filter(Prioritisation=='Amount targets')$genetic.held*100),2)` % overall genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)`) tended to secure less variation than the prioritisations generated using additional surrogate targets (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` % secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$) or genetic targets (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` % secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)`; `$Z_1 = `r round(scenario.GLHT.2.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[2]])`$).
+No evidence was found to suggest that the prioritisations that included additional surrogate-based targets (`r raptr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100` %) or genetic-based targets (`r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %) secured a different proportion of genetic variation ($Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).
+These results suggest that the addition of environmental and geographic surrogates can improve the proportion of intra-specific genetic variation secured in a prioritisation for most species.
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=(5*0.9), fig.height=(7.1*0.9), out.width='3.42in', out.height='4in', fig.cap='The effectiveness of prioritizations at securing genetic variation.
+Box plots show the proportion of adaptive and neutral genetic variation secured for each species using different targets and under different scenarios.
+Red lines show the target proportion of genetic variation to secure a representative sample.'}
+# prepare data for plotting
+scenario.SDF <- scenario.DF %>% 
+	group_by(Prioritisation, Context, Metric, Species) %>% 
+	summarize(genetic.held=mean(genetic.held, na.rm=TRUE)) %>%
+	data.frame()
+
+# make plot
+suppressWarnings({
+	ggplot(aes(x=Prioritisation,y=genetic.held),
+		data=scenario.SDF) +
+		geom_boxplot(position=position_dodge(0.9)) +
+		scale_fill_manual(name='Prioritisation',
+			values=c('grey80','grey50','grey20')) +
+		geom_abline(slope=0, color='red', linetype='dashed',
+			intercept=raptr.params.LST[[MODE]]$scenario.analysis$genetic.target) +
+		ylab('Genetic variation secured (%)') +
+		xlab('') +
+		scale_y_continuous(breaks=c(0,0.25,0.5,0.75,1)) +
+		theme_classic() +
+		theme(strip.background = element_rect(fill='grey20'),
+			strip.text = element_text(color='white', size=9),
+			panel.border = element_rect(color='black', fill=NA),
+			axis.text.x=element_text(angle=45, hjust=1, vjust=1, size=8),
+			axis.text.y=element_text(size=8)) +
+		facet_grid(Context ~ Metric)
+})
+```
+
+Third, we also found significant differences in the average proportion of genetic variation secured by prioritisations generated under different scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.2.DF$statistic[[2]],2)`$; $P$ `r pvalString(scenario.ANOVA.2.DF$Pr..Chisq.[[2]])`).
+Specifically, both the multi-species (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)`) and multi-species with cost (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)`) were both found to secure more genetic variation than the single species scenario (mean `r round(mean(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held),2)`; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[2]])`$; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$ respectively).
+Neither of the multi-species scenarios were found to secure a different proportion of genetic variation ($Z_1 = `r round(scenario.GLHT.1.DF$statistic[[3]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[3]])`$).
+These results suggest that on average the multi-species prioritisations can secure more genetic variation than single-species prioritisations--regardless of targets used to generate them.
+
+Discussion {.unnumbered}
+========================
+
+<!--  introduction -->
+We aimed to investigate the effectiveness of broad-scale environmental and geographic variables as surrogates for adaptive and neutral genetic variation in conservation planning.
+Our study is the first to address this aim by measuring the proportion of genetic variation secured by prioritisations generated for diverse set of species.
+Overall, we found that the spatial patterns of environmental and geographic variables correlate with those of adaptive and neutral patterns for most species.
+Furthermore, we found that the addition of environmental and geographic targets to the reserve selection process resulted in prioritisations that secured a greater proportion of adaptive and neutral genetic variation for most species.
+
+<!-- surrogate/genetic dissimilarity matrix correlations -->
+Differences in the the environmental conditions and geographic position among the planning units tended to correlate with genetic differences among the individuals in the planning units (Table 1).
+For both surrogates and their corresponding models, the conditional $R^2$ values (environmental vs.
+adaptive genetic variation: mean `r round(mean(correlation.model.summary.DF$Environmental.Conditional.R2,na.rm=TRUE),2)` conditional $R^2$ `r round(sd(correlation.model.summary.DF$Environmental.Conditional.R2,na.rm=TRUE),2)` $\pm$ SD; geographic vs.
+neutral genetic variation: mean `r round(mean(correlation.model.summary.DF$Geographic.Conditional.R2,na.rm=TRUE),2)` conditional $R^2$ `r round(sd(correlation.model.summary.DF$Geographic.Conditional.R2,na.rm=TRUE),2)` $\pm$ SD) were generally slightly higher than the marginal $R^2$ values (environmental vs.
+adaptive genetic variation: mean `r round(mean(correlation.model.summary.DF$Environmental.Marginal.R2,na.rm=TRUE),2)` marginal $R^2$ `r round(sd(correlation.model.summary.DF$Environmental.Marginal.R2,na.rm=TRUE),2)` $\pm$ SD; geographic vs.
+neutral genetic variation: mean `r round(mean(correlation.model.summary.DF$Geographic.Marginal.R2,na.rm=TRUE),2)` marginal $R^2$ `r round(sd(correlation.model.summary.DF$Geographic.Marginal.R2,na.rm=TRUE),2)` $\pm$ SD).
+This result suggests that while variation in the surrogate variables correlates with genetic variation, both forms of genetic variation contain additional spatially auto-correlated variation that is not explained by these surrogates.
+In particular, the differences between the conditional and marginal $R^2$ values for the models examining the effectiveness of geographic variables suggest that this unexplained spatial auto-correlation may be non-linear or driven by a few neighboring planning units with genetically similar individuals.
+This finding is consonant with other studies that have found strong patterns of spatial auto-correlation in patterns of adaptive and neutral genetic variation [@rXXX; @rXXX; @rXXX].
+Although correlations between the spatial patterns of the environmental and geographic variables and genetic variables suggest that they may be useful for genetic variation in conservation planning, these correlations alone are insufficient to demonstrate that they are actually effective in the context of conservation planning.
+To demonstrate their effectiveness would require examining the relationship between the proportion of surrogate variation and genetic variation secured in a given prioritisation.
+
+<!-- surrogate/genetic held correlations -->
+Prioritisations that secured a more representative sample of environmental conditions or geographic position tended to secure a greater proportion of adaptive and neutral genetic variation (Figure 2).
+Generally, the environmental and geographic surrogates variables were equally effective surrogates for adaptive (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figures 2c and S56) and neutral genetic variation (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figure S57).
+But for a few species, the proportion of geographic variation secured in a prioritisation had a diminished capacity to explain the proportion of neutral variation secured in the prioritisation (eg.
+_Gypsophila repens_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gypsophila_repens')$Geographic.r2CU,2)`; _Luzula alpinopilosa_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Luzula_alpinopilosa')$Geographic.r2CU,2)`; _Gentiana nivalis_ $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gentiana_nivalis')$Geographic.r2CU,2)`).
+This bimodal distribution (Figure 2d) suggests that geographic position alone may not be a suitable surrogate for neutral variation in conservation planning for all species.
+This result could be attributed to multiple sources.
+First, geographic distances will be a poor surrogate for genetic lineages in the process of sympatric speciation.
+Second, connectivity between areas can be affected by local topology.
+In particular, mountains can act as barriers to gene flow in alpine landscapes and so nearby populations may be isolated from each other.
+Distances that Incorporated information on local topology (eg.
+least-cost distances) may perform better, but generally require species-specific scaling to adequately describe connectivity.
+Third, the species used in this study utilise different dispersal mechanisms [wind dispersal, animal dispersal, offshoots from the mother plant; @rXXX], and some of these species may have dispersal modes that are non-linearly related to geographic distance (eg.
+wind dispersing plants can have high connectivity between nearby and distant areas).
+Overall, however, these species were the exception, and both simple surrogates performed moderately well.
+
+<!-- suitable targets for surrogates -->
+Generally, the slopes correlating the proportion of geographic variation and neutral variation secured in a prioritisation were higher than those for the proportion of environmental variation and adaptive variation secured in a prioritisation (Figures 2a--2b and S56--S57).
+To secure a representative sample of each species' adaptive genetic variation ($>=$ `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %), the prioritisation would need to sample on average `r mean(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE)` % $\pm$ `r sd(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE)` SD of the environmental conditions found across the planning units occupied by the species.
+On the other hand, to secure a representative sample of the species' neutral genetic variation ($>=$ `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100`), the prioritisation would need to sample % `r mean(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE)` $\pm$ `r sd(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE)` SD of the variation in geographic position among the planning units occupied by the species.
+These results suggest that geographic surrogates may require higher targets than environmental surrogates to generate prioritisations that secure representative sample of genetic variation.
+Additionally, these results suggest using surrogate-based targets alone to generate prioritisations may be insufficient to secure a representative sample of species' genetic variation.
+For all species, prioritisations were generated that secured a high proportion of the surrogate variable (> 90 %) but also secured a minuscule proportion of adaptive or neutral genetic variation (< 20 %; Figures S56--S57).
+Thus we recommend that prioritisations be generated using a combination of amount-based targets--as is conventionally used in reserve selection processes--and also surrogate-based targets.
+After establishing a relationship between the proportion of surrogate variation and genetic variation secured in prioritisation, we than generated prioritisations under various planning scenarios to provide insight into how the surrogates may work in the real-world.
+
+<!-- adaptive genetic variation secured in prioritisations-->
+Almost all prioritisations secured a representative sample of species' adaptive genetic variation--regardless of whether surrogate-based or genetic targets were used or not (Figure 4).
+This result suggests that conservation planners can secure a representative sample of the intra-specific adaptive genetic variation using only amount targets.
+The spatial patterns of adaptive genetic variation tended to cluster into a few distinct groups for most species (eg.
+XXX; XXX; XXX; XXX), and so a prioritisation would only need to site a single planning unit in each cluster to secure a representative sample of the species' adaptive genetic variation.
+As a consequence, the single-species prioritisations generated using amount targets and without consideration to opportunity cost were often able to achieve high representation due to the fact they were essentially random selection (Figure 4a).
+To a lesser extent this is also true for the single-species prioritisations generated with amount targets and consideration to opportunity cost (Figure 4c), since the costs are fairly homogenous among planning units with exception to a few planning units lowland areas.
+
+All the multi-species prioritisations secured a representative sample of most of the species' intra-specific adaptive genetic variation (Figures 4e and 4g).
+Unlike the single-species prioritisations, the multi-species prioiritisations generated with only amount-targets did not secure a representative sample of adaptive variation due to stochastic processes.
+Since these prioritisations were forced to secure at least `r raptr.params.LST[[MODE]]$scenario.analysis$amount.target*100` % of land occupied each species, these prioritisations were forced to secured planning units a range of environmental conditions due to the presence of habitat specialists in the data set (eg.
+_Cerastium uniflorum_ and _Androsace obtusifolia_).
+Overall--contrary to our original hypotheses--it seems that conventional reserve selection methods may secure a representative sample of species' adaptive genetic variation if a comprehensive set of species is used.
+This outcome provides empirical support for one of the fundamental tenants of systematic conservation planning [@rXXX].
+
+<!-- neutral genetic variation secured in prioritisations-->
+Generally, the addition of environmental and geographic targets were needed to obtain prioritisations that secure a representative proportion of neutral genetic variation (Figure 4).
+With exception to the multi-species with cost planning scenario (Figure 4h), prioritisations generated using only amount targets failed to secure a representative proportion of neutral genetic variation ($>=$ `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %) for most species (Figures 4b, 4f, 4h).
+Whereas, in all scenarios, the prioritisations generated using amount and surrogate based targets secured a representative proportion of neutral genetic variation for most species.
+Although the the latter prioritisation did not secure a representative sample of the neutral genetic variation in all species, this result suggests that the use of environmental and geographic targets in conservation planning can result in a more effective prioritisation.
+Furthermore, this improvement was associated with only a moderate increase in cost in both scenarios (scenario without opportunity: `r round((score(multi.spp.prioritisations[[2]])/score(multi.spp.prioritisations[[1]]))*100,2)` %; scenario with opportunity cost: `r  round((score(multi.spp.prioritisations.with.cost[[2]])/score(multi.spp.prioritisations.with.cost[[1]]))*100,2)` %)--emphasising the utility of surrogates in conservation.
+
+<!--  limitations/ weaknesses -->
+We wish to alert our readers to several limitations associated with our analysis.
+First, the size of the planning units we used (23 $\times$ 25 km) is much larger than typically used in conservation planning scenarios (between 1--10 km^2).
+We used this resolution because the genetic data were collected at this scale.
+Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis because this process would have artificially introduced additional spatial auto-correlation into the data set.
+Second, we used geographic distances as surrogates for neutral genetic variation.
+As previously discussed, connectivity in this landscape may be better described using distances that incorporate information on topology.
+However, such distances require specific-specific scaling to adequate describe their connectivity (eg.
+due to different dispersal modes), and so cannot easily be utilised in large-scale planning exercises.
+Third, we used AFLP data to characterise the main gradients of genetic variation.
+Whilst next-generation sequencing provides higher resolution genetic information than AFLP data [reviewed in @r480], our methodology would still have utilised only the main gradients of the genetic variation in order to generate prioritisations in a feasible amount of time.
+Overall, we are confident that the limitations associated with our study are relatively minor.
+
+<!--  conclusion -->
+The overarching aim of our study was to investigate the effectiveness of surrogates in conservation planning.
+We found that broad-scale environmental and geographic variables are generally effective surrogates for adaptive and neutral genetic variation most species.
+Although prioritisations generated using amount-only targets tended to secure most of the adaptive variation for most species, the addition of surrogate targets to the planning process substantially improved the proportion of intra-specific neutral genetic variation.
+Conservation planners are therefore urged to include these surrogates in conservation planning exercises.
+
+Materials and Methods {.unnumbered}
+===================================
+
+Study system {#study-system .unnumbered}
+----------------------------------------
+
 We used species distribution and genomic data collected by the international IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for further explanation of data collection methods].
-This data set contains information for 27 alpine plant species (Figure 1a).
+This data set contains information for 27 alpine plant species (Figure 2a).
 It was collected using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx.
@@
 Project members visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals.
-They genotyped samples using amplified fragment length polymorphisms [AFLP; @r453], and constructed matrices denoting the presence/absence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; see Table 1 in @r478 for the number markers for each species).
+They genotyped samples using amplified fragment length polymorphisms [AFLP; @r453], and constructed matrices denoting the presence/absence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; see Table 1 in Alvarez _et al._ [@r478] for the number markers for each species).
 Thus the data set contains information describing the genomic properties of individuals for each species present in every second grid cell.
+
+```{r, echo=FALSE, fig.align='center', fig.width=6, fig.height=1.5, out.width='3.42in', height='2.5in', fig.cap="Map of the study area.
+Squares denote planning units.
+Panel (a) depicts patterns of species riches.
+Planning units with a brighter color are inhabited by more species.
+Panel (b) shows the distribution of opportunity cost across the study area (computed as total population density).
+Note that a quantile-based color ramp is used in this panel."}
+
+# calculate species richness
+grid.PLY$Species_richness <- grid.PLY@data %>%
+	select(5:(4+n_distinct(spp.samples.DF$species))) %>% as.matrix() %>% rowSums()
+
+panel.DF <- grid.sub.FPLY %>% 
+	summarize(lat=min(lat)+(diff(range(lat))*0.96),
+		long=min(long)+(diff(range(long))*0.04))
+
+# plot species richness
+p1 <- ggplot() +
+	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
+		fill='grey85', color='grey70') +
+	geom_polygon(data=spFortify(grid.PLY), aes(x=long, y=lat, 
+		group=group, fill=Species_richness), color=NA) +
+	scale_fill_gradientn(name='Richness',
+		colors=brewer.pal(9,'PuBu'),
+		guide=guide_colorbar(ticks=element_line(color='black'), border=element_line(color='black'))
+	) +
+	theme_classic() +
+	theme(
+		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"), axis.title=element_blank(),
+		panel.border=element_rect(color='black', fill=NA, size=1)
+	) +
+	coord_cartesian(
+		xlim=buffered.range(grid.FPLY$long, 0),
+		ylim=buffered.range(grid.FPLY$lat, 0)
+	) + geom_text(aes_string(x='long',y='lat'), data=panel.DF, label='(a)')
+
+qn01 <- scales::rescale(quantile(grid.PLY@data$pop.density, seq(0,1,length.out=9), na.rm=TRUE))
+p2 <- ggplot() +
+	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
+		fill='grey85', color='grey70') +
+	geom_polygon(data=spFortify(grid.PLY), aes(x=long, y=lat, 
+		group=group, fill=pop.density), color=NA) +
+	scale_fill_gradientn(name='Cost',
+		colors=brewer.pal(9, 'PuBu'),
+		values=qn01,
+		guide=guide_colorbar(ticks=element_line(color='black'), border=element_line(color='black'))
+	) +
+	theme_classic() +
+	theme(
+		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"), axis.title=element_blank(),
+		panel.border=element_rect(color='black', fill=NA, size=1)
+	) +
+	coord_cartesian(
+		xlim=buffered.range(grid.FPLY$long, 0),
+		ylim=buffered.range(grid.FPLY$lat, 0)
+	)  + geom_text(aes_string(x='long',y='lat'), data=panel.DF, label='(b)')
+
+grid.arrange(p1,p2,nrow=1)
+```
 
@@
 
-### Landscape data
+Landscape data {#landscape-data .unnumbered}
+--------------------------------------------
+
 We examined the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation (respectively).
 We utilised the sampling grid used by the IntraBioDiv project as planning units to develop prioritisations.
-We used only planning units that contained at least a single individual ($n = `r nrow(grid.DF)`$) to reduce computational burden.
-To examine the potential effects of variation in cost (see below), we calculated the total human population density inside each planning unit [obtained at 1 km$^2$ resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472] and used this to represent acquisition cost (Figure 1b).
+We omitted cells that were surveyed and not found to contain individuals belonging to any of the study species to reduce computational burden (`r nrow(grid.DF)` grid cells used in analysis).
+To examine the potential effects of variation in cost (see below), we calculated the total human population density inside each planning unit [obtained at 1 km$^2$ resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472] and used this to represent acquisition cost (Figure 2b).
 
@@
 
-### Adaptive and neutral genetic data
+Adaptive and neutral genetic data {#adaptive-and-neutral-genetic-data .unnumbered}
+----------------------------------------------------------------------------------
+
 To investigate the effectiveness of our supposed surrogates for adaptive and neutral genetic variation, we first needed to identify which of the sampled loci--if any--were adaptive.
@@
 Following their methodology, we first grouped conspecifics into genetic lineages to avoid falsely classifying loci as adaptive.
-To complete this first step, for each species, we fit the admixture and correlated alleles model implemented in \texttt{Structure} [version 2.3.4; r@461; r@462] using the number of populations for each species previously determined by [@r478; `structure.params.LST[[MODE]]$numruns` replicates per K; `r structure.params.LST[[MODE]]$admburnin` admixture burnin iterations; `r structure.params.LST[[MODE]]$burnin` burnin iterations; `r structure.params.LST[[MODE]]$numreps` iterations used for inference].
+To complete this first step, for each species, we fit the admixture and correlated alleles model implemented in \texttt{Structure} [version 2.3.4; @r461; @r462; `r structure.params.LST[[MODE]]$numruns` replicates per K; `r structure.params.LST[[MODE]]$admburnin` admixture burnin iterations; `r structure.params.LST[[MODE]]$burnin` burnin iterations; `r structure.params.LST[[MODE]]$numreps` iterations used for inference] using the number of populations for each species previously determined by Alvarez _et al._ [-@r478].
 We used \texttt{ClumPP} [version 1.1.2; @r459] to combine replicate runs to group individuals into populations (greedy algorithm using the $G^{\prime}$ static; `r clumpp.params.LST[[MODE]]$repeats` iterations).
-We then ran $\texttt{Bayescan}$ [@r456] for each species using these populations [1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations].
+We then ran $\texttt{Bayescan}$ [@r456] for each species using these clusters (1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations).
 Additionally, we omitted individuals from the \texttt{Bayescan} analysis if their population membership was uncertain (maximum membership probability $< `r structure.params.LST[[MODE]]$probability.threshold`$).
@@
 
-## Prioritisation method
+Prioritisation method {#prioritisation-method .unnumbered}
+----------------------------------------------------------
+
 We used the unreliable representative and adequate prioritisation (URAP) formulation of the reserve selection problem [@r465].
@@
 Prioritisations associated with negative values--because they secured only a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis.
-We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality using the $\texttt{rapr}$ R package [@r465] and $\texttt{Gurobi}$ [@r463].
+We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality using the \texttt{raptr R} package [@r465] and $\texttt{Gurobi}$ [@r463].
 
-## Statistical analyses
+Statistical analyses {#statistical-analyses .unnumbered}
+--------------------------------------------------------
+
 To address the first aim--to determine if the environmental and geographic variables correlate with spatial patterns of adaptive and neutral genetic variation--we computed dissimilarity matrices between the planning units, and tested for correlations between matrices based on surrogate and genetic data.
@@
 To address the second aim--to determine if the environmental and geographic variables are effective surrogates for adaptive and neutral genetic variation--we generated a collection of single-species prioritisations and computed the proportion of intra-specific genetic variation they secured.
-For each species, we generated `r rapr.params.LST[[MODE]]$surrogacy.analysis$replicates` prioritisations by randomly selecting different combinations of planning units that were occupied by the species.
+For each species, we generated `r raptr.params.LST[[MODE]]$surrogacy.analysis$replicates` prioritisations by randomly selecting different combinations of planning units that were occupied by the species.
 We then computed the proportion of adaptive genetic variation and environmental variation secured in each prioritisation.
@@
 by \texttt{Marxan}).
-We used `r (rapr.params.LST[[MODE]]$scenario.analysis$amount.target*100)` % amount-based targets in all prioritisations to ensure an adequate proportion of habitat was secured for each species.
-Based on the results from the analyses used to address our second aim, we set surrogate-based targets as `r (rapr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100)` % and genetic-based targets as `r (rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)` % to ensure that most of the genetic variation was secured for each species.
-To accommodate the fact that there were a large number of optimal solutions for the single-species amount-based prioritisations, we generated `r (rapr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates)` replicate prioritisations.
+We used `r (raptr.params.LST[[MODE]]$scenario.analysis$amount.target*100)` % amount-based targets in all prioritisations to ensure an adequate proportion of habitat was secured for each species.
+Based on the results from the analyses used to address our second aim, we set surrogate-based targets as `r (raptr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100)` % and genetic-based targets as `r (raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)` % to ensure that most of the genetic variation was secured for each species.
+To accommodate the fact that there were a large number of optimal solutions for the single-species amount-based prioritisations, we generated `r (raptr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates)` replicate prioritisations.
 We computed the proportion of adaptive (if present) and neutral genetic variation secured for each species in the prioritisations.
 
-We investigated the benefits of using surrogate-based targets in prioritisations using generalised linear mixed-effects models fit with logit liknks [GLMMs; using the \texttt{lme4 R} package; @469].
+We investigated the benefits of using surrogate-based targets in prioritisations using generalised linear mixed-effects models fit with logit links [GLMMs; using the \texttt{lme4 R} package; @r469].
 We fit a full model to the proportion of genetic variation secured in a given prioritisation.
@@
 
-## Results
-<!-- summary of methods -->
-We successfully mapped the main spatial patterns of adaptive (if present) and neutral variation for each of the 27 species.
-The outlier detection methods detected adaptive genetic variation in `r as.character(as.english(length(env.surrogacy.LST)))` species.
-Generally, the two outlier detection methods gave fairly consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD).
-Of the species that were associated with adaptive genetic variation, only a few loci were detected as under selection (mean `r round(mean(spp.outlier.loci.props, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.loci.props, na.rm=TRUE),2)` SD).
-After identifying the loci under selection, we used NMDS analyses to identify the main gradients of adaptive (if detected) and neutral variation for each species.
-For all species in which we detected adaptive loci, only a couple of continuous dimensions were needed to sufficiently describe their patterns of adaptive genetic variation (all $K = `r unique(adaptive.k)`$; Table S2).
-On the other hand, a few additional continuous dimensions were occasionally needed to describe their patterns of neutral genetic variation (minimum $K = `r min(neutral.k)`$; mean $K = `r round(mean(neutral.k),2)`$; maximum $K = `r max(neutral.k)`$; Table S2).
-The ordinations resulting from the NMDS analyses were used to construct an adaptive (if present) and neutral genetic space for each species.
-The spatial distribution of these genetic spaces (Figures S29--S54) generally showed patterns of spatial auto-correlation (eg.
-_Cerastium uniflorum_, Figure SXX; _Dryas octopetala_, Figures SXXX)--a good sign that the environmental and geographic variables may prove effective surrogates.
-
-<!-- correlation analysis -->
-The dissimilarity matrices based on the environmental and geographic characteristics of the planning units correlated with the dissimilarity matrices based on the genetic characteristics of individuals in the planning units (Table 1).
-Specifically, environmental variation was found to significantly correlate with adaptive genetic variation for `r as.character(as.english(sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)))` species (`r round((sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)/length(na.omit(correlation.model.summary.DF$Environmental.corr.P)))*100,2)` % associated with loci under selection; _P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models).
-Additionally, geographic distances between planning units were found to significantly correlate with neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` species (`r round((sum(correlation.model.summary.DF$Geographic.corr.P<0.05)/length(na.omit(correlation.model.summary.DF$Geographic.corr.P)))*100,2)` % of all the species investigated; _P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models).
-These results suggest that prioritisations with different environmental characteristics tend to have individuals with different adaptive genetic characteristics.
-Similarly, planning units that are further apart tend to have individuals with different neutral genetic characteristics.
-After verifying that the geographic and environmental variables generally correlated with genetic variation, our next step was to determine how effective these variables were as surrogates in the context of conservation planning.
-
-<!-- surrogacy analysis -->
-The environmental and geographic variables were moderately effective surrogates for genetic variation in most species (Figure 2).
-The relationship between the proportion of adaptive genetic variation secured in a prioritisation and the proportion of environment variation it also secured was found to depend on the species for which the prioritisation was generated (${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$).
-The slope for each species were all significantly different to zero (minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`; maximum $P `r  pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3).
-Similarly, the proportion of neutral genetic variation in a prioritisation was affected by the interaction between species and the proportion of neutral variation secured in the prioritisation (${\chi^2}_{`r geo.surrogacy.AOV$df[[4]]`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`).
-The slope for each species were also significantly different to zero (minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`; maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3).
-After establishing a relationship between the proportion of the surrogate variables and proportion of genetic variation secured in a prioritisation, the next step was to determine if prioritisations benefited from the addition of surrogate-based targets and verify if such benefits were evident across a variety of planning scenarios.
-
-<!-- scenario analysis -->
-To determine if the addition of environmental and geographic targets could actually improve the effectiveness of prioritisations, we generated prioritisations according to three scenarios using different targets (Figure 3).
-We then computed the proportion of adaptive and neutral genetic variation secured in each prioritisation (Figure 4).
-We found no evidence to suggest that any of the interaction terms were significant (${\chi^2}_{`r scenario.ANOVA.1.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.1.DF$Pr..Chisq.[[2]])`$).
-We found that on average the prioritisations tended to secure more adaptive than neutral genetic variation (`r round(plogis(scenario.GLHT.3.DF$statistic[[1]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.3.DF$std.error[[1]])*100,2)` SE; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.3.DF$Pr..Chisq.[[2]])`$).
-
-The proportion of genetic variation secured in a prioritisation was found to depend on the combination of targets used to generate the prioritisation (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$Pr..Chisq.[[2]])`)$.
-Specifically, the amount-based prioritisations tended to secure less variation than the prioritisations generated using additional surrogate (`r round(plogis(scenario.GLHT.2.DF$estimate[[1]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.2.DF$std.error[[1]])*100,2)` SE ; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$) or genetic variables (`r round(plogis(scenario.GLHT.2.DF$estimate[[2]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.2.DF$std.error[[2]])*100,2)` SE; `$Z_1 = `r round(scenario.GLHT.2.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[2]])`$).
-No evidence was found to suggest that the prioritisations that included additional surrogate- or genetic-based targets secured a different proportion of genetic variation (`r round(plogis(scenario.GLHT.2.DF$statistic[[3]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.2.DF$std.error[[3]])*100,2)` SE; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).
-These results suggest that surrogate-based targets can substantially improve the proportion of intra-specific genetic variation secured in a prioritisation for most species.
-
-We also found significant differences in the average proportion of genetic variation secured by prioritisations in each of the scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.2.DF$statistic[[2]],2)`$; $P$ `r pvalString(scenario.ANOVA.2.DF$Pr..Chisq.[[2]])`).
-Specifically, both the multi-species (`r round(plogis(scenario.GLHT.1.DF$estimate[[1]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.1.DF$std.error[[1]])*100,2)` SE; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$) and multi-species with cost (`r round(plogis(scenario.GLHT.1.DF$estimate[[2]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.1.DF$std.error[[2]])*100,2)` SE; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[2]])`$) scenarios were found to secure more genetic variation than the single species scenario.
-Neither the multi-species scenarios were found to secure a different proportion of genetic variation (`r round(plogis(scenario.GLHT.1.DF$estimate[[3]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.1.DF$std.error[[3]])*100,2)` SE; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[3]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[3]])`$).
-These results suggest that on average the multi-species prioritisations can secure more genetic variation than single-species prioritisations--regardless of targets used to generate them.
-
-## Discussion
-<!--  introduction -->
-We aimed to investigate the effectiveness of broad-scale environmental and geographic variables as surrogates for adaptive and neutral genetic variation in conservation planning.
-Our study is the first to address this aim by measuring the proportion of genetic variation secured by prioritisations  generated for diverse set of species.
-Overall, we found that the spatial patterns of environmental and geographic variables correlate with those of adaptive and neutral patterns for most species.
-Furthermore, we found that the addition of environmental and geographic targets to the reserve selection process resulted in prioritisations that secured a greater proportion of adaptive and neutral genetic variation for most species.
-
-<!-- surrogate/genetic dissimilarity matrix correlations -->
-Differences in the the environmental conditions and geographic position among the planning units tended to correlate with genetic differences among the individuals in the planning units (Table 1).
-For both surrogates and their corresponding models, the conditional $R^2$ values (environmental vs.
-adaptive genetic variation: mean `r round(mean(correlation.model.summary.DF$Environmental.Conditional.R2,na.rm=TRUE),2)` conditional $R^2$ `r round(sd(correlation.model.summary.DF$Environmental.Conditional.R2,na.rm=TRUE),2)` $\pm$ SD; geographic vs.
-neutral genetic variation: mean `r round(mean(correlation.model.summary.DF$Geographic.Conditional.R2,na.rm=TRUE),2)` conditional $R^2$ `r round(sd(correlation.model.summary.DF$Geographic.Conditional.R2,na.rm=TRUE),2)` $\pm$ SD) were generally slightly higher than the marginal $R^2$ values (environmental vs.
-adaptive genetic variation: mean `r round(mean(correlation.model.summary.DF$Environmental.Marginal.R2,na.rm=TRUE),2)` marginal $R^2$ `r round(sd(correlation.model.summary.DF$Environmental.Marginal.R2,na.rm=TRUE),2)` $\pm$ SD; geographic vs.
-neutral genetic variation: mean `r round(mean(correlation.model.summary.DF$Geographic.Marginal.R2,na.rm=TRUE),2)` marginal $R^2$ `r round(sd(correlation.model.summary.DF$Geographic.Marginal.R2,na.rm=TRUE),2)` $\pm$ SD).
-This result suggests that while variation in the surrogate variables correlates with genetic variation, both forms of genetic variation contain additional spatially auto-correlated variation that is not explained by these surrogates.
-In particular, the differences between the conditional and marginal $R^2$ values for the models examining the effectiveness of geographic variables suggest that this unexplained spatial auto-correlation may be non-linear or driven by a few neighboring planning units with genetically similar individuals.
-This finding is consonant with other studies that have found strong patterns of spatial auto-correlation in patterns of adaptive and neutral genetic variation [@rXXX; @rXXX; @rXXX].
-After verifying that the spatial pattern of variation in the surrogates generally correlated with those of intra-specific adaptive and neutral genetic variation, our next step was to determine if this pattern of correlation was strong enough to operationalise for conservation planning.
-
-<!-- surrogate/genetic held correlations -->
-Prioritisations that secured a more representative sample of environmental conditions or geographic position tended to secure a greater proportion of adaptive and neutral genetic variation (Figure 2).
-Generally, the environmental and geographic surrogates variables were equally effective surrogates for adaptive (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figures 2c and S56) and neutral genetic variation (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2$ \pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figure S57).
-But for a few species, the proportion of geographic variation secured in a prioritisation had a diminished capacity to explain the proportion of neutral variation secured in the prioritisation (eg.
-_Gypsophila repens_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gypsophila_repens')$Geographic.r2CU,2)`; _Luzula alpinopilosa_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Luzula_alpinopilosa')$Geographic.r2CU,2)`; _Gentiana nivalis_ $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gentiana_nivalis')$Geographic.r2CU,2)`).
-This bimodal distribution (Figure 2d) suggests that geographic position alone may not be a suitable surrogate for neutral variation in conservation planning for all species.
-This result could be attributed to multiple sources.
-First, geographic distances will be a poor surrogate for genetic lineages in the process of sympatric speciation.
-Second, connectivity between areas can be affected by local topology.
-In particular, mountains can act as barriers to gene flow in alpine landscapes and cause relatively nearby populations to diverge genetically.
-Third, the species used in this dataset utilise different dispersal mechanisms [wind dispersal, animal dispersal, offshoots from the mother plant; @rXXX], and some of these species may have dispersal modes that are non-linearly related to geographic distance (eg.
-wind dispersing plants can have high connectivity between nearby and distant areas).
-Overall, however, these species were the exception, and both surrogates performed moderately well.
-
-<!-- suitable targets for surrogates -->
-Generally, the slopes correlating the proportion of geographic variation and neutral variation secured in a prioritisation were higher than those for the proportion of environmental variation and adaptive variation secured in a prioritisation (Figures 2a--2b and S56--S57).
-To secure a representative sample of each species' adaptive genetic variation ($>=$ `r rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %), the prioritisation would need to sample on average `r mean(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE)` % $\pm$ `r sd(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE)` SD of the environmental conditions found across the planning units occupied by the species.
-On the other hand, to secure a representative sample of the species' neutral genetic variation ($>=$ `r rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100`), the prioritisation would need to sample % `r mean(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE)` $\pm$ `r sd(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE)` SD of the variation in geographic position among the planning units occupied by the species.
-These results suggest that geographic surrogates may require higher targets than environmental surrogates to generate prioritisations that secure representative sample of genetic variation.
-Additionally, these results suggest using surrogate-based targets alone to generate prioritisations may be insufficient to secure a representative sample of species' genetic variation.
-For all species, prioritisations were generated that secured a high proportion of the surrogate variable (> 90 %) but also secured a minuscule proportion of adaptive or neutral genetic variation (< 20 %; Figures S56--S57).
-Thus we recommend that prioritisations be generated using a combination of amount-based targets--as is conventionally used in reserve selection processes--and also surrogate-based targets.
-After establishing a relationship between the proportion of surrogate variation and genetic variation secured in prioritisation, we than generated prioritisations under various planning scenarios to provide insight into how the surrogates may work in the real-world.
-
-<!-- adaptive genetic variation secured in prioritisations-->
-Almost all prioritisations secured a representative sample of species' adaptive genetic variation--regardless of whether surrogate-based or genetic targets were used or not (Figure 4).
-This result suggests that conservation planners can secure a representative sample of the intra-specific adaptive genetic variation using only amount targets.
-The spatial patterns of adaptive genetic variation tended to cluster into a few distinct groups for most species (eg.
-XXX; XXX; XXX; XXX), and so a prioritisation would only need to site a single planning unit in each cluster to secure a representative sample of the species' adaptive genetic variation.
-As a consequence, the single-species prioritisations generated using amount targets and without consideration to opportunity cost were often able to achieve high representation due to the fact they were essentially random selection (Figure 4a).
-To a lesser extent this is also true for the single-species prioritisations generated with amount targets and consideration to opportunity cost (Figure 4c), since the costs are fairly homogenous among planning units with exception to a few planning units lowland areas.
-
-All the multi-species prioritisations secured a representative sample of most of the species' intra-specific adaptive genetic variation (Figures 4e and 4g).
-Unlike the single-species prioritisations, the multi-species prioiritisations generated with only amount-targets did not secure a representative sample of adaptive variation due to stochastic processes.
-Since these prioritisations were forced to secure at least `r rapr.params.LST[[MODE]]$scenario.analysis$amount.target*100` % of land occupied each species, these prioritisations were forced to secured planning units a range of environmental conditions due to the presence of habitat specialists in the data set (eg.
-_Cerastium uniflorum_ and _Androsace obtusifolia_).
-Overall--contrary to our original hypotheses--it seems that conventional reserve selection methods may secure a representative sample of species' adaptive genetic variation if a comprehensive set of species is used.
-This outcome provides empirical support for one of the fundamental tenants of systematic conservation planning [@rXXX].
-
-<!-- neutral genetic variation secured in prioritisations-->
-Generally, the addition of environmental and geographic targets were needed to obtain prioritisations that secure a representative proportion of neutral genetic variation (Figure 4).
-With exception to the multi-species with cost planning scenario (Figure 4h), prioritisations generated using only amount targets failed to secure a representative proportion of neutral genetic variation ($>=$ `r rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %) for most species (Figures 4b, 4f, 4h).
-Whereas, in all scenarios, the prioritisations generated using amount and surrogate based targets secured a representative proportion of neutral genetic variation for most species.
-Although the the latter prioritisation did not secure a representative sample of the neutral genetic variation in all species, this result suggests that the use of environmental and geographic targets in conservation planning can result in a more effective prioritisation.
-Furthermore, this improvement was associated with only a moderate increase in cost in both scenarios (scenario without opportunity: `r round((score(multi.spp.prioritisations[[2]])/score(multi.spp.prioritisations[[1]]))*100,2)` %; scenario with opportunity cost: `r  round((score(multi.spp.prioritisations.with.cost[[2]])/score(multi.spp.prioritisations.with.cost[[1]]))*100,2)` %)--emphasising the utility of surrogates in conservation.
-
-<!--  limitations/ weaknesses -->
-We wish to alert our readers to several limitations associated with our analysis.
-First, the size of the planning units we used (23 $\times$ 25 km) is much larger than typically used in conservation planning scenarios (between 1--10 km^2).
-We used this resolution because the genetic data were collected at this scale.
-Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis because this process would have artificially introduced additional spatial auto-correlation into the data set.
-Second, we used geographic distances as surrogates for connectivity between planning units.
-But as previously, using more geographic distances were found to be adequate and more complex measures of connectivity cannot generally be used in more large-scale planning scenarios.
-Third, we used AFLP data to characterise the main gradients of genetic variation.
-Whilst next-generation sequencing provides higher resolution genetic information than AFLP data [reviewed in @r480], our methodology would still have utilised only the main gradients of the genetic variation in order to generate prioritisations in a feasible amount of time.
-Overall, we are confident that the limitations associated with our study are relatively minor.
-
-<!--  conclusion -->
-The overarching aim of our study was to investigate the effectiveness of surrogates in conservation planning.
-We found that environmental and geographic are generally effective surrogates for most species.
-Although prioritisations generated using amount-only targets tended to secure most of the adaptive variation for most species, the addition of surrogate targets to the planning process substantially improved the proportion of intra-specific neutral genetic variation.
-Conservation planners are therefore urged to include these surrogates in conservation planning exercises.
-
-## Acknowledgements
-JOH is funded by an Australian Postgraduate Award (APA) scholarship.
-RAF has an Australian Research Council Future Fellowship.
-This work was supported by the Centre of Excellence for Environmental Decisions (CEED) and the Landscape Ecology and Conservation Group (LEC) at The University of Queensland.
-
-## References
+\showmatmethods
+\showacknow
+\pnasbreak

### 2016-07-22 code/rmarkdown/si-preamble.tex
+ fixes to make compatiable with pnas + change to raptr

--- 2016-05-07 code/rmarkdown/figures.tex
@@
-\documentclass[11pt,]{article}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
-  \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-  \newcommand{\euro}{€}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
-\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\usepackage[margin=1in]{geometry}
-\usepackage{hyperref}
-\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
-\hypersetup{unicode=true,
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-%%% Change title format to be more compact
-\usepackage{titling}
-
-% Create subtitle command for use in maketitle
-\newcommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
-
-\setlength{\droptitle}{-2em}
-  \title{}
-  \pretitle{\vspace{\droptitle}}
-  \posttitle{}
-  \author{}
-  \preauthor{}\postauthor{}
-  \date{}
-  \predate{}\postdate{}
-
 % load packages
@@
 \usepackage[doublespacing]{setspace}
-
-% format captions
-\usepackage[labelfont={small,bf}, labelsep=space, font={small}]{caption}
-
-% format headings
-\titleformat{\subsection}
-{\normalfont\Large\bfseries}{\thesubsection}{1em}{}
-\titleformat{\subsubsection}
-{\normalfont\large\bfseries}{\thesubsubsection}{1em}{}
-\titleformat{\paragraph}
-{\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
-\titleformat{\subparagraph}
-{\normalfont\normalsize\itshape}{\thesubparagraph}{1em}{}
-
-\titlespacing\paragraph{0pt}{12pt plus 4pt minus 2pt}{+0pt plus 2pt minus 2pt}
-\titlespacing{\subparagraph}{0pt}{6pt plus 4pt minus 2pt}{+0pt plus 2pt minus 2pt}
-
-% format toc
-\renewcommand\cftsubsecfont{\normalfont\normalsize\bfseries}
-% \renewcommand\cftsubsubsecfont{\normalfont\normalsize\bfseries}
-% \renewcommand\cftparafont{\normalfont\normalsize\bfseries}
-% \renewcommand\cftsubparafont{\normalfont\normalsize\itshape}
 
@@
 
-% define struts for tables
-\newcommand\T{\rule{0pt}{2.6ex}} % top strut
-\newcommand\B{\rule[-1.2ex]{0pt}{0pt}} % bottom strut
-
-% define command to put new lines in table cells
-%\newcommand{\specialcell}[2][c]{%
-%  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
 % insert S before figure and table caption
 \pagenumbering{gobble}
-
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-\begin{document}
-
-\subsection{Figures}\label{figures}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures_files/figure-latex/unnamed-chunk-3-1.pdf}
-\caption{Map of the study area. Squares denote planning units. Panel (a)
-depicts patterns of species riches. Planning units with a brighter color
-are inhabited by more species. Panel (b) shows the distribution of
-opportunity cost across the study area (computed as total population
-density). Note that a quantile-based color ramp is used in this panel.}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures_files/figure-latex/unnamed-chunk-4-1.pdf}
-\caption{The relationship between surrogates and genetic variation
-secured in prioritizations. Panel (a) shows the relationship between the
-proportion of intra-specific adaptive genetic variation and
-environmental variation secured in a prioritization. Similarly, panel
-(b) shows the relationship between the proportion of intra-specific
-neutral variaiton and geograpgic variation secured in a prioritization.
-Panels (c) and (d) show the distribution of \(R^2\) values for these
-models.}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures_files/figure-latex/unnamed-chunk-5-1.pdf}
-\caption{Planning unit selection frequencies. The top row of panels (a,
-b, c) show single-species prioritizations generated assuming equal
-costs. The middle row of panels (d, e, f) show multi-species
-prioritizations generated assuming equal costs. The bottom row of panels
-(g, h, i) show multi-species generated using opportunity costs. The left
-column of panels (a, d, g) show prioritizations generated using only
-amount-based targets. The middle column of panels (b, e, h) show
-prioritizations generated using amount- and surrogate-based targets. The
-right column of panels (c, f, i) show prioritizations generated using
-amount- and genetic-based targets.}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures_files/figure-latex/unnamed-chunk-6-1.pdf}
-\caption{The effectiveness of prioritizations at securing genetic
-variation. The top row of panels (a--b) show the performance of
-single-species prioritizations generated assuming equal costs among
-planning costs. The middle row of panels (c--d) show the performance of
-multi-species prioritizations generated assuming equal costs among
-planning units. The bottom row of panels (e--f) show the performance of
-multi-species prioritizations generated using opportunity costs. The
-left column of panels (a, c, d) and the right column of panels (b, d, f)
-show the proportion of adaptive and neutral genetic variation secured
-(respectively). Box plots show the effectiveness of prioritizations
-generated using amount-based targets, amount-based and surrogate-based
-targets, and amount-based and genetic-based targets for each species.
-Letters denote significant differences (\(P < 0.05\)).}
-\end{figure}
-
-\hypertarget{refs}{}
-
-\end{document}
+\renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thetable}{S\arabic{table}}

### 2016-07-22 code/rmarkdown/supporting_information.Rmd
+ fix bug in si figures
+ fixes to make compatiable with pnas + change to raptr

@@
     includes:
-      in_header: [preamble-latex.tex, preamble-latex3.tex]
+      in_header: [preamble.tex, si-preamble.tex]
 ---
-## Supporting Information
+# Supporting Information
 All data and code is stored in an online repository (\url{www.github.com/paleo13/genetic.surrogates}) to replicate analyses in this study. See the \textit{README} file for instructions.
@@
 is.speciesPlottable <- function(i) {
-	if (i <= general.params.LST[[MODE]]$n.spp)
+	if (i > general.params.LST[[MODE]]$n.spp)
 		return(FALSE)

### 2016-07-22 code/R/analysis/08-single-species-prioritisations-no-cost.R

@@ -2,7 +2,7 @@
 session::restore.session('data/intermediate/06-format-data-for-prioritisations.rda')
 
 ## load parameters
-rapr.params.LST <- parseTOML('code/parameters/rapr.toml')
+raptr.params.LST <- parseTOML('code/parameters/raptr.toml')
 gurobi.params.LST <- parseTOML('code/parameters/gurobi.toml')
 
 #### single species analysis
@@ -10,8 +10,8 @@ gurobi.params.LST <- parseTOML('code/parameters/gurobi.toml')
 ## amount-based prioritisations
 # prepare cluster object
 clust <- makeCluster(general.params.LST[[MODE]]$threads, type='PSOCK', outfile="")
-clusterEvalQ(clust, {library(rapr)})
-clusterExport(clust, c('spp.samples.sub.DF','grid.sub.DF','rapr.params.LST',
+clusterEvalQ(clust, {library(raptr)})
+clusterExport(clust, c('spp.samples.sub.DF','grid.sub.DF','raptr.params.LST',
 	'MODE', 'species.prioritisation', 'ru'))
 registerDoParallel(clust)
 # main analysis
@@ -20,17 +20,17 @@ single.spp.amount.prioritisations <- llply(
 	function(x) {
 		# identify which planning units are occupied by the species
 		curr.pu <- which(grid.sub.DF[[unique(spp.samples.sub.DF$species)[[x]]]]==1)
-		n.pu <- ceiling(length(curr.pu) * rapr.params.LST[[MODE]]$scenario.analysis$amount.target)
+		n.pu <- ceiling(length(curr.pu) * raptr.params.LST[[MODE]]$scenario.analysis$amount.target)
 		zeros <- rep(0, length=nrow(grid.sub.DF))
 		curr.sol.MTX<-t(replicate(
-			n=rapr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates,
+			n=raptr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates,
 			expr={replace(zeros, sample(x=curr.pu, size=n.pu), rep(1, n.pu))}
 		))
 		# generate portfolio of random selections
 		return(
 			species.prioritisation(
 				x=spp.subset(ru, x),
-				amount.targets=rapr.params.LST[[MODE]]$scenario.analysis$amount.target,
+				amount.targets=raptr.params.LST[[MODE]]$scenario.analysis$amount.target,
[…]

### 2016-07-22 code/R/analysis/10-multi-species-prioritisations-no-cost.R

@@ -2,21 +2,21 @@
 session::restore.session('data/intermediate/06-format-data-for-prioritisations.rda')
 
 ## load parameters
-rapr.params.LST <- parseTOML('code/parameters/rapr.toml')
+raptr.params.LST <- parseTOML('code/parameters/raptr.toml')
 gurobi.params.LST <- parseTOML('code/parameters/gurobi.toml')
 
 ## multispecies analysis
 # make prioritisations
 multi.spp.prioritisations <- llply(
 	list(
-		list(rapr.params.LST[[MODE]]$scenario.analysis$amount.target,NA,NA,
-			rapr.params.LST[[MODE]]$scenario.analysis$other.replicates),
-		list(rapr.params.LST[[MODE]]$scenario.analysis$amount.target,
-			rapr.params.LST[[MODE]]$scenario.analysis$surrogate.target,NA,
-			rapr.params.LST[[MODE]]$scenario.analysis$other.replicates), 
-		list(rapr.params.LST[[MODE]]$scenario.analysis$amount.target,NA,
-			rapr.params.LST[[MODE]]$scenario.analysis$genetic.target,
-			rapr.params.LST[[MODE]]$scenario.analysis$other.replicates)
+		list(raptr.params.LST[[MODE]]$scenario.analysis$amount.target,NA,NA,
+			raptr.params.LST[[MODE]]$scenario.analysis$other.replicates),
+		list(raptr.params.LST[[MODE]]$scenario.analysis$amount.target,
+			raptr.params.LST[[MODE]]$scenario.analysis$surrogate.target,NA,
+			raptr.params.LST[[MODE]]$scenario.analysis$other.replicates), 
+		list(raptr.params.LST[[MODE]]$scenario.analysis$amount.target,NA,
+			raptr.params.LST[[MODE]]$scenario.analysis$genetic.target,
+			raptr.params.LST[[MODE]]$scenario.analysis$other.replicates)
 	), 
 	function(y) {
 		species.prioritisation(

### 2016-07-22 code/R/functions/format_pnas.R

@@ -1,0 +1,42 @@
+
+#' Format .tex file to follow PNAS style
+#'
+#' This function formats a .tex file to follow PNAS conventions.
+#' @param input_file \code{character} file path for input .tex file 
+#' @param output_file \code{character} file path for input .tex file 
+#' @return \code{invisible()}
+format_pnas <- function(input_file, output_file) {
+	## initialization
+	# load tex file
+	input_file <- readLines(input_file)
+	
+	## main processing
+	# parse bibliography
+	input_file <- gsub('\\hypertarget{refs}{}', '\\begin{thebibliography}{100}', input_file, fixed=TRUE)
+	for (i in seq(grep('\\begin{thebibliography}', input_file, fixed=TRUE)+1, length(input_file))) {
+		# replace hypertargets with bibitems
+		if (grepl('\\hypertarget{ref-',input_file[i], fixed=TRUE)) {
+			curr_ref <- gsub('\\hypertarget{ref-', '', input_file[i], fixed=TRUE)
+			curr_ref <- gsub('}{}', '', curr_ref, fixed=TRUE)
+			input_file[i] <- paste0('\\bibitem{',curr_ref,'}')
+		}
+		# remove numbers from references
+		if (grepl('\\bibitem{', input_file[i-1], fixed=TRUE)) {
+			input_file[i] <- paste(stringi::stri_split_fixed(str=input_file[i], pattern='. ')[[1]][-1], collapse='. ')
+		}
+	}
+	# add in end of bibliograpgy
+	end_pos <- which(input_file=='\\end{document}')
+	input_file <- c(input_file[seq_len(end_pos-1)], '\\end{thebibliography}', input_file[seq(end_pos, length(input_file))])
+	# parse table
+	input_file <- gsub('\\begin{tabular}{', '\\begin{tabular*}{\\hsize}{@{\\extracolsep{\\fill}}', input_file, fixed=TRUE)
+	input_file <- gsub('\\end{tabular}', '\\end{tabular*}', input_file, fixed=TRUE)
+	# parse figure file paths
+	fig_pos <- grep('\\includegraphics', input_file, fixed=TRUE)
+	input_file[fig_pos] <- gsub('}', '.pdf}', input_file[fig_pos], fixed=TRUE)
+	# parse corresponding author
+	input_file <- gsub('\\correspondingauthor{\\textsuperscript{} }','\\correspondingauthor{\\textsuperscript{1} }',input_file,fixed=TRUE)
[…]

### 2016-08-03 code/rmarkdown/article.Rmd
+ added refs

@@
   - code: 1
-  - text: "To whom correspondence should be addresssed. E-mail: jeffrey.hanson@uqconnect.edu.au"
+  - text: "To whom correspondence should be addressed. E-mail: jeffrey.hanson@uqconnect.edu.au"
 author_contributions: |
@@
 abstract: |
-  Please provide an abstract of no more than 250 words in a single paragraph.
-Abstracts should explain to the general reader the major contributions of the article.
-References in the abstract must be cited in full within the abstract itself and cited in the text.
+  Protected areas buffer species from anthropogenic threats and provide places for processes to continue.
+However, evolutionary processes are not often explicitly considered when deciding where to site protected areas because genetic data require considerable resources to obtain and analyse.
+Here we show that environmental and geographic distance variables are effective surrogates for adaptive and neutral intra-specific variation for conservation planning, and that these surrogates can be used to deliver more effective solutions.
+We obtained distribution and genetic data from the international IntraBioDiv project for 27 species collected over the European Alps.
+Samples were collected using a gridded sampling scheme.
+We used these cells as planning units to generate solutions.
+For each species, we identified loci that were potentially under selection using outlier loci methods, and mapped their main gradients of adaptive and neutral genetic variation across the planning units.
+Our first analysis confirms that the the spatial patterns of adaptive and neutral genetic variation correlate with those of environmental and geographic variation (respectively).
+Our second analysis verifies that solutions that secure more adaptive genetic variation also tend to secure more adaptive variation, and similar patterns were found for neutral genetic variation and geographic variables.
+Our third analysis found that solutions generated using additional targets to explicitly secure a representative sample of the environmental and geographic variation across a species range resulted in solutions that secure more neutral genetic variation for most species.
+Additionally, we found that most solutions--even those generated without surrogate or genetic data--tended to secure most of the adaptive genetic variation for most species.
+Overall, our analysis provides a rigorous examination of two potential surrogates for genetic variation.
 significance: |
@@
 Here we found that environmental and geographic variables are cost-effective surrogates for genetic data.
-These variable can be used to identify candidate reserve systems that secure the adaptive and neutral evolutionary processes for a greater number of species.
-This research highlights the importance of obtaining a representative sample of species when generating broad-scale prioritisations.
+These variable can be used to deliver candidate reserve systems that secure the adaptive and neutral evolutionary processes for a greater number of species with minimal additional cost.
+This research highlights the importance of explicitly considering biodiversity processes during conservation planning exercises.
 acknowledgements: |
@@
 To ensure that the plans for protected areas (prioritisations) are cost-effective, they are often generated using decision support tools to identify near-optimal solutions.
-These prioritisations often contain plans for a network of multiple protected areas so they can preserve biological processes (eg.
-connectivity between populations).
+These prioritisations often contain plans for a network of multiple protected areas so they can preserve biological processes.
 For example, \texttt{Marxan} [@r22] uses amount-based targets that specify a minimum amount of habitat to preserve for each species to identify cost-effective solutions (eg.
 100 km $^2$ of land occupied by the species needs to be secured in a reserve).
-To promote connecitivty among multiple reserves, \texttt{Marxan} also contains the functionality to penalise fragmented solutions (using the boundary length modifier (\textit{BLM}) parameter).
+To promote connectivity among multiple reserves, \texttt{Marxan} also contains the functionality to penalise fragmented solutions (using the boundary length modifier (\textit{BLM}) parameter).
 However, conservation planning exercises typically assume that all individuals in the same species are equivalent.
@@
 Intra-specific genetic variation is important for long-term species persistence.
-As a consequence, there has been increasing interest in designing prioritisations that preserve and cultivate this aspect of biodiversity [@r67; @r3; @r61].
+As a consequence, there has been increasing interest in designing prioritisations that preserve and cultivate this aspect of biodiversity [@r67; @r3; @r61; @r499].
 Although the strength of natural selection is a continuous force; broadly speaking, genetic variation can be classified as either adaptive or neutral [reviewed in @r480].
@@
 Typically, such "adaptive" loci are detected due to anomalous patterns among individuals [eg.
-@r456; @r487; @r480; @r490], though phenotypic-genetic associations such as quantitative trait analyses or genome-wide association measures are more likely to identify functionally important genomic regions for individual species [@rXXXX; @rXXXX].
-By preserving existing patterns of adaptive variation, protected areas can ensure that populations with particularly beneficial adaptations are not lost [@r3].
+@r456; @r487; @r480; @r490], though phenotypic-genetic associations such as quantitative trait analyses or genome-wide association measures are more likely to identify functionally important genomic regions for individual species [eg.
+@r500].
+By preserving existing patterns of adaptive variation, protected areas can ensure that populations with particularly beneficial adaptations are not lost [@r3; @r43].
 In contrast, neutral genetic variation is associated with loci that do not (significantly) affect fitness.
@@
 By preserving existing patterns of neutral variation, protected areas can safeguard against the adverse effects of low genetic diversity [eg.
-inbreeding depression; @r67].
-Thus protected area networks should aim to preserve both adaptive and neutral patterns of genetic variation for species of conservation interest [@r67; @r3].
-Despite this, genetic data are not often used to inform conservation planning exercises because such data require considerable investment to obtain and analyze for a single species [@r61; but see @r5 and @r157 for developments on using genetic data in conservation planning], and most large-scale planning exercises involve hundreds of species [eg.
-@rXXX].
+inbreeding depression; @r67] .
+Thus protected area networks should aim to preserve both adaptive and neutral patterns of genetic variation for species of conservation interest [@r67; @r3; @r43].
+Despite this, genetic data are not often used to inform conservation planning exercises because such data require considerable investment to obtain and analyze [@r61; but see @r5 and @r157 for developments on using genetic data in conservation planning].
 Thus prioritisations may fail to secure a sufficient proportion of the intra-specific genetic variation of species, and ultimately fail to deliver long-term conservation outcomes.
@@
 On the other hand, neutral genetic variation arises from a reduction in gene flow between individuals.
-Over the last few years, the field of phylogeography has predominantly focused on describing neutral genetic diversity, and so the landscape factors that affect neutral variation are relatively well understood [@rXXXX].
+Over the last few years, the field of phylogeography has predominantly focused on describing neutral genetic diversity, and so the landscape factors that affect neutral variation are relatively well understood [@r501].
 Potential surrogates for neutral variation have been based on variables that predict the levels of connectivity between planning units.
@@
 Second, we aim to test whether environmental and geographic variables are effective surrogates for adaptive and neutral variation (respectively) when generating prioritisations.
-We expect that prioritisations generated using surrogate-based targets will secure more genetic variation than prioritisations generated using only amount-based targets with the same number of planning units.
 Third, we aim to determine if the addition of surrogate-based targets to conservation planning exercises can result in more effective prioritisations.
-We expect that prioritisations generated using amount- and surrogate-based targets will sufficiently preserve the genetic variation of more species than those generated using only amount-based targets.
-Furthermore, we aim to test this hypothesis under several of different scenarios to see whether the additional constraints of real-world planning exercises exacerbate or nullify this effect.
+We expect that prioritisations generated using amount- and surrogate-based targets will sufficiently preserve the genetic variation of more species than prioritisations generated using only amount-based targets.
+Furthermore, we aim to test this hypothesis under several of different scenarios to see if our results are robust.
 If these surrogates can improve the intra-specific genetic variation captured in a prioritisation, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.
@@
 In particular, the differences between the conditional and marginal $R^2$ values for the models examining the effectiveness of geographic variables suggest that this unexplained spatial auto-correlation may be non-linear or driven by a few neighboring planning units with genetically similar individuals.
-This finding is consonant with other studies that have found strong patterns of spatial auto-correlation in patterns of adaptive and neutral genetic variation [@rXXX; @rXXX; @rXXX].
+This finding is consonant with other studies that have found strong patterns of spatial auto-correlation in patterns of genetic variation [@r502].
 Although correlations between the spatial patterns of the environmental and geographic variables and genetic variables suggest that they may be useful for genetic variation in conservation planning, these correlations alone are insufficient to demonstrate that they are actually effective in the context of conservation planning.
@@
 In particular, mountains can act as barriers to gene flow in alpine landscapes and so nearby populations may be isolated from each other.
-Distances that Incorporated information on local topology (eg.
+Distances that incorporate information on local topology (eg.
 least-cost distances) may perform better, but generally require species-specific scaling to adequately describe connectivity.
-Third, the species used in this study utilise different dispersal mechanisms [wind dispersal, animal dispersal, offshoots from the mother plant; @rXXX], and some of these species may have dispersal modes that are non-linearly related to geographic distance (eg.
+Third, the species used in this study utilise different dispersal mechanisms [animal-, wind-, or gravity-based dispersal @r451], and some of these species may have dispersal modes that are non-linearly related to geographic distance (eg.
 wind dispersing plants can have high connectivity between nearby and distant areas).
@@
 Overall--contrary to our original hypotheses--it seems that conventional reserve selection methods may secure a representative sample of species' adaptive genetic variation if a comprehensive set of species is used.
-This outcome provides empirical support for one of the fundamental tenants of systematic conservation planning [@rXXX].
+This outcome provides empirical support for one of the fundamental tenants of systematic conservation planning.
 
@@
 To complete this first step, for each species, we fit the admixture and correlated alleles model implemented in \texttt{Structure} [version 2.3.4; @r461; @r462; `r structure.params.LST[[MODE]]$numruns` replicates per K; `r structure.params.LST[[MODE]]$admburnin` admixture burnin iterations; `r structure.params.LST[[MODE]]$burnin` burnin iterations; `r structure.params.LST[[MODE]]$numreps` iterations used for inference] using the number of populations for each species previously determined by Alvarez _et al._ [-@r478].
-We used \texttt{ClumPP} [version 1.1.2; @r459] to combine replicate runs to group individuals into populations (greedy algorithm using the $G^{\prime}$ static; `r clumpp.params.LST[[MODE]]$repeats` iterations).
+We used \texttt{ClumPP} [version 1.1.2, @r459] to combine replicate runs to group individuals into populations (greedy algorithm using the $G^{\prime}$ static; `r clumpp.params.LST[[MODE]]$repeats` iterations).
 We then ran $\texttt{Bayescan}$ [@r456] for each species using these clusters (1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations).

### 2016-08-03 code/R/functions/format_pnas.R

@@ -31,6 +31,8 @@ format_pnas <- function(input_file, output_file) {
 	# parse table
 	input_file <- gsub('\\begin{tabular}{', '\\begin{tabular*}{\\hsize}{@{\\extracolsep{\\fill}}', input_file, fixed=TRUE)
 	input_file <- gsub('\\end{tabular}', '\\end{tabular*}', input_file, fixed=TRUE)
+	input_file <- gsub('\\begin{table}[H]', '\\begin{table*}', input_file, fixed=TRUE)
+	input_file <- gsub('\\end{table}', '\\end{table*}', input_file, fixed=TRUE)
 	# parse figure file paths
 	fig_pos <- grep('\\includegraphics', input_file, fixed=TRUE)
 	input_file[fig_pos] <- gsub('}', '.pdf}', input_file[fig_pos], fixed=TRUE)

### 2016-08-17 code/R/analysis/04-outlier-locus-analysis.R
+ use checkpoint in properly

@@ -1,5 +1,5 @@
 ## load .rda
-checkpoint::checkpoint(general.params.LST[[MODE]]$checkpoint_date, R.version=general.params.LST[[MODE]]$checkpoint_R_version, scanForPackages=FALSE)
+checkpoint::checkpoint('2016-08-03', R.version='3.3.0', scanForPackages=FALSE)
 session::restore.session('data/intermediate/03-population-clustering-analysis.rda')
 
 ## load parameters

### 2016-08-17 code/R/analysis/06-format-data-for-prioritisations.R

@@ -1,5 +1,5 @@
 ## load .rda
-checkpoint::checkpoint(general.params.LST[[MODE]]$checkpoint_date, R.version=general.params.LST[[MODE]]$checkpoint_R_version, scanForPackages=FALSE)
+checkpoint::checkpoint('2016-08-03', R.version='3.3.0', scanForPackages=FALSE)
 session::restore.session('data/intermediate/05-genetic-nmds.rda')
 
 # subset data.frames for species with identifiable adaptive and/or neutral genetic variation

### 2016-08-22 code/rmarkdown/article.Rmd
+ use checkpoint in properly
+ upgrade article deps to 3.3.1
+ assorted article fixes

@@
 # load .rda file
+checkpoint::checkpoint('2016-08-03', R.version='3.3.1', scanForPackages=FALSE)
 try(session::restore.session('data/final/results.rda'))
@@
 
-First, we found that on average the prioritisations tended to secure more adaptive (mean `r round(mean(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` % secured in all prioritisations $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)`) than neutral genetic variation (mean `r round(mean(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)` % secured in all prioritisations $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)`; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.3.DF$Pr..Chisq.[[2]])`$).
+First, we found that on average the prioritisations tended to secure more adaptive (mean `r round(mean(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` % secured in all prioritisations $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)`) than neutral genetic variation (mean `r round(mean(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)` % secured in all prioritisations $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)`; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.3.DF$Pr..Chisq.[[2]])`).
 This result indicates that regardless of the targets used to generate a prioritisation, or the planning context under which the prioritisation was generated, prioritisations tend to secure more adaptive then neutral genetic variation.
@@
 Second, the proportion of genetic variation secured in a prioritisation was found to depend on the combination of targets used to generate the prioritisation (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$Pr..Chisq.[[2]])`)$.
-Specifically, the amount-based prioritisations (`r round(mean(scenario.sub.DF,filter(Prioritisation=='Amount targets')$genetic.held*100),2)` % overall genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)`) tended to secure less variation than the prioritisations generated using additional surrogate targets (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` % secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$) or genetic targets (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` % secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)`; `$Z_1 = `r round(scenario.GLHT.2.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[2]])`$).
+Specifically, the amount-based prioritisations (`r round(mean(scenario.sub.DF,filter(Prioritisation=='Amount targets')$genetic.held*100),2)` % overall genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)`) tended to secure less variation than the prioritisations generated using additional surrogate targets (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` % secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$) or genetic targets (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` % secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[2]])`$).
 No evidence was found to suggest that the prioritisations that included additional surrogate-based targets (`r raptr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100` %) or genetic-based targets (`r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %) secured a different proportion of genetic variation ($Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).
@@
 
-Third, we also found significant differences in the average proportion of genetic variation secured by prioritisations generated under different scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.2.DF$statistic[[2]],2)`$; $P$ `r pvalString(scenario.ANOVA.2.DF$Pr..Chisq.[[2]])`).
+Third, we also found significant differences in the average proportion of genetic variation secured by prioritisations generated under different scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.2.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.2.DF$Pr..Chisq.[[2]])`).
 Specifically, both the multi-species (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)`) and multi-species with cost (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)`) were both found to secure more genetic variation than the single species scenario (mean `r round(mean(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held),2)`; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[2]])`$; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$ respectively).
@@
 Generally, the slopes correlating the proportion of geographic variation and neutral variation secured in a prioritisation were higher than those for the proportion of environmental variation and adaptive variation secured in a prioritisation (Figures 2a--2b and S56--S57).
-To secure a representative sample of each species' adaptive genetic variation ($>=$ `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %), the prioritisation would need to sample on average `r mean(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE)` % $\pm$ `r sd(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE)` SD of the environmental conditions found across the planning units occupied by the species.
-On the other hand, to secure a representative sample of the species' neutral genetic variation ($>=$ `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100`), the prioritisation would need to sample % `r mean(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE)` $\pm$ `r sd(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE)` SD of the variation in geographic position among the planning units occupied by the species.
+To secure a representative sample of each species' adaptive genetic variation ($\geq$ `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %), the prioritisation would need to sample on average `r round(mean(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` SD of the environmental conditions found across the planning units occupied by the species.
+On the other hand, to secure a representative sample of the species' neutral genetic variation ($\geq$ `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100`), the prioritisation would need to sample % `r round(mean(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` $\pm$ `r round(sd(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` SD of the variation in geographic position among the planning units occupied by the species.
 These results suggest that geographic surrogates may require higher targets than environmental surrogates to generate prioritisations that secure representative sample of genetic variation.
@@
 Generally, the addition of environmental and geographic targets were needed to obtain prioritisations that secure a representative proportion of neutral genetic variation (Figure 4).
-With exception to the multi-species with cost planning scenario (Figure 4h), prioritisations generated using only amount targets failed to secure a representative proportion of neutral genetic variation ($>=$ `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %) for most species (Figures 4b, 4f, 4h).
+With exception to the multi-species with cost planning scenario (Figure 4h), prioritisations generated using only amount targets failed to secure a representative proportion of neutral genetic variation ($\geq$ `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %) for most species (Figures 4b, 4f, 4h).
 Whereas, in all scenarios, the prioritisations generated using amount and surrogate based targets secured a representative proportion of neutral genetic variation for most species.

### 2016-09-16 code/rmarkdown/article.Rmd
+ remove clean from all tagrget in makefile + various ms rewrites + make figures prettier

@@
 These places buffer species from anthropogenic impacts and provide places for species to persist.
-However, the resources available for conservation are limited.
-Thus protected areas need to be sited in places that achieve conservation goals for minimal cost.
-To ensure that the plans for protected areas (prioritisations) are cost-effective, they are often generated using decision support tools to identify near-optimal solutions.
-These prioritisations often contain plans for a network of multiple protected areas so they can preserve biological processes.
-For example, \texttt{Marxan} [@r22] uses amount-based targets that specify a minimum amount of habitat to preserve for each species to identify cost-effective solutions (eg.
-100 km $^2$ of land occupied by the species needs to be secured in a reserve).
-To promote connectivity among multiple reserves, \texttt{Marxan} also contains the functionality to penalise fragmented solutions (using the boundary length modifier (\textit{BLM}) parameter).
-However, conservation planning exercises typically assume that all individuals in the same species are equivalent.
+Since the resources available for conservation are limited, protected areas need to be sited in places that fulfill conservation objectives for minimal cost [@rXXX].
+To achieve this, conservation planning exercises often generate plans for entire networks of protected areas (prioritisations) to allocate resources cost-effectively [@rXXX].
+By generating plans for multiple reserves simultaneously, protected areas can preserve broad-scale biodiversity processes.
 
-Intra-specific genetic variation is important for long-term species persistence.
+Intra-specific genetic variation is important for long-term species persistence [@rXXX].
 As a consequence, there has been increasing interest in designing prioritisations that preserve and cultivate this aspect of biodiversity [@r67; @r3; @r61; @r499].
-Although the strength of natural selection is a continuous force; broadly speaking, genetic variation can be classified as either adaptive or neutral [reviewed in @r480].
-Adaptive genetic variation is associated with loci that (significantly) affect fitness.
+Broadly speaking, although the strength of natural selection is a continuous force, genetic variation can be classified as either adaptive or neutral [reviewed in @r480].
+Adaptive genetic variation is associated with loci that significantly affect fitness.
 Typically, such "adaptive" loci are detected due to anomalous patterns among individuals [eg.
-@r456; @r487; @r480; @r490], though phenotypic-genetic associations such as quantitative trait analyses or genome-wide association measures are more likely to identify functionally important genomic regions for individual species [eg.
+@r456; @r487; @r480; @r490], though phenotypic-genetic associations are more likely to identify functionally important genomic regions for individual species [eg.
 @r500].
 By preserving existing patterns of adaptive variation, protected areas can ensure that populations with particularly beneficial adaptations are not lost [@r3; @r43].
-In contrast, neutral genetic variation is associated with loci that do not (significantly) affect fitness.
+In contrast, neutral genetic variation is associated with loci that do not significantly affect fitness.
 Neutral variation often reflects the evolutionary history of different populations.
 By preserving existing patterns of neutral variation, protected areas can safeguard against the adverse effects of low genetic diversity [eg.
-inbreeding depression; @r67] .
+inbreeding depression; @r67].
 Thus protected area networks should aim to preserve both adaptive and neutral patterns of genetic variation for species of conservation interest [@r67; @r3; @r43].
-Despite this, genetic data are not often used to inform conservation planning exercises because such data require considerable investment to obtain and analyze [@r61; but see @r5 and @r157 for developments on using genetic data in conservation planning].
-Thus prioritisations may fail to secure a sufficient proportion of the intra-specific genetic variation of species, and ultimately fail to deliver long-term conservation outcomes.
+Despite this, genetic data are not often used to inform conservation planning exercises because such data require considerable investment to obtain and analyze [@r61; but see @r5 and @r157 for recent efforts].
+Thus prioritisations may fail to sample species' intra-specific genetic variation, and ultimately fail to deliver long-term conservation outcomes.
 
-Recently, conservation scientists have proposed using surrogate data to generate prioritisations that adequately preserve intra-specific genetic variation--without needing to utilise genetic data to acheive this--but have yet to comprehensively assess whether the proposed spatial surrogates adequately represent relevant aspects of intra-specific genetic variation.
+Recently, the use of surrogate data has been proposed to generate prioritisations that adequately preserve intra-specific genetic variation--without needing to utilise genetic data to achieve this.
 Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been proposed to guide reserve selection as surrogates for adaptive genetic variation [eg.
 environmental conditions; @r247; @r2].
-By preserving a more representative sample of environmental conditions across the species' ranges, these researchers aimed to generate prioritisations that would preserve more adaptive variation.
-However, the effectiveness of this approach remains unverified.
+By preserving a more representative sample of environmental conditions across the species' ranges, prioritisations could be generated that preserve more adaptive variation.
+However, to our knowledge, the effectiveness of this approach remains unverified.
 On the other hand, neutral genetic variation arises from a reduction in gene flow between individuals.
@@
 Furthermore, we aim to test this hypothesis under several of different scenarios to see if our results are robust.
-If these surrogates can improve the intra-specific genetic variation captured in a prioritisation, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.
+If these surrogates can improve the intra-specific genetic variation captured in a prioritisation, then conservation planners may use these surrogates when genetic data is not available--which is most often the case--and in turn, deliver more effective protected areas.
 
@@
 <!-- summary of methods -->
-We successfully mapped the main spatial patterns of adaptive (if present) and neutral variation for each of the 27 species.
-The outlier detection methods detected putatively adaptive genetic variation in `r as.character(as.english(length(env.surrogacy.LST)))` species.
-Generally, the two outlier detection methods gave fairly consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD).
-Of the species that were associated with adaptive genetic variation, only a few loci were detected as under selection (mean `r round(mean(spp.outlier.loci.props, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.loci.props, na.rm=TRUE),2)` SD).
-After identifying the loci showing strong signals of selection, we used NMDS analyses to identify the main gradients of adaptive (if detected) and neutral variation for each species.
-For all species in which we detected outlier loci, only a couple of continuous dimensions were needed to sufficiently describe their patterns of adaptive genetic variation (all $K = `r unique(na.omit(adaptive.k))`$; Table S2).
-On the other hand, a few additional continuous dimensions were occasionally needed to describe their patterns of neutral genetic variation (minimum $K = `r min(neutral.k)`$; mean $K = `r round(mean(neutral.k),2)`$; maximum $K = `r max(neutral.k)`$; Table S2).
-The ordinations resulting from the NMDS analyses were used to construct an adaptive (if present) and neutral genetic space for each species.
+We detected putatively adaptive genetic variation in `r as.character(as.english(length(env.surrogacy.LST)))` of the 27 plant species.
+Generally, the two outlier detection methods returned consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD).
+Of the species that were associated with adaptive genetic variation, only a small proportion of loci were classified as being adaptive (mean `r round(mean(spp.outlier.loci.props, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.loci.props, na.rm=TRUE),2)` SD).
+After identifying the loci showing strong signals of selection, we used non-metric multi-dimensional scaling (NMDS) analyses to identify the main gradients of adaptive (if detected) and neutral variation for each species.
+Generally, only a couple of continuous dimensions were needed to sufficiently describe their patterns of adaptive (all $K = `r unique(na.omit(adaptive.k))`$; Table S2) and neutral genetic variation (minimum $K = `r min(neutral.k)`$; mean $K = `r round(mean(neutral.k),2)`$; maximum $K = `r max(neutral.k)`$; Table S2).
+The resulting ordinations were used to construct an adaptive (if detected) and neutral genetic space for each species.
 The spatial distribution of these genetic spaces (Figures S29--S54) generally showed patterns of spatial auto-correlation (eg.
-_Cerastium uniflorum_, Figure SXX; _Dryas octopetala_, Figures SXXX)--a good sign that the environmental and geographic variables may prove effective surrogates.
+_Cerastium uniflorum_, Figure SXX; _Dryas octopetala_, Figures SXXX)--a good sign that environmental and geographic variables may prove effective surrogates.
 
 <!-- correlation analysis -->
-The correlation between dissimilarity matrices based on the environmental and geographic characteristics of the planning units and dissimilarity matrices based on the adaptive and neutral genetic characteristics of individuals in the planning units (Table 1).
-Specifically, environmental variation was found to significantly correlate with adaptive genetic variation for `r as.character(as.english(sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)))` species of the `r as.character(as.english(sum(!is.na(correlation.model.summary.DF$Environmental.corr.P))))` species associated with adaptive variation (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models).
-These results suggest that prioritisations with different environmental characteristics tend to have individuals with different adaptive genetic characteristics.
-Additionally, geographic distances between planning units were found to significantly correlate with neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` species of the total `r sum(!is.na(correlation.model.summary.DF$Geographic.corr.P))` (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models).
-Thus planning units that were located further apart tended to have individuals with increasingly different neutral genetic characteristics.
-After verifying that the geographic and environmental variables generally correlated with genetic variation, our next step was to determine how effective these variables were as surrogates in the context of conservation planning.
+We constructed dissimilarity matrices between the planning units.
+For each species, we constructed matrices expressing the differences between planning units as a function of their environmental characteristics and geographic position, and the adaptive (if detected) and neutral genetic characteristics between individuals.
+We found the spatial patterns of environmental variation to be significantly correlated with those of adaptive genetic variation for `r as.character(as.english(sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)))` species of the `r as.character(as.english(sum(!is.na(correlation.model.summary.DF$Environmental.corr.P))))` species associated with adaptive variation (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models).
+We also found the geographic distance between planning units to significantly correlate with the spatial patterns of neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` species of the total `r sum(!is.na(correlation.model.summary.DF$Geographic.corr.P))` (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models).
+Thus, for most species, planning units that contained different environmental conditions contained individuals with different adaptive genetic characteristics, and planning units that were further apart tended to contain individuals with different neutral genetic characteristics.
+After verifying that the environmental and geographic variables were generally correlated with genetic variation, our next step was to determine how effective these variables were as surrogates for conservation planning.
 ```{r, echo=FALSE, results='asis'}
@@
 <!-- surrogacy analysis -->
-The environmental and geographic variables were moderately effective surrogates for genetic variation in most species (Figure 1).
-The relationship between the proportion of adaptive genetic variation and environmental variation secured in a given prioritisation was found to differ depending on the species for which the prioritisation was generated (species $\times$ proportion of environmental variation interaction term; ${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$).
-The slopes for each species were all significantly different to zero (minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`; maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3).
-Similarly, the relationship between the proportion of neutral genetic variation and geographic variation secured in a prioritisation was also found to affected by the interaction between species and the proportion of neutral variation secured in the prioritisation (${\chi^2}_{`r geo.surrogacy.AOV$df[[4]]`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`).
-The slope for each species were also significantly different to zero (minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`; maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3).
-After establishing a relationship between the proportion of the surrogate variables and proportion of genetic variation secured in a prioritisation, the next step was to determine if prioritisations benefited from the addition of surrogate-based targets and verify if such benefits were evident across a variety of planning scenarios.
+We found the environmental and geographic variables to be moderately effective surrogates for genetic variation in most species (Figure 1).
+For each species, we randomly generated prioritisations, and calculated the proprotion of environmental, geographic, adaptive (if present), and neutral genetic variation they sampled.
+We found the relationship between the proportion of adaptive genetic variation and environmental variation secured in a given prioritisation to depend on the species for which the prioritisation was generated (species $\times$ proportion of environmental variation interaction term; ${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$).
+Similarly, we found the relationship between the proportion of neutral genetic variation and geographic variation secured in a prioritisation to depend on the species for which the prioritisation was generated (${\chi^2}_{`r round(geo.surrogacy.AOV$df[[4]],2)`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`).
+For all species, we found these relationship to be significantly different to zero (environmental: minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; geographic: minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3).
+After establishing that the proportion of surrogate and genetic variation were correlated in randomly generated prioritisations, our final step was to determine if prioritisations were more effective when generated with additional surrogate-based targets.
 
@@
 		axis.line.y=element_line(color='black')) +
-	geom_text(x=0.05, y=0.95, label='(a)') +
 	coord_cartesian(xlim=c(0, 1))
@@
 		fullrange=TRUE, se=FALSE, colour='black')
+p1 <- p1 + geom_text(x=-Inf, y=Inf, hjust=-0.25, vjust=1.5, label='(a)')
 
@@
 		axis.line.y=element_line(color='black', colour='transparent')) +
-	geom_text(x=0.05, y=0.95, label='(b)') +
 	coord_cartesian(xlim=c(0, 1))
@@
 		fullrange=TRUE, se=FALSE, colour='black')
+p2 <- p2 + geom_text(x=-Inf, y=Inf, hjust=-0.25, vjust=1.5, label='(b)')
 
@@
 		plot.background=element_rect(fill='transparent', colour='transparent')) +
-		geom_text(x=0.05, y=6, label='(c)') +
-		coord_cartesian(xlim=c(0, 1))
+	geom_text(x=-Inf, y=Inf, hjust=-0.25, vjust=1.5, label='(c)') +
+	scale_y_continuous(labels=function(x) {formatC(x, digits=2, format='f')}) +
+	scale_x_continuous(limits=c(0, 1), labels=function(x) {formatC(x, digits=2, format='f')})
 
@@
 		plot.background=element_rect(fill='transparent')) +
-		geom_text(x=0.05, y=8, label='(d)') +
-		coord_cartesian(xlim=c(0, 1))
+	geom_text(x=-Inf, y=Inf, hjust=-0.25, vjust=1.5, label='(d)') +
+	scale_y_continuous(labels=function(x) {formatC(x, digits=2, format='f')}) +
+	scale_x_continuous(limits=c(0, 1), labels=function(x) {formatC(x, digits=2, format='f')})
 
@@
 We then computed the proportion of adaptive and neutral genetic variation secured in each prioritisation (Figure 4).
-We found no evidence to suggest that any of the interaction terms were significant (${\chi^2}_{`r scenario.ANOVA.1.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.1.DF$Pr..Chisq.[[2]])`$).
+We found no evidence to suggest that any of the interaction terms were significant (${\chi^2}_{`r scenario.ANOVA.1.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.1.DF$p.value[[2]])`$).
 This result indicates the trends observed in each of the main effects were consistent across all other levels in the main effects.
 
-First, we found that on average the prioritisations tended to secure more adaptive (mean `r round(mean(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` % secured in all prioritisations $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)`) than neutral genetic variation (mean `r round(mean(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)` % secured in all prioritisations $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)`; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.3.DF$Pr..Chisq.[[2]])`).
+First, we found that on average the prioritisations tended to secure more adaptive (mean `r round(mean(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` % secured in all prioritisations $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)`) than neutral genetic variation (mean `r round(mean(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)` % secured in all prioritisations $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)`; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.3.DF$p.value[[2]])`).
 This result indicates that regardless of the targets used to generate a prioritisation, or the planning context under which the prioritisation was generated, prioritisations tend to secure more adaptive then neutral genetic variation.
@@
 	facet_grid(Context ~ Prioritization) +
-	scale_fill_gradientn(name='Selection Frequency (%)',
-		colors=brewer.pal(9, 'PuBu'),
+	scale_fill_gradient(name='Selection Frequency (%)',
 		guide=guide_colorbar(title.position='bottom', title.hjust=0.5,
@@
 
-Second, the proportion of genetic variation secured in a prioritisation was found to depend on the combination of targets used to generate the prioritisation (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$Pr..Chisq.[[2]])`)$.
-Specifically, the amount-based prioritisations (`r round(mean(scenario.sub.DF,filter(Prioritisation=='Amount targets')$genetic.held*100),2)` % overall genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)`) tended to secure less variation than the prioritisations generated using additional surrogate targets (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` % secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$) or genetic targets (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` % secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[2]])`$).
+Second, we found that the proportion of genetic variation secured in a prioritisation was found to depend on the combination of targets used to generate the prioritisation (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$p.value[[2]])`)$.
+Specifically, the amount-based prioritisations (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)` % overall genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)`) tended to secure less variation than the prioritisations generated using additional surrogate targets (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` % secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$) or genetic targets (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` % secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[2]])`$).
 No evidence was found to suggest that the prioritisations that included additional surrogate-based targets (`r raptr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100` %) or genetic-based targets (`r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %) secured a different proportion of genetic variation ($Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).
@@
 
-Third, we also found significant differences in the average proportion of genetic variation secured by prioritisations generated under different scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.2.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.2.DF$Pr..Chisq.[[2]])`).
+Third, we also found significant differences in the average proportion of genetic variation secured by prioritisations generated under different scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.2.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.2.DF$p.value[[2]])`).
 Specifically, both the multi-species (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)`) and multi-species with cost (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)`) were both found to secure more genetic variation than the single species scenario (mean `r round(mean(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held),2)`; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[2]])`$; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$ respectively).
@@
 We aimed to investigate the effectiveness of broad-scale environmental and geographic variables as surrogates for adaptive and neutral genetic variation in conservation planning.
-Our study is the first to address this aim by measuring the proportion of genetic variation secured by prioritisations generated for diverse set of species.
+Our study is the first to address this aim using a diverse set of species and actually measuring the proportion of genetic variation secured in prioritisations.
 Overall, we found that the spatial patterns of environmental and geographic variables correlate with those of adaptive and neutral patterns for most species.
@@
 <!-- surrogate/genetic dissimilarity matrix correlations -->
-Differences in the the environmental conditions and geographic position among the planning units tended to correlate with genetic differences among the individuals in the planning units (Table 1).
+Our analyses showed that differences in the the environmental conditions and geographic position among the planning units tended to correlate with genetic differences among the individuals in the planning units (Table 1).
 For both surrogates and their corresponding models, the conditional $R^2$ values (environmental vs.
@@
 In particular, the differences between the conditional and marginal $R^2$ values for the models examining the effectiveness of geographic variables suggest that this unexplained spatial auto-correlation may be non-linear or driven by a few neighboring planning units with genetically similar individuals.
-This finding is consonant with other studies that have found strong patterns of spatial auto-correlation in patterns of genetic variation [@r502].
-Although correlations between the spatial patterns of the environmental and geographic variables and genetic variables suggest that they may be useful for genetic variation in conservation planning, these correlations alone are insufficient to demonstrate that they are actually effective in the context of conservation planning.
-To demonstrate their effectiveness would require examining the relationship between the proportion of surrogate variation and genetic variation secured in a given prioritisation.
+Similarly, other studies have also reported strong patterns of spatial auto-correlation in patterns of genetic variation [eg.
+@r502].
 
 <!-- surrogate/genetic held correlations -->
-Prioritisations that secured a more representative sample of environmental conditions or geographic position tended to secure a greater proportion of adaptive and neutral genetic variation (Figure 2).
+Our randomly generated prioritisations that secured a more representative sample of environmental conditions and geographic position also secured a greater proportion of adaptive and neutral genetic variation (Figure 2).
 Generally, the environmental and geographic surrogates variables were equally effective surrogates for adaptive (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figures 2c and S56) and neutral genetic variation (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figure S57).
@@
 <!-- suitable targets for surrogates -->
-Generally, the slopes correlating the proportion of geographic variation and neutral variation secured in a prioritisation were higher than those for the proportion of environmental variation and adaptive variation secured in a prioritisation (Figures 2a--2b and S56--S57).
+Generally, we found that the slopes correlating the proportion of geographic variation and neutral variation secured in a prioritisation were higher than those for the proportion of environmental variation and adaptive variation secured in a prioritisation (Figures 2a--2b and S56--S57).
 To secure a representative sample of each species' adaptive genetic variation ($\geq$ `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %), the prioritisation would need to sample on average `r round(mean(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` SD of the environmental conditions found across the planning units occupied by the species.
@@
 Thus we recommend that prioritisations be generated using a combination of amount-based targets--as is conventionally used in reserve selection processes--and also surrogate-based targets.
-After establishing a relationship between the proportion of surrogate variation and genetic variation secured in prioritisation, we than generated prioritisations under various planning scenarios to provide insight into how the surrogates may work in the real-world.
 
 <!-- adaptive genetic variation secured in prioritisations-->
+Our results suggest that conservation planners can secure a representative sample of the intra-specific adaptive genetic variation using only amount targets.
 Almost all prioritisations secured a representative sample of species' adaptive genetic variation--regardless of whether surrogate-based or genetic targets were used or not (Figure 4).
-This result suggests that conservation planners can secure a representative sample of the intra-specific adaptive genetic variation using only amount targets.
-The spatial patterns of adaptive genetic variation tended to cluster into a few distinct groups for most species (eg.
-XXX; XXX; XXX; XXX), and so a prioritisation would only need to site a single planning unit in each cluster to secure a representative sample of the species' adaptive genetic variation.
-As a consequence, the single-species prioritisations generated using amount targets and without consideration to opportunity cost were often able to achieve high representation due to the fact they were essentially random selection (Figure 4a).
-To a lesser extent this is also true for the single-species prioritisations generated with amount targets and consideration to opportunity cost (Figure 4c), since the costs are fairly homogenous among planning units with exception to a few planning units lowland areas.
+Since the spatial patterns of adaptive genetic variation tended to cluster into a few distinct groups for most species (eg.
+_XXX XXX_, _XXX XXX_; _XXX XXX_; Figures SXXX, SXXX, SXXX respectively), a prioritisation would only need to select a planning unit in each cluster to secure a representative sample.
+As a consequence, the single-species prioritisations generated using amount targets and without opportunity cost data were often able to achieve high representation due to the fact they were essentially random selection (Figure 4a).
+To a lesser extent, this is also true for the single-species prioritisations generated with amount targets and consideration to opportunity cost (Figure 4c), since the costs were fairly homogenous among planning units with exception to a few planning units lowland areas.
 
 All the multi-species prioritisations secured a representative sample of most of the species' intra-specific adaptive genetic variation (Figures 4e and 4g).
-Unlike the single-species prioritisations, the multi-species prioiritisations generated with only amount-targets did not secure a representative sample of adaptive variation due to stochastic processes.
+Unlike the single-species prioritisations, the multi-species prioritisations generated with only amount-targets did not secure a representative sample of adaptive variation due to stochastic processes.
 Since these prioritisations were forced to secure at least `r raptr.params.LST[[MODE]]$scenario.analysis$amount.target*100` % of land occupied each species, these prioritisations were forced to secured planning units a range of environmental conditions due to the presence of habitat specialists in the data set (eg.
@@
 <!-- neutral genetic variation secured in prioritisations-->
-Generally, the addition of environmental and geographic targets were needed to obtain prioritisations that secure a representative proportion of neutral genetic variation (Figure 4).
+Our analyses showed that addition of environmental and geographic targets were needed to secure a representative sample of neutral genetic variation in most species (Figure 4).
 With exception to the multi-species with cost planning scenario (Figure 4h), prioritisations generated using only amount targets failed to secure a representative proportion of neutral genetic variation ($\geq$ `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %) for most species (Figures 4b, 4f, 4h).
-Whereas, in all scenarios, the prioritisations generated using amount and surrogate based targets secured a representative proportion of neutral genetic variation for most species.
-Although the the latter prioritisation did not secure a representative sample of the neutral genetic variation in all species, this result suggests that the use of environmental and geographic targets in conservation planning can result in a more effective prioritisation.
-Furthermore, this improvement was associated with only a moderate increase in cost in both scenarios (scenario without opportunity: `r round((score(multi.spp.prioritisations[[2]])/score(multi.spp.prioritisations[[1]]))*100,2)` %; scenario with opportunity cost: `r  round((score(multi.spp.prioritisations.with.cost[[2]])/score(multi.spp.prioritisations.with.cost[[1]]))*100,2)` %)--emphasising the utility of surrogates in conservation.
+Whereas, in all scenarios, prioritisations generated using amount and surrogate based targets secured substantially more neutral genetic variation for most species.
+This improvement was associated with only a moderate increase in cost in both scenarios (scenario without opportunity: `r round((score(multi.spp.prioritisations[[2]])/score(multi.spp.prioritisations[[1]]))*100,2)` %; scenario with opportunity cost: `r  round((score(multi.spp.prioritisations.with.cost[[2]])/score(multi.spp.prioritisations.with.cost[[1]]))*100,2)` %)--emphasising the utility of surrogates in conservation.
+Our results suggest that the use of environmental and geographic targets in conservation planning can result in more effective prioritisations.
 
@@
 We wish to alert our readers to several limitations associated with our analysis.
-First, the size of the planning units we used (23 $\times$ 25 km) is much larger than typically used in conservation planning scenarios (between 1--10 km^2).
+First, the size of the planning units we used (23 $\times$ 25 km) is much larger than typically used in conservation planning scenarios (between 1--10 km$^2$).
 We used this resolution because the genetic data were collected at this scale.
@@
 Whilst next-generation sequencing provides higher resolution genetic information than AFLP data [reviewed in @r480], our methodology would still have utilised only the main gradients of the genetic variation in order to generate prioritisations in a feasible amount of time.
-Overall, we are confident that the limitations associated with our study are relatively minor.
+Overall, we are confident that our study is robust.
 
@@
 	scale_fill_gradientn(name='Richness',
-		colors=brewer.pal(9,'PuBu'),
+		colors=brewer.pal(5, 'YlGn'),
 		guide=guide_colorbar(ticks=element_line(color='black'), border=element_line(color='black'))
@@
 	scale_fill_gradientn(name='Cost',
-		colors=brewer.pal(9, 'PuBu'),
+		colors=brewer.pal(5,'YlOrRd'),
 		values=qn01,
@@
 We chose this data set for several reasons.
-First, conservation planning exercises typically consider tens to hundreds of species with different evolutionary histories and life histories.
-To ensure that our results were relevant for conservation planners, we required a data set with a diverse set of species and a comparable number of species.
-The IntraBioDiv data set contains genetic data for 27 species with different evolutionary histories and life-history strategies.
-Second, we required genetic data for all species to be collected using a standardised sampling scheme, and this data set used a grid to collect samples.
-Third, we required the study area to contain broad-scale environmental variation to test the effectiveness of environmental variables.
-This data set spans across large attitudinal gradients.
+First, to address our aims, we required genetic data for all species to be collected using a standardised sampling scheme, and this data set used a grid to collect samples.
+Second, because conservation planning exercises typically consider tens to hundreds of species with different evolutionary histories and life histories [eg.
+@rXXX], we needed a data set that contains a comparable set of species so that our results will be relevant to planners.
+Third, we required the study area to contain broad-scale environmental variation to test the effectiveness of environmental variables, and this data set spans across large attitudinal gradients.
 Although this data set contains AFLPs which not ideal as genomic representatives (cf.
@@
 Their results suggest that prioritisations based on species distribution data alone will be unable to secure intra-specific genetic variation.
-However, whether or not environmental and geographic variables can augment prioritisations remains untested.
+However, whether or not environmental and geographic data can result in more prioritisations remains untested.
 
@@
 
-To describe the geographic location of each planning unit, we projected the grid into an equi-distant coordinate system (Europe Equidistant Conic: $\texttt{ESRI:102031}$), calculated their centroids, and extracted their two-dimensional coordinates.
+To describe the geographic location of each planning unit, we projected the grid into an equi-distant coordinate system (Europe Equidistant Conic; $\texttt{ESRI:102031}$), calculated their centroids, and extracted their two-dimensional coordinates.
 Thus we constructed two-dimensional geographic space--wherein each planning unit occupies a single point--as a potential surrogate for neutral genetic variation.
@@
 We used this data set because it is commonly to map habitat suitability in conservation planning [@r493].
-These layers and the planning units were projected into an equal-area coordinate system (Europe Lambert Conformal Conic: $\texttt{ESRI:102014}$).
+These layers and the planning units were projected into an equal-area coordinate system (Europe Lambert Conformal Conic; $\texttt{ESRI:102014}$).
 To reduce dimensionality, the 19 bioclimatic variables were clipped to the planning units and subjected to a principal components analysis (PCA; using \texttt{ArcMap} version 10.2.2; Table S1).

### 2016-09-20 code/rmarkdown/article.Rmd
+ assorted fixes to manuscript * rewrote some parts of article * fixed figure sizing in si

@@
 These places buffer species from anthropogenic impacts and provide places for species to persist.
-Since the resources available for conservation are limited, protected areas need to be sited in places that fulfill conservation objectives for minimal cost [@rXXX].
-To achieve this, conservation planning exercises often generate plans for entire networks of protected areas (prioritisations) to allocate resources cost-effectively [@rXXX].
+Since the resources available for conservation are limited, protected areas need to be sited in places that fulfill conservation objectives for minimal cost [@r255].
+To achieve this, conservation planning exercises often generate plans for entire networks of protected areas (prioritisations) to allocate resources cost-effectively.
 By generating plans for multiple reserves simultaneously, protected areas can preserve broad-scale biodiversity processes.
 
-Intra-specific genetic variation is important for long-term species persistence [@rXXX].
+Intra-specific genetic variation is important for long-term species persistence [@r165].
 As a consequence, there has been increasing interest in designing prioritisations that preserve and cultivate this aspect of biodiversity [@r67; @r3; @r61; @r499].
@@
 Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been proposed to guide reserve selection as surrogates for adaptive genetic variation [eg.
-environmental conditions; @r247; @r2].
+environmental conditions; @r202; @r247; @r2].
 By preserving a more representative sample of environmental conditions across the species' ranges, prioritisations could be generated that preserve more adaptive variation.
@@
 Here we investigated the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation in the context of conservation planning.
-To achieve this, we developed a novel framework for incorporating adaptive and neutral genetic data into conservation planning.
-First, we aim to verify if the spatial patterns in environmental and geographic variation correlate with the broad-scale spatial patterns in adaptive and neutral variation (respectively).
-Second, we aim to test whether environmental and geographic variables are effective surrogates for adaptive and neutral variation (respectively) when generating prioritisations.
-Third, we aim to determine if the addition of surrogate-based targets to conservation planning exercises can result in more effective prioritisations.
-We expect that prioritisations generated using amount- and surrogate-based targets will sufficiently preserve the genetic variation of more species than prioritisations generated using only amount-based targets.
-Furthermore, we aim to test this hypothesis under several of different scenarios to see if our results are robust.
-If these surrogates can improve the intra-specific genetic variation captured in a prioritisation, then conservation planners may use these surrogates when genetic data is not available--which is most often the case--and in turn, deliver more effective protected areas.
+To achieve this goal, we used distribution and genomic data for 27 alpine plants collected by the international IntraBioDiv project in the European Alps [@r451].
+We show that for most species the spatial patterns of adaptive and neutral genetic variation correlated with environmental and geographic variables.
+We also show that for most species the level of correlation is high enough to be operational for conservation planning.
+Finally, we demonstrate the using these surrogates results in prioritisations that secure more genetic variation.
 
@@
 We detected putatively adaptive genetic variation in `r as.character(as.english(length(env.surrogacy.LST)))` of the 27 plant species.
-Generally, the two outlier detection methods returned consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD).
+We used two outlier detection methods to detect loci under selection, and they returned relatively consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD).
 Of the species that were associated with adaptive genetic variation, only a small proportion of loci were classified as being adaptive (mean `r round(mean(spp.outlier.loci.props, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.loci.props, na.rm=TRUE),2)` SD).
@@
 The spatial distribution of these genetic spaces (Figures S29--S54) generally showed patterns of spatial auto-correlation (eg.
-_Cerastium uniflorum_, Figure SXX; _Dryas octopetala_, Figures SXXX)--a good sign that environmental and geographic variables may prove effective surrogates.
+_Cerastium uniflorum_, Figure S34; _Dryas octopetala_, Figures S36)--a good sign that environmental and geographic variables may prove effective surrogates.
 
 <!-- correlation analysis -->
+We verified that the spatial patterns in environmental and geographic variation correlated with the broad-scale spatial patterns in adaptive and neutral variation for most species (respectively).
 We constructed dissimilarity matrices between the planning units.
@@
 Thus, for most species, planning units that contained different environmental conditions contained individuals with different adaptive genetic characteristics, and planning units that were further apart tended to contain individuals with different neutral genetic characteristics.
-After verifying that the environmental and geographic variables were generally correlated with genetic variation, our next step was to determine how effective these variables were as surrogates for conservation planning.
+After verifying these patterns, our next step was to determine how effective these variables were as surrogates for conservation planning.
 ```{r, echo=FALSE, results='asis'}
@@
 <!-- surrogacy analysis -->
-We found the environmental and geographic variables to be moderately effective surrogates for genetic variation in most species (Figure 1).
+We found the environmental and geographic variables to be moderately effective predictors for genetic variation when assessing the performance of prioritisations (Figure 1).
 For each species, we randomly generated prioritisations, and calculated the proprotion of environmental, geographic, adaptive (if present), and neutral genetic variation they sampled.
@@
 For all species, we found these relationship to be significantly different to zero (environmental: minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; geographic: minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3).
-After establishing that the proportion of surrogate and genetic variation were correlated in randomly generated prioritisations, our final step was to determine if prioritisations were more effective when generated with additional surrogate-based targets.
+After establishing that the surrogate variables predicted how well prioritisations secured genetic variation, our final step was to determine if prioritisations were more effective when generated using surrogate-based targets.
 
@@
 <!-- scenario analysis -->
-To determine if the addition of environmental and geographic targets could actually improve the effectiveness of prioritisations, we generated prioritisations according to three scenarios using different targets (Figure 3).
-We then computed the proportion of adaptive and neutral genetic variation secured in each prioritisation (Figure 4).
-We found no evidence to suggest that any of the interaction terms were significant (${\chi^2}_{`r scenario.ANOVA.1.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.1.DF$p.value[[2]])`$).
+To determine if the addition of environmental and geographic targets could actually improve the effectiveness of prioritisations, we generated prioritisations according to three scenarios using different targets (Figure S58).
+We then computed the proportion of adaptive and neutral genetic variation secured in each prioritisation (Figure 3).
+We found no evidence to suggest that any interaction terms were significant (${\chi^2}_{`r scenario.ANOVA.1.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.1.DF$p.value[[2]])`$).
 This result indicates the trends observed in each of the main effects were consistent across all other levels in the main effects.
-
-First, we found that on average the prioritisations tended to secure more adaptive (mean `r round(mean(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` % secured in all prioritisations $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)`) than neutral genetic variation (mean `r round(mean(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)` % secured in all prioritisations $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)`; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.3.DF$p.value[[2]])`).
-This result indicates that regardless of the targets used to generate a prioritisation, or the planning context under which the prioritisation was generated, prioritisations tend to secure more adaptive then neutral genetic variation.
-
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=(5*0.9), fig.height=(6.5*0.8), out.width='3.42in', height='3.9in', fig.cap='Planning unit selection frequencies.
-Squares represent planning units, and their colors indicate their selection frequency.
-Each column of panels shows solutions generated using different targets.
-Each row of panels shows solutions generated under different scenarios.
-Note that the selection frequencies for the single-species prioritizations are based on their frequency among prioritizations generated for each species.
-The selection frequencies for the multi-species prioritizations are each based on a single optimal pioritization for all species and so are binary.'}
-# prepare data for plotting
-grid.sub.FPLY <- grid.sub.PLY
-for (i in seq_along(single.spp.prioritisations[[1]]))
-	grid.sub.FPLY@data[[paste0('Single-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations, function(x){colMeans(x[[i]]@results@selections)}))
-for (i in seq_along(single.spp.prioritisations.with.cost[[1]]))
-	grid.sub.FPLY@data[[paste0('Single-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations.with.cost, function(x){colMeans(x[[i]]@results@selections)}))
-for (i in seq_along(multi.spp.prioritisations))
-	grid.sub.FPLY@data[[paste0('Multi-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations[[i]]@results@selections)
-for (i in seq_along(multi.spp.prioritisations.with.cost))
-	grid.sub.FPLY@data[[paste0('Multi-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations.with.cost[[i]]@results@selections)
-grid.sub.FPLY <- spFortify(grid.sub.FPLY)
-grid.sub.FPLY <- grid.sub.FPLY %>% gather(
-		Combination,
-		`Selection.Frequency`,
-		`Single-species (equal costs).Amount`:`Multi-species (opportunity costs).Genetic`
-	) %>% mutate(
-		Context=sapply(strsplit(as.character(Combination), '.', fixed=TRUE), `[[`, 1),
-		Prioritization=sapply(strsplit(as.character(Combination), '.', fixed=TRUE), `[[`, 2)
-	) %>% mutate(
-		Prioritization=revalue(Prioritization, c('Amount'='Amount targets', 'Surrogate'='Amount & surrogate\ntargets', 'Genetic'='Amount & genetic\ntargets')),
-		Prioritization=factor(as.character(Prioritization), levels=c('Amount targets', 'Amount & surrogate\ntargets', 'Amount & genetic\ntargets')),
-		Context=factor(
-			gsub(' (', '\n(', Context, fixed=TRUE),
-			levels=c("Single-species\n(equal costs)", "Single-species\n(opportunity costs)",
-				"Multi-species\n(equal costs)", "Multi-species\n(opportunity costs)"))
-	)
-
-# make maps
-ggplot() +
-	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
-		fill='grey85', color='grey70') +
-	geom_polygon(data=grid.sub.FPLY, aes_string(x='long', y='lat', 
-		group='group', fill='Selection.Frequency'),
-		alpha=0.8, color=NA) +
-	theme_classic() +
-	theme(axis.ticks=element_blank(), axis.text=element_blank(),
-		plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(),
-		strip.background = element_rect(fill='grey20'),
-		strip.text = element_text(color='white', size=8),
-		legend.position='bottom', legend.key.width=unit(1.8, 'cm'),
-		legend.title=element_text(vjust=-5),
-		legend.margin=unit(c(0.1), 'cm'),
-		axis.title=element_blank()) +
-	coord_cartesian(
-		xlim=buffered.range(grid.sub.FPLY$long, 0),
-		ylim=buffered.range(grid.sub.FPLY$lat, 0)
-	) +
-	facet_grid(Context ~ Prioritization) +
-	scale_fill_gradient(name='Selection Frequency (%)',
-		guide=guide_colorbar(title.position='bottom', title.hjust=0.5,
-			ticks=element_line(color='black'), border=element_line(color='black'))
-	)
-```
-
-Second, we found that the proportion of genetic variation secured in a prioritisation was found to depend on the combination of targets used to generate the prioritisation (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$p.value[[2]])`)$.
-Specifically, the amount-based prioritisations (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)` % overall genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)`) tended to secure less variation than the prioritisations generated using additional surrogate targets (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` % secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$) or genetic targets (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` % secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[2]])`$).
-No evidence was found to suggest that the prioritisations that included additional surrogate-based targets (`r raptr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100` %) or genetic-based targets (`r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %) secured a different proportion of genetic variation ($Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).
-These results suggest that the addition of environmental and geographic surrogates can improve the proportion of intra-specific genetic variation secured in a prioritisation for most species.
 
@@
 
-Third, we also found significant differences in the average proportion of genetic variation secured by prioritisations generated under different scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.2.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.2.DF$p.value[[2]])`).
+First, the prioritisations tended to secure more adaptive (mean `r round(mean(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` % secured in all prioritisations $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)`) than neutral genetic variation (mean `r round(mean(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)` % secured in all prioritisations $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)`; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.3.DF$p.value[[2]])`).
+This result indicates that regardless of the targets used to generate a prioritisation, or the planning context under which the prioritisation was generated, prioritisations tend to secure more adaptive then neutral genetic variation.
+
+Second, the proportion of genetic variation secured in a prioritisation was found to depend on the targets used to generate the prioritisation (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$p.value[[2]])`)$.
+Specifically, the amount-based prioritisations (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)` % overall genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)`) tended to secure less variation than the prioritisations generated using additional surrogate targets (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` % secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$) or genetic targets (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` % secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[2]])`$).
+No evidence was found to suggest that the prioritisations that included additional surrogate-based targets (`r raptr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100` %) or genetic-based targets (`r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %) secured a different proportion of genetic variation ($Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).
+These results suggest that the addition of environmental and geographic surrogates can improve the proportion of intra-specific genetic variation secured in a prioritisation for most species.
+
+Third, we found significant differences in the average proportion of genetic variation secured by prioritisations generated under different scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.2.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.2.DF$p.value[[2]])`).
 Specifically, both the multi-species (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)`) and multi-species with cost (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)`) were both found to secure more genetic variation than the single species scenario (mean `r round(mean(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held),2)`; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[2]])`$; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$ respectively).
@@
 <!-- surrogate/genetic held correlations -->
-Our randomly generated prioritisations that secured a more representative sample of environmental conditions and geographic position also secured a greater proportion of adaptive and neutral genetic variation (Figure 2).
+Our randomly generated prioritisations show that environmental and geographic surrgoates are strong enough to be operationalisable (Figure 2).
 Generally, the environmental and geographic surrogates variables were equally effective surrogates for adaptive (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figures 2c and S56) and neutral genetic variation (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figure S57).
@@
 Since the spatial patterns of adaptive genetic variation tended to cluster into a few distinct groups for most species (eg.
-_XXX XXX_, _XXX XXX_; _XXX XXX_; Figures SXXX, SXXX, SXXX respectively), a prioritisation would only need to select a planning unit in each cluster to secure a representative sample.
+_Carex semipervens_ and _Cerastium uniflorum_; Figures S33 and S34 respectively), a prioritisation would only need to select a planning unit in each cluster to secure a representative sample.
 As a consequence, the single-species prioritisations generated using amount targets and without opportunity cost data were often able to achieve high representation due to the fact they were essentially random selection (Figure 4a).
 To a lesser extent, this is also true for the single-species prioritisations generated with amount targets and consideration to opportunity cost (Figure 4c), since the costs were fairly homogenous among planning units with exception to a few planning units lowland areas.
-
-All the multi-species prioritisations secured a representative sample of most of the species' intra-specific adaptive genetic variation (Figures 4e and 4g).
-Unlike the single-species prioritisations, the multi-species prioritisations generated with only amount-targets did not secure a representative sample of adaptive variation due to stochastic processes.
-Since these prioritisations were forced to secure at least `r raptr.params.LST[[MODE]]$scenario.analysis$amount.target*100` % of land occupied each species, these prioritisations were forced to secured planning units a range of environmental conditions due to the presence of habitat specialists in the data set (eg.
-_Cerastium uniflorum_ and _Androsace obtusifolia_).
 Overall--contrary to our original hypotheses--it seems that conventional reserve selection methods may secure a representative sample of species' adaptive genetic variation if a comprehensive set of species is used.
@@
 
-We chose this data set for several reasons.
-First, to address our aims, we required genetic data for all species to be collected using a standardised sampling scheme, and this data set used a grid to collect samples.
-Second, because conservation planning exercises typically consider tens to hundreds of species with different evolutionary histories and life histories [eg.
-@rXXX], we needed a data set that contains a comparable set of species so that our results will be relevant to planners.
-Third, we required the study area to contain broad-scale environmental variation to test the effectiveness of environmental variables, and this data set spans across large attitudinal gradients.
+This data set has previously been used to address a range of different questions.
+Many studies have used this data set to explore patterns of adaptive [@r484; @r451; @r475; @r455; @r485] and neutral genetic variation [@r478; @r451; @r486].
+However, to our knowledge, only a single study has used this data set to explore the effectiveness of a surrogate in the context of conservation planning [@r486].
+We chose this data set for several reasons: (1) we required genetic data for all species to be collected using a standardised sampling scheme, and this data set used a grid to collect samples, (2) conservation planning exercises typically consider tens to hundreds of species with different evolutionary histories and life histories [reviewed in @r255], and we needed a data set containing a comparable set of species, and (3) we required the study area to contain broad-scale environmental variation to test the effectiveness of environmental variables, and this data set spans across large attitudinal gradients.
 Although this data set contains AFLPs which not ideal as genomic representatives (cf.
@@
 Thus this data set was well suited for our study.
-
-This data set has previously been used to address a range of different questions.
-Many studies have used this data set to explore patterns of adaptive [@r484; @r451; @r475; @r455; @r485] and neutral genetic variation [@r478; @r451; @r486].
-However, to our knowledge, only a single study has used this data set to explore the effectiveness of a surrogate in the context of conservation planning.
-Taberlet _et al._ [-@r486] investigated whether prioritisations generated using the spatial distribution of a comprehensive set species (> 300) could adequately secure the intra-specific genetic variation for 27 species.
-They found that the prioritisation based on species distributions was substantially different to the prioritisation based on genomic data.
-Their results suggest that prioritisations based on species distribution data alone will be unable to secure intra-specific genetic variation.
-However, whether or not environmental and geographic data can result in more prioritisations remains untested.
 

### 2016-09-20 code/rmarkdown/supporting_information.Rmd
+ upgrade article deps to 3.3.1
+ assorted fixes to manuscript * rewrote some parts of article * fixed figure sizing in si

@@
 					strip.background = element_rect(fill='grey20'),
-					strip.text = element_text(color='white', size=5)
+					strip.text = element_text(color='white', size=11)
 				) +
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, include=FALSE}
 plot.spp.nmds <- function(i) {
@@
 				panel.border=element_rect(color='black', fill=NA, size=1),
-				strip.background = element_rect(fill='grey20'),
-				strip.text = element_text(color='white', size=20)
+				strip.background = element_rect(fill='grey20'), strip.text = element_text(color='white', size=20)
 			) +
@@
 				guide=guide_colorbar(ticks=element_line(color='black'),
-					border=element_line(color='black'), barheight=unit(7, 'cm'))
+					border=element_line(color='black'), barheight=unit(3.0, 'cm'))
 			) + facet_wrap(~d)
@@
 	# exports
-	return((((n.plots-1) %/% 3)+1) * 2.5)
+	if (n.plots >= 3) return((((n.plots-1) %/% 3)+1) * 2)
+	return(1.85)
 }
@@
 	# exports
-	if (n.plots > 3) n.plots <- 3
-	return((10/3) * n.plots)
+	if (n.plots >= 3) return(10)
+	return(4 * n.plots)
 }
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(1), fig.height=calc.fig.height(1), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[1]),'}.
+```{r, echo=FALSE, fig.width=calc.fig.width(1), fig.height=calc.fig.height(1), fig.width=calc.fig.width(1), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[1]),'}.
 Each panel shows a different dimension of the variation as indicated by its title.
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(2), fig.height=calc.fig.height(2), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[2]),'}.
-See Figure S6 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(2), fig.height=calc.fig.height(2), fig.width=calc.fig.width(2), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[2]),'}.
+See Figure S29 caption for conventions.')}
 if (is.speciesPlottable(2)) plot.spp.nmds(2)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(3), fig.height=calc.fig.height(3), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[3]),'}.
-See Figure S6 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(3), fig.height=calc.fig.height(3), fig.width=calc.fig.width(3), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[3]),'}.
+See Figure S29 caption for conventions.')}
 if (is.speciesPlottable(3)) plot.spp.nmds(3)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(4), fig.height=calc.fig.height(4), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[4]),'}.
-See Figure S6 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(4), fig.height=calc.fig.height(4), fig.width=calc.fig.width(4), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[4]),'}.
+See Figure S29 caption for conventions.')}
 if (is.speciesPlottable(4)) plot.spp.nmds(4)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(5), fig.height=calc.fig.height(5), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[5]),'}.
-See Figure S6 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(5), fig.height=calc.fig.height(5), fig.width=calc.fig.width(5), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[5]),'}.
+See Figure S29 caption for conventions.')}
 if (is.speciesPlottable(5)) plot.spp.nmds(5)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(6), fig.height=calc.fig.height(6), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[6]),'}.
-See Figure S6 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(6), fig.height=calc.fig.height(6), fig.width=calc.fig.width(6), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[6]),'}.
+See Figure S29 caption for conventions.')}
 if (is.speciesPlottable(6)) plot.spp.nmds(6)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(7), fig.height=calc.fig.height(7), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[7]),'}.
-See Figure S6 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(7), fig.height=calc.fig.height(7), fig.width=calc.fig.width(7), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[7]),'}.
+See Figure S29 caption for conventions.')}
 if (is.speciesPlottable(7)) plot.spp.nmds(7)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(8), fig.height=calc.fig.height(8), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[8]),'}.
-See Figure S6 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(8), fig.height=calc.fig.height(8), fig.width=calc.fig.width(8), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[8]),'}.
+See Figure S29 caption for conventions.')}
 if (is.speciesPlottable(8)) plot.spp.nmds(8)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(9), fig.height=calc.fig.height(9), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[9]),'}.
-See Figure S6 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(9), fig.height=calc.fig.height(9), fig.width=calc.fig.width(9), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[9]),'}.
+See Figure S29 caption for conventions.')}
 if (is.speciesPlottable(9)) plot.spp.nmds(9)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(10), fig.height=calc.fig.height(10), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[10]),'}.
-See Figure S6 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(10), fig.height=calc.fig.height(10), fig.width=calc.fig.width(10), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[10]),'}.
+See Figure S29 caption for conventions.')}
 if (is.speciesPlottable(10)) plot.spp.nmds(10)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(11), fig.height=calc.fig.height(11), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[11]),'}.
-See Figure S6 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(11), fig.height=calc.fig.height(11), fig.width=calc.fig.width(11), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[11]),'}.
+See Figure S29 caption for conventions.')}
 if (is.speciesPlottable(11)) plot.spp.nmds(11)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(12), fig.height=calc.fig.height(12), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[12]),'}.
-See Figure S6 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(12), fig.height=calc.fig.height(12), fig.width=calc.fig.width(12), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[12]),'}.
+See Figure S29 caption for conventions.')}
 if (is.speciesPlottable(12)) plot.spp.nmds(12)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(13), fig.height=calc.fig.height(13), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[13]),'}.
-See Figure S6 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(13), fig.height=calc.fig.height(13), fig.width=calc.fig.width(13), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[13]),'}.
+See Figure S29 caption for conventions.')}
 if (is.speciesPlottable(13)) plot.spp.nmds(13)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(14), fig.height=calc.fig.height(14), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[14]),'}.
-See Figure S6 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(14), fig.height=calc.fig.height(14), fig.width=calc.fig.width(14), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[14]),'}.
+See Figure S29 caption for conventions.')}
 if (is.speciesPlottable(14)) plot.spp.nmds(14)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(15), fig.height=calc.fig.height(15), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[15]),'}.
-See Figure S6 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(15), fig.height=calc.fig.height(15), fig.width=calc.fig.width(15), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[15]),'}.
+See Figure S29 caption for conventions.')}
 if (is.speciesPlottable(15)) plot.spp.nmds(15)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(16), fig.height=calc.fig.height(16), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[16]),'}.
-See Figure S6 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(16), fig.height=calc.fig.height(16), fig.width=calc.fig.width(16), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[16]),'}.
+See Figure S29 caption for conventions.')}
 if (is.speciesPlottable(16)) plot.spp.nmds(16)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(17), fig.height=calc.fig.height(17), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[17]),'}.
-See Figure S6 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(17), fig.height=calc.fig.height(17), fig.width=calc.fig.width(17), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[17]),'}.
+See Figure S29 caption for conventions.')}
 if (is.speciesPlottable(17)) plot.spp.nmds(17)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(18), fig.height=calc.fig.height(18), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[18]),'}.
-See Figure S6 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(18), fig.height=calc.fig.height(18), fig.width=calc.fig.width(18), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[18]),'}.
+See Figure S29 caption for conventions.')}
 if (is.speciesPlottable(18)) plot.spp.nmds(18)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(19), fig.height=calc.fig.height(19), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[19]),'}.
-See Figure S6 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(19), fig.height=calc.fig.height(19), fig.width=calc.fig.width(19), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[19]),'}.
+See Figure S29 caption for conventions.')}
 if (is.speciesPlottable(19)) plot.spp.nmds(19)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(20), fig.height=calc.fig.height(20), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[20]),'}.
-See Figure S6 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(20), fig.height=calc.fig.height(20), fig.width=calc.fig.width(20), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[20]),'}.
+See Figure S29 caption for conventions.')}
 if (is.speciesPlottable(20)) plot.spp.nmds(20)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(21), fig.height=calc.fig.height(21), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[21]),'}.
-See Figure S6 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(21), fig.height=calc.fig.height(21), fig.width=calc.fig.width(21), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[21]),'}.
+See Figure S29 caption for conventions.')}
 if (is.speciesPlottable(21)) plot.spp.nmds(21)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(22), fig.height=calc.fig.height(22), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[22]),'}.
-See Figure S6 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(22), fig.height=calc.fig.height(22), fig.width=calc.fig.width(22), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[22]),'}.
+See Figure S29 caption for conventions.')}
 if (is.speciesPlottable(22)) plot.spp.nmds(22)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(23), fig.height=calc.fig.height(23), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[23]),'}.
-See Figure S6 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(23), fig.height=calc.fig.height(23), fig.width=calc.fig.width(23), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[23]),'}.
+See Figure S29 caption for conventions.')}
 if (is.speciesPlottable(23)) plot.spp.nmds(23)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(24), fig.height=calc.fig.height(24), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[24]),'}.
-See Figure S6 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(24), fig.height=calc.fig.height(24), fig.width=calc.fig.width(24), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[24]),'}.
+See Figure S29 caption for conventions.')}
 if (is.speciesPlottable(24)) plot.spp.nmds(24)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(25), fig.height=calc.fig.height(25), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[25]),'}.
-See Figure S6 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(25), fig.height=calc.fig.height(25), fig.width=calc.fig.width(25), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[25]),'}.
+See Figure S29 caption for conventions.')}
 if (is.speciesPlottable(25)) plot.spp.nmds(25)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(26), fig.height=calc.fig.height(26), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[26]),'}.
-See Figure S6 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(26), fig.height=calc.fig.height(26), fig.width=calc.fig.width(26), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[26]),'}.
+See Figure S29 caption for conventions.')}
 if (is.speciesPlottable(26)) plot.spp.nmds(26)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(27), fig.height=calc.fig.height(27), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[27]),'}.
-See Figure S6 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(27), fig.height=calc.fig.height(27), fig.width=calc.fig.width(27), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[27]),'}.
+See Figure S29 caption for conventions.')}
 if (is.speciesPlottable(27)) plot.spp.nmds(27)
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=(2.5*4), fig.cap='The relationship between the proportion of adaptive genetic variation secured in a prioritizaiton and the proportion of environmental variation it also secures.
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=(2.5*4), fig.cap='The relationship between the proportion of adaptive genetic variation secured in a prioritization and the proportion of environmental variation it also secures.
 Points represent values associated with randomly generated prioritizations.
@@
 		ylab('Adaptive genetic variation secured (%)') +
-		geom_point(fill='transparent') +
+		geom_hex() +
 		suppressWarnings(stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
 			se=FALSE, fill='blue', alpha=0.8)) +
-		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=8, color='darkred') +
+		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=5, color='red') +
 		facet_wrap(~ Species, ncol=4)
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=calc.fig.height3(n_distinct(geo.surrogacy.DF$Species)), fig.cap='The relationship between the proportion of neutral genetic variation secured in a prioritizaiton and the proportion of geographic variation it also secures.
-Conventions are detailed in Figure S55.'}
+Conventions are detailed in Figure S56.'}
 # format data
@@
 		ylab('Neutral genetic variation secured (%)') +
-		geom_point(fill='transparent') +
+		geom_hex() +
 		suppressWarnings(stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
 			se=FALSE, fill='blue', alpha=0.8)) +
-		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=8, color='darkred') +
+		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=5, color='red') +
 		facet_wrap(~ Species, ncol=4)
+)
+```
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=(5*0.9), fig.height=(6.5*0.8), out.width='3.42in', height='3.9in', fig.cap='Planning unit selection frequencies.
+Squares represent planning units, and their colors indicate their selection frequency.
+Each column of panels shows solutions generated using different targets.
+Each row of panels shows solutions generated under different scenarios.
+Note that the selection frequencies for the single-species prioritizations are based on their frequency among prioritizations generated for each species.
+The selection frequencies for the multi-species prioritizations are each based on a single optimal pioritization for all species and so are binary.'}
+# prepare data for plotting
+grid.sub.FPLY <- grid.sub.PLY
+for (i in seq_along(single.spp.prioritisations[[1]]))
+	grid.sub.FPLY@data[[paste0('Single-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations, function(x){colMeans(x[[i]]@results@selections)}))
+for (i in seq_along(single.spp.prioritisations.with.cost[[1]]))
+	grid.sub.FPLY@data[[paste0('Single-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations.with.cost, function(x){colMeans(x[[i]]@results@selections)}))
+for (i in seq_along(multi.spp.prioritisations))
+	grid.sub.FPLY@data[[paste0('Multi-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations[[i]]@results@selections)
+for (i in seq_along(multi.spp.prioritisations.with.cost))
+	grid.sub.FPLY@data[[paste0('Multi-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations.with.cost[[i]]@results@selections)
+grid.sub.FPLY <- spFortify(grid.sub.FPLY)
+grid.sub.FPLY <- grid.sub.FPLY %>% gather(
+		Combination,
+		`Selection.Frequency`,
+		`Single-species (equal costs).Amount`:`Multi-species (opportunity costs).Genetic`
+	) %>% mutate(
+		Context=sapply(strsplit(as.character(Combination), '.', fixed=TRUE), `[[`, 1),
+		Prioritization=sapply(strsplit(as.character(Combination), '.', fixed=TRUE), `[[`, 2)
+	) %>% mutate(
+		Prioritization=revalue(Prioritization, c('Amount'='Amount targets', 'Surrogate'='Amount & surrogate\ntargets', 'Genetic'='Amount & genetic\ntargets')),
+		Prioritization=factor(as.character(Prioritization), levels=c('Amount targets', 'Amount & surrogate\ntargets', 'Amount & genetic\ntargets')),
+		Context=factor(
+			gsub(' (', '\n(', Context, fixed=TRUE),
+			levels=c("Single-species\n(equal costs)", "Single-species\n(opportunity costs)",
+				"Multi-species\n(equal costs)", "Multi-species\n(opportunity costs)"))
+	)
+
+# make maps
+ggplot() +
+	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
+		fill='grey85', color='grey70') +
+	geom_polygon(data=grid.sub.FPLY, aes_string(x='long', y='lat', 
+		group='group', fill='Selection.Frequency'),
+		alpha=0.8, color=NA) +
+	theme_classic() +
+	theme(axis.ticks=element_blank(), axis.text=element_blank(),
+		plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(),
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white', size=8),
+		legend.position='bottom', legend.key.width=unit(1.8, 'cm'),
+		legend.title=element_text(vjust=-5),
+		legend.margin=unit(c(0.1), 'cm'),
+		axis.title=element_blank()) +
+	coord_cartesian(
+		xlim=buffered.range(grid.sub.FPLY$long, 0),
+		ylim=buffered.range(grid.sub.FPLY$lat, 0)
+	) +
+	facet_grid(Context ~ Prioritization) +
+	scale_fill_gradient(name='Selection Frequency (%)',
+		guide=guide_colorbar(title.position='bottom', title.hjust=0.5,
+			ticks=element_line(color='black'), border=element_line(color='black'))
 	)

### 2016-09-28 code/rmarkdown/article.Rmd
+ incorporated rich's comments on the ms

@@
 Protected areas spearhead conservation efforts [@r440].
-These places buffer species from anthropogenic impacts and provide places for species to persist.
+They buffer species from anthropogenic impacts, providing places for species to persist.
 Since the resources available for conservation are limited, protected areas need to be sited in places that fulfill conservation objectives for minimal cost [@r255].
-To achieve this, conservation planning exercises often generate plans for entire networks of protected areas (prioritisations) to allocate resources cost-effectively.
-By generating plans for multiple reserves simultaneously, protected areas can preserve broad-scale biodiversity processes.
+To achieve this, conservation planning exercises often generate plans for entire networks of protected areas (prioritisations) to allocate resources cost-effectively and preserve broad-scale biodiversity processes.
+Intra-specific genetic variation is fundamental to long-term species persistence [@r165], and as a consequence, there has been increasing interest in designing prioritisations that capture and cultivate this aspect of biodiversity [@r67; @r3; @r61; @r499].
 
-Intra-specific genetic variation is important for long-term species persistence [@r165].
-As a consequence, there has been increasing interest in designing prioritisations that preserve and cultivate this aspect of biodiversity [@r67; @r3; @r61; @r499].
-Broadly speaking, although the strength of natural selection is a continuous force, genetic variation can be classified as either adaptive or neutral [reviewed in @r480].
+Broadly speaking, though the strength of natural selection is a continuous force, genetic variation can be classified as either adaptive or neutral [reviewed in @r480].
 Adaptive genetic variation is associated with loci that significantly affect fitness.
@@
 @r500].
-By preserving existing patterns of adaptive variation, protected areas can ensure that populations with particularly beneficial adaptations are not lost [@r3; @r43].
-In contrast, neutral genetic variation is associated with loci that do not significantly affect fitness.
-Neutral variation often reflects the evolutionary history of different populations.
-By preserving existing patterns of neutral variation, protected areas can safeguard against the adverse effects of low genetic diversity [eg.
-inbreeding depression; @r67].
-Thus protected area networks should aim to preserve both adaptive and neutral patterns of genetic variation for species of conservation interest [@r67; @r3; @r43].
-Despite this, genetic data are not often used to inform conservation planning exercises because such data require considerable investment to obtain and analyze [@r61; but see @r5 and @r157 for recent efforts].
-Thus prioritisations may fail to sample species' intra-specific genetic variation, and ultimately fail to deliver long-term conservation outcomes.
+If they represent the full range of adaptive variation, protected areas can help ensure that populations with particularly beneficial adaptations are not lost [@r3; @r43].
+In contrast, neutral genetic variation is associated with loci that do not significantly affect fitness, but instead reflects the evolutionary history of different populations.
+By representing the full range of neutral variation, protected areas can safeguard against the adverse effects of low genetic diversity [@rXXX].
+Thus optimally sited protected area networks would aim to represent both adaptive and neutral patterns of genetic variation for species of conservation interest [@r67; @r3; @r43].
 
-Recently, the use of surrogate data has been proposed to generate prioritisations that adequately preserve intra-specific genetic variation--without needing to utilise genetic data to achieve this.
+Recently, the use of surrogate data has been proposed to generate prioritisations that adequately preserve intra-specific genetic variation--without needing to directly utilise genetic data to achieve this.
 Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been proposed to guide reserve selection as surrogates for adaptive genetic variation [eg.
 environmental conditions; @r202; @r247; @r2].
-By preserving a more representative sample of environmental conditions across the species' ranges, prioritisations could be generated that preserve more adaptive variation.
-However, to our knowledge, the effectiveness of this approach remains unverified.
-On the other hand, neutral genetic variation arises from a reduction in gene flow between individuals.
-Over the last few years, the field of phylogeography has predominantly focused on describing neutral genetic diversity, and so the landscape factors that affect neutral variation are relatively well understood [@r501].
-Potential surrogates for neutral variation have been based on variables that predict the levels of connectivity between planning units.
-For instance, the isolation by distance theory [IBD; @r259] predicts that populations further apart will have less connectivity, exchange less genetic material, and in turn have less genetic variation in common.
-Based on this theory, geographic distance-based surrogates have been found to improve prioritisations in an island network [@r200].
+By preserving a more representative sample of environmental conditions across the species' ranges, prioritisations could be generated that also capture more adaptive variation.
+However, the effectiveness of environmental surrogates in representing adaptive genetic variation remains unverified.
+Neutral genetic variation arises from a reduction in gene flow between individuals.
+Over the last few years, the field of phylogeography has predominantly focused on describing neutral genetic diversity, and so the landscape factors that affect neutral variation are relatively well understood [@r501], and potential surrogates for neutral variation have been based on variables that predict the levels of connectivity between planning units.
+For instance, isolation by distance [IBD; @r259] predicts that populations further apart will be more weakly connected, exchange less genetic material, and in turn have less genetic variation in common.
+Based on this idea, geographic surrogates based on distance have been found to improve conservation prioritisations in an island network [@r200].
 However this remains untested in spatially contiguous systems where connectivity is complicated by additional factors [eg.
-terrain or habitat type; @r474; @r497]--the very systems commonly featured in real-world conservation planning exercises [eg.
+terrain or habitat type; @r474; @r497]--the typical situations in most conservation planning exercises [eg.
 @r197].
-Before conservation planners can reliably use surrogates for genetic data in conservation planning, the effectiveness of these surrogates need to be verified.
-Otherwise, conservation planners may waste precious resources on protecting places that provide little benefit to biodiversity.
 
-Here we investigated the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation in the context of conservation planning.
-To achieve this goal, we used distribution and genomic data for 27 alpine plants collected by the international IntraBioDiv project in the European Alps [@r451].
+Here we investigate the effectiveness of environmental and geographic surrogates for capturing adaptive and neutral genetic variation respectively in the context of conservation planning.
+We used distribution and genomic data for 27 alpine plants collected by the international IntraBioDiv project in the European Alps [@r451].
 We show that for most species the spatial patterns of adaptive and neutral genetic variation correlated with environmental and geographic variables.
-We also show that for most species the level of correlation is high enough to be operational for conservation planning.
-Finally, we demonstrate the using these surrogates results in prioritisations that secure more genetic variation.
+We also show that for most species the level of association is strong enough to be operational for conservation planning.
+Finally, we demonstrate the using these surrogates results in prioritisations that secure more genetic variation than expected by chance.
 
@@
 We detected putatively adaptive genetic variation in `r as.character(as.english(length(env.surrogacy.LST)))` of the 27 plant species.
-We used two outlier detection methods to detect loci under selection, and they returned relatively consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD).
+We used two outlier detection methods to detect loci under selection, and they returned reasonably consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species assigned the same classification $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD).
 Of the species that were associated with adaptive genetic variation, only a small proportion of loci were classified as being adaptive (mean `r round(mean(spp.outlier.loci.props, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.loci.props, na.rm=TRUE),2)` SD).
 After identifying the loci showing strong signals of selection, we used non-metric multi-dimensional scaling (NMDS) analyses to identify the main gradients of adaptive (if detected) and neutral variation for each species.
-Generally, only a couple of continuous dimensions were needed to sufficiently describe their patterns of adaptive (all $K = `r unique(na.omit(adaptive.k))`$; Table S2) and neutral genetic variation (minimum $K = `r min(neutral.k)`$; mean $K = `r round(mean(neutral.k),2)`$; maximum $K = `r max(neutral.k)`$; Table S2).
+Generally, only a small number of continuous dimensions was needed to sufficiently describe their patterns of adaptive (all $K = `r unique(na.omit(adaptive.k))`$; Table S2) and neutral genetic variation (minimum $K = `r min(neutral.k)`$; mean $K = `r round(mean(neutral.k),2)`$; maximum $K = `r max(neutral.k)`$; Table S2).
 The resulting ordinations were used to construct an adaptive (if detected) and neutral genetic space for each species.
-The spatial distribution of these genetic spaces (Figures S29--S54) generally showed patterns of spatial auto-correlation (eg.
-_Cerastium uniflorum_, Figure S34; _Dryas octopetala_, Figures S36)--a good sign that environmental and geographic variables may prove effective surrogates.
+The spatial distribution of these genetic spaces (Figures S29--S54) generally showed substiantial spatial auto-correlation (eg.
+_Cerastium uniflorum_, Figure S34; _Dryas octopetala_, Figures S36), suggesting that the environmental and geographic variables may prove to be effective surrogates.
 
@@
 For each species, we constructed matrices expressing the differences between planning units as a function of their environmental characteristics and geographic position, and the adaptive (if detected) and neutral genetic characteristics between individuals.
-We found the spatial patterns of environmental variation to be significantly correlated with those of adaptive genetic variation for `r as.character(as.english(sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)))` species of the `r as.character(as.english(sum(!is.na(correlation.model.summary.DF$Environmental.corr.P))))` species associated with adaptive variation (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models).
-We also found the geographic distance between planning units to significantly correlate with the spatial patterns of neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` species of the total `r sum(!is.na(correlation.model.summary.DF$Geographic.corr.P))` (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models).
+The spatial patterns of environmental variation was significantly correlated with that of adaptive genetic variation for `r as.character(as.english(sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)))` of the `r as.character(as.english(sum(!is.na(correlation.model.summary.DF$Environmental.corr.P))))` species associated with adaptive variation (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models).
+Geographic distance between planning units also correlated significantly with the spatial pattern of neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` of the `r sum(!is.na(correlation.model.summary.DF$Geographic.corr.P))` species (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models).
 Thus, for most species, planning units that contained different environmental conditions contained individuals with different adaptive genetic characteristics, and planning units that were further apart tended to contain individuals with different neutral genetic characteristics.
-After verifying these patterns, our next step was to determine how effective these variables were as surrogates for conservation planning.
+After verifying these patterns, our next step was to determine whether these variables are effective surrogates for conservation planning.
 ```{r, echo=FALSE, results='asis'}
@@
 <!-- surrogacy analysis -->
-We found the environmental and geographic variables to be moderately effective predictors for genetic variation when assessing the performance of prioritisations (Figure 1).
-For each species, we randomly generated prioritisations, and calculated the proprotion of environmental, geographic, adaptive (if present), and neutral genetic variation they sampled.
-We found the relationship between the proportion of adaptive genetic variation and environmental variation secured in a given prioritisation to depend on the species for which the prioritisation was generated (species $\times$ proportion of environmental variation interaction term; ${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$).
-Similarly, we found the relationship between the proportion of neutral genetic variation and geographic variation secured in a prioritisation to depend on the species for which the prioritisation was generated (${\chi^2}_{`r round(geo.surrogacy.AOV$df[[4]],2)`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`).
-For all species, we found these relationship to be significantly different to zero (environmental: minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; geographic: minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3).
+For each species, we generated a suite of `r raptr.params.LST[[MODE]]$surrogacy.analysis$random.replicates*2` random prioritisations, and calcuated the proportion of environmental variation and adaptive (if present) genetic variation they sampled.
+We then repeated this process, and calculated the proportion of geographic variation and neutral genetic variation they sampled.
+Environmental and geographic variables were moderately effective predictors for genetic variation for the genetic variation represented by the randomly generated prioritisations (Figure 1).
+The relationship between the proportion of genetic variation and surrogate variation secured in the prioritisations varied among species for environmental surrogates of adaptive genetic variation (species $\times$ proportion of environmental variation interaction term; ${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$), and geographic surrogates of neutral variation (${\chi^2}_{`r round(geo.surrogacy.AOV$df[[4]],2)`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`).
+_Post-hoc_ analyses showed that the relationships for all species were significantly different to zero (environmental: minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; geographic: minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3).
 After establishing that the surrogate variables predicted how well prioritisations secured genetic variation, our final step was to determine if prioritisations were more effective when generated using surrogate-based targets.
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=6, fig.height=4.5, out.width='3.42in', out.height='2.5in', fig.cap='The relationship between surrogates and genetic variation secured in prioritizations.
-Panel (a) shows the relationship between the proportion of intra-specific adaptive genetic variation and environmental variation secured in a prioritization.
-Similarly, panel (b) shows the relationship between the proportion of intra-specific neutral variaiton and geograpgic variation secured in a prioritization.
-Panels (c) and (d) show the distribution of $R^2$ values for these models.'}
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=6, fig.height=4.5, out.width='3.42in', out.height='2.5in', fig.cap='Adaptive (a) and neutral (b) genetic variation represented in a suite of randomly generated prioritisations as a function of the environmental and geographic surrogates they also secured (respectively).
+Panels (c) and (d) show the distribution of $R^2$ values for models using each surrogate to explain genetic variation.'}
 # make figure
@@
 		axis.line.y=element_line(color='black')) +
-	coord_cartesian(xlim=c(0, 1))
+	coord_cartesian(xlim=c(0, 1)) +
+	scale_x_continuous(limits=c(0,1), labels=function(x) {x*100}) +
+	scale_y_continuous(limits=c(0,1), labels=function(x) {x*100})
 for (i in seq_along(env.surrogacy.LST))
@@
 		axis.line.y=element_line(color='black', colour='transparent')) +
-	coord_cartesian(xlim=c(0, 1))
+	coord_cartesian(xlim=c(0, 1)) +
+	scale_x_continuous(limits=c(0,1), labels=function(x) {x*100}) +
+	scale_y_continuous(limits=c(0,1), labels=function(x) {x*100})
 for (i in seq_along(geo.surrogacy.LST))
@@
 <!-- scenario analysis -->
-To determine if the addition of environmental and geographic targets could actually improve the effectiveness of prioritisations, we generated prioritisations according to three scenarios using different targets (Figure S58).
-We then computed the proportion of adaptive and neutral genetic variation secured in each prioritisation (Figure 3).
-We found no evidence to suggest that any interaction terms were significant (${\chi^2}_{`r scenario.ANOVA.1.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.1.DF$p.value[[2]])`$).
-This result indicates the trends observed in each of the main effects were consistent across all other levels in the main effects.
+To determine whether using environmental and geographic targets could improve the effectiveness of prioritisations in representing genetic variation, we generated prioritisations using (i) "amount targets" reflecting the traditional approach of representing a certain proportion of each species' goegraphic distribution, (ii) "amount and surrogate targets" where we also targeted representation of the environmental and geographic surrogates as well as a certain proportion of each species' geographic distribution, and (iii) "amount and genetic targets" where we targeted representation of the directly measured genetic variation as well as a certain proportion of each species' geographic distribution (Figure 3).
+To determine if these patterns were robust, for each of the three combinations of the targets, we generated prioritsations under four scenarios: (i) single-species with equal costs, (ii) single species with opportunity costs, (iii) multi-species with equal costs, and (iv) multi-species with opportunity costs.
 
@@
 
-First, the prioritisations tended to secure more adaptive (mean `r round(mean(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` % secured in all prioritisations $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)`) than neutral genetic variation (mean `r round(mean(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)` % secured in all prioritisations $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)`; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.3.DF$p.value[[2]])`).
-This result indicates that regardless of the targets used to generate a prioritisation, or the planning context under which the prioritisation was generated, prioritisations tend to secure more adaptive then neutral genetic variation.
+The proportion of genetic variation secured in a prioritisation depended on the targets used to generate the prioritisation (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$p.value[[2]])`)$.
+Most notably, prioritisations based on the surrogate variables represented significantly more genetic variation than amount-based prioritisations (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` overall genetic variation secured versus `r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$), and were not distinguishable from prioritisations based on measured genetic data (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).
+These results indicate that environmental and geographic surrogates in this case far outperform traditional amount-based conservation planning, and perform as well as a conservation plan based directly on measured genetic data.
 
-Second, the proportion of genetic variation secured in a prioritisation was found to depend on the targets used to generate the prioritisation (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$p.value[[2]])`)$.
-Specifically, the amount-based prioritisations (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)` % overall genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)`) tended to secure less variation than the prioritisations generated using additional surrogate targets (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` % secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$) or genetic targets (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` % secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[2]])`$).
-No evidence was found to suggest that the prioritisations that included additional surrogate-based targets (`r raptr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100` %) or genetic-based targets (`r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %) secured a different proportion of genetic variation ($Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).
-These results suggest that the addition of environmental and geographic surrogates can improve the proportion of intra-specific genetic variation secured in a prioritisation for most species.
+Overall, the prioritisations tended to secure more adaptive then neutral genetic variation (adaptive: mean `r round(mean(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` secured across all prioritisations, neutral: mean `r round(mean(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)` $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)`; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.3.DF$p.value[[2]])`).
+Thus, regardless of the targets used to generate a prioritisation, or the planning context under which the prioritisation was generated, prioritisations tend to secure more adaptive than neutral genetic genetic variation.
 
-Third, we found significant differences in the average proportion of genetic variation secured by prioritisations generated under different scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.2.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.2.DF$p.value[[2]])`).
-Specifically, both the multi-species (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)`) and multi-species with cost (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)`) were both found to secure more genetic variation than the single species scenario (mean `r round(mean(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held),2)`; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[2]])`$; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$ respectively).
-Neither of the multi-species scenarios were found to secure a different proportion of genetic variation ($Z_1 = `r round(scenario.GLHT.1.DF$statistic[[3]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[3]])`$).
+The average proportion of genetic variation secured by prioritisations varied under different scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.2.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.2.DF$p.value[[2]])`).
+Both the multi-species (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)`) and multi-species with cost (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)`) were both found to secure more genetic variation than the single species scenario (mean `r round(mean(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[2]])`$; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$ respectively).
 These results suggest that on average the multi-species prioritisations can secure more genetic variation than single-species prioritisations--regardless of targets used to generate them.
@@
 We aimed to investigate the effectiveness of broad-scale environmental and geographic variables as surrogates for adaptive and neutral genetic variation in conservation planning.
-Our study is the first to address this aim using a diverse set of species and actually measuring the proportion of genetic variation secured in prioritisations.
-Overall, we found that the spatial patterns of environmental and geographic variables correlate with those of adaptive and neutral patterns for most species.
-Furthermore, we found that the addition of environmental and geographic targets to the reserve selection process resulted in prioritisations that secured a greater proportion of adaptive and neutral genetic variation for most species.
+Our study is the first to address this aim using a diverse set of species and calibrated against field measurements of genetic variation.
+The spatial patterns of environmental and geographic surrogates was strongly correlated with the patterns of adaptive and neutral genetic variation for most species, and environmental and geographic targets resulted in prioritisations that secured a greater proportion of adaptive and neutral genetic variation than traditional conservation planning for most species.
+These environmental and geographic surrogates are based on freely available data, and could be calculated for any study region across the world.
 
 <!-- surrogate/genetic dissimilarity matrix correlations -->
-Our analyses showed that differences in the the environmental conditions and geographic position among the planning units tended to correlate with genetic differences among the individuals in the planning units (Table 1).
+Our analyses showed that differences in the environmental conditions and geographic position among the planning units tended to correlate with genetic differences among the individuals in the planning units (Table 1).
 For both surrogates and their corresponding models, the conditional $R^2$ values (environmental vs.
@@
 <!-- surrogate/genetic held correlations -->
-Our randomly generated prioritisations show that environmental and geographic surrgoates are strong enough to be operationalisable (Figure 2).
-Generally, the environmental and geographic surrogates variables were equally effective surrogates for adaptive (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figures 2c and S56) and neutral genetic variation (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figure S57).
-But for a few species, the proportion of geographic variation secured in a prioritisation had a diminished capacity to explain the proportion of neutral variation secured in the prioritisation (eg.
+Our results demonstrate that environmental and geographic surrogates are strong enough to be operational (Figure 2).
+Generally, the environmental and geographic surrogates variables were equally effective surrogates for adaptive (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figures 2c and S56) and neutral genetic variation (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figures 2d and Figure S57), although for a few species, the geograsphic surrogate performed poorly (eg.
 _Gypsophila repens_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gypsophila_repens')$Geographic.r2CU,2)`; _Luzula alpinopilosa_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Luzula_alpinopilosa')$Geographic.r2CU,2)`; _Gentiana nivalis_ $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gentiana_nivalis')$Geographic.r2CU,2)`).
-This bimodal distribution (Figure 2d) suggests that geographic position alone may not be a suitable surrogate for neutral variation in conservation planning for all species.
-This result could be attributed to multiple sources.
-First, geographic distances will be a poor surrogate for genetic lineages in the process of sympatric speciation.
-Second, connectivity between areas can be affected by local topology.
-In particular, mountains can act as barriers to gene flow in alpine landscapes and so nearby populations may be isolated from each other.
-Distances that incorporate information on local topology (eg.
-least-cost distances) may perform better, but generally require species-specific scaling to adequately describe connectivity.
-Third, the species used in this study utilise different dispersal mechanisms [animal-, wind-, or gravity-based dispersal @r451], and some of these species may have dispersal modes that are non-linearly related to geographic distance (eg.
-wind dispersing plants can have high connectivity between nearby and distant areas).
-Overall, however, these species were the exception, and both simple surrogates performed moderately well.
+The bimodal distribution of explained variation (Figure 2d) suggests that geographic position alone may not be a suitable surrogate for neutral variation in conservation planning for all species.
+Indeed, geographic distances will be a poor surrogate for genetic lineages in the process of sympatric speciation, connectivity between areas can be strongly affected by local topology, and some of species have dispersal modes that are non-linearly related to geographic distance (eg.
+wind dispersing plants can have high connectivity between nearby and distant areas; @r451).
+Overall, however, such species were the exception in our analysis, and both simple surrogates performed moderately well.
 
 <!-- suitable targets for surrogates -->
-Generally, we found that the slopes correlating the proportion of geographic variation and neutral variation secured in a prioritisation were higher than those for the proportion of environmental variation and adaptive variation secured in a prioritisation (Figures 2a--2b and S56--S57).
-To secure a representative sample of each species' adaptive genetic variation ($\geq$ `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %), the prioritisation would need to sample on average `r round(mean(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` SD of the environmental conditions found across the planning units occupied by the species.
-On the other hand, to secure a representative sample of the species' neutral genetic variation ($\geq$ `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100`), the prioritisation would need to sample % `r round(mean(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` $\pm$ `r round(sd(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` SD of the variation in geographic position among the planning units occupied by the species.
+In general, adaptive variation accumulated more quickly with environmental representation than neutral variation did with geographic representation (Figures 2a--2b and S56--S57).
+To secure at least `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` % of each species' adaptive genetic variation, prioritisations needed to sample on average `r round(mean(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` SD of the environmental conditions found across the planning units occupied by the species.
+On the other hand, to secure a representative sample of the species' neutral genetic variation ($\geq$ `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %), prioritisations needed to sample `r round(mean(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` SD of the variation in geographic position among the planning units occupied by the species.
 These results suggest that geographic surrogates may require higher targets than environmental surrogates to generate prioritisations that secure representative sample of genetic variation.
 Additionally, these results suggest using surrogate-based targets alone to generate prioritisations may be insufficient to secure a representative sample of species' genetic variation.
-For all species, prioritisations were generated that secured a high proportion of the surrogate variable (> 90 %) but also secured a minuscule proportion of adaptive or neutral genetic variation (< 20 %; Figures S56--S57).
-Thus we recommend that prioritisations be generated using a combination of amount-based targets--as is conventionally used in reserve selection processes--and also surrogate-based targets.
+For all species, it was possible to generate solution that secured a high proportion of the surrogate variable (> 90 %) but only a small proportion of adaptive or neutral genetic variation (< 20 %; Figures S56--S57).
+Thus prioritisations should be generated using a combination of amount-based targets--as is conventionally used in reserve selection processes--and also surrogate-based targets.
 
 <!-- adaptive genetic variation secured in prioritisations-->
-Our results suggest that conservation planners can secure a representative sample of the intra-specific adaptive genetic variation using only amount targets.
-Almost all prioritisations secured a representative sample of species' adaptive genetic variation--regardless of whether surrogate-based or genetic targets were used or not (Figure 4).
-Since the spatial patterns of adaptive genetic variation tended to cluster into a few distinct groups for most species (eg.
+Our results suggest that conservation planners can secure a reasonably representative sample of the intra-specific adaptive genetic variation using only amount-based targets.
+Since the spatial patterns of adaptive genetic variation clustered into a few distinct groups for most species (eg.
 _Carex semipervens_ and _Cerastium uniflorum_; Figures S33 and S34 respectively), a prioritisation would only need to select a planning unit in each cluster to secure a representative sample.
-As a consequence, the single-species prioritisations generated using amount targets and without opportunity cost data were often able to achieve high representation due to the fact they were essentially random selection (Figure 4a).
-To a lesser extent, this is also true for the single-species prioritisations generated with amount targets and consideration to opportunity cost (Figure 4c), since the costs were fairly homogenous among planning units with exception to a few planning units lowland areas.
-Overall--contrary to our original hypotheses--it seems that conventional reserve selection methods may secure a representative sample of species' adaptive genetic variation if a comprehensive set of species is used.
-This outcome provides empirical support for one of the fundamental tenants of systematic conservation planning.
-
-<!-- neutral genetic variation secured in prioritisations-->
-Our analyses showed that addition of environmental and geographic targets were needed to secure a representative sample of neutral genetic variation in most species (Figure 4).
-With exception to the multi-species with cost planning scenario (Figure 4h), prioritisations generated using only amount targets failed to secure a representative proportion of neutral genetic variation ($\geq$ `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %) for most species (Figures 4b, 4f, 4h).
-Whereas, in all scenarios, prioritisations generated using amount and surrogate based targets secured substantially more neutral genetic variation for most species.
-This improvement was associated with only a moderate increase in cost in both scenarios (scenario without opportunity: `r round((score(multi.spp.prioritisations[[2]])/score(multi.spp.prioritisations[[1]]))*100,2)` %; scenario with opportunity cost: `r  round((score(multi.spp.prioritisations.with.cost[[2]])/score(multi.spp.prioritisations.with.cost[[1]]))*100,2)` %)--emphasising the utility of surrogates in conservation.
-Our results suggest that the use of environmental and geographic targets in conservation planning can result in more effective prioritisations.
+As a consequence, the single-species prioritisations generated using amount targets and without opportunity cost data were often able to achieve high representation because they were essentially random selection (Figure 4a).
+To a lesser extent, this is also true for the single-species prioritisations generated with amount targets and consideration to opportunity cost (Figure 4c), since the costs were fairly homogenous among planning units with the exception of a few planning units in lowland areas.
+Overall--and contrary to our original hypotheses--it seems that conventional reserve selection methods may secure a representative sample of species' adaptive genetic variation if a comprehensive set of species is used.
 
 <!--  limitations/ weaknesses -->
-We wish to alert our readers to several limitations associated with our analysis.
-First, the size of the planning units we used (23 $\times$ 25 km) is much larger than typically used in conservation planning scenarios (between 1--10 km$^2$).
+There are several limitations associated with our analysis.
+First, the size of the planning units we used (23 $\times$ 25 km) is much larger than typically used in regional conservation planning scenarios (1--10 km$^2$).
 We used this resolution because the genetic data were collected at this scale.
-Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis because this process would have artificially introduced additional spatial auto-correlation into the data set.
-Second, we used geographic distances as surrogates for neutral genetic variation.
-As previously discussed, connectivity in this landscape may be better described using distances that incorporate information on topology.
-However, such distances require specific-specific scaling to adequate describe their connectivity (eg.
-due to different dispersal modes), and so cannot easily be utilised in large-scale planning exercises.
-Third, we used AFLP data to characterise the main gradients of genetic variation.
-Whilst next-generation sequencing provides higher resolution genetic information than AFLP data [reviewed in @r480], our methodology would still have utilised only the main gradients of the genetic variation in order to generate prioritisations in a feasible amount of time.
-Overall, we are confident that our study is robust.
+Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis by artificially introducing additional spatial auto-correlation.
+Second, we used geographic distances as surrogates for neutral genetic variation, but connectivity in this landscape may be better described using distances that incorporate information on topography.
+However, such distances require specific-specific scaling to adequate describe their connectivity [eg.
+due to different dispersal modes; @r497], and so cannot easily be utilised in large-scale planning exercises.
+Third, we used amplified fragment length polymorphism (AFLP) data to characterise the main gradients of genetic variation.
+Whilst next-generation sequencing provides higher resolution genetic information than AFLP data [reviewed in @r480], no multi-species genomic data set with similar properties currently exists, and our methodology would still have utilised only the main gradients of the genetic variation to generate prioritisations in a feasible amount of time.
 
 <!--  conclusion -->
-The overarching aim of our study was to investigate the effectiveness of surrogates in conservation planning.
-We found that broad-scale environmental and geographic variables are generally effective surrogates for adaptive and neutral genetic variation most species.
-Although prioritisations generated using amount-only targets tended to secure most of the adaptive variation for most species, the addition of surrogate targets to the planning process substantially improved the proportion of intra-specific neutral genetic variation.
-Conservation planners are therefore urged to include these surrogates in conservation planning exercises.
+Broad-scale environmental and geographic variables are generally effective surrogates for representing adaptive and neutral genetic variation for most species in our study system.
+Moreover, data to calculate the surrogate variables are freely available for any location in the world.
+Careful use of such surrogates in conservation planning could vastly improve the chances of long-term biodiversity persistence for relatively additional cost.
 
@@
 This data set contains information for 27 alpine plant species (Figure 2a).
+This data set has previously been used to explore patterns of adaptive [@r484; @r451; @r475; @r455; @r485] and neutral genetic variation [@r478; @r451; @r486], and to show that species richness is not an effective surrogate for genetic diversity in conservation planning [@r486].
 It was collected using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx.
@@
 Thus the data set contains information describing the genomic properties of individuals for each species present in every second grid cell.
+We used this data set because it comprised of comparable genetic data for a range of species with different evolutionary and life histories collected using a standardised sampling scheme, and pertained to a study area with substantial environmental variation to test the effectiveness of environmental variables.
 
@@
 Planning units with a brighter color are inhabited by more species.
-Panel (b) shows the distribution of opportunity cost across the study area (computed as total population density).
-Note that a quantile-based color ramp is used in this panel."}
+Panel (b) shows the distribution of opportunity cost across the study area (computed as total population density) plotted using a quantile-based color ramp."}
 
@@
 
-This data set has previously been used to address a range of different questions.
-Many studies have used this data set to explore patterns of adaptive [@r484; @r451; @r475; @r455; @r485] and neutral genetic variation [@r478; @r451; @r486].
-However, to our knowledge, only a single study has used this data set to explore the effectiveness of a surrogate in the context of conservation planning [@r486].
-We chose this data set for several reasons: (1) we required genetic data for all species to be collected using a standardised sampling scheme, and this data set used a grid to collect samples, (2) conservation planning exercises typically consider tens to hundreds of species with different evolutionary histories and life histories [reviewed in @r255], and we needed a data set containing a comparable set of species, and (3) we required the study area to contain broad-scale environmental variation to test the effectiveness of environmental variables, and this data set spans across large attitudinal gradients.
-Although this data set contains AFLPs which not ideal as genomic representatives (cf.
-single nucleotide polymorphisms; SNPs), no multi-species genomic data set with similar properties currently exists.
-Thus this data set was well suited for our study.
-
 Landscape data {#landscape-data .unnumbered}
@@
 
-We examined the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation (respectively).
-We utilised the sampling grid used by the IntraBioDiv project as planning units to develop prioritisations.
-We omitted cells that were surveyed and not found to contain individuals belonging to any of the study species to reduce computational burden (`r nrow(grid.DF)` grid cells used in analysis).
+We adopted the sampling grid used by the IntraBioDiv project as planning units to develop prioritisations, omitting cells that were surveyed and not found to contain individuals belonging to any of the study species to reduce computational burden (`r nrow(grid.DF)` grid cells used in analysis).
 To examine the potential effects of variation in cost (see below), we calculated the total human population density inside each planning unit [obtained at 1 km$^2$ resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472] and used this to represent acquisition cost (Figure 2b).
+Unless otherwise specified, all computational, spatial, and statistical analysis were conducted in R [version `r paste0(R.version$major, '.', R.version$minor)`; @r192].
 
-To describe the geographic location of each planning unit, we projected the grid into an equi-distant coordinate system (Europe Equidistant Conic; $\texttt{ESRI:102031}$), calculated their centroids, and extracted their two-dimensional coordinates.
+To describe the geographic location of each planning unit, we projected the grid into an equi-distant coordinate system (Europe Equidistant Conic; $\texttt{ESRI:102031}$), calculated the centroid of each grid cell, and extracted their two-dimensional coordinates.
 Thus we constructed two-dimensional geographic space--wherein each planning unit occupies a single point--as a potential surrogate for neutral genetic variation.
@@
 
-To describe the environmental characteristics of each planning unit, we used broad-scale climatic variation.
-We obtained data for 19 bioclimatic variables across the extent of the study area [at $1^{\prime\prime}$ resolution; @r33].
+To describe the environmental characteristics of each planning unit, we obtained data for 19 bioclimatic variables across the extent of the study area [at $30^{\prime\prime}$ resolution; @r33].
 We used this data set because it is commonly to map habitat suitability in conservation planning [@r493].
@@
 The first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components (PCs) cumulatively explained `r round(pca.DF[[4]][surrogate.params.LST[[MODE]]$number.components],1)` % of the variation and were used to generate new layers.
-We calculated the average value for each PC layer in each planning unit, and used these values to characterise climatic variation among the planning units (Figure S2).
-Thus we constructed a three-dimensional environmental space--wherein each planning unit occupies a single point--as a potential surrogate for adaptive variation.
+We calculated the average value for each of the `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components in each planning unit, and used these values to characterise climatic variation among the planning units (Figure S2).
+Thus we constructed a `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))`-dimensional environmental space--wherein each planning unit occupies a single point--as a potential surrogate for adaptive variation.
 We expect planning units further apart from each other in this environmental space to contain individuals with increasingly different patterns in their adaptive loci.
@@
 
-To investigate the effectiveness of our supposed surrogates for adaptive and neutral genetic variation, we first needed to identify which of the sampled loci--if any--were adaptive.
+To investigate the effectiveness of our supposed surrogates for adaptive and neutral genetic variation, we first needed to identify which of the sampled loci are adaptive.
 We used two outlier detection methods to achieve this [an individual- and a population-level method as recommended by @r495].
-The basic premise underpinning such methods is that neutral loci are expected to exhibit a specific level of variation among populations, and loci that deviate from this expectation should be under selection [reviewed in @r480].
-The advantage of these methods--in contrast with environmental association analyses [@r489]--is that they do not use environmental data to identify loci under selection, which would have biased our analysis.
-Loci identified as outliers in both analyses were treated as adaptive and the rest were treated as neutral.
-To avoid false-positives, we omitted loci from both outlier detection methods where the global frequency of the minor allele was $> 0.1$ and assumed they were neutral.
+The basic premise underpinning such methods is that neutral loci are expected to exhibit a specific level of variation among populations, and loci that deviate from this expectation are likely to be under selection [reviewed in @r480].
+The advantage of these methods--in contrast with environmental association analyses [@r489]--is that they do not use environmental data to identify loci under selection, which would have introduced circularity into our analysis.
+Loci identified as outliers in both analyses were treated as adaptive and the remainder were treated as neutral.
+To minimise false-positives, we omitted loci from both outlier detection methods where the global frequency of the minor allele was less than 10 % and assumed they were neutral.
 
@@
 For a given species, each planning unit occupied by the species was associated with a multi-dimensional point in the species' neutral genetic space.
-Planning units that were closer together in this space are occupied by individuals with similar polymorphisms in their neutral loci.
+Planning units that were closer together in this space are occupied by individuals with more similar polymorphisms in their neutral loci.
 Similarly, if the species was associated with adaptive loci, each planning unit occupied by the species was associated with a multi-dimensional point in the species' adaptive genetic space.
@@
 
-Statistical analyses {#statistical-analyses .unnumbered}
+Statistical and computational analyses {#statistical-analyses .unnumbered}
 --------------------------------------------------------
 
-To address the first aim--to determine if the environmental and geographic variables correlate with spatial patterns of adaptive and neutral genetic variation--we computed dissimilarity matrices between the planning units, and tested for correlations between matrices based on surrogate and genetic data.
+To determine if the environmental and geographic variables correlate with spatial patterns of adaptive and neutral genetic variation, we computed dissimilarity matrices between the planning units, and tested for correlations between matrices based on surrogate and genetic data.
 For each species, we computed two dissimilarity matrices between the planning units occupied by the species in terms of their environmental conditions (their coordinates in the environmental space) and geographic position (their coordinates in the geographic space).
@@
 To address the second aim--to determine if the environmental and geographic variables are effective surrogates for adaptive and neutral genetic variation--we generated a collection of single-species prioritisations and computed the proportion of intra-specific genetic variation they secured.
-For each species, we generated `r raptr.params.LST[[MODE]]$surrogacy.analysis$replicates` prioritisations by randomly selecting different combinations of planning units that were occupied by the species.
+For each species, we generated `r raptr.params.LST[[MODE]]$surrogacy.analysis$random.replicates` prioritisations by randomly selecting different combinations of planning units that were occupied by the species.
 We then computed the proportion of adaptive genetic variation and environmental variation secured in each prioritisation.
@@
 
-We fit two full generalised linear models (GLM) with a logit links.
+We fit two full generalised linear models (GLM) with logit links.
 The first model was fit to the proportion of adaptive genetic variation secured in a prioritisation, and contained a continuous predictor variable measuring the proportion of environmental variation also secured in the prioritisation, a categorical variable indicating the species for which the prioritisation was generated, and an interaction term between these predictor variables.
@@
 
-To address the third aim--to determine if the use of surrogate-based targets can actually improve prioritisations--we investigated the effectiveness of surrogate-based targets under three scenarios with increasing levels of complexity.
+To determine whether the use of surrogate-based targets improves prioritisations, we investigated the effectiveness of surrogate-based targets under three scenarios with increasing levels of complexity.
 To assess the benefits of surrogate-based targets under the simplest of circumstances, our first scenario involved generating single-species prioritisations for each species.

### 2016-10-06 code/rmarkdown/article.Rmd
+ bug fixes with incorporating separate pcas for each species

@@
 These layers and the planning units were projected into an equal-area coordinate system (Europe Lambert Conformal Conic; $\texttt{ESRI:102014}$).
-To reduce dimensionality, the 19 bioclimatic variables were clipped to the planning units and subjected to a principal components analysis (PCA; using \texttt{ArcMap} version 10.2.2; Table S1).
-The first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components (PCs) cumulatively explained `r round(pca.DF[[4]][surrogate.params.LST[[MODE]]$number.components],1)` % of the variation and were used to generate new layers.
-We calculated the average value for each of the `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components in each planning unit, and used these values to characterise climatic variation among the planning units (Figure S2).
-Thus we constructed a `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))`-dimensional environmental space--wherein each planning unit occupies a single point--as a potential surrogate for adaptive variation.
+The average value in each planning unit for each of the 19 layers was calculated.
+To reduce dimensionality, for each species, the climatic values associated with the planning units occupied by the species were subject to principal components analysis (PCA; Table S1).
+On average, the first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components (PCs) cumulatively explained `r round(mean(spp.pca.DF[[2]]*100),2)` % $\pm$ `r round(sd(spp.pca.DF[[2]]*100),2)` of the variation.
+We then used the first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components to characterise climatic variation among the planning units for each species (Figure S2).
+Thus we constructed a `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))`-dimensional environmental space for each species--wherein each planning unit occupies a single point--as a potential surrogate for adaptive variation for that species.
 We expect planning units further apart from each other in this environmental space to contain individuals with increasingly different patterns in their adaptive loci.

### 2016-10-06 code/rmarkdown/supporting_information.Rmd
+ bug fixes with incorporating separate pcas for each species

@@
 ```{r, fig.width=15, fig.height=3, fig.cap='Climatic variation.
-Each panel depicts variation based on a different principle component (PC).
+Each panel depicts variation based on a principle component (PC) for a different species.
 Squares represent planning units.
 The color of each planning unit denotes the average PC value of pixels inside it.
-Planning units with more similar colors have more similar climate regimes.', echo=FALSE}
-do.call(
-	grid.arrange,
-		append(
-		llply(grep('^env\\_.*$', names(grid.DF), value=TRUE), function(x) {
-			grid.FPLY$plot_title <- paste0('PC ', substr(x, nchar(x), nchar(x)))
-			ggplot() +
-				geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
-					fill='grey85', color='grey70') +
-				geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
-					group='group', fill=x), color='grey10') +
-				theme_classic() +
-				theme(
-					axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"), axis.title=element_blank(),
-					panel.border=element_rect(color='black', fill=NA, size=1),
-					strip.background = element_rect(fill='grey20'),
-					strip.text = element_text(color='white', size=11)
-				) +
-				coord_cartesian(
-					xlim=buffered.range(grid.FPLY$long, 0),
-					ylim=buffered.range(grid.FPLY$lat, 0)
-				) +
-				scale_fill_gradientn(name='',
-					colors=brewer.pal(9, 'RdYlBu'),
-					guide=guide_colorbar(ticks=element_line(color='black'), border=element_line(color='black'))
-				) +
-				facet_wrap(~ plot_title)
-		}),
-		list(ncol=4)
-	)
-)
+For a given species, the planning units with more similar colors have more similar climate regimes.', echo=FALSE}
+plot(1)
 ```
@@
 ```{r, echo=FALSE, results='asis'}
-## make results table showing Eigen values
+## make table showing pca resuls
 curr.file <- tempfile(fileext='.tex')
-pca.DF[['Eigen Value']] <- c(round(pca.DF[['Eigen Value']][-19],2), pvalString(pca.DF[['Eigen Value']][19], digits=2))
-pca.DF[['Variation explained (%)']] <- c(round(pca.DF[['Variation explained (%)']][1:3],2), pvalString(pca.DF[['Variation explained (%)']][-1:-3], digits=2))
-pca.DF[['Accumulative variation explained (%)']] <- as.character(round(pca.DF[['Accumulative variation explained (%)']],2))
-x <- pca.DF %>%
-	latex(file=curr.file, 
-		rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, col.just=rep('c', ncol(pca.DF)), here=TRUE,
-		colheads=c("\\parbox[t]{2.5cm}{\\centering \\bfseries{Principle Component}}", "\\parbox[t]{2cm}{\\centering \\bfseries{Eigen value}}",
-			"\\parbox[t]{3cm}{\\centering \\bfseries{Variation explained (\\%)}}", "\\parbox[t]{4.5cm}{\\centering \\bfseries{Cumulative variation explained (\\%)}}"),
-		caption=paste0('Summary of principle components analysis (PCA) on bioclimatic variation across the study area.
-The first ',as.character(as.english(surrogate.params.LST[[MODE]]$number.components)),' principle components (PCs) were used for subsequent analysis.')
+x <- spp.pca.DF %>%
+	mutate(species=gsub('_', ' ', species, fixed=TRUE), percent_variation_explained=percent_variation_explained*100) %>%
+	latex(
+		file=curr.file,
+		rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, col.just=rep('c', 2),
+		colheads=c('\\bfseries{Species}', '\\bfseries{Variation explained (\\%)}'), here=TRUE,
+		caption=paste0('Summary of principal component analyses fit to the climatic variation in the species\' ranges.
+Percentages correspond to the percent of variation described by the first ',as.character(as.english(surrogate.params.LST[[MODE]]$number.components)),' principle components.')
 	)
 readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
+	gsub(pattern='\\cline{1-', replacement='\\cline{3-4} \\cline{6-', fixed=TRUE) %>% 
 	gsub(pattern='toprule', replacement='toprule[1pt]', fixed=TRUE) %>%

### 2016-10-06 code/R/analysis/02-surrogate-data.R

@@ -23,7 +23,7 @@ grid.PPLY <- spTransform(grid.PLY, europeEA)
 # sample data as SpatialPoints
 spp.sample.PTS <- SpatialPointsDataFrame(
 	coords=as.matrix(spp.samples.DF[,5:6]),
-	data=spp.s	amples.DF,
+	data=spp.samples.DF,
 	proj4string=wgs1984
 )
 spp.sample.PPTS <- spTransform(spp.sample.PTS, europeEA)
@@ -57,10 +57,17 @@ grid.PPLY$id <- seq_len(nrow(grid.DF))
 
 ## extract climatic data
 # load climatic data
-bioclim.STK <- stack(dir('data/raw/BioClim_variables', '^.*\\.tif$' full.names=TRUE))
+bioclim.STK <- stack(dir('data/raw/BioClim_variables', '^.*\\.tif$', full.names=TRUE))
+studyarea.PLY <- grid.PPLY %>%
+	extent() %>%
+	as('SpatialPolygons') %>%
+	gBuffer(width=20000, byid=FALSE)  %>%
+	`slot<-`(name='proj4string', value=grid.PPLY@proj4string) %>%
+	spTransform(bioclim.STK@crs)
+bioclim.STK <- bioclim.STK %>% crop(studyarea.PLY) %>% mask(studyarea.PLY) %>% projectRaster(crs=grid.PPLY@proj4string, res=1000)
 # extract mean for each cell for each principle component
 extract.DF <- grid.PPLY %>% rasterize(bioclim.STK, field='id') %>% 
-	zonal(x=bioclim.STK) %>%  as.data.frame() %>% select(-1) %>%
+	zonal(x=bioclim.STK, fun='mean') %>%  as.data.frame() %>% select(-1) %>%
 	`names<-`(paste0('bio_d',seq_len(nlayers(bioclim.STK))))
 # merge with grid.DF
 grid.DF <- cbind(grid.DF, extract.DF)
@@ -68,22 +75,23 @@ grid.DF <- cbind(grid.DF, extract.DF)
 spp.pca.DF <- data.frame(species=character(0), percent_variation_explained=numeric(0))
 for (i in seq_along(unique(spp.samples.DF$species))) {
 	# extract bio vars for species
-	curr.sub.DF <- filter(spp.samples.DF, species==unique(spp.samples.DF$species)[i])$cell
[…]

### 2016-10-11 code/rmarkdown/article.Rmd
+ article within page pnas limits

--- 2016-09-28 code/rmarkdown/article.Rmd
@@
 However, evolutionary processes are not often explicitly considered when deciding where to site protected areas because genetic data require considerable resources to obtain and analyse.
-Here we show that environmental and geographic distance variables are effective surrogates for adaptive and neutral intra-specific variation for conservation planning, and that these surrogates can be used to deliver more effective solutions.
-We obtained distribution and genetic data from the international IntraBioDiv project for 27 species collected over the European Alps.
-Samples were collected using a gridded sampling scheme.
-We used these cells as planning units to generate solutions.
-For each species, we identified loci that were potentially under selection using outlier loci methods, and mapped their main gradients of adaptive and neutral genetic variation across the planning units.
-Our first analysis confirms that the the spatial patterns of adaptive and neutral genetic variation correlate with those of environmental and geographic variation (respectively).
-Our second analysis verifies that solutions that secure more adaptive genetic variation also tend to secure more adaptive variation, and similar patterns were found for neutral genetic variation and geographic variables.
-Our third analysis found that solutions generated using additional targets to explicitly secure a representative sample of the environmental and geographic variation across a species range resulted in solutions that secure more neutral genetic variation for most species.
-Additionally, we found that most solutions--even those generated without surrogate or genetic data--tended to secure most of the adaptive genetic variation for most species.
-Overall, our analysis provides a rigorous examination of two potential surrogates for genetic variation.
+Here we show that environmental and geographic distance variables are effective surrogates for adaptive and neutral intra-specific variation for conservation planning.
+We obtained occurrence and genetic data from the IntraBioDiv project for 27 species collected over the European Alps using a gridded sampling scheme.
+For each species, we identified loci that were potentially under selection using outlier loci methods, and mapped their main gradients of adaptive and neutral genetic variation across the grid cells.
+We then used the cells as planning units to generate prioritisations.
+Our first analysis verified that the spatial patterns of environmental and geographic variation correlated with those adaptive and neutral genetic variation (respectively).
+Our second analysis showed that these surrogates can predict the proportion of genetic variation secured in randomly generated solutions.
+Our third analysis found that solutions generated using additional surrogate-based targets secured more neutral genetic variation.
+We discovered that solutions generated using traditional amount-based targets tend to secure most species' adaptive genetic variation.
+Overall, our work provides a rigorous examination of two potential surrogates for genetic variation.
 significance: |
@@
 Since the resources available for conservation are limited, protected areas need to be sited in cost-effective locations.
-However, despite the stated emphasis on biodiversity processes, evolutionary processes are not often used to guide reserve selection.
-This is due to the fact that genetic data is often too costly to obtain and process.
-Here we found that environmental and geographic variables are cost-effective surrogates for genetic data.
-These variable can be used to deliver candidate reserve systems that secure the adaptive and neutral evolutionary processes for a greater number of species with minimal additional cost.
-This research highlights the importance of explicitly considering biodiversity processes during conservation planning exercises.
+However, despite their importance, evolutionary processes are not often used to guide reserve selection.
+This is primarily because genetic data require extensive resources to obtain and process.
+Here we show that freely available environmental and geographic variables can be used as surrogates for genetic data in conservation planning.
 acknowledgements: |
@@
 RAF has an Australian Research Council Future Fellowship.
-This work was supported by the Centre of Excellence for Environmental Decisions (CEED) and the Landscape Ecology and Conservation Group (LEC) at The University of Queensland.
+This work was supported by the Centre of Biodiversity and Conservation Science.
 keywords:
@@
 Protected areas spearhead conservation efforts [@r440].
-They buffer species from anthropogenic impacts, providing places for species to persist.
+They buffer species from anthropogenic impacts; providing places for them to persist.
 Since the resources available for conservation are limited, protected areas need to be sited in places that fulfill conservation objectives for minimal cost [@r255].
@@
 Typically, such "adaptive" loci are detected due to anomalous patterns among individuals [eg.
-@r456; @r487; @r480; @r490], though phenotypic-genetic associations are more likely to identify functionally important genomic regions for individual species [eg.
+@r456; @r487], though phenotypic-genetic associations are more likely to identify functionally important genomic regions for individual species [eg.
 @r500].
-If they represent the full range of adaptive variation, protected areas can help ensure that populations with particularly beneficial adaptations are not lost [@r3; @r43].
+By representing the full range of adaptive variation, protected areas can help ensure that populations with particularly beneficial adaptations are not lost [@r3; @r43].
 In contrast, neutral genetic variation is associated with loci that do not significantly affect fitness, but instead reflects the evolutionary history of different populations.
-By representing the full range of neutral variation, protected areas can safeguard against the adverse effects of low genetic diversity [@rXXX].
-Thus optimally sited protected area networks would aim to represent both adaptive and neutral patterns of genetic variation for species of conservation interest [@r67; @r3; @r43].
+By representing the full range of neutral variation, protected areas can safeguard against the adverse effects of low genetic diversity [@r67].
+Thus optimally sited protected area networks would aim to represent both adaptive and neutral patterns of genetic variation [@r67; @r3; @r43].
 
-Recently, the use of surrogate data has been proposed to generate prioritisations that adequately preserve intra-specific genetic variation--without needing to directly utilise genetic data to achieve this.
+Recently, the use of surrogate data has been proposed to generate prioritisations that sufficiently intra-specific genetic variation--without needing to directly utilise genetic data to achieve this.
 Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been proposed to guide reserve selection as surrogates for adaptive genetic variation [eg.
-environmental conditions; @r202; @r247; @r2].
-By preserving a more representative sample of environmental conditions across the species' ranges, prioritisations could be generated that also capture more adaptive variation.
+@r202; @r2].
+By generating prioritisations that preserve a more representative sample of environmental conditions across species' ranges, these prioritisations could also capture more adaptive variation.
 However, the effectiveness of environmental surrogates in representing adaptive genetic variation remains unverified.
 Neutral genetic variation arises from a reduction in gene flow between individuals.
-Over the last few years, the field of phylogeography has predominantly focused on describing neutral genetic diversity, and so the landscape factors that affect neutral variation are relatively well understood [@r501], and potential surrogates for neutral variation have been based on variables that predict the levels of connectivity between planning units.
-For instance, isolation by distance [IBD; @r259] predicts that populations further apart will be more weakly connected, exchange less genetic material, and in turn have less genetic variation in common.
+Over the last few years, the field of phylogeography has predominantly focused on describing neutral genetic diversity, and so the landscape factors that affect neutral variation are relatively well understood [@r501].
+Potential surrogates for neutral variation have been based on variables that predict the levels of connectivity between planning units.
+For instance, isolation by distance [IBD; @r259] predicts that populations further apart will experience less gene flow, and in turn, share less neutral loci.
 Based on this idea, geographic surrogates based on distance have been found to improve conservation prioritisations in an island network [@r200].
 However this remains untested in spatially contiguous systems where connectivity is complicated by additional factors [eg.
-terrain or habitat type; @r474; @r497]--the typical situations in most conservation planning exercises [eg.
-@r197].
+habitat type @r497]--the typical situations in most conservation planning exercises.
 
@@
 We detected putatively adaptive genetic variation in `r as.character(as.english(length(env.surrogacy.LST)))` of the 27 plant species.
-We used two outlier detection methods to detect loci under selection, and they returned reasonably consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species assigned the same classification $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD).
+We used two outlier detection methods to detect loci under selection, and they returned reasonably consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species assigned the same classification $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD; Table S1).
 Of the species that were associated with adaptive genetic variation, only a small proportion of loci were classified as being adaptive (mean `r round(mean(spp.outlier.loci.props, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.loci.props, na.rm=TRUE),2)` SD).
 After identifying the loci showing strong signals of selection, we used non-metric multi-dimensional scaling (NMDS) analyses to identify the main gradients of adaptive (if detected) and neutral variation for each species.
-Generally, only a small number of continuous dimensions was needed to sufficiently describe their patterns of adaptive (all $K = `r unique(na.omit(adaptive.k))`$; Table S2) and neutral genetic variation (minimum $K = `r min(neutral.k)`$; mean $K = `r round(mean(neutral.k),2)`$; maximum $K = `r max(neutral.k)`$; Table S2).
+Generally, only a small number of continuous dimensions was needed to sufficiently describe their patterns of adaptive (all $K = `r unique(na.omit(adaptive.k))`$; Table S2) and neutral genetic variation ($K = `r min(neutral.k)`$--$`r max(neutral.k)`$; Table S1).
 The resulting ordinations were used to construct an adaptive (if detected) and neutral genetic space for each species.
-The spatial distribution of these genetic spaces (Figures S29--S54) generally showed substiantial spatial auto-correlation (eg.
+The spatial distribution of these genetic spaces (Figures S29--S54) generally showed substantial spatial auto-correlation (eg.
 _Cerastium uniflorum_, Figure S34; _Dryas octopetala_, Figures S36), suggesting that the environmental and geographic variables may prove to be effective surrogates.
@@
 We verified that the spatial patterns in environmental and geographic variation correlated with the broad-scale spatial patterns in adaptive and neutral variation for most species (respectively).
-We constructed dissimilarity matrices between the planning units.
-For each species, we constructed matrices expressing the differences between planning units as a function of their environmental characteristics and geographic position, and the adaptive (if detected) and neutral genetic characteristics between individuals.
-The spatial patterns of environmental variation was significantly correlated with that of adaptive genetic variation for `r as.character(as.english(sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)))` of the `r as.character(as.english(sum(!is.na(correlation.model.summary.DF$Environmental.corr.P))))` species associated with adaptive variation (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models).
-Geographic distance between planning units also correlated significantly with the spatial pattern of neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` of the `r sum(!is.na(correlation.model.summary.DF$Geographic.corr.P))` species (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models).
+For each species,we constructed dissimilarity matrices based on the planning units they were found in.
+These matrices expressed differences between the planning units in terms of their (i) environmental characteristics, (ii) geographic position, and also the (iii) adaptive (if detected), and (iv) neutral genetic characteristics of individuals found inside them.
+By correlating these matrices, the spatial patterns of environmental variation were found to be significantly correlated with that of adaptive genetic variation for `r as.character(as.english(sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)))` of the `r as.character(as.english(sum(!is.na(correlation.model.summary.DF$Environmental.corr.P))))` species associated with adaptive variation (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models; Table S3).
+Similarly, geographic distance between planning units was also correlated significantly with the spatial pattern of neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` of the `r sum(!is.na(correlation.model.summary.DF$Geographic.corr.P))` species (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models; Table S3).
 Thus, for most species, planning units that contained different environmental conditions contained individuals with different adaptive genetic characteristics, and planning units that were further apart tended to contain individuals with different neutral genetic characteristics.
-After verifying these patterns, our next step was to determine whether these variables are effective surrogates for conservation planning.
-```{r, echo=FALSE, results='asis'}
-curr.file <- tempfile(fileext='.tex')
-x <- correlation.model.summary.DF %>%
-	select(-contains('raw')) %>%
-	mutate(
-		Environmental.corr.P=pvalString(Environmental.corr.P),
-		Geographic.corr.P=pvalString(Geographic.corr.P),
-		Geographic.Marginal.R2=pvalString(Geographic.Marginal.R2),
-		Geographic.Conditional.R2=pvalString(Geographic.Conditional.R2),
-		Environmental.Marginal.R2=pvalString(Environmental.Marginal.R2),
-		Environmental.Conditional.R2=pvalString(Environmental.Conditional.R2),
-		blank=rep('', length(Species)),
-		Species=paste0('\\textit{', gsub('\\ ', ' ', Species), '}')) %>%
-	mutate(Environmental.Chisq=sapply(round(Environmental.Chisq,2), as.character)) %>%
-	mutate(Geographic.Chisq=sapply(round(Geographic.Chisq,2), as.character)) %>%
-	select(Species:Environmental.corr.P, Geographic.Marginal.R2:Geographic.corr.P) %>%
-	latex(file=curr.file, 
-		cgroup=c('', 'Environmental', 'Geographic'), n.cgroup=c(1, 4, 4), col.just=rep('c', 9),
-		digits=2, rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, here=TRUE,
-		colheads=c('\\bfseries{Species}', '$MR^2$', '$CR^2$', '$\\chi^2$', '\\textit{P}', '$MR^2$', '$CR^2$', '$\\chi^2$', '\\textit{P}'),
-		caption='Summary of maximum likelihood population-effects (MLPE) models.
-Briefly, for each species associated with potentially adaptive loci, a model was fit to correlate dissimilarities among planning units where the species in terms of the adaptive genetic characteristics of individuals in them with dissimilarities in terms of the environmental conditions inside the planning units (denoted in the environmental column group).
-Similarly, for each species, a model was fit to correlate dissimilarities among the planning units in terms of the neutral genetic characteristics of indiviudals in them with dissimilarities in terms of the geographic position of the planning units (denoted in the geographic column group).
-Data shows the results from comparing each model to its corresponding null model using $\\mathbf{\\chi^2}$ tests (1 degree of freedom), and the $\\mathbf{R^2}$ ($\\mathbf{MR^2}$) and conditional $\\mathbf{R^2}$ ($\\mathbf{CR^2}$) for each model.'
-	)
-readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
-	gsub(pattern='\\cline{1-', replacement='\\cline{2-6} \\cline{8-', fixed=TRUE) %>% 
-	gsub(pattern='{c}{\\textit{P}}\\tabularnewline', replacement='{c}{\\textit{P}}\\T\\B\\tabularnewline', fixed=TRUE) %>% 
-	gsub(pattern='\\toprule', replacement='', fixed=TRUE) %>%
-	gsub(pattern='\\midrule', replacement='\\midrule', fixed=TRUE) %>%
-	gsub(pattern='\\bottomrule', replacement='\\midrule', fixed=TRUE) %>%
-	cat()
-```
+After verifying these patterns were correlated, our next step was to determine whether they were correlated enough to be effective for conservation planning.
 
 <!-- surrogacy analysis -->
-For each species, we generated a suite of `r raptr.params.LST[[MODE]]$surrogacy.analysis$random.replicates*2` random prioritisations, and calcuated the proportion of environmental variation and adaptive (if present) genetic variation they sampled.
+For each species, we generated a suite of `r format(raptr.params.LST[[MODE]]$surrogacy.analysis$random.replicates, big.mark=',', scientific=FALSE)` random prioritisations, and calculated the proportion of environmental variation and adaptive (if present) genetic variation they sampled.
 We then repeated this process, and calculated the proportion of geographic variation and neutral genetic variation they sampled.
-Environmental and geographic variables were moderately effective predictors for genetic variation for the genetic variation represented by the randomly generated prioritisations (Figure 1).
-The relationship between the proportion of genetic variation and surrogate variation secured in the prioritisations varied among species for environmental surrogates of adaptive genetic variation (species $\times$ proportion of environmental variation interaction term; ${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$), and geographic surrogates of neutral variation (${\chi^2}_{`r round(geo.surrogacy.AOV$df[[4]],2)`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`).
-_Post-hoc_ analyses showed that the relationships for all species were significantly different to zero (environmental: minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; geographic: minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3).
-After establishing that the surrogate variables predicted how well prioritisations secured genetic variation, our final step was to determine if prioritisations were more effective when generated using surrogate-based targets.
+The environmental and geographic variables were moderately effective predictors for the genetic variation represented by the randomly generated prioritisations (Figure 1).
+The relationship between the proportion of genetic and surrogate variation secured in the prioritisations varied among the species (species $\times$ proportion of environmental variation interaction term: ${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$; species $\times$ proportion of geographic variation interaction term: ${\chi^2}_{`r round(geo.surrogacy.AOV$df[[4]],2)`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`).
+_Post-hoc_ analyses showed that these relationships were positive for all species (environmental: minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; geographic: minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S4).
+After establishing that the the proportion of surrogate variation secured in a prioritisation also predicted the proportion of genetic variation it secured, our final step was to determine if prioritisations were more effective when generated using targets based on the surrogates.
 
@@
 <!-- scenario analysis -->
-To determine whether using environmental and geographic targets could improve the effectiveness of prioritisations in representing genetic variation, we generated prioritisations using (i) "amount targets" reflecting the traditional approach of representing a certain proportion of each species' goegraphic distribution, (ii) "amount and surrogate targets" where we also targeted representation of the environmental and geographic surrogates as well as a certain proportion of each species' geographic distribution, and (iii) "amount and genetic targets" where we targeted representation of the directly measured genetic variation as well as a certain proportion of each species' geographic distribution (Figure 3).
-To determine if these patterns were robust, for each of the three combinations of the targets, we generated prioritsations under four scenarios: (i) single-species with equal costs, (ii) single species with opportunity costs, (iii) multi-species with equal costs, and (iv) multi-species with opportunity costs.
+To determine whether using environmental and geographic targets could improve the effectiveness of prioritisations in representing genetic variation, we generated prioritisations using (i) "amount targets" reflecting the traditional approach of representing a certain proportion of each species' geographic distribution, (ii) "amount and surrogate targets" where we also targeted representation of the environmental and geographic surrogates as well as a certain proportion of each species' geographic distribution, and (iii) "amount and genetic targets" where we targeted representation of the directly measured genetic variation as well as a certain proportion of each species' geographic distribution (Figure 3).
+To determine if these patterns were robust, for each of the three combinations of the targets, we generated prioritisations under four scenarios: (i) single-species with equal costs, (ii) single species with opportunity costs, (iii) multi-species with equal costs, and (iv) multi-species with opportunity costs.
 
@@
 The proportion of genetic variation secured in a prioritisation depended on the targets used to generate the prioritisation (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$p.value[[2]])`)$.
-Most notably, prioritisations based on the surrogate variables represented significantly more genetic variation than amount-based prioritisations (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` overall genetic variation secured versus `r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$), and were not distinguishable from prioritisations based on measured genetic data (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).
+Most notably, prioritisations based on the surrogate variables represented significantly more genetic variation than amount-based prioritisations (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` SD overall genetic variation secured versus `r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)` SD; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$), and were not distinguishable from prioritisations based on measured genetic data (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` SD; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).
 These results indicate that environmental and geographic surrogates in this case far outperform traditional amount-based conservation planning, and perform as well as a conservation plan based directly on measured genetic data.
@@
 We aimed to investigate the effectiveness of broad-scale environmental and geographic variables as surrogates for adaptive and neutral genetic variation in conservation planning.
-Our study is the first to address this aim using a diverse set of species and calibrated against field measurements of genetic variation.
-The spatial patterns of environmental and geographic surrogates was strongly correlated with the patterns of adaptive and neutral genetic variation for most species, and environmental and geographic targets resulted in prioritisations that secured a greater proportion of adaptive and neutral genetic variation than traditional conservation planning for most species.
+Our study is the first to address this aim using a broad range of species and calibrated against field measurements of genetic variation.
+The spatial patterns of environmental and geographic surrogates were strongly correlated with the patterns of adaptive and neutral genetic variation for most species, and environmental and geographic targets resulted in prioritisations that secured a greater proportion of adaptive and neutral genetic variation than traditional conservation planning for most species.
 These environmental and geographic surrogates are based on freely available data, and could be calculated for any study region across the world.
-
-<!-- surrogate/genetic dissimilarity matrix correlations -->
-Our analyses showed that differences in the environmental conditions and geographic position among the planning units tended to correlate with genetic differences among the individuals in the planning units (Table 1).
-For both surrogates and their corresponding models, the conditional $R^2$ values (environmental vs.
-adaptive genetic variation: mean `r round(mean(correlation.model.summary.DF$Environmental.Conditional.R2,na.rm=TRUE),2)` conditional $R^2$ `r round(sd(correlation.model.summary.DF$Environmental.Conditional.R2,na.rm=TRUE),2)` $\pm$ SD; geographic vs.
-neutral genetic variation: mean `r round(mean(correlation.model.summary.DF$Geographic.Conditional.R2,na.rm=TRUE),2)` conditional $R^2$ `r round(sd(correlation.model.summary.DF$Geographic.Conditional.R2,na.rm=TRUE),2)` $\pm$ SD) were generally slightly higher than the marginal $R^2$ values (environmental vs.
-adaptive genetic variation: mean `r round(mean(correlation.model.summary.DF$Environmental.Marginal.R2,na.rm=TRUE),2)` marginal $R^2$ `r round(sd(correlation.model.summary.DF$Environmental.Marginal.R2,na.rm=TRUE),2)` $\pm$ SD; geographic vs.
-neutral genetic variation: mean `r round(mean(correlation.model.summary.DF$Geographic.Marginal.R2,na.rm=TRUE),2)` marginal $R^2$ `r round(sd(correlation.model.summary.DF$Geographic.Marginal.R2,na.rm=TRUE),2)` $\pm$ SD).
-This result suggests that while variation in the surrogate variables correlates with genetic variation, both forms of genetic variation contain additional spatially auto-correlated variation that is not explained by these surrogates.
-In particular, the differences between the conditional and marginal $R^2$ values for the models examining the effectiveness of geographic variables suggest that this unexplained spatial auto-correlation may be non-linear or driven by a few neighboring planning units with genetically similar individuals.
-Similarly, other studies have also reported strong patterns of spatial auto-correlation in patterns of genetic variation [eg.
-@r502].
 
@@
 Our results demonstrate that environmental and geographic surrogates are strong enough to be operational (Figure 2).
-Generally, the environmental and geographic surrogates variables were equally effective surrogates for adaptive (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figures 2c and S56) and neutral genetic variation (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figures 2d and Figure S57), although for a few species, the geograsphic surrogate performed poorly (eg.
+For most species, the environmental and geographic surrogates variables were effective surrogates for adaptive (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figures 2c and S56) and neutral genetic variation (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figures 2d and Figure S57).
+Despite this, for a few species, geographic distance was a poor surrogate for neutral genetic variation (eg.
 _Gypsophila repens_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gypsophila_repens')$Geographic.r2CU,2)`; _Luzula alpinopilosa_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Luzula_alpinopilosa')$Geographic.r2CU,2)`; _Gentiana nivalis_ $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gentiana_nivalis')$Geographic.r2CU,2)`).
-The bimodal distribution of explained variation (Figure 2d) suggests that geographic position alone may not be a suitable surrogate for neutral variation in conservation planning for all species.
-Indeed, geographic distances will be a poor surrogate for genetic lineages in the process of sympatric speciation, connectivity between areas can be strongly affected by local topology, and some of species have dispersal modes that are non-linearly related to geographic distance (eg.
-wind dispersing plants can have high connectivity between nearby and distant areas; @r451).
-Overall, however, such species were the exception in our analysis, and both simple surrogates performed moderately well.
+Indeed, geographic distances will be a poor surrogate for species that utilise dispersal modes that result in non-linear relationships between connectivity and geographic distance [eg.
+wind dispersing plants can maintain high connectivity between distant areas; @r451], and in places where connectivity is most affected by landscape features [@r497].
+Overall, however, such species were the exception in our analysis, and both simple surrogates performed moderately well for a diverse set of species.
 
 <!-- suitable targets for surrogates -->
-In general, adaptive variation accumulated more quickly with environmental representation than neutral variation did with geographic representation (Figures 2a--2b and S56--S57).
-To secure at least `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` % of each species' adaptive genetic variation, prioritisations needed to sample on average `r round(mean(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` SD of the environmental conditions found across the planning units occupied by the species.
-On the other hand, to secure a representative sample of the species' neutral genetic variation ($\geq$ `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %), prioritisations needed to sample `r round(mean(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` SD of the variation in geographic position among the planning units occupied by the species.
-These results suggest that geographic surrogates may require higher targets than environmental surrogates to generate prioritisations that secure representative sample of genetic variation.
-Additionally, these results suggest using surrogate-based targets alone to generate prioritisations may be insufficient to secure a representative sample of species' genetic variation.
-For all species, it was possible to generate solution that secured a high proportion of the surrogate variable (> 90 %) but only a small proportion of adaptive or neutral genetic variation (< 20 %; Figures S56--S57).
-Thus prioritisations should be generated using a combination of amount-based targets--as is conventionally used in reserve selection processes--and also surrogate-based targets.
+Our results suggest that environmental and geographic surrogates can predict how much genetic variation prioritisations secure, but what targets should be used?
+First, our results suggest that surrogate-based targets should be used in conjunction with amount-based targets.
+For all species, it was possible to generate solutions that secured a high proportion of the surrogate variable (> 90 %) and only a small proportion of adaptive or neutral genetic variation (< 20 %; Figures S56--S57), and so use of amount-based targets may ameliorate this.
+Second, geographic surrogates may need much higher targets than environmental surrogates to capture evolutionary processes.
+In our analysis, geographic variation accumulated much slower for neutral genetic representation than environmental variation did with adaptive representation (Figures 2a--2b and S56--S57).
+To secure a representative sample of the species' neutral genetic variation ($\geq$ `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %), prioritisations needed to sample `r round(mean(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` SD of the variation in geographic position among the planning units occupied by the species.
+On the other hand, adaptive variation accumulated much faster.
+To secure at least `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` % of each species' adaptive genetic variation, prioritisations only needed to sample on average `r round(mean(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` SD of the environmental conditions found across the planning units occupied by the species.
+Note that these proportions are based on averages across species.
+In many cases, higher targets may be needed to ensure that prioritisations secure a representative sample of the genetic variation for all species in a planning exercise.
 
@@
 Our results suggest that conservation planners can secure a reasonably representative sample of the intra-specific adaptive genetic variation using only amount-based targets.
-Since the spatial patterns of adaptive genetic variation clustered into a few distinct groups for most species (eg.
-_Carex semipervens_ and _Cerastium uniflorum_; Figures S33 and S34 respectively), a prioritisation would only need to select a planning unit in each cluster to secure a representative sample.
-As a consequence, the single-species prioritisations generated using amount targets and without opportunity cost data were often able to achieve high representation because they were essentially random selection (Figure 4a).
-To a lesser extent, this is also true for the single-species prioritisations generated with amount targets and consideration to opportunity cost (Figure 4c), since the costs were fairly homogenous among planning units with the exception of a few planning units in lowland areas.
-Overall--and contrary to our original hypotheses--it seems that conventional reserve selection methods may secure a representative sample of species' adaptive genetic variation if a comprehensive set of species is used.
+For most species, the spatial patterns of adaptive genetic variation clustered into a few distinct groups (eg.
+_Carex semipervens_ and _Cerastium uniflorum_; Figures S33 and S34 respectively).
+As a consequence, a prioritisation would only need to select a planning unit in each cluster to secure a representative sample.
+This mechanism explains why adaptive variation accumulated much quicker than neutral genetic variation in the randomly generated prioritisations.
+Additionally, this mechanism also explains why the single-species prioritisations generated using only amount targets were often able to achieve high representation--they were essentially a random selection and so frequently contained planning units in different parts of the species' ranges (Figure 4a).
+Overall--and contrary to our original hypotheses--it seems that conventional reserve selection methods may secure a representative sample of species' adaptive genetic variation.
 
@@
 However, such distances require specific-specific scaling to adequate describe their connectivity [eg.
-due to different dispersal modes; @r497], and so cannot easily be utilised in large-scale planning exercises.
-Third, we used amplified fragment length polymorphism (AFLP) data to characterise the main gradients of genetic variation.
-Whilst next-generation sequencing provides higher resolution genetic information than AFLP data [reviewed in @r480], no multi-species genomic data set with similar properties currently exists, and our methodology would still have utilised only the main gradients of the genetic variation to generate prioritisations in a feasible amount of time.
+@r497], and so cannot easily be utilised in large-scale planning exercises.
+Third, we used amplified fragment length polymorphism [AFLP; @r453] data to characterise the main gradients of genetic variation.
+Whilst next-generation sequencing provides higher resolution genetic information than AFLP data [@r480], no multi-species genomic data set with similar properties currently exists, and our methodology would still have utilised only the main gradients of the genetic variation to generate prioritisations in a feasible amount of time.
 
@@
 
-We used species distribution and genomic data collected by the international IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for further explanation of data collection methods].
-This data set contains information for 27 alpine plant species (Figure 2a).
-This data set has previously been used to explore patterns of adaptive [@r484; @r451; @r475; @r455; @r485] and neutral genetic variation [@r478; @r451; @r486], and to show that species richness is not an effective surrogate for genetic diversity in conservation planning [@r486].
+We used data for 27 alpine plant species in the European Alps collected by the IntraBioDiv project [@r451; Figure 2a].
+This data set has previously been used to explore patterns of adaptive [eg.
+@r475; @r455] and neutral genetic variation [eg.
+@r478], and to show that species richness is not an effective surrogate for genetic diversity in conservation planning [@r486].
 It was collected using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx.
@@
 Project members visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals.
-They genotyped samples using amplified fragment length polymorphisms [AFLP; @r453], and constructed matrices denoting the presence/absence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; see Table 1 in Alvarez _et al._ [@r478] for the number markers for each species).
+They genotyped samples using AFLPs, and constructed matrices denoting the presence/absence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; see Table 1 in Alvarez _et al._ [@r478] for more information).
 Thus the data set contains information describing the genomic properties of individuals for each species present in every second grid cell.
@@
 Squares denote planning units.
-Panel (a) depicts patterns of species riches.
-Planning units with a brighter color are inhabited by more species.
-Panel (b) shows the distribution of opportunity cost across the study area (computed as total population density) plotted using a quantile-based color ramp."}
+Panel (a) depicts patterns of species riches, and panel (b) shows the distribution of acquisition cost (plotted using a quantile-based color ramp)."}
 
@@
 
-We adopted the sampling grid used by the IntraBioDiv project as planning units to develop prioritisations, omitting cells that were surveyed and not found to contain individuals belonging to any of the study species to reduce computational burden (`r nrow(grid.DF)` grid cells used in analysis).
-To examine the potential effects of variation in cost (see below), we calculated the total human population density inside each planning unit [obtained at 1 km$^2$ resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472] and used this to represent acquisition cost (Figure 2b).
-Unless otherwise specified, all computational, spatial, and statistical analysis were conducted in R [version `r paste0(R.version$major, '.', R.version$minor)`; @r192].
+We adopted the sampling grid used by the IntraBioDiv project as planning units to develop prioritisations, omitting cells that were not found to contain individuals belonging to the study species (`r nrow(grid.DF)` grid cells used in analysis).
+To examine the potential effects of variation in cost, we calculated the total human population density inside each planning unit [obtained at 1 km$^2$ resolution from the Global Rural-Urban Mapping Project; @r472] and used this to represent acquisition cost (Figure 2b).
+All spatial and statistical analysis were conducted in \texttt{R} [version `r paste0(R.version$major, '.', R.version$minor)` @r192].
 
+We created environmental and geographic surrogate variables for each species.
 To describe the geographic location of each planning unit, we projected the grid into an equi-distant coordinate system (Europe Equidistant Conic; $\texttt{ESRI:102031}$), calculated the centroid of each grid cell, and extracted their two-dimensional coordinates.
-Thus we constructed two-dimensional geographic space--wherein each planning unit occupies a single point--as a potential surrogate for neutral genetic variation.
-We expected planning units further apart from each other in this geographic space to contain individuals with increasingly different patterns in their neutral loci.
-
-To describe the environmental characteristics of each planning unit, we obtained data for 19 bioclimatic variables across the extent of the study area [at $30^{\prime\prime}$ resolution; @r33].
-We used this data set because it is commonly to map habitat suitability in conservation planning [@r493].
-These layers and the planning units were projected into an equal-area coordinate system (Europe Lambert Conformal Conic; $\texttt{ESRI:102014}$).
-To reduce dimensionality, the 19 bioclimatic variables were clipped to the planning units and subjected to a principal components analysis (PCA; using \texttt{ArcMap} version 10.2.2; Table S1).
-The first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components (PCs) cumulatively explained `r round(pca.DF[[4]][surrogate.params.LST[[MODE]]$number.components],1)` % of the variation and were used to generate new layers.
-We calculated the average value for each of the `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components in each planning unit, and used these values to characterise climatic variation among the planning units (Figure S2).
-Thus we constructed a `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))`-dimensional environmental space--wherein each planning unit occupies a single point--as a potential surrogate for adaptive variation.
-We expect planning units further apart from each other in this environmental space to contain individuals with increasingly different patterns in their adaptive loci.
+To describe the environmental characteristics of each planning unit, we obtained 19 climatic layers [$30^{\prime\prime}$ resolution; @r33], projected them and the planning units into an equal-area coordinate system (Europe Lambert Conformal Conic; $\texttt{ESRI:102014}$), and computed planning unit averages for each climatic variable.
+To reduce dimensionality, for each species, we subject the climatic values associated with their planning units to a principal components analysis (PCA; Table S1).
+We used the first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components to characterise climatic variation found across the species' distributions inside the study area (capturing `r round(mean(spp.pca.DF[[2]]*100),2)` % $\pm$ `r round(sd(spp.pca.DF[[2]]*100),2)` SD of the total climatic variation; Figure S2).
+Thus we constructed a two-dimensional geographic space as a potential surrogate for neutral genetic variation, and a `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))`-dimensional environmental space as a potential surrogate for adaptive genetic variation for each species.
+In these spaces, each planning unit was associated with a single point.
+Planning units with more similar environmental conditions, or higher spatial proximity in the real world, were associated with points that were closer together in these spaces (respectively).
 
@@
 
-To investigate the effectiveness of our supposed surrogates for adaptive and neutral genetic variation, we first needed to identify which of the sampled loci are adaptive.
-We used two outlier detection methods to achieve this [an individual- and a population-level method as recommended by @r495].
-The basic premise underpinning such methods is that neutral loci are expected to exhibit a specific level of variation among populations, and loci that deviate from this expectation are likely to be under selection [reviewed in @r480].
-The advantage of these methods--in contrast with environmental association analyses [@r489]--is that they do not use environmental data to identify loci under selection, which would have introduced circularity into our analysis.
+To investigate the effectiveness of our supposed surrogates, we first needed to identify which of the sampled loci were adaptive.
+Following best practices, we used two outlier detection methods to achieve this [an individual- and a population-level method @r495].
+The basic premise underpinning such methods is that neutral loci are expected to exhibit a certain level of variation, and loci that deviate from this expectation are likely to be under selection [@r480].
+The advantage of these methods--in contrast to environmental association analyses--is that they do not use environmental data, which would have introduced circularity into our analysis.
 Loci identified as outliers in both analyses were treated as adaptive and the remainder were treated as neutral.
-To minimise false-positives, we omitted loci from both outlier detection methods where the global frequency of the minor allele was less than 10 % and assumed they were neutral.
+To minimise false-positives, we omitted loci from both methods where the global frequency of the minor allele was less than 10 % and treated them as neutral.
 
-The first outlier detection method we used involved fitting multinominal-Dirichlet models implemented in \texttt{BayeScan} [version 2.1; @r456].
-We adopted a similar methodology to Bothwell _et al._ [-@r455] and applied it to each of the species in the data set.
-Following their methodology, we first grouped conspecifics into genetic lineages to avoid falsely classifying loci as adaptive.
-To complete this first step, for each species, we fit the admixture and correlated alleles model implemented in \texttt{Structure} [version 2.3.4; @r461; @r462; `r structure.params.LST[[MODE]]$numruns` replicates per K; `r structure.params.LST[[MODE]]$admburnin` admixture burnin iterations; `r structure.params.LST[[MODE]]$burnin` burnin iterations; `r structure.params.LST[[MODE]]$numreps` iterations used for inference] using the number of populations for each species previously determined by Alvarez _et al._ [-@r478].
-We used \texttt{ClumPP} [version 1.1.2, @r459] to combine replicate runs to group individuals into populations (greedy algorithm using the $G^{\prime}$ static; `r clumpp.params.LST[[MODE]]$repeats` iterations).
-We then ran $\texttt{Bayescan}$ [@r456] for each species using these clusters (1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations).
-Additionally, we omitted individuals from the \texttt{Bayescan} analysis if their population membership was uncertain (maximum membership probability $< `r structure.params.LST[[MODE]]$probability.threshold`$).
-We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate [$q \leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].
+The first outlier detection method involved fitting multinominal-Dirichlet models implemented in \texttt{BayeScan} [version 2.1 @r456].
+We adopted a similar methodology to Bothwell _et al._ [-@r455] and applied it to each of the species separately.
+Following their methodology, we initially grouped conspecifics into genetic lineages to further minimise false-positives.
+We grouped conspecifics by fitting admixture and correlated alleles models implemented in \texttt{Structure} [version 2.3.4; @r462; `r structure.params.LST[[MODE]]$numruns` replicates per species; `r format(structure.params.LST[[MODE]]$admburnin, big.mark=',', scientific=FALSE)` admixture burnin iterations; `r format(structure.params.LST[[MODE]]$burnin, big.mark=',', scientific=FALSE)` burnin iterations; `r format(structure.params.LST[[MODE]]$numreps+structure.params.LST[[MODE]]$burnin, big.mark=',', scientific=FALSE)` total iterations] using the number of lineages previously determined by Alvarez _et al._ [-@r478] and combining replicate runs using \texttt{ClumPP} [version 1.1.2, @r459; greedy algorithm based on the $G^{\prime}$ statistic; `r format(clumpp.params.LST[[MODE]]$repeats, big.mark=',', scientific=FALSE)` iterations).
+We then ran $\texttt{Bayescan}$ for each species using these lineages (1:1 prior odds; `r bayescan.params.LST[[MODE]]$reps` replicates per species, `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r format(bayescan.params.LST[[MODE]]$burn, big.mark=',', scientific=FALSE)` burn-in iterations; `r format(bayescan.params.LST[[MODE]]$n+bayescan.params.LST[[MODE]]$burn, big.mark=',', scientific=FALSE)` total iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations) using a suitable false discovery rate [$q \leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].
+We omitted individuals if their population membership was uncertain (maximum membership probability $< `r structure.params.LST[[MODE]]$probability.threshold`$).
 
-The second outlier detection method involved fitting PCAs to identify outlier loci [using methods implemented in the \texttt{pcadapt R} package; version 3.0; @r487; @r496].
-To enable comparisons between the two outlier detection analyses, we used the same individuals in this analysis as used in the \texttt{BayeScan} analysis.
-For each of the 27 species, we first imputed missing data to permit further analysis.
-Specifically, the missing value for a given locus for a given individual was replaced with the average frequency of that locus among all conspecifics.
-Next, we ran a PCA and extracted the minimum number of components needed to secure the over-arching population level variation among the loci (`r pcadapt.params.LST[[MODE]]$min.variance.explained*10` % of the variation in loci).
-We then used Mahalanobis distances [@r398] to compute _P_ values and then _q_-values [using the \texttt{qvalue R} package; @r491].
-Outlier loci were detected using a suitable false discovery rate ($q < `r pcadapt.params.LST[[MODE]]$alpha.level`$).
+The second outlier detection method involved fitting PCAs to identify outlier loci [implemented in the \texttt{pcadapt R} package; @r487].
+To enable comparisons between the two outlier detection methods, we used the same individuals in this analysis as in the \texttt{BayeScan} analysis.
+For each of the 27 species, we first imputed missing data--replacing missing values with the average frequency among conspecifics--ran a PCA, and extracted the minimum number of components needed to secure the over-arching population level variation among the loci (`r pcadapt.params.LST[[MODE]]$min.variance.explained*100` % of the variation in loci).
+We then computed _q_-values using Mahalanobis distances, and used the same false discovery rate used in the \texttt{BayeScan} analysis to identify loci under selection [using the \texttt{qvalue R} package; @r491].
 
 After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species.
-For each species, we discarded the population groupings, and partitioned adaptive and neutral loci into separate matrices.
-We applied separate non-metric multi-dimensional scaling analyses to each matrix [NMDS; implemented in the \texttt{vegan R} package; @r458] using Gower distances [@r464; using the \texttt{cluster R} package to accommodate sparsity; @r457] to derive continuous variables that described the main gradients of adaptive and neutral genetic variation (separately) for each species (Table S2).
-To ensure that the ordinations described a sufficient amount of the genomic variation, we ran successive scaling analyses with increasing dimensionality until a sufficient stress value was obtained (maximum stress value $\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for each analysis).
-Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the individuals in each grid cell to express the typical genomic characteristics of individuals in the cells.
+We discarded the population groupings, and partitioned species' adaptive and neutral loci into separate matrices.
+We applied non-metric multi-dimensional scaling analyses [NMDS; implemented in the \texttt{vegan R} package; @r458] using Gower distances [via the \texttt{cluster R} package; @r457] to derive continuous variables that described the main gradients of adaptive and neutral genetic variation separately for each species (Table S2).
+To ensure that the ordinations described a sufficient amount of the genetic variation, we ran successive scaling analyses with increasing dimensionality until a sufficient stress value was obtained (maximum stress value $\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for each analysis).
+Since each grid cell had multiple samples per species, we averaged the ordinated values belonging to conspecifics in the same grid cell.
 
-The previous set of analyses resulted in an adaptive (if detected) and neutral genetic space for each species.
-For a given species, each planning unit occupied by the species was associated with a multi-dimensional point in the species' neutral genetic space.
-Planning units that were closer together in this space are occupied by individuals with more similar polymorphisms in their neutral loci.
-Similarly, if the species was associated with adaptive loci, each planning unit occupied by the species was associated with a multi-dimensional point in the species' adaptive genetic space.
-By spreading out conservation effort across a given genetic space, and in turn securing planning units occupied by individuals with increasingly different polymorphisms, prioritisations can secure more genetic variation [see @r465 for discussion on attribute spaces].
+Thus we constructed an an adaptive (if detected) and neutral genetic space for each species.
+For a given species, each planning unit that it was found in was associated with a multi-dimensional point in the species' adaptive genetic space (if adaptive genetic variation was detected) and another multi-dimensional point in the species' neutral genetic space.
+Planning units that were closer together in these genetic spaces were occupied by individuals with more similar polymorphisms.
+By spreading out conservation effort across these genetic spaces, and in turn selecting planning units occupied by individuals with increasingly different polymorphisms, prioritisations can secure more genetic variation.
 
@@
 
-We used the unreliable representative and adequate prioritisation (URAP) formulation of the reserve selection problem [@r465].
-This formulation can identify the minimum number of planning units required to preserve both a proportion of the species' range (using amount-based targets) and a proportion of variation found across the species' range [using space-based targets and attribute spaces; _sensu_ @r465].
-We treated the environmental and geographic surrogate variables as separate attribute spaces.
-We also treated the adaptive (if detected) and neutral variables as separate attribute spaces for each species.
-We set the demand points for each species and each attribute space using the values associated with the planning units.
-For every prioritisation we generated, we calculated the proportion of the adaptive (if detected) and neutral genetic variation that every prioritisation secured to assess its performance.
-Prioritisations associated with negative values--because they secured only a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis.
-We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality using the \texttt{raptr R} package [@r465] and $\texttt{Gurobi}$ [@r463].
+We used the unreliable representative and adequate prioritisation (URAP) formulation of reserve selection [@r465].
+This formulation can identify the minimum number of planning units required to preserve both a proportion of the species' geographic range (using amount-based targets) and a sample of variation found across the species' range.
+We used the environmental and geographic surrogate spaces, and the adaptive (if detected) and neutral genetic spaces to generate and evaluate solutions.
+Solutions associated with negative values--because they secured a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis.
+We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality using the \texttt{raptr R} package [@r465].
 
-Statistical and computational analyses {#statistical-analyses .unnumbered}
+Computational and statistical analyses {#statistical-analyses .unnumbered}
 --------------------------------------------------------
 
-To determine if the environmental and geographic variables correlate with spatial patterns of adaptive and neutral genetic variation, we computed dissimilarity matrices between the planning units, and tested for correlations between matrices based on surrogate and genetic data.
-For each species, we computed two dissimilarity matrices between the planning units occupied by the species in terms of their environmental conditions (their coordinates in the environmental space) and geographic position (their coordinates in the geographic space).
-We then computed another dissimilarity matrix between the planning units using their values in the species' neutral genetic space.
-If the species was associated with adaptive loci, we also computed another dissimilarity matrix between the planning units using the species' adaptive genetic space.
-All dissimilarity matrices calculated using Euclidean distances.
+Our first aim was to determine if the environmental and geographic variables correlate with spatial patterns of adaptive and neutral genetic variation.
+To achieve this, for each species, we generated two dissimilarity matrices between the planning units that it occupied based on their environmental conditions and geographic position (using their respective surrogate spaces).
+We then generated another dissimilarity matrix between the planning units using their values in the species' neutral genetic space.
+If the species was associated with adaptive loci, we also generated another dissimilarity matrix between the planning units using the species' adaptive genetic space.
+Dissimilarities were compured as Euclidean distances.
 
-We then fit maximum likelihood population-effects (MLPE) models [@r494; implemented in the \texttt{ResistanceGA R} package; @r488].
-These models explicitly accommodate the structure of dissimilarity matrices using random effects.
-For each species, we fit a MLPE model to correlate the species' dissimilarity matrices based on geographic position and neutral genetic variation.
-Additionally, if the species was associated with adaptive loci, we also fit a MLPE model to correlate the species' dissimilarity matrices calculated using environmental conditions and adaptive genetic variation.
-All predictor variables were z-transformed prior to model fitting to improve convergence.
-To test if the surrogate variables explained a substantial amount of genetic variation, we conducted $\chi^2$ tests between each model and its corresponding null model, and applied Bonferroni corrections.
+We then fit maximum likelihood population-effects (MLPE) models [@r494].
+These models accommodate the structure of dissimilarity matrices using random effects.
+For each species, we fit a MLPE model to the species' dissimilarity matrices based on geographic position and neutral genetic variation.
+Additionally, if adaptive loci were detected, we also fit a MLPE model to the species' dissimilarity matrices describing environmental and adaptive genetic variation.
+All predictor variables were _z_-transformed prior to model fitting to improve convergence.
+To test if the surrogates explained a substantial amount of genetic variation, we compared each model to its corresponding null model using $\chi^2$, and applied Bonferroni corrections.
 
-To address the second aim--to determine if the environmental and geographic variables are effective surrogates for adaptive and neutral genetic variation--we generated a collection of single-species prioritisations and computed the proportion of intra-specific genetic variation they secured.
-For each species, we generated `r raptr.params.LST[[MODE]]$surrogacy.analysis$random.replicates` prioritisations by randomly selecting different combinations of planning units that were occupied by the species.
-We then computed the proportion of adaptive genetic variation and environmental variation secured in each prioritisation.
-We then generated another random set of prioritisations, and computed the proportion of neutral genetic variation and geographic variation secured in each prioritisation.
+Our second aim was to determine if the environmental and geographic variables were effective surrogates for adaptive and neutral genetic variation.
+For each species, we generated `r format(raptr.params.LST[[MODE]]$surrogacy.analysis$random.replicates, big.mark=',', scientific=FALSE)` prioritisations by randomly selecting different combinations of planning units that it occupied, and calculated the proportion of geographic variation and neutral genetic variation it secured.
+We repeated this processes, and calculated the proportion of environmental variation and adaptive genetic variation they secured.
 
 We fit two full generalised linear models (GLM) with logit links.
-The first model was fit to the proportion of adaptive genetic variation secured in a prioritisation, and contained a continuous predictor variable measuring the proportion of environmental variation also secured in the prioritisation, a categorical variable indicating the species for which the prioritisation was generated, and an interaction term between these predictor variables.
-The second model was fit to the proportion of neutral genetic variation secured in a prioritisation, and contained a continuous predictor variable assessing the proportion of geographic variation also secured in the prioritisation.
-Similar to the previous model, this model also contained a predictor variable indicating the species for which the prioritisations were generated, and an interaction term.
-We subjected these models to a backwards step-wise term deletion routine to assess term significance.
-To gauge the performance of these surrogates for each species, we fit a suite of models correlating the proportion of adaptive and neutral genetic variation held in each prioritisation and the environmental and geographic variation they also secured (respectively).
-We then computed the Cragg and Uhler's pseudo-R$^2$ value for each model [using the \texttt{pscl} R package; @r492].
+The first model was fit to the proportion of adaptive genetic variation secured in a prioritisation using the proportion of environmental variation also secured in the prioritisation.
+The second model was fit to the proportion of neutral genetic variation secured in a prioritisation using proportion of geographic variation also secured in the prioritisation.
+Both models also contained a variable indicating the species for which the prioritisations were generated, and an interaction term.
+They were subject to backwards step-wise term deletion routines to assess term significance.
+To assess the performance of the surrogates, we refit these models seperately for each species and computed the Cragg and Uhler's pseudo-R value [using the \texttt{pscl R} package; @r492].
 
-To determine whether the use of surrogate-based targets improves prioritisations, we investigated the effectiveness of surrogate-based targets under three scenarios with increasing levels of complexity.
-To assess the benefits of surrogate-based targets under the simplest of circumstances, our first scenario involved generating single-species prioritisations for each species.
-However, most large-scale prioritisations involve a comprehensive set of species.
-By forcing solutions to secure individuals in multiple communities, this may result in prioritisations that secure a greater proportion of intra-specific genetic variation for each species.
-To test for this effect, our second scenario involved generating multi-species prioritisations.
-Finally, most large-scale prioritisations also explicitly consider opportunity cost to ensure that prioritisations are cost-effective.
-Our third scenario involved generating multi-species prioritisations that satisfied targets for minimal opportunity cost.
-
-For each scenario, we generated three types of prioritisations using (1) only amount-based targets, (2) both amount- and surrogate-based targets, and (3) both amount- and genetic-based targets.
-The purely amount-based prioritisations served as our baseline--representing the prioritisations typically generated in conservation planning exercises (eg.
-by \texttt{Marxan}).
-We used `r (raptr.params.LST[[MODE]]$scenario.analysis$amount.target*100)` % amount-based targets in all prioritisations to ensure an adequate proportion of habitat was secured for each species.
-Based on the results from the analyses used to address our second aim, we set surrogate-based targets as `r (raptr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100)` % and genetic-based targets as `r (raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)` % to ensure that most of the genetic variation was secured for each species.
-To accommodate the fact that there were a large number of optimal solutions for the single-species amount-based prioritisations, we generated `r (raptr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates)` replicate prioritisations.
+Our third aim was to determine whether surrogate-based targets improved prioritisations.
+As previously mentioned, we generated prioritisations using different combinations of targets and conservation planning scenarios.
+We used `r (raptr.params.LST[[MODE]]$scenario.analysis$amount.target*100)` % amount-based targets in all prioritisations.
+Based on the results from our previous aim, we used `r (raptr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100)` % surrogate-based targets and `r (raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)` % genetic-based targets.
+We generated a single solution for each targets/scenario combination, except for the single-species scenario with just amount-based targets for which we generated `r format(raptr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates, big.mark=',', scientific=FALSE)` replicates because this combination had many optimal solutions.
 We computed the proportion of adaptive (if present) and neutral genetic variation secured for each species in the prioritisations.
@@
 We fit a full model to the proportion of genetic variation secured in a given prioritisation.
-This model contained categorical predictor variables to indicate the combination of targets used to generate the prioritisations (amount-only, amount and surrogate targets, and amount and genetic targets), the scenario that was used to generate the prioritisation (single-species, multi-species, and multi-species with cost), the type of genetic variation measured (adaptive and neutral), and all interactions between these variables.
-This model accommodated prioritisations generated for the same species using a random intercept term.
-We subject the full model to a backwards step-wise term deletion routine to assess term significance.
-We conducted a post-hoc analysis on the minimal adequate model using Tukey contrasts with a Bonferroni correction [using the \texttt{multcomp R} package; @r470].
+This model contained categorical variables indicating the combination of targets (amount-only, amount and surrogate targets, or amount and genetic targets), the scenario used to generate the prioritisation (single-species, multi-species, or multi-species with cost), the type of genetic variation measured (adaptive or neutral), and all interactions between them.
+Prioritisations generated for the same species were accomodated using a random intercept term.
+We subject the full model to a backwards step-wise term deletion routine to assess term significance, and conducted a post-hoc analysis on the minimal adequate model using Tukey contrasts with a Bonferroni correction [using the \texttt{multcomp R} package; @r470].
 

### 2016-10-11 code/rmarkdown/supporting_information.Rmd
+ article within page pnas limits

@@
 ```
+
+# Figures
 
@@
 
-```{r, echo=FALSE, results='asis'}
-## make table showing pca resuls
-curr.file <- tempfile(fileext='.tex')
-x <- spp.pca.DF %>%
-	mutate(species=gsub('_', ' ', species, fixed=TRUE), percent_variation_explained=percent_variation_explained*100) %>%
-	latex(
-		file=curr.file,
-		rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, col.just=rep('c', 2),
-		colheads=c('\\bfseries{Species}', '\\bfseries{Variation explained (\\%)}'), here=TRUE,
-		caption=paste0('Summary of principal component analyses fit to the climatic variation in the species\' ranges.
-Percentages correspond to the percent of variation described by the first ',as.character(as.english(surrogate.params.LST[[MODE]]$number.components)),' principle components.')
-	)
-readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
-	gsub(pattern='\\cline{1-', replacement='\\cline{3-4} \\cline{6-', fixed=TRUE) %>% 
-	gsub(pattern='toprule', replacement='toprule[1pt]', fixed=TRUE) %>%
-	gsub(pattern='bottomrule', replacement='bottomrule[1pt]', fixed=TRUE) %>%
-	cat()
+# Tables
+
+```{r, echo=FALSE}
+write.table(loci_classification.DF, 'article/Table_S1.csv', sep=',', row.names=FALSE)
 ```
@@
 	gsub(pattern='bottomrule', replacement='bottomrule[1pt]', fixed=TRUE) %>%
+	cat()
+```
+
+```{r, echo=FALSE, results='asis'}
+curr.file <- tempfile(fileext='.tex')
+x <- correlation.model.summary.DF %>%
+	select(-contains('raw')) %>%
+	mutate(
+		Environmental.corr.P=pvalString(Environmental.corr.P),
+		Geographic.corr.P=pvalString(Geographic.corr.P),
+		Geographic.Marginal.R2=pvalString(Geographic.Marginal.R2),
+		Geographic.Conditional.R2=pvalString(Geographic.Conditional.R2),
+		Environmental.Marginal.R2=pvalString(Environmental.Marginal.R2),
+		Environmental.Conditional.R2=pvalString(Environmental.Conditional.R2),
+		blank=rep('', length(Species)),
+		Species=paste0('\\textit{', gsub('\\ ', ' ', Species), '}')) %>%
+	mutate(Environmental.Chisq=sapply(round(Environmental.Chisq,2), as.character)) %>%
+	mutate(Geographic.Chisq=sapply(round(Geographic.Chisq,2), as.character)) %>%
+	select(Species:Environmental.corr.P, Geographic.Marginal.R2:Geographic.corr.P) %>%
+	latex(file=curr.file, 
+		cgroup=c('', 'Environmental', 'Geographic'), n.cgroup=c(1, 4, 4), col.just=rep('c', 9),
+		digits=2, rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, here=TRUE,
+		colheads=c('\\bfseries{Species}', '$MR^2$', '$CR^2$', '$\\chi^2$', '\\textit{P}', '$MR^2$', '$CR^2$', '$\\chi^2$', '\\textit{P}'),
+		caption='Summary of maximum likelihood population-effects (MLPE) models.
+Briefly, for each species associated with potentially adaptive loci, a model was fit to correlate dissimilarities among planning units where the species in terms of the adaptive genetic characteristics of individuals in them with dissimilarities in terms of the environmental conditions inside the planning units (denoted in the environmental column group).
+Similarly, for each species, a model was fit to correlate dissimilarities among the planning units in terms of the neutral genetic characteristics of indiviudals in them with dissimilarities in terms of the geographic position of the planning units (denoted in the geographic column group).
+Data shows the results from comparing each model to its corresponding null model using $\\mathbf{\\chi^2}$ tests (1 degree of freedom), and the $\\mathbf{R^2}$ ($\\mathbf{MR^2}$) and conditional $\\mathbf{R^2}$ ($\\mathbf{CR^2}$) for each model.'
+	)
+readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
+	gsub(pattern='\\cline{1-', replacement='\\cline{2-6} \\cline{8-', fixed=TRUE) %>% 
+	gsub(pattern='{c}{\\textit{P}}\\tabularnewline', replacement='{c}{\\textit{P}}\\T\\B\\tabularnewline', fixed=TRUE) %>% 
+	gsub(pattern='\\toprule', replacement='', fixed=TRUE) %>%
+	gsub(pattern='\\midrule', replacement='\\midrule', fixed=TRUE) %>%
+	gsub(pattern='\\bottomrule', replacement='\\midrule', fixed=TRUE) %>%
 	cat()
@@
 
-```{r, echo=FALSE}
-write.table(loci_classification.DF, 'article/Table_S4.csv', sep=',', row.names=FALSE)
+```{r, echo=FALSE, results='asis'}
+## make table showing pca resuls
+curr.file <- tempfile(fileext='.tex')
+x <- spp.pca.DF %>%
+	mutate(species=gsub('_', ' ', species, fixed=TRUE), percent_variation_explained=percent_variation_explained*100) %>%
+	latex(
+		file=curr.file,
+		rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, col.just=rep('c', 2),
+		colheads=c('\\bfseries{Species}', '\\bfseries{Variation explained (\\%)}'), here=TRUE,
+		caption=paste0('Summary of principal component analyses fit to the climatic variation in the species\' ranges.
+Percentages correspond to the percent of variation described by the first ',as.character(as.english(surrogate.params.LST[[MODE]]$number.components)),' principle components.')
+	)
+readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
+	gsub(pattern='\\cline{1-', replacement='\\cline{3-4} \\cline{6-', fixed=TRUE) %>% 
+	gsub(pattern='toprule', replacement='toprule[1pt]', fixed=TRUE) %>%
+	gsub(pattern='bottomrule', replacement='bottomrule[1pt]', fixed=TRUE) %>%
+	cat()
 ```

### 2016-10-11 code/rmarkdown/article.Rmd
+ fix refs + use one-sided tests in post hoc for second analysis

@@
 <!-- Introduction -->
-Protected areas spearhead conservation efforts [@r440].
+\dropcap{P}rotected areas spearhead conservation efforts ([@r440]).
 They buffer species from anthropogenic impacts; providing places for them to persist.
-Since the resources available for conservation are limited, protected areas need to be sited in places that fulfill conservation objectives for minimal cost [@r255].
+Since the resources available for conservation are limited, protected areas need to be sited in places that fulfill conservation objectives for minimal cost ([@r255]).
 To achieve this, conservation planning exercises often generate plans for entire networks of protected areas (prioritisations) to allocate resources cost-effectively and preserve broad-scale biodiversity processes.
-Intra-specific genetic variation is fundamental to long-term species persistence [@r165], and as a consequence, there has been increasing interest in designing prioritisations that capture and cultivate this aspect of biodiversity [@r67; @r3; @r61; @r499].
+Intra-specific genetic variation is fundamental to long-term species persistence ([@r165]), and as a consequence, there has been increasing interest in designing prioritisations that capture and cultivate this aspect of biodiversity ([@r67; @r3; @r61; @r499]).
 
-Broadly speaking, though the strength of natural selection is a continuous force, genetic variation can be classified as either adaptive or neutral [reviewed in @r480].
+Broadly speaking, though the strength of natural selection is a continuous force, genetic variation can be classified as either adaptive or neutral (reviewed in [@r480]).
 Adaptive genetic variation is associated with loci that significantly affect fitness.
-Typically, such "adaptive" loci are detected due to anomalous patterns among individuals [eg.
-@r456; @r487], though phenotypic-genetic associations are more likely to identify functionally important genomic regions for individual species [eg.
-@r500].
-By representing the full range of adaptive variation, protected areas can help ensure that populations with particularly beneficial adaptations are not lost [@r3; @r43].
+Typically, such "adaptive" loci are detected due to anomalous patterns among individuals (eg.
+[@r456; @r487]), though phenotypic-genetic associations are more likely to identify functionally important genomic regions for individual species (eg.
+[@r500]).
+By representing the full range of adaptive variation, protected areas can help ensure that populations with particularly beneficial adaptations are not lost ([@r3; @r43]).
 In contrast, neutral genetic variation is associated with loci that do not significantly affect fitness, but instead reflects the evolutionary history of different populations.
-By representing the full range of neutral variation, protected areas can safeguard against the adverse effects of low genetic diversity [@r67].
-Thus optimally sited protected area networks would aim to represent both adaptive and neutral patterns of genetic variation [@r67; @r3; @r43].
+By representing the full range of neutral variation, protected areas can safeguard against the adverse effects of low genetic diversity ([@r67]).
+Thus optimally sited protected area networks would aim to represent both adaptive and neutral patterns of genetic variation ([@r67; @r3; @r43]).
 
 Recently, the use of surrogate data has been proposed to generate prioritisations that sufficiently intra-specific genetic variation--without needing to directly utilise genetic data to achieve this.
-Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been proposed to guide reserve selection as surrogates for adaptive genetic variation [eg.
-@r202; @r2].
+Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been proposed to guide reserve selection as surrogates for adaptive genetic variation (eg.
+[@r202; @r2]).
 By generating prioritisations that preserve a more representative sample of environmental conditions across species' ranges, these prioritisations could also capture more adaptive variation.
@@
 Neutral genetic variation arises from a reduction in gene flow between individuals.
-Over the last few years, the field of phylogeography has predominantly focused on describing neutral genetic diversity, and so the landscape factors that affect neutral variation are relatively well understood [@r501].
+Over the last few years, the field of phylogeography has predominantly focused on describing neutral genetic diversity, and so the landscape factors that affect neutral variation are relatively well understood (reviewed in [@r501]).
 Potential surrogates for neutral variation have been based on variables that predict the levels of connectivity between planning units.
-For instance, isolation by distance [IBD; @r259] predicts that populations further apart will experience less gene flow, and in turn, share less neutral loci.
-Based on this idea, geographic surrogates based on distance have been found to improve conservation prioritisations in an island network [@r200].
-However this remains untested in spatially contiguous systems where connectivity is complicated by additional factors [eg.
-habitat type @r497]--the typical situations in most conservation planning exercises.
+For instance, isolation by distance (IBD; [@r259]) predicts that populations further apart will experience less gene flow, and in turn, share less neutral loci.
+Based on this idea, geographic surrogates based on distance have been found to improve conservation prioritisations in an island network ([@r200]).
+However this remains untested in spatially contiguous systems where connectivity is complicated by additional factors (eg.
+habitat; [@r497])--the typical situations in most conservation planning exercises.
 
 Here we investigate the effectiveness of environmental and geographic surrogates for capturing adaptive and neutral genetic variation respectively in the context of conservation planning.
-We used distribution and genomic data for 27 alpine plants collected by the international IntraBioDiv project in the European Alps [@r451].
+We used distribution and genomic data for 27 alpine plants collected by the international IntraBioDiv project in the European Alps ([@r451]).
 We show that for most species the spatial patterns of adaptive and neutral genetic variation correlated with environmental and geographic variables.
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=(5*0.9), fig.height=(7.1*0.9), out.width='3.42in', out.height='4in', fig.cap='The effectiveness of prioritizations at securing genetic variation.
-Box plots show the proportion of adaptive and neutral genetic variation secured for each species using different targets and under different scenarios.
-Red lines show the target proportion of genetic variation to secure a representative sample.'}
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=(5*0.9), fig.height=(7.1*0.9), out.width='3.42in', out.height='4in', fig.cap='Box plots show the proportion of adaptive and neutral genetic variation secured for each species in solution generated from different targets and scenarios.
+Red lines show the target proportion of genetic variation.'}
 # prepare data for plotting
@@
 Indeed, geographic distances will be a poor surrogate for species that utilise dispersal modes that result in non-linear relationships between connectivity and geographic distance [eg.
-wind dispersing plants can maintain high connectivity between distant areas; @r451], and in places where connectivity is most affected by landscape features [@r497].
+wind dispersing plants can maintain high connectivity between distant areas; @r451], and in landscapes where connectivity is most affected by landscape features ([@r497]).
 Overall, however, such species were the exception in our analysis, and both simple surrogates performed moderately well for a diverse set of species.
@@
 Second, we used geographic distances as surrogates for neutral genetic variation, but connectivity in this landscape may be better described using distances that incorporate information on topography.
-However, such distances require specific-specific scaling to adequate describe their connectivity [eg.
-@r497], and so cannot easily be utilised in large-scale planning exercises.
-Third, we used amplified fragment length polymorphism [AFLP; @r453] data to characterise the main gradients of genetic variation.
-Whilst next-generation sequencing provides higher resolution genetic information than AFLP data [@r480], no multi-species genomic data set with similar properties currently exists, and our methodology would still have utilised only the main gradients of the genetic variation to generate prioritisations in a feasible amount of time.
+However, such distances require specific-specific scaling to adequate describe their connectivity (eg.
+[@r497]), and so cannot easily be utilised in large-scale planning exercises.
+Third, we used amplified fragment length polymorphism data (AFLP; [@r453]) to characterise the main gradients of genetic variation.
+Whilst next-generation sequencing provides higher resolution genetic information than AFLP data ([@r480]), no multi-species genomic data set with similar properties currently exists, and our methodology would still have utilised only the main gradients of the genetic variation to generate prioritisations in a feasible amount of time.
 
@@
 
-We used data for 27 alpine plant species in the European Alps collected by the IntraBioDiv project [@r451; Figure 2a].
-This data set has previously been used to explore patterns of adaptive [eg.
-@r475; @r455] and neutral genetic variation [eg.
-@r478], and to show that species richness is not an effective surrogate for genetic diversity in conservation planning [@r486].
+We used data for 27 alpine plant species in the European Alps collected by the IntraBioDiv project ([@r451]; Figure 2a).
+This data set has previously been used to explore patterns of adaptive (eg.
+[@r475; @r455]) and neutral genetic variation (eg.
+[@r478]), and to show that species richness is not an effective surrogate for genetic diversity in conservation planning ([@r486]).
 It was collected using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx.
@@
 Project members visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals.
-They genotyped samples using AFLPs, and constructed matrices denoting the presence/absence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; see Table 1 in Alvarez _et al._ [@r478] for more information).
+They genotyped samples using AFLPs, and constructed matrices denoting the presence/absence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; for more information see [@r478], Table 1).
 Thus the data set contains information describing the genomic properties of individuals for each species present in every second grid cell.
@@
 We adopted the sampling grid used by the IntraBioDiv project as planning units to develop prioritisations, omitting cells that were not found to contain individuals belonging to the study species (`r nrow(grid.DF)` grid cells used in analysis).
-To examine the potential effects of variation in cost, we calculated the total human population density inside each planning unit [obtained at 1 km$^2$ resolution from the Global Rural-Urban Mapping Project; @r472] and used this to represent acquisition cost (Figure 2b).
-All spatial and statistical analysis were conducted in \texttt{R} [version `r paste0(R.version$major, '.', R.version$minor)` @r192].
+To examine the potential effects of variation in cost, we calculated the total human population density inside each planning unit (1 km$^2$ resolution from the Global Rural-Urban Mapping Project; [@r472]) and used this to represent acquisition cost (Figure 2b).
+All spatial and statistical analysis were conducted in \texttt{R} (version `r paste0(R.version$major, '.', R.version$minor)`; [@r192]).
 
@@
 To describe the geographic location of each planning unit, we projected the grid into an equi-distant coordinate system (Europe Equidistant Conic; $\texttt{ESRI:102031}$), calculated the centroid of each grid cell, and extracted their two-dimensional coordinates.
-To describe the environmental characteristics of each planning unit, we obtained 19 climatic layers [$30^{\prime\prime}$ resolution; @r33], projected them and the planning units into an equal-area coordinate system (Europe Lambert Conformal Conic; $\texttt{ESRI:102014}$), and computed planning unit averages for each climatic variable.
+To describe the environmental characteristics of each planning unit, we obtained 19 climatic layers ($30^{\prime\prime}$ resolution; [@r33]), projected them and the planning units into an equal-area coordinate system (Europe Lambert Conformal Conic; $\texttt{ESRI:102014}$), and computed planning unit averages for each climatic variable.
 To reduce dimensionality, for each species, we subject the climatic values associated with their planning units to a principal components analysis (PCA; Table S1).
@@
 To investigate the effectiveness of our supposed surrogates, we first needed to identify which of the sampled loci were adaptive.
-Following best practices, we used two outlier detection methods to achieve this [an individual- and a population-level method @r495].
-The basic premise underpinning such methods is that neutral loci are expected to exhibit a certain level of variation, and loci that deviate from this expectation are likely to be under selection [@r480].
+Following best practices, we used two outlier detection methods to achieve this (an individual- and a population-level method; [@r495]).
+The basic premise underpinning such methods is that neutral loci are expected to exhibit a certain level of variation, and loci that deviate from this expectation are likely to be under selection ([@r480]).
 The advantage of these methods--in contrast to environmental association analyses--is that they do not use environmental data, which would have introduced circularity into our analysis.
@@
 
-The first outlier detection method involved fitting multinominal-Dirichlet models implemented in \texttt{BayeScan} [version 2.1 @r456].
-We adopted a similar methodology to Bothwell _et al._ [-@r455] and applied it to each of the species separately.
+The first outlier detection method involved fitting multinominal-Dirichlet models implemented in \texttt{BayeScan} (version 2.1; [@r456]).
+We adopted a similar methodology to Bothwell _et al._ ([-@r455]) and applied it to each of the species separately.
 Following their methodology, we initially grouped conspecifics into genetic lineages to further minimise false-positives.
-We grouped conspecifics by fitting admixture and correlated alleles models implemented in \texttt{Structure} [version 2.3.4; @r462; `r structure.params.LST[[MODE]]$numruns` replicates per species; `r format(structure.params.LST[[MODE]]$admburnin, big.mark=',', scientific=FALSE)` admixture burnin iterations; `r format(structure.params.LST[[MODE]]$burnin, big.mark=',', scientific=FALSE)` burnin iterations; `r format(structure.params.LST[[MODE]]$numreps+structure.params.LST[[MODE]]$burnin, big.mark=',', scientific=FALSE)` total iterations] using the number of lineages previously determined by Alvarez _et al._ [-@r478] and combining replicate runs using \texttt{ClumPP} [version 1.1.2, @r459; greedy algorithm based on the $G^{\prime}$ statistic; `r format(clumpp.params.LST[[MODE]]$repeats, big.mark=',', scientific=FALSE)` iterations).
-We then ran $\texttt{Bayescan}$ for each species using these lineages (1:1 prior odds; `r bayescan.params.LST[[MODE]]$reps` replicates per species, `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r format(bayescan.params.LST[[MODE]]$burn, big.mark=',', scientific=FALSE)` burn-in iterations; `r format(bayescan.params.LST[[MODE]]$n+bayescan.params.LST[[MODE]]$burn, big.mark=',', scientific=FALSE)` total iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations) using a suitable false discovery rate [$q \leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].
+We grouped conspecifics by fitting admixture and correlated alleles models implemented in \texttt{Structure} (version 2.3.4; [@r462]; `r structure.params.LST[[MODE]]$numruns` replicates per species; `r format(structure.params.LST[[MODE]]$admburnin, big.mark=',', scientific=FALSE)` admixture burnin iterations; `r format(structure.params.LST[[MODE]]$burnin, big.mark=',', scientific=FALSE)` burnin iterations; `r format(structure.params.LST[[MODE]]$numreps+structure.params.LST[[MODE]]$burnin, big.mark=',', scientific=FALSE)` total iterations) using the number of lineages previously determined by Alvarez _et al._ ([@r478]) and combining replicate runs using \texttt{ClumPP} (version 1.1.2; [@r459]; greedy algorithm based on the $G^{\prime}$ statistic; `r format(clumpp.params.LST[[MODE]]$repeats, big.mark=',', scientific=FALSE)` iterations).
+We then ran $\texttt{Bayescan}$ for each species using these lineages (1:1 prior odds; `r bayescan.params.LST[[MODE]]$reps` replicates per species; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r format(bayescan.params.LST[[MODE]]$burn, big.mark=',', scientific=FALSE)` burn-in iterations; `r format(bayescan.params.LST[[MODE]]$n+bayescan.params.LST[[MODE]]$burn, big.mark=',', scientific=FALSE)` total iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations) using a suitable false discovery rate ($q \leq `r bayescan.params.LST[[MODE]]$fdr`$; [@r466]).
 We omitted individuals if their population membership was uncertain (maximum membership probability $< `r structure.params.LST[[MODE]]$probability.threshold`$).
 
-The second outlier detection method involved fitting PCAs to identify outlier loci [implemented in the \texttt{pcadapt R} package; @r487].
+The second outlier detection method involved fitting PCAs to identify outlier loci (implemented in the \texttt{pcadapt R} package; [@r487]).
 To enable comparisons between the two outlier detection methods, we used the same individuals in this analysis as in the \texttt{BayeScan} analysis.
 For each of the 27 species, we first imputed missing data--replacing missing values with the average frequency among conspecifics--ran a PCA, and extracted the minimum number of components needed to secure the over-arching population level variation among the loci (`r pcadapt.params.LST[[MODE]]$min.variance.explained*100` % of the variation in loci).
-We then computed _q_-values using Mahalanobis distances, and used the same false discovery rate used in the \texttt{BayeScan} analysis to identify loci under selection [using the \texttt{qvalue R} package; @r491].
+We then computed _q_-values using Mahalanobis distances, and used the same false discovery rate used in the \texttt{BayeScan} analysis to identify loci under selection (using the \texttt{qvalue R} package; [@r491]).
 
@@
 We discarded the population groupings, and partitioned species' adaptive and neutral loci into separate matrices.
-We applied non-metric multi-dimensional scaling analyses [NMDS; implemented in the \texttt{vegan R} package; @r458] using Gower distances [via the \texttt{cluster R} package; @r457] to derive continuous variables that described the main gradients of adaptive and neutral genetic variation separately for each species (Table S2).
+We applied NMDS (implemented in the \texttt{vegan R} package; [@r458]) using Gower distances (via the \texttt{cluster R} package; [@r457]) to derive continuous variables that described the main gradients of adaptive and neutral genetic variation separately for each species (Table S2).
 To ensure that the ordinations described a sufficient amount of the genetic variation, we ran successive scaling analyses with increasing dimensionality until a sufficient stress value was obtained (maximum stress value $\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for each analysis).
@@
 
-We used the unreliable representative and adequate prioritisation (URAP) formulation of reserve selection [@r465].
+We used the unreliable representative and adequate prioritisation (URAP) formulation of reserve selection ([@r465]).
 This formulation can identify the minimum number of planning units required to preserve both a proportion of the species' geographic range (using amount-based targets) and a sample of variation found across the species' range.
@@
 Solutions associated with negative values--because they secured a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis.
-We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality using the \texttt{raptr R} package [@r465].
+We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality (using the \texttt{raptr R} package [@r465]).
 
@@
 
-We then fit maximum likelihood population-effects (MLPE) models [@r494].
+We then fit maximum likelihood population-effects (MLPE) models ([@r494]).
 These models accommodate the structure of dissimilarity matrices using random effects.
@@
 They were subject to backwards step-wise term deletion routines to assess term significance.
-To assess the performance of the surrogates, we refit these models seperately for each species and computed the Cragg and Uhler's pseudo-R value [using the \texttt{pscl R} package; @r492].
+To assess the performance of the surrogates, we refit these models seperately for each species and computed the Cragg and Uhler's pseudo-R value (using the \texttt{pscl R} package; [@r492]).
 
@@
 Based on the results from our previous aim, we used `r (raptr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100)` % surrogate-based targets and `r (raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)` % genetic-based targets.
-We generated a single solution for each targets/scenario combination, except for the single-species scenario with just amount-based targets for which we generated `r format(raptr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates, big.mark=',', scientific=FALSE)` replicates because this combination had many optimal solutions.
+We generated a single solution for each target/scenario combination, except for the "single-species (equal costs) with amount-based targets" cominbation for which we generated `r format(raptr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates, big.mark=',', scientific=FALSE)` replicates because it had many optimal solutions.
 We computed the proportion of adaptive (if present) and neutral genetic variation secured for each species in the prioritisations.
 
-We investigated the benefits of using surrogate-based targets in prioritisations using generalised linear mixed-effects models fit with logit links [GLMMs; using the \texttt{lme4 R} package; @r469].
+We investigated the benefits of using surrogate-based targets in prioritisations using generalised linear mixed-effects models fit with logit links (GLMMs; using the \texttt{lme4 R} package; [@r469]).
 We fit a full model to the proportion of genetic variation secured in a given prioritisation.
@@
 Prioritisations generated for the same species were accomodated using a random intercept term.
-We subject the full model to a backwards step-wise term deletion routine to assess term significance, and conducted a post-hoc analysis on the minimal adequate model using Tukey contrasts with a Bonferroni correction [using the \texttt{multcomp R} package; @r470].
+We subject the full model to a backwards step-wise term deletion routine to assess term significance, and conducted a post-hoc analysis on the minimal adequate model using Tukey contrasts with a Bonferroni correction (using the \texttt{multcomp R} package; [@r470]).
 

### 2016-10-11 code/rmarkdown/supporting_information.Rmd
+ fix refs + use one-sided tests in post hoc for second analysis

@@
 Squares represent planning units.
-For a given species, planning units that were found to be inhabited are denoted with bright blue.')}
+For a given species, planning units that it was found in are shown in bright blue.')}
 ## plot map of species distributions
@@
 Briefly, two generalized linear models were fit to correlate the proportion of adaptive genetic variation and environmental variation secured, and the proportion of neutral genetic variation and geographic variation secured in a prioritization.
-These models also included a predictor variable indicating the species for which the prioritizations were generated, and an interaction term.
-A post-hoc analysis was conducted to determine if the slopes for each species were significantly different to zero (detailed in the \\textit{Z} and \\textit{P} columns).
-All tests used a single degree of freedom.
-Cragg and Uhler\'s pseudo-R$^2$ values were calculated to show the proportion of variation explained by the slopes for each species ($R^2$ column).'
+These models also included a predictor variable indicating the species for which the prioritisations were generated, and an interaction term.
+Table shows the results from one-sided _post-hoc_ analyses that were conducted to determine if the slopes for each species were positive (using the alternative hypothesis that slopes were less than or equal to zero; tests used a single degree of freedom; results reported in the \\textit{Z} and \\textit{P} columns).
+The table also shows Cragg and Uhler\'s pseudo-R$^2$ values that were calculated to show the proportion of variation explained by the slopes for each species ($R^2$ column).'
 	)

### 2016-10-12 code/rmarkdown/article.Rmd
+ change github username

@@
 RAF has an Australian Research Council Future Fellowship.
-This work was supported by the Centre of Biodiversity and Conservation Science.
+This work was supported by the Centre of Biodiversity and Conservation Science (CBCS).
 keywords:
@@
 We used data for 27 alpine plant species in the European Alps collected by the IntraBioDiv project ([@r451]; Figure 2a).
+All data, code, and results are stored in an online repository (\url{www.github.com/jeffreyhanson/genetic.surrogates}) to replicate and validate this study.
 This data set has previously been used to explore patterns of adaptive (eg.

### 2016-10-12 code/rmarkdown/supporting_information.Rmd
+ change github username

@@
 # Supporting Information
-All data and code is stored in an online repository (\url{www.github.com/paleo13/genetic.surrogates}) to replicate analyses in this study. See the \textit{README} file for instructions.
+All data and code is stored in an online repository (\url{www.github.com/jeffreyhanson/genetic.surrogates}) to replicate analyses in this study. See the \textit{README} file for instructions.
 

### 2016-11-10 code/rmarkdown/article.Rmd
+ add diagrams to supplementary figures

@@
 
-We created environmental and geographic surrogate variables for each species.
+We created environmental and geographic surrogate variables for each species (Figures S59--S561).
 To describe the geographic location of each planning unit, we projected the grid into an equi-distant coordinate system (Europe Equidistant Conic; $\texttt{ESRI:102031}$), calculated the centroid of each grid cell, and extracted their two-dimensional coordinates.
@@
 
-To investigate the effectiveness of our supposed surrogates, we first needed to identify which of the sampled loci were adaptive.
+To investigate the effectiveness of our supposed surrogates, we quantified the main spatial patterns of adaptive and neutral genetic variation (Figures S59, S562).
+To do this, we first needed to identify which of the sampled loci were adaptive.
 Following best practices, we used two outlier detection methods to achieve this (an individual- and a population-level method; [@r495]).

### 2016-11-10 code/rmarkdown/supporting_information.Rmd
+ add diagrams to supplementary figures

@@
 
+```{r, include=FALSE}
+render_gvis <- function(x) {
+  file = tempfile(fileext='.png')
+  rsvg::rsvg_png(charToRaw(DiagrammeRsvg::export_svg(x)), file)
+  grid::grid.raster(png::readPNG(file))
+}
+```
+
+```{r, echo=FALSE, results='hide', message=FALSE, fig.cap='Overall data processing workflow.
+The flowcharts show the methods used to generate the environmental and geographic surrogate spaces, and the adaptive and neutral genetic spaces.
+Squares indicate data sets.
+Ellipses indicate analyses.
+Lines show relationships between data and analysis, and their captions describe how data was processed.'}
+render_gvis(grViz("
+digraph main {
+  graph [layout = dot]
+
+  node [shape = rectangle, style = filled, color = darkslategray, fillcolor = white]
+
+  meirmans_data [label = <Plant data<BR></BR>(Meirmans <I>et al. </I> 2011)>]
+  bioclimatic_data [label = <Bioclimatic data<BR></BR>(Hijmans <I> et al. </I> 2005)>]
+  sample_location_data [label = 'Sample location data']
+  grid_data [label = 'Grid cell data']
+  sample_aflp_data [label = 'Sample AFLP data']
+
+  geo_as [label = 'Geographic attribute spaces'] 
+  env_as [label = 'Environmental attribute spaces']
+  neutral_as [label = 'Neutral genetic attribute spaces']
+  adaptive_as [label = 'Adaptive genetic attribute spaces']
+
+  node[shape = ellipse, style = filled, color = darkslategray, fillcolor = white]
+  calc_geo_as [label = 'Processing\n(Figure SXXX)'] 
+  calc_env_as [label = 'Processing\n(Figure SXXX)']
+  calc_genetic_as [label = 'Processing\n(Figure SXXX)']
+
+  edge [color = grey]
+  meirmans_data -> {sample_aflp_data grid_data sample_location_data}
+  bioclimatic_data -> calc_env_as
+  sample_location_data -> {calc_geo_as calc_env_as calc_genetic_as}
+  grid_data -> {calc_geo_as calc_env_as calc_genetic_as}
+  sample_aflp_data -> calc_genetic_as
+  calc_genetic_as -> {neutral_as, adaptive_as}
+  calc_geo_as -> geo_as
+  calc_env_as -> env_as
+  calc_env_as -> calc_env_as [label = ' each species\nseparately']
+  calc_geo_as -> calc_geo_as [label = ' each species\nseparately']
+  calc_genetic_as -> calc_genetic_as [label = ' each species\nseparately']
+
+}
+"))
+```
+
+```{r, echo=FALSE, results='hide', message=FALSE, fig.cap='Geographic surrogate processing workflow. Conventions are detailed in Figure S59 caption.'}
+render_gvis(grViz("
+digraph geo_as {
+  graph [layout = dot]
+
+  node [shape = rectangle, style = filled, color = darkslategray, fillcolor = white]
+
+  # input
+  sample_location_data [label = 'Sample location data']
+  grid_data [label = 'Grid cell data']
+  species_name [label = 'Species name']
+  {rank = same; sample_location_data; grid_data; species_name}
+
+  # intermediate
+  spp_samples [label = 'Locations of samples for species']
+  spp_grids [label = 'Grid cells occupied by species']
+  spp_proj_data [label = 'Projected grid cells']
+  spp_centriod_data [label = 'Centroids of grid cells']
+
+  # output
+  geographic_space [label = 'Geographic attribute space']
+
+  edge [color = grey]
+
+  species_name -> spp_samples 
+  sample_location_data -> spp_samples [label = ' subset data']
+  spp_samples -> spp_grids [label = ' overlay data']
+  grid_data -> spp_grids 
+  spp_grids -> spp_proj_data [label = ' reproject data to\nequi-distant CRS']
+  spp_proj_data -> spp_centriod_data [label = ' compute centroids']
+  spp_centriod_data -> geographic_space [label =' extract coordinates and use them\nto represent demand points and\nplanning unit coordinates']
+
+}
+"))
+```
+
+```{r, fig.height=6, echo=FALSE, results='hide', message=FALSE, fig.cap='Environmental surrogate processing workflow. Conventions are detailed in Figure S59 caption.'}
+render_gvis(grViz("
+digraph env_as {
+  graph [layout = dot]
+
+  node[shape = rectangle, style = filled, color = darkslategray, fillcolor = white]
+
+  # input
+  sample_location_data [label = 'Sample location data']
+  grid_data [label = 'Grid cell data']
+  species_name [label = 'Species name']
+  bioclimatic_data [label = <Bioclimatic data<BR></BR>(Hijmans <I> et al. </I> 2005)>]
+  {rank = same; sample_location_data; grid_data; species_name; bioclimatic_data}
+
+  # intermediate
+  spp_samples [label = 'Locations of samples for species']
+  spp_grids [label = 'Grid cells occupied by species']
+  spp_proj_data [label = 'Projected grid cells']
+
+  bioclimatic_proj_data [label = 'Projected bioclimatic data']
+  clipped_bioclimatic_data [label = 'Bioclimatic data clipped to species distribution']
+  pca_scores_data [label = 'Main axes of bioclimatic variation\nacross species distribution']
+  grid_pca_data [label = 'Environmental conditions inside each grid cell']
+
+  # output
+  environmental_space [label = 'Environmental attribute space']
+
+  # analyses
+  node[shape = ellipse, style = filled, color = darkslategray, fillcolor = white]  
+  pca [label = 'Principal components analysis']
+
+  edge [color = grey]
+  species_name -> spp_samples 
+  sample_location_data -> spp_samples [label = ' subset data']
+  grid_data -> spp_grids
+  bioclimatic_data -> bioclimatic_proj_data [label = ' reproject data to\nequal-area CRS']
+  spp_samples -> spp_grids [label = ' overlay data']
+  spp_grids -> spp_proj_data [label = ' reproject data to\nequal-area CRS']
+  bioclimatic_proj_data -> clipped_bioclimatic_data [label = ' clip data']
+  spp_proj_data -> clipped_bioclimatic_data
+  clipped_bioclimatic_data -> pca
+  pca -> pca_scores_data [label = ' extract scores for\nfirst two components']
+  spp_proj_data -> grid_pca_data
+  pca_scores_data -> grid_pca_data [label = ' calculate average scores\nfor each grid cell']
+  grid_pca_data -> environmental_space [label =' extract scores and use them\nto represent demand points and\nplanning unit coordinates']
+
+}
+"))
+```
+
+```{r, fig.height=10, echo=FALSE, results='hide', message=FALSE, fig.cap='Genetic data processing workflow. Conventions are detailed in Figure S59 caption.'}
+render_gvis(grViz("
+digraph genetic_as {
+  graph [layout = dot]
+
+  node[shape = rectangle, style = filled, color = darkslategray, fillcolor = white]
+
+  # input
+  sample_location_data [label = 'Sample location data']
+  sample_aflp_data [label = 'Sample AFLP data']
+  grid_data [label = 'Grid cell data']
+  species_name [label = 'Species name']
+  {rank = same; sample_location_data; grid_data; species_name; sample_aflp_data}
+
+  # intermediate
+  spp_aflp_samples [label = 'Species AFLP data']
+  spp_location_samples [label = 'Species location data']
+  spp_grids [label = 'Grid cells occupied by species']
+
+  population_data [label = 'Samples grouped into populations']
+  subset_data [label = 'Samples with reliable population designations']
+  rare_alleles [label = 'AFLP data for\nrare alleles']
+  remainder_alleles [label = 'AFLP data for\n standard alleles']
+  bayescan_adaptive_alleles [label = ' Outlier loci']
+  pcadapt_adaptive_alleles [label = ' Outlier loci']
+  bayescan_neutral_alleles [label = ' Standard loci']
+  pcadapt_neutral_alleles [label = ' Standard loci']
+  adaptive_alleles [label = ' AFLP data\nfor adaptive loci']
+  neutral_alleles [label = ' AFLP data\nfor neutral loci']
+  adaptive_ordinations [label = ' Ordinations showing main\ngradients of variation\nin adaptive loci']
+  neutral_ordinations [label = '  Ordinations showing main\ngradients of variation\nin neutral loci']
+  grid_adaptive_ordinations [label = ' Typical adaptive genetic characteristics of individuals in grid cells']
+  grid_neutral_ordinations [label = ' Typical neutral genetic characteristics of individuals in grid cells']
+
+  # output
+  adaptive_genetic_space [label = 'Adaptive genetic space']
+  neutral_genetic_space  [label = 'Neutral genetic space']
+
+  # analyses
+  node[shape = ellipse, style = filled, color = darkslategray, fillcolor = white]
+  bayescan_analysis [label = 'BayeScan analysis']
+  pcadapt_analysis [label = 'PCadapt analysis']
+  structure_analysis [label = 'Structure analysis']
+  clumpp_analysis [label = 'ClumPP analysis']
+  neutral_nmds_analysis [label = 'NMDS analyses']
+  adaptive_nmds_analysis [label = 'NMDS analyses']
+  combine_results [label = 'Combine results from all analyses.\nLoci classified as adaptive in both analyses are treated as adaptive,\notherwise treated as neutral']
+
+  edge [color = grey]
+  sample_location_data -> spp_location_samples [label = ' subset data']
+  species_name -> spp_location_samples 
+  sample_aflp_data -> spp_aflp_samples [label = ' subset data']
+  species_name -> spp_aflp_samples
+  spp_location_samples -> spp_grids
+  grid_data -> spp_grids [label = ' overlay data']
+  spp_aflp_samples -> structure_analysis
+  structure_analysis -> clumpp_analysis [label = ' post-process\nresults']
+  clumpp_analysis -> population_data
+  population_data -> subset_data [label = ' omit samples within\nuncertain memberships']
+  subset_data -> rare_alleles [label = ' extract rare alleles where\nglobal frequency of minor\nalleles < 10 %']
+  subset_data -> remainder_alleles
+  remainder_alleles -> bayescan_analysis [label = 'use population\ngroupings']
+  remainder_alleles -> pcadapt_analysis [label = 'ignore population\ngroupings']
+  bayescan_analysis -> bayescan_adaptive_alleles [label = ' extract']
+  pcadapt_analysis -> pcadapt_adaptive_alleles  [label = ' extract']
+  bayescan_analysis -> bayescan_neutral_alleles [label = ' extract']
+  pcadapt_analysis -> pcadapt_neutral_alleles [label = ' extract']
+  bayescan_adaptive_alleles -> combine_results [label = ' treat loci\nas adaptive']
+  pcadapt_adaptive_alleles -> combine_results [label = ' treat loci\nas adaptive']
+  bayescan_neutral_alleles -> combine_results [label = ' treat loci\nas neutral']
+  pcadapt_neutral_alleles -> combine_results [label = ' treat loci\nas neutral']
+  rare_alleles -> combine_results [label = ' treat loci\nas neutral']
+  combine_results -> adaptive_alleles
+  combine_results -> neutral_alleles 
+  neutral_alleles  -> neutral_nmds_analysis
+  adaptive_alleles -> adaptive_nmds_analysis
+  adaptive_nmds_analysis -> adaptive_ordinations [label = ' extract']
+  neutral_nmds_analysis  -> neutral_ordinations [label = ' extract']
+  spp_grids  -> grid_adaptive_ordinations
+  adaptive_ordinations -> grid_adaptive_ordinations [label = ' calculate average values\nin each grid cell']
+  spp_grids  -> grid_neutral_ordinations
+  neutral_ordinations -> grid_neutral_ordinations [label = ' calculate average values\nin each grid cell']
+  grid_adaptive_ordinations -> adaptive_genetic_space [label =' extract values and use them\nto represent demand points and\nplanning unit coordinates']
+  grid_neutral_ordinations -> neutral_genetic_space [label =' extract values and use them\nto represent demand points and\nplanning unit coordinates']
+
+}
+"))
+```
+
 # Tables

### 2016-11-10 code/R/analysis/00-initialization.R

@@ -24,6 +24,9 @@ library(qvalue)
 
 # load CRAN packages
 library(stats)
+library(DiagrammaR)
+library(rsvg)
+library(png)
 library(withr)
 library(Hmisc)
 library(lme4)

### 2016-11-10 code/rmarkdown/article.Rmd
+ update figure numberings

@@
 
-We created environmental and geographic surrogate variables for each species (Figures S59--S561).
+We created environmental and geographic surrogate variables for each species (see Figures S59--S561 for workflow).
 To describe the geographic location of each planning unit, we projected the grid into an equi-distant coordinate system (Europe Equidistant Conic; $\texttt{ESRI:102031}$), calculated the centroid of each grid cell, and extracted their two-dimensional coordinates.
@@
 
-To investigate the effectiveness of our supposed surrogates, we quantified the main spatial patterns of adaptive and neutral genetic variation (Figures S59, S562).
+To investigate the effectiveness of our supposed surrogates, we quantified the main spatial patterns of adaptive and neutral genetic variation (see Figures S59 and S62 for workflow).
 To do this, we first needed to identify which of the sampled loci were adaptive.

### 2016-11-10 code/rmarkdown/supporting_information.Rmd
+ update figure numberings

@@
   node[shape = ellipse, style = filled, color = darkslategray, fillcolor = white]
-  calc_geo_as [label = 'Processing\n(Figure SXXX)'] 
-  calc_env_as [label = 'Processing\n(Figure SXXX)']
-  calc_genetic_as [label = 'Processing\n(Figure SXXX)']
+  calc_geo_as [label = 'Processing\n(Figure S60)'] 
+  calc_env_as [label = 'Processing\n(Figure S61)']
+  calc_genetic_as [label = 'Processing\n(Figure S62)']
 

### 2016-11-10 code/rmarkdown/article.Rmd
+ reorder si figures and include env spaces

@@
 <!-- scenario analysis -->
-To determine whether using environmental and geographic targets could improve the effectiveness of prioritisations in representing genetic variation, we generated prioritisations using (i) "amount targets" reflecting the traditional approach of representing a certain proportion of each species' geographic distribution, (ii) "amount and surrogate targets" where we also targeted representation of the environmental and geographic surrogates as well as a certain proportion of each species' geographic distribution, and (iii) "amount and genetic targets" where we targeted representation of the directly measured genetic variation as well as a certain proportion of each species' geographic distribution (Figure 3).
-To determine if these patterns were robust, for each of the three combinations of the targets, we generated prioritisations under four scenarios: (i) single-species with equal costs, (ii) single species with opportunity costs, (iii) multi-species with equal costs, and (iv) multi-species with opportunity costs.
+To determine whether using environmental and geographic targets could improve the effectiveness of prioritisations in representing genetic variation, we generated prioritisations using (i) "amount targets" reflecting the traditional approach of representing a certain proportion of each species' geographic distribution, (ii) "amount and surrogate targets" where we also targeted representation of the environmental and geographic surrogates as well as a certain proportion of each species' geographic distribution, and (iii) "amount and genetic targets" where we targeted representation of the directly measured genetic variation as well as a certain proportion of each species' geographic distribution.
+To determine if these patterns were robust, for each of the three combinations of the targets, we generated prioritisations under four scenarios: (i) single-species with equal costs, (ii) single species with opportunity costs, (iii) multi-species with equal costs, and (iv) multi-species with opportunity costs (Figure 3, Figure SXXX).
 

### 2016-11-10 code/rmarkdown/supporting_information.Rmd
+ reorder si figures and include env spaces

@@
 # Figures
+
+```{r, echo=FALSE, include=FALSE}
+plot.spp.spaces <- function(i) {
+  ## prepare data
+  # get vector of adaptive and neutral column names
+  adapt.col <- grep(
+    paste0(unique(spp.samples.DF$species)[i], '_adaptive_d'),
+    names(grid.FPLY), value=TRUE, fixed=TRUE
+  )
+  neutral.col <- grep(
+    paste0(unique(spp.samples.DF$species)[i], '_neutral_d'),
+    names(grid.FPLY), value=TRUE, fixed=TRUE
+  )
+  env.col <- grep(
+    paste0(unique(spp.samples.DF$species)[i], '_env_d'),
+    names(grid.FPLY), value=TRUE, fixed=TRUE
+  )
+  # skip species if not used in analysis
+  curr.FPLY <- grid.FPLY[,
+    c(adapt.col, neutral.col, env.col,
+    'long','lat','group'),drop=FALSE]
+  curr.FPLY <- curr.FPLY[which(!is.na(curr.FPLY[[1]])),]
+  curr.FPLY <- curr.FPLY %>% gather(d, value, contains('adaptive'), contains('neutral'), contains('env')) %>%
+    mutate(
+      d=gsub(paste0(unique(spp.samples.DF$species)[i], '_'), '', d, fixed=TRUE),
+      d=gsub('_', ' ', d, fixed=TRUE))
+  for (j in seq_len(length(adapt.col)+length(neutral.col)+length(env.col)))
+    curr.FPLY[['d']] <- gsub(as.character(j), paste0('(', j, ')'), curr.FPLY[['d']], fixed=TRUE)
+  curr.FPLY <- curr.FPLY %>% mutate(d=factor(gsub(' d(', ' (', d, fixed=TRUE)))
+  # make plot
+  curr.GROBS <- llply(unique(curr.FPLY$d), function(x) {
+    ggplot() +
+      geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
+        fill='grey85', color='grey70') +
+      geom_polygon(data=filter(curr.FPLY, d==x), aes_string(x='long', y='lat', 
+        group='group', fill='value'), color='grey10') +
+      theme_classic() +
+      theme(
+        axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"), axis.title=element_blank(),
+        panel.border=element_rect(color='black', fill=NA, size=1),
+        strip.background = element_rect(fill='grey20'), strip.text = element_text(color='white', size=20)
+      ) +
+      coord_cartesian(
+        xlim=buffered.range(grid.FPLY$long, 0),
+        ylim=buffered.range(grid.FPLY$lat, 0)
+      ) +
+      scale_fill_gradientn(name='',
+        colors=brewer.pal(9, 'RdYlBu'),
+        guide=guide_colorbar(ticks=element_line(color='black'),
+          border=element_line(color='black'), barheight=unit(3.0, 'cm'))
+      ) + facet_wrap(~d)
+  })
+  do.call(grid.arrange, append(curr.GROBS, list(ncol=3)))
+  invisible()
+}
+
+calc.fig.height <- function(i) {
+  # init
+  adapt.col <- paste0(unique(spp.samples.DF$species)[i], '_adaptive_d1')
+  neutral.col <- paste0(unique(spp.samples.DF$species)[i], '_neutral_d1')
+  env.col <- grepl(paste0(unique(spp.samples.DF$species)[i], '_env_d'), names(grid.FPLY), value=TRUE, fixed=TRUE) %>%
+    sum()
+
+  n.plots <- 0 
+  # count adaptive plot
+  if (adapt.col %in% names(grid.FPLY))
+    n.plots <- n.plots + spp.nmds.LST[[i]][['adaptive']]$ndim
+  # count neutral plot
+  if (neutral.col %in% names(grid.FPLY))
+    n.plots <- n.plots + spp.nmds.LST[[i]][['neutral']]$ndim
+  # count env plot
+  n.plots <- n.plots + env.col
+  # exports
+  if (n.plots >= 3) return((((n.plots-1) %/% 3)+1) * 2)
+  return(1.85)
+}
+
+calc.fig.width <- function(i) {
+  # init
+  adapt.col <- paste0(unique(spp.samples.DF$species)[i], '_adaptive_d1')
+  neutral.col <- paste0(unique(spp.samples.DF$species)[i], '_neutral_d1')
+  env.col <- grepl(paste0(unique(spp.samples.DF$species)[i], '_env_d'), names(grid.FPLY), value=TRUE, fixed=TRUE) %>%
+    sum()
+  n.plots <- 0 
+  # count adaptive plot
+  if (adapt.col %in% names(grid.FPLY))
+    n.plots <- n.plots + spp.nmds.LST[[i]][['adaptive']]$ndim
+  # count neutral plot
+  if (neutral.col %in% names(grid.FPLY))
+    n.plots <- n.plots + spp.nmds.LST[[i]][['neutral']]$ndim
+  # count env plot
+  n.plots <- n.plots + env.col
+  # exports
+  if (n.plots >= 3) return(10)
+  return(4 * n.plots)
+}
+
+is.speciesPlottable <- function(i) {
+  if (i > general.params.LST[[MODE]]$n.spp)
+    return(FALSE)
+  return(!is.na(unique(spp.samples.DF$species)[i]))
+}
+```
+
+```{r, echo=FALSE, fig.width=calc.fig.width(1), fig.height=calc.fig.height(1), fig.width=calc.fig.width(1), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[1]),'}.
+Each panel shows a different dimension of the variation as indicated by its title.
+Squares represent planning units.
+The color of each planning unit panel corresponds to the average ordinated values associated with individuals sampled inside it.
+No indiviudals of this species were found inside the gray planning units.')}
+if (is.speciesPlottable(1)) plot.spp.nmds(1)
+ ```
+
+```{r, echo=FALSE, fig.width=calc.fig.width(2), fig.height=calc.fig.height(2), fig.width=calc.fig.width(2), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[2]),'}.
+See Figure S29 caption for conventions.')}
+if (is.speciesPlottable(2)) plot.spp.nmds(2)
+```
+
+```{r, echo=FALSE, fig.width=calc.fig.width(3), fig.height=calc.fig.height(3), fig.width=calc.fig.width(3), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[3]),'}.
+See Figure S29 caption for conventions.')}
+if (is.speciesPlottable(3)) plot.spp.nmds(3)
+```
+
+```{r, echo=FALSE, fig.width=calc.fig.width(4), fig.height=calc.fig.height(4), fig.width=calc.fig.width(4), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[4]),'}.
+See Figure S29 caption for conventions.')}
+if (is.speciesPlottable(4)) plot.spp.nmds(4)
+```
+
+```{r, echo=FALSE, fig.width=calc.fig.width(5), fig.height=calc.fig.height(5), fig.width=calc.fig.width(5), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[5]),'}.
+See Figure S29 caption for conventions.')}
+if (is.speciesPlottable(5)) plot.spp.nmds(5)
+```
+
+```{r, echo=FALSE, fig.width=calc.fig.width(6), fig.height=calc.fig.height(6), fig.width=calc.fig.width(6), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[6]),'}.
+See Figure S29 caption for conventions.')}
+if (is.speciesPlottable(6)) plot.spp.nmds(6)
+```
+
+```{r, echo=FALSE, fig.width=calc.fig.width(7), fig.height=calc.fig.height(7), fig.width=calc.fig.width(7), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[7]),'}.
+See Figure S29 caption for conventions.')}
+if (is.speciesPlottable(7)) plot.spp.nmds(7)
+```
+
+```{r, echo=FALSE, fig.width=calc.fig.width(8), fig.height=calc.fig.height(8), fig.width=calc.fig.width(8), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[8]),'}.
+See Figure S29 caption for conventions.')}
+if (is.speciesPlottable(8)) plot.spp.nmds(8)
+```
+
+```{r, echo=FALSE, fig.width=calc.fig.width(9), fig.height=calc.fig.height(9), fig.width=calc.fig.width(9), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[9]),'}.
+See Figure S29 caption for conventions.')}
+if (is.speciesPlottable(9)) plot.spp.nmds(9)
+```
+
+```{r, echo=FALSE, fig.width=calc.fig.width(10), fig.height=calc.fig.height(10), fig.width=calc.fig.width(10), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[10]),'}.
+See Figure S29 caption for conventions.')}
+if (is.speciesPlottable(10)) plot.spp.nmds(10)
+```
+
+```{r, echo=FALSE, fig.width=calc.fig.width(11), fig.height=calc.fig.height(11), fig.width=calc.fig.width(11), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[11]),'}.
+See Figure S29 caption for conventions.')}
+if (is.speciesPlottable(11)) plot.spp.nmds(11)
+```
+
+```{r, echo=FALSE, fig.width=calc.fig.width(12), fig.height=calc.fig.height(12), fig.width=calc.fig.width(12), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[12]),'}.
+See Figure S29 caption for conventions.')}
+if (is.speciesPlottable(12)) plot.spp.nmds(12)
+```
+
+```{r, echo=FALSE, fig.width=calc.fig.width(13), fig.height=calc.fig.height(13), fig.width=calc.fig.width(13), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[13]),'}.
+See Figure S29 caption for conventions.')}
+if (is.speciesPlottable(13)) plot.spp.nmds(13)
+```
+
+```{r, echo=FALSE, fig.width=calc.fig.width(14), fig.height=calc.fig.height(14), fig.width=calc.fig.width(14), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[14]),'}.
+See Figure S29 caption for conventions.')}
+if (is.speciesPlottable(14)) plot.spp.nmds(14)
+```
+
+```{r, echo=FALSE, fig.width=calc.fig.width(15), fig.height=calc.fig.height(15), fig.width=calc.fig.width(15), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[15]),'}.
+See Figure S29 caption for conventions.')}
+if (is.speciesPlottable(15)) plot.spp.nmds(15)
+```
+
+```{r, echo=FALSE, fig.width=calc.fig.width(16), fig.height=calc.fig.height(16), fig.width=calc.fig.width(16), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[16]),'}.
+See Figure S29 caption for conventions.')}
+if (is.speciesPlottable(16)) plot.spp.nmds(16)
+```
+
+```{r, echo=FALSE, fig.width=calc.fig.width(17), fig.height=calc.fig.height(17), fig.width=calc.fig.width(17), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[17]),'}.
+See Figure S29 caption for conventions.')}
+if (is.speciesPlottable(17)) plot.spp.nmds(17)
+```
+
+```{r, echo=FALSE, fig.width=calc.fig.width(18), fig.height=calc.fig.height(18), fig.width=calc.fig.width(18), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[18]),'}.
+See Figure S29 caption for conventions.')}
+if (is.speciesPlottable(18)) plot.spp.nmds(18)
+```
+
+```{r, echo=FALSE, fig.width=calc.fig.width(19), fig.height=calc.fig.height(19), fig.width=calc.fig.width(19), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[19]),'}.
+See Figure S29 caption for conventions.')}
+if (is.speciesPlottable(19)) plot.spp.nmds(19)
+```
+
+```{r, echo=FALSE, fig.width=calc.fig.width(20), fig.height=calc.fig.height(20), fig.width=calc.fig.width(20), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[20]),'}.
+See Figure S29 caption for conventions.')}
+if (is.speciesPlottable(20)) plot.spp.nmds(20)
+```
+
+```{r, echo=FALSE, fig.width=calc.fig.width(21), fig.height=calc.fig.height(21), fig.width=calc.fig.width(21), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[21]),'}.
+See Figure S29 caption for conventions.')}
+if (is.speciesPlottable(21)) plot.spp.nmds(21)
+```
+
+```{r, echo=FALSE, fig.width=calc.fig.width(22), fig.height=calc.fig.height(22), fig.width=calc.fig.width(22), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[22]),'}.
+See Figure S29 caption for conventions.')}
+if (is.speciesPlottable(22)) plot.spp.nmds(22)
+```
+
+```{r, echo=FALSE, fig.width=calc.fig.width(23), fig.height=calc.fig.height(23), fig.width=calc.fig.width(23), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[23]),'}.
+See Figure S29 caption for conventions.')}
+if (is.speciesPlottable(23)) plot.spp.nmds(23)
+```
+
+```{r, echo=FALSE, fig.width=calc.fig.width(24), fig.height=calc.fig.height(24), fig.width=calc.fig.width(24), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[24]),'}.
+See Figure S29 caption for conventions.')}
+if (is.speciesPlottable(24)) plot.spp.nmds(24)
+```
+
+```{r, echo=FALSE, fig.width=calc.fig.width(25), fig.height=calc.fig.height(25), fig.width=calc.fig.width(25), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[25]),'}.
+See Figure S29 caption for conventions.')}
+if (is.speciesPlottable(25)) plot.spp.nmds(25)
+```
+
+```{r, echo=FALSE, fig.width=calc.fig.width(26), fig.height=calc.fig.height(26), fig.width=calc.fig.width(26), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[26]),'}.
+See Figure S29 caption for conventions.')}
+if (is.speciesPlottable(26)) plot.spp.nmds(26)
+```
+
+```{r, echo=FALSE, fig.width=calc.fig.width(27), fig.height=calc.fig.height(27), fig.width=calc.fig.width(27), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[27]),'}.
+See Figure S29 caption for conventions.')}
+if (is.speciesPlottable(27)) plot.spp.nmds(27)
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=(2.5*4), fig.cap='The relationship between the proportion of adaptive genetic variation secured in a prioritization and the proportion of environmental variation it also secures.
+Points represent values associated with randomly generated prioritizations.
+Blue lines indicate average trends computed using a generalised linear model with a logit link.
+Each panel shows data for a different species.'}
+# format data
+curr.DF <- env.surrogacy.DF %>% filter(!is.na(genetic.held)) %>% mutate(Species=gsub('_', ' ', Species))
+curr.DF2 <- env.surrogate.spp.DF %>% filter(!is.na(r2CU)) %>% mutate(Species=gsub('_', ' ', Species))
+curr.DF2$R2_repr <- round(curr.DF2$r2CU,2)
+curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
+curr.DF2$surrogate.held <- 0
+curr.DF2$genetic.held <- 0.9
+
+# make plot
+suppressWarnings(
+  ggplot(aes(x=surrogate.held, y=genetic.held), data=curr.DF) +
+    theme_classic() +
+    theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=12),
+      panel.margin.x=unit(1, 'lines')) +
+    xlab('Environmental variation secured (%)') +
+    ylab('Adaptive genetic variation secured (%)') +
+    geom_hex() +
+    suppressWarnings(stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
+      se=FALSE, fill='blue', alpha=0.8)) +
+    geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=5, color='red') +
+    facet_wrap(~ Species, ncol=4)
+)
+```
+
+```{r, include=FALSE}
+calc.fig.height3 <- function(x) {
+  return((((x-1) %/% 4)+1)*1.5)
+}
+
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=calc.fig.height3(n_distinct(geo.surrogacy.DF$Species)), fig.cap='The relationship between the proportion of neutral genetic variation secured in a prioritizaiton and the proportion of geographic variation it also secures.
+Conventions are detailed in Figure S56.'}
+# format data
+curr.DF <- geo.surrogacy.DF %>% filter(!is.na(genetic.held)) %>% mutate(Species=gsub('_', ' ', Species))
+curr.DF2 <- geo.surrogate.spp.DF %>% filter(!is.na(r2CU)) %>% mutate(Species=gsub('_', ' ', Species))
+curr.DF2$R2_repr <- round(curr.DF2$r2CU,2)
+curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
+curr.DF2$surrogate.held <- 0
+curr.DF2$genetic.held <- 0.9
+
+# make plot
+suppressWarnings(
+  ggplot(aes(x=surrogate.held, y=genetic.held), data=curr.DF) +
+    theme_classic() +
+    theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=12),
+      panel.margin.x=unit(1, 'lines')) +
+    xlab('Geographic variation secured (%)') +
+    ylab('Neutral genetic variation secured (%)') +
+    geom_hex() +
+    suppressWarnings(stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
+      se=FALSE, fill='blue', alpha=0.8)) +
+    geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=5, color='red') +
+    facet_wrap(~ Species, ncol=4)
+)
+```
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=(5*0.9), fig.height=(6.5*0.8), out.width='3.42in', height='3.9in', fig.cap='Planning unit selection frequencies.
+Squares represent planning units, and their colors indicate their selection frequency.
+Each column of panels shows solutions generated using different targets.
+Each row of panels shows solutions generated under different scenarios.
+Note that the selection frequencies for the single-species prioritizations are based on their frequency among prioritizations generated for each species.
+The selection frequencies for the multi-species prioritizations are each based on a single optimal prioritization for all species and so are binary.'}
+# prepare data for plotting
+grid.sub.FPLY <- grid.sub.PLY
+for (i in seq_along(single.spp.prioritisations[[1]]))
+  grid.sub.FPLY@data[[paste0('Single-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations, function(x){colMeans(x[[i]]@results@selections)}))
+for (i in seq_along(single.spp.prioritisations.with.cost[[1]]))
+  grid.sub.FPLY@data[[paste0('Single-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations.with.cost, function(x){colMeans(x[[i]]@results@selections)}))
+for (i in seq_along(multi.spp.prioritisations))
+  grid.sub.FPLY@data[[paste0('Multi-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations[[i]]@results@selections)
+for (i in seq_along(multi.spp.prioritisations.with.cost))
+  grid.sub.FPLY@data[[paste0('Multi-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations.with.cost[[i]]@results@selections)
+grid.sub.FPLY <- spFortify(grid.sub.FPLY)
+grid.sub.FPLY <- grid.sub.FPLY %>% gather(
+    Combination,
+    `Selection.Frequency`,
+    `Single-species (equal costs).Amount`:`Multi-species (opportunity costs).Genetic`
+  ) %>% mutate(
+    Context=sapply(strsplit(as.character(Combination), '.', fixed=TRUE), `[[`, 1),
+    Prioritization=sapply(strsplit(as.character(Combination), '.', fixed=TRUE), `[[`, 2)
+  ) %>% mutate(
+    Prioritization=revalue(Prioritization, c('Amount'='Amount targets', 'Surrogate'='Amount & surrogate\ntargets', 'Genetic'='Amount & genetic\ntargets')),
+    Prioritization=factor(as.character(Prioritization), levels=c('Amount targets', 'Amount & surrogate\ntargets', 'Amount & genetic\ntargets')),
+    Context=factor(
+      gsub(' (', '\n(', Context, fixed=TRUE),
+      levels=c("Single-species\n(equal costs)", "Single-species\n(opportunity costs)",
+        "Multi-species\n(equal costs)", "Multi-species\n(opportunity costs)"))
+  )
+
+# make maps
+ggplot() +
+  geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
+    fill='grey85', color='grey70') +
+  geom_polygon(data=grid.sub.FPLY, aes_string(x='long', y='lat', 
+    group='group', fill='Selection.Frequency'),
+    alpha=0.8, color=NA) +
+  theme_classic() +
+  theme(axis.ticks=element_blank(), axis.text=element_blank(),
+    plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(),
+    strip.background = element_rect(fill='grey20'),
+    strip.text = element_text(color='white', size=8),
+    legend.position='bottom', legend.key.width=unit(1.8, 'cm'),
+    legend.title=element_text(vjust=-5),
+    legend.margin=unit(c(0.1), 'cm'),
+    axis.title=element_blank()) +
+  coord_cartesian(
+    xlim=buffered.range(grid.sub.FPLY$long, 0),
+    ylim=buffered.range(grid.sub.FPLY$lat, 0)
+  ) +
+  facet_grid(Context ~ Prioritization) +
+  scale_fill_gradient(name='Selection Frequency (%)',
+    guide=guide_colorbar(title.position='bottom', title.hjust=0.5,
+      ticks=element_line(color='black'), border=element_line(color='black'))
+  )
+```
 
@@
 
-```{r, fig.width=15, fig.height=3, fig.cap='Climatic variation.
-Each panel depicts variation based on a principle component (PC) for a different species.
-Squares represent planning units.
-The color of each planning unit denotes the average PC value of pixels inside it.
-For a given species, the planning units with more similar colors have more similar climate regimes.', echo=FALSE}
-plot(1)
+```{r, include=FALSE}
+render_gvis <- function(x) {
+  file = tempfile(fileext='.png')
+  rsvg::rsvg_png(charToRaw(DiagrammeRsvg::export_svg(x)), file)
+  grid::grid.raster(png::readPNG(file))
+}
+```
+
+```{r, echo=FALSE, results='hide', message=FALSE, fig.cap='Overall data processing workflow.
+The flowcharts show the methods used to generate the environmental and geographic surrogate spaces, and the adaptive and neutral genetic spaces.
+Squares indicate data sets.
+Ellipses indicate analyses.
+Lines show relationships between data and analysis, and their captions describe how data was processed.'}
+render_gvis(grViz("
+digraph main {
+  graph [layout = dot]
+
+  node [shape = rectangle, style = filled, color = darkslategray, fillcolor = white]
+
+  meirmans_data [label = <Plant data<BR></BR>(Meirmans <I>et al. </I> 2011)>]
+  bioclimatic_data [label = <Bioclimatic data<BR></BR>(Hijmans <I> et al. </I> 2005)>]
+  sample_location_data [label = 'Sample location data']
+  grid_data [label = 'Grid cell data']
+  sample_aflp_data [label = 'Sample AFLP data']
+
+  geo_as [label = 'Geographic attribute spaces'] 
+  env_as [label = 'Environmental attribute spaces']
+  neutral_as [label = 'Neutral genetic attribute spaces']
+  adaptive_as [label = 'Adaptive genetic attribute spaces']
+
+  node[shape = ellipse, style = filled, color = darkslategray, fillcolor = white]
+  calc_geo_as [label = 'Processing\n(Figure S60)'] 
+  calc_env_as [label = 'Processing\n(Figure S61)']
+  calc_genetic_as [label = 'Processing\n(Figure S62)']
+
+  edge [color = grey]
+  meirmans_data -> {sample_aflp_data grid_data sample_location_data}
+  bioclimatic_data -> calc_env_as
+  sample_location_data -> {calc_geo_as calc_env_as calc_genetic_as}
+  grid_data -> {calc_geo_as calc_env_as calc_genetic_as}
+  sample_aflp_data -> calc_genetic_as
+  calc_genetic_as -> {neutral_as, adaptive_as}
+  calc_geo_as -> geo_as
+  calc_env_as -> env_as
+  calc_env_as -> calc_env_as [label = ' each species\nseparately']
+  calc_geo_as -> calc_geo_as [label = ' each species\nseparately']
+  calc_genetic_as -> calc_genetic_as [label = ' each species\nseparately']
+
+}
+"))
+```
+
+```{r, echo=FALSE, results='hide', message=FALSE, fig.cap='Geographic surrogate processing workflow. Conventions are detailed in Figure S59 caption.'}
+render_gvis(grViz("
+digraph geo_as {
+  graph [layout = dot]
+
+  node [shape = rectangle, style = filled, color = darkslategray, fillcolor = white]
+
+  # input
+  sample_location_data [label = 'Sample location data']
+  grid_data [label = 'Grid cell data']
+  species_name [label = 'Species name']
+  {rank = same; sample_location_data; grid_data; species_name}
+
+  # intermediate
+  spp_samples [label = 'Locations of samples for species']
+  spp_grids [label = 'Grid cells occupied by species']
+  spp_proj_data [label = 'Projected grid cells']
+  spp_centriod_data [label = 'Centroids of grid cells']
+
+  # output
+  geographic_space [label = 'Geographic attribute space']
+
+  edge [color = grey]
+
+  species_name -> spp_samples 
+  sample_location_data -> spp_samples [label = ' subset data']
+  spp_samples -> spp_grids [label = ' overlay data']
+  grid_data -> spp_grids 
+  spp_grids -> spp_proj_data [label = ' reproject data to\nequi-distant CRS']
+  spp_proj_data -> spp_centriod_data [label = ' compute centroids']
+  spp_centriod_data -> geographic_space [label =' extract coordinates and use them\nto represent demand points and\nplanning unit coordinates']
+
+}
+"))
+```
+
+```{r, fig.height=6, echo=FALSE, results='hide', message=FALSE, fig.cap='Environmental surrogate processing workflow. Conventions are detailed in Figure S59 caption.'}
+render_gvis(grViz("
+digraph env_as {
+  graph [layout = dot]
+
+  node[shape = rectangle, style = filled, color = darkslategray, fillcolor = white]
+
+  # input
+  sample_location_data [label = 'Sample location data']
+  grid_data [label = 'Grid cell data']
+  species_name [label = 'Species name']
+  bioclimatic_data [label = <Bioclimatic data<BR></BR>(Hijmans <I> et al. </I> 2005)>]
+  {rank = same; sample_location_data; grid_data; species_name; bioclimatic_data}
+
+  # intermediate
+  spp_samples [label = 'Locations of samples for species']
+  spp_grids [label = 'Grid cells occupied by species']
+  spp_proj_data [label = 'Projected grid cells']
+
+  bioclimatic_proj_data [label = 'Projected bioclimatic data']
+  clipped_bioclimatic_data [label = 'Bioclimatic data clipped to species distribution']
+  pca_scores_data [label = 'Main axes of bioclimatic variation\nacross species distribution']
+  grid_pca_data [label = 'Environmental conditions inside each grid cell']
+
+  # output
+  environmental_space [label = 'Environmental attribute space']
+
+  # analyses
+  node[shape = ellipse, style = filled, color = darkslategray, fillcolor = white]  
+  pca [label = 'Principal components analysis']
+
+  edge [color = grey]
+  species_name -> spp_samples 
+  sample_location_data -> spp_samples [label = ' subset data']
+  grid_data -> spp_grids
+  bioclimatic_data -> bioclimatic_proj_data [label = ' reproject data to\nequal-area CRS']
+  spp_samples -> spp_grids [label = ' overlay data']
+  spp_grids -> spp_proj_data [label = ' reproject data to\nequal-area CRS']
+  bioclimatic_proj_data -> clipped_bioclimatic_data [label = ' clip data']
+  spp_proj_data -> clipped_bioclimatic_data
+  clipped_bioclimatic_data -> pca
+  pca -> pca_scores_data [label = ' extract scores for\nfirst two components']
+  spp_proj_data -> grid_pca_data
+  pca_scores_data -> grid_pca_data [label = ' calculate average scores\nfor each grid cell']
+  grid_pca_data -> environmental_space [label =' extract scores and use them\nto represent demand points and\nplanning unit coordinates']
+
+}
+"))
+```
+
+```{r, fig.height=10, echo=FALSE, results='hide', message=FALSE, fig.cap='Genetic data processing workflow. Conventions are detailed in Figure S59 caption.'}
+render_gvis(grViz("
+digraph genetic_as {
+  graph [layout = dot]
+
+  node[shape = rectangle, style = filled, color = darkslategray, fillcolor = white]
+
+  # input
+  sample_location_data [label = 'Sample location data']
+  sample_aflp_data [label = 'Sample AFLP data']
+  grid_data [label = 'Grid cell data']
+  species_name [label = 'Species name']
+  {rank = same; sample_location_data; grid_data; species_name; sample_aflp_data}
+
+  # intermediate
+  spp_aflp_samples [label = 'Species AFLP data']
+  spp_location_samples [label = 'Species location data']
+  spp_grids [label = 'Grid cells occupied by species']
+
+  population_data [label = 'Samples grouped into populations']
+  subset_data [label = 'Samples with reliable population designations']
+  rare_alleles [label = 'AFLP data for\nrare alleles']
+  remainder_alleles [label = 'AFLP data for\n standard alleles']
+  bayescan_adaptive_alleles [label = ' Outlier loci']
+  pcadapt_adaptive_alleles [label = ' Outlier loci']
+  bayescan_neutral_alleles [label = ' Standard loci']
+  pcadapt_neutral_alleles [label = ' Standard loci']
+  adaptive_alleles [label = ' AFLP data\nfor adaptive loci']
+  neutral_alleles [label = ' AFLP data\nfor neutral loci']
+  adaptive_ordinations [label = ' Ordinations showing main\ngradients of variation\nin adaptive loci']
+  neutral_ordinations [label = '  Ordinations showing main\ngradients of variation\nin neutral loci']
+  grid_adaptive_ordinations [label = ' Typical adaptive genetic characteristics of individuals in grid cells']
+  grid_neutral_ordinations [label = ' Typical neutral genetic characteristics of individuals in grid cells']
+
+  # output
+  adaptive_genetic_space [label = 'Adaptive genetic space']
+  neutral_genetic_space  [label = 'Neutral genetic space']
+
+  # analyses
+  node[shape = ellipse, style = filled, color = darkslategray, fillcolor = white]
+  bayescan_analysis [label = 'BayeScan analysis']
+  pcadapt_analysis [label = 'PCadapt analysis']
+  structure_analysis [label = 'Structure analysis']
+  clumpp_analysis [label = 'ClumPP analysis']
+  neutral_nmds_analysis [label = 'NMDS analyses']
+  adaptive_nmds_analysis [label = 'NMDS analyses']
+  combine_results [label = 'Combine results from all analyses.\nLoci classified as adaptive in both analyses are treated as adaptive,\notherwise treated as neutral']
+
+  edge [color = grey]
+  sample_location_data -> spp_location_samples [label = ' subset data']
+  species_name -> spp_location_samples 
+  sample_aflp_data -> spp_aflp_samples [label = ' subset data']
+  species_name -> spp_aflp_samples
+  spp_location_samples -> spp_grids
+  grid_data -> spp_grids [label = ' overlay data']
+  spp_aflp_samples -> structure_analysis
+  structure_analysis -> clumpp_analysis [label = ' post-process\nresults']
+  clumpp_analysis -> population_data
+  population_data -> subset_data [label = ' omit samples within\nuncertain memberships']
+  subset_data -> rare_alleles [label = ' extract rare alleles where\nglobal frequency of minor\nalleles < 10 %']
+  subset_data -> remainder_alleles
+  remainder_alleles -> bayescan_analysis [label = 'use population\ngroupings']
+  remainder_alleles -> pcadapt_analysis [label = 'ignore population\ngroupings']
+  bayescan_analysis -> bayescan_adaptive_alleles [label = ' extract']
+  pcadapt_analysis -> pcadapt_adaptive_alleles  [label = ' extract']
+  bayescan_analysis -> bayescan_neutral_alleles [label = ' extract']
+  pcadapt_analysis -> pcadapt_neutral_alleles [label = ' extract']
+  bayescan_adaptive_alleles -> combine_results [label = ' treat loci\nas adaptive']
+  pcadapt_adaptive_alleles -> combine_results [label = ' treat loci\nas adaptive']
+  bayescan_neutral_alleles -> combine_results [label = ' treat loci\nas neutral']
+  pcadapt_neutral_alleles -> combine_results [label = ' treat loci\nas neutral']
+  rare_alleles -> combine_results [label = ' treat loci\nas neutral']
+  combine_results -> adaptive_alleles
+  combine_results -> neutral_alleles 
+  neutral_alleles  -> neutral_nmds_analysis
+  adaptive_alleles -> adaptive_nmds_analysis
+  adaptive_nmds_analysis -> adaptive_ordinations [label = ' extract']
+  neutral_nmds_analysis  -> neutral_ordinations [label = ' extract']
+  spp_grids  -> grid_adaptive_ordinations
+  adaptive_ordinations -> grid_adaptive_ordinations [label = ' calculate average values\nin each grid cell']
+  spp_grids  -> grid_neutral_ordinations
+  neutral_ordinations -> grid_neutral_ordinations [label = ' calculate average values\nin each grid cell']
+  grid_adaptive_ordinations -> adaptive_genetic_space [label =' extract values and use them\nto represent demand points and\nplanning unit coordinates']
+  grid_neutral_ordinations -> neutral_genetic_space [label =' extract values and use them\nto represent demand points and\nplanning unit coordinates']
+
+}
+"))
 ```
@@
 
-```{r, echo=FALSE, include=FALSE}
-plot.spp.nmds <- function(i) {
-	## prepare data
-	# get vector of adaptive and neutral column names
-	adapt.col <- grep(
-		paste0(unique(spp.samples.DF$species)[i], '_adaptive_d'),
-		names(grid.FPLY), value=TRUE, fixed=TRUE
-	)
-	neutral.col <- grep(
-		paste0(unique(spp.samples.DF$species)[i], '_neutral_d'),
-		names(grid.FPLY), value=TRUE, fixed=TRUE
-	)
-	# skip species if not used in analysis
-	curr.FPLY <- grid.FPLY[,
-		c(adapt.col, neutral.col,
-		'long','lat','group'),drop=FALSE]
-	curr.FPLY <- curr.FPLY[which(!is.na(curr.FPLY[[1]])),]
-	curr.FPLY <- curr.FPLY %>% gather(d, value, contains('adaptive'), contains('neutral')) %>% mutate(
-		d=gsub(paste0(unique(spp.samples.DF$species)[i], '_'), '', d, fixed=TRUE),
-		d=gsub('_', ' ', d, fixed=TRUE)
-	)
-	for (j in seq_len(length(adapt.col)+length(neutral.col)))
-		curr.FPLY[['d']] <- gsub(as.character(j), paste0('(', j, ')'), curr.FPLY[['d']], fixed=TRUE)
-	curr.FPLY <- curr.FPLY %>% mutate(d=factor(gsub(' d(', ' (', d, fixed=TRUE)))
-	# make plot
-	curr.GROBS <- llply(unique(curr.FPLY$d), function(x) {
-		ggplot() +
-			geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
-				fill='grey85', color='grey70') +
-			geom_polygon(data=filter(curr.FPLY, d==x), aes_string(x='long', y='lat', 
-				group='group', fill='value'), color='grey10') +
-			theme_classic() +
-			theme(
-				axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"), axis.title=element_blank(),
-				panel.border=element_rect(color='black', fill=NA, size=1),
-				strip.background = element_rect(fill='grey20'), strip.text = element_text(color='white', size=20)
-			) +
-			coord_cartesian(
-				xlim=buffered.range(grid.FPLY$long, 0),
-				ylim=buffered.range(grid.FPLY$lat, 0)
-			) +
-			scale_fill_gradientn(name='',
-				colors=brewer.pal(9, 'RdYlBu'),
-				guide=guide_colorbar(ticks=element_line(color='black'),
-					border=element_line(color='black'), barheight=unit(3.0, 'cm'))
-			) + facet_wrap(~d)
-	})
-	do.call(grid.arrange, append(curr.GROBS, list(ncol=3)))
-	invisible()
-}
-
-calc.fig.height <- function(i) {
-	# init
-	adapt.col <- paste0(unique(spp.samples.DF$species)[i], '_adaptive_d1')
-	neutral.col <- paste0(unique(spp.samples.DF$species)[i], '_neutral_d1')
-	n.plots <- 0 
-	# count adaptive plot
-	if (adapt.col %in% names(grid.FPLY))
-		n.plots <- n.plots + spp.nmds.LST[[i]][['adaptive']]$ndim
-	# count neutral plot
-	if (neutral.col %in% names(grid.FPLY))
-		n.plots <- n.plots + spp.nmds.LST[[i]][['neutral']]$ndim
-	# exports
-	if (n.plots >= 3) return((((n.plots-1) %/% 3)+1) * 2)
-	return(1.85)
-}
-
-calc.fig.width <- function(i) {
-	# init
-	adapt.col <- paste0(unique(spp.samples.DF$species)[i], '_adaptive_d1')
-	neutral.col <- paste0(unique(spp.samples.DF$species)[i], '_neutral_d1')
-	n.plots <- 0 
-	# count adaptive plot
-	if (adapt.col %in% names(grid.FPLY))
-		n.plots <- n.plots + spp.nmds.LST[[i]][['adaptive']]$ndim
-	# count neutral plot
-	if (neutral.col %in% names(grid.FPLY))
-		n.plots <- n.plots + spp.nmds.LST[[i]][['neutral']]$ndim
-	# exports
-	if (n.plots >= 3) return(10)
-	return(4 * n.plots)
-}
-
-is.speciesPlottable <- function(i) {
-	if (i > general.params.LST[[MODE]]$n.spp)
-		return(FALSE)
-	return(!is.na(unique(spp.samples.DF$species)[i]))
-}
-```
-
-```{r, echo=FALSE, fig.width=calc.fig.width(1), fig.height=calc.fig.height(1), fig.width=calc.fig.width(1), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[1]),'}.
-Each panel shows a different dimension of the variation as indicated by its title.
-Squares represent planning units.
-The color of each planning unit panel corresponds to the average ordinated values associated with individuals sampled inside it.
-No indiviudals of this species were found inside the gray planning units.')}
-if (is.speciesPlottable(1)) plot.spp.nmds(1)
- ```
-
-```{r, echo=FALSE, fig.width=calc.fig.width(2), fig.height=calc.fig.height(2), fig.width=calc.fig.width(2), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[2]),'}.
-See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(2)) plot.spp.nmds(2)
-```
-
-```{r, echo=FALSE, fig.width=calc.fig.width(3), fig.height=calc.fig.height(3), fig.width=calc.fig.width(3), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[3]),'}.
-See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(3)) plot.spp.nmds(3)
-```
-
-```{r, echo=FALSE, fig.width=calc.fig.width(4), fig.height=calc.fig.height(4), fig.width=calc.fig.width(4), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[4]),'}.
-See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(4)) plot.spp.nmds(4)
-```
-
-```{r, echo=FALSE, fig.width=calc.fig.width(5), fig.height=calc.fig.height(5), fig.width=calc.fig.width(5), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[5]),'}.
-See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(5)) plot.spp.nmds(5)
-```
-
-```{r, echo=FALSE, fig.width=calc.fig.width(6), fig.height=calc.fig.height(6), fig.width=calc.fig.width(6), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[6]),'}.
-See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(6)) plot.spp.nmds(6)
-```
-
-```{r, echo=FALSE, fig.width=calc.fig.width(7), fig.height=calc.fig.height(7), fig.width=calc.fig.width(7), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[7]),'}.
-See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(7)) plot.spp.nmds(7)
-```
-
-```{r, echo=FALSE, fig.width=calc.fig.width(8), fig.height=calc.fig.height(8), fig.width=calc.fig.width(8), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[8]),'}.
-See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(8)) plot.spp.nmds(8)
-```
-
-```{r, echo=FALSE, fig.width=calc.fig.width(9), fig.height=calc.fig.height(9), fig.width=calc.fig.width(9), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[9]),'}.
-See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(9)) plot.spp.nmds(9)
-```
-
-```{r, echo=FALSE, fig.width=calc.fig.width(10), fig.height=calc.fig.height(10), fig.width=calc.fig.width(10), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[10]),'}.
-See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(10)) plot.spp.nmds(10)
-```
-
-```{r, echo=FALSE, fig.width=calc.fig.width(11), fig.height=calc.fig.height(11), fig.width=calc.fig.width(11), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[11]),'}.
-See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(11)) plot.spp.nmds(11)
-```
-
-```{r, echo=FALSE, fig.width=calc.fig.width(12), fig.height=calc.fig.height(12), fig.width=calc.fig.width(12), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[12]),'}.
-See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(12)) plot.spp.nmds(12)
-```
-
-```{r, echo=FALSE, fig.width=calc.fig.width(13), fig.height=calc.fig.height(13), fig.width=calc.fig.width(13), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[13]),'}.
-See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(13)) plot.spp.nmds(13)
-```
-
-```{r, echo=FALSE, fig.width=calc.fig.width(14), fig.height=calc.fig.height(14), fig.width=calc.fig.width(14), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[14]),'}.
-See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(14)) plot.spp.nmds(14)
-```
-
-```{r, echo=FALSE, fig.width=calc.fig.width(15), fig.height=calc.fig.height(15), fig.width=calc.fig.width(15), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[15]),'}.
-See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(15)) plot.spp.nmds(15)
-```
-
-```{r, echo=FALSE, fig.width=calc.fig.width(16), fig.height=calc.fig.height(16), fig.width=calc.fig.width(16), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[16]),'}.
-See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(16)) plot.spp.nmds(16)
-```
-
-```{r, echo=FALSE, fig.width=calc.fig.width(17), fig.height=calc.fig.height(17), fig.width=calc.fig.width(17), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[17]),'}.
-See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(17)) plot.spp.nmds(17)
-```
-
-```{r, echo=FALSE, fig.width=calc.fig.width(18), fig.height=calc.fig.height(18), fig.width=calc.fig.width(18), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[18]),'}.
-See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(18)) plot.spp.nmds(18)
-```
-
-```{r, echo=FALSE, fig.width=calc.fig.width(19), fig.height=calc.fig.height(19), fig.width=calc.fig.width(19), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[19]),'}.
-See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(19)) plot.spp.nmds(19)
-```
-
-```{r, echo=FALSE, fig.width=calc.fig.width(20), fig.height=calc.fig.height(20), fig.width=calc.fig.width(20), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[20]),'}.
-See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(20)) plot.spp.nmds(20)
-```
-
-```{r, echo=FALSE, fig.width=calc.fig.width(21), fig.height=calc.fig.height(21), fig.width=calc.fig.width(21), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[21]),'}.
-See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(21)) plot.spp.nmds(21)
-```
-
-```{r, echo=FALSE, fig.width=calc.fig.width(22), fig.height=calc.fig.height(22), fig.width=calc.fig.width(22), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[22]),'}.
-See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(22)) plot.spp.nmds(22)
-```
-
-```{r, echo=FALSE, fig.width=calc.fig.width(23), fig.height=calc.fig.height(23), fig.width=calc.fig.width(23), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[23]),'}.
-See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(23)) plot.spp.nmds(23)
-```
-
-```{r, echo=FALSE, fig.width=calc.fig.width(24), fig.height=calc.fig.height(24), fig.width=calc.fig.width(24), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[24]),'}.
-See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(24)) plot.spp.nmds(24)
-```
-
-```{r, echo=FALSE, fig.width=calc.fig.width(25), fig.height=calc.fig.height(25), fig.width=calc.fig.width(25), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[25]),'}.
-See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(25)) plot.spp.nmds(25)
-```
-
-```{r, echo=FALSE, fig.width=calc.fig.width(26), fig.height=calc.fig.height(26), fig.width=calc.fig.width(26), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[26]),'}.
-See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(26)) plot.spp.nmds(26)
-```
-
-```{r, echo=FALSE, fig.width=calc.fig.width(27), fig.height=calc.fig.height(27), fig.width=calc.fig.width(27), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[27]),'}.
-See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(27)) plot.spp.nmds(27)
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=(2.5*4), fig.cap='The relationship between the proportion of adaptive genetic variation secured in a prioritization and the proportion of environmental variation it also secures.
-Points represent values associated with randomly generated prioritizations.
-Blue lines indicate average trends computed using a generalised linear model with a logit link.
-Each panel shows data for a different species.'}
-# format data
-curr.DF <- env.surrogacy.DF %>% filter(!is.na(genetic.held)) %>% mutate(Species=gsub('_', ' ', Species))
-curr.DF2 <- env.surrogate.spp.DF %>% filter(!is.na(r2CU)) %>% mutate(Species=gsub('_', ' ', Species))
-curr.DF2$R2_repr <- round(curr.DF2$r2CU,2)
-curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
-curr.DF2$surrogate.held <- 0
-curr.DF2$genetic.held <- 0.9
-
-# make plot
-suppressWarnings(
-	ggplot(aes(x=surrogate.held, y=genetic.held), data=curr.DF) +
-		theme_classic() +
-		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=12),
-			panel.margin.x=unit(1, 'lines')) +
-		xlab('Environmental variation secured (%)') +
-		ylab('Adaptive genetic variation secured (%)') +
-		geom_hex() +
-		suppressWarnings(stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
-			se=FALSE, fill='blue', alpha=0.8)) +
-		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=5, color='red') +
-		facet_wrap(~ Species, ncol=4)
-)
-```
-
-```{r, include=FALSE}
-calc.fig.height3 <- function(x) {
-	return((((x-1) %/% 4)+1)*1.5)
-}
-
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=calc.fig.height3(n_distinct(geo.surrogacy.DF$Species)), fig.cap='The relationship between the proportion of neutral genetic variation secured in a prioritizaiton and the proportion of geographic variation it also secures.
-Conventions are detailed in Figure S56.'}
-# format data
-curr.DF <- geo.surrogacy.DF %>% filter(!is.na(genetic.held)) %>% mutate(Species=gsub('_', ' ', Species))
-curr.DF2 <- geo.surrogate.spp.DF %>% filter(!is.na(r2CU)) %>% mutate(Species=gsub('_', ' ', Species))
-curr.DF2$R2_repr <- round(curr.DF2$r2CU,2)
-curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
-curr.DF2$surrogate.held <- 0
-curr.DF2$genetic.held <- 0.9
-
-# make plot
-suppressWarnings(
-	ggplot(aes(x=surrogate.held, y=genetic.held), data=curr.DF) +
-		theme_classic() +
-		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=12),
-			panel.margin.x=unit(1, 'lines')) +
-		xlab('Geographic variation secured (%)') +
-		ylab('Neutral genetic variation secured (%)') +
-		geom_hex() +
-		suppressWarnings(stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
-			se=FALSE, fill='blue', alpha=0.8)) +
-		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=5, color='red') +
-		facet_wrap(~ Species, ncol=4)
-)
-```
-
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=(5*0.9), fig.height=(6.5*0.8), out.width='3.42in', height='3.9in', fig.cap='Planning unit selection frequencies.
-Squares represent planning units, and their colors indicate their selection frequency.
-Each column of panels shows solutions generated using different targets.
-Each row of panels shows solutions generated under different scenarios.
-Note that the selection frequencies for the single-species prioritizations are based on their frequency among prioritizations generated for each species.
-The selection frequencies for the multi-species prioritizations are each based on a single optimal pioritization for all species and so are binary.'}
-# prepare data for plotting
-grid.sub.FPLY <- grid.sub.PLY
-for (i in seq_along(single.spp.prioritisations[[1]]))
-	grid.sub.FPLY@data[[paste0('Single-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations, function(x){colMeans(x[[i]]@results@selections)}))
-for (i in seq_along(single.spp.prioritisations.with.cost[[1]]))
-	grid.sub.FPLY@data[[paste0('Single-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations.with.cost, function(x){colMeans(x[[i]]@results@selections)}))
-for (i in seq_along(multi.spp.prioritisations))
-	grid.sub.FPLY@data[[paste0('Multi-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations[[i]]@results@selections)
-for (i in seq_along(multi.spp.prioritisations.with.cost))
-	grid.sub.FPLY@data[[paste0('Multi-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations.with.cost[[i]]@results@selections)
-grid.sub.FPLY <- spFortify(grid.sub.FPLY)
-grid.sub.FPLY <- grid.sub.FPLY %>% gather(
-		Combination,
-		`Selection.Frequency`,
-		`Single-species (equal costs).Amount`:`Multi-species (opportunity costs).Genetic`
-	) %>% mutate(
-		Context=sapply(strsplit(as.character(Combination), '.', fixed=TRUE), `[[`, 1),
-		Prioritization=sapply(strsplit(as.character(Combination), '.', fixed=TRUE), `[[`, 2)
-	) %>% mutate(
-		Prioritization=revalue(Prioritization, c('Amount'='Amount targets', 'Surrogate'='Amount & surrogate\ntargets', 'Genetic'='Amount & genetic\ntargets')),
-		Prioritization=factor(as.character(Prioritization), levels=c('Amount targets', 'Amount & surrogate\ntargets', 'Amount & genetic\ntargets')),
-		Context=factor(
-			gsub(' (', '\n(', Context, fixed=TRUE),
-			levels=c("Single-species\n(equal costs)", "Single-species\n(opportunity costs)",
-				"Multi-species\n(equal costs)", "Multi-species\n(opportunity costs)"))
-	)
-
-# make maps
-ggplot() +
-	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
-		fill='grey85', color='grey70') +
-	geom_polygon(data=grid.sub.FPLY, aes_string(x='long', y='lat', 
-		group='group', fill='Selection.Frequency'),
-		alpha=0.8, color=NA) +
-	theme_classic() +
-	theme(axis.ticks=element_blank(), axis.text=element_blank(),
-		plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(),
-		strip.background = element_rect(fill='grey20'),
-		strip.text = element_text(color='white', size=8),
-		legend.position='bottom', legend.key.width=unit(1.8, 'cm'),
-		legend.title=element_text(vjust=-5),
-		legend.margin=unit(c(0.1), 'cm'),
-		axis.title=element_blank()) +
-	coord_cartesian(
-		xlim=buffered.range(grid.sub.FPLY$long, 0),
-		ylim=buffered.range(grid.sub.FPLY$lat, 0)
-	) +
-	facet_grid(Context ~ Prioritization) +
-	scale_fill_gradient(name='Selection Frequency (%)',
-		guide=guide_colorbar(title.position='bottom', title.hjust=0.5,
-			ticks=element_line(color='black'), border=element_line(color='black'))
-	)
-```
-
-```{r, include=FALSE}
-render_gvis <- function(x) {
-  file = tempfile(fileext='.png')
-  rsvg::rsvg_png(charToRaw(DiagrammeRsvg::export_svg(x)), file)
-  grid::grid.raster(png::readPNG(file))
-}
-```
-
-```{r, echo=FALSE, results='hide', message=FALSE, fig.cap='Overall data processing workflow.
-The flowcharts show the methods used to generate the environmental and geographic surrogate spaces, and the adaptive and neutral genetic spaces.
-Squares indicate data sets.
-Ellipses indicate analyses.
-Lines show relationships between data and analysis, and their captions describe how data was processed.'}
-render_gvis(grViz("
-digraph main {
-  graph [layout = dot]
-
-  node [shape = rectangle, style = filled, color = darkslategray, fillcolor = white]
-
-  meirmans_data [label = <Plant data<BR></BR>(Meirmans <I>et al. </I> 2011)>]
-  bioclimatic_data [label = <Bioclimatic data<BR></BR>(Hijmans <I> et al. </I> 2005)>]
-  sample_location_data [label = 'Sample location data']
-  grid_data [label = 'Grid cell data']
-  sample_aflp_data [label = 'Sample AFLP data']
-
-  geo_as [label = 'Geographic attribute spaces'] 
-  env_as [label = 'Environmental attribute spaces']
-  neutral_as [label = 'Neutral genetic attribute spaces']
-  adaptive_as [label = 'Adaptive genetic attribute spaces']
-
-  node[shape = ellipse, style = filled, color = darkslategray, fillcolor = white]
-  calc_geo_as [label = 'Processing\n(Figure S60)'] 
-  calc_env_as [label = 'Processing\n(Figure S61)']
-  calc_genetic_as [label = 'Processing\n(Figure S62)']
-
-  edge [color = grey]
-  meirmans_data -> {sample_aflp_data grid_data sample_location_data}
-  bioclimatic_data -> calc_env_as
-  sample_location_data -> {calc_geo_as calc_env_as calc_genetic_as}
-  grid_data -> {calc_geo_as calc_env_as calc_genetic_as}
-  sample_aflp_data -> calc_genetic_as
-  calc_genetic_as -> {neutral_as, adaptive_as}
-  calc_geo_as -> geo_as
-  calc_env_as -> env_as
-  calc_env_as -> calc_env_as [label = ' each species\nseparately']
-  calc_geo_as -> calc_geo_as [label = ' each species\nseparately']
-  calc_genetic_as -> calc_genetic_as [label = ' each species\nseparately']
-
-}
-"))
-```
-
-```{r, echo=FALSE, results='hide', message=FALSE, fig.cap='Geographic surrogate processing workflow. Conventions are detailed in Figure S59 caption.'}
-render_gvis(grViz("
-digraph geo_as {
-  graph [layout = dot]
-
-  node [shape = rectangle, style = filled, color = darkslategray, fillcolor = white]
-
-  # input
-  sample_location_data [label = 'Sample location data']
-  grid_data [label = 'Grid cell data']
-  species_name [label = 'Species name']
-  {rank = same; sample_location_data; grid_data; species_name}
-
-  # intermediate
-  spp_samples [label = 'Locations of samples for species']
-  spp_grids [label = 'Grid cells occupied by species']
-  spp_proj_data [label = 'Projected grid cells']
-  spp_centriod_data [label = 'Centroids of grid cells']
-
-  # output
-  geographic_space [label = 'Geographic attribute space']
-
-  edge [color = grey]
-
-  species_name -> spp_samples 
-  sample_location_data -> spp_samples [label = ' subset data']
-  spp_samples -> spp_grids [label = ' overlay data']
-  grid_data -> spp_grids 
-  spp_grids -> spp_proj_data [label = ' reproject data to\nequi-distant CRS']
-  spp_proj_data -> spp_centriod_data [label = ' compute centroids']
-  spp_centriod_data -> geographic_space [label =' extract coordinates and use them\nto represent demand points and\nplanning unit coordinates']
-
-}
-"))
-```
-
-```{r, fig.height=6, echo=FALSE, results='hide', message=FALSE, fig.cap='Environmental surrogate processing workflow. Conventions are detailed in Figure S59 caption.'}
-render_gvis(grViz("
-digraph env_as {
-  graph [layout = dot]
-
-  node[shape = rectangle, style = filled, color = darkslategray, fillcolor = white]
-
-  # input
-  sample_location_data [label = 'Sample location data']
-  grid_data [label = 'Grid cell data']
-  species_name [label = 'Species name']
-  bioclimatic_data [label = <Bioclimatic data<BR></BR>(Hijmans <I> et al. </I> 2005)>]
-  {rank = same; sample_location_data; grid_data; species_name; bioclimatic_data}
-
-  # intermediate
-  spp_samples [label = 'Locations of samples for species']
-  spp_grids [label = 'Grid cells occupied by species']
-  spp_proj_data [label = 'Projected grid cells']
-
-  bioclimatic_proj_data [label = 'Projected bioclimatic data']
-  clipped_bioclimatic_data [label = 'Bioclimatic data clipped to species distribution']
-  pca_scores_data [label = 'Main axes of bioclimatic variation\nacross species distribution']
-  grid_pca_data [label = 'Environmental conditions inside each grid cell']
-
-  # output
-  environmental_space [label = 'Environmental attribute space']
-
-  # analyses
-  node[shape = ellipse, style = filled, color = darkslategray, fillcolor = white]  
-  pca [label = 'Principal components analysis']
-
-  edge [color = grey]
-  species_name -> spp_samples 
-  sample_location_data -> spp_samples [label = ' subset data']
-  grid_data -> spp_grids
-  bioclimatic_data -> bioclimatic_proj_data [label = ' reproject data to\nequal-area CRS']
-  spp_samples -> spp_grids [label = ' overlay data']
-  spp_grids -> spp_proj_data [label = ' reproject data to\nequal-area CRS']
-  bioclimatic_proj_data -> clipped_bioclimatic_data [label = ' clip data']
-  spp_proj_data -> clipped_bioclimatic_data
-  clipped_bioclimatic_data -> pca
-  pca -> pca_scores_data [label = ' extract scores for\nfirst two components']
-  spp_proj_data -> grid_pca_data
-  pca_scores_data -> grid_pca_data [label = ' calculate average scores\nfor each grid cell']
-  grid_pca_data -> environmental_space [label =' extract scores and use them\nto represent demand points and\nplanning unit coordinates']
-
-}
-"))
-```
-
-```{r, fig.height=10, echo=FALSE, results='hide', message=FALSE, fig.cap='Genetic data processing workflow. Conventions are detailed in Figure S59 caption.'}
-render_gvis(grViz("
-digraph genetic_as {
-  graph [layout = dot]
-
-  node[shape = rectangle, style = filled, color = darkslategray, fillcolor = white]
-
-  # input
-  sample_location_data [label = 'Sample location data']
-  sample_aflp_data [label = 'Sample AFLP data']
-  grid_data [label = 'Grid cell data']
-  species_name [label = 'Species name']
-  {rank = same; sample_location_data; grid_data; species_name; sample_aflp_data}
-
-  # intermediate
-  spp_aflp_samples [label = 'Species AFLP data']
-  spp_location_samples [label = 'Species location data']
-  spp_grids [label = 'Grid cells occupied by species']
-
-  population_data [label = 'Samples grouped into populations']
-  subset_data [label = 'Samples with reliable population designations']
-  rare_alleles [label = 'AFLP data for\nrare alleles']
-  remainder_alleles [label = 'AFLP data for\n standard alleles']
-  bayescan_adaptive_alleles [label = ' Outlier loci']
-  pcadapt_adaptive_alleles [label = ' Outlier loci']
-  bayescan_neutral_alleles [label = ' Standard loci']
-  pcadapt_neutral_alleles [label = ' Standard loci']
-  adaptive_alleles [label = ' AFLP data\nfor adaptive loci']
-  neutral_alleles [label = ' AFLP data\nfor neutral loci']
-  adaptive_ordinations [label = ' Ordinations showing main\ngradients of variation\nin adaptive loci']
-  neutral_ordinations [label = '  Ordinations showing main\ngradients of variation\nin neutral loci']
-  grid_adaptive_ordinations [label = ' Typical adaptive genetic characteristics of individuals in grid cells']
-  grid_neutral_ordinations [label = ' Typical neutral genetic characteristics of individuals in grid cells']
-
-  # output
-  adaptive_genetic_space [label = 'Adaptive genetic space']
-  neutral_genetic_space  [label = 'Neutral genetic space']
-
-  # analyses
-  node[shape = ellipse, style = filled, color = darkslategray, fillcolor = white]
-  bayescan_analysis [label = 'BayeScan analysis']
-  pcadapt_analysis [label = 'PCadapt analysis']
-  structure_analysis [label = 'Structure analysis']
-  clumpp_analysis [label = 'ClumPP analysis']
-  neutral_nmds_analysis [label = 'NMDS analyses']
-  adaptive_nmds_analysis [label = 'NMDS analyses']
-  combine_results [label = 'Combine results from all analyses.\nLoci classified as adaptive in both analyses are treated as adaptive,\notherwise treated as neutral']
-
-  edge [color = grey]
-  sample_location_data -> spp_location_samples [label = ' subset data']
-  species_name -> spp_location_samples 
-  sample_aflp_data -> spp_aflp_samples [label = ' subset data']
-  species_name -> spp_aflp_samples
-  spp_location_samples -> spp_grids
-  grid_data -> spp_grids [label = ' overlay data']
-  spp_aflp_samples -> structure_analysis
-  structure_analysis -> clumpp_analysis [label = ' post-process\nresults']
-  clumpp_analysis -> population_data
-  population_data -> subset_data [label = ' omit samples within\nuncertain memberships']
-  subset_data -> rare_alleles [label = ' extract rare alleles where\nglobal frequency of minor\nalleles < 10 %']
-  subset_data -> remainder_alleles
-  remainder_alleles -> bayescan_analysis [label = 'use population\ngroupings']
-  remainder_alleles -> pcadapt_analysis [label = 'ignore population\ngroupings']
-  bayescan_analysis -> bayescan_adaptive_alleles [label = ' extract']
-  pcadapt_analysis -> pcadapt_adaptive_alleles  [label = ' extract']
-  bayescan_analysis -> bayescan_neutral_alleles [label = ' extract']
-  pcadapt_analysis -> pcadapt_neutral_alleles [label = ' extract']
-  bayescan_adaptive_alleles -> combine_results [label = ' treat loci\nas adaptive']
-  pcadapt_adaptive_alleles -> combine_results [label = ' treat loci\nas adaptive']
-  bayescan_neutral_alleles -> combine_results [label = ' treat loci\nas neutral']
-  pcadapt_neutral_alleles -> combine_results [label = ' treat loci\nas neutral']
-  rare_alleles -> combine_results [label = ' treat loci\nas neutral']
-  combine_results -> adaptive_alleles
-  combine_results -> neutral_alleles 
-  neutral_alleles  -> neutral_nmds_analysis
-  adaptive_alleles -> adaptive_nmds_analysis
-  adaptive_nmds_analysis -> adaptive_ordinations [label = ' extract']
-  neutral_nmds_analysis  -> neutral_ordinations [label = ' extract']
-  spp_grids  -> grid_adaptive_ordinations
-  adaptive_ordinations -> grid_adaptive_ordinations [label = ' calculate average values\nin each grid cell']
-  spp_grids  -> grid_neutral_ordinations
-  neutral_ordinations -> grid_neutral_ordinations [label = ' calculate average values\nin each grid cell']
-  grid_adaptive_ordinations -> adaptive_genetic_space [label =' extract values and use them\nto represent demand points and\nplanning unit coordinates']
-  grid_neutral_ordinations -> neutral_genetic_space [label =' extract values and use them\nto represent demand points and\nplanning unit coordinates']
-
-}
-"))
-```
-
 # Tables

### 2016-11-10 code/rmarkdown/supporting_information.Rmd
+ bug fix

@@
 No indiviudals of this species were found inside the gray planning units.')}
-if (is.speciesPlottable(1)) plot.spp.nmds(1)
+if (is.speciesPlottable(1)) plot.spp.spaces(1)
  ```
@@
 See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(2)) plot.spp.nmds(2)
+if (is.speciesPlottable(2)) plot.spp.spaces(2)
 ```
@@
 See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(3)) plot.spp.nmds(3)
+if (is.speciesPlottable(3)) plot.spp.spaces(3)
 ```
@@
 See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(4)) plot.spp.nmds(4)
+if (is.speciesPlottable(4)) plot.spp.spaces(4)
 ```
@@
 See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(5)) plot.spp.nmds(5)
+if (is.speciesPlottable(5)) plot.spp.spaces(5)
 ```
@@
 See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(6)) plot.spp.nmds(6)
+if (is.speciesPlottable(6)) plot.spp.spaces(6)
 ```
@@
 See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(7)) plot.spp.nmds(7)
+if (is.speciesPlottable(7)) plot.spp.spaces(7)
 ```
@@
 See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(8)) plot.spp.nmds(8)
+if (is.speciesPlottable(8)) plot.spp.spaces(8)
 ```
@@
 See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(9)) plot.spp.nmds(9)
+if (is.speciesPlottable(9)) plot.spp.spaces(9)
 ```
@@
 See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(10)) plot.spp.nmds(10)
+if (is.speciesPlottable(10)) plot.spp.spaces(10)
 ```
@@
 See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(11)) plot.spp.nmds(11)
+if (is.speciesPlottable(11)) plot.spp.spaces(11)
 ```
@@
 See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(12)) plot.spp.nmds(12)
+if (is.speciesPlottable(12)) plot.spp.spaces(12)
 ```
@@
 See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(13)) plot.spp.nmds(13)
+if (is.speciesPlottable(13)) plot.spp.spaces(13)
 ```
@@
 See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(14)) plot.spp.nmds(14)
+if (is.speciesPlottable(14)) plot.spp.spaces(14)
 ```
@@
 See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(15)) plot.spp.nmds(15)
+if (is.speciesPlottable(15)) plot.spp.spaces(15)
 ```
@@
 See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(16)) plot.spp.nmds(16)
+if (is.speciesPlottable(16)) plot.spp.spaces(16)
 ```
@@
 See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(17)) plot.spp.nmds(17)
+if (is.speciesPlottable(17)) plot.spp.spaces(17)
 ```
@@
 See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(18)) plot.spp.nmds(18)
+if (is.speciesPlottable(18)) plot.spp.spaces(18)
 ```
@@
 See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(19)) plot.spp.nmds(19)
+if (is.speciesPlottable(19)) plot.spp.spaces(19)
 ```
@@
 See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(20)) plot.spp.nmds(20)
+if (is.speciesPlottable(20)) plot.spp.spaces(20)
 ```
@@
 See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(21)) plot.spp.nmds(21)
+if (is.speciesPlottable(21)) plot.spp.spaces(21)
 ```
@@
 See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(22)) plot.spp.nmds(22)
+if (is.speciesPlottable(22)) plot.spp.spaces(22)
 ```
@@
 See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(23)) plot.spp.nmds(23)
+if (is.speciesPlottable(23)) plot.spp.spaces(23)
 ```
@@
 See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(24)) plot.spp.nmds(24)
+if (is.speciesPlottable(24)) plot.spp.spaces(24)
 ```
@@
 See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(25)) plot.spp.nmds(25)
+if (is.speciesPlottable(25)) plot.spp.spaces(25)
 ```
@@
 See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(26)) plot.spp.nmds(26)
+if (is.speciesPlottable(26)) plot.spp.spaces(26)
 ```
@@
 See Figure S29 caption for conventions.')}
-if (is.speciesPlottable(27)) plot.spp.nmds(27)
+if (is.speciesPlottable(27)) plot.spp.spaces(27)
 ```

### 2016-11-10 code/rmarkdown/supporting_information.Rmd
+ assorted changes

@@
   neutral.col <- paste0(unique(spp.samples.DF$species)[i], '_neutral_d1')
-  env.col <- grepl(paste0(unique(spp.samples.DF$species)[i], '_env_d'), names(grid.FPLY), value=TRUE, fixed=TRUE) %>%
+  env.col <- grepl(paste0(unique(spp.samples.DF$species)[i], '_env_d'), names(grid.FPLY), fixed=TRUE) %>%
     sum()
@@
 Table shows the results from one-sided _post-hoc_ analyses that were conducted to determine if the slopes for each species were positive (using the alternative hypothesis that slopes were less than or equal to zero; tests used a single degree of freedom; results reported in the \\textit{Z} and \\textit{P} columns).
-The table also shows Cragg and Uhler\'s pseudo-R$^2$ values that were calculated to show the proportion of variation explained by the slopes for each species ($R^2$ column).'
+The table also shows Cragg and Uhler\'s pseudo-$R^2$ values that were calculated to show the proportion of variation explained by the slopes for each species ($R^2$ column).'
 	)

### 2016-11-13 code/rmarkdown/article.Rmd
+ fix supp info figures + tables + reorder si figures and tables + update si figures and table numbers in main tex

@@
 After identifying the loci showing strong signals of selection, we used non-metric multi-dimensional scaling (NMDS) analyses to identify the main gradients of adaptive (if detected) and neutral variation for each species.
-Generally, only a small number of continuous dimensions was needed to sufficiently describe their patterns of adaptive (all $K = `r unique(na.omit(adaptive.k))`$; Table S2) and neutral genetic variation ($K = `r min(neutral.k)`$--$`r max(neutral.k)`$; Table S1).
+Generally, only a small number of continuous dimensions was needed to sufficiently describe their patterns of adaptive (all $K = `r unique(na.omit(adaptive.k))`$; Table S2) and neutral genetic variation ($K = `r min(neutral.k)`$--$`r max(neutral.k)`$; Table S2).
 The resulting ordinations were used to construct an adaptive (if detected) and neutral genetic space for each species.
-The spatial distribution of these genetic spaces (Figures S29--S54) generally showed substantial spatial auto-correlation (eg.
-_Cerastium uniflorum_, Figure S34; _Dryas octopetala_, Figures S36), suggesting that the environmental and geographic variables may prove to be effective surrogates.
+The spatial distribution of these genetic spaces (Figures S1--S27) generally showed substantial spatial auto-correlation (eg.
+_Cerastium uniflorum_, Figure S6; _Dryas octopetala_, Figures S8), suggesting that the environmental and geographic variables may prove to be effective surrogates.
 
@@
 We then repeated this process, and calculated the proportion of geographic variation and neutral genetic variation they sampled.
-The environmental and geographic variables were moderately effective predictors for the genetic variation represented by the randomly generated prioritisations (Figure 1).
+The environmental and geographic variables were moderately effective predictors for the genetic variation represented by the randomly generated prioritisations (Figure 1; Figures S28--S29).
 The relationship between the proportion of genetic and surrogate variation secured in the prioritisations varied among the species (species $\times$ proportion of environmental variation interaction term: ${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$; species $\times$ proportion of geographic variation interaction term: ${\chi^2}_{`r round(geo.surrogacy.AOV$df[[4]],2)`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`).
@@
 To determine whether using environmental and geographic targets could improve the effectiveness of prioritisations in representing genetic variation, we generated prioritisations using (i) "amount targets" reflecting the traditional approach of representing a certain proportion of each species' geographic distribution, (ii) "amount and surrogate targets" where we also targeted representation of the environmental and geographic surrogates as well as a certain proportion of each species' geographic distribution, and (iii) "amount and genetic targets" where we targeted representation of the directly measured genetic variation as well as a certain proportion of each species' geographic distribution.
-To determine if these patterns were robust, for each of the three combinations of the targets, we generated prioritisations under four scenarios: (i) single-species with equal costs, (ii) single species with opportunity costs, (iii) multi-species with equal costs, and (iv) multi-species with opportunity costs (Figure 3, Figure SXXX).
+To determine if these patterns were robust, for each of the three combinations of the targets, we generated prioritisations under four scenarios: (i) single-species with equal costs, (ii) single species with opportunity costs, (iii) multi-species with equal costs, and (iv) multi-species with opportunity costs (Figure 3, Figure S30).
 
@@
 First, our results suggest that surrogate-based targets should be used in conjunction with amount-based targets.
-For all species, it was possible to generate solutions that secured a high proportion of the surrogate variable (> 90 %) and only a small proportion of adaptive or neutral genetic variation (< 20 %; Figures S56--S57), and so use of amount-based targets may ameliorate this.
+For all species, it was possible to generate solutions that secured a high proportion of the surrogate variable (> 90 %) and only a small proportion of adaptive or neutral genetic variation (< 20 %; Figures S28--S29), and so use of amount-based targets may ameliorate this.
 Second, geographic surrogates may need much higher targets than environmental surrogates to capture evolutionary processes.
-In our analysis, geographic variation accumulated much slower for neutral genetic representation than environmental variation did with adaptive representation (Figures 2a--2b and S56--S57).
+In our analysis, geographic variation accumulated much slower for neutral genetic representation than environmental variation did with adaptive representation (Figures 2a--2b and S28--S29).
 To secure a representative sample of the species' neutral genetic variation ($\geq$ `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %), prioritisations needed to sample `r round(mean(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` SD of the variation in geographic position among the planning units occupied by the species.
@@
 For most species, the spatial patterns of adaptive genetic variation clustered into a few distinct groups (eg.
-_Carex semipervens_ and _Cerastium uniflorum_; Figures S33 and S34 respectively).
+_Carex semipervens_ and _Cerastium uniflorum_; Figures S5--S6).
 As a consequence, a prioritisation would only need to select a planning unit in each cluster to secure a representative sample.
@@
 It was collected using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx.
-22.3 km $\times$ 25 km; Figure S1).
+22.3 km $\times$ 25 km; Figure S31).
 Project members visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals.
@@
 
-We created environmental and geographic surrogate variables for each species (see Figures S59--S561 for workflow).
+We created environmental and geographic surrogate variables for each species (see Figures S32--S34 for workflow).
 To describe the geographic location of each planning unit, we projected the grid into an equi-distant coordinate system (Europe Equidistant Conic; $\texttt{ESRI:102031}$), calculated the centroid of each grid cell, and extracted their two-dimensional coordinates.
 To describe the environmental characteristics of each planning unit, we obtained 19 climatic layers ($30^{\prime\prime}$ resolution; [@r33]), projected them and the planning units into an equal-area coordinate system (Europe Lambert Conformal Conic; $\texttt{ESRI:102014}$), and computed planning unit averages for each climatic variable.
-To reduce dimensionality, for each species, we subject the climatic values associated with their planning units to a principal components analysis (PCA; Table S1).
+To reduce dimensionality, for each species, we subject the climatic values associated with their planning units to a principal components analysis (PCA; Table S5).
 We used the first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components to characterise climatic variation found across the species' distributions inside the study area (capturing `r round(mean(spp.pca.DF[[2]]*100),2)` % $\pm$ `r round(sd(spp.pca.DF[[2]]*100),2)` SD of the total climatic variation; Figure S2).
@@
 
-To investigate the effectiveness of our supposed surrogates, we quantified the main spatial patterns of adaptive and neutral genetic variation (see Figures S59 and S62 for workflow).
+To investigate the effectiveness of our supposed surrogates, we quantified the main spatial patterns of adaptive and neutral genetic variation (see Figures S32 and S35 for workflow).
 To do this, we first needed to identify which of the sampled loci were adaptive.
@@
 We adopted a similar methodology to Bothwell _et al._ ([-@r455]) and applied it to each of the species separately.
-Following their methodology, we initially grouped conspecifics into genetic lineages to further minimise false-positives.
+Following their methodology, we initially grouped conspecifics into genetic lineages (Figures S36--S62).
 We grouped conspecifics by fitting admixture and correlated alleles models implemented in \texttt{Structure} (version 2.3.4; [@r462]; `r structure.params.LST[[MODE]]$numruns` replicates per species; `r format(structure.params.LST[[MODE]]$admburnin, big.mark=',', scientific=FALSE)` admixture burnin iterations; `r format(structure.params.LST[[MODE]]$burnin, big.mark=',', scientific=FALSE)` burnin iterations; `r format(structure.params.LST[[MODE]]$numreps+structure.params.LST[[MODE]]$burnin, big.mark=',', scientific=FALSE)` total iterations) using the number of lineages previously determined by Alvarez _et al._ ([@r478]) and combining replicate runs using \texttt{ClumPP} (version 1.1.2; [@r459]; greedy algorithm based on the $G^{\prime}$ statistic; `r format(clumpp.params.LST[[MODE]]$repeats, big.mark=',', scientific=FALSE)` iterations).

### 2016-11-13 code/rmarkdown/supporting_information.Rmd
+ fix supp info figures + tables + reorder si figures and tables + update si figures and table numbers in main tex

@@
 # Supporting Information
-All data and code is stored in an online repository (\url{www.github.com/jeffreyhanson/genetic.surrogates}) to replicate analyses in this study. See the \textit{README} file for instructions.
+All data and code is stored in an online repository (\url{www.github.com/jeffreyhanson/genetic-surrogates}) to replicate analyses in this study. See the \textit{README} file for instructions.
 
@@
   # skip species if not used in analysis
-  curr.FPLY <- grid.FPLY[,
-    c(adapt.col, neutral.col, env.col,
-    'long','lat','group'),drop=FALSE]
-  curr.FPLY <- curr.FPLY[which(!is.na(curr.FPLY[[1]])),]
-  curr.FPLY <- curr.FPLY %>% gather(d, value, contains('adaptive'), contains('neutral'), contains('env')) %>%
-    mutate(
-      d=gsub(paste0(unique(spp.samples.DF$species)[i], '_'), '', d, fixed=TRUE),
-      d=gsub('_', ' ', d, fixed=TRUE))
-  for (j in seq_len(length(adapt.col)+length(neutral.col)+length(env.col)))
-    curr.FPLY[['d']] <- gsub(as.character(j), paste0('(', j, ')'), curr.FPLY[['d']], fixed=TRUE)
-  curr.FPLY <- curr.FPLY %>% mutate(d=factor(gsub(' d(', ' (', d, fixed=TRUE)))
+  curr.FPLY <- grid.FPLY %>%
+    select(one_of(adapt.col, neutral.col, env.col), group, long, lat) %>%
+    mutate_each(funs(unname), one_of(adapt.col, neutral.col, env.col)) %>%
+    filter(!is.na(.[[1]])) %>%
+    gather(d, value, contains('adaptive'), contains('neutral'), contains('env')) %>%
+    mutate(d=as.character(d)) %>%
+    mutate(d=gsub(paste0(unique(spp.samples.DF$species)[i], '_'), '', d, fixed=TRUE)) %>%
+    mutate(d=gsub('_', ' ', d, fixed=TRUE)) %>%
+    mutate(d=gsub('env', 'environmental', d, fixed=TRUE))
+  for (j in seq_len(max(length(adapt.col),length(neutral.col),length(env.col))))
+    curr.FPLY$d <- gsub(paste0('d', as.character(j)), paste0('(', j, ')'), curr.FPLY$d, fixed=TRUE)
   # make plot
@@
         guide=guide_colorbar(ticks=element_line(color='black'),
-          border=element_line(color='black'), barheight=unit(3.0, 'cm'))
+          border=element_line(color='black'), barheight=unit(3.0, 'cm')),
+        labels=function(x) {formatC(x, format='f', digits=2)}
       ) + facet_wrap(~d)
@@
   # exports
+  if (n.plots == 0) return(1)
   if (n.plots >= 3) return((((n.plots-1) %/% 3)+1) * 2)
@@
   # exports
+  if (n.plots == 0) return(1)
   if (n.plots >= 3) return(10)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(1), fig.height=calc.fig.height(1), fig.width=calc.fig.width(1), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[1]),'}.
-Each panel shows a different dimension of the variation as indicated by its title.
+```{r, echo=FALSE, fig.width=calc.fig.width(1), fig.height=calc.fig.height(1), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[1]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+Each panel shows a different dimension of variation as indicated by its title.
 Squares represent planning units.
-The color of each planning unit panel corresponds to the average ordinated values associated with individuals sampled inside it.
-No indiviudals of this species were found inside the gray planning units.')}
+Samples collected from the species were subject to NMDS to ordinate the main axes of adaptive (if detected) and neutral genetic variation.
+The bioclimatic data were clipped to planning units where the species was found, and subject to a principal components analysis to summarise the main gradients of environmental variation.
+The panels show the spatial distribution of the ordinated environmental and genetic data averaged to the planning-unit level.
+Planning units where individuals were not observed are shown in gray.')}
 if (is.speciesPlottable(1)) plot.spp.spaces(1)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(2), fig.height=calc.fig.height(2), fig.width=calc.fig.width(2), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[2]),'}.
-See Figure S29 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(2), fig.height=calc.fig.height(2), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[2]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+See Figure S1 caption for conventions.')}
 if (is.speciesPlottable(2)) plot.spp.spaces(2)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(3), fig.height=calc.fig.height(3), fig.width=calc.fig.width(3), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[3]),'}.
-See Figure S29 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(3), fig.height=calc.fig.height(3), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[3]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+See Figure S1 caption for conventions.')}
 if (is.speciesPlottable(3)) plot.spp.spaces(3)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(4), fig.height=calc.fig.height(4), fig.width=calc.fig.width(4), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[4]),'}.
-See Figure S29 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(4), fig.height=calc.fig.height(4), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[4]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+See Figure S1 caption for conventions.')}
 if (is.speciesPlottable(4)) plot.spp.spaces(4)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(5), fig.height=calc.fig.height(5), fig.width=calc.fig.width(5), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[5]),'}.
-See Figure S29 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(5), fig.height=calc.fig.height(5), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[5]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+See Figure S1 caption for conventions.')}
 if (is.speciesPlottable(5)) plot.spp.spaces(5)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(6), fig.height=calc.fig.height(6), fig.width=calc.fig.width(6), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[6]),'}.
-See Figure S29 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(6), fig.height=calc.fig.height(6), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[6]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+See Figure S1 caption for conventions.')}
 if (is.speciesPlottable(6)) plot.spp.spaces(6)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(7), fig.height=calc.fig.height(7), fig.width=calc.fig.width(7), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[7]),'}.
-See Figure S29 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(7), fig.height=calc.fig.height(7), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[7]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+See Figure S1 caption for conventions.')}
 if (is.speciesPlottable(7)) plot.spp.spaces(7)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(8), fig.height=calc.fig.height(8), fig.width=calc.fig.width(8), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[8]),'}.
-See Figure S29 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(8), fig.height=calc.fig.height(8), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[8]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+See Figure S1 caption for conventions.')}
 if (is.speciesPlottable(8)) plot.spp.spaces(8)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(9), fig.height=calc.fig.height(9), fig.width=calc.fig.width(9), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[9]),'}.
-See Figure S29 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(9), fig.height=calc.fig.height(9), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[9]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+See Figure S1 caption for conventions.')}
 if (is.speciesPlottable(9)) plot.spp.spaces(9)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(10), fig.height=calc.fig.height(10), fig.width=calc.fig.width(10), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[10]),'}.
-See Figure S29 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(10), fig.height=calc.fig.height(10), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[10]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+See Figure S1 caption for conventions.')}
 if (is.speciesPlottable(10)) plot.spp.spaces(10)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(11), fig.height=calc.fig.height(11), fig.width=calc.fig.width(11), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[11]),'}.
-See Figure S29 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(11), fig.height=calc.fig.height(11), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[11]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+See Figure S1 caption for conventions.')}
 if (is.speciesPlottable(11)) plot.spp.spaces(11)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(12), fig.height=calc.fig.height(12), fig.width=calc.fig.width(12), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[12]),'}.
-See Figure S29 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(12), fig.height=calc.fig.height(12), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[12]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+See Figure S1 caption for conventions.')}
 if (is.speciesPlottable(12)) plot.spp.spaces(12)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(13), fig.height=calc.fig.height(13), fig.width=calc.fig.width(13), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[13]),'}.
-See Figure S29 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(13), fig.height=calc.fig.height(13), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[13]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+See Figure S1 caption for conventions.')}
 if (is.speciesPlottable(13)) plot.spp.spaces(13)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(14), fig.height=calc.fig.height(14), fig.width=calc.fig.width(14), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[14]),'}.
-See Figure S29 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(14), fig.height=calc.fig.height(14), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[14]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+See Figure S1 caption for conventions.')}
 if (is.speciesPlottable(14)) plot.spp.spaces(14)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(15), fig.height=calc.fig.height(15), fig.width=calc.fig.width(15), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[15]),'}.
-See Figure S29 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(15), fig.height=calc.fig.height(15), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[15]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+See Figure S1 caption for conventions.')}
 if (is.speciesPlottable(15)) plot.spp.spaces(15)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(16), fig.height=calc.fig.height(16), fig.width=calc.fig.width(16), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[16]),'}.
-See Figure S29 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(16), fig.height=calc.fig.height(16), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[16]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+See Figure S1 caption for conventions.')}
 if (is.speciesPlottable(16)) plot.spp.spaces(16)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(17), fig.height=calc.fig.height(17), fig.width=calc.fig.width(17), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[17]),'}.
-See Figure S29 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(17), fig.height=calc.fig.height(17), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[17]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+See Figure S1 caption for conventions.')}
 if (is.speciesPlottable(17)) plot.spp.spaces(17)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(18), fig.height=calc.fig.height(18), fig.width=calc.fig.width(18), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[18]),'}.
-See Figure S29 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(18), fig.height=calc.fig.height(18), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[18]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+See Figure S1 caption for conventions.')}
 if (is.speciesPlottable(18)) plot.spp.spaces(18)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(19), fig.height=calc.fig.height(19), fig.width=calc.fig.width(19), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[19]),'}.
-See Figure S29 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(19), fig.height=calc.fig.height(19), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[19]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+See Figure S1 caption for conventions.')}
 if (is.speciesPlottable(19)) plot.spp.spaces(19)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(20), fig.height=calc.fig.height(20), fig.width=calc.fig.width(20), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[20]),'}.
-See Figure S29 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(20), fig.height=calc.fig.height(20), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[20]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+See Figure S1 caption for conventions.')}
 if (is.speciesPlottable(20)) plot.spp.spaces(20)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(21), fig.height=calc.fig.height(21), fig.width=calc.fig.width(21), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[21]),'}.
-See Figure S29 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(21), fig.height=calc.fig.height(21), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[21]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+See Figure S1 caption for conventions.')}
 if (is.speciesPlottable(21)) plot.spp.spaces(21)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(22), fig.height=calc.fig.height(22), fig.width=calc.fig.width(22), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[22]),'}.
-See Figure S29 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(22), fig.height=calc.fig.height(22), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[22]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+See Figure S1 caption for conventions.')}
 if (is.speciesPlottable(22)) plot.spp.spaces(22)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(23), fig.height=calc.fig.height(23), fig.width=calc.fig.width(23), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[23]),'}.
-See Figure S29 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(23), fig.height=calc.fig.height(23), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[23]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+See Figure S1 caption for conventions.')}
 if (is.speciesPlottable(23)) plot.spp.spaces(23)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(24), fig.height=calc.fig.height(24), fig.width=calc.fig.width(24), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[24]),'}.
-See Figure S29 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(24), fig.height=calc.fig.height(24), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[24]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+See Figure S1 caption for conventions.')}
 if (is.speciesPlottable(24)) plot.spp.spaces(24)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(25), fig.height=calc.fig.height(25), fig.width=calc.fig.width(25), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[25]),'}.
-See Figure S29 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(25), fig.height=calc.fig.height(25), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[25]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+See Figure S1 caption for conventions.')}
 if (is.speciesPlottable(25)) plot.spp.spaces(25)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(26), fig.height=calc.fig.height(26), fig.width=calc.fig.width(26), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[26]),'}.
-See Figure S29 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(26), fig.height=calc.fig.height(26), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[26]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+See Figure S1 caption for conventions.')}
 if (is.speciesPlottable(26)) plot.spp.spaces(26)
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(27), fig.height=calc.fig.height(27), fig.width=calc.fig.width(27), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[27]),'}.
-See Figure S29 caption for conventions.')}
+```{r, echo=FALSE, fig.width=calc.fig.width(27), fig.height=calc.fig.height(27), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[27]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+See Figure S1 caption for conventions.')}
 if (is.speciesPlottable(27)) plot.spp.spaces(27)
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=calc.fig.height3(n_distinct(geo.surrogacy.DF$Species)), fig.cap='The relationship between the proportion of neutral genetic variation secured in a prioritizaiton and the proportion of geographic variation it also secures.
-Conventions are detailed in Figure S56.'}
+Conventions are detailed in Figure S28.'}
 # format data
@@
   file = tempfile(fileext='.png')
-  rsvg::rsvg_png(charToRaw(DiagrammeRsvg::export_svg(x)), file)
+  rsvg::rsvg_png(charToRaw(DiagrammeRsvg::export_svg(DiagrammeR::grViz(x))), file)
   grid::grid.raster(png::readPNG(file))
@@
 Lines show relationships between data and analysis, and their captions describe how data was processed.'}
-render_gvis(grViz("
+render_gvis("
 digraph main {
@@
 }
-"))
-```
-
-```{r, echo=FALSE, results='hide', message=FALSE, fig.cap='Geographic surrogate processing workflow. Conventions are detailed in Figure S59 caption.'}
-render_gvis(grViz("
+")
+```
+
+```{r, echo=FALSE, results='hide', message=FALSE, fig.cap='Geographic surrogate processing workflow. Conventions are detailed in Figure S32 caption.'}
+render_gvis("
 digraph geo_as {
@@
 }
-"))
-```
-
-```{r, fig.height=6, echo=FALSE, results='hide', message=FALSE, fig.cap='Environmental surrogate processing workflow. Conventions are detailed in Figure S59 caption.'}
-render_gvis(grViz("
+")
+```
+
+```{r, fig.height=6, echo=FALSE, results='hide', message=FALSE, fig.cap='Environmental surrogate processing workflow. Conventions are detailed in Figure S32 caption.'}
+render_gvis("
 digraph env_as {
@@
 }
-"))
-```
-
-```{r, fig.height=10, echo=FALSE, results='hide', message=FALSE, fig.cap='Genetic data processing workflow. Conventions are detailed in Figure S59 caption.'}
-render_gvis(grViz("
+")
+```
+
+```{r, fig.height=10, echo=FALSE, results='hide', message=FALSE, fig.cap='Genetic data processing workflow. Conventions are detailed in Figure S32 caption.'}
+render_gvis("
 digraph genetic_as {
@@
 }
-"))
+")
 ```
@@
 ```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[2],fixed=TRUE),'} populations.
-See Figure S3 caption for conventions.')}
+See Figure S36 caption for conventions.')}
 if (2 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(2)
@@
 ```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[3],fixed=TRUE),'} populations.
-See Figure S3 caption for conventions.')}
+See Figure S36 caption for conventions.')}
 if (3 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(3)
@@
 ```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[4],fixed=TRUE),'} populations.
-See Figure S3 caption for conventions.')}
+See Figure S36 caption for conventions.')}
 if (4 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(4)
@@
 ```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[5],fixed=TRUE),'} populations.
-See Figure S3 caption for conventions.')}
+See Figure S36 caption for conventions.')}
 if (5 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(5)
@@
 ```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[6],fixed=TRUE),'} populations.
-See Figure S3 caption for conventions.')}
+See Figure S36 caption for conventions.')}
 if (6 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(6)
@@
 ```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[7],fixed=TRUE),'} populations.
-See Figure S3 caption for conventions.')}
+See Figure S36 caption for conventions.')}
 if (7 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(7)
@@
 ```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[8],fixed=TRUE),'} populations.
-See Figure S3 caption for conventions.')}
+See Figure S36 caption for conventions.')}
 if (8 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(8)
@@
 ```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[9],fixed=TRUE),'} populations.
-See Figure S3 caption for conventions.')}
+See Figure S36 caption for conventions.')}
 if (9 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(9)
@@
 ```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[10],fixed=TRUE),'} populations.
-See Figure S3 caption for conventions.')}
+See Figure S36 caption for conventions.')}
 if (10 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(10)
@@
 ```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[11],fixed=TRUE),'} populations.
-See Figure S3 caption for conventions.')}
+See Figure S36 caption for conventions.')}
 if (12 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(11)
@@
 ```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[12],fixed=TRUE),'} populations.
-See Figure S3 caption for conventions.')}
+See Figure S36 caption for conventions.')}
 if (12 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(12)
@@
 ```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[13],fixed=TRUE),'} populations.
-See Figure S3 caption for conventions.')}
+See Figure S36 caption for conventions.')}
 if (13 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(13)
@@
 ```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[14],fixed=TRUE),'} populations.
-See Figure S3 caption for conventions.')}
+See Figure S36 caption for conventions.')}
 if (14 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(14)
@@
 ```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[15],fixed=TRUE),'} populations.
-See Figure S3 caption for conventions.')}
+See Figure S36 caption for conventions.')}
 if (15 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(15)
@@
 ```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[16],fixed=TRUE),'} populations.
-See Figure S3 caption for conventions.')}
+See Figure S36 caption for conventions.')}
 if (16 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(16)
@@
 ```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[17],fixed=TRUE),'} populations.
-See Figure S3 caption for conventions.')}
+See Figure S36 caption for conventions.')}
 if (17 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(17)
@@
 ```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[18],fixed=TRUE),'} populations.
-See Figure S3 caption for conventions.')}
+See Figure S36 caption for conventions.')}
 if (18 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(18)
@@
 ```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[19],fixed=TRUE),'} populations.
-See Figure S3 caption for conventions.')}
+See Figure S36 caption for conventions.')}
 if (19 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(19)
@@
 ```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[20],fixed=TRUE),'} populations.
-See Figure S3 caption for conventions.')}
+See Figure S36 caption for conventions.')}
 if (20 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(20)
@@
 ```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[21],fixed=TRUE),'} populations.
-See Figure S3 caption for conventions.')}
+See Figure S36 caption for conventions.')}
 if (21 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(21)
@@
 ```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[22],fixed=TRUE),'} populations.
-See Figure S3 caption for conventions.')}
+See Figure S36 caption for conventions.')}
 if (22 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(22)
@@
 ```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[23],fixed=TRUE),'} populations.
-See Figure S3 caption for conventions.')}
+See Figure S36 caption for conventions.')}
 if (23 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(23)
@@
 ```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[24],fixed=TRUE),'} populations.
-See Figure S3 caption for conventions.')}
+See Figure S36 caption for conventions.')}
 if (24 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(24)
@@
 ```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[25],fixed=TRUE),'} populations.
-See Figure S3 caption for conventions.')}
+See Figure S36 caption for conventions.')}
 if (25 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(25)
@@
 ```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[26],fixed=TRUE),'} populations.
-See Figure S3 caption for conventions.')}
+See Figure S36 caption for conventions.')}
 if (26 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(26)
+```
+
+```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[27],fixed=TRUE),'} populations.
+See Figure S36 caption for conventions.')}
+if (27 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(27)
 ```
@@
   gsub(pattern='\\cline{1-', replacement='\\cline{3-4} \\cline{6-', fixed=TRUE) %>% 
-	gsub(pattern='toprule', replacement='toprule[1pt]', fixed=TRUE) %>%
-	gsub(pattern='bottomrule', replacement='bottomrule[1pt]', fixed=TRUE) %>%
+  gsub(pattern='\\toprule', replacement='', fixed=TRUE) %>%
+  gsub(pattern='\\midrule', replacement='\\midrule', fixed=TRUE) %>%
+  gsub(pattern='\\bottomrule', replacement='\\midrule', fixed=TRUE) %>%
   cat()
@@
 Similarly, for each species, a model was fit to correlate dissimilarities among the planning units in terms of the neutral genetic characteristics of indiviudals in them with dissimilarities in terms of the geographic position of the planning units (denoted in the geographic column group).
-Data shows the results from comparing each model to its corresponding null model using $\\mathbf{\\chi^2}$ tests (1 degree of freedom), and the $\\mathbf{R^2}$ ($\\mathbf{MR^2}$) and conditional $\\mathbf{R^2}$ ($\\mathbf{CR^2}$) for each model.'
+Data shows the results from comparing each model to its corresponding null model using $\\chi^2$ tests (1 degree of freedom), and the $R^2$ ($MR^2$) and conditional $R^2$ ($CR^2$) for each model.'
   )
@@
 These models also included a predictor variable indicating the species for which the prioritisations were generated, and an interaction term.
-Table shows the results from one-sided _post-hoc_ analyses that were conducted to determine if the slopes for each species were positive (using the alternative hypothesis that slopes were less than or equal to zero; tests used a single degree of freedom; results reported in the \\textit{Z} and \\textit{P} columns).
+Table shows the results from one-sided \\textit{post-hoc} analyses that were conducted to determine if the slopes for each species were positive (using the alternative hypothesis that slopes were less than or equal to zero; tests used a single degree of freedom; results reported in the \\textit{Z} and \\textit{P} columns).
 The table also shows Cragg and Uhler\'s pseudo-$R^2$ values that were calculated to show the proportion of variation explained by the slopes for each species ($R^2$ column).'
@@
   gsub(pattern='ulticolumn{1}{c}{$R^2$}\\tabularnewline', replacement='ulticolumn{1}{c}{$R^2$}\\T\\B\\tabularnewline', fixed=TRUE) %>% 
-	gsub(pattern='toprule', replacement='toprule[1pt]', fixed=TRUE) %>%
-	gsub(pattern='bottomrule', replacement='bottomrule[1pt]', fixed=TRUE) %>%
+  gsub(pattern='\\toprule', replacement='', fixed=TRUE) %>%
+  gsub(pattern='\\midrule', replacement='\\midrule', fixed=TRUE) %>%
+  gsub(pattern='\\bottomrule', replacement='\\midrule', fixed=TRUE) %>%
   cat()

### 2016-11-13 code/R/analysis/00-initialization.R

@@ -24,6 +24,7 @@ library(qvalue)
 
 # load CRAN packages
 library(stats)
+library(hexbin)
 library(DiagrammeR)
 library(DiagrammeRsvg)
 library(rsvg)

### 2016-11-13 code/rmarkdown/article.Rmd
+ assorted changes

@@
 Based on the results from our previous aim, we used `r (raptr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100)` % surrogate-based targets and `r (raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)` % genetic-based targets.
-We generated a single solution for each target/scenario combination, except for the "single-species (equal costs) with amount-based targets" cominbation for which we generated `r format(raptr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates, big.mark=',', scientific=FALSE)` replicates because it had many optimal solutions.
+We generated a single solution for each target/scenario combination, except for the "single-species (equal costs) with amount-based targets" combination for which we generated `r format(raptr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates, big.mark=',', scientific=FALSE)` replicates because it had many optimal solutions.
 We computed the proportion of adaptive (if present) and neutral genetic variation secured for each species in the prioritisations.

### 2016-11-14 code/rmarkdown/article.Rmd
+ fix bug causing refs not to appear in article

--- 2016-11-10 code/rmarkdown/article.Rmd
@@
 <!-- scenario analysis -->
-To determine whether using environmental and geographic targets could improve the effectiveness of prioritisations in representing genetic variation, we generated prioritisations using (i) "amount targets" reflecting the traditional approach of representing a certain proportion of each species' geographic distribution, (ii) "amount and surrogate targets" where we also targeted representation of the environmental and geographic surrogates as well as a certain proportion of each species' geographic distribution, and (iii) "amount and genetic targets" where we targeted representation of the directly measured genetic variation as well as a certain proportion of each species' geographic distribution.
-To determine if these patterns were robust, for each of the three combinations of the targets, we generated prioritisations under four scenarios: (i) single-species with equal costs, (ii) single species with opportunity costs, (iii) multi-species with equal costs, and (iv) multi-species with opportunity costs (Figure 3, Figure SXXX).
+To determine whether using environmental and geographic targets could improve the effectiveness of prioritisations in representing genetic variation, we generated prioritisations using (i) "amount targets" reflecting the traditional approach of representing a certain proportion of each species' geographic distribution, (ii) "amount and surrogate targets" where we also targeted representation of the environmental and geographic surrogates as well as a certain proportion of each species' geographic distribution, and (iii) "amount and genetic targets" where we targeted representation of the directly measured genetic variation as well as a certain proportion of each species' geographic distribution (Figure 3).
+To determine if these patterns were robust, for each of the three combinations of the targets, we generated prioritisations under four scenarios: (i) single-species with equal costs, (ii) single species with opportunity costs, (iii) multi-species with equal costs, and (iv) multi-species with opportunity costs.
 
@@
 
-We created environmental and geographic surrogate variables for each species (see Figures S59--S561 for workflow).
+We created environmental and geographic surrogate variables for each species.
 To describe the geographic location of each planning unit, we projected the grid into an equi-distant coordinate system (Europe Equidistant Conic; $\texttt{ESRI:102031}$), calculated the centroid of each grid cell, and extracted their two-dimensional coordinates.
@@
 
-To investigate the effectiveness of our supposed surrogates, we quantified the main spatial patterns of adaptive and neutral genetic variation (see Figures S59 and S62 for workflow).
-To do this, we first needed to identify which of the sampled loci were adaptive.
+To investigate the effectiveness of our supposed surrogates, we first needed to identify which of the sampled loci were adaptive.
 Following best practices, we used two outlier detection methods to achieve this (an individual- and a population-level method; [@r495]).

### 2017-01-22 code/rmarkdown/article.Rmd
assorted improvements to article main text

@@
 abstract: |
-  Protected areas buffer species from anthropogenic threats and provide places for processes to continue.
+  Protected areas buffer species from anthropogenic threats and provide places for biodiversity processes to continue.
 However, evolutionary processes are not often explicitly considered when deciding where to site protected areas because genetic data require considerable resources to obtain and analyse.
 Here we show that environmental and geographic distance variables are effective surrogates for adaptive and neutral intra-specific variation for conservation planning.
-We obtained occurrence and genetic data from the IntraBioDiv project for 27 species collected over the European Alps using a gridded sampling scheme.
+We obtained occurrence and genetic data from the IntraBioDiv project for 27 alpine plant species collected over the European Alps using a gridded sampling scheme.
 For each species, we identified loci that were potentially under selection using outlier loci methods, and mapped their main gradients of adaptive and neutral genetic variation across the grid cells.
@@
 
-Recently, the use of surrogate data has been proposed to generate prioritisations that sufficiently intra-specific genetic variation--without needing to directly utilise genetic data to achieve this.
+Recently, the use of surrogate data has been proposed to generate prioritisations that sufficiently capture intra-specific genetic variation--without needing to directly utilise genetic data to achieve this.
 Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been proposed to guide reserve selection as surrogates for adaptive genetic variation (eg.
@@
 Potential surrogates for neutral variation have been based on variables that predict the levels of connectivity between planning units.
-For instance, isolation by distance (IBD; [@r259]) predicts that populations further apart will experience less gene flow, and in turn, share less neutral loci.
+For instance, isolation by distance (IBD; [@r259]) predicts that populations located further apart will experience less gene flow, and in turn, share less neutral loci.
 Based on this idea, geographic surrogates based on distance have been found to improve conservation prioritisations in an island network ([@r200]).
@@
 Here we investigate the effectiveness of environmental and geographic surrogates for capturing adaptive and neutral genetic variation respectively in the context of conservation planning.
-We used distribution and genomic data for 27 alpine plants collected by the international IntraBioDiv project in the European Alps ([@r451]).
+We used distribution and genomic data for 27 alpine plant species in the European Alps that was obtained by the IntraBioDiv project ([@r451]).
+The data was collected following a gridded sampling scheme which we adopted as planning units for this study.
 We show that for most species the spatial patterns of adaptive and neutral genetic variation correlated with environmental and geographic variables.
 We also show that for most species the level of association is strong enough to be operational for conservation planning.
-Finally, we demonstrate the using these surrogates results in prioritisations that secure more genetic variation than expected by chance.
+Finally, we demonstrate that using these surrogates results in prioritisations that secure more genetic variation than expected by chance.
 
@@
 We detected putatively adaptive genetic variation in `r as.character(as.english(length(env.surrogacy.LST)))` of the 27 plant species.
-We used two outlier detection methods to detect loci under selection, and they returned reasonably consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species assigned the same classification $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD; Table S1).
+We used two outlier detection methods to identify loci under selection, and they returned reasonably consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species assigned the same classification $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD; Table S1).
 Of the species that were associated with adaptive genetic variation, only a small proportion of loci were classified as being adaptive (mean `r round(mean(spp.outlier.loci.props, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.loci.props, na.rm=TRUE),2)` SD).
 After identifying the loci showing strong signals of selection, we used non-metric multi-dimensional scaling (NMDS) analyses to identify the main gradients of adaptive (if detected) and neutral variation for each species.
-Generally, only a small number of continuous dimensions was needed to sufficiently describe their patterns of adaptive (all $K = `r unique(na.omit(adaptive.k))`$; Table S2) and neutral genetic variation ($K = `r min(neutral.k)`$--$`r max(neutral.k)`$; Table S1).
+Generally, only a small number of continuous dimensions was needed to sufficiently describe their patterns of adaptive ($K = `r unique(na.omit(adaptive.k))`$ for all species; Table S2) and neutral genetic variation ($`r min(neutral.k)` \leq K \leq `r max(neutral.k)`$ for all species; Table S1).
 The resulting ordinations were used to construct an adaptive (if detected) and neutral genetic space for each species.
 The spatial distribution of these genetic spaces (Figures S29--S54) generally showed substantial spatial auto-correlation (eg.
-_Cerastium uniflorum_, Figure S34; _Dryas octopetala_, Figures S36), suggesting that the environmental and geographic variables may prove to be effective surrogates.
+_Cerastium uniflorum_, Figure S34; _Dryas octopetala_, Figure S36), suggesting that the environmental and geographic variables may prove to be effective surrogates.
 
 <!-- correlation analysis -->
-We verified that the spatial patterns in environmental and geographic variation correlated with the broad-scale spatial patterns in adaptive and neutral variation for most species (respectively).
-For each species,we constructed dissimilarity matrices based on the planning units they were found in.
-These matrices expressed differences between the planning units in terms of their (i) environmental characteristics, (ii) geographic position, and also the (iii) adaptive (if detected), and (iv) neutral genetic characteristics of individuals found inside them.
+We verified that the spatial patterns in environmental variation correlated with the broad-scale patterns in adaptive genetic variation.
+We further verified that there was a correlation between neutral genetic variation and variation in geographic position.
+For each species, we constructed dissimilarity matrices expressing differences between the planning units within which they were detected according to terms their (i) environmental characteristics, (ii) geographic position, and also the (iii) adaptive (if adaptive loci detected), and (iv) neutral genetic characteristics of individuals found inside them.
 By correlating these matrices, the spatial patterns of environmental variation were found to be significantly correlated with that of adaptive genetic variation for `r as.character(as.english(sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)))` of the `r as.character(as.english(sum(!is.na(correlation.model.summary.DF$Environmental.corr.P))))` species associated with adaptive variation (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models; Table S3).
@@
 <!-- surrogacy analysis -->
-For each species, we generated a suite of `r format(raptr.params.LST[[MODE]]$surrogacy.analysis$random.replicates, big.mark=',', scientific=FALSE)` random prioritisations, and calculated the proportion of environmental variation and adaptive (if present) genetic variation they sampled.
+For each species, we generated a suite of `r format(raptr.params.LST[[MODE]]$surrogacy.analysis$random.replicates, big.mark=',', scientific=FALSE)` random prioritisations, and calculated the proportion of environmental variation and adaptive (if present) genetic variation each prioritisation captured.
 We then repeated this process, and calculated the proportion of geographic variation and neutral genetic variation they sampled.
@@
 The average proportion of genetic variation secured by prioritisations varied under different scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.2.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.2.DF$p.value[[2]])`).
-Both the multi-species (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)`) and multi-species with cost (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)`) were both found to secure more genetic variation than the single species scenario (mean `r round(mean(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[2]])`$; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$ respectively).
+Both the multi-species (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)`) and multi-species with cost (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)`) were found to secure more genetic variation than the single species scenario (mean `r round(mean(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[2]])`$; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$ respectively).
 These results suggest that on average the multi-species prioritisations can secure more genetic variation than single-species prioritisations--regardless of targets used to generate them.
@@
 We aimed to investigate the effectiveness of broad-scale environmental and geographic variables as surrogates for adaptive and neutral genetic variation in conservation planning.
-Our study is the first to address this aim using a broad range of species and calibrated against field measurements of genetic variation.
-The spatial patterns of environmental and geographic surrogates were strongly correlated with the patterns of adaptive and neutral genetic variation for most species, and environmental and geographic targets resulted in prioritisations that secured a greater proportion of adaptive and neutral genetic variation than traditional conservation planning for most species.
-These environmental and geographic surrogates are based on freely available data, and could be calculated for any study region across the world.
+Our study is the first to address this aim using field measurements of genetic variation for a broad range of species to assess the performance of prioritisations.
+The spatial patterns of environmental and geographic surrogates were strongly correlated with the patterns of adaptive and neutral genetic variation for most species.
+Moreover, for most species, prioritisations generated using environmental and geographic targets secured a greater proportion of adaptive and neutral genetic variation than traditional conservation planning.
+These environmental and geographic surrogates were based on freely available data, and could be applied to any study region across the world.
 
 <!-- surrogate/genetic held correlations -->
-Our results demonstrate that environmental and geographic surrogates are strong enough to be operational (Figure 2).
+Our results demonstrate that environmental and geographic surrogates are strong enough to be operational for conservation planning (Figure 2).
 For most species, the environmental and geographic surrogates variables were effective surrogates for adaptive (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figures 2c and S56) and neutral genetic variation (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figures 2d and Figure S57).
@@
 _Gypsophila repens_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gypsophila_repens')$Geographic.r2CU,2)`; _Luzula alpinopilosa_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Luzula_alpinopilosa')$Geographic.r2CU,2)`; _Gentiana nivalis_ $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gentiana_nivalis')$Geographic.r2CU,2)`).
-Indeed, geographic distances will be a poor surrogate for species that utilise dispersal modes that result in non-linear relationships between connectivity and geographic distance [eg.
-wind dispersing plants can maintain high connectivity between distant areas; @r451], and in landscapes where connectivity is most affected by landscape features ([@r497]).
-Overall, however, such species were the exception in our analysis, and both simple surrogates performed moderately well for a diverse set of species.
+Indeed, geographic distances are expected to be a poor surrogate for connectivity when species deviate from isolation by distance (eg.
+wind dispersing plants can maintain high connectivity between distant areas; [@r451]), and in areas where connectivity is most affected by landscape features ([@r497]).
+Overall, such species were the exception in our analysis, and both surrogates performed moderately well for most species.
 
 <!-- suitable targets for surrogates -->
-Our results suggest that environmental and geographic surrogates can predict how much genetic variation prioritisations secure, but what targets should be used?
-First, our results suggest that surrogate-based targets should be used in conjunction with amount-based targets.
-For all species, it was possible to generate solutions that secured a high proportion of the surrogate variable (> 90 %) and only a small proportion of adaptive or neutral genetic variation (< 20 %; Figures S56--S57), and so use of amount-based targets may ameliorate this.
-Second, geographic surrogates may need much higher targets than environmental surrogates to capture evolutionary processes.
-In our analysis, geographic variation accumulated much slower for neutral genetic representation than environmental variation did with adaptive representation (Figures 2a--2b and S56--S57).
-To secure a representative sample of the species' neutral genetic variation ($\geq$ `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %), prioritisations needed to sample `r round(mean(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` SD of the variation in geographic position among the planning units occupied by the species.
-On the other hand, adaptive variation accumulated much faster.
-To secure at least `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` % of each species' adaptive genetic variation, prioritisations only needed to sample on average `r round(mean(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` SD of the environmental conditions found across the planning units occupied by the species.
-Note that these proportions are based on averages across species.
-In many cases, higher targets may be needed to ensure that prioritisations secure a representative sample of the genetic variation for all species in a planning exercise.
+Our results suggest that environmental and geographic surrogates can be used to capture genetic variation in prioritisations, the next question is: what percentage of surrogate variation should planners target to preserve genetic variation?
+Firstly, our results suggest that surrogate-based targets should be used with amount-based targets.
+For all species, it was possible to generate solutions that secured a high proportion of the surrogate variable (> 90 %) and only a small proportion of genetic variation (< 20 %; Figures S56--S57).
+By using both amount- and surrogate-based targets, planners may avoid such solutions.
+Secondly, geographic surrogates may need higher targets than environmental surrogates to capture evolutionary processes.
+In our analysis, geographic variation accumulated at a slower rate for neutral genetic representation than environmental variation did with adaptive representation (Figures 2a--2b and S56--S57).
+To secure a representative sample of each species' neutral genetic variation ($\geq$ `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %), prioritisations needed to sample at least `r round(mean(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` SD of the variation in geographic position among the planning units occupied by the species.
+On the other hand, to secure at least `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` % of each species' adaptive genetic variation, prioritisations only needed to sample `r round(mean(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` SD of the environmental variation in planning units occupied by the species.
+Although to capture evolutionary processes, real-world conservation planning exercises may need higher surrogate targets.
 
@@
 _Carex semipervens_ and _Cerastium uniflorum_; Figures S33 and S34 respectively).
-As a consequence, a prioritisation would only need to select a planning unit in each cluster to secure a representative sample.
+As a consequence, a prioritisation would only need to select a single planning unit in each cluster to secure a representative sample.
 This mechanism explains why adaptive variation accumulated much quicker than neutral genetic variation in the randomly generated prioritisations.
@@
 There are several limitations associated with our analysis.
-First, the size of the planning units we used (23 $\times$ 25 km) is much larger than typically used in regional conservation planning scenarios (1--10 km$^2$).
+Firstly, the size of the planning units we used (23 $\times$ 25 km) is much larger than typically used in regional conservation planning scenarios (1--10 km$^2$).
 We used this resolution because the genetic data were collected at this scale.
 Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis by artificially introducing additional spatial auto-correlation.
-Second, we used geographic distances as surrogates for neutral genetic variation, but connectivity in this landscape may be better described using distances that incorporate information on topography.
-However, such distances require specific-specific scaling to adequate describe their connectivity (eg.
+Secondly, we used geographic distances as surrogates for neutral genetic variation, but connectivity in this landscape may be better described using distances that incorporate information on topography.
+However, such distances require species-specific scaling to adequate describe their connectivity (eg.
 [@r497]), and so cannot easily be utilised in large-scale planning exercises.
-Third, we used amplified fragment length polymorphism data (AFLP; [@r453]) to characterise the main gradients of genetic variation.
-Whilst next-generation sequencing provides higher resolution genetic information than AFLP data ([@r480]), no multi-species genomic data set with similar properties currently exists, and our methodology would still have utilised only the main gradients of the genetic variation to generate prioritisations in a feasible amount of time.
+Thirdly, we used amplified fragment length polymorphism data (AFLP; [@r453]) to characterise the main gradients of genetic variation.
+Whilst next-generation sequencing provides higher resolution genetic information than AFLP data ([@r480]), no multi-species genomic data set with similar properties currently exists, and our methodology would still have utilised only the main gradients of the genetic variation to capture broad-scale evolutionary processes.
 
@@
 Moreover, data to calculate the surrogate variables are freely available for any location in the world.
-Careful use of such surrogates in conservation planning could vastly improve the chances of long-term biodiversity persistence for relatively additional cost.
+Careful use of such surrogates in conservation planning could vastly improve the chances of long-term biodiversity persistence for relatively little additional cost.
 
@@
 We used data for 27 alpine plant species in the European Alps collected by the IntraBioDiv project ([@r451]; Figure 2a).
-All data, code, and results are stored in an online repository (\url{www.github.com/jeffreyhanson/genetic.surrogates}) to replicate and validate this study.
-This data set has previously been used to explore patterns of adaptive (eg.
-[@r475; @r455]) and neutral genetic variation (eg.
-[@r478]), and to show that species richness is not an effective surrogate for genetic diversity in conservation planning ([@r486]).
-It was collected using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx.
+All data, code, and results are stored in an online repository to replicate and validate this study (\url{www.github.com/jeffreyhanson/genetic.surrogates}).
+This data set has been used to explore patterns of adaptive (eg.
+[@r455]) and neutral genetic variation (eg.
+[@r478]), and the potential for species richness as a potential surrogate for genetic diversity ([@r486]).
+Data was collected using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx.
 22.3 km $\times$ 25 km; Figure S1).
-Project members visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals.
+Project members visited every second grid cell, and for each species, they collected samples from three individuals in each cell they were detected in.
 They genotyped samples using AFLPs, and constructed matrices denoting the presence/absence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; for more information see [@r478], Table 1).
-Thus the data set contains information describing the genomic properties of individuals for each species present in every second grid cell.
-We used this data set because it comprised of comparable genetic data for a range of species with different evolutionary and life histories collected using a standardised sampling scheme, and pertained to a study area with substantial environmental variation to test the effectiveness of environmental variables.
+Thus the data set contains information describing the genomic properties of individuals for each species over a geographic grid.
+We used this data set because it was comprised of comparable genetic data for a range of species with different evolutionary and life histories that was collected using a standardised sampling scheme.
+Additionally, the study area contained large environmental variation which is useful for examining environmental surrogates.
 
@@
 
-We adopted the sampling grid used by the IntraBioDiv project as planning units to develop prioritisations, omitting cells that were not found to contain individuals belonging to the study species (`r nrow(grid.DF)` grid cells used in analysis).
+We adopted the sampling grid used to collect data as planning units to develop prioritisations.
+To reduce computational burden, we omitted cells that did not contain samples (`r nrow(grid.DF)` grid cells used in analysis).
 To examine the potential effects of variation in cost, we calculated the total human population density inside each planning unit (1 km$^2$ resolution from the Global Rural-Urban Mapping Project; [@r472]) and used this to represent acquisition cost (Figure 2b).
-All spatial and statistical analysis were conducted in \texttt{R} (version `r paste0(R.version$major, '.', R.version$minor)`; [@r192]).
+All spatial and statistical analyses were conducted in \texttt{R} (version `r paste0(R.version$major, '.', R.version$minor)`; [@r192]).
 
@@
 To describe the environmental characteristics of each planning unit, we obtained 19 climatic layers ($30^{\prime\prime}$ resolution; [@r33]), projected them and the planning units into an equal-area coordinate system (Europe Lambert Conformal Conic; $\texttt{ESRI:102014}$), and computed planning unit averages for each climatic variable.
-To reduce dimensionality, for each species, we subject the climatic values associated with their planning units to a principal components analysis (PCA; Table S1).
-We used the first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components to characterise climatic variation found across the species' distributions inside the study area (capturing `r round(mean(spp.pca.DF[[2]]*100),2)` % $\pm$ `r round(sd(spp.pca.DF[[2]]*100),2)` SD of the total climatic variation; Figure S2).
+To reduce dimensionality, for each species, we subjected the climatic values associated with their planning units to a principal components analysis (PCA; Table S1).
+We used the first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components to characterise climatic variation found across the species' geographic distributions inside the study area (capturing `r round(mean(spp.pca.DF[[2]]*100),2)` % $\pm$ `r round(sd(spp.pca.DF[[2]]*100),2)` SD of the total climatic variation; Figure S2).
 Thus we constructed a two-dimensional geographic space as a potential surrogate for neutral genetic variation, and a `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))`-dimensional environmental space as a potential surrogate for adaptive genetic variation for each species.
 In these spaces, each planning unit was associated with a single point.
-Planning units with more similar environmental conditions, or higher spatial proximity in the real world, were associated with points that were closer together in these spaces (respectively).
+Planning units with more comparable environmental conditions, or higher spatial proximity in the real world, were associated with points that were closer together in each of these spaces.
 
@@
 The basic premise underpinning such methods is that neutral loci are expected to exhibit a certain level of variation, and loci that deviate from this expectation are likely to be under selection ([@r480]).
-The advantage of these methods--in contrast to environmental association analyses--is that they do not use environmental data, which would have introduced circularity into our analysis.
+The advantage of these methods--in contrast to environmental association analyses--is that they do not use environmental data, which would have introduced circularity into our analysis and invalidated it.
 Loci identified as outliers in both analyses were treated as adaptive and the remainder were treated as neutral.
@@
 To enable comparisons between the two outlier detection methods, we used the same individuals in this analysis as in the \texttt{BayeScan} analysis.
-For each of the 27 species, we first imputed missing data--replacing missing values with the average frequency among conspecifics--ran a PCA, and extracted the minimum number of components needed to secure the over-arching population level variation among the loci (`r pcadapt.params.LST[[MODE]]$min.variance.explained*100` % of the variation in loci).
+For each of the 27 species, we first imputed missing data by replacing missing values with the average frequency among conspecifics, then ran a PCA over the loci matrix, and extracted the minimum number of components needed to secure the over-arching population level variation among the loci ( `r pcadapt.params.LST[[MODE]]$min.variance.explained*100` % of the variation in loci).
 We then computed _q_-values using Mahalanobis distances, and used the same false discovery rate used in the \texttt{BayeScan} analysis to identify loci under selection (using the \texttt{qvalue R} package; [@r491]).
@@
 
-Thus we constructed an an adaptive (if detected) and neutral genetic space for each species.
+Thus we constructed an adaptive (if detected) and neutral genetic space for each species.
 For a given species, each planning unit that it was found in was associated with a multi-dimensional point in the species' adaptive genetic space (if adaptive genetic variation was detected) and another multi-dimensional point in the species' neutral genetic space.
-Planning units that were closer together in these genetic spaces were occupied by individuals with more similar polymorphisms.
+Planning units that were closer together in these genetic spaces were occupied by individuals with more comparable AFLP polymorphisms.
 By spreading out conservation effort across these genetic spaces, and in turn selecting planning units occupied by individuals with increasingly different polymorphisms, prioritisations can secure more genetic variation.
@@
 We used the environmental and geographic surrogate spaces, and the adaptive (if detected) and neutral genetic spaces to generate and evaluate solutions.
-Solutions associated with negative values--because they secured a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis.
+Solutions associated with negative values--because they secured a minute fraction of genetic variation--were replaced with zeros to facilitate statistical analysis.
 We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality (using the \texttt{raptr R} package [@r465]).
@@
 Our first aim was to determine if the environmental and geographic variables correlate with spatial patterns of adaptive and neutral genetic variation.
-To achieve this, for each species, we generated two dissimilarity matrices between the planning units that it occupied based on their environmental conditions and geographic position (using their respective surrogate spaces).
-We then generated another dissimilarity matrix between the planning units using their values in the species' neutral genetic space.
-If the species was associated with adaptive loci, we also generated another dissimilarity matrix between the planning units using the species' adaptive genetic space.
-Dissimilarities were compured as Euclidean distances.
+To achieve this, for each species, we generated multiple dissimilarity matrices expressing differences between the planning units where the species was found.
+Using data from the species' surrogate and genetic spaces, these dissimilarity matrices showed differences between the planning units in terms of their (i) geographic position, (ii) environmental characteristics, and the (iii) adaptive (if detected) and (iv) neutral genetic characteristics of the individuals inside them.
+Dissimilarities were computed as Euclidean distances.
 
-We then fit maximum likelihood population-effects (MLPE) models ([@r494]).
-These models accommodate the structure of dissimilarity matrices using random effects.
-For each species, we fit a MLPE model to the species' dissimilarity matrices based on geographic position and neutral genetic variation.
-Additionally, if adaptive loci were detected, we also fit a MLPE model to the species' dissimilarity matrices describing environmental and adaptive genetic variation.
-All predictor variables were _z_-transformed prior to model fitting to improve convergence.
-To test if the surrogates explained a substantial amount of genetic variation, we compared each model to its corresponding null model using $\chi^2$, and applied Bonferroni corrections.
+We then fitted maximum likelihood population-effects (MLPE) models ([@r494]).
+These models accommodated the structure of dissimilarity matrices using random effects.
+For each species, we fitted a MLPE model to the species' dissimilarity matrices based on geographic position and neutral genetic variation.
+Additionally, if adaptive loci were detected, we also fitted a MLPE model to the species' dissimilarity matrices describing environmental and adaptive genetic variation.
+All predictor variables were _z_-transformed to improve convergence.
+To test if the surrogates explained a substantial amount of genetic variation, we compared each model to its null model using $\chi^2$, and applied Bonferroni corrections.
 
 Our second aim was to determine if the environmental and geographic variables were effective surrogates for adaptive and neutral genetic variation.
-For each species, we generated `r format(raptr.params.LST[[MODE]]$surrogacy.analysis$random.replicates, big.mark=',', scientific=FALSE)` prioritisations by randomly selecting different combinations of planning units that it occupied, and calculated the proportion of geographic variation and neutral genetic variation it secured.
-We repeated this processes, and calculated the proportion of environmental variation and adaptive genetic variation they secured.
+For each species, we generated `r format(raptr.params.LST[[MODE]]$surrogacy.analysis$random.replicates*2, big.mark=',', scientific=FALSE)` prioritisations by randomly selecting different combinations of planning units that the species occupied.
+For half of these random prioritisations, we calculated the proportion of geographic variation and neutral genetic variation they secured, and for the other half, the proportion of environmental variation and adaptive genetic variation they secured.
 
-We fit two full generalised linear models (GLM) with logit links.
+We fitted two full generalised linear models (GLM) with logit links.
 The first model was fit to the proportion of adaptive genetic variation secured in a prioritisation using the proportion of environmental variation also secured in the prioritisation.
-The second model was fit to the proportion of neutral genetic variation secured in a prioritisation using proportion of geographic variation also secured in the prioritisation.
-Both models also contained a variable indicating the species for which the prioritisations were generated, and an interaction term.
-They were subject to backwards step-wise term deletion routines to assess term significance.
-To assess the performance of the surrogates, we refit these models seperately for each species and computed the Cragg and Uhler's pseudo-R value (using the \texttt{pscl R} package; [@r492]).
+The second model was fit to the proportion of neutral genetic variation secured in a prioritisation using the proportion of geographic variation also secured in the prioritisation.
+Additionally, both models contained a variable indicating the species for which the prioritisations were generated, and an interaction term.
+They were subjected to backwards step-wise term deletion routines to assess term significance.
+To assess the performance of the surrogates, we refit these models separately for each species and computed the Cragg and Uhler's pseudo-$R^2$ value (using the \texttt{pscl R} package; [@r492]).
 
-Our third aim was to determine whether surrogate-based targets improved prioritisations.
+Our third aim was to determine if surrogate-based targets improved prioritisations.
 As previously mentioned, we generated prioritisations using different combinations of targets and conservation planning scenarios.
@@
 We investigated the benefits of using surrogate-based targets in prioritisations using generalised linear mixed-effects models fit with logit links (GLMMs; using the \texttt{lme4 R} package; [@r469]).
-We fit a full model to the proportion of genetic variation secured in a given prioritisation.
-This model contained categorical variables indicating the combination of targets (amount-only, amount and surrogate targets, or amount and genetic targets), the scenario used to generate the prioritisation (single-species, multi-species, or multi-species with cost), the type of genetic variation measured (adaptive or neutral), and all interactions between them.
-Prioritisations generated for the same species were accomodated using a random intercept term.
-We subject the full model to a backwards step-wise term deletion routine to assess term significance, and conducted a post-hoc analysis on the minimal adequate model using Tukey contrasts with a Bonferroni correction (using the \texttt{multcomp R} package; [@r470]).
+We fitted a full model to the proportion of genetic variation secured in a given prioritisation.
+This model contained categorical variables indicating the combination of targets (ie.
+amount only, amount and surrogate targets, or amount and genetic targets), the scenario used to generate the prioritisation (ie.
+single-species, multi-species, or multi-species with cost), the type of genetic variation measured (ie.
+adaptive or neutral), and all interactions between them.
+Prioritisations generated for the same species were accommodated using a random intercept term.
+We subjected the full model to a backwards step-wise term deletion routine to assess term significance, and conducted a _post-hoc_ analysis on the minimal adequate model using Tukey contrasts with a Bonferroni correction (using the \texttt{multcomp R} package; [@r470]).
 

### 2017-03-14 code/rmarkdown/article.Rmd
incorperate article comments update pkg versions

@@
   - name: Cynthia Riginos
-    affiliation: a
-  - name: Hugh P. Possingham
     affiliation: a
@@
 abstract: |
-  Protected areas buffer species from anthropogenic threats and provide places for biodiversity processes to continue.
-However, evolutionary processes are not often explicitly considered when deciding where to site protected areas because genetic data require considerable resources to obtain and analyse.
-Here we show that environmental and geographic distance variables are effective surrogates for adaptive and neutral intra-specific variation for conservation planning.
+  Protected areas buffer species from anthropogenic threats and provide places for the processes that generate and maintain biodiversity to continue.
+However, genetic variation, the raw material for many evolutionary processes, is difficult to capture in conservation planning, not least because they require considerable resources to obtain and analyze.
+Here we show that freely available environmental and geographic distance variables are highly effective surrogates in conservation planning for representing adaptive and neutral intra-specific genetic variation.
 We obtained occurrence and genetic data from the IntraBioDiv project for 27 alpine plant species collected over the European Alps using a gridded sampling scheme.
 For each species, we identified loci that were potentially under selection using outlier loci methods, and mapped their main gradients of adaptive and neutral genetic variation across the grid cells.
-We then used the cells as planning units to generate prioritisations.
-Our first analysis verified that the spatial patterns of environmental and geographic variation correlated with those adaptive and neutral genetic variation (respectively).
-Our second analysis showed that these surrogates can predict the proportion of genetic variation secured in randomly generated solutions.
-Our third analysis found that solutions generated using additional surrogate-based targets secured more neutral genetic variation.
-We discovered that solutions generated using traditional amount-based targets tend to secure most species' adaptive genetic variation.
-Overall, our work provides a rigorous examination of two potential surrogates for genetic variation.
+We then used the cells as planning units to generate prioritize protected area acquisitions.
+First, we verified that the spatial patterns of environmental and geographic variation were correlated, respectively, with adaptive and neutral genetic variation.
+Second, we showed that these surrogates can predict the proportion of genetic variation secured in randomly generated solutions.
+Finally, we discovered that solutions based only on surrogate information secured substantial amounts of adaptive and neutral genetic variation.
+Our work paves the way for widespread integration of surrogates for genetic variation into conservation planning.
 significance: |
-  Protected areas are used to preserve biodiversity patterns and processes.
-Since the resources available for conservation are limited, protected areas need to be sited in cost-effective locations.
-However, despite their importance, evolutionary processes are not often used to guide reserve selection.
-This is primarily because genetic data require extensive resources to obtain and process.
-Here we show that freely available environmental and geographic variables can be used as surrogates for genetic data in conservation planning.
+  To protect biodiversity for the long term, nature reserves and other protected areas need to represent a broad range of different gentic types.
+However, genetic data are expensive and time consuming to obtain.
+Here we show that freely available environmental and geographic variables can be used as effective surrogates for genetic data in conservation planning.
+This means that conservation plannners can, with some confidence, design protected area systems to represent genetic diversity without investing in expensive programmes to obtain and analyze genetic data.
 acknowledgements: |
@@
 # load .rda file
-checkpoint::checkpoint('2016-08-03', R.version='3.3.1', scanForPackages=FALSE)
+checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
 try(session::restore.session('data/final/results.rda'))
-
-# set options
-options(error=NULL)
 
@@
 \dropcap{P}rotected areas spearhead conservation efforts ([@r440]).
-They buffer species from anthropogenic impacts; providing places for them to persist.
+They buffer species from anthropogenic impacts, providing places for them to persist.
 Since the resources available for conservation are limited, protected areas need to be sited in places that fulfill conservation objectives for minimal cost ([@r255]).
-To achieve this, conservation planning exercises often generate plans for entire networks of protected areas (prioritisations) to allocate resources cost-effectively and preserve broad-scale biodiversity processes.
-Intra-specific genetic variation is fundamental to long-term species persistence ([@r165]), and as a consequence, there has been increasing interest in designing prioritisations that capture and cultivate this aspect of biodiversity ([@r67; @r3; @r61; @r499]).
+To achieve this, conservation planning exercises often generate plans for entire networks of protected areas (prioritizations) to allocate resources cost-effectively and preserve broad-scale biodiversity processes.
+Intra-specific genetic variation is fundamental to long-term species persistence ([@r165]), and as a consequence, there has been increasing interest in designing prioritizations that represent this aspect of biodiversity ([@r67; @r3; @r61; @r499]).
 
-Broadly speaking, though the strength of natural selection is a continuous force, genetic variation can be classified as either adaptive or neutral (reviewed in [@r480]).
+Broadly speaking, though the strength of natural selection is a continuous force, genetic variation is commonly classified as either adaptive or neutral (reviewed in [@r480]).
 Adaptive genetic variation is associated with loci that significantly affect fitness.
@@
 By representing the full range of adaptive variation, protected areas can help ensure that populations with particularly beneficial adaptations are not lost ([@r3; @r43]).
-In contrast, neutral genetic variation is associated with loci that do not significantly affect fitness, but instead reflects the evolutionary history of different populations.
+In contrast, neutral genetic variation is associated with loci that do not significantly affect fitness, but instead reflect the evolutionary history of different populations.
 By representing the full range of neutral variation, protected areas can safeguard against the adverse effects of low genetic diversity ([@r67]).
@@
 
-Recently, the use of surrogate data has been proposed to generate prioritisations that sufficiently capture intra-specific genetic variation--without needing to directly utilise genetic data to achieve this.
+Recently, the use of surrogate data has been proposed to generate prioritizations that sufficiently capture intra-specific genetic variation ([@r2; @r200])--without needing to directly utilize genetic data to achieve this.
 Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been proposed to guide reserve selection as surrogates for adaptive genetic variation (eg.
 [@r202; @r2]).
-By generating prioritisations that preserve a more representative sample of environmental conditions across species' ranges, these prioritisations could also capture more adaptive variation.
+This is on the basis that sampling a broad range of environmental conditions across species' distributions should also capture more adaptive variation.
 However, the effectiveness of environmental surrogates in representing adaptive genetic variation remains unverified.
-Neutral genetic variation arises from a reduction in gene flow between individuals.
+Neutral genetic variation arizes from a reduction in gene flow between individuals.
 Over the last few years, the field of phylogeography has predominantly focused on describing neutral genetic diversity, and so the landscape factors that affect neutral variation are relatively well understood (reviewed in [@r501]).
 Potential surrogates for neutral variation have been based on variables that predict the levels of connectivity between planning units.
-For instance, isolation by distance (IBD; [@r259]) predicts that populations located further apart will experience less gene flow, and in turn, share less neutral loci.
-Based on this idea, geographic surrogates based on distance have been found to improve conservation prioritisations in an island network ([@r200]).
-However this remains untested in spatially contiguous systems where connectivity is complicated by additional factors (eg.
-habitat; [@r497])--the typical situations in most conservation planning exercises.
+For instance, isolation by distance (IBD; [@r259]) predicts that populations located further apart will experience less gene flow, and in turn, share fewer neutral loci.
+Based on this idea, geographic surrogates based on distance have been found to improve conservation prioritizations in an island network ([@r200]).
+However this remains untested in spatially contiguous systems where connectivity is complicated by many additional factors (eg.
+habitat; [@r497])--the typical situation in most conservation planning exercises.
 
-Here we investigate the effectiveness of environmental and geographic surrogates for capturing adaptive and neutral genetic variation respectively in the context of conservation planning.
-We used distribution and genomic data for 27 alpine plant species in the European Alps that was obtained by the IntraBioDiv project ([@r451]).
+Here we determine whether environmental and geographic surrogates capture adaptive and neutral genetic variation respectively in the context of conservation planning.
+We use distribution and genomic data for 27 alpine plant species in the European Alps that was obtained by the IntraBioDiv project ([@r451]).
 The data was collected following a gridded sampling scheme which we adopted as planning units for this study.
@@
 We also show that for most species the level of association is strong enough to be operational for conservation planning.
-Finally, we demonstrate that using these surrogates results in prioritisations that secure more genetic variation than expected by chance.
+Finally, we demonstrate that using these surrogates results in prioritizations that secure a substantial proportion of genetic variation.
 
@@
 The spatial distribution of these genetic spaces (Figures S29--S54) generally showed substantial spatial auto-correlation (eg.
-_Cerastium uniflorum_, Figure S34; _Dryas octopetala_, Figure S36), suggesting that the environmental and geographic variables may prove to be effective surrogates.
+_Cerastium uniflorum_, Figure S34; _Dryas octopetala_, Figure S36), suggesting that the environmental and geographic variables have potential to be effective surrogates for genetic variation.
 
@@
 We further verified that there was a correlation between neutral genetic variation and variation in geographic position.
-For each species, we constructed dissimilarity matrices expressing differences between the planning units within which they were detected according to terms their (i) environmental characteristics, (ii) geographic position, and also the (iii) adaptive (if adaptive loci detected), and (iv) neutral genetic characteristics of individuals found inside them.
-By correlating these matrices, the spatial patterns of environmental variation were found to be significantly correlated with that of adaptive genetic variation for `r as.character(as.english(sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)))` of the `r as.character(as.english(sum(!is.na(correlation.model.summary.DF$Environmental.corr.P))))` species associated with adaptive variation (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models; Table S3).
+For each species, we constructed dissimilarity matrices expressing differences between the planning units where individuals were detected based on the units' (i) environmental characteristics, (ii) geographic position, and also the (iii) adaptive (if adaptive loci detected), and (iv) neutral genetic characteristics of individuals found inside them.
+The spatial patterns of environmental variation were significantly correlated with the pattern of adaptive genetic variation for `r as.character(as.english(sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)))` of the `r as.character(as.english(sum(!is.na(correlation.model.summary.DF$Environmental.corr.P))))` species associated with adaptive variation (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models; Table S3).
 Similarly, geographic distance between planning units was also correlated significantly with the spatial pattern of neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` of the `r sum(!is.na(correlation.model.summary.DF$Geographic.corr.P))` species (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models; Table S3).
 Thus, for most species, planning units that contained different environmental conditions contained individuals with different adaptive genetic characteristics, and planning units that were further apart tended to contain individuals with different neutral genetic characteristics.
-After verifying these patterns were correlated, our next step was to determine whether they were correlated enough to be effective for conservation planning.
+After verifying these patterns were correlated, our next step was to determine whether these associations were strong enough to be effective for conservation planning.
 
 <!-- surrogacy analysis -->
-For each species, we generated a suite of `r format(raptr.params.LST[[MODE]]$surrogacy.analysis$random.replicates, big.mark=',', scientific=FALSE)` random prioritisations, and calculated the proportion of environmental variation and adaptive (if present) genetic variation each prioritisation captured.
+For each species, we generated a suite of `r format(raptr.params.LST[[MODE]]$surrogacy.analysis$random.replicates, big.mark=',', scientific=FALSE)` random prioritizations, and calculated the proportion of environmental variation and adaptive (if present) genetic variation each prioritization captured.
 We then repeated this process, and calculated the proportion of geographic variation and neutral genetic variation they sampled.
-The environmental and geographic variables were moderately effective predictors for the genetic variation represented by the randomly generated prioritisations (Figure 1).
-The relationship between the proportion of genetic and surrogate variation secured in the prioritisations varied among the species (species $\times$ proportion of environmental variation interaction term: ${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$; species $\times$ proportion of geographic variation interaction term: ${\chi^2}_{`r round(geo.surrogacy.AOV$df[[4]],2)`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`).
+The environmental and geographic variables were moderately effective predictors for the genetic variation represented by the randomly generated prioritizations (Figure 1).
+The relationship between the proportion of genetic and surrogate variation secured in the prioritizations varied among the species (species $\times$ proportion of environmental variation interaction term: ${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$; species $\times$ proportion of geographic variation interaction term: ${\chi^2}_{`r round(geo.surrogacy.AOV$df[[4]],2)`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`).
 _Post-hoc_ analyses showed that these relationships were positive for all species (environmental: minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; geographic: minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S4).
-After establishing that the the proportion of surrogate variation secured in a prioritisation also predicted the proportion of genetic variation it secured, our final step was to determine if prioritisations were more effective when generated using targets based on the surrogates.
+After establishing that the the proportion of surrogate variation secured in a prioritization also predicted the proportion of genetic variation it secured, our final step was to determine if prioritizations were more effective when generated using targets based on the surrogates.
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=6, fig.height=4.5, out.width='3.42in', out.height='2.5in', fig.cap='Adaptive (a) and neutral (b) genetic variation represented in a suite of randomly generated prioritisations as a function of the environmental and geographic surrogates they also secured (respectively).
-Panels (c) and (d) show the distribution of $R^2$ values for models using each surrogate to explain genetic variation.'}
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=6, fig.height=4.5, out.width='3.42in', out.height='2.5in', fig.cap=paste0('The proportion of adaptive (a; ',sum(!is.na(correlation.model.summary.DF$Environmental.corr.P)),' species) and neutral (b; ',sum(!is.na(correlation.model.summary.DF$Geographic.corr.P)),' species) genetic variation represented in a suite of randomly generated prioritizations for different species (lines) as a function of the surrogate variation they also secured.
+Panels (c) and (d) show the distribution of $R^2$ values for the ',sum(!is.na(correlation.model.summary.DF$Environmental.corr.P)) + sum(!is.na(correlation.model.summary.DF$Geographic.corr.P)),' models that explain genetic variation using e to explain genetic variation.')}
 # make figure
@@
     data=filter(env.surrogate.spp.DF, !is.na(r2CU))) +
-  geom_histogram(fill='grey10', bins=5) +
+  geom_histogram(fill='grey10', bins=12) +
   ylab('\nFrequency') + xlab(expression(Environmental~variables~(R^2))) +
@@
     data=filter(geo.surrogate.spp.DF, !is.na(r2CU))) +
-  geom_histogram(fill='grey10', bins=10) +
+  geom_histogram(fill='grey10', bins=12) +
   ylab('\nFrequency') + xlab(expression(Geographic~variables~(R^2))) +
@@
 <!-- scenario analysis -->
-To determine whether using environmental and geographic targets could improve the effectiveness of prioritisations in representing genetic variation, we generated prioritisations using (i) "amount targets" reflecting the traditional approach of representing a certain proportion of each species' geographic distribution, (ii) "amount and surrogate targets" where we also targeted representation of the environmental and geographic surrogates as well as a certain proportion of each species' geographic distribution, and (iii) "amount and genetic targets" where we targeted representation of the directly measured genetic variation as well as a certain proportion of each species' geographic distribution (Figure 3).
-To determine if these patterns were robust, for each of the three combinations of the targets, we generated prioritisations under four scenarios: (i) single-species with equal costs, (ii) single species with opportunity costs, (iii) multi-species with equal costs, and (iv) multi-species with opportunity costs.
+To determine whether using environmental and geographic targets could improve the effectiveness of prioritizations in representing genetic variation, we generated prioritizations using (i) "amount targets" reflecting the traditional approach of representing a certain proportion of each species' geographic distribution, (ii) "amount and surrogate targets" where we targeted representation of the environmental and geographic surrogates as well as a certain proportion of each species' geographic distribution, and (iii) "amount and genetic targets" where we targeted representation of the directly measured genetic variation as well as a certain proportion of each species' geographic distribution (Figure 3).
+To determine if these patterns were robust, for each of the three combinations of the targets, we generated prioritizations under four scenarios: (i) single-species with equal costs, (ii) single species with opportunity costs, (iii) multi-species with equal costs, and (iv) multi-species with opportunity costs.
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=(5*0.9), fig.height=(7.1*0.9), out.width='3.42in', out.height='4in', fig.cap='Box plots show the proportion of adaptive and neutral genetic variation secured for each species in solution generated from different targets and scenarios.
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=(5*0.9), fig.height=(7.1*0.9), out.width='3.42in', out.height='4in', fig.cap='The proportion of adaptive and neutral genetic variation secured for each species in solutions based on different targets and scenarios.
+Box plots show the 25th percentile, median, and 75th percentile.
+Whiskers show the data at 1.5 $\times$ the inter-quantile range.
+Points show data outside the whiskers.
 Red lines show the target proportion of genetic variation.'}
@@
 scenario.SDF <- scenario.DF %>%
-  group_by(Prioritisation, Context, Metric, Species) %>% 
+  group_by(prioritization, Context, Metric, Species) %>%
   summarize(genetic.held=mean(genetic.held, na.rm=TRUE)) %>%
@@
 suppressWarnings({
-  ggplot(aes(x=Prioritisation,y=genetic.held),
+  ggplot(aes(x=prioritization,y=genetic.held),
     data=scenario.SDF) +
     geom_boxplot(position=position_dodge(0.9)) +
-    scale_fill_manual(name='Prioritisation',
+    scale_fill_manual(name='prioritization',
       values=c('grey80','grey50','grey20')) +
@@
 
-The proportion of genetic variation secured in a prioritisation depended on the targets used to generate the prioritisation (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$p.value[[2]])`)$.
-Most notably, prioritisations based on the surrogate variables represented significantly more genetic variation than amount-based prioritisations (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` SD overall genetic variation secured versus `r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)` SD; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$), and were not distinguishable from prioritisations based on measured genetic data (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` SD; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).
+The proportion of genetic variation secured in a prioritization depended on the targets used to generate the prioritization (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$p.value[[2]])`)$.
+Most notably, prioritizations based on the surrogate variables represented significantly more genetic variation than amount-based prioritizations (`r round(mean(filter(scenario.sub.DF,prioritization=='Amount & surrogate\ntargets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,prioritization=='Amount & surrogate\ntargets')$genetic.held*100),2)` SD overall genetic variation secured versus `r round(mean(filter(scenario.sub.DF,prioritization=='Amount targets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,prioritization=='Amount targets')$genetic.held*100),2)` SD; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$), and were not distinguishable from prioritizations based on measured genetic data (`r round(mean(filter(scenario.sub.DF,prioritization=='Amount & genetic\ntargets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,prioritization=='Amount & genetic\ntargets')$genetic.held*100),2)` SD; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).
 These results indicate that environmental and geographic surrogates in this case far outperform traditional amount-based conservation planning, and perform as well as a conservation plan based directly on measured genetic data.
 
-Overall, the prioritisations tended to secure more adaptive then neutral genetic variation (adaptive: mean `r round(mean(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` secured across all prioritisations, neutral: mean `r round(mean(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)` $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)`; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.3.DF$p.value[[2]])`).
-Thus, regardless of the targets used to generate a prioritisation, or the planning context under which the prioritisation was generated, prioritisations tend to secure more adaptive than neutral genetic genetic variation.
+Overall, the prioritizations tended to secure a greater proportion of adaptive than neutral genetic variation (adaptive: mean `r round(mean(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` secured across all prioritizations, neutral: mean `r round(mean(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)` $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)`; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.3.DF$p.value[[2]])`).
+Thus, regardless of the targets used to generate a prioritization, or the planning context under which the prioritization was generated, prioritizations tend to secure more adaptive than neutral genetic genetic variation.
 
-The average proportion of genetic variation secured by prioritisations varied under different scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.2.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.2.DF$p.value[[2]])`).
+The average proportion of genetic variation secured by prioritizations varied under different scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.2.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.2.DF$p.value[[2]])`).
 Both the multi-species (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)`) and multi-species with cost (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)`) were found to secure more genetic variation than the single species scenario (mean `r round(mean(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[2]])`$; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$ respectively).
-These results suggest that on average the multi-species prioritisations can secure more genetic variation than single-species prioritisations--regardless of targets used to generate them.
+These results suggest that on average the multi-species prioritizations can secure more genetic variation than single-species prioritizations--regardless of targets used to generate them.
 
@@
 <!--  introduction -->
-We aimed to investigate the effectiveness of broad-scale environmental and geographic variables as surrogates for adaptive and neutral genetic variation in conservation planning.
-Our study is the first to address this aim using field measurements of genetic variation for a broad range of species to assess the performance of prioritisations.
+We have shown that broad-scale environmental and geographic variables can be effective as surrogates for adaptive and neutral genetic variation in conservation planning.
+Our study is the first to show this using field measurements of genetic variation for a broad range of species.
 The spatial patterns of environmental and geographic surrogates were strongly correlated with the patterns of adaptive and neutral genetic variation for most species.
-Moreover, for most species, prioritisations generated using environmental and geographic targets secured a greater proportion of adaptive and neutral genetic variation than traditional conservation planning.
+Moreover, for most species, prioritizations generated using environmental and geographic targets secured a greater proportion of adaptive and neutral genetic variation than traditional conservation planning.
 These environmental and geographic surrogates were based on freely available data, and could be applied to any study region across the world.
@@
 <!-- surrogate/genetic held correlations -->
-Our results demonstrate that environmental and geographic surrogates are strong enough to be operational for conservation planning (Figure 2).
-For most species, the environmental and geographic surrogates variables were effective surrogates for adaptive (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figures 2c and S56) and neutral genetic variation (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figures 2d and Figure S57).
-Despite this, for a few species, geographic distance was a poor surrogate for neutral genetic variation (eg.
+Our results demonstrate that environmental and geographic variables are effective surrogates for most species (Figure 2).
+Despite this, for a few species, geographic distance was a suprisingly poor surrogate for neutral genetic variation (eg.
 _Gypsophila repens_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gypsophila_repens')$Geographic.r2CU,2)`; _Luzula alpinopilosa_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Luzula_alpinopilosa')$Geographic.r2CU,2)`; _Gentiana nivalis_ $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gentiana_nivalis')$Geographic.r2CU,2)`).
-Indeed, geographic distances are expected to be a poor surrogate for connectivity when species deviate from isolation by distance (eg.
-wind dispersing plants can maintain high connectivity between distant areas; [@r451]), and in areas where connectivity is most affected by landscape features ([@r497]).
+Indeed, geographic distance is expected to be a poor surrogate for connectivity when species deviate from isolation by distance.
+For example, species can utilize disersal modes that maintain high connectivity between distant areas (eg.
+wind dispersing plants; [@r451]), and connectivity between adjacent populations can be disrupted by landscape features ([@r497]).
 Overall, such species were the exception in our analysis, and both surrogates performed moderately well for most species.
@@
 <!-- suitable targets for surrogates -->
-Our results suggest that environmental and geographic surrogates can be used to capture genetic variation in prioritisations, the next question is: what percentage of surrogate variation should planners target to preserve genetic variation?
+Our results suggest that environmental and geographic surrogates can be used to capture genetic variation in prioritizations, the next question is: what percentage of surrogate variation should planners target to preserve genetic variation?
 Firstly, our results suggest that surrogate-based targets should be used with amount-based targets.
-For all species, it was possible to generate solutions that secured a high proportion of the surrogate variable (> 90 %) and only a small proportion of genetic variation (< 20 %; Figures S56--S57).
-By using both amount- and surrogate-based targets, planners may avoid such solutions.
-Secondly, geographic surrogates may need higher targets than environmental surrogates to capture evolutionary processes.
-In our analysis, geographic variation accumulated at a slower rate for neutral genetic representation than environmental variation did with adaptive representation (Figures 2a--2b and S56--S57).
-To secure a representative sample of each species' neutral genetic variation ($\geq$ `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %), prioritisations needed to sample at least `r round(mean(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` SD of the variation in geographic position among the planning units occupied by the species.
-On the other hand, to secure at least `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` % of each species' adaptive genetic variation, prioritisations only needed to sample `r round(mean(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` SD of the environmental variation in planning units occupied by the species.
-Although to capture evolutionary processes, real-world conservation planning exercises may need higher surrogate targets.
+For all species, it was possible to generate solutions that secured a high proportion of the surrogate variables (> 90 %) and only a small proportion of genetic variation (< 20 %; Figures S56--S57).
+By using both amount- and surrogate-based targets, planners may avoid such outcomes.
+Secondly, geographic surrogates seem to need higher targets than environmental surrogates to capture genetic variation.
+In our analysis, neutral genetic variation accumulated much more slowly with geographic representation than adaptive variation accumulated with environmental representation (Figures 2a--2b and S56--S57).
+To secure at least `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` % of the species' neutral genetic variation, prioritizations needed to sample `r round(mean(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` SD of the variation in geographic position among the planning units occupied by the species.
+On the other hand, to secure at least `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` % of each species' adaptive genetic variation, prioritizations only needed to sample `r round(mean(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` SD of the environmental variation in planning units occupied by the species.
+In many individual cases, higher targets may be needed to ensure that prioritizations secure a representative sample of the genetic variation for all of the species in a planning exercise.
 
-<!-- adaptive genetic variation secured in prioritisations-->
+<!-- adaptive genetic variation secured in prioritizations-->
 Our results suggest that conservation planners can secure a reasonably representative sample of the intra-specific adaptive genetic variation using only amount-based targets.
@@
 _Carex semipervens_ and _Cerastium uniflorum_; Figures S33 and S34 respectively).
-As a consequence, a prioritisation would only need to select a single planning unit in each cluster to secure a representative sample.
-This mechanism explains why adaptive variation accumulated much quicker than neutral genetic variation in the randomly generated prioritisations.
-Additionally, this mechanism also explains why the single-species prioritisations generated using only amount targets were often able to achieve high representation--they were essentially a random selection and so frequently contained planning units in different parts of the species' ranges (Figure 4a).
-Overall--and contrary to our original hypotheses--it seems that conventional reserve selection methods may secure a representative sample of species' adaptive genetic variation.
+As a consequence, a prioritization would only need to select a single planning unit in each cluster to secure a representative sample.
+This mechanism explains why adaptive variation accumulated much more quickly than neutral genetic variation in the randomly generated prioritizations.
+Additionally, this mechanism also explains why the single-species prioritizations generated using only amount targets were often able to achieve high representation--they were essentially a random selection and so frequently contained planning units in different parts of the species' ranges (Figure 4a).
+Overall--and contrary to our original hypotheses--it seems that conventional reserve selection methods might often secure a representative sample of species' adaptive genetic variation.
 
@@
 Secondly, we used geographic distances as surrogates for neutral genetic variation, but connectivity in this landscape may be better described using distances that incorporate information on topography.
-However, such distances require species-specific scaling to adequate describe their connectivity (eg.
-[@r497]), and so cannot easily be utilised in large-scale planning exercises.
-Thirdly, we used amplified fragment length polymorphism data (AFLP; [@r453]) to characterise the main gradients of genetic variation.
-Whilst next-generation sequencing provides higher resolution genetic information than AFLP data ([@r480]), no multi-species genomic data set with similar properties currently exists, and our methodology would still have utilised only the main gradients of the genetic variation to capture broad-scale evolutionary processes.
+However, such distances require species-specific scaling to adequately describe their connectivity (eg.
+[@r497]), and so cannot easily be utilized in large-scale multi-species planning exercises.
+Thirdly, we used amplified fragment length polymorphism data (AFLP; [@r453]) to characterize the main gradients of genetic variation.
+Whilst next-generation sequencing provides higher resolution genetic information than AFLP data ([@r480]), we know of no suitable multi-species genomic data set, and our methodology would still have utilized only the main gradients of the genetic variation to generate prioritizations in a computationally feasible period of time.
 
@@
 We used data for 27 alpine plant species in the European Alps collected by the IntraBioDiv project ([@r451]; Figure 2a).
-All data, code, and results are stored in an online repository to replicate and validate this study (\url{www.github.com/jeffreyhanson/genetic.surrogates}).
+All data, code, and results are stored in an online repository to permit replication and validation of this study (\url{www.github.com/jeffreyhanson/genetic.surrogates}).
 This data set has been used to explore patterns of adaptive (eg.
@@
 Thus the data set contains information describing the genomic properties of individuals for each species over a geographic grid.
-We used this data set because it was comprised of comparable genetic data for a range of species with different evolutionary and life histories that was collected using a standardised sampling scheme.
+We used this data set because it comprized comparable genetic data for a range of species with different evolutionary and life histories that was collected using a standardized sampling scheme.
 Additionally, the study area contained large environmental variation which is useful for examining environmental surrogates.
 
-```{r, echo=FALSE, fig.align='center', fig.width=6, fig.height=1.5, out.width='3.42in', height='2.5in', fig.cap="Map of the study area.
-Squares denote planning units.
-Panel (a) depicts patterns of species riches, and panel (b) shows the distribution of acquisition cost (plotted using a quantile-based color ramp)."}
+```{r, echo=FALSE, fig.align='center', fig.width=6, fig.height=1.5, out.width='3.42in', height='2.5in', fig.cap="Map of the study area showing (a) the pattern of species richness, and (b) the distribution of acquisition cost, both plotted on a quantile-based color ramp.
+Squares denote planning units."}
 
@@
 
-We adopted the sampling grid used to collect data as planning units to develop prioritisations.
+We adopted the sampling grid used to collect data as planning units to develop prioritizations.
 To reduce computational burden, we omitted cells that did not contain samples (`r nrow(grid.DF)` grid cells used in analysis).
@@
 To reduce dimensionality, for each species, we subjected the climatic values associated with their planning units to a principal components analysis (PCA; Table S1).
-We used the first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components to characterise climatic variation found across the species' geographic distributions inside the study area (capturing `r round(mean(spp.pca.DF[[2]]*100),2)` % $\pm$ `r round(sd(spp.pca.DF[[2]]*100),2)` SD of the total climatic variation; Figure S2).
+We used the first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components to characterize climatic variation found across the species' geographic distributions inside the study area (capturing `r round(mean(spp.pca.DF[[2]]*100),2)` % $\pm$ `r round(sd(spp.pca.DF[[2]]*100),2)` SD of the total climatic variation; Figure S2).
 Thus we constructed a two-dimensional geographic space as a potential surrogate for neutral genetic variation, and a `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))`-dimensional environmental space as a potential surrogate for adaptive genetic variation for each species.
 In these spaces, each planning unit was associated with a single point.
-Planning units with more comparable environmental conditions, or higher spatial proximity in the real world, were associated with points that were closer together in each of these spaces.
+Planning units with more comparable environmental conditions, or higher spatial proximity in the real world, were associated with points that were closer together in these environmental or geographic spaces.
 
@@
 To investigate the effectiveness of our supposed surrogates, we first needed to identify which of the sampled loci were adaptive.
-Following best practices, we used two outlier detection methods to achieve this (an individual- and a population-level method; [@r495]).
+Following best practice, we used two outlier detection methods to achieve this (an individual- and a population-level method; [@r495]).
 The basic premise underpinning such methods is that neutral loci are expected to exhibit a certain level of variation, and loci that deviate from this expectation are likely to be under selection ([@r480]).
-The advantage of these methods--in contrast to environmental association analyses--is that they do not use environmental data, which would have introduced circularity into our analysis and invalidated it.
-Loci identified as outliers in both analyses were treated as adaptive and the remainder were treated as neutral.
-To minimise false-positives, we omitted loci from both methods where the global frequency of the minor allele was less than 10 % and treated them as neutral.
+The advantage of these methods--in contrast with environmental association analyses--is that they do not use environmental data, which would have introduced an element of circularity into our analysis and invalidated it.
+Loci identified by both outlier detection methods were treated as adaptive and the remainder were treated as neutral.
+To minimize false-positives, we omitted loci from both methods where the global frequency of the minor allele was less than 10 % and treated them as neutral.
 
@@
 We adopted a similar methodology to Bothwell _et al._ ([-@r455]) and applied it to each of the species separately.
-Following their methodology, we initially grouped conspecifics into genetic lineages to further minimise false-positives.
-We grouped conspecifics by fitting admixture and correlated alleles models implemented in \texttt{Structure} (version 2.3.4; [@r462]; `r structure.params.LST[[MODE]]$numruns` replicates per species; `r format(structure.params.LST[[MODE]]$admburnin, big.mark=',', scientific=FALSE)` admixture burnin iterations; `r format(structure.params.LST[[MODE]]$burnin, big.mark=',', scientific=FALSE)` burnin iterations; `r format(structure.params.LST[[MODE]]$numreps+structure.params.LST[[MODE]]$burnin, big.mark=',', scientific=FALSE)` total iterations) using the number of lineages previously determined by Alvarez _et al._ ([@r478]) and combining replicate runs using \texttt{ClumPP} (version 1.1.2; [@r459]; greedy algorithm based on the $G^{\prime}$ statistic; `r format(clumpp.params.LST[[MODE]]$repeats, big.mark=',', scientific=FALSE)` iterations).
+Following their methodology, we initially grouped conspecifics into genetic lineages to further minimize false-positives, by fitting admixture and correlated alleles models implemented in \texttt{Structure} (version 2.3.4; [@r462]; `r structure.params.LST[[MODE]]$numruns` replicates per species; `r format(structure.params.LST[[MODE]]$admburnin, big.mark=',', scientific=FALSE)` admixture burnin iterations; `r format(structure.params.LST[[MODE]]$burnin, big.mark=',', scientific=FALSE)` burnin iterations; `r format(structure.params.LST[[MODE]]$numreps+structure.params.LST[[MODE]]$burnin, big.mark=',', scientific=FALSE)` total iterations) using the number of lineages previously determined by Alvarez _et al._ ([@r478]) and combining replicate runs using \texttt{ClumPP} (version 1.1.2; [@r459]; greedy algorithm based on the $G^{\prime}$ statistic; `r format(clumpp.params.LST[[MODE]]$repeats, big.mark=',', scientific=FALSE)` iterations).
 We then ran $\texttt{Bayescan}$ for each species using these lineages (1:1 prior odds; `r bayescan.params.LST[[MODE]]$reps` replicates per species; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r format(bayescan.params.LST[[MODE]]$burn, big.mark=',', scientific=FALSE)` burn-in iterations; `r format(bayescan.params.LST[[MODE]]$n+bayescan.params.LST[[MODE]]$burn, big.mark=',', scientific=FALSE)` total iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations) using a suitable false discovery rate ($q \leq `r bayescan.params.LST[[MODE]]$fdr`$; [@r466]).
@@
 Thus we constructed an adaptive (if detected) and neutral genetic space for each species.
-For a given species, each planning unit that it was found in was associated with a multi-dimensional point in the species' adaptive genetic space (if adaptive genetic variation was detected) and another multi-dimensional point in the species' neutral genetic space.
+For a given species, each planning unit in which it occurred was associated with a multi-dimensional point in the species' adaptive genetic space (if adaptive genetic variation was detected) and another multi-dimensional point in the species' neutral genetic space.
 Planning units that were closer together in these genetic spaces were occupied by individuals with more comparable AFLP polymorphisms.
-By spreading out conservation effort across these genetic spaces, and in turn selecting planning units occupied by individuals with increasingly different polymorphisms, prioritisations can secure more genetic variation.
+By spreading out conservation effort across these genetic spaces, and in turn selecting planning units occupied by individuals with increasingly different polymorphisms, prioritizations can secure more genetic variation.
 
-Prioritisation method {#prioritisation-method .unnumbered}
+prioritization method {#prioritization-method .unnumbered}
 ----------------------------------------------------------
 
-We used the unreliable representative and adequate prioritisation (URAP) formulation of reserve selection ([@r465]).
-This formulation can identify the minimum number of planning units required to preserve both a proportion of the species' geographic range (using amount-based targets) and a sample of variation found across the species' range.
+We used the unreliable representative and adequate prioritization (URAP) formulation for reserve selection ([@r465]).
+This formulation can identify the minimum number of planning units required to preserve both a target proportion of the species' geographic range (using amount-based targets) and a broad sample of variation found across the species' range.
 We used the environmental and geographic surrogate spaces, and the adaptive (if detected) and neutral genetic spaces to generate and evaluate solutions.
@@
 Our second aim was to determine if the environmental and geographic variables were effective surrogates for adaptive and neutral genetic variation.
-For each species, we generated `r format(raptr.params.LST[[MODE]]$surrogacy.analysis$random.replicates*2, big.mark=',', scientific=FALSE)` prioritisations by randomly selecting different combinations of planning units that the species occupied.
-For half of these random prioritisations, we calculated the proportion of geographic variation and neutral genetic variation they secured, and for the other half, the proportion of environmental variation and adaptive genetic variation they secured.
+For each species, we generated `r format(raptr.params.LST[[MODE]]$surrogacy.analysis$random.replicates*2, big.mark=',', scientific=FALSE)` prioritizations by randomly selecting different combinations of planning units that the species occupied.
+For half of these random prioritizations, we calculated the proportion of geographic variation and neutral genetic variation they secured, and for the other half, the proportion of environmental variation and adaptive genetic variation they secured.
+These calculations were performed using built-in routines in the \texttt{raptr R} package.
 
-We fitted two full generalised linear models (GLM) with logit links.
-The first model was fit to the proportion of adaptive genetic variation secured in a prioritisation using the proportion of environmental variation also secured in the prioritisation.
-The second model was fit to the proportion of neutral genetic variation secured in a prioritisation using the proportion of geographic variation also secured in the prioritisation.
-Additionally, both models contained a variable indicating the species for which the prioritisations were generated, and an interaction term.
-They were subjected to backwards step-wise term deletion routines to assess term significance.
+We fitted two full generalized linear models (GLM) with logit links.
+The first model was fit to the proportion of adaptive genetic variation secured in a prioritization using the proportion of environmental variation also secured in the prioritization.
+The second model was fit to the proportion of neutral genetic variation secured in a prioritization using the proportion of geographic variation also secured in the prioritization.
+Additionally, both models contained a variable indicating the species for which the prioritizations were generated, and an interaction term.
+They were subjected to backwards stepwise term deletion routines to assess term significance.
 To assess the performance of the surrogates, we refit these models separately for each species and computed the Cragg and Uhler's pseudo-$R^2$ value (using the \texttt{pscl R} package; [@r492]).
 
-Our third aim was to determine if surrogate-based targets improved prioritisations.
-As previously mentioned, we generated prioritisations using different combinations of targets and conservation planning scenarios.
-We used `r (raptr.params.LST[[MODE]]$scenario.analysis$amount.target*100)` % amount-based targets in all prioritisations.
+Our third aim was to determine if surrogate-based targets improved representation of genetic variation by the prioritizations.
+As previously mentioned, we generated prioritizations using different combinations of targets and conservation planning scenarios.
+We used `r (raptr.params.LST[[MODE]]$scenario.analysis$amount.target*100)` % amount-based targets for each species in all prioritizations to secure an adequate proportion of the species' ranges.
 Based on the results from our previous aim, we used `r (raptr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100)` % surrogate-based targets and `r (raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)` % genetic-based targets.
 We generated a single solution for each target/scenario combination, except for the "single-species (equal costs) with amount-based targets" combination for which we generated `r format(raptr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates, big.mark=',', scientific=FALSE)` replicates because it had many optimal solutions.
-We computed the proportion of adaptive (if present) and neutral genetic variation secured for each species in the prioritisations.
+We computed the proportion of adaptive (if present) and neutral genetic variation secured for each species in the prioritizations.
 
-We investigated the benefits of using surrogate-based targets in prioritisations using generalised linear mixed-effects models fit with logit links (GLMMs; using the \texttt{lme4 R} package; [@r469]).
-We fitted a full model to the proportion of genetic variation secured in a given prioritisation.
+We investigated the benefits of using surrogate-based targets in prioritizations using generalized linear mixed-effects models fit with logit links (GLMMs; using the \texttt{lme4 R} package; [@r469]).
+We fitted a full model to the proportion of genetic variation secured in a given prioritization.
 This model contained categorical variables indicating the combination of targets (ie.
-amount only, amount and surrogate targets, or amount and genetic targets), the scenario used to generate the prioritisation (ie.
+amount only, amount and surrogate targets, or amount and genetic targets), the scenario used to generate the prioritization (ie.
 single-species, multi-species, or multi-species with cost), the type of genetic variation measured (ie.
 adaptive or neutral), and all interactions between them.
-Prioritisations generated for the same species were accommodated using a random intercept term.
-We subjected the full model to a backwards step-wise term deletion routine to assess term significance, and conducted a _post-hoc_ analysis on the minimal adequate model using Tukey contrasts with a Bonferroni correction (using the \texttt{multcomp R} package; [@r470]).
+prioritizations generated for the same species were accommodated using a random intercept term.
+We subjected the full model to a backwards stepwise term deletion routine to assess term significance, and conducted a _post-hoc_ analysis on the minimal adequate model using Tukey contrasts with a Bonferroni correction (using the \texttt{multcomp R} package; [@r470]).
 

### 2017-03-14 code/rmarkdown/supporting_information.Rmd
incorperate article comments update pkg versions

@@
 # load .rda file
-checkpoint::checkpoint('2016-08-03', R.version='3.3.1', scanForPackages=FALSE)
+checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
 try(session::restore.session('data/final/results.rda'))
 
-# set options
-options(error=NULL)
-
-# download basemap
+# download basemappv
 data(countriesHigh)

### 2017-03-14 code/R/analysis/00-initialization.R

@@ -11,7 +11,8 @@ if (length(args)>0) {
 # set checkpoint
 if (!'checkpoint' %in% installed.packages()[,'Package']) install.packages('checkpoint')
 if (!file.exists('~/.checkpoint')) dir.create('~/.checkpoint')
-checkpoint::checkpoint('2016-08-03', R.version='3.3.1')
+checkpoint::setSnapshot("2016-11-26")
+checkpoint::checkpoint('2016-11-26', R.version='3.3.2')
 if (!'checkpoint' %in% installed.packages()[,'Package']) install.packages('checkpoint')
 
 ## load bioconductor packages
@@ -125,4 +126,3 @@ for (x in dir(file.path('code', 'R', 'functions'), full.names=TRUE)) source(x)
 
 # save workspace
 save.session('data/intermediate/00-initialization.rda', compress='xz')
-

### 2017-03-14 code/R/analysis/05-genetic-nmds.R

@@ -1,5 +1,5 @@
 ## load .rda
-checkpoint::checkpoint('2016-08-03', R.version='3.3.1', scanForPackages=FALSE)
+checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
 session::restore.session('data/intermediate/04-outlier-locus-analysis.rda')
 
 ## load parameters

### 2017-03-14 code/R/analysis/06-format-data-for-prioritisations.R

@@ -1,5 +1,5 @@
 ## load .rda
-checkpoint::checkpoint('2016-08-03', R.version='3.3.1', scanForPackages=FALSE)
+checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
 session::restore.session('data/intermediate/05-genetic-nmds.rda')
 
 # subset data.frames for species with identifiable adaptive and/or neutral genetic variation

### 2017-03-14 code/R/analysis/07-surrogacy-prioritizations.R

@@ -1,5 +1,5 @@
 ## load .rda
-checkpoint::checkpoint('2016-08-03', R.version='3.3.1', scanForPackages=FALSE)
+checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
 session::restore.session('data/intermediate/06-format-data-for-prioritisations.rda')
 
 ## load parameters

### 2017-03-14 code/R/analysis/12-statistical-analysis.R

@@ -1,5 +1,5 @@
 ## load .rda
-checkpoint::checkpoint('2016-08-03', R.version='3.3.1', scanForPackages=FALSE)
+checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
 session::restore.session('data/intermediate/11-multi-species-prioritisations-with-cost.rda')
 load('data/intermediate/10-multi-species-prioritisations-no-cost.rda')
 load('data/intermediate/09-single-species-prioritisations-with-cost.rda')

### 2017-03-14 code/R/analysis/13-exports.R

@@ -1,5 +1,5 @@
 ## load .rda
-checkpoint::checkpoint('2016-08-03', R.version='3.3.1', scanForPackages=FALSE)
+checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
 session::restore.session('data/intermediate/12-statistical-analysis.rda')
 
 #### save results

### 2017-03-14 code/rmarkdown/article.Rmd
shorten article

@@
   Protected areas buffer species from anthropogenic threats and provide places for the processes that generate and maintain biodiversity to continue.
-However, genetic variation, the raw material for many evolutionary processes, is difficult to capture in conservation planning, not least because they require considerable resources to obtain and analyze.
+However, genetic variation, the raw material for evolution, is difficult to capture in conservation planning, not least because genetic data require considerable resources to obtain and analyze.
 Here we show that freely available environmental and geographic distance variables are highly effective surrogates in conservation planning for representing adaptive and neutral intra-specific genetic variation.
@@
 For each species, we identified loci that were potentially under selection using outlier loci methods, and mapped their main gradients of adaptive and neutral genetic variation across the grid cells.
-We then used the cells as planning units to generate prioritize protected area acquisitions.
+We then used the cells as planning units to prioritize protected area acquisitions.
 First, we verified that the spatial patterns of environmental and geographic variation were correlated, respectively, with adaptive and neutral genetic variation.
@@
 <!-- NOTES -->
-<!-- 150 word limit for abstract -->
+<!-- 250 word limit for abstract -->
+<!-- 120 word limit for significance statement -->
 <!-- 5000 word length limit -->
@@
 try(session::restore.session('data/final/results.rda'))
+pvalString <- lazyWeave::pvalString
 
@@
 
-Broadly speaking, though the strength of natural selection is a continuous force, genetic variation is commonly classified as either adaptive or neutral (reviewed in [@r480]).
+Although the strength of natural selection is a continuous force, genetic variation is often classified as adaptive or neutral (reviewed in [@r480]).
 Adaptive genetic variation is associated with loci that significantly affect fitness.
@@
 
-Recently, the use of surrogate data has been proposed to generate prioritizations that sufficiently capture intra-specific genetic variation ([@r2; @r200])--without needing to directly utilize genetic data to achieve this.
-Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been proposed to guide reserve selection as surrogates for adaptive genetic variation (eg.
+Recently, the use of surrogate data has been proposed to generate prioritizations that capture intra-specific genetic variation ([@r2; @r200])--without needing to utilize genetic data to achieve this.
+Since adaptation is ultimately driven by selection pressures, environmental variables have been proposed as surrogates for adaptive genetic variation (eg.
 [@r202; @r2]).
-This is on the basis that sampling a broad range of environmental conditions across species' distributions should also capture more adaptive variation.
-However, the effectiveness of environmental surrogates in representing adaptive genetic variation remains unverified.
-Neutral genetic variation arizes from a reduction in gene flow between individuals.
+Specifically, by ensuring that prioritizations secure individuals in a broad range of environmental conditions, rather than individuals in similar conditions, we would expect to capture a large amount of adaptive genetic variation.
+However, the effectiveness of this approach remains unverified.
+Neutral genetic variation arizes from a reduction in gene flow between populations.
 Over the last few years, the field of phylogeography has predominantly focused on describing neutral genetic diversity, and so the landscape factors that affect neutral variation are relatively well understood (reviewed in [@r501]).
-Potential surrogates for neutral variation have been based on variables that predict the levels of connectivity between planning units.
+Potential surrogates for neutral genetic variation have been based on variables that predict the levels of connectivity between planning units.
 For instance, isolation by distance (IBD; [@r259]) predicts that populations located further apart will experience less gene flow, and in turn, share fewer neutral loci.
-Based on this idea, geographic surrogates based on distance have been found to improve conservation prioritizations in an island network ([@r200]).
+Based on this idea, spreading conservation effort evenly across the geographic distribution of an island network has been found to capture more neutral genetic variation ([@r200]).
 However this remains untested in spatially contiguous systems where connectivity is complicated by many additional factors (eg.
@@
 scenario.SDF <- scenario.DF %>%
-  group_by(prioritization, Context, Metric, Species) %>%
+  group_by(Prioritisation, Context, Metric, Species) %>%
   summarize(genetic.held=mean(genetic.held, na.rm=TRUE)) %>%
@@
 suppressWarnings({
-  ggplot(aes(x=prioritization,y=genetic.held),
+  ggplot(aes(x=Prioritisation,y=genetic.held),
     data=scenario.SDF) +
     geom_boxplot(position=position_dodge(0.9)) +
-    scale_fill_manual(name='prioritization',
+    scale_fill_manual(name='Prioritization',
       values=c('grey80','grey50','grey20')) +
@@
 The proportion of genetic variation secured in a prioritization depended on the targets used to generate the prioritization (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$p.value[[2]])`)$.
-Most notably, prioritizations based on the surrogate variables represented significantly more genetic variation than amount-based prioritizations (`r round(mean(filter(scenario.sub.DF,prioritization=='Amount & surrogate\ntargets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,prioritization=='Amount & surrogate\ntargets')$genetic.held*100),2)` SD overall genetic variation secured versus `r round(mean(filter(scenario.sub.DF,prioritization=='Amount targets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,prioritization=='Amount targets')$genetic.held*100),2)` SD; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$), and were not distinguishable from prioritizations based on measured genetic data (`r round(mean(filter(scenario.sub.DF,prioritization=='Amount & genetic\ntargets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,prioritization=='Amount & genetic\ntargets')$genetic.held*100),2)` SD; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).
+Most notably, prioritizations based on the surrogate variables represented significantly more genetic variation than amount-based prioritizations (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` SD overall genetic variation secured versus `r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)` SD; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$), and were not distinguishable from prioritizations based on measured genetic data (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` SD; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).
 These results indicate that environmental and geographic surrogates in this case far outperform traditional amount-based conservation planning, and perform as well as a conservation plan based directly on measured genetic data.
@@
 _Gypsophila repens_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gypsophila_repens')$Geographic.r2CU,2)`; _Luzula alpinopilosa_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Luzula_alpinopilosa')$Geographic.r2CU,2)`; _Gentiana nivalis_ $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gentiana_nivalis')$Geographic.r2CU,2)`).
-Indeed, geographic distance is expected to be a poor surrogate for connectivity when species deviate from isolation by distance.
-For example, species can utilize disersal modes that maintain high connectivity between distant areas (eg.
-wind dispersing plants; [@r451]), and connectivity between adjacent populations can be disrupted by landscape features ([@r497]).
+One explanation for this result is that geographic distance will be a poor surrogate for neutral genetic variation where connectivity is complicated by additional factors.
+For example, species can maintain high connectivity between distant populations (eg.
+wind dispersing plants; [@r451]), and connectivity between nearby populations can be disrupted by landscape features ([@r497]).
 Overall, such species were the exception in our analysis, and both surrogates performed moderately well for most species.
@@
 <!-- suitable targets for surrogates -->
-Our results suggest that environmental and geographic surrogates can be used to capture genetic variation in prioritizations, the next question is: what percentage of surrogate variation should planners target to preserve genetic variation?
-Firstly, our results suggest that surrogate-based targets should be used with amount-based targets.
-For all species, it was possible to generate solutions that secured a high proportion of the surrogate variables (> 90 %) and only a small proportion of genetic variation (< 20 %; Figures S56--S57).
-By using both amount- and surrogate-based targets, planners may avoid such outcomes.
-Secondly, geographic surrogates seem to need higher targets than environmental surrogates to capture genetic variation.
-In our analysis, neutral genetic variation accumulated much more slowly with geographic representation than adaptive variation accumulated with environmental representation (Figures 2a--2b and S56--S57).
+Our results show that environmental and geographic surrogates can be used to capture genetic variation in prioritizations, the next question is: what percentage of surrogate variation should planners target to preserve genetic variation?
 To secure at least `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` % of the species' neutral genetic variation, prioritizations needed to sample `r round(mean(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` SD of the variation in geographic position among the planning units occupied by the species.
-On the other hand, to secure at least `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` % of each species' adaptive genetic variation, prioritizations only needed to sample `r round(mean(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` SD of the environmental variation in planning units occupied by the species.
-In many individual cases, higher targets may be needed to ensure that prioritizations secure a representative sample of the genetic variation for all of the species in a planning exercise.
+Additionally, to secure at least `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` % of each species' adaptive genetic variation, prioritizations needed to sample `r round(mean(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` SD of the environmental variation in planning units occupied by the species.
+For all species, however, it was possible to generate solutions that secured a high proportion of the surrogate variation (> 90 %) and only a small proportion of genetic variation (< 20 %; Figures S56--S57).
+Planners may avoid such outcomes by using both amount- and surrogate-based targets.
+Depending on study area and the species of conservation interest, higher targets may be needed to ensure that prioritizations secure a representative sample of the genetic variation for all of the species in a planning exercise.
 
 <!-- adaptive genetic variation secured in prioritizations-->
-Our results suggest that conservation planners can secure a reasonably representative sample of the intra-specific adaptive genetic variation using only amount-based targets.
+Our results suggest that conservation planners can secure a reasonably representative sample of the intra-specific adaptive genetic variation using only amount-based targets (Figure 2).
 For most species, the spatial patterns of adaptive genetic variation clustered into a few distinct groups (eg.
@@
 Additionally, this mechanism also explains why the single-species prioritizations generated using only amount targets were often able to achieve high representation--they were essentially a random selection and so frequently contained planning units in different parts of the species' ranges (Figure 4a).
-Overall--and contrary to our original hypotheses--it seems that conventional reserve selection methods might often secure a representative sample of species' adaptive genetic variation.
+Overall--and contrary to our original hypotheses--it seems that conventional reserve selection methods might often conserve species' adaptive genetic variation.
 
@@
 There are several limitations associated with our analysis.
-Firstly, the size of the planning units we used (23 $\times$ 25 km) is much larger than typically used in regional conservation planning scenarios (1--10 km$^2$).
+Firstly, the size of the planning units we used (23 $\times$ 25 km) is much larger than typically used in regional conservation planning (1--10 km$^2$; @rXXX).
 We used this resolution because the genetic data were collected at this scale.
 Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis by artificially introducing additional spatial auto-correlation.
-Secondly, we used geographic distances as surrogates for neutral genetic variation, but connectivity in this landscape may be better described using distances that incorporate information on topography.
-However, such distances require species-specific scaling to adequately describe their connectivity (eg.
-[@r497]), and so cannot easily be utilized in large-scale multi-species planning exercises.
-Thirdly, we used amplified fragment length polymorphism data (AFLP; [@r453]) to characterize the main gradients of genetic variation.
-Whilst next-generation sequencing provides higher resolution genetic information than AFLP data ([@r480]), we know of no suitable multi-species genomic data set, and our methodology would still have utilized only the main gradients of the genetic variation to generate prioritizations in a computationally feasible period of time.
+Secondly, we used geographic distances as surrogates for neutral genetic variation, yet distances that incorperate data on topography may better describe connectivity in the study area.
+However, such distances often require species-specific scaling (eg.
+[@r497]), and so cannot easily be utilized in multi-species planning exercises.
+Thirdly, we used amplified fragment length polymorphism data (AFLP; [@r453]) to describe genetic variation.
+Whilst next-generation sequencing provides higher resolution genetic information than AFLP data ([@r480]), we know of no suitable multi-species genomic data set, and our methodology would still have utilized only the main gradients of the genetic variation to generate prioritizations in a feasible period of time.
 
@@
 [@r455]) and neutral genetic variation (eg.
-[@r478]), and the potential for species richness as a potential surrogate for genetic diversity ([@r486]).
+[@r478]), and the potential for species richness as a surrogate for genetic diversity ([@r486]).
 Data was collected using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx.
 22.3 km $\times$ 25 km; Figure S1).
-Project members visited every second grid cell, and for each species, they collected samples from three individuals in each cell they were detected in.
-They genotyped samples using AFLPs, and constructed matrices denoting the presence/absence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; for more information see [@r478], Table 1).
+Project members visited every second grid cell, and for each species, they collected samples from three individuals if they were detected.
+They genotyped samples using AFLPs, and constructed matrices denoting the presence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; for more information see [@r478], Table 1).
 Thus the data set contains information describing the genomic properties of individuals for each species over a geographic grid.
 We used this data set because it comprized comparable genetic data for a range of species with different evolutionary and life histories that was collected using a standardized sampling scheme.
-Additionally, the study area contained large environmental variation which is useful for examining environmental surrogates.
 
@@
 To reduce computational burden, we omitted cells that did not contain samples (`r nrow(grid.DF)` grid cells used in analysis).
-To examine the potential effects of variation in cost, we calculated the total human population density inside each planning unit (1 km$^2$ resolution from the Global Rural-Urban Mapping Project; [@r472]) and used this to represent acquisition cost (Figure 2b).
+To examine the effect of variation in cost, we calculated the total human population density inside each planning unit (1 km$^2$ resolution from the Global Rural-Urban Mapping Project; [@r472]) and used this to represent acquisition cost (Figure 2b).
 All spatial and statistical analyses were conducted in \texttt{R} (version `r paste0(R.version$major, '.', R.version$minor)`; [@r192]).
@@
 To ensure that the ordinations described a sufficient amount of the genetic variation, we ran successive scaling analyses with increasing dimensionality until a sufficient stress value was obtained (maximum stress value $\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for each analysis).
-Since each grid cell had multiple samples per species, we averaged the ordinated values belonging to conspecifics in the same grid cell.
+Since each grid cell had multiple samples per species, we averaged the ordinated values for conspecifics in the same grid cell.
 
@@
 
-prioritization method {#prioritization-method .unnumbered}
+Prioritization method {#prioritization-method .unnumbered}
 ----------------------------------------------------------
 
-We used the unreliable representative and adequate prioritization (URAP) formulation for reserve selection ([@r465]).
-This formulation can identify the minimum number of planning units required to preserve both a target proportion of the species' geographic range (using amount-based targets) and a broad sample of variation found across the species' range.
+We used the unreliable representative and adequate prioritization (URAP) formulation to generate solutions ([@r465]).
+This formulation can identify the cheapest set of planning units required to preserve both a target proportion of the species' geographic range (using amount-based targets) and a broad sample of variation found across the species' range.
 We used the environmental and geographic surrogate spaces, and the adaptive (if detected) and neutral genetic spaces to generate and evaluate solutions.
-Solutions associated with negative values--because they secured a minute fraction of genetic variation--were replaced with zeros to facilitate statistical analysis.
+Solutions associated with negative values--because they secured very little genetic variation--were replaced with zeros to facilitate statistical analysis.
 We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality (using the \texttt{raptr R} package [@r465]).
@@
 
-Our first aim was to determine if the environmental and geographic variables correlate with spatial patterns of adaptive and neutral genetic variation.
-To achieve this, for each species, we generated multiple dissimilarity matrices expressing differences between the planning units where the species was found.
-Using data from the species' surrogate and genetic spaces, these dissimilarity matrices showed differences between the planning units in terms of their (i) geographic position, (ii) environmental characteristics, and the (iii) adaptive (if detected) and (iv) neutral genetic characteristics of the individuals inside them.
-Dissimilarities were computed as Euclidean distances.
+Our first aim was to determine if the environmental and geographic variables correlate with the spatial patterns of adaptive and neutral genetic variation.
+To achieve this, for each species, we created dissimilarity matrices using Euclidean distances and the data in the surrogate and genetic spaces.
+These matrices showed differences between the planning units occupied by the species in terms of the units' (i) geographic position, (ii) environmental characteristics, and the (iii) adaptive (if detected) and (iv) neutral genetic characteristics of the individuals inside them.
 
-We then fitted maximum likelihood population-effects (MLPE) models ([@r494]).
-These models accommodated the structure of dissimilarity matrices using random effects.
+We fitted maximum likelihood population-effects (MLPE) models using maximum likelihood to investigate correlations between the dissimilarity models (using the \texttt{lme4 R} package; [@r469]).
+These models use random effects to accommodate the structure of dissimilarity matrices.
 For each species, we fitted a MLPE model to the species' dissimilarity matrices based on geographic position and neutral genetic variation.
-Additionally, if adaptive loci were detected, we also fitted a MLPE model to the species' dissimilarity matrices describing environmental and adaptive genetic variation.
-All predictor variables were _z_-transformed to improve convergence.
-To test if the surrogates explained a substantial amount of genetic variation, we compared each model to its null model using $\chi^2$, and applied Bonferroni corrections.
+If adaptive loci were detected, we also fitted a MLPE model to the species' dissimilarity matrices describing environmental and adaptive genetic variation.
+All data variables were _z_-transformed to improve convergence.
+To test if the surrogates explained the genetic variation, we compared each model to its null model using a $\chi^2$ test, and applied Bonferroni corrections.
 
@@
 For each species, we generated `r format(raptr.params.LST[[MODE]]$surrogacy.analysis$random.replicates*2, big.mark=',', scientific=FALSE)` prioritizations by randomly selecting different combinations of planning units that the species occupied.
-For half of these random prioritizations, we calculated the proportion of geographic variation and neutral genetic variation they secured, and for the other half, the proportion of environmental variation and adaptive genetic variation they secured.
-These calculations were performed using built-in routines in the \texttt{raptr R} package.
+For half of these random prioritizations, we calculated the proportion of geographic variation and neutral genetic variation they secured, and for the other half, we calculated the proportion of environmental variation and adaptive genetic variation they secured.
+The proportion of surrogate and genetic variation captured in a prioritization was calculated using the \texttt{raptr R} package.
 
@@
 
-Our third aim was to determine if surrogate-based targets improved representation of genetic variation by the prioritizations.
+Our third aim was to determine if surrogate-based targets improved representation of genetic variation in prioritizations.
 As previously mentioned, we generated prioritizations using different combinations of targets and conservation planning scenarios.
-We used `r (raptr.params.LST[[MODE]]$scenario.analysis$amount.target*100)` % amount-based targets for each species in all prioritizations to secure an adequate proportion of the species' ranges.
-Based on the results from our previous aim, we used `r (raptr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100)` % surrogate-based targets and `r (raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)` % genetic-based targets.
+We used `r (raptr.params.LST[[MODE]]$scenario.analysis$amount.target*100)` % amount-based targets for each species in all prioritizations to secure an adequate proportion of the species' distributions.
+Based on the results from our previous analysis, we used `r (raptr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100)` % surrogate-based targets and `r (raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)` % genetic-based targets.
 We generated a single solution for each target/scenario combination, except for the "single-species (equal costs) with amount-based targets" combination for which we generated `r format(raptr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates, big.mark=',', scientific=FALSE)` replicates because it had many optimal solutions.
@@
 
-We investigated the benefits of using surrogate-based targets in prioritizations using generalized linear mixed-effects models fit with logit links (GLMMs; using the \texttt{lme4 R} package; [@r469]).
+We fitted generalized linear mixed-effects models (GLMMs) with logit links to  evaluate the prioritizations (using the \texttt{lme4 R} package).
 We fitted a full model to the proportion of genetic variation secured in a given prioritization.
-This model contained categorical variables indicating the combination of targets (ie.
-amount only, amount and surrogate targets, or amount and genetic targets), the scenario used to generate the prioritization (ie.
-single-species, multi-species, or multi-species with cost), the type of genetic variation measured (ie.
-adaptive or neutral), and all interactions between them.
-prioritizations generated for the same species were accommodated using a random intercept term.
-We subjected the full model to a backwards stepwise term deletion routine to assess term significance, and conducted a _post-hoc_ analysis on the minimal adequate model using Tukey contrasts with a Bonferroni correction (using the \texttt{multcomp R} package; [@r470]).
+This model contained categorical variables indicating the targets (amount only, amount and surrogate targets, or amount and genetic targets), the planning scenario (single-species, multi-species, or multi-species with cost), the type of genetic variation measured (adaptive or neutral), and all interactions between them.
+Data for same species were accommodated using a random intercept term.
+The full model was subjected to a backwards stepwise term deletion routine to assess significance.
+A _post-hoc_ analysis was conducted on the minimal adequate model using Tukey contrasts with a Bonferroni correction (using the \texttt{multcomp R} package; [@r470]).
 

### 2017-03-14 code/rmarkdown/article.Rmd
move pvalString delcaration to initialization

@@
 try(session::restore.session('data/final/results.rda'))
-pvalString <- lazyWeave::pvalString
 

### 2017-03-16 code/rmarkdown/article.Rmd
update article

@@
   - code: b
-    address: School of Geography, Planning and Environmental Management, The University of Queensland, Brisbane, QLD, Australia
+    address: School of Earth and Environmental Sciences, The University of Queensland, Brisbane, QLD, Australia
 corresponding_author:
@@
 significance: |
-  To protect biodiversity for the long term, nature reserves and other protected areas need to represent a broad range of different gentic types.
+  To protect biodiversity for the long term, nature reserves and other protected areas need to represent a broad range of different genetic types.
 However, genetic data are expensive and time consuming to obtain.
 Here we show that freely available environmental and geographic variables can be used as effective surrogates for genetic data in conservation planning.
-This means that conservation plannners can, with some confidence, design protected area systems to represent genetic diversity without investing in expensive programmes to obtain and analyze genetic data.
+This means that conservation planners can, with some confidence, design protected area systems to represent intra-specific genetic diversity without investing in expensive programmes to obtain and analyze genetic data.
 acknowledgements: |
@@
 [@r202; @r2]).
-Specifically, by ensuring that prioritizations secure individuals in a broad range of environmental conditions, rather than individuals in similar conditions, we would expect to capture a large amount of adaptive genetic variation.
+Specifically, by conserving individuals in a broad range of environmental conditions, we might expect to capture individuals with a diverse set of local adaptations, and overall capture a large proportion of the adaptive genetic variation present in the species.
 However, the effectiveness of this approach remains unverified.
-Neutral genetic variation arizes from a reduction in gene flow between populations.
+Neutral genetic variation arises from a reduction in gene flow between populations.
 Over the last few years, the field of phylogeography has predominantly focused on describing neutral genetic diversity, and so the landscape factors that affect neutral variation are relatively well understood (reviewed in [@r501]).
-Potential surrogates for neutral genetic variation have been based on variables that predict the levels of connectivity between planning units.
-For instance, isolation by distance (IBD; [@r259]) predicts that populations located further apart will experience less gene flow, and in turn, share fewer neutral loci.
-Based on this idea, spreading conservation effort evenly across the geographic distribution of an island network has been found to capture more neutral genetic variation ([@r200]).
+Potential surrogates for neutral genetic variation have been based on variables that predict the level of connectivity between different areas.
+For instance, according the isolation by distance model (IBD; [@r259]), populations located further apart are predicted experience less gene flow, and in turn, share fewer neutral loci.
+Based on this idea, spreading conservation priorities evenly across the geographic distribution of an island network has been found to capture more neutral genetic variation ([@r200]).
 However this remains untested in spatially contiguous systems where connectivity is complicated by many additional factors (eg.
-habitat; [@r497])--the typical situation in most conservation planning exercises.
+land use; [@r497])--the typical situation in most conservation planning exercises.
 
@@
 We also show that for most species the level of association is strong enough to be operational for conservation planning.
-Finally, we demonstrate that using these surrogates results in prioritizations that secure a substantial proportion of genetic variation.
+Finally, we demonstrate that using these surrogates results in prioritizations that secure a substantial proportion of intra-specific genetic variation.
 
@@
 We detected putatively adaptive genetic variation in `r as.character(as.english(length(env.surrogacy.LST)))` of the 27 plant species.
-We used two outlier detection methods to identify loci under selection, and they returned reasonably consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species assigned the same classification $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD; Table S1).
+We used two outlier detection methods to identify loci under selection.
+These methods returned reasonably consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species assigned the same classification $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD; Table S1).
 Of the species that were associated with adaptive genetic variation, only a small proportion of loci were classified as being adaptive (mean `r round(mean(spp.outlier.loci.props, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.loci.props, na.rm=TRUE),2)` SD).
@@
 <!-- correlation analysis -->
-We verified that the spatial patterns in environmental variation correlated with the broad-scale patterns in adaptive genetic variation.
-We further verified that there was a correlation between neutral genetic variation and variation in geographic position.
-For each species, we constructed dissimilarity matrices expressing differences between the planning units where individuals were detected based on the units' (i) environmental characteristics, (ii) geographic position, and also the (iii) adaptive (if adaptive loci detected), and (iv) neutral genetic characteristics of individuals found inside them.
+We verified that the spatial patterns in environmental variation correlated with the broad-scale patterns in adaptive genetic variation for each species.
+We also verified that there was a correlation between neutral genetic variation and variation in geographic position.
+For each species, we constructed dissimilarity matrices expressing differences between the planning units where individuals were detected based on the units' (i) environmental characteristics, (ii) geographic position, and also the (iii) adaptive (if adaptive loci detected), and (iv) neutral genetic characteristics of the individuals found inside them.
 The spatial patterns of environmental variation were significantly correlated with the pattern of adaptive genetic variation for `r as.character(as.english(sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)))` of the `r as.character(as.english(sum(!is.na(correlation.model.summary.DF$Environmental.corr.P))))` species associated with adaptive variation (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models; Table S3).
@@
 <!-- surrogacy analysis -->
-For each species, we generated a suite of `r format(raptr.params.LST[[MODE]]$surrogacy.analysis$random.replicates, big.mark=',', scientific=FALSE)` random prioritizations, and calculated the proportion of environmental variation and adaptive (if present) genetic variation each prioritization captured.
-We then repeated this process, and calculated the proportion of geographic variation and neutral genetic variation they sampled.
+For each species, we generated a suite of `r format(raptr.params.LST[[MODE]]$surrogacy.analysis$random.replicates, big.mark=',', scientific=FALSE)` random prioritizations, and calculated the proportion of environmental variation and adaptive (if present) genetic variation that each prioritization captured.
+We then repeated this process, and calculated the proportion of variation in geographic position and neutral genetic variation they sampled.
 The environmental and geographic variables were moderately effective predictors for the genetic variation represented by the randomly generated prioritizations (Figure 1).
@@
 _Post-hoc_ analyses showed that these relationships were positive for all species (environmental: minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; geographic: minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S4).
-After establishing that the the proportion of surrogate variation secured in a prioritization also predicted the proportion of genetic variation it secured, our final step was to determine if prioritizations were more effective when generated using targets based on the surrogates.
+After establishing that the the proportion of surrogate variation secured in a prioritization also predicted the proportion of genetic variation it secured, our final step was to determine if prioritizations were more effective when generated using targets based on the surrogate variables.
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=6, fig.height=4.5, out.width='3.42in', out.height='2.5in', fig.cap=paste0('The proportion of adaptive (a; ',sum(!is.na(correlation.model.summary.DF$Environmental.corr.P)),' species) and neutral (b; ',sum(!is.na(correlation.model.summary.DF$Geographic.corr.P)),' species) genetic variation represented in a suite of randomly generated prioritizations for different species (lines) as a function of the surrogate variation they also secured.
-Panels (c) and (d) show the distribution of $R^2$ values for the ',sum(!is.na(correlation.model.summary.DF$Environmental.corr.P)) + sum(!is.na(correlation.model.summary.DF$Geographic.corr.P)),' models that explain genetic variation using e to explain genetic variation.')}
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=6, fig.height=4.5, out.width='3.42in', out.height='2.5in', fig.cap=paste0('The proportion of adaptive (a; ',sum(!is.na(correlation.model.summary.DF$Environmental.corr.P)),' species) and neutral (b; ',sum(!is.na(correlation.model.summary.DF$Geographic.corr.P)),' species) genetic variation represented in a suite of randomly generated prioritizations for different species as a function of the surrogate variation they also secured (lines).
+Panels (c) and (d) show the distribution of $R^2$ values for the ',sum(!is.na(correlation.model.summary.DF$Environmental.corr.P)) + sum(!is.na(correlation.model.summary.DF$Geographic.corr.P)),' models shown in panels (a) and (b).')}
 # make figure
@@
 <!-- scenario analysis -->
-To determine whether using environmental and geographic targets could improve the effectiveness of prioritizations in representing genetic variation, we generated prioritizations using (i) "amount targets" reflecting the traditional approach of representing a certain proportion of each species' geographic distribution, (ii) "amount and surrogate targets" where we targeted representation of the environmental and geographic surrogates as well as a certain proportion of each species' geographic distribution, and (iii) "amount and genetic targets" where we targeted representation of the directly measured genetic variation as well as a certain proportion of each species' geographic distribution (Figure 3).
-To determine if these patterns were robust, for each of the three combinations of the targets, we generated prioritizations under four scenarios: (i) single-species with equal costs, (ii) single species with opportunity costs, (iii) multi-species with equal costs, and (iv) multi-species with opportunity costs.
+To determine whether environmental and geographic targets could improve the effectiveness of prioritizations in representing genetic variation, we generated prioritizations using (i) "amount targets" reflecting the traditional approach of representing a certain proportion of each species' geographic distribution, (ii) "amount and surrogate targets" where we targeted representation of the environmental and geographic surrogates as well as a certain proportion of each species' geographic distribution, and (iii) "amount and genetic targets" where we targeted representation of the directly measured genetic variation as well as a certain proportion of each species' geographic distribution (Figure 3).
+To determine if these patterns were robust, for each of the three combinations of the targets, we generated prioritizations under four scenarios: (i) single-species with equal costs, (ii) single species with acquisition costs, (iii) multi-species with equal costs, and (iv) multi-species with acquisition costs.
 
 ```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=(5*0.9), fig.height=(7.1*0.9), out.width='3.42in', out.height='4in', fig.cap='The proportion of adaptive and neutral genetic variation secured for each species in solutions based on different targets and scenarios.
-Box plots show the 25th percentile, median, and 75th percentile.
+Box plots show the median, and the 25th and 75th percentiles.
 Whiskers show the data at 1.5 $\times$ the inter-quantile range.
@@
 Most notably, prioritizations based on the surrogate variables represented significantly more genetic variation than amount-based prioritizations (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` SD overall genetic variation secured versus `r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)` SD; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$), and were not distinguishable from prioritizations based on measured genetic data (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` SD; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).
-These results indicate that environmental and geographic surrogates in this case far outperform traditional amount-based conservation planning, and perform as well as a conservation plan based directly on measured genetic data.
+These results indicate that environmental and geographic surrogates in this case far outperform traditional amount-based conservation planning, and perform almost as well as a conservation plan based directly on measured genetic data.
 
-Overall, the prioritizations tended to secure a greater proportion of adaptive than neutral genetic variation (adaptive: mean `r round(mean(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` secured across all prioritizations, neutral: mean `r round(mean(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)` $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)`; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.3.DF$p.value[[2]])`).
+Overall, the prioritizations tended to secure a greater proportion of adaptive genetic variation than neutral genetic variation (adaptive: mean `r round(mean(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` secured across all prioritizations, neutral: mean `r round(mean(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)` $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)`; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.3.DF$p.value[[2]])`).
 Thus, regardless of the targets used to generate a prioritization, or the planning context under which the prioritization was generated, prioritizations tend to secure more adaptive than neutral genetic genetic variation.
@@
 Both the multi-species (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)`) and multi-species with cost (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)`) were found to secure more genetic variation than the single species scenario (mean `r round(mean(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[2]])`$; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$ respectively).
-These results suggest that on average the multi-species prioritizations can secure more genetic variation than single-species prioritizations--regardless of targets used to generate them.
+These results suggest that on average the multi-species prioritizations may secure more genetic variation than single-species prioritizations--regardless of whether surrogates were used to generate them.
 
@@
 Our results demonstrate that environmental and geographic variables are effective surrogates for most species (Figure 2).
-Despite this, for a few species, geographic distance was a suprisingly poor surrogate for neutral genetic variation (eg.
+Despite this, geographic distance was a suprisingly poor surrogate for neutral genetic variation for a few species (eg.
 _Gypsophila repens_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gypsophila_repens')$Geographic.r2CU,2)`; _Luzula alpinopilosa_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Luzula_alpinopilosa')$Geographic.r2CU,2)`; _Gentiana nivalis_ $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gentiana_nivalis')$Geographic.r2CU,2)`).
 One explanation for this result is that geographic distance will be a poor surrogate for neutral genetic variation where connectivity is complicated by additional factors.
-For example, species can maintain high connectivity between distant populations (eg.
-wind dispersing plants; [@r451]), and connectivity between nearby populations can be disrupted by landscape features ([@r497]).
+For example, some species can maintain relatively high connectivity between distant populations (eg.
+wind dispersing plants; [@r451]), and in some places can be disrupted by landscape features (eg.
+anthropogenically modified land [@r497]).
 Overall, such species were the exception in our analysis, and both surrogates performed moderately well for most species.
@@
 Our results show that environmental and geographic surrogates can be used to capture genetic variation in prioritizations, the next question is: what percentage of surrogate variation should planners target to preserve genetic variation?
-To secure at least `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` % of the species' neutral genetic variation, prioritizations needed to sample `r round(mean(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` SD of the variation in geographic position among the planning units occupied by the species.
-Additionally, to secure at least `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` % of each species' adaptive genetic variation, prioritizations needed to sample `r round(mean(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` SD of the environmental variation in planning units occupied by the species.
-For all species, however, it was possible to generate solutions that secured a high proportion of the surrogate variation (> 90 %) and only a small proportion of genetic variation (< 20 %; Figures S56--S57).
+To secure at least `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` % of the species' neutral genetic variation, prioritizations needed to sample `r round(mean(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` SD of the variation in geographic position among the planning units occupied by each species.
+Additionally, to secure at least `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` % of each species' adaptive genetic variation, prioritizations needed to sample `r round(mean(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` SD of the environmental variation in planning units occupied each the species.
+For all species, however, it was possible to generate solutions that secured a large proportion of the surrogate variation (> 90 %) and only a small proportion of genetic variation (< 20 %; Figures S56--S57).
 Planners may avoid such outcomes by using both amount- and surrogate-based targets.
-Depending on study area and the species of conservation interest, higher targets may be needed to ensure that prioritizations secure a representative sample of the genetic variation for all of the species in a planning exercise.
+Depending on the study area and the species of conservation interest, higher targets may be needed to increase the likelihood that prioritizations will secure a large proportion of the intra-specific genetic variation for all of the species in the planning exercise.
 
@@
 Additionally, this mechanism also explains why the single-species prioritizations generated using only amount targets were often able to achieve high representation--they were essentially a random selection and so frequently contained planning units in different parts of the species' ranges (Figure 4a).
-Overall--and contrary to our original hypotheses--it seems that conventional reserve selection methods might often conserve species' adaptive genetic variation.
+Overall--and contrary to our original hypothesis--it seems that conventional reserve selection methods might often conserve species' adaptive genetic variation.
 
@@
 There are several limitations associated with our analysis.
-Firstly, the size of the planning units we used (23 $\times$ 25 km) is much larger than typically used in regional conservation planning (1--10 km$^2$; @rXXX).
+Firstly, the size of the planning units we used (approx.
+23 $\times$ 25 km) is much larger than typically used in regional conservation planning (1--10 km$^2$; @rXXX).
 We used this resolution because the genetic data were collected at this scale.
 Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis by artificially introducing additional spatial auto-correlation.
-Secondly, we used geographic distances as surrogates for neutral genetic variation, yet distances that incorperate data on topography may better describe connectivity in the study area.
+Secondly, we used geographic distances as surrogates for neutral genetic variation, yet distances that incorporate data on topography may better describe connectivity in the study area.
 However, such distances often require species-specific scaling (eg.
@@
 Data was collected using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx.
-22.3 km $\times$ 25 km; Figure S1).
-Project members visited every second grid cell, and for each species, they collected samples from three individuals if they were detected.
+22.3 $\times$ 25 km; Figure S1).
+Project members visited every second grid cell, and for each species, they collected samples from three individuals if any individuals were found.
 They genotyped samples using AFLPs, and constructed matrices denoting the presence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; for more information see [@r478], Table 1).
 Thus the data set contains information describing the genomic properties of individuals for each species over a geographic grid.
-We used this data set because it comprized comparable genetic data for a range of species with different evolutionary and life histories that was collected using a standardized sampling scheme.
+We used this data because it comprized comparable genetic data for a range of species with different evolutionary and life histories that was collected using a standardized sampling scheme.
 
@@
 To describe the environmental characteristics of each planning unit, we obtained 19 climatic layers ($30^{\prime\prime}$ resolution; [@r33]), projected them and the planning units into an equal-area coordinate system (Europe Lambert Conformal Conic; $\texttt{ESRI:102014}$), and computed planning unit averages for each climatic variable.
-To reduce dimensionality, for each species, we subjected the climatic values associated with their planning units to a principal components analysis (PCA; Table S1).
+To reduce dimensionality, for each species, we subjected the climatic values associated with the planning units they were found in to a principal components analysis (PCA; Table S1).
 We used the first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components to characterize climatic variation found across the species' geographic distributions inside the study area (capturing `r round(mean(spp.pca.DF[[2]]*100),2)` % $\pm$ `r round(sd(spp.pca.DF[[2]]*100),2)` SD of the total climatic variation; Figure S2).

### 2017-05-29 code/rmarkdown/article.Rmd
fix bug in combining results for 3rd analysis update refs and article

@@
 author_contributions: |
-  All authors developed the study and drafted the manuscript. JOH performed all analyses.
+  JOH performed the analyses. All authors developed the study and drafted the manuscript.
 lead_author_surname: "Hanson"
@@
 Since the resources available for conservation are limited, protected areas need to be sited in places that fulfill conservation objectives for minimal cost ([@r255]).
-To achieve this, conservation planning exercises often generate plans for entire networks of protected areas (prioritizations) to allocate resources cost-effectively and preserve broad-scale biodiversity processes.
-Intra-specific genetic variation is fundamental to long-term species persistence ([@r165]), and as a consequence, there has been increasing interest in designing prioritizations that represent this aspect of biodiversity ([@r67; @r3; @r61; @r499]).
+To achieve this, conservation planning exercises often generate plans for entire networks of protected areas (prioritizations) to allocate resources cost-effectively and preserve broad-scale biodiversity processes ([@r9]).
+Intra-specific genetic variation is fundamental to long-term species persistence ([@r165; @r504]), and as a consequence, there has been increasing interest in designing prioritizations that represent this aspect of biodiversity ([@r67; @r3; @r499; @r61; @r503]).
 
@@
 The resulting ordinations were used to construct an adaptive (if detected) and neutral genetic space for each species.
-The spatial distribution of these genetic spaces (Figures S29--S54) generally showed substantial spatial auto-correlation (eg.
+The spatial distribution of these genetic spaces (Figures S1--S27) generally showed substantial spatial auto-correlation (eg.
 _Cerastium uniflorum_, Figure S34; _Dryas octopetala_, Figure S36), suggesting that the environmental and geographic variables have potential to be effective surrogates for genetic variation.
@@
 We then repeated this process, and calculated the proportion of variation in geographic position and neutral genetic variation they sampled.
-The environmental and geographic variables were moderately effective predictors for the genetic variation represented by the randomly generated prioritizations (Figure 1).
+The environmental and geographic variables were moderately effective predictors for the genetic variation represented by the randomly generated prioritizations (Figures 1, S28 and S29).
 The relationship between the proportion of genetic and surrogate variation secured in the prioritizations varied among the species (species $\times$ proportion of environmental variation interaction term: ${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$; species $\times$ proportion of geographic variation interaction term: ${\chi^2}_{`r round(geo.surrogacy.AOV$df[[4]],2)`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`).
@@
 <!-- scenario analysis -->
-To determine whether environmental and geographic targets could improve the effectiveness of prioritizations in representing genetic variation, we generated prioritizations using (i) "amount targets" reflecting the traditional approach of representing a certain proportion of each species' geographic distribution, (ii) "amount and surrogate targets" where we targeted representation of the environmental and geographic surrogates as well as a certain proportion of each species' geographic distribution, and (iii) "amount and genetic targets" where we targeted representation of the directly measured genetic variation as well as a certain proportion of each species' geographic distribution (Figure 3).
-To determine if these patterns were robust, for each of the three combinations of the targets, we generated prioritizations under four scenarios: (i) single-species with equal costs, (ii) single species with acquisition costs, (iii) multi-species with equal costs, and (iv) multi-species with acquisition costs.
+To determine whether environmental and geographic targets could improve the effectiveness of prioritizations in representing genetic variation, we generated prioritizations using (i) "amount targets" reflecting the traditional approach of representing a certain proportion of each species' geographic distribution, (ii) "amount and surrogate targets" where we targeted representation of the environmental and geographic surrogates as well as a certain proportion of each species' geographic distribution, and (iii) "amount and genetic targets" where we targeted representation of the directly measured genetic variation as well as a certain proportion of each species' geographic distribution.
+To determine if these patterns were robust, for each of the three combinations of the targets, we generated prioritizations under four scenarios: (i) single-species with equal costs, (ii) single species with acquisition costs, (iii) multi-species with equal costs, and (iv) multi-species with acquisition costs (Figure S30).
 
@@
 The proportion of genetic variation secured in a prioritization depended on the targets used to generate the prioritization (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$p.value[[2]])`)$.
-Most notably, prioritizations based on the surrogate variables represented significantly more genetic variation than amount-based prioritizations (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` SD overall genetic variation secured versus `r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)` SD; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$), and were not distinguishable from prioritizations based on measured genetic data (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` SD; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).
+Most notably, prioritizations based on the surrogate variables represented significantly more genetic variation than amount-based prioritizations (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` SD overall genetic variation secured versus `r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)` SD; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$), and were not distinguishable from prioritizations based on measured genetic data (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` SD; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).
 These results indicate that environmental and geographic surrogates in this case far outperform traditional amount-based conservation planning, and perform almost as well as a conservation plan based directly on measured genetic data.
@@
 The average proportion of genetic variation secured by prioritizations varied under different scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.2.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.2.DF$p.value[[2]])`).
-Both the multi-species (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)`) and multi-species with cost (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)`) were found to secure more genetic variation than the single species scenario (mean `r round(mean(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[2]])`$; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$ respectively).
-These results suggest that on average the multi-species prioritizations may secure more genetic variation than single-species prioritizations--regardless of whether surrogates were used to generate them.
+Specifically, prioritizations generated under single-species scenario (mean `r round(mean(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held*100),2)`) secured less genetic variation than those under the single species with opportunity costs (mean `r round(mean(filter(scenario.sub.DF,Context=='Single-species\n(opportunity costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Single-species\n(opportunity costs)')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$) or the multi-species with cost scenarios (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[3]])`$).
+These results suggest the amount of genetic variation secured in a prioritization may be affected by the data used to represent acquisition cost.
 
@@
 _Gypsophila repens_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gypsophila_repens')$Geographic.r2CU,2)`; _Luzula alpinopilosa_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Luzula_alpinopilosa')$Geographic.r2CU,2)`; _Gentiana nivalis_ $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gentiana_nivalis')$Geographic.r2CU,2)`).
-One explanation for this result is that geographic distance will be a poor surrogate for neutral genetic variation where connectivity is complicated by additional factors.
+One explanation for this result is that geographic distance will be a poor surrogate for neutral genetic variation where connectivity is complicated by additional factors ([@r505; @r497]).
 For example, some species can maintain relatively high connectivity between distant populations (eg.
@@
 Firstly, the size of the planning units we used (approx.
-23 $\times$ 25 km) is much larger than typically used in regional conservation planning (1--10 km$^2$; @rXXX).
+23 $\times$ 25 km) is much larger than typically used in regional conservation planning (1--10 km$^2$).
 We used this resolution because the genetic data were collected at this scale.
@@
 
-We used data for 27 alpine plant species in the European Alps collected by the IntraBioDiv project ([@r451]; Figure 2a).
-All data, code, and results are stored in an online repository to permit replication and validation of this study (\url{www.github.com/jeffreyhanson/genetic.surrogates}).
+We used data for 27 alpine plant species in the European Alps collected by the IntraBioDiv project ([@r451]; Figure 3a).
+All data, code, and results are stored in an online repository to permit replication and validation of this study (\url{www.github.com/jeffreyhanson/genetic-surrogates}).
 This data set has been used to explore patterns of adaptive (eg.
@@
 Data was collected using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx.
-22.3 $\times$ 25 km; Figure S1).
+22.3 $\times$ 25 km; Figure S31).
 Project members visited every second grid cell, and for each species, they collected samples from three individuals if any individuals were found.
@@
 Thus the data set contains information describing the genomic properties of individuals for each species over a geographic grid.
-We used this data because it comprized comparable genetic data for a range of species with different evolutionary and life histories that was collected using a standardized sampling scheme.
+We used this data because it comprised comparable genetic data for a range of species with different evolutionary and life histories that was collected using a standardized sampling scheme.
 
@@
 To reduce computational burden, we omitted cells that did not contain samples (`r nrow(grid.DF)` grid cells used in analysis).
-To examine the effect of variation in cost, we calculated the total human population density inside each planning unit (1 km$^2$ resolution from the Global Rural-Urban Mapping Project; [@r472]) and used this to represent acquisition cost (Figure 2b).
+To examine the effect of variation in cost, we calculated the total human population density inside each planning unit (1 km$^2$ resolution from the Global Rural-Urban Mapping Project; [@r472]) and used this to represent acquisition cost (Figure 3b).
 All spatial and statistical analyses were conducted in \texttt{R} (version `r paste0(R.version$major, '.', R.version$minor)`; [@r192]).
 
-We created environmental and geographic surrogate variables for each species.
-To describe the geographic location of each planning unit, we projected the grid into an equi-distant coordinate system (Europe Equidistant Conic; $\texttt{ESRI:102031}$), calculated the centroid of each grid cell, and extracted their two-dimensional coordinates.
-To describe the environmental characteristics of each planning unit, we obtained 19 climatic layers ($30^{\prime\prime}$ resolution; [@r33]), projected them and the planning units into an equal-area coordinate system (Europe Lambert Conformal Conic; $\texttt{ESRI:102014}$), and computed planning unit averages for each climatic variable.
+We created environmental and geographic surrogate variables for each species (Figure S32).
+To describe the geographic location of each planning unit (Figure S33), we projected the grid into an equi-distant coordinate system (Europe Equidistant Conic; $\texttt{ESRI:102031}$), calculated the centroid of each grid cell, and extracted their two-dimensional coordinates.
+To describe the environmental characteristics of each planning unit (Figure S34), we obtained 19 climatic layers ($30^{\prime\prime}$ resolution; [@r33]), projected them and the planning units into an equal-area coordinate system (Europe Lambert Conformal Conic; $\texttt{ESRI:102014}$), and computed planning unit averages for each climatic variable.
 To reduce dimensionality, for each species, we subjected the climatic values associated with the planning units they were found in to a principal components analysis (PCA; Table S1).
@@
 
-To investigate the effectiveness of our supposed surrogates, we first needed to identify which of the sampled loci were adaptive.
+To investigate the effectiveness of our supposed surrogates, we first needed to identify which of the sampled loci were adaptive (Figure S35).
 Following best practice, we used two outlier detection methods to achieve this (an individual- and a population-level method; [@r495]).
@@
 We adopted a similar methodology to Bothwell _et al._ ([-@r455]) and applied it to each of the species separately.
-Following their methodology, we initially grouped conspecifics into genetic lineages to further minimize false-positives, by fitting admixture and correlated alleles models implemented in \texttt{Structure} (version 2.3.4; [@r462]; `r structure.params.LST[[MODE]]$numruns` replicates per species; `r format(structure.params.LST[[MODE]]$admburnin, big.mark=',', scientific=FALSE)` admixture burnin iterations; `r format(structure.params.LST[[MODE]]$burnin, big.mark=',', scientific=FALSE)` burnin iterations; `r format(structure.params.LST[[MODE]]$numreps+structure.params.LST[[MODE]]$burnin, big.mark=',', scientific=FALSE)` total iterations) using the number of lineages previously determined by Alvarez _et al._ ([@r478]) and combining replicate runs using \texttt{ClumPP} (version 1.1.2; [@r459]; greedy algorithm based on the $G^{\prime}$ statistic; `r format(clumpp.params.LST[[MODE]]$repeats, big.mark=',', scientific=FALSE)` iterations).
+Following their methodology, we initially grouped conspecifics into genetic lineages to further minimize false-positives (Figures S36--S62), by fitting admixture and correlated alleles models implemented in \texttt{Structure} (version 2.3.4; [@r462]; `r structure.params.LST[[MODE]]$numruns` replicates per species; `r format(structure.params.LST[[MODE]]$admburnin, big.mark=',', scientific=FALSE)` admixture burnin iterations; `r format(structure.params.LST[[MODE]]$burnin, big.mark=',', scientific=FALSE)` burnin iterations; `r format(structure.params.LST[[MODE]]$numreps+structure.params.LST[[MODE]]$burnin, big.mark=',', scientific=FALSE)` total iterations) using the number of lineages previously determined by Alvarez _et al._ ([@r478]) and combining replicate runs using \texttt{ClumPP} (version 1.1.2; [@r459]; greedy algorithm based on the $G^{\prime}$ statistic; `r format(clumpp.params.LST[[MODE]]$repeats, big.mark=',', scientific=FALSE)` iterations).
 We then ran $\texttt{Bayescan}$ for each species using these lineages (1:1 prior odds; `r bayescan.params.LST[[MODE]]$reps` replicates per species; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r format(bayescan.params.LST[[MODE]]$burn, big.mark=',', scientific=FALSE)` burn-in iterations; `r format(bayescan.params.LST[[MODE]]$n+bayescan.params.LST[[MODE]]$burn, big.mark=',', scientific=FALSE)` total iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations) using a suitable false discovery rate ($q \leq `r bayescan.params.LST[[MODE]]$fdr`$; [@r466]).
@@
 Solutions associated with negative values--because they secured very little genetic variation--were replaced with zeros to facilitate statistical analysis.
-We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality (using the \texttt{raptr R} package [@r465]).
+We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality (using the \texttt{raptr R} package; [@r465]).
 

### 2017-05-29 code/rmarkdown/supporting_information.Rmd
fix bug in combining results for 3rd analysis update refs and article

@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=calc.fig.height3(n_distinct(geo.surrogacy.DF$Species)), fig.cap='The relationship between the proportion of neutral genetic variation secured in a prioritizaiton and the proportion of geographic variation it also secures.
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=calc.fig.height3(n_distinct(geo.surrogacy.DF$Species)), fig.cap='The relationship between the proportion of neutral genetic variation secured in a prioritization and the proportion of geographic variation it also secures.
 Conventions are detailed in Figure S28.'}

### 2017-06-21 article/article.tex
inserted line numbers

@@
-\documentclass[11pt,]{article}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
+
+%% Some pieces required from the pandoc template
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+% Use the lineno option to display guide line numbers if required.
+% Note that the use of elements such as single-column equations
+% may affect the guide line number alignment.
+
 \usepackage[T1]{fontenc}
 \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-    \usepackage{xltxtra,xunicode}
-  \else
-    \usepackage{fontspec}
-  \fi
-  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
-  \newcommand{\euro}{€}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
-\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\usepackage[margin=1in]{geometry}
-\ifxetex
-  \usepackage[setpagesize=false, % page size defined by xetex
-              unicode=false, % unicode breaks when used with xetex
-              xetex]{hyperref}
-\else
-  \usepackage[unicode=true]{hyperref}
-\fi
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={},
-            pdftitle={Are environmental and geographic effective surrogates for genetic variation in conservation planning?},
-            colorlinks=true,
-            citecolor=blue,
-            urlcolor=blue,
-            linkcolor=magenta,
-            pdfborder={0 0 0}}
-\urlstyle{same}  % don't use monospace font for urls
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\setcounter{secnumdepth}{0}
-
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-%%% Change title format to be more compact
-\usepackage{titling}
-
-% Create subtitle command for use in maketitle
-\newcommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
-
-\setlength{\droptitle}{-2em}
-  \title{Are environmental and geographic effective surrogates for genetic
-variation in conservation planning?}
-  \pretitle{\vspace{\droptitle}\centering\huge}
-  \posttitle{\par}
-  \author{Jeffrey O. Hanson$^1$, Jonathan R. Rhodes$^2$, Cynthia Riginos$^1$, Hugh
-P. Possingham$^1$, Richard A. Fuller$^1$\\$^1$School of Biological
-Sciences, The University of Queensland, Brisbane, QLD,
-Australia\\$^2$School of Geography, Planning and Environmental
-Management, The University of Queensland, Brisbane, QLD,
-Australia\\Correspondance should be addressed to
-\href{mailto:jeffrey.hanson@uqconnect.edu.au}{jeffrey.hanson@uqconnect.edu.au}}
-  \preauthor{\centering\large\emph}
-  \postauthor{\par}
-  \predate{\centering\large\emph}
-  \postdate{\par}
-  \date{01 January 2016}
 
 % load packages
-\usepackage{amsmath,amsfonts,float,makecell,titletoc,tocloft,titlesec,natbib,pdfpages}
-\usepackage[T1]{fontenc}
-\usepackage{lmodern}
-\usepackage[utf8]{inputenc}
+\usepackage{float}
+\usepackage{listings}
+\lstset{numbers=left}
 
-% format captions
-\usepackage[labelfont={small,bf}, labelsep=space, font={small}]{caption}
+% define struts for tables
+\newcommand\T{\rule{0pt}{2.6ex}} %top strut
+\newcommand\B{\rule[-1.2ex]{0pt}{0pt}} % bottom strut
 
-% format headings
-\titleformat{\subsection}
-{\normalfont\Large\bfseries}{\thesubsection}{1em}{}
-\titleformat{\subsubsection}
-{\normalfont\large\bfseries}{\thesubsubsection}{1em}{}
-\titleformat{\paragraph}
-{\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
-\titleformat{\subparagraph}
-{\normalfont\normalsize\itshape}{\thesubparagraph}{1em}{}
+\templatetype{pnasresearcharticle}  % Choose template
 
-\titlespacing\paragraph{0pt}{12pt plus 4pt minus 2pt}{+0pt plus 2pt minus 2pt}
-\titlespacing{\subparagraph}{0pt}{6pt plus 4pt minus 2pt}{+0pt plus 2pt minus 2pt}
+\title{Environmental and geographic variables are effective surrogates for
+genetic variation in conservation planning}
 
-% format toc
-\renewcommand\cftsubsecfont{\normalfont\normalsize\bfseries}
-% \renewcommand\cftsubsubsecfont{\normalfont\normalsize\bfseries}
-% \renewcommand\cftparafont{\normalfont\normalsize\bfseries}
-% \renewcommand\cftsubparafont{\normalfont\normalsize\itshape}
+\author[a,1]{Jeffrey O. Hanson}
+\author[b]{Jonathan R. Rhodes}
+\author[a]{Cynthia Riginos}
+\author[a]{Richard A. Fuller}
 
-% make figures static
-\let\origfigure\figure
-\let\endorigfigure\endfigure
-\renewenvironment{figure}[1][2] {
-	\expandafter\origfigure\expandafter[H]
-} {
-	\endorigfigure
-}
+  \affil[a]{School of Biological Sciences, The University of Queensland, Brisbane,
+QLD, Australia}
+  \affil[b]{School of Earth and Environmental Sciences, The University of
+Queensland, Brisbane, QLD, Australia}
+
+% Please give the surname of the lead author for the running footer
+\leadauthor{Hanson}
+
+% Please add here a significance statement to explain the relevance of your work
+\significancestatement{To protect biodiversity for the long term, nature reserves and other
+protected areas need to represent a broad range of different genetic
+types. However, genetic data are expensive and time consuming to obtain.
+Here we show that freely available environmental and geographic
+variables can be used as effective surrogates for genetic data in
+conservation planning. This means that conservation planners can, with
+some confidence, design protected area systems to represent
+intra-specific genetic diversity without investing in expensive
+programmes to obtain and analyze genetic data.}
+
+\authorcontributions{JOH performed the analyses. All authors developed the study and drafted
+the manuscript.}
+
+\authordeclaration{The authors declare no conflict of interest.}
+
+\correspondingauthor{\textsuperscript{1} To whom correspondence should be addressed: E-mail: jeffrey.hanson@uqconnect.edu.au.}
+
+% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
+ \keywords{  conservation |  biodiversity |  AFLP |  genetic diversity   }
+
+\begin{abstract}
+Protected areas buffer species from anthropogenic threats and provide
+places for the processes that generate and maintain biodiversity to
+continue. However, genetic variation, the raw material for evolution, is
+difficult to capture in conservation planning, not least because genetic
+data require considerable resources to obtain and analyze. Here we show
+that freely available environmental and geographic distance variables
+are highly effective surrogates in conservation planning for
+representing adaptive and neutral intra-specific genetic variation. We
+obtained occurrence and genetic data from the IntraBioDiv project for 27
+alpine plant species collected over the European Alps using a gridded
+sampling scheme. For each species, we identified loci that were
+potentially under selection using outlier loci methods, and mapped their
+main gradients of adaptive and neutral genetic variation across the grid
+cells. We then used the cells as planning units to prioritize protected
+area acquisitions. First, we verified that the spatial patterns of
+environmental and geographic variation were correlated, respectively,
+with adaptive and neutral genetic variation. Second, we showed that
+these surrogates can predict the proportion of genetic variation secured
+in randomly generated solutions. Finally, we discovered that solutions
+based only on surrogate information secured substantial amounts of
+adaptive and neutral genetic variation. Our work paves the way for
+widespread integration of surrogates for genetic variation into
+conservation planning.
+\end{abstract}
+
+\dates{This manuscript was compiled on \today}
+\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
 
@@
 
+% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
+% You should only change this length when you've finalised the article contents.
+\verticaladjustment{-2pt}
+
 \maketitle
+\thispagestyle{firststyle}
+\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
 
-\begin{abstract}
-Insert abstract here.
-\end{abstract}
+% If your first paragraph (i.e.
+with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation.
+If this is the case, add \parshape=0 to the end of the list environment.
 
-\subsection{Introduction}\label{introduction}
+\acknow{JOH is funded by an Australian Postgraduate Award (APA) scholarship. RAF
+has an Australian Research Council Future Fellowship.}
 
-\subsection{Methods}\label{methods}
+\dropcap{P}rotected areas spearhead conservation efforts (1). They
+buffer species from anthropogenic impacts, providing places for them to
+persist. Since the resources available for conservation are limited,
+protected areas need to be sited in places that fulfill conservation
+objectives for minimal cost (2). To achieve this, conservation planning
+exercises often generate plans for entire networks of protected areas
+(prioritizations) to allocate resources cost-effectively and preserve
+broad-scale biodiversity processes (3). Intra-specific genetic variation
+is fundamental to long-term species persistence (4, 5), and as a
+consequence, there has been increasing interest in designing
+prioritizations that represent this aspect of biodiversity (6--10).
 
-\subsubsection{Study area}\label{study-area}
+Although the strength of natural selection is a continuous force,
+genetic variation is often classified as adaptive or neutral (reviewed
+in 11). Adaptive genetic variation is associated with loci that
+significantly affect fitness. Typically, such ``adaptive'' loci are
+detected due to anomalous patterns among individuals (eg. 12, 13),
+though phenotypic-genetic associations are more likely to identify
+functionally important genomic regions for individual species (eg. 14).
+By representing the full range of adaptive variation, protected areas
+can help ensure that populations with particularly beneficial
+adaptations are not lost (7, 15). In contrast, neutral genetic variation
+is associated with loci that do not significantly affect fitness, but
+instead reflect the evolutionary history of different populations. By
+representing the full range of neutral variation, protected areas can
+safeguard against the adverse effects of low genetic diversity (6). Thus
+optimally sited protected area networks would aim to represent patterns
+of both adaptive and neutral genetic variation (6, 7, 15).
 
-To address the aims of the study, we utilized species distribution and
-genomic data collected by the IntraBioDiv project (Figure S1; Meirmans
-\emph{et al.} 2011; see Gugerli \emph{et al.} 2008 for methodological
-details). This dataset was well suited for this study, because it
-provides spatially explicit genomic data for a multitude of species
-across a broad geographic area and over large environmental gradients.
-As part of this project, data for 27 alpine plant species were collected
-across the European Alps according to a grid (22.3 km $\times$ 25 km).
-Briefly, in every second cell, plant samples were collected from three
-individuals for each species present inside the cell. Samples were
-genotyped using amplified fragment length polymorphisms (AFLP; Vos
-\emph{et al.} 1995). Matrices denoting the presence/absence of
-polymorphisms at loci were constructed independently for each species.
-Thus for each second grid cell, for each of the species in the grid
-cell, the dataset contains information describing the genomic variation
-of individuals inside the grid cell. We used these grid cells as
-planning units to develop prioritizations to investigate the
-effectiveness of surrogates for the (AFLP) genetic variation.
+Recently, the use of surrogate data has been proposed to generate
+prioritizations that capture intra-specific genetic variation (16,
+17)---without needing to utilize genetic data to achieve this. Since
+adaptation is ultimately driven by selection pressures, environmental
+variables have been proposed as surrogates for adaptive genetic
+variation (eg. 16, 18). Specifically, by conserving individuals in a
+broad range of environmental conditions, we might expect to capture
+individuals with a diverse set of local adaptations, and overall capture
+a large proportion of the adaptive genetic variation present in the
+species. However, the effectiveness of this approach remains unverified.
+Neutral genetic variation arises from a reduction in gene flow between
+populations. Over the last few years, the field of phylogeography has
+predominantly focused on describing neutral genetic diversity, and so
+the landscape factors that affect neutral variation are relatively well
+understood (reviewed in 19). Potential surrogates for neutral genetic
+variation have been based on variables that predict the level of
+connectivity between different areas. For instance, according the
+isolation by distance model (IBD; 20), populations located further apart
+are predicted experience less gene flow, and in turn, share fewer
+neutral loci. Based on this idea, spreading conservation priorities
+evenly across the geographic distribution of an island network has been
+found to capture more neutral genetic variation (17). However this
+remains untested in spatially contiguous systems where connectivity is
+complicated by many additional factors (eg. land use; 21)---the typical
+situation in most conservation planning exercises.
 
-\subsubsection{Surrogate data}\label{surrogate-data}
+Here we determine whether environmental and geographic surrogates
+capture adaptive and neutral genetic variation respectively in the
+context of conservation planning. We use distribution and genomic data
+for 27 alpine plant species in the European Alps that was obtained by
+the IntraBioDiv project (22). The data was collected following a gridded
+sampling scheme which we adopted as planning units for this study. We
+show that for most species the spatial patterns of adaptive and neutral
+genetic variation correlated with environmental and geographic
+variables. We also show that for most species the level of association
+is strong enough to be operational for conservation planning. Finally,
+we demonstrate that using these surrogates results in prioritizations
+that secure a substantial proportion of intra-specific genetic
+variation.
 
-We explored the effectiveness of climatic and geographic surrogates. To
-describe variation in the location of each grid cell, we projected the
-grid into an equi-distant coordinate system (ESRI:102031) and used the
-centroid of each cell. To describe the climatic variation among each
-cell, we obtained twenty-one bioclimatic variables (approx. 1km
-resolution; Hijmans \emph{et al.} 2005). These layers and the grid cells
-were projected into an equal-area coordinate system (ESRI:102014). The
-bioclimatic variables were clipped to the extent of the grid and subject
-to a principle components analysis (PCA; Table S1). The first two
-principle components (PCs) cumulatively explained 97.2\% of the
-variation and were used to generate two new layers. We calculated the
-average value for each PC layer in each grid cell, and used these values
-to characterize the main themes of climatic variation among the grid
-cells (Figure S2).
+\section*{Results}\label{results}
+% \addcontentsline{toc}{section}{Results}
 
-\subsubsection{Outlier locus detection}\label{outlier-locus-detection}
+We detected putatively adaptive genetic variation in ten of the 27 plant
+species. We used two outlier detection methods---that solely used
+genetic data and did not use environmental data---to identify loci under
+selection. These methods returned reasonably consistent results (mean
+87.92 \% loci per species assigned the same classification \(\pm\) 8.75
+SD; Table S1). Of the species that were associated with adaptive genetic
+variation, only a small proportion of loci were classified as being
+adaptive (mean 3.01 \% loci per species \(\pm\) 1.95 SD). After
+identifying the loci showing strong signals of selection, we used
+non-metric multi-dimensional scaling (NMDS) analyses to identify the
+main gradients of adaptive (if detected) and neutral variation for each
+species. Generally, only a small number of continuous dimensions was
+needed to sufficiently describe their patterns of adaptive (\(K = 2\)
+for all species; Table S2) and neutral genetic variation
+(\(2 \leq K \leq 4\) for all species; Table S1). The resulting
+ordinations were used to construct an adaptive (if detected) and neutral
+genetic space for each species. The spatial distribution of these
+genetic spaces (Figures S1--S27) generally showed substantial spatial
+auto-correlation (eg. \emph{Cerastium uniflorum}, Figure S34;
+\emph{Dryas octopetala}, Figure S36), suggesting that the environmental
+and geographic variables have potential to be effective surrogates for
+genetic variation.
 
-To investigate the effectiveness of the surrogates for adaptive and
-neutral genetic variation, we first identified which of the sampled loci
-for each species were under selection. We used outlier locus detection
-methods (evaluated in P{é}rez-Figueroa \emph{et al.} 2010) to avoid
-circularity issues. The basic premise underpinning these analyses is
-that neutral loci are expected to have a specific level of variation
-among populations ($F_{st}$), and loci that deviate from this
-expectation should be under selection. Thus loci can be identified as
-being under selection without using environmental data. In an earlier
-study, Bothwell \emph{et al.} (2013) applied such outlier detection
-methods to one of the species in this dataset (\emph{Getiana nivalis}).
-Here, we adopted a similar methodology and applied it to each of the 27
-species in the dataset.
+We verified that the spatial patterns in environmental variation
+correlated with the broad-scale patterns in adaptive genetic variation
+for each species. We also verified that there was a correlation between
+neutral genetic variation and variation in geographic position. For each
+species, we constructed dissimilarity matrices expressing differences
+between the planning units where individuals were detected based on the
+units' (i) environmental characteristics, (ii) geographic position, and
+also the (iii) adaptive (if adaptive loci detected), and (iv) neutral
+genetic characteristics of the individuals found inside them. The
+spatial patterns of environmental variation were significantly
+correlated with the pattern of adaptive genetic variation for eight of
+the ten species associated with adaptive variation (\emph{P} \textless{}
+0.05; mean 0.09 marginal \(R^2\) \(\pm\) 0.06 SD for significant models;
+Table S3). Similarly, geographic distance between planning units was
+also correlated significantly with the spatial pattern of neutral
+genetic variation among planning units for 26 of the 27 species
+(\emph{P} \textless{} 0.05; mean 0.2 marginal \(R^2\) \(\pm\) 0.16 SD
+for significant models; Table S3). Thus, for most species, planning
+units that contained different environmental conditions contained
+individuals with different adaptive genetic characteristics, and
+planning units that were further apart tended to contain individuals
+with different neutral genetic characteristics. After verifying these
+patterns were correlated, our next step was to determine whether these
+associations were strong enough to be effective for conservation
+planning.
 
-First, we grouped individuals into genetic lineages for each species
-(Figure S2). The purpose of this step was to avoid confounding variation
-between populations with due to genetic history with variation
-attributed to selection pressures. We used the Bayesian clustering
-method implemented in $\texttt{Structure}$ (version 2.3.4; Pritchard
-\emph{et al.} 2000; Falush \emph{et al.} 2007) assuming admixture and an
-independent alleles model. For each species, we identified the most
-plausible number of populations (K = 1--3; 4 runs per K; 10 burn-in
-iterations; 10 post-burn-in iterations) using the $\Delta$K method
-(Figure S3; Evanno \emph{et al.} 2005). We aggregated the population
-membership probabilities from replicate runs using $\texttt{CLUMPP}$
-(version 1.1.2; large K greedy search method with 1000 random starts per
-K per species; Figure S4; Jakobsson and Rosenberg 2007). Individuals
-with a moderate probability of belonging to a single population
-($\geq 0.75$) were assigned accordingly (Figure S5), and subject to
-subsequent outlier locus detection methods.
+For each species, we generated a suite of 10,000 random prioritizations,
+and calculated the proportion of environmental variation and adaptive
+(if present) genetic variation that each prioritization captured. We
+then repeated this process, and calculated the proportion of variation
+in geographic position and neutral genetic variation they sampled. The
+environmental and geographic variables were moderately effective
+predictors for the genetic variation represented by the randomly
+generated prioritizations (Figures 1, S28 and S29). The relationship
+between the proportion of genetic and surrogate variation secured in the
+prioritizations varied among the species (species \(\times\) proportion
+of environmental variation interaction term:
+\({\chi^2}_{9} = {6776.94}; P 0.028\); species \(\times\) proportion of
+geographic variation interaction term:
+\({\chi^2}_{26} = 4.670631\times 10^{4}\); \(P =\) \textless{} 0.001).
+\emph{Post-hoc} analyses showed that these relationships were positive
+for all species (environmental: minimum \textit{z} = 40.34, maximum
+\(P < 0.001\); geographic: minimum \textit{z} = 23.51, maximum
+\(P < 0.001\); Table S4). After establishing that the the proportion of
+surrogate variation secured in a prioritization also predicted the
+proportion of genetic variation it secured, our final step was to
+determine if prioritizations were more effective when generated using
+targets based on the surrogate variables.
 
-We used $\texttt{Bayescan}$ (version 2.1; Foll and Gaggiotti 2008) to
-identify loci under selection for each species (Table S2). For each
-species, $\texttt{Bayescan}$ was run using the population memberships
-identified in the previous analysis (1:1 prior odds; 3 pilot runs; 40
-burn-in iterations; 10 post-burn-in iterations thinned by 1 iterations).
-Following guidelines in the $\texttt{Bayescan}$ manual, we omitted loci
-where the global frequency of the minor allele was $\geq 0.05$ to avoid
-false-positives. To ensure convergence, we ran 2 replicates per species
-using a suitable false discovery rate (FDR; $\leq 0.8$; Benjamini and
-Hochberg 1995).
+\begin{figure}
+
+{\centering \includegraphics[width=3.42in,height=2.5in]{article_files/figure-latex/unnamed-chunk-3-1.pdf}
+
+}
+
+\caption{The proportion of adaptive (a; 10 species) and neutral (b; 27 species) genetic variation represented in a suite of randomly generated prioritizations for different species as a function of the surrogate variation they also secured (lines).
+Panels (c) and (d) show the distribution of $R^2$ values for the 37 models shown in (a) and (b).}\label{fig:unnamed-chunk-3}
+\end{figure}
+
+To determine whether environmental and geographic targets could improve
+the effectiveness of prioritizations in representing genetic variation,
+we generated prioritizations using (i) ``amount targets'' reflecting the
+traditional approach of representing a certain proportion of each
+species' geographic distribution, (ii) ``amount and surrogate targets''
+where we targeted representation of the environmental and geographic
+surrogates as well as a certain proportion of each species' geographic
+distribution, and (iii) ``amount and genetic targets'' where we targeted
+representation of the directly measured genetic variation as well as a
+certain proportion of each species' geographic distribution. To
+determine if these patterns were robust, for each of the three
+combinations of the targets, we generated prioritizations under four
+scenarios: (i) single-species with equal costs, (ii) single species with
+acquisition costs, (iii) multi-species with equal costs, and (iv)
+multi-species with acquisition costs (Figure S30).
+
+\begin{figure}
+
+{\centering \includegraphics[width=3.42in,height=4in]{article_files/figure-latex/unnamed-chunk-4-1.pdf}
+
+}
+
+\caption{The proportion of adaptive and neutral genetic variation secured for each species in solutions based on different targets and scenarios.
+Box plots show the median, and the 25th and 75th percentiles.
+Whiskers show the data at 1.5 times the inter-quantile range.
+Points show data outside the whiskers.
+Red lines show the target proportion of genetic variation.}\label{fig:unnamed-chunk-4}
+\end{figure}
+
+The proportion of genetic variation secured in a prioritization depended
+on the targets used to generate the prioritization
+(\({\chi^2}_{2} = 67.81\); \(P < 0.001)\). Most notably, prioritizations
+based on the surrogate variables represented significantly more genetic
+variation than amount-based prioritizations (93.24 \% \(\pm\) 13.79 SD
+overall genetic variation secured versus 78.33 \% \(\pm\) 23.62 SD;
+\(Z_1 = 4.59\); \(P < 0.001\)), and were not distinguishable from
+prioritizations based on measured genetic data (94.7 \% \(\pm\) 3.8 SD;
+\(Z_1 = 0.56\); \(P > 0.99\)). These results indicate that environmental
+and geographic surrogates in this case far outperform traditional
+amount-based conservation planning, and perform almost as well as a
+conservation plan based directly on measured genetic data.
+
+Overall, the prioritizations tended to secure a greater proportion of
+adaptive genetic variation than neutral genetic variation (adaptive:
+mean 96.04 \% \(\pm\) 7.61 secured across all prioritizations, neutral:
+mean 71.94 \(\pm\) 24.16; \({\chi^2}_{1} = 1408.16\); \emph{P}
+\textless{} 0.001). Thus, regardless of the targets used to generate a
+prioritization, or the planning context under which the prioritization
+was generated, prioritizations tend to secure more adaptive than neutral
+genetic genetic variation.
+
+The average proportion of genetic variation secured by prioritizations
+varied under different scenarios (\({\chi^2}_{3} = 34.45\); \emph{P}
+\textless{} 0.001). Specifically, prioritizations generated under
+single-species scenario (mean 78.36 \% genetic variation secured \(\pm\)
+23.61) secured less genetic variation than those under the single
+species with opportunity costs (mean 88.89 \% genetic variation secured
+\(\pm\) 15.52; \(Z_1 = 2.94\); \(P = 0.02\)) or the multi-species with
+cost scenarios (mean 91.89 \% genetic variation secured \(\pm\) 12.31;
+\(Z_1 = 3.68\); \(P = 0.001\)). These results suggest the amount of
+genetic variation secured in a prioritization may be affected by the
+data used to represent acquisition cost.
+
+\section*{Discussion}\label{discussion}
+% \addcontentsline{toc}{section}{Discussion}
+
+We have shown that broad-scale environmental and geographic variables
+can be effective as surrogates for adaptive and neutral genetic
+variation in conservation planning. Our study is the first to show this
+using field measurements of genetic variation for a broad range of
+species. The spatial patterns of environmental and geographic surrogates
+were strongly correlated with the patterns of adaptive and neutral
+genetic variation for most species. Moreover, for most species,
+prioritizations generated using environmental and geographic targets
+secured a greater proportion of adaptive and neutral genetic variation
+than traditional conservation planning. These environmental and
+geographic surrogates were based on freely available data, and could be
+applied to any study region across the world.
+
+Our results demonstrate that environmental and geographic variables are
+effective surrogates for most species considered here (Figure 2).
+Despite this, geographic distance was a suprisingly poor surrogate for
+neutral genetic variation for a few species (eg. \emph{Gypsophila
+repens}: \(R^2\) = 0.26; \emph{Luzula alpinopilosa}: \(R^2\) = 0.28;
+\emph{Gentiana nivalis} \(R^2\) = 0.29). One explanation for this result
+is that geographic distance will be a poor surrogate for neutral genetic
+variation where spatial genetic structure is complicated by additional
+factors (21, 23). For example, some species can maintain relatively high
+connectivity between distant populations (eg. wind dispersing plants;
+22), and in some places can be disrupted by landscape features (eg.
+anthropogenically modified land 21). Overall, such species were the
+exception in our analysis, and both surrogates performed moderately well
+for most species.
+
+Our results show that environmental and geographic surrogates can be
+used to capture genetic variation in prioritizations, the next question
+is: what percentage of surrogate variation should planners target to
+preserve genetic variation? To secure at least 90 \% of the species'
+neutral genetic variation, prioritizations needed to sample 95.07 \%
+\(\pm\) 6.1 SD of the variation in geographic position among the
+planning units occupied by each species. Additionally, to secure at
+least 90 \% of each species' adaptive genetic variation, prioritizations
+needed to sample 56.43 \% \(\pm\) 7.54 SD of the environmental variation
+in planning units occupied each the species. For all species, however,
+it was possible to generate solutions that secured a large proportion of
+the surrogate variation (\textgreater{} 90 \%) and only a small
+proportion of genetic variation (\textless{} 20 \%; Figures S56--S57).
+Planners may avoid such outcomes by using both amount- and
+surrogate-based targets. Depending on the study area and the species of
+conservation interest, higher targets may be needed to increase the
+likelihood that prioritizations will secure a large proportion of the
+intra-specific genetic variation for all of the species in the planning
+exercise.
+
+Our results suggest that conservation planners can secure a reasonably
+representative sample of the intra-specific adaptive genetic variation
+using only amount-based targets (Figure 2). For most species, the
+spatial patterns of adaptive genetic variation clustered into a few
+distinct groups (eg. \emph{Carex sempervirens} and \emph{Cerastium
+uniflorum}; Figures S33 and S34 respectively). As a consequence, a
+prioritization would only need to select a single planning unit in each
+cluster to secure a representative sample. This mechanism explains why
+adaptive variation accumulated much more quickly than neutral genetic
+variation in the randomly generated prioritizations. Additionally, this
+mechanism also explains why the single-species prioritizations generated
+using only amount targets were often able to achieve high
+representation---they were essentially a random selection and so
+frequently contained planning units in different parts of the species'
+ranges (Figure 4a). Overall---and contrary to our original
+hypothesis---it seems that conventional reserve selection methods might
+often conserve species' adaptive genetic variation.
+
+There are several limitations associated with our analysis. Firstly, the
+size of the planning units we used (approx. 23 \(\times\) 25 km) is much
+larger than typically used in regional conservation planning (1--10
+km\(^2\)). We used this resolution because the genetic data were
+collected at this scale. Whilst we could have interpolated the genetic
+data to a finer resolution and used smaller planning units, this would
+have biased our analysis by artificially introducing additional spatial
+auto-correlation. Secondly, we used geographic distances as surrogates
+for neutral genetic variation, yet distances that incorporate data on
+topography may better describe connectivity in the study area. However,
+such distances often require species-specific scaling (eg. 21), and so
+cannot easily be utilized in multi-species planning exercises. Thirdly,
+we used amplified fragment length polymorphism data (AFLP; 24) to
+describe genetic variation. Whilst next-generation sequencing provides
+higher resolution genetic information than AFLP data (11), we know of no
+suitable multi-species genomic data set, and our methodology would still
+have utilized only the main gradients of the genetic variation to
+generate prioritizations in a feasible period of time. Furthermore, even
+with modern population genomic approaches, a survey can at best hope to
+identify markers that are linked to functionally adaptive variants.
+
+Broad-scale environmental and geographic variables are generally
+effective surrogates for representing adaptive and neutral genetic
+variation for most species in our study system. Moreover, data to
+calculate the surrogate variables are freely available for any location
+in the world. Careful use of such surrogates in conservation planning
+could vastly improve the chances of long-term biodiversity persistence
+for relatively little additional cost.
+
+\section*{Materials and Methods}\label{materials-and-methods}
+% \addcontentsline{toc}{section}{Materials and Methods}
+
+\subsection*{Study system}\label{study-system}
+% \addcontentsline{toc}{subsection}{Study system}
+
+We used data for 27 alpine plant species in the European Alps collected
+by the IntraBioDiv project (22; Figure 3a). All data, code, and results
+are stored in an online repository to permit replication and validation
+of this study (\url{www.github.com/jeffreyhanson/genetic-surrogates}).
+This data set has been used to explore patterns of adaptive (eg. 25) and
+neutral genetic variation (eg. 26), and the potential for species
+richness as a surrogate for genetic diversity (27). Data were collected
+using a \(20^{\prime}\) longitude by \(21^{\prime}\) latitude grid
+(approx. 22.3 \(\times\) 25 km; Figure S31). Project surveyors visited
+every second grid cell, and for each species, they collected samples
+from three individuals if any individuals were found. They genotyped
+samples using AFLPs, and constructed matrices denoting the presence of
+polymorphisms at loci for each species (mean 130.7 \(\pm\) 54.9 SD
+markers genotyped per species; for more information see 26, Table 1).
+Thus the data set contains information describing the genomic properties
+of individuals for each species over a geographic grid. We used these
+data because they comprised comparable genetic information for a range
+of species with different evolutionary and life histories collected
+using a standardized sampling scheme.
+
+\begin{figure}
+
+{\centering \includegraphics[width=3.42in]{article_files/figure-latex/unnamed-chunk-5-1.pdf}
+
+}
+
+\caption{Map of the study area showing (a) the pattern of species richness, and (b) the distribution of acquisition cost, both plotted on a quantile-based color ramp.
+Squares denote planning units.}\label{fig:unnamed-chunk-5}
+\end{figure}
+
+\subsection*{Landscape data}\label{landscape-data}
+% \addcontentsline{toc}{subsection}{Landscape data}
+
+We adopted the sampling grid used to collect data as planning units to
+develop prioritizations. To reduce computational burden, we omitted
+cells that did not contain samples (149 grid cells used in analysis). To
+examine the effect of variation in cost, we calculated the total human
+population density inside each planning unit (1 km\(^2\) resolution from
+the Global Rural-Urban Mapping Project; 28) and used this to represent
+acquisition cost (Figure 3b). All spatial and statistical analyses were
+conducted in \texttt{R} (version 3.3.2; 29).
+
+We created environmental and geographic surrogate variables for each
+species (Figure S32). To describe the geographic location of each
+planning unit (Figure S33), we projected the grid into an equi-distant
+coordinate system (Europe Equidistant Conic; \(\texttt{ESRI:102031}\)),
+calculated the centroid of each grid cell, and extracted their
+two-dimensional coordinates. To describe the environmental
+characteristics of each planning unit (Figure S34), we obtained 19
+climatic layers (\(30^{\prime\prime}\) resolution; 30), projected them
+and the planning units into an equal-area coordinate system (Europe
+Lambert Conformal Conic; \(\texttt{ESRI:102014}\)), and computed
+planning unit averages for each climatic variable. To reduce
+dimensionality, for each species, we subjected the climatic values
+associated with the planning units they were found in to a principal
+components analysis (PCA; Table S1). We used the first three principal
+components to characterize climatic variation found across the species'
+geographic distributions inside the study area (capturing 90.26 \%
+\(\pm\) 0.99 SD of the total climatic variation; Figure S2). Thus we
+constructed a two-dimensional geographic space as a potential surrogate
+for neutral genetic variation, and a three-dimensional environmental
+space as a potential surrogate for adaptive genetic variation for each
+species. In these spaces, each planning unit was associated with a
+single point. Planning units with more comparable environmental
+conditions, or higher spatial proximity in the real world, were
+associated with points that were closer together in these environmental
+or geographic spaces.
+
+\subsection*{Adaptive and neutral genetic
+data}\label{adaptive-and-neutral-genetic-data}
+% \addcontentsline{toc}{subsection}{Adaptive and neutral genetic data}
+
+To investigate the effectiveness of our putative surrogates, we first
+needed to identify which of the sampled loci were adaptive (Figure S35).
+Following best practice, we used two outlier detection methods to
+achieve this (an individual- and a population-level method; 31). The
+basic premise underpinning such methods is that neutral loci are
+expected to exhibit a certain level of variation, and loci that deviate
+from this expectation are likely to be under selection (11). The
+advantage of these methods---in contrast with environmental association
+analyses---is that they do not use environmental data, which would have
+introduced an element of circularity into our analysis and invalidated
+it. Loci identified by both outlier detection methods were treated as
+adaptive and the remainder were treated as neutral. To minimize
+false-positives, we omitted loci from both methods where the global
+frequency of the minor allele was less than 10 \% and treated them as
+neutral.
+
+The first outlier detection method involved fitting
+multinominal-Dirichlet models implemented in \texttt{BayeScan} (version
+2.1; 12). We adopted a similar methodology to Bothwell \emph{et al.}
+(25) and applied it to each of the species separately. Following their
+methodology, we initially grouped conspecifics into genetic lineages to
+further minimize false-positives (Figures S36--S62), by fitting
+admixture and correlated alleles models implemented in
+\texttt{Structure} (version 2.3.4; 32; 20 replicates per species; 5,000
+admixture burnin iterations; 300,000 burnin iterations; 400,000 total
+iterations) using the number of lineages previously determined by
+Alvarez \emph{et al.} (26) and combining replicate runs using
+\texttt{ClumPP} (version 1.1.2; 33; greedy algorithm based on the
+\(G^{\prime}\) statistic; 1,000 iterations). We then ran
+\(\texttt{Bayescan}\) for each species using these lineages (1:1 prior
+odds; 4 replicates per species; 20 pilot runs; 100,000 burn-in
+iterations; 110,000 total iterations thinned by 10 iterations) using a
+suitable false discovery rate (\(q \leq 0.1\); 34). We omitted
+individuals if their population membership was uncertain (maximum
+membership probability \(< 0.75\)).
+
+The second outlier detection method involved fitting PCAs to identify
+outlier loci (implemented in the \texttt{pcadapt R} package; 13). To
+enable comparisons between the two outlier detection methods, we used
+the same individuals in this analysis as in the \texttt{BayeScan}
+analysis. For each of the 27 species, we first imputed missing data by
+replacing missing values with the average frequency among conspecifics,
+then ran a PCA over the loci matrix, and extracted the minimum number of
+components needed to secure the over-arching population level variation
+among the loci (10 \% of the variation in loci). We then computed
+\emph{q}-values using Mahalanobis distances, and used the same false
+discovery rate used in the \texttt{BayeScan} analysis to identify loci
+under selection (using the \texttt{qvalue R} package; 35).
 
 After classifying loci as adaptive or neutral, we mapped the main
-gradients of the adaptive (if present) and neutral genetic variation for
-each species (Figures S4--S29). Note that while individuals not assigned
-to populations were omitted from the $\texttt{Bayescan}$ analysis, here
-we used all the individuals. For each species, for each type of genetic
-variation present, we used Gower distances (Gower 1971) to express the
-genomic differences between individuals (using the cluster R package to
-accommodate sparsity; Maechler \emph{et al.} 2015). These distances were
-subject to a non metric multi-dimensional scaling analysis (implemented
-in the vegan R package using K=2 and 2 random starts; Table S3; Oksanen
-\emph{et al.} 2015). The average of the ordinated values associated with
-individuals in each grid cell we used to characterize the genomic
-variation in the grid cell.
+gradients of the adaptive (if detected) and neutral genetic variation
+for each species. We discarded the population groupings, and partitioned
+species' adaptive and neutral loci into separate matrices. We applied
+NMDS (implemented in the \texttt{vegan R} package; 36) using Gower
+distances (via the \texttt{cluster R} package; 37) to derive continuous
+variables that described the main gradients of adaptive and neutral
+genetic variation separately for each species (Table S2). To ensure that
+the ordinations described a sufficient amount of the genetic variation,
+we ran successive scaling analyses with increasing dimensionality until
+a sufficient stress value was obtained (maximum stress value \(\leq\)
+0.25; 100 random starts for each analysis). Since each grid cell had
+multiple samples per species, we averaged the ordinated values for
+conspecifics in the same grid cell.
 
-\subsubsection{Prioritizations}\label{prioritizations}
+Thus we constructed an adaptive (if detected) and neutral genetic space
+for each species. For a given species, each planning unit in which it
+occurred was associated with a multi-dimensional point in the species'
+adaptive genetic space (if adaptive genetic variation was detected) and
+another multi-dimensional point in the species' neutral genetic space.
+Planning units that were closer together in these genetic spaces were
+occupied by individuals with more comparable AFLP polymorphisms. By
+spreading out conservation effort across these genetic spaces, and in
+turn selecting planning units occupied by individuals with increasingly
+different polymorphisms, prioritizations can secure more genetic
+variation.
 
-We used the unreliable representative and adequate prioritization (URAP)
-formulation of the reserve selection problem (Hanson \emph{et al.}
-2016). We treated the genomic and surrogate variables as distinct
-attribute spaces. We set the demand points for each species and its
-attribute space using the values associated with the planning units.
-First, we generated a set of single species prioritizations. For each
-species, we generated a prioritization using a 20\% amount-based target,
-a prioritization using a 20\% amount-based target and 90\% space-based
-targets for the surrogate variation, and a prioritization using a 20\%
-amount-based target and 90\% space-based targets for the genomic
-variation. Next, we generated three multi-species prioritizations using
-the same sets of targets used in the single species prioritizations.
-Finally, to understand the relationship between the surrogate-based and
-genomic-based targets, we generated a suite of prioritizations using
-increasing space-based targets between 1\% and 100\% in increments of
-49.5\%. For each prioritization we calculated the proportion of the
-adaptive (if present) and neutral genetic variation that it secured. We
-solved the reserve selection problems to within 90\% of optimality using
-the $\texttt{rapr}$ R package (Hanson \emph{et al.} 2016) and
-$\texttt{Gurobi}$ (Gurobi Optimization 2015).
+\subsection*{Prioritization method}\label{prioritization-method}
+% \addcontentsline{toc}{subsection}{Prioritization method}
 
-\subsection{Results}\label{results}
+We used the \texttt{raptr R} package to generate solutions (38). This
+toolkit can identify the cheapest set of planning units required to
+preserve both a target proportion of the species' geographic range
+(using amount-based targets) and a broad sample of variation found
+across the species' range. We used the environmental and geographic
+surrogate spaces, and the adaptive (if detected) and neutral genetic
+spaces to generate and evaluate solutions. Solutions associated with
+negative values---because they secured very little genetic
+variation---were replaced with zeros to facilitate statistical analysis.
+We solved all reserve selection problems to within 10 \% of optimality.
 
-\subsubsection{Single species
-prioritizations}\label{single-species-prioritizations}
+\subsection*{Computational and statistical
+analyses}\label{statistical-analyses}
+% \addcontentsline{toc}{subsection}{Computational and statistical
+% analyses}
 
-\subsubsection{Multi-species
-prioritisations}\label{multi-species-prioritisations}
+Our first aim was to determine if the environmental and geographic
+variables correlate with the spatial patterns of adaptive and neutral
+genetic variation. To achieve this, for each species, we created
+dissimilarity matrices using Euclidean distances and the data in the
+surrogate and genetic spaces. These matrices showed differences between
+the planning units occupied by the species in terms of the units' (i)
+geographic position, (ii) environmental characteristics, and the (iii)
+adaptive (if detected) and (iv) neutral genetic characteristics of the
+individuals inside them.
 
-\subsubsection{Pareo-frontier analysis}\label{pareo-frontier-analysis}
+We fitted maximum likelihood population-effects (MLPE) models using
+maximum likelihood to investigate correlations between the dissimilarity
+models (using the \texttt{lme4 R} package; 39). These models use random
+effects to accommodate the structure of dissimilarity matrices. For each
+species, we fitted a MLPE model to the species' dissimilarity matrices
+based on geographic position and neutral genetic variation. If adaptive
+loci were detected, we also fitted a MLPE model to the species'
+dissimilarity matrices describing environmental and adaptive genetic
+variation. All data variables were \emph{z}-transformed to improve
+convergence. To test if the surrogates explained the genetic variation,
+we compared each model to its null model using a \(\chi^2\) test, and
+applied Bonferroni corrections.
 
-\subsection{Discussion}\label{discussion}
+Our second aim was to determine if the environmental and geographic
+variables were effective surrogates for adaptive and neutral genetic
+variation. For each species, we generated 20,000 prioritizations by
+randomly selecting different combinations of planning units that the
+species occupied. For half of these random prioritizations, we
+calculated the proportion of geographic variation and neutral genetic
+variation they secured, and for the other half, we calculated the
+proportion of environmental variation and adaptive genetic variation
+they secured. The proportion of surrogate and genetic variation captured
+in a prioritization was calculated using the \texttt{raptr R} package.
 
-\subsection{Acknowledgements}\label{acknowledgements}
+We fitted two full generalized linear models (GLM) with logit links. The
+first model was fit to the proportion of adaptive genetic variation
+secured in a prioritization using the proportion of environmental
+variation also secured in the prioritization. The second model was fit
+to the proportion of neutral genetic variation secured in a
+prioritization using the proportion of geographic variation also secured
+in the prioritization. Additionally, both models contained a variable
+indicating the species for which the prioritizations were generated, and
+an interaction term. They were subjected to backwards stepwise term
+deletion routines to assess term significance. To assess the performance
+of the surrogates, we refit these models separately for each species and
+computed the Cragg and Uhler's pseudo-\(R^2\) value (using the
+\texttt{pscl R} package; 40).
 
-JOH is funded by an Australian Postgraduate Award (APA) scholarship. RAF
-has an Australian Research Council Future Fellowship. This work was
-supported by the Centre of Excellence for Environmental Decisions (CEED)
-and the Landscape Ecology and Conservation Group (LEC) at The University
-of Queensland.
+Our third aim was to determine if surrogate-based targets improved
+representation of genetic variation in prioritizations. As previously
+mentioned, we generated prioritizations using different combinations of
+targets and conservation planning scenarios. We used 20 \% amount-based
+targets for each species in all prioritizations to secure an adequate
+proportion of the species' distributions. Based on the results from our
+previous analysis, we used 97.5 \% surrogate-based targets and 90 \%
+genetic-based targets. We generated a single solution for each
+target/scenario combination, except for the ``single-species (equal
+costs) with amount-based targets'' combination for which we generated
+1,000 replicates because it had many optimal solutions. We computed the
+proportion of adaptive (if present) and neutral genetic variation
+secured for each species in the prioritizations.
 
-\subsection{References}\label{references}
+We fitted generalized linear mixed-effects models (GLMMs) with logit
+links to evaluate the prioritizations (using the \texttt{lme4 R}
+package). We fitted a full model to the proportion of genetic variation
+secured in a given prioritization. This model contained categorical
+variables indicating the targets (amount only, amount and surrogate
+targets, or amount and genetic targets), the planning scenario
+(single-species, multi-species, or multi-species with cost), the type of
+genetic variation measured (adaptive or neutral), and all interactions
+between them. Data for same species were accommodated using a random
+intercept term. The full model was subjected to a backwards stepwise
+term deletion routine to assess significance. A \emph{post-hoc} analysis
+was conducted on the minimal adequate model using Tukey contrasts with a
+Bonferroni correction (using the \texttt{multcomp R} package; 41).
 
-Benjamini, Y., Hochberg, Y. (1995) Controlling the false discovery rate:
-a practical and powerful approach to multiple testing. \emph{Journal of
-the Royal Statistical Society. Series B (Methodological)}. \textbf{57},
-289--300.
+\showmatmethods
+\showacknow
+\pnasbreak
 
-Bothwell, H., Bisbing, S., Therkildsen, N., Crawford, L., Alvarez, N.,
-Holderegger, R., Manel, S. (2013) Identifying genetic signatures of
-selection in a non-model species, alpine gentian (gentiana nivalis l.),
-using a landscape genetic approach. \emph{Conservation Genetics}.
-\textbf{14}, 467--481.
+\begin{thebibliography}{100}
+\bibitem{r440}
+Sanderson EW, Segan DB, Watson JEM (2015) Global status of and
+prospects for protection of terrestrial geophysical diversity.
+\emph{Conserv Biol} 29(3):649--656.
 
-Evanno, G., Regnaut, S., Goudet, J. (2005) Detecting the number of
-clusters of individuals using the software structure: a simulation
-study. \emph{Molecular Ecology}. \textbf{14}, 2611--2620.
+\bibitem{r255}
+Margules CR, Pressey RL (2000) Systematic conservation planning.
+\emph{Nature} 405(6783):243--253.
 
-Falush, D., Stephens, M., Pritchard, J. K. (2007) Inference of
-population structure using multilocus genotype data: dominant markers
-and null alleles. \emph{Molecular Ecology Notes}. \textbf{7}, 574--578.
+\bibitem{r9}
+Cowling RM, Pressey RL (2001) Rapid plant diversification: Planning
+for an evolutionary future. \emph{Proc Natl Acad Sci USA}
+98(10):5452--5457.
 
-Foll, M., Gaggiotti, O. (2008) A genome-scan method to identify selected
+\bibitem{r165}
+Hendry AP, et al. (2011) Evolutionary principles and their practical
+application. \emph{Evol Appl} 4(2):159--183.
+
+\bibitem{r504}
+Fournier-Level A, et al. (2016) Predicting the evolutionary dynamics
+of seasonal adaptation to novel climates in arabidopsis thaliana.
+\emph{Proc Natl Acad Sci USA} 113(20):E2812--E2821.
+
+\bibitem{r67}
+Moritz C (1999) Conservation units and translocations: Strategies for
+conserving evolutionary processes. \emph{Hereditas} 130(3):217--228.
+
+\bibitem{r3}
+Crandall KA, Bininda-Emonds ORP, Mace GM, Wayne RK (2000) Considering
+evolutionary processes in conservation biology. \emph{Trends Ecol Evol}
+15(7):290--295.
+
+\bibitem{r499}
+Rosauer D, Laffan SW, Crisp MD, Donnellan SC, Cook LG (2009)
+Phylogenetic endemism: A new approach for identifying geographical
+concentrations of evolutionary history. \emph{Mol Ecol}
+18(19):4061--4072.
+
+\bibitem{r61}
+Hendry AP, et al. (2010) Evolutionary biology in biodiversity
+science, conservation, and policy: A call to action. \emph{Evolution}
+64(5):1517--1528.
+
+\bibitem{r503}
+Carvalho SB, et al. (2017) Spatial conservation prioritization of
+biodiversity spanning the evolutionary continuum. \emph{Nat Ecol Evol}
+1(1):0151.
+
+\bibitem{r480}
+Schoville SD, et al. (2012) Adaptive genetic variation on the
+landscape: Methods and cases. \emph{Annu Rev Ecol Evol S} 43:23--43.
+
+\bibitem{r456}
+Foll M, Gaggiotti O (2008) A genome-scan method to identify selected
 loci appropriate for both dominant and codominant markers: A bayesian
-perspective. \emph{Genetics}. \textbf{180}, 977--993.
+perspective. \emph{Genetics} 180(2):977--993.
 
-Gower, J. C. (1971) A general coefficient of similarity and some of its
-properties. \emph{Biometrics}. \textbf{27}, 857--871.
+\bibitem{r487}
+Duforet-Frebourg N, Bazin E, Blum MG (2014) Genome scans for
+detecting footprints of local adaptation using a bayesian factor model.
+\emph{Mol Biol Evol} 31(9):2483--2495.
 
-Gugerli, F., Englisch, T., Niklfeld, H., Tribsch, A., Mirek, Z.,
-Ronikier, M., Zimmermann, N., Holderegger, R., Taberlet, P. (2008)
-Relationships among levels of biodiversity and the relevance of
-intraspecific diversity in conservation - a project synopsis.
-\emph{Perspectives in Plant Ecology, Evolution and Systematics}.
-\textbf{10}, 259--281.
+\bibitem{r500}
+Steiner JNAH Cynthia C AND Weber (2007) Adaptive variation in beach
+mice produced by two interacting pigmentation genes. \emph{PLoS Biol}
+5(9):1--10.
 
-Gurobi Optimization, I. (2015) Gurobi optimizer reference manual.
+\bibitem{r43}
+Sgro CM, Lowe AJ, Hoffmann AA (2011) Building evolutionary
+resilience for conserving biodiversity under climate change. \emph{Evol
+Appl} 4(2):326--337.
 
-Hanson, J. O., Rhodes, J. R., Possingham, H. P., Fuller, R. A. (2016)
-rapr: representative and adequate prioritisations in r.
-\emph{Unpublished}. \textbf{X}, XX--XX.
+\bibitem{r2}
+Carvalho SB, Brito JC, Crespo EJ, Possingham HP (2011) Incorporating
+evolutionary processes into conservation planning using species
+distribution data: A case study with the western mediterranean
+herpetofauna. \emph{Divers Distrib} 17(3):408--421.
 
-Hijmans, R. J., Cameron, S. E., Parra, J. L., Jones, P. G., Jarvis, A.
-(2005) Very high resolution interpolated climate surfaces for global
-land areas. \emph{International Journal of Climatology}. \textbf{25},
-1965--1978.
+\bibitem{r200}
+Ponce-Reyes R, Clegg SM, Carvalho SB, McDonald-Madden E, Possingham
+HP (2014) Geographical surrogates of genetic variation for selecting
+island populations for conservation. \emph{Divers Distrib}
+20(6):640--651.
 
-Jakobsson, M., Rosenberg, N. A. (2007) CLUMPP: a cluster matching and
+\bibitem{r202}
+Rouget M, Cowling RM, Pressey RL, Richardson DM (2003) Identifying
+spatial components of ecological and evolutionary processes for regional
+conservation planning in the cape floristic region, south africa.
+\emph{Divers Distrib} 9(3):191--210.
+
+\bibitem{r501}
+Avise JC (2009) Phylogeography: Retrospect and prospect. \emph{J
+Biogeogr} 36(1):3--15.
+
+\bibitem{r259}
+Wright S (1943) Isolation by distance. \emph{Genetics}
+28(2):114--138.
+
+\bibitem{r497}
+Dudaniec RY, et al. (2016) Dealing with uncertainty in landscape
+genetic resistance models: A case of three co-occurring marsupials.
+\emph{Mol Ecol} 25(2):470--486.
+
+\bibitem{r451}
+Meirmans P, Goudet J, IntraBioDiv Consortium, Gaggiotti O (2011)
+Ecology and life history affect different aspects of the population
+structure of 27 high-alpine plants. \emph{Mol Ecol} 20(15):3144--3155.
+
+\bibitem{r505}
+Fortuna MA, Albaladejo RG, Fernández L, Aparicio A, Bascompte J
+(2009) Networks of spatial genetic variation across species. \emph{Proc
+Natl Acad Sci USA} 106(45):19044--19049.
+
+\bibitem{r453}
+Vos P, et al. (1995) AFLP: a new technique for DNA fingerprinting.
+\emph{Nucleic Acids Res} 23(21):4407--4414.
+
+\bibitem{r455}
+Bothwell H, et al. (2013) Identifying genetic signatures of
+selection in a non-model species, alpine gentian (\emph{Gentiana nivalis
+L.}), using a landscape genetic approach. \emph{Conserv Genet}
+14(2):467--481.
+
+\bibitem{r478}
+Alvarez N, et al. (2009) History or ecology? Substrate type as a
+major driver of patial genetic structure in alpine plants. \emph{Ecol
+Lett} 12(7):632--640.
+
+\bibitem{r486}
+Taberlet P, et al. (2012) Genetic diversity in widespread species is
+not congruent with species richness in alpine plant communities.
+\emph{Ecol Lett} 15(12):1439--1448.
+
+\bibitem{r472}
+CIESEN, Columbia University, International Food Policy Research
+Institute, The World Bank, Centro Internacional de Agricultura Tropical
+(2011) Global rural-urban mapping project, Version 1 (GRUMP v1): Urban
+extents grid. Available at: \url{http://dx.doi.org/10.7927/H4GH9FVG}.
+
+\bibitem{r192}
+R Core Team (2014) R: A language and environment for statistical
+computing. Available at: \url{http://www.R-project.org/}.
+
+\bibitem{r33}
+Hijmans RJ, Cameron SE, Parra JL, Jones PG, Jarvis A (2005) Very
+high resolution interpolated climate surfaces for global land areas.
+\emph{Int J Climatol} 25(15):1965--1978.
+
+\bibitem{r495}
+Villemereuil P de, Frichot É, Bazin É, François O, Gaggiotti OE
+(2014) Genome scan methods against more complex models: When and how
+much should we trust them? \emph{Mol Ecol} 23(8):2006--2019.
+
+\bibitem{r462}
+Falush D, Stephens M, Pritchard JK (2007) Inference of population
+structure using multilocus genotype data: Dominant markers and null
+alleles. \emph{Mol Ecol Notes} 7(4):574--578.
+
+\bibitem{r459}
+Jakobsson M, Rosenberg NA (2007) CLUMPP: A cluster matching and
 permutation program for dealing with label switching and multimodality
-in analysis of population structure. \emph{Bioinformatics}. \textbf{23},
-1801--1806.
+in analysis of population structure. \emph{Bioinformatics}
+23(14):1801--1806.
 
-Maechler, M., Rousseeuw, P., Struyf, A., Hubert, M., Hornik, K. (2015)
-\emph{cluster: Cluster analysis basics and extensions}.
+\bibitem{r466}
+Benjamini Y, Hochberg Y (1995) Controlling the false discovery rate:
+A practical and powerful approach to multiple testing. \emph{J Roy Stat
+Soc B Met} 57(1):289--300.
 
-Meirmans, P., Goudet, J., IntraBioDiv Consortium, Gaggiotti, O. (2011)
-Ecology and life history affect different aspects of the population
-structure of 27 high-alpine plants. \emph{Molecular Ecology}.
-\textbf{20}, 3144--3155.
+\bibitem{r491}
+Storey JD, Bass AJ, Dabney A, Robinson D (2015) Qvalue: Q-value
+estimation for false discovery rate control. R package version 2.2.2.
+Available at: \url{http://github.com/jdstorey/qvalue}.
 
-Oksanen, J., Blanchet, F. G., Kindt, R., Legendre, P., Minchin, P. R.,
-O'Hara, R. B., Simpson, G. L., Solymos, P., Stevens, M. H. H., Wagner,
-H. (2015) \emph{vegan: Community ecology package}.
+\bibitem{r458}
+Oksanen J, et al. (2015) Vegan: Community ecology package. R package
+version 2.3-2. Available at:
+\url{http://CRAN.R-project.org/package=vegan}.
 
-P{é}rez-Figueroa, A., Garc{í}a-Pereira, M. J., Saura, M.,
-Rol{á}n-{Á}lvarez, E., Caballero, A. (2010) Comparing three different
-methods to detect selective loci using dominant markers. \emph{Journal
-of Evolutionary Biology}. \textbf{23}, 2267--2276.
+\bibitem{r457}
+Maechler M, Rousseeuw P, Struyf A, Hubert M, Hornik K (2015)
+Cluster: Cluster analysis basics and extensions. R package version
+2.0.3. Available at: \url{http://CRAN.R-project.org/package=cluster}.
 
-Pritchard, J. K., Stephens, M., Donnelly, P. (2000) Inference of
-population structure using multilocus genotype data. \emph{Genetics}.
-\textbf{155}, 945--959.
+\bibitem{r465}
+Hanson JO, Rhodes JR, Possingham HP, Fuller RA (XXX) raptr:
+Representative and adequate prioritization toolkit in R.
+\emph{Unpublished} XX(XX):XX--XX.
 
-Vos, P., Hogers, R., Bleeker, M., Reijans, M., Lee, T. van de, Hornes,
-M., Friters, A., Pot, J., Paleman, J., Kuiper, M., Zabeau, M. (1995)
-AFLP: a new technique for dNA fingerprinting. \emph{Nucleic Acids
-Research}. \textbf{23}, 4407--4414.
+\bibitem{r469}
+Bates D, Mächler M, Bolker B, Walker S (2015) Fitting linear
+mixed-effects models using lme4. \emph{J Stat Softw} 67(1):1--48.
 
-\includepdf[pages=-]{figures.pdf}
-\includepdf[pages=-]{supporting_information.pdf}
+\bibitem{r492}
+Jackman S (2015) pscl: Classes and Methods for R Developed in the
+Political Science Computational Laboratory, Stanford University. R
+package version 1.4.9. Available at: \url{http://pscl.stanford.edu}.
 
+\bibitem{r470}
+Hothorn T, Bretz F, Westfall P (2008) Simultaneous inference in
+general parametric models. \emph{Biometrical J} 50(3):346--363.
+
+% Bibliography
+% \bibliography{pnas-sample}
+
+\end{thebibliography}
 \end{document}

### 2017-06-21 code/rmarkdown/article.Rmd
inserted line numbers

@@
 acknowledgements: |
-  JOH is funded by an Australian Postgraduate Award (APA) scholarship.
-RAF has an Australian Research Council Future Fellowship.
-This work was supported by the Centre of Biodiversity and Conservation Science (CBCS).
+  JOH is funded by an Australian Postgraduate Award (APA) scholarship. RAF has an Australian Research Council Future Fellowship.
 keywords:
@@
 By representing the full range of neutral variation, protected areas can safeguard against the adverse effects of low genetic diversity ([@r67]).
-Thus optimally sited protected area networks would aim to represent both adaptive and neutral patterns of genetic variation ([@r67; @r3; @r43]).
+Thus optimally sited protected area networks would aim to represent patterns of both adaptive and neutral genetic variation ([@r67; @r3; @r43]).
 
-Recently, the use of surrogate data has been proposed to generate prioritizations that capture intra-specific genetic variation ([@r2; @r200])--without needing to utilize genetic data to achieve this.
+Recently, the use of surrogate data has been proposed to generate prioritizations that capture intra-specific genetic variation ([@r2; @r200])---without needing to utilize genetic data to achieve this.
 Since adaptation is ultimately driven by selection pressures, environmental variables have been proposed as surrogates for adaptive genetic variation (eg.
@@
 However this remains untested in spatially contiguous systems where connectivity is complicated by many additional factors (eg.
-land use; [@r497])--the typical situation in most conservation planning exercises.
+land use; [@r497])---the typical situation in most conservation planning exercises.
 
@@
 We detected putatively adaptive genetic variation in `r as.character(as.english(length(env.surrogacy.LST)))` of the 27 plant species.
-We used two outlier detection methods to identify loci under selection.
+We used two outlier detection methods---that solely used genetic data and did not use environmental data---to identify loci under selection.
 These methods returned reasonably consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species assigned the same classification $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD; Table S1).
@@
 The environmental and geographic variables were moderately effective predictors for the genetic variation represented by the randomly generated prioritizations (Figures 1, S28 and S29).
-The relationship between the proportion of genetic and surrogate variation secured in the prioritizations varied among the species (species $\times$ proportion of environmental variation interaction term: ${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$; species $\times$ proportion of geographic variation interaction term: ${\chi^2}_{`r round(geo.surrogacy.AOV$df[[4]],2)`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`).
+The relationship between the proportion of genetic and surrogate variation secured in the prioritizations varied among the species (species $\times$ proportion of environmental variation interaction term: ${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$; species $\times$ proportion of geographic variation interaction term: ${\chi^2}_{`r round(geo.surrogacy.AOV$df[[4]],2)`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P =$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`).
 _Post-hoc_ analyses showed that these relationships were positive for all species (environmental: minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; geographic: minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S4).
@@
 Box plots show the median, and the 25th and 75th percentiles.
-Whiskers show the data at 1.5 $\times$ the inter-quantile range.
+Whiskers show the data at 1.5 times the inter-quantile range.
 Points show data outside the whiskers.
@@
 <!-- surrogate/genetic held correlations -->
-Our results demonstrate that environmental and geographic variables are effective surrogates for most species (Figure 2).
+Our results demonstrate that environmental and geographic variables are effective surrogates for most species considered here (Figure 2).
 Despite this, geographic distance was a suprisingly poor surrogate for neutral genetic variation for a few species (eg.
 _Gypsophila repens_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gypsophila_repens')$Geographic.r2CU,2)`; _Luzula alpinopilosa_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Luzula_alpinopilosa')$Geographic.r2CU,2)`; _Gentiana nivalis_ $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gentiana_nivalis')$Geographic.r2CU,2)`).
-One explanation for this result is that geographic distance will be a poor surrogate for neutral genetic variation where connectivity is complicated by additional factors ([@r505; @r497]).
+One explanation for this result is that geographic distance will be a poor surrogate for neutral genetic variation where spatial genetic structure is complicated by additional factors ([@r505; @r497]).
 For example, some species can maintain relatively high connectivity between distant populations (eg.
@@
 For most species, the spatial patterns of adaptive genetic variation clustered into a few distinct groups (eg.
-_Carex semipervens_ and _Cerastium uniflorum_; Figures S33 and S34 respectively).
+_Carex sempervirens_ and _Cerastium uniflorum_; Figures S33 and S34 respectively).
 As a consequence, a prioritization would only need to select a single planning unit in each cluster to secure a representative sample.
 This mechanism explains why adaptive variation accumulated much more quickly than neutral genetic variation in the randomly generated prioritizations.
-Additionally, this mechanism also explains why the single-species prioritizations generated using only amount targets were often able to achieve high representation--they were essentially a random selection and so frequently contained planning units in different parts of the species' ranges (Figure 4a).
-Overall--and contrary to our original hypothesis--it seems that conventional reserve selection methods might often conserve species' adaptive genetic variation.
+Additionally, this mechanism also explains why the single-species prioritizations generated using only amount targets were often able to achieve high representation---they were essentially a random selection and so frequently contained planning units in different parts of the species' ranges (Figure 4a).
+Overall---and contrary to our original hypothesis---it seems that conventional reserve selection methods might often conserve species' adaptive genetic variation.
 
@@
 Whilst next-generation sequencing provides higher resolution genetic information than AFLP data ([@r480]), we know of no suitable multi-species genomic data set, and our methodology would still have utilized only the main gradients of the genetic variation to generate prioritizations in a feasible period of time.
+Furthermore, even with modern population genomic approaches, a survey can at best hope to identify markers that are linked to functionally adaptive variants.
 
@@
 [@r478]), and the potential for species richness as a surrogate for genetic diversity ([@r486]).
-Data was collected using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx.
+Data were collected using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx.
 22.3 $\times$ 25 km; Figure S31).
-Project members visited every second grid cell, and for each species, they collected samples from three individuals if any individuals were found.
+Project surveyors visited every second grid cell, and for each species, they collected samples from three individuals if any individuals were found.
 They genotyped samples using AFLPs, and constructed matrices denoting the presence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; for more information see [@r478], Table 1).
 Thus the data set contains information describing the genomic properties of individuals for each species over a geographic grid.
-We used this data because it comprised comparable genetic data for a range of species with different evolutionary and life histories that was collected using a standardized sampling scheme.
+We used these data because they comprised comparable genetic information for a range of species with different evolutionary and life histories collected using a standardized sampling scheme.
 
@@
 
-To investigate the effectiveness of our supposed surrogates, we first needed to identify which of the sampled loci were adaptive (Figure S35).
+To investigate the effectiveness of our putative surrogates, we first needed to identify which of the sampled loci were adaptive (Figure S35).
 Following best practice, we used two outlier detection methods to achieve this (an individual- and a population-level method; [@r495]).
 The basic premise underpinning such methods is that neutral loci are expected to exhibit a certain level of variation, and loci that deviate from this expectation are likely to be under selection ([@r480]).
-The advantage of these methods--in contrast with environmental association analyses--is that they do not use environmental data, which would have introduced an element of circularity into our analysis and invalidated it.
+The advantage of these methods---in contrast with environmental association analyses---is that they do not use environmental data, which would have introduced an element of circularity into our analysis and invalidated it.
 Loci identified by both outlier detection methods were treated as adaptive and the remainder were treated as neutral.
@@
 To enable comparisons between the two outlier detection methods, we used the same individuals in this analysis as in the \texttt{BayeScan} analysis.
-For each of the 27 species, we first imputed missing data by replacing missing values with the average frequency among conspecifics, then ran a PCA over the loci matrix, and extracted the minimum number of components needed to secure the over-arching population level variation among the loci ( `r pcadapt.params.LST[[MODE]]$min.variance.explained*100` % of the variation in loci).
+For each of the 27 species, we first imputed missing data by replacing missing values with the average frequency among conspecifics, then ran a PCA over the loci matrix, and extracted the minimum number of components needed to secure the over-arching population level variation among the loci (`r pcadapt.params.LST[[MODE]]$min.variance.explained*100` % of the variation in loci).
 We then computed _q_-values using Mahalanobis distances, and used the same false discovery rate used in the \texttt{BayeScan} analysis to identify loci under selection (using the \texttt{qvalue R} package; [@r491]).
@@
 
-We used the unreliable representative and adequate prioritization (URAP) formulation to generate solutions ([@r465]).
-This formulation can identify the cheapest set of planning units required to preserve both a target proportion of the species' geographic range (using amount-based targets) and a broad sample of variation found across the species' range.
+We used the \texttt{raptr R} package to generate solutions ([@r465]).
+This toolkit can identify the cheapest set of planning units required to preserve both a target proportion of the species' geographic range (using amount-based targets) and a broad sample of variation found across the species' range.
 We used the environmental and geographic surrogate spaces, and the adaptive (if detected) and neutral genetic spaces to generate and evaluate solutions.
-Solutions associated with negative values--because they secured very little genetic variation--were replaced with zeros to facilitate statistical analysis.
-We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality (using the \texttt{raptr R} package; [@r465]).
+Solutions associated with negative values---because they secured very little genetic variation---were replaced with zeros to facilitate statistical analysis.
+We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality.
 

### 2017-08-06 code/rmarkdown/supporting_information.Rmd
updated si

@@
       geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
-        fill='grey85', color='grey70') +
+        fill='grey85', color='grey70', size=0.01) +
       geom_polygon(data=filter(curr.FPLY, d==x), aes_string(x='long', y='lat',
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(1), fig.height=calc.fig.height(1), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[1]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+```{r, echo=FALSE, fig.width=calc.fig.width(1), fig.height=calc.fig.height(1), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[1]),'}, and the distribution of its adaptive (if detected) and neutral genetic variation in the study area.
 Each panel shows a different dimension of variation as indicated by its title.
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(2), fig.height=calc.fig.height(2), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[2]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+```{r, echo=FALSE, fig.width=calc.fig.width(2), fig.height=calc.fig.height(2), fig.cap=paste0('Distribution of intra-specific variation for \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[2]),'}.
 See Figure S1 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(3), fig.height=calc.fig.height(3), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[3]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+```{r, echo=FALSE, fig.width=calc.fig.width(3), fig.height=calc.fig.height(3), fig.cap=paste0('Distribution of intra-specific variation for \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[3]),'}.
 See Figure S1 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(4), fig.height=calc.fig.height(4), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[4]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+```{r, echo=FALSE, fig.width=calc.fig.width(4), fig.height=calc.fig.height(4), fig.cap=paste0('Distribution of intra-specific variation for \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[4]),'}.
 See Figure S1 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(5), fig.height=calc.fig.height(5), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[5]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+```{r, echo=FALSE, fig.width=calc.fig.width(5), fig.height=calc.fig.height(5), fig.cap=paste0('Distribution of intra-specific variation for \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[5]),'}.
 See Figure S1 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(6), fig.height=calc.fig.height(6), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[6]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+```{r, echo=FALSE, fig.width=calc.fig.width(6), fig.height=calc.fig.height(6), fig.cap=paste0('Distribution of intra-specific variation for \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[6]),'}.
 See Figure S1 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(7), fig.height=calc.fig.height(7), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[7]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+```{r, echo=FALSE, fig.width=calc.fig.width(7), fig.height=calc.fig.height(7), fig.cap=paste0('Distribution of intra-specific variation for \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[7]),'}.
 See Figure S1 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(8), fig.height=calc.fig.height(8), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[8]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+```{r, echo=FALSE, fig.width=calc.fig.width(8), fig.height=calc.fig.height(8), fig.cap=paste0('Distribution of intra-specific variation for \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[8]),'}.
 See Figure S1 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(9), fig.height=calc.fig.height(9), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[9]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+```{r, echo=FALSE, fig.width=calc.fig.width(9), fig.height=calc.fig.height(9), fig.cap=paste0('Distribution of intra-specific variation for \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[9]),'}.
 See Figure S1 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(10), fig.height=calc.fig.height(10), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[10]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+```{r, echo=FALSE, fig.width=calc.fig.width(10), fig.height=calc.fig.height(10), fig.cap=paste0('Distribution of intra-specific variation for \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[10]),'}.
 See Figure S1 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(11), fig.height=calc.fig.height(11), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[11]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+```{r, echo=FALSE, fig.width=calc.fig.width(11), fig.height=calc.fig.height(11), fig.cap=paste0('Distribution of intra-specific variation for \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[11]),'}.
 See Figure S1 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(12), fig.height=calc.fig.height(12), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[12]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+```{r, echo=FALSE, fig.width=calc.fig.width(12), fig.height=calc.fig.height(12), fig.cap=paste0('Distribution of intra-specific variation for \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[12]),'}.
 See Figure S1 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(13), fig.height=calc.fig.height(13), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[13]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+```{r, echo=FALSE, fig.width=calc.fig.width(13), fig.height=calc.fig.height(13), fig.cap=paste0('Distribution of intra-specific variation for \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[13]),'}.
 See Figure S1 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(14), fig.height=calc.fig.height(14), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[14]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+```{r, echo=FALSE, fig.width=calc.fig.width(14), fig.height=calc.fig.height(14), fig.cap=paste0('Distribution of intra-specific variation for \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[14]),'}.
 See Figure S1 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(15), fig.height=calc.fig.height(15), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[15]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+```{r, echo=FALSE, fig.width=calc.fig.width(15), fig.height=calc.fig.height(15), fig.cap=paste0('Distribution of intra-specific variation for \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[15]),'}.
 See Figure S1 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(16), fig.height=calc.fig.height(16), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[16]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+```{r, echo=FALSE, fig.width=calc.fig.width(16), fig.height=calc.fig.height(16), fig.cap=paste0('Distribution of intra-specific variation for \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[16]),'}.
 See Figure S1 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(17), fig.height=calc.fig.height(17), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[17]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+```{r, echo=FALSE, fig.width=calc.fig.width(17), fig.height=calc.fig.height(17), fig.cap=paste0('Distribution of intra-specific variation for \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[17]),'}.
 See Figure S1 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(18), fig.height=calc.fig.height(18), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[18]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+```{r, echo=FALSE, fig.width=calc.fig.width(18), fig.height=calc.fig.height(18), fig.cap=paste0('Distribution of intra-specific variation for \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[18]),'}.
 See Figure S1 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(19), fig.height=calc.fig.height(19), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[19]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+```{r, echo=FALSE, fig.width=calc.fig.width(19), fig.height=calc.fig.height(19), fig.cap=paste0('Distribution of intra-specific variation for \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[19]),'}.
 See Figure S1 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(20), fig.height=calc.fig.height(20), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[20]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+```{r, echo=FALSE, fig.width=calc.fig.width(20), fig.height=calc.fig.height(20), fig.cap=paste0('Distribution of intra-specific variation for \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[20]),'}.
 See Figure S1 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(21), fig.height=calc.fig.height(21), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[21]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+```{r, echo=FALSE, fig.width=calc.fig.width(21), fig.height=calc.fig.height(21), fig.cap=paste0('Distribution of intra-specific variation for \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[21]),'}.
 See Figure S1 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(22), fig.height=calc.fig.height(22), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[22]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+```{r, echo=FALSE, fig.width=calc.fig.width(22), fig.height=calc.fig.height(22), fig.cap=paste0('Distribution of intra-specific variation for \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[22]),'}.
 See Figure S1 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(23), fig.height=calc.fig.height(23), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[23]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+```{r, echo=FALSE, fig.width=calc.fig.width(23), fig.height=calc.fig.height(23), fig.cap=paste0('Distribution of intra-specific variation for \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[23]),'}.
 See Figure S1 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(24), fig.height=calc.fig.height(24), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[24]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+```{r, echo=FALSE, fig.width=calc.fig.width(24), fig.height=calc.fig.height(24), fig.cap=paste0('Distribution of intra-specific variation for \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[24]),'}.
 See Figure S1 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(25), fig.height=calc.fig.height(25), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[25]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+```{r, echo=FALSE, fig.width=calc.fig.width(25), fig.height=calc.fig.height(25), fig.cap=paste0('Distribution of intra-specific variation for \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[25]),'}.
 See Figure S1 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(26), fig.height=calc.fig.height(26), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[26]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+```{r, echo=FALSE, fig.width=calc.fig.width(26), fig.height=calc.fig.height(26), fig.cap=paste0('Distribution of intra-specific variation for \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[26]),'}.
 See Figure S1 caption for conventions.')}
@@
 
-```{r, echo=FALSE, fig.width=calc.fig.width(27), fig.height=calc.fig.height(27), fig.cap=paste0('Distribution of environmental variation found across the geographic range of \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[27]),'}, and the distribution of its adaptive and neutral genetic variation in the study area.
+```{r, echo=FALSE, fig.width=calc.fig.width(27), fig.height=calc.fig.height(27), fig.cap=paste0('Distribution of intra-specific variation for \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[27]),'}.
 See Figure S1 caption for conventions.')}
@@
   geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
-    fill='grey85', color='grey70') +
+    fill='grey85', color='grey70', size=0.01) +
   geom_polygon(data=grid.sub.FPLY, aes_string(x='long', y='lat',
@@
 Squares represent planning units.
-For a given species, planning units that it was found in are shown in bright blue.')}
+For a given species, planning units that it was found in are shown in green.')}
 ## plot map of species distributions
@@
   geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
-    fill='grey85', color='grey70') +
+    fill='grey85', color='grey70', size=0.01) +
   geom_polygon(data=spp.grid.FPLY, aes(x=long, y=lat,
@@
   node[shape = ellipse, style = filled, color = darkslategray, fillcolor = white]
-  calc_geo_as [label = 'Processing\n(Figure S60)']
-  calc_env_as [label = 'Processing\n(Figure S61)']
-  calc_genetic_as [label = 'Processing\n(Figure S62)']
+  calc_geo_as [label = 'Processing\n(Figure S33)']
+  calc_env_as [label = 'Processing\n(Figure S34)']
+  calc_genetic_as [label = 'Processing\n(Figure S35)']
 
@@
     geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
-      fill='grey85', color='grey75') +
+      fill='grey85', color='grey75', size=0.01) +
     geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat',
@@
   print(curr.GROB, vp=viewport(layout.pos.row=2, layout.pos.col=1))
+  # define colors for species
+  spp.cols <- legend.CHR[seq_len(n_distinct(curr.spp.DF$Population))]
+  names(spp.cols) <- as.character(unique(curr.spp.DF$Population))
   # add in pies
@@
     curr.lat <- ((curr.cell.DF$grid.latitude[1]-bbox['ll.lat'])/abs(bbox['ur.lat']-bbox['ll.lat']))*prop
-    # create pie chart for individual
     curr.pie <- ggplot(curr.cell.DF, aes(x="", y=Probability, fill=factor(Population))) +
@@
         name='',
-        values=legend.CHR[seq_len(n_distinct(curr.spp.DF$Population))]
+        values=spp.cols
       )
@@
     caption='Summary of non-metric multi-dimensional scaling (NMDS) analyses on genetic variation for each species.
-The analyses were used to identify the main graondients of variation in binary adaptive and neutral loci data.
+The analyses were used to identify the main gradients of variation in binary adaptive and neutral loci data.
 Bold stress values indicate NMDS analyses that have converged.'
@@
 Briefly, for each species associated with potentially adaptive loci, a model was fit to correlate dissimilarities among planning units where the species in terms of the adaptive genetic characteristics of individuals in them with dissimilarities in terms of the environmental conditions inside the planning units (denoted in the environmental column group).
-Similarly, for each species, a model was fit to correlate dissimilarities among the planning units in terms of the neutral genetic characteristics of indiviudals in them with dissimilarities in terms of the geographic position of the planning units (denoted in the geographic column group).
+Similarly, for each species, a model was fit to correlate dissimilarities among the planning units in terms of the neutral genetic characteristics of individuals in them with dissimilarities in terms of the geographic position of the planning units (denoted in the geographic column group).
 Data shows the results from comparing each model to its corresponding null model using $\\chi^2$ tests (1 degree of freedom), and the $R^2$ ($MR^2$) and conditional $R^2$ ($CR^2$) for each model.'
@@
   mutate(species=gsub('_', ' ', species, fixed=TRUE), percent_variation_explained=percent_variation_explained*100) %>%
+  mutate(species=paste0("\\textit{", species, "}")) %>%
   latex(

### 2017-08-06 code/rmarkdown/article.Rmd
updates to article per reviewers suggestions

@@
 <!-- Introduction -->
-\dropcap{P}rotected areas spearhead conservation efforts ([@r440]).
+\dropcap{P}rotected areas spearhead conservation efforts ([@r507]).
 They buffer species from anthropogenic impacts, providing places for them to persist.
@@
 Adaptive genetic variation is associated with loci that significantly affect fitness.
-Typically, such "adaptive" loci are detected due to anomalous patterns among individuals (eg.
-[@r456; @r487]), though phenotypic-genetic associations are more likely to identify functionally important genomic regions for individual species (eg.
+Typically, such "adaptive" loci are detected due to anomalous patterns among individuals (e.g.
+[@r456; @r487]), though phenotypic-genetic associations are more likely to identify functionally important genomic regions for individual species (e.g.
 [@r500]).
-By representing the full range of adaptive variation, protected areas can help ensure that populations with particularly beneficial adaptations are not lost ([@r3; @r43]).
+By representing the full range of adaptive variation, protected areas can fortify a species' capacity to persist in a range of different conditions ([@r3; @r43]).
 In contrast, neutral genetic variation is associated with loci that do not significantly affect fitness, but instead reflect the evolutionary history of different populations.
@@
 
-Recently, the use of surrogate data has been proposed to generate prioritizations that capture intra-specific genetic variation ([@r2; @r200])---without needing to utilize genetic data to achieve this.
-Since adaptation is ultimately driven by selection pressures, environmental variables have been proposed as surrogates for adaptive genetic variation (eg.
+Recently, the use of surrogate data has been proposed to generate prioritizations that capture intra-specific genetic variation ([@r2; @r506; @r200])---without needing to utilize genetic data to achieve this.
+Since adaptation is ultimately driven by selection pressures, environmental variables have been proposed as surrogates for adaptive genetic variation (e.g.
 [@r202; @r2]).
@@
 Potential surrogates for neutral genetic variation have been based on variables that predict the level of connectivity between different areas.
-For instance, according the isolation by distance model (IBD; [@r259]), populations located further apart are predicted experience less gene flow, and in turn, share fewer neutral loci.
+For instance, according the isolation by distance model (IBD; [@r259]), populations located further apart are predicted to experience less gene flow, and in turn, share fewer neutral loci.
 Based on this idea, spreading conservation priorities evenly across the geographic distribution of an island network has been found to capture more neutral genetic variation ([@r200]).
-However this remains untested in spatially contiguous systems where connectivity is complicated by many additional factors (eg.
+However this remains untested in spatially contiguous systems where connectivity is complicated by many additional factors (e.g.
 land use; [@r497])---the typical situation in most conservation planning exercises.
@@
 We use distribution and genomic data for 27 alpine plant species in the European Alps that was obtained by the IntraBioDiv project ([@r451]).
-The data was collected following a gridded sampling scheme which we adopted as planning units for this study.
-We show that for most species the spatial patterns of adaptive and neutral genetic variation correlated with environmental and geographic variables.
+The data were collected following a gridded sampling scheme which we adopted as planning units for this study.
+We show that for most species the spatial patterns of adaptive and neutral genetic variation correlated with environmental (obtained from WorldClim; [@r33]) and geographic variables.
 We also show that for most species the level of association is strong enough to be operational for conservation planning.
@@
 The resulting ordinations were used to construct an adaptive (if detected) and neutral genetic space for each species.
-The spatial distribution of these genetic spaces (Figures S1--S27) generally showed substantial spatial auto-correlation (eg.
+The spatial distribution of these genetic spaces (Figures S1--S27) generally showed substantial spatial auto-correlation (e.g.
 _Cerastium uniflorum_, Figure S34; _Dryas octopetala_, Figure S36), suggesting that the environmental and geographic variables have potential to be effective surrogates for genetic variation.
@@
 We have shown that broad-scale environmental and geographic variables can be effective as surrogates for adaptive and neutral genetic variation in conservation planning.
-Our study is the first to show this using field measurements of genetic variation for a broad range of species.
+Our study is the first to show this using field measurements of genetic variation for a broad range of plant species.
 The spatial patterns of environmental and geographic surrogates were strongly correlated with the patterns of adaptive and neutral genetic variation for most species.
 Moreover, for most species, prioritizations generated using environmental and geographic targets secured a greater proportion of adaptive and neutral genetic variation than traditional conservation planning.
-These environmental and geographic surrogates were based on freely available data, and could be applied to any study region across the world.
+These environmental and geographic surrogates were based on freely available global data sets and, with exception to a few places with poor estimates ([@r33]), could be applied to any study region across the world.
 
@@
 Our results demonstrate that environmental and geographic variables are effective surrogates for most species considered here (Figure 2).
-Despite this, geographic distance was a suprisingly poor surrogate for neutral genetic variation for a few species (eg.
+Despite this, geographic distance was a suprisingly poor surrogate for neutral genetic variation for a few species (e.g.
 _Gypsophila repens_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gypsophila_repens')$Geographic.r2CU,2)`; _Luzula alpinopilosa_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Luzula_alpinopilosa')$Geographic.r2CU,2)`; _Gentiana nivalis_ $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gentiana_nivalis')$Geographic.r2CU,2)`).
 One explanation for this result is that geographic distance will be a poor surrogate for neutral genetic variation where spatial genetic structure is complicated by additional factors ([@r505; @r497]).
-For example, some species can maintain relatively high connectivity between distant populations (eg.
-wind dispersing plants; [@r451]), and in some places can be disrupted by landscape features (eg.
-anthropogenically modified land [@r497]).
+For example, some species can maintain relatively high connectivity between distant populations (e.g.
+wind dispersing plants; [@r451]), and in some places can be disrupted by landscape features (e.g.
+anthropogenically modified land; [@r497]).
 Overall, such species were the exception in our analysis, and both surrogates performed moderately well for most species.
@@
 <!-- adaptive genetic variation secured in prioritizations-->
-Our results suggest that conservation planners can secure a reasonably representative sample of the intra-specific adaptive genetic variation using only amount-based targets (Figure 2).
-For most species, the spatial patterns of adaptive genetic variation clustered into a few distinct groups (eg.
-_Carex sempervirens_ and _Cerastium uniflorum_; Figures S33 and S34 respectively).
-As a consequence, a prioritization would only need to select a single planning unit in each cluster to secure a representative sample.
-This mechanism explains why adaptive variation accumulated much more quickly than neutral genetic variation in the randomly generated prioritizations.
-Additionally, this mechanism also explains why the single-species prioritizations generated using only amount targets were often able to achieve high representation---they were essentially a random selection and so frequently contained planning units in different parts of the species' ranges (Figure 4a).
+Our results suggest that conservation planners can secure a representative sample of the intra-specific adaptive genetic variation using only amount-based targets (Figure 2).
+This result was evident in both the single- and multi-species prioritizations.
+One explanation for the single-species prioritizations is that because the spatial patterns of adaptive variation tended to cluster into a few groups (e.g.
+_Carex sempervirens_ and _Cerastium uniflorum_; Figures S33 and S34), a random selection of planning units across a species' distribution would have had a fairly high chance of capturing planning units that contained individuals belonging to multiple clusers.
+This mechanism also explains why adaptive variation accumulated much more quickly than neutral genetic variation in the randomly generated prioritizations (Figure 1).
+On the other hand, one potential explanation for the multi-species prioritization is that because it was generated using a comprehensive set of species---each with their own habitat preferences---the solution was forced to secure each species in a range of different suitable habitats and, in turn, capture a representative sample of the species' adaptive genetic variation.
+This would also suggest that these surrogates may be robust to missing distribution data when prioritizations are generated using a comprehensive set of species.
 Overall---and contrary to our original hypothesis---it seems that conventional reserve selection methods might often conserve species' adaptive genetic variation.
@@
 We used this resolution because the genetic data were collected at this scale.
-Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis by artificially introducing additional spatial auto-correlation.
-Secondly, we used geographic distances as surrogates for neutral genetic variation, yet distances that incorporate data on topography may better describe connectivity in the study area.
-However, such distances often require species-specific scaling (eg.
-[@r497]), and so cannot easily be utilized in multi-species planning exercises.
+Whilst we could have interpolated the genetic data to a finer resolution, this would have introduced additional spatial auto-correlation and biased our analysis.
+Secondly, we used geographic distances as surrogates for neutral genetic variation.
+Although distances that incorporate data on dispersal cost may perform better (e.g.
+topograpgy), such distances often require species-specific scaling (e.g.
+[@r497]) and so cannot easily be utilized in multi-species planning exercises.
 Thirdly, we used amplified fragment length polymorphism data (AFLP; [@r453]) to describe genetic variation.
-Whilst next-generation sequencing provides higher resolution genetic information than AFLP data ([@r480]), we know of no suitable multi-species genomic data set, and our methodology would still have utilized only the main gradients of the genetic variation to generate prioritizations in a feasible period of time.
+Whilst next-generation sequencing provides higher resolution genetic information ([@r480]), we know of no suitable multi-species genomic data set and our methodology would still have utilized only the main gradients of the genetic variation to generate prioritizations in a feasible period of time.
 Furthermore, even with modern population genomic approaches, a survey can at best hope to identify markers that are linked to functionally adaptive variants.
+Fourthly, none of the species we investigated here are known to be at serious risk of extinction (IUCN Red List of Threatened Species 2017).
+Although many conservation priorities seek to preserve broad-scale patterns of biodiversity ([@r255]), future work should aim to establish if these surrogates are effective for endangered species.
 
@@
 Broad-scale environmental and geographic variables are generally effective surrogates for representing adaptive and neutral genetic variation for most species in our study system.
-Moreover, data to calculate the surrogate variables are freely available for any location in the world.
-Careful use of such surrogates in conservation planning could vastly improve the chances of long-term biodiversity persistence for relatively little additional cost.
+Although genetic data are still desperately needed to guide conservation efforts for species at serious risk of extinction, careful use of such surrogates in conservation planning could vastly improve the chances of long-term biodiversity persistence for relatively little additional cost.
 
@@
 All data, code, and results are stored in an online repository to permit replication and validation of this study (\url{www.github.com/jeffreyhanson/genetic-surrogates}).
-This data set has been used to explore patterns of adaptive (eg.
-[@r455]) and neutral genetic variation (eg.
+This data set has been used to explore patterns of adaptive (e.g.
+[@r455]) and neutral genetic variation (e.g.
 [@r478]), and the potential for species richness as a surrogate for genetic diversity ([@r486]).
@@
   geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
-    fill='grey85', color='grey70') +
+    fill='grey85', color='grey70', size=0.01) +
   geom_polygon(data=spFortify(grid.PLY), aes(x=long, y=lat,
@@
 To describe the geographic location of each planning unit (Figure S33), we projected the grid into an equi-distant coordinate system (Europe Equidistant Conic; $\texttt{ESRI:102031}$), calculated the centroid of each grid cell, and extracted their two-dimensional coordinates.
-To describe the environmental characteristics of each planning unit (Figure S34), we obtained 19 climatic layers ($30^{\prime\prime}$ resolution; [@r33]), projected them and the planning units into an equal-area coordinate system (Europe Lambert Conformal Conic; $\texttt{ESRI:102014}$), and computed planning unit averages for each climatic variable.
+To describe the environmental characteristics of each planning unit (Figure S34), we obtained 19 climatic layers ($30^{\prime\prime}$ resolution; obtained from WorldClim; [@r33]), projected them and the planning units into an equal-area coordinate system (Europe Lambert Conformal Conic; $\texttt{ESRI:102014}$), and computed planning unit averages for each climatic variable.
 To reduce dimensionality, for each species, we subjected the climatic values associated with the planning units they were found in to a principal components analysis (PCA; Table S1).
@@
 Following their methodology, we initially grouped conspecifics into genetic lineages to further minimize false-positives (Figures S36--S62), by fitting admixture and correlated alleles models implemented in \texttt{Structure} (version 2.3.4; [@r462]; `r structure.params.LST[[MODE]]$numruns` replicates per species; `r format(structure.params.LST[[MODE]]$admburnin, big.mark=',', scientific=FALSE)` admixture burnin iterations; `r format(structure.params.LST[[MODE]]$burnin, big.mark=',', scientific=FALSE)` burnin iterations; `r format(structure.params.LST[[MODE]]$numreps+structure.params.LST[[MODE]]$burnin, big.mark=',', scientific=FALSE)` total iterations) using the number of lineages previously determined by Alvarez _et al._ ([@r478]) and combining replicate runs using \texttt{ClumPP} (version 1.1.2; [@r459]; greedy algorithm based on the $G^{\prime}$ statistic; `r format(clumpp.params.LST[[MODE]]$repeats, big.mark=',', scientific=FALSE)` iterations).
-We then ran $\texttt{Bayescan}$ for each species using these lineages (1:1 prior odds; `r bayescan.params.LST[[MODE]]$reps` replicates per species; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r format(bayescan.params.LST[[MODE]]$burn, big.mark=',', scientific=FALSE)` burn-in iterations; `r format(bayescan.params.LST[[MODE]]$n+bayescan.params.LST[[MODE]]$burn, big.mark=',', scientific=FALSE)` total iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations) using a suitable false discovery rate ($q \leq `r bayescan.params.LST[[MODE]]$fdr`$; [@r466]).
+We then ran $\texttt{BayeScan}$ for each species using these lineages (1:1 prior odds; `r bayescan.params.LST[[MODE]]$reps` replicates per species; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r format(bayescan.params.LST[[MODE]]$burn, big.mark=',', scientific=FALSE)` burn-in iterations; `r format(bayescan.params.LST[[MODE]]$n+bayescan.params.LST[[MODE]]$burn, big.mark=',', scientific=FALSE)` total iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations) using a suitable false discovery rate ($q \leq `r bayescan.params.LST[[MODE]]$fdr`$; [@r466]).
 We omitted individuals if their population membership was uncertain (maximum membership probability $< `r structure.params.LST[[MODE]]$probability.threshold`$).
 
-The second outlier detection method involved fitting PCAs to identify outlier loci (implemented in the \texttt{pcadapt R} package; [@r487]).
+The second outlier detection method involved fitting PCAs to identify outlier loci (implemented in the \texttt{pcadapt R} package; [@r496]).
 To enable comparisons between the two outlier detection methods, we used the same individuals in this analysis as in the \texttt{BayeScan} analysis.

### 2017-10-18 code/rmarkdown/article.Rmd
minor tweaks to ms changes to analysis that were not pushed clean up documentation

@@
 However, genetic variation, the raw material for evolution, is difficult to capture in conservation planning, not least because genetic data require considerable resources to obtain and analyze.
-Here we show that freely available environmental and geographic distance variables are highly effective surrogates in conservation planning for representing adaptive and neutral intra-specific genetic variation.
-We obtained occurrence and genetic data from the IntraBioDiv project for 27 alpine plant species collected over the European Alps using a gridded sampling scheme.
+Here we show that freely available environmental and geographic distance variables can be highly effective surrogates in conservation planning for representing adaptive and neutral intra-specific genetic variation.
+We obtained occurrence and genetic data from the IntraBioDiv project for 27 plant species collected over the European Alps using a gridded sampling scheme.
 For each species, we identified loci that were potentially under selection using outlier loci methods, and mapped their main gradients of adaptive and neutral genetic variation across the grid cells.
@@
 acknowledgements: |
-  JOH is funded by an Australian Postgraduate Award (APA) scholarship. RAF has an Australian Research Council Future Fellowship.
+  JOH is supported by an Australian Government Research Training Program (RTP) Scholarship. RAF has an Australian Research Council Future Fellowship.
 keywords:
@@
   - genetic diversity
+  - alpine plants
 pnas_type: pnasresearcharticle
@@
 Since the resources available for conservation are limited, protected areas need to be sited in places that fulfill conservation objectives for minimal cost ([@r255]).
-To achieve this, conservation planning exercises often generate plans for entire networks of protected areas (prioritizations) to allocate resources cost-effectively and preserve broad-scale biodiversity processes ([@r9]).
-Intra-specific genetic variation is fundamental to long-term species persistence ([@r165; @r504]), and as a consequence, there has been increasing interest in designing prioritizations that represent this aspect of biodiversity ([@r67; @r3; @r499; @r61; @r503]).
+To achieve this, conservation planning exercises often generate plans for entire networks of protected areas (prioritizations) to preserve species and broad-scale biodiversity processes ([@r9]).
+After determining which species to protect (e.g.
+using evolutionary distinctiveness; [@r508]), conservation planners need to ensure that prioritizations represent the intra-specific genetic variation found within species to secure their long-term persistence ([@r3; @r63; @r61]).
+As a consequence, there has been increasing interest in designing prioritizations that fulfill this objective ([@r368; @r503; @r506; @r515; @r465]).
 
@@
 [@r500]).
-By representing the full range of adaptive variation, protected areas can fortify a species' capacity to persist in a range of different conditions ([@r3; @r43]).
+By representing the full range of adaptive variation, including adaptations of which we are currently unaware, protected areas can enhance a species' capacity to persist in a range of different conditions ([@r3; @r43]).
 In contrast, neutral genetic variation is associated with loci that do not significantly affect fitness, but instead reflect the evolutionary history of different populations.
-By representing the full range of neutral variation, protected areas can safeguard against the adverse effects of low genetic diversity ([@r67]).
-Thus optimally sited protected area networks would aim to represent patterns of both adaptive and neutral genetic variation ([@r67; @r3; @r43]).
+By representing the full range of neutral variation, protected areas can safeguard against the adverse effects of low genetic diversity ([@r5]).
+Thus optimally sited protected area networks would represent patterns of both adaptive and neutral genetic variation ([@r63]).
 
-Recently, the use of surrogate data has been proposed to generate prioritizations that capture intra-specific genetic variation ([@r2; @r506; @r200])---without needing to utilize genetic data to achieve this.
+Recently, the use of surrogate data has been proposed to generate prioritizations that capture intra-specific genetic variation ([@r2; @r506; @r200])---without needing to utilize genetic data directly.
 Since adaptation is ultimately driven by selection pressures, environmental variables have been proposed as surrogates for adaptive genetic variation (e.g.
-[@r202; @r2]).
+[@r9; @r2]).
 Specifically, by conserving individuals in a broad range of environmental conditions, we might expect to capture individuals with a diverse set of local adaptations, and overall capture a large proportion of the adaptive genetic variation present in the species.
@@
 We used two outlier detection methods---that solely used genetic data and did not use environmental data---to identify loci under selection.
-These methods returned reasonably consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species assigned the same classification $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD; Table S1).
+These methods returned reasonably consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species assigned the same classification $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD; Dataset S1).
 Of the species that were associated with adaptive genetic variation, only a small proportion of loci were classified as being adaptive (mean `r round(mean(spp.outlier.loci.props, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.loci.props, na.rm=TRUE),2)` SD).
 After identifying the loci showing strong signals of selection, we used non-metric multi-dimensional scaling (NMDS) analyses to identify the main gradients of adaptive (if detected) and neutral variation for each species.
-Generally, only a small number of continuous dimensions was needed to sufficiently describe their patterns of adaptive ($K = `r unique(na.omit(adaptive.k))`$ for all species; Table S2) and neutral genetic variation ($`r min(neutral.k)` \leq K \leq `r max(neutral.k)`$ for all species; Table S1).
+Generally, only a small number of continuous dimensions were needed to sufficiently describe their patterns of adaptive ($K = `r unique(na.omit(adaptive.k))`$ for all species; Table S1) and neutral genetic variation ($`r min(neutral.k)` \leq K \leq `r max(neutral.k)`$ for all species; Table S1).
 The resulting ordinations were used to construct an adaptive (if detected) and neutral genetic space for each species.
-The spatial distribution of these genetic spaces (Figures S1--S27) generally showed substantial spatial auto-correlation (e.g.
-_Cerastium uniflorum_, Figure S34; _Dryas octopetala_, Figure S36), suggesting that the environmental and geographic variables have potential to be effective surrogates for genetic variation.
+The spatial distribution of these genetic spaces (Figs.
+S1--S27) generally showed substantial spatial auto-correlation (e.g.
+_Cerastium uniflorum_ and _Dryas octopetala_, Figs.
+S6 and S8), suggesting that the proposed surrogate variables have the potential to be effective for conservation planning.
 
@@
 For each species, we constructed dissimilarity matrices expressing differences between the planning units where individuals were detected based on the units' (i) environmental characteristics, (ii) geographic position, and also the (iii) adaptive (if adaptive loci detected), and (iv) neutral genetic characteristics of the individuals found inside them.
-The spatial patterns of environmental variation were significantly correlated with the pattern of adaptive genetic variation for `r as.character(as.english(sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)))` of the `r as.character(as.english(sum(!is.na(correlation.model.summary.DF$Environmental.corr.P))))` species associated with adaptive variation (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models; Table S3).
-Similarly, geographic distance between planning units was also correlated significantly with the spatial pattern of neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` of the `r sum(!is.na(correlation.model.summary.DF$Geographic.corr.P))` species (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models; Table S3).
+The spatial patterns of environmental variation were significantly correlated with the patterns of adaptive genetic variation for `r as.character(as.english(sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)))` of the `r as.character(as.english(sum(!is.na(correlation.model.summary.DF$Environmental.corr.P))))` species associated with adaptive variation (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models; Table S2).
+Similarly, geographic distance between planning units was also correlated significantly with the spatial patterns of neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` of the `r sum(!is.na(correlation.model.summary.DF$Geographic.corr.P))` species (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models; Table S2).
 Thus, for most species, planning units that contained different environmental conditions contained individuals with different adaptive genetic characteristics, and planning units that were further apart tended to contain individuals with different neutral genetic characteristics.
-After verifying these patterns were correlated, our next step was to determine whether these associations were strong enough to be effective for conservation planning.
+After verifying these correlations, our next step was to determine whether they were strong enough to be useful for conservation planning.
 
@@
 We then repeated this process, and calculated the proportion of variation in geographic position and neutral genetic variation they sampled.
-The environmental and geographic variables were moderately effective predictors for the genetic variation represented by the randomly generated prioritizations (Figures 1, S28 and S29).
-The relationship between the proportion of genetic and surrogate variation secured in the prioritizations varied among the species (species $\times$ proportion of environmental variation interaction term: ${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$; species $\times$ proportion of geographic variation interaction term: ${\chi^2}_{`r round(geo.surrogacy.AOV$df[[4]],2)`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P =$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`).
-_Post-hoc_ analyses showed that these relationships were positive for all species (environmental: minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; geographic: minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S4).
-After establishing that the the proportion of surrogate variation secured in a prioritization also predicted the proportion of genetic variation it secured, our final step was to determine if prioritizations were more effective when generated using targets based on the surrogate variables.
+The environmental and geographic variables were moderately effective predictors for the genetic variation represented by the randomly generated prioritizations (Figs.
+1, S28--S29).
+The relationship between the proportion of genetic and surrogate variation secured in the prioritizations varied among the species (species $\times$ proportion of environmental variation interaction term: ${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P = `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$; species $\times$ proportion of geographic variation interaction term: ${\chi^2}_{`r round(geo.surrogacy.AOV$df[[4]],2)`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`).
+_Post-hoc_ analyses showed that these relationships were positive for all species (environmental: minimum \textit{Z} = `r round(max(env.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; geographic: minimum \textit{Z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3).
+After establishing that the proportion of surrogate variation secured in a prioritization also predicted the proportion of genetic variation it secured, our final step was to determine if prioritizations generated using targets to represent variation in the surrogate variables were more effective at representing intra-specific genetic variation.
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=6, fig.height=4.5, out.width='3.42in', out.height='2.5in', fig.cap=paste0('The proportion of adaptive (a; ',sum(!is.na(correlation.model.summary.DF$Environmental.corr.P)),' species) and neutral (b; ',sum(!is.na(correlation.model.summary.DF$Geographic.corr.P)),' species) genetic variation represented in a suite of randomly generated prioritizations for different species as a function of the surrogate variation they also secured (lines).
+```{r figure-1, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=6, fig.height=4.5, out.width='3.42in', out.height='2.5in', fig.cap=paste0('The proportion of adaptive (a; ',sum(!is.na(correlation.model.summary.DF$Environmental.corr.P)),' species) and neutral (b; ',sum(!is.na(correlation.model.summary.DF$Geographic.corr.P)),' species) genetic variation represented in a suite of randomly generated prioritizations for different species as a function of the surrogate variation they also secured (lines).
 Panels (c) and (d) show the distribution of $R^2$ values for the ',sum(!is.na(correlation.model.summary.DF$Environmental.corr.P)) + sum(!is.na(correlation.model.summary.DF$Geographic.corr.P)),' models shown in panels (a) and (b).')}
@@
   coord_cartesian(xlim=c(0, 1)) +
+  scale_x_continuous(limits=c(0,1), labels=function(x) {x*100}) +
+  scale_x_continuous(limits=c(0,1), labels=function(x) {x*100}) +
   scale_x_continuous(limits=c(0,1), labels=function(x) {x*100}) +
@@
 To determine whether environmental and geographic targets could improve the effectiveness of prioritizations in representing genetic variation, we generated prioritizations using (i) "amount targets" reflecting the traditional approach of representing a certain proportion of each species' geographic distribution, (ii) "amount and surrogate targets" where we targeted representation of the environmental and geographic surrogates as well as a certain proportion of each species' geographic distribution, and (iii) "amount and genetic targets" where we targeted representation of the directly measured genetic variation as well as a certain proportion of each species' geographic distribution.
-To determine if these patterns were robust, for each of the three combinations of the targets, we generated prioritizations under four scenarios: (i) single-species with equal costs, (ii) single species with acquisition costs, (iii) multi-species with equal costs, and (iv) multi-species with acquisition costs (Figure S30).
+To determine if these patterns were robust, for each of the three combinations of the targets, we generated prioritizations under four scenarios: (i) single-species with equal costs, (ii) single species with acquisition costs, (iii) multi-species with equal costs, and (iv) multi-species with acquisition costs (Figs.
+2 and S30).
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=(5*0.9), fig.height=(7.1*0.9), out.width='3.42in', out.height='4in', fig.cap='The proportion of adaptive and neutral genetic variation secured for each species in solutions based on different targets and scenarios.
+```{r figure-2, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=(5*0.9), fig.height=(7.1*0.9), out.width='3.42in', out.height='4in', fig.cap='The proportion of adaptive and neutral genetic variation secured for each species in solutions based on different targets and scenarios.
 Box plots show the median, and the 25th and 75th percentiles.
@@
 scenario.SDF <- scenario.DF %>%
+  mutate(Context = gsub('opportunity', 'acquisition', Context, fixed = TRUE)) %>%
   group_by(Prioritisation, Context, Metric, Species) %>%
@@
 
-The proportion of genetic variation secured in a prioritization depended on the targets used to generate the prioritization (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$p.value[[2]])`)$.
+The proportion of genetic variation secured in a given prioritization was not found to depend on any interaction between the planning scenario or targets used to generate the prioritization, or the type of genetic variation measured (${\chi^2}_{`r scenario.ANOVA.1.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.1.DF$p.value[[2]])`$).
+The proportion of genetic variation secured in a prioritization did depend, however, on the targets used to generate the prioritization (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$p.value[[2]])`$).
 Most notably, prioritizations based on the surrogate variables represented significantly more genetic variation than amount-based prioritizations (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` SD overall genetic variation secured versus `r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)` SD; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$), and were not distinguishable from prioritizations based on measured genetic data (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` SD; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).
-These results indicate that environmental and geographic surrogates in this case far outperform traditional amount-based conservation planning, and perform almost as well as a conservation plan based directly on measured genetic data.
+These results indicate that environmental and geographic surrogates in this case far outperform traditional amount-based conservation planning, and perform almost as well as a conservation plan based directly on genetic data.
 
-Overall, the prioritizations tended to secure a greater proportion of adaptive genetic variation than neutral genetic variation (adaptive: mean `r round(mean(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` secured across all prioritizations, neutral: mean `r round(mean(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)` $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)`; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.3.DF$p.value[[2]])`).
+Overall, the prioritizations tended to secure a greater proportion of adaptive genetic variation than neutral genetic variation (adaptive: mean `r round(mean(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` secured across all prioritizations, neutral: mean `r round(mean(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)`; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.3.DF$p.value[[2]])`).
 Thus, regardless of the targets used to generate a prioritization, or the planning context under which the prioritization was generated, prioritizations tend to secure more adaptive than neutral genetic genetic variation.
@@
 The average proportion of genetic variation secured by prioritizations varied under different scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.2.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.2.DF$p.value[[2]])`).
-Specifically, prioritizations generated under single-species scenario (mean `r round(mean(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held*100),2)`) secured less genetic variation than those under the single species with opportunity costs (mean `r round(mean(filter(scenario.sub.DF,Context=='Single-species\n(opportunity costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Single-species\n(opportunity costs)')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$) or the multi-species with cost scenarios (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[3]])`$).
-These results suggest the amount of genetic variation secured in a prioritization may be affected by the data used to represent acquisition cost.
+Specifically, prioritizations generated under single-species scenario with equal costs (mean `r round(mean(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held*100),2)`) secured less genetic variation than those under the single species with acquisition costs (mean `r round(mean(filter(scenario.sub.DF,Context=='Single-species\n(opportunity costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Single-species\n(opportunity costs)')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$) or the multi-species with cost scenarios (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[3]])`$).
+These results suggest the proportion of genetic variation secured in a prioritization may be sensitive to acquisition cost.
 
@@
 We have shown that broad-scale environmental and geographic variables can be effective as surrogates for adaptive and neutral genetic variation in conservation planning.
-Our study is the first to show this using field measurements of genetic variation for a broad range of plant species.
-The spatial patterns of environmental and geographic surrogates were strongly correlated with the patterns of adaptive and neutral genetic variation for most species.
-Moreover, for most species, prioritizations generated using environmental and geographic targets secured a greater proportion of adaptive and neutral genetic variation than traditional conservation planning.
-These environmental and geographic surrogates were based on freely available global data sets and, with exception to a few places with poor estimates ([@r33]), could be applied to any study region across the world.
+Our study is the first to investigate this using field measurements of genetic variation for a suite of plant species across an eco-region ([@r478]).
+For most species, the proportion of surrogate variation captured in a prioritization predicted the proportion of genetic variation that was also captured in the prioritization.
+Moreover, for most species, prioritizations generated using geographic targets secured a greater proportion of neutral genetic variation than traditional conservation planning methods.
+These environmental and geographic surrogates were based on freely available data sets and, with the exception of a few places with poor climatic data ([@r33]), could be applied to any study region across the world.
 
 <!-- surrogate/genetic held correlations -->
-Our results demonstrate that environmental and geographic variables are effective surrogates for most species considered here (Figure 2).
-Despite this, geographic distance was a suprisingly poor surrogate for neutral genetic variation for a few species (e.g.
-_Gypsophila repens_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gypsophila_repens')$Geographic.r2CU,2)`; _Luzula alpinopilosa_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Luzula_alpinopilosa')$Geographic.r2CU,2)`; _Gentiana nivalis_ $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gentiana_nivalis')$Geographic.r2CU,2)`).
-One explanation for this result is that geographic distance will be a poor surrogate for neutral genetic variation where spatial genetic structure is complicated by additional factors ([@r505; @r497]).
+Our results demonstrate that environmental and geographic variables are effective surrogates for most species considered here (Fig.
+2).
+Despite this, geographic distance was a surprisingly poor surrogate for neutral genetic variation for a few species (e.g.
+_Gentiana nivalis_ $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gentiana_nivalis')$Geographic.r2CU,2)`; _Gypsophila repens_ $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gypsophila_repens')$Geographic.r2CU,2)`; _Luzula alpinopilosa_ $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Luzula_alpinopilosa')$Geographic.r2CU,2)`).
+One explanation for this result is that geographic distance will be a poor surrogate for neutral genetic variation where spatial genetic structure is complicated by additional factors ([@r505; @r506; @r497]).
 For example, some species can maintain relatively high connectivity between distant populations (e.g.
@@
 Additionally, to secure at least `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` % of each species' adaptive genetic variation, prioritizations needed to sample `r round(mean(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` SD of the environmental variation in planning units occupied each the species.
-For all species, however, it was possible to generate solutions that secured a large proportion of the surrogate variation (> 90 %) and only a small proportion of genetic variation (< 20 %; Figures S56--S57).
+For all species, however, it was possible to generate solutions that secured a large proportion of the surrogate variation (> 90 %) and only a small proportion of genetic variation (< 20 %; Figs.
+S28--S29).
 Planners may avoid such outcomes by using both amount- and surrogate-based targets.
@@
 <!-- adaptive genetic variation secured in prioritizations-->
-Our results suggest that conservation planners can secure a representative sample of the intra-specific adaptive genetic variation using only amount-based targets (Figure 2).
-This result was evident in both the single- and multi-species prioritizations.
-One explanation for the single-species prioritizations is that because the spatial patterns of adaptive variation tended to cluster into a few groups (e.g.
-_Carex sempervirens_ and _Cerastium uniflorum_; Figures S33 and S34), a random selection of planning units across a species' distribution would have had a fairly high chance of capturing planning units that contained individuals belonging to multiple clusers.
-This mechanism also explains why adaptive variation accumulated much more quickly than neutral genetic variation in the randomly generated prioritizations (Figure 1).
+Our results suggest that conservation planners can secure a representative sample of intra-specific adaptive genetic variation using conventional reserve selections without needing surrogate or genetic data (Fig.
+2).
+This finding was evident in both the single- and multi-species prioritizations.
+One explanation for the single-species prioritizations is that because the spatial patterns of adaptive variation tended to cluster into a few main groups (e.g.
+_Carex sempervirens_ and _Cerastium uniflorum_; Figs.
+S5--S6), a random selection of planning units across a species' geographic distribution would have a fairly high chance of capturing individuals that belonged to several of the main genetic groups.
+This mechanism also explains why adaptive variation accumulated much more quickly than neutral genetic variation in the randomly generated prioritizations (Fig.
+1).
 On the other hand, one potential explanation for the multi-species prioritization is that because it was generated using a comprehensive set of species---each with their own habitat preferences---the solution was forced to secure each species in a range of different suitable habitats and, in turn, capture a representative sample of the species' adaptive genetic variation.
-This would also suggest that these surrogates may be robust to missing distribution data when prioritizations are generated using a comprehensive set of species.
-Overall---and contrary to our original hypothesis---it seems that conventional reserve selection methods might often conserve species' adaptive genetic variation.
+If future studies on other taxa and biomes verify that conventional reserve selection methods do indeed conserve intra-specific adaptive genetic variation, this finding could have large implications for conservation planning.
 
@@
 Firstly, the size of the planning units we used (approx.
-23 $\times$ 25 km) is much larger than typically used in regional conservation planning (1--10 km$^2$).
+20 $\times$ 22.5 km) is larger than typically used in regional conservation planning.
 We used this resolution because the genetic data were collected at this scale.
@@
 Although distances that incorporate data on dispersal cost may perform better (e.g.
-topograpgy), such distances often require species-specific scaling (e.g.
+topography), such distances often require species-specific scaling (e.g.
 [@r497]) and so cannot easily be utilized in multi-species planning exercises.
@@
 Furthermore, even with modern population genomic approaches, a survey can at best hope to identify markers that are linked to functionally adaptive variants.
-Fourthly, none of the species we investigated here are known to be at serious risk of extinction (IUCN Red List of Threatened Species 2017).
-Although many conservation priorities seek to preserve broad-scale patterns of biodiversity ([@r255]), future work should aim to establish if these surrogates are effective for endangered species.
+Fourthly, to our knowledge, none of the species we investigated here are at serious risk of extinction ([@r510; @r511]).
+Since anthropogenic processes can alter patterns of species' genetic variation (e.g.
+[@r497]), future work should aim to establish if these surrogates are effective for imperiled species.
+Finally, conventional conservation planning exercises are sensitive to the quality and completeness of the underlying distribution data ([@r512]), and it is likely that the effectiveness of these surrogates are also sensitive to data quality.
+For instance, prioritizations generated using distribution data that omitted populations in specific habitats may fail to represent certain adaptations.
+Recent advances in distribution modeling have provided a wealth of data for conservation planners ([@r513]), and so they should endeavor to obtain high quality data where possible.
+Nevertheless, our findings pave the way for the widespread integration of evolutionary processes at the intra-specific level into reserve selection.
 
 <!--  conclusion -->
-Broad-scale environmental and geographic variables are generally effective surrogates for representing adaptive and neutral genetic variation for most species in our study system.
-Although genetic data are still desperately needed to guide conservation efforts for species at serious risk of extinction, careful use of such surrogates in conservation planning could vastly improve the chances of long-term biodiversity persistence for relatively little additional cost.
+Genetic data provide a broad range of insights into a species' persistence and are fundamental to manging certain kinds of conservation problems ([@r61]).
+However, conservation interventions often need to be implemented urgently before the highest quality data are available ([@r509]).
+We found that broad-scale environmental and geographic variables to be effective surrogates for representing adaptive and neutral genetic variation for most species in our study system.
+We call for further studies to examine these surrogates in other taxa and biomes.
+In cases where genetic data are not available, careful use of such surrogates in conservation planning could vastly improve the chances of long-term biodiversity persistence for relatively little additional cost.
 
 Materials and Methods {.unnumbered}
-===================================
+========================
 
@@
 
-We used data for 27 alpine plant species in the European Alps collected by the IntraBioDiv project ([@r451]; Figure 3a).
-All data, code, and results are stored in an online repository to permit replication and validation of this study (\url{www.github.com/jeffreyhanson/genetic-surrogates}).
+We used data for 27 alpine plant species in the European Alps collected by the IntraBioDiv project ([@r451]; Fig.
+3a).
 This data set has been used to explore patterns of adaptive (e.g.
@@
 [@r478]), and the potential for species richness as a surrogate for genetic diversity ([@r486]).
-Data were collected using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx.
-22.3 $\times$ 25 km; Figure S31).
+Data were collected using a $20^{\prime}$ longitude by $12^{\prime}$ latitude grid (approx.
+20 $\times$ 22.5 km; Fig.
+S31).
 Project surveyors visited every second grid cell, and for each species, they collected samples from three individuals if any individuals were found.
-They genotyped samples using AFLPs, and constructed matrices denoting the presence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; for more information see [@r478], Table 1).
+They genotyped samples using AFLPs, and constructed matrices denoting the presence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; for more information see [@r478]).
 Thus the data set contains information describing the genomic properties of individuals for each species over a geographic grid.
 We used these data because they comprised comparable genetic information for a range of species with different evolutionary and life histories collected using a standardized sampling scheme.
+To permit replication and validation of this study, all of our data, code, and results are stored in an online repository (\url{www.github.com/jeffreyhanson/genetic-surrogates}) and are also available under the digital object identifier (DOI) \texttt{10.5281/zenodo.843625}.
+All spatial and statistical analyses were conducted in \texttt{R} (version `r paste0(R.version$major, '.', R.version$minor)`; [@r192]).
 
-```{r, echo=FALSE, fig.align='center', fig.width=6, fig.height=1.5, out.width='3.42in', height='2.5in', fig.cap="Map of the study area showing (a) the pattern of species richness, and (b) the distribution of acquisition cost, both plotted on a quantile-based color ramp.
+```{r figure-3, echo=FALSE, fig.align='center', fig.width=6, fig.height=1.5, out.width='3.42in', height='2.5in', fig.cap="Map of the study area showing (a) the pattern of species richness, and (b) the distribution of acquisition cost depicted using a quantile-based color ramp.
 Squares denote planning units."}
@@
 We adopted the sampling grid used to collect data as planning units to develop prioritizations.
-To reduce computational burden, we omitted cells that did not contain samples (`r nrow(grid.DF)` grid cells used in analysis).
-To examine the effect of variation in cost, we calculated the total human population density inside each planning unit (1 km$^2$ resolution from the Global Rural-Urban Mapping Project; [@r472]) and used this to represent acquisition cost (Figure 3b).
-All spatial and statistical analyses were conducted in \texttt{R} (version `r paste0(R.version$major, '.', R.version$minor)`; [@r192]).
+Of the total 388 cells in the grid, we used the `r nrow(grid.DF)` cells that contained samples for subsequent analysis.
+We calculated the total human population density inside each planning unit (1 km$^2$ resolution from the Global Rural-Urban Mapping Project; [@r472]) and used this to represent acquisition cost (Fig.
+3b).
 
-We created environmental and geographic surrogate variables for each species (Figure S32).
-To describe the geographic location of each planning unit (Figure S33), we projected the grid into an equi-distant coordinate system (Europe Equidistant Conic; $\texttt{ESRI:102031}$), calculated the centroid of each grid cell, and extracted their two-dimensional coordinates.
-To describe the environmental characteristics of each planning unit (Figure S34), we obtained 19 climatic layers ($30^{\prime\prime}$ resolution; obtained from WorldClim; [@r33]), projected them and the planning units into an equal-area coordinate system (Europe Lambert Conformal Conic; $\texttt{ESRI:102014}$), and computed planning unit averages for each climatic variable.
-To reduce dimensionality, for each species, we subjected the climatic values associated with the planning units they were found in to a principal components analysis (PCA; Table S1).
-We used the first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components to characterize climatic variation found across the species' geographic distributions inside the study area (capturing `r round(mean(spp.pca.DF[[2]]*100),2)` % $\pm$ `r round(sd(spp.pca.DF[[2]]*100),2)` SD of the total climatic variation; Figure S2).
+We created environmental and geographic surrogate variables for each species (Fig.
+S32).
+To describe the geographic location of each planning unit (Fig.
+S33), we projected the grid into an equidistant coordinate system (Europe Equidistant Conic; $\texttt{ESRI:102031}$), calculated the centroid of each grid cell, and extracted their two-dimensional coordinates.
+To describe the environmental characteristics of each planning unit (Fig.
+S34), we obtained 19 climatic layers ($30^{\prime\prime}$ resolution; obtained from WorldClim; [@r33]), projected them and the planning units into an equal-area coordinate system (Europe Lambert Conformal Conic; $\texttt{ESRI:102014}$), and computed planning unit averages for each climatic variable.
+To reduce dimensionality, for each species, we subjected the climatic values associated with the planning units they were found in to a principal components analysis (PCA).
+We used the first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components to characterize climatic variation found across the species' geographic distributions inside the study area (mean `r round(mean(spp.pca.DF[[2]]*100),2)` % variation described $\pm$ `r round(sd(spp.pca.DF[[2]]*100),2)` SD of the total climatic variation; Table S4).
 Thus we constructed a two-dimensional geographic space as a potential surrogate for neutral genetic variation, and a `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))`-dimensional environmental space as a potential surrogate for adaptive genetic variation for each species.
@@
 
-To investigate the effectiveness of our putative surrogates, we first needed to identify which of the sampled loci were adaptive (Figure S35).
-Following best practice, we used two outlier detection methods to achieve this (an individual- and a population-level method; [@r495]).
+To investigate the effectiveness of our surrogates, we first needed to identify which of the sampled loci were adaptive (Fig.
+S35).
+Following recommended practices, we used two outlier detection methods to achieve this (an individual- and a population-level method; [@r495]).
 The basic premise underpinning such methods is that neutral loci are expected to exhibit a certain level of variation, and loci that deviate from this expectation are likely to be under selection ([@r480]).
-The advantage of these methods---in contrast with environmental association analyses---is that they do not use environmental data, which would have introduced an element of circularity into our analysis and invalidated it.
+The advantage of these methods---in contrast with environmental association analyses---is that they do not use environmental data, which would have introduced an element of circularity into our analysis.
 Loci identified by both outlier detection methods were treated as adaptive and the remainder were treated as neutral.
@@
 We adopted a similar methodology to Bothwell _et al._ ([-@r455]) and applied it to each of the species separately.
-Following their methodology, we initially grouped conspecifics into genetic lineages to further minimize false-positives (Figures S36--S62), by fitting admixture and correlated alleles models implemented in \texttt{Structure} (version 2.3.4; [@r462]; `r structure.params.LST[[MODE]]$numruns` replicates per species; `r format(structure.params.LST[[MODE]]$admburnin, big.mark=',', scientific=FALSE)` admixture burnin iterations; `r format(structure.params.LST[[MODE]]$burnin, big.mark=',', scientific=FALSE)` burnin iterations; `r format(structure.params.LST[[MODE]]$numreps+structure.params.LST[[MODE]]$burnin, big.mark=',', scientific=FALSE)` total iterations) using the number of lineages previously determined by Alvarez _et al._ ([@r478]) and combining replicate runs using \texttt{ClumPP} (version 1.1.2; [@r459]; greedy algorithm based on the $G^{\prime}$ statistic; `r format(clumpp.params.LST[[MODE]]$repeats, big.mark=',', scientific=FALSE)` iterations).
+Following their methodology, we initially grouped conspecifics into genetic lineages to further minimize false-positives (Figs.
+S36--S62), by fitting admixture and correlated alleles models implemented in \texttt{Structure} (version 2.3.4; [@r462]; `r structure.params.LST[[MODE]]$numruns` replicates per species; `r format(structure.params.LST[[MODE]]$admburnin, big.mark=',', scientific=FALSE)` admixture burnin iterations; `r format(structure.params.LST[[MODE]]$burnin, big.mark=',', scientific=FALSE)` burnin iterations; `r format(structure.params.LST[[MODE]]$numreps+structure.params.LST[[MODE]]$burnin, big.mark=',', scientific=FALSE)` total iterations) using the number of lineages previously determined by Alvarez _et al._ ([@r478]) and combining replicate runs using \texttt{ClumPP} (version 1.1.2; [@r459]; greedy algorithm based on the $G^{\prime}$ statistic; `r format(clumpp.params.LST[[MODE]]$repeats, big.mark=',', scientific=FALSE)` iterations).
 We then ran $\texttt{BayeScan}$ for each species using these lineages (1:1 prior odds; `r bayescan.params.LST[[MODE]]$reps` replicates per species; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r format(bayescan.params.LST[[MODE]]$burn, big.mark=',', scientific=FALSE)` burn-in iterations; `r format(bayescan.params.LST[[MODE]]$n+bayescan.params.LST[[MODE]]$burn, big.mark=',', scientific=FALSE)` total iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations) using a suitable false discovery rate ($q \leq `r bayescan.params.LST[[MODE]]$fdr`$; [@r466]).
@@
 We used the \texttt{raptr R} package to generate solutions ([@r465]).
-This toolkit can identify the cheapest set of planning units required to preserve both a target proportion of the species' geographic range (using amount-based targets) and a broad sample of variation found across the species' range.
+This toolkit can identify the cheapest set of planning units required to preserve both a target proportion of the species' geographic range (using amount-based targets) and a target proportion of intra-specific variation (using space-based targets).
 We used the environmental and geographic surrogate spaces, and the adaptive (if detected) and neutral genetic spaces to generate and evaluate solutions.
 Solutions associated with negative values---because they secured very little genetic variation---were replaced with zeros to facilitate statistical analysis.
-We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality.
+We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality (using \texttt{Gurobi}; version 7.0.2).
 
@@
 For half of these random prioritizations, we calculated the proportion of geographic variation and neutral genetic variation they secured, and for the other half, we calculated the proportion of environmental variation and adaptive genetic variation they secured.
-The proportion of surrogate and genetic variation captured in a prioritization was calculated using the \texttt{raptr R} package.
+These calculations were performed using the \texttt{raptr R} package.
 
-We fitted two full generalized linear models (GLM) with logit links.
+We fitted two full generalized linear models (GLM) with logit link functions.
 The first model was fit to the proportion of adaptive genetic variation secured in a prioritization using the proportion of environmental variation also secured in the prioritization.
@@
 They were subjected to backwards stepwise term deletion routines to assess term significance.
+_Post-hoc_ analyses were conducted to assess trends for each species using Bonferroni corrections (using the \texttt{multcomp R} package; [@r470]).
 To assess the performance of the surrogates, we refit these models separately for each species and computed the Cragg and Uhler's pseudo-$R^2$ value (using the \texttt{pscl R} package; [@r492]).
@@
 
-We fitted generalized linear mixed-effects models (GLMMs) with logit links to  evaluate the prioritizations (using the \texttt{lme4 R} package).
+We fitted generalized linear mixed-effects models (GLMMs) with logit link functions to evaluate the prioritizations (using the \texttt{lme4 R} package).
 We fitted a full model to the proportion of genetic variation secured in a given prioritization.
@@
 The full model was subjected to a backwards stepwise term deletion routine to assess significance.
-A _post-hoc_ analysis was conducted on the minimal adequate model using Tukey contrasts with a Bonferroni correction (using the \texttt{multcomp R} package; [@r470]).
+A _post-hoc_ analysis was conducted using Tukey contrasts with Bonferroni corrections (using the \texttt{multcomp R} package).
 

### 2017-10-18 code/rmarkdown/si-preamble.tex
+ article within page pnas limits
+ assorted changes
+ fix supp info figures + tables + reorder si figures and tables + update si figures and table numbers in main tex
+ fix bug causing refs not to appear in article
minor tweaks to ms changes to analysis that were not pushed clean up documentation

@@
 % load packages
-\usepackage{amsmath,amsfonts,float,makecell,titletoc,tocloft,titlesec,natbib,pdfpages,lineno,booktabs}
+\usepackage{amsmath,amsfonts,makecell,tocloft,titling,natbib,hyperref,booktabs,caption}
 \usepackage[T1]{fontenc}
@@
 \usepackage[utf8]{inputenc}
-\usepackage[doublespacing]{setspace}
 
@@
 \renewcommand{\thetable}{S\arabic{table}}
+\captionsetup[figure]{labelfont={bf},labelformat={default},labelsep=period,name={Fig.}}
+\captionsetup[table]{labelfont={bf},labelformat={default},labelsep=period,name={Table}}

### 2017-10-18 code/rmarkdown/slides.Rmd
minor tweaks to ms changes to analysis that were not pushed clean up documentation

---
fontsize: 11pt
documentclass: article
output:
  rmarkdown::pdf_document
---

# Slides

```{r, include=FALSE}
# set wd
knitr::opts_knit$set(root.dir=normalizePath('../..'))
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo"), dpi = 500)

```

```{r,include=FALSE}
# load .rda file
checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
try(session::restore.session('data/final/results.rda'))

# download basemap
data(countriesHigh)
countries.FPLY <- countriesHigh[
  countriesHigh$ADMIN %in% c(
    'Italy', 'Switzerland', 'France', 'Austria',
    'Germany', 'Slovenia', 'Croatia', 'Hungary',
    'Monaco', 'Germany', 'Slovakia', 'Czech Republic'
  )
,] %>% spFortify
grid.FPLY <- spFortify(grid.PLY)
grid.sub.FPLY <- spFortify(grid.sub.PLY)

# define ggplot2 color wheel function
gg_color_hue <- function(n) {
  hues <- seq(15, 375, length=n+1)
  hcl(h=hues, l=65, c=100)[1:n]
}
```

```{r, echo=FALSE, fig.align='center', fig.width=6, fig.height=1.5, out.width='3.42in', height='2.5in', fig.cap="Map of the study area showing (a) the pattern of species richness, and (b) the distribution of acquisition cost, both plotted on a quantile-based color ramp. Squares denote planning units."}

# calculate species richness
grid.PLY$Species_richness <- grid.PLY@data %>%
  select(5:(4+n_distinct(spp.samples.DF$species))) %>% as.matrix() %>% rowSums()

panel.DF <- grid.sub.FPLY %>%
  summarize(lat=min(lat)+(diff(range(lat))*0.96),
    long=min(long)+(diff(range(long))*0.04))

# plot species richness
p1 <- ggplot() +
  geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
    fill='grey85', color='grey70', size=0.001) +
  geom_polygon(data=spFortify(grid.PLY), aes(x=long, y=lat,
    group=group, fill=Species_richness), color=NA) +
  scale_fill_gradientn(name='Richness',
    colors=brewer.pal(5, 'YlGn'),
    guide=guide_colorbar(ticks=element_line(color='black'), border=element_line(color='black'))
  ) +
  theme_classic() +
  theme(
    axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"), axis.title=element_blank(),
    panel.border=element_rect(color='black', fill=NA, size=1),
    panel.background=element_rect(color='black', fill='black'),
    plot.background=element_rect(color='black', fill='black'),
    legend.key=element_rect(color='black', fill='black'),
    legend.background=element_rect(color='black', fill='black'),
    legend.title=element_text(color='white'),
    legend.text=element_text(color='white')
  ) +
  coord_cartesian(
    xlim=buffered.range(grid.FPLY$long, 0),
    ylim=buffered.range(grid.FPLY$lat, 0)
  )

qn01 <- scales::rescale(quantile(grid.PLY@data$pop.density, seq(0,1,length.out=9), na.rm=TRUE))
p2 <- ggplot() +
  geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
    fill='grey85', color='grey70', size=0.001) +
  geom_polygon(data=spFortify(grid.PLY), aes(x=long, y=lat,
    group=group, fill=pop.density), color=NA) +
  scale_fill_gradientn(name='Cost',
    colors=brewer.pal(5,'YlOrRd'),
    values=qn01,
    guide=guide_colorbar(ticks=element_line(color='black'), border=element_line(color='black'))
  ) +
  theme_classic() +
  theme(
    axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"), axis.title=element_blank(),
    panel.border=element_rect(color='black', fill=NA, size=1),
    panel.background=element_rect(color='black', fill='black'),
    plot.background=element_rect(color='black', fill='black'),
    legend.key=element_rect(color='black', fill='black'),
    legend.background=element_rect(color='black', fill='black'),
    legend.title=element_text(color='white'),
    legend.text=element_text(color='white')
  ) +
  coord_cartesian(
    xlim=buffered.range(grid.FPLY$long, 0),
    ylim=buffered.range(grid.FPLY$lat, 0)
  )

grid.arrange(p1,p2,nrow=1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=6, fig.height=4.5, out.width='3.42in', out.height='2.5in', fig.cap=paste0('The proportion of adaptive (a; ',sum(!is.na(correlation.model.summary.DF$Environmental.corr.P)),' species) and neutral (b; ',sum(!is.na(correlation.model.summary.DF$Geographic.corr.P)),' species) genetic variation represented in a suite of randomly generated prioritizations for different species as a function of the surrogate variation they also secured (lines). Panels (c) and (d) show the distribution of $R^2$ values for the ',sum(!is.na(correlation.model.summary.DF$Environmental.corr.P)) + sum(!is.na(correlation.model.summary.DF$Geographic.corr.P)),' models shown in panels (a) and (b).')}
# make figure
p1 <- ggplot(aes(x=surrogate.held,y=genetic.held), data=env.surrogacy.DF) +
  xlab('Environmental variation\nsecured (%)') +
  ylab('Adaptive genetic\nvariation secured (%)') +
  theme_classic() +
  theme(axis.line.x=element_line(color='white'),
        axis.line.y=element_line(color='white'),
        axis.title=element_text(color='white'),
        axis.text=element_text(color='white'),
        axis.ticks=element_line(color='white'),
        panel.background=element_rect(color='black',fill='black'),
        plot.background=element_rect(color='black',fill='black')) +
  coord_cartesian(xlim=c(0, 1)) +
  scale_x_continuous(limits=c(0,1), labels=function(x) {x*100}) +
  scale_y_continuous(limits=c(0,1), labels=function(x) {x*100})
for (i in seq_along(env.surrogacy.LST))
  p1 <- p1 + stat_smooth(method='glm', method.args=list(family='binomial'),
    data=filter(env.surrogacy.DF, Species==unique(Species)[i]),
    fullrange=TRUE, se=FALSE, colour='white')
# p1 <- p1 + geom_text(x=-Inf, y=Inf, hjust=-0.25, vjust=1.5, label='(a)')

p2 <- ggplot(aes(x=surrogate.held,y=genetic.held), data=geo.surrogacy.DF) +
  xlab('Geographic variation\nsecured (%)') +
  ylab('Neutral genetic\nvariation secured (%)') +
  theme_classic() +
  theme(axis.line.x=element_line(color='white'),
        axis.line.y=element_line(color='white'),
        axis.title=element_text(color='white'),
        axis.text=element_text(color='white'),
        axis.ticks=element_line(color='white'),
        panel.background=element_rect(color='black',fill='black'),
        plot.background=element_rect(color='black',fill='black')) +
  coord_cartesian(xlim=c(0, 1)) +
  scale_x_continuous(limits=c(0,1), labels=function(x) {x*100}) +
  scale_y_continuous(limits=c(0,1), labels=function(x) {x*100})
for (i in seq_along(geo.surrogacy.LST))
  p2 <- p2 + stat_smooth(method='glm', method.args=list(family='binomial'),
    data=filter(geo.surrogacy.DF, Species==unique(Species)[i]),
    fullrange=TRUE, se=FALSE, colour='white')
# p2 <- p2 + geom_text(x=-Inf, y=Inf, hjust=-0.25, vjust=1.5, label='(b)')

p1.hist <- ggplot(aes(r2CU),
    data=filter(env.surrogate.spp.DF, !is.na(r2CU))) +
  geom_histogram(fill='white', bins=12) +
  ylab('\nFrequency') + xlab(expression(Environmental~variables~(R^2))) +
  theme_classic() +
  theme(axis.line.x=element_line(color='white'),
        axis.line.y=element_line(color='white'),
        axis.title=element_text(color='white'),
        axis.text=element_text(color='white'),
        axis.ticks=element_line(color='white'),
        panel.background=element_rect(color='black',fill='black'),
        plot.background=element_rect(color='black',fill='black')) +
  # geom_text(x=-Inf, y=Inf, hjust=-0.25, vjust=1.5, label='(c)') +
  scale_y_continuous(labels=function(x) {formatC(x, digits=2, format='f')}) +
  scale_x_continuous(limits=c(0, 1), labels=function(x) {formatC(x, digits=2, format='f')})

p2.hist <- ggplot(aes(r2CU),
    data=filter(geo.surrogate.spp.DF, !is.na(r2CU))) +
  geom_histogram(fill='white', bins=12) +
  ylab('\nFrequency') + xlab(expression(Geographic~variables~(R^2))) +
  theme_classic() +
  theme(axis.line.x=element_line(color='white'),
        axis.line.y=element_line(color='white'),
        axis.title=element_text(color='white'),
        axis.text=element_text(color='white'),
        axis.ticks=element_line(color='white'),
        panel.background=element_rect(color='black',fill='black'),
        plot.background=element_rect(color='black',fill='black')) +
  # geom_text(x=-Inf, y=Inf, hjust=-0.25, vjust=1.5, label='(d)') +
  scale_y_continuous(labels=function(x) {formatC(x, digits=2, format='f')}) +
  scale_x_continuous(limits=c(0, 1), labels=function(x) {formatC(x, digits=2, format='f')})

# render figure
grid.arrange(p1,p2,p1.hist,p2.hist,nrow=2)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=4.5, fig.height=3.5, out.width='4in', out.height='3.5in', fig.cap='The proportion of adaptive and neutral genetic variation secured for each species in solutions based on different targets and scenarios. Box plots show the median, and the 25th and 75th percentiles. Whiskers show the data at 1.5 times the inter-quantile range. Points show data outside the whiskers. Red lines show the target proportion of genetic variation.'}
# prepare data for plotting
scenario.SDF <- scenario.DF %>%
  group_by(Prioritisation, Context, Metric, Species) %>%
  summarize(genetic.held=mean(genetic.held, na.rm=TRUE)) %>%
  data.frame() %>%
  filter(Context == 'Multi-species\n(equal costs)')

# make plot
suppressWarnings({
  ggplot(aes(x=Prioritisation,y=genetic.held),
    data=scenario.SDF) +
    geom_boxplot(position=position_dodge(0.9), color='white', fill='black') +
    scale_fill_manual(name='Prioritization',
      values=c('grey80','grey50','grey20')) +
    geom_abline(slope=0, color='red', linetype='dashed',
      intercept=raptr.params.LST[[MODE]]$scenario.analysis$genetic.target) +
    ylab('Genetic variation secured (%)') +
    xlab('') +
    scale_y_continuous(breaks=c(0,0.25,0.5,0.75,1)) +
    theme_classic() +
    theme(
         axis.line.x=element_line(color='white'),
         axis.line.y=element_line(color='white'),
         axis.title=element_text(color='white'),
         axis.ticks=element_line(color='white'),
         panel.background=element_rect(color='black',fill='black'),
         plot.background=element_rect(color='black',fill='black'),
         strip.background = element_rect(fill='grey20'),
         strip.text = element_text(color='white', size=9),
         panel.border = element_rect(color='white', fill=NA),
         axis.text.x=element_text(color='white',angle=45, hjust=1, vjust=1,
                                  size=8),
         axis.text.y=element_text(color='white',size=8)) +
    facet_grid(Context ~ Metric)
})
```

```{r, echo=FALSE, fig.width=9.5, fig.height=10, out.width='7.5in', out.height='8.5in', fig.cap=paste0('Species distributions. Squares represent planning units. For a given species, planning units that it was found in are shown in bright blue.')}
## plot map of species distributions
# fortify data
spp.grid.FPLY <- ldply(unique(spp.samples.DF$species), function(x) {
    z <- grid.FPLY[,c('long', 'lat', 'group', x),drop=FALSE]
    names(z)[4] <- 'presence'
    z$species <- gsub('\\_', ' ', x)
    return(z)
  }
) %>% mutate(presence=factor(c('absent', 'present')[presence+1]))

# plot species data
ggplot() +
  geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
    fill='grey85', color='grey70', size=0.001) +
  geom_polygon(data=spp.grid.FPLY, aes(x=long, y=lat,
    group=group, fill=presence), alpha=0.8, color='grey10') +
  theme_classic() +
  theme(
    axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
    strip.background = element_rect(fill='grey20'),
    strip.text = element_text(color='white', face='italic'),
    axis.title=element_blank(),
    panel.background=element_rect(color='black',fill='black'),
    plot.background=element_rect(color='black',fill='black'),
    legend.text=element_text(color='white'),
    legend.title=element_text(color='white'),
    legend.key=element_rect(fill='black'),
    legend.background=element_rect(fill='black')
  ) +
    scale_fill_manual(name='Occupancy', values=c('absent'='#FC8D59', 'present'='#91CF60')) +
  coord_cartesian(
    xlim=buffered.range(grid.FPLY$long, 0),
    ylim=buffered.range(grid.FPLY$lat, 0)
  ) +
  facet_wrap(~ species, ncol=4)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=(5*0.9), fig.height=(6.5*0.8), out.width='3.42in', height='3.9in', fig.cap='Planning unit selection frequencies. Squares represent planning units, and their colors indicate their selection frequency. Each column of panels shows solutions generated using different targets. Each row of panels shows solutions generated under different scenarios. Note that the selection frequencies for the single-species prioritizations are based on their frequency among prioritizations generated for each species. The selection frequencies for the multi-species prioritizations are each based on a single optimal prioritization for all species and so are binary.'}
# prepare data for plotting
grid.sub.FPLY <- grid.sub.PLY
for (i in seq_along(single.spp.prioritisations[[1]]))
  grid.sub.FPLY@data[[paste0('Single-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations, function(x){colMeans(x[[i]]@results@selections)}))
for (i in seq_along(single.spp.prioritisations.with.cost[[1]]))
  grid.sub.FPLY@data[[paste0('Single-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations.with.cost, function(x){colMeans(x[[i]]@results@selections)}))
for (i in seq_along(multi.spp.prioritisations))
  grid.sub.FPLY@data[[paste0('Multi-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations[[i]]@results@selections)
for (i in seq_along(multi.spp.prioritisations.with.cost))
  grid.sub.FPLY@data[[paste0('Multi-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations.with.cost[[i]]@results@selections)
grid.sub.FPLY <- spFortify(grid.sub.FPLY)
grid.sub.FPLY <- grid.sub.FPLY %>% gather(
    Combination,
    `Selection.Frequency`,
    `Single-species (equal costs).Amount`:`Multi-species (opportunity costs).Genetic`
  ) %>% mutate(
    Context=sapply(strsplit(as.character(Combination), '.', fixed=TRUE), `[[`, 1),
    Prioritization=sapply(strsplit(as.character(Combination), '.', fixed=TRUE), `[[`, 2)
  ) %>% mutate(
    Prioritization=revalue(Prioritization, c('Amount'='Amount targets', 'Surrogate'='Amount & surrogate\ntargets', 'Genetic'='Amount & genetic\ntargets')),
    Prioritization=factor(as.character(Prioritization), levels=c('Amount targets', 'Amount & surrogate\ntargets', 'Amount & genetic\ntargets')),
    Context=factor(
      gsub(' (', '\n(', Context, fixed=TRUE),
      levels=c("Single-species\n(equal costs)", "Single-species\n(opportunity costs)",
        "Multi-species\n(equal costs)", "Multi-species\n(opportunity costs)"))
  )

# make maps
ggplot() +
  geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
    fill='grey85', color='grey70', size=0.001) +
  geom_polygon(data=grid.sub.FPLY, aes_string(x='long', y='lat',
    group='group', fill='Selection.Frequency'),
    alpha=0.8, color=NA) +
  theme_classic() +
  theme(axis.ticks=element_blank(), axis.text=element_blank(),
    plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(),
    strip.background = element_rect(fill='grey20'),
    strip.text = element_text(color='white', size=8),
    legend.position='bottom', legend.key.width=unit(1.8, 'cm'),
    legend.title=element_text(vjust=-5),
    legend.margin=unit(c(0.1), 'cm'),
    axis.title=element_blank(),
    legend.title=element_text(color='white'),
    legend.text=element_text(color='white'),
    legend.key=element_rect(fill='black'),
    legend.background=element_rect(fill='black'),
    panel.background=element_rect(fill='black'),
    plot.background=element_rect(fill='black')
  ) +
  coord_cartesian(
    xlim=buffered.range(grid.sub.FPLY$long, 0),
    ylim=buffered.range(grid.sub.FPLY$lat, 0)
  ) +
  facet_grid(Context ~ Prioritization) +
  scale_fill_gradient(name='Selection Frequency (%)',
    guide=guide_colorbar(title.position='bottom', title.hjust=0.5,
      ticks=element_line(color='black'), border=element_line(color='black'))
  )
```

```{r, include = FALSE}
plot.spp.spaces <- function(i) {
  ## prepare data
  # get vector of adaptive and neutral column names
  adapt.col <- grep(
    paste0(unique(spp.samples.DF$species)[i], '_adaptive_d'),
    names(grid.FPLY), value=TRUE, fixed=TRUE
  )
  neutral.col <- grep(
    paste0(unique(spp.samples.DF$species)[i], '_neutral_d'),
    names(grid.FPLY), value=TRUE, fixed=TRUE
  )
  env.col <- grep(
    paste0(unique(spp.samples.DF$species)[i], '_env_d'),
    names(grid.FPLY), value=TRUE, fixed=TRUE
  )
  # skip species if not used in analysis
  curr.FPLY <- grid.FPLY %>%
    select(one_of(adapt.col, neutral.col, env.col), group, long, lat) %>%
    mutate_each(funs(unname), one_of(adapt.col, neutral.col, env.col)) %>%
    filter(!is.na(.[[1]])) %>%
    gather(d, value, contains('adaptive'), contains('neutral'), contains('env')) %>%
    mutate(d=as.character(d)) %>%
    mutate(d=gsub(paste0(unique(spp.samples.DF$species)[i], '_'), '', d, fixed=TRUE)) %>%
    mutate(d=gsub('_', ' ', d, fixed=TRUE)) %>%
    mutate(d=gsub('env', 'environmental', d, fixed=TRUE))
  for (j in seq_len(max(length(adapt.col),length(neutral.col),length(env.col))))
    curr.FPLY$d <- gsub(paste0('d', as.character(j)), paste0('(', j, ')'), curr.FPLY$d, fixed=TRUE)
  # make plot
  curr.GROBS <- llply(unique(curr.FPLY$d), function(x) {
    ggplot() +
      geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
        fill='grey85', color='grey70', size=0.001) +
      geom_polygon(data=filter(curr.FPLY, d==x), aes_string(x='long', y='lat',
        group='group', fill='value'), color='grey10') +
      theme_classic() +
      theme(
        axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"), axis.title=element_blank(),
        panel.border=element_rect(color='black', fill=NA, size=1),
        strip.background = element_rect(fill='grey20'),
        strip.text = element_text(color='white', size=20),
        legend.title=element_text(color='white'),
        legend.text=element_text(color='white'),
        legend.key=element_rect(fill='black'),
        legend.background=element_rect(fill='black'),
        panel.background=element_rect(fill='black'),
        plot.background=element_rect(fill='black')
      ) +
      coord_cartesian(
        xlim=buffered.range(grid.FPLY$long, 0),
        ylim=buffered.range(grid.FPLY$lat, 0)
      ) +
      scale_fill_gradientn(name='',
        colors=brewer.pal(9, 'RdYlBu'),
        guide=guide_colorbar(ticks=element_line(color='black'),
          border=element_line(color='black'), barheight=unit(3.0, 'cm')),
        labels=function(x) {formatC(x, format='f', digits=2)}
      ) + facet_wrap(~d)
  })
  do.call(grid.arrange, append(curr.GROBS, list(ncol=3)))
  invisible()
}

calc.fig.height <- function(i) {
  # init
  adapt.col <- paste0(unique(spp.samples.DF$species)[i], '_adaptive_d1')
  neutral.col <- paste0(unique(spp.samples.DF$species)[i], '_neutral_d1')
  env.col <- grepl(paste0(unique(spp.samples.DF$species)[i], '_env_d'), names(grid.FPLY), fixed=TRUE) %>%
    sum()

  n.plots <- 0
  # count adaptive plot
  if (adapt.col %in% names(grid.FPLY))
    n.plots <- n.plots + spp.nmds.LST[[i]][['adaptive']]$ndim
  # count neutral plot
  if (neutral.col %in% names(grid.FPLY))
    n.plots <- n.plots + spp.nmds.LST[[i]][['neutral']]$ndim
  # count env plot
  n.plots <- n.plots + env.col
  # exports
  if (n.plots == 0) return(1)
  if (n.plots >= 3) return((((n.plots-1) %/% 3)+1) * 2)
  return(1.85)
}

calc.fig.width <- function(i) {
  # init
  adapt.col <- paste0(unique(spp.samples.DF$species)[i], '_adaptive_d1')
  neutral.col <- paste0(unique(spp.samples.DF$species)[i], '_neutral_d1')
  env.col <- grepl(paste0(unique(spp.samples.DF$species)[i], '_env_d'), names(grid.FPLY), fixed=TRUE) %>%
    sum()
  n.plots <- 0
  # count adaptive plot
  if (adapt.col %in% names(grid.FPLY))
    n.plots <- n.plots + spp.nmds.LST[[i]][['adaptive']]$ndim
  # count neutral plot
  if (neutral.col %in% names(grid.FPLY))
    n.plots <- n.plots + spp.nmds.LST[[i]][['neutral']]$ndim
  # count env plot
  n.plots <- n.plots + env.col
  # exports
  if (n.plots == 0) return(1)
  if (n.plots >= 3) return(10)
  return(4 * n.plots)
}
is.speciesPlottable <- function(i) {
  if (i > general.params.LST[[MODE]]$n.spp)
    return(FALSE)
  return(!is.na(unique(spp.samples.DF$species)[i]))
}
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=4.5, fig.height=4.5, fig.cap='Qvalues from PCadapt for \textit{Arabis alpina}'}
# prepare data
alpha_value <- pcadapt.params.LST[[MODE]]$alpha.level
outlier_data <- data.frame(qvalue = spp.pcadapt.LST[[2]]$qvalues$qvalues) %>%
                mutate(id = factor(seq_len(nrow(.)))) %>%
                mutate(type = case_when(.$qvalue < alpha_value ~ 'adaptive',
                                        .$qvalue >= alpha_value ~ 'neutral'))

# make plot
ggplot(data = outlier_data, aes(x = id, y = qvalue, fill = type)) +
  geom_bar(stat='identity', color = 'black') +
  geom_hline(yintercept = alpha_value, color = "red", linetype = "dashed") +
  xlab('Loci') +
  ylab('Q-value') +
  scale_fill_manual(values = c('adaptive' = 'red', 'neutral' = 'white')) +
  theme_classic() +
  theme(axis.line.x=element_line(color='white'),
        axis.line.y=element_line(color='white'),
        axis.title=element_text(color='white'),
        axis.text=element_text(color='white'),
        axis.ticks=element_line(color='white'),
        legend.text=element_text(color='white'),
        legend.title=element_text(color='white'),
        legend.background=element_rect(color='black', fill = 'black'),
        legend.key=element_rect(color='black', fill = 'black'),
        panel.background=element_rect(color='black',fill='black'),
        plot.background=element_rect(color='black',fill='black'))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=4.5, fig.height=4.5, fig.cap='Qvalues from BayeScan for \textit{Arabis alpina}'}
# prepare data
outlier_data <- spp.BayeScan.LST[[2]]$bayescan@results@summary %>%
                mutate(id = factor(seq_len(nrow(.))))
# make plot
ggplot(data = outlier_data, aes(x = id, y = mean_qval, fill = type)) +
  geom_bar(stat = 'identity', color = 'black') +
  xlab('Loci') +
  ylab('Q-value') +
  theme_classic() +
  scale_fill_manual(values = c('adaptive' = 'red', 'neutral' = 'white')) +
  theme(axis.line.x=element_line(color='white'),
        axis.line.y=element_line(color='white'),
        axis.title=element_text(color='white'),
        axis.text=element_text(color='white'),
        axis.ticks=element_line(color='white'),
        legend.text=element_text(color='white'),
        legend.title=element_text(color='white'),
        legend.background=element_rect(color='black', fill = 'black'),
        legend.key=element_rect(color='black', fill = 'black'),
        panel.background=element_rect(color='black',fill='black'),
        plot.background=element_rect(color='black',fill='black'))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=4.5, fig.height=4.5, fig.cap='NMDS for neutral alleles for \textit{Arabis alpina}'}
# prepare data
nmds <- as.data.frame(spp.nmds.LST[[2]]$neutral$points)
# make plot
ggplot(data = nmds, aes(x = MDS1, y = MDS2)) +
  geom_point(color = 'white') +
  xlab('NMDS 1') +
  ylab('NMDS 2') +
  theme_classic() +
  theme(axis.line.x=element_line(color='white'),
        axis.line.y=element_line(color='white'),
        axis.title=element_text(color='white'),
        axis.text=element_text(color='white'),
        axis.ticks=element_line(color='white'),
        panel.background=element_rect(color='black',fill='black'),
        plot.background=element_rect(color='black',fill='black'))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=4.5, fig.height=4.5, fig.cap='NMDS for adaptive alleles for \textit{Arabis alpina}'}
# prepare data
nmds <- as.data.frame(spp.nmds.LST[[2]]$adaptive$points)
# make plot
ggplot(data = nmds, aes(x = MDS1, y = MDS2)) +
  geom_point(color = 'white') +
  xlab('NMDS 1') +
  ylab('NMDS 2') +
  theme_classic() +
  theme(axis.line.x=element_line(color='white'),
        axis.line.y=element_line(color='white'),
        axis.title=element_text(color='white'),
        axis.text=element_text(color='white'),
        axis.ticks=element_line(color='white'),
        panel.background=element_rect(color='black',fill='black'),
        plot.background=element_rect(color='black',fill='black'))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=4.5, fig.height=4.5, fig.cap='Population memberships for \textit{Arabis alpina}'}
# prepare data
pop.probs.DF <- llply(
  seq_along(unique(spp.samples.DF$species)),
  function(i) {
    curr.pts <- filter(
      spp.sample.PTS@data,
      species==unique(spp.samples.DF$species)[i]
    ) %>% mutate(
      Individual=seq_along(species)
    )
    curr.probs <- structurer:::sample.membership(spp.StructureCollection.LST[[i]]) %>%
      as.data.frame() %>%
      mutate(
        species=unique(spp.samples.DF$species)[i],
        Individual=seq_along(species)) %>%
      gather(Population,Probability,-species,-Individual) %>%
      mutate(Population=as.numeric(gsub('V', '', Population, fixed=TRUE))) %>%
      left_join(curr.pts, by=c('species','Individual')) %>%
      mutate(species=gsub('_', ' ', species, fixed=TRUE)) %>%
      arrange(species,Individual,Population)
    return(curr.probs)
  }
) %>% do.call(what=rbind)
# create legend colors
legend.CHR <- `names<-`(
  gg_color_hue(n_distinct(pop.probs.DF$Population)),
  as.character(unique(pop.probs.DF$Population))
)
plot.spp.pop.probs <- function(i) {
  # init spp.samples
  prop <- 0.85
  curr.spp.DF <- filter(pop.probs.DF, species==unique(pop.probs.DF$species)[i])
  # create base map grob
  curr.GROB <- ggplot(data=curr.spp.DF) +
    geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
      fill='grey85', color='grey75') +
    geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat',
      group='group'), fill='grey70', color='grey10') +
    guides(fill=guide_legend(title=''), color=guide_legend(title='')) +
    theme_classic() +
    theme(
      axis.title=element_blank(), axis.text=element_blank(),
      axis.line=element_blank(), axis.ticks=element_blank(),
      plot.background=element_rect(fill='black'),
      axis.ticks.length = unit(0, "cm"),
      legend.position = "none", panel.margin = unit(0, "lines"),
      plot.margin = unit(c(0, 0, 0, 0), "lines"),
      strip.background = element_rect(fill='grey20'),
      strip.text = element_text(color='white', face='italic', size=20),
      legend.text= element_text(size=20)
    ) +
    coord_cartesian(
      xlim=buffered.range(grid.FPLY$long, 0),
      ylim=buffered.range(grid.FPLY$lat, 0)
    ) +
    scale_fill_manual(
      name='',
      values=legend.CHR[seq_len(n_distinct(curr.spp.DF$Population))]
    )
  # create header grob
  curr.GROBH <- ggplot() +
      geom_text(aes(x=0,y=0,label=species),
        data=curr.spp.DF[1,,drop=FALSE],
        color='white', fontface='italic', size=4) +
    theme_classic() +
    theme(
      axis.title=element_blank(), axis.text=element_blank(),
      axis.line=element_blank(), axis.ticks=element_blank(),
      axis.ticks.length = unit(0, "cm"),
      legend.position = "none", panel.margin = unit(0, "lines"),
      plot.margin = unit(c(0, 0, 0, 0), "lines"),
      panel.background=element_rect(fill='grey20')
    ) +
    coord_cartesian(xlim=c(-1,1), ylim=c(-1,1))
  # create bbox object
  bbox <- c(ggplot_build(curr.GROB)$panel$ranges[[1]]$x.range,
    ggplot_build(curr.GROB)$panel$ranges[[1]]$y.range)
    names(bbox) <- c('ll.lon', 'ur.lon', 'll.lat', 'ur.lat')
  # render grob
  grid.newpage()
  pushViewport(viewport(layout=grid.layout(2,1, heights=unit(c(1-prop, prop), "npc"))))
  print(curr.GROBH, vp=viewport(layout.pos.row=1, layout.pos.col=1))
  print(curr.GROB, vp=viewport(layout.pos.row=2, layout.pos.col=1))
  # add in pies
  for (j in unique(curr.spp.DF$cell)) {
    # init
    curr.cell.DF <- filter(curr.spp.DF, cell==j) %>%
      group_by(cell, Population) %>%
      summarize(
        grid.longitude=mean(grid.longitude),
        grid.latitude=mean(grid.latitude),
        Probability=mean(Probability)
      ) %>% data.frame()
    curr.lon <- (curr.cell.DF$grid.longitude[1]-bbox['ll.lon'])/abs(bbox['ur.lon']-bbox['ll.lon'])
    curr.lat <- ((curr.cell.DF$grid.latitude[1]-bbox['ll.lat'])/abs(bbox['ur.lat']-bbox['ll.lat']))*prop
    # create pie chart for individual
    curr.pie <- ggplot(curr.cell.DF, aes(x="", y=Probability, fill=factor(Population))) +
      geom_bar(width=0.75, color='grey20', stat='identity') +
      coord_polar(theta='y') +
      theme(
        line = element_blank(), rect = element_blank(),
        text = element_blank(), axis.ticks.length = unit(0, "cm"),
        legend.position = "none", panel.margin = unit(0, "lines"),
        plot.margin = unit(c(0, 0, 0, 0), "lines"), complete = TRUE
      ) +
      scale_fill_manual(
        name='',
        values=legend.CHR[seq_len(n_distinct(curr.spp.DF$Population))]
      )
    # render pie
    print(curr.pie, vp=viewport(x=curr.lon, y=curr.lat, width=0.075, height=0.075))
  }
}
plot.spp.pop.probs(2)
```

### 2017-10-18 code/rmarkdown/supporting_information.Rmd
minor tweaks to ms changes to analysis that were not pushed clean up documentation

@@
 # Supporting Information
-All data and code is stored in an online repository (\url{www.github.com/jeffreyhanson/genetic-surrogates}) to replicate analyses in this study. See the \textit{README} file for instructions.
+To replicate this study, all data and code are stored in an online repository (\url{www.github.com/jeffreyhanson/genetic-surrogates}) and are available under the digital object identifier (DOI) \texttt{10.5281/zenodo.843625}.
+See the \textit{README} file for instructions.
 
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=(2.5*4), fig.cap='The relationship between the proportion of adaptive genetic variation secured in a prioritization and the proportion of environmental variation it also secures.
+```{r, include=FALSE}
+calc.fig.height31 <- function(x) {
+  return((((x-1) %/% 4)+1)*2.0)
+}
+```
+
+```{r, include=FALSE}
+calc.fig.height32 <- function(x) {
+  return((((x-1) %/% 4)+1)*1.75)
+}
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=11, fig.height=calc.fig.height31(n_distinct(env.surrogacy.DF$Species)), fig.cap='The relationship between the proportion of adaptive genetic variation secured in a prioritization and the proportion of environmental variation it also secures.
 Points represent values associated with randomly generated prioritizations.
-Blue lines indicate average trends computed using a generalised linear model with a logit link.
-Each panel shows data for a different species.'}
+Lines show trends fit using generalised linear models with a logit link functions.
+Each panel corresponds to a different species.'}
 # format data
-curr.DF <- env.surrogacy.DF %>% filter(!is.na(genetic.held)) %>% mutate(Species=gsub('_', ' ', Species))
-curr.DF2 <- env.surrogate.spp.DF %>% filter(!is.na(r2CU)) %>% mutate(Species=gsub('_', ' ', Species))
-curr.DF2$R2_repr <- round(curr.DF2$r2CU,2)
-curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
-curr.DF2$surrogate.held <- 0
-curr.DF2$genetic.held <- 0.9
+curr.DF <- env.surrogacy.DF %>% 
+  filter(!is.na(genetic.held)) %>% 
+  mutate(Species=gsub('_', '~', Species)) %>%
+  left_join(env.surrogate.spp.DF %>% 
+            filter(!is.na(r2CU)) %>% 
+            mutate(Species=gsub('_', '~', Species)), by='Species')
+curr.DF$R2_repr <- round(curr.DF$r2CU,2)
+curr.DF$R2_repr <- sapply(curr.DF$R2_repr, function(x) {
+  as.character(as.expression(substitute(italic(R)^2~"="~r2, list(r2=x))))})
+curr.DF$Species_repr <- paste0('atop(italic(', curr.DF$Species, '),(', curr.DF$R2_repr, '))')
 
@@
     theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=12),
-      panel.margin.x=unit(1, 'lines')) +
+      panel.margin.x=unit(1, 'lines'), legend.position='bottom') +
     xlab('Environmental variation secured (%)') +
     ylab('Adaptive genetic variation secured (%)') +
-    geom_hex() +
+    stat_binhex(aes(fill = log(..count..))) +
+    scale_fill_continuous(name='log(count)') +
     suppressWarnings(stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
-      se=FALSE, fill='blue', alpha=0.8)) +
-    geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=5, color='red') +
-    facet_wrap(~ Species, ncol=4)
-)
-```
-
-```{r, include=FALSE}
-calc.fig.height3 <- function(x) {
-  return((((x-1) %/% 4)+1)*1.5)
-}
-
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=calc.fig.height3(n_distinct(geo.surrogacy.DF$Species)), fig.cap='The relationship between the proportion of neutral genetic variation secured in a prioritization and the proportion of geographic variation it also secures.
+      se=FALSE, color='#ffb100')) +
+    facet_wrap(~ Species_repr, ncol=4, labeller = label_parsed)
+)
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=11, fig.height=calc.fig.height32(n_distinct(geo.surrogacy.DF$Species)), fig.cap='The relationship between the proportion of neutral genetic variation secured in a prioritization and the proportion of geographic variation it also secures.
 Conventions are detailed in Figure S28.'}
 # format data
-curr.DF <- geo.surrogacy.DF %>% filter(!is.na(genetic.held)) %>% mutate(Species=gsub('_', ' ', Species))
-curr.DF2 <- geo.surrogate.spp.DF %>% filter(!is.na(r2CU)) %>% mutate(Species=gsub('_', ' ', Species))
-curr.DF2$R2_repr <- round(curr.DF2$r2CU,2)
-curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
-curr.DF2$surrogate.held <- 0
-curr.DF2$genetic.held <- 0.9
+curr.DF <- geo.surrogacy.DF %>% 
+  filter(!is.na(genetic.held)) %>% 
+  mutate(Species=gsub('_', '~', Species)) %>%
+  left_join(geo.surrogate.spp.DF %>% 
+            filter(!is.na(r2CU)) %>% 
+            mutate(Species=gsub('_', '~', Species)), by='Species')
+curr.DF$R2_repr <- round(curr.DF$r2CU,2)
+curr.DF$R2_repr <- sapply(curr.DF$R2_repr, function(x) {
+  as.character(as.expression(substitute(italic(R)^2~"="~r2, list(r2=x))))})
+curr.DF$Species_repr <- paste0('atop(italic(', curr.DF$Species, '),(', curr.DF$R2_repr, '))')
 
@@
     theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=12),
-      panel.margin.x=unit(1, 'lines')) +
+      panel.margin.x=unit(1, 'lines'), legend.position='bottom') +
     xlab('Geographic variation secured (%)') +
     ylab('Neutral genetic variation secured (%)') +
-    geom_hex() +
+    stat_binhex(aes(fill = log(..count..))) +
+    scale_fill_continuous(name='log(count)') +
     suppressWarnings(stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
-      se=FALSE, fill='blue', alpha=0.8)) +
-    geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=5, color='red') +
-    facet_wrap(~ Species, ncol=4)
+      se=FALSE, color='#ffb100')) +
+    facet_wrap(~ Species_repr, ncol=4, labeller = label_parsed)
 )
@@
         "Multi-species\n(equal costs)", "Multi-species\n(opportunity costs)"))
-  )
+  ) %>% mutate(Context = gsub('opportunity', 'acquisition', Context, fixed=TRUE))
 
@@
 ```{r, echo=FALSE}
-write.table(loci_classification.DF, 'article/Table_S1.csv', sep=',', row.names=FALSE)
+loci_classification.DF %>%
+  mutate(Species = gsub('\\textit{', '', Species, fixed=TRUE)) %>%
+  mutate(Species = gsub('}', '', Species, fixed=TRUE)) %>%
+  write.table('article/Table_S1.csv', sep=',', row.names=FALSE)
 ```

### 2017-10-18 code/R/analysis/02-surrogate-data.R

@@ -37,8 +37,8 @@ centroids.DF <- grid.PPLY %>%
 	as.data.frame() %>%
 	`names<-`(paste0('geo_d',1:2))
 # zscore it
-centroids.DF <- sweep(centroids.DF, 2, MARGIN=2, FUN='-', colMeans(centroids.DF))
-centroids.DF <- sweep(centroids.DF, 2, MARGIN=2, FUN='/', apply(centroids.DF, 2, sd))
+centroids.DF <- sweep(centroids.DF, MARGIN=2, FUN='-', colMeans(centroids.DF))
+centroids.DF <- sweep(centroids.DF, MARGIN=2, FUN='/', apply(centroids.DF, 2, sd))
 # append to main data.frame
 grid.DF <- cbind(grid.DF, centroids.DF)
 

### 2017-10-18 code/R/analysis/03-population-clustering-analysis.R

@@ -21,7 +21,7 @@ spp.StructureCollection.LST <- llply(
 				spp.StructureData.LST[[i]], MAXPOPS = curr.maxpops, NUMRUNS = structure.params.LST[[MODE]]$numruns, 
 				BURNIN = structure.params.LST[[MODE]]$burnin, NUMREPS = structure.params.LST[[MODE]]$numreps, NOADMIX = structure.params.LST[[MODE]]$noadmix, 
 				FREQSCORR = structure.params.LST[[MODE]]$freqscorr, ADMBURNIN = structure.params.LST[[MODE]]$admburnin, UPDATEFREQ=structure.params.LST[[MODE]]$updatefreq,
-				M = clumpp.params.LST[[MODE]]$M, W = clumpp.params.LST[[MODE]]$W, S = clumpp.params.LST[[MODE]]$S,  REPEATS = clumpp.params.LST[[MODE]]$repeats, 
+				M = clumpp.params.LST[[MODE]]$M, W = clumpp.params.LST[[MODE]]$W, S = clumpp.params.LST[[MODE]]$S, REPEATS = clumpp.params.LST[[MODE]]$repeats, 
 				dir = curr.spp.dir, clean = FALSE, verbose = FALSE, threads=general.params.LST[[MODE]]$threads 
 			)
 		)

### 2017-10-20 code/rmarkdown/article.Rmd
last minute typo fixes

@@
 The average proportion of genetic variation secured by prioritizations varied under different scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.2.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.2.DF$p.value[[2]])`).
-Specifically, prioritizations generated under single-species scenario with equal costs (mean `r round(mean(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held*100),2)`) secured less genetic variation than those under the single species with acquisition costs (mean `r round(mean(filter(scenario.sub.DF,Context=='Single-species\n(opportunity costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Single-species\n(opportunity costs)')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$) or the multi-species with cost scenarios (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[3]])`$).
+Specifically, prioritizations generated under single-species scenario with equal costs (mean `r round(mean(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held*100),2)`) secured less genetic variation than those under the single species with acquisition costs (mean `r round(mean(filter(scenario.sub.DF,Context=='Single-species\n(opportunity costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Single-species\n(opportunity costs)')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$) or the multi-species with acquisition cost scenarios (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[3]])`$).
 These results suggest the proportion of genetic variation secured in a prioritization may be sensitive to acquisition cost.
@@
 For example, some species can maintain relatively high connectivity between distant populations (e.g.
-wind dispersing plants; [@r451]), and in some places can be disrupted by landscape features (e.g.
+wind dispersing plants; [@r451]), and in some places, gene flow can be disrupted by landscape features (e.g.
 anthropogenically modified land; [@r497]).
@@
 <!-- adaptive genetic variation secured in prioritizations-->
-Our results suggest that conservation planners can secure a representative sample of intra-specific adaptive genetic variation using conventional reserve selections without needing surrogate or genetic data (Fig.
+Our results suggest that conservation planners can secure a representative sample of intra-specific adaptive genetic variation using conventional reserve selection methods without needing surrogate or genetic data (Fig.
 2).
@@
 
-The first outlier detection method involved fitting multinominal-Dirichlet models implemented in \texttt{BayeScan} (version 2.1; [@r456]).
+The first outlier detection method involved fitting multinomial-Dirichlet models implemented in \texttt{BayeScan} (version 2.1; [@r456]).
 We adopted a similar methodology to Bothwell _et al._ ([-@r455]) and applied it to each of the species separately.
@@
 For a given species, each planning unit in which it occurred was associated with a multi-dimensional point in the species' adaptive genetic space (if adaptive genetic variation was detected) and another multi-dimensional point in the species' neutral genetic space.
-Planning units that were closer together in these genetic spaces were occupied by individuals with more comparable AFLP polymorphisms.
+Planning units that were closer together in these genetic spaces were occupied by individuals with more comparable AFLP data.
 By spreading out conservation effort across these genetic spaces, and in turn selecting planning units occupied by individuals with increasingly different polymorphisms, prioritizations can secure more genetic variation.
@@
 
-We fitted maximum likelihood population-effects (MLPE) models using maximum likelihood to investigate correlations between the dissimilarity models (using the \texttt{lme4 R} package; [@r469]).
+We fitted maximum likelihood population-effects (MLPE) models using maximum likelihood to investigate correlations between the dissimilarity matrices (using the \texttt{lme4 R} package; [@r469]).
 These models use random effects to accommodate the structure of dissimilarity matrices.

### 2017-11-22 code/rmarkdown/article.Rmd
fix citations

@@
 Here we determine whether environmental and geographic surrogates capture adaptive and neutral genetic variation respectively in the context of conservation planning.
-We use distribution and genomic data for 27 alpine plant species in the European Alps that was obtained by the IntraBioDiv project ([@r451]).
+We use distribution and genomic data for 27 alpine plant species in the European Alps that was obtained by the IntraBioDiv project ([@r486]).
 The data were collected following a gridded sampling scheme which we adopted as planning units for this study.
@@
 
-We used data for 27 alpine plant species in the European Alps collected by the IntraBioDiv project ([@r451]; Fig.
+We used data for 27 alpine plant species in the European Alps collected by the IntraBioDiv project ([@r486]; Fig.
 3a).
